### 2022-05-05 Supplemental-Material-A-Annotated-Analysis.Rmd
initial commit

---
title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
subtitle: "Supplemental Material for 'Qualitative Goals'"
author:
- ██████ ██████████^1^
- ██████████ ████████^1^
- █████ █████████^1^
- █████ ██ █████^1^
- ███ ███████^1^
- ^1^██████████ ██ █████████, Department of Psychology
- "Author Information:"
- "Correspondence concerning this article should be addressed to ██████ ██████████, ██████████ ██ ██████████, █████████ ██ ██████████, █████ ██████████ ███, ████ ██ █████████ ████████████████. E-mail: █████████████████"
- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
- 'The data repository for this manuscript is available at <a href="https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970" target="_blank">osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970</a>'
- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    fig_caption: yes
    md_extensions: +footnotes
    code_folding: hide
    mathjax: default
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: false
    css: style.css
    includes:
      in_header: "_includes/head-custom-rmd.html" 
editor_options:
  chunk_output_type: console
#bibliography: references.bib
csl: apa.csl
header-includes:
   - \usepackage{amsmath, nccmath}
---

<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
.table {
  margin-left:auto; 
  margin-right:auto;
}
</style>

```{r setup, include=FALSE}
# R Studio Clean-Up
cat("\014") # clear console
# rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
gc() # garbage collector

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  "brms",
  "psych",
  "ggplot2",
  "ggthemes",
  "haven",
  "RColorBrewer",
  "plotly",
  "gridExtra",
  "ggpattern",
  "lme4",
  "nlme",
  "jtools",
  "gtsummary",
  "sessioninfo",
  "tibble",
  "pander",
  "devtools",
  "mada",
  "data.table",
  "plyr",
  "dplyr",
  "tidyr",
  "Hmisc",
  "kableExtra",
  "papaja",
  "stringr",
  "stringi",
  "reshape2",
  "lubridate",
  "purrr",
  "metafor",
  "dygraphs"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

# Load Custom Packages
source("./scripts/functions/fun.panel.R")
source("./scripts/functions/themes.R")
source("./scripts/functions/binaryCor.R")
source("./scripts/functions/MlCorMat.R")
source("./scripts/functions/MlTbl.R")
source("./scripts/functions/metaLmer.R")

# Markdown Options
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
knitr::opts_knit$get("root.dir") # check working directory
options(
  scipen = 999,
  digits = 4,
  width = 400
) # removes scientific quotation
# knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings
knitr::knit_hooks$set(
  error = function(x, options) {
    paste('\n\n<div class="alert alert-danger">',
      gsub("##", "\n", gsub("^##\ Error", "**Error**", x)),
      "</div>",
      sep = "\n"
    )
  },
  warning = function(x, options) {
    paste('\n\n<div class="alert alert-warning">',
      gsub("##", "\n", gsub("^##\ Warning:", "**Warning**", x)),
      "</div>",
      sep = "\n"
    )
  },
  message = function(x, options) {
    paste('\n\n<div class="alert alert-info">',
      gsub("##", "\n", x),
      "</div>",
      sep = "\n"
    )
  }
)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "Figures/",
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

<br/>

<i class="fas fa-exclamation-circle"></i> Note. Boxplots display the interquartile range (IQR, center box), and the whiskers extend 1.5*IQR from the lower and upper hinge. The white point indicates the mean and the white center line indicates the median.

<br/>

# **Data Preparation**

In an initial preparatory step, we import the data into the R project environment and prepare the variables for further processing and later analyses.

## Data Import

The data were collected using two different survey tools. For the study with sojourners (Study 1: worker) we used the survey platform Qualtrics XM, whereas the studies with international students (Study 2: student), and the international medical professionals (Study 3: medical) were conducted using the survey framework FormR. This means that the datasets had inconsistent file formats and naming conventions. For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '_data/S1_Workers/processed/cleaned_' and for the raw data files see '_data/S1_Workers/raw_'). For the two other studies, we import the raw csv files from their respective folders.

```{r formrImport}
# workers
# initial data cleaning was done in SPSS (syntax files are available in "")
dtWorker <- list(
  raw.pre = read_spss("data/S1_Workers/processed/cleaned/MT - Pre-Measure - 06-15-2018.sav"),
  raw.post = read_spss("data/S1_Workers/processed/cleaned/MT - Post-Measure - 06-15-2018.sav"),
  raw.morning = read_spss("data/S1_Workers/processed/cleaned/MT - Morning - 06-15-2018.sav"),
  raw.afternoon = read_spss("data/S1_Workers/processed/cleaned/MT - Afternoon - 06-15-2018.sav")
)

# students
dtStudents <- list(
  raw.pre = read.csv(file = "data/S2_Students/raw/AOTS_Pre.csv", header = T, sep = ","),
  raw.post = read.csv(file = "data/S2_Students/raw/AOTS_Post.csv", header = T, sep = ","),
  raw.daily = read.csv(file = "data/S2_Students/raw/AOTS_Daily.csv", header = T, sep = ",")
)

# young medical professionals
dtMedical <- list(
  raw.eligibility = read.csv("data/S3_Medical/raw/AOTM_Eligibility.csv"),
  raw.pre = read.csv("data/S3_Medical/raw/AOTM_Pre.csv"),
  raw.post = read.csv("data/S3_Medical/raw/AOTM_Post.csv"),
  raw.daily = read.csv("data/S3_Medical/raw/AOTM_Daily.csv")
)
```

## Data Cleaning & Data Exclusions

### Worker

For the sojourner sample data was collected in four separate surveys: (1) the pre-measurement, (2) the daily morning survey, (3) the daily afternoon survey, as well as (4) a post-measurement. We combine the four individual surveys into one cohesive dataframe and drop superfluous variables that are not relevant to the analyses relevant here. We then format the time and date variables and add person- and measurement indices (for easy and meaningful addressing of the data). We also exclude our own test data.  
<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtWorker_' list.

```{r cleanWorker}
#  important names for Morning and Afternoon
names.m <- c(
  "StartDate",
  "EndDate",
  "Finished",
  "Duration__in_seconds_",
  "RecordedDate",
  "ExternalReference",
  "Meta_Operating_System",
  "Contact_dum",
  "number",
  "time",
  "duration_1",
  "dyad.group",
  "gr_size",
  "gr_type_1",
  "gr_type_2",
  "gr_type_3",
  "gr_type_4",
  "gr_type_5",
  "gr_type_6",
  "gr_type_7",
  "gr_type_8",
  "gr_type_9",
  "gr_type_10",
  "gr_type_11",
  "gr_type_12",
  "gr_type_13",
  "gr_type_14",
  "gr_type_15",
  "gr_type_16",
  "gr_type_17_TEXT",
  "gr_context_1",
  "gr_context_2",
  "gr_context_3",
  "gr_context_4",
  "gr_context_5",
  "gr_context_6",
  "gr_context_7",
  "gr_context_8",
  "gr_context_9",
  "gr_context_10",
  "gr_context_11",
  "gr_context_12",
  "gr_context_13_TEXT",
  "gr_context_14_TEXT",
  "gr_dutchness",
  "dyad_type_1",
  "dyad_type_2",
  "dyad_type_3",
  "dyad_type_4",
  "dyad_type_5",
  "dyad_type_6",
  "dyad_type_7",
  "dyad_type_8",
  "dyad_type_9",
  "dyad_type_10",
  "dyad_type_11",
  "dyad_type_12",
  "dyad_type_13",
  "dyad_type_14",
  "dyad_type_15",
  "dyad_type_16",
  "dyad_type_17_TEXT",
  "Context_1",
  "Context_2",
  "Context_3",
  "Context_4",
  "Context_5",
  "Context_6",
  "Context_7",
  "Context_8",
  "Context_9",
  "Context_10",
  "Context_11",
  "Context_12",
  "Context_13_TEXT",
  "Context_14_TEXT",
  "keyMotive",
  "keymotive_fulfillemt_1",
  "keyMotive_Dutch_1",
  "autonomy_1",
  "competence_1",
  "relatedness_self_1",
  "relatedness_other_1",
  "qualityAccidental_1",
  "qualityVoluntary_1",
  "qualityCooperative_1",
  "qualityDutchy_1",
  "quality_overall_1",
  "quality_meaning_1",
  "quality_star_1",
  "wantInt",
  "desire_type_1",
  "desire_type_2",
  "desire_type_3",
  "desire_type_4",
  "desire_type_5",
  "desire_type_6",
  "desire_type_7",
  "desire_type_8",
  "desire_type_9",
  "desire_type_10",
  "desire_type_11",
  "desire_type_12",
  "desire_type_13",
  "desire_type_14",
  "desire_type_15",
  "desire_type_16",
  "desire_type_17_TEXT",
  "desire_context_1",
  "desire_context_2",
  "desire_context_3",
  "desire_context_4",
  "desire_context_5",
  "desire_context_6",
  "desire_context_7",
  "desire_context_8",
  "desire_context_9",
  "desire_context_10",
  "desire_context_11",
  "desire_context_12",
  "desire_context_13_TEXT",
  "desire_context_14_TEXT",
  "Reason_nodesire",
  "keyMotive_noInt",
  "keyMotive_noInt_fulf_1",
  "autonomy_NoInt_1",
  "competence_NoInt_1",
  "relatedness_1_NoInt_1",
  "thermometerDutch_1",
  "thermometerDutchInt_2",
  "ExWB_1",
  "alertness1",
  "calmness1",
  "valence1",
  "alertness2",
  "calmness2",
  "valence2",
  "inNonDutch",
  "NonDutchNum",
  "NonDutchType_1",
  "NonDutchType_2",
  "NonDutchType_3",
  "NonDutchType_4",
  "NonDutchType_5",
  "NonDutchType_6",
  "NonDutchType_7",
  "NonDutchType_8",
  "NonDutchType_9",
  "NonDutchType_10",
  "NonDutchType_11",
  "NonDutchType_12",
  "NonDutchType_13",
  "NonDutchType_14",
  "NonDutchType_15_TEXT",
  "date",
  "time.0",
  "LocationLatitude",
  "LocationLongitude"
)

names.a <- c(names.m, "keyInteraction_1", "keyInteractionTime")

# Create reduced data sets for morning and afternoon
dat.mo <- dtWorker$raw.morning[, names.m]
dat.mo$daytime <- "morning"

dat.af <- dtWorker$raw.afternoon[, names.a]
dat.af$daytime <- "afternoon"

# merge morning and afternoon measurements with indicator [+ clean up]
daily.dat <- rbind.fill(dat.mo, dat.af)
daily.dat <- daily.dat[daily.dat$ExternalReference != 55951, ]
dtWorker$daily <- daily.dat
rm(dat.mo, dat.af, names.m, names.a, daily.dat)

# names for pre-measurement
names.pre <- c(
  "Finished",
  "age",
  "Gender",
  "Living",
  "roommate_1",
  "roommate_2",
  "roommate_3",
  "nationality",
  "SecondNationality",
  "timeNL_1",
  "Reason_2",
  "Reason_5",
  "Reason_7",
  "Reason_8_TEXT",
  "DutchLang",
  "occupation_1",
  "occupation_2",
  "occupation_3",
  "occupation_4",
  "occupation_7",
  "CurrentEducation_1",
  "education_level",
  "EduLang_2",
  "RUG_faculty",
  "Study.0",
  "association",
  "DutchMeetNum",
  "DutchFriends_1",
  "assimilation",
  "separation",
  "integration",
  "marginalization",
  "VIA_heritage",
  "VIA_Dutch",
  "SSAS_surrounding",
  "SSAS_privat",
  "SSAS_public",
  "autonomy",
  "relatedness",
  "competence",
  "anxiety",
  "swl",
  "alertness",
  "calmness",
  "valence",
  "date",
  "time",
  "City",
  "ZIP",
  "id"
)

# reduced data set for pre measurement
dat.pre.red <- dtWorker$raw.pre[, names.pre]

# merge with daily data [+ clean up]
df.pre <- merge(
  x = dtWorker$daily,
  y = dat.pre.red,
  by.x = "ExternalReference",
  by.y = "id",
  all = T
)
rm(names.pre)

# adjust duplicate names to fit to indicate daily or pre measurement
names(df.pre) <- gsub("[[:punct:]]x", ".daily", names(df.pre))
names(df.pre) <- gsub("[[:punct:]]y", ".pre", names(df.pre))

# names for post measurement
names.post <- c(
  "ExternalReference",
  "assimilation",
  "separation",
  "integration",
  "marginalization",
  "VIA_heritage",
  "VIA_Dutch",
  "anxiety",
  "swl",
  "rosenberg",
  "social_support",
  "stress",
  "discrimination",
  "discrimination_month",
  "NLE_1month",
  "NLE_6month",
  "NLE_12month"
)

# reduced data set for post-measurement
dat.post.red <- dtWorker$raw.post[, names.post]

# merge post measurement with pre- and daily data
df <- merge(
  x = df.pre,
  y = dat.post.red,
  by.x = "ExternalReference",
  by.y = "ExternalReference",
  all = T
)

# adjust duplicate names to indicate pre or post
names(df) <- gsub("[[:punct:]]x", ".pre", names(df))
names(df) <- gsub("[[:punct:]]y", ".post", names(df))

# add to list
dtWorker$combined <- df

# create data frame with cleaned data
df <- dtWorker$combined %>%
  filter(
    Finished.pre == 1,
    Finished.daily == 1,
    !is.na(ExternalReference)
  )

# add running number as measurement ID within participants
df$measureID <- rowidv(df, cols = c("ExternalReference"))

df <- df %>%
  mutate(
    PID = as.numeric(factor(ExternalReference)),
    # participant ID
    TID = measureID - 1,
    # time ID with t0 = 0 for meaningfull intercept interpretations
    date = substr(StartDate, 1, 10),
    # awkward way of extracting date (best converted to )
    time = substr(StartDate, 12, 19),
    # awkward way of extracting time
    daynum = as.numeric(factor(date)),
    # all days as numeric for ordering
    daycor = ifelse(
      daytime == "morning" &
        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
        daytime == "afternoon" &
          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
      daynum - 1,
      daynum
    ),
    # correctly identify which date the questionnaire is about
    daycor.lead = sprintf("%02d", daycor),
    daytime.lt = ifelse(daytime == "morning", "a", "b"),
    # morning / afternoon to a / b
    day_time = paste(daycor.lead, daytime.lt, sep = "_"),
    # combine day id with morning / afternoon
    session = as.numeric(factor(day_time)),
    # day and time identifier as numeric id
    SubTime = chron::times(time.0),
    time.daily = as.character(time.daily),
    PPDate = as.Date(df$date.daily),
    number = replace_na(number, 0),
    NonDutchNum = replace_na(NonDutchNum, 0)
  )

dtWorker$clean <- df

# clean up
rm(df.pre, names.post, dat.post.red, dat.pre.red, df)

# Export reduced Data
# write.csv(dtWorker$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
# save(dtWorker$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
```

### Student

For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement. We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here. We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement). We also reformat missing values and format core ID variables.  
<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtStudents_' list.

```{r cleanStudents}
# our own test IDs
ownIDs <- c(
  "beautifulLionfishXXXR5rcgVBzGu8hPvOqrK8UBJBw4owvi9nfRFSFu3lMzYhE",
  "niceDogoXXXmB8JI5SFu78SF3DVof84mGUPPNUr14p2HYFTtp31a6D1OwAzM6F-K",
  "amusedQuailXXXmhuc_fpTp8vPkMwDH1BzjaH1d1kHSO1bsPEfsnaEYk4WeVBfPi",
  "juwGAbtXX0_1kmZtSVqKh3PGaHOICqUyU4iBkrT3nDsI_uifuD1gzKcZerxaM5FL"
)

# Prepare dfs for Cleaning
df.pre <- dtStudents$raw.pre %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!is.na(ended)) %>% # remove all who did not finish
  filter(!e_mail %in% .$e_mail[duplicated(.$e_mail)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
  filter(!session %in% ownIDs) %>% # remove our own test
  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)

df.post <- dtStudents$raw.post %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!is.na(session)) %>% # remove own test runs
  filter(!session %in% ownIDs) %>% # remove our own test
  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
  filter(!is.na(ended)) %>% # remove all who never finished
  filter(!session %in% .$session[duplicated(.$session)]) %>% # remove all duplicate sessions
  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)

df.daily <- dtStudents$raw.daily %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!session %in% ownIDs) %>% # remove our own test
  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
  filter(!is.na(ended)) %>% # remove all who never finished
  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)

# merge daily with pre
dfPreDaily <- merge(
  x = df.daily,
  y = df.pre,
  by = "session",
  suffixes = c(".daily", ".pre"),
  all = F
)

# merge daily with post
dfCombined <- merge(
  x = dfPreDaily,
  y = df.post,
  by = "session",
  suffixes = c(".pre", ".post"),
  all = F
)

# add to list
dtStudents$clean <- dfCombined

# clean up workspace
rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
```

### Medical

For the medical professionals sample data was, again, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement. We combine the three individual surveys into one large dataframe. We exclude our own test data. We also reformat missing values and format core ID variables.  
<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtMedical_' list.

```{r cleanMedical}
# our own test IDs
ownIDs <- c(
  "test_LeonieXXXSklxecPLW0-FBPM4796o3pUwUhAY5jb9KGw8jQsKxWmGpa1Jiy", 
  "test_MaxXXXtOp_5dTNefIq0yKXtXt2IN6eEKxeHoPY9mlyvdsqPpLp1B0NGg4UL",
  "test_JannisXXXBsNqk62fOpX6chbd2tMWPptUdjjnhAqnQ3uBqckZ7gLIEoPlfZ",
  "quaintLeopardCatXXXAJ9cfSj-_SZLwNwMDxv_xv_iyr1Bg5YFLTlYdrjW0UXZY",
  "blue-eyedIndianElephantXXXLf5zPMpQCDGS3umFzIj-YVky7ivTItvvozW49m"
)

# Prepare dfs for Cleaning
df.pre <- dtMedical$raw.pre %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!is.na(ended)) %>% # remove all who did not finish
  filter(!session %in% ownIDs) %>% # remove our own test
  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)

df.post <- dtMedical$raw.post %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>% 
  filter(!is.na(session)) %>% # remove own test runs
  filter(!session %in% ownIDs) %>% # remove our own test
  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
  filter(!session %in% .$session[duplicated(.$session)]) %>% # remove all duplicate sessions
  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)

df.daily <- dtMedical$raw.daily %>%
  mutate_all(na_if, "") %>%
  mutate_all(na_if, "NA") %>%
  filter(!session %in% ownIDs) %>% # remove our own test
  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)

# merge daily with pre
dfPreDaily <- merge(
  x = df.daily,
  y = df.pre,
  by = "session",
  suffixes = c(".daily", ".pre"),
  all = F
)

# merge daily with post
dfCombined <- merge(
  x = dfPreDaily,
  y = df.post,
  by = "session",
  suffixes = c(".pre", ".post"),
  all = F
)

# add to list
dtMedical$clean <- dfCombined

# clean up workspace
rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
```

## Calculate needed transformations

### Worker

For the worker sample, the data transformation stage had three main aims:

1. We first corrected time indicators within the surveys. In some cases participants completed their daily diary surveys for the afternoon after midnight. In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.   
2. We then created indices of scales. Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).   
3. Finally, we calculated several basic participant summaries (averages across all measurement occasions).

```{r newVarsWorkers}
df <- dtWorker$clean

# Time and Date Variables
# remove seconds from afternoon time
df$SubTime[df$daytime == "afternoon"] <- paste0(substring(as.character(df$time.0[df$daytime == "afternoon"]), 4, 8), ":00")
df$time.daily[df$daytime == "afternoon" &
  !is.na(df$time.daily != "<NA>")] <- paste0(substring(as.character(df$time.daily[df$daytime == "afternoon" &
  !is.na(df$time.daily != "<NA>")]), 4, 8), ":00")

# Correct morning / afternoon date where survey was collected the day after to indicate the correct date that was targeted
df$PPDate[df$SubTime < "11:50:00" &
  df$daytime == "morning"] <- df$PPDate[df$SubTime < "11:50:00" &
  df$daytime == "morning"] - 1
df$PPDate[df$SubTime < "18:50:00" &
  df$daytime == "afternoon"] <- df$PPDate[df$SubTime < "18:50:00" &
  df$daytime == "afternoon"] - 1

# Need scales
df$keyMotiveFulfilled <- rowSums(df[, c("keymotive_fulfillemt_1", "keyMotive_noInt_fulf_1")], na.rm = T)
df$autonomy.daily.all <- rowSums(df[, c("autonomy_1", "autonomy_NoInt_1")], na.rm = T)
df$competence.daily.all <- rowSums(df[, c("competence_1", "competence_NoInt_1")], na.rm = T)
# cor(df$relatedness_other_1, df$relatedness_self_1,use="complete.obs")
df$relatedness.daily.all <- rowMeans(df[, c(
  "relatedness_other_1",
  "relatedness_self_1",
  "relatedness_1_NoInt_1"
)], na.rm = T)

pairs.panels.new(
  df[c("relatedness_self_1", "relatedness_other_1")],
  labels = c(
    "I shared information about myself.",
    "X shared information about themselves."
  )
)
df$relatedness_1 <- rowMeans(df[, c("relatedness_other_1", "relatedness_self_1")], na.rm = T)

# summarize by participant (check that everything is within pp might not be the case for )
between <- df %>%
  group_by(ExternalReference) %>%
  mutate(
    CtContactNL = sum(Contact_dum),
    CtContactNonNl = sum(inNonDutch),
    CtContactNLAll = sum(number),
    CtContactNonNlAll = sum(NonDutchNum),
    AvKeyNeed = mean(keyMotiveFulfilled, na.rm = T),
    AvKeyNeedInt = mean(keymotive_fulfillemt_1, na.rm = T),
    AvKeyNeedNoInt = mean(keyMotive_noInt_fulf_1, na.rm = T),
    AvAutonomy = mean(autonomy.daily.all, na.rm = T),
    AvCompetence = mean(competence.daily.all, na.rm = T),
    AvRelatedness = mean(relatedness.daily.all, na.rm = T),
    AvThermo = mean(thermometerDutch_1, na.rm = T),
    AvWB = mean(ExWB_1, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(
    CtContactNL_c = scale(CtContactNL, scale = FALSE),
    AvKeyNeedInt_c = scale(AvKeyNeedInt, scale = FALSE),
    AvKeyNeed_c = scale(AvKeyNeed, scale = FALSE),
    CtContactNL_z = scale(CtContactNL, scale = TRUE),
    AvKeyNeedInt_z = scale(AvKeyNeedInt, scale = TRUE),
    AvKeyNeed_z = scale(AvKeyNeed, scale = TRUE)
  )

warning(
  "some variable transformations (esp. _c and _z) might be across all participants (i.e., not within PP). See next step."
)

dtWorker$full <- between
rm(df, between)

# dataframe where interaction types are recoded
workerInteractionType <- dtWorker$full %>%
  mutate(
    OutgroupInteraction = as_factor(Contact_dum),
    NonOutgroupInteraction = as_factor(inNonDutch)
  )

# Create variables centered and standardized within Participant
# i.e., divide into trait and state
workerWithinBetween <-
  MlTraitState(
    data = workerInteractionType,
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "competence.daily.all",
        "autonomy.daily.all",
        "relatedness.daily.all",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1", 
        "quality_overall_1", 
        "OutgroupInteraction",
        "NonOutgroupInteraction"
      )
  )

workerOutWithinBetween <-
  MlTraitState(
    data = workerInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1", 
        "quality_overall_1"
      )
  )

# Between participants contact frequency
workerContactFreq <- dtWorker$full %>%
  group_by(PID) %>%
  summarise(
    n = n(),
    SumContactNL = sum(Contact_dum),
    PercContactNL = SumContactNL / n * 100,
    SumContactNLAll = sum(number),
    AvAttitude = mean(thermometerDutch_1, na.rm = T)
  ) %>%
  mutate(
    WinSumContactNL = DescTools::Winsorize(SumContactNL),
    WinSumContactNLAll = DescTools::Winsorize(SumContactNLAll)
  )

# save cleaned data
# save(df.btw, file = "data/processed/df.btw.RData")
# write_sav(df.btw, "data/processed/MT_clean-merged_pre-post.sav")

# export data to Mplus
# df.mplus = remove_all_labels(select(df,
#                                     PID, session,
#                                     thermometerDutch_1, inNonDutch, Contact_dum,
#                                     keyMotiveFulfilled, autonomy.daily.all, competence.daily.all, relatedness.daily.all))
# names(df.mplus)= c("PID", "session", "att", "intin", "intout", "keymot", "aut", "comp", "rel")
# mplus = df.mplus[order(df.mplus$PID, df.mplus$session),]
# mplus.intcont = mplus[mplus$intout==1,]
# prepareMplusData(mplus.intcont, "data/processed/dynamic-subset-intonly.dat")
```

### Student

For the student sample, the data transformation stage had five main aims:

1. We first create person, survey type, and measurement ID variables.   
2. We then created indices of scales. Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).   
3. We add information about the interaction partner to the beep during which a person was selected as an interaction partner.   
4. We cluster mean-center key variables within participants.   
5. Finally, we calculated several basic participant summaries (averages across all measurement occasions).

```{r newVarsStudents}
df <- dtStudents$clean

# Add ID variables
df$PID <- as.numeric(factor(df$session)) # participant ID

# order time
df$TID <-
  factor(df$date_period, levels = unique(dtStudents$raw.daily$date_period))
df$TIDnum <- as.numeric(df$TID) # get numeric TID

# check whether time ordering worked
df <- df %>%
  arrange(PID, TID) # %>%
# View()

# Interaction as Factor
df$interaction.f <-
  factor(df$Interaction,
    levels = c("no interaction", "Dutch", "Non-Dutch")
  )
df$intNL <- ifelse(df$Interaction == "Dutch", 1, 0)
df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)

# -------------------------------------------------------------------------------------------------------------
#                                       Combine Variables
# -------------------------------------------------------------------------------------------------------------
# Relatedness
pairs.panels.new(
  df[c("RelatednessSelf", "RelatednessOther")],
  labels = c(
    "I shared information about myself.",
    "X shared information about themselves."
  )
)
df$RelatednessInteraction <-
  rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = T)
df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <-
  NA
# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
df$Relatedness <-
  rowMeans(df[, c("RelatednessInteraction", "RelatednessNoInteraction")],
    na.rm =
      T
  )
# Pro-Sociality
df$ProSo <-
  rowMeans(df[, c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm = T)
# Anti-Sociality
df$AntiSo <-
  rowMeans(df[, c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm = T)

# -------------------------------------------------------------------------------------------------------------
#                                 Add Variables related to interaction partner
# -------------------------------------------------------------------------------------------------------------
# create function for later lapply
createIntPartDf <- function(inp) {
  # prepare the dataframe so that we can forloop over it later
  tmp <- data.frame(
    CC = as.character(inp$CC),
    NewCC = as.character(inp$NewCC),
    NewName = as.character(inp$NewName),
    NewCloseness = inp$NewCloseness,
    NewGender = inp$NewGender,
    NewEthnicity = as.character(inp$NewEthnicity),
    NewRelationship = as.character(inp$NewRelationship)
  )

  tmp$CC2 <- recode(tmp$CC, "SOMEONE ELSE" = "NA")
  tmp$CC2 <-
    ifelse(
      tmp$CC == 1 |
        tmp$CC == "SOMEONE ELSE",
      as.character(tmp$NewName),
      as.character(tmp$CC2)
    )
  # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
  tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
  tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)

  # open the variables that will be filled up in the foor-loop
  tmp$closeness <- rep(NA, nrow(tmp))
  tmp$gender <- rep(NA, nrow(tmp))
  tmp$ethnicity <- rep(NA, nrow(tmp))
  tmp$relationship <- rep(NA, nrow(tmp))

  # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
  # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
  for (i in 1:nrow(tmp)) {
    if (is.na(tmp$CC2[i])) {
      next
    } else {
      tmp$closeness[i] <-
        na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # find closeness where CC2 matches NewName (na.omit + [1] to get the number)
      tmp$gender[i] <-
        na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # (na.omit + [1] to get the number and not the rest of the na.omit list)
      tmp$ethnicity[i] <-
        na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] # PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
      tmp$relationship[i] <-
        na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
    }
  }

  out <- tmp
  out
}

# split df per participants and run function
PP <- split(df, df$PID)
PP <- lapply(PP, createIntPartDf)
rm(createIntPartDf)

# add variables back to df
remergePP <- do.call(rbind.data.frame, PP)
colnames(remergePP) <-
  paste(colnames(remergePP), "_Calc", sep = "")
df <- cbind(df, remergePP)
rm(remergePP, PP)

# -------------------------------------------------------------------------------------------------------------
#                                 Center Relevant Variables
# -------------------------------------------------------------------------------------------------------------

df <- df %>%
  group_by(PID) %>%
  mutate(
    KeyNeedFullfillment.cm = mean(KeyNeedFullfillment, na.rm = TRUE),
    # cluster mean (mean of PP)
    KeyNeedFullfillment.cwc = KeyNeedFullfillment - KeyNeedFullfillment.cm,
    # cluster mean centered (within PP centered)
    closeness.cm = mean(closeness_Calc, na.rm = TRUE),
    closeness.cwc = closeness_Calc - closeness.cm
  ) %>%
  ungroup()

# store
dtStudents$full <- df
rm(df)

# Between participants contact frequency
studentContactFreq <- dtStudents$full %>%
  group_by(PID) %>%
  summarise(
    n = n(),
    SumContactNL = sum(InteractionDumDutch),
    PercContactNL = SumContactNL / n * 100,
    SumContactNLAll = sum(ContactNum[InteractionDumDutch == 1], na.rm = TRUE),
    AvAttitude = mean(AttitudesDutch, na.rm = TRUE),
    AvQuality = mean(quality_overall, na.rm = TRUE)
  ) %>%
  mutate(
    WinSumContactNL = DescTools::Winsorize(SumContactNL),
    WinSumContactNLAll = DescTools::Winsorize(SumContactNLAll)
  )

# dataframe where interaction types are recoded
studentInteractionType <- dtStudents$full %>%
  mutate(
    NonDutchContact = tidyr::replace_na(NonDutchContact, 2), # make second non-Dutch countable
    NonDutchContact = NonDutchContact*-1+2 # recode (yes = 1 -> 1, no = 2 -> 0)
  ) %>%
  mutate(
    OutgroupInteraction = factor(
      InteractionDumDutch,
      levels = c(0, 1),
      labels = c("No", "Yes")
    ),
    NonOutgroupInteraction = factor(
      rowSums(select(., c(InteractionDumNonDutch, NonDutchContact))), # combine the two non-Dutch Q.,
      levels = c(0, 1),
      labels = c("No", "Yes")
    )
  )

# select a subset of IDs to display in plots
studentPltIDs <-
  studentInteractionType %>%
  group_by(PID) %>%
  summarise(n = n()) %>%
  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
  select(PID) %>%
  as.matrix %>%
  as.vector

# select a subset of IDs to display in plots (only outgroup interactions)
studentOutPltIDs <-
  studentInteractionType %>%
  filter(OutgroupInteraction == "Yes") %>%
  group_by(PID) %>%
  summarise(n = n()) %>%
  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
  select(PID) %>%
  as.matrix %>%
  as.vector

# Center within and between
# divide into trait and state
studentWithinBetween <-
  MlTraitState(
    data = studentInteractionType,
    id = "PID",
    selection =
      c(
        "KeyNeedFullfillment",
        "Competence",
        "Autonomy",
        "Relatedness",
        "AttitudesDutch",
        "quality_overall",
        "OutgroupInteraction",
        "NonOutgroupInteraction"
      )
  )
studentOutWithinBetween <-
  MlTraitState(
    data = studentInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection =
      c(
        "KeyNeedFullfillment",
        "Competence",
        "Autonomy",
        "Relatedness",
        "AttitudesDutch",
        "quality_overall"
      )
  )

```

### Medical

For the medical professional sample, the data transformation stage had five main aims:

1. We first create person, survey type, and measurement ID variables.   
2. We then created indices of scales. Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).   
3. We cluster mean-center key variables within participants.   
4. Finally, we calculated several basic participant summaries (averages across all measurement occasions).

```{r newVarsMedical}
df <- dtMedical$clean

# Add ID variables
df$PID <- as.numeric(factor(df$session)) # participant ID

# order time
df$TID <-
  factor(df$date_period, levels = unique(dtMedical$raw.daily$date_period))
df$TIDnum <- as.numeric(df$TID) # get numeric TID

# check whether time ordering worked
df <- df %>%
  arrange(PID, TID) # %>%
# View()

# Interaction as Factor
df$interaction.f <-
  factor(df$Interaction,
    levels = c("no interaction", "Dutch", "Non-Dutch")
  )
df$intNL <- ifelse(df$Interaction == "Dutch", 1, 0)
df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)

df <- df %>%
  mutate(
    NonDutchContact = replace_na(NonDutchNum, 0), # make second non-Dutch countable
    NonDutchContact = ifelse(NonDutchContact > 1, 1, 0) # recode (yes = 1 -> 1, no = 2 -> 0)
  ) %>%
  mutate(
    OutgroupInteraction = factor(
      InteractionDumDutch,
      levels = c(0, 1),
      labels = c("No", "Yes")
    ),
    NonOutgroupInteraction = factor(
      rowSums(select(., c(InteractionDumNonDutch, NonDutchContact)), na.rm = TRUE), # combine the two non-Dutch Q.,
      levels = c(0, 1),
      labels = c("No", "Yes")
    )
  )

# -------------------------------------------------------------------------------------------------------------
#                                       Combine Variables
# -------------------------------------------------------------------------------------------------------------
# Relatedness
pairs.panels.new(
  df[c("RelatednessSelf", "RelatednessOther")],
  labels = c(
    "I shared information about myself.",
    "X shared information about themselves."
  )
)
df$RelatednessInteraction <-
  rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = T)
df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <-
  NA
# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
df$Relatedness <-
  rowMeans(df[, c("RelatednessInteraction", "RelatednessNoInteraction")],
           na.rm = TRUE) %>%
  ifelse(is.nan(.), NA, .)

df$DaytimeNeedFullfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
  df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]

df$DaytimeNeedImportance <- NA 
df$DaytimeNeedImportance[df$InteractionDum == 1 & !is.na(df$DaytimeNeedFulfillment)] <- 
  df$DaytimeNeedFulfillment[df$InteractionDum == 1 & !is.na(df$DaytimeNeedFulfillment)]

df$InteractionNeedFullfillment <- NA 
df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)] <- 
  df$KeyNeedFulfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)]

df$InteractionNeedImportance <- NA 
df$InteractionNeedImportance[df$InteractionDum == 1 & !is.na(df$KeyNeedImp)] <- 
  df$KeyNeedImp[df$InteractionDum == 1 & !is.na(df$KeyNeedImp)]

# Pro-Sociality
df$ProSo <-
  rowMeans(df[, c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm = T)
# Anti-Sociality
df$AntiSo <-
  rowMeans(df[, c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm = T)

# Allport's Conditions
df %>%
  #filter(OutgroupInteraction == "Yes") %>%
  select(
    InteractionContextEqualStatus,
    KeyNeedShared,
    InteractionContextCooperative,
    InteractionContextvoluntary
  ) %>%
  pairs.panels.new

df %>%
  #filter(OutgroupInteraction == "Yes") %>%
  select(
    InteractionContextEqualStatus,
    KeyNeedShared,
    InteractionContextCooperative,
    InteractionContextvoluntary
  ) %>%
  psych::describe(., skew=F,ranges=T) %>%
  as.data.frame() %>%
  select(-vars) %>%
  kable(., caption = "Descriptives of Allport's Condition items") %>% 
  kable_styling("hover", full_width = F, latex_options = "hold_position")

iaWorkerAllport <- 
  df %>%
  #filter(OutgroupInteraction == "Yes") %>%
  select(
    InteractionContextEqualStatus,
    KeyNeedShared,
    InteractionContextCooperative,
    InteractionContextvoluntary
  )

sjPlot::tab_itemscale(iaWorkerAllport)

pca <- parameters::principal_components(iaWorkerAllport)
factor.groups <- parameters::closest_component(pca)

sjPlot::tab_itemscale(iaWorkerAllport, factor.groups)

ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)

data <- 
  df %>%
  select(
    PID,
    TIDnum,
    InteractionContextEqualStatus,
    KeyNeedShared,
    InteractionContextCooperative,
    InteractionContextvoluntary
  ) %>%
  drop_na %>%
  melt(
    ., 
    id.vars = c("PID", "TIDnum")
  )

horst::nestedAlpha(item.level.1 = "value",
                   level.2      = "TIDnum",
                   level.3      = "PID",
                   data         = data)
rm(data)

iaWorkerAllportScale <- 
  iaWorkerAllport %>%
  Scale::Scale() %>%
  Scale::ItemAnalysis()

df$AllportsCondition <-
  scoreItems(
    keys = c(1, 1, 1, 1),
    items = df %>% select(
      InteractionContextEqualStatus,
      KeyNeedShared,
      InteractionContextCooperative,
      InteractionContextvoluntary
    ),
    min = 0,
    max = 100
  )$scores

as.data.frame(psych::describe(df$AllportsCondition, skew=T)) %>%
  mutate(vars = "Allport's Conditions Index") %>%
  kable(., caption = "Allport's Conditions: Scale Descriptives", row.names = FALSE) %>% 
  kable_styling("hover", full_width = F, latex_options = "hold_position")

ggplot(df, aes(x = AllportsCondition)) +
  geom_histogram()

# -------------------------------------------------------------------------------------------------------------
#                                 Add Variables related to interaction partner
# -------------------------------------------------------------------------------------------------------------
# create function for later lapply
createIntPartDf <- function(inp) {
  # prepare the dataframe so that we can forloop over it later
  tmp <- data.frame(
    CC = as.character(inp$CC),
    NewCC = as.character(inp$NewCC),
    NewName = as.character(inp$NewName),
    NewCloseness = inp$NewCloseness,
    NewGender = inp$NewGender,
    NewEthnicity = as.character(inp$NewEthnicity),
    NewRelationship = as.character(inp$NewRelationship)
  )

  tmp$CC2 <- recode(tmp$CC, "SOMEONE ELSE" = "NA")
  tmp$CC2 <-
    ifelse(
      tmp$CC == 1 |
        tmp$CC == "SOMEONE ELSE",
      as.character(tmp$NewName),
      as.character(tmp$CC2)
    )
  # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
  tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
  tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)

  # open the variables that will be filled up in the foor-loop
  tmp$closeness <- rep(NA, nrow(tmp))
  tmp$gender <- rep(NA, nrow(tmp))
  tmp$ethnicity <- rep(NA, nrow(tmp))
  tmp$relationship <- rep(NA, nrow(tmp))

  # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
  # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
  for (i in 1:nrow(tmp)) {
    if (is.na(tmp$CC2[i])) {
      next
    } else {
      tmp$closeness[i] <-
        na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # find closeness where CC2 matches NewName (na.omit + [1] to get the number)
      tmp$gender[i] <-
        na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # (na.omit + [1] to get the number and not the rest of the na.omit list)
      tmp$ethnicity[i] <-
        na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] # PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
      tmp$relationship[i] <-
        na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
    }
  }

  out <- tmp
  out
}

# split df per participants and run function
PP <- split(df, df$PID)
PP <- lapply(PP, createIntPartDf)
rm(createIntPartDf)

# add variables back to df
remergePP <- do.call(rbind.data.frame, PP)
colnames(remergePP) <-
  paste(colnames(remergePP), "_Calc", sep = "")
df <- cbind(df, remergePP)
rm(remergePP, PP)

# -------------------------------------------------------------------------------------------------------------
#                                 Center Relevant Variables
# -------------------------------------------------------------------------------------------------------------
# divide into trait and state
medicalOutWithinBetween <-
  MlTraitState(
    data = df %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection =
      c(
        "KeyNeedFulfillment",
        "Competence",
        "Autonomy",
        "Relatedness",
        "AllportsCondition",
        "AttitudesDutch",
        "qualityOverall"
      )
  )

medicalWithinBetween <-
  MlTraitState(
    data = df,
    id = "PID",
    selection =
      c(
        "KeyNeedFulfillment",
        "Competence",
        "Autonomy",
        "Relatedness",
        "AllportsCondition",
        "AttitudesDutch",
        "qualityOverall",
        "OutgroupInteraction",
        "NonOutgroupInteraction"
      )
  )

df <- # keep only for compatibility of old framgents
  MlTraitState(
    data = df,
    id = "PID",
    selection =
      c(
        "KeyNeedFulfillment",
        "Competence",
        "Autonomy",
        "Relatedness",
        "AllportsCondition",
        "AttitudesDutch",
        "qualityOverall"
      )
  )

# store
dtMedical$full <- df
rm(df)

# Between participants contact frequency
medicalContactFreq <- 
  dtMedical$full %>%
  group_by(PID) %>%
  summarise(
    n = n(),
    SumContactNL = sum(InteractionDumDutch, na.rm = TRUE),
    PercContactNL = SumContactNL / n * 100,
    SumContactNLAll = sum(ContactNum[InteractionDumDutch == 1], na.rm = TRUE),
    AvAttitude = mean(AttitudesDutch, na.rm = TRUE),
    AvQuality = mean(qualityOverall, na.rm = TRUE)
  ) %>%
  mutate(
    WinSumContactNL = DescTools::Winsorize(SumContactNL),
    WinSumContactNLAll = DescTools::Winsorize(SumContactNLAll)
  )

# select a subset of IDs to display in plots
medicalPltIDs <-
  dtMedical$full %>%
  group_by(PID) %>%
  summarise(n = n()) %>%
  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
  select(PID) %>%
  as.matrix %>%
  as.vector

# select a subset of IDs to display in plots (only outgroup interactions)
medicalOutPltIDs <-
  dtMedical$full %>%
  filter(OutgroupInteraction == "Yes") %>%
  group_by(PID) %>%
  summarise(n = n()) %>%
  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
  select(PID) %>%
  as.matrix %>%
  as.vector
```

## Data Availability and Time Scales

```{r workerDataAvailability}
dtworkerMissing <- dtWorker$full %>%
  select(
    PID,
    Finished.daily,
    PPDate,
    daytime,
    TID
  ) %>%
  mutate(
    date_period = paste(PPDate, daytime, sep = " "),
    available = 1
  )

library(reshape)
dtworkerAvailability <- reshape::cast(dtworkerMissing, PID ~ TID)
write.csv(dtworkerAvailability, "data/S1_Workers/processed/workerAvailability.csv")
```

```{r studentDataAvailability}
dtstudentMissing <- dtStudents$full %>%
  select(
    PID,
    date.daily,
    periodMA,
    date_period,
    TID
  ) %>%
  mutate(
    TID = as.numeric(TID),
    available = 1
  )

dtstudentAvailability <- reshape::cast(dtstudentMissing, PID ~ TID)
write.csv(dtstudentAvailability, "data/S2_Students/processed/studentAvailability.csv")
```

```{r medicalDataAvailability}
dtMedicalMissing <- dtMedical$full %>%
  select(
    PID,
    date.daily,
    periodMA,
    date_period,
    TID
  ) %>%
  mutate(
    TID = as.numeric(TID),
    available = 1
  )

dtMedicalAvailability <- reshape::cast(dtMedicalMissing, PID ~ TID)
write.csv(dtMedicalAvailability, "data/S3_Medical/processed/medicalAvailability.csv")
```

# Mean Plots per Item

## Study 1 (Worker):

```{r workerDataSelection}

```

## Study 2 (Student):

```{r studentDataSelection}

```

## Study 3 (Medical):

```{r medicalDataSelection}
varListMedical <- 
  c(
    # "KeyNeedFulfillment",
    # "KeyNeedImp",
    "DaytimeNeedFullfillment",
    "DaytimeNeedImportance",
    "InteractionNeedFullfillment",
    "InteractionNeedImportance",
    "AntiSo1",
    "AntiSo2",
    "AntiSo3",
    "AntiSo4",
    "AntiSo5",
    "AntiSo6",
    "AntiSo7",
    "ProSo1",
    "ProSo2",
    "ProSo3",
    "ProSo4",
    "StudentGoal01",
    "StudentGoal02",
    "StudentGoal03",
    "StudentGoal04",
    "StudentGoal05",
    "StudentGoal06",
    "StudentGoal07",
    "StudentGoal08",
    "StudentGoal09",
    "StudentGoal10",
    "Autonomy",
    "Competence",
    # Break relatedness up into interaction vs. no interaction(?)
    # vars are already calculated
    "Relatedness",
    "AttitudesDutch",
    "AttitudesPartner",
    "exWB",
    "angry",
    "afraid",
    "energy",
    "lonelyAlways",
    "Event"
  )

varNameListMedical <- 
  c(
    # "KeyNeedFulfillment" = "Key Need Fulfillment",
    # "KeyNeedImp" = "Key Need Importance",
    "DaytimeNeedFullfillment" = "Daytime Core Motive Fulfillment",
    "DaytimeNeedImportance" = "Daytime Core Motive Importance",
    "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
    "InteractionNeedImportance" = "Interaction Core Motive Importance",
    "AntiSo1" = "Put someone down",
    "AntiSo2" = "Show little attention in someones opinion",
    "AntiSo3" = "Demeaning remarks",
    "AntiSo4" = "Inpropperly addressing someone",
    "AntiSo5" = "Ignored or excluded someone",
    "AntiSo6" = "Doubt someones judgement",
    "AntiSo7" = "Unwanted attempts of personal matters",
    "ProSo1" = "Listen to someones problems",
    "ProSo2" = "Cheer someone up",
    "ProSo3" = "Help someone get things done",
    "ProSo4" = "Help someone with responsibilities",
    "StudentGoal01" = "Social support and connectedness",
    "StudentGoal02" = "Romantic or sexual relationship",
    "StudentGoal03" = "Academic",
    "StudentGoal04" = "Career",
    "StudentGoal05" = "Financial",
    "StudentGoal06" = "Health and fitness",
    "StudentGoal07" = "Leasure and fun",
    "StudentGoal08" = "Personal improvement and growth",
    "StudentGoal09" = "Service and help",
    "StudentGoal10" = "Spiritual or religious",
    "Autonomy" = "Autonomy",
    "Competence" = "Competence",
    # Break relatedness up into interaction vs. no interaction(?)
    # vars are already calculated
    "Relatedness" = "Relatedness",
    "AttitudesDutch" = "Outgroup Attitude",
    "AttitudesPartner" = "Attitude Interaction Partner",
    "exWB" = "Sadness Happiness",
    "angry" = "Anger",
    "afraid" = "Anxiety",
    "energy" = "Energy",
    "lonelyAlways" = "Loneliness",
    "Event" = "Positive or negative event present"
  )

dtMedical$viz <- list()

dtMedical$vizall <- dtMedical$full %>%
  select(
    PID,
    TID,
    TIDnum,
    varListMedical
  )

dtMedical$viz$DaytimeNeed <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    starts_with("DaytimeNeed")
  )

dtMedical$viz$InteractionNeed <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    starts_with("InteractionNeed")
  )

dtMedical$viz$AntiSocialBehaviorPt1 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    AntiSo1,
    AntiSo2,
    AntiSo3
  )

dtMedical$viz$AntiSocialBehaviorPt2 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    AntiSo4,
    AntiSo5
  )

dtMedical$viz$AntiSocialBehaviorPt3 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    AntiSo6,
    AntiSo7
  )

dtMedical$viz$ProSocialBehaviorPt1 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    ProSo1,
    ProSo2
  )

dtMedical$viz$ProSocialBehaviorPt2 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    ProSo3,
    ProSo4
  )

dtMedical$viz$GoalsPt1 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    StudentGoal01,
    StudentGoal02
  )

dtMedical$viz$GoalsPt2 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    StudentGoal03,
    StudentGoal04
  )

dtMedical$viz$GoalsPt3 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    StudentGoal05,
    StudentGoal06
  )

dtMedical$viz$GoalsPt4 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    StudentGoal07,
    StudentGoal08
  )

dtMedical$viz$GoalsPt5 <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    StudentGoal09,
    StudentGoal10
  )

dtMedical$viz$SDT <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    Autonomy,
    Relatedness, 
    Competence
  )

dtMedical$viz$Attitudes <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    starts_with("Attitudes")
  )

dtMedical$viz$WB <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    exWB
  )

dtMedical$viz$Emotion <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    angry,
    afraid,
    energy,
    lonelyAlways
  )

dtMedical$viz$Event <- dtMedical$vizall %>%
  select(
    PID,
    TID,
    TIDnum,
    Event
  )

lWideToLong <- function(data, ...){
  reshape2::melt(
    data, 
    id.vars = c("PID", "TIDnum")
  )
}

dtMedical$vizLong <- lapply(
  dtMedical$viz,
  lWideToLong
)

```

```{r MeanVizFun}

lWideToLong <- function(data, ...){
  reshape2::melt(
    data, 
    id.vars = c("PID", "TID", "TIDnum")
  )
}

meanVarPlot <- function(dataIn, name, namList,...){
  # dataIn <- dtMedical$viz$InteractionNeed
  # name <- "Interaction Core Need"
  # namList <- varNameListMedical
  dataLong <- dataIn %>%
    mutate(TID = stri_replace_all_regex(
      TID,
      pattern = c('Morning', 'Afternoon'),
      replacement = c('12:00:00', '19:00:00'),
      vectorize = FALSE
    ) %>% as.POSIXct) %>%
    lWideToLong %>%
    mutate(variable = recode(variable, !!!namList))

  pTidMain <-
    ggplot(dataLong,
           aes(x = TIDnum, y = value, color = variable)) +
    geom_point(alpha = 0) +
    stat_summary(fun = mean, geom = "line") +
    labs(y = "Mean per Day",
         x = "Time Index") +
    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
    ggthemes::scale_colour_calc() +
    scale_x_continuous(breaks = seq(0, max(dataLong$TIDnum), 25)) +
    theme_Publication() +
    theme(
      panel.border = element_rect(colour = "black"),
      plot.margin = margin(0, 10, 10, 10, "mm")
    )
  pTidTop <-
    ggplot(dataLong %>% filter(!is.na(value)), aes(x = TIDnum, y = (..count.. / max(count)) * 100),
           color = variable) +
    geom_freqpoly(binwidth = 2) +
    geom_hline(yintercept = 80,
               linetype = "dashed",
               color = "black") +
    annotate(
      # add white background
      "label",
      x = max(dataLong$TIDnum),
      y = 80,
      label = "80%",
      vjust = 0.5,
      hjust = -1,
      label.size = NA
    )  +
    coord_cartesian(clip = "off") +
    labs(y = "Percentage") +
    #scale_colour_manual(values = rep("black", 20)) +
    theme_classic() +
    theme(
      legend.position = "none",
      axis.line.x = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(0, 10, 0, 10, "mm")
    )

  pTidTitle <- cowplot::ggdraw() +
    cowplot::draw_label(
      paste("Variable Set:", name, "[Time Id Plot]"),
      fontface = 'bold',
      x = 0.5,
      hjust = 0.5
    )

  pTid <- cowplot::plot_grid(
    pTidTitle,
    pTidTop,
    pTidMain,
    align = "v",
    nrow = 3,
    rel_heights = c(0.1, 1 / 5, 4 / 5)
  )

  pDateMain <-
    ggplot(dataLong,
           aes(x = TID, y = value, color = variable)) +
    geom_point(alpha = 0) +
    stat_summary(fun = mean, geom = "line") +
    labs(y = "Mean per Day",
         x = "Date") +
    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
    ggthemes::scale_colour_calc() +
    scale_x_datetime(breaks = scales::date_breaks("10 day")) +
    theme_Publication() +
    theme(
      panel.border = element_rect(colour = "black"),
      plot.margin = margin(0, 10, 10, 10, "mm")
    )
  pDateTop <-
    ggplot(dataLong,
           aes(
             x = TID,
             y = (..count.. / max(count)) * 100,
             color = variable
           )) +
    geom_freqpoly(binwidth = 1 * 3600 * 24) +
    geom_hline(yintercept = 80,
               linetype = "dashed",
               color = "black") +
    annotate(
      # add white background
      "label",
      x = max(unique(dataLong$TID)),
      y = 80,
      label = "80%",
      vjust = 0.5,
      hjust = -1,
      label.size = NA
    )  +
    coord_cartesian(clip = "off") +
    labs(y = "Percentage") +
    scale_colour_manual(values = rep("black", 20)) +
    theme_classic() +
    theme(
      legend.position = "none",
      axis.line.x = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(0, 10, 0, 10, "mm")
    )

  pDateTitle <- cowplot::ggdraw() +
    cowplot::draw_label(
      paste("Variable Set: Key-Need", "[Date Plot]"),
      fontface = 'bold',
      x = 0.5,
      hjust = 0.5
    )

  pDate <- cowplot::plot_grid(
    pDateTitle,
    pDateTop,
    pDateMain,
    align = "v",
    nrow = 3,
    rel_heights = c(0.1, 1 / 5, 4 / 5)
  )

  list(pTid, pDate)
  pTid
}
```

```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
#, warning=FALSE, message=FALSE, error=FALSE
# meanVarPlot(dtMedical$viz$DaytimeNeed, "Key-Need")
# meanVarPlot(dtMedical$viz$AntiSocialBehavior, "Anti-Social Behavior")

dtMedical$vizPlots <-
  lapply(seq_along(dtMedical$viz),
         function(i)
           meanVarPlot(dataIn = dtMedical$viz[[i]], 
                       name = names(dtMedical$viz)[[i]], 
                       namList = varNameListMedical))
dtMedical$vizPlots
```

# **Software Information**  
The full session information with all relevant system information and all loaded and installed packages is available in the collapsible section below.

<details>
  <summary>System Info</summary>

  \renewcommand{\arraystretch}{0.8} <!-- decrease line spacing for the table -->
```{r Reproducibility-SessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
df_session_platform <- devtools::session_info()$platform %>%
  unlist(.) %>%
  as.data.frame(.) %>%
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kbl(
  df_session_platform,
  booktabs = T,
  align = "l",
  caption = "R environment session info for reproducibility of results" # complete caption for main document
) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```
  \renewcommand{\arraystretch}{1} <!-- reset row height/line spacing -->
 </details>
 <br>
 <details>
  <summary>Package Info</summary>

\renewcommand{\arraystretch}{0.6} <!-- decrease line spacing for the table -->
```{r Reproducibility-SessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
df_session_packages <- devtools::session_info()$packages %>%
  as.data.frame(.) %>%
  filter(attached == TRUE) %>%
  dplyr::select(loadedversion, date, source) %>%
  rownames_to_column()

colnames(df_session_packages) <- c("Package", "Loaded version", "Date", "Source")

kbl(
  df_session_packages,
  booktabs = T,
  align = "l",
  caption = "Package info for reproducibility of results" # complete caption for main document
) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```
\renewcommand{\arraystretch}{1} <!-- reset row height/line spacing -->
</details>
<br>
<details>
  <summary>Full Session Info (including loaded but unattached packages --- for troubleshooting only)</summary>
    `r pander(sessionInfo(), compact = FALSE)`
</details>

</br>

--------------------------------------------------------------------

</br>

# **References**

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

### 2022-05-05 scripts/functions/MlCorMat.R

@@ -1,0 +1,323 @@
+MlCorMat <-
+  function(data,
+           id,
+           selection,
+           labels = selection,
+           method = c("pearson", "spearman"),
+           removeTriangle = c("upper", "lower"),
+           result = c("none", "html", "latex")) {
+    # for testing
+    # data <- dtWorkerSupp$workerInteractionType
+    # id <- "PID"
+    # selection <- varSelection
+    # labels <- varlabels
+    
+    # calculate trait scores (participant means)
+    data <- 
+      data %>%
+      select(id, selection) %>%
+      group_by_at(vars(matches(id))) %>%
+      mutate_at(selection, list(trait = ~mean(., na.rm=TRUE))) %>%
+      ungroup 
+    
+    data <-
+      map_dfc(selection,
+              ~ data %>%
+                transmute(
+                  !!str_c(.x, '_state') :=
+                    !!rlang::sym(.x)-!!rlang::sym(str_c(.x, "_trait"))
+                )) %>%
+      bind_cols(data, .)
+    
+    # Fid out why own is different
+    # raw <-
+    #   data %>%
+    #   select(selection) %>%
+    #   as.matrix
+    # 
+    # corRaw <- rcorr(raw, type=method[1])
+    # rRaw <- corRaw$r # Matrix of correlation coeficients
+    # pRaw <- corRaw$P # Matrix of p-value 
+    # 
+    # ## Define notions for significance levels; spacing is important.
+    # starsRaw <- ifelse(pRaw < .001, "***", ifelse(pRaw < .01, "**", ifelse(pRaw < .05, "* ", "   ")))
+    # 
+    # ## trunctuate the correlation matrix to two decimal
+    # rRaw <- format(round(cbind(rep(-1.11, ncol(raw)), rRaw), 2))[,-1]
+    # 
+    # ## build a new matrix that includes the correlations with their apropriate stars
+    # rRawStar <- matrix(paste(rRaw, starsRaw, sep=""), ncol=ncol(raw))
+    # 
+    # 
+    # trait <-
+    #   data %>%
+    #   select(paste0(selection, "_trait")) %>%
+    #   as.matrix
+    # 
+    # corTrait <- rcorr(trait, type=method[1])
+    # rTrait <- corTrait$r # Matrix of correlation coeficients
[…]

### 2022-05-05 scripts/functions/MlTbl.R

@@ -1,0 +1,328 @@
+#obj <- mdlWorkerLmerSlopesAttCoreInt
+#objz <- mdlWorkerLmerSlopesAttCoreIntZ
+
+lmerTblPrep <- function (obj, objz = NULL, alpha = 0.05, ...) {
+  # Summarize models
+  smry <- summ(obj)
+  smryZ <- summ(objz)
+  
+  # save grouing variable
+  groupNam <- as.character(as.data.frame(smry$gvars)["Group"])
+  
+  # Get standardized coefficients 
+  coefZ <- smryZ$coeftable %>% 
+    as.data.frame %>%
+    tibble::rownames_to_column(., "coef")
+  confZ <- confint(objz, 
+                   method = "Wald") %>% 
+                   #method = "boot", 
+                   #parallel = "multicore", 
+                   #ncpus = parallel::detectCores(logical = TRUE)) %>% 
+    as.data.frame %>%
+    tibble::rownames_to_column(., "coef")
+  coefZCI <- merge(coefZ, confZ)
+  colnames(coefZCI) <- c("coef", "est", "se", "tval", "df", "p", "lwr", "upr")
+  
+  coefZR2 <- r2glmm::r2beta(
+    model = objz,
+    partial = TRUE,
+    # 'sgv' standardized generalized variance approach, 
+    # 'kr' (Kenward Roger approach; only with lme),
+    # 'nsj' (Nakagawa and Schielzeth approach)
+    method = 'sgv' 
+  ) %>%
+    as.data.frame %>%
+    select(
+      coef = Effect,
+      Rsq,
+      RsqLower = lower.CL,
+      RsqUpper = upper.CL
+    )
+  coefZOut <- merge(coefZCI, coefZR2)
+  
+  coefZOut$Beta <- paste0(
+    format(round(coefZOut$est, 2), nsmall = 2),
+    " [", 
+    format(round(coefZOut$lwr, 2), nsmall = 2),
+    ", ",
+    format(round(coefZOut$upr, 2), nsmall = 2),
+    "]"
+  )
+  
+  # Get unstandardized coefficients
+  coef <- smry$coeftable %>% 
+    as.data.frame %>%
+    tibble::rownames_to_column(., "coef")
+  conf <- confint(obj,
+                  method = "Wald") %>% 
+                  #method = "profile") %>% 
+    as.data.frame %>%
[…]

### 2022-05-05 scripts/functions/metaLmer.R

@@ -1,0 +1,575 @@
+strLabFix <- function(varIn, ...) {
+  stri_replace_all_fixed(
+    varIn,
+    pattern = c(
+      "SumContactNL", 
+      "AvQuality", 
+      "SumContactNL.AvQuality",
+      "NonOutgroupInteraction",
+      "OutgroupInteraction",
+      "CoreNeedZ",
+      "CoreNeed",
+      "QualityZ",
+      "_zwc",
+      "."
+    ),
+    replacement = c(
+      "Number of Outgroup Contacts",
+      "Average Interaction Quality",
+      "Contact X Quality",
+      "Non-Outgroup Interaction",
+      "Outgroup Interaction",
+      "Core Need Fulfillment",
+      "Core Need Fulfillment",
+      "Interaction Quality",
+      "",
+      " X "
+    ),
+    vectorize_all = FALSE
+  )
+}
+
+strStudyFix <- function(studyIn, ...) {
+  stri_replace_all_fixed(
+    studyIn,
+    pattern = c(
+      "Worker", 
+      "worker", 
+      "Student",
+      "student",
+      "Medical",
+      "medical"
+    ),
+    replacement = c(
+      "Study 1",
+      "Study 1",
+      "Study 2",
+      "Study 2",
+      "Study 3",
+      "Study 3"
+    ),
+    vectorize_all = FALSE
+  )
+}
+
+metaLmerOut <- function(lmerDataTbl, type = "FE", name, title, ...) {
+  # lmerDataTbl <- GeneralLmerMetaTbl
+  # name <- "Theory"
+  # title <- "Need Fullfillment — Mixed Effects Regression"
+  
+  # extract parameter names
+  var <- as.character(unique(lmerDataTbl$coef))[!as.character(unique(lmerDataTbl$coef)) %in% c("(Intercept)")]
+  names <- strLabFix(var)
+  studies <- as.character(unique(lmerDataTbl$sample))
+  
+  
+  ### Parametric Meta-Analysis
+  meta.parametric <- list()
+  
+  sum.names <- c("effect", 
+                 "name",
+                 paste(c("beta", "lwr", "upr", "CI"),rep(studies, each = 4),sep="."),
+                 "beta.meta",
+                 "lwr.meta",
+                 "upr.meta",
+                 "CI.meta",
+                 "pval.meta",
+                 "star.meta")
+  meta.parametric.sum <-
+    data.frame(matrix(
+      ncol = length(sum.names),
[…]

### 2022-05-05 Supplemental-Material-A-Annotated-Analysis.Rmd
student means added + varNam fixes

@@
 - 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
-- 'The data repository for this manuscript is available at <a href="https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970" target="_blank">osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970</a>'
+- 'The data repository for this manuscript is available at <a href="https://osf.io/TBD" target="_blank">osf.io/TBD</a>'
 - 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
@@
 source("./scripts/functions/metaLmer.R")
+source("./scripts/functions/meanViz.R")
 
@@
 # order time
-df$TID <-
-  factor(df$date_period, levels = unique(dtStudents$raw.daily$date_period))
+df$TID <- factor(df$date_period, levels = unique(dtStudents$raw.daily$date_period))
 df$TIDnum <- as.numeric(df$TID) # get numeric TID
@@
 df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)
+
+df$IntergroupContact <- (df$IntergroupContact-2)*-1
+
+# Fix naming issues
+df$AttitudesPartner <- df$AttitdutuesPartnerstick
 
@@
 )
-df$RelatednessInteraction <-
-  rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = T)
-df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <-
-  NA
+df$RelatednessInteraction <- rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = TRUE)
+df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
 # Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
@@
   rowMeans(df[, c("RelatednessInteraction", "RelatednessNoInteraction")],
-    na.rm =
-      T
-  )
+           na.rm = TRUE) %>%
+  ifelse(is.nan(.), NA, .)
+
+# Core Need
+df$DaytimeNeedFullfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFullfillment)] <-
+  df$KeyNeedFullfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFullfillment)]
+
+df$InteractionNeedFullfillment <- NA 
+df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFullfillment)] <- 
+  df$KeyNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFullfillment)]
+
 # Pro-Sociality
@@
 
+df$IntergroupContact <- (df$IntergroupContact-2)*-1
+
 # -------------------------------------------------------------------------------------------------------------
@@
 df$RelatednessInteraction <-
-  rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = T)
-df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <-
-  NA
+  rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = TRUE)
+df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
 # Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
@@
 df$ProSo <-
-  rowMeans(df[, c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm = T)
+  rowMeans(df[, c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm = TRUE)
 # Anti-Sociality
 df$AntiSo <-
-  rowMeans(df[, c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm = T)
+  rowMeans(df[, c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm = TRUE)
 
@@
 ```{r studentDataSelection}
-
+varListStudent <- 
+  c(
+    "IntergroupContact",
+    "InteractionDum",
+    # "KeyNeedFullfillment",
+    # "KeyNeedImp",
+    "DaytimeNeedFullfillment",
+    #"DaytimeNeedImportance",
+    "InteractionNeedFullfillment",
+    #"InteractionNeedImportance",
+    "AntiSo1",
+    "AntiSo2",
+    "AntiSo3",
+    "AntiSo4",
+    "AntiSo5",
+    "AntiSo6",
+    "AntiSo7",
+    "ProSo1",
+    "ProSo2",
+    "ProSo3",
+    "ProSo4",
+    "StudentGoal01",
+    "StudentGoal02",
+    "StudentGoal03",
+    "StudentGoal04",
+    "StudentGoal05",
+    "StudentGoal06",
+    "StudentGoal07",
+    "StudentGoal08",
+    "StudentGoal09",
+    "StudentGoal10",
+    "Autonomy",
+    "Competence",
+    # Break relatedness up into interaction vs. no interaction(?)
+    # vars are already calculated
+    "Relatedness",
+    "AttitudesDutch",
+    "AttitudesPartner",
+    "ExWB",
+    #"angry",
+    #"afraid",
+    #"energy",
+    #"lonelyAlways",
+    "Event"
+  )
+
+varNameListStudent <- 
+  c(
+    "IntergroupContact" = "Dutch Interaction",
+    "InteractionDum" = "Any Interaction",
+    # "KeyNeedFulfillment" = "Key Need Fulfillment",
+    # "KeyNeedImp" = "Key Need Importance",
+    "DaytimeNeedFullfillment" = "Daytime Core Motive Fulfillment",
+    #"DaytimeNeedImportance" = "Daytime Core Motive Importance",
+    "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
+    #"InteractionNeedImportance" = "Interaction Core Motive Importance",
+    "AntiSo1" = "Put someone down",
+    "AntiSo2" = "Show little attention in someones opinion",
+    "AntiSo3" = "Demeaning remarks",
+    "AntiSo4" = "Inpropperly addressing someone",
+    "AntiSo5" = "Ignored or excluded someone",
+    "AntiSo6" = "Doubt someones judgement",
+    "AntiSo7" = "Unwanted attempts of personal matters",
+    "ProSo1" = "Listen to someones problems",
+    "ProSo2" = "Cheer someone up",
+    "ProSo3" = "Help someone get things done",
+    "ProSo4" = "Help someone with responsibilities",
+    "StudentGoal01" = "Social support and connectedness",
+    "StudentGoal02" = "Romantic or sexual relationship",
+    "StudentGoal03" = "Academic",
+    "StudentGoal04" = "Career",
+    "StudentGoal05" = "Financial",
+    "StudentGoal06" = "Health and fitness",
+    "StudentGoal07" = "Leasure and fun",
+    "StudentGoal08" = "Personal improvement and growth",
+    "StudentGoal09" = "Service and help",
+    "StudentGoal10" = "Spiritual or religious",
+    "Autonomy" = "Autonomy",
+    "Competence" = "Competence",
+    # Break relatedness up into interaction vs. no interaction(?)
+    # vars are already calculated
+    "Relatedness" = "Relatedness",
+    "AttitudesDutch" = "Outgroup Attitude",
+    "AttitudesPartner" = "Attitude Interaction Partner",
+    "ExWB" = "Sadness Happiness",
+    #"angry" = "Anger",
+    #"afraid" = "Anxiety",
+    #"energy" = "Energy",
+    #"lonelyAlways" = "Loneliness",
+    "Event" = "Positive or negative event present"
+  )
+
+dtStudents$viz <- list()
+
+dtStudents$vizall <- dtStudents$full %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    all_of(varListStudent)
+  )
+
+dtStudents$viz$Interaction <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    IntergroupContact,
+    InteractionDum
+  )
+
+dtStudents$viz$KeyNeed <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    DaytimeNeedFullfillment,
+    InteractionNeedFullfillment
+  )
+
+dtStudents$viz$AntiSocialBehaviorPt1 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    AntiSo1,
+    AntiSo2,
+    AntiSo3
+  )
+
+dtStudents$viz$AntiSocialBehaviorPt2 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    AntiSo4,
+    AntiSo5
+  )
+
+dtStudents$viz$AntiSocialBehaviorPt3 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    AntiSo6,
+    AntiSo7
+  )
+
+dtStudents$viz$ProSocialBehaviorPt1 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    ProSo1,
+    ProSo2
+  )
+
+dtStudents$viz$ProSocialBehaviorPt2 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    ProSo3,
+    ProSo4
+  )
+
+dtStudents$viz$GoalsPt1 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    StudentGoal01,
+    StudentGoal02
+  )
+
+dtStudents$viz$GoalsPt2 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    StudentGoal03,
+    StudentGoal04
+  )
+
+dtStudents$viz$GoalsPt3 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    StudentGoal05,
+    StudentGoal06
+  )
+
+dtStudents$viz$GoalsPt4 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    StudentGoal07,
+    StudentGoal08
+  )
+
+dtStudents$viz$GoalsPt5 <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    StudentGoal09,
+    StudentGoal10
+  )
+
+dtStudents$viz$SDT <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    Autonomy,
+    Relatedness, 
+    Competence
+  )
+
+dtStudents$viz$Attitudes <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    starts_with("Attitudes")
+  )
+
+dtStudents$viz$WB <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    ExWB
+  )
+
+dtStudents$viz$Event <- dtStudents$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    Event
+  )
+
+```
+
+```{r studentMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
+dtStudents$vizPlots <-
+  lapply(seq_along(dtStudents$viz),
+         function(i)
+           meanVarPlot(dataIn = dtStudents$viz[[i]], 
+                       name = names(dtStudents$viz)[[i]], 
+                       namList = varNameListStudent))
+dtStudents$vizPlots
 ```
@@
   c(
+    "IntergroupContact",
+    "InteractionDum",
     # "KeyNeedFulfillment",
@@
   c(
+    "IntergroupContact" = "Dutch Interaction",
+    "InteractionDum" = "Any Interaction",
     # "KeyNeedFulfillment" = "Key Need Fulfillment",
@@
     TIDnum,
-    varListMedical
-  )
-
-dtMedical$viz$DaytimeNeed <- dtMedical$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    starts_with("DaytimeNeed")
-  )
-
-dtMedical$viz$InteractionNeed <- dtMedical$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    starts_with("InteractionNeed")
+    all_of(varListMedical)
+  )
+
+dtMedical$viz$Interaction <- dtMedical$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    IntergroupContact,
+    InteractionDum
+  )
+
+dtMedical$viz$KeyNeedFulfillment <- dtMedical$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    DaytimeNeedFullfillment,
+    InteractionNeedFullfillment
+  )
+
+dtMedical$viz$KeyNeedImportance <- dtMedical$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    DaytimeNeedImportance,
+    InteractionNeedImportance
   )
@@
 
-lWideToLong <- function(data, ...){
-  reshape2::melt(
-    data, 
-    id.vars = c("PID", "TIDnum")
-  )
-}
-
-dtMedical$vizLong <- lapply(
-  dtMedical$viz,
-  lWideToLong
-)
-
 ```
 
-```{r MeanVizFun}
-
-lWideToLong <- function(data, ...){
-  reshape2::melt(
-    data, 
-    id.vars = c("PID", "TID", "TIDnum")
-  )
-}
-
-meanVarPlot <- function(dataIn, name, namList,...){
-  # dataIn <- dtMedical$viz$InteractionNeed
-  # name <- "Interaction Core Need"
-  # namList <- varNameListMedical
-  dataLong <- dataIn %>%
-    mutate(TID = stri_replace_all_regex(
-      TID,
-      pattern = c('Morning', 'Afternoon'),
-      replacement = c('12:00:00', '19:00:00'),
-      vectorize = FALSE
-    ) %>% as.POSIXct) %>%
-    lWideToLong %>%
-    mutate(variable = recode(variable, !!!namList))
-
-  pTidMain <-
-    ggplot(dataLong,
-           aes(x = TIDnum, y = value, color = variable)) +
-    geom_point(alpha = 0) +
-    stat_summary(fun = mean, geom = "line") +
-    labs(y = "Mean per Day",
-         x = "Time Index") +
-    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
-    ggthemes::scale_colour_calc() +
-    scale_x_continuous(breaks = seq(0, max(dataLong$TIDnum), 25)) +
-    theme_Publication() +
-    theme(
-      panel.border = element_rect(colour = "black"),
-      plot.margin = margin(0, 10, 10, 10, "mm")
-    )
-  pTidTop <-
-    ggplot(dataLong %>% filter(!is.na(value)), aes(x = TIDnum, y = (..count.. / max(count)) * 100),
-           color = variable) +
-    geom_freqpoly(binwidth = 2) +
-    geom_hline(yintercept = 80,
-               linetype = "dashed",
-               color = "black") +
-    annotate(
-      # add white background
-      "label",
-      x = max(dataLong$TIDnum),
-      y = 80,
-      label = "80%",
-      vjust = 0.5,
-      hjust = -1,
-      label.size = NA
-    )  +
-    coord_cartesian(clip = "off") +
-    labs(y = "Percentage") +
-    #scale_colour_manual(values = rep("black", 20)) +
-    theme_classic() +
-    theme(
-      legend.position = "none",
-      axis.line.x = element_blank(),
-      axis.title.x = element_blank(),
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank(),
-      plot.margin = margin(0, 10, 0, 10, "mm")
-    )
-
-  pTidTitle <- cowplot::ggdraw() +
-    cowplot::draw_label(
-      paste("Variable Set:", name, "[Time Id Plot]"),
-      fontface = 'bold',
-      x = 0.5,
-      hjust = 0.5
-    )
-
-  pTid <- cowplot::plot_grid(
-    pTidTitle,
-    pTidTop,
-    pTidMain,
-    align = "v",
-    nrow = 3,
-    rel_heights = c(0.1, 1 / 5, 4 / 5)
-  )
-
-  pDateMain <-
-    ggplot(dataLong,
-           aes(x = TID, y = value, color = variable)) +
-    geom_point(alpha = 0) +
-    stat_summary(fun = mean, geom = "line") +
-    labs(y = "Mean per Day",
-         x = "Date") +
-    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
-    ggthemes::scale_colour_calc() +
-    scale_x_datetime(breaks = scales::date_breaks("10 day")) +
-    theme_Publication() +
-    theme(
-      panel.border = element_rect(colour = "black"),
-      plot.margin = margin(0, 10, 10, 10, "mm")
-    )
-  pDateTop <-
-    ggplot(dataLong,
-           aes(
-             x = TID,
-             y = (..count.. / max(count)) * 100,
-             color = variable
-           )) +
-    geom_freqpoly(binwidth = 1 * 3600 * 24) +
-    geom_hline(yintercept = 80,
-               linetype = "dashed",
-               color = "black") +
-    annotate(
-      # add white background
-      "label",
-      x = max(unique(dataLong$TID)),
-      y = 80,
-      label = "80%",
-      vjust = 0.5,
-      hjust = -1,
-      label.size = NA
-    )  +
-    coord_cartesian(clip = "off") +
-    labs(y = "Percentage") +
-    scale_colour_manual(values = rep("black", 20)) +
-    theme_classic() +
-    theme(
-      legend.position = "none",
-      axis.line.x = element_blank(),
-      axis.title.x = element_blank(),
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank(),
-      plot.margin = margin(0, 10, 0, 10, "mm")
-    )
-
-  pDateTitle <- cowplot::ggdraw() +
-    cowplot::draw_label(
-      paste("Variable Set: Key-Need", "[Date Plot]"),
-      fontface = 'bold',
-      x = 0.5,
-      hjust = 0.5
-    )
-
-  pDate <- cowplot::plot_grid(
-    pDateTitle,
-    pDateTop,
-    pDateMain,
-    align = "v",
-    nrow = 3,
-    rel_heights = c(0.1, 1 / 5, 4 / 5)
-  )
-
-  list(pTid, pDate)
-  pTid
-}
-```
-
 ```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
-#, warning=FALSE, message=FALSE, error=FALSE
-# meanVarPlot(dtMedical$viz$DaytimeNeed, "Key-Need")
-# meanVarPlot(dtMedical$viz$AntiSocialBehavior, "Anti-Social Behavior")
-
 dtMedical$vizPlots <-

### 2022-05-05 scripts/functions/meanViz.R

@@ -1,0 +1,155 @@
+lWideToLong <- function(data, ...){
+  reshape2::melt(
+    data, 
+    id.vars = c("PID", "TID", "TIDnum")
+  )
+}
+
+
+meanVarPlot <- function(dataIn, name, namList,...){
+  # dataIn <- dtMedical$viz$InteractionNeed
+  # name <- "Interaction Core Need"
+  # namList <- varNameListMedical
+  dataLong <- dataIn %>%
+    mutate(TID = stri_replace_all_regex(
+      TID,
+      pattern = c('Morning', 'Afternoon'),
+      replacement = c('12:00:00', '19:00:00'),
+      vectorize = FALSE
+    ) %>% as.POSIXct) %>%
+    lWideToLong %>%
+    mutate(variable = recode(variable, !!!namList))
+  
+  
+  
+  pTidMain <-
+    ggplot(dataLong,
+           aes(x = TIDnum, y = value, color = variable)) +
+    geom_point(alpha = 0) +
+    stat_summary(fun = mean, geom = "line") +
+    labs(y = "Mean per Day",
+         x = "Time Index") +
+    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
+    ggthemes::scale_colour_calc() +
+    scale_x_continuous(breaks = seq(0, max(dataLong$TIDnum), 25)) +
+    theme_Publication() +
+    theme(
+      panel.border = element_rect(colour = "black"),
+      plot.margin = margin(0, 10, 10, 10, "mm")
+    )
+  pTidTop <-
+    ggplot(dataLong %>% filter(!is.na(value)), aes(x = TIDnum, y = (..count.. / max(count)) * 100),
+           color = variable) +
+    geom_freqpoly(binwidth = 2) +
+    geom_hline(yintercept = 80,
+               linetype = "dashed",
+               color = "black") +
+    annotate(
+      # add white background
+      "label",
+      x = max(dataLong$TIDnum),
+      y = 80,
+      label = "80%",
+      vjust = 0.5,
+      hjust = -1,
+      label.size = NA
+    )  +
+    coord_cartesian(clip = "off") +
+    labs(y = "Percentage") +
+    #scale_colour_manual(values = rep("black", 20)) +
+    theme_classic() +
+    theme(
+      legend.position = "none",
+      axis.line.x = element_blank(),
[…]

### 2022-05-05 Supplemental-Material-A-Annotated-Analysis.Rmd
sd plots added

@@
 
-# Mean Plots per Item
+# **Plots per Item**
 
@@
 
+### Mean Plots
+
 ```{r studentMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtStudents$vizPlots <-
+dtStudents$vizPlotsMean <-
   lapply(seq_along(dtStudents$viz),
          function(i)
-           meanVarPlot(dataIn = dtStudents$viz[[i]], 
+           statVarPlot(dataIn = dtStudents$viz[[i]], 
                        name = names(dtStudents$viz)[[i]], 
-                       namList = varNameListStudent))
-dtStudents$vizPlots
+                       namList = varNameListStudent,
+                       stat = "mean")
+         )
+dtStudents$vizPlotsMean
+```
+
+### SD Plots
+
+```{r studentSdViz, echo=FALSE, results='hide',fig.keep='all'} 
+dtStudents$vizPlotsSd <-
+  lapply(seq_along(dtStudents$viz),
+         function(i)
+           statVarPlot(dataIn = dtStudents$viz[[i]], 
+                       name = names(dtStudents$viz)[[i]], 
+                       namList = varNameListStudent,
+                       stat = "sd")
+         )
+dtStudents$vizPlotsSd
 ```
@@
 
+### Mean Plots
+
 ```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtMedical$vizPlots <-
+dtMedical$vizPlotsMean <-
   lapply(seq_along(dtMedical$viz),
          function(i)
-           meanVarPlot(dataIn = dtMedical$viz[[i]], 
+           statVarPlot(dataIn = dtMedical$viz[[i]], 
                        name = names(dtMedical$viz)[[i]], 
-                       namList = varNameListMedical))
-dtMedical$vizPlots
+                       namList = varNameListMedical,
+                       stat = "mean")
+         )
+dtMedical$vizPlotsMean
+```
+
+### SD Plots
+
+```{r medicalSdViz, echo=FALSE, results='hide',fig.keep='all'} 
+dtMedical$vizPlotsSd <-
+  lapply(seq_along(dtMedical$viz),
+         function(i)
+           statVarPlot(dataIn = dtMedical$viz[[i]], 
+                       name = names(dtMedical$viz)[[i]], 
+                       namList = varNameListMedical,
+                       stat = "sd")
+         )
+dtMedical$vizPlotsSd
 ```
@@
 
-# **References**
+<!-- # **References**   -->
 

### 2022-05-05 scripts/functions/meanViz.R

@@ -6,10 +6,14 @@ lWideToLong <- function(data, ...){
 }
 
 
-meanVarPlot <- function(dataIn, name, namList,...){
-  # dataIn <- dtMedical$viz$InteractionNeed
-  # name <- "Interaction Core Need"
+statVarPlot <- function(dataIn, name, namList, stat = "mean", ...){
+  # dataIn <- dtMedical$viz$KeyNeedFulfillment
+  # name <- "Core Need Fulfillment"
   # namList <- varNameListMedical
+  # stat <- "sd"
+  
+  varNames <- names(dataIn)[!names(dataIn) %in% c("PID", "TID", "TIDnum")]
+  
   dataLong <- dataIn %>%
     mutate(TID = stri_replace_all_regex(
       TID,
@@ -20,27 +24,32 @@ meanVarPlot <- function(dataIn, name, namList,...){
     lWideToLong %>%
     mutate(variable = recode(variable, !!!namList))
   
+  dataLongNaRm <- dataLong %>% filter(!is.na(value))
+  
+  nTimeLong <- dataLongNaRm %>%
+    select(-PID) %>%
+    group_by(TIDnum, TID, variable) %>%
+    summarise(
+      n = n()
+    ) %>%
+    mutate(
+      perc = n/length(unique(dataLongNaRm$PID))*100
+    ) 
   
   
-  pTidMain <-
-    ggplot(dataLong,
-           aes(x = TIDnum, y = value, color = variable)) +
-    geom_point(alpha = 0) +
-    stat_summary(fun = mean, geom = "line") +
-    labs(y = "Mean per Day",
-         x = "Time Index") +
-    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
+  
+  pTidTop <- ggplot(nTimeLong, aes(x = TIDnum, y = perc, color = variable, linetype = variable)) +
+    geom_line(
+      stat="smooth",
+      method = "loess",
+      formula = 'y ~ x',
+      span = 0.1,
+      se = FALSE,
+      #size = 1,
+      #color = "black",
+      alpha = 0.75
+    ) +
     ggthemes::scale_colour_calc() +
-    scale_x_continuous(breaks = seq(0, max(dataLong$TIDnum), 25)) +
-    theme_Publication() +
-    theme(
-      panel.border = element_rect(colour = "black"),
-      plot.margin = margin(0, 10, 10, 10, "mm")
-    )
-  pTidTop <-
[…]

### 2022-05-05 scripts/functions/sdViz.R

@@ -1,0 +1,160 @@
+
+sdVarPlot <- function(dataIn, name, namList,...){
+  # dataIn <- dtMedical$viz$KeyNeedFulfillment
+  # name <- "Core Need Fulfillment"
+  # namList <- varNameListMedical
+  
+  varNames <- names(dataIn)[!names(dataIn) %in% c("PID", "TID", "TIDnum")]
+  
+  dataIn %>%
+    mutate(TID = stri_replace_all_regex(
+      TID,
+      pattern = c('Morning', 'Afternoon'),
+      replacement = c('12:00:00', '19:00:00'),
+      vectorize = FALSE
+    ) %>% as.POSIXct) %>%
+    select(-PID) %>%
+    #filter_at(vars(varNames), all_vars(!is.na(.))) %>%
+    group_by(
+      TIDnum, TID
+    ) %>%
+    summarise(
+      n = n(),
+      #across(varNames, ~mean(.x, na.rm = TRUE), .names = "mean_{.col}"),
+      across(varNames, ~n(.x), .names = "n_{.col}")
+    ) 
+
+  
+  
+  
+  pTidMain <-
+    ggplot(dataLong,
+           aes(x = TIDnum, y = value, color = variable)) +
+    geom_point(alpha = 0) +
+    stat_summary(fun = mean, geom = "line") +
+    labs(y = "Mean per Day",
+         x = "Time Index") +
+    #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
+    ggthemes::scale_colour_calc() +
+    scale_x_continuous(breaks = seq(0, max(dataLong$TIDnum), 25)) +
+    theme_Publication() +
+    theme(
+      panel.border = element_rect(colour = "black"),
+      plot.margin = margin(0, 10, 10, 10, "mm")
+    )
+  pTidTop <-
+    ggplot(dataLong %>% filter(!is.na(value)), aes(x = TIDnum, y = (..count.. / max(count)) * 100),
+           color = variable) +
+    geom_freqpoly(binwidth = 2) +
+    geom_hline(yintercept = 80,
+               linetype = "dashed",
+               color = "black") +
+    annotate(
+      # add white background
+      "label",
+      x = max(dataLong$TIDnum),
+      y = 80,
+      label = "80%",
+      vjust = 0.5,
+      hjust = -1,
+      label.size = NA
+    )  +
+    coord_cartesian(clip = "off") +
[…]

### 2022-05-06 Supplemental-Material-A-Annotated-Analysis.Rmd
typo
study one and annotations added

@@
 title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
-subtitle: "Supplemental Material for 'Qualitative Goals'"
+subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
 author:
-- ██████ ██████████^1^
-- ██████████ ████████^1^
-- █████ █████████^1^
-- █████ ██ █████^1^
-- ███ ███████^1^
-- ^1^██████████ ██ █████████, Department of Psychology
+- Jannis Kreienkamp^1,3^
+- Kai Epstude^1,3^
+- Laura F. Bringmann^1,3^
+- Maximilian Agostini^1,3^
+- Peter de Jonge^1,3^
+- Rei Tendeiro-Monden^2,3^
+- ^1^University of Groningen, Department of Psychology
+- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
+- ^3^Author order TBD [currently in first name alphabetical order]
 - "Author Information:"
-- "Correspondence concerning this article should be addressed to ██████ ██████████, ██████████ ██ ██████████, █████████ ██ ██████████, █████ ██████████ ███, ████ ██ █████████ ████████████████.
-E-mail: █████████████████"
+- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
+E-mail: j.kreienkamp@rug.nl"
 - 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
-- 'The data repository for this manuscript is available at <a href="https://osf.io/TBD" target="_blank">osf.io/TBD</a>'
+- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
 - 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
@@
 
-<br/>
+</br>
+
+--------------------------------------------------------------------
+
+</br>
 
@@
   "ExWB_1",
+  "MDMQ.v1_1",
+  "MDMQ.v1_2",
+  "MDMQ.v1_3",
+  #"MDMQ.v1_3r",
+  "MDMQ.v1_4",
+  #"MDMQ.v1_4r",
+  "MDMQ.v1_5",
+  #"MDMQ.v1_5r",
+  "MDMQ.v1_6",
+  "MDMQ.v1_7",
+  #"MDMQ.v1_7r",
+  "MDMQ.v1_8",
+  "MDMQ.v1_9",
+  #"MDMQ.v1_9r",
+  "MDMQ.v1_10",
+  "MDMQ.v1_11",
+  #"MDMQ.v1_11r",
+  "MDMQ.v1_12",
+  "MDMQ.v1_13",
+  "MDMQ.v2_1",
+  #"MDMQ.v2_1r",
+  "MDMQ.v2_2",
+  "MDMQ.v2_3",
+  "MDMQ.v2_4",
+  #"MDMQ.v2_4r",
+  "MDMQ.v2_5",
+  "MDMQ.v2_6",
+  #"MDMQ.v2_6r",
+  "MDMQ.v2_7",
+  #"MDMQ.v2_7r",
+  "MDMQ.v2_8",
+  "MDMQ.v2_9",
+  "MDMQ.v2_10",
+  #"MDMQ.v2_10r",
+  "MDMQ.v2_11",
+  #"MDMQ.v2_11r",
+  "MDMQ.v2_12",
+  "MDMQ.v2_13",
   "alertness1",
@@
 
+# Make time IDs consistent with later studies
+df$TID <- paste(df$PPDate, str_to_title(df$daytime))
+df$TIDnum <- as.numeric(factor(df$TID %>% gsub(" Morning", "-A", .) %>% gsub(" Afternoon", "-B", .)))
+
+# Make contact dummies consistent with later studies
+df$IntergroupContact <- df$Contact_dum
+
+df <- df %>%
+  rowwise() %>%
+  mutate(
+    InteractionDum = sum(Contact_dum, inNonDutch, na.rm = FALSE),
+    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum),
+    alertness = sum(alertness1, alertness2, na.rm = TRUE),
+    calmness = sum(calmness1, calmness2, na.rm = TRUE),
+    valence = sum(valence1, valence2, na.rm = TRUE)
+  ) %>%
+  ungroup()
+
 # Need scales
+df$InteractionNeedFullfillment <- df$keymotive_fulfillemt_1
+df$DaytimeNeedFullfillment <- df$keyMotive_noInt_fulf_1
+
 df$keyMotiveFulfilled <- rowSums(df[, c("keymotive_fulfillemt_1", "keyMotive_noInt_fulf_1")], na.rm = T)
@@
 df$relatedness_1 <- rowMeans(df[, c("relatedness_other_1", "relatedness_self_1")], na.rm = T)
+
+df$Autonomy <- df$autonomy.daily.all
+df$Competence <- df$competence.daily.all
+df$Relatedness <- df$relatedness.daily.all
+
+# Make Outgroup Attitudes consistent with later studies
+df$AttitudesDutch <- df$thermometerDutch_1
+df$AttitudesPartner <- df$thermometerDutchInt_2
+
+# Make experienced well-being consistent with later studies
+df$ExWB <- df$ExWB_1
 
@@
 ggplot(df, aes(x = AllportsCondition)) +
-  geom_histogram()
+  geom_histogram() +
+  theme_Publication()
 
@@
 
+As one of our main analyses is a three-mode principal component analysis (3MPCA) we begin by assessing the amount of missing data in each of the ESM studies.
+To assess the missingness in detail, we prepare a data availability table of whether data is available or missing for each possible measurement point (data + morning/afternoon) for all individual participants.
+We then export the data availability tables as comma separated value files (.csv) to be assessed in detail using a spreadsheet program (such as MS Excel).
+
 ```{r workerDataAvailability}
@@
 
+</br>
+
+--------------------------------------------------------------------
+
+</br>
+
 # **Plots per Item**
 
-## Study 1 (Worker):
+The three-mode PCA ultimately aims to explain the most variance with the fewest components (by maximizing the variance of data mapped in a lower dimensional space --- often by maximally separating groups of cases with similar variable patterns).
+Importantly, this procedure depends on variances and covariances to be present within the data.
+As one of our three modes is time and variance over time is not always given (e.g., because a variable is stable over the measured time span, such as personality over a month), it becomes important to assess whether a dimension reduction is possible for the variables measured during the measured time period.
+We, thus, start by visually inspecting the average changes and variance developments of all potentially relevant variables.
+We hope so see changes within the means and standard deviations over the measurement period.
+
+## Study 1 (Worker)
+
+The first study, utilized not only the smallest sample but also a slightly different set of measured variables compared to the later two studies.
+Several variables are consistent across all three studies (e.g., types of social interactions, interaction needs, self-determiantion theory needs, outgroup attitudes, and experienced well-being).
+Study 1 additionally included several emotion and mood measurements but fewer cognitive and behavioral measures than studies two and three.
+
+We first select all potentially relevant variables and their labels, and then group the variables in small groups of one to three variables for plotting.
 
 ```{r workerDataSelection}
-
+varListWorker <- 
+  c(
+    "IntergroupContact",
+    "InteractionDum",
+    "DaytimeNeedFullfillment",
+    "InteractionNeedFullfillment",
+    "Autonomy",
+    "Competence",
+    "Relatedness",
+    "AttitudesDutch",
+    "AttitudesPartner",
+    "ExWB",
+    "alertness",
+    "calmness",
+    "valence",
+    "MDMQ.v1_1",
+    "MDMQ.v1_2",
+    "MDMQ.v1_3",
+    "MDMQ.v1_4",
+    "MDMQ.v1_5",
+    "MDMQ.v1_6",
+    "MDMQ.v1_7",
+    "MDMQ.v1_8",
+    "MDMQ.v1_9",
+    "MDMQ.v1_10",
+    "MDMQ.v1_11",
+    "MDMQ.v1_12",
+    "MDMQ.v1_13",
+    "MDMQ.v2_1",
+    "MDMQ.v2_2",
+    "MDMQ.v2_3",
+    "MDMQ.v2_4",
+    "MDMQ.v2_5",
+    "MDMQ.v2_6",
+    "MDMQ.v2_7",
+    "MDMQ.v2_8",
+    "MDMQ.v2_9",
+    "MDMQ.v2_10",
+    "MDMQ.v2_11",
+    "MDMQ.v2_12",
+    "MDMQ.v2_13"
+  )
+
+varNameListWorker <- 
+  c(
+    "IntergroupContact" = "Dutch Interaction",
+    "InteractionDum" = "Any Interaction",
+    "DaytimeNeedFullfillment" = "Daytime Core Motive Fulfillment",
+    "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
+    "Autonomy" = "Autonomy",
+    "Competence" = "Competence",
+    "Relatedness" = "Relatedness",
+    "AttitudesDutch" = "Outgroup Attitude",
+    "AttitudesPartner" = "Attitude Interaction Partner",
+    "ExWB" = "Sadness Happiness",
+    "alertness" = "Alertness",
+    "calmness" = "Calmness",
+    "valence" = "Emotional Valence",
+    "MDMQ.v1_1" = "content",
+    "MDMQ.v1_2" = "rested",
+    "MDMQ.v1_3" = "restless",
+    "MDMQ.v1_4" = "bad",
+    "MDMQ.v1_5" = "worn-out",
+    "MDMQ.v1_6" = "composed",
+    "MDMQ.v1_7" = "tired",
+    "MDMQ.v1_8" = "great",
+    "MDMQ.v1_9" = "uneasy",
+    "MDMQ.v1_10" = "energetic",
+    "MDMQ.v1_11" = "uncomfortable",
+    "MDMQ.v1_12" = "relaxed",
+    "MDMQ.v1_13" = "alive and vital",
+    "MDMQ.v2_1" = "sleepy",
+    "MDMQ.v2_2" = "good",
+    "MDMQ.v2_3" = "at ease",
+    "MDMQ.v2_4" = "unhappy",
+    "MDMQ.v2_5" = "alert",
+    "MDMQ.v2_6" = "discontent",
+    "MDMQ.v2_7" = "tense",
+    "MDMQ.v2_8" = "fresh",
+    "MDMQ.v2_9" = "happy",
+    "MDMQ.v2_10" = "nervous",
+    "MDMQ.v2_11" = "exhausted",
+    "MDMQ.v2_12" = "calm",
+    "MDMQ.v2_13" = "alive and vital"
+  )
+
+dtWorker$viz <- list()
+
+dtWorker$vizall <- dtWorker$full %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    all_of(varListWorker)
+  )
+
+dtWorker$viz$Interaction <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    IntergroupContact,
+    InteractionDum
+  )
+
+dtWorker$viz$KeyNeed <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    DaytimeNeedFullfillment,
+    InteractionNeedFullfillment
+  )
+
+dtWorker$viz$SDT <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    Autonomy,
+    Relatedness, 
+    Competence
+  )
+
+dtWorker$viz$Attitudes <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    starts_with("Attitudes")
+  )
+
+dtWorker$viz$WB <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    ExWB
+  )
+
+dtWorker$viz$MoodOverall <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    alertness,
+    calmness,
+    valence
+  )
+
+dtWorker$viz$MoodV01Pt01 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v1_1,
+    MDMQ.v1_2
+  )
+
+dtWorker$viz$MoodV01Pt02 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v1_3,
+    MDMQ.v1_4
+  )
+
+dtWorker$viz$MoodV01Pt03 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v1_5,
+    MDMQ.v1_6
+  )
+
+dtWorker$viz$MoodV01Pt04 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v1_7,
+    MDMQ.v1_8
+  )
+
+dtWorker$viz$MoodV01Pt05 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v1_9,
+    MDMQ.v1_10
+  )
+
+dtWorker$viz$MoodV01Pt06 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v1_11,
+    MDMQ.v1_12,
+    MDMQ.v1_13
+  )
+
+dtWorker$viz$MoodV02Pt01 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v2_1,
+    MDMQ.v2_2
+  )
+
+dtWorker$viz$MoodV02Pt02 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v2_3,
+    MDMQ.v2_4
+  )
+
+dtWorker$viz$MoodV02Pt03 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v2_5,
+    MDMQ.v2_6
+  )
+
+dtWorker$viz$MoodV02Pt04 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v2_7,
+    MDMQ.v2_8
+  )
+
+dtWorker$viz$MoodV02Pt05 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v2_9,
+    MDMQ.v2_10
+  )
+
+dtWorker$viz$MoodV02Pt06 <- dtWorker$vizall %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    MDMQ.v2_11,
+    MDMQ.v2_12,
+    MDMQ.v2_13
+  )
 ```
 
-## Study 2 (Student):
+### Mean Plots
+
+We begin by plotting the item means of all variables.
+We always only display one to three variables jointly to improve visual inspection.
+We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point.
+We additionally plot the data availability of each variable as a marginal plot above the main plot.
+The displayed time axis is the measurement index rather than the actual date to ensure comparability of the changes between each measurement.
+Plots with the raw date-time axis are saved and can be displayed if necessary.
+
+```{r workerMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
+dtWorker$vizPlotsMean <-
+  lapply(seq_along(dtWorker$viz),
+         function(i)
+           statVarPlot(dataIn = dtWorker$viz[[i]], 
+                       name = names(dtWorker$viz)[[i]], 
+                       namList = varNameListWorker,
+                       stat = "mean")
+         )
+dtWorker$vizPlotsMean
+```
+
+### SD Plots
+
+To illustrate the developments of variance around the previously displayed means, we then plot the item standard deviations of all variables.
+We, thus, plot the average variation from the mean for each individual variable, again summarizing the responses of all participants at any given time point.
+We retain the same variable grouping, the data availability line plot, and the measurement index as the x axis.
+
+```{r WorkerSdViz, echo=FALSE, results='hide',fig.keep='all'} 
+dtWorker$vizPlotsSd <-
+  lapply(seq_along(dtWorker$viz),
+         function(i)
+           statVarPlot(dataIn = dtWorker$viz[[i]], 
+                       name = names(dtWorker$viz)[[i]], 
+                       namList = varNameListWorker,
+                       stat = "sd")
+         )
+dtWorker$vizPlotsSd
+```
+
+## Study 2 (Student)
+
+During the second study we collected the largest sample and additionally collected a range of motivations as well as pro-social and anti-social behaviors.
+
+We, again, select all potentially relevant variables and their labels, and then group the variables in small groups of one to three variables for plotting.
 
@@
 
+We again begin by plotting the item means of all variables. The plotting method is identical to that of Study 1.
+
 ```{r studentMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
@@
 
+We then also plot the average variation around the mean for each time point and variable. The plotting method is identical to that of Study 1.
+
 ```{r studentSdViz, echo=FALSE, results='hide',fig.keep='all'} 
@@
 
-## Study 3 (Medical):
+## Study 3 (Medical)
+
+The third study uses a similar set up as Study 2 but targets a more vulnerable sample (of young medical professionals).
+While most key variables are identical to Study 2, in this last study we additionally collected several cognitive evaluations (e.g., importance ratings) and emotional status measures (e.g., anger, nervousness, energy, and loneliness).
+
+We, again, select all potentially relevant variables and their labels, and then group the variables in small groups of one to three variables for plotting.
 
@@
 
+As with the previous studies, we begin by plotting the item means of all variables. The plotting method is identical to that of Studies 1 and 2.
+
 ```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
@@
 
+Lastly, also again plot the item standard deviations of all variables at each time point. The plotting procedures are again identical to Studies 1 and 2.
+
 ```{r medicalSdViz, echo=FALSE, results='hide',fig.keep='all'} 
@@
 ```
+
+</br>
+
+--------------------------------------------------------------------
+
+</br>
 

### 2022-05-06 scripts/functions/meanViz.R

@@ -7,10 +7,10 @@ lWideToLong <- function(data, ...){
 
 
 statVarPlot <- function(dataIn, name, namList, stat = "mean", ...){
-  # dataIn <- dtMedical$viz$KeyNeedFulfillment
-  # name <- "Core Need Fulfillment"
-  # namList <- varNameListMedical
-  # stat <- "sd"
+  # dataIn <- dtWorker$viz$MoodOverall
+  # name <- "Mood"
+  # namList <- varNameListWorker
+  # stat <- "mean"
   
   varNames <- names(dataIn)[!names(dataIn) %in% c("PID", "TID", "TIDnum")]
   
@@ -81,7 +81,7 @@ statVarPlot <- function(dataIn, name, namList, stat = "mean", ...){
            aes(x = TIDnum, y = value, color = variable)) +
     #geom_point(alpha = 0) +
     stat_summary(fun = stat, geom = "line") +
-    labs(y = paste0(stat, " per Day"),
+    labs(y = paste0(stat),
          x = "Time Index") +
     #scale_colour_manual(values = RColorBrewer::brewer.pal(length(unique(dataLong$variable)), "Set3")) +
     ggthemes::scale_colour_calc() +
@@ -155,7 +155,7 @@ statVarPlot <- function(dataIn, name, namList, stat = "mean", ...){
            aes(x = TID, y = value, color = variable)) +
     #geom_point(alpha = 0) +
     stat_summary(fun = stat, geom = "line") +
-    labs(y = paste0(stat, " per day"),
+    labs(y = paste0(stat),
          x = "Date") +
     ggthemes::scale_colour_calc() +
     scale_x_datetime(breaks = scales::date_breaks("10 day")) +

### 2022-05-12 preregistration/Preregistration3MPCA.Rmd
preregistration added

---
title           : "Migration Experience Trajectories: A Three Mode Principle Component Analysis"
shorttitle      : "Preregistration: Migration Experience Trajectories"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
author: 
  - name        : Jannis Kreienkamp
    affiliation : "1"
  - name        : Kai Epstude
    affiliation : "1"
  - name        : Rei Tendeiro-Monden
    affiliation : "2"
  - name        : Maximilian Agostini
    affiliation : "1"
  - name        : Peter de Jonge
    affiliation : "1"
  - name        : Laura F. Bringmann
    affiliation : "1"
affiliation:
  - id          : 1
    institution : University of Groningen, Department of Psychology
  - id          : 2
    institution : Hiroshima University, Graduate School of Advanced Science and Engineering
bibliography: "../references.bib"
csl: apa.csl
biblio-style: "apalike" 
output: 
  prereg::cos_prereg:
    citation_package: natbib
editor_options: 
  chunk_output_type: console
header-includes:
   - \usepackage{amsmath, nccmath}
---

```{r setup, include=FALSE}
# R Studio Clean-Up
  cat("\014") # clear console
  #rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
  gc() # garbage collector

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file 
lib <- c("rmarkdown", "knitr", "remedy", "bookdown", "psych",
         "ggplot2", "ggthemes", "haven", "RColorBrewer", "plotly", "gridExtra",
         "sessioninfo", "tibble", "pander", "devtools",
         "data.table", "dplyr", "tidyr", "Hmisc", "kableExtra", 
         "stringr", "stringi", "reshape2", "lubridate")
invisible(lapply(lib, library, character.only = TRUE))  
rm(lib)

# Markdown Options
  knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory

# Global Chunk Options
  knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figures/',
                        echo=TRUE, warning=FALSE, message=FALSE)
```

# Study Information

## Title
<!-- Provide the working title of your study. It may be the same title that you submit for publication of your final manuscript, but it is not a requirement. The title should be a specific and informative description of a project. Vague titles such as 'Fruit fly preregistration plan' are not appropriate.

Example: Effect of sugar on brownie tastiness. -->

`r rmarkdown::metadata$title`

## Description
<!-- Please give a brief description of your study, including some background, the purpose of the of the study, or broad research questions. The description should be no longer than the length of an abstract. It can give some context for the proposed study, but great detail is not needed here for your preregistration.

Example: Though there is strong evidence to suggest that sugar affects taste preferences, the effect has never been demonstrated in brownies. Therefore, we will measure taste preference for four different levels of sugar concentration in a standard brownie recipe to determine if the effect exists in this pastry. -->

In its most essential interpretation, the intergroup contact hypothesis postulates that frequent and positive contact with an out-group reduces prejudice and increases favorable attitudes towards the other group [@Hewstone1996; @Pettigrew1998]. A key condition for these contact benefits has been that the interaction is perceived as positive --- making the interaction quality a crucial mechanism of inter-group contact [e.g., @MacInnis2015]. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for a positive contact [Allport's Optimal Contact conditions, @Allport1954b]. And indeed a major meta-analytic review showed, that intergroup contact benefits were larger when Allport's conditions were met. However, the meta-analysis also showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met [@Pettigrew2006]. It, thus, remains unclear why and how exactly Allport's conditions work.

We propose that Allport's optimal contact conditions are effective in creating positive contact because they constitute common psychological needs of the interacting individuals. If this is indeed the case, the impact of fulfilling other fundamental needs should produce similar beneficial contact results. And more importantly, the satisfaction of key situation needs during the interaction should predict the positive contact effects most strongly.

## Hypotheses
<!-- List specific, concise, and testable hypotheses. Please state if the hypotheses are directional or non-directional. If directional, state the direction. A predicted effect is also appropriate here. If a specific interaction or moderation is important to your research, you can list that as a separate hypothesis.

Example: If taste affects preference, then mean preference indices will be higher with higher concentrations of sugar. -->

1. \label{hyp:contact}Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes.
    a. \label{hyp:contactFreq}Participants with more intergroup interactions should have more favorable outgroup attitudes.
    b. \label{hyp:contactDummy}Outgroup attitudes should be more positive after an intergroup interaction compared to a non-outgroup interaction.
    c. \label{hyp:contactFreqQual}Participants with more intergroup interactions should have more favorable outgroup attitudes depending on the average interaction quality.
2. \label{hyp:AllportsConditions}Based on Allport's optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
    a. \label{hyp:AllportsPred}Based on Allport's optimal contact conditions, outgroup attitudes should be more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
    b. \label{hyp:AllportsQuality}Based on past research on the role of interaction quality, interaction quality should be perceived as more positive after intergroup interactions with equal status, common goals, collaboration, and structural support.
    c. \label{hyp:AllportsQualityMediation}Taken together, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extend be explained by interaction quality.
3. \label{hyp:keyNeed}Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
    a. \label{hyp:keyNeedPred}Outgroup attitudes should be more favorable after intergroup interactions with high key need fulfillment.
    b. \label{hyp:keyNeedQual}Interaction Quality should be perceived as more positive after intergroup interactions with higher key need fulfillment.
    c. \label{hyp:keyNeedMediation}The variance explained in outgroup attitudes by key need fulfillment should to a large extend be explained by interaction quality.
    d. \label{hyp:keyNeedContactType}The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should be stronger for intergroup interact than for non-outgroup interactions (or need fulfillment during a time of no interaction).
    e. \label{hyp:keyNeedSDT}The effect of key need fulfillment on outgroup attitudes should persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
4. \label{hyp:comparison}Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict outgroup attitudes at least as well as Allport's conditions.
    a. \label{hyp:compModel}The need model (H\ref{hyp:keyNeedPred}) should predict more variance in outgroup attitudes than the model based on Allport's conditions (H\ref{hyp:AllportsPred}).
    b. \label{hyp:compTogether}The  effect of key need fulfillment on outgroup attitudes should  persist even when taking Allport's conditions into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for equal status, common goals, collaboration, and structural support.

# Design Plan
<!-- In this section, you will be asked to describe the overall design of your study. Remember that this research plan is designed to register a single study, so if you have multiple experimental designs, please complete a separate preregistration. -->

## Study type

**Observational Study**. Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, natural experiments, and regression discontinuity designs.

## Blinding
<!-- Blinding describes who is aware of the experimental manipulations within a study. Select all that apply. Is there any additional blinding in this study? -->

No participant blinding is involved in this study.

## Study design
<!-- Describe your study design. Examples include two-group, factorial, randomized block, and repeated measures. Is it a between (unpaired), within-subject (paired), or mixed design? Describe any counterbalancing required. Typical study designs for observation studies include cohort, cross sectional, and case-control studies.

This question has a variety of possible answers. The key is for a researcher to be as detailed as is necessary given the specifics of their design. Be careful to determine if every parameter has been specified in the description of the study design. There may be some overlap between this question and the following questions. That is OK, as long as sufficient detail is given in one of the areas to provide all of the requested information. For example, if the study design describes a complete factorial, 2 X 3 design and the treatments and levels are specified previously, you do not have to repeat that information.

Example: We have a between subjects design with 1 factor (sugar by mass) with 4 levels. -->

This study used an extensive longitudinal design. Using a daily diary format, for at least 30 days participants received a short survey twice per day (at around 12pm and 7pm). We additionally included a longer pre- and post measurement survey the days before and after the extensive longitudinal data collection.

## Randomization
<!-- If you are doing a randomized study, how will you randomize, and at what level? Typical randomization techniques include: simple, block, stratified, and adaptive covariate randomization. If randomization is required for the study, the method should be specified here, not simply the source of random numbers.

Example: We will use block randomization, where each participant will be randomly assigned to one of the four equally sized, predetermined blocks. The random number list used to create these four blocks will be created using the web applications available at https://random.org. -->

No randomization is involved in this study.

# Sampling Plan
<!-- In this section we’ll ask you to describe how you plan to collect samples, as well as the number of samples you plan to collect and your rationale for this decision. Please keep in mind that the data described in this section should be the actual data used for analysis, so if you are using a subset of a larger dataset, please describe the subset that will actually be used in your study. -->

## Existing data
<!-- Preregistration is designed to make clear the distinction between confirmatory tests, specified prior to seeing the data, and exploratory analyses conducted after observing the data. Therefore, creating a research plan in which existing data will be used presents unique challenges. Please select the description that best describes your situation. Please do not hesitate to contact us if you have questions about how to answer this question (prereg@cos.io). -->

**Registration prior to accessing the data**. As of the date of submission, the data exist, but have not been accessed by you or your collaborators. Commonly, this includes data that has been collected by another researcher or institution.

## Explanation of existing data
<!-- If you indicate that you will be using some data that already exist in this study, please describe the steps you have taken to assure that you are unaware of any patterns or summary statistics in the data. This may include an explanation of how access to the data has been limited, who has observed the data, or how you have avoided observing any analysis of the specific data you will use in your study.

An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point.

Example: An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point. -->

The data was collected as part of a larger collaboration on daily intergroup relations. The researchers have had no direct access to the data, as the data was collected using an automated system using the FormR framework [@Arslan2020]. The data were then downloaded and stored securely on an encrypted device until the eventual data analysis. None of the researchers have accessed the data thus far.

## Data collection procedures
<!-- Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that donÍt include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

The answer to this question requires a specific set of instructions so that another person could repeat the data collection procedures and recreate the study population. Alternatively, if the study population would be unable to be reproduced because it relies on a specific set of circumstances unlikely to be recreated (e.g., a community of people from a specific time and location), the criteria and methods for creating the group and the rationale for this unique set of subjects should be clear.

Example: Participants will be recruited through advertisements at local pastry shops. Participants will be paid $10 for agreeing to participate (raised to $30 if our sample size is not reached within 15 days of beginning recruitment). Participants must be at least 18 years old and be able to eat the ingredients of the pastries. -->

The data was collected in a three step procedure:

1. Entry Survey: A pre-measurement questionnaire (appr. 25 minutes) including demographic information, and relations to the Dutch majority (payment: 2 Euros).
2. Experience Recaps: At least 30 days of short reflection surveys (appr. 3—5 minutes) on intergroup interactions twice a day (payment: 1 Euro per Recap; up to 2 Euros per day).
3. Conclusion Survey: On the last day, we conclude with a post-measurement questionnaire (appr. 25 minutes) with some questions on habits and reflections on the study (payment: 2 Euros).

After the initial 30 day duration, participants are offered the possibility to continue participating in the study either with payment if daily diary measures were missed during the initial study phase or without payment after a total of 60 daily diary measurements were completed. After the initial 30-day period, participants receive automated feedback visualizing the development of their own well-being, attitudes, and motive responses as an additional initiative and to give participants access to their own data and to compensate study participation.

## Sample size
<!-- Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis? For some studies, this will simply be the number of samples or the number of clusters. For others, this could be an expected range, minimum, or maximum number.

Example: Our target sample size is 280 participants. We will attempt to recruit up to 320, assuming that not all will complete the total task. -->

Our target sample size is a sum of 4,000 daily diary measurements. With 100% completion rate that would be archived with 67 participants (60 daily diary responses each). Given that we expect some incomplete daily diary measurements, we aimed to recruit 80 participants.

## Sample size rationale
<!-- This could include a power analysis or an arbitrary constraint such as time, money, or personnel. This gives you an opportunity to specifically state how the sample size will be determined. A wide range of possible answers is acceptable; remember that transparency is more important than principled justifications. If you state any reason for a sample size upfront, it is better than stating no reason and leaving the reader to "fill in the blanks." Acceptable rationales include: a power analysis, an arbitrary number of subjects, or a number based on time or monetary constraints.

Example: We used the software program G*Power to conduct a power analysis. Our goal was to obtain .95 power to detect a medium effect size of .25 at the standard .05 alpha error probability. -->

The targeted sample size depended on a combination of different factors. Different analyses were planned as part of the collaboration and budgeting was a practical constraint. Some analyses were planned based on (1) the pre- to post measurements, (2) the dynamic developments over the daily diary measurements, or (3) the contemporary effects within the daily diary measurements.

Power considerations of mixed effects model such as with extensive longitudinal data are difficult to estimate because of the complex covariance structures. Simulation studies based on the first two samples within this project indicated that with well-distributed scales, and small to medium effect sizes, 70-80 participants with at least seven daily diary measurements and a simple pre--post survey were sufficiently powerful (power = .8, alpha = .05) to answer most of the key research questions of the collaboration.

The ultimate sampling procedure decision was made as a practical balancing of the number of participants and the number of measurements provided by each participants. Given that this study primarily focuses on the contemporaneous effects within the daily diary measurements, power should be high enough.

## Stopping rule
<!-- If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection.

You may specify a stopping rule based on p-values only in the specific case of sequential analyses with pre-specified checkpoints, alphas levels, and stopping rules. Unacceptable rationales include stopping based on p-values if checkpoints and stopping rules are not specified. If you have control over your sample size, then including a stopping rule is not necessary, though it must be clear in this question or a previous question how an exact sample size is attained.

Example: We will post participant sign-up slots by week on the preceding Friday night, with 20 spots posted per week. We will post 20 new slots each week if, on that Friday night, we are below 320 participants. -->

Participants will be recruited until 80 participants finish the pre-measurement. Invitations to complete additional daily diary measurements will be extended until participants chose to leave the study or at the most until two months (i.e., 64 days) after the initial entry survey (i.e., from the pre measurement survey).

# Variables
<!-- In this section you can describe all variables (both manipulated and measured variables) that will later be used in your confirmatory analysis plan. In your analysis plan, you will have the opportunity to describe how each variable will be used. If you have variables which you are measuring for exploratory analyses, you are not required to list them, though you are permitted to do so. -->

## Manipulated variables
<!-- Describe all variables you plan to manipulate and the levels or treatment arms of each variable. This is not applicable to any observational study. For any experimental manipulation, you should give a precise definition of each manipulated variable. This must include a precise description of the levels at which each variable will be set, or a specific definition for each categorical treatment. For example, “loud or quiet,” should instead give either a precise decibel level or a means of recreating each level. 'Presence/absence' or 'positive/negative' is an acceptable description if the variable is precisely described.

Example: We manipulated the percentage of sugar by mass added to brownies. The four levels of this categorical variable are: 15%, 20%, 25%, or 40% cane sugar by mass. -->

Not applicable given that the study design is observational.

## Measured variables
<!-- Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

Observational studies and meta-analyses will include only measured variables. As with the previous questions, the answers here must be precise. For example, 'intelligence,' 'accuracy,' 'aggression,' and 'color' are too vague. Acceptable alternatives could be 'IQ as measured by Wechsler Adult Intelligence Scale' 'percent correct,' 'number of threat displays,' and 'percent reflectance at 400 nm.'

Example: The single outcome variable will be the perceived tastiness of the single brownie each participant will eat. We will measure this by asking participants ‘How much did you enjoy eating the brownie’ (on a scale of 1-7, 1 being 'not at all', 7 being 'a great deal') and 'How good did the brownie taste' (on a scale of 1-7, 1 being 'very bad', 7 being 'very good'). -->

**Key variables:**

1. Whether an interaction happened during the past daytime period
    a. Did you meet a Dutch person this [morning/afternoon]?
    a. Did you meet a Non-Dutch person [morning/afternoon]?
1. Number of interactions during the past daytime period
    a. How many interactions with Dutch people did you have this [morning/afternoon]?
    a. How many interactions with Non-Dutch people did you have this [morning/afternoon]?
1. Key need assessment
    a. What was your main goal during the interaction with [name interaction partner]?
    a. During the interaction with [name interaction partner] your goal [free-text entry interaction key need] was fulfilled.
1. Allport's Conditions (equal status, common goal, collaboration, and structural support)
    a. The interaction with [name interaction partner] was on equal footing (same status)
    a. [name interaction partner] shared your goal ([free-text entry interaction key need])
    a. The interaction with [name interaction partner] was cooperative
    a. The interaction with [name interaction partner] was voluntary
1. Interaction Quality
    a. Overall, the interaction with [name interaction partner] was: Unpleasant to Pleasant
1. Outgroup Attitudes
    a. After the interaction, how favorably do you feel towards the Dutch.

**Control variables:**

1. Self Determination Theory needs ("During the interaction with [name interaction partner]")
    a. I was myself.
    a. I felt competent.
    a. I shared information about myself. 
    a. [name interaction partner] shared information about themselves.
1. General key need during daytime
    a. What was your main goal during this [morning/afternoon]?
    a. During this [morning/afternoon] your goal ([free-text entry daytime key need]) was fulfilled
    a. How important was your goal ([free-text entry daytime key need]) to you:
1. Student Goals ("Please rate the importance of each goal during the interaction with [name interaction partner]"):
    a. Social support / social connection goals
    a. Romantic or sexual relationship goals
    a. Academic goals
    a. Career goals
    a. Financial goals
    a. Health / fitness goals
    a. Leisure / fun goals
    a. Personal improvement / growth goals
    a. Service / helping goals
    a. Spiritual / religious goals

**Additional Analyses:**

1. Generalization from interaction partner to outgroup
    a. After the interaction, how favorably do you feel towards [name interaction partner].
    a. The interaction with [name interaction partner] was representative of the Dutch.
1. Alternative interaction quality definition ("Overall, the interaction with [name interaction partner] was ...")
    a. Unpleasant to Pleasant
    a. Superficial to Meaningful
    a. Ineffective to Effective
    a. Unimportant to Important
1. Potential moderation by the relationship with interaction partner
    a. How close do you feel to [name interaction partner]?
    a. How much do you trust [name interaction partner]?
    a. What is your relationship with [name interaction partner]?
1. Potential moderation by key need characteristics
    a. [name interaction partner] helped fulfill your goal [free-text entry interaction key need].
    a. How important was your goal ([free-text interaction key need]) to you?
1. Potential moderation by the interaction characteristics
    a. How long was the interaction? (in minutes)
    a. What was the main language during the interaction?
    a. The interaction was: With a single other / In a group

_Full variable information is available in the codebook ('Codebook_AOT-M_ItemsPerSection.xlsx')_

## Indices
<!-- If any measurements are  going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If your are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.

If you are using multiple pieces of data to construct a single variable, how will this occur? Both the data that are included and the formula or weights for each measure must be specified. Standard summary statistics, such as "means" do not require a formula, though more complicated indices require either the exact formula or, if it is an established index in the field, the index must be unambiguously defined. For example, "biodiversity index" is too broad, whereas "Shannon’s biodiversity index" is appropriate.

Example: We will take the mean of the two questions above to create a single measure of 'brownie enjoyment.'  -->

1. Mean Allport's conditions. We create a mean-averaged index of Allport's conditions in response to past findings indicating that the conditions are best conceptualized jointly and as functioning together rather than as fully independent factors [@Pettigrew2006, p. 766]. Similar to past studies we thus hope to build a global indicator [e.g., see @Pettigrew2006]. As with other indices we will ensure that the individual items indeed relate to a common latent construct and are  meaningfully combine in an index. If this is not possible we will create sub-indices and/or assess the impact of the conditions separately.
    a. The interaction with [name interaction partner] was on equal footing (same status)
    a. [name interaction partner] shared your goal ([free-text entry interaction key need])
    a. The interaction with [name interaction partner] was cooperative
    a. The interaction with [name interaction partner] was voluntary
1. Mean belongingness during intergroup contact
    a. I shared information about myself. 
    a. [name interaction partner] shared information about themselves.
1. Mean alternative interaction quality definition ("Overall, the interaction with [name interaction partner] was ...")
    a. Unpleasant to Pleasant
    a. Superficial to Meaningful
    a. Ineffective to Effective
    a. Unimportant to Important

# Analysis Plan
<!-- You may describe one or more confirmatory analysis in this preregistration. Please remember that all analyses specified below must be reported in the final article, and any additional analyses must be noted as exploratory or hypothesis generating.

A confirmatory analysis plan must state up front which variables are predictors (independent) and which are the outcomes (dependent), otherwise it is an exploratory analysis. You are allowed to describe any exploratory work here, but a clear confirmatory analysis is required. -->

## Statistical models
<!-- What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions, subgroup analyses, pairwise or complex contrasts, or follow-up tests from omnibus tests. If you plan on using any positive controls, negative controls, or manipulation checks you may mention that here. Remember that any test not included here must be noted as an exploratory test in your final article.

This is perhaps the most important and most complicated question within the preregistration. As with all of the other questions, the key is to provide a specific recipe for analyzing the collected data. Ask yourself: is enough detail provided to run the same analysis again with the information provided by the user? Be aware for instances where the statistical models appear specific, but actually leave openings for the precise test. See the following examples:

- If someone specifies a 2x3 ANOVA with both factors within subjects, there is still flexibility with the various types of ANOVAs that could be run. Either a repeated measures ANOVA (RMANOVA) or a multivariate ANOVA (MANOVA) could be used for that design, which are two different tests. 
- If you are going to perform a sequential analysis and check after 50, 100, and 150 samples, you must also specify the p-values you’ll test against at those three points.

Example:  We will use a one-way between subjects ANOVA to analyze our results. The manipulated, categorical independent variable is 'sugar' whereas the dependent variable is our taste index. -->

We will use a sequential analysis strategy in line with our proposed hypotheses. Given the nested structure of much of our data we test many of our hypotheses using a multilevel approach, where $y_{ti}$ denotes the response at measurement occasion $t$ ($t = 1, ...,T_i$; level 1) for individual $i$ ($i = 1, ..., n$; level 2). It should be noted that we will follow a hierarchical modeling approach. We follow the common four-step procedure [e.g., @Bliese2013]: (1) Test whether enough variation exists within and between participants to justify a multilevel structure, (2) (sequentially) add key predictors, (3) check whether a random slope explains an adequate amount of variance, and (4) check for autocorrelations and heteroscedasticity. For brevity we will only present the full multilevel regression formulas below.

\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{0pt}

1. Contact Hypothesis (partially using between participant aggregates to meaningfully include interaction frequency)
    a. Correlation:
      \begin{fleqn}[\parindent] 
        \begin{equation}
          r_{ContactFreq, Attitude} \neq 0
        \end{equation}
      \end{fleqn}
    a. Regression:
        <!-- 
          Explanation: 
          - 'fleqn' aligns whole math left 
          - 'equation' adds number
          - 'split' aligns after level header
          - 'aligned[t]' aligns after equal sign
        -->
        \begin{fleqn}[\parindent] 
        \begin{equation}
          \begin{split}
            \textrm{Level 1:} &
              \begin{aligned}[t]
                \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}OutgroupInteraction_{ti} + \\
                                   &\ \beta_{2i}NonOutgroupInteraction_{ti} + e_{ti}
              \end{aligned} \\
            \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                              &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                              &\ \beta_{2i} = \gamma_{20} + u_{2i}
          \end{split} 
        \end{equation}
        \end{fleqn}
    a. Regression: 
      \begin{fleqn}[\parindent] 
        \begin{equation}
          Attitude = ContactFreq \times AverageQual
        \end{equation}
      \end{fleqn}
1. Allport's Conditions
    a. Regression:
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttAllport}
            \begin{split}
                \textrm{Level 1:} &\ Attitude_{ti} = \beta_{0i} + \beta_{1i}AllportConditions_{ti} + e_{ti}\\
                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
            \end{split}
          \end{equation}
        \end{fleqn}
    a. Regression:
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesQualAllport}
            \begin{split}
                \textrm{Level 1:} &\ InteractionQuality_{ti} = \beta_{0i} + \beta_{1i}AllportConditions_{ti} + e_{ti}\\
                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
            \end{split}
          \end{equation}
        \end{fleqn}
    a. Regression:
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttAllportQual}
            \begin{split}
              \textrm{Level 1:} &
                \begin{aligned}[t]
                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}AllportConditions_{ti} + \\
                                     &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
                \end{aligned} \\
              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                                &\ \beta_{2i} = \gamma_{20} + u_{2i}
            \end{split} 
          \end{equation}
        \end{fleqn}  
1. Key Need fulfillment
    a. Regression: 
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttCore}
            \begin{split}
                \textrm{Level 1:} &\ Attitude_{ti} = \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}\\
                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
            \end{split}
          \end{equation}
        \end{fleqn}
    a. Regression:
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesQltCore}
            \begin{split}
                \textrm{Level 1:} &\ InteractionQuality_{ti} = \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}\\
                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
            \end{split}
          \end{equation}
        \end{fleqn}
    a. Regression:
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttCoreQual}
            \begin{split}
              \textrm{Level 1:} &
                \begin{aligned}[t]
                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                                     &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
                \end{aligned} \\
              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                                &\ \beta_{2i} = \gamma_{20} + u_{2i}
            \end{split} 
          \end{equation}
        \end{fleqn}
    a. Regression: 
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttCoreQualX}
            \begin{split}
              \textrm{Level 1:} &
                \begin{aligned}[t]
                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                                     &\ \beta_{2i}InteractionQuality_{ti} + \\
                                     &\ \beta_{3i}KeyNeedFulfill*InteractionQuality_{ti} + e_{ti}
                \end{aligned} \\
              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                                &\ \beta_{2i} = \gamma_{20} + u_{2i} \\
                                &\ \beta_{3i} = \gamma_{30} + u_{3i}
            \end{split} 
          \end{equation}
        \end{fleqn}
    a. Regression: 
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttCoreSdt}
            \begin{split}
              \textrm{Level 1:} &
                \begin{aligned}[t]
                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \beta_{2i}Autonomy_{ti} + \\
                                     &\ \beta_{3i}Competence_{ti} + \beta_{4i}Relatedness_{ti} + e_{ti}
                \end{aligned} \\
              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                                &\ \beta_{2i} = \gamma_{20} + u_{2i} \\
                                &\ \beta_{3i} = \gamma_{30} + u_{3i} \\
                                &\ \beta_{4i} = \gamma_{40} + u_{4i}
            \end{split} 
          \end{equation}
        \end{fleqn}
1. Comparison with Allport's Conditions
    a. Model Comparison:
      \begin{fleqn}[\parindent] 
        \begin{equation}
          AIC_{KeyNeedModel} < AIC_{AllportModel}
        \end{equation}
      \end{fleqn}
    a. Regression: 
        \begin{fleqn}[\parindent]
          \begin{equation} \label{eq:SlopesAttCoreAllport}
            \begin{split}
              \textrm{Level 1:} &
                \begin{aligned}[t]
                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                                     &\ \beta_{2i}AllportConditions_{ti} + e_{ti}
                \end{aligned} \\
              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                                &\ \beta_{2i} = \gamma_{20} + u_{2i}
            \end{split} 
          \end{equation}
        \end{fleqn}

All multilevel assumptions are tested as usual including (e.g., for random slopes model with _j_ within person predictors):

\begin{flalign}
  &\textrm{Level 1 Variance:}\ e_{ti} \sim \mathcal{N}(0,\sigma^2) \\
  &\textrm{Level 2 Variance:}\ \begin{bmatrix} u_{0i}\\ \vdots\\ u_{ji}\end{bmatrix} 
  \sim \mathcal{N}
  \begin{pmatrix}
    \begin{bmatrix} 
      0 \\ 
      \vdots \\
      0
    \end{bmatrix}, 
    \begin{bmatrix} 
      \tau_{00}^2 &  & \\ 
      \vdots & \ddots & \\
      \tau_{j0} & \ldots & \tau_{jj}^2
    \end{bmatrix}
  \end{pmatrix}
\end{flalign}

## Transformations
<!-- If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process. If any categorical predictors are included in a regression, indicate how those variables will be coded (e.g. dummy coding, summation coding, etc.) and what the reference category will be.

Example: The "Effect of sugar on brownie tastiness" does not require any additional transformations. However, if it were using a regression analysis and each level of sweet had been categorically described (e.g. not sweet, somewhat sweet, sweet, and very sweet), 'sweet' could be dummy coded with 'not sweet' as the reference category. -->

We will mean-center all level-one predictors within participants (i.e., cluster mean centering). This is relevant for any predictor variable during the interaction (e.g., key need fulfillment, interaction quality, ...). We mean-center within-person predictors, to meaningfully distinguish within and between person variances and because we obtain a meaningful intercept when having a random slope model [e.g., @Hamaker2014].

## Inference criteria
<!-- What criteria will you use to make inferences? Please describe the information youÍll use (e.g. p-values, bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?

p-values, confidence intervals, and effect sizes are standard means for making an inference, and any level is acceptable, though some criteria must be specified in this or previous fields. Bayesian analyses should specify a Bayes factor or a credible interval. If you are selecting models, then how will you determine the relative quality of each? In regards to multiple comparisons, this is a question with few "wrong" answers. In other words, transparency is more important than any specific method of controlling the false discovery rate or false error rate. One may state an intention to report all tests conducted or one may conduct a specific correction procedure; either strategy is acceptable.

Example: We will use the standard p<.05 criteria for determining if the ANOVA and the post hoc test suggest that the results are significantly different from those expected if the null hypothesis were correct. The post-hoc Tukey-Kramer test adjusts for multiple comparisons. -->

We will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant. However, given the large sample size we will place a particular emphasis on effect sizes in our interpretations of the results.

## Data exclusion
<!-- How will you determine what data or samples, if any, to exclude from your analyses? How will outliers be handled? Will you use any awareness check? Any rule for excluding a particular set of data is acceptable. One may describe rules for excluding a participant or for identifying outlier data.

Example: No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the three tastiness indices. Outliers will be included in the analysis. -->

No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the variables of interest for a given analysis. Outliers will generally be included in analyses, however we will use sensitivity analyses to assess the robustness of the results to outliers.

## Missing data
<!-- How will you deal with incomplete or missing data? Any relevant explanation is acceptable. As a final reminder, remember that the final analysis must follow the specified plan, and deviations must be either strongly justified or included as a separate, exploratory analysis.

Example: If a subject does not complete any of the three indices of tastiness, that subject will not be included in the analysis. -->

Given the (expectedly) large sample size of contemporaneous measurements, we will remove all instances during which a participant did not respond to any of the relevant variables. Given that we are not necessarily interested in individual trajectories in our analyses, we will use all available data (i.e., will not remove entire participants but rather individual incomplete daily diary measurements).

## Exploratory analyses (optional)
<!-- If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.

Example: We expect that certain demographic traits may be related to taste preferences. Therefore, we will look for relationships between demographic variables (age, gender, income, and marital status) and the primary outcome measures of taste preferences. -->

@Brown2005 proposed in their integrative theory of intergroup contact that "group salience" should moderate the positive effect of intergroup contact on outgroup attitudes. They argue that especially when the outgroup membership is salient, the positive attitudes should generalize from the outgroup individual to the general outgroup. We have data to test this hypothesis within a large intensive longitudinal dataset of daily intergroup interactions. We do this by separately assessing attitudes towards the outgroup interaction partner and the outgroup generally. And by assessing whether perceiving the interaction as representative of the outgroup moderates the positive contact effects.

To build confidence in our conclusions about the role of contact quality we calculate an alternative measure of interaction quality that includes perceived pleasantness, meaning, effectiveness, and importance of the interaction.

We expect that certain contextual characteristics may be related to the positive effects of high quality outgroup interactions. Therefore, we will look for the effect of key need fulfillment on outgroup attitudes while also including information about the interaction partner (perceived closeness, trust, relationship), the interaction characteristics (duration, interaction language, group vs. dyadic), as well as characteristics of the key need (importance of the need, fulfillment due to the interaction partner).

# Other

## Other (Optional)
<!-- If there is any additional information that you feel needs to be included in your preregistration, please enter it here. Literature cited, disclosures of any related work such as replications or work that uses the same data, or other context that will be helpful for future readers would be appropriate here. -->

Not applicable.

## 
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
\noindent

### 2022-05-12 TitlePage.tex
Update TitlePage.tex

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper, donotrepeattitle]{apa7}

% change margins (e.g., for margin comments):
%\usepackage{geometry}
% \geometry{
% a4paper,
% marginparwidth=30mm,
% right=50mm,
%}

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome}
\usepackage[normalem]{ulem}
% \usepackage{longtable}
\usepackage{amsmath}
\usepackage{afterpage}
\usepackage{float}
\usepackage{array}
\usepackage{censor}

% formatting links in the PDF file
\hypersetup{
pdfpagemode={UseOutlines},
bookmarksopen=true,
bookmarksopenlevel=0,
hypertexnames=false,
colorlinks   = true, %Colours links instead of ugly boxes
urlcolor     = blue, %Colour for external hyperlinks
linkcolor    = blue, %Colour of internal links
citecolor   = cyan, %Colour of citations
pdfstartview={FitV},
unicode,
breaklinks=true,
}

% language settings
\DeclareLanguageMapping{american}{american-apa}

% add reference library file
\addbibresource{references.bib}

% add reference library file
\addbibresource{references.bib}

% Title and header
\title{Migration Experience Trajectories: Subtitle TBD}
\shorttitle{Migration Experience Trajectories}

% Authors
\author[*,1,3]{Jannis Kreienkamp}
\author[1,3]{Laura F. Bringmann}
\author[1,3]{Kai Epstude}
\author[1,3]{Maximilian Agostini}
\author[1,3]{Peter de Jonge}
\author[2,3]{Rei Tendeiro-Monden}
\affiliation{\hfill}

\affil[1]{University of Groningen, Department of Psychology}
\affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
\affil[3]{Author order TBD [currently in first name alphabetical order]}

\authornote{
   \addORCIDlink{* Jannis Kreienkamp}{0000-0002-1831-5604}\\
   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}\\
   \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
   \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
   \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software is available at \url{https://janniscodes.github.io/migration-trajectories/}  \citep{KreienkampTBD}. Protocols, materials, data, and code are available at \url{https://osf.io/TBA} \citep{KreienkampTBD}. The preregistration of our analysis can be accessed as part of our Open Science Framework repository \citep{KreienkampTBD}.

Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: j.kreienkamp@rug.nl}

\leftheader{Kreienkamp}

\setlength\parindent{1.27cm}

% Start of the main document:
\begin{document}

% add title information (incl. title page and abstract)
\maketitle

\printbibliography

\end{document}

### 2022-05-12 preregistration/Preregistration3MPCA.Rmd

@@
 -->
+
+Recent calls for:   
+- ABCD model (esp., affect and desire)   
+- longitudinal data (for both acculturation and intergroup literature)   
+- real-life data (for both acculturation and intergroup literature)
+
+Recent developments:  
+In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
+However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}
+- data collection possible (e.g., ContactLogger [@Keil2020], FormR [@Arslan2020])  
+- analyses become more accessible [@ODonnell2021]
+
+Problem: HOWEVER, ...   
+- unclear how do deal with this more complex data (many variables, persons, and time points)   
+    - many different variables could be important    
+    - how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome)
+- heterogeneity between people, over time, and across variables   
+- **unclear how to identify core/important developments.** Especially across multiple variables at he same time (i.e., multiple aspects).  
+- unclear which variables, time scales, and methods are useful in practice
+
+Solution:   
+- 3MPCA uses the data cube
+
+not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
+Allows us to identify clusters of similar individuals and developments, and whether these clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+
+We recently proposed that migration experiences and the resulting psychological changes (i.e., acculturation) can be structured in terms of affect, behavior, cognition, and desire developments (ABCD of acculturation).
+However, while there have been several calls for intensive longitudinal and real-life data on acculturation, little data has thus far investigated the development of migration experiences and no research has assessed the co-development of all four experience aspects.
+Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
+Simultaneously clustering the person-, experience aspect-, and time level of migration experiences, using a three mode PCA, may allow to identify clusters of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
 
@@
 
-**Registration prior to accessing the data**.
-As of the date of submission, the data exist, but have not been accessed by you or your collaborators.
-Commonly, this includes data that has been collected by another researcher or institution.
+<!-- **Registration prior to analysis of the data**.
+As of the date of submission, the data exist and you have accessed it, though no analysis has been conducted related to the research plan (including calculation of summary statistics).
+A common situation for this scenario when a large dataset exists that is used for many different studies over time, or when a data set is randomly split into a sample for exploratory analyses, and the other section of data is reserved for later confirmatory data analysis.
+-->
+
+**Registration following analysis of the data**: As of the date of submission, you have accessed and analyzed some of the data relevant to the research plan.
+This includes preliminary analysis of variables, calculation of descriptive statistics, and observation of data distributions.
+Please see cos.io/prereg for more information.
 

### 2022-05-12 preregistration/Preregistration3MPCA.Rmd

@@
 
-Recent calls for:   
-- ABCD model (esp., affect and desire)   
-- longitudinal data (for both acculturation and intergroup literature)   
-- real-life data (for both acculturation and intergroup literature)
+Two important developments, within the psychological sciences in general and in inter-ethnic/migrantion research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
+For the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations [@Kreienkamp2022d, @Ward2019, @Cottam2010, @Hogg2005, @Jhangiani2014].
+At the same time prominent figures in the field have called for studies that collect longitudinal [@Pettigrew1998, @Pettigrew2008, @Pettigrew2008b, @Pettigrew2011] and real-life experience-sampling data outside the lab [@MacInnis2015, @McKeown2017, @Dixon2005].
 
-Recent developments:  
 In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
-However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}
-- data collection possible (e.g., ContactLogger [@Keil2020], FormR [@Arslan2020])  
-- analyses become more accessible [@ODonnell2021]
+However, recent technological developments allow us to easily collect experience sampling data on mobile devices [e.g., @Keil2020] or using web-based applications [@Arslan2020].
+At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable [@ODonnell2021].
+Given these, technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.
+
+* data collection possible (e.g., ContactLogger [@Keil2020], FormR [@Arslan2020])
+* analyses become more accessible [@ODonnell2021]
 
 Problem: HOWEVER, ...
-- unclear how do deal with this more complex data (many variables, persons, and time points)   
-    - many different variables could be important    
-    - how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome)
-- heterogeneity between people, over time, and across variables   
-- **unclear how to identify core/important developments.** Especially across multiple variables at he same time (i.e., multiple aspects).  
-- unclear which variables, time scales, and methods are useful in practice
+
+* unclear how do deal with this more complex data (many variables, persons, and time points).
+  + many different variables could be important.
+  + how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
+* heterogeneity between people, over time, and across variables.
+* **unclear how to identify core/important developments.** Especially across multiple variables at he same time (i.e., multiple aspects).
+* unclear which variables, time scales, and methods are useful in practice.
 
 Solution:
-- 3MPCA uses the data cube
+
+* 3MPCA uses the data cube
 
@@
 Simultaneously clustering the person-, experience aspect-, and time level of migration experiences, using a three mode PCA, may allow to identify clusters of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
-
-In its most essential interpretation, the intergroup contact hypothesis postulates that frequent and positive contact with an out-group reduces prejudice and increases favorable attitudes towards the other group [@Hewstone1996; @Pettigrew1998].
-A key condition for these contact benefits has been that the interaction is perceived as positive --- making the interaction quality a crucial mechanism of inter-group contact [e.g., @MacInnis2015].
-It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for a positive contact [Allport's Optimal Contact conditions, @Allport1954b].
-And indeed a major meta-analytic review showed, that intergroup contact benefits were larger when Allport's conditions were met.
-However, the meta-analysis also showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met [@Pettigrew2006].
-It, thus, remains unclear why and how exactly Allport's conditions work.
-
-We propose that Allport's optimal contact conditions are effective in creating positive contact because they constitute common psychological needs of the interacting individuals.
-If this is indeed the case, the impact of fulfilling other fundamental needs should produce similar beneficial contact results.
-And more importantly, the satisfaction of key situation needs during the interaction should predict the positive contact effects most strongly.
 

### 2022-05-12 preregistration/Preregistration3MPCA.Rmd
typo

@@
 
-Two important developments, within the psychological sciences in general and in inter-ethnic/migrantion research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
-For the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations [@Kreienkamp2022d, @Ward2019, @Cottam2010, @Hogg2005, @Jhangiani2014].
-At the same time prominent figures in the field have called for studies that collect longitudinal [@Pettigrew1998, @Pettigrew2008, @Pettigrew2008b, @Pettigrew2011] and real-life experience-sampling data outside the lab [@MacInnis2015, @McKeown2017, @Dixon2005].
+Two important developments, within the psychological sciences in general and in migration research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
+For the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations [@Kreienkamp2022d; @Ward2019; @Cottam2010; @Hogg2005; @Jhangiani2014].
+At the same time prominent figures in the field have called for studies that collect longitudinal [@Pettigrew1998; @Pettigrew2008; @Pettigrew2008b; @Pettigrew2011] and real-life experience-sampling data outside the lab [@MacInnis2015; @McKeown2017; @Dixon2005].
 
@@
 Given these, technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.
-
-* data collection possible (e.g., ContactLogger [@Keil2020], FormR [@Arslan2020])
-* analyses become more accessible [@ODonnell2021]
 

### 2022-05-12 preregistration/Preregistration3MPCA.Rmd
initial problem description added

@@
 Two important developments, within the psychological sciences in general and in migration research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
-For the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations [@Kreienkamp2022d; @Ward2019; @Cottam2010; @Hogg2005; @Jhangiani2014].
+Within the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations [@Kreienkamp2022d; @Ward2019].
 At the same time prominent figures in the field have called for studies that collect longitudinal [@Pettigrew1998; @Pettigrew2008; @Pettigrew2008b; @Pettigrew2011] and real-life experience-sampling data outside the lab [@MacInnis2015; @McKeown2017; @Dixon2005].
@@
 In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
-However, recent technological developments allow us to easily collect experience sampling data on mobile devices [e.g., @Keil2020] or using web-based applications [@Arslan2020].
-At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable [@ODonnell2021].
-Given these, technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.
+However, recent technological developments now allow us to easily collect large amounts of experience sampling data on mobile devices [e.g., @Keil2020] or using web-based applications [@Arslan2020].
+And while generally speaking analytical methods for such, more complex, data have become more readily available [e.g., @ODonnell2021], it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time (i.e., multiple aspects).
+In essence, the new extensive longitudinal data comes with new forms of heterogeneity when jointly considering differences between people, over time, and across variables.
+Yet, past analytical advances have almost exclusively pushed for stationary lagged regression models^[meaning models that assume stable means and variances over time; e.g., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses] or basic trajectory models^[e.g., mixed effects models, spline regression models, and latent growth curve modeling.].
+And while these two approaches are important for inferential model testing, both approaches ignore the more fundamental task of describing and unraveling the developmental data patterns.  
+And while these two approaches are important for inferential model testing, we still miss methods for the more fundamental task of describing and unraveling the developmental data patterns.
+
+Such descriptive methods are not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
+We we need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+
+We recently conducted three experience sampling studies, following the migration experiences of recent migrants to the Netherlands.
+In this manuscript, we aim to assess the utility of an analytical methods that has recently received increasing attention for simultaneously decomposing/considering person-, variable-, and time heterogeneity.
+The proposed three mode principle component analysis is of particular interest because its approach mirrors the famous data box of participants, variables, and measurement occasions (Catell's data box) for understanding psychological data.
 
@@
   + how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
+At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable [@ODonnell2021].
 * heterogeneity between people, over time, and across variables.
@@
 
-not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
-Allows us to identify clusters of similar individuals and developments, and whether these clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
-
-We recently proposed that migration experiences and the resulting psychological changes (i.e., acculturation) can be structured in terms of affect, behavior, cognition, and desire developments (ABCD of acculturation).
-However, while there have been several calls for intensive longitudinal and real-life data on acculturation, little data has thus far investigated the development of migration experiences and no research has assessed the co-development of all four experience aspects.
+little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
 Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
@@
 
-1.
-\label{hyp:contact}Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes.
-    a. \label{hyp:contactFreq}Participants with more intergroup interactions should have more favorable outgroup attitudes.
-    b. \label{hyp:contactDummy}Outgroup attitudes should be more positive after an intergroup interaction compared to a non-outgroup interaction.
-    c. \label{hyp:contactFreqQual}Participants with more intergroup interactions should have more favorable outgroup attitudes depending on the average interaction quality.
-2.
-\label{hyp:AllportsConditions}Based on Allport's optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
-    a.
-\label{hyp:AllportsPred}Based on Allport's optimal contact conditions, outgroup attitudes should be more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
-    b.
-\label{hyp:AllportsQuality}Based on past research on the role of interaction quality, interaction quality should be perceived as more positive after intergroup interactions with equal status, common goals, collaboration, and structural support.
-    c. \label{hyp:AllportsQualityMediation}Taken together, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extend be explained by interaction quality.
-3.
-\label{hyp:keyNeed}Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
-    a. \label{hyp:keyNeedPred}Outgroup attitudes should be more favorable after intergroup interactions with high key need fulfillment.
-    b. \label{hyp:keyNeedQual}Interaction Quality should be perceived as more positive after intergroup interactions with higher key need fulfillment.
-    c. \label{hyp:keyNeedMediation}The variance explained in outgroup attitudes by key need fulfillment should to a large extend be explained by interaction quality.
-    d.
-\label{hyp:keyNeedContactType}The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general.
-Thus, the effect of key need fulfillment on outgroup attitudes should be stronger for intergroup interact than for non-outgroup interactions (or need fulfillment during a time of no interaction).
-    e.
-\label{hyp:keyNeedSDT}The effect of key need fulfillment on outgroup attitudes should persist even when taking other fundamental psychological needs into account.
-Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory).
-4. \label{hyp:comparison}Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict outgroup attitudes at least as well as Allport's conditions.
-    a. \label{hyp:compModel}The need model (H\ref{hyp:keyNeedPred}) should predict more variance in outgroup attitudes than the model based on Allport's conditions (H\ref{hyp:AllportsPred}).
-    b.
-\label{hyp:compTogether}The  effect of key need fulfillment on outgroup attitudes should  persist even when taking Allport's conditions into account.
-Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for equal status, common goals, collaboration, and structural support.
+We do not have any hypotheses in the traditional sense. Our analysis plan is based on ... .
 
@@
 
-This study used an extensive longitudinal design.
+All three studies used an extensive longitudinal design.
 Using a daily diary format, for at least 30 days participants received a short survey twice per day (at around 12pm and 7pm).
@@
 The data was collected as part of a larger collaboration on daily intergroup relations.
-The researchers have had no direct access to the data, as the data was collected using an automated system using the FormR framework [@Arslan2020].
-The data were then downloaded and stored securely on an encrypted device until the eventual data analysis.
-None of the researchers have accessed the data thus far.
+A sub-sample of variables has recently been accessed by the research team for an unrelated analysis.
+Not relevant, no insight for the current analysis.
 
@@
 
-The data was collected in a three step procedure:
+For all three studies, the data was collected in a three-step procedure:
 
@@
 
-After the initial 30 day duration, participants are offered the possibility to continue participating in the study either with payment if daily diary measures were missed during the initial study phase or without payment after a total of 60 daily diary measurements were completed.
+For the third study, participants had the option of continuing the study for an unspecified amount of time.
+After the initial 30 day duration, participants were offered the possibility to continue participating in the study either with payment if daily diary measures were missed during the initial study phase or without payment after a total of 60 daily diary measurements were completed.
 After the initial 30-day period, participants receive automated feedback visualizing the development of their own well-being, attitudes, and motive responses as an additional initiative and to give participants access to their own data and to compensate study participation.

### 2022-05-13 preregistration/Preregistration3MPCA.Rmd
analysis plan added

@@
 
-We do not have any hypotheses in the traditional sense. Our analysis plan is based on ... .
+We do not have hypotheses in the traditional sense.
+Our analysis plan is based on the aim of describing and understanding a complex data set of extensive longitudinal data.
+We propose that the dimension reduction procedure we employ will identify meaningful patterns and developments within the data, which are useful to migration researchers and practitioners.
 
@@
 The data was collected as part of a larger collaboration on daily intergroup relations.
-A sub-sample of variables has recently been accessed by the research team for an unrelated analysis.
-Not relevant, no insight for the current analysis.
+A sub-sample of variables has recently been accessed by the research team for an unrelated analysis [@Kreienkamp2022b].
+Additionally, several of the variables were prepared for a graphical presentation of the dataset.
+Thus far, none of the proposed analyses have been conducted and none of the previous analyses have been related to dimension reductions or longitudinal trajectories.
 
@@
 
-Our target sample size is a sum of 4,000 daily diary measurements.
+For Study 1 (initial preliminary study) our target sample size is a sum of 1,000 daily diary measurements.
+As we expected a completion rate of around 80% we aimed to recruite 20 participants (20 participants X 50 measurements).
+We ultimately over-sampled slightly to compensate for potential drop outs given the length and intensity of the study.
+We then used several key relationships (relevant to the broader research team) to estimate sufficient power for a range of different analyses (i.e., using power simulations).
+
+For Studies 2 and 3, our target sample size is a sum of 4,000 daily diary measurements.
 With 100% completion rate that would be archived with 67 participants (60 daily diary responses each).
 Given that we expect some incomplete daily diary measurements, we aimed to recruit 80 participants.
+For both studies we again over-sampled slightly to account for expected drop outs within each target population.
 
@@
 Power considerations of mixed effects model such as with extensive longitudinal data are difficult to estimate because of the complex covariance structures.
-Simulation studies based on the first two samples within this project indicated that with well-distributed scales, and small to medium effect sizes, 70-80 participants with at least seven daily diary measurements and a simple pre--post survey were sufficiently powerful (power = .8, alpha = .05) to answer most of the key research questions of the collaboration.
+Simulation studies based on the first sample within this project indicated that with well-distributed scales, and small to medium effect sizes, 70-80 participants with at least seven daily diary measurements and a simple pre--post survey were sufficiently powerful (power = .8, alpha = .05) to answer most of the key research questions of the collaboration.
 
 The ultimate sampling procedure decision was made as a practical balancing of the number of participants and the number of measurements provided by each participants.
-Given that this study primarily focuses on the contemporaneous effects within the daily diary measurements, power should be high enough.
 
@@
 
-Participants will be recruited until 80 participants finish the pre-measurement.
-Invitations to complete additional daily diary measurements will be extended until participants chose to leave the study or at the most until two months (i.e., 64 days) after the initial entry survey (i.e., from the pre measurement survey).
+Participants were recruited until the targeted participant number finish the pre-measurement.
+Invitations to complete additional daily diary measurements (in Study 3) were extended until participants chose to leave the study or at the most until two months (i.e., 64 days) after the initial entry survey (i.e., from the pre measurement survey).
 
@@
 
-We will use a sequential analysis strategy in line with our proposed hypotheses.
-Given the nested structure of much of our data we test many of our hypotheses using a multilevel approach, where $y_{ti}$ denotes the response at measurement occasion $t$ ($t = 1, ...,T_i$; level 1) for individual $i$ ($i = 1, ..., n$; level 2).
-It should be noted that we will follow a hierarchical modeling approach.
-We follow the common four-step procedure [e.g., @Bliese2013]: (1) Test whether enough variation exists within and between participants to justify a multilevel structure, (2) (sequentially) add key predictors, (3) check whether a random slope explains an adequate amount of variance, and (4) check for autocorrelations and heteroscedasticity.
-For brevity we will only present the full multilevel regression formulas below.
+In our analysis, we follow the procedure outlined by @Monden2015
 
+1.
+**Sample Selection.** Our sample selection consists of three main steps.
+We (1) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data.
+We then address the selection of time point and participants.
+Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
+Thus, we then (2) select the time points for which we have responses from 55% of the participants and (3) select participants who have responded to at least 55% of the surveys across the selected time range.
+    1.
+Time point aggregation: Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to choose the appropriate time scales using variance decomposition [e.g., see @Ram2014].
+This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
+For example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.
+    \fbox{
+    \parbox{0.5\linewidth}{%
+    Potentially nested measurements:\\
+    1 - 2 - 4 - 8 - 16 - 32 - 64\\
+    1 - 2 - 6 - 12 - 24 - 48\\
+    1 - 2 - 10 - 30 - 60\\
+    1 - 2 - 14 - 28 - 56\\
+    Weird but possible (uneven):\\
+    1 - 3 - 6 - 12 - 24 - 48\\
+    1 - 5 - 10 - 30 - 60\\
+    1 - 7 - 14 - 28 - 56
+    }}  
+    \vspace{12pt}  
+    We then chose the ideal time scale for each of the key variables.  
+\newlength{\mylength}
+\[
+\setlength{\fboxsep}{15pt}
+\setlength{\mylength}{\linewidth}
+\addtolength{\mylength}{-2\fboxsep}
+\addtolength{\mylength}{-2\fboxrule}
+\fbox{%
+\parbox{\mylength}{
 \setlength{\abovedisplayskip}{0pt}
 \setlength{\belowdisplayskip}{0pt}
-\setlength{\abovedisplayshortskip}{0pt}
-\setlength{\belowdisplayshortskip}{0pt}
-
-1. Contact Hypothesis (partially using between participant aggregates to meaningfully include interaction frequency)
-    a. Correlation:
-      \begin{fleqn}[\parindent] 
-        \begin{equation}
-          r_{ContactFreq, Attitude} \neq 0
-        \end{equation}
-      \end{fleqn}
-    a. Regression:
-        <!-- 
-          Explanation: 
-          - 'fleqn' aligns whole math left 
-          - 'equation' adds number
-          - 'split' aligns after level header
-          - 'aligned[t]' aligns after equal sign
+\textbf{Question:}  Should the aggregation happen after the multiple imputation?
+The current procedure would reduce missing data but the aggregated values would be based on differing amounts of data (depending on the amount of missingness).
+Yet again, a daily or weekly summary this is probably still more accurate then a single later recall response.
+}}
+\]
+2.
+**Multiple Imputation.** We then create 20 imputed datasets to perform the analyses on.
+As outlined by @Monden2015, we will use the key variables themself as well as auxiliary variables to impute missing values.
+Auxiliary variables are any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
+3.
+**Three-way ANOVA.** We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset.
+Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions.
+We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the interdependence of the three modes.
+\[
+\fbox{%
+\parbox{\linewidth}{
+\textbf{Question:}  What is the DV of the three-way ANOVA?
+}}
+\]
+4.
+**Data preprocessing.** For the three mode PCA, we center (across participants but within time point) and to normalize (within variable but across time points) all key variables.
+The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering).
+The normalization within variable are important for ensuring equal variances across variables, which ensures equal weighting of the variables in the 3MPCA.
+For methodological detail see [Transformations].
+5.
+**Selection of 3MPCA model complexity** We then run the 3MPCA.
+We use a generalized scree plot to select the appropriate number of components for each of the three modes (i.e., person, variable, time).
+To test the stability of the solution results are compared across the 20 imputed datasets.
+\[
+\fbox{%
+\parbox{\linewidth}{
+\textbf{Question:}  Additionally split-half procedures? How does this work is this the same as split-half reliability?
+}}
+\]
+6.
+**3MPCA model fitting.** For the selected complexities we then extract simple component structures (i.e., for each mode individually) as well as the array core (the full 3MPCA array, which holds the weights of all combinations of the three modes).
+A Joint Orthomax orthogonal rotation and standard weights (when necessary) are used to extract human interpretable component scores.
+This procedure is done on all 20 imputed data sets.
+\[
+\fbox{%
+\parbox{\linewidth}{
+\textbf{Question:}  What exactly does the core array entail?
+In the @Monden2015 article it says: "the core-array, which describes the patterns of interaction between the symptom- and time-components in each person-component.".
+Additionally, in a more mathematical piece by Jorge it reads: "a three-way array called the core array, which holds the weights for the joint impact of any triple of components (one from each mode)." Does this mean that this is basically an array of loading scores in the lower dimensional space (similar to fitted values in regressions)?
+Or does it store the weights (i.e., interactions scores) that combines the loading scores of all three dimension reductions (not sure I express this right :-D )?
+I think this comes down to: What is the interpretation of a core element?
+}}
+\]
+7.
+**Generalized Procrustes rotation.** In order to combine the three individual component structure from the 20 data sets, we use a generalized Procrustes rotation --- calculating the average of each compontent and core array.
+<!-- The 3PCA model has been proposed by Tucker [114].
+Given an array X of order I×J×K, 3PCA determines component matrices for each of the three modes (individuals, variables and occasions) and a three-way array called the core array, which holds the weights for the joint impact of any triple of components (one from each mode).
 -->
-        \begin{fleqn}[\parindent] 
-        \begin{equation}
-          \begin{split}
-            \textrm{Level 1:} &
-              \begin{aligned}[t]
-                \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}OutgroupInteraction_{ti} + \\
-                                   &\ \beta_{2i}NonOutgroupInteraction_{ti} + e_{ti}
-              \end{aligned} \\
-            \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                              &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
-                              &\ \beta_{2i} = \gamma_{20} + u_{2i}
-          \end{split} 
-        \end{equation}
-        \end{fleqn}
-    a. Regression: 
-      \begin{fleqn}[\parindent] 
-        \begin{equation}
-          Attitude = ContactFreq \times AverageQual
-        \end{equation}
-      \end{fleqn}
-1. Allport's Conditions
-    a. Regression:
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttAllport}
-            \begin{split}
-                \textrm{Level 1:} &\ Attitude_{ti} = \beta_{0i} + \beta_{1i}AllportConditions_{ti} + e_{ti}\\
-                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
-            \end{split}
-          \end{equation}
-        \end{fleqn}
-    a. Regression:
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesQualAllport}
-            \begin{split}
-                \textrm{Level 1:} &\ InteractionQuality_{ti} = \beta_{0i} + \beta_{1i}AllportConditions_{ti} + e_{ti}\\
-                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
-            \end{split}
-          \end{equation}
-        \end{fleqn}
-    a. Regression:
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttAllportQual}
-            \begin{split}
-              \textrm{Level 1:} &
-                \begin{aligned}[t]
-                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}AllportConditions_{ti} + \\
-                                     &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
-                \end{aligned} \\
-              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
-                                &\ \beta_{2i} = \gamma_{20} + u_{2i}
-            \end{split} 
-          \end{equation}
-        \end{fleqn}  
-1. Key Need fulfillment
-    a. Regression: 
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttCore}
-            \begin{split}
-                \textrm{Level 1:} &\ Attitude_{ti} = \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}\\
-                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
-            \end{split}
-          \end{equation}
-        \end{fleqn}
-    a. Regression:
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesQltCore}
-            \begin{split}
-                \textrm{Level 1:} &\ InteractionQuality_{ti} = \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}\\
-                \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                  &\ \beta_{1i} = \gamma_{10} + u_{1i}
-            \end{split}
-          \end{equation}
-        \end{fleqn}
-    a. Regression:
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttCoreQual}
-            \begin{split}
-              \textrm{Level 1:} &
-                \begin{aligned}[t]
-                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
-                                     &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
-                \end{aligned} \\
-              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
-                                &\ \beta_{2i} = \gamma_{20} + u_{2i}
-            \end{split} 
-          \end{equation}
-        \end{fleqn}
-    a. Regression: 
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttCoreQualX}
-            \begin{split}
-              \textrm{Level 1:} &
-                \begin{aligned}[t]
-                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
-                                     &\ \beta_{2i}InteractionQuality_{ti} + \\
-                                     &\ \beta_{3i}KeyNeedFulfill*InteractionQuality_{ti} + e_{ti}
-                \end{aligned} \\
-              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
-                                &\ \beta_{2i} = \gamma_{20} + u_{2i} \\
-                                &\ \beta_{3i} = \gamma_{30} + u_{3i}
-            \end{split} 
-          \end{equation}
-        \end{fleqn}
-    a. Regression: 
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttCoreSdt}
-            \begin{split}
-              \textrm{Level 1:} &
-                \begin{aligned}[t]
-                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \beta_{2i}Autonomy_{ti} + \\
-                                     &\ \beta_{3i}Competence_{ti} + \beta_{4i}Relatedness_{ti} + e_{ti}
-                \end{aligned} \\
-              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
-                                &\ \beta_{2i} = \gamma_{20} + u_{2i} \\
-                                &\ \beta_{3i} = \gamma_{30} + u_{3i} \\
-                                &\ \beta_{4i} = \gamma_{40} + u_{4i}
-            \end{split} 
-          \end{equation}
-        \end{fleqn}
-1. Comparison with Allport's Conditions
-    a. Model Comparison:
-      \begin{fleqn}[\parindent] 
-        \begin{equation}
-          AIC_{KeyNeedModel} < AIC_{AllportModel}
-        \end{equation}
-      \end{fleqn}
-    a. Regression: 
-        \begin{fleqn}[\parindent]
-          \begin{equation} \label{eq:SlopesAttCoreAllport}
-            \begin{split}
-              \textrm{Level 1:} &
-                \begin{aligned}[t]
-                  \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
-                                     &\ \beta_{2i}AllportConditions_{ti} + e_{ti}
-                \end{aligned} \\
-              \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
-                                &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
-                                &\ \beta_{2i} = \gamma_{20} + u_{2i}
-            \end{split} 
-          \end{equation}
-        \end{fleqn}
-
-All multilevel assumptions are tested as usual including (e.g., for random slopes model with _j_ within person predictors):
-
-\begin{flalign}
-  &\textrm{Level 1 Variance:}\ e_{ti} \sim \mathcal{N}(0,\sigma^2) \\
-  &\textrm{Level 2 Variance:}\ \begin{bmatrix} u_{0i}\\ \vdots\\ u_{ji}\end{bmatrix} 
-  \sim \mathcal{N}
-  \begin{pmatrix}
-    \begin{bmatrix} 
-      0 \\ 
-      \vdots \\
-      0
-    \end{bmatrix}, 
-    \begin{bmatrix} 
-      \tau_{00}^2 &  & \\ 
-      \vdots & \ddots & \\
-      \tau_{j0} & \ldots & \tau_{jj}^2
-    \end{bmatrix}
-  \end{pmatrix}
-\end{flalign}
+8.
+**Explained variance.** As fit indices we then calculate (1) fit percentage of the estimated array for each imputed data set (i.e., explained variance around the removed general trend) as well as (2) an _overall_ fit percentage of the estimated array (i.e., explained variance including the general trend).
+Both these fit metrices should give an indication of how much of the original variance the reduced component structure was able to capture.
+9.
+**Component interpretation** To embed and interpret the resulting components (especially the person components), we calculate correlations of the component scores with key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
+If components remain unclear still we can additionally perform k-means clustering on the lower dimensional space to identify clearer groups which can be compared on the adaptation markers if necessary.
 
@@
 
-We will mean-center all level-one predictors within participants (i.e., cluster mean centering).
+<!-- We will mean-center all level-one predictors within participants (i.e., cluster mean centering).
 This is relevant for any predictor variable during the interaction (e.g., key need fulfillment, interaction quality, ...).
 We mean-center within-person predictors, to meaningfully distinguish within and between person variances and because we obtain a meaningful intercept when having a random slope model [e.g., @Hamaker2014].
+-->
+
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{0pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{0pt}
+
+For three mode PCA it is custom to center (across participants but within time point) and to normalize (within variable but across time points).
+This is done to create a meaningful zero value for each time point (to compare participants) and a variance of 1 across all time points of a variable (removing any overall trends).
+
+We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable $k$ ($k = 1,\ ...,\ K$) at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points are within the dataset.
+For further information as well as an example illustration see Supplemental Material 3 in @Monden2015.
+
+\begin{equation} \label{eq:Mean}
+  \overline{x}_{kt} =  \sum_{i=1}^I x_{ikt}
+\end{equation}
+
+\begin{equation} \label{eq:NormalizationSd}
+  \sigma_{k} =  \sqrt{\sum_{i=1}^I \sum_{k=1}^K \frac{(x_{ikt}-\overline{x}_{kt})^2}{I*T}}
+\end{equation}
+
+\begin{equation} \label{eq:NormalizedScore}
+  z_{ikt} =  \frac{x_{ikt}-\overline{x}_{kt}}{\sigma_{k}}
+\end{equation}
 
@@
 
-We will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant.
-However, given the large sample size we will place a particular emphasis on effect sizes in our interpretations of the results.
+Many of the proposed procedures are more descriptive rather then inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
+If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant.
+We will place particular emphasis on effect sizes in our interpretations of the results.
 
@@
 
-Given the (expectedly) large sample size of contemporaneous measurements, we will remove all instances during which a participant did not respond to any of the relevant variables.
-Given that we are not necessarily interested in individual trajectories in our analyses, we will use all available data (i.e., will not remove entire participants but rather individual incomplete daily diary measurements).
+The sample selection based on proportions of missing data and the multiple imputation procedures are described in [Statistical models].
 
@@
 
-@Brown2005 proposed in their integrative theory of intergroup contact that "group salience" should moderate the positive effect of intergroup contact on outgroup attitudes.
-They argue that especially when the outgroup membership is salient, the positive attitudes should generalize from the outgroup individual to the general outgroup.
-We have data to test this hypothesis within a large intensive longitudinal dataset of daily intergroup interactions.
-We do this by separately assessing attitudes towards the outgroup interaction partner and the outgroup generally.
-And by assessing whether perceiving the interaction as representative of the outgroup moderates the positive contact effects.
-
-To build confidence in our conclusions about the role of contact quality we calculate an alternative measure of interaction quality that includes perceived pleasantness, meaning, effectiveness, and importance of the interaction.
-
-We expect that certain contextual characteristics may be related to the positive effects of high quality outgroup interactions.
-Therefore, we will look for the effect of key need fulfillment on outgroup attitudes while also including information about the interaction partner (perceived closeness, trust, relationship), the interaction characteristics (duration, interaction language, group vs.
-dyadic), as well as characteristics of the key need (importance of the need, fulfillment due to the interaction partner).
+A recent alternative for trajectory dimension reduction has been the use of LSTM auto encoders.
+The encoder decoder classification procedure has the advantage of being extremely flexible in the use of the data set and could be interpreted in a similar way as the 3MPCA (see [Statistical models] 9).
 

### 2022-05-14 preregistration/Preregistration3MPCA.Rmd
Variable selection added

@@
 
-**Key variables:**
+Given the circumstance that we utilize three independent experience sampling studies for our analyses, the variables are at times slightly different between studies.
+Additionally, given that the analysis we aim to undertake (i.e., 3MPCA) uses a large number of variables for the multiple imputations, dimension reduction, and correlational analyses, we will not list all items individually.
+Instead, we provide full variable information is available in the following data sheets: 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'.
 
-1. Whether an interaction happened during the past daytime period
-    a. Did you meet a Dutch person this [morning/afternoon]?
-    a. Did you meet a Non-Dutch person [morning/afternoon]?
-1. Number of interactions during the past daytime period
-    a. How many interactions with Dutch people did you have this [morning/afternoon]?
-    a. How many interactions with Non-Dutch people did you have this [morning/afternoon]?
-1. Key need assessment
-    a. What was your main goal during the interaction with [name interaction partner]?
-    a. During the interaction with [name interaction partner] your goal [free-text entry interaction key need] was fulfilled.
-1. Allport's Conditions (equal status, common goal, collaboration, and structural support)
-    a. The interaction with [name interaction partner] was on equal footing (same status)
-    a. [name interaction partner] shared your goal ([free-text entry interaction key need])
-    a. The interaction with [name interaction partner] was cooperative
-    a. The interaction with [name interaction partner] was voluntary
-1. Interaction Quality
-    a. Overall, the interaction with [name interaction partner] was: Unpleasant to Pleasant
-1. Outgroup Attitudes
-    a. After the interaction, how favorably do you feel towards the Dutch.
+**Key variables**  
+Most of the variables are identical across all studies, however some differ slightly in content or wording.
+However, for each study we aimed to collect the following types of variables.
+We collected substantially more motivational and affecive variables, as they have been understudied in the past.
 
-**Control variables:**
+1. Motivational variables
+    a. core motive fulfillment (i.e., "_During your interaction with -X- (/this morning) your goal (-TEXT-) was fulfilled._")
+    a. goal importance ratings of 10 individual goals (e.g., "_career goals_", "_health / fitness goals_")
+    a. self determination theory needs (i.e., autonomy, relatedness, competence)
+1. Affective states
+    a. experienced well-being (happiness, energy)
+    a. general emotional state scale (e.g., "_How do you feel right now? not angry at all to very angry_")
+1. Behavioral self-reports (Studies 2 and 3)
+    a. pro-social behavior (e.g., "_Made demeaning, rude or derogatory remarks about someone._")
+    a. anti-social behavior (e.g., "_I was there to listen to someone's problems._")
+1. Cognitive measures
+    a. outgroup attitude ("_After the interaction, how favorably do you feel towards … / At the moment, how favorably do you feel towards …  the Dutch._")
+    a. Allport's contact condition perceptions (i.e., equal status, shared goal, cooperative, voluntary)
+    a. interaction quality ratings (e.g., "_Overall, the interaction was... Unpleasant to Pleasant_")
 
-1. Self Determination Theory needs ("During the interaction with [name interaction partner]")
-    a. I was myself.
-    a. I felt competent.
-    a. I shared information about myself. 
-    a. [name interaction partner] shared information about themselves.
-1. General key need during daytime
-    a. What was your main goal during this [morning/afternoon]?
-    a. During this [morning/afternoon] your goal ([free-text entry daytime key need]) was fulfilled
-    a. How important was your goal ([free-text entry daytime key need]) to you:
-1. Student Goals ("Please rate the importance of each goal during the interaction with [name interaction partner]"):
-    a. Social support / social connection goals
-    a. Romantic or sexual relationship goals
-    a. Academic goals
-    a. Career goals
-    a. Financial goals
-    a. Health / fitness goals
-    a. Leisure / fun goals
-    a. Personal improvement / growth goals
-    a. Service / helping goals
-    a. Spiritual / religious goals
+**Interpretation correlation variables**  
+To interpret the components identified by the 3MPCA, we would like to correlate the component scores with a range of person- or time point specific variables.
+While the full list of variables is available in the accompanying data sheets (see 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'), we provide a few examples of the targeted categories below.
 
-**Additional Analyses:**
+1. demographic variables
+    a. age
+    a. gender
+    a. ...
+1. societal participation
+    a. language comprehension and -use
+    a. work inclusion
+    a. ...
+1. adaptation indicators
+    a. social identification
+    a. intergroup anxiety
+    a. ...
+1. everyday life obstacles
+    a. discrimination experiences
+    a. negative life events
+    a. ...
+1. individual differences
+    a. big five personality scores
+    a. attachment style
+    a. ...
 
-1. Generalization from interaction partner to outgroup
-    a. After the interaction, how favorably do you feel towards [name interaction partner].
-    a. The interaction with [name interaction partner] was representative of the Dutch.
-1. Alternative interaction quality definition ("Overall, the interaction with [name interaction partner] was ...")
-    a. Unpleasant to Pleasant
-    a. Superficial to Meaningful
-    a. Ineffective to Effective
-    a. Unimportant to Important
-1. Potential moderation by the relationship with interaction partner
-    a. How close do you feel to [name interaction partner]?
-    a. How much do you trust [name interaction partner]?
-    a. What is your relationship with [name interaction partner]?
-1. Potential moderation by key need characteristics
-    a. [name interaction partner] helped fulfill your goal [free-text entry interaction key need].
-    a. How important was your goal ([free-text interaction key need]) to you?
-1. Potential moderation by the interaction characteristics
-    a. How long was the interaction? (in minutes)
-    a. What was the main language during the interaction?
-    a. The interaction was: With a single other / In a group
+**Auxiliary variables for multiple imputation**  
+For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
+For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx').
+Additional information about the multiple imputation procedure is also available at [Statistical models] 2.
 
-_Full variable information is available in the codebook ('Codebook_AOT-M_ItemsPerSection.xlsx')_
+_We also provide an exemplary codebook (using Study 3 as an example; see 'Codebook_AOT-M_ItemsPerSection.xlsx').
+The full survey files will be available as part of the main OSF repository connected to this preregistration._
 
@@
     a. Unimportant to Important
+1. Mood Subscales (MDMQ)
+    a. alertness
+    a. calmness
+    a. valence
+1. Mean anti-social behaviors
+    a. Put someone down
+    a. Show little attention in someones opinion
+    a. Demeaning remarks
+    a. Inappropriately addressing someone
+    a. Ignored or excluded someone
+    a. Doubt someones judgement
+    a. Unwanted attempts of personal matters
+1. Mean pro-social behaviors
+    a. Listen to someones problems
+    a. Cheer someone up
+    a. Help someone get things done
+    a. Help someone with responsibilities
 

### 2022-05-14 preregistration/Preregistration3MPCA.Rmd
first draft

@@
 
-Two important developments, within the psychological sciences in general and in migration research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
-Within the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations [@Kreienkamp2022d; @Ward2019].
-At the same time prominent figures in the field have called for studies that collect longitudinal [@Pettigrew1998; @Pettigrew2008; @Pettigrew2008b; @Pettigrew2011] and real-life experience-sampling data outside the lab [@MacInnis2015; @McKeown2017; @Dixon2005].
+Recent reviews have called for more comprehensive assessment of human experiences and for more longitudinal real-life data, within the psychological sciences more broadly and in migration research in particular [e.g., @Kreienkamp2022d; @MacInnis2015; @McKeown2017; @Pettigrew2011; @Ward2019].
+However, while generally speaking analytical methods for such, more complex, data have become more readily available [e.g., @ODonnell2021], it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time.
 
-In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
-However, recent technological developments now allow us to easily collect large amounts of experience sampling data on mobile devices [e.g., @Keil2020] or using web-based applications [@Arslan2020].
-And while generally speaking analytical methods for such, more complex, data have become more readily available [e.g., @ODonnell2021], it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time (i.e., multiple aspects).
-In essence, the new extensive longitudinal data comes with new forms of heterogeneity when jointly considering differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for stationary lagged regression models^[meaning models that assume stable means and variances over time; e.g., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses] or basic trajectory models^[e.g., mixed effects models, spline regression models, and latent growth curve modeling.].
-And while these two approaches are important for inferential model testing, both approaches ignore the more fundamental task of describing and unraveling the developmental data patterns.  
-And while these two approaches are important for inferential model testing, we still miss methods for the more fundamental task of describing and unraveling the developmental data patterns.
+In essence, the novel extensive longitudinal datasets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
+Yet, past analytical advances have almost exclusively pushed for top-down inferential modeling procedures^[For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).].
+And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the developmental data patterns.
 
-Such descriptive methods are not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
+As an example, little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
+Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
 We we need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
 
-We recently conducted three experience sampling studies, following the migration experiences of recent migrants to the Netherlands.
-In this manuscript, we aim to assess the utility of an analytical methods that has recently received increasing attention for simultaneously decomposing/considering person-, variable-, and time heterogeneity.
-The proposed three mode principle component analysis is of particular interest because its approach mirrors the famous data box of participants, variables, and measurement occasions (Catell's data box) for understanding psychological data.
-
-Problem: HOWEVER, ...
-
-* unclear how do deal with this more complex data (many variables, persons, and time points).
-  + many different variables could be important.
-  + how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
-At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable [@ODonnell2021].
-* heterogeneity between people, over time, and across variables.
-* **unclear how to identify core/important developments.** Especially across multiple variables at he same time (i.e., multiple aspects).
-* unclear which variables, time scales, and methods are useful in practice.
-
-Solution:
-
-* 3MPCA uses the data cube
-
-little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
-Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
-Simultaneously clustering the person-, experience aspect-, and time level of migration experiences, using a three mode PCA, may allow to identify clusters of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
+In this manuscript, we aim to assess the utility of one such promising analysis for the use of social psychological experience sampling data.
+Recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [i.e., three-mode principle component analyses, 3MPCA; e.g., @Monden2015].
+We will, particularly, analyse three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
 
@@
 We do not have hypotheses in the traditional sense.
-Our analysis plan is based on the aim of describing and understanding a complex data set of extensive longitudinal data.
+Our analysis plan is based on the aim of describing, summarizing, and understanding a complex data set of extensive longitudinal data.
 We propose that the dimension reduction procedure we employ will identify meaningful patterns and developments within the data, which are useful to migration researchers and practitioners.
@@
 As we expected a completion rate of around 80% we aimed to recruite 20 participants (20 participants X 50 measurements).
-We ultimately over-sampled slightly to compensate for potential drop outs given the length and intensity of the study.
+We further over-sampled slightly to compensate for potential drop outs given the length and intensity of the study.
+
 We then used several key relationships (relevant to the broader research team) to estimate sufficient power for a range of different analyses (i.e., using power simulations).
-
-For Studies 2 and 3, our target sample size is a sum of 4,000 daily diary measurements.
+Based on these power simulations, for Studies 2 and 3, our target sample size was a sum of 4,000 daily diary measurements.
 With 100% completion rate that would be archived with 67 participants (60 daily diary responses each).
@@
 
-Participants were recruited until the targeted participant number finish the pre-measurement.
+Participants were recruited until the targeted number of participant finished the pre-measurement.
 Invitations to complete additional daily diary measurements (in Study 3) were extended until participants chose to leave the study or at the most until two months (i.e., 64 days) after the initial entry survey (i.e., from the pre measurement survey).
@@
 Additionally, given that the analysis we aim to undertake (i.e., 3MPCA) uses a large number of variables for the multiple imputations, dimension reduction, and correlational analyses, we will not list all items individually.
-Instead, we provide full variable information is available in the following data sheets: 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'.
+Instead, we provide full variable information in the following data sheets: 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'.
 
 **Key variables**  
-Most of the variables are identical across all studies, however some differ slightly in content or wording.
-However, for each study we aimed to collect the following types of variables.
-We collected substantially more motivational and affecive variables, as they have been understudied in the past.
+Most of the variables are identical across all studies, however some differ slightly in content or wording. However, for each study we aimed to collect the following types of variables.
 
@@
     a. interaction quality ratings (e.g., "_Overall, the interaction was... Unpleasant to Pleasant_")
+
+It should be noted, that we collected substantially more motivational and affecive variables, as they have been understudied in the past.
 
@@
 **Auxiliary variables for multiple imputation**  
-For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
+For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with either the key variables or missingness on the key variables.
 For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx').
@@
 
-In our analysis, we follow the procedure outlined by @Monden2015
+For our analysis, we adapt the procedures outlined by @Monden2015:
 
@@
 **Sample Selection.** Our sample selection consists of three main steps.
-We (1) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data.
+We (a) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data.
 We then address the selection of time point and participants.
 Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
-Thus, we then (2) select the time points for which we have responses from 55% of the participants and (3) select participants who have responded to at least 55% of the surveys across the selected time range.
-    1.
-Time point aggregation: Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to choose the appropriate time scales using variance decomposition [e.g., see @Ram2014].
+Thus, we then (b) select the time points for which we have responses from 55% of the participants and (c) select participants who have responded to at least 55% of the surveys across the selected time range.
+    a.
+**Time point aggregation:** Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
 This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
-For example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.
+We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.
     \fbox{
@@
 4.
-**Data preprocessing.** For the three mode PCA, we center (across participants but within time point) and to normalize (within variable but across time points) all key variables.
+**Data preprocessing.** For the three mode PCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
 The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering).
@@
 **Component interpretation** To embed and interpret the resulting components (especially the person components), we calculate correlations of the component scores with key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
-If components remain unclear still we can additionally perform k-means clustering on the lower dimensional space to identify clearer groups which can be compared on the adaptation markers if necessary.
+If components remain unclear still we can additionally perform k-means clustering on the lower dimensional space to identify clearer groups, which can be compared on the adaptation markers if necessary.
 
@@
 
-For three mode PCA it is custom to center (across participants but within time point) and to normalize (within variable but across time points).
-This is done to create a meaningful zero value for each time point (to compare participants) and a variance of 1 across all time points of a variable (removing any overall trends).
+For three-mode PCA it is custom to center (across participants but within time point) and to normalize (within variable but across time points).
+This is done to create a meaningful zero value for each time point (to compare participants) and a variance of 1 across all time points of a variable (ensuring equal weightin of variables in the 3MPCA).
 
-We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable $k$ ($k = 1,\ ...,\ K$) at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points are within the dataset.
+We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable $k$ ($k = 1,\ ...,\ K$) at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points within the dataset.
 For further information as well as an example illustration see Supplemental Material 3 in @Monden2015.
@@
 
-Many of the proposed procedures are more descriptive rather then inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
+Many of the proposed procedures are descriptive rather then inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
 If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant.
-We will place particular emphasis on effect sizes in our interpretations of the results.
+We will place particular emphasis on effect sizes in our interpretations of the results wherever necessary.
 

### 2022-05-20 preregistration/Preregistration3MPCA.tex
track changes added

\documentclass[]{article}
    \usepackage{lmodern}
    \usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
  \else % if luatex or xelatex
\ifxetex
\usepackage{mathspec}
\usepackage{xltxtra,xunicode}
\else
  \usepackage{fontspec}
\fi
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
\newcommand{\euro}{€}
        \fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
  \usepackage{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
  \usepackage[left=2.5in,bottom=1.25in,top=1.25in,right=1in]{geometry}
  \ifxetex
\usepackage[setpagesize=false, % page size defined by xetex
            unicode=false, % unicode breaks when used with xetex
            xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
bookmarks=true,
pdfauthor={},
pdftitle={Migration Experience Trajectories: A Three Mode Principle Component Analysis},
colorlinks=true,
citecolor=blue,
urlcolor=blue,
linkcolor=magenta,
pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% Customization for cos_prereg
\usepackage{longtable,booktabs,threeparttable,tabularx}
\linespread{1.5}
\newcounter{question}
\setcounter{question}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
\posttitle{
\begin{center}\large#1\end{center}
}
}

\setlength{\droptitle}{-2em}
\title{Migration Experience Trajectories: A Three Mode Principle
Component Analysis}
\pretitle{\begin{changemargin}{-8pc}{0pc} \centering\large Preregistration\\ \Huge}
\posttitle{\end{changemargin}}
  \author{
          Jannis Kreienkamp\textsuperscript{1},
          Kai Epstude\textsuperscript{1},
          Rei Monden\textsuperscript{2},
          Maximilian Agostini\textsuperscript{1},
          Peter de Jonge\textsuperscript{1},
          Laura F.
Bringmann\textsuperscript{1}          \\ \vspace{0.5cm}
              \textsuperscript{1} University of Groningen, Department of
Psychology\\
              \textsuperscript{2} Hiroshima University, Graduate School
of Advanced Science and Engineering      }

  \def\affdep{{"", "", "", "", "", ""}}%
  \def\affcity{{"", "", "", "", "", ""}}%
  \preauthor{\begin{changemargin}{-8pc}{0pc} \centering\large}
  \postauthor{\end{changemargin}}
\date{20. May 2022}
\predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
\postdate{\end{changemargin}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{14.0pt}
\renewcommand{\headrulewidth}{0pt}
\lhead{}
\rhead{\large\textsc{\MakeLowercase{Preregistration: Migration
Experience Trajectories}}}

\usepackage{amsmath}
\usepackage{nccmath}
\usepackage{LatexPackage/trackchanges}
\usepackage[usenames,dvipsnames]{xcolor}
\addeditor{Jannis}

% Title settings
\usepackage{titlesec}
\titleformat{\section}[display]{\bfseries\Large}{\thesection}{}{}[]
\titlespacing{\section}{0pc}{*3}{*1.5}
\titleformat{\subsection}[leftmargin]{\titlerule\bfseries\filleft}{\thesubsection}{.5em}{}
\titlespacing{\subsection}{8pc}{5ex plus .1ex minus .2ex}{1.5pc}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle
\vspace{2pc}

\newcommand\Question[2]{%
   \leavevmode\par
   \stepcounter{question}
   \noindent
   \textbf{\thequestion. #1}. #2\par}

\newcommand\Answer[1]{%
    \noindent
    \textit{Registered response}: #1\par}

\newlength{\mylength}
\setlength{\fboxsep}{15pt}
\setlength{\mylength}{\linewidth}
\addtolength{\mylength}{-2\fboxsep}
\addtolength{\mylength}{-2\fboxrule}
\definecolor{editPurple}{RGB}{84, 28, 117}

\hypertarget{study-information}{%
\section{Study Information}\label{study-information}}

\hypertarget{title}{%
\subsection{Title}\label{title}}

Migration Experience Trajectories: A Three Mode Principle Component
Analysis

\hypertarget{description}{%
\subsection{Description}\label{description}}

Recent reviews have called for more comprehensive assessment of human
experiences and for more longitudinal real-life data, within the
psychological sciences more broadly and in migration research in
particular
\citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
However, while generally speaking analytical methods for such, more
complex, data have become more readily available
\citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify
key developmental patterns --- especially across multiple variables at
the same time.

In essence, the novel extensive longitudinal datasets come with new
forms of heterogeneity, where we have to consider differences between
people, over time, and across variables. Yet, past analytical advances
have almost exclusively pushed for top-down inferential modeling
procedures\footnote{For example, stationary lagged regression models
  that assume stable means and variances over time (incl., vector
  \change{autoregression}{autoregressive} models, dynamic structural
  equation models, autoregressive integrated moving average models, and
  cross-lagged panel analyses) or basic trajectory models (e.g., mixed
  effects models, spline regression models, and latent growth curve
  modeling).}. And while inferential model testing is certainly
important, we still miss discussions of methods for the more fundamental
task of describing, summarizing, and understanding the developmental
data patterns.

As an example, little data has thus far investigated the development of
migration experiences and no research has assessed the co-development of
multiple experience aspects. Yet, understanding how people differ in
their migration trajectories, can be crucial in understanding adaptive
and maladaptive patterns. \remove{We we need}\change{t}{T}o identify
which variables are most important in the adaptation of migrants
\add{over time}, we need methods to
\change{identify clusters of similar individuals and developments}{break down the data heterogeneity into its core components (in terms of important variables and developments)}
and we need \change{
to assess whether such clusters differ in}{ways to identify how these core components relate to}
key adaptation markers (including, well-being, intergroup anxiety,
outgroup trust, or societal participation). There is, thus, a clear need
to assess the utility of analysis procedures focused on the description
and understanding of complex dynamical data.

In this manuscript, we aim to assess the utility of one such promising
analysis for the use of social psychological experience sampling data.
Recent studies have laid out the potential effectiveness of dimension
reduction procedures, which can address the new forms of person-,
variable-, and time point heterogeneity jointly \citep[i.e., three-mode
principle component analyses, 3MPCA; e.g.,][]{Monden2015}.

\change{We will, particularly,}{To this aim, we will} analyse the data
from three experience sampling studies, which followed the migration
experiences of recent migrants to the Netherlands.
\add{All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation). However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences}
\citep{Kreienkamp2022d}\add{, the three studies have places a particular emphasis on the needs, goals, and motives of young migrants.}

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

We do not have hypotheses in the traditional sense. Our analysis plan is
based on the aim of describing, summarizing, and understanding a complex
data set of extensive longitudinal data. We propose that the dimension
reduction procedure we employ will identify meaningful patterns and
developments within the data, which are useful to migration researchers
and practitioners.

\hypertarget{design-plan}{%
\section{Design Plan}\label{design-plan}}

\hypertarget{study-type}{%
\subsection{Study type}\label{study-type}}

\textbf{Observational Study}. Data is collected from study subjects that
are not randomly assigned to a treatment. This includes surveys, natural
experiments, and regression discontinuity designs.
\note{Note: This is part of the OSF dropdown answer and we can't change the description.}

\hypertarget{blinding}{%
\subsection{Blinding}\label{blinding}}

No participant blinding is involved in this study.

\hypertarget{study-design}{%
\subsection{Study design}\label{study-design}}

All three
studies\note{Note: I now extended the introduction of the three studies in the end of the description section. So I hope this doesn't come out of the blue as much anymore.}
used an extensive longitudinal design. Using a daily diary format, for
at least 30 days participants received a short survey twice per day (at
around 12pm and 7pm). We additionally included a longer pre- and post
measurement survey the days before and after the extensive longitudinal
data collection.

\hypertarget{randomization}{%
\subsection{Randomization}\label{randomization}}

No randomization is involved in this study.

\hypertarget{sampling-plan}{%
\section{Sampling Plan}\label{sampling-plan}}

\hypertarget{existing-data}{%
\subsection{Existing data}\label{existing-data}}

\textbf{Registration following analysis of the data}: As of the date of
submission, you have accessed and analyzed some of the data relevant to
the research plan. This includes preliminary analysis of variables,
calculation of descriptive statistics, and observation of data
distributions. Please see cos.io/prereg for more information.

\hypertarget{explanation-of-existing-data}{%
\subsection{Explanation of existing
data}\label{explanation-of-existing-data}}

The data was collected as part of a larger collaboration on daily
intergroup relations. A sub-sample of variables has recently been
accessed by the research team for an unrelated analysis
\citep{Kreienkamp2022b}. Additionally, several of the variables were
prepared for a graphical presentation of the dataset. Thus far, none of
the proposed analyses have been conducted and none of the previous
analyses have been related to dimension reductions
\change{or longitudinal trajectories}{nor have any of the past analyses analysed the longitudinal nature of the data}.

\hypertarget{data-collection-procedures}{%
\subsection{Data collection
procedures}\label{data-collection-procedures}}

\add{We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands. }For
all three studies, the data was collected in a three-step procedure:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Entry Survey: A pre-measurement questionnaire (appr. 25 minutes)
  including demographic information, and relations to the Dutch majority
  (payment: 2 Euros).
\item
  Experience Recaps: At least 30 days of short reflection surveys (appr.
  3---5 minutes) on intergroup interactions twice a day (payment: 1 Euro
  per Recap; up to 2 Euros per day).
\item
  Conclusion Survey: On the last day, we conclude with a
  post-measurement questionnaire (appr. 25 minutes) with some questions
  on habits and reflections on the study (payment: 2 Euros).
\end{enumerate}

For the third study, participants had the option of continuing the study
for an unspecified amount of time. After the initial 30 day duration,
participants were offered the possibility to continue participating in
the study either with payment if daily diary measures were missed during
the initial study phase or without payment after a total of 60 daily
diary measurements were completed. After the initial 30-day period,
participants receive automated feedback visualizing the development of
their own well-being, attitudes, and motive responses as an additional
initiative and to give participants access to their own data and to
compensate study participation.

\hypertarget{sample-size}{%
\subsection{Sample size}\label{sample-size}}

For Study 1 (initial preliminary study) our target sample size is a sum
of 1,000 daily diary measurements. As we expected a completion rate of
around 80\% we aimed to recruite 20 participants (20 participants X 50
measurements). We further over-sampled slightly to compensate for
potential drop outs given the length and intensity of the study.
\add{The total collected sample size was thus 23 participants.}

We then used several key relationships (relevant to the broader research
team) to estimate sufficient power for a range of different analyses
(i.e., using power simulations). Based on these power simulations, for
Studies 2 and 3, our target sample size was a sum of 4,000 daily diary
measurements. With 100\% completion rate that would be archived with 67
participants (60 daily diary responses each). Given that we expect some
incomplete daily diary measurements, we aimed to recruit 80
participants. For \change{both studies}{Studies 2 and 3,} we again aimed
to over-sampled slightly to account for expected drop outs.
\add{We were ultimately able to recruited a total of 113 particpants for Study 2 and 71 participants for Study 3.}

\hypertarget{sample-size-rationale}{%
\subsection{Sample size rationale}\label{sample-size-rationale}}

The targeted sample size depended on a combination of different factors.
Different analyses were planned as part of the collaboration and
budgeting was a practical constraint. Some analyses were planned based
on (1) the pre- to post measurements, (2) the dynamic developments over
the daily diary measurements, or (3) the contemporary effects within the
daily diary measurements.

Power considerations of mixed effects model such as with extensive
longitudinal data are difficult to estimate because of the complex
covariance structures. Simulation studies based on the first sample
within this project indicated that with well-distributed scales, and
small to medium effect sizes, 70-80 participants with at least seven
daily diary measurements and a simple pre--post survey were sufficiently
powerful (power = .8, alpha = .05) to answer most of the key research
questions of the collaboration.

The ultimate sampling procedure decision was made as a practical
balancing of the number of participants and the number of measurements
provided by each participants.

\hypertarget{stopping-rule}{%
\subsection{Stopping rule}\label{stopping-rule}}

Participants were recruited until the targeted number of participant
finished the pre-measurement. Invitations to complete additional daily
diary measurements (in Study 3) were extended until participants chose
to leave the study or at the most until two months (i.e., 64 days) after
the initial entry survey (i.e., from the pre measurement survey).

\hypertarget{variables}{%
\section{Variables}\label{variables}}

\hypertarget{manipulated-variables}{%
\subsection{Manipulated variables}\label{manipulated-variables}}

Not applicable given that the study design is observational.

\hypertarget{measured-variables}{%
\subsection{Measured variables}\label{measured-variables}}

Given the circumstance that we utilize three independent experience
sampling studies for our analyses, the variables are at times slightly
different between studies. Additionally, given that the analysis we aim
to undertake (i.e., 3MPCA) uses a large number of variables for the
multiple imputations, dimension reduction, and correlational analyses,
we will not list all items individually. Instead, we provide full
variable information in the following data sheets:
`VariableSelectionDaily.xlsx' and `VariableSelectionPrePost.xlsx'.

\textbf{Key variables}\\
\add{While} most of the variables are identical across all studies,
\remove{however} some differ slightly in content or wording. However,
for each study we aimed to collect the following types of variables.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Motivational variables

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    core motive fulfillment (i.e., ``\emph{During your interaction with
    -X- (/this morning) your goal (-TEXT-) was fulfilled.}'')
  \item
    goal importance ratings of 10 individual goals (e.g., ``\emph{career
    goals}'', ``\emph{health / fitness goals}'')
  \item
    self determination theory needs (i.e., autonomy, relatedness,
    competence)
  \end{enumerate}
\item
  Affective states

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    experienced well-being (happiness, energy)
  \item
    general emotional state scale (e.g., ``\emph{How do you feel right
    now? not angry at all to very angry}'')
  \end{enumerate}
\item
  Behavioral self-reports (Studies 2 and 3)

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    pro-social behavior (e.g., ``\emph{Made demeaning, rude or
    derogatory remarks about someone.}'')
  \item
    anti-social behavior (e.g., ``\emph{I was there to listen to
    someone's problems.}'')
  \end{enumerate}
\item
  Cognitive measures

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    outgroup attitude (``\emph{After the interaction, how favorably do
    you feel towards \ldots{} / At the moment, how favorably do you feel
    towards \ldots{} the Dutch.}'')
  \item
    Allport's contact condition perceptions (i.e., equal status, shared
    goal, cooperative, voluntary)
  \item
    interaction quality ratings (e.g., ``\emph{Overall, the interaction
    was\ldots{} Unpleasant to Pleasant}'')
  \end{enumerate}
\end{enumerate}

It should be noted, that we collected substantially more motivational
and affecive variables, as they have been understudied in the past.

\textbf{Interpretation correlation variables}\\
To interpret the \add{person-mode} components identified by the 3MPCA,
we would like to correlate the component scores with a range of person-
or time point specific variables. While the full list of variables is
available in the accompanying data sheets (see
`VariableSelectionDaily.xlsx' and `VariableSelectionPrePost.xlsx'), we
provide a few examples of the targeted categories below.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  demographic variables

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    age
  \item
    gender
  \item
    \ldots{}
  \end{enumerate}
\item
  societal participation

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    language comprehension and -use
  \item
    work inclusion
  \item
    \ldots{}
  \end{enumerate}
\item
  adaptation indicators

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    social identification
  \item
    intergroup anxiety
  \item
    \ldots{}
  \end{enumerate}
\item
  everyday life obstacles

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    discrimination experiences
  \item
    negative life events
  \item
    \ldots{}
  \end{enumerate}
\item
  individual differences

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    big five personality scores
  \item
    attachment style
  \item
    \ldots{}
  \end{enumerate}
\end{enumerate}

\textbf{Auxiliary variables for multiple imputation}\\
For the missing data imputation we follow the procedure reported by
\citet{Monden2015}, and include any pre-, post-, or daily dairy
variables that are significantly (\(p\) \textless{} .01) and
meaningfully (\(r\) \textgreater{} .3) correlated with either the key
variables or missingness on the key variables. For an overview of the
variables that qualify as auxiliary variables see the enclosed data
sheets (see `VariableSelectionDaily.xlsx' and
`VariableSelectionPrePost.xlsx'). Additional information about the
multiple imputation procedure is also available at
\protect\hyperlink{statistical-models}{Statistical models} 2.

\emph{We also provide an exemplary codebook (using Study 3 as an
example; see `Codebook\_AOT-M\_ItemsPerSection.xlsx'). The full survey
files will be available as part of the main OSF repository connected to
this preregistration.}

\hypertarget{indices}{%
\subsection{Indices}\label{indices}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Mean Allport's conditions. We create a mean-averaged index of
  Allport's conditions in response to past findings indicating that the
  conditions are best conceptualized jointly and as functioning together
  rather than as fully independent factors
  \citep[p.~766]{Pettigrew2006}. Similar to past studies we thus hope to
  build a global indicator \citep[e.g., see][]{Pettigrew2006}. As with
  other indices we will ensure that the individual items indeed relate
  to a common latent construct and are meaningfully combine\add{d} in an
  index. If this is not possible we will create sub-indices and/or
  assess the impact of the conditions separately.

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    The interaction with {[}name interaction partner{]} was on equal
    footing (same status)
  \item
    {[}name interaction partner{]} shared your goal ({[}free-text entry
    interaction key need{]})
  \item
    The interaction with {[}name interaction partner{]} was cooperative
  \item
    The interaction with {[}name interaction partner{]} was voluntary
  \end{enumerate}
\item
  Mean belongingness during intergroup contact

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    I shared information about myself.
  \item
    {[}name interaction partner{]} shared information about themselves.
  \end{enumerate}
\item
  Mean alternative interaction quality definition (``Overall, the
  interaction with {[}name interaction partner{]} was \ldots{}'')

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Unpleasant to Pleasant
  \item
    Superficial to Meaningful
  \item
    Ineffective to Effective
  \item
    Unimportant to Important
  \end{enumerate}
\item
  Mood Subscales (MDMQ)

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    alertness
  \item
    calmness
  \item
    valence
  \end{enumerate}
\item
  Mean anti-social behaviors

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Put someone down
  \item
    Show little attention in someones opinion
  \item
    Demeaning remarks
  \item
    Inappropriately addressing someone
  \item
    Ignored or excluded someone
  \item
    Doubt someones judgement
  \item
    Unwanted attempts of personal matters
  \end{enumerate}
\item
  Mean pro-social behaviors

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Listen to someones problems
  \item
    Cheer someone up
  \item
    Help someone get things done
  \item
    Help someone with responsibilities
  \end{enumerate}
\end{enumerate}

Additionally, in the pre- and post-measurement surveys are several
validated scales that we will combine into their sub-scales. Indices
will be created in line with the accompanying validation literature:

\add{Added:}

\fbox{\parbox{\mylength}{
  \color{editPurple}
  \textbf{Note.}\\
  We have a group of six variables that are consistent across all three studies. Additionally, we have a number of variables that are either only relevant to a single study or a subset of the studies. If we agree on the study consolidation procedure as it is described below, I would suggest that we focus on the variables we have in all studies, and will include the other variables in the follow-up analyses and report only significant correlations (after alpha correction). I would also suggest that we add the list below to one of the variable selection files (Excel) instead of listing them all here (to avoid unnecessary confusion).
  \color{black}
}}

\color{editPurple}

Study 1 only:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Acculturation, Habits, and Interests Multicultural Scale for
  Adolescents (AHIMSA; S1)
\item
  Vancouver Index of Acculturation (VIA; S1)
\item
  Situational Self Awareness Scale (SAS; S1)
\end{enumerate}

Study 3 only:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Outgroup Trust
\item
  Work Group Inclusion
\item
  Meaningful Life Scale
\item
  Big Five Personality Questionnaire
\item
  Attachment Styles Scale
\item
  Paranoia Scale
\end{enumerate}

Studies 1 and 2:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Self Determination Theory Scale (SDT; S1, S2)
\item
  Interpersonal Support Evaluation List (ISEL; S1, S2)
\item
  Perceived Stress Scale (PSS; S1, S2)
\item
  Negative Life Events Scale (NLES; S1, S2)
\end{enumerate}

Studies 2 and 3:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Social Identification Scale (S2, S3)
\item
  Multidimensional Mood State Questionnaire (MDMQ; S2, S3)
\item
  Homesickness Scale (S2, S3)
\end{enumerate}

All Studies:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Intergroup Anxiety Scale (IAS)
\item
  The Satisfaction with Life Scale (SWL)
\item
  Rosenberg Self-Esteem Scale (SES)
\item
  Everyday Discrimination Scale
\item
  Language Proficiency
\item
  Association Participation
\end{enumerate}

\color{black}

\hypertarget{analysis-plan}{%
\section{Analysis Plan}\label{analysis-plan}}

\hypertarget{statistical-models}{%
\subsection{Statistical models}\label{statistical-models}}

For our analysis, we adapt the procedures outlined by
\citet{Monden2015}:

\fbox{\parbox{\mylength}{
  \color{editPurple}
  \textbf{Main Data Splits:}
  \begin{enumerate}
    \item \textbf{Three studies (S1, S2, S3)}
      \begin{enumerate}
        \item Combined for most power in correlations (including, correlation with study indicator)
        \item Separate in follow up analysis to ensure that we do not have a simpson's paradox
        \item Explore additional ideosyncratic variables in separated follow up
      \end{enumerate}
    \item \textbf{Two question types (interaction, no interaction)}:
      \begin{enumerate}
        \item Combined for most power in correlations (including, correlation with interaction indicator)
        \item Separate in subgroup follow up analysis to ensure that we do not have a simpson's paradox
      \end{enumerate}
    \item \textbf{Three time scales (half-day, day, week)}
      \begin{enumerate}
        \item If all variables clearly show single most variance on the same time-scale, single analysis only
        \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
        \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
      \end{enumerate}
    \item \textbf{Multiple Imputation (20 imputations)}
      \begin{enumerate}
        \item Test stability
        \item Split-half analysis
        \item Combine after analysis
      \end{enumerate}
  \end{enumerate}
  \color{black}
}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Sample Selection.} \add{For} our sample selection
  \remove{consists of three main steps. We (a) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data. }we
  address both the selection of time points and participants. Given that
  multiple imputation procedures even work well with large proportions
  of missing data {[}assuming missingness at random;
  \citet{Madley-Dowd2019}{]}, we decided on a general criterion of less
  than 45\% missingness to balance sample size retention and bias in the
  multiple imputation model. Thus, we then \remove{(b)} select the time
  points for which we have \textbackslash change\{responses from 55\% of
  the participants\}\{less than 45\% missingness\} and \remove{(c)}
  select participants who have \textbackslash change\{responded to at
  least 55\% of the surveys\}\{less than 45\% missingness\} across the
  selected time range.
  \add{Because the studies are comparable but not identical in their sampling strategy, question selection, and question wording, we also have to consider, whether to combine the samples. }
\item
  \textbf{Time point aggregation:}
  \add{We then aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.}
  Given that little data is available on the meaningful time scales of
  the selected psychological variables, we chose to determine the
  appropriate time scales using variance decomposition \citep[e.g.,
  see][]{Ram2014}. This is to say that we create multi-level
  unconditional means models (without predictors) that include possible
  nested time scales as levels.
  \change{We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.}{We chose to select time scales that align with common human cycles. We thus compare the variances of bi-daily, daily, and weekly aggregations. Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.}
  We then chose the ideal time scale for each of the key variables.
\end{enumerate}

\fbox{\parbox{\mylength}{
  \color{editPurple}
  \textbf{Note:} I talked about the individual time scales with Rei. And one issue that came up was that although the time points do not have to be equidistal between measurement ocasions, each variable needs to have data at each measured time point and the time scales should be equal across variables. This would mean that we will need to have the same time scale (e.g., daily) for all variables. Given that we have now only three time scales (i.e., bi-daily, daily, weekly) we could run the 3MPCA for all three time scales and assess differences (in a multi-verse style analysis).\\
  \textbf{Question:} I am not entirely sure I fully understand yet why it is not possible to aggregate items across different time windows as long as data is available for all time points. For example, if energy levels are kept at bi-daily measurements but for well-being we calculate a daily average, we could then replace both daily time points of well-being with the daily average. I understand that this sounds a bit silly because we give up the morning/afternoon variance but it might still be useful in terms of missingess reduction. Here my confused thoughts: Does this mean that 3MPCA cannot deal with variables that change at different speeds? Is this because the time components are estimated across all variables and might be mis-specified if we have different time scales? This seems counter my current understanding of the core array (i.e., interactions between the components) where time components could interact with variable components.\\
  Long story short: Maybe we can discuss this issue again, I feel like I might have misunderstood something somewhere :-D\\
  Sorry...
  \color{black}
}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \textbf{Multiple Imputation.} We then create 20 imputed datasets to
  perform the analyses on. As outlined by \citet{Monden2015}, we will
  use the key variables themself as well as auxiliary variables to
  impute missing values. Auxiliary variables are any pre-, post-, or
  daily dairy variables that are significantly (\(p\) \textless{} .01)
  and meaningfully (\(r\) \textgreater{} .3) correlated with (1) the key
  variables or (2) missingness on the key variables.
\item
  \textbf{Three-way ANOVA.} We assess the percentages of explained
  variance for the person-, variable-, and time aspects --- which offers
  an indication of whether a 3MPCA is useful for the dataset. Based on
  the grand mean centered variables we conduct a fixed-effects three-way
  ANOVA of the person-, variable-, and time modes as well as their
  various interactions. We are particularly interested in the highest
  order interaction term Person * Variable * Time + error, as a large
  amount of variance in this effect would speak towards the
  interdependence of the three modes.
\item
  \textbf{Data preprocessing.} For the \change{three mode }{3M}PCA, we
  center (across participants but within time point) and normalize
  (within variable but across time points) all key variables. The
  between-person centering, ensures that all variations are around the
  mean trend (which is removed by the centering). The normalization
  within variable are important for ensuring equal variances across
  variables, which ensures equal weighting of the variables in the
  3MPCA. For methodological detail see
  \protect\hyperlink{transformations}{Transformations}.
\item
  \textbf{Selection of 3MPCA model complexity} We then run the 3MPCA. We
  use a generalized scree plot to select the appropriate number of
  components for each of the three modes (i.e., person, variable, time).
  To test the stability of the solution results are compared across the
  20 imputed datasets \add{and split-half stability is assessed}.
\item
  \textbf{3MPCA model fitting.} For the selected complexities we then
  extract simple component structures (i.e., for each mode individually)
  as well as the \change{array core}{core array} (the full 3MPCA array,
  which
  \change{holds the weights of all combinations of the three modes}{specifies all combinations of the three mode components}).
  A Joint Orthomax orthogonal rotation and standard weights
  \remove{(when necessary)} are used to extract human interpretable
  component scores. This procedure is done on all 20 imputed data sets.
\item
  \textbf{Generalized Procrustes rotation.} In order to combine the
  three individual component structure from the 20 data sets, we use a
  generalized Procrustes rotation --- calculating the average of each
  compontent and core array.
\item
  \textbf{Explained variance.} As fit indices we then calculate (1) fit
  percentage of the estimated array for each imputed data set (i.e.,
  explained variance around the removed general trend) as well as (2) an
  \emph{overall} fit percentage of the estimated array (i.e., explained
  variance including the general trend). Both these fit metrices should
  give an indication of how much of the original variance the reduced
  component structure was able to capture.
\item
  \textbf{Component interpretation} To embed and interpret the resulting
  components (especially the person components), we calculate
  correlations of the component scores with key adaptation markers
  (including, well-being, anxiety, trust, or societal inclusion). If
  components remain unclear still we can additionally perform k-means
  clustering on the lower dimensional space to identify clearer groups,
  which can be compared on the adaptation markers if necessary.
\end{enumerate}

\hypertarget{transformations}{%
\subsection{Transformations}\label{transformations}}

\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{0pt}

For three-mode PCA it is custom to center (across participants but
within time point) and to normalize (within variable but across time
points). This is done to create a meaningful zero value for each time
point (to compare participants) and a variance of 1 across all time
points of a variable (ensuring equal weightin of variables in the
3MPCA).

We provide the main equations of the procedure below, where individual
\(i\) (\(i = 1,\ ...,\ I\)) reported
variable\change{ $k$ ($k = 1,\ ...,\ K$)}{ $x$ and its naturalized form $z$ (where $[k = 1,\ ...,\ K]$ indexes all available variables)}
at time point \(t\) (\(t = 1,\ ...,\ T\)), so that \(I\)=number of
persons, \(K\)=number of items, \(T\)=number of time points within the
dataset. For further information as well as an example illustration see
Supplemental Material 3 in \citet{Monden2015}.

\begin{equation} \label{eq:Mean}
  \overline{x}_{kt} =  \sum_{i=1}^I x_{ikt}
\end{equation}

\begin{equation} \label{eq:NormalizationSd}
  \sigma_{k} =  \sqrt{\sum_{i=1}^I \sum_{k=1}^K \frac{(x_{ikt}-\overline{x}_{kt})^2}{I*T}}
\end{equation}

\begin{equation} \label{eq:NormalizedScore}
  z_{ikt} =  \frac{x_{ikt}-\overline{x}_{kt}}{\sigma_{k}}
\end{equation}

\hypertarget{inference-criteria}{%
\subsection{Inference criteria}\label{inference-criteria}}

Many of the proposed procedures are descriptive rather then inferential.
Importantly, we will use amounts of variances explained as our main fit
indices. If we perform inferential analyses, we will use the standard
p\textless.05 criteria for determining whether the correlation and
regression coefficients are statistically significant.
\add{We will use appropriate multiple-comparison corrections whenever multiple tests are performed.}
We will place particular emphasis on effect sizes
\add{(e.g., correlations) }in our interpretations of the results
wherever \change{necessary}{possible}.

\hypertarget{data-exclusion}{%
\subsection{Data exclusion}\label{data-exclusion}}

No checks will be performed to determine eligibility for inclusion
besides verification that each subject answered each of the variables of
interest for a given analysis. Outliers will generally be included in
analyses, however we will use sensitivity analyses to assess the
robustness of the results to outliers.

\hypertarget{missing-data}{%
\subsection{Missing data}\label{missing-data}}

The sample selection based on proportions of missing data and the
multiple imputation procedures are described in
\protect\hyperlink{statistical-models}{Statistical models}.

\hypertarget{exploratory-analyses-optional}{%
\subsection{Exploratory analyses
(optional)}\label{exploratory-analyses-optional}}

A recent alternative for trajectory dimension reduction has been the use
of LSTM auto encoders. The encoder decoder classification procedure has
the advantage of being extremely flexible in the use of the data set and
could be interpreted in a similar way as the 3MPCA (see
\protect\hyperlink{statistical-models}{Statistical models} 9).

\hypertarget{other}{%
\section{Other}\label{other}}

\hypertarget{other-optional}{%
\subsection{Other (Optional)}\label{other-optional}}

Not applicable.

\hypertarget{section}{%
\subsection{}\label{section}}

\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}

\noindent

\bibliography{../references.bib}

\end{document}

### 2022-05-20 preregistration/Preregistration3MPCA.Rmd
pre post vars S1 added
feedback added

@@
     affiliation : "1"
-  - name        : Rei Tendeiro-Monden
+  - name        : Rei Monden
     affiliation : "2"
@@
     citation_package: natbib
+    keep_tex:  true
 editor_options: 
@@
 header-includes:
-   - \usepackage{amsmath, nccmath}
+   - \usepackage{amsmath}
+   - \usepackage{nccmath}
+   - \usepackage{LatexPackage/trackchanges}
+   # - \usepackage[finalnew]{LatexPackage/trackchanges}
+   # - \usepackage[finalold]{LatexPackage/trackchanges}
+   # - \usepackage[inline]{LatexPackage/trackchanges}
+   - \usepackage[usenames,dvipsnames]{xcolor}
+   - \addeditor{Jannis}
 ---
+
+\newlength{\mylength}
+\setlength{\fboxsep}{15pt}
+\setlength{\mylength}{\linewidth}
+\addtolength{\mylength}{-2\fboxsep}
+\addtolength{\mylength}{-2\fboxrule}
+\definecolor{editPurple}{RGB}{84, 28, 117}
 
@@
 In essence, the novel extensive longitudinal datasets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for top-down inferential modeling procedures^[For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).].
+Yet, past analytical advances have almost exclusively pushed for top-down inferential modeling procedures^[For example, stationary lagged regression models that assume stable means and variances over time (incl., vector \change{autoregression}{autoregressive} models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).].
 And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the developmental data patterns.
@@
 Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
-We we need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+\remove{We we need}\change{t}{T}o identify which variables are most important in the adaptation of migrants \add{over time}, we need methods to \change{identify clusters of similar individuals and developments}{break down the data heterogeneity into its core components (in terms of important variables and developments)} and we need \change{
+to assess whether such clusters differ in}{ways to identify how these core components relate to} key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
 There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
@@
 Recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [i.e., three-mode principle component analyses, 3MPCA; e.g., @Monden2015].
-We will, particularly, analyse three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+
+\change{We will, particularly,}{To this aim, we will} analyse the data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+\add{All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
+However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences} [@Kreienkamp2022d]\add{, the three studies have places a particular emphasis on the needs, goals, and motives of young migrants.}
 
@@
 **Observational Study**. Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, natural experiments, and regression discontinuity designs.
+\note{Note: This is part of the OSF dropdown answer and we can't change the description.}
 
@@
 
-All three studies used an extensive longitudinal design.
+All three studies\note{Note: I now extended the introduction of the three studies in the end of the description section.
+So I hope this doesn't come out of the blue as much anymore.} used an extensive longitudinal design.
 Using a daily diary format, for at least 30 days participants received a short survey twice per day (at around 12pm and 7pm).
@@
 Additionally, several of the variables were prepared for a graphical presentation of the dataset.
-Thus far, none of the proposed analyses have been conducted and none of the previous analyses have been related to dimension reductions or longitudinal trajectories.
+Thus far, none of the proposed analyses have been conducted and none of the previous analyses have been related to dimension reductions \change{or longitudinal trajectories}{nor have any of the past analyses analysed the longitudinal nature of the data}.
 
@@
 
-For all three studies, the data was collected in a three-step procedure:
+\add{We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands.
+}For all three studies, the data was collected in a three-step procedure:
 
@@
 We further over-sampled slightly to compensate for potential drop outs given the length and intensity of the study.
+\add{The total collected sample size was thus 23 participants.}
 
@@
 Given that we expect some incomplete daily diary measurements, we aimed to recruit 80 participants.
-For both studies we again over-sampled slightly to account for expected drop outs within each target population.
+For \change{both studies}{Studies 2 and 3,} we again aimed to over-sampled slightly to account for expected drop outs.
+\add{We were ultimately able to recruited a total of 113 particpants for Study 2 and 71 participants for Study 3.}
 
@@
 **Key variables**  
-Most of the variables are identical across all studies, however some differ slightly in content or wording. However, for each study we aimed to collect the following types of variables.
+\add{While} most of the variables are identical across all studies, \remove{however} some differ slightly in content or wording.
+However, for each study we aimed to collect the following types of variables.
 
@@
 **Interpretation correlation variables**  
-To interpret the components identified by the 3MPCA, we would like to correlate the component scores with a range of person- or time point specific variables.
+To interpret the \add{person-mode} components identified by the 3MPCA, we would like to correlate the component scores with a range of person- or time point specific variables.
 While the full list of variables is available in the accompanying data sheets (see 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'), we provide a few examples of the targeted categories below.
@@
 Similar to past studies we thus hope to build a global indicator [e.g., see @Pettigrew2006].
-As with other indices we will ensure that the individual items indeed relate to a common latent construct and are  meaningfully combine in an index.
+As with other indices we will ensure that the individual items indeed relate to a common latent construct and are  meaningfully combine\add{d} in an index.
 If this is not possible we will create sub-indices and/or assess the impact of the conditions separately.
@@
 
+Additionally, in the pre- and post-measurement surveys are several validated scales that we will combine into their sub-scales.
+Indices will be created in line with the accompanying validation literature:
+
+\add{Added:}
+
+\fbox{\parbox{\mylength}{
+  \color{editPurple}
+  \textbf{Note.}\\
+  We have a group of six variables that are consistent across all three studies.
+Additionally, we have a number of variables that are either only relevant to a single study or a subset of the studies.
+If we agree on the study consolidation procedure as it is described below, I would suggest that we focus on the variables we have in all studies, and will include the other variables in the follow-up analyses and report only significant correlations (after alpha correction).
+I would also suggest that we add the list below to one of the variable selection files (Excel) instead of listing them all here (to avoid unnecessary confusion).
+  \color{black}
+}}
+
+\color{editPurple}
+
+Study 1 only:
+
+1. Acculturation, Habits, and Interests Multicultural Scale for Adolescents (AHIMSA; S1)
+1. Vancouver Index of Acculturation (VIA; S1)
+1. Situational Self Awareness Scale (SAS; S1)
+
+Study 3 only:
+
+1. Outgroup Trust
+1. Work Group Inclusion
+1. Meaningful Life Scale
+1. Big Five Personality Questionnaire
+1. Attachment Styles Scale
+1. Paranoia Scale
+
+Studies 1 and 2:
+
+1. Self Determination Theory Scale (SDT; S1, S2)
+1. Interpersonal Support Evaluation List (ISEL; S1, S2)
+1. Perceived Stress Scale (PSS; S1, S2)
+1. Negative Life Events Scale (NLES; S1, S2)
+
+Studies 2 and 3:
+
+1. Social Identification Scale (S2, S3)
+1. Multidimensional Mood State Questionnaire (MDMQ; S2, S3)
+1. Homesickness Scale (S2, S3)
+
+All Studies:
+
+1. Intergroup Anxiety Scale (IAS)
+1. The Satisfaction with Life Scale (SWL)
+1. Rosenberg Self-Esteem Scale (SES)
+1. Everyday Discrimination Scale
+1. Language Proficiency
+1. Association Participation
+
+\color{black}
+
 # Analysis Plan
@@
 
+\fbox{\parbox{\mylength}{
+  \color{editPurple}
+  \textbf{Main Data Splits:}
+  \begin{enumerate}
+    \item \textbf{Three studies (S1, S2, S3)}
+      \begin{enumerate}
+        \item Combined for most power in correlations (including, correlation with study indicator)
+        \item Separate in follow up analysis to ensure that we do not have a simpson's paradox
+        \item Explore additional ideosyncratic variables in separated follow up
+      \end{enumerate}
+    \item \textbf{Two question types (interaction, no interaction)}:
+      \begin{enumerate}
+        \item Combined for most power in correlations (including, correlation with interaction indicator)
+        \item Separate in subgroup follow up analysis to ensure that we do not have a simpson's paradox
+      \end{enumerate}
+    \item \textbf{Three time scales (half-day, day, week)}
+      \begin{enumerate}
+        \item If all variables clearly show single most variance on the same time-scale, single analysis only
+        \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
+        \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
+      \end{enumerate}
+    \item \textbf{Multiple Imputation (20 imputations)}
+      \begin{enumerate}
+        \item Test stability
+        \item Split-half analysis
+        \item Combine after analysis
+      \end{enumerate}
+  \end{enumerate}
+  \color{black}
+}}
+
 1.
-**Sample Selection.** Our sample selection consists of three main steps.
+**Sample Selection.** \add{For} our sample selection \remove{consists of three main steps.
 We (a) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data.
-We then address the selection of time point and participants.
+}we address both the selection of time points and participants.
 Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
-Thus, we then (b) select the time points for which we have responses from 55% of the participants and (c) select participants who have responded to at least 55% of the surveys across the selected time range.
-    a.
-**Time point aggregation:** Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
+Thus, we then \remove{(b)} select the time points for which we have \change{responses from 55% of the participants}{less than 45% missingness} and \remove{(c)} select participants who have \change{responded to at least 55% of the surveys}{less than 45% missingness} across the selected time range.
+\add{Because the studies are comparable but not identical in their sampling strategy, question selection, and question wording, we also have to consider, whether to combine the samples.
+}
+2.
+**Time point aggregation:** \add{We then aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.} Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
 This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
-We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.
-    \fbox{
-    \parbox{0.5\linewidth}{%
-    Potentially nested measurements:\\
-    1 - 2 - 4 - 8 - 16 - 32 - 64\\
-    1 - 2 - 6 - 12 - 24 - 48\\
-    1 - 2 - 10 - 30 - 60\\
-    1 - 2 - 14 - 28 - 56\\
-    Weird but possible (uneven):\\
-    1 - 3 - 6 - 12 - 24 - 48\\
-    1 - 5 - 10 - 30 - 60\\
-    1 - 7 - 14 - 28 - 56
+\change{We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.}{We chose to select time scales that align with common human cycles.
+We thus compare the variances of bi-daily, daily, and weekly aggregations.
+Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.}
+We then chose the ideal time scale for each of the key variables.
+
+\fbox{\parbox{\mylength}{
+  \color{editPurple}
+  \textbf{Note:} I talked about the individual time scales with Rei.
+And one issue that came up was that although the time points do not have to be equidistal between measurement ocasions, each variable needs to have data at each measured time point and the time scales should be equal across variables.
+This would mean that we will need to have the same time scale (e.g., daily) for all variables.
+Given that we have now only three time scales (i.e., bi-daily, daily, weekly) we could run the 3MPCA for all three time scales and assess differences (in a multi-verse style analysis).\\
+  \textbf{Question:} I am not entirely sure I fully understand yet why it is not possible to aggregate items across different time windows as long as data is available for all time points.
+For example, if energy levels are kept at bi-daily measurements but for well-being we calculate a daily average, we could then replace both daily time points of well-being with the daily average.
+I understand that this sounds a bit silly because we give up the morning/afternoon variance but it might still be useful in terms of missingess reduction.
+Here my confused thoughts: Does this mean that 3MPCA cannot deal with variables that change at different speeds?
+Is this because the time components are estimated across all variables and might be mis-specified if we have different time scales?
+This seems counter my current understanding of the core array (i.e., interactions between the components) where time components could interact with variable components.\\
+  Long story short: Maybe we can discuss this issue again, I feel like I might have misunderstood something somewhere :-D\\
+  Sorry...
+  \color{black}
 }}
-    \vspace{12pt}  
-    We then chose the ideal time scale for each of the key variables.  
-\newlength{\mylength}
-\[
-\setlength{\fboxsep}{15pt}
-\setlength{\mylength}{\linewidth}
-\addtolength{\mylength}{-2\fboxsep}
-\addtolength{\mylength}{-2\fboxrule}
-\fbox{%
-\parbox{\mylength}{
-\setlength{\abovedisplayskip}{0pt}
-\setlength{\belowdisplayskip}{0pt}
-\textbf{Question:}  Should the aggregation happen after the multiple imputation?
-The current procedure would reduce missing data but the aggregated values would be based on differing amounts of data (depending on the amount of missingness).
-Yet again, a daily or weekly summary this is probably still more accurate then a single later recall response.
-}}
-\]
-2.
+
+3.
 **Multiple Imputation.** We then create 20 imputed datasets to perform the analyses on.
@@
 Auxiliary variables are any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
-3.
+4.
 **Three-way ANOVA.** We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset.
@@
 We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the interdependence of the three modes.
-\[
-\fbox{%
-\parbox{\linewidth}{
-\textbf{Question:}  What is the DV of the three-way ANOVA?
-}}
-\]
-4.
-**Data preprocessing.** For the three mode PCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
+5.
+**Data preprocessing.** For the \change{three mode }{3M}PCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
 The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering).
@@
 For methodological detail see [Transformations].
-5.
+6.
 **Selection of 3MPCA model complexity** We then run the 3MPCA.
 We use a generalized scree plot to select the appropriate number of components for each of the three modes (i.e., person, variable, time).
-To test the stability of the solution results are compared across the 20 imputed datasets.
-\[
-\fbox{%
-\parbox{\linewidth}{
-\textbf{Question:}  Additionally split-half procedures? How does this work is this the same as split-half reliability?
-}}
-\]
-6.
-**3MPCA model fitting.** For the selected complexities we then extract simple component structures (i.e., for each mode individually) as well as the array core (the full 3MPCA array, which holds the weights of all combinations of the three modes).
-A Joint Orthomax orthogonal rotation and standard weights (when necessary) are used to extract human interpretable component scores.
+To test the stability of the solution results are compared across the 20 imputed datasets \add{and split-half stability is assessed}.
+7.
+**3MPCA model fitting.** For the selected complexities we then extract simple component structures (i.e., for each mode individually) as well as the \change{array core}{core array} (the full 3MPCA array, which \change{holds the weights of all combinations of the three modes}{specifies all combinations of the three mode components}).
+A Joint Orthomax orthogonal rotation and standard weights \remove{(when necessary)} are used to extract human interpretable component scores.
 This procedure is done on all 20 imputed data sets.
-\[
-\fbox{%
-\parbox{\linewidth}{
-\textbf{Question:}  What exactly does the core array entail?
-In the @Monden2015 article it says: "the core-array, which describes the patterns of interaction between the symptom- and time-components in each person-component.".
-Additionally, in a more mathematical piece by Jorge it reads: "a three-way array called the core array, which holds the weights for the joint impact of any triple of components (one from each mode)." Does this mean that this is basically an array of loading scores in the lower dimensional space (similar to fitted values in regressions)?
-Or does it store the weights (i.e., interactions scores) that combines the loading scores of all three dimension reductions (not sure I express this right :-D )?
-I think this comes down to: What is the interpretation of a core element?
-}}
-\]
-7.
+8.
 **Generalized Procrustes rotation.** In order to combine the three individual component structure from the 20 data sets, we use a generalized Procrustes rotation --- calculating the average of each compontent and core array.
-<!-- The 3PCA model has been proposed by Tucker [114].
-Given an array X of order I×J×K, 3PCA determines component matrices for each of the three modes (individuals, variables and occasions) and a three-way array called the core array, which holds the weights for the joint impact of any triple of components (one from each mode).
--->
-8.
+9.
 **Explained variance.** As fit indices we then calculate (1) fit percentage of the estimated array for each imputed data set (i.e., explained variance around the removed general trend) as well as (2) an _overall_ fit percentage of the estimated array (i.e., explained variance including the general trend).
 Both these fit metrices should give an indication of how much of the original variance the reduced component structure was able to capture.
-9.
+10.
 **Component interpretation** To embed and interpret the resulting components (especially the person components), we calculate correlations of the component scores with key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
@@
 
-We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable $k$ ($k = 1,\ ...,\ K$) at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points within the dataset.
+We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable\change{ $k$ ($k = 1,\ ...,\ K$)}{ $x$ and its naturalized form $z$ (where $[k = 1,\ ...,\ K]$ indexes all available variables)} at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points within the dataset.
 For further information as well as an example illustration see Supplemental Material 3 in @Monden2015.
@@
 If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant.
-We will place particular emphasis on effect sizes in our interpretations of the results wherever necessary.
+\add{We will use appropriate multiple-comparison corrections whenever multiple tests are performed.} We will place particular emphasis on effect sizes \add{(e.g., correlations) }in our interpretations of the results wherever \change{necessary}{possible}.
 

### 2022-05-27 preregistration/Preregistration3MPCA.Rmd
Update Preregistration3MPCA.Rmd
feedback LFB included

--- 2022-05-14 preregistration/Preregistration3MPCA.Rmd
@@
     citation_package: natbib
+    keep_tex:  true
+urlcolor: BlueViolet
+citecolor: BlueViolet
+linkcolor: BlueViolet
 editor_options: 
@@
 In essence, the novel extensive longitudinal datasets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for top-down inferential modeling procedures^[For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).].
-And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the developmental data patterns.
+Yet, past analytical advances have almost exclusively pushed for inferential modeling procedures^[For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).].
+And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the initial developmental data patterns.
 
-As an example, little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
+As an example, a recent review of migration experiences has pointed out that despite many complex and dynamic theories, investigations of migrant adaptation have undervalued developmental data --- especially when it comes to the more internal experiences of motivations and emotions [@Kreienkamp2022d].
 Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
-We we need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+To identify which variables are most important in the adaptation of migrants over time, we need methods to break down the data heterogeneity into its core components (in terms of important variables and developments) and we need ways to identify how these core components relate to key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
 There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
 
-In this manuscript, we aim to assess the utility of one such promising analysis for the use of social psychological experience sampling data.
-Recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [i.e., three-mode principle component analyses, 3MPCA; e.g., @Monden2015].
-We will, particularly, analyse three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principle component analysis} (3MPCA).
+We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015].
+We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to decompose social psychological experiences.
+This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects [e.g., @Kreienkamp2022d; @MacInnis2015; @McKeown2017; @Pettigrew2011; @Ward2019].
+
+To untangle the heterogeneity in real-life migrant adaptations, we will analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
+However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences [@Kreienkamp2022d], the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
+To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.
 
@@
 We do not have hypotheses in the traditional sense.
-Our analysis plan is based on the aim of describing, summarizing, and understanding a complex data set of extensive longitudinal data.
+Our analysis plan is based on the aim of describing, summarizing, and understanding a complex set of extensive longitudinal data.
 We propose that the dimension reduction procedure we employ will identify meaningful patterns and developments within the data, which are useful to migration researchers and practitioners.
+An additional aim is to test the feasibility and utility of simultaneous dimension reduction procedures in experience sampling data.
 
@@
 Additionally, several of the variables were prepared for a graphical presentation of the dataset.
-Thus far, none of the proposed analyses have been conducted and none of the previous analyses have been related to dimension reductions or longitudinal trajectories.
+Thus far, none of the proposed analyses have been conducted and none of the previous analyses have been related to dimension reductions or time-series modeling.
 
@@
 
+We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands.
 For all three studies, the data was collected in a three-step procedure:
@@
 We further over-sampled slightly to compensate for potential drop outs given the length and intensity of the study.
+The total collected sample size of Study 1 was thus 1,225 survey responses from 23 participants.
 
-We then used several key relationships (relevant to the broader research team) to estimate sufficient power for a range of different analyses (i.e., using power simulations).
-Based on these power simulations, for Studies 2 and 3, our target sample size was a sum of 4,000 daily diary measurements.
+For Studies 2 and 3, our target sample size was a sum of 4,000 daily diary measurements each.
 With 100% completion rate that would be archived with 67 participants (60 daily diary responses each).
-Given that we expect some incomplete daily diary measurements, we aimed to recruit 80 participants.
-For both studies we again over-sampled slightly to account for expected drop outs within each target population.
+Given that we expect some incomplete daily diary measurements, we again over-sampled in both studies.
+In Study 2, we were ultimately able to collect 4,965 survey responses from 113 participants.
+In Study 3 we collected 4,107 survey responses from 71 participants.
 
@@
 Power considerations of mixed effects model such as with extensive longitudinal data are difficult to estimate because of the complex covariance structures.
-Simulation studies based on the first sample within this project indicated that with well-distributed scales, and small to medium effect sizes, 70-80 participants with at least seven daily diary measurements and a simple pre--post survey were sufficiently powerful (power = .8, alpha = .05) to answer most of the key research questions of the collaboration.
+Simulation studies based on the first sample indicated that with well-distributed scales, and small to medium effect sizes, 70-80 participants with at least seven daily diary measurements and a simple pre--post survey were sufficiently powerful (power = .8, alpha = .05) to answer most of the key research questions of the collaboration.
 
@@
 **Key variables**  
-Most of the variables are identical across all studies, however some differ slightly in content or wording. However, for each study we aimed to collect the following types of variables.
+While most of the variables are identical across all studies, some differ slightly in content or wording. However, for each study we aimed to collect the following types of variables.
 
@@
 
+For the detailed variable selection of each analysis see 'VariableSelectionAnalyses.xlsx'.
+
 **Interpretation correlation variables**  
-To interpret the components identified by the 3MPCA, we would like to correlate the component scores with a range of person- or time point specific variables.
-While the full list of variables is available in the accompanying data sheets (see 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'), we provide a few examples of the targeted categories below.
+To interpret the person-mode components identified by the 3MPCA, we would like to correlate the component scores with a range of person- or time point specific variables.
+While the full list of variables is available in the accompanying data sheets (see 'VariableSelectionAnalyses.xlsx'), we provide a few examples of the targeted categories below.
 
@@
 **Auxiliary variables for multiple imputation**  
-For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with either the key variables or missingness on the key variables.
-For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx').
+For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully correlated ($r$ > .3) with either the key variables or missingness on the key variables.
+For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see 'VariableSelectionAnalyses.xlsx').
 Additional information about the multiple imputation procedure is also available at [Statistical models] 2.
@@
 Example: We will take the mean of the two questions above to create a single measure of 'brownie enjoyment.'  -->
-
+˜
 1.
@@
 Similar to past studies we thus hope to build a global indicator [e.g., see @Pettigrew2006].
-As with other indices we will ensure that the individual items indeed relate to a common latent construct and are  meaningfully combine in an index.
+As with other indices we will ensure that the individual items indeed relate to a common latent construct and are  meaningfully combined in an index.
 If this is not possible we will create sub-indices and/or assess the impact of the conditions separately.
@@
     a. Help someone with responsibilities
+
+Additionally, in the pre- and post-measurement surveys are several validated scales that we will combine into their sub-scales.
+Indices will be created in line with their respective validation literature.
+For a full list of all variables and indices see 'VariableSelectionAnalyses.xlsx'.
+
+\color{red}
+
+\textbf{TO BE REMOVED ONCE 'VariableSelectionAnalyses.xlsx' IS DONE!}
+
+\color{black}
+
+\fbox{\parbox{\mylength}{
+  \color{editPurple}
+  \textbf{Note.}\\
+  We have a group of six variables that are consistent across all three studies.
+Additionally, we have a number of variables that are either only relevant to a single study or a subset of the studies.
+If we agree on the study consolidation procedure as it is described below, I would suggest that we focus on the variables we have in all studies, and will include the other variables in the follow-up analyses and report only significant correlations (after alpha correction).
+I would also suggest that we add the list below to one of the variable selection files (Excel) instead of listing them all here (to avoid unnecessary confusion).
+  \color{black}
+}}
+
+\color{editPurple}
+
+Study 1 only:
+
+1. Acculturation, Habits, and Interests Multicultural Scale for Adolescents (AHIMSA; S1)
+1. Vancouver Index of Acculturation (VIA; S1)
+1. Situational Self Awareness Scale (SAS; S1)
+
+Study 3 only:
+
+1. Outgroup Trust
+1. Work Group Inclusion
+1. Meaningful Life Scale
+1. Big Five Personality Questionnaire
+1. Attachment Styles Scale
+1. Paranoia Scale
+
+Studies 1 and 2:
+
+1. Self Determination Theory Scale (SDT; S1, S2)
+1. Interpersonal Support Evaluation List (ISEL; S1, S2)
+1. Perceived Stress Scale (PSS; S1, S2)
+1. Negative Life Events Scale (NLES; S1, S2)
+
+Studies 2 and 3:
+
+1. Social Identification Scale (S2, S3)
+1. Multidimensional Mood State Questionnaire (MDMQ; S2, S3)
+1. Homesickness Scale (S2, S3)
+
+All Studies:
+
+1. Intergroup Anxiety Scale (IAS)
+1. The Satisfaction with Life Scale (SWL)
+1. Rosenberg Self-Esteem Scale (SES)
+1. Everyday Discrimination Scale
+1. Language Proficiency
+1. Association Participation
+
+\color{black}
 
@@
 1.
-**Sample Selection.** Our sample selection consists of three main steps.
-We (a) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data.
-We then address the selection of time point and participants.
+**Sample Selection.** For our sample selection we address both the selection of time points and participants.
 Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
-Thus, we then (b) select the time points for which we have responses from 55% of the participants and (c) select participants who have responded to at least 55% of the surveys across the selected time range.
-    a.
-**Time point aggregation:** Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
+Thus, we then select the time points for which we have less than 45% missingness and select participants who have less than 45% missingness across the selected time range.
+Because the studies are comparable but not identical in their sampling strategy, question selection, and question wording, we also have to consider, whether to combine the samples.
+2.
+**Time point aggregation.** We then aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.
+Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
 This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
-We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.
-    \fbox{
-    \parbox{0.5\linewidth}{%
-    Potentially nested measurements:\\
-    1 - 2 - 4 - 8 - 16 - 32 - 64\\
-    1 - 2 - 6 - 12 - 24 - 48\\
-    1 - 2 - 10 - 30 - 60\\
-    1 - 2 - 14 - 28 - 56\\
-    Weird but possible (uneven):\\
-    1 - 3 - 6 - 12 - 24 - 48\\
-    1 - 5 - 10 - 30 - 60\\
-    1 - 7 - 14 - 28 - 56
-    }}  
-    \vspace{12pt}  
+We chose to select time scales that align with common human cycles.
+We thus compare the variances of bi-daily, daily, and weekly aggregations.
+Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.
 We then chose the ideal time scale for each of the key variables.
-\newlength{\mylength}
-\[
-\setlength{\fboxsep}{15pt}
-\setlength{\mylength}{\linewidth}
-\addtolength{\mylength}{-2\fboxsep}
-\addtolength{\mylength}{-2\fboxrule}
-\fbox{%
-\parbox{\mylength}{
-\setlength{\abovedisplayskip}{0pt}
-\setlength{\belowdisplayskip}{0pt}
-\textbf{Question:}  Should the aggregation happen after the multiple imputation?
-The current procedure would reduce missing data but the aggregated values would be based on differing amounts of data (depending on the amount of missingness).
-Yet again, a daily or weekly summary this is probably still more accurate then a single later recall response.
-}}
-\]
-2.
+3.
 **Multiple Imputation.** We then create 20 imputed datasets to perform the analyses on.
@@
 Auxiliary variables are any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
-3.
-**Three-way ANOVA.** We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset.
+4.
+**Three-way ANOVA\*.** We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset.
 Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions.
 We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the interdependence of the three modes.
-\[
-\fbox{%
-\parbox{\linewidth}{
-\textbf{Question:}  What is the DV of the three-way ANOVA?
-}}
-\]
-4.
-**Data preprocessing.** For the three mode PCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
+5.
+**Data preprocessing\*.** For the 3MPCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
 The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering).
@@
 For methodological detail see [Transformations].
-5.
-**Selection of 3MPCA model complexity** We then run the 3MPCA.
+6.
+**Selection of 3MPCA model complexity\*.** We then run the 3MPCA.
 We use a generalized scree plot to select the appropriate number of components for each of the three modes (i.e., person, variable, time).
-To test the stability of the solution results are compared across the 20 imputed datasets.
-\[
-\fbox{%
-\parbox{\linewidth}{
-\textbf{Question:}  Additionally split-half procedures? How does this work is this the same as split-half reliability?
-}}
-\]
-6.
-**3MPCA model fitting.** For the selected complexities we then extract simple component structures (i.e., for each mode individually) as well as the array core (the full 3MPCA array, which holds the weights of all combinations of the three modes).
-A Joint Orthomax orthogonal rotation and standard weights (when necessary) are used to extract human interpretable component scores.
+To test the stability of the solution results are compared across the 20 imputed datasets and split-half stability is assessed.
+7.
+**3MPCA model fitting\*.** For the selected complexities we then extract simple component structures (i.e., for each mode individually) as well as the core array (the full 3MPCA array, which specifies all combinations of the three mode components).
+A Joint Orthomax orthogonal rotation and standard weights are used to extract human interpretable component scores.
 This procedure is done on all 20 imputed data sets.
-\[
-\fbox{%
-\parbox{\linewidth}{
-\textbf{Question:}  What exactly does the core array entail?
-In the @Monden2015 article it says: "the core-array, which describes the patterns of interaction between the symptom- and time-components in each person-component.".
-Additionally, in a more mathematical piece by Jorge it reads: "a three-way array called the core array, which holds the weights for the joint impact of any triple of components (one from each mode)." Does this mean that this is basically an array of loading scores in the lower dimensional space (similar to fitted values in regressions)?
-Or does it store the weights (i.e., interactions scores) that combines the loading scores of all three dimension reductions (not sure I express this right :-D )?
-I think this comes down to: What is the interpretation of a core element?
-}}
-\]
-7.
-**Generalized Procrustes rotation.** In order to combine the three individual component structure from the 20 data sets, we use a generalized Procrustes rotation --- calculating the average of each compontent and core array.
-<!-- The 3PCA model has been proposed by Tucker [114].
-Given an array X of order I×J×K, 3PCA determines component matrices for each of the three modes (individuals, variables and occasions) and a three-way array called the core array, which holds the weights for the joint impact of any triple of components (one from each mode).
--->
 8.
-**Explained variance.** As fit indices we then calculate (1) fit percentage of the estimated array for each imputed data set (i.e., explained variance around the removed general trend) as well as (2) an _overall_ fit percentage of the estimated array (i.e., explained variance including the general trend).
+**Generalized Procrustes rotation\*.** In order to combine the three individual component structure from the 20 data sets, we use a generalized Procrustes rotation --- calculating the average of each compontent and core array.
+9.
+**Explained variance\*.** As fit indices we then calculate (1) fit percentage of the estimated array for each imputed data set (i.e., explained variance around the removed general trend) as well as (2) an _overall_ fit percentage of the estimated array (i.e., explained variance including the general trend).
 Both these fit metrices should give an indication of how much of the original variance the reduced component structure was able to capture.
-9.
-**Component interpretation** To embed and interpret the resulting components (especially the person components), we calculate correlations of the component scores with key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
+10.
+**Component interpretation\*.** To embed and interpret the resulting components (especially the person components), we calculate correlations of the component scores with key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
 If components remain unclear still we can additionally perform k-means clustering on the lower dimensional space to identify clearer groups, which can be compared on the adaptation markers if necessary.
+
+\textbf{*Step-wise Follow-up Analyses:}
+\begin{enumerate}
+  \item \textbf{Three studies (S1, S2, S3)}
+    \begin{enumerate}
+      \item Main Analysis: Combined for most power in correlations (including, correlation with study indicator)
+      \item Separate in follow up analysis to ensure that we do not have a Simpson's paradox
+      \item Explore additional idiosyncratic variables in separated follow up
+    \end{enumerate}
+  \item \textbf{Two question types (interaction, no interaction)}:
+    \begin{enumerate}
+      \item Main Analysis: Combined for most power in correlations (including, correlation with interaction indicator)
+      \item Separate in subgroup follow up analysis to ensure that we do not have a Simpson's paradox
+    \end{enumerate}
+  \item \textbf{Three time scales (half-day, day, week)}
+    \begin{enumerate}
+      \item Main Analysis: If all variables clearly show single most variance on the same time-scale, single analysis only
+      \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
+      \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
+    \end{enumerate}
+\end{enumerate}
 
@@
 
-We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable $k$ ($k = 1,\ ...,\ K$) at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points within the dataset.
+We provide the main equations of the procedure below, where individual $i$ ($i = 1,\ ...,\ I$) reported variable $x$ and its naturalized form $z$ (where $[k = 1,\ ...,\ K]$ indexes all available variables) at time point $t$ ($t = 1,\ ...,\ T$), so that $I$=number of persons, $K$=number of items, $T$=number of time points within the dataset.
 For further information as well as an example illustration see Supplemental Material 3 in @Monden2015.
@@
 If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant.
-We will place particular emphasis on effect sizes in our interpretations of the results wherever necessary.
+We will use appropriate multiple-comparison corrections whenever multiple tests are performed.
+We will place particular emphasis on effect sizes (e.g., correlations) in our interpretations of the results wherever possible.
 

### 2022-05-27 preregistration/Preregistration3MPCA.tex
updated
feedback LFB included

@@
 colorlinks=true,
-citecolor=blue,
-urlcolor=blue,
-linkcolor=magenta,
+citecolor=BlueViolet,
+urlcolor=BlueViolet,
+linkcolor=BlueViolet,
 pdfborder={0 0 0}}
@@
   \postauthor{\end{changemargin}}
-\date{20. May 2022}
+\date{27. May 2022}
 \predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
@@
 people, over time, and across variables. Yet, past analytical advances
-have almost exclusively pushed for top-down inferential modeling
+have almost exclusively pushed for inferential modeling
 procedures\footnote{For example, stationary lagged regression models
   that assume stable means and variances over time (incl., vector
-  \change{autoregression}{autoregressive} models, dynamic structural
-  equation models, autoregressive integrated moving average models, and
-  cross-lagged panel analyses) or basic trajectory models (e.g., mixed
-  effects models, spline regression models, and latent growth curve
-  modeling).}. And while inferential model testing is certainly
-important, we still miss discussions of methods for the more fundamental
-task of describing, summarizing, and understanding the developmental
-data patterns.
+  autoregressive models, dynamic structural equation models,
+  autoregressive integrated moving average models, and cross-lagged
+  panel analyses) or basic trajectory models (e.g., mixed effects
+  models, spline regression models, and latent growth curve modeling).}.
+And while inferential model testing is certainly important, we still
+miss discussions of methods for the more fundamental task of describing,
+summarizing, and understanding the initial developmental data patterns.
 
-As an example, little data has thus far investigated the development of
-migration experiences and no research has assessed the co-development of
-multiple experience aspects. Yet, understanding how people differ in
-their migration trajectories, can be crucial in understanding adaptive
-and maladaptive patterns. \remove{We we need}\change{t}{T}o identify
-which variables are most important in the adaptation of migrants
-\add{over time}, we need methods to
-\change{identify clusters of similar individuals and developments}{break down the data heterogeneity into its core components (in terms of important variables and developments)}
-and we need \change{
-to assess whether such clusters differ in}{ways to identify how these core components relate to}
-key adaptation markers (including, well-being, intergroup anxiety,
-outgroup trust, or societal participation). There is, thus, a clear need
-to assess the utility of analysis procedures focused on the description
-and understanding of complex dynamical data.
+As an example, a recent review of migration experiences has pointed out
+that despite many complex and dynamic theories, investigations of
+migrant adaptation have undervalued developmental data --- especially
+when it comes to the more internal experiences of motivations and
+emotions \citep{Kreienkamp2022d}. Yet, understanding how people differ
+in their migration trajectories, can be crucial in understanding
+adaptive and maladaptive patterns. To identify which variables are most
+important in the adaptation of migrants over time, we need methods to
+break down the data heterogeneity into its core components (in terms of
+important variables and developments) and we need ways to identify how
+these core components relate to key adaptation markers (including,
+well-being, intergroup anxiety, outgroup trust, or societal
+participation). There is, thus, a clear need to assess the utility of
+analysis procedures focused on the description and understanding of
+complex dynamical data.
 
 In this manuscript, we aim to assess the utility of one such promising
-analysis for the use of social psychological experience sampling data.
-Recent studies have laid out the potential effectiveness of dimension
-reduction procedures, which can address the new forms of person-,
-variable-, and time point heterogeneity jointly \citep[i.e., three-mode
-principle component analyses, 3MPCA; e.g.,][]{Monden2015}.
+analysis technique --- \textit{three-mode principle component analysis}
+(3MPCA). We use 3MPCA specifically because recent studies have laid out
+the potential effectiveness of dimension reduction procedures, which can
+address the new forms of person-, variable-, and time point
+heterogeneity jointly \citep[e.g.,][]{Monden2015}. We are among the
+first to apply this analysis to experience sampling data and, to the
+best of our knowledge, we are the first to decompose social
+psychological experiences. This stands in stark contrast, to a renewed
+recognition that social psychological phenomena unfold over time, a
+rapid increase of extensive longitudinal data collections, and a growing
+interest in understanding the co-development of multiple experience
+aspects
+\citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
 
-\change{We will, particularly,}{To this aim, we will} analyse the data
-from three experience sampling studies, which followed the migration
-experiences of recent migrants to the Netherlands.
-\add{All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
-However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences}
-\citep{Kreienkamp2022d}\add{, the three studies have places a particular emphasis on the needs, goals, and motives of young migrants.}
+To untangle the heterogeneity in real-life migrant adaptations, we will
+analyse data from three experience sampling studies, which followed the
+migration experiences of recent migrants to the Netherlands. All three
+studies focus on the psychological adaptation of migrants (i.e.,
+psychological acculturation). However, given that past investigations of
+psychological acculturation have underexplored the crucial aspect of
+motivational experiences \citep{Kreienkamp2022d}, the three studies have
+placed a particular emphasis on the needs, goals, and motives of young
+migrants. To make full use of all three data sets and to guide future
+use of dimension reduction procedures with extensive psychological data,
+we will analyze the data in a descriptive and step wise manner, where we
+first consider the data jointly and will then assess potential
+differences between datasets, variables, and time scales.
 
@@
 based on the aim of describing, summarizing, and understanding a complex
-data set of extensive longitudinal data. We propose that the dimension
+set of extensive longitudinal data. We propose that the dimension
 reduction procedure we employ will identify meaningful patterns and
 developments within the data, which are useful to migration researchers
-and practitioners.
+and practitioners. An additional aim is to test the feasibility and
+utility of simultaneous dimension reduction procedures in experience
+sampling data.
 
@@
 experiments, and regression discontinuity designs.
-\note{Note: This is part of the OSF dropdown answer and we can't change the description.}
 
@@
 
-All three
-studies\note{Note: I now extended the introduction of the three studies in the end of the description section. So I hope this doesn't come out of the blue as much anymore.}
-used an extensive longitudinal design. Using a daily diary format, for
-at least 30 days participants received a short survey twice per day (at
-around 12pm and 7pm). We additionally included a longer pre- and post
-measurement survey the days before and after the extensive longitudinal
-data collection.
+All three studies used an extensive longitudinal design. Using a daily
+diary format, for at least 30 days participants received a short survey
+twice per day (at around 12pm and 7pm). We additionally included a
+longer pre- and post measurement survey the days before and after the
+extensive longitudinal data collection.
 
@@
 the proposed analyses have been conducted and none of the previous
-analyses have been related to dimension reductions
-\change{or longitudinal trajectories}{nor have any of the past analyses analysed the longitudinal nature of the data}.
+analyses have been related to dimension reductions or time-series
+modeling.
 
@@
 
-\add{We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands. }For
-all three studies, the data was collected in a three-step procedure:
+We collected three experience sampling studies, following the daily
+migration experiences of young migrants, who had recently arrived in the
+Netherlands. For all three studies, the data was collected in a
+three-step procedure:
 
@@
 measurements). We further over-sampled slightly to compensate for
-potential drop outs given the length and intensity of the study.
-\add{The total collected sample size was thus 23 participants.}
+potential drop outs given the length and intensity of the study. The
+total collected sample size of Study 1 was thus 1,225 survey responses
+from 23 participants.
 
-We then used several key relationships (relevant to the broader research
-team) to estimate sufficient power for a range of different analyses
-(i.e., using power simulations). Based on these power simulations, for
-Studies 2 and 3, our target sample size was a sum of 4,000 daily diary
-measurements. With 100\% completion rate that would be archived with 67
-participants (60 daily diary responses each). Given that we expect some
-incomplete daily diary measurements, we aimed to recruit 80
-participants. For \change{both studies}{Studies 2 and 3,} we again aimed
-to over-sampled slightly to account for expected drop outs.
-\add{We were ultimately able to recruited a total of 113 particpants for Study 2 and 71 participants for Study 3.}
+For Studies 2 and 3, our target sample size was a sum of 4,000 daily
+diary measurements each. With 100\% completion rate that would be
+archived with 67 participants (60 daily diary responses each). Given
+that we expect some incomplete daily diary measurements, we again
+over-sampled in both studies. In Study 2, we were ultimately able to
+collect 4,965 survey responses from 113 participants. In Study 3 we
+collected 4,107 survey responses from 71 participants.
 
@@
 covariance structures. Simulation studies based on the first sample
-within this project indicated that with well-distributed scales, and
-small to medium effect sizes, 70-80 participants with at least seven
-daily diary measurements and a simple pre--post survey were sufficiently
-powerful (power = .8, alpha = .05) to answer most of the key research
-questions of the collaboration.
+indicated that with well-distributed scales, and small to medium effect
+sizes, 70-80 participants with at least seven daily diary measurements
+and a simple pre--post survey were sufficiently powerful (power = .8,
+alpha = .05) to answer most of the key research questions of the
+collaboration.
 
@@
 \textbf{Key variables}\\
-\add{While} most of the variables are identical across all studies,
-\remove{however} some differ slightly in content or wording. However,
-for each study we aimed to collect the following types of variables.
+While most of the variables are identical across all studies, some
+differ slightly in content or wording. However, for each study we aimed
+to collect the following types of variables.
 
@@
 
+For the detailed variable selection of each analysis see
+`VariableSelectionAnalyses.xlsx'.
+
 \textbf{Interpretation correlation variables}\\
-To interpret the \add{person-mode} components identified by the 3MPCA,
-we would like to correlate the component scores with a range of person-
-or time point specific variables. While the full list of variables is
+To interpret the person-mode components identified by the 3MPCA, we
+would like to correlate the component scores with a range of person- or
+time point specific variables. While the full list of variables is
 available in the accompanying data sheets (see
-`VariableSelectionDaily.xlsx' and `VariableSelectionPrePost.xlsx'), we
-provide a few examples of the targeted categories below.
+`VariableSelectionAnalyses.xlsx'), we provide a few examples of the
+targeted categories below.
 
@@
 variables that are significantly (\(p\) \textless{} .01) and
-meaningfully (\(r\) \textgreater{} .3) correlated with either the key
+meaningfully correlated (\(r\) \textgreater{} .3) with either the key
 variables or missingness on the key variables. For an overview of the
 variables that qualify as auxiliary variables see the enclosed data
-sheets (see `VariableSelectionDaily.xlsx' and
-`VariableSelectionPrePost.xlsx'). Additional information about the
-multiple imputation procedure is also available at
+sheets (see `VariableSelectionAnalyses.xlsx'). Additional information
+about the multiple imputation procedure is also available at
 \protect\hyperlink{statistical-models}{Statistical models} 2.
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Mean Allport's conditions. We create a mean-averaged index of
+˜ 1. Mean Allport's conditions. We create a mean-averaged index of
 Allport's conditions in response to past findings indicating that the
 conditions are best conceptualized jointly and as functioning together
-  rather than as fully independent factors
-  \citep[p.~766]{Pettigrew2006}. Similar to past studies we thus hope to
-  build a global indicator \citep[e.g., see][]{Pettigrew2006}. As with
-  other indices we will ensure that the individual items indeed relate
-  to a common latent construct and are meaningfully combine\add{d} in an
-  index. If this is not possible we will create sub-indices and/or
-  assess the impact of the conditions separately.
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    The interaction with {[}name interaction partner{]} was on equal
-    footing (same status)
-  \item
-    {[}name interaction partner{]} shared your goal ({[}free-text entry
-    interaction key need{]})
-  \item
-    The interaction with {[}name interaction partner{]} was cooperative
-  \item
-    The interaction with {[}name interaction partner{]} was voluntary
-  \end{enumerate}
-\item
-  Mean belongingness during intergroup contact
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    I shared information about myself.
-  \item
-    {[}name interaction partner{]} shared information about themselves.
-  \end{enumerate}
-\item
-  Mean alternative interaction quality definition (``Overall, the
-  interaction with {[}name interaction partner{]} was \ldots{}'')
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    Unpleasant to Pleasant
-  \item
-    Superficial to Meaningful
-  \item
-    Ineffective to Effective
-  \item
-    Unimportant to Important
-  \end{enumerate}
-\item
-  Mood Subscales (MDMQ)
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    alertness
-  \item
-    calmness
-  \item
-    valence
-  \end{enumerate}
-\item
-  Mean anti-social behaviors
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    Put someone down
-  \item
-    Show little attention in someones opinion
-  \item
-    Demeaning remarks
-  \item
-    Inappropriately addressing someone
-  \item
-    Ignored or excluded someone
-  \item
-    Doubt someones judgement
-  \item
-    Unwanted attempts of personal matters
-  \end{enumerate}
-\item
-  Mean pro-social behaviors
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    Listen to someones problems
-  \item
-    Cheer someone up
-  \item
-    Help someone get things done
-  \item
-    Help someone with responsibilities
-  \end{enumerate}
-\end{enumerate}
+rather than as fully independent factors \citep[p.~766]{Pettigrew2006}.
+Similar to past studies we thus hope to build a global indicator
+\citep[e.g., see][]{Pettigrew2006}. As with other indices we will ensure
+that the individual items indeed relate to a common latent construct and
+are meaningfully combined in an index. If this is not possible we will
+create sub-indices and/or assess the impact of the conditions
+separately. a. The interaction with {[}name interaction partner{]} was
+on equal footing (same status) a. {[}name interaction partner{]} shared
+your goal ({[}free-text entry interaction key need{]}) a. The
+interaction with {[}name interaction partner{]} was cooperative a. The
+interaction with {[}name interaction partner{]} was voluntary 1. Mean
+belongingness during intergroup contact a. I shared information about
+myself. a. {[}name interaction partner{]} shared information about
+themselves. 1. Mean alternative interaction quality definition
+(``Overall, the interaction with {[}name interaction partner{]} was
+\ldots{}'') a. Unpleasant to Pleasant a. Superficial to Meaningful a.
+Ineffective to Effective a. Unimportant to Important 1. Mood Subscales
+(MDMQ) a. alertness a. calmness a. valence 1. Mean anti-social behaviors
+a. Put someone down a. Show little attention in someones opinion a.
+Demeaning remarks a. Inappropriately addressing someone a. Ignored or
+excluded someone a. Doubt someones judgement a. Unwanted attempts of
+personal matters 1. Mean pro-social behaviors a. Listen to someones
+problems a. Cheer someone up a. Help someone get things done a. Help
+someone with responsibilities
 
@@
 validated scales that we will combine into their sub-scales. Indices
-will be created in line with the accompanying validation literature:
+will be created in line with their respective validation literature. For
+a full list of all variables and indices see
+`VariableSelectionAnalyses.xlsx'.
 
-\add{Added:}
+\color{red}
+
+\textbf{TO BE REMOVED ONCE 'VariableSelectionAnalyses.xlsx' IS DONE!}
+
+\color{black}
 
@@
 
-\fbox{\parbox{\mylength}{
-  \color{editPurple}
-  \textbf{Main Data Splits:}
-  \begin{enumerate}
-    \item \textbf{Three studies (S1, S2, S3)}
-      \begin{enumerate}
-        \item Combined for most power in correlations (including, correlation with study indicator)
-        \item Separate in follow up analysis to ensure that we do not have a simpson's paradox
-        \item Explore additional ideosyncratic variables in separated follow up
-      \end{enumerate}
-    \item \textbf{Two question types (interaction, no interaction)}:
-      \begin{enumerate}
-        \item Combined for most power in correlations (including, correlation with interaction indicator)
-        \item Separate in subgroup follow up analysis to ensure that we do not have a simpson's paradox
-      \end{enumerate}
-    \item \textbf{Three time scales (half-day, day, week)}
-      \begin{enumerate}
-        \item If all variables clearly show single most variance on the same time-scale, single analysis only
-        \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
-        \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
-      \end{enumerate}
-    \item \textbf{Multiple Imputation (20 imputations)}
-      \begin{enumerate}
-        \item Test stability
-        \item Split-half analysis
-        \item Combine after analysis
-      \end{enumerate}
-  \end{enumerate}
-  \color{black}
-}}
-
 \begin{enumerate}
@@
 \item
-  \textbf{Sample Selection.} \add{For} our sample selection
-  \remove{consists of three main steps. We (a) aggregate the key variables over time to archive a reasonably interpretable number of time points and the remove a first proportion of missing data. }we
-  address both the selection of time points and participants. Given that
-  multiple imputation procedures even work well with large proportions
-  of missing data {[}assuming missingness at random;
-  \citet{Madley-Dowd2019}{]}, we decided on a general criterion of less
-  than 45\% missingness to balance sample size retention and bias in the
-  multiple imputation model. Thus, we then \remove{(b)} select the time
-  points for which we have \textbackslash change\{responses from 55\% of
-  the participants\}\{less than 45\% missingness\} and \remove{(c)}
-  select participants who have \textbackslash change\{responded to at
-  least 55\% of the surveys\}\{less than 45\% missingness\} across the
-  selected time range.
-  \add{Because the studies are comparable but not identical in their sampling strategy, question selection, and question wording, we also have to consider, whether to combine the samples. }
+  \textbf{Sample Selection.} For our sample selection we address both
+  the selection of time points and participants. Given that multiple
+  imputation procedures even work well with large proportions of missing
+  data {[}assuming missingness at random; \citet{Madley-Dowd2019}{]}, we
+  decided on a general criterion of less than 45\% missingness to
+  balance sample size retention and bias in the multiple imputation
+  model. Thus, we then select the time points for which we have less
+  than 45\% missingness and select participants who have less than 45\%
+  missingness across the selected time range. Because the studies are
+  comparable but not identical in their sampling strategy, question
+  selection, and question wording, we also have to consider, whether to
+  combine the samples.
 \item
-  \textbf{Time point aggregation:}
-  \add{We then aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.}
-  Given that little data is available on the meaningful time scales of
-  the selected psychological variables, we chose to determine the
-  appropriate time scales using variance decomposition \citep[e.g.,
-  see][]{Ram2014}. This is to say that we create multi-level
-  unconditional means models (without predictors) that include possible
-  nested time scales as levels.
-  \change{We, for example, assess the explained variances for measurements nested in days, nested in weeks, nested in months.}{We chose to select time scales that align with common human cycles.
-We thus compare the variances of bi-daily, daily, and weekly aggregations.
-Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.}
-  We then chose the ideal time scale for each of the key variables.
-\end{enumerate}
-
-\fbox{\parbox{\mylength}{
-  \color{editPurple}
-  \textbf{Note:} I talked about the individual time scales with Rei.
-And one issue that came up was that although the time points do not have to be equidistal between measurement ocasions, each variable needs to have data at each measured time point and the time scales should be equal across variables.
-This would mean that we will need to have the same time scale (e.g., daily) for all variables.
-Given that we have now only three time scales (i.e., bi-daily, daily, weekly) we could run the 3MPCA for all three time scales and assess differences (in a multi-verse style analysis).\\
-  \textbf{Question:} I am not entirely sure I fully understand yet why it is not possible to aggregate items across different time windows as long as data is available for all time points.
-For example, if energy levels are kept at bi-daily measurements but for well-being we calculate a daily average, we could then replace both daily time points of well-being with the daily average.
-I understand that this sounds a bit silly because we give up the morning/afternoon variance but it might still be useful in terms of missingess reduction.
-Here my confused thoughts: Does this mean that 3MPCA cannot deal with variables that change at different speeds?
-Is this because the time components are estimated across all variables and might be mis-specified if we have different time scales?
-This seems counter my current understanding of the core array (i.e., interactions between the components) where time components could interact with variable components.\\
-  Long story short: Maybe we can discuss this issue again, I feel like I might have misunderstood something somewhere :-D\\
-  Sorry...
-  \color{black}
-}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
+  \textbf{Time point aggregation.} We then aggregate the key variables
+  over time to archive a reasonably interpretable number of time points
+  and remove a first proportion of missing data. Given that little data
+  is available on the meaningful time scales of the selected
+  psychological variables, we chose to determine the appropriate time
+  scales using variance decomposition \citep[e.g., see][]{Ram2014}. This
+  is to say that we create multi-level unconditional means models
+  (without predictors) that include possible nested time scales as
+  levels. We chose to select time scales that align with common human
+  cycles. We thus compare the variances of bi-daily, daily, and weekly
+  aggregations. Additional aggregations of two weeks or the full four
+  weeks might be possible but would most likely reduce the variance too
+  much for any meaningul further reduction during the 3MPCA. We then
+  chose the ideal time scale for each of the key variables.
 \item
@@
 \item
-  \textbf{Three-way ANOVA.} We assess the percentages of explained
+  \textbf{Three-way ANOVA*.} We assess the percentages of explained
   variance for the person-, variable-, and time aspects --- which offers
@@
 \item
-  \textbf{Data preprocessing.} For the \change{three mode }{3M}PCA, we
-  center (across participants but within time point) and normalize
-  (within variable but across time points) all key variables. The
-  between-person centering, ensures that all variations are around the
-  mean trend (which is removed by the centering). The normalization
-  within variable are important for ensuring equal variances across
-  variables, which ensures equal weighting of the variables in the
-  3MPCA. For methodological detail see
-  \protect\hyperlink{transformations}{Transformations}.
+  \textbf{Data preprocessing*.} For the 3MPCA, we center (across
+  participants but within time point) and normalize (within variable but
+  across time points) all key variables. The between-person centering,
+  ensures that all variations are around the mean trend (which is
+  removed by the centering). The normalization within variable are
+  important for ensuring equal variances across variables, which ensures
+  equal weighting of the variables in the 3MPCA. For methodological
+  detail see \protect\hyperlink{transformations}{Transformations}.
 \item
-  \textbf{Selection of 3MPCA model complexity} We then run the 3MPCA. We
-  use a generalized scree plot to select the appropriate number of
+  \textbf{Selection of 3MPCA model complexity*.} We then run the 3MPCA.
+  We use a generalized scree plot to select the appropriate number of
   components for each of the three modes (i.e., person, variable, time).
   To test the stability of the solution results are compared across the
-  20 imputed datasets \add{and split-half stability is assessed}.
+  20 imputed datasets and split-half stability is assessed.
 \item
-  \textbf{3MPCA model fitting.} For the selected complexities we then
+  \textbf{3MPCA model fitting*.} For the selected complexities we then
   extract simple component structures (i.e., for each mode individually)
-  as well as the \change{array core}{core array} (the full 3MPCA array,
-  which
-  \change{holds the weights of all combinations of the three modes}{specifies all combinations of the three mode components}).
-  A Joint Orthomax orthogonal rotation and standard weights
-  \remove{(when necessary)} are used to extract human interpretable
-  component scores. This procedure is done on all 20 imputed data sets.
+  as well as the core array (the full 3MPCA array, which specifies all
+  combinations of the three mode components). A Joint Orthomax
+  orthogonal rotation and standard weights are used to extract human
+  interpretable component scores. This procedure is done on all 20
+  imputed data sets.
 \item
-  \textbf{Generalized Procrustes rotation.} In order to combine the
+  \textbf{Generalized Procrustes rotation*.} In order to combine the
   three individual component structure from the 20 data sets, we use a
@@
 \item
-  \textbf{Explained variance.} As fit indices we then calculate (1) fit
+  \textbf{Explained variance*.} As fit indices we then calculate (1) fit
   percentage of the estimated array for each imputed data set (i.e.,
@@
 \item
-  \textbf{Component interpretation} To embed and interpret the resulting
-  components (especially the person components), we calculate
+  \textbf{Component interpretation*.} To embed and interpret the
+  resulting components (especially the person components), we calculate
   correlations of the component scores with key adaptation markers
@@
   which can be compared on the adaptation markers if necessary.
+\end{enumerate}
+
+\textbf{*Step-wise Follow-up Analyses:}
+
+\begin{enumerate}
+  \item \textbf{Three studies (S1, S2, S3)}
+    \begin{enumerate}
+      \item Main Analysis: Combined for most power in correlations (including, correlation with study indicator)
+      \item Separate in follow up analysis to ensure that we do not have a Simpson's paradox
+      \item Explore additional idiosyncratic variables in separated follow up
+    \end{enumerate}
+  \item \textbf{Two question types (interaction, no interaction)}:
+    \begin{enumerate}
+      \item Main Analysis: Combined for most power in correlations (including, correlation with interaction indicator)
+      \item Separate in subgroup follow up analysis to ensure that we do not have a Simpson's paradox
+    \end{enumerate}
+  \item \textbf{Three time scales (half-day, day, week)}
+    \begin{enumerate}
+      \item Main Analysis: If all variables clearly show single most variance on the same time-scale, single analysis only
+      \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
+      \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
+    \end{enumerate}
 \end{enumerate}
@@
 We provide the main equations of the procedure below, where individual
-\(i\) (\(i = 1,\ ...,\ I\)) reported
-variable\change{ $k$ ($k = 1,\ ...,\ K$)}{ $x$ and its naturalized form $z$ (where $[k = 1,\ ...,\ K]$ indexes all available variables)}
+\(i\) (\(i = 1,\ ...,\ I\)) reported variable \(x\) and its naturalized
+form \(z\) (where \([k = 1,\ ...,\ K]\) indexes all available variables)
 at time point \(t\) (\(t = 1,\ ...,\ T\)), so that \(I\)=number of
@@
 p\textless.05 criteria for determining whether the correlation and
-regression coefficients are statistically significant.
-\add{We will use appropriate multiple-comparison corrections whenever multiple tests are performed.}
-We will place particular emphasis on effect sizes
-\add{(e.g., correlations) }in our interpretations of the results
-wherever \change{necessary}{possible}.
+regression coefficients are statistically significant. We will use
+appropriate multiple-comparison corrections whenever multiple tests are
+performed. We will place particular emphasis on effect sizes (e.g.,
+correlations) in our interpretations of the results wherever possible.
 

### 2022-05-27 preregistration/Preregistration3MPCA.Rmd
feedback added

@@
    - \usepackage{nccmath}
+   - \usepackage{enumerate}
    - \usepackage{LatexPackage/trackchanges}
@@
 -->
+
+\textbf{Shortened Description:}
+
+Recent reviews have called for more comprehensive assessment of human experiences and for more longitudinal real-life data, within the psychological sciences more broadly and in migration research in particular [e.g., @Kreienkamp2022d; @MacInnis2015; @McKeown2017; @Pettigrew2011; @Ward2019].
+However, while generally speaking analytical methods for such, more complex, data have become more readily available [e.g., @ODonnell2021], it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time.
+
+In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principle component analysis} (3MPCA).
+We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015].
+We analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
+However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences [@Kreienkamp2022d], the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
+To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.
+
+---
+
+\textbf{Full Description:}
 
@@
 Additionally, given that the analysis we aim to undertake (i.e., 3MPCA) uses a large number of variables for the multiple imputations, dimension reduction, and correlational analyses, we will not list all items individually.
-Instead, we provide full variable information in the following data sheets: 'VariableSelectionDaily.xlsx' and 'VariableSelectionPrePost.xlsx'.
+Instead, we provide full variable information in the following data sheets in `VariableSelectionAnalyses.xlsx'.
 
@@
 
-For the detailed variable selection of each analysis see 'VariableSelectionAnalyses.xlsx'.
+For the detailed variable selection of each analysis see `VariableSelectionAnalyses.xlsx'.
 
@@
 To interpret the person-mode components identified by the 3MPCA, we would like to correlate the component scores with a range of person- or time point specific variables.
-While the full list of variables is available in the accompanying data sheets (see 'VariableSelectionAnalyses.xlsx'), we provide a few examples of the targeted categories below.
+While the full list of variables is available in the accompanying data sheets (see `VariableSelectionAnalyses.xlsx'), we provide a few examples of the targeted categories below.
 
@@
 For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully correlated ($r$ > .3) with either the key variables or missingness on the key variables.
-For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see 'VariableSelectionAnalyses.xlsx').
+For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see `VariableSelectionAnalyses.xlsx').
 Additional information about the multiple imputation procedure is also available at [Statistical models] 2.
@@
 Example: We will take the mean of the two questions above to create a single measure of 'brownie enjoyment.'  -->
-˜
+
 1.
@@
 Indices will be created in line with their respective validation literature.
-For a full list of all variables and indices see 'VariableSelectionAnalyses.xlsx'.
-
-\color{red}
-
-\textbf{TO BE REMOVED ONCE 'VariableSelectionAnalyses.xlsx' IS DONE!}
-
-\color{black}
-
-\fbox{\parbox{\mylength}{
-  \color{editPurple}
-  \textbf{Note.}\\
-  We have a group of six variables that are consistent across all three studies.
-Additionally, we have a number of variables that are either only relevant to a single study or a subset of the studies.
-If we agree on the study consolidation procedure as it is described below, I would suggest that we focus on the variables we have in all studies, and will include the other variables in the follow-up analyses and report only significant correlations (after alpha correction).
-I would also suggest that we add the list below to one of the variable selection files (Excel) instead of listing them all here (to avoid unnecessary confusion).
-  \color{black}
-}}
-
-\color{editPurple}
-
-Study 1 only:
-
-1. Acculturation, Habits, and Interests Multicultural Scale for Adolescents (AHIMSA; S1)
-1. Vancouver Index of Acculturation (VIA; S1)
-1. Situational Self Awareness Scale (SAS; S1)
-
-Study 3 only:
-
-1. Outgroup Trust
-1. Work Group Inclusion
-1. Meaningful Life Scale
-1. Big Five Personality Questionnaire
-1. Attachment Styles Scale
-1. Paranoia Scale
-
-Studies 1 and 2:
-
-1. Self Determination Theory Scale (SDT; S1, S2)
-1. Interpersonal Support Evaluation List (ISEL; S1, S2)
-1. Perceived Stress Scale (PSS; S1, S2)
-1. Negative Life Events Scale (NLES; S1, S2)
-
-Studies 2 and 3:
-
-1. Social Identification Scale (S2, S3)
-1. Multidimensional Mood State Questionnaire (MDMQ; S2, S3)
-1. Homesickness Scale (S2, S3)
-
-All Studies:
-
-1. Intergroup Anxiety Scale (IAS)
-1. The Satisfaction with Life Scale (SWL)
-1. Rosenberg Self-Esteem Scale (SES)
-1. Everyday Discrimination Scale
-1. Language Proficiency
-1. Association Participation
-
-\color{black}
+For a full list of all variables and indices see `VariableSelectionAnalyses.xlsx'.
 
@@
 Thus, we then select the time points for which we have less than 45% missingness and select participants who have less than 45% missingness across the selected time range.
-Because the studies are comparable but not identical in their sampling strategy, question selection, and question wording, we also have to consider, whether to combine the samples.
 2.
-**Time point aggregation.** We then aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.
+**Time point aggregation.** We aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.
 Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
@@
 Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.
-We then chose the ideal time scale for each of the key variables.
+We then chose the time scales that have the most variance.
 3.
@@
 
-\textbf{*Step-wise Follow-up Analyses:}
-\begin{enumerate}
-  \item \textbf{Three studies (S1, S2, S3)}
-    \begin{enumerate}
-      \item Main Analysis: Combined for most power in correlations (including, correlation with study indicator)
-      \item Separate in follow up analysis to ensure that we do not have a Simpson's paradox
-      \item Explore additional idiosyncratic variables in separated follow up
-    \end{enumerate}
-  \item \textbf{Two question types (interaction, no interaction)}:
-    \begin{enumerate}
-      \item Main Analysis: Combined for most power in correlations (including, correlation with interaction indicator)
-      \item Separate in subgroup follow up analysis to ensure that we do not have a Simpson's paradox
-    \end{enumerate}
-  \item \textbf{Three time scales (half-day, day, week)}
-    \begin{enumerate}
-      \item Main Analysis: If all variables clearly show single most variance on the same time-scale, single analysis only
-      \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
-      \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
-    \end{enumerate}
+\textbf{*Step-wise Analysis Approach:}
+\begin{enumerate}[I.]
+  \item \textbf{Studies (Study 1, Study 2, Study 3)}
+    \begin{itemize}
+      \item \textbf{Main Analysis:} All studies combined to have the most power.
+      \item \textbf{Follow-up Analysis \#01:} To make sure there is no Simpson's paradox, we also separate the three data sets in separate follow-up analyses.
+      \item \textbf{Follow-up Analysis\#02:} We add idiosyncratic variables (i.e., variables only available in a subset of studies) within the separated follow-up analyses.
+With these additional variables we explore whether our main analysis is robust to other psychological experiences or whether the additional variables might provide a qualitatively different understanding of migration processes.
+    \end{itemize}
+  \item \textbf{Response types (interaction, no interaction)}\footnote{Some questions were only available or had a slightly different focus depending on whether participants had an interaction or not.}
+    \begin{itemize}
+      \item \textbf{Main Analysis:} All available questions combined to have the most power.
+      \item \textbf{Follow-up Analysis\#03:} To make sure there is no Simpson's paradox, we separate the data into two subgroup follow-up analyses (i.e., responses following an interaction and responses without interactions).
+This analysis will again test the robustness of the main analysis and will explore whether the core dimensions are substantially different for interaction specific phenomena.
+    \end{itemize}
+  \item \textbf{Time scales (half-day, day, week)}
+    \begin{itemize}
+      \item \textbf{Main Analysis:} For the main analysis, we will use the time scale that shows the most variance across the included items (e.g., daily; see \textit{Time point aggregation} above).
+      \item \textbf{Follow-up Analysis\#04:} To explore the impact of time scales in (social) psychological data, we will also run the 3MPCA on the other time scale options (e.g., bi-daily, weekly). 
+    \end{itemize}
 \end{enumerate}

### 2022-05-27 preregistration/Preregistration3MPCA.tex
feedback added

@@
   \postauthor{\end{changemargin}}
-\date{27. May 2022}
+\date{28. May 2022}
 \predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
@@
 \usepackage{nccmath}
+\usepackage{enumerate}
 \usepackage{LatexPackage/trackchanges}
@@
 \subsection{Description}\label{description}}
+
+\textbf{Shortened Description:}
+
+Recent reviews have called for more comprehensive assessment of human
+experiences and for more longitudinal real-life data, within the
+psychological sciences more broadly and in migration research in
+particular
+\citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
+However, while generally speaking analytical methods for such, more
+complex, data have become more readily available
+\citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify
+key developmental patterns --- especially across multiple variables at
+the same time.
+
+In this manuscript, we aim to assess the utility of a promising analysis
+technique for describing, summarizing, and understanding the initial
+developmental data patterns ---
+\textit{three-mode principle component analysis} (3MPCA). We use 3MPCA
+specifically because recent studies have laid out the potential
+effectiveness of dimension reduction procedures, which can address the
+new forms of person-, variable-, and time point heterogeneity jointly
+\citep[e.g.,][]{Monden2015}. We analyse data from three experience
+sampling studies, which followed the migration experiences of recent
+migrants to the Netherlands. All three studies focus on the
+psychological adaptation of migrants (i.e., psychological
+acculturation). However, given that past investigations of psychological
+acculturation have underexplored the crucial aspect of motivational
+experiences \citep{Kreienkamp2022d}, the three studies have placed a
+particular emphasis on the needs, goals, and motives of young migrants.
+To make full use of all three data sets and to guide future use of
+dimension reduction procedures with extensive psychological data, we
+will analyze the data in a descriptive and step wise manner, where we
+first consider the data jointly and will then assess potential
+differences between datasets, variables, and time scales.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\textbf{Full Description:}
 
@@
 we will not list all items individually. Instead, we provide full
-variable information in the following data sheets:
-`VariableSelectionDaily.xlsx' and `VariableSelectionPrePost.xlsx'.
+variable information in the following data sheets in
+`VariableSelectionAnalyses.xlsx'.
 
@@
 
-˜ 1. Mean Allport's conditions. We create a mean-averaged index of
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Mean Allport's conditions. We create a mean-averaged index of
   Allport's conditions in response to past findings indicating that the
   conditions are best conceptualized jointly and as functioning together
-rather than as fully independent factors \citep[p.~766]{Pettigrew2006}.
-Similar to past studies we thus hope to build a global indicator
-\citep[e.g., see][]{Pettigrew2006}. As with other indices we will ensure
-that the individual items indeed relate to a common latent construct and
-are meaningfully combined in an index. If this is not possible we will
-create sub-indices and/or assess the impact of the conditions
-separately. a. The interaction with {[}name interaction partner{]} was
-on equal footing (same status) a. {[}name interaction partner{]} shared
-your goal ({[}free-text entry interaction key need{]}) a. The
-interaction with {[}name interaction partner{]} was cooperative a. The
-interaction with {[}name interaction partner{]} was voluntary 1. Mean
-belongingness during intergroup contact a. I shared information about
-myself. a. {[}name interaction partner{]} shared information about
-themselves. 1. Mean alternative interaction quality definition
-(``Overall, the interaction with {[}name interaction partner{]} was
-\ldots{}'') a. Unpleasant to Pleasant a. Superficial to Meaningful a.
-Ineffective to Effective a. Unimportant to Important 1. Mood Subscales
-(MDMQ) a. alertness a. calmness a. valence 1. Mean anti-social behaviors
-a. Put someone down a. Show little attention in someones opinion a.
-Demeaning remarks a. Inappropriately addressing someone a. Ignored or
-excluded someone a. Doubt someones judgement a. Unwanted attempts of
-personal matters 1. Mean pro-social behaviors a. Listen to someones
-problems a. Cheer someone up a. Help someone get things done a. Help
-someone with responsibilities
+  rather than as fully independent factors
+  \citep[p.~766]{Pettigrew2006}. Similar to past studies we thus hope to
+  build a global indicator \citep[e.g., see][]{Pettigrew2006}. As with
+  other indices we will ensure that the individual items indeed relate
+  to a common latent construct and are meaningfully combined in an
+  index. If this is not possible we will create sub-indices and/or
+  assess the impact of the conditions separately.
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    The interaction with {[}name interaction partner{]} was on equal
+    footing (same status)
+  \item
+    {[}name interaction partner{]} shared your goal ({[}free-text entry
+    interaction key need{]})
+  \item
+    The interaction with {[}name interaction partner{]} was cooperative
+  \item
+    The interaction with {[}name interaction partner{]} was voluntary
+  \end{enumerate}
+\item
+  Mean belongingness during intergroup contact
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    I shared information about myself.
+  \item
+    {[}name interaction partner{]} shared information about themselves.
+  \end{enumerate}
+\item
+  Mean alternative interaction quality definition (``Overall, the
+  interaction with {[}name interaction partner{]} was \ldots{}'')
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    Unpleasant to Pleasant
+  \item
+    Superficial to Meaningful
+  \item
+    Ineffective to Effective
+  \item
+    Unimportant to Important
+  \end{enumerate}
+\item
+  Mood Subscales (MDMQ)
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    alertness
+  \item
+    calmness
+  \item
+    valence
+  \end{enumerate}
+\item
+  Mean anti-social behaviors
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    Put someone down
+  \item
+    Show little attention in someones opinion
+  \item
+    Demeaning remarks
+  \item
+    Inappropriately addressing someone
+  \item
+    Ignored or excluded someone
+  \item
+    Doubt someones judgement
+  \item
+    Unwanted attempts of personal matters
+  \end{enumerate}
+\item
+  Mean pro-social behaviors
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    Listen to someones problems
+  \item
+    Cheer someone up
+  \item
+    Help someone get things done
+  \item
+    Help someone with responsibilities
+  \end{enumerate}
+\end{enumerate}
 
@@
 `VariableSelectionAnalyses.xlsx'.
-
-\color{red}
-
-\textbf{TO BE REMOVED ONCE 'VariableSelectionAnalyses.xlsx' IS DONE!}
-
-\color{black}
-
-\fbox{\parbox{\mylength}{
-  \color{editPurple}
-  \textbf{Note.}\\
-  We have a group of six variables that are consistent across all three studies.
-Additionally, we have a number of variables that are either only relevant to a single study or a subset of the studies.
-If we agree on the study consolidation procedure as it is described below, I would suggest that we focus on the variables we have in all studies, and will include the other variables in the follow-up analyses and report only significant correlations (after alpha correction).
-I would also suggest that we add the list below to one of the variable selection files (Excel) instead of listing them all here (to avoid unnecessary confusion).
-  \color{black}
-}}
-
-\color{editPurple}
-
-Study 1 only:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Acculturation, Habits, and Interests Multicultural Scale for
-  Adolescents (AHIMSA; S1)
-\item
-  Vancouver Index of Acculturation (VIA; S1)
-\item
-  Situational Self Awareness Scale (SAS; S1)
-\end{enumerate}
-
-Study 3 only:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Outgroup Trust
-\item
-  Work Group Inclusion
-\item
-  Meaningful Life Scale
-\item
-  Big Five Personality Questionnaire
-\item
-  Attachment Styles Scale
-\item
-  Paranoia Scale
-\end{enumerate}
-
-Studies 1 and 2:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Self Determination Theory Scale (SDT; S1, S2)
-\item
-  Interpersonal Support Evaluation List (ISEL; S1, S2)
-\item
-  Perceived Stress Scale (PSS; S1, S2)
-\item
-  Negative Life Events Scale (NLES; S1, S2)
-\end{enumerate}
-
-Studies 2 and 3:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Social Identification Scale (S2, S3)
-\item
-  Multidimensional Mood State Questionnaire (MDMQ; S2, S3)
-\item
-  Homesickness Scale (S2, S3)
-\end{enumerate}
-
-All Studies:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Intergroup Anxiety Scale (IAS)
-\item
-  The Satisfaction with Life Scale (SWL)
-\item
-  Rosenberg Self-Esteem Scale (SES)
-\item
-  Everyday Discrimination Scale
-\item
-  Language Proficiency
-\item
-  Association Participation
-\end{enumerate}
-
-\color{black}
 
@@
   than 45\% missingness and select participants who have less than 45\%
-  missingness across the selected time range. Because the studies are
-  comparable but not identical in their sampling strategy, question
-  selection, and question wording, we also have to consider, whether to
-  combine the samples.
+  missingness across the selected time range.
 \item
-  \textbf{Time point aggregation.} We then aggregate the key variables
-  over time to archive a reasonably interpretable number of time points
-  and remove a first proportion of missing data. Given that little data
-  is available on the meaningful time scales of the selected
-  psychological variables, we chose to determine the appropriate time
-  scales using variance decomposition \citep[e.g., see][]{Ram2014}. This
-  is to say that we create multi-level unconditional means models
-  (without predictors) that include possible nested time scales as
-  levels. We chose to select time scales that align with common human
-  cycles. We thus compare the variances of bi-daily, daily, and weekly
+  \textbf{Time point aggregation.} We aggregate the key variables over
+  time to archive a reasonably interpretable number of time points and
+  remove a first proportion of missing data. Given that little data is
+  available on the meaningful time scales of the selected psychological
+  variables, we chose to determine the appropriate time scales using
+  variance decomposition \citep[e.g., see][]{Ram2014}. This is to say
+  that we create multi-level unconditional means models (without
+  predictors) that include possible nested time scales as levels. We
+  chose to select time scales that align with common human cycles. We
+  thus compare the variances of bi-daily, daily, and weekly
   aggregations. Additional aggregations of two weeks or the full four
@@
   much for any meaningul further reduction during the 3MPCA. We then
-  chose the ideal time scale for each of the key variables.
+  chose the time scales that have the most variance.
 \item
@@
 
-\textbf{*Step-wise Follow-up Analyses:}
+\textbf{*Step-wise Analysis Approach:}
 
-\begin{enumerate}
-  \item \textbf{Three studies (S1, S2, S3)}
-    \begin{enumerate}
-      \item Main Analysis: Combined for most power in correlations (including, correlation with study indicator)
-      \item Separate in follow up analysis to ensure that we do not have a Simpson's paradox
-      \item Explore additional idiosyncratic variables in separated follow up
-    \end{enumerate}
-  \item \textbf{Two question types (interaction, no interaction)}:
-    \begin{enumerate}
-      \item Main Analysis: Combined for most power in correlations (including, correlation with interaction indicator)
-      \item Separate in subgroup follow up analysis to ensure that we do not have a Simpson's paradox
-    \end{enumerate}
-  \item \textbf{Three time scales (half-day, day, week)}
-    \begin{enumerate}
-      \item Main Analysis: If all variables clearly show single most variance on the same time-scale, single analysis only
-      \item If multiple time-scales are possible within single analysis, single analysis only (see question box below)
-      \item If variance split across multiple time scales for substantial number of variables (> 1/4 of all variables), multiple analyses for the different time scales 
-    \end{enumerate}
+\begin{enumerate}[I.]
+  \item \textbf{Studies (Study 1, Study 2, Study 3)}
+    \begin{itemize}
+      \item \textbf{Main Analysis:} All studies combined to have the most power.
+      \item \textbf{Follow-up Analysis \#01:} To make sure there is no Simpson's paradox, we also separate the three data sets in separate follow-up analyses.
+      \item \textbf{Follow-up Analysis\#02:} We add idiosyncratic variables (i.e., variables only available in a subset of studies) within the separated follow-up analyses.
+With these additional variables we explore whether our main analysis is robust to other psychological experiences or whether the additional variables might provide a qualitatively different understanding of migration processes.
+    \end{itemize}
+  \item \textbf{Response types (interaction, no interaction)}\footnote{Some questions were only available or had a slightly different focus depending on whether participants had an interaction or not.}
+    \begin{itemize}
+      \item \textbf{Main Analysis:} All available questions combined to have the most power.
+      \item \textbf{Follow-up Analysis\#03:} To make sure there is no Simpson's paradox, we separate the data into two subgroup follow-up analyses (i.e., responses following an interaction and responses without interactions).
+This analysis will again test the robustness of the main analysis and will explore whether the core dimensions are substantially different for interaction specific phenomena.
+    \end{itemize}
+  \item \textbf{Time scales (half-day, day, week)}
+    \begin{itemize}
+      \item \textbf{Main Analysis:} For the main analysis, we will use the time scale that shows the most variance across the included items (e.g., daily; see \textit{Time point aggregation} above).
+      \item \textbf{Follow-up Analysis\#04:} To explore the impact of time scales in (social) psychological data, we will also run the 3MPCA on the other time scale options (e.g., bi-daily, weekly). 
+    \end{itemize}
 \end{enumerate}

### 2022-06-16 preregistration/Preregistration3MPCA.Rmd
Prereg final

@@
 ---
-title           : "Migration Experience Trajectories: A Three Mode Principle Component Analysis"
+title           : "Migration Experience Trajectories: A Three-Mode Principal Component Analysis"
 shorttitle      : "Preregistration: Migration Experience Trajectories"
@@
 
-In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principle component analysis} (3MPCA).
+In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principal component analysis} (3MPCA).
 We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015].
@@
 
-In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principle component analysis} (3MPCA).
+In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principal component analysis} (3MPCA).
 We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015].
@@
 
-We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands.
+We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands (median time in the Netherlands = 3 Months).
 For all three studies, the data was collected in a three-step procedure:
@@
 **Auxiliary variables for multiple imputation**  
-For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully correlated ($r$ > .3) with either the key variables or missingness on the key variables.
+For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly correlated with either the key variables or missingness on the key variables (at $p$ < .01).
+In case we encounter a convergence problem during the multiple imputation, we only include variable with a correlation larger than .3.
 For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see `VariableSelectionAnalyses.xlsx').
@@
 Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions.
-We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the interdependence of the three modes.
+We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.
 5.
@@
 
-Many of the proposed procedures are descriptive rather then inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
+Many of the proposed procedures are descriptive rather than inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
 If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant.
-We will use appropriate multiple-comparison corrections whenever multiple tests are performed.
+We will use appropriate multiple-comparison corrections whenever multiple tests are performed (such as Bonferroni correction).
 We will place particular emphasis on effect sizes (e.g., correlations) in our interpretations of the results wherever possible.
@@
 No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the variables of interest for a given analysis.
-Outliers will generally be included in analyses, however we will use sensitivity analyses to assess the robustness of the results to outliers.
+Outliers will generally be included in analyses.
+However, we will use sensitivity analyses to assess the robustness of the results to outliers.
 

### 2022-06-16 preregistration/Preregistration3MPCA.tex
Prereg final

@@
 pdfauthor={},
-pdftitle={Migration Experience Trajectories: A Three Mode Principle Component Analysis},
+pdftitle={Migration Experience Trajectories: A Three-Mode Principal Component Analysis},
 colorlinks=true,
@@
 \setlength{\droptitle}{-2em}
-\title{Migration Experience Trajectories: A Three Mode Principle
+\title{Migration Experience Trajectories: A Three-Mode Principal
 Component Analysis}
@@
   \postauthor{\end{changemargin}}
-\date{28. May 2022}
+\date{01. June 2022}
 \predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
@@
 
-Migration Experience Trajectories: A Three Mode Principle Component
+Migration Experience Trajectories: A Three-Mode Principal Component
 Analysis
@@
 developmental data patterns ---
-\textit{three-mode principle component analysis} (3MPCA). We use 3MPCA
+\textit{three-mode principal component analysis} (3MPCA). We use 3MPCA
 specifically because recent studies have laid out the potential
@@
 In this manuscript, we aim to assess the utility of one such promising
-analysis technique --- \textit{three-mode principle component analysis}
+analysis technique --- \textit{three-mode principal component analysis}
 (3MPCA). We use 3MPCA specifically because recent studies have laid out
@@
 migration experiences of young migrants, who had recently arrived in the
-Netherlands. For all three studies, the data was collected in a
-three-step procedure:
+Netherlands (median time in the Netherlands = 3 Months). For all three
+studies, the data was collected in a three-step procedure:
 
@@
 \citet{Monden2015}, and include any pre-, post-, or daily dairy
-variables that are significantly (\(p\) \textless{} .01) and
-meaningfully correlated (\(r\) \textgreater{} .3) with either the key
-variables or missingness on the key variables. For an overview of the
-variables that qualify as auxiliary variables see the enclosed data
-sheets (see `VariableSelectionAnalyses.xlsx'). Additional information
-about the multiple imputation procedure is also available at
-\protect\hyperlink{statistical-models}{Statistical models} 2.
+variables that are significantly correlated with either the key
+variables or missingness on the key variables (at \(p\) \textless{}
+.01). In case we encounter a convergence problem during the multiple
+imputation, we only include variable with a correlation larger than .3.
+For an overview of the variables that qualify as auxiliary variables see
+the enclosed data sheets (see `VariableSelectionAnalyses.xlsx').
+Additional information about the multiple imputation procedure is also
+available at \protect\hyperlink{statistical-models}{Statistical models}
+2.
 
@@
   order interaction term Person * Variable * Time + error, as a large
-  amount of variance in this effect would speak towards the
+  amount of variance in this effect would speak towards the possible
   interdependence of the three modes.
@@
 
-Many of the proposed procedures are descriptive rather then inferential.
+Many of the proposed procedures are descriptive rather than inferential.
 Importantly, we will use amounts of variances explained as our main fit
@@
 appropriate multiple-comparison corrections whenever multiple tests are
-performed. We will place particular emphasis on effect sizes (e.g.,
-correlations) in our interpretations of the results wherever possible.
+performed (such as Bonferroni correction). We will place particular
+emphasis on effect sizes (e.g., correlations) in our interpretations of
+the results wherever possible.
 
@@
 interest for a given analysis. Outliers will generally be included in
-analyses, however we will use sensitivity analyses to assess the
+analyses. However, we will use sensitivity analyses to assess the
 robustness of the results to outliers.

### 2022-06-16 Supplemental-Material-A-Annotated-Analysis.Rmd
sample selection test

@@
   "metafor",
-  "dygraphs"
+  "dygraphs",
+  "readxl", 
+  "reshape"
 )
@@
 ```{r formrImport}
+# Load variable name lookup table
+varNames <- readxl::read_excel("preregistration/varNames.xlsx")
+
+# rename function
+reNam <- function(data = NA, names = varNames, study = "S1", survey = "pre") {
+  # for testing
+  # data = dtS1$raw.pre
+  # names = varNames
+  # study = "S1"
+  # survey = "pre"
+
+  var <- paste0("survey", study)
+  varNamOld <- paste0("varNam", study)
+
+  nameTbl <- names %>%
+    filter((!!sym(var)) == survey) %>%
+    select(varNam, varNamOld = (!!sym(varNamOld)))
+
+  data.table::setnames(data, as.character(nameTbl$varNamOld), as.character(nameTbl$varNam), skip_absent=TRUE)
+
+  return(data)
+}
+
 # workers
 # initial data cleaning was done in SPSS (syntax files are available in "")
-dtWorker <- list(
-  raw.pre = read_spss("data/S1_Workers/processed/cleaned/MT - Pre-Measure - 06-15-2018.sav"),
-  raw.post = read_spss("data/S1_Workers/processed/cleaned/MT - Post-Measure - 06-15-2018.sav"),
-  raw.morning = read_spss("data/S1_Workers/processed/cleaned/MT - Morning - 06-15-2018.sav"),
-  raw.afternoon = read_spss("data/S1_Workers/processed/cleaned/MT - Afternoon - 06-15-2018.sav")
+dtS1 <- list(
+  raw.pre = read_spss("data/S1_Workers/processed/cleaned/MT - Pre-Measure - 06-15-2018.sav") %>% 
+    reNam(data = ., names = varNames, study = "S1", survey = "pre"),
+  raw.post = read_spss("data/S1_Workers/processed/cleaned/MT - Post-Measure - 06-15-2018.sav") %>% 
+    reNam(data = ., names = varNames, study = "S1", survey = "post"),
+  raw.morning = read_spss("data/S1_Workers/processed/cleaned/MT - Morning - 06-15-2018.sav") %>% 
+    reNam(data = ., names = varNames, study = "S1", survey = "daily"),
+  raw.afternoon = read_spss("data/S1_Workers/processed/cleaned/MT - Afternoon - 06-15-2018.sav") %>% 
+    reNam(data = ., names = varNames, study = "S1", survey = "daily")
 )
@@
 # students
-dtStudents <- list(
-  raw.pre = read.csv(file = "data/S2_Students/raw/AOTS_Pre.csv", header = T, sep = ","),
-  raw.post = read.csv(file = "data/S2_Students/raw/AOTS_Post.csv", header = T, sep = ","),
-  raw.daily = read.csv(file = "data/S2_Students/raw/AOTS_Daily.csv", header = T, sep = ",")
+dtS2 <- list(
+  raw.pre = read.csv(file = "data/S2_Students/raw/AOTS_Pre.csv", header = T, sep = ",") %>% 
+    reNam(data = ., names = varNames, study = "S2", survey = "pre"),
+  raw.post = read.csv(file = "data/S2_Students/raw/AOTS_Post.csv", header = T, sep = ",") %>% 
+    reNam(data = ., names = varNames, study = "S2", survey = "post"),
+  raw.daily = read.csv(file = "data/S2_Students/raw/AOTS_Daily.csv", header = T, sep = ",") %>% 
+    reNam(data = ., names = varNames, study = "S2", survey = "daily")
 )
@@
 # young medical professionals
-dtMedical <- list(
-  raw.eligibility = read.csv("data/S3_Medical/raw/AOTM_Eligibility.csv"),
-  raw.pre = read.csv("data/S3_Medical/raw/AOTM_Pre.csv"),
-  raw.post = read.csv("data/S3_Medical/raw/AOTM_Post.csv"),
-  raw.daily = read.csv("data/S3_Medical/raw/AOTM_Daily.csv")
-)
+dtS3 <- list(
+  raw.eligibility = read.csv("data/S3_Medical/raw/AOTM_Eligibility.csv") %>% 
+    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"), # works but should be 'survey = "pre_elig"'
+  raw.pre = read.csv("data/S3_Medical/raw/AOTM_Pre.csv") %>% 
+    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"),
+  raw.post = read.csv("data/S3_Medical/raw/AOTM_Post.csv") %>% 
+    reNam(data = ., names = varNames, study = "S3", survey = "post"),
+  raw.daily = read.csv("data/S3_Medical/raw/AOTM_Daily.csv") %>% 
+    reNam(data = ., names = varNames, study = "S3", survey = "daily")
+)
+
+# # FROM INITIAL VAR NAME EXPORT [LEGACY]
+# # export var names
+# S1Nam <- rbind(
+#   data.frame(
+#     study = "S1",
+#     survey = "pre",
+#     varNam = names(dtS1$raw.pre)
+#   ),
+#   data.frame(
+#     study = "S1",
+#     survey = "daily",
+#     varNam = names(dtS1$raw.afternoon)
+#   ),
+#   data.frame(
+#     study = "S1",
+#     survey = "post",
+#     varNam = names(dtS1$raw.post)
+#   )
+# )
+# S2Nam <- rbind(
+#   data.frame(
+#     study = "S2",
+#     survey = "pre",
+#     varNam = names(dtS2$raw.pre)
+#   ),
+#   data.frame(
+#     study = "S2",
+#     survey = "daily",
+#     varNam = names(dtS2$raw.daily)
+#   ),
+#   data.frame(
+#     study = "S2",
+#     survey = "post",
+#     varNam = names(dtS2$raw.post)
+#   )
+# )
+# S3Nam <- rbind(
+#   data.frame(
+#     study = "S3",
+#     survey = "pre_elig",
+#     varNam = names(dtS3$raw.eligibility)
+#   ),
+#   data.frame(
+#     study = "S3",
+#     survey = "pre_entry",
+#     varNam = names(dtS3$raw.pre)
+#   ),
+#   data.frame(
+#     study = "S3",
+#     survey = "daily",
+#     varNam = names(dtS3$raw.daily)
+#   ),
+#   data.frame(
+#     study = "S3",
+#     survey = "post",
+#     varNam = names(dtS3$raw.post)
+#   )
+# )
+# write.csv(
+#   x = rbind(
+#     S1Nam,
+#     S2Nam,
+#     S3Nam
+#   ),
+#   file = "preregistration/varNames.csv"
+# )
 ```
@@
 We also exclude our own test data.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtWorker_' list.
+<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS1_' list.
 
 ```{r cleanWorker}
-#  important names for Morning and Afternoon
-names.m <- c(
-  "StartDate",
-  "EndDate",
-  "Finished",
-  "Duration__in_seconds_",
-  "RecordedDate",
-  "ExternalReference",
-  "Meta_Operating_System",
-  "Contact_dum",
-  "number",
-  "time",
-  "duration_1",
-  "dyad.group",
-  "gr_size",
-  "gr_type_1",
-  "gr_type_2",
-  "gr_type_3",
-  "gr_type_4",
-  "gr_type_5",
-  "gr_type_6",
-  "gr_type_7",
-  "gr_type_8",
-  "gr_type_9",
-  "gr_type_10",
-  "gr_type_11",
-  "gr_type_12",
-  "gr_type_13",
-  "gr_type_14",
-  "gr_type_15",
-  "gr_type_16",
-  "gr_type_17_TEXT",
-  "gr_context_1",
-  "gr_context_2",
-  "gr_context_3",
-  "gr_context_4",
-  "gr_context_5",
-  "gr_context_6",
-  "gr_context_7",
-  "gr_context_8",
-  "gr_context_9",
-  "gr_context_10",
-  "gr_context_11",
-  "gr_context_12",
-  "gr_context_13_TEXT",
-  "gr_context_14_TEXT",
-  "gr_dutchness",
-  "dyad_type_1",
-  "dyad_type_2",
-  "dyad_type_3",
-  "dyad_type_4",
-  "dyad_type_5",
-  "dyad_type_6",
-  "dyad_type_7",
-  "dyad_type_8",
-  "dyad_type_9",
-  "dyad_type_10",
-  "dyad_type_11",
-  "dyad_type_12",
-  "dyad_type_13",
-  "dyad_type_14",
-  "dyad_type_15",
-  "dyad_type_16",
-  "dyad_type_17_TEXT",
-  "Context_1",
-  "Context_2",
-  "Context_3",
-  "Context_4",
-  "Context_5",
-  "Context_6",
-  "Context_7",
-  "Context_8",
-  "Context_9",
-  "Context_10",
-  "Context_11",
-  "Context_12",
-  "Context_13_TEXT",
-  "Context_14_TEXT",
-  "keyMotive",
-  "keymotive_fulfillemt_1",
-  "keyMotive_Dutch_1",
-  "autonomy_1",
-  "competence_1",
-  "relatedness_self_1",
-  "relatedness_other_1",
-  "qualityAccidental_1",
-  "qualityVoluntary_1",
-  "qualityCooperative_1",
-  "qualityDutchy_1",
-  "quality_overall_1",
-  "quality_meaning_1",
-  "quality_star_1",
-  "wantInt",
-  "desire_type_1",
-  "desire_type_2",
-  "desire_type_3",
-  "desire_type_4",
-  "desire_type_5",
-  "desire_type_6",
-  "desire_type_7",
-  "desire_type_8",
-  "desire_type_9",
-  "desire_type_10",
-  "desire_type_11",
-  "desire_type_12",
-  "desire_type_13",
-  "desire_type_14",
-  "desire_type_15",
-  "desire_type_16",
-  "desire_type_17_TEXT",
-  "desire_context_1",
-  "desire_context_2",
-  "desire_context_3",
-  "desire_context_4",
-  "desire_context_5",
-  "desire_context_6",
-  "desire_context_7",
-  "desire_context_8",
-  "desire_context_9",
-  "desire_context_10",
-  "desire_context_11",
-  "desire_context_12",
-  "desire_context_13_TEXT",
-  "desire_context_14_TEXT",
-  "Reason_nodesire",
-  "keyMotive_noInt",
-  "keyMotive_noInt_fulf_1",
-  "autonomy_NoInt_1",
-  "competence_NoInt_1",
-  "relatedness_1_NoInt_1",
-  "thermometerDutch_1",
-  "thermometerDutchInt_2",
-  "ExWB_1",
-  "MDMQ.v1_1",
-  "MDMQ.v1_2",
-  "MDMQ.v1_3",
-  #"MDMQ.v1_3r",
-  "MDMQ.v1_4",
-  #"MDMQ.v1_4r",
-  "MDMQ.v1_5",
-  #"MDMQ.v1_5r",
-  "MDMQ.v1_6",
-  "MDMQ.v1_7",
-  #"MDMQ.v1_7r",
-  "MDMQ.v1_8",
-  "MDMQ.v1_9",
-  #"MDMQ.v1_9r",
-  "MDMQ.v1_10",
-  "MDMQ.v1_11",
-  #"MDMQ.v1_11r",
-  "MDMQ.v1_12",
-  "MDMQ.v1_13",
-  "MDMQ.v2_1",
-  #"MDMQ.v2_1r",
-  "MDMQ.v2_2",
-  "MDMQ.v2_3",
-  "MDMQ.v2_4",
-  #"MDMQ.v2_4r",
-  "MDMQ.v2_5",
-  "MDMQ.v2_6",
-  #"MDMQ.v2_6r",
-  "MDMQ.v2_7",
-  #"MDMQ.v2_7r",
-  "MDMQ.v2_8",
-  "MDMQ.v2_9",
-  "MDMQ.v2_10",
-  #"MDMQ.v2_10r",
-  "MDMQ.v2_11",
-  #"MDMQ.v2_11r",
-  "MDMQ.v2_12",
-  "MDMQ.v2_13",
-  "alertness1",
-  "calmness1",
-  "valence1",
-  "alertness2",
-  "calmness2",
-  "valence2",
-  "inNonDutch",
-  "NonDutchNum",
-  "NonDutchType_1",
-  "NonDutchType_2",
-  "NonDutchType_3",
-  "NonDutchType_4",
-  "NonDutchType_5",
-  "NonDutchType_6",
-  "NonDutchType_7",
-  "NonDutchType_8",
-  "NonDutchType_9",
-  "NonDutchType_10",
-  "NonDutchType_11",
-  "NonDutchType_12",
-  "NonDutchType_13",
-  "NonDutchType_14",
-  "NonDutchType_15_TEXT",
-  "date",
-  "time.0",
-  "LocationLatitude",
-  "LocationLongitude"
-)
-
-names.a <- c(names.m, "keyInteraction_1", "keyInteractionTime")
-
 # Create reduced data sets for morning and afternoon
-dat.mo <- dtWorker$raw.morning[, names.m]
+dat.mo <- dtS1$raw.morning %>%
+  select(-starts_with("t_"))
+#setdiff(names(dtS1$raw.morning), names(dat.mo))
 dat.mo$daytime <- "morning"
 
-dat.af <- dtWorker$raw.afternoon[, names.a]
+dat.af <- dtS1$raw.afternoon %>%
+  select(-starts_with("t_"))
 dat.af$daytime <- "afternoon"
@@
 # merge morning and afternoon measurements with indicator [+ clean up]
-daily.dat <- rbind.fill(dat.mo, dat.af)
-daily.dat <- daily.dat[daily.dat$ExternalReference != 55951, ]
-dtWorker$daily <- daily.dat
-rm(dat.mo, dat.af, names.m, names.a, daily.dat)
-
-# names for pre-measurement
-names.pre <- c(
-  "Finished",
-  "age",
-  "Gender",
-  "Living",
-  "roommate_1",
-  "roommate_2",
-  "roommate_3",
-  "nationality",
-  "SecondNationality",
-  "timeNL_1",
-  "Reason_2",
-  "Reason_5",
-  "Reason_7",
-  "Reason_8_TEXT",
-  "DutchLang",
-  "occupation_1",
-  "occupation_2",
-  "occupation_3",
-  "occupation_4",
-  "occupation_7",
-  "CurrentEducation_1",
-  "education_level",
-  "EduLang_2",
-  "RUG_faculty",
-  "Study.0",
-  "association",
-  "DutchMeetNum",
-  "DutchFriends_1",
-  "assimilation",
-  "separation",
-  "integration",
-  "marginalization",
-  "VIA_heritage",
-  "VIA_Dutch",
-  "SSAS_surrounding",
-  "SSAS_privat",
-  "SSAS_public",
-  "autonomy",
-  "relatedness",
-  "competence",
-  "anxiety",
-  "swl",
-  "alertness",
-  "calmness",
-  "valence",
-  "date",
-  "time",
-  "City",
-  "ZIP",
-  "id"
-)
+daily.dat <- plyr::rbind.fill(dat.mo, dat.af)
+daily.dat <- daily.dat[daily.dat$last_outside_referrer != 55951, ]
+dtS1$daily <- daily.dat
+rm(dat.mo, dat.af, daily.dat)
 
 # reduced data set for pre measurement
-dat.pre.red <- dtWorker$raw.pre[, names.pre]
+dat.pre.red <- dtS1$raw.pre %>%
+  select(-starts_with("t_"))
 
@@
 df.pre <- merge(
-  x = dtWorker$daily,
+  x = dtS1$daily,
   y = dat.pre.red,
-  by.x = "ExternalReference",
-  by.y = "id",
+  by.x = "last_outside_referrer",
+  by.y = "platformId",
   all = T
 )
-rm(names.pre)
 
@@
 
-# names for post measurement
-names.post <- c(
-  "ExternalReference",
-  "assimilation",
-  "separation",
-  "integration",
-  "marginalization",
-  "VIA_heritage",
-  "VIA_Dutch",
-  "anxiety",
-  "swl",
-  "rosenberg",
-  "social_support",
-  "stress",
-  "discrimination",
-  "discrimination_month",
-  "NLE_1month",
-  "NLE_6month",
-  "NLE_12month"
-)
-
 # reduced data set for post-measurement
-dat.post.red <- dtWorker$raw.post[, names.post]
+dat.post.red <- dtS1$raw.post %>%
+  select(-starts_with("t_"))
 
@@
   y = dat.post.red,
-  by.x = "ExternalReference",
-  by.y = "ExternalReference",
+  by.x = "last_outside_referrer",
+  by.y = "last_outside_referrer",
   all = T
@@
 # add to list
-dtWorker$combined <- df
+dtS1$combined <- df
+rm(df, df.pre, dat.post.red, dat.pre.red)
 
 # create data frame with cleaned data
-df <- dtWorker$combined %>%
+df <- dtS1$combined %>%
   filter(
@@
     Finished.daily == 1,
-    !is.na(ExternalReference)
+    !is.na(last_outside_referrer)
   )
@@
 # add running number as measurement ID within participants
-df$measureID <- rowidv(df, cols = c("ExternalReference"))
+df$measureID <- data.table::rowidv(df, cols = c("last_outside_referrer"))
 
@@
   mutate(
-    PID = as.numeric(factor(ExternalReference)),
+    PID = as.numeric(factor(last_outside_referrer)),
     # participant ID
@@
     # time ID with t0 = 0 for meaningfull intercept interpretations
-    date = substr(StartDate, 1, 10),
+    date = substr(created, 1, 10),
     # awkward way of extracting date (best converted to )
-    time = substr(StartDate, 12, 19),
+    time = substr(created, 12, 19),
     # awkward way of extracting time
-    daynum = as.numeric(factor(date)),
+    daynum = as.numeric(factor(dateQualtrics)),
     # all days as numeric for ordering
@@
     # combine day id with morning / afternoon
-    session = as.numeric(factor(day_time)),
+    ResponseId = as.numeric(factor(day_time)),
     # day and time identifier as numeric id
@@
     time.daily = as.character(time.daily),
-    PPDate = as.Date(df$date.daily),
-    number = replace_na(number, 0),
+    PPDate = as.Date(df$dateQualtrics.daily),
+    number = replace_na(ContactNum, 0),
     NonDutchNum = replace_na(NonDutchNum, 0)
@@
 
-dtWorker$clean <- df
+dtS1$clean <- df
 
 # clean up
-rm(df.pre, names.post, dat.post.red, dat.pre.red, df)
+rm(df)
 
 # Export reduced Data
-# write.csv(dtWorker$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
-# save(dtWorker$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
+# write.csv(dtS1$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
+# save(dtS1$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
 ```
@@
 We also reformat missing values and format core ID variables.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtStudents_' list.
+<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS2_' list.
 
@@
 # Prepare dfs for Cleaning
-df.pre <- dtStudents$raw.pre %>%
+df.pre <- dtS2$raw.pre %>%
   mutate_all(na_if, "") %>%
@@
   filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!e_mail %in% .$e_mail[duplicated(.$e_mail)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
-  filter(!session %in% ownIDs) %>% # remove our own test
-  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)
-
-df.post <- dtStudents$raw.post %>%
+  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+
+df.post <- dtS2$raw.post %>%
   mutate_all(na_if, "") %>%
   mutate_all(na_if, "NA") %>%
-  filter(!is.na(session)) %>% # remove own test runs
-  filter(!session %in% ownIDs) %>% # remove our own test
-  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
+  filter(!is.na(ResponseId)) %>% # remove own test runs
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
   filter(!is.na(ended)) %>% # remove all who never finished
-  filter(!session %in% .$session[duplicated(.$session)]) %>% # remove all duplicate sessions
-  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)
-
-df.daily <- dtStudents$raw.daily %>%
+  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+
+df.daily <- dtS2$raw.daily %>%
   mutate_all(na_if, "") %>%
   mutate_all(na_if, "NA") %>%
-  filter(!session %in% ownIDs) %>% # remove our own test
-  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
   filter(!is.na(ended)) %>% # remove all who never finished
-  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
 
@@
   y = df.pre,
-  by = "session",
+  by = "ResponseId",
   suffixes = c(".daily", ".pre"),
@@
   y = df.post,
-  by = "session",
+  by = "ResponseId",
   suffixes = c(".pre", ".post"),
@@
 # add to list
-dtStudents$clean <- dfCombined
+dtS2$clean <- dfCombined
 
@@
 We also reformat missing values and format core ID variables.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtMedical_' list.
+<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS3_' list.
 
@@
 # Prepare dfs for Cleaning
-df.pre <- dtMedical$raw.pre %>%
+df.pre <- dtS3$raw.pre %>%
   mutate_all(na_if, "") %>%
@@
   filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!session %in% ownIDs) %>% # remove our own test
-  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)
-
-df.post <- dtMedical$raw.post %>%
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+
+df.post <- dtS3$raw.post %>%
   mutate_all(na_if, "") %>%
   mutate_all(na_if, "NA") %>% 
-  filter(!is.na(session)) %>% # remove own test runs
-  filter(!session %in% ownIDs) %>% # remove our own test
-  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
+  filter(!is.na(ResponseId)) %>% # remove own test runs
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
   #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  filter(!session %in% .$session[duplicated(.$session)]) %>% # remove all duplicate sessions
-  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)
-
-df.daily <- dtMedical$raw.daily %>%
+  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+
+df.daily <- dtS3$raw.daily %>%
   mutate_all(na_if, "") %>%
   mutate_all(na_if, "NA") %>%
-  filter(!session %in% ownIDs) %>% # remove our own test
-  filter(session %in% df.pre$session) %>% # remove anyone who wasn't in the pre
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
   #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  mutate(session = as.character(session)) # turn factor into character strings (probably just precaution)
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
 
@@
 # add to list
-dtMedical$clean <- dfCombined
+dtS3$clean <- dfCombined
 
@@
 ```{r newVarsWorkers}
-df <- dtWorker$clean
+df <- dtS1$clean
 
@@
 
-# Make contact dummies consistent with later studies
-df$IntergroupContact <- df$Contact_dum
-
 df <- df %>%
@@
   mutate(
-    InteractionDum = sum(Contact_dum, inNonDutch, na.rm = FALSE),
+    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = FALSE),
     InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum),
@@
 # Need scales
-df$InteractionNeedFullfillment <- df$keymotive_fulfillemt_1
-df$DaytimeNeedFullfillment <- df$keyMotive_noInt_fulf_1
-
-df$keyMotiveFulfilled <- rowSums(df[, c("keymotive_fulfillemt_1", "keyMotive_noInt_fulf_1")], na.rm = T)
-df$autonomy.daily.all <- rowSums(df[, c("autonomy_1", "autonomy_NoInt_1")], na.rm = T)
-df$competence.daily.all <- rowSums(df[, c("competence_1", "competence_NoInt_1")], na.rm = T)
-# cor(df$relatedness_other_1, df$relatedness_self_1,use="complete.obs")
+df$keyMotiveFulfilled <- rowSums(df[, c("KeyNeedFulfillment", "DaytimeNeedFulfillment")], na.rm = T)
+df$autonomy.daily.all <- rowSums(df[, c("autonomy", "autonomy_NoInt_1")], na.rm = T)
+df$competence.daily.all <- rowSums(df[, c("competence", "competence_NoInt_1")], na.rm = T)
+# cor(df$relatednessOther, df$relatedness_self_1,use="complete.obs")
 df$relatedness.daily.all <- rowMeans(df[, c(
-  "relatedness_other_1",
-  "relatedness_self_1",
-  "relatedness_1_NoInt_1"
+  "relatednessOther",
+  "relatednessSelf",
+  "relatednessNoInteraction"
 )], na.rm = T)
@@
 pairs.panels.new(
-  df[c("relatedness_self_1", "relatedness_other_1")],
+  df[c("relatednessSelf", "relatednessOther")],
   labels = c(
@@
 )
-df$relatedness_1 <- rowMeans(df[, c("relatedness_other_1", "relatedness_self_1")], na.rm = T)
+df$relatedness <- rowMeans(df[, c("relatednessOther", "relatednessSelf")], na.rm = T)
 
@@
 
-# Make Outgroup Attitudes consistent with later studies
-df$AttitudesDutch <- df$thermometerDutch_1
-df$AttitudesPartner <- df$thermometerDutchInt_2
-
-# Make experienced well-being consistent with later studies
-df$ExWB <- df$ExWB_1
-
 # summarize by participant (check that everything is within pp might not be the case for )
 between <- df %>%
-  group_by(ExternalReference) %>%
+  group_by(last_outside_referrer) %>%
   mutate(
-    CtContactNL = sum(Contact_dum),
-    CtContactNonNl = sum(inNonDutch),
-    CtContactNLAll = sum(number),
+    CtContactNL = sum(IntergroupContact),
+    CtContactNonNl = sum(IngroupContact),
+    CtContactNLAll = sum(ContactNum),
     CtContactNonNlAll = sum(NonDutchNum),
     AvKeyNeed = mean(keyMotiveFulfilled, na.rm = T),
-    AvKeyNeedInt = mean(keymotive_fulfillemt_1, na.rm = T),
-    AvKeyNeedNoInt = mean(keyMotive_noInt_fulf_1, na.rm = T),
+    AvKeyNeedInt = mean(KeyNeedFulfillment, na.rm = T),
+    AvKeyNeedNoInt = mean(DaytimeNeedFulfillment, na.rm = T),
     AvAutonomy = mean(autonomy.daily.all, na.rm = T),
@@
     AvRelatedness = mean(relatedness.daily.all, na.rm = T),
-    AvThermo = mean(thermometerDutch_1, na.rm = T),
-    AvWB = mean(ExWB_1, na.rm = T)
+    AvThermo = mean(AttitudesDutch, na.rm = T),
+    AvWB = mean(exWB, na.rm = T)
   ) %>%
@@
 
-dtWorker$full <- between
+dtS1$full <- between
 rm(df, between)
@@
 # dataframe where interaction types are recoded
-workerInteractionType <- dtWorker$full %>%
+workerInteractionType <- dtS1$full %>%
   mutate(
-    OutgroupInteraction = as_factor(Contact_dum),
-    NonOutgroupInteraction = as_factor(inNonDutch)
+    OutgroupInteraction = as_factor(IntergroupContact),
+    NonOutgroupInteraction = as_factor(IngroupContact)
   )
@@
         "relatedness.daily.all",
-        "thermometerDutch_1",
-        "keymotive_fulfillemt_1",
-        "competence_1",
-        "autonomy_1",
-        "relatedness_1", 
-        "quality_overall_1", 
+        "AttitudesDutch",
+        "KeyNeedFulfillment",
+        "competence",
+        "autonomy",
+        "relatednessNoInteraction", 
+        "qualityOverall", 
         "OutgroupInteraction",
@@
         "keyMotiveFulfilled",
-        "thermometerDutch_1",
-        "keymotive_fulfillemt_1",
-        "competence_1",
-        "autonomy_1",
-        "relatedness_1", 
-        "quality_overall_1"
+        "AttitudesDutch",
+        "KeyNeedFulfillment",
+        "competence",
+        "autonomy",
+        "relatedness", 
+        "qualityOverall"
       )
@@
 # Between participants contact frequency
-workerContactFreq <- dtWorker$full %>%
+workerContactFreq <- dtS1$full %>%
   group_by(PID) %>%
@@
     n = n(),
-    SumContactNL = sum(Contact_dum),
+    SumContactNL = sum(IntergroupContact),
     PercContactNL = SumContactNL / n * 100,
     SumContactNLAll = sum(number),
-    AvAttitude = mean(thermometerDutch_1, na.rm = T)
+    AvAttitude = mean(AttitudesDutch, na.rm = T)
   ) %>%
@@
 # df.mplus = remove_all_labels(select(df,
-#                                     PID, session,
+#                                     PID, ResponseId,
 #                                     thermometerDutch_1, inNonDutch, Contact_dum,
 #                                     keyMotiveFulfilled, autonomy.daily.all, competence.daily.all, relatedness.daily.all))
-# names(df.mplus)= c("PID", "session", "att", "intin", "intout", "keymot", "aut", "comp", "rel")
-# mplus = df.mplus[order(df.mplus$PID, df.mplus$session),]
+# names(df.mplus)= c("PID", "ResponseId", "att", "intin", "intout", "keymot", "aut", "comp", "rel")
+# mplus = df.mplus[order(df.mplus$PID, df.mplus$ResponseId),]
 # mplus.intcont = mplus[mplus$intout==1,]
@@
 ```{r newVarsStudents}
-df <- dtStudents$clean
+df <- dtS2$clean
 
 # Add ID variables
-df$PID <- as.numeric(factor(df$session)) # participant ID
+df$PID <- as.numeric(factor(df$ResponseId)) # participant ID
 
 # order time
-df$TID <- factor(df$date_period, levels = unique(dtStudents$raw.daily$date_period))
+df$TID <- factor(df$date_period, levels = unique(dtS2$raw.daily$date_period))
 df$TIDnum <- as.numeric(df$TID) # get numeric TID
@@
 df$IntergroupContact <- (df$IntergroupContact-2)*-1
-
-# Fix naming issues
-df$AttitudesPartner <- df$AttitdutuesPartnerstick
 
@@
 pairs.panels.new(
-  df[c("RelatednessSelf", "RelatednessOther")],
+  df[c("relatednessSelf", "relatednessOther")],
   labels = c(
@@
 )
-df$RelatednessInteraction <- rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = TRUE)
+df$RelatednessInteraction <- rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
 df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
@@
 df$Relatedness <-
-  rowMeans(df[, c("RelatednessInteraction", "RelatednessNoInteraction")],
+  rowMeans(df[, c("RelatednessInteraction", "relatednessNoInteraction")],
            na.rm = TRUE) %>%
@@
 # Core Need
-df$DaytimeNeedFullfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFullfillment)] <-
-  df$KeyNeedFullfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFullfillment)]
+df$DaytimeNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
+  df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]
 
 df$InteractionNeedFullfillment <- NA 
-df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFullfillment)] <- 
-  df$KeyNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFullfillment)]
+df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)] <- 
+  df$KeyNeedFulfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)]
 
@@
   mutate(
-    KeyNeedFullfillment.cm = mean(KeyNeedFullfillment, na.rm = TRUE),
+    KeyNeedFulfillment.cm = mean(KeyNeedFulfillment, na.rm = TRUE),
     # cluster mean (mean of PP)
-    KeyNeedFullfillment.cwc = KeyNeedFullfillment - KeyNeedFullfillment.cm,
+    KeyNeedFulfillment.cwc = KeyNeedFulfillment - KeyNeedFulfillment.cm,
     # cluster mean centered (within PP centered)
@@
 # store
-dtStudents$full <- df
+dtS2$full <- df
 rm(df)
@@
 # Between participants contact frequency
-studentContactFreq <- dtStudents$full %>%
+studentContactFreq <- dtS2$full %>%
   group_by(PID) %>%
@@
     AvAttitude = mean(AttitudesDutch, na.rm = TRUE),
-    AvQuality = mean(quality_overall, na.rm = TRUE)
+    AvQuality = mean(qualityOverall, na.rm = TRUE)
   ) %>%
@@
 # dataframe where interaction types are recoded
-studentInteractionType <- dtStudents$full %>%
+studentInteractionType <- dtS2$full %>%
   mutate(
@@
   mutate(
+    NonDutchNum = ifelse(NonDutchContact == 0, 0, NonDutchNum), # measurement error --- only measured when NonDutchContact==1
     OutgroupInteraction = factor(
@@
       c(
-        "KeyNeedFullfillment",
-        "Competence",
-        "Autonomy",
+        "KeyNeedFulfillment",
+        "competence",
+        "autonomy",
         "Relatedness",
         "AttitudesDutch",
-        "quality_overall",
+        "qualityOverall",
         "OutgroupInteraction",
@@
       c(
-        "KeyNeedFullfillment",
-        "Competence",
-        "Autonomy",
+        "KeyNeedFulfillment",
+        "competence",
+        "autonomy",
         "Relatedness",
         "AttitudesDutch",
-        "quality_overall"
+        "qualityOverall"
       )
@@
 ```{r newVarsMedical}
-df <- dtMedical$clean
+df <- dtS3$clean
 
 # Add ID variables
-df$PID <- as.numeric(factor(df$session)) # participant ID
+df$PID <- as.numeric(factor(df$ResponseId)) # participant ID
 
@@
 df$TID <-
-  factor(df$date_period, levels = unique(dtMedical$raw.daily$date_period))
+  factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
 df$TIDnum <- as.numeric(df$TID) # get numeric TID
@@
 df$RelatednessInteraction <-
-  rowMeans(df[c("RelatednessSelf", "RelatednessOther")], na.rm = TRUE)
+  rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
 df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
@@
 
-df$DaytimeNeedFullfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
+df$DaytimeNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
   df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]
-
-df$DaytimeNeedImportance <- NA 
-df$DaytimeNeedImportance[df$InteractionDum == 1 & !is.na(df$DaytimeNeedFulfillment)] <- 
-  df$DaytimeNeedFulfillment[df$InteractionDum == 1 & !is.na(df$DaytimeNeedFulfillment)]
 
@@
         "KeyNeedFulfillment",
-        "Competence",
-        "Autonomy",
+        "competence",
+        "autonomy",
         "Relatedness",
@@
 # store
-dtMedical$full <- df
+dtS3$full <- df
 rm(df)
@@
 medicalContactFreq <- 
-  dtMedical$full %>%
+  dtS3$full %>%
   group_by(PID) %>%
@@
 medicalPltIDs <-
-  dtMedical$full %>%
+  dtS3$full %>%
   group_by(PID) %>%
@@
 medicalOutPltIDs <-
-  dtMedical$full %>%
+  dtS3$full %>%
   filter(OutgroupInteraction == "Yes") %>%
@@
 
-## Data Availability and Time Scales
+## Data Availability and Sample Selection  {.tabset}
 
@@
 
+### Study 1
+
 ```{r workerDataAvailability}
-dtworkerMissing <- dtWorker$full %>%
-  select(
-    PID,
+dtS1Missing <- dtS1$full %>%
+  select(
+    PID,
+    TIDnum,
     Finished.daily,
@@
 
-library(reshape)
-dtworkerAvailability <- reshape::cast(dtworkerMissing, PID ~ TID)
-write.csv(dtworkerAvailability, "data/S1_Workers/processed/workerAvailability.csv")
+dtS1Availability <- dtS1$full %>%
+  select(
+    PID,
+    TIDnum
+  ) %>%
+  arrange(PID, TIDnum) %>%
+  mutate(data = 1)
+dtS1Availability <- reshape::cast(dtS1Availability, PID ~ TIDnum) %>%
+  select(-PID)
+# sum(colMeans(dtS1Availability)*100 >= 66)
+# sum(rowMeans(dtS1Availability)*100 >= 66)
+# sum(colMeans(dtS1Availability[-5,])*100 >= 66)
+# sum(rowMeans(dtS1Availability[,-8])*100 >= 66)
+
+rownames(dtS1Availability) <- paste("PP", 1:nrow(dtS1Availability), sep = "_")
+colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
+dtS1Availability[1:ncol(dtS1Availability)] <- lapply(dtS1Availability[1:ncol(dtS1Availability)], function(x) {
+    cell_spec(x,
+              bold = FALSE,
+              color = "white",
+              background = ifelse(x == 1, "green", "red")
+              )
+})
+kbl(
+  dtS1Availability,
+  format = "html",
+  escape = FALSE,
+  align = "c",
+  booktabs = TRUE,
+  caption = "Study 1: Data Availability" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
+
+write.csv(dtS1Availability, "data/S1_Workers/processed/workerAvailability.csv")
+rm(dtS1Missing)
 ```
 
+```{r testOptimizer, include=FALSE}
+set.seed(42)
+df <- as.data.table(matrix(rbinom(10*5,1,.66), ncol=10, nrow=5))
+df
+
+best.list.row.df = list()
+for (i in 1:nrow(df)) {
+  # get best subset for rows based on how many columns have more than 66% data
+  rowlist = combn(nrow(df), i)
+  numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
+  cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
+  best = which(numobs == max(numobs))[1]
+  best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
+}
+
+best.list.col.df = list()
+for (i in 1:ncol(df)) {
+  # get best subset for columns based on how many rows have more than 66% data
+  collist = combn(ncol(df), i)
+  numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
+  cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
+  best = which(numobs == max(numobs))[1]
+  best.list.col.df = c(best.list.col.df, list(collist[, best]))
+}
+rm(df, best.list.row.df, best.list.col.df)
+```
+
+### Study 2
+
 ```{r studentDataAvailability}
-dtstudentMissing <- dtStudents$full %>%
+dtS2Missing <- dtS2$full %>%
   select(
@@
 
-dtstudentAvailability <- reshape::cast(dtstudentMissing, PID ~ TID)
-write.csv(dtstudentAvailability, "data/S2_Students/processed/studentAvailability.csv")
+dtS2Availability <- dtS2$full %>%
+  select(
+    PID,
+    TIDnum
+  ) %>%
+  arrange(PID, TIDnum) %>%
+  mutate(data = 1)
+dtS2Availability <- reshape::cast(dtS2Availability, PID ~ TIDnum) %>%
+  select(-PID)
+
+rownames(dtS2Availability) <- paste("PP", 1:nrow(dtS2Availability), sep = "_")
+colnames(dtS2Availability) <- paste("t", 1:ncol(dtS2Availability), sep = "_")
+
+write.csv(dtS2Availability, "data/S2_Students/processed/studentAvailability.csv")
+rm(dtS2Missing)
+
+dtS2Availability[1:ncol(dtS2Availability)] <- lapply(dtS2Availability[1:ncol(dtS2Availability)], function(x) {
+    cell_spec(x,
+              bold = FALSE,
+              color = "white",
+              background = ifelse(x == 1, "green", "red")
+              )
+})
+kbl(
+  dtS2Availability,
+  format = "html",
+  escape = FALSE,
+  align = "c",
+  booktabs = TRUE,
+  caption = "Study 2: Data Availability" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
 ```
 
+### Study 3
+
 ```{r medicalDataAvailability}
-dtMedicalMissing <- dtMedical$full %>%
+dtS3Missing <- dtS3$full %>%
   select(
@@
 
-dtMedicalAvailability <- reshape::cast(dtMedicalMissing, PID ~ TID)
-write.csv(dtMedicalAvailability, "data/S3_Medical/processed/medicalAvailability.csv")
+dtS3Availability <- dtS3$full %>%
+  select(
+    PID,
+    TIDnum
+  ) %>%
+  arrange(PID, TIDnum) %>%
+  mutate(data = 1)
+dtS3Availability <- reshape::cast(dtS3Availability, PID ~ TIDnum) %>%
+  select(-PID)
+
+rownames(dtS3Availability) <- paste("PP", 1:nrow(dtS3Availability), sep = "_")
+colnames(dtS3Availability) <- paste("t", 1:ncol(dtS3Availability), sep = "_")
+
+write.csv(dtS3Availability, "data/S3_Medical/processed/medicalAvailability.csv")
+rm(dtS3Missing)
+
+dtS3Availability[1:ncol(dtS3Availability)] <- lapply(dtS3Availability[1:ncol(dtS3Availability)], function(x) {
+    cell_spec(x,
+              bold = FALSE,
+              color = "white",
+              background = ifelse(x == 1, "green", "red")
+              )
+})
+kbl(
+  dtS3Availability,
+  format = "html",
+  escape = FALSE,
+  align = "c",
+  booktabs = TRUE,
+  caption = "Study 2: Data Availability" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
 ```
+
+## Time Scales and Timepoint Aggregation
 
@@
 
-```{r workerDataSelection}
+```{r timeInNL}
+#median(rbind(rbind(dtS1$raw.pre$timeNL_1, dtS2$raw.pre$timeNL), dtS3$raw.pre$timeNL), na.rm=TRUE)
+```
+
+```{r workerDataSelection, eval=FALSE}
 varListWorker <- 
@@
 
-dtWorker$viz <- list()
-
-dtWorker$vizall <- dtWorker$full %>%
+dtS1$viz <- list()
+
+dtS1$vizall <- dtS1$full %>%
   select(
@@
 
-dtWorker$viz$Interaction <- dtWorker$vizall %>%
+dtS1$viz$Interaction <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$KeyNeed <- dtWorker$vizall %>%
+dtS1$viz$KeyNeed <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$SDT <- dtWorker$vizall %>%
+dtS1$viz$SDT <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$Attitudes <- dtWorker$vizall %>%
+dtS1$viz$Attitudes <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$WB <- dtWorker$vizall %>%
+dtS1$viz$WB <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodOverall <- dtWorker$vizall %>%
+dtS1$viz$MoodOverall <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV01Pt01 <- dtWorker$vizall %>%
+dtS1$viz$MoodV01Pt01 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV01Pt02 <- dtWorker$vizall %>%
+dtS1$viz$MoodV01Pt02 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV01Pt03 <- dtWorker$vizall %>%
+dtS1$viz$MoodV01Pt03 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV01Pt04 <- dtWorker$vizall %>%
+dtS1$viz$MoodV01Pt04 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV01Pt05 <- dtWorker$vizall %>%
+dtS1$viz$MoodV01Pt05 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV01Pt06 <- dtWorker$vizall %>%
+dtS1$viz$MoodV01Pt06 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV02Pt01 <- dtWorker$vizall %>%
+dtS1$viz$MoodV02Pt01 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV02Pt02 <- dtWorker$vizall %>%
+dtS1$viz$MoodV02Pt02 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV02Pt03 <- dtWorker$vizall %>%
+dtS1$viz$MoodV02Pt03 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV02Pt04 <- dtWorker$vizall %>%
+dtS1$viz$MoodV02Pt04 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV02Pt05 <- dtWorker$vizall %>%
+dtS1$viz$MoodV02Pt05 <- dtS1$vizall %>%
   select(
@@
 
-dtWorker$viz$MoodV02Pt06 <- dtWorker$vizall %>%
+dtS1$viz$MoodV02Pt06 <- dtS1$vizall %>%
   select(
@@
 
-```{r workerMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtWorker$vizPlotsMean <-
-  lapply(seq_along(dtWorker$viz),
+```{r workerMeanViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
+dtS1$vizPlotsMean <-
+  lapply(seq_along(dtS1$viz),
          function(i)
-           statVarPlot(dataIn = dtWorker$viz[[i]], 
-                       name = names(dtWorker$viz)[[i]], 
+           statVarPlot(dataIn = dtS1$viz[[i]], 
+                       name = names(dtS1$viz)[[i]], 
                        namList = varNameListWorker,
@@
          )
-dtWorker$vizPlotsMean
+dtS1$vizPlotsMean
 ```
@@
 
-```{r WorkerSdViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtWorker$vizPlotsSd <-
-  lapply(seq_along(dtWorker$viz),
+```{r WorkerSdViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
+dtS1$vizPlotsSd <-
+  lapply(seq_along(dtS1$viz),
          function(i)
-           statVarPlot(dataIn = dtWorker$viz[[i]], 
-                       name = names(dtWorker$viz)[[i]], 
+           statVarPlot(dataIn = dtS1$viz[[i]], 
+                       name = names(dtS1$viz)[[i]], 
                        namList = varNameListWorker,
@@
          )
-dtWorker$vizPlotsSd
+dtS1$vizPlotsSd
 ```
@@
 
-```{r studentDataSelection}
+```{r studentDataSelection, eval=FALSE}
 varListStudent <- 
@@
 
-dtStudents$viz <- list()
-
-dtStudents$vizall <- dtStudents$full %>%
+dtS2$viz <- list()
+
+dtS2$vizall <- dtS2$full %>%
   select(
@@
 
-dtStudents$viz$Interaction <- dtStudents$vizall %>%
+dtS2$viz$Interaction <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$KeyNeed <- dtStudents$vizall %>%
+dtS2$viz$KeyNeed <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$AntiSocialBehaviorPt1 <- dtStudents$vizall %>%
+dtS2$viz$AntiSocialBehaviorPt1 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$AntiSocialBehaviorPt2 <- dtStudents$vizall %>%
+dtS2$viz$AntiSocialBehaviorPt2 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$AntiSocialBehaviorPt3 <- dtStudents$vizall %>%
+dtS2$viz$AntiSocialBehaviorPt3 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$ProSocialBehaviorPt1 <- dtStudents$vizall %>%
+dtS2$viz$ProSocialBehaviorPt1 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$ProSocialBehaviorPt2 <- dtStudents$vizall %>%
+dtS2$viz$ProSocialBehaviorPt2 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$GoalsPt1 <- dtStudents$vizall %>%
+dtS2$viz$GoalsPt1 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$GoalsPt2 <- dtStudents$vizall %>%
+dtS2$viz$GoalsPt2 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$GoalsPt3 <- dtStudents$vizall %>%
+dtS2$viz$GoalsPt3 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$GoalsPt4 <- dtStudents$vizall %>%
+dtS2$viz$GoalsPt4 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$GoalsPt5 <- dtStudents$vizall %>%
+dtS2$viz$GoalsPt5 <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$SDT <- dtStudents$vizall %>%
+dtS2$viz$SDT <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$Attitudes <- dtStudents$vizall %>%
+dtS2$viz$Attitudes <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$WB <- dtStudents$vizall %>%
+dtS2$viz$WB <- dtS2$vizall %>%
   select(
@@
 
-dtStudents$viz$Event <- dtStudents$vizall %>%
+dtS2$viz$Event <- dtS2$vizall %>%
   select(
@@
 
-```{r studentMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtStudents$vizPlotsMean <-
-  lapply(seq_along(dtStudents$viz),
+```{r studentMeanViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
+dtS2$vizPlotsMean <-
+  lapply(seq_along(dtS2$viz),
          function(i)
-           statVarPlot(dataIn = dtStudents$viz[[i]], 
-                       name = names(dtStudents$viz)[[i]], 
+           statVarPlot(dataIn = dtS2$viz[[i]], 
+                       name = names(dtS2$viz)[[i]], 
                        namList = varNameListStudent,
@@
          )
-dtStudents$vizPlotsMean
+dtS2$vizPlotsMean
 ```
@@
 
-```{r studentSdViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtStudents$vizPlotsSd <-
-  lapply(seq_along(dtStudents$viz),
+```{r studentSdViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
+dtS2$vizPlotsSd <-
+  lapply(seq_along(dtS2$viz),
          function(i)
-           statVarPlot(dataIn = dtStudents$viz[[i]], 
-                       name = names(dtStudents$viz)[[i]], 
+           statVarPlot(dataIn = dtS2$viz[[i]], 
+                       name = names(dtS2$viz)[[i]], 
                        namList = varNameListStudent,
@@
          )
-dtStudents$vizPlotsSd
+dtS2$vizPlotsSd
 ```
@@
 
-```{r medicalDataSelection}
+```{r medicalDataSelection, eval=FALSE}
 varListMedical <- 
@@
 
-dtMedical$viz <- list()
-
-dtMedical$vizall <- dtMedical$full %>%
+dtS3$viz <- list()
+
+dtS3$vizall <- dtS3$full %>%
   select(
@@
 
-dtMedical$viz$Interaction <- dtMedical$vizall %>%
+dtS3$viz$Interaction <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$KeyNeedFulfillment <- dtMedical$vizall %>%
+dtS3$viz$KeyNeedFulfillment <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$KeyNeedImportance <- dtMedical$vizall %>%
+dtS3$viz$KeyNeedImportance <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$AntiSocialBehaviorPt1 <- dtMedical$vizall %>%
+dtS3$viz$AntiSocialBehaviorPt1 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$AntiSocialBehaviorPt2 <- dtMedical$vizall %>%
+dtS3$viz$AntiSocialBehaviorPt2 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$AntiSocialBehaviorPt3 <- dtMedical$vizall %>%
+dtS3$viz$AntiSocialBehaviorPt3 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$ProSocialBehaviorPt1 <- dtMedical$vizall %>%
+dtS3$viz$ProSocialBehaviorPt1 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$ProSocialBehaviorPt2 <- dtMedical$vizall %>%
+dtS3$viz$ProSocialBehaviorPt2 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$GoalsPt1 <- dtMedical$vizall %>%
+dtS3$viz$GoalsPt1 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$GoalsPt2 <- dtMedical$vizall %>%
+dtS3$viz$GoalsPt2 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$GoalsPt3 <- dtMedical$vizall %>%
+dtS3$viz$GoalsPt3 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$GoalsPt4 <- dtMedical$vizall %>%
+dtS3$viz$GoalsPt4 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$GoalsPt5 <- dtMedical$vizall %>%
+dtS3$viz$GoalsPt5 <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$SDT <- dtMedical$vizall %>%
+dtS3$viz$SDT <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$Attitudes <- dtMedical$vizall %>%
+dtS3$viz$Attitudes <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$WB <- dtMedical$vizall %>%
+dtS3$viz$WB <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$Emotion <- dtMedical$vizall %>%
+dtS3$viz$Emotion <- dtS3$vizall %>%
   select(
@@
 
-dtMedical$viz$Event <- dtMedical$vizall %>%
+dtS3$viz$Event <- dtS3$vizall %>%
   select(
@@
 
-```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtMedical$vizPlotsMean <-
-  lapply(seq_along(dtMedical$viz),
+```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
+dtS3$vizPlotsMean <-
+  lapply(seq_along(dtS3$viz),
          function(i)
-           statVarPlot(dataIn = dtMedical$viz[[i]], 
-                       name = names(dtMedical$viz)[[i]], 
+           statVarPlot(dataIn = dtS3$viz[[i]], 
+                       name = names(dtS3$viz)[[i]], 
                        namList = varNameListMedical,
@@
          )
-dtMedical$vizPlotsMean
+dtS3$vizPlotsMean
 ```
@@
 
-```{r medicalSdViz, echo=FALSE, results='hide',fig.keep='all'} 
-dtMedical$vizPlotsSd <-
-  lapply(seq_along(dtMedical$viz),
+```{r medicalSdViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
+dtS3$vizPlotsSd <-
+  lapply(seq_along(dtS3$viz),
          function(i)
-           statVarPlot(dataIn = dtMedical$viz[[i]], 
-                       name = names(dtMedical$viz)[[i]], 
+           statVarPlot(dataIn = dtS3$viz[[i]], 
+                       name = names(dtS3$viz)[[i]], 
                        namList = varNameListMedical,
@@
          )
-dtMedical$vizPlotsSd
+dtS3$vizPlotsSd
 ```
@@
 # **Software Information**  
-The full session information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
+The full ResponseId information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
 
@@
   \renewcommand{\arraystretch}{0.8} <!-- decrease line spacing for the table -->
-```{r Reproducibility-SessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
+```{r Reproducibility-sessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
 df_session_platform <- devtools::session_info()$platform %>%
@@
 \renewcommand{\arraystretch}{0.6} <!-- decrease line spacing for the table -->
-```{r Reproducibility-SessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
+```{r Reproducibility-sessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
 df_session_packages <- devtools::session_info()$packages %>%

### 2022-06-17 Supplemental-Material-A-Annotated-Analysis.Rmd
daily and weekly added

@@
   select(-starts_with("t_"))
+names(dat.pre.red) <- paste(names(dat.pre.red), "pre", sep = ".")
 
@@
   by.x = "last_outside_referrer",
-  by.y = "platformId",
+  by.y = "platformId.pre",
   all = T
@@
   select(-starts_with("t_"))
+names(dat.post.red) <- paste(names(dat.post.red), "post", sep = ".")
 
@@
   by.x = "last_outside_referrer",
-  by.y = "last_outside_referrer",
+  by.y = "last_outside_referrer.post",
   all = T
@@
     Finished.pre == 1,
-    Finished.daily == 1,
+    Finished == 1,
     !is.na(last_outside_referrer)
@@
     SubTime = chron::times(time.0),
-    time.daily = as.character(time.daily),
-    PPDate = as.Date(df$dateQualtrics.daily),
+    time.daily = as.character(time),
+    PPDate = as.Date(df$dateQualtrics),
     number = replace_na(ContactNum, 0),
@@
   mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
 
@@
   mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+names(df.post) <- paste(names(df.post), "post", sep = ".")
 
@@
   y = df.pre,
-  by = "ResponseId",
+  by.x = "ResponseId",
+  by.y = "ResponseId.pre", 
   suffixes = c(".daily", ".pre"),
-  all = F
+  all = FALSE
 )
@@
   y = df.post,
-  by = "ResponseId",
+  by.x = "ResponseId",
+  by.y = "ResponseId.post", 
   suffixes = c(".pre", ".post"),
-  all = F
+  all = FALSE
 )
@@
   y = df.pre,
-  by = "ResponseId",
+  by.x = "ResponseId",
+  by.y = "ResponseId.pre", 
   suffixes = c(".daily", ".pre"),
@@
   y = df.post,
-  by = "ResponseId",
+  by.x = "ResponseId",
+  by.y = "ResponseId.post", 
   suffixes = c(".pre", ".post"),
@@
 df$keyMotiveFulfilled <- rowSums(df[, c("KeyNeedFulfillment", "DaytimeNeedFulfillment")], na.rm = T)
-df$autonomy.daily.all <- rowSums(df[, c("autonomy", "autonomy_NoInt_1")], na.rm = T)
-df$competence.daily.all <- rowSums(df[, c("competence", "competence_NoInt_1")], na.rm = T)
+df$autonomy.daily.all <- rowSums(df[, c("autonomy_Int", "autonomy_NoInt")], na.rm = T)
+df$competence.daily.all <- rowSums(df[, c("competence_Int", "competence_NoInt")], na.rm = T)
 # cor(df$relatednessOther, df$relatedness_self_1,use="complete.obs")
@@
 
-df$Autonomy <- df$autonomy.daily.all
-df$Competence <- df$competence.daily.all
-df$Relatedness <- df$relatedness.daily.all
+df$autonomy <- df$autonomy.daily.all
+df$competence <- df$competence.daily.all
+df$relatedness <- df$relatedness.daily.all
 
@@
     AvKeyNeedNoInt = mean(DaytimeNeedFulfillment, na.rm = T),
-    AvAutonomy = mean(autonomy.daily.all, na.rm = T),
-    AvCompetence = mean(competence.daily.all, na.rm = T),
-    AvRelatedness = mean(relatedness.daily.all, na.rm = T),
+    AvAutonomy = mean(autonomy, na.rm = T),
+    AvCompetence = mean(competence, na.rm = T),
+    AvRelatedness = mean(relatedness, na.rm = T),
     AvThermo = mean(AttitudesDutch, na.rm = T),
@@
         "keyMotiveFulfilled",
-        "competence.daily.all",
-        "autonomy.daily.all",
-        "relatedness.daily.all",
+        "competence",
+        "autonomy",
+        "relatedness",
         "AttitudesDutch",
@@
 )
-df$RelatednessInteraction <- rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
-df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
+df$relatednessInteraction <- rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
+df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
 # Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
-df$Relatedness <-
-  rowMeans(df[, c("RelatednessInteraction", "relatednessNoInteraction")],
+df$relatedness <-
+  rowMeans(df[, c("relatednessInteraction", "relatednessNoInteraction")],
            na.rm = TRUE) %>%
@@
         "autonomy",
-        "Relatedness",
+        "relatedness",
         "AttitudesDutch",
@@
 )
-df$RelatednessInteraction <-
+df$relatednessInteraction <-
   rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
-df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
+df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
 # Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
-df$Relatedness <-
-  rowMeans(df[, c("RelatednessInteraction", "relatednessNoInteraction")],
+df$relatedness <-
+  rowMeans(df[, c("relatednessInteraction", "relatednessNoInteraction")],
            na.rm = TRUE) %>%
@@
         "autonomy",
-        "Relatedness",
+        "relatedness",
         "AllportsCondition",
@@
 ```{r workerDataAvailability}
-dtS1Missing <- dtS1$full %>%
-  select(
-    PID,
-    TIDnum,
-    Finished.daily,
-    PPDate,
-    daytime,
-    TID
-  ) %>%
-  mutate(
-    date_period = paste(PPDate, daytime, sep = " "),
-    available = 1
-  )
-
 dtS1Availability <- dtS1$full %>%
@@
 dtS1Availability <- reshape::cast(dtS1Availability, PID ~ TIDnum) %>%
-  select(-PID)
+  select(-PID) %>%
+  mutate_all(function(x) ifelse(x>1,1,x))
 # sum(colMeans(dtS1Availability)*100 >= 66)
@@
 write.csv(dtS1Availability, "data/S1_Workers/processed/workerAvailability.csv")
-rm(dtS1Missing)
 ```
@@
 set.seed(42)
-df <- as.data.table(matrix(rbinom(10*5,1,.66), ncol=10, nrow=5))
+df <- as.data.table(matrix(rbinom(5*10,1,.66), nrow=5, ncol=10))
 df
@@
 ```{r studentDataAvailability}
-dtS2Missing <- dtS2$full %>%
-  select(
-    PID,
-    date.daily,
-    periodMA,
-    date_period,
-    TID
-  ) %>%
-  mutate(
-    TID = as.numeric(TID),
-    available = 1
-  )
-
 dtS2Availability <- dtS2$full %>%
@@
 write.csv(dtS2Availability, "data/S2_Students/processed/studentAvailability.csv")
-rm(dtS2Missing)
 
@@
 ```{r medicalDataAvailability}
-dtS3Missing <- dtS3$full %>%
-  select(
-    PID,
-    date.daily,
-    periodMA,
-    date_period,
-    TID
-  ) %>%
-  mutate(
-    TID = as.numeric(TID),
-    available = 1
-  )
-
 dtS3Availability <- dtS3$full %>%
@@
 write.csv(dtS3Availability, "data/S3_Medical/processed/medicalAvailability.csv")
-rm(dtS3Missing)
 
@@
 
+```{r checkVarAvailability}
+varNamesPre <- varNames %>%
+  filter(
+    surveyS1 == "pre",
+    surveyS2 == "pre",
+    str_detect(surveyS3, "^pre")
+  ) %>%
+  select(varNam)
+varNamesPreFull <-
+  data.frame(varNam = Reduce(intersect, list(
+    names(dtS1$full), names(dtS3$full), names(dtS3$full)
+  ))) %>%
+  filter(str_detect(varNam, ".pre$"))
+
+varNamesDaily <- varNames %>%
+  filter(
+    surveyS1 == "daily",
+    surveyS2 == "daily",
+    surveyS3 == "daily"
+  ) %>%
+  select(varNam)
+varNamesDailyFull <-
+  data.frame(varNam = Reduce(intersect, list(
+    names(dtS1$full), names(dtS3$full), names(dtS3$full)
+  ))) %>%
+  filter(!str_detect(varNam, ".pre$|.post$"))
+
+varNamesPost <- varNames %>%
+  filter(
+    surveyS1 == "post",
+    surveyS2 == "post",
+    surveyS3 == "post"
+  ) %>%
+  select(varNam)
+varNamesPostFull <-
+  data.frame(varNam = Reduce(intersect, list(
+    names(dtS1$full), names(dtS3$full), names(dtS3$full)
+  ))) %>%
+  filter(str_detect(varNam, ".post$"))
+```
+
+```{r analysisDFs}
+# Main Analysis: 3MPCA (common variables across studies)
+dropVarPcaMain <- c(
+  # Meta Data
+  "last_outside_referrer",
+  "created",
+  "ended",
+  "ip_address",
+  "date",
+  "ResponseId",
+  "server_language",
+  "Meta_Browser",
+  # Interaction Dummies (non-continuous)
+  "IntergroupContact",
+  "IngroupContact",
+  "InteractionDum",
+  # Interaction Counts (non-continuous)
+  "ContactNum",
+  "NonDutchNum",
+  # Interaction Descriptives (non-continuous)
+  "duration",
+  "dyadGroup",
+  "groupSize",
+  # Keyneed Freetexts (non-continuous)
+  "KeyNeed",
+  "DaytimeNeed",
+  # Relatedness individual items
+  "relatednessSelf",
+  "relatednessOther",
+  "relatednessNoInteraction"
+)
+
+varNam3mpcaMain <- varNamesDailyFull %>%
+  filter(!varNam %in% dropVarPcaMain) %>%
+  pull
+
+dtMain3mpca <- rbind(
+  dtS1$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
+  dtS3$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
+  dtS3$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
+  ) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+    ) %>%
+  select(
+    ID,
+    PID,
+    TID,
+    date,
+    week,
+    TIDnum,
+    study,
+    everything()
+  ) %>%
+  arrange(ID, TIDnum)
+
+dtMain3mpcaDaily <- dtMain3mpca %>%
+  group_by(ID, date, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(date))) %>%
+  ungroup %>%
+  select(ID, date, TIDnum, everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .))
+
+dtMain3mpcaWeekly <- dtMain3mpca %>%
+  group_by(ID, week, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(week))) %>%
+  ungroup %>%
+  select(ID, week, TIDnum, everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .))
+
+dtMain3mpcaWeekly
+
+stats::reshape(dtMain3mpca, idvar = "ID", timevar = "TIDnum", direction = "wide")
+
+```
+
 </br>
@@
     "InteractionNeedFullfillment",
-    "Autonomy",
-    "Competence",
-    "Relatedness",
+    "autonomy",
+    "competence",
+    "relatedness",
     "AttitudesDutch",
@@
     "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
-    "Autonomy" = "Autonomy",
-    "Competence" = "Competence",
-    "Relatedness" = "Relatedness",
+    "autonomy" = "Autonomy",
+    "competence" = "Competence",
+    "relatedness" = "Relatedness",
     "AttitudesDutch" = "Outgroup Attitude",
@@
     TIDnum,
-    Autonomy,
-    Relatedness, 
-    Competence
+    autonomy,
+    relatedness, 
+    competence
   )
@@
     "StudentGoal10",
-    "Autonomy",
-    "Competence",
+    "autonomy",
+    "competence",
     # Break relatedness up into interaction vs. no interaction(?)
     # vars are already calculated
-    "Relatedness",
+    "relatedness",
     "AttitudesDutch",
@@
     "StudentGoal10" = "Spiritual or religious",
-    "Autonomy" = "Autonomy",
-    "Competence" = "Competence",
+    "autonomy" = "Autonomy",
+    "competence" = "Competence",
     # Break relatedness up into interaction vs. no interaction(?)
     # vars are already calculated
-    "Relatedness" = "Relatedness",
+    "relatedness" = "Relatedness",
     "AttitudesDutch" = "Outgroup Attitude",
@@
     # vars are already calculated
-    "Relatedness",
+    "relatedness",
     "AttitudesDutch",
@@
     # vars are already calculated
-    "Relatedness" = "Relatedness",
+    "relatedness" = "Relatedness",
     "AttitudesDutch" = "Outgroup Attitude",

### 2022-06-20 Supplemental-Material-A-Annotated-Analysis.Rmd
sample selection v.02

@@
 colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
-dtS1Availability[1:ncol(dtS1Availability)] <- lapply(dtS1Availability[1:ncol(dtS1Availability)], function(x) {
+dtS1AvailabilityKbl <- dtS1Availability
+
+dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)] <- lapply(dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)], function(x) {
     cell_spec(x,
@@
 kbl(
-  dtS1Availability,
+  dtS1AvailabilityKbl,
   format = "html",
@@
 ```{r testOptimizer, include=FALSE}
-set.seed(42)
-df <- as.data.table(matrix(rbinom(5*10,1,.66), nrow=5, ncol=10))
-df
-
-best.list.row.df = list()
-for (i in 1:nrow(df)) {
-  # get best subset for rows based on how many columns have more than 66% data
-  rowlist = combn(nrow(df), i)
-  numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
-  cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
-  best = which(numobs == max(numobs))[1]
-  best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
+# set.seed(42)
+# nr <- 5
+# nc <- 10
+# df <- as.data.table(matrix(rbinom(nr*nc,1,.66), nrow=nr, ncol=nc))
+# df
+# 
+# best.list.row.df = list()
+# for (i in 1:nrow(df)) {
+#   # get best subset for rows based on how many columns have more than 66% data
+#   rowlist = combn(nrow(df), i)
+#   numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
+#   cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
+#   best = which(numobs == max(numobs))[1]
+#   best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
+# }
+# 
+# best.list.col.df = list()
+# for (i in 1:ncol(df)) {
+#   # get best subset for columns based on how many rows have more than 66% data
+#   collist = combn(ncol(df), i)
+#   numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
+#   cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
+#   best = which(numobs == max(numobs))[1]
+#   best.list.col.df = c(best.list.col.df, list(collist[, best]))
+# }
+# rm(df, best.list.row.df, best.list.col.df)
+```
+
+```{r testOptimizer03, include=FALSE}
+# library(dplyr)
+# library(ROI)
+# library(ROI.plugin.glpk)
+# library(ompr)
+# library(ompr.roi)
+# 
+# set.seed(42)
+# tf <- matrix(sample(c(TRUE, FALSE), 1488, replace=TRUE), 31)
+# M3 <- t(replicate(31, 1969:2016, simplify=TRUE))
+# M3[tf] <- NA
+# 
+# m <- +!is.na(M3) # gets logical matrix; 0 if NA else 1    
+# #m <- as.data.frame(dtS1AvailabilityTest)
+# nr <- nrow(m)
+# nc <- ncol(m)
+# n_years <- 15 
+# 
+# 
+# model <- MIPModel() %>% 
+#   # keep[i,j] is 1 if matrix cell [i,j] is to be kept else 0
+#   add_variable(keep[i,j], i = 1:nr, j = 1:nc, typ = "binary") %>% 
+#   # rm_row[i] is 1 if row i is selected for removal else 0
+#   add_variable(rm_row[i], i = 1:nr, type = "binary") %>% 
+#   # rm_col[j] is 1 if column j is selected for removal else 0
+#   add_variable(rm_col[j], j = 1:nc, type = "binary") %>% 
+#   # maximize good cells kept
+#   set_objective(sum_expr(keep[i,j], i = 1:nr, j = 1:nc), "max") %>% 
+#   # cell can be kept only when row is not selected for removal
+#   add_constraint(sum_expr(keep[i,j], j = 1:nc) <= 1 - rm_row[i], i = 1:nr) %>%
+#   # cell can be kept only when column is not selected for removal
+#   add_constraint(sum_expr(keep[i,j], i = 1:nr) <= 1 - rm_col[j], j = 1:nc) %>%
+#   # keep at most n_years columns i.e. remove at least (nc - n_years) columns
+#   # only non-NA values can be kept
+#   add_constraint(m[i,j] + rm_row[i] + rm_col[j] >= 1, i = 1:nr, j = 1:nc) %>% 
+#   #add_constraint(sum_expr(m[,j], j = 1:nc)/nc*100 >= 66) %>% 
+#   #add_constraint(sum_expr(m[i,], i = 1:nr)/nr*100 >= 66) %>% 
+#   # keep cols with more 
+#   #add_constraint(colMeans(m[i,j], i = 1:nr, j = 1:nc) >= 0.66) %>%
+#   # I used >= instead of == to avoid infeasiblity
+#   #add_constraint(sum_expr(rm_col[j], j = 1:nc) >= nc - n_years) %>% 
+#   # solve using free glpk solver
+#   solve_model(with_ROI(solver = "glpk", verbose = TRUE))
+# 
+# solver_status(model)
+# rm_rows <- model %>% 
+#   get_solution(rm_row[i]) %>% 
+#   filter(value > 0) %>% 
+#   pull(i)
+# 
+# rm_cols <- model %>% 
+#   get_solution(rm_col[j]) %>% 
+#   filter(value > 0) %>% 
+#   pull(j) 
+# 
+# result <- m[-rm_rows, -rm_cols, drop = F]
+# result
+```
+
+```{r testOptimizer05}
+cleanM <- function(M, c = 66) {
+  check <- list()
+  rowNamesOut <- c()
+  colNamesOut <- c()
+  for (i in 1:length(c(row.names(M), names(M)))) {
+    rowMeans <- rowMeans(M) * 100
+    colMeans <- colMeans(M) * 100
+    rcMeans <- c(rowMeans, colMeans)
+    rm <- which.min(rcMeans) %>% names
+    rc <- ifelse(startsWith(rm, "PP_"), "row", "col")
+
+    check[[i]] <- rcMeans
+
+    if (!all(rcMeans >= c) && rc == "row") {
+      M <- M[!(row.names(M) %in% rm), ]
+      rowNamesOut <- append(rowNamesOut, rm)
+      cat(i, ": Row ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
+    } else if (!all(rcMeans >= c) && rc == "col") {
+      M <- M[, !(names(M) %in% rm)]
+      colNamesOut <- append(colNamesOut, rm)
+      cat(i, ": Column ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
+    } else {
+      cat(i, ": All row- and column means are over ", c, "%. The final matrix has ", nrow(M), " rows and ", ncol(M), " columns.", sep = "")
+      return(
+        list(
+          reducedMatrix = M,
+          rowNamesIn = row.names(M),
+          colNamesIn = names(M),
+          rowNamesOut = rowNamesOut,
+          colNamesOut = colNamesOut,
+          rowMeans = rowMeans,
+          colMeans = colMeans
+        )
+      )
     }
-
-best.list.col.df = list()
-for (i in 1:ncol(df)) {
-  # get best subset for columns based on how many rows have more than 66% data
-  collist = combn(ncol(df), i)
-  numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
-  cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
-  best = which(numobs == max(numobs))[1]
-  best.list.col.df = c(best.list.col.df, list(collist[, best]))
   }
-rm(df, best.list.row.df, best.list.col.df)
+}
+
+dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 66)
+
+PIDout <- gsub("PP_", "", dtS1RedInfo$rowNamesOut) %>% as.numeric
+TIDout <- gsub("t_", "", dtS1RedInfo$colNamesOut) %>% as.numeric
+
+dtS1Red <- dtS1$full %>%
+  filter(
+    !PID %in% PIDout,
+    !TIDnum %in% TIDout
+  ) %>%
+  mutate(TIDnum = as.numeric(factor(TIDnum)))
+rm(PIDout, TIDout)
+
+dtS1AvailabilityRedKbl <- dtS1RedInfo$reducedMatrix
+dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)] <- lapply(dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)], function(x) {
+    cell_spec(x,
+              bold = FALSE,
+              color = "white",
+              background = ifelse(x == 1, "green", "red")
+              )
+})
+kbl(
+  dtS1AvailabilityRedKbl,
+  format = "html",
+  escape = FALSE,
+  align = "c",
+  booktabs = TRUE,
+  caption = "Study 1: FinalData Availability" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
 ```
@@
 
-dtS2Availability[1:ncol(dtS2Availability)] <- lapply(dtS2Availability[1:ncol(dtS2Availability)], function(x) {
+dtS2AvailabilityKbl <- dtS2Availability
+dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)] <- lapply(dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)], function(x) {
     cell_spec(x,
@@
 kbl(
-  dtS2Availability,
+  dtS2AvailabilityKbl,
   format = "html",
@@
 
+```{r s2SampleReduction}
+dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 66)
+
+PIDout <- gsub("PP_", "", dtS2RedInfo$rowNamesOut) %>% as.numeric
+TIDout <- gsub("t_", "", dtS2RedInfo$colNamesOut) %>% as.numeric
+
+dtS2Red <- dtS2$full %>%
+  filter(
+    !PID %in% PIDout,
+    !TIDnum %in% TIDout
+  ) %>%
+  mutate(TIDnum = as.numeric(factor(TIDnum)))
+rm(PIDout, TIDout)
+
+dtS2AvailabilityRedKbl <- dtS2RedInfo$reducedMatrix
+dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)] <- lapply(dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)], function(x) {
+    cell_spec(x,
+              bold = FALSE,
+              color = "white",
+              background = ifelse(x == 1, "green", "red")
+              )
+})
+kbl(
+  dtS2AvailabilityRedKbl,
+  format = "html",
+  escape = FALSE,
+  align = "c",
+  booktabs = TRUE,
+  caption = "Study 2: Final Data Availability" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
 ### Study 3
@@
 
-dtS3Availability[1:ncol(dtS3Availability)] <- lapply(dtS3Availability[1:ncol(dtS3Availability)], function(x) {
+dtS3AvailabilityKbl <- dtS3Availability
+dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)] <- lapply(dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)], function(x) {
     cell_spec(x,
@@
 kbl(
-  dtS3Availability,
+  dtS3AvailabilityKbl,
   format = "html",
@@
   caption = "Study 2: Data Availability" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
+rm(dtS3AvailabilityKbl)
+```
+
+```{r s3SampleReduction}
+dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 66)
+
+PIDout <- gsub("PP_", "", dtS3RedInfo$rowNamesOut) %>% as.numeric
+TIDout <- gsub("t_", "", dtS3RedInfo$colNamesOut) %>% as.numeric
+
+dtS3Red <- dtS3$full %>%
+  filter(
+    !PID %in% PIDout,
+    !TIDnum %in% TIDout
+  ) %>%
+  mutate(TIDnum = as.numeric(factor(TIDnum)))
+rm(PIDout, TIDout)
+
+dtS3AvailabilityRedKbl <- dtS3RedInfo$reducedMatrix
+dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)] <- lapply(dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)], function(x) {
+    cell_spec(x,
+              bold = FALSE,
+              color = "white",
+              background = ifelse(x == 1, "green", "red")
+              )
+})
+kbl(
+  dtS3AvailabilityRedKbl,
+  format = "html",
+  escape = FALSE,
+  align = "c",
+  booktabs = TRUE,
+  caption = "Study 3: Final Data Availability" # complete caption for main document
 ) %>%
@@
   data.frame(varNam = Reduce(intersect, list(
-    names(dtS1$full), names(dtS3$full), names(dtS3$full)
+    names(dtS1Red), names(dtS3Red), names(dtS3Red)
   ))) %>%
@@
 dtMain3mpca <- rbind(
-  dtS1$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
-  dtS3$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
-  dtS3$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
+  dtS1Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
+  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
   ) %>%
@@
   mutate_all(~ifelse(is.nan(.), NA, .))
-
-dtMain3mpcaWeekly
-
-stats::reshape(dtMain3mpca, idvar = "ID", timevar = "TIDnum", direction = "wide")
-
-```
+```
+
+### Timescale ML test
+
+```{r }
+dtMain3mpca <- dtMain3mpca %>%
+  mutate(
+    TIDdate = as.numeric(factor(date)),
+    TIDweek = as.numeric(factor(week))
+  )
+
+test <-
+  lme4::lmer(KeyNeedFulfillment ~ 1 + (1 + TIDnum + TIDdate + TIDweek | PID),
+    data = dtMain3mpca
+  ) # use optim if it does not converge
+
+test2 <-
+  lme(
+    KeyNeedFulfillment ~ 1,
+    random = ~ 1 + TIDnum + TIDdate + TIDweek | PID,
+    data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
+    control = list(opt = "nlmimb")
+  ) # use optim if it does not converge
+
+summary(test2)
+#summ(test, digits = 3, center = FALSE)
+anova(test2)
+
+```
+
+## Multiple Imputation
+
+```{}
+
+## Center and Normalize
+
+```{r mainCentered}
+varNam3mpcaMainAnalysis <- varNamesDailyFull %>%
+  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum")) %>%
+  pull
+
+dtMain3mpcaNorm <-
+  PCAnorm(data = dtMain3mpca,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = varNam3mpcaMainAnalysis)
+```
+
+## Long to Wide
+
+```{r longToWide}
+
+dtMain3mpcaWideGmc <- dtMain3mpcaNorm %>% 
+  select(
+    ID,
+    TIDnum,
+    ends_with("_gmc")
+  ) %>%
+  reshape2::melt(.,
+              id.vars = c("ID", "TIDnum")) %>%
+  mutate(variable = paste(variable, TIDnum, sep = "_")) %>%
+  select(-TIDnum) %>% 
+  arrange(variable, ID) %>%
+  reshape(., idvar = "ID", timevar = "variable", direction = "wide")
+  #pivot_wider(names_from = variable, values_from = value)
+
+```
+
+## Three-Way ANOVA
+
+```{r threeWayAnova}
+library(ThreeWay)
+# threewayanova(
+#   Y = dtMain3mpcaWideGmc,
+#   n = nrow(dtMain3mpcaWideGmc), # Nr. ppt.s
+#   m = length(unique(dtMain3mpcaNorm$TIDnum)), # Nr. time points
+#   p = length(varNam3mpcaMainAnalysis) # Nr. variables
+# )
+```
+
+## Plot (x = var, y = z, group/line = t)
 

### 2022-06-20 scripts/functions/MlCorMat.R

@@ -245,6 +245,42 @@ MlTraitState <-
     data
   }
 
+PCAnorm <-
+  function(data,
+           pid,
+           tid,
+           selection) {
+    
+    # for testing
+    # data <- dtMain3mpca
+    # pid <- "PID"
+    # tid <- "TIDnum"
+    # selection <- varNam3mpcaMainAnalysis
+    
+    data02 <- data %>%
+      mutate_at(selection, as.numeric) %>%
+      group_by(!!rlang::sym(tid)) %>%
+      mutate_at(selection, list(cwt = ~ . - mean(., na.rm = TRUE))) %>%
+      ungroup %>%
+      filter_at(selection, all_vars(!is.na(.))) %>%
+      mutate_at(selection, list(gmc = ~ . - mean(., na.rm = TRUE))) %>%
+      mutate_at(selection, list(sigma = ~ sqrt(sum(. ^ 2)/(length(unique(!!rlang::sym(pid)))*length(unique(!!rlang::sym(tid)))))))
+    
+    cat("@Jannis: Don't forget to fix the filter once sample section is finalized.")
+    
+    data03 <- map_dfc(selection,
+                      ~ data02 %>%
+                        transmute(
+                          !!str_c(.x, '_z') :=
+                            !!rlang::sym(str_c(.x, "_cwt"))/!!rlang::sym(str_c(.x, "_sigma"))
+                        )) %>% 
+      bind_cols(data02, .)
+    
+    out <- data03
+    
+    out
+  }
+
 
 var_reduction = function(m0, m1) {
   library(tidyverse)

### 2022-06-21 Supplemental-Material-A-Annotated-Analysis.Rmd
initial amelia implementation
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 TIDout <- gsub("t_", "", dtS1RedInfo$colNamesOut) %>% as.numeric
+TIDInRed <- gsub("t_", "", dtS1RedInfo$colNamesIn) %>% as.numeric
+TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
 
@@
     !PID %in% PIDout,
-    !TIDnum %in% TIDout
-  ) %>%
-  mutate(TIDnum = as.numeric(factor(TIDnum)))
-rm(PIDout, TIDout)
+    TIDnum %in% TIDIn
+  ) %>% 
+  select(TIDnum) %>%
+  mutate(TIDnum = TIDnum - min(TIDnum))
+rm(PIDout, TIDout, TIDInRed, TIDIn)
 
@@
   booktabs = TRUE,
-  caption = "Study 1: FinalData Availability" # complete caption for main document
+  caption = "Study 1: Final Data Availability" # complete caption for main document
 ) %>%
@@
 TIDout <- gsub("t_", "", dtS2RedInfo$colNamesOut) %>% as.numeric
+TIDInRed <- gsub("t_", "", dtS2RedInfo$colNamesIn) %>% as.numeric
+TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
 
@@
 
-```{}
+```{r MiMain3mpca}
+library(Amelia)
+
+dtMain3mpcaMI<- dtMain3mpca %>%
+  select(
+    -date,
+    -TID,
+    -week
+  ) %>%
+  as.data.frame
+
+a.out <- amelia(dtMain3mpcaMI, m = 20, ts = "TIDnum", idvars = c("ID", "PID", "study"))
+
+```
 

### 2022-06-27 Supplemental-Material-A-Annotated-Analysis.Rmd
munge clean up

@@
   "readxl", 
-  "reshape"
+  "reshape",
+  "factoextra"
 )
@@
 
-### Worker
+### Study 1
 
@@
 
+dtS1$clean <- df
+
+# clean up
+rm(df)
+
+# Export reduced Data
+# write.csv(dtS1$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
+# save(dtS1$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
+```
+
+### Study 2
+
+For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
+We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
+We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement).
+We also reformat missing values and format core ID variables.
+<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS2_' list.
+
+```{r cleanStudents}
+# our own test IDs
+ownIDs <- c(
+  "beautifulLionfishXXXR5rcgVBzGu8hPvOqrK8UBJBw4owvi9nfRFSFu3lMzYhE",
+  "niceDogoXXXmB8JI5SFu78SF3DVof84mGUPPNUr14p2HYFTtp31a6D1OwAzM6F-K",
+  "amusedQuailXXXmhuc_fpTp8vPkMwDH1BzjaH1d1kHSO1bsPEfsnaEYk4WeVBfPi",
+  "juwGAbtXX0_1kmZtSVqKh3PGaHOICqUyU4iBkrT3nDsI_uifuD1gzKcZerxaM5FL"
+)
+
+# Prepare dfs for Cleaning
+df.pre <- dtS2$raw.pre %>%
+  mutate_all(na_if, "") %>%
+  mutate_all(na_if, "NA") %>%
+  filter(!is.na(ended)) %>% # remove all who did not finish
+  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
+
+df.post <- dtS2$raw.post %>%
+  mutate_all(na_if, "") %>%
+  mutate_all(na_if, "NA") %>%
+  filter(!is.na(ResponseId)) %>% # remove own test runs
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
+  filter(!is.na(ended)) %>% # remove all who never finished
+  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+names(df.post) <- paste(names(df.post), "post", sep = ".")
+
+df.daily <- dtS2$raw.daily %>%
+  mutate_all(na_if, "") %>%
+  mutate_all(na_if, "NA") %>%
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
+  filter(!is.na(ended)) %>% # remove all who never finished
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+
+# merge daily with pre
+dfPreDaily <- merge(
+  x = df.daily,
+  y = df.pre,
+  by.x = "ResponseId",
+  by.y = "ResponseId.pre", 
+  suffixes = c(".daily", ".pre"),
+  all = FALSE
+)
+
+# merge daily with post
+dfCombined <- merge(
+  x = dfPreDaily,
+  y = df.post,
+  by.x = "ResponseId",
+  by.y = "ResponseId.post", 
+  suffixes = c(".pre", ".post"),
+  all = FALSE
+)
+
+# add to list
+dtS2$clean <- dfCombined
+
+# clean up workspace
+rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
+```
+
+### Study 3
+
+For the medical professionals sample data was, again, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
+We combine the three individual surveys into one large dataframe.
+We exclude our own test data.
+We also reformat missing values and format core ID variables.
+<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS3_' list.
+
+```{r cleanMedical}
+# our own test IDs
+ownIDs <- c(
+  "test_LeonieXXXSklxecPLW0-FBPM4796o3pUwUhAY5jb9KGw8jQsKxWmGpa1Jiy", 
+  "test_MaxXXXtOp_5dTNefIq0yKXtXt2IN6eEKxeHoPY9mlyvdsqPpLp1B0NGg4UL",
+  "test_JannisXXXBsNqk62fOpX6chbd2tMWPptUdjjnhAqnQ3uBqckZ7gLIEoPlfZ",
+  "quaintLeopardCatXXXAJ9cfSj-_SZLwNwMDxv_xv_iyr1Bg5YFLTlYdrjW0UXZY",
+  "blue-eyedIndianElephantXXXLf5zPMpQCDGS3umFzIj-YVky7ivTItvvozW49m"
+)
+
+# Prepare dfs for Cleaning
+df.pre <- dtS3$raw.pre %>%
+  mutate_all(na_if, "") %>%
+  mutate_all(na_if, "NA") %>%
+  filter(!is.na(ended)) %>% # remove all who did not finish
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
+
+df.post <- dtS3$raw.post %>%
+  mutate_all(na_if, "") %>%
+  mutate_all(na_if, "NA") %>% 
+  filter(!is.na(ResponseId)) %>% # remove own test runs
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
+  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
+  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+names(df.post) <- paste(names(df.post), "post", sep = ".")
+
+df.daily <- dtS3$raw.daily %>%
+  mutate_all(na_if, "") %>%
+  mutate_all(na_if, "NA") %>%
+  filter(!ResponseId %in% ownIDs) %>% # remove our own test
+  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
+  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
+  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
+
+# merge daily with pre
+dfPreDaily <- merge(
+  x = df.daily,
+  y = df.pre,
+  by.x = "ResponseId",
+  by.y = "ResponseId.pre", 
+  suffixes = c(".daily", ".pre"),
+  all = F
+)
+
+# merge daily with post
+dfCombined <- merge(
+  x = dfPreDaily,
+  y = df.post,
+  by.x = "ResponseId",
+  by.y = "ResponseId.post", 
+  suffixes = c(".pre", ".post"),
+  all = F
+)
+
+# add to list
+dtS3$clean <- dfCombined
+
+# clean up workspace
+rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
+```
+
+## Calculate needed transformations
+
+### Study 1
+
+For the worker sample, the data transformation stage had three main aims:
+
+1.
+We first corrected time indicators within the surveys.
+In some cases participants completed their daily diary surveys for the afternoon after midnight.
+In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.
+2.
+We then created indices of scales.
+Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).
+3. Finally, we calculated several basic participant summaries (averages across all measurement occasions).
+
+```{r newVarsWorkers}
+df <- dtS1$clean
+
+# Correct date_period ()
 df <- df %>%
   mutate(
-    PID = as.numeric(factor(last_outside_referrer)),
-    # participant ID
+    startDate = as.Date(created),
+    startTime = format(as.POSIXct(created), format = "%H:%M:%S")
+  )
+
+# order time
+# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
+# df$TIDnum <- as.numeric(df$TID) # get numeric TID
+df <- df %>%
+  mutate(
+    PID = as.numeric(factor(last_outside_referrer))
+  )
+
+# Time and Date Variables
+df <- df %>%
+  mutate(
     TID = measureID - 1,
@@
     ),
+    daycorTest = ifelse(
+      daytime == "morning" &
+        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
+        daytime == "afternoon" &
+          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
+      1,
+      0
+    ),
     # correctly identify which date the questionnaire is about
@@
 
-dtS1$clean <- df
-
-# clean up
-rm(df)
-
-# Export reduced Data
-# write.csv(dtS1$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
-# save(dtS1$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
-```
-
-### Student
-
-For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement).
-We also reformat missing values and format core ID variables.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS2_' list.
-
-```{r cleanStudents}
-# our own test IDs
-ownIDs <- c(
-  "beautifulLionfishXXXR5rcgVBzGu8hPvOqrK8UBJBw4owvi9nfRFSFu3lMzYhE",
-  "niceDogoXXXmB8JI5SFu78SF3DVof84mGUPPNUr14p2HYFTtp31a6D1OwAzM6F-K",
-  "amusedQuailXXXmhuc_fpTp8vPkMwDH1BzjaH1d1kHSO1bsPEfsnaEYk4WeVBfPi",
-  "juwGAbtXX0_1kmZtSVqKh3PGaHOICqUyU4iBkrT3nDsI_uifuD1gzKcZerxaM5FL"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS2$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS2$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS2$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = FALSE
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = FALSE
-)
-
-# add to list
-dtS2$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-### Medical
-
-For the medical professionals sample data was, again, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe.
-We exclude our own test data.
-We also reformat missing values and format core ID variables.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS3_' list.
-
-```{r cleanMedical}
-# our own test IDs
-ownIDs <- c(
-  "test_LeonieXXXSklxecPLW0-FBPM4796o3pUwUhAY5jb9KGw8jQsKxWmGpa1Jiy", 
-  "test_MaxXXXtOp_5dTNefIq0yKXtXt2IN6eEKxeHoPY9mlyvdsqPpLp1B0NGg4UL",
-  "test_JannisXXXBsNqk62fOpX6chbd2tMWPptUdjjnhAqnQ3uBqckZ7gLIEoPlfZ",
-  "quaintLeopardCatXXXAJ9cfSj-_SZLwNwMDxv_xv_iyr1Bg5YFLTlYdrjW0UXZY",
-  "blue-eyedIndianElephantXXXLf5zPMpQCDGS3umFzIj-YVky7ivTItvvozW49m"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS3$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS3$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>% 
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS3$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = F
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = F
-)
-
-# add to list
-dtS3$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-## Calculate needed transformations
-
-### Worker
-
-For the worker sample, the data transformation stage had three main aims:
-
-1.
-We first corrected time indicators within the surveys.
-In some cases participants completed their daily diary surveys for the afternoon after midnight.
-In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).
-3. Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsWorkers}
-df <- dtS1$clean
-
-# Time and Date Variables
 # remove seconds from afternoon time
@@
 
-# summarize by participant (check that everything is within pp might not be the case for )
-between <- df %>%
-  group_by(last_outside_referrer) %>%
+#
+df <- df %>%
   mutate(
-    CtContactNL = sum(IntergroupContact),
-    CtContactNonNl = sum(IngroupContact),
-    CtContactNLAll = sum(ContactNum),
-    CtContactNonNlAll = sum(NonDutchNum),
-    AvKeyNeed = mean(keyMotiveFulfilled, na.rm = T),
-    AvKeyNeedInt = mean(KeyNeedFulfillment, na.rm = T),
-    AvKeyNeedNoInt = mean(DaytimeNeedFulfillment, na.rm = T),
-    AvAutonomy = mean(autonomy, na.rm = T),
-    AvCompetence = mean(competence, na.rm = T),
-    AvRelatedness = mean(relatedness, na.rm = T),
-    AvThermo = mean(AttitudesDutch, na.rm = T),
-    AvWB = mean(exWB, na.rm = T)
-  ) %>%
-  ungroup() %>%
-  mutate(
-    CtContactNL_c = scale(CtContactNL, scale = FALSE),
-    AvKeyNeedInt_c = scale(AvKeyNeedInt, scale = FALSE),
-    AvKeyNeed_c = scale(AvKeyNeed, scale = FALSE),
-    CtContactNL_z = scale(CtContactNL, scale = TRUE),
-    AvKeyNeedInt_z = scale(AvKeyNeedInt, scale = TRUE),
-    AvKeyNeed_z = scale(AvKeyNeed, scale = TRUE)
-  )
-
-warning(
-  "some variable transformations (esp. _c and _z) might be across all participants (i.e., not within PP). See next step."
-)
-
-dtS1$full <- between
-rm(df, between)
-
-# dataframe where interaction types are recoded
-workerInteractionType <- dtS1$full %>%
-  mutate(
-    OutgroupInteraction = as_factor(IntergroupContact),
-    NonOutgroupInteraction = as_factor(IngroupContact)
-  )
-
-# Create variables centered and standardized within Participant
-# i.e., divide into trait and state
-workerWithinBetween <-
-  MlTraitState(
-    data = workerInteractionType,
-    id = "PID",
-    selection =
-      c(
-        "keyMotiveFulfilled",
-        "competence",
-        "autonomy",
-        "relatedness",
-        "AttitudesDutch",
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatednessNoInteraction", 
-        "qualityOverall", 
-        "OutgroupInteraction",
-        "NonOutgroupInteraction"
-      )
-  )
-
-workerOutWithinBetween <-
-  MlTraitState(
-    data = workerInteractionType %>% filter(OutgroupInteraction == "Yes"),
-    id = "PID",
-    selection =
-      c(
-        "keyMotiveFulfilled",
-        "AttitudesDutch",
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatedness", 
-        "qualityOverall"
-      )
-  )
-
-# Between participants contact frequency
-workerContactFreq <- dtS1$full %>%
-  group_by(PID) %>%
-  summarise(
-    n = n(),
-    SumContactNL = sum(IntergroupContact),
-    PercContactNL = SumContactNL / n * 100,
-    SumContactNLAll = sum(number),
-    AvAttitude = mean(AttitudesDutch, na.rm = T)
-  ) %>%
-  mutate(
-    WinSumContactNL = DescTools::Winsorize(SumContactNL),
-    WinSumContactNLAll = DescTools::Winsorize(SumContactNLAll)
-  )
-
-# save cleaned data
-# save(df.btw, file = "data/processed/df.btw.RData")
-# write_sav(df.btw, "data/processed/MT_clean-merged_pre-post.sav")
-
-# export data to Mplus
-# df.mplus = remove_all_labels(select(df,
-#                                     PID, ResponseId,
-#                                     thermometerDutch_1, inNonDutch, Contact_dum,
-#                                     keyMotiveFulfilled, autonomy.daily.all, competence.daily.all, relatedness.daily.all))
-# names(df.mplus)= c("PID", "ResponseId", "att", "intin", "intout", "keymot", "aut", "comp", "rel")
-# mplus = df.mplus[order(df.mplus$PID, df.mplus$ResponseId),]
-# mplus.intcont = mplus[mplus$intout==1,]
-# prepareMplusData(mplus.intcont, "data/processed/dynamic-subset-intonly.dat")
-```
-
-### Student
+    exWB = scales::rescale(exWB, from = range(exWB, na.rm = TRUE), to = c(-100, 100)),
+    relatedness = relatedness + 50,
+    autonomy = autonomy + 50,
+    competence = competence + 50,
+    KeyNeedFulfillment = KeyNeedFulfillment + 50,
+    KeyNeedDueToPartner = KeyNeedDueToPartner + 50,
+    InteractionContextAccidental = InteractionContextAccidental + 50,
+    InteractionContextvoluntary = InteractionContextvoluntary + 50,
+    InteractionContextCooperative = InteractionContextCooperative + 50,
+    InteractionContextRepresentativeNL = InteractionContextRepresentativeNL + 50,
+    qualityOverall = qualityOverall + 50,
+    qualityMeaning = qualityMeaning + 50,
+    DaytimeNeedFulfillment = DaytimeNeedFulfillment + 50
+  )
+dtS1$full <- df
+rm(df)
+```
+
+### Study 2
 
@@
 
+# Correct date_period ()
+df <- df %>%
+  mutate(
+    startDate = as.Date(created),
+    startTime = format(as.POSIXct(created), format = "%H:%M:%S"),
+    date_period_old = date_period,
+    date_period = ifelse(
+      startTime >= c("12:00") & startTime <= c("18:59"),
+      paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Morning"),
+      ifelse(
+        startTime >= c("19:00") & startTime <= c("23:59"),
+        paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Afternoon"),
+        ifelse(
+          startTime >= c("00:00") & startTime <= c("11:59"),
+          paste(as.Date(created) - 1, "Afternoon"),
+          NA
+        )
+      )
+    )
+  )
+
 # order time
-df$TID <- factor(df$date_period, levels = unique(dtS2$raw.daily$date_period))
-df$TIDnum <- as.numeric(df$TID) # get numeric TID
+# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
+# df$TIDnum <- as.numeric(df$TID) # get numeric TID
+df <- df %>%
+  mutate(
+    dateOrder = stri_replace_all_regex(
+      date_period,
+      pattern = c(' Morning', ' Afternoon'),
+      replacement = c('_A', '_B'),
+      vectorize = FALSE
+    ),
+    TID = factor(date_period),
+    TIDnum = as.numeric(factor(dateOrder))
+  )
 
@@
 df <- df %>%
-  arrange(PID, TID) # %>%
-# View()
+  arrange(PID, TID)
 
@@
 
-# Pro-Sociality
-df$ProSo <-
-  rowMeans(df[, c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm = T)
-# Anti-Sociality
-df$AntiSo <-
-  rowMeans(df[, c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm = T)
+# Randomly selected variables
+df <- df %>%
+  rowwise() %>%
+  mutate(
+    ProSo = mean(c(ProSo1, ProSo2, ProSo3, ProSo4), na.rm = TRUE),
+    AntiSo = mean(c(AntiSo1, AntiSo2, AntiSo3, AntiSo4, AntiSo5, AntiSo6, AntiSo7), na.rm = TRUE)
+  )
 
@@
 
-# -------------------------------------------------------------------------------------------------------------
-#                                 Center Relevant Variables
-# -------------------------------------------------------------------------------------------------------------
-
-df <- df %>%
-  group_by(PID) %>%
-  mutate(
-    KeyNeedFulfillment.cm = mean(KeyNeedFulfillment, na.rm = TRUE),
-    # cluster mean (mean of PP)
-    KeyNeedFulfillment.cwc = KeyNeedFulfillment - KeyNeedFulfillment.cm,
-    # cluster mean centered (within PP centered)
-    closeness.cm = mean(closeness_Calc, na.rm = TRUE),
-    closeness.cwc = closeness_Calc - closeness.cm
-  ) %>%
-  ungroup()
-
-# store
 dtS2$full <- df
 rm(df)
-
-# Between participants contact frequency
-studentContactFreq <- dtS2$full %>%
-  group_by(PID) %>%
-  summarise(
-    n = n(),
-    SumContactNL = sum(InteractionDumDutch),
-    PercContactNL = SumContactNL / n * 100,
-    SumContactNLAll = sum(ContactNum[InteractionDumDutch == 1], na.rm = TRUE),
-    AvAttitude = mean(AttitudesDutch, na.rm = TRUE),
-    AvQuality = mean(qualityOverall, na.rm = TRUE)
-  ) %>%
-  mutate(
-    WinSumContactNL = DescTools::Winsorize(SumContactNL),
-    WinSumContactNLAll = DescTools::Winsorize(SumContactNLAll)
-  )
-
-# dataframe where interaction types are recoded
-studentInteractionType <- dtS2$full %>%
-  mutate(
-    NonDutchContact = tidyr::replace_na(NonDutchContact, 2), # make second non-Dutch countable
-    NonDutchContact = NonDutchContact*-1+2 # recode (yes = 1 -> 1, no = 2 -> 0)
-  ) %>%
-  mutate(
-    NonDutchNum = ifelse(NonDutchContact == 0, 0, NonDutchNum), # measurement error --- only measured when NonDutchContact==1
-    OutgroupInteraction = factor(
-      InteractionDumDutch,
-      levels = c(0, 1),
-      labels = c("No", "Yes")
-    ),
-    NonOutgroupInteraction = factor(
-      rowSums(select(., c(InteractionDumNonDutch, NonDutchContact))), # combine the two non-Dutch Q.,
-      levels = c(0, 1),
-      labels = c("No", "Yes")
-    )
-  )
-
-# select a subset of IDs to display in plots
-studentPltIDs <-
-  studentInteractionType %>%
-  group_by(PID) %>%
-  summarise(n = n()) %>%
-  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
-  select(PID) %>%
-  as.matrix %>%
-  as.vector
-
-# select a subset of IDs to display in plots (only outgroup interactions)
-studentOutPltIDs <-
-  studentInteractionType %>%
-  filter(OutgroupInteraction == "Yes") %>%
-  group_by(PID) %>%
-  summarise(n = n()) %>%
-  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
-  select(PID) %>%
-  as.matrix %>%
-  as.vector
-
-# Center within and between
-# divide into trait and state
-studentWithinBetween <-
-  MlTraitState(
-    data = studentInteractionType,
-    id = "PID",
-    selection =
-      c(
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatedness",
-        "AttitudesDutch",
-        "qualityOverall",
-        "OutgroupInteraction",
-        "NonOutgroupInteraction"
-      )
-  )
-studentOutWithinBetween <-
-  MlTraitState(
-    data = studentInteractionType %>% filter(OutgroupInteraction == "Yes"),
-    id = "PID",
-    selection =
-      c(
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatedness",
-        "AttitudesDutch",
-        "qualityOverall"
-      )
-  )
-
-```
-
-### Medical
+```
+
+### Study 3
 
@@
 
+# Correct date_period ()
+df <- df %>%
+  mutate(
+    startDate = as.Date(created),
+    startTime = format(as.POSIXct(created), format = "%H:%M:%S"),
+    date_period_old = date_period,
+    date_period = ifelse(
+      startTime >= c("12:00") & startTime <= c("18:59"),
+      paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Morning"),
+      ifelse(
+        startTime >= c("19:00") & startTime <= c("23:59"),
+        paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Afternoon"),
+        ifelse(
+          startTime >= c("00:00") & startTime <= c("11:59"),
+          paste(as.Date(created) - 1, "Afternoon"),
+          NA
+        )
+      )
+    )
+  )
+
 # order time
-df$TID <-
-  factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-df$TIDnum <- as.numeric(df$TID) # get numeric TID
+# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
+# df$TIDnum <- as.numeric(df$TID) # get numeric TID
+df <- df %>%
+  mutate(
+    dateOrder = stri_replace_all_regex(
+      date_period,
+      pattern = c(' Morning', ' Afternoon'),
+      replacement = c('_A', '_B'),
+      vectorize = FALSE
+    ),
+    TID = factor(date_period),
+    TIDnum = as.numeric(factor(dateOrder))
+  )
 
@@
 
-# Pro-Sociality
-df$ProSo <-
-  rowMeans(df[, c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm = TRUE)
-# Anti-Sociality
-df$AntiSo <-
-  rowMeans(df[, c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm = TRUE)
+# Randomly selected variables
+df <- df %>%
+  rowwise() %>%
+  mutate(
+    ProSo = mean(c(ProSo1, ProSo2, ProSo3, ProSo4), na.rm = TRUE),
+    AntiSo = mean(c(AntiSo1, AntiSo2, AntiSo3, AntiSo4, AntiSo5, AntiSo6, AntiSo7), na.rm = TRUE),
+    agency = mean(c(agency1, agency2, agency3), na.rm = TRUE),
+    autoFrust = mean(c(autoFrust1, autoFrust2, autoFrust3, autoFrust4), na.rm = TRUE),
+    autoSat = mean(c(autoSat1, autoSat2, autoSat3, autoSat4), na.rm = TRUE),
+    relatFrust = mean(c(relatFrust1, relatFrust2, relatFrust3, relatFrust4), na.rm = TRUE),
+    relatSat = mean(c(relatSat1, relatSat2, relatSat3, relatSat4), na.rm = TRUE),
+    compFrust = mean(c(compFrust1, compFrust2, compFrust3, compFrust4), na.rm = TRUE),
+    compSat = mean(c(compSat1, compSat2, compSat3, compSat4), na.rm = TRUE),
+    lonely = mean(c(lonely4, lonely4, lonely4, lonely4), na.rm = TRUE),
+    emotRegPos = mean(c(emotRegPos01, emotRegPos02), na.rm = TRUE),
+    emotRegNeg = mean(c(emotRegNeg01, emotRegNeg02), na.rm = TRUE)
+  )
 
@@
   drop_na %>%
-  melt(
+  reshape2::melt(
     ., 
@@
 
-# -------------------------------------------------------------------------------------------------------------
-#                                 Center Relevant Variables
-# -------------------------------------------------------------------------------------------------------------
-# divide into trait and state
-medicalOutWithinBetween <-
-  MlTraitState(
-    data = df %>% filter(OutgroupInteraction == "Yes"),
-    id = "PID",
-    selection =
-      c(
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatedness",
-        "AllportsCondition",
-        "AttitudesDutch",
-        "qualityOverall"
-      )
-  )
-
-medicalWithinBetween <-
-  MlTraitState(
-    data = df,
-    id = "PID",
-    selection =
-      c(
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatedness",
-        "AllportsCondition",
-        "AttitudesDutch",
-        "qualityOverall",
-        "OutgroupInteraction",
-        "NonOutgroupInteraction"
-      )
-  )
-
-df <- # keep only for compatibility of old framgents
-  MlTraitState(
-    data = df,
-    id = "PID",
-    selection =
-      c(
-        "KeyNeedFulfillment",
-        "competence",
-        "autonomy",
-        "relatedness",
-        "AllportsCondition",
-        "AttitudesDutch",
-        "qualityOverall"
-      )
-  )
-
-# store
 dtS3$full <- df
-rm(df)
-
-# Between participants contact frequency
-medicalContactFreq <- 
-  dtS3$full %>%
-  group_by(PID) %>%
-  summarise(
-    n = n(),
-    SumContactNL = sum(InteractionDumDutch, na.rm = TRUE),
-    PercContactNL = SumContactNL / n * 100,
-    SumContactNLAll = sum(ContactNum[InteractionDumDutch == 1], na.rm = TRUE),
-    AvAttitude = mean(AttitudesDutch, na.rm = TRUE),
-    AvQuality = mean(qualityOverall, na.rm = TRUE)
-  ) %>%
-  mutate(
-    WinSumContactNL = DescTools::Winsorize(SumContactNL),
-    WinSumContactNLAll = DescTools::Winsorize(SumContactNLAll)
-  )
-
-# select a subset of IDs to display in plots
-medicalPltIDs <-
-  dtS3$full %>%
-  group_by(PID) %>%
-  summarise(n = n()) %>%
-  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
-  select(PID) %>%
-  as.matrix %>%
-  as.vector
-
-# select a subset of IDs to display in plots (only outgroup interactions)
-medicalOutPltIDs <-
-  dtS3$full %>%
-  filter(OutgroupInteraction == "Yes") %>%
-  group_by(PID) %>%
-  summarise(n = n()) %>%
-  slice_max(n, n = 20) %>% # chose the 20 with the most number of measurements
-  select(PID) %>%
-  as.matrix %>%
-  as.vector
 ```
@@
   ) %>% 
-  select(TIDnum) %>%
   mutate(TIDnum = TIDnum - min(TIDnum))
 rm(PIDout, TIDout, TIDInRed, TIDIn)
+
+# change glitch where survey was sent too early
+dtS1Red$TIDnum[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- 10
+dtS1Red$TID[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- "2018-05-18 Morning"
 
@@
   ) %>%
+  #na.omit %>%
   arrange(PID, TIDnum) %>%
@@
   ) %>% 
-  select(TIDnum) %>%
   mutate(TIDnum = TIDnum - min(TIDnum))
 rm(PIDout, TIDout, TIDInRed, TIDIn)
+
+# remove glitch entry 
+dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 57 & dtS2Red$ended == "2018-12-20 21:58:31"),]
 
@@
   mutate(data = 1)
-dtS3Availability <- reshape::cast(dtS3Availability, PID ~ TIDnum) %>%
+dtS3Availability <- reshape::cast(dtS3Availability, PID ~ TIDnum, fill = 0) %>%
   select(-PID)
@@
 TIDout <- gsub("t_", "", dtS3RedInfo$colNamesOut) %>% as.numeric
+TIDInRed <- gsub("t_", "", dtS3RedInfo$colNamesIn) %>% as.numeric
+TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
 
@@
     !PID %in% PIDout,
-    !TIDnum %in% TIDout
-  ) %>%
-  mutate(TIDnum = as.numeric(factor(TIDnum)))
-rm(PIDout, TIDout)
+    TIDnum %in% TIDIn
+  ) %>% 
+  mutate(TIDnum = TIDnum - min(TIDnum))
+rm(PIDout, TIDout, TIDInRed, TIDIn)
 
@@
   data.frame(varNam = Reduce(intersect, list(
-    names(dtS1Red), names(dtS3Red), names(dtS3Red)
+    names(dtS1Red), names(dtS2Red), names(dtS3Red)
   ))) %>%
@@
   "NonDutchNum",
+  "groupDutch",
   # Interaction Descriptives (non-continuous)
@@
 
+### Studies 1-3
+
 ```{r MiMain3mpca}
@@
 
-dtMain3mpcaMI<- dtMain3mpca %>%
+varNamInMiMain <- varNames %>%
+  filter(
+    aux == 1,
+    studyS1 == "S1",
+    studyS2 == "S2",
+    studyS3 == "S3"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+varNamInMiMainPsbl <- c(
+  paste(varNamInMiMain, rep(".pre", length(varNamInMiMain)), sep = ""),
+  varNamInMiMain,
+  paste(varNamInMiMain, rep(".post", length(varNamInMiMain)), sep = "")
+)
+
+varNamInMiMainRed <- Reduce(
+  intersect,
+  list(
+    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
+    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
+    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names
+  )
+)
+
+varNamMiMainNoms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS1 == "S1",
+    studyS2 == "S2",
+    studyS3 == "S3"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+dtMiMain <- rbind(
+  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% mutate(across(!TID &
+                                                                                             !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  select(ID,
+         study,
+         date,
+         week,
+         everything(),
+         -PID,
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>%
+  select_if( ~ sum(!is.na(.)) > 1) # only include variables that have any data (i.e., not all NA)
+
+MiMain <-
+  amelia(
+    dtMiMain,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("TID", "date", "week", "study"), #"PID", 
+    noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMain, file = "data/imputed/MainImputed.RData")
+```
+
+### Study 1
+
+```{r MiS1}
+varNamInMiS1 <- varNames %>%
+  filter(
+    aux == 1,
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+varNamInMiS1Psbl <- c(
+  paste(varNamInMiS1, rep(".pre", length(varNamInMiS1)), sep = ""),
+  varNamInMiS1,
+  paste(varNamInMiS1, rep(".post", length(varNamInMiS1)), sep = "")
+)
+
+varNamdtS1Noms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+dtMiS1 <- dtS1Red %>%
   select(
-    -date,
-    -TID,
-    -week
-  ) %>%
-  as.data.frame
-
-a.out <- amelia(dtMain3mpcaMI, m = 20, ts = "TIDnum", idvars = c("ID", "PID", "study"))
-
-```
+    PID,
+    TID,
+    TIDnum,
+    date,
+    week,
+    any_of(varNamInMiS1Psbl),
+    -time,
+    -starts_with("ResponseId")
+  ) %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+    ) %>%
+  as.data.frame %>%
+  select_if(~sum(!is.na(.)) > 1) #%>% # only include variables that have any data (i.e., not all NA)
+
+  which(!is.numeric(dtMiS1) | sapply(dtMiS1, sd, na.rm=TRUE)>0)
+
+  select_if(~sd(., na.rm = TRUE) > 0) # only include variables that vary (i.e., sd != 0)
+
+#dtS1MiCs <- names(dtMiS1)[endsWith(names(dtMiS1), c(".pre", ".post"))] # cross sectional vars
+
+MiS1 <-
+  amelia(
+    dtMiS1,
+    m = 2,
+    ts = "TIDnum",
+    cs = "PID",
+    idvars =  c("TID", "date", "week"),
+    noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+write.amelia(MiS1, file.stem = "data/imputed/S1Imputed.Amelia")
+```
+
+### Study 2
+
+### Study 3
 
@@
 
+```{r varNames}
+# varNames
+varNamInS123 <- varNames %>%
+  filter(
+    pca != 0,
+    surveyS1 == "daily",
+    surveyS2 == "daily",
+    surveyS3 == "daily"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamInS1 <- varNames %>%
+  filter(
+    pca != 0,
+    surveyS1 == "daily"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamInS2 <- varNames %>%
+  filter(
+    pca != 0,
+    surveyS2 == "daily"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamInS3 <- varNames %>%
+  filter(
+    pca != 0,
+    surveyS3 == "daily"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamAdd <- c(
+  "PID",
+  "TID",
+  "TIDnum",
+  "relatedness",
+  "autonomy",
+  "competence"
+)
+varNamOut <- c(
+  "ResponseId",
+  "relatednessNoInteraction",
+  "relatednessSelf",
+  "relatednessOther",
+  "autonomy_Int",
+  "autonomy_NoInt",
+  "competence_Int",
+  "competence_NoInt",
+  varNamInS1[grepl('^MDMQ', varNamInS1)],
+  varNamInS2[grepl('^ProSo|^AntiSo', varNamInS2)],
+  varNamInS3[grepl(
+    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
+    varNamInS3
+  )]
+)
+varNamNewS1 <- c(
+  "alertness", 
+  "calmness", 
+  "valence"
+)
+varNamNewS2 <- c(
+  "ProSo", 
+  "AntiSo"
+)
+varNamNewS3 <- c(
+  "ProSo", 
+  "AntiSo",
+  "agency",
+  "autoFrust",
+  "autoSat",
+  "relatFrust",
+  "relatSat",
+  "compFrust",
+  "compSat",
+  "lonely",
+  "emotRegPos",
+  "emotRegNeg"
+)
+varNamIntDep <- c(
+  varNamInS123[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamInS123)]
+)
+
+varNamPcaS123 <- c(varNamAdd, varNamInS123[!varNamInS123 %in% varNamOut])
+varNamPcaS123Indep <- varNamPcaS123[!varNamPcaS123 %in% varNamIntDep]
+varNamPcaS1 <- c(varNamAdd, varNamNewS1, varNamInS1[!varNamInS1 %in% varNamOut])
+varNamPcaS2 <- c(varNamAdd, varNamNewS2, varNamInS2[!varNamInS2 %in% varNamOut])
+varNamPcaS3 <- c(varNamAdd, varNamNewS3, varNamInS3[!varNamInS3 %in% varNamOut])
+
+idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
+```
+
 ```{r longToWide}
-
-dtMain3mpcaWideGmc <- dtMain3mpcaNorm %>% 
+dtPcaS123 <- rbind(
+  dtS1Red %>% select(any_of(varNamPcaS123)) %>% mutate(study = "S1"),
+  dtS2Red %>% select(any_of(varNamPcaS123)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamPcaS123)) %>% mutate(study = "S3")
+  ) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+    ) %>%
+  select(
+    ID,
+    PID,
+    TID,
+    date,
+    week,
+    TIDnum,
+    study,
+    everything()
+  ) %>%
+  arrange(ID, TIDnum) %>% 
+  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
+
+# test <- PCAnorm(data = dtPcaS123,
+#           pid = "ID",
+#           tid = "TIDnum",
+#           selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
+#   select(
+#     ID,
+#     PID,
+#     TIDnum,
+#     study,
+#     ends_with("_gmc")
+#   ) %>%
+#   data.frame %>%
+#   melt(
+#     .,
+#     id=c("ID", "PID", "TIDnum", "study")
+#   ) %>%
+#   mutate(
+#     variable = paste(gsub("_gmc", "", variable), TIDnum, sep = "_")
+#   ) %>%
+#   #select(-TIDnum) %>% 
+#   arrange(variable, ID) %>%
+#   mutate(
+#     idVar = paste(ID, variable, sep = "_")
+#   ) 
+# plyr::join(test, data.frame(table(idVar = test$idVar)))
+
+test <- PCAnorm(data = dtPcaS123,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>%
+  select(
+    ID,
+    PID,
+    TIDnum,
+    study,
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>% 
+  melt(
+    .,
+    id=c("ID", "PID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste(gsub("_gmc", "", variable), stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+  ) %>%
+  select(-TIDnum) %>%
+  arrange(variable, ID) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  replace(is.na(.), 0) %>%
+  select(-ID, -PID, -study)
+  #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
+  #pivot_wider(names_from = variable, values_from = value)
+
+test2 <- dtPcaS123 %>%
+  select(
+    -TID,
+    -date,
+    -week
+  ) %>%
+  data.frame %>% 
+  melt(
+    .,
+    id=c("ID", "PID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+  ) %>%
+  select(-TIDnum) %>%
+  arrange(variable, ID) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  replace(is.na(.), 0) %>%
+  select(-ID, -PID, -study)
+
+```
+
+## Three-Way ANOVA
+
+```{r threeWayAnova}
+library(ThreeWay)
+threeWayAno <- threewayanova(
+  Y = test2,
+  n = nrow(test2), # Nr. ppt.s (A)
+  m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
+  p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
+)
+
+```
+
+# **Plots per Item** {.tabset}
+
+The three-mode PCA ultimately aims to explain the most variance with the fewest components (by maximizing the variance of data mapped in a lower dimensional space --- often by maximally separating groups of cases with similar variable patterns).
+Importantly, this procedure depends on variances and covariances to be present within the data.
+As one of our three modes is time and variance over time is not always given (e.g., because a variable is stable over the measured time span, such as personality over a month), it becomes important to assess whether a dimension reduction is possible for the variables measured during the measured time period.
+We, thus, start by visually inspecting the average changes and variance developments of all potentially relevant variables.
+We hope so see changes within the means and standard deviations over the measurement periods.
+
+## Means
+
+### Studies 1-3 (Means)
+
+```{r plotAllMean}
+varClS123 <- dtPcaS123 %>%
+  PCAnorm(data = .,
+          pid = "PID",
+          tid = "TIDnum",
+          selection = names(.)[!names(.) %in% idVars]) %>%
+  select(
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>%
+  summarise_all(
+    sd, na.rm = TRUE
+  ) %>%
+  t %>% 
+  as.data.frame %>%
+  kmeans(., centers = 2)
+
+PCAnorm(data = dtPcaS123,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
   select(
@@
   ) %>%
-  reshape2::melt(.,
-              id.vars = c("ID", "TIDnum")) %>%
-  mutate(variable = paste(variable, TIDnum, sep = "_")) %>%
-  select(-TIDnum) %>% 
-  arrange(variable, ID) %>%
-  reshape(., idvar = "ID", timevar = "variable", direction = "wide")
-  #pivot_wider(names_from = variable, values_from = value)
-
-```
-
-## Three-Way ANOVA
-
-```{r threeWayAnova}
-library(ThreeWay)
-# threewayanova(
-#   Y = dtMain3mpcaWideGmc,
-#   n = nrow(dtMain3mpcaWideGmc), # Nr. ppt.s
-#   m = length(unique(dtMain3mpcaNorm$TIDnum)), # Nr. time points
-#   p = length(varNam3mpcaMainAnalysis) # Nr. variables
-# )
-```
-
-## Plot (x = var, y = z, group/line = t)
-
-</br>
-
---------------------------------------------------------------------
-
-</br>
-
-# **Plots per Item**
-
-The three-mode PCA ultimately aims to explain the most variance with the fewest components (by maximizing the variance of data mapped in a lower dimensional space --- often by maximally separating groups of cases with similar variable patterns).
-Importantly, this procedure depends on variances and covariances to be present within the data.
-As one of our three modes is time and variance over time is not always given (e.g., because a variable is stable over the measured time span, such as personality over a month), it becomes important to assess whether a dimension reduction is possible for the variables measured during the measured time period.
-We, thus, start by visually inspecting the average changes and variance developments of all potentially relevant variables.
-We hope so see changes within the means and standard deviations over the measurement period.
-
-## Study 1 (Worker)
+  data.frame %>%
+  melt(
+    .,
+    id=c("ID","TIDnum")
+  ) %>%
+  mutate(
+    variable = gsub("_gmc", "", variable)
+  ) %>%
+  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
+  #geom_jitter() +
+  stat_summary(fun=mean, geom="line") +
+  labs(
+    title = "Studies 1-3 Means over time [bi-daily]",
+    y = "Grand Mean Centered Average",
+    x = "Variable",
+    color = "Timepoint",
+    group = "Timepoint"
+  ) +
+  #coord_flip() +
+  theme_Publication() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust=1),
+    legend.key.size = unit(1, 'cm'),
+    legend.key.height = unit(0.4, 'cm'),
+  )
+
+dtPcaS123 %>%
+  group_by(ID, date, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(date))) %>%
+  ungroup %>%
+  select(ID, date, TIDnum, everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  PCAnorm(data = .,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(.)[!names(.) %in% idVars]) %>% 
+  select(
+    ID,
+    TIDnum,
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>%
+  melt(
+    .,
+    id=c("ID","TIDnum")
+  ) %>%
+  mutate(
+    variable = gsub("_gmc", "", variable)
+  ) %>%
+  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
+  #geom_jitter() +
+  stat_summary(fun=mean, geom="line") +
+  labs(
+    title = "Studies 1-3 Means over time [daily]",
+    y = "Grand Mean Centered Average",
+    x = "Variable",
+    color = "Timepoint",
+    group = "Timepoint"
+  ) +
+  #coord_flip() +
+  theme_Publication() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust=1),
+    legend.key.size = unit(1, 'cm'),
+    legend.key.height = unit(0.4, 'cm'),
+  )
+
+dtPcaS123 %>%
+  group_by(ID, week, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(week))) %>%
+  ungroup %>%
+  select(ID, week, TIDnum, everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  PCAnorm(data = .,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(.)[!names(.) %in% idVars]) %>% 
+  select(
+    ID,
+    TIDnum,
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>%
+  melt(
+    .,
+    id=c("ID","TIDnum")
+  ) %>%
+  mutate(
+    variable = gsub("_gmc", "", variable)
+  ) %>%
+  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
+  #geom_jitter() +
+  stat_summary(fun=mean, geom="line") +
+  labs(
+    title = "Studies 1-3 Means over time [weekly]",
+    y = "Grand Mean Centered Average",
+    x = "Variable",
+    color = "Timepoint",
+    group = "Timepoint"
+  ) +
+  #coord_flip() +
+  theme_Publication() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust=1),
+    legend.key.size = unit(1, 'cm'),
+    legend.key.height = unit(0.4, 'cm'),
+  )
+```
+
+### Study 1 (Means)
 
@@
 
-We first select all potentially relevant variables and their labels, and then group the variables in small groups of one to three variables for plotting.
-
-```{r timeInNL}
-#median(rbind(rbind(dtS1$raw.pre$timeNL_1, dtS2$raw.pre$timeNL), dtS3$raw.pre$timeNL), na.rm=TRUE)
-```
-
-```{r workerDataSelection, eval=FALSE}
-varListWorker <- 
-  c(
-    "IntergroupContact",
-    "InteractionDum",
-    "DaytimeNeedFullfillment",
-    "InteractionNeedFullfillment",
-    "autonomy",
-    "competence",
-    "relatedness",
-    "AttitudesDutch",
-    "AttitudesPartner",
-    "ExWB",
-    "alertness",
-    "calmness",
-    "valence",
-    "MDMQ.v1_1",
-    "MDMQ.v1_2",
-    "MDMQ.v1_3",
-    "MDMQ.v1_4",
-    "MDMQ.v1_5",
-    "MDMQ.v1_6",
-    "MDMQ.v1_7",
-    "MDMQ.v1_8",
-    "MDMQ.v1_9",
-    "MDMQ.v1_10",
-    "MDMQ.v1_11",
-    "MDMQ.v1_12",
-    "MDMQ.v1_13",
-    "MDMQ.v2_1",
-    "MDMQ.v2_2",
-    "MDMQ.v2_3",
-    "MDMQ.v2_4",
-    "MDMQ.v2_5",
-    "MDMQ.v2_6",
-    "MDMQ.v2_7",
-    "MDMQ.v2_8",
-    "MDMQ.v2_9",
-    "MDMQ.v2_10",
-    "MDMQ.v2_11",
-    "MDMQ.v2_12",
-    "MDMQ.v2_13"
-  )
-
-varNameListWorker <- 
-  c(
-    "IntergroupContact" = "Dutch Interaction",
-    "InteractionDum" = "Any Interaction",
-    "DaytimeNeedFullfillment" = "Daytime Core Motive Fulfillment",
-    "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
-    "autonomy" = "Autonomy",
-    "competence" = "Competence",
-    "relatedness" = "Relatedness",
-    "AttitudesDutch" = "Outgroup Attitude",
-    "AttitudesPartner" = "Attitude Interaction Partner",
-    "ExWB" = "Sadness Happiness",
-    "alertness" = "Alertness",
-    "calmness" = "Calmness",
-    "valence" = "Emotional Valence",
-    "MDMQ.v1_1" = "content",
-    "MDMQ.v1_2" = "rested",
-    "MDMQ.v1_3" = "restless",
-    "MDMQ.v1_4" = "bad",
-    "MDMQ.v1_5" = "worn-out",
-    "MDMQ.v1_6" = "composed",
-    "MDMQ.v1_7" = "tired",
-    "MDMQ.v1_8" = "great",
-    "MDMQ.v1_9" = "uneasy",
-    "MDMQ.v1_10" = "energetic",
-    "MDMQ.v1_11" = "uncomfortable",
-    "MDMQ.v1_12" = "relaxed",
-    "MDMQ.v1_13" = "alive and vital",
-    "MDMQ.v2_1" = "sleepy",
-    "MDMQ.v2_2" = "good",
-    "MDMQ.v2_3" = "at ease",
-    "MDMQ.v2_4" = "unhappy",
-    "MDMQ.v2_5" = "alert",
-    "MDMQ.v2_6" = "discontent",
-    "MDMQ.v2_7" = "tense",
-    "MDMQ.v2_8" = "fresh",
-    "MDMQ.v2_9" = "happy",
-    "MDMQ.v2_10" = "nervous",
-    "MDMQ.v2_11" = "exhausted",
-    "MDMQ.v2_12" = "calm",
-    "MDMQ.v2_13" = "alive and vital"
-  )
-
-dtS1$viz <- list()
-
-dtS1$vizall <- dtS1$full %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    all_of(varListWorker)
-  )
-
-dtS1$viz$Interaction <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    IntergroupContact,
-    InteractionDum
-  )
-
-dtS1$viz$KeyNeed <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    DaytimeNeedFullfillment,
-    InteractionNeedFullfillment
-  )
-
-dtS1$viz$SDT <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    autonomy,
-    relatedness, 
-    competence
-  )
-
-dtS1$viz$Attitudes <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    starts_with("Attitudes")
-  )
-
-dtS1$viz$WB <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    ExWB
-  )
-
-dtS1$viz$MoodOverall <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    alertness,
-    calmness,
-    valence
-  )
-
-dtS1$viz$MoodV01Pt01 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v1_1,
-    MDMQ.v1_2
-  )
-
-dtS1$viz$MoodV01Pt02 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v1_3,
-    MDMQ.v1_4
-  )
-
-dtS1$viz$MoodV01Pt03 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v1_5,
-    MDMQ.v1_6
-  )
-
-dtS1$viz$MoodV01Pt04 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v1_7,
-    MDMQ.v1_8
-  )
-
-dtS1$viz$MoodV01Pt05 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v1_9,
-    MDMQ.v1_10
-  )
-
-dtS1$viz$MoodV01Pt06 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v1_11,
-    MDMQ.v1_12,
-    MDMQ.v1_13
-  )
-
-dtS1$viz$MoodV02Pt01 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v2_1,
-    MDMQ.v2_2
-  )
-
-dtS1$viz$MoodV02Pt02 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v2_3,
-    MDMQ.v2_4
-  )
-
-dtS1$viz$MoodV02Pt03 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v2_5,
-    MDMQ.v2_6
-  )
-
-dtS1$viz$MoodV02Pt04 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v2_7,
-    MDMQ.v2_8
-  )
-
-dtS1$viz$MoodV02Pt05 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v2_9,
-    MDMQ.v2_10
-  )
-
-dtS1$viz$MoodV02Pt06 <- dtS1$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    MDMQ.v2_11,
-    MDMQ.v2_12,
-    MDMQ.v2_13
-  )
-```
-
-### Mean Plots
-
 We begin by plotting the item means of all variables.
-We always only display one to three variables jointly to improve visual inspection.
 We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point.
-We additionally plot the data availability of each variable as a marginal plot above the main plot.
 The displayed time axis is the measurement index rather than the actual date to ensure comparability of the changes between each measurement.
-Plots with the raw date-time axis are saved and can be displayed if necessary.
-
-```{r workerMeanViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
-dtS1$vizPlotsMean <-
-  lapply(seq_along(dtS1$viz),
-         function(i)
-           statVarPlot(dataIn = dtS1$viz[[i]], 
-                       name = names(dtS1$viz)[[i]], 
-                       namList = varNameListWorker,
-                       stat = "mean")
-         )
-dtS1$vizPlotsMean
-```
-
-### SD Plots
-
-To illustrate the developments of variance around the previously displayed means, we then plot the item standard deviations of all variables.
-We, thus, plot the average variation from the mean for each individual variable, again summarizing the responses of all participants at any given time point.
-We retain the same variable grouping, the data availability line plot, and the measurement index as the x axis.
-
-```{r WorkerSdViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
-dtS1$vizPlotsSd <-
-  lapply(seq_along(dtS1$viz),
-         function(i)
-           statVarPlot(dataIn = dtS1$viz[[i]], 
-                       name = names(dtS1$viz)[[i]], 
-                       namList = varNameListWorker,
-                       stat = "sd")
-         )
-dtS1$vizPlotsSd
-```
-
-## Study 2 (Student)
+We plot the means for bi-daily, daily, and weekly timescales.
+
+```{r plotS1Mean, fig.height=16}
+plotStat(
+  dt = dtS1Red,
+  varNams = varNamPcaS1,
+  id = "PID",
+  timescale = "bi-daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 1 Means over time [bi-daily]"
+)
+
+plotStat(
+  dt = dtS1Red,
+  varNams = varNamPcaS1,
+  id = "PID",
+  timescale = "daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 1 Means over time [daily]"
+)
+
+plotStat(
+  dt = dtS1Red,
+  varNams = varNamPcaS1,
+  id = "PID",
+  timescale = "weekly",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 1 Means over time [weekly]"
+)
+```
+
+### Study 2 (Means)
 
@@
 
-We, again, select all potentially relevant variables and their labels, and then group the variables in small groups of one to three variables for plotting.
-
-```{r studentDataSelection, eval=FALSE}
-varListStudent <- 
-  c(
-    "IntergroupContact",
-    "InteractionDum",
-    # "KeyNeedFullfillment",
-    # "KeyNeedImp",
-    "DaytimeNeedFullfillment",
-    #"DaytimeNeedImportance",
-    "InteractionNeedFullfillment",
-    #"InteractionNeedImportance",
-    "AntiSo1",
-    "AntiSo2",
-    "AntiSo3",
-    "AntiSo4",
-    "AntiSo5",
-    "AntiSo6",
-    "AntiSo7",
-    "ProSo1",
-    "ProSo2",
-    "ProSo3",
-    "ProSo4",
-    "StudentGoal01",
-    "StudentGoal02",
-    "StudentGoal03",
-    "StudentGoal04",
-    "StudentGoal05",
-    "StudentGoal06",
-    "StudentGoal07",
-    "StudentGoal08",
-    "StudentGoal09",
-    "StudentGoal10",
-    "autonomy",
-    "competence",
-    # Break relatedness up into interaction vs. no interaction(?)
-    # vars are already calculated
-    "relatedness",
-    "AttitudesDutch",
-    "AttitudesPartner",
-    "ExWB",
-    #"angry",
-    #"afraid",
-    #"energy",
-    #"lonelyAlways",
-    "Event"
-  )
-
-varNameListStudent <- 
-  c(
-    "IntergroupContact" = "Dutch Interaction",
-    "InteractionDum" = "Any Interaction",
-    # "KeyNeedFulfillment" = "Key Need Fulfillment",
-    # "KeyNeedImp" = "Key Need Importance",
-    "DaytimeNeedFullfillment" = "Daytime Core Motive Fulfillment",
-    #"DaytimeNeedImportance" = "Daytime Core Motive Importance",
-    "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
-    #"InteractionNeedImportance" = "Interaction Core Motive Importance",
-    "AntiSo1" = "Put someone down",
-    "AntiSo2" = "Show little attention in someones opinion",
-    "AntiSo3" = "Demeaning remarks",
-    "AntiSo4" = "Inpropperly addressing someone",
-    "AntiSo5" = "Ignored or excluded someone",
-    "AntiSo6" = "Doubt someones judgement",
-    "AntiSo7" = "Unwanted attempts of personal matters",
-    "ProSo1" = "Listen to someones problems",
-    "ProSo2" = "Cheer someone up",
-    "ProSo3" = "Help someone get things done",
-    "ProSo4" = "Help someone with responsibilities",
-    "StudentGoal01" = "Social support and connectedness",
-    "StudentGoal02" = "Romantic or sexual relationship",
-    "StudentGoal03" = "Academic",
-    "StudentGoal04" = "Career",
-    "StudentGoal05" = "Financial",
-    "StudentGoal06" = "Health and fitness",
-    "StudentGoal07" = "Leasure and fun",
-    "StudentGoal08" = "Personal improvement and growth",
-    "StudentGoal09" = "Service and help",
-    "StudentGoal10" = "Spiritual or religious",
-    "autonomy" = "Autonomy",
-    "competence" = "Competence",
-    # Break relatedness up into interaction vs. no interaction(?)
-    # vars are already calculated
-    "relatedness" = "Relatedness",
-    "AttitudesDutch" = "Outgroup Attitude",
-    "AttitudesPartner" = "Attitude Interaction Partner",
-    "ExWB" = "Sadness Happiness",
-    #"angry" = "Anger",
-    #"afraid" = "Anxiety",
-    #"energy" = "Energy",
-    #"lonelyAlways" = "Loneliness",
-    "Event" = "Positive or negative event present"
-  )
-
-dtS2$viz <- list()
-
-dtS2$vizall <- dtS2$full %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    all_of(varListStudent)
-  )
-
-dtS2$viz$Interaction <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    IntergroupContact,
-    InteractionDum
-  )
-
-dtS2$viz$KeyNeed <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    DaytimeNeedFullfillment,
-    InteractionNeedFullfillment
-  )
-
-dtS2$viz$AntiSocialBehaviorPt1 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    AntiSo1,
-    AntiSo2,
-    AntiSo3
-  )
-
-dtS2$viz$AntiSocialBehaviorPt2 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    AntiSo4,
-    AntiSo5
-  )
-
-dtS2$viz$AntiSocialBehaviorPt3 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    AntiSo6,
-    AntiSo7
-  )
-
-dtS2$viz$ProSocialBehaviorPt1 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    ProSo1,
-    ProSo2
-  )
-
-dtS2$viz$ProSocialBehaviorPt2 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    ProSo3,
-    ProSo4
-  )
-
-dtS2$viz$GoalsPt1 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal01,
-    StudentGoal02
-  )
-
-dtS2$viz$GoalsPt2 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal03,
-    StudentGoal04
-  )
-
-dtS2$viz$GoalsPt3 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal05,
-    StudentGoal06
-  )
-
-dtS2$viz$GoalsPt4 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal07,
-    StudentGoal08
-  )
-
-dtS2$viz$GoalsPt5 <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal09,
-    StudentGoal10
-  )
-
-dtS2$viz$SDT <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    autonomy,
-    relatedness, 
-    competence
-  )
-
-dtS2$viz$Attitudes <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    starts_with("Attitudes")
-  )
-
-dtS2$viz$WB <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    ExWB
-  )
-
-dtS2$viz$Event <- dtS2$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    Event
-  )
-
-```
-
-### Mean Plots
+We, again, select all potentially relevant variables and their labels, and then plot their developments on a bi-daily, daily, and weekly level.
 
@@
 
-```{r studentMeanViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
-dtS2$vizPlotsMean <-
-  lapply(seq_along(dtS2$viz),
-         function(i)
-           statVarPlot(dataIn = dtS2$viz[[i]], 
-                       name = names(dtS2$viz)[[i]], 
-                       namList = varNameListStudent,
-                       stat = "mean")
-         )
-dtS2$vizPlotsMean
-```
-
-### SD Plots
-
-We then also plot the average variation around the mean for each time point and variable. The plotting method is identical to that of Study 1.
-
-```{r studentSdViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
-dtS2$vizPlotsSd <-
-  lapply(seq_along(dtS2$viz),
-         function(i)
-           statVarPlot(dataIn = dtS2$viz[[i]], 
-                       name = names(dtS2$viz)[[i]], 
-                       namList = varNameListStudent,
-                       stat = "sd")
-         )
-dtS2$vizPlotsSd
-```
-
-## Study 3 (Medical)
+```{r plotS2Mean, fig.height=16}
+plotStat(
+  dt = dtS2Red,
+  varNams = varNamPcaS2,
+  id = "PID",
+  timescale = "bi-daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 2 Means over time [bi-daily]"
+)
+
+plotStat(
+  dt = dtS2Red,
+  varNams = varNamPcaS2,
+  id = "PID",
+  timescale = "daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 2 Means over time [daily]"
+)
+
+plotStat(
+  dt = dtS2Red,
+  varNams = varNamPcaS2,
+  id = "PID",
+  timescale = "weekly",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 2 Means over time [weekly]"
+)
+```
+
+### Study 3 (Means)
 
@@
 
-We, again, select all potentially relevant variables and their labels, and then group the variables in small groups of one to three variables for plotting.
-
-```{r medicalDataSelection, eval=FALSE}
-varListMedical <- 
-  c(
-    "IntergroupContact",
-    "InteractionDum",
-    # "KeyNeedFulfillment",
-    # "KeyNeedImp",
-    "DaytimeNeedFullfillment",
-    "DaytimeNeedImportance",
-    "InteractionNeedFullfillment",
-    "InteractionNeedImportance",
-    "AntiSo1",
-    "AntiSo2",
-    "AntiSo3",
-    "AntiSo4",
-    "AntiSo5",
-    "AntiSo6",
-    "AntiSo7",
-    "ProSo1",
-    "ProSo2",
-    "ProSo3",
-    "ProSo4",
-    "StudentGoal01",
-    "StudentGoal02",
-    "StudentGoal03",
-    "StudentGoal04",
-    "StudentGoal05",
-    "StudentGoal06",
-    "StudentGoal07",
-    "StudentGoal08",
-    "StudentGoal09",
-    "StudentGoal10",
-    "Autonomy",
-    "Competence",
-    # Break relatedness up into interaction vs. no interaction(?)
-    # vars are already calculated
-    "relatedness",
-    "AttitudesDutch",
-    "AttitudesPartner",
-    "exWB",
-    "angry",
-    "afraid",
-    "energy",
-    "lonelyAlways",
-    "Event"
-  )
-
-varNameListMedical <- 
-  c(
-    "IntergroupContact" = "Dutch Interaction",
-    "InteractionDum" = "Any Interaction",
-    # "KeyNeedFulfillment" = "Key Need Fulfillment",
-    # "KeyNeedImp" = "Key Need Importance",
-    "DaytimeNeedFullfillment" = "Daytime Core Motive Fulfillment",
-    "DaytimeNeedImportance" = "Daytime Core Motive Importance",
-    "InteractionNeedFullfillment" = "Interaction Core Motive Fulfillment",
-    "InteractionNeedImportance" = "Interaction Core Motive Importance",
-    "AntiSo1" = "Put someone down",
-    "AntiSo2" = "Show little attention in someones opinion",
-    "AntiSo3" = "Demeaning remarks",
-    "AntiSo4" = "Inpropperly addressing someone",
-    "AntiSo5" = "Ignored or excluded someone",
-    "AntiSo6" = "Doubt someones judgement",
-    "AntiSo7" = "Unwanted attempts of personal matters",
-    "ProSo1" = "Listen to someones problems",
-    "ProSo2" = "Cheer someone up",
-    "ProSo3" = "Help someone get things done",
-    "ProSo4" = "Help someone with responsibilities",
-    "StudentGoal01" = "Social support and connectedness",
-    "StudentGoal02" = "Romantic or sexual relationship",
-    "StudentGoal03" = "Academic",
-    "StudentGoal04" = "Career",
-    "StudentGoal05" = "Financial",
-    "StudentGoal06" = "Health and fitness",
-    "StudentGoal07" = "Leasure and fun",
-    "StudentGoal08" = "Personal improvement and growth",
-    "StudentGoal09" = "Service and help",
-    "StudentGoal10" = "Spiritual or religious",
-    "Autonomy" = "Autonomy",
-    "Competence" = "Competence",
-    # Break relatedness up into interaction vs. no interaction(?)
-    # vars are already calculated
-    "relatedness" = "Relatedness",
-    "AttitudesDutch" = "Outgroup Attitude",
-    "AttitudesPartner" = "Attitude Interaction Partner",
-    "exWB" = "Sadness Happiness",
-    "angry" = "Anger",
-    "afraid" = "Anxiety",
-    "energy" = "Energy",
-    "lonelyAlways" = "Loneliness",
-    "Event" = "Positive or negative event present"
-  )
-
-dtS3$viz <- list()
-
-dtS3$vizall <- dtS3$full %>%
+We, again, select all potentially relevant variables and their labels, and then plot their developments on a bi-daily, daily, and weekly level.
+
+As with the previous studies, we begin by plotting the item means of all variables. The plotting method is identical to that of Studies 1 and 2.
+
+```{r plotS3Mean, fig.height=16}
+plotStat(
+  dt = dtS3Red,
+  varNams = varNamPcaS3,
+  id = "PID",
+  timescale = "bi-daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 3 Means over time [bi-daily]"
+)
+
+plotStat(
+  dt = dtS3Red,
+  varNams = varNamPcaS3,
+  id = "PID",
+  timescale = "daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 3 Means over time [daily]"
+)
+
+plotStat(
+  dt = dtS3Red,
+  varNams = varNamPcaS3,
+  id = "PID",
+  timescale = "weekly",
+  clNum = 2,
+  idVars = idVars,
+  stat = "mean",
+  title = "Study 3 Means over time [weekly]"
+)
+```
+
+## SD
+
+<font size="1"> This is my text number 1</font>
+
+<font size="2"> This is my text number 2</font>
+
+<font size="3"> This is my text number 3</font>
+
+<font size="4"> This is my text number 4</font>
+
+<font size="5"> This is my text number 5</font>
+
+<font size="6"> This is my text number 6</font>
+
+To illustrate the developments of variance around the means, we also plot the item standard deviations of all variables.
+We, thus, plot the average variation from the mean for each individual variable, again summarizing the responses of all participants at any given time point.
+We retain the same measurement indices as the x axis.
+
+### Studies 1-3 (SDs)
+
+```{r plotAllSd}
+PCAnorm(data = dtPcaS123,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
   select(
-    PID,
-    TID,
+    ID,
     TIDnum,
-    all_of(varListMedical)
-  )
-
-dtS3$viz$Interaction <- dtS3$vizall %>%
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>%
+  melt(
+    .,
+    id=c("ID","TIDnum")
+  ) %>%
+  mutate(
+    variable = gsub("_gmc", "", variable)
+  ) %>%
+  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
+  #geom_jitter() +
+  stat_summary(fun=sd, geom="line") +
+  labs(
+    title = "Studies 1-3 SDs over time [bi-daily]",
+    y = "Standard Deviation",
+    x = "Variable",
+    color = "Timepoint",
+    group = "Timepoint"
+  ) +
+  #coord_flip() +
+  theme_Publication() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust=1),
+    legend.key.size = unit(1, 'cm'),
+    legend.key.height = unit(0.4, 'cm'),
+  )
+
+dtPcaS123 %>%
+  group_by(ID, date, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(date))) %>%
+  ungroup %>%
+  select(ID, date, TIDnum, everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  PCAnorm(data = .,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(.)[!names(.) %in% idVars]) %>% 
   select(
-    PID,
-    TID,
+    ID,
     TIDnum,
-    IntergroupContact,
-    InteractionDum
-  )
-
-dtS3$viz$KeyNeedFulfillment <- dtS3$vizall %>%
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>%
+  melt(
+    .,
+    id=c("ID","TIDnum")
+  ) %>%
+  mutate(
+    variable = gsub("_gmc", "", variable)
+  ) %>%
+  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
+  #geom_jitter() +
+  stat_summary(fun=sd, geom="line") +
+  labs(
+    title = "Studies 1-3 SDs over time [daily]",
+    y = "Standard Deviation",
+    x = "Variable",
+    color = "Timepoint",
+    group = "Timepoint"
+  ) +
+  #coord_flip() +
+  theme_Publication() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust=1),
+    legend.key.size = unit(1, 'cm'),
+    legend.key.height = unit(0.4, 'cm'),
+  )
+
+dtPcaS123 %>%
+  group_by(ID, week, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(week))) %>%
+  ungroup %>%
+  select(ID, week, TIDnum, everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  PCAnorm(data = .,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = names(.)[!names(.) %in% idVars]) %>% 
   select(
-    PID,
-    TID,
+    ID,
     TIDnum,
-    DaytimeNeedFullfillment,
-    InteractionNeedFullfillment
-  )
-
-dtS3$viz$KeyNeedImportance <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    DaytimeNeedImportance,
-    InteractionNeedImportance
-  )
-
-dtS3$viz$AntiSocialBehaviorPt1 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    AntiSo1,
-    AntiSo2,
-    AntiSo3
-  )
-
-dtS3$viz$AntiSocialBehaviorPt2 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    AntiSo4,
-    AntiSo5
-  )
-
-dtS3$viz$AntiSocialBehaviorPt3 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    AntiSo6,
-    AntiSo7
-  )
-
-dtS3$viz$ProSocialBehaviorPt1 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    ProSo1,
-    ProSo2
-  )
-
-dtS3$viz$ProSocialBehaviorPt2 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    ProSo3,
-    ProSo4
-  )
-
-dtS3$viz$GoalsPt1 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal01,
-    StudentGoal02
-  )
-
-dtS3$viz$GoalsPt2 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal03,
-    StudentGoal04
-  )
-
-dtS3$viz$GoalsPt3 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal05,
-    StudentGoal06
-  )
-
-dtS3$viz$GoalsPt4 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal07,
-    StudentGoal08
-  )
-
-dtS3$viz$GoalsPt5 <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    StudentGoal09,
-    StudentGoal10
-  )
-
-dtS3$viz$SDT <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    autonomy,
-    relatedness, 
-    competence
-  )
-
-dtS3$viz$Attitudes <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    starts_with("Attitudes")
-  )
-
-dtS3$viz$WB <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    exWB
-  )
-
-dtS3$viz$Emotion <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    angry,
-    afraid,
-    energy,
-    lonelyAlways
-  )
-
-dtS3$viz$Event <- dtS3$vizall %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    Event
-  )
-
-```
-
-### Mean Plots
-
-As with the previous studies, we begin by plotting the item means of all variables. The plotting method is identical to that of Studies 1 and 2.
-
-```{r medicalMeanViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
-dtS3$vizPlotsMean <-
-  lapply(seq_along(dtS3$viz),
-         function(i)
-           statVarPlot(dataIn = dtS3$viz[[i]], 
-                       name = names(dtS3$viz)[[i]], 
-                       namList = varNameListMedical,
-                       stat = "mean")
-         )
-dtS3$vizPlotsMean
-```
-
-### SD Plots
-
-Lastly, also again plot the item standard deviations of all variables at each time point. The plotting procedures are again identical to Studies 1 and 2.
-
-```{r medicalSdViz, echo=FALSE, results='hide',fig.keep='all', eval=FALSE} 
-dtS3$vizPlotsSd <-
-  lapply(seq_along(dtS3$viz),
-         function(i)
-           statVarPlot(dataIn = dtS3$viz[[i]], 
-                       name = names(dtS3$viz)[[i]], 
-                       namList = varNameListMedical,
-                       stat = "sd")
-         )
-dtS3$vizPlotsSd
+    ends_with("_gmc")
+  ) %>%
+  data.frame %>%
+  melt(
+    .,
+    id=c("ID","TIDnum")
+  ) %>%
+  mutate(
+    variable = gsub("_gmc", "", variable)
+  ) %>%
+  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
+  #geom_jitter() +
+  stat_summary(fun=sd, geom="line") +
+  labs(
+    title = "Studies 1-3 SDs over time [weekly]",
+    y = "Standard Deviation",
+    x = "Variable",
+    color = "Timepoint",
+    group = "Timepoint"
+  ) +
+  #coord_flip() +
+  theme_Publication() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust=1),
+    legend.key.size = unit(1, 'cm'),
+    legend.key.height = unit(0.4, 'cm'),
+  )
+```
+
+### Study 1 (SDs)
+
+```{r plotS1Sd, fig.height=16}
+plotStat(
+  dt = dtS1Red,
+  varNams = varNamPcaS1,
+  id = "PID",
+  timescale = "bi-daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 1 SDs over time [bi-daily]"
+)
+
+plotStat(
+  dt = dtS1Red,
+  varNams = varNamPcaS1,
+  id = "PID",
+  timescale = "daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 1 SDs over time [daily]"
+)
+
+plotStat(
+  dt = dtS1Red,
+  varNams = varNamPcaS1,
+  id = "PID",
+  timescale = "weekly",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 1 SDs over time [weekly]"
+)
+```
+
+### Study 2 (SDs)
+
+```{r plotS2Sd, fig.height=16}
+plotStat(
+  dt = dtS2Red,
+  varNams = varNamPcaS2,
+  id = "PID",
+  timescale = "bi-daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 2 SDs over time [bi-daily]"
+)
+
+plotStat(
+  dt = dtS2Red,
+  varNams = varNamPcaS2,
+  id = "PID",
+  timescale = "daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 2 SDs over time [daily]"
+)
+
+plotStat(
+  dt = dtS2Red,
+  varNams = varNamPcaS2,
+  id = "PID",
+  timescale = "weekly",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 2 SDs over time [weekly]"
+)
+```
+
+### Study 3 (SDs)
+
+```{r plotS3Sd, fig.height=16}
+plotStat(
+  dt = dtS3Red,
+  varNams = varNamPcaS3,
+  id = "PID",
+  timescale = "bi-daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 3 SDs over time [bi-daily]"
+)
+
+plotStat(
+  dt = dtS3Red,
+  varNams = varNamPcaS3,
+  id = "PID",
+  timescale = "daily",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 3 SDs over time [daily]"
+)
+
+plotStat(
+  dt = dtS3Red,
+  varNams = varNamPcaS3,
+  id = "PID",
+  timescale = "weekly",
+  clNum = 2,
+  idVars = idVars,
+  stat = "sd",
+  title = "Study 3 SDs over time [weekly]"
+)
 ```

### 2022-06-27 scripts/functions/MlCorMat.R
Update MlCorMat.R

@@ -262,11 +262,12 @@ PCAnorm <-
       group_by(!!rlang::sym(tid)) %>%
       mutate_at(selection, list(cwt = ~ . - mean(., na.rm = TRUE))) %>%
       ungroup %>%
-      filter_at(selection, all_vars(!is.na(.))) %>%
+      #filter_at(selection, all_vars(!is.na(.))) %>%
       mutate_at(selection, list(gmc = ~ . - mean(., na.rm = TRUE))) %>%
       mutate_at(selection, list(sigma = ~ sqrt(sum(. ^ 2)/(length(unique(!!rlang::sym(pid)))*length(unique(!!rlang::sym(tid)))))))
     
-    cat("@Jannis: Don't forget to fix the filter once sample section is finalized.")
+    #cat("@Jannis: Don't forget to fix the filter once sample section is finalized.")
+    warning("@Jannis: Don't forget to fix the filter once sample section is finalized.")
     
     data03 <- map_dfc(selection,
                       ~ data02 %>%

### 2022-06-27 scripts/functions/meanViz.R
plotStat added

@@ -186,3 +186,131 @@ statVarPlot <- function(dataIn, name, namList, stat = "mean", ...){
   list(pTid, pDate)
   pTid
 }
+
+plotStat <- function(dt, varNams, id, timescale, clNum, idVars, stat, title) {
+  # dt = dtS3Red
+  # varNams = varNamPcaS3
+  # id = "PID"
+  # timescale = "bi-daily"
+  # clNum = 2
+  # idVars = idVars
+  # stat = "sd"
+  # title = "Study 3 Means over time [bi-daily]"
+  
+  varCl <- dt %>%
+    select(any_of(varNams)) %>%
+    mutate(
+      date = as.Date(gsub(" .*", "", TID)),
+      week = strftime(date, format = "%Y-W%V")
+    ) %>%
+    select(
+      any_of(id),
+      TID,
+      date,
+      week,
+      TIDnum,
+      everything()
+    ) %>%
+    arrange(get(id), TIDnum) %>%
+    PCAnorm(data = .,
+            pid = id,
+            tid = "TIDnum",
+            selection = names(.)[!names(.) %in% idVars]) %>%
+    select(
+      ends_with("_gmc")
+    ) %>%
+    data.frame %>%
+    summarise_all(
+      sd, na.rm = TRUE
+    ) %>%
+    t %>% 
+    as.data.frame %>%
+    kmeans(., centers = clNum)
+  
+  if (timescale == "bi-daily") {
+    dtTimescale <- dt %>%
+      select(any_of(varNams)) %>%
+      mutate(date = as.Date(gsub(" .*", "", TID)),
+             week = strftime(date, format = "%Y-W%V")) %>%
+      select(any_of(id),
+             TID,
+             date,
+             week,
+             TIDnum,
+             everything()) %>%
+      arrange(get(id), TIDnum) %>%
+      mutate_all( ~ ifelse(is.nan(.), NA, .))
+  } else if (timescale == "daily") {
+    dtTimescale <- dt %>%
+      select(any_of(varNams)) %>%
+      mutate(date = as.Date(gsub(" .*", "", TID)),
+             week = strftime(date, format = "%Y-W%V")) %>%
+      group_by(get(id), date) %>%
+      summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+      ungroup %>%
+      mutate(TIDnum = as.numeric(factor(date))) %>%
+      select(any_of(id), date, TIDnum, everything()) %>%
[…]

### 2022-06-27 scripts/functions/meanViz.R
mode function added

@@ -1,3 +1,15 @@
+calc_mode <- function(x){
+  
+  # List the distinct / unique values
+  distinct_values <- unique(x)
+  
+  # Count the occurrence of each distinct value
+  distinct_tabulate <- tabulate(match(x, distinct_values))
+  
+  # Return the value with the highest occurrence
+  distinct_values[which.max(distinct_tabulate)]
+}
+
 lWideToLong <- function(data, ...){
   reshape2::melt(
     data, 

### 2022-06-27 Supplemental-Material-A-Annotated-Analysis.Rmd
MI, three-way ANOVA, Plots added + clean up

@@
 - Peter de Jonge^1,3^
-- Rei Tendeiro-Monden^2,3^
+- Rei Monden^2,3^
 - ^1^University of Groningen, Department of Psychology
@@
   chunk_output_type: console
-#bibliography: references.bib
+bibliography: references.bib
 csl: apa.csl
@@
 
-df %>%
+AlportDescr <- df %>%
   #filter(OutgroupInteraction == "Yes") %>%
@@
 
-sjPlot::tab_itemscale(iaWorkerAllport)
+itemScaleAllport01 <- sjPlot::tab_itemscale(iaWorkerAllport)
 
@@
 
-sjPlot::tab_itemscale(iaWorkerAllport, factor.groups)
-
-ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)
+sjPlot::tab_itemscale(iaWorkerAllport, factor.groups, show.kurtosis = TRUE)
+
+AllportAlpha <- ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)
 
@@
 
-horst::nestedAlpha(item.level.1 = "value",
+AllportNestedAlpha <- horst::nestedAlpha(item.level.1 = "value",
                    level.2      = "TIDnum",
@@
 
-## Data Availability and Sample Selection  {.tabset}
+## Data Availability and Sample Selection {.tabset .tabset-fade}
 
@@
 We then export the data availability tables as comma separated value files (.csv) to be assessed in detail using a spreadsheet program (such as MS Excel).
+
+For our sample selection we address both the selection of time points and participants.
+Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
+Thus, we then select the time points for which we have less than 45% missingness and select participants who have less than 45% missingness across the selected time range.
+
+To approximate an ideal sample selection that maximized the number of participants and timepoints that fit the 55% data availability criterion.
+We select sequentially remove the rows or columns with the lowest data availability rate untill all rows and columns have more than 55% data available (i.e., less than 45% missingness).
+
+<div class="alert alert-info alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-info-circle"></i> <b>Note: </b><br/>
+  This needs to be adjusted current missingness criterion set to more conservative 33%.
+</div>
 
@@
 
-### Timescale ML test
-
-```{r }
+We aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.
+Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
+This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
+We chose to select time scales that align with common human cycles.
+We thus compare the variances of bi-daily, daily, and weekly aggregations.
+Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.
+We then chose the time scales that have the most variance.
+
+```{r keyNeedTest}
+
+# WILL USE IMPUTED DATA AND FOR-LOOP ACROSS ALL RELEVANT DVs
+# PLOT EFFECT SIZES
+
 dtMain3mpca <- dtMain3mpca %>%
+  group_by(study) %>%
   mutate(
@@
     TIDweek = as.numeric(factor(week))
-  )
-
-test <-
-  lme4::lmer(KeyNeedFulfillment ~ 1 + (1 + TIDnum + TIDdate + TIDweek | PID),
+  ) %>%
+  ungroup
+
+# Not sure whether TIDs as predictors or empty model?
+keyNeedMlTest01 <-
+  lme4::lmer(KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek + (1 + TIDnum + TIDdate + TIDweek | ID),
     data = dtMain3mpca
-  ) # use optim if it does not converge
-
-test2 <-
+  )
+
+keyNeedMlTest02 <-
   lme(
-    KeyNeedFulfillment ~ 1,
-    random = ~ 1 + TIDnum + TIDdate + TIDweek | PID,
+    KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek,
+    random = ~ 1 + TIDnum + TIDdate + TIDweek | ID,
     data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
@@
 
-summary(test2)
-#summ(test, digits = 3, center = FALSE)
-anova(test2)
+# summary(keyNeedMlTest02)
+# summ(keyNeedMlTest01, digits = 3, center = FALSE)
+# anova(keyNeedMlTest02)
 
@@
 
-### Studies 1-3
-
-```{r MiMain3mpca}
-library(Amelia)
-
-varNamInMiMain <- varNames %>%
-  filter(
-    aux == 1,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamInMiMainPsbl <- c(
-  paste(varNamInMiMain, rep(".pre", length(varNamInMiMain)), sep = ""),
-  varNamInMiMain,
-  paste(varNamInMiMain, rep(".post", length(varNamInMiMain)), sep = "")
-)
-
-varNamInMiMainRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names
-  )
-)
-
-varNamMiMainNoms <- varNames %>%
-  filter(
-    aux == 1,
-    format == "nominal",
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-dtMiMain <- rbind(
-  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% mutate(across(!TID &
-                                                                                             !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         study,
-         date,
-         week,
-         everything(),
-         -PID,
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>%
-  select_if( ~ sum(!is.na(.)) > 1) # only include variables that have any data (i.e., not all NA)
-
-MiMain <-
-  amelia(
-    dtMiMain,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("TID", "date", "week", "study"), #"PID", 
-    noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMain, file = "data/imputed/MainImputed.RData")
-```
-
-### Study 1
-
-```{r MiS1}
-varNamInMiS1 <- varNames %>%
-  filter(
-    aux == 1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamInMiS1Psbl <- c(
-  paste(varNamInMiS1, rep(".pre", length(varNamInMiS1)), sep = ""),
-  varNamInMiS1,
-  paste(varNamInMiS1, rep(".post", length(varNamInMiS1)), sep = "")
-)
-
-varNamdtS1Noms <- varNames %>%
-  filter(
-    aux == 1,
-    format == "nominal",
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-dtMiS1 <- dtS1Red %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
-    date,
-    week,
-    any_of(varNamInMiS1Psbl),
-    -time,
-    -starts_with("ResponseId")
-  ) %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-    ) %>%
-  as.data.frame %>%
-  select_if(~sum(!is.na(.)) > 1) #%>% # only include variables that have any data (i.e., not all NA)
-
-  which(!is.numeric(dtMiS1) | sapply(dtMiS1, sd, na.rm=TRUE)>0)
-
-  select_if(~sd(., na.rm = TRUE) > 0) # only include variables that vary (i.e., sd != 0)
-
-#dtS1MiCs <- names(dtMiS1)[endsWith(names(dtMiS1), c(".pre", ".post"))] # cross sectional vars
-
-MiS1 <-
-  amelia(
-    dtMiS1,
-    m = 2,
-    ts = "TIDnum",
-    cs = "PID",
-    idvars =  c("TID", "date", "week"),
-    noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-write.amelia(MiS1, file.stem = "data/imputed/S1Imputed.Amelia")
-```
-
-### Study 2
-
-### Study 3
-
-## Center and Normalize
-
-```{r mainCentered}
-varNam3mpcaMainAnalysis <- varNamesDailyFull %>%
-  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum")) %>%
-  pull
-
-dtMain3mpcaNorm <-
-  PCAnorm(data = dtMain3mpca,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = varNam3mpcaMainAnalysis)
-```
-
-## Long to Wide
+We ultimately aim to create 20 imputed datasets to perform the analyses on.
+As outlined by @Monden2015, we will use the key variables themself as well as auxiliary variables to impute missing values.
+Auxiliary variables are any pre-, post-, or daily dairy variables other than the main variables.
+If there are any modeling issues we will limit the auxiliary variables to all variables that that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
+
+We create the imputed data set for each of the analyses.
+This includes the main analyses with across all three studies, using the set of common variables, as well as follow-up analyses for study-specific, interaction-specific and time scale specific datasets.
 
@@
 
+### Studies 1-3
+
+We begin by preparing the data for the main analysis. Here, we:
+
+1. Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
+2. Create the combined dataset in the correct format (long format with all missed time points as NAs).
+3. Impute the data and save the imputed datasets for later recall.
+
+```{r MiMain3mpca}
+library(Amelia)
+
+varNamInMiMain <- varNames %>%
+  filter(
+    aux == 1,
+    studyS1 == "S1",
+    studyS2 == "S2",
+    studyS3 == "S3"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+varNamMiMain <- c(varNamAdd, varNamInMiMain[!varNamInMiMain %in% varNamOut])
+
+varNamInMiMainPsbl <- c(
+  paste(varNamMiMain, rep(".pre", length(varNamInMiMain)), sep = ""),
+  varNamMiMain,
+  paste(varNamMiMain, rep(".post", length(varNamInMiMain)), sep = "")
+)
+
+varNamInMiMainRed <- Reduce(
+  intersect,
+  list(
+    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
+    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
+    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names
+  )
+)
+
+varNamMiMainNoms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS1 == "S1",
+    studyS2 == "S2",
+    studyS3 == "S3"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+dtMiMain <- rbind(
+  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% mutate(across(!TID &
+                                                                                             !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V"),
+    roomate.pre = as.numeric(roomate.pre),
+    Reason.pre = as.numeric(Reason.pre),
+    association.pre = as.numeric(association.pre)
+  ) %>%
+  select(ID,
+         study,
+         date,
+         week,
+         everything(),
+         -PID,
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(ID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  mutate(study = replace_na(study, calc_mode(study))) %>%
+  ungroup %>%
+  group_by(study, TIDnum) %>%
+  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>%
+  as.data.frame
+
+# MiMain <-
+#   amelia(
+#     dtMiMain,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMain, file = "data/imputed/MainImputed.RData")
+load("data/imputed/MainImputed.RData")
+```
+
+### Study 1
+
+To be done once initial tests approved.
+
+```{r MiS1}
+varNamInMiS1 <- varNames %>%
+  filter(
+    aux == 1,
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+varNamMiS1 <- c(varNamAdd, varNamInMiS1[!varNamInMiS1 %in% varNamOut])
+
+varNamInMiS1Psbl <- c(
+  paste(varNamMiS1, rep(".pre", length(varNamMiS1)), sep = ""),
+  varNamMiS1,
+  paste(varNamMiS1, rep(".post", length(varNamMiS1)), sep = "")
+)
+
+varNamdtS1Noms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# TO BE ADAPTED:
+dtMiS1 <- dtS1Red %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+    ) %>%
+  select(
+    PID,
+    TID,
+    TIDnum,
+    date,
+    week,
+    any_of(varNamInMiS1Psbl),
+    -time,
+    -starts_with("ResponseId")
+  ) %>%
+  as.data.frame %>%
+  mutate_at(vars(-c("TID", "date", "week")), as.numeric) %>%
+  select_if(~sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+  as.data.frame
+
+#dtS1MiCs <- names(dtMiS1)[endsWith(names(dtMiS1), c(".pre", ".post"))] # cross sectional vars
+
+# MiS1 <-
+#   amelia(
+#     dtMiS1,
+#     m = 1,
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars =  c("TID", "date", "week"),
+#     #noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     #parallel = "multicore",
+#     #ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS1, file = "data/imputed/S1Imputed.RData")
+# load("data/imputed/S1Imputed.RData")
+```
+
+### Study 2
+
+To be done once initial tests approved.
+
+```{r MiS2}
+# BASED ON MiS1
+```
+
+### Study 3
+
+To be done once initial tests approved.
+
+```{r MiS3}
+# BASED ON MiS1
+```
+
+## Center and Normalize
+
+For the 3MPCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
+The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering).
+The normalization within variable are important for ensuring equal variances across variables, which ensures equal weighting of the variables in the 3MPCA.
+
+```{r mainCentered}
+varNam3mpcaMainAnalysis <- varNamesDailyFull %>%
+  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum")) %>%
+  pull
+
+dtMain3mpcaNorm <-
+  PCAnorm(data = dtMain3mpca,
+          pid = "ID",
+          tid = "TIDnum",
+          selection = varNam3mpcaMainAnalysis)
+```
+
+## Long to Wide
+
+For the Three-way ANOVA and the Three-mode PCA, the datasets need to be in a $n * mp$ matrix format, where $n$ is are the participants, $m$ are the time points, and $p$ are the variables.
+We do this by taking the prepared datasets, melting them into a id-variable-value table and casting the table into a wide format of variables_timepoint columns.
+
 ```{r longToWide}
@@
 
-# test <- PCAnorm(data = dtPcaS123,
-#           pid = "ID",
-#           tid = "TIDnum",
-#           selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
-#   select(
-#     ID,
-#     PID,
-#     TIDnum,
-#     study,
-#     ends_with("_gmc")
-#   ) %>%
-#   data.frame %>%
-#   melt(
-#     .,
-#     id=c("ID", "PID", "TIDnum", "study")
-#   ) %>%
-#   mutate(
-#     variable = paste(gsub("_gmc", "", variable), TIDnum, sep = "_")
-#   ) %>%
-#   #select(-TIDnum) %>% 
-#   arrange(variable, ID) %>%
-#   mutate(
-#     idVar = paste(ID, variable, sep = "_")
-#   ) 
-# plyr::join(test, data.frame(table(idVar = test$idVar)))
-
-test <- PCAnorm(data = dtPcaS123,
+dtPcaS123Imp <- MiMain$imputations[[1]] %>% 
+  select(ID, date, week, study, any_of(varNamPcaS123)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
+
+dtMainWideC <- PCAnorm(data = dtPcaS123,
           pid = "ID",
@@
 
-test2 <- dtPcaS123 %>%
+dtMainWide <- dtPcaS123 %>%
   select(
@@
 
+dtMainImpWide <- dtPcaS123Imp %>%
+  select(
+    -TID,
+    -date,
+    -week
+  ) %>% 
+  melt(
+    .,
+    id=c("ID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+  ) %>%
+  select(-TIDnum, -study) %>%
+  arrange(variable, ID) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-ID)
+
 ```
@@
 
-```{r threeWayAnova}
+We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset.
+Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions.
+We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.
+
+```{r threeWayAnova, message=FALSE, warning=FALSE}
 library(ThreeWay)
-threeWayAno <- threewayanova(
-  Y = test2,
-  n = nrow(test2), # Nr. ppt.s (A)
-  m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
-  p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
-)
-
-```
-
-# **Plots per Item** {.tabset}
+# threeWayAno <- threewayanova(
+#   Y = dtMainWide,
+#   n = nrow(dtMainWide), # Nr. ppt.s (A)
+#   m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
+#   #p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
+#   p = length(names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) # Nr. variables (C)
+# )
+
+capture.output(
+  threeWayAnoImp <- threewayanova(
+    Y = dtMainImpWide,
+    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+    m = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (B)
+    p = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) # Nr. variables (C)
+  ), 
+file='NUL')
+
+options(knitr.kable.NA = '')
+data.frame(SS = sapply(threeWayAnoImp, c)) %>%
+  rownames_to_column(., var = "EffectNam") %>%
+  mutate(RSqrd = SS / sum(SS) * 100) %>%
+  add_row(
+    EffectNam = "Total", 
+    SS = sum(.$SS),
+    RSqrd = NA
+  ) %>%
+  mutate(
+    Effect = c(
+      "Person",
+      "Time",
+      "Variable",
+      "Person*Time",
+      "Person*Variable",
+      "Time*Variable",
+      "Person*Time*Variable+error",
+      "Total"
+    )
+  ) %>%
+  select(Effect, everything(), -EffectNam) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "r", "c"),
+      digits=2,
+      col.names = c(
+        "Effect",
+        "Sum of Squares",
+        "$R^2$"
+      ),
+      caption = "Three-way ANOVA Main Analysis Single Imputation") %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+
+```
+
+We find that the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA.
+Interestingly, time itself has a relatively small percentage of explained variance and only becomes important in compination with person- or person- and variable-specific differences.
+
+```{r panel-setup, include = FALSE}
+xaringanExtra::use_panelset()
+xaringanExtra::style_panelset_tabs(font_family = "inherit")
+```
+
+# **Plots per Item** {.tabset .tabset-fade}
 
@@
 
+Please note that the items might be measured on different ranges.
+To avoid misinterpretation of variances and changes over time, we cluster variables with similar ranges in different plots where necessary.
+
 ## Means
 
-### Studies 1-3 (Means)
+To illustrate the developments of the means,we plot the item means of all variables.
+We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point.
+The displayed time grouping is based on the measurement index rather than the actual date to ensure comparability of the changes between each measurement as well as studies.
+We plot the means for bi-daily, daily, and weekly timescales.
+
+### Studies 1-3 (Means) {.unlisted .unnumbered}
+
+We beginn by assessing the imputed dataset of the main analysis, where participants from all three studies are combined, using the relevant variables that are available for all three studies.
 
 ```{r plotAllMean}
-varClS123 <- dtPcaS123 %>%
+varClS123 <- dtPcaS123Imp %>%
   PCAnorm(data = .,
-          pid = "PID",
+          pid = "ID",
           tid = "TIDnum",
@@
 
-PCAnorm(data = dtPcaS123,
+PCAnorm(data = dtPcaS123Imp,
           pid = "ID",
@@
 
-### Study 1 (Means)
+### Study 1 (Means) {.unlisted .unnumbered}
 
@@
 Study 1 additionally included several emotion and mood measurements but fewer cognitive and behavioral measures than studies two and three.
-
-We begin by plotting the item means of all variables.
-We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point.
-The displayed time axis is the measurement index rather than the actual date to ensure comparability of the changes between each measurement.
-We plot the means for bi-daily, daily, and weekly timescales.
 
@@
 
-### Study 2 (Means)
+### Study 2 (Means) {.unlisted .unnumbered}
 
@@
 
-### Study 3 (Means)
+### Study 3 (Means) {.unlisted .unnumbered}
 
@@
 
-<font size="1"> This is my text number 1</font>
-
-<font size="2"> This is my text number 2</font>
-
-<font size="3"> This is my text number 3</font>
-
-<font size="4"> This is my text number 4</font>
-
-<font size="5"> This is my text number 5</font>
-
-<font size="6"> This is my text number 6</font>
-
 To illustrate the developments of variance around the means, we also plot the item standard deviations of all variables.
@@
 
-### Studies 1-3 (SDs)
+### Studies 1-3 (SDs) {.unlisted .unnumbered}
+<!-- <font size="5">Studies 1-3 (SDs)</font>  -->
 
@@
 
-### Study 1 (SDs)
+### Study 1 (SDs) {.unlisted .unnumbered}
 
@@
 
-### Study 2 (SDs)
+### Study 2 (SDs) {.unlisted .unnumbered}
 
@@
 
-### Study 3 (SDs)
+### Study 3 (SDs) {.unlisted .unnumbered}
 
@@
 
-<!-- # **References**   -->
+# **References**
 

### 2022-06-28 Supplemental-Material-A-Annotated-Analysis.Rmd
MI & 3-way ANOVA daily weekly added

@@
 
-```{r MiMain3mpca}
+```{r MiMain}
 library(Amelia)
-
 varNamInMiMain <- varNames %>%
@@
 
+dtMiMainDaily <- rbind(
+  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% 
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V"),
+    roomate.pre = as.numeric(roomate.pre),
+    Reason.pre = as.numeric(Reason.pre),
+    association.pre = as.numeric(association.pre)
+  ) %>%
+  group_by(ID, date, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(date))) %>%
+  ungroup %>%
+  select(ID,
+         study,
+         date,
+         everything(),
+         -PID,
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(ID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  mutate(study = replace_na(study, calc_mode(study))) %>%
+  ungroup %>%
+  group_by(study, TIDnum) %>%
+  mutate(across(c("date"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate_if(is.numeric, ~ ifelse(is.nan(.), NA, .)) %>%
+  as.data.frame
+
+dtMiMainWeeekly <- rbind(
+  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% 
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V"),
+    roomate.pre = as.numeric(roomate.pre),
+    Reason.pre = as.numeric(Reason.pre),
+    association.pre = as.numeric(association.pre)
+  ) %>%
+  group_by(ID, week, study) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>%
+  group_by(study) %>%
+  mutate(TIDnum = as.numeric(factor(week))) %>%
+  ungroup %>%
+  select(ID,
+         study,
+         week,
+         everything(),
+         -PID,
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(ID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  mutate(study = replace_na(study, calc_mode(study))) %>%
+  ungroup %>%
+  group_by(study, TIDnum) %>%
+  mutate(across(c("week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate_if(is.numeric, ~ ifelse(is.nan(.), NA, .)) %>%
+  as.data.frame
+
 # MiMain <-
@@
 load("data/imputed/MainImputed.RData")
+
+# MiMainDaily <-
+#   amelia(
+#     dtMiMainDaily,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("date", "study"), #"PID",
+#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
+load("data/imputed/MainDailyImputed.RData")
+
+# MiMainWeekly <-
+#   amelia(
+#     dtMiMainWeeekly,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("week", "study"), #"PID",
+#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
+#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
+load("data/imputed/MainWeeklyImputed.RData")
 ```
@@
   filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
+dtPcaS123DailyImp <- MiMainDaily$imputations[[1]] %>%
+  select(ID, date, study, any_of(varNamPcaS123)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(
+    TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
+  )
+dtPcaS123WeeklyImp <- MiMainWeekly$imputations[[1]] %>%
+  select(ID, week, study, any_of(varNamPcaS123)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(
+    TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
+  )
 
@@
 
+dtMainDailyImpWide <- dtPcaS123DailyImp %>%
+  select(
+    -date
+  ) %>% 
+  melt(
+    .,
+    id=c("ID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+  ) %>%
+  select(-TIDnum, -study) %>%
+  arrange(variable, ID) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-ID)
+
+dtMainWeeklyImpWide <- dtPcaS123WeeklyImp %>%
+  select(
+    -week
+  ) %>% 
+  melt(
+    .,
+    id=c("ID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+  ) %>%
+  select(-TIDnum, -study) %>%
+  arrange(variable, ID) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-ID)
 ```
@@
 
-```{r threeWayAnova, message=FALSE, warning=FALSE}
+```{r threeWayAnovaMainBiDaily, message=FALSE, warning=FALSE}
 library(ThreeWay)
-# threeWayAno <- threewayanova(
-#   Y = dtMainWide,
-#   n = nrow(dtMainWide), # Nr. ppt.s (A)
-#   m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
-#   #p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
-#   p = length(names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) # Nr. variables (C)
-# )
-
 capture.output(
-  threeWayAnoImp <- threewayanova(
+  threeWayAnoBiDailyImp <- threewayanova(
     Y = dtMainImpWide,
@@
 options(knitr.kable.NA = '')
-data.frame(SS = sapply(threeWayAnoImp, c)) %>%
+data.frame(SS = sapply(threeWayAnoBiDailyImp, c)) %>%
   rownames_to_column(., var = "EffectNam") %>%
@@
       ),
-      caption = "Three-way ANOVA Main Analysis Single Imputation") %>%
+      caption = "Three-way ANOVA Main Analysis Bi-daily (Single Imputation)") %>%
   kable_classic(
@@
   )
-
-```
-
-We find that the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA.
+```
+
+<br>
+
+We find that for the bi-daily time scale the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA.
 Interestingly, time itself has a relatively small percentage of explained variance and only becomes important in compination with person- or person- and variable-specific differences.
 
-```{r panel-setup, include = FALSE}
-xaringanExtra::use_panelset()
-xaringanExtra::style_panelset_tabs(font_family = "inherit")
-```
+```{r threeWayAnovaMainDaily, message=FALSE, warning=FALSE}
+capture.output(
+  threeWayAnoDailyImp <- threewayanova(
+    Y = dtMainDailyImpWide,
+    n = nrow(dtMainDailyImpWide), # Nr. ppt.s (A)
+    m = length(unique(dtPcaS123DailyImp$TIDnum)), # Nr. time points (B)
+    p = length(names(dtPcaS123DailyImp)[!names(dtPcaS123DailyImp) %in% idVars]) # Nr. variables (C)
+  ), 
+file='NUL')
+
+options(knitr.kable.NA = '')
+data.frame(SS = sapply(threeWayAnoDailyImp, c)) %>%
+  rownames_to_column(., var = "EffectNam") %>%
+  mutate(RSqrd = SS / sum(SS) * 100) %>%
+  add_row(
+    EffectNam = "Total", 
+    SS = sum(.$SS),
+    RSqrd = NA
+  ) %>%
+  mutate(
+    Effect = c(
+      "Person",
+      "Time",
+      "Variable",
+      "Person*Time",
+      "Person*Variable",
+      "Time*Variable",
+      "Person*Time*Variable+error",
+      "Total"
+    )
+  ) %>%
+  select(Effect, everything(), -EffectNam) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "r", "c"),
+      digits=2,
+      col.names = c(
+        "Effect",
+        "Sum of Squares",
+        "$R^2$"
+      ),
+      caption = "Three-way ANOVA Main Analysis Daily (Single Imputation)") %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+<br>
+
+We additionally see that for the daily time scale aggregation, the $Person*Time*Variable+error$ sums of squares are reduced but remain the largest sums of squares and are still substantial.
+
+```{r threeWayAnovaMainWweekly, message=FALSE, warning=FALSE}
+capture.output(
+  threeWayAnoWeeklyImp <- threewayanova(
+    Y = dtMainWeeklyImpWide,
+    n = nrow(dtMainWeeklyImpWide), # Nr. ppt.s (A)
+    m = length(unique(dtPcaS123WeeklyImp$TIDnum)), # Nr. time points (B)
+    p = length(names(dtPcaS123WeeklyImp)[!names(dtPcaS123WeeklyImp) %in% idVars]) # Nr. variables (C)
+  ), 
+file='NUL')
+
+options(knitr.kable.NA = '')
+data.frame(SS = sapply(threeWayAnoWeeklyImp, c)) %>%
+  rownames_to_column(., var = "EffectNam") %>%
+  mutate(RSqrd = SS / sum(SS) * 100) %>%
+  add_row(
+    EffectNam = "Total", 
+    SS = sum(.$SS),
+    RSqrd = NA
+  ) %>%
+  mutate(
+    Effect = c(
+      "Person",
+      "Time",
+      "Variable",
+      "Person*Time",
+      "Person*Variable",
+      "Time*Variable",
+      "Person*Time*Variable+error",
+      "Total"
+    )
+  ) %>%
+  select(Effect, everything(), -EffectNam) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "r", "c"),
+      digits=2,
+      col.names = c(
+        "Effect",
+        "Sum of Squares",
+        "$R^2$"
+      ),
+      caption = "Three-way ANOVA Main Analysis Weekly (Single Imputation)") %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+<br>
+
+However, with the aggregation to `r length(unique(dtPcaS123WeeklyImp$TIDnum))` weekly measurements, the $Person*Time*Variable+error$ sums of squares become less prominent and the variable differences explain substantially more variance.
 

### 2022-07-12 scripts/functions/meanViz.R
testing post renv upgrade

@@ -200,14 +200,14 @@ statVarPlot <- function(dataIn, name, namList, stat = "mean", ...){
 }
 
 plotStat <- function(dt, varNams, id, timescale, clNum, idVars, stat, title) {
-  # dt = dtS3Red
-  # varNams = varNamPcaS3
+  # dt = dtS2Red
+  # varNams = varNamS2PCA
   # id = "PID"
   # timescale = "bi-daily"
   # clNum = 2
   # idVars = idVars
   # stat = "sd"
-  # title = "Study 3 Means over time [bi-daily]"
+  # title = "Study 2 Means over time [bi-daily]"
   
   varCl <- dt %>%
     select(any_of(varNams)) %>%
@@ -231,13 +231,13 @@ plotStat <- function(dt, varNams, id, timescale, clNum, idVars, stat, title) {
     select(
       ends_with("_gmc")
     ) %>%
-    data.frame %>%
+    data.frame %>% 
     summarise_all(
       sd, na.rm = TRUE
     ) %>%
     t %>% 
-    as.data.frame %>%
-    kmeans(., centers = clNum)
+    as.data.frame %>% 
+    stats::kmeans(., centers = clNum)
   
   if (timescale == "bi-daily") {
     dtTimescale <- dt %>%

### 2022-07-12 Supplemental-Material-A-Annotated-Analysis.Rmd
three-way anova heading adjusted
S1 scale indices added

@@
 
+```{=html}
 <style type="text/css">
@@
 </style>
-
+```
 ```{r setup, include=FALSE}
@@
   "reshape",
-  "factoextra"
+  "factoextra",
+  "Amelia", 
+  "ThreeWay"
 )
@@
 
---------------------------------------------------------------------
+------------------------------------------------------------------------
 
@@
 <i class="fas fa-exclamation-circle"></i> Note.
-Boxplots display the interquartile range (IQR, center box), and the whiskers extend 1.5*IQR from the lower and upper hinge.
+Boxplots display the interquartile range (IQR, center box), and the whiskers extend 1.5\*IQR from the lower and upper hinge.
 The white point indicates the mean and the white center line indicates the median.
@@
 This means that the datasets had inconsistent file formats and naming conventions.
-For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '_data/S1_Workers/processed/cleaned_' and for the raw data files see '_data/S1_Workers/raw_').
+For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*').
 For the two other studies, we import the raw csv files from their respective folders.
@@
 We then format the time and date variables and add person- and measurement indices (for easy and meaningful addressing of the data).
-We also exclude our own test data.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS1_' list.
+We also exclude our own test data.\
+<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS1*' list.
 
@@
 We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement).
-We also reformat missing values and format core ID variables.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS2_' list.
+We also reformat missing values and format core ID variables.\
+<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS2*' list.
 
@@
 We exclude our own test data.
-We also reformat missing values and format core ID variables.
-<i class="fas fa-info-circle"></i> _Note:_ All data preparation steps are saved in the '_dtS3_' list.
+We also reformat missing values and format core ID variables.\
+<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS3*' list.
 
@@
 In some cases participants completed their daily diary surveys for the afternoon after midnight.
-In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.
+In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.\
 2.
 We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).
+Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
 3.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
@@
 
+varNamIndicesS1 <- c(
+  "autonomy",
+  "competence",
+  "relatedness",
+  "education_level.pre",
+  "associationMerged.pre",
+  "assimilation.pre",
+  "separation.pre",
+  "integration.pre",
+  "marginalization.pre",
+  "VIA_heritage.pre",
+  "VIA_Dutch.pre",
+  "SSAS_surrounding.pre",
+  "SSAS_privat.pre",
+  "SSAS_public.pre",
+  "assimilation.post",
+  "separation.post",
+  "integration.post",
+  "marginalization.post",
+  "VIA_heritage.post",
+  "VIA_Dutch.post",
+  "rosenberg.post",
+  "social_support.post",
+  "stress.post",
+  "discrimination.post",
+  "discrimination_month.post",
+  "NLE_1month.post",
+  "NLE_6month.post",
+  "NLE_12month.post"
+  )
+
+df$roommate.pre_calc <- df %>%
+  select(
+    starts_with("roommate")
+  ) %>% 
+  mutate_all(as_factor) %>%
+  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>%
+  pull
+
+df$Reason.pre_calc <- df %>%
+  select(
+    starts_with("Reason"),
+    -Reason_nodesire, 
+    -ReasonOther.pre 
+  ) %>% 
+  mutate_all(as_factor) %>% 
+  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>%
+  pull
+
+df$occupation.pre_calc <- df %>%
+  select(
+    starts_with("occupation"),
+    -occupation_8_TEXT.pre
+  ) %>%
+  mutate_all(as_factor) %>% 
+  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>% 
+  pull
+
+df$CurrentEducation.pre_calc <-
+  df %>%
+  select(starts_with("CurrentEducation"),
+         -CurrentEducation_6_TEXT.pre) %>%
+  mutate_all(as_factor) %>%
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$groupType.pre_calc <-
+  df %>%
+  select(starts_with("gr_type"),
+          -ends_with("TEXT")) %>%
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$groupContext.pre_calc <-
+  df %>%
+  select(starts_with("gr_context"),
+          -ends_with("TEXT")) %>%
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$dyadType.pre_calc <-
+  df %>%
+  select(starts_with("dyad_type"),
+          -ends_with("TEXT")) %>%
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$context.pre_calc <-
+  df %>%
+  select(starts_with("Context"),
+          -ends_with("TEXT")) %>% 
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$desireType.pre_calc <-
+  df %>%
+  select(starts_with("desire_type"),
+          -ends_with("TEXT")) %>% 
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$desireType.pre_calc <-
+  df %>%
+  select(starts_with("desire_context"),
+          -ends_with("TEXT")) %>% 
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$nonDutchType.pre_calc <-
+  df %>%
+  select(starts_with("NonDutchType"),
+          -ends_with("TEXT")) %>% 
+  mutate_all(as_factor) %>% 
+  unite(.,
+        "Comb",
+        sep = ", ",
+        remove = TRUE,
+        na.rm = TRUE) %>%
+  pull
+
+df$autonomySat.pre_calc <-
+  psych::scoreItems(
+    keys = c(1, 1, 1, 1),
+    items = df %>% select(matches("^SDT0[1-4].pre")),
+    min = 1,
+    max = 7
+  )$scores %>% 
+  as.numeric
+df$autonomyFrust.pre_calc <-
+  psych::scoreItems(
+    keys = c(1, 1, 1, 1),
+    items = df %>% select(matches("^SDT0[5-8].pre")),
+    min = 1,
+    max = 7
+  )$scores %>% 
+  as.numeric
+
+df$relatednessSat.pre_calc <-
+  psych::scoreItems(
+    keys = c(1, 1, 1, 1),
+    items = df %>% select(matches("^SDT(09|1[0-2]).pre")),
+    min = 1,
+    max = 7
+  )$scores %>% 
+  as.numeric
+df$relatednessFrust.pre_calc <-
+  psych::scoreItems(
+    keys = c(1, 1, 1, 1),
+    items = df %>% select(matches("^SDT1[3-6].pre")),
+    min = 1,
+    max = 7
+  )$scores %>% 
+  as.numeric
+
+df$competenceSat.pre_calc <-
+  psych::scoreItems(
+    keys = c(1, 1, 1, 1),
+    items = df %>% select(matches("^SDT(1[7-9]|20).pre")),
+    min = 1,
+    max = 7
+  )$scores %>% 
+  as.numeric
+df$competenceFrust.pre_calc <-
+  psych::scoreItems(
+    keys = c(1, 1, 1, 1),
+    items = df %>% select(matches("^SDT2[1-4].pre")),
+    min = 1,
+    max = 7
+  )$scores %>% 
+  as.numeric
+
+df$IntGrAnx.pre_calc <-
+  psych::scoreItems(
+    keys = c(rep(1,6),rep(-1,6)),
+    items = df %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
+    min = 1,
+    max = 10
+  )$scores %>% 
+  as.numeric
+
+df$swl.pre_calc <-
+  psych::scoreItems(
+    keys = rep(1,5),
+    items = df %>% select(matches("^SWL[0-9]{2}\\.pre")),
+    min = 1,
+    max = 7,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>% 
+  as.numeric
+
+df$mdmq.alertness.pre_calc <- 
+  psych::scoreItems(
+    keys = c(1, -1, -1, 1),
+    items = df %>% select(MDMQ02.pre, MDMQ05.pre, MDMQ07.pre, MDMQ10.pre),
+    min = 1,
+    max = 6,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>% 
+  as.numeric
+df$mdmq.calmness.pre_calc <- 
+  psych::scoreItems(
+    keys = c(-1, 1, -1, 1),
+    items = df %>% select(MDMQ03.pre, MDMQ06.pre, MDMQ09.pre, MDMQ12.pre),
+    min = 1,
+    max = 6,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>% 
+  as.numeric
+df$mdmq.valence.pre_calc <- 
+  psych::scoreItems(
+    keys = c(1, -1, 1, -1),
+    items = df %>% select(MDMQ01.pre, MDMQ04.pre, MDMQ08.pre, MDMQ11.pre),
+    min = 1,
+    max = 6,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>% 
+  as.numeric
+
+df$IntGrAnx.post_calc <-
+  psych::scoreItems(
+    keys = c(rep(1,6),rep(-1,6)),
+    items = df %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
+    min = 1,
+    max = 10
+  )$scores %>% 
+  as.numeric
+
+df$swl.post_calc <-
+  psych::scoreItems(
+    keys = rep(1,5),
+    items = df %>% select(matches("^SWL[0-9]{2}\\.post")),
+    min = 1,
+    max = 7,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>% 
+  as.numeric
+
+varNamIndicesS1 <-
+  append(varNamIndicesS1,
+         df %>% select(ends_with("_calc")) %>% names)
+
 #
@@
 
-1. We first create person, survey type, and measurement ID variables.   
+1.  We first create person, survey type, and measurement ID variables.\
 2.
 We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).
-3. We add information about the interaction partner to the beep during which a person was selected as an interaction partner.   
-4. We cluster mean-center key variables within participants.   
+Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
+3.  We add information about the interaction partner to the beep during which a person was selected as an interaction partner.\
+4.  We cluster mean-center key variables within participants.\
 5.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
@@
 df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
-# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
+# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; JK: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
 df$relatedness <-
@@
 
-1. We first create person, survey type, and measurement ID variables.   
+1.  We first create person, survey type, and measurement ID variables.\
 2.
 We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).
-3. We cluster mean-center key variables within participants.   
+Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
+3.  We cluster mean-center key variables within participants.\
 4.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
@@
 
-<div class="alert alert-info alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-info-circle"></i> <b>Note: </b><br/>
-  This needs to be adjusted current missingness criterion set to more conservative 33%.
-</div>
+::: {.alert .alert-info .alert-dismissible .fade .in role="alert"}
+<a class="close" data-dismiss="alert" aria-label="close">×</a> <i class="fas fa-info-circle"></i> <b>Note: </b><br/> This needs to be adjusted current missingness criterion set to more conservative 33%.
+:::
 
@@
 ```{r testOptimizer05}
-cleanM <- function(M, c = 66) {
+cleanM <- function(M, c = 55) {
   check <- list()
@@
 
-dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 66)
+dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 55)
 
@@
 ```{r s2SampleReduction}
-dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 66)
+dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 55)
 
@@
 
-# remove glitch entry 
+# remove glitch entries 
 dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 57 & dtS2Red$ended == "2018-12-20 21:58:31"),]
+dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 59 & dtS2Red$ended == "2018-12-20 21:58:31"),]
 
@@
 ```{r s3SampleReduction}
-dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 66)
+dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 55)
 
@@
 
-We ultimately aim to create 20 imputed datasets to perform the analyses on.
-As outlined by @Monden2015, we will use the key variables themself as well as auxiliary variables to impute missing values.
+We ultimately aim to create 20 imputed data sets to perform the analyses on.
+As outlined by @Monden2015, we will use the key variables themselves as well as auxiliary variables to impute missing values.
 Auxiliary variables are any pre-, post-, or daily dairy variables other than the main variables.
-If there are any modeling issues we will limit the auxiliary variables to all variables that that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.
+If there are any modeling issues we will limit the auxiliary variables to all variables that that are significantly ($p$ \< .01) and meaningfully ($r$ \> .3) correlated with (1) the key variables or (2) missingness on the key variables.
 
 We create the imputed data set for each of the analyses.
-This includes the main analyses with across all three studies, using the set of common variables, as well as follow-up analyses for study-specific, interaction-specific and time scale specific datasets.
+This includes the main analyses with across all three studies, using the set of common variables, as well as follow-up analyses for study-specific, interaction-specific and time scale specific data sets.
 
 ```{r varNames}
-# varNames
-varNamInS123 <- varNames %>%
+varNamS123Aux <- varNames %>%
+  filter(
+    aux != 0,
+    studyS1 == "S1",
+    studyS2 == "S2",
+    studyS3 == "S3"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamS1Aux <- varNames %>%
+  filter(
+    aux != 0,
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamS2Aux <- varNames %>%
+  filter(
+    aux != 0,
+    studyS2 == "S2"
+  ) %>%
+  select(varNam) %>%
+  pull
+varNamS3Aux <- varNames %>%
+  filter(
+    aux != 0,
+    studyS3 == "S3"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+varNamS123PCA <- varNames %>%
   filter(
     pca != 0,
-    surveyS1 == "daily",
-    surveyS2 == "daily",
-    surveyS3 == "daily"
+    studyS1 == "S1",
+    studyS2 == "S2",
+    studyS3 == "S3"
   ) %>%
@@
   pull
-varNamInS1 <- varNames %>%
+varNamS1PCA <- varNames %>%
   filter(
     pca != 0,
-    surveyS1 == "daily"
+    studyS1 == "S1"
   ) %>%
@@
   pull
-varNamInS2 <- varNames %>%
+varNamS2PCA <- varNames %>%
   filter(
     pca != 0,
-    surveyS2 == "daily"
+    studyS2 == "S2"
   ) %>%
@@
   pull
-varNamInS3 <- varNames %>%
+varNamS3PCA <- varNames %>%
   filter(
     pca != 0,
-    surveyS3 == "daily"
+    studyS3 == "S3"
   ) %>%
@@
   pull
-varNamAdd <- c(
-  "PID",
-  "TID",
-  "TIDnum",
-  "relatedness",
-  "autonomy",
-  "competence"
-)
+
 varNamOut <- c(
@@
   "competence_NoInt",
-  varNamInS1[grepl('^MDMQ', varNamInS1)],
-  varNamInS2[grepl('^ProSo|^AntiSo', varNamInS2)],
-  varNamInS3[grepl(
+  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
+  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
+  varNamS3PCA[grepl(
     '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamInS3
+    varNamS3PCA
   )]
 )
+varNamCore <- c(
+  "PID",
+  "TID",
+  "TIDnum"
+)
 varNamNewS1 <- c(
+  "relatedness",
+  "autonomy",
+  "competence",
   "alertness", 
@@
 varNamNewS2 <- c(
+  "relatedness",
+  "autonomy",
+  "competence",
   "ProSo", 
@@
 varNamNewS3 <- c(
+  "relatedness",
+  "autonomy",
+  "competence",
   "ProSo", 
@@
 )
+
+varNamNewCat <- c(
+  "closeness_Calc",
+  "gender_Calc",
+  "ethnicity_Calc",
+  "relationship_Calc"
+)
+
 varNamIntDep <- c(
-  varNamInS123[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamInS123)]
-)
-
-varNamPcaS123 <- c(varNamAdd, varNamInS123[!varNamInS123 %in% varNamOut])
-varNamPcaS123Indep <- varNamPcaS123[!varNamPcaS123 %in% varNamIntDep]
-varNamPcaS1 <- c(varNamAdd, varNamNewS1, varNamInS1[!varNamInS1 %in% varNamOut])
-varNamPcaS2 <- c(varNamAdd, varNamNewS2, varNamInS2[!varNamInS2 %in% varNamOut])
-varNamPcaS3 <- c(varNamAdd, varNamNewS3, varNamInS3[!varNamInS3 %in% varNamOut])
+  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
+)
+
+varNamIndiceItemsS1 <- varNames %>%
+  filter(
+    aux == -1,
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull %>%
+  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
+  unique
+
+varNamS123MI <- c(varNamCore, varNamS123Aux)
+varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
+varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
+varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
+
+varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
+varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
+varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
+varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
 
@@
 ```{r MiMain}
-library(Amelia)
-varNamInMiMain <- varNames %>%
-  filter(
-    aux == 1,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamMiMain <- c(varNamAdd, varNamInMiMain[!varNamInMiMain %in% varNamOut])
-
-varNamInMiMainPsbl <- c(
-  paste(varNamMiMain, rep(".pre", length(varNamInMiMain)), sep = ""),
-  varNamMiMain,
-  paste(varNamMiMain, rep(".post", length(varNamInMiMain)), sep = "")
-)
-
-varNamInMiMainRed <- Reduce(
+
+varNamS123MIPsbl <- c(
+  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
+  varNamS123MI,
+  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
+)
+
+varNamS123MiRed <- Reduce(
   intersect,
   list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamInMiMainPsbl)) %>% names
+    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
+    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
+    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
   )
@@
 
+# ##################################
+# EXPORT FOR STM ML
+# ##################################
+#
+# varNamPosAdd <- c("relatedness", "autonomy", "competence")
+# 
+# dtAllStm <- rbind(
+#   dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>% 
+#     mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
+#   dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
+#   dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
+# ) %>%
+#   group_by(study, PID) %>%
+#   mutate(ID = cur_group_id()) %>%
+#   ungroup %>%
+#   mutate(
+#     date = as.Date(gsub(" .*", "", TID)),
+#     week = strftime(date, format = "%Y-W%V")
+#   ) %>%
+#   select(ID,
+#          PID,
+#          study,
+#          TID,
+#          TIDnum,
+#          date,
+#          week,
+#          IntergroupContact,
+#          IngroupContact,
+#          everything(),
+#          -starts_with("ResponseId")) %>%
+#   as.data.frame 
+# 
+# dtAllStm2 <- dtAllStm
+# 
+# dtAllStm2$IntGrAnx.pre_calc <-
+#   psych::scoreItems(
+#     keys = c(rep(1,6),rep(-1,6)),
+#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
+#     min = 1,
+#     max = 10
+#   )$scores %>% 
+#   as.numeric
+# 
+# dtAllStm2$IntGrAnx.post_calc <-
+#   psych::scoreItems(
+#     keys = c(rep(1,6),rep(-1,6)),
+#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
+#     min = 1,
+#     max = 10
+#   )$scores %>% 
+#   as.numeric
+# 
+# dtAllStm2$swl.pre_calc <-
+#   psych::scoreItems(
+#     keys = rep(1,5),
+#     items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
+#     min = 1,
+#     max = 7,
+#     missing = TRUE,
+#     impute = "none",
+#   )$scores %>% 
+#   as.numeric
+# 
+# dtAllStm2$discrimination.post_calc <-
+#   psych::scoreItems(
+#     keys = rep(1,9),
+#     items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
+#     min = 1,
+#     max = 6,
+#     missing = TRUE,
+#     impute = "none",
+#   )$scores %>% 
+#   as.numeric
+# 
+# dtAllStm3 <- dtAllStm2 %>%
+#   select(
+#     -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
+#     -matches("^IntGrAnx[0-9]{2}.?\\.post"),
+#     -matches("^SWL[0-9]{2}\\.pre"),
+#     -matches("^EvDayDiscr[0-9]{2}\\.post"),
+#     -relatednessSelf,
+#     -relatednessOther,
+#     -relatednessNoInteraction
+#   )
+# 
+# write.csv2(dtAllStm3, 
+#            file = "data/keyMotivesSTM.csv")
+# #######################################
+
 dtMiMain <- rbind(
-  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% mutate(across(!TID &
-                                                                                             !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
 ) %>%
@@
     date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V"),
-    roomate.pre = as.numeric(roomate.pre),
-    Reason.pre = as.numeric(Reason.pre),
-    association.pre = as.numeric(association.pre)
+    week = strftime(date, format = "%Y-W%V")
   ) %>%
@@
          -PID,
-         -starts_with("ResponseId")) %>%
+         -starts_with("ResponseId"), 
+         -starts_with("roommate"),
+         -starts_with("Reason")) %>%
   as.data.frame %>%
@@
 dtMiMainDaily <- rbind(
-  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% 
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
     mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
 ) %>%
@@
 dtMiMainWeeekly <- rbind(
-  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% 
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
     mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
 ) %>%
@@
 
-# MiMain <-
-#   amelia(
-#     dtMiMain,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMain, file = "data/imputed/MainImputed.RData")
+MiMain <-
+  amelia(
+    dtMiMain,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("TID", "date", "week", "study"), #"PID",
+    #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+    #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+    #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMain, file = "data/imputed/MainImputed.RData")
 load("data/imputed/MainImputed.RData")
 
-# MiMainDaily <-
-#   amelia(
-#     dtMiMainDaily,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("date", "study"), #"PID",
-#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
+MiMainDaily <-
+  amelia(
+    dtMiMainDaily,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("date", "study"), #"PID",
+    #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+    #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+    #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
 load("data/imputed/MainDailyImputed.RData")
 
-# MiMainWeekly <-
-#   amelia(
-#     dtMiMainWeeekly,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("week", "study"), #"PID",
-#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
-#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
+MiMainWeekly <-
+  amelia(
+    dtMiMainWeeekly,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("week", "study"), #"PID",
+    #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
+    #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
+    #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
 load("data/imputed/MainWeeklyImputed.RData")
@@
 
-```{r MiS1}
-varNamInMiS1 <- varNames %>%
-  filter(
-    aux == 1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamMiS1 <- c(varNamAdd, varNamInMiS1[!varNamInMiS1 %in% varNamOut])
-
-varNamInMiS1Psbl <- c(
-  paste(varNamMiS1, rep(".pre", length(varNamMiS1)), sep = ""),
-  varNamMiS1,
-  paste(varNamMiS1, rep(".post", length(varNamMiS1)), sep = "")
+```{r MiS1Prep}
+varNamS1MiPsbl <- c(
+  paste(varNamS1MI, rep(".pre", length(varNamS1MI)), sep = ""),
+  varNamS1MI,
+  paste(varNamS1MI, rep(".post", length(varNamS1MI)), sep = "")
 )
@@
 
-# TO BE ADAPTED:
 dtMiS1 <- dtS1Red %>% 
+  select(TID, any_of(varNamS1MiPsbl)) %>% 
+  mutate(across(!TID, as.numeric)) %>%
   mutate(
@@
   ) %>%
-  select(
-    PID,
-    TID,
-    TIDnum,
+  select(PID,
          date,
          week,
-    any_of(varNamInMiS1Psbl),
-    -time,
-    -starts_with("ResponseId")
-  ) %>%
+         everything(),
+         -starts_with("ResponseId")) %>%
   as.data.frame %>% 
-  mutate_at(vars(-c("TID", "date", "week")), as.numeric) %>%
-  select_if(~sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(PID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  ungroup %>% 
+  group_by(TIDnum) %>%
+  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
   select(PID,
@@
 
-#dtS1MiCs <- names(dtMiS1)[endsWith(names(dtMiS1), c(".pre", ".post"))] # cross sectional vars
+dtMiS1Daily <- dtS1Red %>% 
+  select(TID, any_of(varNamS1MiPsbl)) %>% 
+  mutate(across(!TID, as.numeric)) %>%
+  mutate(date = as.Date(gsub(" .*", "", TID))) %>%
+  group_by(PID, date) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>% 
+  mutate(TIDnum = as.numeric(factor(date)),
+         week = strftime(date, format = "%Y-W%V")) %>% 
+  select(PID,
+         date,
+         week,
+         everything(),
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>% 
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(PID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  ungroup %>% 
+  group_by(TIDnum) %>%
+  mutate(across(c("date", "week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
+  select(PID,
+         TIDnum,
+         date,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+  as.data.frame
+
+dtMiS1Weekly <- dtS1Red %>% 
+  select(TID, any_of(varNamS1MiPsbl)) %>% 
+  mutate(across(!TID, as.numeric)) %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  group_by(PID, week) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  ungroup %>% 
+  mutate(TIDnum = as.numeric(factor(week))) %>% 
+  select(PID,
+         week,
+         everything(),
+         -starts_with("ResponseId")) %>% 
+  as.data.frame %>% 
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(PID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  ungroup %>% 
+  group_by(TIDnum) %>%
+  mutate(across(c("week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
+  select(PID,
+         TIDnum,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+  as.data.frame 
+```
+
+```{r MiS1}
+allDuplicated <- function(vec){
+  front <- duplicated(vec)
+  back <- duplicated(vec, fromLast = TRUE)
+  all_dup <- front + back > 0
+  return(all_dup)
+}
+
+varNamDifS1 <- names(dtMiS1)[gsub(".pre|.post|_calc", "", names(dtMiS1)) %>% allDuplicated]
+
+dtMiS1 <- dtMiS1 %>%
+  mutate(
+    assimilation.dif = assimilation.post - assimilation.pre,
+    integration.dif = integration.post - integration.pre,
+    marginalization.dif = marginalization.post - marginalization.pre,
+    separation.dif = separation.post - separation.pre,
+    VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+    VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+    IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+    swl.dif = swl.post_calc - swl.pre_calc
+  ) %>% 
+  select(
+    -any_of(varNamDifS1)
+  ) %>%
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>%
+  as.data.frame
+
+dtMiS1Daily <- dtMiS1Daily %>%
+  mutate(
+    assimilation.dif = assimilation.post - assimilation.pre,
+    integration.dif = integration.post - integration.pre,
+    marginalization.dif = marginalization.post - marginalization.pre,
+    separation.dif = separation.post - separation.pre,
+    VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+    VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+    IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+    swl.dif = swl.post_calc - swl.pre_calc
+  ) %>% 
+  select(
+    -any_of(varNamDifS1)
+  ) %>%
+  select(PID,
+         TIDnum,
+         date,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>%
+  as.data.frame
+
+dtMiS1Weekly <- dtMiS1Weekly %>%
+  mutate(
+    assimilation.dif = assimilation.post - assimilation.pre,
+    integration.dif = integration.post - integration.pre,
+    marginalization.dif = marginalization.post - marginalization.pre,
+    separation.dif = separation.post - separation.pre,
+    VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+    VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+    IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+    swl.dif = swl.post_calc - swl.pre_calc
+  ) %>% 
+  select(
+    -any_of(varNamDifS1)
+  ) %>%
+  select(PID,
+         TIDnum,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>%
+  as.data.frame
+
+lares::corr_cross(
+  dtMiS1 %>% select_if(is.numeric), # name of dataset
+  max_pvalue = 0.05, # display only significant correlations (at 5% level)
+  top = 20 # display top 20 couples of variables (by correlation coefficient)
+)
+
+##----------------------------------------------------------------
+##                    IMPUTATIONS
+##----------------------------------------------------------------
 
@@
 #     dtMiS1,
-#     m = 1,
+#     m = 1, # 20
 #     ts = "TIDnum",
@@
 #     idvars = c("TID", "date", "week"),
-#     #noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     #parallel = "multicore",
-#     #ncpus = parallel::detectCores(),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
 #     p2s = 2
@@
 # load("data/imputed/S1Imputed.RData")
+# 
+# MiS1Daily <-
+#   amelia(
+#     dtMiS1Daily,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("date", "week"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS1Daily, file = "data/imputed/S1DailyImputed.RData")
+# load("data/imputed/S1DailyImputed.RData")
+# 
+# MiS1Weekly <-
+#   amelia(
+#     dtMiS1Weekly,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("week"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiS1Weekly, file = "data/imputed/S1WeeklyImputed.RData")
+# load("data/imputed/S1WeeklyImputed.RData")
 ```
@@
 dtPcaS123 <- rbind(
-  dtS1Red %>% select(any_of(varNamPcaS123)) %>% mutate(study = "S1"),
-  dtS2Red %>% select(any_of(varNamPcaS123)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamPcaS123)) %>% mutate(study = "S3")
+  dtS1Red %>% select(any_of(varNamS123PCA)) %>% mutate(study = "S1"),
+  dtS2Red %>% select(any_of(varNamS123PCA)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123PCA)) %>% mutate(study = "S3")
   ) %>%
@@
 dtPcaS123Imp <- MiMain$imputations[[1]] %>% 
-  select(ID, date, week, study, any_of(varNamPcaS123)) %>%
+  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
   arrange(ID, TIDnum) %>%
@@
 dtPcaS123DailyImp <- MiMainDaily$imputations[[1]] %>%
-  select(ID, date, study, any_of(varNamPcaS123)) %>%
+  select(ID, date, study, any_of(varNamS123PCA)) %>%
   arrange(ID, TIDnum) %>%
@@
 dtPcaS123WeeklyImp <- MiMainWeekly$imputations[[1]] %>%
-  select(ID, week, study, any_of(varNamPcaS123)) %>%
+  select(ID, week, study, any_of(varNamS123PCA)) %>%
   arrange(ID, TIDnum) %>%
@@
   ) %>%
+  arrange(variable, ID) %>%
+  #arrange(ID, variable, TIDnum) %>%
+  select(-TIDnum) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  replace(is.na(.), 0) %>%
+  select(-ID, -PID, -study)
+  #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
+  #pivot_wider(names_from = variable, values_from = value)
+
+# dtMainWideC %>%
+#   filter(
+#     ID == 54
+#   ) %>% 
+#   t
+# 
+# dtMainWideC$AttitudesDutch_59[dtMainWideC$ID == 54]
+
+dtMainWide <- dtPcaS123 %>%
+  select(
+    -TID,
+    -date,
+    -week
+  ) %>%
+  data.frame %>% 
+  melt(
+    .,
+    id=c("ID", "PID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+  ) %>%
   select(-TIDnum) %>%
@@
   select(-ID, -PID, -study)
-  #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
-  #pivot_wider(names_from = variable, values_from = value)
-
-dtMainWide <- dtPcaS123 %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  data.frame %>% 
-  melt(
-    .,
-    id=c("ID", "PID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
-  ) %>%
-  select(-TIDnum) %>%
-  arrange(variable, ID) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  replace(is.na(.), 0) %>%
-  select(-ID, -PID, -study)
 
@@
 
-## Three-Way ANOVA
+# **Three-Way ANOVA**
 
@@
 Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions.
-We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.
+We are particularly interested in the highest order interaction term Person \* Variable \* Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.
 
 ```{r threeWayAnovaMainBiDaily, message=FALSE, warning=FALSE}
-library(ThreeWay)
 capture.output(
@@
   dt = dtS1Red,
-  varNams = varNamPcaS1,
+  varNams = varNamS1PCA,
   id = "PID",
@@
   dt = dtS2Red,
-  varNams = varNamPcaS2,
+  varNams = varNamS2PCA,
   id = "PID",
@@
 
+sum(apply(dtS2Red, 2, is.finite))
+
 plotStat(
   dt = dtS2Red,
-  varNams = varNamPcaS2,
+  varNams = varNamS2PCA,
   id = "PID",
@@
   dt = dtS3Red,
-  varNams = varNamPcaS3,
+  varNams = varNamS3PCA,
   id = "PID",
@@
 ### Studies 1-3 (SDs) {.unlisted .unnumbered}
+
 <!-- <font size="5">Studies 1-3 (SDs)</font>  -->
@@
 # **Software Information**
+
 The full ResponseId information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
@@
 <details>
-  <summary>System Info</summary>
-
-  \renewcommand{\arraystretch}{0.8} <!-- decrease line spacing for the table -->
+
+<summary>
+
+System Info
+
+</summary>
+
+\renewcommand{\arraystretch}{0.8}
+
+<!-- decrease line spacing for the table -->
+
 ```{r Reproducibility-sessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
@@
 ```
-  \renewcommand{\arraystretch}{1} <!-- reset row height/line spacing -->
+
+\renewcommand{\arraystretch}{1}
+
+<!-- reset row height/line spacing -->
+
 </details>
+
 <br>
+
 <details>
-  <summary>Package Info</summary>
-
-\renewcommand{\arraystretch}{0.6} <!-- decrease line spacing for the table -->
+
+<summary>
+
+Package Info
+
+</summary>
+
+\renewcommand{\arraystretch}{0.6}
+
+<!-- decrease line spacing for the table -->
+
 ```{r Reproducibility-sessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
@@
 ```
-\renewcommand{\arraystretch}{1} <!-- reset row height/line spacing -->
+
+\renewcommand{\arraystretch}{1}
+
+<!-- reset row height/line spacing -->
+
 </details>
+
 <br>
+
 <details>
-  <summary>Full Session Info (including loaded but unattached packages --- for troubleshooting only)</summary>
+
+<summary>
+
+Full Session Info (including loaded but unattached packages --- for troubleshooting only)
+
+</summary>
+
 `r pander(sessionInfo(), compact = FALSE)`
+
 </details>
@@
 
-<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
+::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
+:::

### 2022-08-05 Supplemental-Material-A-Annotated-Analysis.Rmd
s1VarOut added for MI

@@
 
-MiMain <-
-  amelia(
-    dtMiMain,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("TID", "date", "week", "study"), #"PID",
-    #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-    #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-    #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMain, file = "data/imputed/MainImputed.RData")
+# MiMain <-
+#   amelia(
+#     dtMiMain,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMain, file = "data/imputed/MainImputed.RData")
 load("data/imputed/MainImputed.RData")
 
-MiMainDaily <-
-  amelia(
-    dtMiMainDaily,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("date", "study"), #"PID",
-    #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-    #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-    #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
+# MiMainDaily <-
+#   amelia(
+#     dtMiMainDaily,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("date", "study"), #"PID",
+#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
 load("data/imputed/MainDailyImputed.RData")
 
-MiMainWeekly <-
-  amelia(
-    dtMiMainWeeekly,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("week", "study"), #"PID",
-    #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
-    #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
-    #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
+# MiMainWeekly <-
+#   amelia(
+#     dtMiMainWeeekly,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("week", "study"), #"PID",
+#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
+#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
 load("data/imputed/MainWeeklyImputed.RData")
@@
 
+# Variables removed for study 1 due to high multicollinearity:
+varNamS1Out <- dtMiS1 %>%
+  select(
+    any_of(varNamDifS1),
+    mdmq.calmness.pre_calc,
+    mdmq.alertness.pre_calc,
+    mdmq.valence.pre_calc,
+    relatednessFrust.pre_calc,
+    competenceFrust.pre_calc,
+    competenceSat.pre_calc,
+    relatednessSat.pre_calc,
+    DutchFriends.pre,
+    secNatDum.pre,
+    timeNL.pre,
+    stress.post,
+    discrimination_month.post,
+    rosenberg.post,
+    qualityStar,
+    starts_with("SSAS")
+  ) %>%
+  names
+
 dtMiS1 <- dtMiS1 %>%
-  mutate(
-    assimilation.dif = assimilation.post - assimilation.pre,
-    integration.dif = integration.post - integration.pre,
-    marginalization.dif = marginalization.post - marginalization.pre,
-    separation.dif = separation.post - separation.pre,
-    VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-    VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-    IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-    swl.dif = swl.post_calc - swl.pre_calc
-  ) %>% 
+  # mutate(
+  #   assimilation.dif = assimilation.post - assimilation.pre,
+  #   integration.dif = integration.post - integration.pre,
+  #   marginalization.dif = marginalization.post - marginalization.pre,
+  #   separation.dif = separation.post - separation.pre,
+  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+  #   swl.dif = swl.post_calc - swl.pre_calc
+  # ) %>% 
   select(
-    -any_of(varNamDifS1)
+    -any_of(varNamS1Out)
   ) %>%
@@
 dtMiS1Daily <- dtMiS1Daily %>%
-  mutate(
-    assimilation.dif = assimilation.post - assimilation.pre,
-    integration.dif = integration.post - integration.pre,
-    marginalization.dif = marginalization.post - marginalization.pre,
-    separation.dif = separation.post - separation.pre,
-    VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-    VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-    IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-    swl.dif = swl.post_calc - swl.pre_calc
-  ) %>% 
+  # mutate(
+  #   assimilation.dif = assimilation.post - assimilation.pre,
+  #   integration.dif = integration.post - integration.pre,
+  #   marginalization.dif = marginalization.post - marginalization.pre,
+  #   separation.dif = separation.post - separation.pre,
+  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+  #   swl.dif = swl.post_calc - swl.pre_calc
+  # ) %>% 
   select(
-    -any_of(varNamDifS1)
+    -any_of(varNamS1Out)
   ) %>%
@@
 dtMiS1Weekly <- dtMiS1Weekly %>%
-  mutate(
-    assimilation.dif = assimilation.post - assimilation.pre,
-    integration.dif = integration.post - integration.pre,
-    marginalization.dif = marginalization.post - marginalization.pre,
-    separation.dif = separation.post - separation.pre,
-    VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-    VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-    IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-    swl.dif = swl.post_calc - swl.pre_calc
-  ) %>% 
+  # mutate(
+  #   assimilation.dif = assimilation.post - assimilation.pre,
+  #   integration.dif = integration.post - integration.pre,
+  #   marginalization.dif = marginalization.post - marginalization.pre,
+  #   separation.dif = separation.post - separation.pre,
+  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+  #   swl.dif = swl.post_calc - swl.pre_calc
+  # ) %>% 
   select(
-    -any_of(varNamDifS1)
+    -any_of(varNamS1Out)
   ) %>%
@@
 
+lares::corr_cross(
+  dtMiS1Daily %>% select_if(is.numeric), # name of dataset
+  max_pvalue = 0.05, # display only significant correlations (at 5% level)
+  top = 20 # display top 20 couples of variables (by correlation coefficient)
+)
+
 ##----------------------------------------------------------------
@@
 
-# MiS1 <-
-#   amelia(
-#     dtMiS1,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS1, file = "data/imputed/S1Imputed.RData")
-# load("data/imputed/S1Imputed.RData")
-# 
-# MiS1Daily <-
-#   amelia(
-#     dtMiS1Daily,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("date", "week"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS1Daily, file = "data/imputed/S1DailyImputed.RData")
-# load("data/imputed/S1DailyImputed.RData")
-# 
-# MiS1Weekly <-
-#   amelia(
-#     dtMiS1Weekly,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("week"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiS1Weekly, file = "data/imputed/S1WeeklyImputed.RData")
-# load("data/imputed/S1WeeklyImputed.RData")
+MiS1 <-
+  amelia(
+    dtMiS1,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "PID",
+    idvars = c("TID", "date", "week"),
+    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+    # leads = ...,
+    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+    # parallel = "multicore",
+    # ncpus = parallel::detectCores(),
+    p2s = 2
+  )
+save(MiS1, file = "data/imputed/S1Imputed.RData")
+load("data/imputed/S1Imputed.RData")
+
+MiS1Daily <-
+  amelia(
+    dtMiS1Daily,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "PID",
+    idvars = c("date", "week"),
+    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+    # leads = ...,
+    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+    # parallel = "multicore",
+    # ncpus = parallel::detectCores(),
+    p2s = 2
+  )
+save(MiS1Daily, file = "data/imputed/S1DailyImputed.RData")
+load("data/imputed/S1DailyImputed.RData")
+
+MiS1Weekly <-
+  amelia(
+    dtMiS1Weekly,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "PID",
+    idvars = c("week"),
+    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+    # leads = ...,
+    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+    # parallel = "multicore",
+    # ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiS1Weekly, file = "data/imputed/S1WeeklyImputed.RData")
+load("data/imputed/S1WeeklyImputed.RData")
 ```

### 2022-08-11 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 df$IntergroupContact <- (df$IntergroupContact-2)*-1
+df$IngroupContact <- (df$IngroupContact-2)*-1
 
@@
 )
+varNamNewAll <- c(
+  "relatedness",
+  "autonomy",
+  "competence"
+)
 varNamNewS1 <- c(
@@
 
-varNamS123MI <- c(varNamCore, varNamS123Aux)
+varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
 varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
@@
 
-```{r MiMain}
-
+```{r MiMainPrep}
+#################################################################
+##                    Variable Name Vectors                    ##
+#################################################################
 varNamS123MIPsbl <- c(
@@
 
-# ##################################
-# EXPORT FOR STM ML
-# ##################################
-#
-# varNamPosAdd <- c("relatedness", "autonomy", "competence")
-# 
-# dtAllStm <- rbind(
-#   dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>% 
-#     mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
-#   dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
-#   dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
-# ) %>%
-#   group_by(study, PID) %>%
-#   mutate(ID = cur_group_id()) %>%
-#   ungroup %>%
-#   mutate(
-#     date = as.Date(gsub(" .*", "", TID)),
-#     week = strftime(date, format = "%Y-W%V")
-#   ) %>%
-#   select(ID,
-#          PID,
-#          study,
-#          TID,
-#          TIDnum,
-#          date,
-#          week,
-#          IntergroupContact,
-#          IngroupContact,
-#          everything(),
-#          -starts_with("ResponseId")) %>%
-#   as.data.frame 
-# 
-# dtAllStm2 <- dtAllStm
-# 
-# dtAllStm2$IntGrAnx.pre_calc <-
-#   psych::scoreItems(
-#     keys = c(rep(1,6),rep(-1,6)),
-#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
-#     min = 1,
-#     max = 10
-#   )$scores %>% 
-#   as.numeric
-# 
-# dtAllStm2$IntGrAnx.post_calc <-
-#   psych::scoreItems(
-#     keys = c(rep(1,6),rep(-1,6)),
-#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
-#     min = 1,
-#     max = 10
-#   )$scores %>% 
-#   as.numeric
-# 
-# dtAllStm2$swl.pre_calc <-
-#   psych::scoreItems(
-#     keys = rep(1,5),
-#     items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
-#     min = 1,
-#     max = 7,
-#     missing = TRUE,
-#     impute = "none",
-#   )$scores %>% 
-#   as.numeric
-# 
-# dtAllStm2$discrimination.post_calc <-
-#   psych::scoreItems(
-#     keys = rep(1,9),
-#     items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
-#     min = 1,
-#     max = 6,
-#     missing = TRUE,
-#     impute = "none",
-#   )$scores %>% 
-#   as.numeric
-# 
-# dtAllStm3 <- dtAllStm2 %>%
-#   select(
-#     -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
-#     -matches("^IntGrAnx[0-9]{2}.?\\.post"),
-#     -matches("^SWL[0-9]{2}\\.pre"),
-#     -matches("^EvDayDiscr[0-9]{2}\\.post"),
-#     -relatednessSelf,
-#     -relatednessOther,
-#     -relatednessNoInteraction
-#   )
-# 
-# write.csv2(dtAllStm3, 
-#            file = "data/keyMotivesSTM.csv")
-# #######################################
-
-dtMiMain <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
+#################################################################
+##               EXPORT FOR STM MACHINE LEARNING               ##
+#################################################################
+# bannerCommenter::banner("EXPORT FOR STM MACHINE LEARNING", centre = TRUE, bandChar = "#")
+varNamPosAdd <- c("relatedness", "autonomy", "competence")
+
+dtAllStm <- rbind(
+  dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>%
+    mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
+  dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
+  dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
 ) %>%
@@
   ) %>%
+  select(ID,
+         PID,
+         study,
+         TID,
+         TIDnum,
+         date,
+         week,
+         IntergroupContact,
+         IngroupContact,
+         everything(),
+         -starts_with("ResponseId")) %>%
+  as.data.frame
+
+dtAllStm2 <- dtAllStm
+
+dtAllStm2$IntGrAnx.pre_calc <-
+  psych::scoreItems(
+    keys = c(rep(1,6),rep(-1,6)),
+    items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
+    min = 1,
+    max = 10
+  )$scores %>%
+  as.numeric
+
+dtAllStm2$IntGrAnx.post_calc <-
+  psych::scoreItems(
+    keys = c(rep(1,6),rep(-1,6)),
+    items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
+    min = 1,
+    max = 10
+  )$scores %>%
+  as.numeric
+
+dtAllStm2$swl.pre_calc <-
+  psych::scoreItems(
+    keys = rep(1,5),
+    items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
+    min = 1,
+    max = 7,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>%
+  as.numeric
+
+dtAllStm2$discrimination.post_calc <-
+  psych::scoreItems(
+    keys = rep(1,9),
+    items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
+    min = 1,
+    max = 6,
+    missing = TRUE,
+    impute = "none",
+  )$scores %>%
+  as.numeric
+
+dtAllStm3 <- dtAllStm2 %>%
+  select(
+    -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
+    -matches("^IntGrAnx[0-9]{2}.?\\.post"),
+    -matches("^SWL[0-9]{2}\\.pre"),
+    -matches("^EvDayDiscr[0-9]{2}\\.post"),
+    -relatednessSelf,
+    -relatednessOther,
+    -relatednessNoInteraction
+  )
+
+write.csv2(dtAllStm3,
+           file = "data/keyMotivesSTM.csv")
+#################################################################
+
+#################################################################
+##                    Dataframe Preparation                    ##
+#################################################################
+dtMiMain <- rbind(
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+  ) %>%
+  rowwise() %>%
+  mutate(
+    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
+    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum)
+  ) %>%
+  ungroup() %>%
   select(ID,
@@
   ) %>%
+  rowwise() %>%
+  mutate(
+    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
+    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum)
+  ) %>%
+  ungroup() %>%
   group_by(ID, date, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
+  summarise_if(is.numeric, mean, na.rm = TRUE) %>% # contact dum should be summarized as sum()
   ungroup %>%
@@
   ) %>%
+  rowwise() %>%
+  mutate(
+    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
+    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum)
+  ) %>%
+  ungroup() %>%
   group_by(ID, week, study) %>%
@@
   as.data.frame 
-
-# MiMain <-
-#   amelia(
-#     dtMiMain,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMain, file = "data/imputed/MainImputed.RData")
+```
+
+```{r MainImp}
+#################################################################
+##                     Imputation Bi-daily                     ##
+#################################################################
+MiMain <-
+  amelia(
+    dtMiMain,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+    #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+    #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+    #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+    noms = c("InteractionDum"),
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMain, file = "data/imputed/MainImputed.RData")
 load("data/imputed/MainImputed.RData")
 
-# MiMainDaily <-
-#   amelia(
-#     dtMiMainDaily,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("date", "study"), #"PID",
-#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
+##---------------------------------------------------------------
+##                Restore Missingness by Design                --
+##---------------------------------------------------------------
+
+# SHOULD STILL BE DONE? 
+# (remove imputed values that were missing by design. E.g., partner attitudes when there were no interactions)
+
+#################################################################
+##               Imputation at Daily Aggregation               ##
+#################################################################
+MiMainDaily <-
+  amelia(
+    dtMiMainDaily,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
+    #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+    #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+    #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
 load("data/imputed/MainDailyImputed.RData")
 
-# MiMainWeekly <-
-#   amelia(
-#     dtMiMainWeeekly,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("week", "study"), #"PID",
-#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
-#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
+##################################################################
+##               Imputation at Weekly Aggregation               ##
+##################################################################
+MiMainWeekly <-
+  amelia(
+    dtMiMainWeeekly,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "ID",
+    idvars = c("week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+    #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
+    #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
+    #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
+    parallel = "multicore",
+    ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
 load("data/imputed/MainWeeklyImputed.RData")
@@
 
-To be done once initial tests approved.
+In progress. Weekly might not be possible with all aux variables (i.e., too many variables for too few observations).
 
 ```{r MiS1Prep}
+#################################################################
+##                    Variable Name Vectors                    ##
+#################################################################
 varNamS1MiPsbl <- c(
@@
 
+#################################################################
+##                    Dataframe Preparation                    ##
+#################################################################
 dtMiS1 <- dtS1Red %>% 
@@
          which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
   as.data.frame
@@
   select(
-    -any_of(varNamS1Out)
+    -any_of(varNamS1Out),
+    -ends_with("_calc"),
+    -ContactNum,
+    -NonDutchNum,
+    -keyInteractionTime,
+    -InteractionContextCooperative,
+    -StudyFeedHonest.post,
+    -starts_with("NLE")
   ) %>%
@@
 
-##----------------------------------------------------------------
-##                    IMPUTATIONS
-##----------------------------------------------------------------
-
+lares::corr_cross(
+  dtMiS1Weekly %>% select_if(is.numeric), # name of dataset
+  max_pvalue = 0.05, # display only significant correlations (at 5% level)
+  top = 20 # display top 20 couples of variables (by correlation coefficient)
+)
+
+# Missingness Percentage
+# (colMeans(is.na(dtMiS1Weekly)))*100
+
+#################################################################
+##                     Imputation Bi-daily                     ##
+#################################################################
 MiS1 <-
@@
     cs = "PID",
-    idvars = c("TID", "date", "week"),
+    idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
@@
 
+#################################################################
+##               Imputation at Daily Aggregation               ##
+#################################################################
 MiS1Daily <-
@@
     cs = "PID",
-    idvars = c("date", "week"),
+    idvars = c("date", "week", "IntergroupContact", "IngroupContact"),
     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
@@
 
+##################################################################
+##               Imputation at Weekly Aggregation               ##
+##################################################################
 MiS1Weekly <-
@@
     cs = "PID",
-    idvars = c("week"),
+    idvars = c("week", "IntergroupContact", "IngroupContact"),
     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
@@
 varNam3mpcaMainAnalysis <- varNamesDailyFull %>%
-  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum")) %>%
+  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum", "startDate", "startTime")) %>%
   pull
 
+dtMainMi <- MiMain$imputations$imp1 %>%
+  rowwise() %>%
+  mutate(
+    InteractionDumOg = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
+    InteractionDumOg = if_else(InteractionDum > 0, 1, InteractionDum),
+    InteractionDumOg = if_else(sum(is.na(c(IntergroupContact, IngroupContact))) == 2, NA_real_, InteractionDum),
+    relatednessInteraction = mean(c(relatednessSelf, relatednessOther), na.rm = TRUE),
+    relatedness = if_else(InteractionDum == 1, relatednessInteraction, relatednessNoInteraction)
+    #relatednessNoInteraction = if_else(is.na(InteractionDumOg), NA_real_, relatednessNoInteraction)
+  ) %>%
+  ungroup()
+
 dtMain3mpcaNorm <-
-  PCAnorm(data = dtMain3mpca,
+  PCAnorm(data = MiMain$imputations$imp1,
           pid = "ID",
@@
   mutate(
-    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
   ) %>%
   select(-TIDnum, -study) %>%
-  arrange(variable, ID) %>%
+  arrange(ID, variable) %>%
   pivot_wider(names_from = variable, values_from = value) %>%
@@
     n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (B)
-    p = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) # Nr. variables (C)
+    m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
+    p = length(unique(dtPcaS123Imp$TIDnum)) # Nr. time points (C)
   ), 
@@
       "Person",
+      "Variable",
       "Time",
-      "Variable",
+      "Person*Variable",
       "Person*Time",
-      "Person*Variable",
-      "Time*Variable",
-      "Person*Time*Variable+error",
+      "Variable*Time",
+      "Person*Variable*Time+error",
       "Total"
@@
     n = nrow(dtMainDailyImpWide), # Nr. ppt.s (A)
-    m = length(unique(dtPcaS123DailyImp$TIDnum)), # Nr. time points (B)
-    p = length(names(dtPcaS123DailyImp)[!names(dtPcaS123DailyImp) %in% idVars]) # Nr. variables (C)
+    m = length(names(dtPcaS123DailyImp)[!names(dtPcaS123DailyImp) %in% idVars]), # Nr. variables (B)
+    p = length(unique(dtPcaS123DailyImp$TIDnum)) # Nr. time points (C)
   ), 
@@
     n = nrow(dtMainWeeklyImpWide), # Nr. ppt.s (A)
-    m = length(unique(dtPcaS123WeeklyImp$TIDnum)), # Nr. time points (B)
-    p = length(names(dtPcaS123WeeklyImp)[!names(dtPcaS123WeeklyImp) %in% idVars]) # Nr. variables (C)
+    m = length(names(dtPcaS123WeeklyImp)[!names(dtPcaS123WeeklyImp) %in% idVars]), # Nr. variables (B)
+    p = length(unique(dtPcaS123WeeklyImp$TIDnum)) # Nr. time points (C)
   ), 
@@
 
+# **Three-Mode PCA**
+
+```{r 3mpcaMain}
+#library("ThreeWay")
+
+nrow(dtMainImpWide)
+length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars])
+length(unique(dtPcaS123Imp$TIDnum))
+
+t3MainFull <- ThreeWay::T3(
+  dtMainImpWide, 
+  paste("PP", 1:nrow(dtMainImpWide), sep = "_"), 
+  names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
+  paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
+)
+
+t3Main <- ThreeWay::T3func(
+  X = dtMainImpWide,
+  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+  m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
+  p = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
+  r1 = 3,
+  r2 = 3,
+  r3 = 2,
+  start = 0,
+  conv = 1e-6
+)
+
+FitT3Main <- T3fitpartitioning(
+  Xprep = dtMainImpWide,
+  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+  m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
+  p = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
+  AS = t3Main$A, 
+  BT = t3Main$B, 
+  CU = t3Main$C, 
+  K = t3Main$H, 
+  renormmode = 0,
+  laba = paste("PP", 1:nrow(dtMainImpWide), sep = "_"), 
+  labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
+  labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
+)
+
+capture.output(
+  threeWayAnoBiDailyImp <- threewayanova(
+    Y = dtMainImpWide,
+    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+    m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
+    p = length(unique(dtPcaS123Imp$TIDnum)) # Nr. time points (C)
+  ), 
+file='NUL')
+
+rrcov3way::Tucker3(
+  X = dtMainImpWide, 
+  P = nrow(dtMainImpWide), # Nr. ppt.s (A)
+  Q = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
+  R = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
+  center = FALSE, #center.mode = c("A", "B", "C", "AB", "AC", "BC", "ABC"), 
+  scale = FALSE, #scale.mode = c("B", "A", "C"), 
+  conv = 1e-06, start="svd",
+  robust = FALSE, coda.transform=c("none", "ilr", "clr"), 
+  ncomp.rpca = 0, alpha = 0.75, robiter=100, crit=0.975, trace = FALSE
+)
+
+```
+
 # **Software Information**

### 2022-08-12 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 # bannerCommenter::banner("EXPORT FOR STM MACHINE LEARNING", centre = TRUE, bandChar = "#")
-varNamPosAdd <- c("relatedness", "autonomy", "competence")
-
-dtAllStm <- rbind(
-  dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
-  dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
-  dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  select(ID,
-         PID,
-         study,
-         TID,
-         TIDnum,
-         date,
-         week,
-         IntergroupContact,
-         IngroupContact,
-         everything(),
-         -starts_with("ResponseId")) %>%
-  as.data.frame
-
-dtAllStm2 <- dtAllStm
-
-dtAllStm2$IntGrAnx.pre_calc <-
-  psych::scoreItems(
-    keys = c(rep(1,6),rep(-1,6)),
-    items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
-    min = 1,
-    max = 10
-  )$scores %>%
-  as.numeric
-
-dtAllStm2$IntGrAnx.post_calc <-
-  psych::scoreItems(
-    keys = c(rep(1,6),rep(-1,6)),
-    items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
-    min = 1,
-    max = 10
-  )$scores %>%
-  as.numeric
-
-dtAllStm2$swl.pre_calc <-
-  psych::scoreItems(
-    keys = rep(1,5),
-    items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
-    min = 1,
-    max = 7,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>%
-  as.numeric
-
-dtAllStm2$discrimination.post_calc <-
-  psych::scoreItems(
-    keys = rep(1,9),
-    items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>%
-  as.numeric
-
-dtAllStm3 <- dtAllStm2 %>%
-  select(
-    -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
-    -matches("^IntGrAnx[0-9]{2}.?\\.post"),
-    -matches("^SWL[0-9]{2}\\.pre"),
-    -matches("^EvDayDiscr[0-9]{2}\\.post"),
-    -relatednessSelf,
-    -relatednessOther,
-    -relatednessNoInteraction
-  )
-
-write.csv2(dtAllStm3,
-           file = "data/keyMotivesSTM.csv")
+# varNamPosAdd <- c("relatedness", "autonomy", "competence")
+# 
+# dtAllStm <- rbind(
+#   dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>%
+#     mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
+#   dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
+#   dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
+# ) %>%
+#   group_by(study, PID) %>%
+#   mutate(ID = cur_group_id()) %>%
+#   ungroup %>%
+#   mutate(
+#     date = as.Date(gsub(" .*", "", TID)),
+#     week = strftime(date, format = "%Y-W%V")
+#   ) %>%
+#   select(ID,
+#          PID,
+#          study,
+#          TID,
+#          TIDnum,
+#          date,
+#          week,
+#          IntergroupContact,
+#          IngroupContact,
+#          everything(),
+#          -starts_with("ResponseId")) %>%
+#   as.data.frame
+# 
+# dtAllStm2 <- dtAllStm
+# 
+# dtAllStm2$IntGrAnx.pre_calc <-
+#   psych::scoreItems(
+#     keys = c(rep(1,6),rep(-1,6)),
+#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
+#     min = 1,
+#     max = 10
+#   )$scores %>%
+#   as.numeric
+# 
+# dtAllStm2$IntGrAnx.post_calc <-
+#   psych::scoreItems(
+#     keys = c(rep(1,6),rep(-1,6)),
+#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
+#     min = 1,
+#     max = 10
+#   )$scores %>%
+#   as.numeric
+# 
+# dtAllStm2$swl.pre_calc <-
+#   psych::scoreItems(
+#     keys = rep(1,5),
+#     items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
+#     min = 1,
+#     max = 7,
+#     missing = TRUE,
+#     impute = "none",
+#   )$scores %>%
+#   as.numeric
+# 
+# dtAllStm2$discrimination.post_calc <-
+#   psych::scoreItems(
+#     keys = rep(1,9),
+#     items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
+#     min = 1,
+#     max = 6,
+#     missing = TRUE,
+#     impute = "none",
+#   )$scores %>%
+#   as.numeric
+# 
+# dtAllStm3 <- dtAllStm2 %>%
+#   select(
+#     -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
+#     -matches("^IntGrAnx[0-9]{2}.?\\.post"),
+#     -matches("^SWL[0-9]{2}\\.pre"),
+#     -matches("^EvDayDiscr[0-9]{2}\\.post"),
+#     -relatednessSelf,
+#     -relatednessOther,
+#     -relatednessNoInteraction
+#   )
+# 
+# write.csv2(dtAllStm3,
+#            file = "data/keyMotivesSTM.csv")
 #################################################################
@@
 
+1.  Center across participants (within time point and variable).\
+2.  Normalize across time points and participants (within variable).\
+3.
+Select number of components for each mode (i.e., mode complexity; using Convex Hull procedure heuristic based on total number of components and scree test value taken from a PCA-SUP analysis [where, PCASUP analyses are PCA's of supermatrices with slices of the 3way array next to each other, thus 3 supermatrices are analyed by PCA, and for each we get a component matrix, and eigenvalues.
+The eigenvalues may give an indication as to the required number of components for each mode.]).
+4.  Run 3MPCA
+5.  Test
+
 ```{r 3mpcaMain}
 #library("ThreeWay")
+
+capture.output(
+  threeWayAnoBiDailyImp <- threewayanova(
+    Y = dtMainImpWide,
+    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainImpWide)))) # Nr. time points (C)
+  ), 
+file='NUL')
 
@@
 
-t3MainFull <- ThreeWay::T3(
-  dtMainImpWide, 
-  paste("PP", 1:nrow(dtMainImpWide), sep = "_"), 
-  names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
-  paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
-)
-
-t3Main <- ThreeWay::T3func(
+# t3MainFull <- ThreeWay::T3(
+#   dtMainImpWide,
+#   paste("PP", 1:nrow(dtMainImpWide), sep = "_"),
+#   names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
+#   paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
+# )
+
+dtMainImpWideC <- ThreeWay::cent3(
   X = dtMainImpWide, 
   n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-  m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
-  p = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
-  r1 = 3,
-  r2 = 3,
-  r3 = 2,
+  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
+  p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
+  mode = 1
+) %>%
+  as.data.frame
+#apply(dtMainImpWideC,2,sum)
+
+dtMainImpWideCZ <- ThreeWay::norm3(
+  X = dtMainImpWideC, 
+  n = nrow(dtMainImpWideC), # Nr. ppt.s (A)
+  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideC)))), # Nr. variables (B)
+  p = length(unique(sub("_.*", "", names(dtMainImpWideC)))), # Nr. time points (C)
+  mode = 2
+) %>%
+  as.data.frame
+names(dtMainImpWideCZ) <- names(dtMainImpWideC)
+#apply(dtMainImpWideC,2,sum)
+# dtMainImpWideCZ %>%
+#   select(ends_with("_AttitudesDutch")) %>%
+#   unlist %>%
+#   sd
+# psych::describe(dtMainImpWideCZ)
+
+MainPcaSup <- ThreeWay::T3runsApproxFit(
+  X = dtMainImpWideCZ, 
+  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
+  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
+  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
+  maxa = 20,
+  maxb = 20,
+  maxc = 20
+)
+MainDimSelector <- capture.output(
+  ThreeWay::DimSelector(
+    out = MainPcaSup,
+    n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
+    model = 2
+  )
+)
+
+MainDim <- tail(MainDimSelector, n = 1) %>%
+  str_extract("\\(.*\\)") %>%
+  stringr::str_extract_all('\\d+',simplify=TRUE) %>%
+  as.data.frame %>% 
+  mutate_all(as.numeric) %>%
+  dplyr::rename(c(
+    A = V1,
+    B = V2,
+    C = V3
+  ))
+MainDim
+
+t3Main <- ThreeWay::T3func(
+  X = dtMainImpWideCZ,
+  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
+  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
+  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
+  r1 = MainDim$A,
+  r2 = MainDim$B,
+  r3 = MainDim$C,
   start = 0,
@@
 FitT3Main <- T3fitpartitioning(
-  Xprep = dtMainImpWide,
-  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-  m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
-  p = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
+  Xprep = dtMainImpWideCZ,
+  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
+  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
+  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
   AS = t3Main$A, 
@@
   renormmode = 0,
-  laba = paste("PP", 1:nrow(dtMainImpWide), sep = "_"), 
+  laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"), 
   labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
@@
 
-capture.output(
-  threeWayAnoBiDailyImp <- threewayanova(
-    Y = dtMainImpWide,
-    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
-    p = length(unique(dtPcaS123Imp$TIDnum)) # Nr. time points (C)
-  ), 
-file='NUL')
-
-rrcov3way::Tucker3(
-  X = dtMainImpWide, 
-  P = nrow(dtMainImpWide), # Nr. ppt.s (A)
-  Q = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
-  R = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
-  center = FALSE, #center.mode = c("A", "B", "C", "AB", "AC", "BC", "ABC"), 
-  scale = FALSE, #scale.mode = c("B", "A", "C"), 
-  conv = 1e-06, start="svd",
-  robust = FALSE, coda.transform=c("none", "ilr", "clr"), 
-  ncomp.rpca = 0, alpha = 0.75, robiter=100, crit=0.975, trace = FALSE
-)
+t3MainRot <- varimcoco(
+  A = t3Main$A, 
+  B = t3Main$B, 
+  C = t3Main$C, 
+  H = t3Main$H,
+  wa_rel = 0,
+  wb_rel = 0,
+  wc_rel = 0
+)
+
+plot(t3MainRot$C[, 1], type = "n", ylim = range(t3MainRot$C), xlab = "Time occasion", ylab = "Component score")
+points(t3MainRot$C[, 1], pch = 16)
+lines(t3MainRot$C[, 1])
+points(t3MainRot$C[, 2], pch = 17)
+lines(t3MainRot$C[, 2])
+legend("topleft", legend = c("C1", "C2"), pch = c(16, 17), bty = "n")
+
+# library(factoextra)
+# fviz_nbclust(t3MainRot$A[,1:2], kmeans, method = "wss")
+# clusterdata <- t3MainRot$A[,1:2]
+# km.res <- kmeans(clusterdata, 6, nstart = 25)
+# fviz_cluster(object = km.res, # kmeans object 
+#              data = clusterdata, # data used for clustering
+#              ellipse.type = "norm",
+#              geom = "point",
+#              palette = "jco",
+#              main = "",
+#              xlab = "PC1",
+#              ylab = "PC2",
+#              ggtheme = theme_minimal())
+
+# lares::corr_cross(
+#   dtMainImpWideCZ, # name of dataset
+#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
+#   top = 50 # display top 50 couples of variables (by correlation coefficient)
+# )
 

### 2022-08-12 Supplemental-Material-A-Annotated-Analysis.Rmd
3MPCA rendered

@@
 #################################################################
-MiMain <-
-  amelia(
-    dtMiMain,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-    #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-    #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-    #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-    noms = c("InteractionDum"),
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMain, file = "data/imputed/MainImputed.RData")
+# set.seed(123)
+# MiMain <-
+#   amelia(
+#     dtMiMain,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMain, file = "data/imputed/MainImputed.RData")
 load("data/imputed/MainImputed.RData")
@@
 #################################################################
-MiMainDaily <-
-  amelia(
-    dtMiMainDaily,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
-    #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-    #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-    #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
+# MiMainDaily <-
+#   amelia(
+#     dtMiMainDaily,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
 load("data/imputed/MainDailyImputed.RData")
@@
 ##################################################################
-MiMainWeekly <-
-  amelia(
-    dtMiMainWeeekly,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "ID",
-    idvars = c("week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-    #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
-    #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
-    #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
-    parallel = "multicore",
-    ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
+# MiMainWeekly <-
+#   amelia(
+#     dtMiMainWeeekly,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
+#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
 load("data/imputed/MainWeeklyImputed.RData")
@@
 
-lares::corr_cross(
-  dtMiS1 %>% select_if(is.numeric), # name of dataset
-  max_pvalue = 0.05, # display only significant correlations (at 5% level)
-  top = 20 # display top 20 couples of variables (by correlation coefficient)
-)
-
-lares::corr_cross(
-  dtMiS1Daily %>% select_if(is.numeric), # name of dataset
-  max_pvalue = 0.05, # display only significant correlations (at 5% level)
-  top = 20 # display top 20 couples of variables (by correlation coefficient)
-)
-
-lares::corr_cross(
-  dtMiS1Weekly %>% select_if(is.numeric), # name of dataset
-  max_pvalue = 0.05, # display only significant correlations (at 5% level)
-  top = 20 # display top 20 couples of variables (by correlation coefficient)
-)
+# lares::corr_cross(
+#   dtMiS1 %>% select_if(is.numeric), # name of dataset
+#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
+#   top = 20 # display top 20 couples of variables (by correlation coefficient)
+# )
+# 
+# lares::corr_cross(
+#   dtMiS1Daily %>% select_if(is.numeric), # name of dataset
+#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
+#   top = 20 # display top 20 couples of variables (by correlation coefficient)
+# )
+# 
+# lares::corr_cross(
+#   dtMiS1Weekly %>% select_if(is.numeric), # name of dataset
+#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
+#   top = 20 # display top 20 couples of variables (by correlation coefficient)
+# )
 
@@
 #################################################################
-MiS1 <-
-  amelia(
-    dtMiS1,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "PID",
-    idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-    # leads = ...,
-    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-    # parallel = "multicore",
-    # ncpus = parallel::detectCores(),
-    p2s = 2
-  )
-save(MiS1, file = "data/imputed/S1Imputed.RData")
+# MiS1 <-
+#   amelia(
+#     dtMiS1,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS1, file = "data/imputed/S1Imputed.RData")
 load("data/imputed/S1Imputed.RData")
@@
 #################################################################
-MiS1Daily <-
-  amelia(
-    dtMiS1Daily,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "PID",
-    idvars = c("date", "week", "IntergroupContact", "IngroupContact"),
-    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-    # leads = ...,
-    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-    # parallel = "multicore",
-    # ncpus = parallel::detectCores(),
-    p2s = 2
-  )
-save(MiS1Daily, file = "data/imputed/S1DailyImputed.RData")
+# MiS1Daily <-
+#   amelia(
+#     dtMiS1Daily,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS1Daily, file = "data/imputed/S1DailyImputed.RData")
 load("data/imputed/S1DailyImputed.RData")
@@
 ##################################################################
-MiS1Weekly <-
-  amelia(
-    dtMiS1Weekly,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "PID",
-    idvars = c("week", "IntergroupContact", "IngroupContact"),
-    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-    # leads = ...,
-    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-    # parallel = "multicore",
-    # ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiS1Weekly, file = "data/imputed/S1WeeklyImputed.RData")
+# MiS1Weekly <-
+#   amelia(
+#     dtMiS1Weekly,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiS1Weekly, file = "data/imputed/S1WeeklyImputed.RData")
 load("data/imputed/S1WeeklyImputed.RData")
@@
 
-sum(apply(dtS2Red, 2, is.finite))
-
 plotStat(
@@
 
-1.  Center across participants (within time point and variable).\
-2.  Normalize across time points and participants (within variable).\
-3.
-Select number of components for each mode (i.e., mode complexity; using Convex Hull procedure heuristic based on total number of components and scree test value taken from a PCA-SUP analysis [where, PCASUP analyses are PCA's of supermatrices with slices of the 3way array next to each other, thus 3 supermatrices are analyed by PCA, and for each we get a component matrix, and eigenvalues.
-The eigenvalues may give an indication as to the required number of components for each mode.]).
-4.  Run 3MPCA
-5.  Test
-
-```{r 3mpcaMain}
-#library("ThreeWay")
-
-capture.output(
-  threeWayAnoBiDailyImp <- threewayanova(
-    Y = dtMainImpWide,
-    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWide)))) # Nr. time points (C)
-  ), 
-file='NUL')
-
-nrow(dtMainImpWide)
-length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars])
-length(unique(dtPcaS123Imp$TIDnum))
-
-# t3MainFull <- ThreeWay::T3(
-#   dtMainImpWide,
-#   paste("PP", 1:nrow(dtMainImpWide), sep = "_"),
-#   names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
-#   paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
-# )
-
+## Center and Normalize
+
+We begin by centering across participants (within time point and variable) and normalize across time points and participants (within variable).
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Question: </b><br/>
+  Are the centering and especially the normalization done correctly?
+</div>
+
+```{r 3mpcaMainPrep}
+##################################################################
+##                  Center across Participants                  ##
+##################################################################
 dtMainImpWideC <- ThreeWay::cent3(
@@
 
+##################################################################
+##                 Normalize across Time Points                 ##
+##################################################################
 dtMainImpWideCZ <- ThreeWay::norm3(
@@
 # psych::describe(dtMainImpWideCZ)
-
+```
+
+## Select Component Complexity
+
+We then select the optimal number of components for each mode (i.e., mode complexity).
+To do so, we use Convex Hull procedure heuristic based on total number of components and scree test value taken from a PCA-SUP analysis (where, PCASUP analyses are PCA's of supermatrices with slices of the 3way array next to each other, thus 3 supermatrices are analyed by PCA, and for each we get a component matrix, and eigenvalues.
+The eigenvalues may give an indication as to the required number of components for each mode.).
+
+```{r 3mpcaMainCompNum}
+# Run PCA SUP to determine best combination
+capture.output(
   MainPcaSup <- ThreeWay::T3runsApproxFit(
@@
     maxc = 20
-)
+  ), 
+file='NUL')
+
+# Select Dimension Complexity
 MainDimSelector <- capture.output(
@@
 )
-
 MainDim <- tail(MainDimSelector, n = 1) %>%
@@
   ))
-MainDim
-
+#MainDim
+```
+
+Based on the Convex Hull heuristic, we select `r MainDim$A` components for the participant mode, `r MainDim$B` principle components for the variable mode, and `r MainDim$C` components for the time mode.
+It should be noted that these heuristics are sensitive to variations in the multiple imputation procedure.
+
+## Model Fitting
+
+We then run the actual three-mode PCA with the inputs from the dimension selection procedure.
+
+```{r 3mpcaMainRun}
+
+# nrow(dtMainImpWide)
+# length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars])
+# length(unique(dtPcaS123Imp$TIDnum))
+# 
+# t3MainFull <- ThreeWay::T3(
+#   dtMainImpWide,
+#   paste("PP", 1:nrow(dtMainImpWide), sep = "_"),
+#   names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
+#   paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
+# )
+
+# 3MPCA with general trend
+capture.output(
+  t3MainRaw <- ThreeWay::T3func(
+    X = dtMainImpWide,
+    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
+    r1 = MainDim$A,
+    r2 = MainDim$B,
+    r3 = MainDim$C,
+    start = 0,
+    conv = 1e-6
+  ), 
+file='NUL')
+
+# 3MPCA without general trend
+capture.output(
   t3Main <- ThreeWay::T3func(
@@
     conv = 1e-6
-)
-
+  ), 
+file='NUL')
+
+# individual fit partitioning
 FitT3Main <- T3fitpartitioning(
@@
 
+# Split half analysis
+capture.output(
+  SplitT3Main <- splithalfT3(
+    X = dtMainImpWide,
+    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
+    r1 = MainDim$A,
+    r2 = MainDim$B,
+    r3 = MainDim$C, 
+    centopt = 1,
+    normopt = 2,
+    conv = 1e-6,
+    renormmode = 0,
+    addanal = 5,
+    wa_rel = 0,
+    wb_rel = 0,
+    wc_rel = 0,
+    laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"), 
+    labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
+    labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
+  ), 
+file='NUL')
+# 0
+# 0
+```
+
+In terms of fit percentage, we find that the three-mode PCA with the general trend included captures `r format(round(t3MainRaw$fp,2), nsmall = 2)`\% of the original variance, and the three-mode PCA without the general trend captures `r format(round(t3Main$fp,2), nsmall = 2)`\% of the variance.
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Question: </b><br/>
+  What is the interpretation of these Percentages? What exactly do the differences mean?
+</div>
+
+## Rotation
+
+For interpretability of the components we perform an orthomax rotation of the component matrices and the core array.
+
+```{r 3mpcaMainRotation}
+capture.output(
   t3MainRot <- varimcoco(
@@
     wc_rel = 0
-)
-
-plot(t3MainRot$C[, 1], type = "n", ylim = range(t3MainRot$C), xlab = "Time occasion", ylab = "Component score")
-points(t3MainRot$C[, 1], pch = 16)
-lines(t3MainRot$C[, 1])
-points(t3MainRot$C[, 2], pch = 17)
-lines(t3MainRot$C[, 2])
-legend("topleft", legend = c("C1", "C2"), pch = c(16, 17), bty = "n")
-
-# library(factoextra)
-# fviz_nbclust(t3MainRot$A[,1:2], kmeans, method = "wss")
-# clusterdata <- t3MainRot$A[,1:2]
-# km.res <- kmeans(clusterdata, 6, nstart = 25)
-# fviz_cluster(object = km.res, # kmeans object 
-#              data = clusterdata, # data used for clustering
-#              ellipse.type = "norm",
-#              geom = "point",
-#              palette = "jco",
-#              main = "",
-#              xlab = "PC1",
-#              ylab = "PC2",
-#              ggtheme = theme_minimal())
-
-# lares::corr_cross(
-#   dtMainImpWideCZ, # name of dataset
-#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
-#   top = 50 # display top 50 couples of variables (by correlation coefficient)
+  ),
+  file = 'NULL'
+)
+```
+
+```{r 3mpcaMainAMat}
+t3MainRot$AS %>%
+  as.data.frame %>%
+  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
+  mutate(Participant = paste("PP", row_number(), sep = "_")) %>%
+  select(
+    Participant, 
+    everything()
+  ) %>%
+  kbl(
+  .,
+  format = "html",
+  escape = FALSE,
+  align = c("l", rep("c",ncol(.)-1)),
+  booktabs = TRUE,
+  digits = 3,
+  caption = "Rotated Component Matrix: Person Mode" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+```{r 3mpcaMainBMat}
+t3MainRot$BT %>%
+  as.data.frame %>%
+  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
+  mutate(Variable = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) %>%
+  select(
+    Variable, 
+    everything()
+  ) %>%
+  kbl(
+  .,
+  format = "html",
+  escape = FALSE,
+  align = c("l", rep("c",ncol(.)-1)),
+  booktabs = TRUE,
+  digits = 3,
+  caption = "Rotated Component Matrix: Variable Mode" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+```{r 3mpcaMainCMat}
+t3MainRot$CU %>%
+  as.data.frame %>%
+  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
+  mutate(Time = paste("T", row_number()-1, sep = "")) %>%
+  select(
+    Time, 
+    everything()
+  ) %>%
+  kbl(
+  .,
+  format = "html",
+  escape = FALSE,
+  align = c("l", rep("c",ncol(.)-1)),
+  booktabs = TRUE,
+  digits = 3,
+  caption = "Rotated Component Matrix: Time Mode" # complete caption for main document
+) %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  ) %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+## Visualization: Time Components
+
+We visualize the `r ncol(t3MainRot$C)` time components in their component scores over time.
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  Not sure whether this interpretation is correct.
+</div>
+
+```{r 3mpcaMainCompVizTime}
+t3MainRot$C %>%
+  as.data.frame %>%
+  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
+  mutate(tid = row_number()) %>%
+  melt(., id.vars = "tid") %>%
+  ggplot(
+    .,
+    aes(
+      x = tid,
+      y = value,
+      linetype = variable,
+      shape = variable,
+      color = variable
+    )
+  ) +
+  geom_point() +
+  geom_line() +
+  geom_line(
+    stat = "smooth",
+    method = "loess",
+    se = FALSE,
+    span = .25,
+    alpha = 0.75
+  ) +
+  labs(x = "Time occasion",
+       y = "Component score",
+       linetype = "Component",
+       shape = "Component",
+       color = "Component") +
+  theme_Publication()
+
+```
+
+## Visualization: Person Components
+
+We visualize how well the first two person components separate the participants.
+We focus on the first two components because they capture the most variance, while retaining easy two dimensional visualization options.
+We additionally color the participants using k means clustering to further showcase the separation.
+
+```{r 3mpcaMainCompVizPerson}
+# # Unsure how this should be interpreted 
+# jointplotgen(
+#   K = t3MainRot$K,
+#   A = t3MainRot$A,
+#   B = t3MainRot$B,
+#   C = t3MainRot$C,
+#   fixmode = 1,
+#   fixunit = 3,
+#   laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"),
+#   labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
+#   labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
 # )
 
-```
+library(factoextra)
+fviz_nbclust(t3MainRot$A[,1:2], kmeans, method = "wss")
+clusterdata <- t3MainRot$A[,1:2]
+km.res <- kmeans(clusterdata, 3, nstart = 25)
+fviz_cluster(object = km.res, # kmeans object
+             data = clusterdata, # data used for clustering
+             ellipse.type = "norm",
+             geom = "point",
+             palette = "jco",
+             main = "",
+             xlab = "Person Component 1",
+             ylab = "Person Component 2",
+             ggtheme = theme_Publication())
+```
+
+## Visualization: Variable Components
+
+To visualize how the principle components separate the different variables, we draw the component scores for all variables.
+
+```{r 3mpcaMainCompVizVariable}
+t3MainRot$B %>%
+  as.data.frame %>%
+  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
+  mutate(var = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) %>%
+  melt(., id.vars = "var") %>%
+  ggplot(
+    .,
+    aes(
+      x = var,
+      y = value,
+      group = variable,
+      linetype = variable,
+      shape = variable,
+      color = variable
+    )
+  ) +
+  geom_point() +
+  geom_line() +
+  coord_flip() +
+  labs(x = "Variable",
+       y = "Component score",
+       linetype = "Component",
+       shape = "Component",
+       color = "Component") +
+  theme_Publication()
+```
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  How do we best interpret and visualize the joint decomposition? 
+</div>
 

### 2022-08-20 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 
-The full ResponseId information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
+The full session information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
 

### 2022-08-22 Supplemental-Material-A-Annotated-Analysis.Rmd
initial stability checks added

@@
 ```{r testOptimizer05}
-cleanM <- function(M, c = 55) {
+cleanM <- function(M, c = 55) { #c = 66
   check <- list()
@@
 
-### Studies 1-3
-
-We begin by preparing the data for the main analysis. Here, we:
+### Main Analysis
+
+We begin by preparing the data for the main analysis, which combines the common variables of all three studies. Here, we:
 
@@
   as.data.frame 
-
+```
+
+```{r MainImp}
+#################################################################
+##                     Imputation Bi-daily                     ##
+#################################################################
+# set.seed(123)
+# MiMain <-
+#   amelia(
+#     dtMiMain,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMain, file = "data/imputed/MainImputed.RData")
+load("data/imputed/MainImputed.RData")
+
+# # Stability to imputations:
+# set.seed(123)
+# MiMainStability <-
+#   amelia(
+#     dtMiMain,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c(
+#       "TID",
+#       "date",
+#       "week",
+#       "study",
+#       "IntergroupContact",
+#       "IngroupContact"
+#     ),
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainStability, file = "data/imputed/MainImputedStability.RData")
+load("data/imputed/MainImputedStability.RData")
+
+##---------------------------------------------------------------
+##                Restore Missingness by Design                --
+##---------------------------------------------------------------
+
+# SHOULD STILL BE DONE? 
+# (remove imputed values that were missing by design. E.g., partner attitudes when there were no interactions)
+
+```
+
+### Follow-up: Studies
+
+#### Study 1
+
+```{r MiS1Prep}
+#################################################################
+##                    Variable Name Vectors                    ##
+#################################################################
+varNamS1MiPsbl <- c(
+  paste(varNamS1MI, rep(".pre", length(varNamS1MI)), sep = ""),
+  varNamS1MI,
+  paste(varNamS1MI, rep(".post", length(varNamS1MI)), sep = "")
+)
+
+varNamdtS1Noms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS1 == "S1"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+#################################################################
+##                    Dataframe Preparation                    ##
+#################################################################
+dtMiS1 <- dtS1Red %>% 
+  select(TID, any_of(varNamS1MiPsbl)) %>% 
+  mutate(across(!TID, as.numeric)) %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+  ) %>%
+  select(PID,
+         date,
+         week,
+         everything(),
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>% 
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(PID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  ungroup %>% 
+  group_by(TIDnum) %>%
+  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+  as.data.frame
+
+allDuplicated <- function(vec){
+  front <- duplicated(vec)
+  back <- duplicated(vec, fromLast = TRUE)
+  all_dup <- front + back > 0
+  return(all_dup)
+}
+
+varNamDifS1 <- names(dtMiS1)[gsub(".pre|.post|_calc", "", names(dtMiS1)) %>% allDuplicated]
+
+# Variables removed for study 1 due to high multicollinearity:
+varNamS1Out <- dtMiS1 %>%
+  select(
+    any_of(varNamDifS1),
+    mdmq.calmness.pre_calc,
+    mdmq.alertness.pre_calc,
+    mdmq.valence.pre_calc,
+    relatednessFrust.pre_calc,
+    competenceFrust.pre_calc,
+    competenceSat.pre_calc,
+    relatednessSat.pre_calc,
+    DutchFriends.pre,
+    secNatDum.pre,
+    timeNL.pre,
+    stress.post,
+    discrimination_month.post,
+    rosenberg.post,
+    qualityStar,
+    starts_with("SSAS")
+  ) %>%
+  names
+
+dtMiS1 <- dtMiS1 %>%
+  # mutate(
+  #   assimilation.dif = assimilation.post - assimilation.pre,
+  #   integration.dif = integration.post - integration.pre,
+  #   marginalization.dif = marginalization.post - marginalization.pre,
+  #   separation.dif = separation.post - separation.pre,
+  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+  #   swl.dif = swl.post_calc - swl.pre_calc
+  # ) %>% 
+  select(
+    -any_of(varNamS1Out)
+  ) %>%
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  as.data.frame
+
+# lares::corr_cross(
+#   dtMiS1 %>% select_if(is.numeric), # name of dataset
+#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
+#   top = 20 # display top 20 couples of variables (by correlation coefficient)
+# )
+```
+
+```{r MiS1}
+#################################################################
+##                     Imputation Bi-daily                     ##
+#################################################################
+# set.seed(123)
+# MiS1 <-
+#   amelia(
+#     dtMiS1,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiS1, file = "data/imputed/S1Imputed.RData")
+load("data/imputed/S1Imputed.RData")
+
+# # Imputation Stability:
+# set.seed(123)
+# MiS1Stability <-
+#   amelia(
+#     dtMiS1,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS1Stability, file = "data/imputed/S1ImputedStability.RData")
+load("data/imputed/S1ImputedStability.RData")
+```
+
+#### Study 2
+
+To be done once initial tests approved.
+
+```{r MiS2Prep}
+#################################################################
+##                    Variable Name Vectors                    ##
+#################################################################
+varNamS2MiPsbl <- c(
+  paste(varNamS2MI, rep(".pre", length(varNamS2MI)), sep = ""),
+  varNamS2MI,
+  paste(varNamS2MI, rep(".post", length(varNamS2MI)), sep = "")
+)
+
+varNamdtS2Noms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS2 == "S2"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+#################################################################
+##                    Dataframe Preparation                    ##
+#################################################################
+dtMiS2 <- dtS2Red %>% 
+  select(TID, any_of(varNamS2MiPsbl)) %>% 
+  mutate(across(!TID, as.numeric)) %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+  ) %>%
+  select(PID,
+         date,
+         week,
+         everything(),
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>% 
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(PID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  ungroup %>% 
+  group_by(TIDnum) %>%
+  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         any_of(names(which(sapply(dtMiS2 %>% select(where(is.numeric)), sd, na.rm = TRUE) > 0)))) %>% # only vars that sd != 0
+         #which(sapply(. %>% select(where(is.numeric)), sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+  as.data.frame
+
+varNamDifS2 <- names(dtMiS2)[gsub(".pre|.post|_calc", "", names(dtMiS2)) %>% allDuplicated]
+
+# # Variables removed for study 2 due to high multicollinearity:
+# varNamS1Out <- dtMiS1 %>%
+#   select(
+#     any_of(varNamDifS1),
+#     mdmq.calmness.pre_calc,
+#     mdmq.alertness.pre_calc,
+#     mdmq.valence.pre_calc,
+#     relatednessFrust.pre_calc,
+#     competenceFrust.pre_calc,
+#     competenceSat.pre_calc,
+#     relatednessSat.pre_calc,
+#     DutchFriends.pre,
+#     secNatDum.pre,
+#     timeNL.pre,
+#     stress.post,
+#     discrimination_month.post,
+#     rosenberg.post,
+#     qualityStar,
+#     starts_with("SSAS")
+#   ) %>%
+#   names
+
+# dtMiS2Fin <- dtMiS2 %>%
+#   # mutate(
+#   #   assimilation.dif = assimilation.post - assimilation.pre,
+#   #   integration.dif = integration.post - integration.pre,
+#   #   marginalization.dif = marginalization.post - marginalization.pre,
+#   #   separation.dif = separation.post - separation.pre,
+#   #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+#   #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+#   #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+#   #   swl.dif = swl.post_calc - swl.pre_calc
+#   # ) %>% 
+#   # select(
+#   #   -any_of(varNamS2Out)
+#   # ) %>%
+#   select(PID,
+#          TID,
+#          TIDnum,
+#          date,
+#          week,
+#          which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
+#   mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+#   as.data.frame
+
+# lares::corr_cross(
+#   dtMiS2 %>% select_if(is.numeric), # name of dataset
+#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
+#   top = 20 # display top 20 couples of variables (by correlation coefficient)
+# )
+```
+
+```{r MiS1}
+#################################################################
+##                     Imputation Bi-daily                     ##
+#################################################################
+set.seed(123)
+MiS2 <-
+  amelia(
+    dtMiS2,
+    m = 1, # 20
+    ts = "TIDnum",
+    cs = "PID",
+    idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+    # leads = ...,
+    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+    # parallel = "multicore",
+    # ncpus = parallel::detectCores(),
+    p2s = 1
+  )
+save(MiS2, file = "data/imputed/S2Imputed.RData")
+load("data/imputed/S2Imputed.RData")
+
+# # Imputation Stability:
+# set.seed(123)
+# MiS1Stability <-
+#   amelia(
+#     dtMiS1,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS1Stability, file = "data/imputed/S1ImputedStability.RData")
+load("data/imputed/S1ImputedStability.RData")
+```
+
+#### Study 3
+
+To be done once initial tests approved.
+
+```{r MiS3}
+# BASED ON MiS1
+```
+
+### Follow-up: Time Scales
+
+#### Daily
+
+Preparing the data for the time scale follow-up, which combines the common variables of all three studies on a daily level. Here, we:
+
+1.  Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
+2.  Create the combined dataset in the correct format (long format with all missed time points as NAs).
+3.  Impute the data and save the imputed datasets for later recall.
+
+```{r MiDailyPrep}
+#################################################################
+##                    Dataframe Preparation                    ##
+#################################################################
 dtMiMainDaily <- rbind(
@@
   as.data.frame 
-
+```
+
+```{r DailyImp}
+#################################################################
+##               Imputation at Daily Aggregation               ##
+#################################################################
+# MiMainDaily <-
+#   amelia(
+#     dtMiMainDaily,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
+load("data/imputed/MainDailyImputed.RData")
+
+# # Stability to imputations:
+# set.seed(123)
+# MiMainDailyStability <-
+#   amelia(
+#     dtMiMainDaily,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
+#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainDailyStability, file = "data/imputed/MainDailyImputedStability.RData")
+load("data/imputed/MainDailyImputedStability.RData")
+```
+
+#### Weekly
+
+Preparing the data for the time scale follow-up, which combines the common variables of all three studies on a weekly level. Here, we:
+
+1.  Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
+2.  Create the combined dataset in the correct format (long format with all missed time points as NAs).
+3.  Impute the data and save the imputed datasets for later recall.
+
+```{r MiWeeklyPrep}
 dtMiMainWeeekly <- rbind(
@@
 
-```{r MainImp}
-#################################################################
-##                     Imputation Bi-daily                     ##
-#################################################################
-# set.seed(123)
-# MiMain <-
-#   amelia(
-#     dtMiMain,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMain, file = "data/imputed/MainImputed.RData")
-load("data/imputed/MainImputed.RData")
-
-##---------------------------------------------------------------
-##                Restore Missingness by Design                --
-##---------------------------------------------------------------
-
-# SHOULD STILL BE DONE? 
-# (remove imputed values that were missing by design. E.g., partner attitudes when there were no interactions)
-
-#################################################################
-##               Imputation at Daily Aggregation               ##
-#################################################################
-# MiMainDaily <-
-#   amelia(
-#     dtMiMainDaily,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
-load("data/imputed/MainDailyImputed.RData")
-
+```{r WeeklyImp}
 ##################################################################
@@
 load("data/imputed/MainWeeklyImputed.RData")
-```
-
-### Study 1
-
-In progress. Weekly might not be possible with all aux variables (i.e., too many variables for too few observations).
-
-```{r MiS1Prep}
-#################################################################
-##                    Variable Name Vectors                    ##
-#################################################################
-varNamS1MiPsbl <- c(
-  paste(varNamS1MI, rep(".pre", length(varNamS1MI)), sep = ""),
-  varNamS1MI,
-  paste(varNamS1MI, rep(".post", length(varNamS1MI)), sep = "")
-)
-
-varNamdtS1Noms <- varNames %>%
-  filter(
-    aux == 1,
-    format == "nominal",
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-
+
+# # Stability to imputations:
+# set.seed(123)
+# MiMainWeeklyStability <-
+#   amelia(
+#     dtMiMainWeeekly,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
+#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiMainWeeklyStability, file = "data/imputed/MainWeeklyImputedStability.RData")
+load("data/imputed/MainWeeklyImputedStability.RData")
+```
+
+### Follow-up: Response Type
+
+#### Interaction Responses
+
+```{r MiInteractionPrep}
 #################################################################
@@
 #################################################################
-dtMiS1 <- dtS1Red %>% 
-  select(TID, any_of(varNamS1MiPsbl)) %>% 
-  mutate(across(!TID, as.numeric)) %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  select(PID,
-         date,
-         week,
-         everything(),
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>% 
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(PID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  ungroup %>% 
-  group_by(TIDnum) %>%
-  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
-  as.data.frame
-
-dtMiS1Daily <- dtS1Red %>% 
-  select(TID, any_of(varNamS1MiPsbl)) %>% 
-  mutate(across(!TID, as.numeric)) %>%
-  mutate(date = as.Date(gsub(" .*", "", TID))) %>%
-  group_by(PID, date) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>% 
-  mutate(TIDnum = as.numeric(factor(date)),
-         week = strftime(date, format = "%Y-W%V")) %>% 
-  select(PID,
-         date,
-         week,
-         everything(),
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>% 
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(PID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  ungroup %>% 
-  group_by(TIDnum) %>%
-  mutate(across(c("date", "week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
-  select(PID,
-         TIDnum,
-         date,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
-  as.data.frame
-
-dtMiS1Weekly <- dtS1Red %>% 
-  select(TID, any_of(varNamS1MiPsbl)) %>% 
-  mutate(across(!TID, as.numeric)) %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  group_by(PID, week) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>% 
-  mutate(TIDnum = as.numeric(factor(week))) %>% 
-  select(PID,
-         week,
-         everything(),
-         -starts_with("ResponseId")) %>% 
-  as.data.frame %>% 
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(PID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  ungroup %>% 
-  group_by(TIDnum) %>%
-  mutate(across(c("week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
-  select(PID,
-         TIDnum,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
-  as.data.frame 
-```
-
-```{r MiS1}
-allDuplicated <- function(vec){
-  front <- duplicated(vec)
-  back <- duplicated(vec, fromLast = TRUE)
-  all_dup <- front + back > 0
-  return(all_dup)
-}
-
-varNamDifS1 <- names(dtMiS1)[gsub(".pre|.post|_calc", "", names(dtMiS1)) %>% allDuplicated]
-
-# Variables removed for study 1 due to high multicollinearity:
-varNamS1Out <- dtMiS1 %>%
-  select(
-    any_of(varNamDifS1),
-    mdmq.calmness.pre_calc,
-    mdmq.alertness.pre_calc,
-    mdmq.valence.pre_calc,
-    relatednessFrust.pre_calc,
-    competenceFrust.pre_calc,
-    competenceSat.pre_calc,
-    relatednessSat.pre_calc,
-    DutchFriends.pre,
-    secNatDum.pre,
-    timeNL.pre,
-    stress.post,
-    discrimination_month.post,
-    rosenberg.post,
-    qualityStar,
-    starts_with("SSAS")
-  ) %>%
-  names
-
-dtMiS1 <- dtMiS1 %>%
-  # mutate(
-  #   assimilation.dif = assimilation.post - assimilation.pre,
-  #   integration.dif = integration.post - integration.pre,
-  #   marginalization.dif = marginalization.post - marginalization.pre,
-  #   separation.dif = separation.post - separation.pre,
-  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-  #   swl.dif = swl.post_calc - swl.pre_calc
-  # ) %>% 
-  select(
-    -any_of(varNamS1Out)
-  ) %>%
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame
-
-dtMiS1Daily <- dtMiS1Daily %>%
-  # mutate(
-  #   assimilation.dif = assimilation.post - assimilation.pre,
-  #   integration.dif = integration.post - integration.pre,
-  #   marginalization.dif = marginalization.post - marginalization.pre,
-  #   separation.dif = separation.post - separation.pre,
-  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-  #   swl.dif = swl.post_calc - swl.pre_calc
-  # ) %>% 
-  select(
-    -any_of(varNamS1Out)
-  ) %>%
-  select(PID,
-         TIDnum,
-         date,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame
-
-dtMiS1Weekly <- dtMiS1Weekly %>%
-  # mutate(
-  #   assimilation.dif = assimilation.post - assimilation.pre,
-  #   integration.dif = integration.post - integration.pre,
-  #   marginalization.dif = marginalization.post - marginalization.pre,
-  #   separation.dif = separation.post - separation.pre,
-  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-  #   swl.dif = swl.post_calc - swl.pre_calc
-  # ) %>% 
-  select(
-    -any_of(varNamS1Out),
-    -ends_with("_calc"),
-    -ContactNum,
-    -NonDutchNum,
-    -keyInteractionTime,
-    -InteractionContextCooperative,
-    -StudyFeedHonest.post,
-    -starts_with("NLE")
-  ) %>%
-  select(PID,
-         TIDnum,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame
-
-# lares::corr_cross(
-#   dtMiS1 %>% select_if(is.numeric), # name of dataset
-#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
-#   top = 20 # display top 20 couples of variables (by correlation coefficient)
-# )
+
+# # any interaction
+# dtMiMain %>%
+#   filter(is.na(InteractionDum) | InteractionDum != 0) %>%
+#   #complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
+#   nrow
 # 
-# lares::corr_cross(
-#   dtMiS1Daily %>% select_if(is.numeric), # name of dataset
-#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
-#   top = 20 # display top 20 couples of variables (by correlation coefficient)
-# )
+# # Intergroup Contact
+# dtMiMain %>%
+#   filter(is.na(IntergroupContact) | IntergroupContact != 0) %>%
+#   #complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
+#   nrow
 # 
-# lares::corr_cross(
-#   dtMiS1Weekly %>% select_if(is.numeric), # name of dataset
-#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
-#   top = 20 # display top 20 couples of variables (by correlation coefficient)
-# )
-
-# Missingness Percentage
-# (colMeans(is.na(dtMiS1Weekly)))*100
-
-#################################################################
-##                     Imputation Bi-daily                     ##
-#################################################################
-# MiS1 <-
-#   amelia(
-#     dtMiS1,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS1, file = "data/imputed/S1Imputed.RData")
-load("data/imputed/S1Imputed.RData")
-
-#################################################################
-##               Imputation at Daily Aggregation               ##
-#################################################################
-# MiS1Daily <-
-#   amelia(
-#     dtMiS1Daily,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS1Daily, file = "data/imputed/S1DailyImputed.RData")
-load("data/imputed/S1DailyImputed.RData")
-
-##################################################################
-##               Imputation at Weekly Aggregation               ##
-##################################################################
-# MiS1Weekly <-
-#   amelia(
-#     dtMiS1Weekly,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiS1Weekly, file = "data/imputed/S1WeeklyImputed.RData")
-load("data/imputed/S1WeeklyImputed.RData")
-```
-
-### Study 2
-
-To be done once initial tests approved.
-
-```{r MiS2}
-# BASED ON MiS1
-```
-
-### Study 3
-
-To be done once initial tests approved.
-
-```{r MiS3}
-# BASED ON MiS1
+# # Ingroup Contact
+# dtMiMain %>%
+#   filter(is.na(IngroupContact) | IngroupContact != 0) %>%
+#   #complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
+#   nrow
+
+```
+
+```{r InteractionImp}
+# TBD
+```
+
+#### Non-Interaction Responses
+
+```{r MiNoInteractionPrep}
+# TBD
+```
+
+```{r NoInteractionImp}
+# TBD
 ```
@@
 
-```{r longToWide}
+### Main Analysis
+
+```{r longToWideMain}
 dtPcaS123 <- rbind(
@@
   filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
+# 
+# dtMainWideC <- PCAnorm(data = dtPcaS123,
+#           pid = "ID",
+#           tid = "TIDnum",
+#           selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>%
+#   select(
+#     ID,
+#     PID,
+#     TIDnum,
+#     study,
+#     ends_with("_gmc")
+#   ) %>%
+#   data.frame %>% 
+#   melt(
+#     .,
+#     id=c("ID", "PID", "TIDnum", "study")
+#   ) %>%
+#   mutate(
+#     variable = paste(gsub("_gmc", "", variable), stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+#   ) %>%
+#   arrange(variable, ID) %>%
+#   #arrange(ID, variable, TIDnum) %>%
+#   select(-TIDnum) %>%
+#   pivot_wider(names_from = variable, values_from = value) %>%
+#   replace(is.na(.), 0) %>%
+#   select(-ID, -PID, -study)
+#   #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
+#   #pivot_wider(names_from = variable, values_from = value)
+# 
+# # dtMainWideC %>%
+# #   filter(
+# #     ID == 54
+# #   ) %>% 
+# #   t
+# # 
+# # dtMainWideC$AttitudesDutch_59[dtMainWideC$ID == 54]
+# 
+# dtMainWide <- dtPcaS123 %>%
+#   select(
+#     -TID,
+#     -date,
+#     -week
+#   ) %>%
+#   data.frame %>% 
+#   melt(
+#     .,
+#     id=c("ID", "PID", "TIDnum", "study")
+#   ) %>%
+#   mutate(
+#     variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
+#   ) %>%
+#   select(-TIDnum) %>%
+#   arrange(variable, ID) %>%
+#   pivot_wider(names_from = variable, values_from = value) %>%
+#   replace(is.na(.), 0) %>%
+#   select(-ID, -PID, -study)
 
@@
   filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
-dtPcaS123DailyImp <- MiMainDaily$imputations[[1]] %>%
-  select(ID, date, study, any_of(varNamS123PCA)) %>%
+
+dtMainImpWide <- MiMain$imputations[[1]] %>% 
+  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
   arrange(ID, TIDnum) %>%
-  filter(
-    TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
-  )
-dtPcaS123WeeklyImp <- MiMainWeekly$imputations[[1]] %>%
-  select(ID, week, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  filter(
-    TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
-  )
-
-dtMainWideC <- PCAnorm(data = dtPcaS123,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>%
-  select(
-    ID,
-    PID,
-    TIDnum,
-    study,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>% 
-  melt(
-    .,
-    id=c("ID", "PID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste(gsub("_gmc", "", variable), stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
-  ) %>%
-  arrange(variable, ID) %>%
-  #arrange(ID, variable, TIDnum) %>%
-  select(-TIDnum) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  replace(is.na(.), 0) %>%
-  select(-ID, -PID, -study)
-  #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
-  #pivot_wider(names_from = variable, values_from = value)
-
-# dtMainWideC %>%
-#   filter(
-#     ID == 54
-#   ) %>% 
-#   t
-# 
-# dtMainWideC$AttitudesDutch_59[dtMainWideC$ID == 54]
-
-dtMainWide <- dtPcaS123 %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  data.frame %>% 
-  melt(
-    .,
-    id=c("ID", "PID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
-  ) %>%
-  select(-TIDnum) %>%
-  arrange(variable, ID) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  replace(is.na(.), 0) %>%
-  select(-ID, -PID, -study)
-
-dtMainImpWide <- dtPcaS123Imp %>%
+  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
   select(
@@
   select(-ID)
-
-dtMainDailyImpWide <- dtPcaS123DailyImp %>%
-  select(
-    -date
-  ) %>% 
-  melt(
-    .,
-    id=c("ID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
+```
+
+```{r longToWideMainStability}
+# Long to Wide Main Bi-daily
+MiMainStabilityWide <- list()
+for (i in 1:length(MiMainStability$imputations)) {
+  MiMainStabilityWide[[i]] <-
+    MiMainStability$imputations[[i]]  %>%
+    # Variable selection
+    select(ID, date, week, study, any_of(varNamS123PCA)) %>%
+    # Sorting
+    arrange(ID, TIDnum) %>%
+    # Filter minimum common num of measurements
+    filter(TIDnum <= min(
+      max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)
+    )) %>%
+    # Rm unnecessary variables for restructuring
+    select(-TID,
+           -date,
+           -week) %>%
+    # Long to wide
+    melt(.,
+         id = c("ID", "TIDnum", "study")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
     select(-TIDnum, -study) %>%
@@
     select(-ID)
-
-dtMainWeeklyImpWide <- dtPcaS123WeeklyImp %>%
+}
+```
+
+### Follow-up: Studies
+
+#### Study 1
+
+##### Common Variables
+
+```{r longToWideS1Common}
+dtS1ImpWide <- MiS1$imputations[[1]] %>% 
+  select(PID, date, week, any_of(varNamS123PCA)) %>%
+  arrange(PID, TIDnum) %>%
+  #filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
   select(
+    -TID,
+    -date,
     -week
+  ) %>%
+  melt(
+    .,
+    id=c("PID", "TIDnum")
+  ) %>%
+  mutate(
+    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
+  ) %>%
+  select(-TIDnum) %>%
+  arrange(PID, variable) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-PID)
+```
+
+```{r longToWideS1StabilityCommon}
+# Long to Wide S1 Bi-daily
+MiS1StabilityWide <- list()
+for (i in 1:length(MiS1Stability$imputations)) {
+  MiS1StabilityWide[[i]] <-
+    MiS1Stability$imputations[[i]]  %>%
+    # Variable selection
+    select(PID, date, week, any_of(varNamS123PCA)) %>%
+    # Sorting
+    arrange(PID, TIDnum) %>%
+    # Rm unnecessary variables for restructuring
+    select(-TID,
+           -date,
+           -week) %>%
+    # Long to wide
+    melt(.,
+         id = c("PID", "TIDnum")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
+    select(-TIDnum) %>%
+    arrange(PID, variable) %>%
+    pivot_wider(names_from = variable, values_from = value) %>%
+    #replace(is.na(.), 0) %>%
+    select(-PID)
+}
+```
+
+##### Idiosyncratic Variables
+
+```{r longToWideS1Idiosyncratic}
+dtS1ImpWideIdio <- MiS1$imputations[[1]] %>% 
+  select(PID, date, week, any_of(varNamS1PCA)) %>%
+  arrange(PID, TIDnum) %>%
+  #filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
+  select(
+    -TID,
+    -date,
+    -week
+  ) %>%
+  melt(
+    .,
+    id=c("PID", "TIDnum")
+  ) %>%
+  mutate(
+    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
+  ) %>%
+  select(-TIDnum) %>%
+  arrange(PID, variable) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-PID)
+```
+
+```{r longToWideS1StabilityIdiosyncratic}
+# Long to Wide S1 Bi-daily
+MiS1StabilityWideIdio <- list()
+for (i in 1:length(MiS1Stability$imputations)) {
+  MiS1StabilityWideIdio[[i]] <-
+    MiS1Stability$imputations[[i]]  %>%
+    # Variable selection
+    select(PID, date, week, any_of(varNamS1PCA)) %>%
+    # Sorting
+    arrange(PID, TIDnum) %>%
+    # Rm unnecessary variables for restructuring
+    select(-TID,
+           -date,
+           -week) %>%
+    # Long to wide
+    melt(.,
+         id = c("PID", "TIDnum")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
+    select(-TIDnum) %>%
+    arrange(PID, variable) %>%
+    pivot_wider(names_from = variable, values_from = value) %>%
+    #replace(is.na(.), 0) %>%
+    select(-PID)
+}
+```
+
+#### Study 2
+
+#### Study 3
+
+### Follow-up: Response Type
+
+#### Interaction Responses
+
+#### Non-Interaction Responses
+
+### Follow-up: Time Scales
+
+#### Daily
+
+```{r longToWideDaily}
+dtMainDailyImpWide <- MiMainDaily$imputations[[1]] %>%
+  select(ID, date, study, any_of(varNamS123PCA)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(
+    TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
+  ) %>%
+  select(
+    -date
   ) %>% 
@@
 
+```{r longToWideDailyStability}
+# Long to Wide Main Daily
+MiMainDailyStabilityWide <- list()
+for (i in 1:length(MiMainDailyStability$imputations)) {
+  MiMainDailyStabilityWide[[i]] <-
+    MiMainDailyStability$imputations[[i]] %>%
+    # Variable selection
+    select(ID, date, study, any_of(varNamS123PCA)) %>%
+    # Sorting
+    arrange(ID, TIDnum) %>%
+    # Filter minimum common num of measurements
+    filter(
+      TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
+    ) %>%
+    # Rm unnecessary variables for restructuring
+    select(-date) %>%
+    # Long to wide
+    melt(.,
+         id = c("ID", "TIDnum", "study")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
+    select(-TIDnum, -study) %>%
+    arrange(ID, variable) %>%
+    pivot_wider(names_from = variable, values_from = value) %>%
+    #replace(is.na(.), 0) %>%
+    select(-ID)
+}
+```
+
+#### Weekly
+
+```{r longToWideWeekly}
+dtMainWeeklyImpWide <- MiMainWeekly$imputations[[1]] %>%
+  select(ID, week, study, any_of(varNamS123PCA)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(
+    TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
+  ) %>%
+  select(
+    -week
+  ) %>% 
+  melt(
+    .,
+    id=c("ID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
+  ) %>%
+  select(-TIDnum, -study) %>%
+  arrange(ID, variable) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-ID)
+```
+
+```{r longToWideWeeklyStability}
+# Long to Wide Main Weekly
+MiMainWeeklyStabilityWide <- list()
+for (i in 1:length(MiMainWeeklyStability$imputations)) {
+  MiMainWeeklyStabilityWide[[i]] <-
+    MiMainWeeklyStability$imputations[[i]] %>%
+    # Variable selection
+    select(ID, week, study, any_of(varNamS123PCA)) %>%
+    # Sorting
+    arrange(ID, TIDnum) %>%
+    # Filter minimum common num of measurements
+    filter(
+      TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
+    ) %>%
+    # Rm unnecessary variables for restructuring
+    select(-week) %>%
+    # Long to wide
+    melt(.,
+         id = c("ID", "TIDnum", "study")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
+    select(-TIDnum, -study) %>%
+    arrange(ID, variable) %>%
+    pivot_wider(names_from = variable, values_from = value) %>%
+    #replace(is.na(.), 0) %>%
+    select(-ID)
+}
+```
+
 # **Three-Way ANOVA**
@@
     n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
-    p = length(unique(dtPcaS123Imp$TIDnum)) # Nr. time points (C)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainImpWide)))) # Nr. time points (C)
   ), 
@@
     n = nrow(dtMainDailyImpWide), # Nr. ppt.s (A)
-    m = length(names(dtPcaS123DailyImp)[!names(dtPcaS123DailyImp) %in% idVars]), # Nr. variables (B)
-    p = length(unique(dtPcaS123DailyImp$TIDnum)) # Nr. time points (C)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainDailyImpWide)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainDailyImpWide)))) # Nr. time points (C)
   ), 
@@
     n = nrow(dtMainWeeklyImpWide), # Nr. ppt.s (A)
-    m = length(names(dtPcaS123WeeklyImp)[!names(dtPcaS123WeeklyImp) %in% idVars]), # Nr. variables (B)
-    p = length(unique(dtPcaS123WeeklyImp$TIDnum)) # Nr. time points (C)
+    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainWeeklyImpWide)))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(dtMainWeeklyImpWide)))) # Nr. time points (C)
   ), 
@@
 
-However, with the aggregation to `r length(unique(dtPcaS123WeeklyImp$TIDnum))` weekly measurements, the $Person*Time*Variable+error$ sums of squares become less prominent and the variable differences explain substantially more variance.
+However, with the aggregation to `r length(unique(sub("_.*", "", names(dtMainWeeklyImpWide))))` weekly measurements, the $Person*Time*Variable+error$ sums of squares become less prominent and the variable differences explain substantially more variance.
 
@@
 
-We beginn by assessing the imputed dataset of the main analysis, where participants from all three studies are combined, using the relevant variables that are available for all three studies.
+We begin by assessing the imputed dataset of the main analysis, where participants from all three studies are combined, using the relevant variables that are available for all three studies.
 
@@
 
-## Center and Normalize
+## Stability Test
+
+### Main Analysis
+
+- all three studies combined
+- variables shared across studies
+- interaction and no interaction responses
+- bi-daily time scale
+
+#### Center and Normalize
+
+```{r MainStabCZ}
+##################################################################
+##                  Center across Participants                  ##
+##################################################################
+MiMainStabilityWideC <- list()
+for (i in 1:length(MiMainStabilityWide)) {
+  MiMainStabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiMainStabilityWide[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiMainStabilityWide[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWide[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiMainStabilityWide[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##################################################################
+##                 Normalize across Time Points                 ##
+##################################################################
+MiMainStabilityWideCZ <- list()
+for (i in 1:length(MiMainStabilityWideC)) {
+  MiMainStabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiMainStabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiMainStabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiMainStabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiMainStabilityWideCZ[[i]]) <- names(MiMainStabilityWideC[[i]])
+}
+```
+
+#### Component Complexity
+
+```{r MainStabCompNum}
+# Run PCA SUP to determine best combination
+MainStabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiMainStabilityWideCZ)) {
+    MainStabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiMainStabilityWideCZ[[i]], 
+    n = nrow(MiMainStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+MainStabDimSelector <- list()
+for (i in 1:length(MainStabPcaSup)) {
+  MainStabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = MainStabPcaSup[[i]],
+      n = nrow(MiMainStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+MainStabDim <- list()
+for (i in 1:length(MainStabDimSelector)) {
+  MainStabDim[[i]] <-
+    tail(MainStabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+MainStabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+### Follow-up: Studies (common variables)
+
+- **individual studies**
+- variables shared across studies
+- interaction and no interaction responses
+- bi-daily time scale
+
+#### Study 1
+
+##### Center and Normalize
+
+```{r S1CommonStabCZ}
+#################################################################
+##                           Study 1                           ##
+#################################################################
+
+##----------------------------------------------------------------
+##                  Center across Participants                  --
+##----------------------------------------------------------------
+MiS1StabilityWideC <- list()
+for (i in 1:length(MiS1StabilityWide)) {
+  MiS1StabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiS1StabilityWide[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiS1StabilityWide[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWide[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiS1StabilityWide[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##----------------------------------------------------------------
+##                 Normalize across Time Points                 --
+##----------------------------------------------------------------
+MiS1StabilityWideCZ <- list()
+for (i in 1:length(MiS1StabilityWideC)) {
+  MiS1StabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiS1StabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiS1StabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiS1StabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiS1StabilityWideCZ[[i]]) <- names(MiS1StabilityWideC[[i]])
+}
+```
+
+##### Component Complexity
+
+```{r S1CommonStabCompNum}
+# Run PCA SUP to determine best combination
+S1StabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiS1StabilityWideCZ)) {
+    S1StabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiS1StabilityWideCZ[[i]], 
+    n = nrow(MiS1StabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+S1StabDimSelector <- list()
+for (i in 1:length(S1StabPcaSup)) {
+  S1StabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = S1StabPcaSup[[i]],
+      n = nrow(MiS1StabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+S1StabDim <- list()
+for (i in 1:length(S1StabDimSelector)) {
+  S1StabDim[[i]] <-
+    tail(S1StabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+S1StabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Follow-up (S1 Common): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+#### Study 2
+
+##### Center and Normalize
+
+##### Component Complexity
+
+#### Study 3
+
+##### Center and Normalize
+
+##### Component Complexity
+
+### Follow-up: Studies (ideosyncratic variables)
+
+- **individual studies**
+- **variables idiosyncratic to studies**
+- interaction and no interaction responses
+- bi-daily time scale
+
+#### Study 1
+
+##### Center and Normalize
+
+```{r S1IdiosyncraticStabCZ}
+#################################################################
+##                           Study 1                           ##
+#################################################################
+
+##----------------------------------------------------------------
+##                  Center across Participants                  --
+##----------------------------------------------------------------
+MiS1IdioStabilityWideC <- list()
+for (i in 1:length(MiS1StabilityWideIdio)) {
+  MiS1IdioStabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiS1StabilityWideIdio[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiS1StabilityWideIdio[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideIdio[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiS1StabilityWideIdio[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##----------------------------------------------------------------
+##                 Normalize across Time Points                 --
+##----------------------------------------------------------------
+MiS1IdioStabilityWideCZ <- list()
+for (i in 1:length(MiS1IdioStabilityWideC)) {
+  MiS1IdioStabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiS1IdioStabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiS1IdioStabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1IdioStabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiS1IdioStabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiS1IdioStabilityWideCZ[[i]]) <- names(MiS1IdioStabilityWideC[[i]])
+}
+```
+
+##### Component Complexity
+
+```{r S1IdiosyncraticStabCompNum}
+# Run PCA SUP to determine best combination
+S1IdioStabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiS1IdioStabilityWideCZ)) {
+    S1IdioStabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiS1IdioStabilityWideCZ[[i]], 
+    n = nrow(MiS1IdioStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+S1IdioStabDimSelector <- list()
+for (i in 1:length(S1IdioStabPcaSup)) {
+  S1IdioStabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = S1IdioStabPcaSup[[i]],
+      n = nrow(MiS1IdioStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+S1IdioStabDim <- list()
+for (i in 1:length(S1IdioStabDimSelector)) {
+  S1IdioStabDim[[i]] <-
+    tail(S1IdioStabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+S1IdioStabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Follow-up (S1 Idiosyncratic): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+#### Study 2
+
+##### Center and Normalize
+
+##### Component Complexity
+
+#### Study 3
+
+##### Center and Normalize
+
+##### Component Complexity
+
+### Follow-up: Response Type (interaction vs. no interaction)
+
+- all three studies combined
+- variables shared across studies
+- **interaction vs. no interaction responses**
+- bi-daily time scale
+
+#### Center and Normalize
+
+#### Component Complexity
+
+### Follow-up: Time Scales (daily, weekly)
+
+- all three studies combined
+- variables shared across studies
+- interaction and no interaction responses
+- **daily and weekly time scales**
+
+#### Daily
+
+##### Center and Normalize
+
+```{r DailyStabCZ}
+##################################################################
+##                  Center across Participants                  ##
+##################################################################
+MiMainDailyStabilityWideC <- list()
+for (i in 1:length(MiMainDailyStabilityWide)) {
+  MiMainDailyStabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiMainDailyStabilityWide[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiMainDailyStabilityWide[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWide[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWide[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##################################################################
+##                 Normalize across Time Points                 ##
+##################################################################
+MiMainDailyStabilityWideCZ <- list()
+for (i in 1:length(MiMainDailyStabilityWideC)) {
+  MiMainDailyStabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiMainDailyStabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiMainDailyStabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiMainDailyStabilityWideCZ[[i]]) <- names(MiMainDailyStabilityWideC[[i]])
+}
+```
+
+##### Component Complexity
+
+```{r MainDailyStabCompNum}
+# Run PCA SUP to determine best combination
+MainDailyStabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiMainDailyStabilityWideCZ)) {
+    MainDailyStabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiMainDailyStabilityWideCZ[[i]], 
+    n = nrow(MiMainDailyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+MainDailyStabDimSelector <- list()
+for (i in 1:length(MainDailyStabPcaSup)) {
+  MainDailyStabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = MainDailyStabPcaSup[[i]],
+      n = nrow(MiMainDailyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+MainDailyStabDim <- list()
+for (i in 1:length(MainDailyStabDimSelector)) {
+  MainDailyStabDim[[i]] <-
+    tail(MainDailyStabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+MainDailyStabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Follow-up (Daily): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+#### Weekly
+
+##### Center and Normalize
+
+```{r MainWeeklyStabCZ}
+##################################################################
+##                  Center across Participants                  ##
+##################################################################
+MiMainWeeklyStabilityWideC <- list()
+for (i in 1:length(MiMainWeeklyStabilityWide)) {
+  MiMainWeeklyStabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiMainWeeklyStabilityWide[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiMainWeeklyStabilityWide[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWide[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWide[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##################################################################
+##                 Normalize across Time Points                 ##
+##################################################################
+MiMainWeeklyStabilityWideCZ <- list()
+for (i in 1:length(MiMainWeeklyStabilityWideC)) {
+  MiMainWeeklyStabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiMainWeeklyStabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiMainWeeklyStabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiMainWeeklyStabilityWideCZ[[i]]) <- names(MiMainWeeklyStabilityWideC[[i]])
+}
+```
+
+##### Component Complexity
+
+```{r MainWeeklyStabCompNum}
+# Run PCA SUP to determine best combination
+MainWeeklyStabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiMainWeeklyStabilityWideCZ)) {
+    MainWeeklyStabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiMainWeeklyStabilityWideCZ[[i]], 
+    n = nrow(MiMainWeeklyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+MainWeeklyStabDimSelector <- list()
+for (i in 1:length(MainWeeklyStabPcaSup)) {
+  MainWeeklyStabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = MainWeeklyStabPcaSup[[i]],
+      n = nrow(MiMainWeeklyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+MainWeeklyStabDim <- list()
+for (i in 1:length(MainWeeklyStabDimSelector)) {
+  MainWeeklyStabDim[[i]] <-
+    tail(MainWeeklyStabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+MainWeeklyStabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Follow-up (Weekly): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+## Main Analysis
+
+### Center and Normalize
 
@@
 
-## Select Component Complexity
+### Select Component Complexity
 
@@
 
-## Model Fitting
+### Model Fitting
 
@@
 
-## Rotation
+### Rotation
 
@@
 
-## Visualization: Time Components
+### Visualization: Time Components
 
@@
 
-## Visualization: Person Components
+### Visualization: Person Components
 
@@
 
-## Visualization: Variable Components
+### Visualization: Variable Components
 

### 2022-08-25 Supplemental-Material-A-Annotated-Analysis.Rmd
interaction follow-up stability added

@@
 #################################################################
-dtMiS2 <- dtS2Red %>% 
+dtMiS2 <- 
+  dtS2Red %>% 
   select(TID, any_of(varNamS2MiPsbl)) %>% 
@@
   mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
+  as.data.frame %>%
+  mutate(date = as.character(date)) %>%
   select(PID,
@@
          week,
-         any_of(names(which(sapply(dtMiS2 %>% select(where(is.numeric)), sd, na.rm = TRUE) > 0)))) %>% # only vars that sd != 0
-         #which(sapply(. %>% select(where(is.numeric)), sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+         everything()) %>%
+  select_if(function(x) is.character(x) | (is.numeric(x) && sd(x) > 0)) %>%
+         #which(sapply(. %>% select(where(is.numeric)), sd, na.rm = TRUE) > 0)) %>% # only vars that sd != 0
+         #) %>% # only include variables that vary (i.e., sd != 0)
+         #which(sd(., na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
   as.data.frame
@@
 
-```{r MiS1}
+```{r MiS2}
 #################################################################
@@
 #################################################################
-set.seed(123)
-MiS2 <-
-  amelia(
-    dtMiS2,
-    m = 1, # 20
-    ts = "TIDnum",
-    cs = "PID",
-    idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-    # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-    # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-    # leads = ...,
-    # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-    # parallel = "multicore",
-    # ncpus = parallel::detectCores(),
-    p2s = 1
-  )
-save(MiS2, file = "data/imputed/S2Imputed.RData")
-load("data/imputed/S2Imputed.RData")
+# set.seed(123)
+# MiS2 <-
+#   amelia(
+#     dtMiS2,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     # parallel = "multicore",
+#     # ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiS2, file = "data/imputed/S2Imputed.RData")
+#load("data/imputed/S2Imputed.RData")
 
@@
 # set.seed(123)
-# MiS1Stability <-
+# MiS2Stability <-
 #   amelia(
-#     dtMiS1,
+#     dtMiS2,
 #     m = 20,
@@
 #     # ncpus = parallel::detectCores(),
-#     p2s = 2
+#     p2s = 1
 #   )
-# save(MiS1Stability, file = "data/imputed/S1ImputedStability.RData")
-load("data/imputed/S1ImputedStability.RData")
+# save(MiS2Stability, file = "data/imputed/S2ImputedStability.RData")
+#load("data/imputed/S2ImputedStability.RData")
 ```
@@
 
-# # any interaction
-# dtMiMain %>%
-#   filter(is.na(InteractionDum) | InteractionDum != 0) %>%
-#   #complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
-#   nrow
-# 
-# # Intergroup Contact
-# dtMiMain %>%
-#   filter(is.na(IntergroupContact) | IntergroupContact != 0) %>%
-#   #complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
-#   nrow
-# 
-# # Ingroup Contact
-# dtMiMain %>%
-#   filter(is.na(IngroupContact) | IngroupContact != 0) %>%
-#   #complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
-#   nrow
-
+# any interaction
+dtMiInteraction <- 
+  dtMiMain %>%
+    mutate_at(vars(any_of(varNamS123PCA), -any_of(idVars)), funs(ifelse(dtMiMain$InteractionDum != 1, NA, .))) %>%
+    #mutate_at(vars(-any_of(idVars), -ends_with(".pre"), -ends_with(".post"), -InteractionDum), funs(ifelse(dtMiMain$InteractionDum != 1, NA, .))) %>%
+    as.data.frame
+
+# Intergroup Contact
+dtMiIntergroup <- 
+  dtMiMain %>%
+  mutate_at(vars(any_of(varNamS123PCA), -any_of(idVars)), funs(ifelse(dtMiMain$IntergroupContact != 1, NA, .))) %>%
+  #mutate_at(vars(-any_of(idVars), -ends_with(".pre"), -ends_with(".post"), -IntergroupContact), funs(ifelse(dtMiMain$IntergroupContact != 1, NA, .))) %>%
+  #filter(is.na(relatednessNoInteraction)) %>%
+  select(-relatednessNoInteraction) %>% # remove completely missing col
+  as.data.frame
+
+# dtMiIntergroup %>% 
+#   select(ID, TIDnum, IntergroupContact, relatednessNoInteraction) %>%
+#   arrange(ID, TIDnum, IntergroupContact)
+
+# Ingroup Contact
+dtMiIngroup <- 
+  dtMiMain %>%
+    #mutate_at(vars(any_of(varNamS123PCA), -any_of(idVars)), funs(ifelse(dtMiMain$IngroupContact != 1, NA, .))) %>%
+    mutate_at(vars(-any_of(idVars), -ends_with(".pre"), -ends_with(".post"), -IngroupContact), funs(ifelse(dtMiMain$IngroupContact != 1, NA, .))) %>%
+    as.data.frame
 ```
@@
 ```{r InteractionImp}
-# TBD
+#################################################################
+##                       Any Interaction                       ##
+#################################################################
+
+##---------------------------------------------------------------
+##                      Single Imputation                      --
+##---------------------------------------------------------------
+# set.seed(123)
+# MiInteraction <-
+#   amelia(
+#     dtMiInteraction,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     #noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiInteraction, file = "data/imputed/InteractionImputed.RData")
+load("data/imputed/InteractionImputed.RData")
+
+##---------------------------------------------------------------
+##                     Multiple Imputation                     --
+##---------------------------------------------------------------
+
+# # Stability to imputations:
+# set.seed(123)
+# MiInteractionStab <-
+#   amelia(
+#     dtMiInteraction,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     #noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiInteractionStab, file = "data/imputed/InteractionImputedStability.RData")
+load("data/imputed/InteractionImputedStability.RData")
+```
+
+```{r intergroupImp}
+##################################################################
+##                    Intergroup Interaction                    ##
+##################################################################
+
+##---------------------------------------------------------------
+##                      Single Imputation                      --
+##---------------------------------------------------------------
+# set.seed(123)
+# MiIntergroup <-
+#   amelia(
+#     dtMiIntergroup,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     #noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiIntergroup, file = "data/imputed/IntergroupImputed.RData")
+load("data/imputed/IntergroupImputed.RData")
+
+##---------------------------------------------------------------
+##                     Multiple Imputation                     --
+##---------------------------------------------------------------
+
+# # Stability to imputations:
+# set.seed(123)
+# MiIntergroupStab <-
+#   amelia(
+#     dtMiIntergroup,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     #noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiIntergroupStab, file = "data/imputed/IntergroupImputedStability.RData")
+load("data/imputed/IntergroupImputedStability.RData")
+```
+
+```{r ingroupImp}
+#################################################################
+##                     Ingroup Interaction                     ##
+##                        NOT CONVERGING                       ##
+#################################################################
+
+##---------------------------------------------------------------
+##                      Single Imputation                      --
+##---------------------------------------------------------------
+# set.seed(123)
+# MiIngroup <-
+#   amelia(
+#     dtMiIngroup,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     #noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiIngroup, file = "data/imputed/IngroupImputed.RData")
+# load("data/imputed/IngroupImputed.RData")
+
+##---------------------------------------------------------------
+##                     Multiple Imputation                     --
+##---------------------------------------------------------------
+
+# # Stability to imputations:
+# set.seed(123)
+# MiIngroupStab <-
+#   amelia(
+#     dtMiIngroup,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "ID",
+#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
+#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
+#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
+#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
+#     #noms = c("InteractionDum"),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiIngroupStab, file = "data/imputed/IngroupImputedStability.RData")
+# load("data/imputed/IngroupImputedStability.RData")
 ```
@@
 
-#### Interaction Responses
+#### All Interaction Responses
+
+```{r longToWideInteraction}
+
+dtInteractionImpWide <- MiInteraction$imputations[[1]] %>% 
+  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
+  select(
+    -TID,
+    -date,
+    -week
+  ) %>%
+  melt(
+    .,
+    id=c("ID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
+  ) %>%
+  select(-TIDnum, -study) %>%
+  arrange(ID, variable) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-ID)
+```
+
+```{r longToWidInteractionStability}
+# Long to Wide Main Bi-daily
+MiInteractionStabilityWide <- list()
+for (i in 1:length(MiInteractionStab$imputations)) {
+  MiInteractionStabilityWide[[i]] <-
+    MiInteractionStab$imputations[[i]]  %>%
+    # Variable selection
+    select(ID, date, week, study, any_of(varNamS123PCA)) %>%
+    # Sorting
+    arrange(ID, TIDnum) %>%
+    # Filter minimum common num of measurements
+    filter(TIDnum <= min(
+      max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)
+    )) %>%
+    # Rm unnecessary variables for restructuring
+    select(-TID,
+           -date,
+           -week) %>%
+    # Long to wide
+    melt(.,
+         id = c("ID", "TIDnum", "study")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
+    select(-TIDnum, -study) %>%
+    arrange(ID, variable) %>%
+    pivot_wider(names_from = variable, values_from = value) %>%
+    #replace(is.na(.), 0) %>%
+    select(-ID)
+}
+```
+
+#### Intergroup Interaction Responses
+
+```{r longToWideIntergroup}
+dtIntergroupImpWide <- MiIntergroup$imputations[[1]] %>% 
+  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
+  arrange(ID, TIDnum) %>%
+  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
+  select(
+    -TID,
+    -date,
+    -week
+  ) %>%
+  melt(
+    .,
+    id=c("ID", "TIDnum", "study")
+  ) %>%
+  mutate(
+    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
+  ) %>%
+  select(-TIDnum, -study) %>%
+  arrange(ID, variable) %>%
+  pivot_wider(names_from = variable, values_from = value) %>%
+  #replace(is.na(.), 0) %>%
+  select(-ID)
+```
+
+```{r longToWidIntergroupStability}
+# Long to Wide Main Bi-daily
+MiIntergroupStabilityWide <- list()
+for (i in 1:length(MiIntergroupStab$imputations)) {
+  MiIntergroupStabilityWide[[i]] <-
+    MiIntergroupStab$imputations[[i]]  %>%
+    # Variable selection
+    select(ID, date, week, study, any_of(varNamS123PCA)) %>%
+    # Sorting
+    arrange(ID, TIDnum) %>%
+    # Filter minimum common num of measurements
+    filter(TIDnum <= min(
+      max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)
+    )) %>%
+    # Rm unnecessary variables for restructuring
+    select(-TID,
+           -date,
+           -week) %>%
+    # Long to wide
+    melt(.,
+         id = c("ID", "TIDnum", "study")) %>%
+    mutate(variable = paste(
+      "T",
+      stringr::str_pad(TIDnum, width = 2, pad = 0),
+      "_",
+      variable,
+      sep = ""
+    )) %>%
+    select(-TIDnum, -study) %>%
+    arrange(ID, variable) %>%
+    pivot_wider(names_from = variable, values_from = value) %>%
+    #replace(is.na(.), 0) %>%
+    select(-ID)
+}
+```
 
@@
 
-#### Center and Normalize
-
-#### Component Complexity
+#### All Interactions
+
+##### Center and Normalize
+
+```{r InteractionStabCZ}
+##################################################################
+##                  Center across Participants                  ##
+##################################################################
+MiInteractionStabilityWideC <- list()
+for (i in 1:length(MiInteractionStabilityWide)) {
+  MiInteractionStabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiInteractionStabilityWide[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiInteractionStabilityWide[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWide[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiInteractionStabilityWide[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##################################################################
+##                 Normalize across Time Points                 ##
+##################################################################
+MiInteractionStabilityWideCZ <- list()
+for (i in 1:length(MiInteractionStabilityWideC)) {
+  MiInteractionStabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiInteractionStabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiInteractionStabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiInteractionStabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiInteractionStabilityWideCZ[[i]]) <- names(MiInteractionStabilityWideC[[i]])
+}
+```
+
+##### Component Complexity
+
+```{r InteractionStabCompNum}
+
+# Run PCA SUP to determine best combination
+InteractionStabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiInteractionStabilityWideCZ)) {
+    InteractionStabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiInteractionStabilityWideCZ[[i]], 
+    n = nrow(MiInteractionStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+InteractionStabDimSelector <- list()
+for (i in 1:length(InteractionStabPcaSup)) {
+  InteractionStabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = InteractionStabPcaSup[[i]],
+      n = nrow(MiInteractionStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+InteractionStabDim <- list()
+for (i in 1:length(InteractionStabDimSelector)) {
+  InteractionStabDim[[i]] <-
+    tail(InteractionStabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+InteractionStabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Follow-up (Interactions): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
+#### Intergroup Interactions
+
+##### Center and Normalize
+
+```{r IntergroupStabCZ}
+##################################################################
+##                  Center across Participants                  ##
+##################################################################
+MiIntergroupStabilityWideC <- list()
+for (i in 1:length(MiIntergroupStabilityWide)) {
+  MiIntergroupStabilityWideC[[i]] <-
+    ThreeWay::cent3(
+      X = MiIntergroupStabilityWide[[i]],
+      # Nr. ppt.s (A)
+      n = nrow(MiIntergroupStabilityWide[[i]]),
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWide[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWide[[i]])))),
+      # Center across Participants
+      mode = 1
+    ) %>%
+    as.data.frame
+}
+
+##################################################################
+##                 Normalize across Time Points                 ##
+##################################################################
+MiIntergroupStabilityWideCZ <- list()
+for (i in 1:length(MiIntergroupStabilityWideC)) {
+  MiIntergroupStabilityWideCZ[[i]] <-
+    ThreeWay::norm3(
+      X = MiIntergroupStabilityWideC[[i]], 
+      # Nr. ppt.s (A)
+      n = nrow(MiIntergroupStabilityWideC[[i]]), 
+      # Nr. variables (B)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWideC[[i]])))),
+      # Nr. time points (C)
+      p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWideC[[i]])))),
+      # Normalize across Time Points
+      mode = 2
+    ) %>%
+    as.data.frame
+  names(MiIntergroupStabilityWideCZ[[i]]) <- names(MiIntergroupStabilityWideC[[i]])
+}
+```
+
+##### Component Complexity
+
+```{r IntergroupStabCompNum}
+
+# Run PCA SUP to determine best combination
+IntergroupStabPcaSup <- list()
+capture.output(
+  for(i in 1:length(MiIntergroupStabilityWideCZ)) {
+    IntergroupStabPcaSup[[i]] <- 
+    ThreeWay::T3runsApproxFit(
+    X = MiIntergroupStabilityWideCZ[[i]], 
+    n = nrow(MiIntergroupStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+    m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. variables (B)
+    p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. time points (C)
+    maxa = 20,
+    maxb = 20,
+    maxc = 20
+  )
+  }, 
+file='NUL')
+
+# Select Dimension Complexity
+IntergroupStabDimSelector <- list()
+for (i in 1:length(IntergroupStabPcaSup)) {
+  IntergroupStabDimSelector[[i]] <- capture.output(
+    ThreeWay::DimSelector(
+      out = IntergroupStabPcaSup[[i]],
+      n = nrow(MiIntergroupStabilityWideCZ[[i]]), # Nr. ppt.s (A)
+      m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. variables (B)
+      p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. time points (C)
+      model = 2
+    )
+  )
+}
+
+IntergroupStabDim <- list()
+for (i in 1:length(IntergroupStabDimSelector)) {
+  IntergroupStabDim[[i]] <-
+    tail(IntergroupStabDimSelector[[i]], n = 1) %>%
+    str_extract("\\(.*\\)") %>%
+    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
+    as.data.frame %>%
+    mutate_all(as.numeric) %>%
+    dplyr::rename(c(A = V1,
+                    B = V2,
+                    C = V3))
+}
+
+IntergroupStabDim %>%
+  bind_rows %>%
+  as.data.frame %>%
+  rownames_to_column(., var = "imputation") %>%
+  group_by(A, B, C) %>%
+  summarise(Frequency = n()) %>%
+  arrange(-Frequency) %>%
+  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
+  kbl(
+    .,
+    escape = FALSE,
+    booktabs = TRUE,
+    align = "c",
+    digits = 2,
+    col.names = c(
+      "Person",
+      "Variable",
+      "Time",
+      "Frequency",
+      "Percentage"
+    ),
+    caption = "Follow-up (Intergroup Interactions): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
+  kable_classic(
+    full_width = FALSE,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
 

### 2022-08-26 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 
-To be done once initial tests approved.
-
 ```{r MiS2Prep}
@@
          everything()) %>%
-  select_if(function(x) is.character(x) | (is.numeric(x) && sd(x) > 0)) %>%
-         #which(sapply(. %>% select(where(is.numeric)), sd, na.rm = TRUE) > 0)) %>% # only vars that sd != 0
-         #) %>% # only include variables that vary (i.e., sd != 0)
-         #which(sd(., na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
+  select_if(function(x) is.character(x) | is.factor(x) | (is.numeric(x) && sd(x, na.rm = TRUE) > 0)) %>%
   as.data.frame
@@
 # # Variables removed for study 2 due to high multicollinearity:
-# varNamS1Out <- dtMiS1 %>%
-#   select(
-#     any_of(varNamDifS1),
+varNamS2Out <- dtMiS2 %>%
+  select(
+    any_of(varNamDifS2),
+    ContactDesire,
+    DesireIntergroup,
+    DesireIngroup,
+    Reason.pre,
     # mdmq.calmness.pre_calc,
@@
     # secNatDum.pre,
-#     timeNL.pre,
+    timeNL.pre,
+    starts_with("HS"),
+    starts_with("Home")
     # stress.post,
@@
     # starts_with("SSAS")
+  ) %>%
+  names
+
+dtMiS2Fin <- dtMiS2 %>%
+  # mutate(
+  #   assimilation.dif = assimilation.post - assimilation.pre,
+  #   integration.dif = integration.post - integration.pre,
+  #   marginalization.dif = marginalization.post - marginalization.pre,
+  #   separation.dif = separation.post - separation.pre,
+  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+  #   swl.dif = swl.post_calc - swl.pre_calc
   # ) %>%
-#   names
-
-# dtMiS2Fin <- dtMiS2 %>%
-#   # mutate(
-#   #   assimilation.dif = assimilation.post - assimilation.pre,
-#   #   integration.dif = integration.post - integration.pre,
-#   #   marginalization.dif = marginalization.post - marginalization.pre,
-#   #   separation.dif = separation.post - separation.pre,
-#   #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-#   #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-#   #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-#   #   swl.dif = swl.post_calc - swl.pre_calc
-#   # ) %>% 
-#   # select(
-#   #   -any_of(varNamS2Out)
-#   # ) %>%
-#   select(PID,
-#          TID,
-#          TIDnum,
-#          date,
-#          week,
-#          which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
-#   mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-#   as.data.frame
+  select(
+    -any_of(varNamS2Out)
+  ) %>%
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         everything()) %>%
+         #which(sapply(., sd, na.rm = TRUE) > 0)) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  as.data.frame
 
 # lares::corr_cross(
-#   dtMiS2 %>% select_if(is.numeric), # name of dataset
+#   dtMiS2Fin %>% select_if(is.numeric), # name of dataset
 #   max_pvalue = 0.05, # display only significant correlations (at 5% level)
@@
 #   amelia(
-#     dtMiS2,
+#     dtMiS2Fin,
 #     m = 1, # 20
@@
 #     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 1
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
 #   )
@@
 #     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
 #     p2s = 1
@@
 
-To be done once initial tests approved.
+```{r MiS3Prep}
+#################################################################
+##                    Variable Name Vectors                    ##
+#################################################################
+varNamS3MiPsbl <- c(
+  paste(varNamS3MI, rep(".pre", length(varNamS3MI)), sep = ""),
+  varNamS3MI,
+  paste(varNamS3MI, rep(".post", length(varNamS3MI)), sep = "")
+)
+
+varNamdtS3Noms <- varNames %>%
+  filter(
+    aux == 1,
+    format == "nominal",
+    studyS3 == "S2"
+  ) %>%
+  select(varNam) %>%
+  pull
+
+#################################################################
+##                    Dataframe Preparation                    ##
+#################################################################
+dtMiS3 <- 
+  dtS3Red %>% 
+  select(TID, any_of(varNamS3MiPsbl)) %>% 
+  mutate(across(!TID, as.numeric)) %>%
+  mutate(
+    date = as.Date(gsub(" .*", "", TID)),
+    week = strftime(date, format = "%Y-W%V")
+  ) %>%
+  select(PID,
+         date,
+         week,
+         everything(),
+         -starts_with("ResponseId")) %>%
+  as.data.frame %>% 
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
+  group_by(PID) %>%
+  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
+  ungroup %>% 
+  group_by(TIDnum) %>%
+  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
+  ungroup %>% 
+  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
+  as.data.frame %>%
+  mutate(date = as.character(date)) %>%
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         everything()) %>%
+  select_if(function(x) is.character(x) | is.factor(x) | (is.numeric(x) && sd(x, na.rm = TRUE) > 0)) %>%
+  as.data.frame
+
+varNamDifS3 <- names(dtMiS3)[gsub(".pre|.post|_calc", "", names(dtMiS3)) %>% allDuplicated]
+
+# # Variables removed for study 2 due to high multicollinearity:
+varNamS3Out <- dtMiS3 %>%
+  select(
+    any_of(varNamDifS3),
+    # ContactDesire,
+    # DesireIntergroup,
+    # DesireIngroup,
+    # Reason.pre,
+    # mdmq.calmness.pre_calc,
+    # mdmq.alertness.pre_calc,
+    # mdmq.valence.pre_calc,
+    # relatednessFrust.pre_calc,
+    # competenceFrust.pre_calc,
+    # competenceSat.pre_calc,
+    # relatednessSat.pre_calc,
+    # DutchFriends.pre,
+    # secNatDum.pre,
+    # timeNL.pre,
+    # starts_with("HS"),
+    # starts_with("Home")
+    # stress.post,
+    # discrimination_month.post,
+    # rosenberg.post,
+    # qualityStar,
+    # starts_with("SSAS")
+  ) %>%
+  names
+
+dtMiS3Fin <- dtMiS3 %>%
+  # mutate(
+  #   assimilation.dif = assimilation.post - assimilation.pre,
+  #   integration.dif = integration.post - integration.pre,
+  #   marginalization.dif = marginalization.post - marginalization.pre,
+  #   separation.dif = separation.post - separation.pre,
+  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
+  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
+  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
+  #   swl.dif = swl.post_calc - swl.pre_calc
+  # ) %>%
+  select(
+    -any_of(varNamS3Out)
+  ) %>%
+  select(PID,
+         TID,
+         TIDnum,
+         date,
+         week,
+         everything()) %>%
+  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
+  as.data.frame
+
+lares::corr_cross(
+  dtMiS3Fin %>% select_if(is.numeric), # name of dataset
+  max_pvalue = 0.05, # display only significant correlations (at 5% level)
+  top = 20 # display top 20 couples of variables (by correlation coefficient)
+)
+```
 
 ```{r MiS3}
-# BASED ON MiS1
+#################################################################
+##                     Imputation Bi-daily                     ##
+#################################################################
+# set.seed(123)
+# MiS3 <-
+#   amelia(
+#     dtMiS3Fin,
+#     m = 1, # 20
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 2
+#   )
+# save(MiS3, file = "data/imputed/S3Imputed.RData")
+#load("data/imputed/S2Imputed.RData")
+
+# Imputation Stability:
+# set.seed(123)
+# MiS3Stability <-
+#   amelia(
+#     dtMiS3Fin,
+#     m = 20,
+#     ts = "TIDnum",
+#     cs = "PID",
+#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
+#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
+#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
+#     # leads = ...,
+#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
+#     parallel = "multicore",
+#     ncpus = parallel::detectCores(),
+#     p2s = 1
+#   )
+# save(MiS3Stability, file = "data/imputed/S3ImputedStability.RData")
+#load("data/imputed/S2ImputedStability.RData")
 ```

### 2022-09-18 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 ```
+
 ```{r setup, include=FALSE}
@@
 #################################################################
-##               EXPORT FOR STM MACHINE LEARNING               ##
-#################################################################
-# bannerCommenter::banner("EXPORT FOR STM MACHINE LEARNING", centre = TRUE, bandChar = "#")
-# varNamPosAdd <- c("relatedness", "autonomy", "competence")
-# 
-# dtAllStm <- rbind(
-#   dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>%
-#     mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
-#   dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
-#   dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
-# ) %>%
-#   group_by(study, PID) %>%
-#   mutate(ID = cur_group_id()) %>%
-#   ungroup %>%
-#   mutate(
-#     date = as.Date(gsub(" .*", "", TID)),
-#     week = strftime(date, format = "%Y-W%V")
-#   ) %>%
-#   select(ID,
-#          PID,
-#          study,
-#          TID,
-#          TIDnum,
-#          date,
-#          week,
-#          IntergroupContact,
-#          IngroupContact,
-#          everything(),
-#          -starts_with("ResponseId")) %>%
-#   as.data.frame
-# 
-# dtAllStm2 <- dtAllStm
-# 
-# dtAllStm2$IntGrAnx.pre_calc <-
-#   psych::scoreItems(
-#     keys = c(rep(1,6),rep(-1,6)),
-#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
-#     min = 1,
-#     max = 10
-#   )$scores %>%
-#   as.numeric
-# 
-# dtAllStm2$IntGrAnx.post_calc <-
-#   psych::scoreItems(
-#     keys = c(rep(1,6),rep(-1,6)),
-#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
-#     min = 1,
-#     max = 10
-#   )$scores %>%
-#   as.numeric
-# 
-# dtAllStm2$swl.pre_calc <-
-#   psych::scoreItems(
-#     keys = rep(1,5),
-#     items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
-#     min = 1,
-#     max = 7,
-#     missing = TRUE,
-#     impute = "none",
-#   )$scores %>%
-#   as.numeric
-# 
-# dtAllStm2$discrimination.post_calc <-
-#   psych::scoreItems(
-#     keys = rep(1,9),
-#     items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
-#     min = 1,
-#     max = 6,
-#     missing = TRUE,
-#     impute = "none",
-#   )$scores %>%
-#   as.numeric
-# 
-# dtAllStm3 <- dtAllStm2 %>%
-#   select(
-#     -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
-#     -matches("^IntGrAnx[0-9]{2}.?\\.post"),
-#     -matches("^SWL[0-9]{2}\\.pre"),
-#     -matches("^EvDayDiscr[0-9]{2}\\.post"),
-#     -relatednessSelf,
-#     -relatednessOther,
-#     -relatednessNoInteraction
-#   )
-# 
-# write.csv2(dtAllStm3,
-#            file = "data/keyMotivesSTM.csv")
-#################################################################
-
-#################################################################
 ##                    Dataframe Preparation                    ##
@@
 # )
+```
+
+```{r s2StlTest}
+# Seasonal Decomposition of Time Series by Loess
+
+dtMiS2 %>% 
+  select(
+    PID,
+    TIDnum,
+    AttitudesDutch
+  ) %>%
+  #filter(PID == 1) %>%
+  #na.exclude(as.ts(.)) %>%
+  as.ts %>%
+  stats::stl(., na.action=na.omit)
+
+testLoess <- loess(AttitudesDutch ~ TIDnum, data = dtMiS2 %>% filter(PID == 1))
+testSpline <- pspline::sm.spline(dtMiS2$AttitudesDutch, dtMiS2$TIDnum)
+testMarkov <- semiMarkov()
+
+stats::stl(dtMiS2Fin)
+
 ```

### 2022-10-02 Supplemental-Material-A-Annotated-Analysis.Rmd
feature extraction added

@@
   "bookdown",
+  #"tidyverse",
+  "MASS",
   "brms",
@@
   "plotly",
+  "grid",
   "gridExtra",
@@
   "reshape2",
+  "anytime",
   "lubridate",
@@
   "Amelia", 
-  "ThreeWay"
+  "ThreeWay",
+  "DescTools",
+  # GAM
+  "gratia",
+  #"MASS",
+  "mgcv",
+  # For moving window models
+  "Kendall",
+  # For the Hamed & Rao correction of tau
+  "modifiedmk", 
+  "mgcViz",
+  "shiny"
 )
@@
 rm(lib)
+
+anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
+anytime::addFormats("%Y-%m-%d %H:%M:%S")
+anytime::addFormats("%d-%m-%Y")  
+#options(rgl.useNULL = TRUE)
 
@@
 source("./scripts/functions/meanViz.R")
+source("./scripts/functions/gam.R")
+source("./scripts/functions/tsFeatureExtractor.R")
 
@@
 
-dtMiS2 %>% 
-  select(
-    PID,
-    TIDnum,
-    AttitudesDutch
-  ) %>%
-  #filter(PID == 1) %>%
-  #na.exclude(as.ts(.)) %>%
-  as.ts %>%
-  stats::stl(., na.action=na.omit)
-
-testLoess <- loess(AttitudesDutch ~ TIDnum, data = dtMiS2 %>% filter(PID == 1))
-testSpline <- pspline::sm.spline(dtMiS2$AttitudesDutch, dtMiS2$TIDnum)
-testMarkov <- semiMarkov()
-
-stats::stl(dtMiS2Fin)
+# dtMiS2 %>% 
+#   select(
+#     PID,
+#     TIDnum,
+#     AttitudesDutch
+#   ) %>%
+#   #filter(PID == 1) %>%
+#   #na.exclude(as.ts(.)) %>%
+#   as.ts %>%
+#   stats::stl(., na.action=na.omit)
+# 
+# testLoess <- loess(AttitudesDutch ~ TIDnum, data = dtMiS2 %>% filter(PID == 1))
+# testSpline <- pspline::sm.spline(dtMiS2$AttitudesDutch, dtMiS2$TIDnum)
+# testMarkov <- semiMarkov()
+# 
+# stats::stl(dtMiS2Fin)
 
@@
 
+# **Feature-based Time-Series Clustering**
+
+1. central tendency
+    + mean ✓
+    + median ✓
+2. variation around the central tendency (i.e., density)
+    + SD ✓
+    + MAD ✓
+3. (in)stability
+    + root mean square of successive differences (rmssd) ✓
+    + mean absolute change ✓
+4. inertia
+    + lag-1 autocorrelation ✓
+    + periodicity (lag-2 & lag-14 autocorrelation) ✓
+5. linear trend
+    + OLS regression parameter ✓
+6. non-linear trend
+    + GAM edf ✓
+
+(!discuss NA procedures!)
+
+## Variable Selection
+
+```{r variable names for all analyses}
+
+# ID Variables (just to re-iterate)
+idVars <- idVars
+
+# Common variables across studies 
+varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
+
+# Variable Groups
+varNames %>%
+  filter(pca != 0) %>%
+  mutate(contactSpecific = recode(.$contactSpecific, 
+                         `-1`="No Interaction Only",
+                         `0`="unspecific",
+                         `1`="Interaction Only")) %>%
+  crosstable::crosstable(., varGroup, by=contactSpecific) %>%
+  select(-c(.id, label)) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "r", "c"),
+      digits=2,
+      caption = "Full Analysis: Variable Groups by Interaction Types") %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+
+varNames %>%
+  filter(
+    pca != 0
+  ) %>%
+  mutate(contactSpecific = recode(.$contactSpecific, 
+                         `-1`="No Interaction Only",
+                         `0`="unspecific",
+                         `1`="Interaction Only")) %>%
+  select(varNam, varGroup, contactSpecific) %>%
+  group_by(varGroup, varNam, contactSpecific) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
+  ungroup %>%
+  t %>%
+  as.data.frame %>%
+  janitor::row_to_names(1) %>%
+  replace(is.na(.), 0) %>%
+  tibble::rownames_to_column(., var = "variable") %>%
+  separate(variable, c("concept", "variable"), " - ") %>%
+  arrange(concept, unspecific, `Interaction Only`) %>%
+  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "l", rep("c", ncol(.)-2)),
+      digits=2,
+      caption = "All Possible Variables by Interaction Types") %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+
+varNames %>%
+  filter(
+    pca != 0,
+    varNam %in% varNamS123PCA
+  ) %>%
+  mutate(contactSpecific = recode(.$contactSpecific, 
+                         `-1`="No Interaction Only",
+                         `0`="unspecific",
+                         `1`="Interaction Only")) %>%
+  select(varNam, varGroup, contactSpecific) %>%
+  group_by(varGroup, varNam, contactSpecific) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
+  ungroup %>%
+  t %>%
+  as.data.frame %>%
+  janitor::row_to_names(1) %>%
+  replace(is.na(.), 0) %>%
+  tibble::rownames_to_column(., var = "variable") %>%
+  separate(variable, c("concept", "variable"), " - ") %>%
+  arrange(concept, unspecific, `Interaction Only`) %>%
+  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "l", rep("c", ncol(.)-2)),
+      digits=2,
+      caption = "Full Analysis: Variables by Interaction Types") %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+
+# Non-Interaction Specific Variables
+varNamS123NoInt <- varNames %>%
+  filter(
+    varNam %in% varNamS123PCA,
+    contactSpecific == 0
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# Interaction Specific Variables
+varNamS123Int <- varNames %>%
+  filter(
+    varNam %in% varNamS123PCA,
+    contactSpecific == 1
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S1 variables
+varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
+
+# S1 Non-Interaction Specific Variables
+varNamS1NoInt <- varNames %>%
+  filter(
+    varNam %in% varNamS1Clust,
+    contactSpecific == 0
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S1 Interaction Specific Variables
+varNamS1Int <- varNames %>%
+  filter(
+    varNam %in% varNamS1Clust,
+    contactSpecific == 1
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S2 variables
+varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
+
+# S2 Non-Interaction Specific Variables
+varNamS2NoInt <- varNames %>%
+  filter(
+    varNam %in% varNamS2Clust,
+    contactSpecific == 0
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S2 Interaction Specific Variables
+varNamS2Int <- varNames %>%
+  filter(
+    varNam %in% varNamS2Clust,
+    contactSpecific == 1
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S3 variables
+varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
+
+# S3 Non-Interaction Specific Variables
+varNamS3NoInt <- varNames %>%
+  filter(
+    varNam %in% varNamS3Clust,
+    contactSpecific == 0
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S3 Interaction Specific Variables
+varNamS3Int <- varNames %>%
+  filter(
+    varNam %in% varNamS3Clust,
+    contactSpecific == 1
+  ) %>%
+  select(varNam) %>%
+  pull
+```
+
+## Feature Extraction
+
+```{r load saved and prepared features}
+load("data/features/featOut.RData")
+
+```
+
+### Across Studies
+
+```{r Feature Extraction Main Analysis}
+# varNames
+#items <- varNamS123PCA[!varNamS123PCA %in% idVars]
+
+# Joined Dataset of all three studies
+featData <- rbind(
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  select(ID,
+         PID,
+         study,
+         date,
+         week,
+         TIDnum,
+         any_of(varNamS123PCA)) %>%
+   mutate(
+     across(all_of(varNamS123PCA), ~ lag(.x, n = 1), .names = "{col}_lag1"),
+     across(all_of(varNamS123PCA), ~ lag(.x, n = 2), .names = "{col}_lag2"),
+     across(all_of(varNamS123PCA), ~ lag(.x, n = 14), .names = "{col}_lag14")
+   ) %>%
+  arrange(ID, TIDnum) %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  as.data.frame
+
+# Don't re-run unless you have a few hours to spare!
+# featFull <-
+#   featureExtractor(
+#     data = featData,
+#     pid = "ID",
+#     tid = "TIDnum",
+#     items = varNamS123
+#   )
+# 
+# featFullContact <-
+#   featureExtractor(
+#     data = featData,
+#     pid = "ID",
+#     tid = "TIDnum",
+#     items = varNamS123Int
+#   )
+# 
+# featFullNoContact <-
+#   featureExtractor(
+#     data = featData,
+#     pid = "ID",
+#     tid = "TIDnum",
+#     items = varNamS123NoInt
+#   )
+# 
+# 
+# featFull$features <- featFull$features %>%
+#     filter(
+#       ID != 19
+#     )
+
+pMissFeat(
+  all = featFull$features,
+  contact = featFullContact$features,
+  nocontact = featFullNoContact$features,
+  title = "Feature-wise Missingess Across all Studies"
+)
+
+# featFullImp <- featureImputer(featFull)
+# featFullNoContactImp <- featureImputer(featFullNoContact)
+# featFullContactImp <- featureImputer(featFullContact)
+
+```
+
+### Study 1
+
+```{r Feature Extraction Study 1 specific variables}
+# varNames
+#items <- varNamS1Clust
+
+# Prepare Dataframe
+featDataS1 <- 
+  dtS1Red %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  select(PID,
+         date,
+         week,
+         TIDnum,
+         any_of(varNamS1Clust)) %>%
+   mutate(
+     across(all_of(varNamS1Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
+     across(all_of(varNamS1Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
+     across(all_of(varNamS1Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
+   ) %>%
+  arrange(PID, TIDnum) %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  as.data.frame
+
+# Don't re-run unless you have a few hours to spare!
+# featS1 <-
+#   featureExtractor(
+#     data = featDataS1,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS1Clust
+#   )
+# 
+# featS1Contact <-
+#   featureExtractor(
+#     data = featDataS1,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS1Int
+#   )
+# 
+# featS1NoContact <-
+#   featureExtractor(
+#     data = featDataS1,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS1NoInt
+#   )
+
+pMissFeat(
+  all = featS1$features,
+  contact = featS1Contact$features,
+  nocontact = featS1NoContact$features,
+  title = "Feature-wise Missingess Study 1"
+)
+
+# featS1Imp <- featureImputer(featS1)
+# featS1NoContactImp <- featureImputer(featS1NoContact)
+# featS1ContactImp <- featureImputer(featS1Contact)
+
+```
+
+### Study 2
+
+```{r Feature Extraction Study 2 specific variables}
+# varNames
+#items <- varNamS2Clust
+
+# Prepare Dataframe
+featDataS2 <- 
+  dtS2Red %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  select(PID,
+         date,
+         week,
+         TIDnum,
+         any_of(varNamS2Clust)) %>%
+   mutate(
+     across(all_of(varNamS2Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
+     across(all_of(varNamS2Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
+     across(all_of(varNamS2Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
+   ) %>%
+  arrange(PID, TIDnum) %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  as.data.frame
+
+# Don't re-run unless you have a few hours to spare!
+# featS2 <-
+#   featureExtractor(
+#     data = featDataS2,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS2Clust
+#   )
+# 
+# featS2Contact <-
+#   featureExtractor(
+#     data = featDataS2,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS2Int
+#   )
+# 
+# featS2NoContact <-
+#   featureExtractor(
+#     data = featDataS2,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS2NoInt
+#   )
+
+pMissFeat(
+  all = featS2$features,
+  contact = featS2Contact$features,
+  nocontact = featS2NoContact$features,
+  title = "Feature-wise Missingess Study 2"
+)
+
+# featS2Imp <- featureImputer(featS2)
+# featS2NoContactImp <- featureImputer(featS2NoContact)
+# featS2ContactImp <- featureImputer(featS2Contact)
+```
+
+### Study 3
+
+```{r Feature Extraction Study 3 specific variables}
+# varNames
+#items <- varNamS3Clust
+
+# Prepare Dataframe
+featDataS3 <- 
+  dtS3Red %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  select(PID,
+         date,
+         week,
+         TIDnum,
+         any_of(varNamS3Clust)) %>%
+   mutate(
+     across(all_of(varNamS3Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
+     across(all_of(varNamS3Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
+     across(all_of(varNamS3Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
+   ) %>%
+  arrange(PID, TIDnum) %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  as.data.frame
+
+# Don't re-run unless you have a few hours to spare!
+# featS3 <-
+#   featureExtractor(
+#     data = featDataS3,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS3Clust
+#   )
+# 
+# featS3Contact <-
+#   featureExtractor(
+#     data = featDataS3,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS3Int
+#   )
+# 
+# featS3NoContact <-
+#   featureExtractor(
+#     data = featDataS3,
+#     pid = "PID",
+#     tid = "TIDnum",
+#     items = varNamS3NoInt
+#   )
+
+pMissFeat(
+  all = featS3$features,
+  contact = featS3Contact$features,
+  nocontact = featS3NoContact$features,
+  title = "Feature-wise Missingess Study 3"
+)
+
+# featS3Imp <- featureImputer(featS3)
+# featS3NoContactImp <- featureImputer(featS3NoContact)
+# featS3ContactImp <- featureImputer(featS3Contact)
+# 
+# save(list = ls(pattern = "feat"), file = "data/features/featOut.RData")
+```
+
+## Clustering
+
+### Across Studies
+
+#### Centroid-based
+
+k-means
+
+```{r kmeans determin k}
+library(factoextra)
+library(NbClust)
+
+scaled_data <- featFull2$featuresImpZMat
+
+# Elbow method
+fviz_nbclust(scaled_data, kmeans, method = "wss") +
+  geom_vline(xintercept = 2, linetype = 2)+
+  labs(subtitle = "Elbow method")
+# Silhouette method
+fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
+  labs(subtitle = "Silhouette method")
+# Gap statistic
+# nboot = 50 to keep the function speedy. 
+# recommended value: nboot= 500 for your analysis.
+# Use verbose = FALSE to hide computing progression.
+set.seed(123)
+fviz_nbclust(scaled_data, kmeans, nstart = 50,  method = "gap_stat", nboot = 500)+
+  labs(subtitle = "Gap statistic method")
+
+kmeansOut <- kmeans(scaled_data, centers = 2, nstart = 100)
+
+fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())
+
+as.data.frame(kmeansOut$cluster)
+
+```
+
+#### Distribution-based
+
+Gaussian mixture model (GMM)
+
+```{r gaussian mixture model}
+library(mclust)
+
+BIC <- mclustBIC(scaled_data)
+plot(BIC)
+summary(BIC)
+
+mod1 <- Mclust(scaled_data, x = BIC)
+
+merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View
+
+summary(mod1, parameters = TRUE)
+table(class, mod1$classification)
+
+mod1dr <- MclustDR(mod1)
+summary(mod1dr)
+plot(mod1dr, what = "pairs")
+```
+
+#### Density-based
+
+dbscan
+
+```{r dbscan}
+library(dbscan)
+library(fpc)
+
+# to plot the eps values
+eps_plot = kNNdistplot(scaled_data, k=3)
+
+# to draw an optimum line
+eps_plot %>% abline(h = 13, lty = 2)
+
+eps <- 13
+minPts <- 3
+
+res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, method="raw")
+res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
+all(res.fpc$cluster == res.db$cluster)
+
+fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
+
+```
+
 # **Three-Way ANOVA**

### 2022-10-02 scripts/functions/gam.R

@@ -1,0 +1,408 @@
+# GAM (Helper) Functions
+
+
+##################################################################
+##                       EWS GAM Function                       ##
+##################################################################
+
+# This function calculates EWS using GAM. It returns a dataframe with:
+# autoregressive coefficient, as well as its upper and lower limit (ll,
+# ul), the intercept upper and lower limit, the mu upper and lower limit.
+# There are also inbuilt checks for the GAM: test of the BIC,
+# significance, edf, whether there is a rise in the model (by comparing
+# the 35th fitted value, to the 35th fitted value from the end).
+# 
+# Scaling denotes whether you want to standardize the variable or not. For
+# the circumplex variables (PA/NA hi/lo), scaling has been done prior to
+# combining (averaging).
+# 
+# Mu denotes the mu/attractor of the proces (= intercept/(1-AR)).
+# 
+# Mean denotes the fitted values using a GAM with time as a predictor.
+# This might represent an 'easier' version of mu (i.e. version that does
+# not depend on the AR).
+# 
+# The sections that are commented out were written by M. J. Schreuder for
+# a different paper, where the rise in AR was determined using the
+# derivative of the two weeks before a transition. We will not apply that
+# here.
+
+set.seed(123) # for reproducible results
+GAM.ews <- function(dw) {
+  set.seed(123) 
+  out <- data.frame("time" = dw$time, "check1.BIC.says.tv" = NA, "check2.sig.p.int" = NA, "check2.sig.p.AR" = NA, "check3.edf.tv" = NA, "check4a.rise" = NA,
+                    "BIC.ar" = NA, "BIC.tvar" = NA, "pval.int.tvar" = NA, "pval.AR.tvar" = NA, "edf" = NA,
+                    "intercept" = NA, 'll.int' = NA, 'ul.int' = NA, "AR" = NA, "ll"=NA, "ul"=NA,"mu" = NA, 'll.mu' = NA, 'ul.mu' = NA, 
[…]

### 2022-10-02 scripts/functions/tsFeatureExtractor.R

@@ -1,0 +1,254 @@
+
+
+featureExtractor <- function(data, items, pid, tid) {
+  # for testing:
+  # data <- featData
+  # pid <- "ID" #c("ID", "PID")
+  # tid <- "TIDnum"
+  # items <- varNamS123PCA[!varNamS123PCA %in% idVars]
+  
+  names(data)[names(data) == pid] <- "ID"
+  names(data)[names(data) == tid] <- "TIDnum"
+  
+  # univariate features
+  featUnivar <- data %>%
+    group_by(ID) %>%
+    summarise(across(
+      any_of(items),
+      list(
+        # central tendency
+        mean = ~ mean(.x, na.rm = TRUE),
+        median = ~ stats::median(.x, na.rm = TRUE),
+        # variance
+        sd = ~ sd(.x, na.rm = TRUE),
+        mad = ~ stats::mad(.x, na.rm = TRUE),
+        # stability
+        rmssd = ~ psych::rmssd(.x, group=ID, lag = 1, na.rm=TRUE) %>% as.numeric,
+        mac = ~ sum(abs(diff(.x, lag = 1)), na.rm = TRUE)/(sum(!is.na(.x)) - 1)
+      )
+    ))
+  
+  # Multivariate and time-based features
+  featMultivar <- featUnivar %>%
+    select(ID)
+  
+  # progress bar
+  Results_GAM <- sapply(as.character(unique(featData$ID)), function(x) NULL)
+  pb <- txtProgressBar(min = 0, max = length(unique(featMultivar$ID)), style = 3) #, width = length(unique(featMultivar$ID))
+  for (i in unique(featMultivar$ID)) {
+    for (j in items) {
+      if (sum(!is.na(featData[[j]][featData$ID == i])) == 0) next # skip if all IV vals NA
+      
+      # linear trend
+      featMultivar[[paste(j, "lin", sep = "_")]][featMultivar$ID == i] <-
+        lm(data = featData %>% filter(ID == i),
+           formula = get(j) ~ TIDnum)$coefficients[["TIDnum"]]
+      
+      # non-linear trend
+      dfGam <- data %>%
+        filter(ID == i) %>%
+        select(time = TIDnum,
+               var = all_of(j) )#,
+      #varLag1 = all_of(as.character(paste0(j, "_lag1"))),
+      #varLag2 = all_of(as.character(paste0(j, "_lag2"))),
[…]

### 2022-10-02 3mpca.Rmd
re-render with main comments added

--- 2022-09-18 Supplemental-Material-A-Annotated-Analysis.Rmd
@@
 ---
-title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
-subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
-author:
-- Jannis Kreienkamp^1,3^
-- Kai Epstude^1,3^
-- Laura F. Bringmann^1,3^
-- Maximilian Agostini^1,3^
-- Peter de Jonge^1,3^
-- Rei Monden^2,3^
-- ^1^University of Groningen, Department of Psychology
-- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
-- ^3^Author order TBD [currently in first name alphabetical order]
-- "Author Information:"
-- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
-E-mail: j.kreienkamp@rug.nl"
-- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
-- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
-- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
-date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
-output: 
-  bookdown::html_document2:
-    fig_caption: yes
-    md_extensions: +footnotes
-    code_folding: hide
-    mathjax: default
-    theme: yeti
-    toc: yes
-    toc_float: yes
-    number_sections: false
-    css: style.css
-    includes:
-      in_header: "_includes/head-custom-rmd.html" 
-editor_options:
-  chunk_output_type: console
-bibliography: references.bib
-csl: apa.csl
-header-includes:
-   - \usepackage{amsmath, nccmath}
+title: "3mpca"
+author: "Jannis Kreienkamp"
+date: "2022-10-02"
+output: html_document
 ---
-
-```{=html}
-<style type="text/css">
-.main-container {
-  max-width: 1300px;
-  margin-left: auto;
-  margin-right: auto;
-}
-.table {
-  margin-left:auto; 
-  margin-right:auto;
-}
-</style>
-```
-
-```{r setup, include=FALSE}
-# R Studio Clean-Up
-cat("\014") # clear console
-# rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
-gc() # garbage collector
-
-# Install and Load Packages
-# !IMPORTANT!
-# BEFORE FIRST RENDER:
-# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
-# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
-lib <- c(
-  "rmarkdown",
-  "knitr",
-  "remedy",
-  "bookdown",
-  "brms",
-  "psych",
-  "ggplot2",
-  "ggthemes",
-  "haven",
-  "RColorBrewer",
-  "plotly",
-  "gridExtra",
-  "ggpattern",
-  "lme4",
-  "nlme",
-  "jtools",
-  "gtsummary",
-  "sessioninfo",
-  "tibble",
-  "pander",
-  "devtools",
-  "mada",
-  "data.table",
-  "plyr",
-  "dplyr",
-  "tidyr",
-  "Hmisc",
-  "kableExtra",
-  "papaja",
-  "stringr",
-  "stringi",
-  "reshape2",
-  "lubridate",
-  "purrr",
-  "metafor",
-  "dygraphs",
-  "readxl", 
-  "reshape",
-  "factoextra",
-  "Amelia", 
-  "ThreeWay"
-)
-invisible(lapply(lib, library, character.only = TRUE))
-rm(lib)
-
-# Load Custom Packages
-source("./scripts/functions/fun.panel.R")
-source("./scripts/functions/themes.R")
-source("./scripts/functions/binaryCor.R")
-source("./scripts/functions/MlCorMat.R")
-source("./scripts/functions/MlTbl.R")
-source("./scripts/functions/metaLmer.R")
-source("./scripts/functions/meanViz.R")
-
-# Markdown Options
-knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
-knitr::opts_knit$get("root.dir") # check working directory
-options(
-  scipen = 999,
-  digits = 4,
-  width = 400
-) # removes scientific quotation
-# knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings
-knitr::knit_hooks$set(
-  error = function(x, options) {
-    paste('\n\n<div class="alert alert-danger">',
-      gsub("##", "\n", gsub("^##\ Error", "**Error**", x)),
-      "</div>",
-      sep = "\n"
-    )
-  },
-  warning = function(x, options) {
-    paste('\n\n<div class="alert alert-warning">',
-      gsub("##", "\n", gsub("^##\ Warning:", "**Warning**", x)),
-      "</div>",
-      sep = "\n"
-    )
-  },
-  message = function(x, options) {
-    paste('\n\n<div class="alert alert-info">',
-      gsub("##", "\n", x),
-      "</div>",
-      sep = "\n"
-    )
-  }
-)
-htmltools::tagList(rmarkdown::html_dependency_font_awesome())
-
-# Global Chunk Options
-knitr::opts_chunk$set(
-  fig.width = 12,
-  fig.height = 8,
-  fig.path = "Figures/",
-  echo = TRUE,
-  warning = FALSE,
-  message = FALSE
-)
-```
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-<i class="fas fa-exclamation-circle"></i> Note.
-Boxplots display the interquartile range (IQR, center box), and the whiskers extend 1.5\*IQR from the lower and upper hinge.
-The white point indicates the mean and the white center line indicates the median.
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-# **Data Preparation**
-
-In an initial preparatory step, we import the data into the R project environment and prepare the variables for further processing and later analyses.
-
-## Data Import
-
-The data were collected using two different survey tools.
-For the study with sojourners (Study 1: worker) we used the survey platform Qualtrics XM, whereas the studies with international students (Study 2: student), and the international medical professionals (Study 3: medical) were conducted using the survey framework FormR.
-This means that the datasets had inconsistent file formats and naming conventions.
-For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*').
-For the two other studies, we import the raw csv files from their respective folders.
-
-```{r formrImport}
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
-
-# rename function
-reNam <- function(data = NA, names = varNames, study = "S1", survey = "pre") {
-  # for testing
-  # data = dtS1$raw.pre
-  # names = varNames
-  # study = "S1"
-  # survey = "pre"
-
-  var <- paste0("survey", study)
-  varNamOld <- paste0("varNam", study)
-
-  nameTbl <- names %>%
-    filter((!!sym(var)) == survey) %>%
-    select(varNam, varNamOld = (!!sym(varNamOld)))
-
-  data.table::setnames(data, as.character(nameTbl$varNamOld), as.character(nameTbl$varNam), skip_absent=TRUE)
-
-  return(data)
-}
-
-# workers
-# initial data cleaning was done in SPSS (syntax files are available in "")
-dtS1 <- list(
-  raw.pre = read_spss("data/S1_Workers/processed/cleaned/MT - Pre-Measure - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "pre"),
-  raw.post = read_spss("data/S1_Workers/processed/cleaned/MT - Post-Measure - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "post"),
-  raw.morning = read_spss("data/S1_Workers/processed/cleaned/MT - Morning - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "daily"),
-  raw.afternoon = read_spss("data/S1_Workers/processed/cleaned/MT - Afternoon - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "daily")
-)
-
-# students
-dtS2 <- list(
-  raw.pre = read.csv(file = "data/S2_Students/raw/AOTS_Pre.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "pre"),
-  raw.post = read.csv(file = "data/S2_Students/raw/AOTS_Post.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "post"),
-  raw.daily = read.csv(file = "data/S2_Students/raw/AOTS_Daily.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "daily")
-)
-
-# young medical professionals
-dtS3 <- list(
-  raw.eligibility = read.csv("data/S3_Medical/raw/AOTM_Eligibility.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"), # works but should be 'survey = "pre_elig"'
-  raw.pre = read.csv("data/S3_Medical/raw/AOTM_Pre.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"),
-  raw.post = read.csv("data/S3_Medical/raw/AOTM_Post.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "post"),
-  raw.daily = read.csv("data/S3_Medical/raw/AOTM_Daily.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "daily")
-)
-
-# # FROM INITIAL VAR NAME EXPORT [LEGACY]
-# # export var names
-# S1Nam <- rbind(
-#   data.frame(
-#     study = "S1",
-#     survey = "pre",
-#     varNam = names(dtS1$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S1",
-#     survey = "daily",
-#     varNam = names(dtS1$raw.afternoon)
-#   ),
-#   data.frame(
-#     study = "S1",
-#     survey = "post",
-#     varNam = names(dtS1$raw.post)
-#   )
-# )
-# S2Nam <- rbind(
-#   data.frame(
-#     study = "S2",
-#     survey = "pre",
-#     varNam = names(dtS2$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S2",
-#     survey = "daily",
-#     varNam = names(dtS2$raw.daily)
-#   ),
-#   data.frame(
-#     study = "S2",
-#     survey = "post",
-#     varNam = names(dtS2$raw.post)
-#   )
-# )
-# S3Nam <- rbind(
-#   data.frame(
-#     study = "S3",
-#     survey = "pre_elig",
-#     varNam = names(dtS3$raw.eligibility)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "pre_entry",
-#     varNam = names(dtS3$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "daily",
-#     varNam = names(dtS3$raw.daily)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "post",
-#     varNam = names(dtS3$raw.post)
-#   )
-# )
-# write.csv(
-#   x = rbind(
-#     S1Nam,
-#     S2Nam,
-#     S3Nam
-#   ),
-#   file = "preregistration/varNames.csv"
-# )
-```
-
-## Data Cleaning & Data Exclusions
-
-### Study 1
-
-For the sojourner sample data was collected in four separate surveys: (1) the pre-measurement, (2) the daily morning survey, (3) the daily afternoon survey, as well as (4) a post-measurement.
-We combine the four individual surveys into one cohesive dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We then format the time and date variables and add person- and measurement indices (for easy and meaningful addressing of the data).
-We also exclude our own test data.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS1*' list.
-
-```{r cleanWorker}
-# Create reduced data sets for morning and afternoon
-dat.mo <- dtS1$raw.morning %>%
-  select(-starts_with("t_"))
-#setdiff(names(dtS1$raw.morning), names(dat.mo))
-dat.mo$daytime <- "morning"
-
-dat.af <- dtS1$raw.afternoon %>%
-  select(-starts_with("t_"))
-dat.af$daytime <- "afternoon"
-
-# merge morning and afternoon measurements with indicator [+ clean up]
-daily.dat <- plyr::rbind.fill(dat.mo, dat.af)
-daily.dat <- daily.dat[daily.dat$last_outside_referrer != 55951, ]
-dtS1$daily <- daily.dat
-rm(dat.mo, dat.af, daily.dat)
-
-# reduced data set for pre measurement
-dat.pre.red <- dtS1$raw.pre %>%
-  select(-starts_with("t_"))
-names(dat.pre.red) <- paste(names(dat.pre.red), "pre", sep = ".")
-
-# merge with daily data [+ clean up]
-df.pre <- merge(
-  x = dtS1$daily,
-  y = dat.pre.red,
-  by.x = "last_outside_referrer",
-  by.y = "platformId.pre",
-  all = T
-)
-
-# adjust duplicate names to fit to indicate daily or pre measurement
-names(df.pre) <- gsub("[[:punct:]]x", ".daily", names(df.pre))
-names(df.pre) <- gsub("[[:punct:]]y", ".pre", names(df.pre))
-
-# reduced data set for post-measurement
-dat.post.red <- dtS1$raw.post %>%
-  select(-starts_with("t_"))
-names(dat.post.red) <- paste(names(dat.post.red), "post", sep = ".")
-
-# merge post measurement with pre- and daily data
-df <- merge(
-  x = df.pre,
-  y = dat.post.red,
-  by.x = "last_outside_referrer",
-  by.y = "last_outside_referrer.post",
-  all = T
-)
-
-# adjust duplicate names to indicate pre or post
-names(df) <- gsub("[[:punct:]]x", ".pre", names(df))
-names(df) <- gsub("[[:punct:]]y", ".post", names(df))
-
-# add to list
-dtS1$combined <- df
-rm(df, df.pre, dat.post.red, dat.pre.red)
-
-# create data frame with cleaned data
-df <- dtS1$combined %>%
-  filter(
-    Finished.pre == 1,
-    Finished == 1,
-    !is.na(last_outside_referrer)
-  )
-
-# add running number as measurement ID within participants
-df$measureID <- data.table::rowidv(df, cols = c("last_outside_referrer"))
-
-dtS1$clean <- df
-
-# clean up
-rm(df)
-
-# Export reduced Data
-# write.csv(dtS1$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
-# save(dtS1$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
-```
-
-### Study 2
-
-For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement).
-We also reformat missing values and format core ID variables.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS2*' list.
-
-```{r cleanStudents}
-# our own test IDs
-ownIDs <- c(
-  "beautifulLionfishXXXR5rcgVBzGu8hPvOqrK8UBJBw4owvi9nfRFSFu3lMzYhE",
-  "niceDogoXXXmB8JI5SFu78SF3DVof84mGUPPNUr14p2HYFTtp31a6D1OwAzM6F-K",
-  "amusedQuailXXXmhuc_fpTp8vPkMwDH1BzjaH1d1kHSO1bsPEfsnaEYk4WeVBfPi",
-  "juwGAbtXX0_1kmZtSVqKh3PGaHOICqUyU4iBkrT3nDsI_uifuD1gzKcZerxaM5FL"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS2$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS2$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS2$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = FALSE
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = FALSE
-)
-
-# add to list
-dtS2$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-### Study 3
-
-For the medical professionals sample data was, again, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe.
-We exclude our own test data.
-We also reformat missing values and format core ID variables.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS3*' list.
-
-```{r cleanMedical}
-# our own test IDs
-ownIDs <- c(
-  "test_LeonieXXXSklxecPLW0-FBPM4796o3pUwUhAY5jb9KGw8jQsKxWmGpa1Jiy", 
-  "test_MaxXXXtOp_5dTNefIq0yKXtXt2IN6eEKxeHoPY9mlyvdsqPpLp1B0NGg4UL",
-  "test_JannisXXXBsNqk62fOpX6chbd2tMWPptUdjjnhAqnQ3uBqckZ7gLIEoPlfZ",
-  "quaintLeopardCatXXXAJ9cfSj-_SZLwNwMDxv_xv_iyr1Bg5YFLTlYdrjW0UXZY",
-  "blue-eyedIndianElephantXXXLf5zPMpQCDGS3umFzIj-YVky7ivTItvvozW49m"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS3$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS3$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>% 
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS3$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = F
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = F
-)
-
-# add to list
-dtS3$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-## Calculate needed transformations
-
-### Study 1
-
-For the worker sample, the data transformation stage had three main aims:
-
-1.
-We first corrected time indicators within the surveys.
-In some cases participants completed their daily diary surveys for the afternoon after midnight.
-In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsWorkers}
-df <- dtS1$clean
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S")
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    PID = as.numeric(factor(last_outside_referrer))
-  )
-
-# Time and Date Variables
-df <- df %>%
-  mutate(
-    TID = measureID - 1,
-    # time ID with t0 = 0 for meaningfull intercept interpretations
-    date = substr(created, 1, 10),
-    # awkward way of extracting date (best converted to )
-    time = substr(created, 12, 19),
-    # awkward way of extracting time
-    daynum = as.numeric(factor(dateQualtrics)),
-    # all days as numeric for ordering
-    daycor = ifelse(
-      daytime == "morning" &
-        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
-        daytime == "afternoon" &
-          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
-      daynum - 1,
-      daynum
-    ),
-    daycorTest = ifelse(
-      daytime == "morning" &
-        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
-        daytime == "afternoon" &
-          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
-      1,
-      0
-    ),
-    # correctly identify which date the questionnaire is about
-    daycor.lead = sprintf("%02d", daycor),
-    daytime.lt = ifelse(daytime == "morning", "a", "b"),
-    # morning / afternoon to a / b
-    day_time = paste(daycor.lead, daytime.lt, sep = "_"),
-    # combine day id with morning / afternoon
-    ResponseId = as.numeric(factor(day_time)),
-    # day and time identifier as numeric id
-    SubTime = chron::times(time.0),
-    time.daily = as.character(time),
-    PPDate = as.Date(df$dateQualtrics),
-    number = replace_na(ContactNum, 0),
-    NonDutchNum = replace_na(NonDutchNum, 0)
-  )
-
-# remove seconds from afternoon time
-df$SubTime[df$daytime == "afternoon"] <- paste0(substring(as.character(df$time.0[df$daytime == "afternoon"]), 4, 8), ":00")
-df$time.daily[df$daytime == "afternoon" &
-  !is.na(df$time.daily != "<NA>")] <- paste0(substring(as.character(df$time.daily[df$daytime == "afternoon" &
-  !is.na(df$time.daily != "<NA>")]), 4, 8), ":00")
-
-# Correct morning / afternoon date where survey was collected the day after to indicate the correct date that was targeted
-df$PPDate[df$SubTime < "11:50:00" &
-  df$daytime == "morning"] <- df$PPDate[df$SubTime < "11:50:00" &
-  df$daytime == "morning"] - 1
-df$PPDate[df$SubTime < "18:50:00" &
-  df$daytime == "afternoon"] <- df$PPDate[df$SubTime < "18:50:00" &
-  df$daytime == "afternoon"] - 1
-
-# Make time IDs consistent with later studies
-df$TID <- paste(df$PPDate, str_to_title(df$daytime))
-df$TIDnum <- as.numeric(factor(df$TID %>% gsub(" Morning", "-A", .) %>% gsub(" Afternoon", "-B", .)))
-
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = FALSE),
-    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum),
-    alertness = sum(alertness1, alertness2, na.rm = TRUE),
-    calmness = sum(calmness1, calmness2, na.rm = TRUE),
-    valence = sum(valence1, valence2, na.rm = TRUE)
-  ) %>%
-  ungroup()
-
-# Need scales
-df$keyMotiveFulfilled <- rowSums(df[, c("KeyNeedFulfillment", "DaytimeNeedFulfillment")], na.rm = T)
-df$autonomy.daily.all <- rowSums(df[, c("autonomy_Int", "autonomy_NoInt")], na.rm = T)
-df$competence.daily.all <- rowSums(df[, c("competence_Int", "competence_NoInt")], na.rm = T)
-# cor(df$relatednessOther, df$relatedness_self_1,use="complete.obs")
-df$relatedness.daily.all <- rowMeans(df[, c(
-  "relatednessOther",
-  "relatednessSelf",
-  "relatednessNoInteraction"
-)], na.rm = T)
-
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatedness <- rowMeans(df[, c("relatednessOther", "relatednessSelf")], na.rm = T)
-
-df$autonomy <- df$autonomy.daily.all
-df$competence <- df$competence.daily.all
-df$relatedness <- df$relatedness.daily.all
-
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-df$roommate.pre_calc <- df %>%
-  select(
-    starts_with("roommate")
-  ) %>% 
-  mutate_all(as_factor) %>%
-  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>%
-  pull
-
-df$Reason.pre_calc <- df %>%
-  select(
-    starts_with("Reason"),
-    -Reason_nodesire, 
-    -ReasonOther.pre 
-  ) %>% 
-  mutate_all(as_factor) %>% 
-  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>%
-  pull
-
-df$occupation.pre_calc <- df %>%
-  select(
-    starts_with("occupation"),
-    -occupation_8_TEXT.pre
-  ) %>%
-  mutate_all(as_factor) %>% 
-  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>% 
-  pull
-
-df$CurrentEducation.pre_calc <-
-  df %>%
-  select(starts_with("CurrentEducation"),
-         -CurrentEducation_6_TEXT.pre) %>%
-  mutate_all(as_factor) %>%
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$groupType.pre_calc <-
-  df %>%
-  select(starts_with("gr_type"),
-          -ends_with("TEXT")) %>%
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$groupContext.pre_calc <-
-  df %>%
-  select(starts_with("gr_context"),
-          -ends_with("TEXT")) %>%
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$dyadType.pre_calc <-
-  df %>%
-  select(starts_with("dyad_type"),
-          -ends_with("TEXT")) %>%
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$context.pre_calc <-
-  df %>%
-  select(starts_with("Context"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$desireType.pre_calc <-
-  df %>%
-  select(starts_with("desire_type"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$desireType.pre_calc <-
-  df %>%
-  select(starts_with("desire_context"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$nonDutchType.pre_calc <-
-  df %>%
-  select(starts_with("NonDutchType"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$autonomySat.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT0[1-4].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-df$autonomyFrust.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT0[5-8].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-
-df$relatednessSat.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT(09|1[0-2]).pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-df$relatednessFrust.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT1[3-6].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-
-df$competenceSat.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT(1[7-9]|20).pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-df$competenceFrust.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT2[1-4].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-
-df$IntGrAnx.pre_calc <-
-  psych::scoreItems(
-    keys = c(rep(1,6),rep(-1,6)),
-    items = df %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
-    min = 1,
-    max = 10
-  )$scores %>% 
-  as.numeric
-
-df$swl.pre_calc <-
-  psych::scoreItems(
-    keys = rep(1,5),
-    items = df %>% select(matches("^SWL[0-9]{2}\\.pre")),
-    min = 1,
-    max = 7,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-
-df$mdmq.alertness.pre_calc <- 
-  psych::scoreItems(
-    keys = c(1, -1, -1, 1),
-    items = df %>% select(MDMQ02.pre, MDMQ05.pre, MDMQ07.pre, MDMQ10.pre),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-df$mdmq.calmness.pre_calc <- 
-  psych::scoreItems(
-    keys = c(-1, 1, -1, 1),
-    items = df %>% select(MDMQ03.pre, MDMQ06.pre, MDMQ09.pre, MDMQ12.pre),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-df$mdmq.valence.pre_calc <- 
-  psych::scoreItems(
-    keys = c(1, -1, 1, -1),
-    items = df %>% select(MDMQ01.pre, MDMQ04.pre, MDMQ08.pre, MDMQ11.pre),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-
-df$IntGrAnx.post_calc <-
-  psych::scoreItems(
-    keys = c(rep(1,6),rep(-1,6)),
-    items = df %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
-    min = 1,
-    max = 10
-  )$scores %>% 
-  as.numeric
-
-df$swl.post_calc <-
-  psych::scoreItems(
-    keys = rep(1,5),
-    items = df %>% select(matches("^SWL[0-9]{2}\\.post")),
-    min = 1,
-    max = 7,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-
-varNamIndicesS1 <-
-  append(varNamIndicesS1,
-         df %>% select(ends_with("_calc")) %>% names)
-
-#
-df <- df %>%
-  mutate(
-    exWB = scales::rescale(exWB, from = range(exWB, na.rm = TRUE), to = c(-100, 100)),
-    relatedness = relatedness + 50,
-    autonomy = autonomy + 50,
-    competence = competence + 50,
-    KeyNeedFulfillment = KeyNeedFulfillment + 50,
-    KeyNeedDueToPartner = KeyNeedDueToPartner + 50,
-    InteractionContextAccidental = InteractionContextAccidental + 50,
-    InteractionContextvoluntary = InteractionContextvoluntary + 50,
-    InteractionContextCooperative = InteractionContextCooperative + 50,
-    InteractionContextRepresentativeNL = InteractionContextRepresentativeNL + 50,
-    qualityOverall = qualityOverall + 50,
-    qualityMeaning = qualityMeaning + 50,
-    DaytimeNeedFulfillment = DaytimeNeedFulfillment + 50
-  )
-dtS1$full <- df
-rm(df)
-```
-
-### Study 2
-
-For the student sample, the data transformation stage had five main aims:
-
-1.  We first create person, survey type, and measurement ID variables.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  We add information about the interaction partner to the beep during which a person was selected as an interaction partner.\
-4.  We cluster mean-center key variables within participants.\
-5.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsStudents}
-df <- dtS2$clean
-
-# Add ID variables
-df$PID <- as.numeric(factor(df$ResponseId)) # participant ID
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S"),
-    date_period_old = date_period,
-    date_period = ifelse(
-      startTime >= c("12:00") & startTime <= c("18:59"),
-      paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Morning"),
-      ifelse(
-        startTime >= c("19:00") & startTime <= c("23:59"),
-        paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Afternoon"),
-        ifelse(
-          startTime >= c("00:00") & startTime <= c("11:59"),
-          paste(as.Date(created) - 1, "Afternoon"),
-          NA
-        )
-      )
-    )
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    dateOrder = stri_replace_all_regex(
-      date_period,
-      pattern = c(' Morning', ' Afternoon'),
-      replacement = c('_A', '_B'),
-      vectorize = FALSE
-    ),
-    TID = factor(date_period),
-    TIDnum = as.numeric(factor(dateOrder))
-  )
-
-# check whether time ordering worked
-df <- df %>%
-  arrange(PID, TID)
-
-# Interaction as Factor
-df$interaction.f <-
-  factor(df$Interaction,
-    levels = c("no interaction", "Dutch", "Non-Dutch")
-  )
-df$intNL <- ifelse(df$Interaction == "Dutch", 1, 0)
-df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)
-
-df$IntergroupContact <- (df$IntergroupContact-2)*-1
-df$IngroupContact <- (df$IngroupContact-2)*-1
-
-# -------------------------------------------------------------------------------------------------------------
-#                                       Combine Variables
-# -------------------------------------------------------------------------------------------------------------
-# Relatedness
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatednessInteraction <- rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
-df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
-# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; JK: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
-df$relatedness <-
-  rowMeans(df[, c("relatednessInteraction", "relatednessNoInteraction")],
-           na.rm = TRUE) %>%
-  ifelse(is.nan(.), NA, .)
-
-# Core Need
-df$DaytimeNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
-  df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]
-
-df$InteractionNeedFullfillment <- NA 
-df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)] <- 
-  df$KeyNeedFulfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)]
-
-# Randomly selected variables
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    ProSo = mean(c(ProSo1, ProSo2, ProSo3, ProSo4), na.rm = TRUE),
-    AntiSo = mean(c(AntiSo1, AntiSo2, AntiSo3, AntiSo4, AntiSo5, AntiSo6, AntiSo7), na.rm = TRUE)
-  )
-
-# -------------------------------------------------------------------------------------------------------------
-#                                 Add Variables related to interaction partner
-# -------------------------------------------------------------------------------------------------------------
-# create function for later lapply
-createIntPartDf <- function(inp) {
-  # prepare the dataframe so that we can forloop over it later
-  tmp <- data.frame(
-    CC = as.character(inp$CC),
-    NewCC = as.character(inp$NewCC),
-    NewName = as.character(inp$NewName),
-    NewCloseness = inp$NewCloseness,
-    NewGender = inp$NewGender,
-    NewEthnicity = as.character(inp$NewEthnicity),
-    NewRelationship = as.character(inp$NewRelationship)
-  )
-
-  tmp$CC2 <- recode(tmp$CC, "SOMEONE ELSE" = "NA")
-  tmp$CC2 <-
-    ifelse(
-      tmp$CC == 1 |
-        tmp$CC == "SOMEONE ELSE",
-      as.character(tmp$NewName),
-      as.character(tmp$CC2)
-    )
-  # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
-  tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
-  tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)
-
-  # open the variables that will be filled up in the foor-loop
-  tmp$closeness <- rep(NA, nrow(tmp))
-  tmp$gender <- rep(NA, nrow(tmp))
-  tmp$ethnicity <- rep(NA, nrow(tmp))
-  tmp$relationship <- rep(NA, nrow(tmp))
-
-  # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
-  # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
-  for (i in 1:nrow(tmp)) {
-    if (is.na(tmp$CC2[i])) {
-      next
-    } else {
-      tmp$closeness[i] <-
-        na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # find closeness where CC2 matches NewName (na.omit + [1] to get the number)
-      tmp$gender[i] <-
-        na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # (na.omit + [1] to get the number and not the rest of the na.omit list)
-      tmp$ethnicity[i] <-
-        na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] # PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
-      tmp$relationship[i] <-
-        na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
-    }
-  }
-
-  out <- tmp
-  out
-}
-
-# split df per participants and run function
-PP <- split(df, df$PID)
-PP <- lapply(PP, createIntPartDf)
-rm(createIntPartDf)
-
-# add variables back to df
-remergePP <- do.call(rbind.data.frame, PP)
-colnames(remergePP) <-
-  paste(colnames(remergePP), "_Calc", sep = "")
-df <- cbind(df, remergePP)
-rm(remergePP, PP)
-
-dtS2$full <- df
-rm(df)
-```
-
-### Study 3
-
-For the medical professional sample, the data transformation stage had five main aims:
-
-1.  We first create person, survey type, and measurement ID variables.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  We cluster mean-center key variables within participants.\
-4.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsMedical}
-df <- dtS3$clean
-
-# Add ID variables
-df$PID <- as.numeric(factor(df$ResponseId)) # participant ID
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S"),
-    date_period_old = date_period,
-    date_period = ifelse(
-      startTime >= c("12:00") & startTime <= c("18:59"),
-      paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Morning"),
-      ifelse(
-        startTime >= c("19:00") & startTime <= c("23:59"),
-        paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Afternoon"),
-        ifelse(
-          startTime >= c("00:00") & startTime <= c("11:59"),
-          paste(as.Date(created) - 1, "Afternoon"),
-          NA
-        )
-      )
-    )
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    dateOrder = stri_replace_all_regex(
-      date_period,
-      pattern = c(' Morning', ' Afternoon'),
-      replacement = c('_A', '_B'),
-      vectorize = FALSE
-    ),
-    TID = factor(date_period),
-    TIDnum = as.numeric(factor(dateOrder))
-  )
-
-# check whether time ordering worked
-df <- df %>%
-  arrange(PID, TID)
-
-# Interaction as Factor
-df$interaction.f <-
-  factor(df$Interaction,
-    levels = c("no interaction", "Dutch", "Non-Dutch")
-  )
-df$intNL <- ifelse(df$Interaction == "Dutch", 1, 0)
-df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)
-
-df <- df %>%
-  mutate(
-    NonDutchContact = replace_na(NonDutchNum, 0), # make second non-Dutch countable
-    NonDutchContact = ifelse(NonDutchContact > 1, 1, 0) # recode (yes = 1 -> 1, no = 2 -> 0)
-  ) %>%
-  mutate(
-    OutgroupInteraction = factor(
-      InteractionDumDutch,
-      levels = c(0, 1),
-      labels = c("No", "Yes")
-    ),
-    NonOutgroupInteraction = factor(
-      rowSums(select(., c(InteractionDumNonDutch, NonDutchContact)), na.rm = TRUE), # combine the two non-Dutch Q.,
-      levels = c(0, 1),
-      labels = c("No", "Yes")
-    )
-  )
-
-df$IntergroupContact <- (df$IntergroupContact-2)*-1
-df$IngroupContact <- (df$IngroupContact-2)*-1
-
-# -------------------------------------------------------------------------------------------------------------
-#                                       Combine Variables
-# -------------------------------------------------------------------------------------------------------------
-# Relatedness
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatednessInteraction <-
-  rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
-df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
-# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
-df$relatedness <-
-  rowMeans(df[, c("relatednessInteraction", "relatednessNoInteraction")],
-           na.rm = TRUE) %>%
-  ifelse(is.nan(.), NA, .)
-
-df$DaytimeNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
-  df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]
-
-df$InteractionNeedFullfillment <- NA 
-df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)] <- 
-  df$KeyNeedFulfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)]
-
-df$InteractionNeedImportance <- NA 
-df$InteractionNeedImportance[df$InteractionDum == 1 & !is.na(df$KeyNeedImp)] <- 
-  df$KeyNeedImp[df$InteractionDum == 1 & !is.na(df$KeyNeedImp)]
-
-# Randomly selected variables
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    ProSo = mean(c(ProSo1, ProSo2, ProSo3, ProSo4), na.rm = TRUE),
-    AntiSo = mean(c(AntiSo1, AntiSo2, AntiSo3, AntiSo4, AntiSo5, AntiSo6, AntiSo7), na.rm = TRUE),
-    agency = mean(c(agency1, agency2, agency3), na.rm = TRUE),
-    autoFrust = mean(c(autoFrust1, autoFrust2, autoFrust3, autoFrust4), na.rm = TRUE),
-    autoSat = mean(c(autoSat1, autoSat2, autoSat3, autoSat4), na.rm = TRUE),
-    relatFrust = mean(c(relatFrust1, relatFrust2, relatFrust3, relatFrust4), na.rm = TRUE),
-    relatSat = mean(c(relatSat1, relatSat2, relatSat3, relatSat4), na.rm = TRUE),
-    compFrust = mean(c(compFrust1, compFrust2, compFrust3, compFrust4), na.rm = TRUE),
-    compSat = mean(c(compSat1, compSat2, compSat3, compSat4), na.rm = TRUE),
-    lonely = mean(c(lonely4, lonely4, lonely4, lonely4), na.rm = TRUE),
-    emotRegPos = mean(c(emotRegPos01, emotRegPos02), na.rm = TRUE),
-    emotRegNeg = mean(c(emotRegNeg01, emotRegNeg02), na.rm = TRUE)
-  )
-
-# Allport's Conditions
-df %>%
-  #filter(OutgroupInteraction == "Yes") %>%
-  select(
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  ) %>%
-  pairs.panels.new
-
-AlportDescr <- df %>%
-  #filter(OutgroupInteraction == "Yes") %>%
-  select(
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  ) %>%
-  psych::describe(., skew=F,ranges=T) %>%
-  as.data.frame() %>%
-  select(-vars) %>%
-  kable(., caption = "Descriptives of Allport's Condition items") %>% 
-  kable_styling("hover", full_width = F, latex_options = "hold_position")
-
-iaWorkerAllport <- 
-  df %>%
-  #filter(OutgroupInteraction == "Yes") %>%
-  select(
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  )
-
-itemScaleAllport01 <- sjPlot::tab_itemscale(iaWorkerAllport)
-
-pca <- parameters::principal_components(iaWorkerAllport)
-factor.groups <- parameters::closest_component(pca)
-
-sjPlot::tab_itemscale(iaWorkerAllport, factor.groups, show.kurtosis = TRUE)
-
-AllportAlpha <- ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)
-
-data <- 
-  df %>%
-  select(
-    PID,
-    TIDnum,
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  ) %>%
-  drop_na %>%
-  reshape2::melt(
-    ., 
-    id.vars = c("PID", "TIDnum")
-  )
-
-AllportNestedAlpha <- horst::nestedAlpha(item.level.1 = "value",
-                   level.2      = "TIDnum",
-                   level.3      = "PID",
-                   data         = data)
-rm(data)
-
-iaWorkerAllportScale <- 
-  iaWorkerAllport %>%
-  Scale::Scale() %>%
-  Scale::ItemAnalysis()
-
-df$AllportsCondition <-
-  scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(
-      InteractionContextEqualStatus,
-      KeyNeedShared,
-      InteractionContextCooperative,
-      InteractionContextvoluntary
-    ),
-    min = 0,
-    max = 100
-  )$scores
-
-as.data.frame(psych::describe(df$AllportsCondition, skew=T)) %>%
-  mutate(vars = "Allport's Conditions Index") %>%
-  kable(., caption = "Allport's Conditions: Scale Descriptives", row.names = FALSE) %>% 
-  kable_styling("hover", full_width = F, latex_options = "hold_position")
-
-ggplot(df, aes(x = AllportsCondition)) +
-  geom_histogram() +
-  theme_Publication()
-
-# -------------------------------------------------------------------------------------------------------------
-#                                 Add Variables related to interaction partner
-# -------------------------------------------------------------------------------------------------------------
-# create function for later lapply
-createIntPartDf <- function(inp) {
-  # prepare the dataframe so that we can forloop over it later
-  tmp <- data.frame(
-    CC = as.character(inp$CC),
-    NewCC = as.character(inp$NewCC),
-    NewName = as.character(inp$NewName),
-    NewCloseness = inp$NewCloseness,
-    NewGender = inp$NewGender,
-    NewEthnicity = as.character(inp$NewEthnicity),
-    NewRelationship = as.character(inp$NewRelationship)
-  )
-
-  tmp$CC2 <- recode(tmp$CC, "SOMEONE ELSE" = "NA")
-  tmp$CC2 <-
-    ifelse(
-      tmp$CC == 1 |
-        tmp$CC == "SOMEONE ELSE",
-      as.character(tmp$NewName),
-      as.character(tmp$CC2)
-    )
-  # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
-  tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
-  tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)
-
-  # open the variables that will be filled up in the foor-loop
-  tmp$closeness <- rep(NA, nrow(tmp))
-  tmp$gender <- rep(NA, nrow(tmp))
-  tmp$ethnicity <- rep(NA, nrow(tmp))
-  tmp$relationship <- rep(NA, nrow(tmp))
-
-  # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
-  # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
-  for (i in 1:nrow(tmp)) {
-    if (is.na(tmp$CC2[i])) {
-      next
-    } else {
-      tmp$closeness[i] <-
-        na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # find closeness where CC2 matches NewName (na.omit + [1] to get the number)
-      tmp$gender[i] <-
-        na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # (na.omit + [1] to get the number and not the rest of the na.omit list)
-      tmp$ethnicity[i] <-
-        na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] # PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
-      tmp$relationship[i] <-
-        na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
-    }
-  }
-
-  out <- tmp
-  out
-}
-
-# split df per participants and run function
-PP <- split(df, df$PID)
-PP <- lapply(PP, createIntPartDf)
-rm(createIntPartDf)
-
-# add variables back to df
-remergePP <- do.call(rbind.data.frame, PP)
-colnames(remergePP) <-
-  paste(colnames(remergePP), "_Calc", sep = "")
-df <- cbind(df, remergePP)
-rm(remergePP, PP)
-
-dtS3$full <- df
-```
-
-## Data Availability and Sample Selection {.tabset .tabset-fade}
-
-As one of our main analyses is a three-mode principal component analysis (3MPCA) we begin by assessing the amount of missing data in each of the ESM studies.
-To assess the missingness in detail, we prepare a data availability table of whether data is available or missing for each possible measurement point (data + morning/afternoon) for all individual participants.
-We then export the data availability tables as comma separated value files (.csv) to be assessed in detail using a spreadsheet program (such as MS Excel).
-
-For our sample selection we address both the selection of time points and participants.
-Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
-Thus, we then select the time points for which we have less than 45% missingness and select participants who have less than 45% missingness across the selected time range.
-
-To approximate an ideal sample selection that maximized the number of participants and timepoints that fit the 55% data availability criterion.
-We select sequentially remove the rows or columns with the lowest data availability rate untill all rows and columns have more than 55% data available (i.e., less than 45% missingness).
-
-::: {.alert .alert-info .alert-dismissible .fade .in role="alert"}
-<a class="close" data-dismiss="alert" aria-label="close">×</a> <i class="fas fa-info-circle"></i> <b>Note: </b><br/> This needs to be adjusted current missingness criterion set to more conservative 33%.
-:::
-
-### Study 1
-
-```{r workerDataAvailability}
-dtS1Availability <- dtS1$full %>%
-  select(
-    PID,
-    TIDnum
-  ) %>%
-  arrange(PID, TIDnum) %>%
-  mutate(data = 1)
-dtS1Availability <- reshape::cast(dtS1Availability, PID ~ TIDnum) %>%
-  select(-PID) %>%
-  mutate_all(function(x) ifelse(x>1,1,x))
-# sum(colMeans(dtS1Availability)*100 >= 66)
-# sum(rowMeans(dtS1Availability)*100 >= 66)
-# sum(colMeans(dtS1Availability[-5,])*100 >= 66)
-# sum(rowMeans(dtS1Availability[,-8])*100 >= 66)
-
-rownames(dtS1Availability) <- paste("PP", 1:nrow(dtS1Availability), sep = "_")
-colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
-dtS1AvailabilityKbl <- dtS1Availability
-
-dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)] <- lapply(dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS1AvailabilityKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 1: Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-
-write.csv(dtS1Availability, "data/S1_Workers/processed/workerAvailability.csv")
-```
-
-```{r testOptimizer, include=FALSE}
-# set.seed(42)
-# nr <- 5
-# nc <- 10
-# df <- as.data.table(matrix(rbinom(nr*nc,1,.66), nrow=nr, ncol=nc))
-# df
-# 
-# best.list.row.df = list()
-# for (i in 1:nrow(df)) {
-#   # get best subset for rows based on how many columns have more than 66% data
-#   rowlist = combn(nrow(df), i)
-#   numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
-#   cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
-#   best = which(numobs == max(numobs))[1]
-#   best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
-# }
-# 
-# best.list.col.df = list()
-# for (i in 1:ncol(df)) {
-#   # get best subset for columns based on how many rows have more than 66% data
-#   collist = combn(ncol(df), i)
-#   numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
-#   cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
-#   best = which(numobs == max(numobs))[1]
-#   best.list.col.df = c(best.list.col.df, list(collist[, best]))
-# }
-# rm(df, best.list.row.df, best.list.col.df)
-```
-
-```{r testOptimizer03, include=FALSE}
-# library(dplyr)
-# library(ROI)
-# library(ROI.plugin.glpk)
-# library(ompr)
-# library(ompr.roi)
-# 
-# set.seed(42)
-# tf <- matrix(sample(c(TRUE, FALSE), 1488, replace=TRUE), 31)
-# M3 <- t(replicate(31, 1969:2016, simplify=TRUE))
-# M3[tf] <- NA
-# 
-# m <- +!is.na(M3) # gets logical matrix; 0 if NA else 1    
-# #m <- as.data.frame(dtS1AvailabilityTest)
-# nr <- nrow(m)
-# nc <- ncol(m)
-# n_years <- 15 
-# 
-# 
-# model <- MIPModel() %>% 
-#   # keep[i,j] is 1 if matrix cell [i,j] is to be kept else 0
-#   add_variable(keep[i,j], i = 1:nr, j = 1:nc, typ = "binary") %>% 
-#   # rm_row[i] is 1 if row i is selected for removal else 0
-#   add_variable(rm_row[i], i = 1:nr, type = "binary") %>% 
-#   # rm_col[j] is 1 if column j is selected for removal else 0
-#   add_variable(rm_col[j], j = 1:nc, type = "binary") %>% 
-#   # maximize good cells kept
-#   set_objective(sum_expr(keep[i,j], i = 1:nr, j = 1:nc), "max") %>% 
-#   # cell can be kept only when row is not selected for removal
-#   add_constraint(sum_expr(keep[i,j], j = 1:nc) <= 1 - rm_row[i], i = 1:nr) %>%
-#   # cell can be kept only when column is not selected for removal
-#   add_constraint(sum_expr(keep[i,j], i = 1:nr) <= 1 - rm_col[j], j = 1:nc) %>%
-#   # keep at most n_years columns i.e. remove at least (nc - n_years) columns
-#   # only non-NA values can be kept
-#   add_constraint(m[i,j] + rm_row[i] + rm_col[j] >= 1, i = 1:nr, j = 1:nc) %>% 
-#   #add_constraint(sum_expr(m[,j], j = 1:nc)/nc*100 >= 66) %>% 
-#   #add_constraint(sum_expr(m[i,], i = 1:nr)/nr*100 >= 66) %>% 
-#   # keep cols with more 
-#   #add_constraint(colMeans(m[i,j], i = 1:nr, j = 1:nc) >= 0.66) %>%
-#   # I used >= instead of == to avoid infeasiblity
-#   #add_constraint(sum_expr(rm_col[j], j = 1:nc) >= nc - n_years) %>% 
-#   # solve using free glpk solver
-#   solve_model(with_ROI(solver = "glpk", verbose = TRUE))
-# 
-# solver_status(model)
-# rm_rows <- model %>% 
-#   get_solution(rm_row[i]) %>% 
-#   filter(value > 0) %>% 
-#   pull(i)
-# 
-# rm_cols <- model %>% 
-#   get_solution(rm_col[j]) %>% 
-#   filter(value > 0) %>% 
-#   pull(j) 
-# 
-# result <- m[-rm_rows, -rm_cols, drop = F]
-# result
-```
-
-```{r testOptimizer05}
-cleanM <- function(M, c = 55) { #c = 66
-  check <- list()
-  rowNamesOut <- c()
-  colNamesOut <- c()
-  for (i in 1:length(c(row.names(M), names(M)))) {
-    rowMeans <- rowMeans(M) * 100
-    colMeans <- colMeans(M) * 100
-    rcMeans <- c(rowMeans, colMeans)
-    rm <- which.min(rcMeans) %>% names
-    rc <- ifelse(startsWith(rm, "PP_"), "row", "col")
-
-    check[[i]] <- rcMeans
-
-    if (!all(rcMeans >= c) && rc == "row") {
-      M <- M[!(row.names(M) %in% rm), ]
-      rowNamesOut <- append(rowNamesOut, rm)
-      cat(i, ": Row ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
-    } else if (!all(rcMeans >= c) && rc == "col") {
-      M <- M[, !(names(M) %in% rm)]
-      colNamesOut <- append(colNamesOut, rm)
-      cat(i, ": Column ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
-    } else {
-      cat(i, ": All row- and column means are over ", c, "%. The final matrix has ", nrow(M), " rows and ", ncol(M), " columns.", sep = "")
-      return(
-        list(
-          reducedMatrix = M,
-          rowNamesIn = row.names(M),
-          colNamesIn = names(M),
-          rowNamesOut = rowNamesOut,
-          colNamesOut = colNamesOut,
-          rowMeans = rowMeans,
-          colMeans = colMeans
-        )
-      )
-    }
-  }
-}
-
-dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 55)
-
-PIDout <- gsub("PP_", "", dtS1RedInfo$rowNamesOut) %>% as.numeric
-TIDout <- gsub("t_", "", dtS1RedInfo$colNamesOut) %>% as.numeric
-TIDInRed <- gsub("t_", "", dtS1RedInfo$colNamesIn) %>% as.numeric
-TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
-
-dtS1Red <- dtS1$full %>%
-  filter(
-    !PID %in% PIDout,
-    TIDnum %in% TIDIn
-  ) %>% 
-  mutate(TIDnum = TIDnum - min(TIDnum))
-rm(PIDout, TIDout, TIDInRed, TIDIn)
-
-# change glitch where survey was sent too early
-dtS1Red$TIDnum[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- 10
-dtS1Red$TID[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- "2018-05-18 Morning"
-
-dtS1AvailabilityRedKbl <- dtS1RedInfo$reducedMatrix
-dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)] <- lapply(dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS1AvailabilityRedKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 1: Final Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-### Study 2
-
-```{r studentDataAvailability}
-dtS2Availability <- dtS2$full %>%
-  select(
-    PID,
-    TIDnum
-  ) %>%
-  #na.omit %>%
-  arrange(PID, TIDnum) %>%
-  mutate(data = 1)
-dtS2Availability <- reshape::cast(dtS2Availability, PID ~ TIDnum) %>%
-  select(-PID)
-
-rownames(dtS2Availability) <- paste("PP", 1:nrow(dtS2Availability), sep = "_")
-colnames(dtS2Availability) <- paste("t", 1:ncol(dtS2Availability), sep = "_")
-
-write.csv(dtS2Availability, "data/S2_Students/processed/studentAvailability.csv")
-
-dtS2AvailabilityKbl <- dtS2Availability
-dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)] <- lapply(dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS2AvailabilityKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 2: Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-```{r s2SampleReduction}
-dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 55)
-
-PIDout <- gsub("PP_", "", dtS2RedInfo$rowNamesOut) %>% as.numeric
-TIDout <- gsub("t_", "", dtS2RedInfo$colNamesOut) %>% as.numeric
-TIDInRed <- gsub("t_", "", dtS2RedInfo$colNamesIn) %>% as.numeric
-TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
-
-dtS2Red <- dtS2$full %>%
-  filter(
-    !PID %in% PIDout,
-    TIDnum %in% TIDIn
-  ) %>% 
-  mutate(TIDnum = TIDnum - min(TIDnum))
-rm(PIDout, TIDout, TIDInRed, TIDIn)
-
-# remove glitch entries 
-dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 57 & dtS2Red$ended == "2018-12-20 21:58:31"),]
-dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 59 & dtS2Red$ended == "2018-12-20 21:58:31"),]
-
-dtS2AvailabilityRedKbl <- dtS2RedInfo$reducedMatrix
-dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)] <- lapply(dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS2AvailabilityRedKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 2: Final Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-### Study 3
-
-```{r medicalDataAvailability}
-dtS3Availability <- dtS3$full %>%
-  select(
-    PID,
-    TIDnum
-  ) %>%
-  arrange(PID, TIDnum) %>%
-  mutate(data = 1)
-dtS3Availability <- reshape::cast(dtS3Availability, PID ~ TIDnum, fill = 0) %>%
-  select(-PID)
-
-rownames(dtS3Availability) <- paste("PP", 1:nrow(dtS3Availability), sep = "_")
-colnames(dtS3Availability) <- paste("t", 1:ncol(dtS3Availability), sep = "_")
-
-write.csv(dtS3Availability, "data/S3_Medical/processed/medicalAvailability.csv")
-
-dtS3AvailabilityKbl <- dtS3Availability
-dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)] <- lapply(dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS3AvailabilityKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 2: Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-rm(dtS3AvailabilityKbl)
-```
-
-```{r s3SampleReduction}
-dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 55)
-
-PIDout <- gsub("PP_", "", dtS3RedInfo$rowNamesOut) %>% as.numeric
-TIDout <- gsub("t_", "", dtS3RedInfo$colNamesOut) %>% as.numeric
-TIDInRed <- gsub("t_", "", dtS3RedInfo$colNamesIn) %>% as.numeric
-TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
-
-dtS3Red <- dtS3$full %>%
-  filter(
-    !PID %in% PIDout,
-    TIDnum %in% TIDIn
-  ) %>% 
-  mutate(TIDnum = TIDnum - min(TIDnum))
-rm(PIDout, TIDout, TIDInRed, TIDIn)
-
-dtS3AvailabilityRedKbl <- dtS3RedInfo$reducedMatrix
-dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)] <- lapply(dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS3AvailabilityRedKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 3: Final Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-## Time Scales and Timepoint Aggregation
-
-```{r checkVarAvailability}
-varNamesPre <- varNames %>%
-  filter(
-    surveyS1 == "pre",
-    surveyS2 == "pre",
-    str_detect(surveyS3, "^pre")
-  ) %>%
-  select(varNam)
-varNamesPreFull <-
-  data.frame(varNam = Reduce(intersect, list(
-    names(dtS1$full), names(dtS3$full), names(dtS3$full)
-  ))) %>%
-  filter(str_detect(varNam, ".pre$"))
-
-varNamesDaily <- varNames %>%
-  filter(
-    surveyS1 == "daily",
-    surveyS2 == "daily",
-    surveyS3 == "daily"
-  ) %>%
-  select(varNam)
-varNamesDailyFull <-
-  data.frame(varNam = Reduce(intersect, list(
-    names(dtS1Red), names(dtS2Red), names(dtS3Red)
-  ))) %>%
-  filter(!str_detect(varNam, ".pre$|.post$"))
-
-varNamesPost <- varNames %>%
-  filter(
-    surveyS1 == "post",
-    surveyS2 == "post",
-    surveyS3 == "post"
-  ) %>%
-  select(varNam)
-varNamesPostFull <-
-  data.frame(varNam = Reduce(intersect, list(
-    names(dtS1Red), names(dtS3Red), names(dtS3Red)
-  ))) %>%
-  filter(str_detect(varNam, ".post$"))
-```
-
-```{r analysisDFs}
-# Main Analysis: 3MPCA (common variables across studies)
-dropVarPcaMain <- c(
-  # Meta Data
-  "last_outside_referrer",
-  "created",
-  "ended",
-  "ip_address",
-  "date",
-  "ResponseId",
-  "server_language",
-  "Meta_Browser",
-  # Interaction Dummies (non-continuous)
-  "IntergroupContact",
-  "IngroupContact",
-  "InteractionDum",
-  # Interaction Counts (non-continuous)
-  "ContactNum",
-  "NonDutchNum",
-  "groupDutch",
-  # Interaction Descriptives (non-continuous)
-  "duration",
-  "dyadGroup",
-  "groupSize",
-  # Keyneed Freetexts (non-continuous)
-  "KeyNeed",
-  "DaytimeNeed",
-  # Relatedness individual items
-  "relatednessSelf",
-  "relatednessOther",
-  "relatednessNoInteraction"
-)
-
-varNam3mpcaMain <- varNamesDailyFull %>%
-  filter(!varNam %in% dropVarPcaMain) %>%
-  pull
-
-dtMain3mpca <- rbind(
-  dtS1Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
-  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
-  ) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-    ) %>%
-  select(
-    ID,
-    PID,
-    TID,
-    date,
-    week,
-    TIDnum,
-    study,
-    everything()
-  ) %>%
-  arrange(ID, TIDnum)
-
-dtMain3mpcaDaily <- dtMain3mpca %>%
-  group_by(ID, date, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(date))) %>%
-  ungroup %>%
-  select(ID, date, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .))
-
-dtMain3mpcaWeekly <- dtMain3mpca %>%
-  group_by(ID, week, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(week))) %>%
-  ungroup %>%
-  select(ID, week, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .))
-```
-
-We aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.
-Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
-This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
-We chose to select time scales that align with common human cycles.
-We thus compare the variances of bi-daily, daily, and weekly aggregations.
-Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.
-We then chose the time scales that have the most variance.
-
-```{r keyNeedTest}
-
-# WILL USE IMPUTED DATA AND FOR-LOOP ACROSS ALL RELEVANT DVs
-# PLOT EFFECT SIZES
-
-dtMain3mpca <- dtMain3mpca %>%
-  group_by(study) %>%
-  mutate(
-    TIDdate = as.numeric(factor(date)),
-    TIDweek = as.numeric(factor(week))
-  ) %>%
-  ungroup
-
-# Not sure whether TIDs as predictors or empty model?
-keyNeedMlTest01 <-
-  lme4::lmer(KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek + (1 + TIDnum + TIDdate + TIDweek | ID),
-    data = dtMain3mpca
-  )
-
-keyNeedMlTest02 <-
-  lme(
-    KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek,
-    random = ~ 1 + TIDnum + TIDdate + TIDweek | ID,
-    data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
-    control = list(opt = "nlmimb")
-  ) # use optim if it does not converge
-
-# summary(keyNeedMlTest02)
-# summ(keyNeedMlTest01, digits = 3, center = FALSE)
-# anova(keyNeedMlTest02)
-
-```
 
@@
 </div>
-
-# **Software Information**
-
-The full session information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
-
-<details>
-
-<summary>
-
-System Info
-
-</summary>
-
-\renewcommand{\arraystretch}{0.8}
-
-<!-- decrease line spacing for the table -->
-
-```{r Reproducibility-sessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
-df_session_platform <- devtools::session_info()$platform %>%
-  unlist(.) %>%
-  as.data.frame(.) %>%
-  rownames_to_column(.)
-
-colnames(df_session_platform) <- c("Setting", "Value")
-
-kbl(
-  df_session_platform,
-  booktabs = T,
-  align = "l",
-  caption = "R environment session info for reproducibility of results" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\renewcommand{\arraystretch}{1}
-
-<!-- reset row height/line spacing -->
-
-</details>
-
-<br>
-
-<details>
-
-<summary>
-
-Package Info
-
-</summary>
-
-\renewcommand{\arraystretch}{0.6}
-
-<!-- decrease line spacing for the table -->
-
-```{r Reproducibility-sessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
-df_session_packages <- devtools::session_info()$packages %>%
-  as.data.frame(.) %>%
-  filter(attached == TRUE) %>%
-  dplyr::select(loadedversion, date, source) %>%
-  rownames_to_column()
-
-colnames(df_session_packages) <- c("Package", "Loaded version", "Date", "Source")
-
-kbl(
-  df_session_packages,
-  booktabs = T,
-  align = "l",
-  caption = "Package info for reproducibility of results" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\renewcommand{\arraystretch}{1}
-
-<!-- reset row height/line spacing -->
-
-</details>
-
-<br>
-
-<details>
-
-<summary>
-
-Full Session Info (including loaded but unattached packages --- for troubleshooting only)
-
-</summary>
-
-`r pander(sessionInfo(), compact = FALSE)`
-
-</details>
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-# **References**
-
-::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
-:::

### 2022-10-02 Supplemental-Material-A-Annotated-Analysis.Rmd
re-render with main comments added

@@
 
-## Multiple Imputation
-
-We ultimately aim to create 20 imputed data sets to perform the analyses on.
-As outlined by @Monden2015, we will use the key variables themselves as well as auxiliary variables to impute missing values.
-Auxiliary variables are any pre-, post-, or daily dairy variables other than the main variables.
-If there are any modeling issues we will limit the auxiliary variables to all variables that that are significantly ($p$ \< .01) and meaningfully ($r$ \> .3) correlated with (1) the key variables or (2) missingness on the key variables.
-
-We create the imputed data set for each of the analyses.
-This includes the main analyses with across all three studies, using the set of common variables, as well as follow-up analyses for study-specific, interaction-specific and time scale specific data sets.
-
-```{r varNames}
+# **Feature-based Time-Series Clustering**
+
+Feature-based clustering has the advantage of being extremely flexible when it comes to dealing with real-world human data.
+Human data is commonly noisy, incomplete, and non-stationary.
+Many descriptive dimension-reduction, modeling, and raw time-series clustering procedures will require complete or stationary data from the researcher, and are thus currently not well-suited to deal with messy intensive-longitudinal data collected by many social science researchers.
+
+We used the following five-step procedure to arrive at a stable, reproducible, and meaningful differentiation of participants ESM trajectories:
+
+1. Variable selection
+2. Feature extraction
+3. (Feature selection)
+4. Clustering
+5. Visualization, interpretation, and validation
+
+## Variable Selection
+
+We begin by selecting the variables, which we would like to include in the clustering procedure.
+For the clustering of time-series characteristics we focus on variables that are specific to the ESM measurements (i.e., variables collected during the various daily dairy pings).
+This means, that variables from the pre- and post-test are excluded from the clustering procedure.
+While the variables can in theory be added as additional features, they are not the focus of this manuscript.
+
+In our case, we begin by collecting the variables that are in the ESM measurements of all three studies, in order to build the most powerful dataset for the main analysis.
+We then additionally, divide the variables by their response type and by study to check for cluster differences in the validation phase:
+
+1. *Main Analysis*: Variables across all three data sets
+2. *Validation*: Variables by response type (interaction-specific or not)
+3. *Validation*: Variables by study
+
+```{r Preparation of Variable Names}
 varNamS123Aux <- varNames %>%
@@
 
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-### Main Analysis
-
-We begin by preparing the data for the main analysis, which combines the common variables of all three studies. Here, we:
-
-1.  Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
-2.  Create the combined dataset in the correct format (long format with all missed time points as NAs).
-3.  Impute the data and save the imputed datasets for later recall.
-
-```{r MiMainPrep}
-#################################################################
-##                    Variable Name Vectors                    ##
-#################################################################
 varNamS123MIPsbl <- c(
@@
 
-varNamMiMainNoms <- varNames %>%
+idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
+```
+
+```{r variable names for all analyses}
+# ID Variables (just to re-iterate)
+idVars <- idVars
+
+# Common variables across studies 
+varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
+
+# Variable Groups
+# varNames %>%
+#   filter(pca != 0) %>%
+#   mutate(contactSpecific = recode(.$contactSpecific, 
+#                          `-1`="No Interaction Only",
+#                          `0`="unspecific",
+#                          `1`="Interaction Only")) %>%
+#   crosstable::crosstable(., varGroup, by=contactSpecific) %>%
+#   select(-c(.id, label)) %>%
+#   kbl(.,
+#       escape = FALSE,
+#       booktabs = T,
+#       align = c("l", "r", "c"),
+#       digits=2,
+#       caption = "Full Analysis: Variable Groups by Interaction Types") %>%
+#   kable_classic(
+#     full_width = F,
+#     lightable_options = "hover",
+#     html_font = "Cambria"
+#   )
+# 
+# varNames %>%
+#   filter(
+#     pca != 0
+#   ) %>%
+#   mutate(contactSpecific = recode(.$contactSpecific, 
+#                          `-1`="No Interaction Only",
+#                          `0`="unspecific",
+#                          `1`="Interaction Only")) %>%
+#   select(varNam, varGroup, contactSpecific) %>%
+#   group_by(varGroup, varNam, contactSpecific) %>%
+#   summarise(n = n()) %>%
+#   pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
+#   ungroup %>%
+#   t %>%
+#   as.data.frame %>%
+#   janitor::row_to_names(1) %>%
+#   replace(is.na(.), 0) %>%
+#   tibble::rownames_to_column(., var = "variable") %>%
+#   separate(variable, c("concept", "variable"), " - ") %>%
+#   arrange(concept, unspecific, `Interaction Only`) %>%
+#   mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+#   kbl(.,
+#       escape = FALSE,
+#       booktabs = T,
+#       align = c("l", "l", rep("c", ncol(.)-2)),
+#       digits=2,
+#       caption = "All Possible Variables by Interaction Types") %>%
+#   kable_classic(
+#     full_width = F,
+#     lightable_options = "hover",
+#     html_font = "Cambria"
+#   ) %>%
+#   scroll_box(width = "100%", height = "500px")
+
+# Non-Interaction Specific Variables
+varNamS123NoInt <- varNames %>%
   filter(
-    aux == 1,
-    format == "nominal",
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
+    varNam %in% varNamS123PCA,
+    contactSpecific == 0
   ) %>%
@@
 
-#################################################################
-##                    Dataframe Preparation                    ##
-#################################################################
-dtMiMain <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  rowwise() %>%
-  mutate(
-    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
-    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum)
-  ) %>%
-  ungroup() %>%
-  select(ID,
-         study,
-         date,
-         week,
-         everything(),
-         -PID,
-         -starts_with("ResponseId"), 
-         -starts_with("roommate"),
-         -starts_with("Reason")) %>%
-  as.data.frame %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(ID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  mutate(study = replace_na(study, calc_mode(study))) %>%
-  ungroup %>%
-  group_by(study, TIDnum) %>%
-  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>%
-  as.data.frame 
-```
-
-```{r MainImp}
-#################################################################
-##                     Imputation Bi-daily                     ##
-#################################################################
-# set.seed(123)
-# MiMain <-
-#   amelia(
-#     dtMiMain,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMain, file = "data/imputed/MainImputed.RData")
-load("data/imputed/MainImputed.RData")
-
-# # Stability to imputations:
-# set.seed(123)
-# MiMainStability <-
-#   amelia(
-#     dtMiMain,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c(
-#       "TID",
-#       "date",
-#       "week",
-#       "study",
-#       "IntergroupContact",
-#       "IngroupContact"
-#     ),
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainStability, file = "data/imputed/MainImputedStability.RData")
-load("data/imputed/MainImputedStability.RData")
-
-##---------------------------------------------------------------
-##                Restore Missingness by Design                --
-##---------------------------------------------------------------
-
-# SHOULD STILL BE DONE? 
-# (remove imputed values that were missing by design. E.g., partner attitudes when there were no interactions)
-
-```
-
-### Follow-up: Studies
-
-#### Study 1
-
-```{r MiS1Prep}
-#################################################################
-##                    Variable Name Vectors                    ##
-#################################################################
-varNamS1MiPsbl <- c(
-  paste(varNamS1MI, rep(".pre", length(varNamS1MI)), sep = ""),
-  varNamS1MI,
-  paste(varNamS1MI, rep(".post", length(varNamS1MI)), sep = "")
-)
-
-varNamdtS1Noms <- varNames %>%
+# Interaction Specific Variables
+varNamS123Int <- varNames %>%
   filter(
-    aux == 1,
-    format == "nominal",
-    studyS1 == "S1"
+    varNam %in% varNamS123PCA,
+    contactSpecific == 1
   ) %>%
@@
 
-#################################################################
-##                    Dataframe Preparation                    ##
-#################################################################
-dtMiS1 <- dtS1Red %>% 
-  select(TID, any_of(varNamS1MiPsbl)) %>% 
-  mutate(across(!TID, as.numeric)) %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  select(PID,
-         date,
-         week,
-         everything(),
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>% 
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(PID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  ungroup %>% 
-  group_by(TIDnum) %>%
-  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
-  as.data.frame
-
-allDuplicated <- function(vec){
-  front <- duplicated(vec)
-  back <- duplicated(vec, fromLast = TRUE)
-  all_dup <- front + back > 0
-  return(all_dup)
-}
-
-varNamDifS1 <- names(dtMiS1)[gsub(".pre|.post|_calc", "", names(dtMiS1)) %>% allDuplicated]
-
-# Variables removed for study 1 due to high multicollinearity:
-varNamS1Out <- dtMiS1 %>%
-  select(
-    any_of(varNamDifS1),
-    mdmq.calmness.pre_calc,
-    mdmq.alertness.pre_calc,
-    mdmq.valence.pre_calc,
-    relatednessFrust.pre_calc,
-    competenceFrust.pre_calc,
-    competenceSat.pre_calc,
-    relatednessSat.pre_calc,
-    DutchFriends.pre,
-    secNatDum.pre,
-    timeNL.pre,
-    stress.post,
-    discrimination_month.post,
-    rosenberg.post,
-    qualityStar,
-    starts_with("SSAS")
-  ) %>%
-  names
-
-dtMiS1 <- dtMiS1 %>%
-  # mutate(
-  #   assimilation.dif = assimilation.post - assimilation.pre,
-  #   integration.dif = integration.post - integration.pre,
-  #   marginalization.dif = marginalization.post - marginalization.pre,
-  #   separation.dif = separation.post - separation.pre,
-  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-  #   swl.dif = swl.post_calc - swl.pre_calc
-  # ) %>% 
-  select(
-    -any_of(varNamS1Out)
-  ) %>%
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         which(sapply(., sd, na.rm = TRUE) > 0)) %>% 
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame
-
-# lares::corr_cross(
-#   dtMiS1 %>% select_if(is.numeric), # name of dataset
-#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
-#   top = 20 # display top 20 couples of variables (by correlation coefficient)
-# )
-```
-
-```{r MiS1}
-#################################################################
-##                     Imputation Bi-daily                     ##
-#################################################################
-# set.seed(123)
-# MiS1 <-
-#   amelia(
-#     dtMiS1,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiS1, file = "data/imputed/S1Imputed.RData")
-load("data/imputed/S1Imputed.RData")
-
-# # Imputation Stability:
-# set.seed(123)
-# MiS1Stability <-
-#   amelia(
-#     dtMiS1,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     # parallel = "multicore",
-#     # ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS1Stability, file = "data/imputed/S1ImputedStability.RData")
-load("data/imputed/S1ImputedStability.RData")
-```
-
-#### Study 2
-
-```{r MiS2Prep}
-#################################################################
-##                    Variable Name Vectors                    ##
-#################################################################
-varNamS2MiPsbl <- c(
-  paste(varNamS2MI, rep(".pre", length(varNamS2MI)), sep = ""),
-  varNamS2MI,
-  paste(varNamS2MI, rep(".post", length(varNamS2MI)), sep = "")
-)
-
-varNamdtS2Noms <- varNames %>%
+# S1 variables
+varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
+
+# S1 Non-Interaction Specific Variables
+varNamS1NoInt <- varNames %>%
   filter(
-    aux == 1,
-    format == "nominal",
-    studyS2 == "S2"
+    varNam %in% varNamS1Clust,
+    contactSpecific == 0
   ) %>%
@@
 
-#################################################################
-##                    Dataframe Preparation                    ##
-#################################################################
-dtMiS2 <- 
-  dtS2Red %>% 
-  select(TID, any_of(varNamS2MiPsbl)) %>% 
-  mutate(across(!TID, as.numeric)) %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  select(PID,
-         date,
-         week,
-         everything(),
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>% 
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(PID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  ungroup %>% 
-  group_by(TIDnum) %>%
-  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
-  as.data.frame %>%
-  mutate(date = as.character(date)) %>%
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         everything()) %>%
-  select_if(function(x) is.character(x) | is.factor(x) | (is.numeric(x) && sd(x, na.rm = TRUE) > 0)) %>%
-  as.data.frame
-
-varNamDifS2 <- names(dtMiS2)[gsub(".pre|.post|_calc", "", names(dtMiS2)) %>% allDuplicated]
-
-# # Variables removed for study 2 due to high multicollinearity:
-varNamS2Out <- dtMiS2 %>%
-  select(
-    any_of(varNamDifS2),
-    ContactDesire,
-    DesireIntergroup,
-    DesireIngroup,
-    Reason.pre,
-    # mdmq.calmness.pre_calc,
-    # mdmq.alertness.pre_calc,
-    # mdmq.valence.pre_calc,
-    # relatednessFrust.pre_calc,
-    # competenceFrust.pre_calc,
-    # competenceSat.pre_calc,
-    # relatednessSat.pre_calc,
-    # DutchFriends.pre,
-    # secNatDum.pre,
-    timeNL.pre,
-    starts_with("HS"),
-    starts_with("Home")
-    # stress.post,
-    # discrimination_month.post,
-    # rosenberg.post,
-    # qualityStar,
-    # starts_with("SSAS")
-  ) %>%
-  names
-
-dtMiS2Fin <- dtMiS2 %>%
-  # mutate(
-  #   assimilation.dif = assimilation.post - assimilation.pre,
-  #   integration.dif = integration.post - integration.pre,
-  #   marginalization.dif = marginalization.post - marginalization.pre,
-  #   separation.dif = separation.post - separation.pre,
-  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-  #   swl.dif = swl.post_calc - swl.pre_calc
-  # ) %>%
-  select(
-    -any_of(varNamS2Out)
-  ) %>%
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         everything()) %>%
-         #which(sapply(., sd, na.rm = TRUE) > 0)) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame
-
-# lares::corr_cross(
-#   dtMiS2Fin %>% select_if(is.numeric), # name of dataset
-#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
-#   top = 20 # display top 20 couples of variables (by correlation coefficient)
-# )
-```
-
-```{r s2StlTest}
-# Seasonal Decomposition of Time Series by Loess
-
-# dtMiS2 %>% 
-#   select(
-#     PID,
-#     TIDnum,
-#     AttitudesDutch
-#   ) %>%
-#   #filter(PID == 1) %>%
-#   #na.exclude(as.ts(.)) %>%
-#   as.ts %>%
-#   stats::stl(., na.action=na.omit)
-# 
-# testLoess <- loess(AttitudesDutch ~ TIDnum, data = dtMiS2 %>% filter(PID == 1))
-# testSpline <- pspline::sm.spline(dtMiS2$AttitudesDutch, dtMiS2$TIDnum)
-# testMarkov <- semiMarkov()
-# 
-# stats::stl(dtMiS2Fin)
-
-```
-
-```{r MiS2}
-#################################################################
-##                     Imputation Bi-daily                     ##
-#################################################################
-# set.seed(123)
-# MiS2 <-
-#   amelia(
-#     dtMiS2Fin,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS2, file = "data/imputed/S2Imputed.RData")
-#load("data/imputed/S2Imputed.RData")
-
-# # Imputation Stability:
-# set.seed(123)
-# MiS2Stability <-
-#   amelia(
-#     dtMiS2,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiS2Stability, file = "data/imputed/S2ImputedStability.RData")
-#load("data/imputed/S2ImputedStability.RData")
-```
-
-#### Study 3
-
-```{r MiS3Prep}
-#################################################################
-##                    Variable Name Vectors                    ##
-#################################################################
-varNamS3MiPsbl <- c(
-  paste(varNamS3MI, rep(".pre", length(varNamS3MI)), sep = ""),
-  varNamS3MI,
-  paste(varNamS3MI, rep(".post", length(varNamS3MI)), sep = "")
-)
-
-varNamdtS3Noms <- varNames %>%
+# S1 Interaction Specific Variables
+varNamS1Int <- varNames %>%
   filter(
-    aux == 1,
-    format == "nominal",
-    studyS3 == "S2"
+    varNam %in% varNamS1Clust,
+    contactSpecific == 1
   ) %>%
@@
 
-#################################################################
-##                    Dataframe Preparation                    ##
-#################################################################
-dtMiS3 <- 
-  dtS3Red %>% 
-  select(TID, any_of(varNamS3MiPsbl)) %>% 
-  mutate(across(!TID, as.numeric)) %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  select(PID,
-         date,
-         week,
-         everything(),
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>% 
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., PID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(PID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  ungroup %>% 
-  group_by(TIDnum) %>%
-  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>% 
-  as.data.frame %>%
-  mutate(date = as.character(date)) %>%
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         everything()) %>%
-  select_if(function(x) is.character(x) | is.factor(x) | (is.numeric(x) && sd(x, na.rm = TRUE) > 0)) %>%
-  as.data.frame
-
-varNamDifS3 <- names(dtMiS3)[gsub(".pre|.post|_calc", "", names(dtMiS3)) %>% allDuplicated]
-
-# # Variables removed for study 2 due to high multicollinearity:
-varNamS3Out <- dtMiS3 %>%
-  select(
-    any_of(varNamDifS3),
-    # ContactDesire,
-    # DesireIntergroup,
-    # DesireIngroup,
-    # Reason.pre,
-    # mdmq.calmness.pre_calc,
-    # mdmq.alertness.pre_calc,
-    # mdmq.valence.pre_calc,
-    # relatednessFrust.pre_calc,
-    # competenceFrust.pre_calc,
-    # competenceSat.pre_calc,
-    # relatednessSat.pre_calc,
-    # DutchFriends.pre,
-    # secNatDum.pre,
-    # timeNL.pre,
-    # starts_with("HS"),
-    # starts_with("Home")
-    # stress.post,
-    # discrimination_month.post,
-    # rosenberg.post,
-    # qualityStar,
-    # starts_with("SSAS")
-  ) %>%
-  names
-
-dtMiS3Fin <- dtMiS3 %>%
-  # mutate(
-  #   assimilation.dif = assimilation.post - assimilation.pre,
-  #   integration.dif = integration.post - integration.pre,
-  #   marginalization.dif = marginalization.post - marginalization.pre,
-  #   separation.dif = separation.post - separation.pre,
-  #   VIA_Dutch.dif = VIA_Dutch.post - VIA_Dutch.post,
-  #   VIA_heritage.dif = VIA_heritage.post - VIA_heritage.pre,
-  #   IntGrAnx.dif = IntGrAnx.post_calc - IntGrAnx.pre_calc,
-  #   swl.dif = swl.post_calc - swl.pre_calc
-  # ) %>%
-  select(
-    -any_of(varNamS3Out)
-  ) %>%
-  select(PID,
-         TID,
-         TIDnum,
-         date,
-         week,
-         everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame
-
-lares::corr_cross(
-  dtMiS3Fin %>% select_if(is.numeric), # name of dataset
-  max_pvalue = 0.05, # display only significant correlations (at 5% level)
-  top = 20 # display top 20 couples of variables (by correlation coefficient)
-)
-```
-
-```{r MiS3}
-#################################################################
-##                     Imputation Bi-daily                     ##
-#################################################################
-# set.seed(123)
-# MiS3 <-
-#   amelia(
-#     dtMiS3Fin,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiS3, file = "data/imputed/S3Imputed.RData")
-#load("data/imputed/S2Imputed.RData")
-
-# Imputation Stability:
-# set.seed(123)
-# MiS3Stability <-
-#   amelia(
-#     dtMiS3Fin,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "PID",
-#     idvars = c("TID", "date", "week", "IntergroupContact", "IngroupContact"),
-#     # lags = names(dtMiS1)[!names(dtMiS1) %in% idVars],
-#     # lags = names(dtMiS1)[names(dtMiS1) %in% varNamS1MiPsbl], # alternative?
-#     # leads = ...,
-#     # noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiS3Stability, file = "data/imputed/S3ImputedStability.RData")
-#load("data/imputed/S2ImputedStability.RData")
-```
-
-### Follow-up: Time Scales
-
-#### Daily
-
-Preparing the data for the time scale follow-up, which combines the common variables of all three studies on a daily level. Here, we:
-
-1.  Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
-2.  Create the combined dataset in the correct format (long format with all missed time points as NAs).
-3.  Impute the data and save the imputed datasets for later recall.
-
-```{r MiDailyPrep}
-#################################################################
-##                    Dataframe Preparation                    ##
-#################################################################
-dtMiMainDaily <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  rowwise() %>%
-  mutate(
-    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
-    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum)
-  ) %>%
-  ungroup() %>%
-  group_by(ID, date, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>% # contact dum should be summarized as sum()
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(date))) %>%
-  ungroup %>%
-  select(ID,
-         study,
-         date,
-         everything(),
-         -PID,
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(ID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  mutate(study = replace_na(study, calc_mode(study))) %>%
-  ungroup %>%
-  group_by(study, TIDnum) %>%
-  mutate(across(c("date"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate_if(is.numeric, ~ ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame 
-```
-
-```{r DailyImp}
-#################################################################
-##               Imputation at Daily Aggregation               ##
-#################################################################
-# MiMainDaily <-
-#   amelia(
-#     dtMiMainDaily,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainDaily, file = "data/imputed/MainDailyImputed.RData")
-load("data/imputed/MainDailyImputed.RData")
-
-# # Stability to imputations:
-# set.seed(123)
-# MiMainDailyStability <-
-#   amelia(
-#     dtMiMainDaily,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("date", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMainDaily)[!names(dtMiMainDaily) %in% idVars],
-#     #lags = names(dtMiMainDaily)[names(dtMiMainDaily) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainDaily)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainDaily))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainDailyStability, file = "data/imputed/MainDailyImputedStability.RData")
-load("data/imputed/MainDailyImputedStability.RData")
-```
-
-#### Weekly
-
-Preparing the data for the time scale follow-up, which combines the common variables of all three studies on a weekly level. Here, we:
-
-1.  Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
-2.  Create the combined dataset in the correct format (long format with all missed time points as NAs).
-3.  Impute the data and save the imputed datasets for later recall.
-
-```{r MiWeeklyPrep}
-dtMiMainWeeekly <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>% 
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-  ) %>%
-  rowwise() %>%
-  mutate(
-    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
-    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum)
-  ) %>%
-  ungroup() %>%
-  group_by(ID, week, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(week))) %>%
-  ungroup %>%
-  select(ID,
-         study,
-         week,
-         everything(),
-         -PID,
-         -starts_with("ResponseId")) %>%
-  as.data.frame %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
-  group_by(ID) %>%
-  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
-  mutate(study = replace_na(study, calc_mode(study))) %>%
-  ungroup %>%
-  group_by(study, TIDnum) %>%
-  mutate(across(c("week"),  ~ replace_na(.x, calc_mode(.x)))) %>%
-  ungroup %>% 
-  mutate_if(is.numeric, ~ ifelse(is.nan(.), NA, .)) %>%
-  as.data.frame 
-```
-
-```{r WeeklyImp}
-##################################################################
-##               Imputation at Weekly Aggregation               ##
-##################################################################
-# MiMainWeekly <-
-#   amelia(
-#     dtMiMainWeeekly,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
-#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainWeekly, file = "data/imputed/MainWeeklyImputed.RData")
-load("data/imputed/MainWeeklyImputed.RData")
-
-# # Stability to imputations:
-# set.seed(123)
-# MiMainWeeklyStability <-
-#   amelia(
-#     dtMiMainWeeekly,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMainWeeekly)[!names(dtMiMainWeeekly) %in% idVars],
-#     #lags = names(dtMiMainWeeekly)[names(dtMiMainWeeekly) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMainWeeekly)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMainWeeekly))],
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 1
-#   )
-# save(MiMainWeeklyStability, file = "data/imputed/MainWeeklyImputedStability.RData")
-load("data/imputed/MainWeeklyImputedStability.RData")
-```
-
-### Follow-up: Response Type
-
-#### Interaction Responses
-
-```{r MiInteractionPrep}
-#################################################################
-##                    Dataframe Preparation                    ##
-#################################################################
-
-# any interaction
-dtMiInteraction <- 
-  dtMiMain %>%
-    mutate_at(vars(any_of(varNamS123PCA), -any_of(idVars)), funs(ifelse(dtMiMain$InteractionDum != 1, NA, .))) %>%
-    #mutate_at(vars(-any_of(idVars), -ends_with(".pre"), -ends_with(".post"), -InteractionDum), funs(ifelse(dtMiMain$InteractionDum != 1, NA, .))) %>%
-    as.data.frame
-
-# Intergroup Contact
-dtMiIntergroup <- 
-  dtMiMain %>%
-  mutate_at(vars(any_of(varNamS123PCA), -any_of(idVars)), funs(ifelse(dtMiMain$IntergroupContact != 1, NA, .))) %>%
-  #mutate_at(vars(-any_of(idVars), -ends_with(".pre"), -ends_with(".post"), -IntergroupContact), funs(ifelse(dtMiMain$IntergroupContact != 1, NA, .))) %>%
-  #filter(is.na(relatednessNoInteraction)) %>%
-  select(-relatednessNoInteraction) %>% # remove completely missing col
-  as.data.frame
-
-# dtMiIntergroup %>% 
-#   select(ID, TIDnum, IntergroupContact, relatednessNoInteraction) %>%
-#   arrange(ID, TIDnum, IntergroupContact)
-
-# Ingroup Contact
-dtMiIngroup <- 
-  dtMiMain %>%
-    #mutate_at(vars(any_of(varNamS123PCA), -any_of(idVars)), funs(ifelse(dtMiMain$IngroupContact != 1, NA, .))) %>%
-    mutate_at(vars(-any_of(idVars), -ends_with(".pre"), -ends_with(".post"), -IngroupContact), funs(ifelse(dtMiMain$IngroupContact != 1, NA, .))) %>%
-    as.data.frame
-```
-
-```{r InteractionImp}
-#################################################################
-##                       Any Interaction                       ##
-#################################################################
-
-##---------------------------------------------------------------
-##                      Single Imputation                      --
-##---------------------------------------------------------------
-# set.seed(123)
-# MiInteraction <-
-#   amelia(
-#     dtMiInteraction,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     #noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiInteraction, file = "data/imputed/InteractionImputed.RData")
-load("data/imputed/InteractionImputed.RData")
-
-##---------------------------------------------------------------
-##                     Multiple Imputation                     --
-##---------------------------------------------------------------
-
-# # Stability to imputations:
-# set.seed(123)
-# MiInteractionStab <-
-#   amelia(
-#     dtMiInteraction,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     #noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiInteractionStab, file = "data/imputed/InteractionImputedStability.RData")
-load("data/imputed/InteractionImputedStability.RData")
-```
-
-```{r intergroupImp}
-##################################################################
-##                    Intergroup Interaction                    ##
-##################################################################
-
-##---------------------------------------------------------------
-##                      Single Imputation                      --
-##---------------------------------------------------------------
-# set.seed(123)
-# MiIntergroup <-
-#   amelia(
-#     dtMiIntergroup,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     #noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiIntergroup, file = "data/imputed/IntergroupImputed.RData")
-load("data/imputed/IntergroupImputed.RData")
-
-##---------------------------------------------------------------
-##                     Multiple Imputation                     --
-##---------------------------------------------------------------
-
-# # Stability to imputations:
-# set.seed(123)
-# MiIntergroupStab <-
-#   amelia(
-#     dtMiIntergroup,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     #noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiIntergroupStab, file = "data/imputed/IntergroupImputedStability.RData")
-load("data/imputed/IntergroupImputedStability.RData")
-```
-
-```{r ingroupImp}
-#################################################################
-##                     Ingroup Interaction                     ##
-##                        NOT CONVERGING                       ##
-#################################################################
-
-##---------------------------------------------------------------
-##                      Single Imputation                      --
-##---------------------------------------------------------------
-# set.seed(123)
-# MiIngroup <-
-#   amelia(
-#     dtMiIngroup,
-#     m = 1, # 20
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     #noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiIngroup, file = "data/imputed/IngroupImputed.RData")
-# load("data/imputed/IngroupImputed.RData")
-
-##---------------------------------------------------------------
-##                     Multiple Imputation                     --
-##---------------------------------------------------------------
-
-# # Stability to imputations:
-# set.seed(123)
-# MiIngroupStab <-
-#   amelia(
-#     dtMiIngroup,
-#     m = 20,
-#     ts = "TIDnum",
-#     cs = "ID",
-#     idvars = c("TID", "date", "week", "study", "IntergroupContact", "IngroupContact", "InteractionDum"), #"PID",
-#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
-#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
-#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
-#     #noms = c("InteractionDum"),
-#     parallel = "multicore",
-#     ncpus = parallel::detectCores(),
-#     p2s = 2
-#   )
-# save(MiIngroupStab, file = "data/imputed/IngroupImputedStability.RData")
-# load("data/imputed/IngroupImputedStability.RData")
-```
-
-#### Non-Interaction Responses
-
-```{r MiNoInteractionPrep}
-# TBD
-```
-
-```{r NoInteractionImp}
-# TBD
-```
-
-## Center and Normalize
-
-For the 3MPCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables.
-The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering).
-The normalization within variable are important for ensuring equal variances across variables, which ensures equal weighting of the variables in the 3MPCA.
-
-```{r mainCentered}
-varNam3mpcaMainAnalysis <- varNamesDailyFull %>%
-  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum", "startDate", "startTime")) %>%
+# S2 variables
+varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
+
+# S2 Non-Interaction Specific Variables
+varNamS2NoInt <- varNames %>%
+  filter(
+    varNam %in% varNamS2Clust,
+    contactSpecific == 0
+  ) %>%
+  select(varNam) %>%
   pull
 
-dtMainMi <- MiMain$imputations$imp1 %>%
-  rowwise() %>%
-  mutate(
-    InteractionDumOg = sum(IntergroupContact, IngroupContact, na.rm = TRUE),
-    InteractionDumOg = if_else(InteractionDum > 0, 1, InteractionDum),
-    InteractionDumOg = if_else(sum(is.na(c(IntergroupContact, IngroupContact))) == 2, NA_real_, InteractionDum),
-    relatednessInteraction = mean(c(relatednessSelf, relatednessOther), na.rm = TRUE),
-    relatedness = if_else(InteractionDum == 1, relatednessInteraction, relatednessNoInteraction)
-    #relatednessNoInteraction = if_else(is.na(InteractionDumOg), NA_real_, relatednessNoInteraction)
-  ) %>%
-  ungroup()
-
-dtMain3mpcaNorm <-
-  PCAnorm(data = MiMain$imputations$imp1,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = varNam3mpcaMainAnalysis)
-```
-
-## Long to Wide
-
-For the Three-way ANOVA and the Three-mode PCA, the datasets need to be in a $n * mp$ matrix format, where $n$ is are the participants, $m$ are the time points, and $p$ are the variables.
-We do this by taking the prepared datasets, melting them into a id-variable-value table and casting the table into a wide format of variables_timepoint columns.
-
-### Main Analysis
-
-```{r longToWideMain}
-dtPcaS123 <- rbind(
-  dtS1Red %>% select(any_of(varNamS123PCA)) %>% mutate(study = "S1"),
-  dtS2Red %>% select(any_of(varNamS123PCA)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123PCA)) %>% mutate(study = "S3")
-  ) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-    ) %>%
-  select(
-    ID,
-    PID,
-    TID,
-    date,
-    week,
-    TIDnum,
-    study,
-    everything()
-  ) %>%
-  arrange(ID, TIDnum) %>%
-  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
-# 
-# dtMainWideC <- PCAnorm(data = dtPcaS123,
-#           pid = "ID",
-#           tid = "TIDnum",
-#           selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>%
-#   select(
-#     ID,
-#     PID,
-#     TIDnum,
-#     study,
-#     ends_with("_gmc")
-#   ) %>%
-#   data.frame %>% 
-#   melt(
-#     .,
-#     id=c("ID", "PID", "TIDnum", "study")
-#   ) %>%
-#   mutate(
-#     variable = paste(gsub("_gmc", "", variable), stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
-#   ) %>%
-#   arrange(variable, ID) %>%
-#   #arrange(ID, variable, TIDnum) %>%
-#   select(-TIDnum) %>%
-#   pivot_wider(names_from = variable, values_from = value) %>%
-#   replace(is.na(.), 0) %>%
-#   select(-ID, -PID, -study)
-#   #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
-#   #pivot_wider(names_from = variable, values_from = value)
-# 
-# # dtMainWideC %>%
-# #   filter(
-# #     ID == 54
-# #   ) %>% 
-# #   t
-# # 
-# # dtMainWideC$AttitudesDutch_59[dtMainWideC$ID == 54]
-# 
-# dtMainWide <- dtPcaS123 %>%
-#   select(
-#     -TID,
-#     -date,
-#     -week
-#   ) %>%
-#   data.frame %>% 
-#   melt(
-#     .,
-#     id=c("ID", "PID", "TIDnum", "study")
-#   ) %>%
-#   mutate(
-#     variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
-#   ) %>%
-#   select(-TIDnum) %>%
-#   arrange(variable, ID) %>%
-#   pivot_wider(names_from = variable, values_from = value) %>%
-#   replace(is.na(.), 0) %>%
-#   select(-ID, -PID, -study)
-
-dtPcaS123Imp <- MiMain$imputations[[1]] %>% 
-  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))
-
-dtMainImpWide <- MiMain$imputations[[1]] %>% 
-  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  melt(
-    .,
-    id=c("ID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum, -study) %>%
-  arrange(ID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-ID)
-```
-
-```{r longToWideMainStability}
-# Long to Wide Main Bi-daily
-MiMainStabilityWide <- list()
-for (i in 1:length(MiMainStability$imputations)) {
-  MiMainStabilityWide[[i]] <-
-    MiMainStability$imputations[[i]]  %>%
-    # Variable selection
-    select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-    # Sorting
-    arrange(ID, TIDnum) %>%
-    # Filter minimum common num of measurements
-    filter(TIDnum <= min(
-      max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)
-    )) %>%
-    # Rm unnecessary variables for restructuring
-    select(-TID,
-           -date,
-           -week) %>%
-    # Long to wide
-    melt(.,
-         id = c("ID", "TIDnum", "study")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum, -study) %>%
-    arrange(ID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-ID)
-}
-```
-
-### Follow-up: Studies
-
-#### Study 1
-
-##### Common Variables
-
-```{r longToWideS1Common}
-dtS1ImpWide <- MiS1$imputations[[1]] %>% 
-  select(PID, date, week, any_of(varNamS123PCA)) %>%
-  arrange(PID, TIDnum) %>%
-  #filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  melt(
-    .,
-    id=c("PID", "TIDnum")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum) %>%
-  arrange(PID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-PID)
-```
-
-```{r longToWideS1StabilityCommon}
-# Long to Wide S1 Bi-daily
-MiS1StabilityWide <- list()
-for (i in 1:length(MiS1Stability$imputations)) {
-  MiS1StabilityWide[[i]] <-
-    MiS1Stability$imputations[[i]]  %>%
-    # Variable selection
-    select(PID, date, week, any_of(varNamS123PCA)) %>%
-    # Sorting
-    arrange(PID, TIDnum) %>%
-    # Rm unnecessary variables for restructuring
-    select(-TID,
-           -date,
-           -week) %>%
-    # Long to wide
-    melt(.,
-         id = c("PID", "TIDnum")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum) %>%
-    arrange(PID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-PID)
-}
-```
-
-##### Idiosyncratic Variables
-
-```{r longToWideS1Idiosyncratic}
-dtS1ImpWideIdio <- MiS1$imputations[[1]] %>% 
-  select(PID, date, week, any_of(varNamS1PCA)) %>%
-  arrange(PID, TIDnum) %>%
-  #filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  melt(
-    .,
-    id=c("PID", "TIDnum")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum) %>%
-  arrange(PID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-PID)
-```
-
-```{r longToWideS1StabilityIdiosyncratic}
-# Long to Wide S1 Bi-daily
-MiS1StabilityWideIdio <- list()
-for (i in 1:length(MiS1Stability$imputations)) {
-  MiS1StabilityWideIdio[[i]] <-
-    MiS1Stability$imputations[[i]]  %>%
-    # Variable selection
-    select(PID, date, week, any_of(varNamS1PCA)) %>%
-    # Sorting
-    arrange(PID, TIDnum) %>%
-    # Rm unnecessary variables for restructuring
-    select(-TID,
-           -date,
-           -week) %>%
-    # Long to wide
-    melt(.,
-         id = c("PID", "TIDnum")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum) %>%
-    arrange(PID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-PID)
-}
-```
-
-#### Study 2
-
-#### Study 3
-
-### Follow-up: Response Type
-
-#### All Interaction Responses
-
-```{r longToWideInteraction}
-
-dtInteractionImpWide <- MiInteraction$imputations[[1]] %>% 
-  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  melt(
-    .,
-    id=c("ID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum, -study) %>%
-  arrange(ID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-ID)
-```
-
-```{r longToWidInteractionStability}
-# Long to Wide Main Bi-daily
-MiInteractionStabilityWide <- list()
-for (i in 1:length(MiInteractionStab$imputations)) {
-  MiInteractionStabilityWide[[i]] <-
-    MiInteractionStab$imputations[[i]]  %>%
-    # Variable selection
-    select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-    # Sorting
-    arrange(ID, TIDnum) %>%
-    # Filter minimum common num of measurements
-    filter(TIDnum <= min(
-      max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)
-    )) %>%
-    # Rm unnecessary variables for restructuring
-    select(-TID,
-           -date,
-           -week) %>%
-    # Long to wide
-    melt(.,
-         id = c("ID", "TIDnum", "study")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum, -study) %>%
-    arrange(ID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-ID)
-}
-```
-
-#### Intergroup Interaction Responses
-
-```{r longToWideIntergroup}
-dtIntergroupImpWide <- MiIntergroup$imputations[[1]] %>% 
-  select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum))) %>%
-  select(
-    -TID,
-    -date,
-    -week
-  ) %>%
-  melt(
-    .,
-    id=c("ID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum, -study) %>%
-  arrange(ID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-ID)
-```
-
-```{r longToWidIntergroupStability}
-# Long to Wide Main Bi-daily
-MiIntergroupStabilityWide <- list()
-for (i in 1:length(MiIntergroupStab$imputations)) {
-  MiIntergroupStabilityWide[[i]] <-
-    MiIntergroupStab$imputations[[i]]  %>%
-    # Variable selection
-    select(ID, date, week, study, any_of(varNamS123PCA)) %>%
-    # Sorting
-    arrange(ID, TIDnum) %>%
-    # Filter minimum common num of measurements
-    filter(TIDnum <= min(
-      max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)
-    )) %>%
-    # Rm unnecessary variables for restructuring
-    select(-TID,
-           -date,
-           -week) %>%
-    # Long to wide
-    melt(.,
-         id = c("ID", "TIDnum", "study")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum, -study) %>%
-    arrange(ID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-ID)
-}
-```
-
-#### Non-Interaction Responses
-
-### Follow-up: Time Scales
-
-#### Daily
-
-```{r longToWideDaily}
-dtMainDailyImpWide <- MiMainDaily$imputations[[1]] %>%
-  select(ID, date, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
+# S2 Interaction Specific Variables
+varNamS2Int <- varNames %>%
   filter(
-    TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
-  ) %>%
-  select(
-    -date
-  ) %>% 
-  melt(
-    .,
-    id=c("ID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum, -study) %>%
-  arrange(ID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-ID)
-```
-
-```{r longToWideDailyStability}
-# Long to Wide Main Daily
-MiMainDailyStabilityWide <- list()
-for (i in 1:length(MiMainDailyStability$imputations)) {
-  MiMainDailyStabilityWide[[i]] <-
-    MiMainDailyStability$imputations[[i]] %>%
-    # Variable selection
-    select(ID, date, study, any_of(varNamS123PCA)) %>%
-    # Sorting
-    arrange(ID, TIDnum) %>%
-    # Filter minimum common num of measurements
+    varNam %in% varNamS2Clust,
+    contactSpecific == 1
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S3 variables
+varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
+
+# S3 Non-Interaction Specific Variables
+varNamS3NoInt <- varNames %>%
   filter(
-      TIDnum <= dtMiMainDaily %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
-    ) %>%
-    # Rm unnecessary variables for restructuring
-    select(-date) %>%
-    # Long to wide
-    melt(.,
-         id = c("ID", "TIDnum", "study")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum, -study) %>%
-    arrange(ID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-ID)
-}
-```
-
-#### Weekly
-
-```{r longToWideWeekly}
-dtMainWeeklyImpWide <- MiMainWeekly$imputations[[1]] %>%
-  select(ID, week, study, any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
+    varNam %in% varNamS3Clust,
+    contactSpecific == 0
+  ) %>%
+  select(varNam) %>%
+  pull
+
+# S3 Interaction Specific Variables
+varNamS3Int <- varNames %>%
   filter(
-    TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
-  ) %>%
-  select(
-    -week
-  ) %>% 
-  melt(
-    .,
-    id=c("ID", "TIDnum", "study")
-  ) %>%
-  mutate(
-    variable = paste("T", stringr::str_pad(TIDnum, width = 2, pad = 0), "_", variable, sep = "")
-  ) %>%
-  select(-TIDnum, -study) %>%
-  arrange(ID, variable) %>%
-  pivot_wider(names_from = variable, values_from = value) %>%
-  #replace(is.na(.), 0) %>%
-  select(-ID)
-```
-
-```{r longToWideWeeklyStability}
-# Long to Wide Main Weekly
-MiMainWeeklyStabilityWide <- list()
-for (i in 1:length(MiMainWeeklyStability$imputations)) {
-  MiMainWeeklyStabilityWide[[i]] <-
-    MiMainWeeklyStability$imputations[[i]] %>%
-    # Variable selection
-    select(ID, week, study, any_of(varNamS123PCA)) %>%
-    # Sorting
-    arrange(ID, TIDnum) %>%
-    # Filter minimum common num of measurements
-    filter(
-      TIDnum <= dtMiMainWeeekly %>% group_by(study) %>% summarise(TIDmax = max(TIDnum)) %>% select(TIDmax) %>% min
-    ) %>%
-    # Rm unnecessary variables for restructuring
-    select(-week) %>%
-    # Long to wide
-    melt(.,
-         id = c("ID", "TIDnum", "study")) %>%
-    mutate(variable = paste(
-      "T",
-      stringr::str_pad(TIDnum, width = 2, pad = 0),
-      "_",
-      variable,
-      sep = ""
-    )) %>%
-    select(-TIDnum, -study) %>%
-    arrange(ID, variable) %>%
-    pivot_wider(names_from = variable, values_from = value) %>%
-    #replace(is.na(.), 0) %>%
-    select(-ID)
-}
-```
-
-# **Feature-based Time-Series Clustering**
-
-1. central tendency
-    + mean ✓
-    + median ✓
-2. variation around the central tendency (i.e., density)
-    + SD ✓
-    + MAD ✓
-3. (in)stability
-    + root mean square of successive differences (rmssd) ✓
-    + mean absolute change ✓
-4. inertia
-    + lag-1 autocorrelation ✓
-    + periodicity (lag-2 & lag-14 autocorrelation) ✓
-5. linear trend
-    + OLS regression parameter ✓
-6. non-linear trend
-    + GAM edf ✓
-
-(!discuss NA procedures!)
-
-## Variable Selection
-
-```{r variable names for all analyses}
-
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Variable Groups
-varNames %>%
-  filter(pca != 0) %>%
-  mutate(contactSpecific = recode(.$contactSpecific, 
-                         `-1`="No Interaction Only",
-                         `0`="unspecific",
-                         `1`="Interaction Only")) %>%
-  crosstable::crosstable(., varGroup, by=contactSpecific) %>%
-  select(-c(.id, label)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "r", "c"),
-      digits=2,
-      caption = "Full Analysis: Variable Groups by Interaction Types") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-
+    varNam %in% varNamS3Clust,
+    contactSpecific == 1
+  ) %>%
+  select(varNam) %>%
+  pull
+```
+
+We also see that a number of variables are specific to interactions.
+These variables will by their very nature have substantially larger missingness (because not every measurement occasion will follow new social interactions) and these variables are also un-imputable in traditional time-series modeling and hypothesis testing procedures.
+
+```{r variable names by response type}
 varNames %>%
   filter(
-    pca != 0
+    pca != 0,
+    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
   ) %>%
@@
       digits=2,
-      caption = "All Possible Variables by Interaction Types") %>%
+      caption = "Variables by Interaction Types") %>%
   kable_classic(
@@
   )
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% varNamS123PCA
-  ) %>%
-  mutate(contactSpecific = recode(.$contactSpecific, 
-                         `-1`="No Interaction Only",
-                         `0`="unspecific",
-                         `1`="Interaction Only")) %>%
-  select(varNam, varGroup, contactSpecific) %>%
-  group_by(varGroup, varNam, contactSpecific) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable"), " - ") %>%
-  arrange(concept, unspecific, `Interaction Only`) %>%
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Full Analysis: Variables by Interaction Types") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
+```
+
+Please note that not all variables are included in the main analysis, as some variables are only found within specific studies.
 
 ## Feature Extraction
+
+Time-series can be described with a plethora of features.
+A good example is the list of 100 time-series features that one can extract with [tsfresh python package](https://tsfresh.readthedocs.io/en/latest/api/tsfresh.feature_extraction.html# "list of features to extract").
+In practice there are usually two steps to the feature engineering process.
+(1) *Feature extraction*, that is summarizing the time series in meaningful descriptors, and (2) *feature selection*, the process of ensuring that only meaningful features are included in the clustering step.
+For the interpretability of our clustering procedure we have made a pre-selection of features, we wish to include to describe our time-series:
+
+1. central tendency
+    + mean ✓
+    + median ✓
+2. variation around the central tendency (i.e., density)
+    + SD ✓
+    + MAD ✓
+3. (in)stability
+    + root mean square of successive differences (rmssd) ✓
+    + mean absolute change (mac) ✓
+4. inertia
+    + lag-1 autocorrelation ✓
+    + periodicity (lag-2 & lag-14 autocorrelation) ✓
+5. linear trend
+    + OLS regression parameter ✓
+6. non-linear trend
+    + GAM edf ✓
+
+(!discuss NA procedures!)
 
@@
 load("data/features/featOut.RData")
-
 ```
@@
 ### Across Studies
+
+We begin by extracting the features for the main analysis that combines the sample of the three studies.
+We do this across all selected variables as well as for the content (non-)specific variables separately.
+During the feature extraction step there can at times be features that cannot be calculated for a small number of participants.
+An example would a participant who never filled in two consecutive measurements that describe an interaction.
+In that case, we cannot extract an auto correlation of lag-1 for some of the interaction specific variables.
 
@@
 
-pMissFeat(
+pMissFull <- pMissFeat(
   all = featFull$features,
@@
 
+pMissFull$pMissPlot
+
 # featFullImp <- featureImputer(featFull)
@@
 
+For the full dataset across the three studies, we find that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r nrow(featFull$features)` participants and `r ncol(featFull$features)` features.
+These missing features are imputed with a single predictive mean matching imputation using the MICE library.
+
 ### Study 1
+
+We then repeat the feature extraction procedure for all selected variables that are specific to Study 1.
 
@@
 
-pMissFeat(
+pMissS1 <- pMissFeat(
   all = featS1$features,
@@
 
+pMissS1$pMissPlot
+
 # featS1Imp <- featureImputer(featS1)
@@
 ### Study 2
+
+In the same manner, the features are extracted for the variables that are specific to Study 2.
 
@@
 
-pMissFeat(
+pMissS2 <- pMissFeat(
   all = featS2$features,
@@
 
+pMissS2$pMissPlot
+
 # featS2Imp <- featureImputer(featS2)
@@
 ### Study 3
+
+And finally, we extract the features for the Study 3 specific variables.
 
@@
 
-pMissFeat(
+pMissS3 <- pMissFeat(
   all = featS3$features,
@@
 )
+pMissS3$pMissPlot
 
@@
 
+## Feature Selection
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  Here we might have to make a selection of the features we extract because some of the clustering algorithms might not be able to deal with the many different features. See below
+</div>
+
 ## Clustering
 
-### Across Studies
+Cluster analyses have been a staple within a variety of sciences for the better part of a century.
+It's applications are wide and because most procedures fall under the "unsupervised learning" umbrella the methodological developments are fast-paced and new methods or refinements are published constantly.
+Among the many options available four types are found most commonly within the modern literature on clustering analyses:
+
+1. centroid-based
+    + e.g., k-means
+    + no assumptions (+)
+    + fast even with large matrices (+)
+    + non-probabilistic (+/-)
+    + sensitive to local minima (-)
+2. distribution-based
+    + e.g., gaussian mixture model (GMM)
+    + flexible shape (+)
+    + probabilistic (+/-)
+    + (gaussian) distributions assumed (-)
+3. density-based
+    + e.g., dbscan / OPTICS
+    + no shape assumption (+)
+    + does not assign outliers/noise (+)
+4. Hierarchical-based (not relevant)
+    + e.g., agglomerative hierarchical clustering
+    + best with hierarchical data (-)
+    + best with small *n* (-)
+
+Each clustering procedures comes with their own sets of advantages and disadvantages.
+We will focus mainly on the k-means procedure because it has the least assumptions and is fast even with large data sets.
+Additionally, the methods are well-established and the results are easy to interpret.
+As such, the procedure should be applicable to most users within the social sciences and for ESM feature data specifically.
+We offer Gaussian mixture modeling and a common density-based approach as supplementary analyses.
+
+### Main Anlaysis: Across Studies
+
+The main clustering procedure, uses the features calculated for the relevant variables across all three studies.
 
@@
 
-k-means
+For our main method, we use the centroid-based k-means clustering.
+The procedure needs the researcher to make a selection on the optimal number of clusters in order to run the main algorithm.
+We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution.
 
 ```{r kmeans determin k}
+# prepare data for main analysis
+raw_data <- featData
+z_data <- featFullImp$featuresImpZ
+scaled_data <- featFullImp$featuresImpZMat
+
+# load libraries
 library(factoextra)
 library(NbClust)
-
-scaled_data <- featFull2$featuresImpZMat
 
@@
   labs(subtitle = "Silhouette method")
-# Gap statistic
+# Gap statistic [takes for ever]
 # nboot = 50 to keep the function speedy. 
@@
 # Use verbose = FALSE to hide computing progression.
-set.seed(123)
-fviz_nbclust(scaled_data, kmeans, nstart = 50,  method = "gap_stat", nboot = 500)+
-  labs(subtitle = "Gap statistic method")
-
+# set.seed(123)
+# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
+#   labs(subtitle = "Gap statistic method")
+```
+
+We use this parameter to run the actual k-means cluster analysis, during which the algorithm iteratively seeks to optimally separate the participants into two clusters based on their feature values.
+During the iterative process the algorithm might at times stop at a non-optimal solution (i.e., local minima).
+To avoid this drawback and to find the (globally) optimal solution we let the algorithm run with 100 separate staring positions.
+
+```{r k-means modeling based on selected k}
 kmeansOut <- kmeans(scaled_data, centers = 2, nstart = 100)
-
+```
+
+Because the clustering happens in a multidimensional space of the `r ncol(featFull$features)` features, we begin by visualizing the data in a two-dimensional space, where we encircle the points of the clusters (i.e., convex hulls) for the first two dimensions of a principle component analysis.
+
+```{r Plot Convex Hulls}
 fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())
-
-as.data.frame(kmeansOut$cluster)
-
+```
+
+We then inspect the cluster-averages of all feature for each variable.
+We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., how cooperative an interaction was) than for other variables (e.g., Attitudes towards the Dutch).
+Additionally, we see that (2) some within the variables some features are better at distinguishing clusters (e.g., mean and median of well-being) than others (e.g., sd and rmssd of well-being).
+
+```{r features for each variable}
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., z_data) %>%
+  group_by(cluster) %>%
+  summarize_all(mean) %>%
+  ungroup %>%
+  select(-ID) %>% 
+  reshape2::melt(id.vars = "cluster") %>% 
+  mutate(lab = variable) %>%
+  mutate(
+    lab = variable,
+    variable = gsub("\\_.*", "", lab),
+    feature = gsub("^.*?\\_", "", lab)
+  ) %>% 
+  ggplot(., aes(x = feature, y = value, group = cluster, color = as.factor(cluster))) +
+  geom_point() +
+  geom_line() +
+  coord_flip() +
+  labs(
+    x = "Features",
+    y = "Standardized Mean",
+    color = "Cluster"
+  ) +
+  facet_wrap(~ variable) +
+  theme_Publication()
+```
+
+Next, we inspect the cluster-averages of all variables for each feature.
+This is the same data as in the plot above.
+However, here we see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median) than others (e.g., auto correlations).
+
+```{r variables for each feature}
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., z_data) %>%
+  group_by(cluster) %>%
+  summarize_all(mean) %>%
+  ungroup %>%
+  select(-ID) %>% 
+  reshape2::melt(id.vars = "cluster") %>% 
+  mutate(
+    lab = variable,
+    variable = gsub("\\_.*", "", lab),
+    feature = gsub("^.*?\\_", "", lab)
+  ) %>% 
+  ggplot(., aes(x = variable, y = value, group = cluster, color = as.factor(cluster))) +
+  geom_point() +
+  geom_line() +
+  coord_flip() +
+  labs(
+    x = "Features",
+    y = "Standardized Mean",
+    color = "Cluster"
+  ) +
+  facet_wrap(~ feature) +
+  theme_Publication()
+```
+
+In a final step we visualize the raw time series for each variable and add the average cluster development to see the features in action.
+
+```{r time-series by variable and cluster}
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., raw_data) %>%
+  select(
+    ID,
+    TIDnum,
+    cluster,
+    any_of(varNamS123)
+  ) %>%
+  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
+  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
+  stat_summary(fun=mean, geom="line") +
+  facet_wrap(~ variable, scales = "free_y") +
+  labs(
+    x = "Time ID",
+    y = "Response (Mean Cluster Response Bold)",
+    color = "Cluster"
+  ) +
+  theme_Publication()
 ```
@@
 
-Gaussian mixture model (GMM)
+For the distribution-based clustering we use the commonly used Gaussian mixture modeling (GMM) procedure.
+This procedure assumes  independent Gaussian distributions for each each cluster.
+These distributions are substantially better at identifying clusters if the data comes from multivariate Gaussian distributions (i.e., ellipses in a two dimensional space).
+
+We again begin by determining the optimal number of clusters. For GMMs this is done using common information criteria, such as the BIC.
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  This is where we run into trouble.
+These plots are relatively uncommon and according to the internet this is most likely a problem of too many (non-distinguishing or highly correlated) features.
+A good paper on this is <a href="https://proceedings.neurips.cc/paper/2002/file/e58aea67b01fa747687f038dfde066f6-Paper.pdf">Law et al.
+(2002)</a>.
+</div>
 
@@
 library(mclust)
+
+scaled_data <- featFullImp$featuresImpZMat
+
+# scaled_data <- featS1Imp$featuresImpZMat  %>% #select(starts_with(varNamS123Int))
+#   select(-any_of(names(which(colSums(is.na(.)) > 0))))
 
@@
 summary(BIC)
-
-mod1 <- Mclust(scaled_data, x = BIC)
-
-merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View
-
-summary(mod1, parameters = TRUE)
-table(class, mod1$classification)
-
-mod1dr <- MclustDR(mod1)
-summary(mod1dr)
-plot(mod1dr, what = "pairs")
-```
-
-#### Density-based
-
-dbscan
-
-```{r dbscan}
-library(dbscan)
-library(fpc)
-
-# to plot the eps values
-eps_plot = kNNdistplot(scaled_data, k=3)
-
-# to draw an optimum line
-eps_plot %>% abline(h = 13, lty = 2)
-
-eps <- 13
-minPts <- 3
-
-res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, method="raw")
-res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
-all(res.fpc$cluster == res.db$cluster)
-
-fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
-
-```
-
-# **Three-Way ANOVA**
-
-We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset.
-Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions.
-We are particularly interested in the highest order interaction term Person \* Variable \* Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.
-
-```{r threeWayAnovaMainBiDaily, message=FALSE, warning=FALSE}
-capture.output(
-  threeWayAnoBiDailyImp <- threewayanova(
-    Y = dtMainImpWide,
-    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWide)))) # Nr. time points (C)
-  ), 
-file='NUL')
-
-options(knitr.kable.NA = '')
-data.frame(SS = sapply(threeWayAnoBiDailyImp, c)) %>%
-  rownames_to_column(., var = "EffectNam") %>%
-  mutate(RSqrd = SS / sum(SS) * 100) %>%
-  add_row(
-    EffectNam = "Total", 
-    SS = sum(.$SS),
-    RSqrd = NA
-  ) %>%
-  mutate(
-    Effect = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Person*Variable",
-      "Person*Time",
-      "Variable*Time",
-      "Person*Variable*Time+error",
-      "Total"
-    )
-  ) %>%
-  select(Effect, everything(), -EffectNam) %>%
+mod1 <- Mclust(scaled_data)
+summary(mod1, parameters = FALSE)
+
+# multivariate <- densityMclust(scaled_data, plot=FALSE)
+# summary(multivariate)            
+# plot(multivariate, what = "BIC")
+
+# merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View
+
+#table(class, mod1$classification)
+
+#mod1dr <- MclustDR(mod1)
+#summary(mod1dr, parameters = FALSE)
+#plot(mod1dr, what = "classification")
+```
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  The feature selection issue is further highlighted when we do the GMMs by feature. Where we find a reasonable number of clusters for each feature. 
+</div>
+
+```{r MClust feature specific cluster numbers}
+nMClust <- list()
+for(f in unique(gsub("^.*?\\_", "", names(featFullImp$featuresImpZMat)))){
+  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
+  mod1 <- Mclust(scaled_data)
+  #summary(mod1, parameters = FALSE)
+  nMClust[[f]] <- mod1$G
+}
+
+nMClust %>%
+  unlist %>%
+  data.frame(feature=names(.), G=., row.names=NULL) %>%
   kbl(.,
@@
       booktabs = T,
-      align = c("l", "r", "c"),
+      align = c("l", "c"),
       digits=2,
       col.names = c(
-        "Effect",
-        "Sum of Squares",
-        "$R^2$"
+        "Feature",
+        "Number of Clusters"
       ),
-      caption = "Three-way ANOVA Main Analysis Bi-daily (Single Imputation)") %>%
+      caption = "Number of MClust Clusters if separated by Feature") %>%
   kable_classic(
@@
 
-<br>
-
-We find that for the bi-daily time scale the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA.
-Interestingly, time itself has a relatively small percentage of explained variance and only becomes important in compination with person- or person- and variable-specific differences.
-
-```{r threeWayAnovaMainDaily, message=FALSE, warning=FALSE}
-capture.output(
-  threeWayAnoDailyImp <- threewayanova(
-    Y = dtMainDailyImpWide,
-    n = nrow(dtMainDailyImpWide), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainDailyImpWide)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainDailyImpWide)))) # Nr. time points (C)
-  ), 
-file='NUL')
-
-options(knitr.kable.NA = '')
-data.frame(SS = sapply(threeWayAnoDailyImp, c)) %>%
-  rownames_to_column(., var = "EffectNam") %>%
-  mutate(RSqrd = SS / sum(SS) * 100) %>%
-  add_row(
-    EffectNam = "Total", 
-    SS = sum(.$SS),
-    RSqrd = NA
-  ) %>%
-  mutate(
-    Effect = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Person*Variable",
-      "Person*Time",
-      "Variable*Time",
-      "Person*Variable*Time+error",
-      "Total"
-    )
-  ) %>%
-  select(Effect, everything(), -EffectNam) %>%
+#### Density-based
+
+For the density-based cluster analysis we focus on the *DBSCAN* algorithm ('Density-Based Spatial Clustering of Applications with Noise') as well as its extension the *OPTICS* procedure ('Ordering Points To Identify the Clustering Structure').
+This procedure is more bottom up than the previous two approaches.
+Instead of providing the number of clusters, a cluster is specified if a minimum number of other points (the MinPts parameter) are within a given radius (the epsilon parameter).
+Together these two parameters determine a density at which a group of data points are considered a cluster.
+While this description is grossly oversimplified, it highlights one of the advantages of the procedures, namely that dense regions can expand to be any shape in the d-dimensional space.
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  Same feature selection issue as above. 
+</div>
+
+```{r dbscan}
+library(dbscan)
+library(fpc)
+
+#scaled_data <- featFullImp$featuresImpZMat %>% select(starts_with(varNamS123Int))
+scaled_data <- featFullImp$featuresImpZMat  # %>% select(starts_with(varNamS123NoInt))#%>%
+  #select(-any_of(names(which(colSums(is.na(scaled_data)) > 0))))
+
+#is.na(scaled_data) %>% table
+
+# to plot the eps values
+eps_plot = kNNdistplot(scaled_data, k=3)
+# to draw an optimum line
+eps_plot %>% abline(h = 15, lty = 2)
+
+potentialMinPts <- round(SciViews::ln(nrow(scaled_data)))
+eps <- 15
+minPts <- 2
+
+res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0)
+res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
+
+resEqual.check <- all(res.fpc$cluster == res.db$cluster)
+nClustDBSCAN <- length(unique(res.fpc$cluster))-1
+
+fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
+#hullplot(scaled_data, res.fpc)
+
+```
+
+<div class="alert alert-warning alert-dismissible fade in" role="alert">
+  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
+  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
+  And again the number of clusters is reasonable if we consider the features separately: 
+</div>
+
+```{r dbscan feature specific cluister numbers}
+minPts <- 3
+xi <- 0.04 #0.05
+nDbscan <- list()
+for(f in unique(gsub("^.*?\\_", "", names(featFullImp$featuresImpZMat)))){
+  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
+  #res.fpc <- dbscan::dbscan(scaled_data, eps=eps_cl, MinPts=minPts)
+
+  res.optics <- dbscan::optics(scaled_data, minPts = 3)
+
+  #plot(res.optics, col = "grey")
+  #dend <- as.dendrogram(res.optics)
+  #plot(dend, ylab = "Reachability dist.", leaflab = "none")
+  #extractDBSCAN(res.optics, eps_cl = 19.12)
+
+  res <- extractXi(res.optics, xi = xi)
+  res$clusters_xi
+  #plot(res)
+  #hullplot(scaled_data, res)
+
+  #summary(res.optics, parameters = FALSE)
+  nDbscan[[f]] <- nrow(res$clusters_xi)
+}
+
+nDbscan %>%
+  unlist %>%
+  data.frame(feature=names(.), G=., row.names=NULL) %>%
   kbl(.,
@@
       booktabs = T,
-      align = c("l", "r", "c"),
+      align = c("l", "c"),
       digits=2,
       col.names = c(
-        "Effect",
-        "Sum of Squares",
-        "$R^2$"
+        "Feature",
+        "Number of Clusters"
       ),
-      caption = "Three-way ANOVA Main Analysis Daily (Single Imputation)") %>%
+      caption = "Number of OPTICS Clusters if separated by Feature") %>%
   kable_classic(
@@
 
-<br>
-
-We additionally see that for the daily time scale aggregation, the $Person*Time*Variable+error$ sums of squares are reduced but remain the largest sums of squares and are still substantial.
-
-```{r threeWayAnovaMainWweekly, message=FALSE, warning=FALSE}
-capture.output(
-  threeWayAnoWeeklyImp <- threewayanova(
-    Y = dtMainWeeklyImpWide,
-    n = nrow(dtMainWeeklyImpWide), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainWeeklyImpWide)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainWeeklyImpWide)))) # Nr. time points (C)
-  ), 
-file='NUL')
-
-options(knitr.kable.NA = '')
-data.frame(SS = sapply(threeWayAnoWeeklyImp, c)) %>%
-  rownames_to_column(., var = "EffectNam") %>%
-  mutate(RSqrd = SS / sum(SS) * 100) %>%
-  add_row(
-    EffectNam = "Total", 
-    SS = sum(.$SS),
-    RSqrd = NA
-  ) %>%
-  mutate(
-    Effect = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Person*Variable",
-      "Person*Time",
-      "Variable*Time",
-      "Person*Variable*Time+error",
-      "Total"
-    )
-  ) %>%
-  select(Effect, everything(), -EffectNam) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "r", "c"),
-      digits=2,
-      col.names = c(
-        "Effect",
-        "Sum of Squares",
-        "$R^2$"
-      ),
-      caption = "Three-way ANOVA Main Analysis Weekly (Single Imputation)") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-<br>
-
-However, with the aggregation to `r length(unique(sub("_.*", "", names(dtMainWeeklyImpWide))))` weekly measurements, the $Person*Time*Variable+error$ sums of squares become less prominent and the variable differences explain substantially more variance.
-
-# **Plots per Item** {.tabset .tabset-fade}
-
-The three-mode PCA ultimately aims to explain the most variance with the fewest components (by maximizing the variance of data mapped in a lower dimensional space --- often by maximally separating groups of cases with similar variable patterns).
-Importantly, this procedure depends on variances and covariances to be present within the data.
-As one of our three modes is time and variance over time is not always given (e.g., because a variable is stable over the measured time span, such as personality over a month), it becomes important to assess whether a dimension reduction is possible for the variables measured during the measured time period.
-We, thus, start by visually inspecting the average changes and variance developments of all potentially relevant variables.
-We hope so see changes within the means and standard deviations over the measurement periods.
-
-Please note that the items might be measured on different ranges.
-To avoid misinterpretation of variances and changes over time, we cluster variables with similar ranges in different plots where necessary.
-
-## Means
-
-To illustrate the developments of the means,we plot the item means of all variables.
-We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point.
-The displayed time grouping is based on the measurement index rather than the actual date to ensure comparability of the changes between each measurement as well as studies.
-We plot the means for bi-daily, daily, and weekly timescales.
-
-### Studies 1-3 (Means) {.unlisted .unnumbered}
-
-We begin by assessing the imputed dataset of the main analysis, where participants from all three studies are combined, using the relevant variables that are available for all three studies.
-
-```{r plotAllMean}
-varClS123 <- dtPcaS123Imp %>%
-  PCAnorm(data = .,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(.)[!names(.) %in% idVars]) %>%
-  select(
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  summarise_all(
-    sd, na.rm = TRUE
-  ) %>%
-  t %>% 
-  as.data.frame %>%
-  kmeans(., centers = 2)
-
-PCAnorm(data = dtPcaS123Imp,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
-  select(
-    ID,
-    TIDnum,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  melt(
-    .,
-    id=c("ID","TIDnum")
-  ) %>%
-  mutate(
-    variable = gsub("_gmc", "", variable)
-  ) %>%
-  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
-  #geom_jitter() +
-  stat_summary(fun=mean, geom="line") +
-  labs(
-    title = "Studies 1-3 Means over time [bi-daily]",
-    y = "Grand Mean Centered Average",
-    x = "Variable",
-    color = "Timepoint",
-    group = "Timepoint"
-  ) +
-  #coord_flip() +
-  theme_Publication() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust=1),
-    legend.key.size = unit(1, 'cm'),
-    legend.key.height = unit(0.4, 'cm'),
-  )
-
-dtPcaS123 %>%
-  group_by(ID, date, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(date))) %>%
-  ungroup %>%
-  select(ID, date, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  PCAnorm(data = .,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(.)[!names(.) %in% idVars]) %>% 
-  select(
-    ID,
-    TIDnum,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  melt(
-    .,
-    id=c("ID","TIDnum")
-  ) %>%
-  mutate(
-    variable = gsub("_gmc", "", variable)
-  ) %>%
-  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
-  #geom_jitter() +
-  stat_summary(fun=mean, geom="line") +
-  labs(
-    title = "Studies 1-3 Means over time [daily]",
-    y = "Grand Mean Centered Average",
-    x = "Variable",
-    color = "Timepoint",
-    group = "Timepoint"
-  ) +
-  #coord_flip() +
-  theme_Publication() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust=1),
-    legend.key.size = unit(1, 'cm'),
-    legend.key.height = unit(0.4, 'cm'),
-  )
-
-dtPcaS123 %>%
-  group_by(ID, week, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(week))) %>%
-  ungroup %>%
-  select(ID, week, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  PCAnorm(data = .,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(.)[!names(.) %in% idVars]) %>% 
-  select(
-    ID,
-    TIDnum,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  melt(
-    .,
-    id=c("ID","TIDnum")
-  ) %>%
-  mutate(
-    variable = gsub("_gmc", "", variable)
-  ) %>%
-  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
-  #geom_jitter() +
-  stat_summary(fun=mean, geom="line") +
-  labs(
-    title = "Studies 1-3 Means over time [weekly]",
-    y = "Grand Mean Centered Average",
-    x = "Variable",
-    color = "Timepoint",
-    group = "Timepoint"
-  ) +
-  #coord_flip() +
-  theme_Publication() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust=1),
-    legend.key.size = unit(1, 'cm'),
-    legend.key.height = unit(0.4, 'cm'),
-  )
-```
-
-### Study 1 (Means) {.unlisted .unnumbered}
-
-The first study, utilized not only the smallest sample but also a slightly different set of measured variables compared to the later two studies.
-Several variables are consistent across all three studies (e.g., types of social interactions, interaction needs, self-determiantion theory needs, outgroup attitudes, and experienced well-being).
-Study 1 additionally included several emotion and mood measurements but fewer cognitive and behavioral measures than studies two and three.
-
-```{r plotS1Mean, fig.height=16}
-plotStat(
-  dt = dtS1Red,
-  varNams = varNamS1PCA,
-  id = "PID",
-  timescale = "bi-daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 1 Means over time [bi-daily]"
-)
-
-plotStat(
-  dt = dtS1Red,
-  varNams = varNamS1PCA,
-  id = "PID",
-  timescale = "daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 1 Means over time [daily]"
-)
-
-plotStat(
-  dt = dtS1Red,
-  varNams = varNamS1PCA,
-  id = "PID",
-  timescale = "weekly",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 1 Means over time [weekly]"
-)
-```
-
-### Study 2 (Means) {.unlisted .unnumbered}
-
-During the second study we collected the largest sample and additionally collected a range of motivations as well as pro-social and anti-social behaviors.
-
-We, again, select all potentially relevant variables and their labels, and then plot their developments on a bi-daily, daily, and weekly level.
-
-We again begin by plotting the item means of all variables. The plotting method is identical to that of Study 1.
-
-```{r plotS2Mean, fig.height=16}
-plotStat(
-  dt = dtS2Red,
-  varNams = varNamS2PCA,
-  id = "PID",
-  timescale = "bi-daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 2 Means over time [bi-daily]"
-)
-
-plotStat(
-  dt = dtS2Red,
-  varNams = varNamS2PCA,
-  id = "PID",
-  timescale = "daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 2 Means over time [daily]"
-)
-
-plotStat(
-  dt = dtS2Red,
-  varNams = varNamS2PCA,
-  id = "PID",
-  timescale = "weekly",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 2 Means over time [weekly]"
-)
-```
-
-### Study 3 (Means) {.unlisted .unnumbered}
-
-The third study uses a similar set up as Study 2 but targets a more vulnerable sample (of young medical professionals).
-While most key variables are identical to Study 2, in this last study we additionally collected several cognitive evaluations (e.g., importance ratings) and emotional status measures (e.g., anger, nervousness, energy, and loneliness).
-
-We, again, select all potentially relevant variables and their labels, and then plot their developments on a bi-daily, daily, and weekly level.
-
-As with the previous studies, we begin by plotting the item means of all variables. The plotting method is identical to that of Studies 1 and 2.
-
-```{r plotS3Mean, fig.height=16}
-plotStat(
-  dt = dtS3Red,
-  varNams = varNamS3PCA,
-  id = "PID",
-  timescale = "bi-daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 3 Means over time [bi-daily]"
-)
-
-plotStat(
-  dt = dtS3Red,
-  varNams = varNamS3PCA,
-  id = "PID",
-  timescale = "daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 3 Means over time [daily]"
-)
-
-plotStat(
-  dt = dtS3Red,
-  varNams = varNamS3PCA,
-  id = "PID",
-  timescale = "weekly",
-  clNum = 2,
-  idVars = idVars,
-  stat = "mean",
-  title = "Study 3 Means over time [weekly]"
-)
-```
-
-## SD
-
-To illustrate the developments of variance around the means, we also plot the item standard deviations of all variables.
-We, thus, plot the average variation from the mean for each individual variable, again summarizing the responses of all participants at any given time point.
-We retain the same measurement indices as the x axis.
-
-### Studies 1-3 (SDs) {.unlisted .unnumbered}
-
-<!-- <font size="5">Studies 1-3 (SDs)</font>  -->
-
-```{r plotAllSd}
-PCAnorm(data = dtPcaS123,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
-  select(
-    ID,
-    TIDnum,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  melt(
-    .,
-    id=c("ID","TIDnum")
-  ) %>%
-  mutate(
-    variable = gsub("_gmc", "", variable)
-  ) %>%
-  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
-  #geom_jitter() +
-  stat_summary(fun=sd, geom="line") +
-  labs(
-    title = "Studies 1-3 SDs over time [bi-daily]",
-    y = "Standard Deviation",
-    x = "Variable",
-    color = "Timepoint",
-    group = "Timepoint"
-  ) +
-  #coord_flip() +
-  theme_Publication() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust=1),
-    legend.key.size = unit(1, 'cm'),
-    legend.key.height = unit(0.4, 'cm'),
-  )
-
-dtPcaS123 %>%
-  group_by(ID, date, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(date))) %>%
-  ungroup %>%
-  select(ID, date, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  PCAnorm(data = .,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(.)[!names(.) %in% idVars]) %>% 
-  select(
-    ID,
-    TIDnum,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  melt(
-    .,
-    id=c("ID","TIDnum")
-  ) %>%
-  mutate(
-    variable = gsub("_gmc", "", variable)
-  ) %>%
-  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
-  #geom_jitter() +
-  stat_summary(fun=sd, geom="line") +
-  labs(
-    title = "Studies 1-3 SDs over time [daily]",
-    y = "Standard Deviation",
-    x = "Variable",
-    color = "Timepoint",
-    group = "Timepoint"
-  ) +
-  #coord_flip() +
-  theme_Publication() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust=1),
-    legend.key.size = unit(1, 'cm'),
-    legend.key.height = unit(0.4, 'cm'),
-  )
-
-dtPcaS123 %>%
-  group_by(ID, week, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(week))) %>%
-  ungroup %>%
-  select(ID, week, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
-  PCAnorm(data = .,
-          pid = "ID",
-          tid = "TIDnum",
-          selection = names(.)[!names(.) %in% idVars]) %>% 
-  select(
-    ID,
-    TIDnum,
-    ends_with("_gmc")
-  ) %>%
-  data.frame %>%
-  melt(
-    .,
-    id=c("ID","TIDnum")
-  ) %>%
-  mutate(
-    variable = gsub("_gmc", "", variable)
-  ) %>%
-  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
-  #geom_jitter() +
-  stat_summary(fun=sd, geom="line") +
-  labs(
-    title = "Studies 1-3 SDs over time [weekly]",
-    y = "Standard Deviation",
-    x = "Variable",
-    color = "Timepoint",
-    group = "Timepoint"
-  ) +
-  #coord_flip() +
-  theme_Publication() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust=1),
-    legend.key.size = unit(1, 'cm'),
-    legend.key.height = unit(0.4, 'cm'),
-  )
-```
-
-### Study 1 (SDs) {.unlisted .unnumbered}
-
-```{r plotS1Sd, fig.height=16}
-plotStat(
-  dt = dtS1Red,
-  varNams = varNamS1PCA,
-  id = "PID",
-  timescale = "bi-daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 1 SDs over time [bi-daily]"
-)
-
-plotStat(
-  dt = dtS1Red,
-  varNams = varNamS1PCA,
-  id = "PID",
-  timescale = "daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 1 SDs over time [daily]"
-)
-
-plotStat(
-  dt = dtS1Red,
-  varNams = varNamS1PCA,
-  id = "PID",
-  timescale = "weekly",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 1 SDs over time [weekly]"
-)
-```
-
-### Study 2 (SDs) {.unlisted .unnumbered}
-
-```{r plotS2Sd, fig.height=16}
-plotStat(
-  dt = dtS2Red,
-  varNams = varNamS2PCA,
-  id = "PID",
-  timescale = "bi-daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 2 SDs over time [bi-daily]"
-)
-
-plotStat(
-  dt = dtS2Red,
-  varNams = varNamS2PCA,
-  id = "PID",
-  timescale = "daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 2 SDs over time [daily]"
-)
-
-plotStat(
-  dt = dtS2Red,
-  varNams = varNamS2PCA,
-  id = "PID",
-  timescale = "weekly",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 2 SDs over time [weekly]"
-)
-```
-
-### Study 3 (SDs) {.unlisted .unnumbered}
-
-```{r plotS3Sd, fig.height=16}
-plotStat(
-  dt = dtS3Red,
-  varNams = varNamS3PCA,
-  id = "PID",
-  timescale = "bi-daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 3 SDs over time [bi-daily]"
-)
-
-plotStat(
-  dt = dtS3Red,
-  varNams = varNamS3PCA,
-  id = "PID",
-  timescale = "daily",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 3 SDs over time [daily]"
-)
-
-plotStat(
-  dt = dtS3Red,
-  varNams = varNamS3PCA,
-  id = "PID",
-  timescale = "weekly",
-  clNum = 2,
-  idVars = idVars,
-  stat = "sd",
-  title = "Study 3 SDs over time [weekly]"
-)
-```
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-# **Three-Mode PCA**
-
-## Stability Test
-
-### Main Analysis
-
-- all three studies combined
-- variables shared across studies
-- interaction and no interaction responses
-- bi-daily time scale
-
-#### Center and Normalize
-
-```{r MainStabCZ}
-##################################################################
-##                  Center across Participants                  ##
-##################################################################
-MiMainStabilityWideC <- list()
-for (i in 1:length(MiMainStabilityWide)) {
-  MiMainStabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiMainStabilityWide[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiMainStabilityWide[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWide[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiMainStabilityWide[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##################################################################
-##                 Normalize across Time Points                 ##
-##################################################################
-MiMainStabilityWideCZ <- list()
-for (i in 1:length(MiMainStabilityWideC)) {
-  MiMainStabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiMainStabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiMainStabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiMainStabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiMainStabilityWideCZ[[i]]) <- names(MiMainStabilityWideC[[i]])
-}
-```
-
-#### Component Complexity
-
-```{r MainStabCompNum}
-# Run PCA SUP to determine best combination
-MainStabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiMainStabilityWideCZ)) {
-    MainStabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiMainStabilityWideCZ[[i]], 
-    n = nrow(MiMainStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-MainStabDimSelector <- list()
-for (i in 1:length(MainStabPcaSup)) {
-  MainStabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = MainStabPcaSup[[i]],
-      n = nrow(MiMainStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiMainStabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-MainStabDim <- list()
-for (i in 1:length(MainStabDimSelector)) {
-  MainStabDim[[i]] <-
-    tail(MainStabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-MainStabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-### Follow-up: Studies (common variables)
-
-- **individual studies**
-- variables shared across studies
-- interaction and no interaction responses
-- bi-daily time scale
-
-#### Study 1
-
-##### Center and Normalize
-
-```{r S1CommonStabCZ}
-#################################################################
-##                           Study 1                           ##
-#################################################################
-
-##----------------------------------------------------------------
-##                  Center across Participants                  --
-##----------------------------------------------------------------
-MiS1StabilityWideC <- list()
-for (i in 1:length(MiS1StabilityWide)) {
-  MiS1StabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiS1StabilityWide[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiS1StabilityWide[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWide[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiS1StabilityWide[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##----------------------------------------------------------------
-##                 Normalize across Time Points                 --
-##----------------------------------------------------------------
-MiS1StabilityWideCZ <- list()
-for (i in 1:length(MiS1StabilityWideC)) {
-  MiS1StabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiS1StabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiS1StabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiS1StabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiS1StabilityWideCZ[[i]]) <- names(MiS1StabilityWideC[[i]])
-}
-```
-
-##### Component Complexity
-
-```{r S1CommonStabCompNum}
-# Run PCA SUP to determine best combination
-S1StabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiS1StabilityWideCZ)) {
-    S1StabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiS1StabilityWideCZ[[i]], 
-    n = nrow(MiS1StabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-S1StabDimSelector <- list()
-for (i in 1:length(S1StabPcaSup)) {
-  S1StabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = S1StabPcaSup[[i]],
-      n = nrow(MiS1StabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiS1StabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-S1StabDim <- list()
-for (i in 1:length(S1StabDimSelector)) {
-  S1StabDim[[i]] <-
-    tail(S1StabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-S1StabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Follow-up (S1 Common): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-#### Study 2
-
-##### Center and Normalize
-
-##### Component Complexity
-
-#### Study 3
-
-##### Center and Normalize
-
-##### Component Complexity
-
-### Follow-up: Studies (ideosyncratic variables)
-
-- **individual studies**
-- **variables idiosyncratic to studies**
-- interaction and no interaction responses
-- bi-daily time scale
-
-#### Study 1
-
-##### Center and Normalize
-
-```{r S1IdiosyncraticStabCZ}
-#################################################################
-##                           Study 1                           ##
-#################################################################
-
-##----------------------------------------------------------------
-##                  Center across Participants                  --
-##----------------------------------------------------------------
-MiS1IdioStabilityWideC <- list()
-for (i in 1:length(MiS1StabilityWideIdio)) {
-  MiS1IdioStabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiS1StabilityWideIdio[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiS1StabilityWideIdio[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1StabilityWideIdio[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiS1StabilityWideIdio[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##----------------------------------------------------------------
-##                 Normalize across Time Points                 --
-##----------------------------------------------------------------
-MiS1IdioStabilityWideCZ <- list()
-for (i in 1:length(MiS1IdioStabilityWideC)) {
-  MiS1IdioStabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiS1IdioStabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiS1IdioStabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1IdioStabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiS1IdioStabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiS1IdioStabilityWideCZ[[i]]) <- names(MiS1IdioStabilityWideC[[i]])
-}
-```
-
-##### Component Complexity
-
-```{r S1IdiosyncraticStabCompNum}
-# Run PCA SUP to determine best combination
-S1IdioStabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiS1IdioStabilityWideCZ)) {
-    S1IdioStabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiS1IdioStabilityWideCZ[[i]], 
-    n = nrow(MiS1IdioStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-S1IdioStabDimSelector <- list()
-for (i in 1:length(S1IdioStabPcaSup)) {
-  S1IdioStabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = S1IdioStabPcaSup[[i]],
-      n = nrow(MiS1IdioStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiS1IdioStabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-S1IdioStabDim <- list()
-for (i in 1:length(S1IdioStabDimSelector)) {
-  S1IdioStabDim[[i]] <-
-    tail(S1IdioStabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-S1IdioStabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Follow-up (S1 Idiosyncratic): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-#### Study 2
-
-##### Center and Normalize
-
-##### Component Complexity
-
-#### Study 3
-
-##### Center and Normalize
-
-##### Component Complexity
-
-### Follow-up: Response Type (interaction vs. no interaction)
-
-- all three studies combined
-- variables shared across studies
-- **interaction vs. no interaction responses**
-- bi-daily time scale
-
-#### All Interactions
-
-##### Center and Normalize
-
-```{r InteractionStabCZ}
-##################################################################
-##                  Center across Participants                  ##
-##################################################################
-MiInteractionStabilityWideC <- list()
-for (i in 1:length(MiInteractionStabilityWide)) {
-  MiInteractionStabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiInteractionStabilityWide[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiInteractionStabilityWide[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWide[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiInteractionStabilityWide[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##################################################################
-##                 Normalize across Time Points                 ##
-##################################################################
-MiInteractionStabilityWideCZ <- list()
-for (i in 1:length(MiInteractionStabilityWideC)) {
-  MiInteractionStabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiInteractionStabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiInteractionStabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiInteractionStabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiInteractionStabilityWideCZ[[i]]) <- names(MiInteractionStabilityWideC[[i]])
-}
-```
-
-##### Component Complexity
-
-```{r InteractionStabCompNum}
-
-# Run PCA SUP to determine best combination
-InteractionStabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiInteractionStabilityWideCZ)) {
-    InteractionStabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiInteractionStabilityWideCZ[[i]], 
-    n = nrow(MiInteractionStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-InteractionStabDimSelector <- list()
-for (i in 1:length(InteractionStabPcaSup)) {
-  InteractionStabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = InteractionStabPcaSup[[i]],
-      n = nrow(MiInteractionStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiInteractionStabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-InteractionStabDim <- list()
-for (i in 1:length(InteractionStabDimSelector)) {
-  InteractionStabDim[[i]] <-
-    tail(InteractionStabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-InteractionStabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Follow-up (Interactions): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-#### Intergroup Interactions
-
-##### Center and Normalize
-
-```{r IntergroupStabCZ}
-##################################################################
-##                  Center across Participants                  ##
-##################################################################
-MiIntergroupStabilityWideC <- list()
-for (i in 1:length(MiIntergroupStabilityWide)) {
-  MiIntergroupStabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiIntergroupStabilityWide[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiIntergroupStabilityWide[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWide[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWide[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##################################################################
-##                 Normalize across Time Points                 ##
-##################################################################
-MiIntergroupStabilityWideCZ <- list()
-for (i in 1:length(MiIntergroupStabilityWideC)) {
-  MiIntergroupStabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiIntergroupStabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiIntergroupStabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiIntergroupStabilityWideCZ[[i]]) <- names(MiIntergroupStabilityWideC[[i]])
-}
-```
-
-##### Component Complexity
-
-```{r IntergroupStabCompNum}
-
-# Run PCA SUP to determine best combination
-IntergroupStabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiIntergroupStabilityWideCZ)) {
-    IntergroupStabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiIntergroupStabilityWideCZ[[i]], 
-    n = nrow(MiIntergroupStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-IntergroupStabDimSelector <- list()
-for (i in 1:length(IntergroupStabPcaSup)) {
-  IntergroupStabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = IntergroupStabPcaSup[[i]],
-      n = nrow(MiIntergroupStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiIntergroupStabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-IntergroupStabDim <- list()
-for (i in 1:length(IntergroupStabDimSelector)) {
-  IntergroupStabDim[[i]] <-
-    tail(IntergroupStabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-IntergroupStabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Follow-up (Intergroup Interactions): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-### Follow-up: Time Scales (daily, weekly)
-
-- all three studies combined
-- variables shared across studies
-- interaction and no interaction responses
-- **daily and weekly time scales**
-
-#### Daily
-
-##### Center and Normalize
-
-```{r DailyStabCZ}
-##################################################################
-##                  Center across Participants                  ##
-##################################################################
-MiMainDailyStabilityWideC <- list()
-for (i in 1:length(MiMainDailyStabilityWide)) {
-  MiMainDailyStabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiMainDailyStabilityWide[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiMainDailyStabilityWide[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWide[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWide[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##################################################################
-##                 Normalize across Time Points                 ##
-##################################################################
-MiMainDailyStabilityWideCZ <- list()
-for (i in 1:length(MiMainDailyStabilityWideC)) {
-  MiMainDailyStabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiMainDailyStabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiMainDailyStabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiMainDailyStabilityWideCZ[[i]]) <- names(MiMainDailyStabilityWideC[[i]])
-}
-```
-
-##### Component Complexity
-
-```{r MainDailyStabCompNum}
-# Run PCA SUP to determine best combination
-MainDailyStabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiMainDailyStabilityWideCZ)) {
-    MainDailyStabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiMainDailyStabilityWideCZ[[i]], 
-    n = nrow(MiMainDailyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-MainDailyStabDimSelector <- list()
-for (i in 1:length(MainDailyStabPcaSup)) {
-  MainDailyStabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = MainDailyStabPcaSup[[i]],
-      n = nrow(MiMainDailyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiMainDailyStabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-MainDailyStabDim <- list()
-for (i in 1:length(MainDailyStabDimSelector)) {
-  MainDailyStabDim[[i]] <-
-    tail(MainDailyStabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-MainDailyStabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Follow-up (Daily): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-#### Weekly
-
-##### Center and Normalize
-
-```{r MainWeeklyStabCZ}
-##################################################################
-##                  Center across Participants                  ##
-##################################################################
-MiMainWeeklyStabilityWideC <- list()
-for (i in 1:length(MiMainWeeklyStabilityWide)) {
-  MiMainWeeklyStabilityWideC[[i]] <-
-    ThreeWay::cent3(
-      X = MiMainWeeklyStabilityWide[[i]],
-      # Nr. ppt.s (A)
-      n = nrow(MiMainWeeklyStabilityWide[[i]]),
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWide[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWide[[i]])))),
-      # Center across Participants
-      mode = 1
-    ) %>%
-    as.data.frame
-}
-
-##################################################################
-##                 Normalize across Time Points                 ##
-##################################################################
-MiMainWeeklyStabilityWideCZ <- list()
-for (i in 1:length(MiMainWeeklyStabilityWideC)) {
-  MiMainWeeklyStabilityWideCZ[[i]] <-
-    ThreeWay::norm3(
-      X = MiMainWeeklyStabilityWideC[[i]], 
-      # Nr. ppt.s (A)
-      n = nrow(MiMainWeeklyStabilityWideC[[i]]), 
-      # Nr. variables (B)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWideC[[i]])))),
-      # Nr. time points (C)
-      p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWideC[[i]])))),
-      # Normalize across Time Points
-      mode = 2
-    ) %>%
-    as.data.frame
-  names(MiMainWeeklyStabilityWideCZ[[i]]) <- names(MiMainWeeklyStabilityWideC[[i]])
-}
-```
-
-##### Component Complexity
-
-```{r MainWeeklyStabCompNum}
-# Run PCA SUP to determine best combination
-MainWeeklyStabPcaSup <- list()
-capture.output(
-  for(i in 1:length(MiMainWeeklyStabilityWideCZ)) {
-    MainWeeklyStabPcaSup[[i]] <- 
-    ThreeWay::T3runsApproxFit(
-    X = MiMainWeeklyStabilityWideCZ[[i]], 
-    n = nrow(MiMainWeeklyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  )
-  }, 
-file='NUL')
-
-# Select Dimension Complexity
-MainWeeklyStabDimSelector <- list()
-for (i in 1:length(MainWeeklyStabPcaSup)) {
-  MainWeeklyStabDimSelector[[i]] <- capture.output(
-    ThreeWay::DimSelector(
-      out = MainWeeklyStabPcaSup[[i]],
-      n = nrow(MiMainWeeklyStabilityWideCZ[[i]]), # Nr. ppt.s (A)
-      m = length(unique(sub("T[0-9]{2}_", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. variables (B)
-      p = length(unique(sub("_.*", "", names(MiMainWeeklyStabilityWideCZ[[i]])))), # Nr. time points (C)
-      model = 2
-    )
-  )
-}
-
-MainWeeklyStabDim <- list()
-for (i in 1:length(MainWeeklyStabDimSelector)) {
-  MainWeeklyStabDim[[i]] <-
-    tail(MainWeeklyStabDimSelector[[i]], n = 1) %>%
-    str_extract("\\(.*\\)") %>%
-    stringr::str_extract_all('\\d+', simplify = TRUE) %>%
-    as.data.frame %>%
-    mutate_all(as.numeric) %>%
-    dplyr::rename(c(A = V1,
-                    B = V2,
-                    C = V3))
-}
-
-MainWeeklyStabDim %>%
-  bind_rows %>%
-  as.data.frame %>%
-  rownames_to_column(., var = "imputation") %>%
-  group_by(A, B, C) %>%
-  summarise(Frequency = n()) %>%
-  arrange(-Frequency) %>%
-  mutate(Percentage = formattable::percent(Frequency/sum(.$Frequency))) %>%
-  kbl(
-    .,
-    escape = FALSE,
-    booktabs = TRUE,
-    align = "c",
-    digits = 2,
-    col.names = c(
-      "Person",
-      "Variable",
-      "Time",
-      "Frequency",
-      "Percentage"
-    ),
-    caption = "Follow-up (Weekly): Frequency Table of Suggested Component Dimensions (from 20 imputed datasets)") %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-## Main Analysis
-
-### Center and Normalize
-
-We begin by centering across participants (within time point and variable) and normalize across time points and participants (within variable).
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Question: </b><br/>
-  Are the centering and especially the normalization done correctly?
-</div>
-
-```{r 3mpcaMainPrep}
-##################################################################
-##                  Center across Participants                  ##
-##################################################################
-dtMainImpWideC <- ThreeWay::cent3(
-  X = dtMainImpWide, 
-  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
-  p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
-  mode = 1
-) %>%
-  as.data.frame
-#apply(dtMainImpWideC,2,sum)
-
-##################################################################
-##                 Normalize across Time Points                 ##
-##################################################################
-dtMainImpWideCZ <- ThreeWay::norm3(
-  X = dtMainImpWideC, 
-  n = nrow(dtMainImpWideC), # Nr. ppt.s (A)
-  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideC)))), # Nr. variables (B)
-  p = length(unique(sub("_.*", "", names(dtMainImpWideC)))), # Nr. time points (C)
-  mode = 2
-) %>%
-  as.data.frame
-names(dtMainImpWideCZ) <- names(dtMainImpWideC)
-#apply(dtMainImpWideC,2,sum)
-# dtMainImpWideCZ %>%
-#   select(ends_with("_AttitudesDutch")) %>%
-#   unlist %>%
-#   sd
-# psych::describe(dtMainImpWideCZ)
-```
-
-### Select Component Complexity
-
-We then select the optimal number of components for each mode (i.e., mode complexity).
-To do so, we use Convex Hull procedure heuristic based on total number of components and scree test value taken from a PCA-SUP analysis (where, PCASUP analyses are PCA's of supermatrices with slices of the 3way array next to each other, thus 3 supermatrices are analyed by PCA, and for each we get a component matrix, and eigenvalues.
-The eigenvalues may give an indication as to the required number of components for each mode.).
-
-```{r 3mpcaMainCompNum}
-# Run PCA SUP to determine best combination
-capture.output(
-  MainPcaSup <- ThreeWay::T3runsApproxFit(
-    X = dtMainImpWideCZ, 
-    n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
-    maxa = 20,
-    maxb = 20,
-    maxc = 20
-  ), 
-file='NUL')
-
-# Select Dimension Complexity
-MainDimSelector <- capture.output(
-  ThreeWay::DimSelector(
-    out = MainPcaSup,
-    n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
-    model = 2
-  )
-)
-MainDim <- tail(MainDimSelector, n = 1) %>%
-  str_extract("\\(.*\\)") %>%
-  stringr::str_extract_all('\\d+',simplify=TRUE) %>%
-  as.data.frame %>% 
-  mutate_all(as.numeric) %>%
-  dplyr::rename(c(
-    A = V1,
-    B = V2,
-    C = V3
-  ))
-#MainDim
-```
-
-Based on the Convex Hull heuristic, we select `r MainDim$A` components for the participant mode, `r MainDim$B` principle components for the variable mode, and `r MainDim$C` components for the time mode.
-It should be noted that these heuristics are sensitive to variations in the multiple imputation procedure.
-
-### Model Fitting
-
-We then run the actual three-mode PCA with the inputs from the dimension selection procedure.
-
-```{r 3mpcaMainRun}
-
-# nrow(dtMainImpWide)
-# length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars])
-# length(unique(dtPcaS123Imp$TIDnum))
-# 
-# t3MainFull <- ThreeWay::T3(
-#   dtMainImpWide,
-#   paste("PP", 1:nrow(dtMainImpWide), sep = "_"),
-#   names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
-#   paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
-# )
-
-# 3MPCA with general trend
-capture.output(
-  t3MainRaw <- ThreeWay::T3func(
-    X = dtMainImpWide,
-    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
-    r1 = MainDim$A,
-    r2 = MainDim$B,
-    r3 = MainDim$C,
-    start = 0,
-    conv = 1e-6
-  ), 
-file='NUL')
-
-# 3MPCA without general trend
-capture.output(
-  t3Main <- ThreeWay::T3func(
-    X = dtMainImpWideCZ,
-    n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
-    r1 = MainDim$A,
-    r2 = MainDim$B,
-    r3 = MainDim$C,
-    start = 0,
-    conv = 1e-6
-  ), 
-file='NUL')
-
-# individual fit partitioning
-FitT3Main <- T3fitpartitioning(
-  Xprep = dtMainImpWideCZ,
-  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
-  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
-  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
-  AS = t3Main$A, 
-  BT = t3Main$B, 
-  CU = t3Main$C, 
-  K = t3Main$H, 
-  renormmode = 0,
-  laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"), 
-  labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
-  labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
-)
-
-# Split half analysis
-capture.output(
-  SplitT3Main <- splithalfT3(
-    X = dtMainImpWide,
-    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
-    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
-    p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
-    r1 = MainDim$A,
-    r2 = MainDim$B,
-    r3 = MainDim$C, 
-    centopt = 1,
-    normopt = 2,
-    conv = 1e-6,
-    renormmode = 0,
-    addanal = 5,
-    wa_rel = 0,
-    wb_rel = 0,
-    wc_rel = 0,
-    laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"), 
-    labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
-    labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
-  ), 
-file='NUL')
-# 0
-# 0
-```
-
-In terms of fit percentage, we find that the three-mode PCA with the general trend included captures `r format(round(t3MainRaw$fp,2), nsmall = 2)`\% of the original variance, and the three-mode PCA without the general trend captures `r format(round(t3Main$fp,2), nsmall = 2)`\% of the variance.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Question: </b><br/>
-  What is the interpretation of these Percentages? What exactly do the differences mean?
-</div>
-
-### Rotation
-
-For interpretability of the components we perform an orthomax rotation of the component matrices and the core array.
-
-```{r 3mpcaMainRotation}
-capture.output(
-  t3MainRot <- varimcoco(
-    A = t3Main$A, 
-    B = t3Main$B, 
-    C = t3Main$C, 
-    H = t3Main$H,
-    wa_rel = 0,
-    wb_rel = 0,
-    wc_rel = 0
-  ),
-  file = 'NULL'
-)
-```
-
-```{r 3mpcaMainAMat}
-t3MainRot$AS %>%
-  as.data.frame %>%
-  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
-  mutate(Participant = paste("PP", row_number(), sep = "_")) %>%
-  select(
-    Participant, 
-    everything()
-  ) %>%
-  kbl(
-  .,
-  format = "html",
-  escape = FALSE,
-  align = c("l", rep("c",ncol(.)-1)),
-  booktabs = TRUE,
-  digits = 3,
-  caption = "Rotated Component Matrix: Person Mode" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-```{r 3mpcaMainBMat}
-t3MainRot$BT %>%
-  as.data.frame %>%
-  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
-  mutate(Variable = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) %>%
-  select(
-    Variable, 
-    everything()
-  ) %>%
-  kbl(
-  .,
-  format = "html",
-  escape = FALSE,
-  align = c("l", rep("c",ncol(.)-1)),
-  booktabs = TRUE,
-  digits = 3,
-  caption = "Rotated Component Matrix: Variable Mode" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r 3mpcaMainCMat}
-t3MainRot$CU %>%
-  as.data.frame %>%
-  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
-  mutate(Time = paste("T", row_number()-1, sep = "")) %>%
-  select(
-    Time, 
-    everything()
-  ) %>%
-  kbl(
-  .,
-  format = "html",
-  escape = FALSE,
-  align = c("l", rep("c",ncol(.)-1)),
-  booktabs = TRUE,
-  digits = 3,
-  caption = "Rotated Component Matrix: Time Mode" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-### Visualization: Time Components
-
-We visualize the `r ncol(t3MainRot$C)` time components in their component scores over time.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Not sure whether this interpretation is correct.
-</div>
-
-```{r 3mpcaMainCompVizTime}
-t3MainRot$C %>%
-  as.data.frame %>%
-  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
-  mutate(tid = row_number()) %>%
-  melt(., id.vars = "tid") %>%
-  ggplot(
-    .,
-    aes(
-      x = tid,
-      y = value,
-      linetype = variable,
-      shape = variable,
-      color = variable
-    )
-  ) +
-  geom_point() +
-  geom_line() +
-  geom_line(
-    stat = "smooth",
-    method = "loess",
-    se = FALSE,
-    span = .25,
-    alpha = 0.75
-  ) +
-  labs(x = "Time occasion",
-       y = "Component score",
-       linetype = "Component",
-       shape = "Component",
-       color = "Component") +
-  theme_Publication()
-
-```
-
-### Visualization: Person Components
-
-We visualize how well the first two person components separate the participants.
-We focus on the first two components because they capture the most variance, while retaining easy two dimensional visualization options.
-We additionally color the participants using k means clustering to further showcase the separation.
-
-```{r 3mpcaMainCompVizPerson}
-# # Unsure how this should be interpreted 
-# jointplotgen(
-#   K = t3MainRot$K,
-#   A = t3MainRot$A,
-#   B = t3MainRot$B,
-#   C = t3MainRot$C,
-#   fixmode = 1,
-#   fixunit = 3,
-#   laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"),
-#   labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
-#   labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
-# )
-
-library(factoextra)
-fviz_nbclust(t3MainRot$A[,1:2], kmeans, method = "wss")
-clusterdata <- t3MainRot$A[,1:2]
-km.res <- kmeans(clusterdata, 3, nstart = 25)
-fviz_cluster(object = km.res, # kmeans object
-             data = clusterdata, # data used for clustering
-             ellipse.type = "norm",
-             geom = "point",
-             palette = "jco",
-             main = "",
-             xlab = "Person Component 1",
-             ylab = "Person Component 2",
-             ggtheme = theme_Publication())
-```
-
-### Visualization: Variable Components
-
-To visualize how the principle components separate the different variables, we draw the component scores for all variables.
-
-```{r 3mpcaMainCompVizVariable}
-t3MainRot$B %>%
-  as.data.frame %>%
-  rename_all(funs(str_replace_all(., "V", "Component_"))) %>%
-  mutate(var = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) %>%
-  melt(., id.vars = "var") %>%
-  ggplot(
-    .,
-    aes(
-      x = var,
-      y = value,
-      group = variable,
-      linetype = variable,
-      shape = variable,
-      color = variable
-    )
-  ) +
-  geom_point() +
-  geom_line() +
-  coord_flip() +
-  labs(x = "Variable",
-       y = "Component score",
-       linetype = "Component",
-       shape = "Component",
-       color = "Component") +
-  theme_Publication()
-```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  How do we best interpret and visualize the joint decomposition? 
-</div>
+### Follow-up: By Response Type
+
+_**currently removed for faster render until feature selection issue is resolved**_
+
+### Follow-up: By Study
+
+_**currently removed for faster render until feature selection issue is resolved**_
 

### 2022-10-02 scripts/functions/tsFeatureExtractor.R

@@ -216,7 +216,7 @@ featureImputer <- function(extractorList) {
   input
 }
 
-pMiss <- function(x){sum(is.na(x))/(ncol(x)*nrow(x))*100}
+pMissTot <- function(x){sum(is.na(x))/(ncol(x)*nrow(x))*100}
 pColMiss <- function(x){sum(is.na(x))/length(x)*100}
 
 pMissFeat <- function(all, contact, nocontact, title) {
@@ -227,16 +227,16 @@ pMissFeat <- function(all, contact, nocontact, title) {
   # title = "Feature-wise Missingess Across all Studies"
   
   pMiss <- list()
-  pMiss$All <- data.frame(set="All", pMissTotal=pMiss(all))
-  pMiss$Contact <- data.frame(set="Contact", pMissTotal=pMiss(contact))
-  pMiss$NoContact <- data.frame(set="No Contact", pMissTotal=pMiss(nocontact))
+  pMiss$All <- data.frame(set="All", pMissTotal=pMissTot(all))
+  pMiss$Contact <- data.frame(set="Contact", pMissTotal=pMissTot(contact))
+  pMiss$NoContact <- data.frame(set="No Contact", pMissTotal=pMissTot(nocontact))
   pMiss <- Reduce(function(x, y) merge(x, y, all=TRUE), pMiss) 
   
   pMissFeature <- list()
   pMissFeature$All <- apply(all, 2, pColMiss) %>% data.frame(feature=names(.), All=., row.names=NULL)
   pMissFeature$Contact <- apply(contact, 2, pColMiss) %>% data.frame(feature=names(.), Contact=., row.names=NULL)
   pMissFeature$NoContact <- apply(nocontact, 2, pColMiss) %>% data.frame(feature=names(.), NoContact=., row.names=NULL)
-  pMissFeature %>% 
+  pMissPlot <- pMissFeature %>% 
     reduce(full_join, by='feature') %>% 
     reshape2::melt(id="feature", variable.name="set", value.name="pMiss") %>% 
     na.omit %>% 
@@ -251,4 +251,10 @@ pMissFeat <- function(all, contact, nocontact, title) {
     ) +
     facet_wrap(~lab) +
     theme_Publication()
+  out <- list(
+    pmiss = pMiss,
+    pMissFeature = pMissFeature,
+    pMissPlot = pMissPlot
+  )
+  out
 }

### 2022-10-14 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 ```{r kmeans determin k}
+
+featureExclusion <- c(
+  "mean",
+  "sd",
+  "rmssd",
+  "ar02",
+  "edf_ar"
+)
+
 # prepare data for main analysis
 raw_data <- featData
-z_data <- featFullImp$featuresImpZ
-scaled_data <- featFullImp$featuresImpZMat
+z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion))
+scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
 
@@
 
-scaled_data <- featFullImp$featuresImpZMat
+scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
 
@@
 ```{r MClust feature specific cluster numbers}
+scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
 nMClust <- list()
-for(f in unique(gsub("^.*?\\_", "", names(featFullImp$featuresImpZMat)))){
+for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){
   scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
@@
   )
+
 ```
@@
 #scaled_data <- featFullImp$featuresImpZMat %>% select(starts_with(varNamS123Int))
-scaled_data <- featFullImp$featuresImpZMat  # %>% select(starts_with(varNamS123NoInt))#%>%
+scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))  # %>% select(starts_with(varNamS123NoInt))#%>%
   #select(-any_of(names(which(colSums(is.na(scaled_data)) > 0))))
@@
 # to draw an optimum line
-eps_plot %>% abline(h = 15, lty = 2)
+eps_plot %>% abline(h = 11, lty = 2)
 
 potentialMinPts <- round(SciViews::ln(nrow(scaled_data)))
-eps <- 15
+eps <- 11
 minPts <- 2
@@
 ```{r dbscan feature specific cluister numbers}
+scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
 minPts <- 3
@@
 nDbscan <- list()
-for(f in unique(gsub("^.*?\\_", "", names(featFullImp$featuresImpZMat)))){
+for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){
   scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))

### 2022-11-02 manuscript.tex
empty

--- 2022-05-12 TitlePage.tex
@@
 % define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
-\documentclass[man, 12pt, a4paper, donotrepeattitle]{apa7}
+\documentclass[man, 12pt, a4paper]{apa7}
 
@@
 \usepackage{scalerel}
-\usepackage{fontawesome}
+\usepackage{fontawesome5}
 \usepackage[normalem]{ulem}
@@
 \usepackage{amsmath}
+\usepackage{ntheorem}
 \usepackage{afterpage}
@@
 \usepackage{censor}
+\usepackage{pdflscape}
+\usepackage{lscape}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{caption}
+\usepackage{adjustbox}
+\usepackage{makecell}
+\usepackage{tabu}
+
+% make warning with red triangle
+\newcommand\Warning[1][2ex]{%
+  \renewcommand\stacktype{L}%
+  \scaleto{\stackon[1.3pt]{\color{red}$\triangle$}{\tiny\bfseries !}}{#1}}%
+
+% make question with red triangle
+\newcommand\Question[1][2ex]{%
+  \renewcommand\stacktype{L}%
+  \scaleto{\stackon[1.3pt]{\color{red}$\triangle$}{\tiny\bfseries ?}}{#1}}%
+
+% add definition sections
+\theoremstyle{break}
+\newtheorem{definition}{Definition}
+
+% add hypothesis sections
+\theoremstyle{plain}
+\theoremseparator{:}
+\newtheorem{hyp}{Hypothesis}
+
+\newtheorem{subhyp}{Hypothesis}
+   \renewcommand\thesubhyp{\thehyp\alph{subhyp}}
+
+% add quote section
+\usepackage{csquotes}
+
+% framed box section
+\usepackage{framed}
+\emergencystretch=1em
 
@@
 
+% custom open science badge height
+\newlength{\badgeheight}
+\setlength{\badgeheight}{1em}
+
 % language settings
@@
 % add reference library file
-\addbibresource{references.bib}
-
-% add reference library file
-\addbibresource{references.bib}
+\addbibresource{referencesZotero.bib}
 
 % Title and header
-\title{Migration Experience Trajectories: Subtitle TBD}
+\title{Migration Experience Trajectories (Working Title)}
 \shorttitle{Migration Experience Trajectories}
@@
 \affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
-\affil[3]{Author order TBD [currently in first name alphabetical order]}
+\affil[3]{Author order TBD (currently in first name alphabetical order)}
 
@@
 Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
-E-mail: j.kreienkamp@rug.nl}
+E-mail: \href{mailto:j.kreienkamp@rug.nl}{j.kreienkamp@rug.nl}
+}
 
@@
 
+% Abstract
+\abstract{
+To be written.
+
+\noindent\textbf{Public significance statement}: To be written.
+
+}
+
+\keywords{
+    TBD\\
+    \vspace{1em}
+    \noindent \textbf{Open Science Practices:}\\
+    \noindent \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registration-color.png}} Preregistration+, 
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data, 
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
+}
+
+% set indentation size
 \setlength\parindent{1.27cm}
@@
 
+% **CHEAT SHEET / LEGEND**
+%
+% Comments:
+% '%' starts a comment in LaTeX (not printed)
+% '\todo[inline]{} makes orange boxes in PDF
+% '\marginpar{}' notes in margins
+% '\footnote{}' footnote
+% '\Warning' important note indicator in PDF (triangle with exclamation mark)
+% '\Question' question note indicator in PDF (triangle with question mark)
+%
+% Citation (with Natbib citation style):
+% '\citep[e.g.][p. 15]{CitationKey}' citation in parentheses "(e.g., Berry, 2003, p. 15)"
+% '\citet{CitationKey}' citation in text "Berry (2003)"
+% '\citealt' and '\citealp' alternate citation without parentheses
+% '\citeauthor' and '\citeyear' only year or author
+% 
+% Headings:
+% '\part{}' and '\chapter{}' only relevant for multi-part or multi-chapter documents
+% '\section{}' heading level 1
+% '\subsection{}' heading level 2
+% '\subsubsection{}' heading level 3
+% '\paragraph{}' heading level 4
+% '\subparagraph{}' heading level 5
+%
+% formatting:
+% '\textbf{}' text bold font
+% '\textit{}' text italic font
+% '\underline{}' text underline
+% '\sout{}' text strike out
+% '\textsc{}' text small caps
+% '\vspace{1em}' add vertical space
+% '\hspace{1em}' add horizontal space
+% '\\' new line (i.e., line break)
+% '\pagebreak' start new page (i.e., page break)
+% '\noindent' do not indent current line (e.g., current paragraph)
+% 'begin{center}...end{center}' center text or object
+%
+% Math mode:
+% '$\alpha = .8$' mathematical equation inline
+% '$$\hat{y} = b_0 + b_1x$$' mathematical equation in its own line
+% '\begin{equation}...\end{equation}' multi-line equation
+% '\approx' approximate symbol
+% '\neq' not equal
+% '\bar' mean bar over letter
+% '\pm' plus minus sign 
+% '^{}' superscript
+% '_{}' subscript
+% '\fraq{numerator}{denominator}' fraction
+% '\sqrt[n]{}' square root
+% '\sum_{k=1}^n' sum for 1 through n
+%
+% Insert things from elsewhere:
+% '\input{filename}' inputs the raw (tex) file as a command (e.g., tables and R-Markdown imports)
+% '\include{filename}' includes section on new page (incl. possible auxiliary info)
+% '\includegraphics[settings]{filename}' add a figure or graph
+% '\caption{}' adds a caption to a table or figure
+% '\label{}' labels sections, tables, figures, etc. so that they can be referred to.
+% '\ref{}' refer to a labelled sections, tables, figures, etc.
+% '\begin{enumerate}...\end{enumerate}' numbered list
+% '\begin{itemize}...\end{itemize}' bullet-ed list
+% '\item' item in list section 
+%
+% Symbols:
+% '\&' and sign
+% '\%' percent sign
+% '\_' three dotes
+% '\#' hash symbol
+% ------------------------------------------------------------------
+
+Recent reviews have called for more comprehensive assessment of human experiences and for more longitudinal real-life data, within the psychological sciences more broadly and in migration research in particular \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
+However, while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time.
+
+In essence, the novel extensive longitudinal datasets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
+Yet, past analytical advances have almost exclusively pushed for inferential modeling procedures\footnote{For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).}.
+And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the initial developmental data patterns.
+
+As an example, a recent review of migration experiences has pointed out that despite many complex and dynamic theories, investigations of migrant adaptation have undervalued developmental data --- especially when it comes to the more internal experiences of motivations and emotions \citep[e.g.,][]{Kreienkamp2022d}.
+Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
+To identify which variables are most important in the adaptation of migrants over time, we need methods to break down the data heterogeneity into its core components (in terms of important variables and developments) and we need ways to identify how these core components relate to key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
+
+In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principle component analysis} (3MPCA).
+We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly \citep[e.g.,][]{Monden2015}.
+We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to decompose social psychological experiences.
+This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
+
+To untangle the heterogeneity in real-life migrant adaptations, we will analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
+However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences \citep[e.g.,][]{Kreienkamp2022d}, the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
+To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.
+
+ALTERNATIVE:
+
+Recent reviews have called for more comprehensive assessment of human experiences and for more longitudinal real-life data, within the psychological sciences more broadly and in migration research in particular \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
+However, while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time.
+
+In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principle component analysis} (3MPCA).
+We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly \citep[e.g.,][]{Monden2015}.
+We analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
+However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences \citep[e.g.,][]{Kreienkamp2022d}, the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
+To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.
+
+% Relevance Pragraph
+
+% Problem Illustration Paragraph
+
+% Aim / Solution / Proposal Paragraph:
+
+ALTERNATIVE:
+
+Two important developments, within the psychological sciences in general and in migration research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
+Within the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations \citep[e.g.,][]{Kreienkamp2022d, Ward2019}.
+At the same time prominent figures in the field have called for studies that collect longitudinal \citep[e.g.,][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[e.g.,][]{MacInnis2015, McKeown2017, Dixon2005}.
+
+In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
+However, recent technological developments now allow us to easily collect large amounts of experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}.
+And while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time (i.e., multiple aspects).
+In essence, the new extensive longitudinal data comes with new forms of heterogeneity when jointly considering differences between people, over time, and across variables.
+Yet, past analytical advances have almost exclusively pushed for stationary lagged regression models\footnote{meaning models that assume stable means and variances over time; e.g., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses} or basic trajectory models\footnote{e.g., mixed effects models, spline regression models, and latent growth curve modeling.}.
+And while these two approaches are important for inferential model testing, both approaches ignore the more fundamental task of describing and unraveling the developmental data patterns.  
+And while these two approaches are important for inferential model testing, we still miss methods for the more fundamental task of describing and unraveling the developmental data patterns.
+
+Such descriptive methods are not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
+We we need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+
+We recently conducted three experience sampling studies, following the migration experiences of recent migrants to the Netherlands.
+In this manuscript, we aim to assess the utility of an analytical methods that has recently received increasing attention for simultaneously decomposing/considering person-, variable-, and time heterogeneity.
+The proposed three mode principle component analysis is of particular interest because its approach mirrors the famous data box of participants, variables, and measurement occasions (Catell's data box) for understanding psychological data.
+
+Problem: HOWEVER, ...
+
+* unclear how do deal with this more complex data (many variables, persons, and time points).
+  + many different variables could be important.
+  + how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
+At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[][]{ODonnell2021}.
+* heterogeneity between people, over time, and across variables.
+* **unclear how to identify core/important developments.** Especially across multiple variables at he same time (i.e., multiple aspects).
+* unclear which variables, time scales, and methods are useful in practice.
+* curse of dimensionality: clustering algorithms fail
+
+Solution:
+
+* 3MPCA uses the data cube
+
+little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
+Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
+Simultaneously clustering the person-, experience aspect-, and time level of migration experiences, using a three mode PCA, may allow to identify clusters of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
+
+\section{Section Heading}
+
+\section{The Present Research}
+To be written.
+
+% Methods and Results from RMarkdown render
+%\input{Methods-and-Results}
+\section{Methods}
+To be written.
+
+\section{Results}
+To be written.
+
+\section{Discussion}
+% aims re-iterated
+To be written.
+
+\subsection{Limitations}
+To be written.
+
+\subsection{Implications}
+To be written.
+
+\subsection{Conclusion}
+To be written.
+
+\section{Notes}
+
+\citep[][]{Madley-Dowd2019} [proportion of missing data and imputations]
+
+\citep[e.g., see][]{Ram2014} [Processes Across Multiple Time-Scales]
+
+% Tables
+% Example
+%\input{Tables/descrFullWide}
+
+% Figures
+
 \printbibliography
 
+\appendix
+
+\section{Appendix A: Title}
+\label{app:AppendixTitle}
+To be written.
+
 \end{document}

### 2022-11-02 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

@@
 
-```{r variable names by response type}
+```{r variable names by response type and study}
 varNames %>%
@@
   ) %>%
-  mutate(contactSpecific = recode(.$contactSpecific, 
-                         `-1`="No Interaction Only",
-                         `0`="unspecific",
-                         `1`="Interaction Only")) %>%
-  select(varNam, varGroup, contactSpecific) %>%
-  group_by(varGroup, varNam, contactSpecific) %>%
+  mutate(
+    contactSpecific = recode(
+      .$contactSpecific,
+      `-1` = "No Interaction Only",
+      `0` = "unspecific",
+      `1` = "Interaction Only"
+    ),
+    study1 = ifelse(!is.na(studyS1), 1, 0),
+    study2 = ifelse(!is.na(studyS2), 1, 0),
+    study3 = ifelse(!is.na(studyS3), 1, 0)
+  ) %>% 
+  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
+  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
   summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
+  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
   ungroup %>%
@@
   tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable"), " - ") %>%
+  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
+  mutate(across(c(study1, study2, study3), as.numeric)) %>%
   arrange(concept, unspecific, `Interaction Only`) %>% 
   mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  relocate(starts_with("study"), .after = last_col()) %>%
+  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
   kbl(.,
@@
       digits=2,
-      caption = "Variables by Interaction Types") %>%
+      caption = "Variables by Interaction Types and Study Availability") %>%
+  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
   kable_classic(
@@
   as.data.frame
+
+apply(dtS1Red, 2, sum(is.na(.))/nrow(.))
+
+your_dataframe %>%
+  group_by(Your_Participantid) %>%
+  summarise(across(everything(), funs(100*mean(is.na(.)))))
 
@@
 
+Feature selection types:
+
+-   manual feature and variable selection
+-   wrapper: procedures that add and/or remove predictors to find the optimal combination that maximizes model performance.
+-   filter: relevance of the predictors outside of the predictive models and subsequently model only the predictors that pass some criterion.
+-   dimension reduction for clustering. No selection more compression of redundant variance.
+
+```{r manual feature selection}
+
+```
+
+```{r automated feature selection}
+
+```
+
+```{r dimension reduction}
+raw_data <- featData
+z_data <- featFullImp$featuresImpZ
+scaled_data <- featFullImp$featuresImpZMat
+
+res.pca <- prcomp(scaled_data, scale = FALSE)
+fviz_eig(res.pca, addlabels = TRUE)
+
+var <- get_pca_var(res.pca)
+corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
+fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
+fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)
+
+res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
+# Description of dimension 1
+res.desc$Dim.1
+
+plot_ly(
+  res.pca$rotation %>% as.data.frame,
+  x = ~ PC1,
+  y = ~ PC2,
+  z = ~ PC3,
+  size = 2,
+  type = 'scatter3d',
+  mode = 'markers',
+  text = ~ paste('Variable:', rownames(res.pca$rotation))
+) %>%
+  layout(title = 'Variables in reduced space')
+plot_ly(
+  res.pca$x %>% as.data.frame,
+  x = ~ PC1,
+  y = ~ PC2,
+  z = ~ PC3,
+  size = 2,
+  type = 'scatter3d',
+  mode = 'markers',
+  text = ~ paste('Participant:', rownames(res.pca$x))
+) %>%
+  layout(title = 'Participants in reduced space')
+
+pca.out <- as.data.frame(res.pca$x[,1:3])
+
+# load libraries
+library(factoextra)
+library(NbClust)
+# elbow method
+fviz_nbclust(pca.out, kmeans, method = "wss") +
+  geom_vline(xintercept = 2, linetype = 2)+
+  labs(subtitle = "Elbow method")
+# Silhouette method
+fviz_nbclust(pca.out, kmeans, method = "silhouette")+
+  labs(subtitle = "Silhouette method")
+# Gap statistic [takes for ever]
+# nboot = 50 to keep the function speedy.
+# recommended value: nboot= 500 for your analysis.
+# Use verbose = FALSE to hide computing progression.
+# set.seed(123)
+# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
+#   labs(subtitle = "Gap statistic method")
+
+kmeansOut <- kmeans(pca.out, centers = 3, nstart = 100)
+fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())
+
+# MCLUST
+library(mclust)
+BIC <- mclustBIC(pca.out)
+plot(BIC)
+summary(BIC)
+mod1 <- Mclust(pca.out)
+summary(mod1, parameters = FALSE)
+
+# DBSCAN
+library(dbscan)
+library(fpc)
+
+# to plot the eps values
+eps_plot = kNNdistplot(pca.out, k=3)
+# to draw an optimum line
+eps_plot %>% abline(h = 3.5, lty = 2)
+
+potentialMinPts <- round(SciViews::ln(nrow(pca.out)))
+eps <- 3.5
+minPts <- 2
+
+res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0)
+res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
+
+resEqual.check <- all(res.fpc$cluster == res.db$cluster)
+nClustDBSCAN <- length(unique(res.fpc$cluster))-1
+
+fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
+
+```
+
 ## Clustering
@@
 
+```{r time-series linear trends by variable and cluster}
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., raw_data) %>%
+  select(
+    ID,
+    TIDnum,
+    cluster,
+    any_of(varNamS123)
+  ) %>%
+  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
+  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
+  geom_smooth(aes(group=cluster)) +
+  #stat_summary(fun=mean, geom="line") +
+  facet_wrap(~ variable, scales = "free_y") +
+  labs(
+    x = "Time ID",
+    y = "Response (Mean Cluster Response Bold)",
+    color = "Cluster"
+  ) +
+  theme_Publication()
+```
+
 #### Distribution-based

### 2022-11-15 TitlePage.tex
Updates from Overleaf

@@
 
-% add reference library file
-\addbibresource{references.bib}
+% custom open science badge height
+\newlength{\badgeheight}
+\setlength{\badgeheight}{1em}
 
 % add reference library file
-\addbibresource{references.bib}
+\addbibresource{referencesZotero.bib}
 
 % Title and header
-\title{Migration Experience Trajectories: Subtitle TBD}
+\title{Migration Experience Trajectories (Working Title)}
 \shorttitle{Migration Experience Trajectories}
@@
 \author[1,3]{Peter de Jonge}
-\author[2,3]{Rei Tendeiro-Monden}
+\author[2,3]{Rei Monden}
 \affiliation{\hfill}
@@
 \affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
-\affil[3]{Author order TBD [currently in first name alphabetical order]}
+\affil[3]{Author order TBD (currently mostly in first name alphabetical order)}
 
@@
    \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}\\
+   \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
    \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
-   \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
    \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
@@
 Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
-E-mail: j.kreienkamp@rug.nl}
+E-mail: \href{mailto:j.kreienkamp@rug.nl}{j.kreienkamp@rug.nl}
+
+\vspace{1em}
+\noindent\textit{Open Science Practices [?]:}\\
+\noindent \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registrationplus-color.png}} Preregistration+, 
+\href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
+\href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
+\href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
+\href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
+
+}
 
@@
 % add title information (incl. title page and abstract)
-\maketitle
+%\maketitle
+
+\makeatletter
+\vspace*{3em}
+\begin{center}
+\noindent \textbf\@title
+\vspace{1em}
+\noindent \@author
+\end{center}
+\vspace{5em}
+\begin{center}%
+    \textbf\acksname%
+\end{center}%
+\vspace{-1em}
+\protect\raggedright
+\setlength{\parindent}{0.5in}
+\indent\par\@acks%
 

### 2022-11-15 manuscript.tex

@@
 % \usepackage{longtable}
-\usepackage{amsmath}
+\usepackage{amsmath, nccmath}
+\usepackage{mdframed}
 \usepackage{ntheorem}
@@
 
+% ref labels
+\newcommand{\fgrref}[2][]{\hyperref[#2]{Figure \ref*{#2}#1}}
+\newcommand{\tblref}[2][]{\hyperref[#2]{Table \ref*{#2}#1}}
+\newcommand{\appref}[2][]{\hyperref[#2]{Appendix \ref*{#2}#1}}
+
 % custom open science badge height
@@
 % Title and header
-\title{Migration Experience Trajectories (Working Title)}
+\title{Describe and Explore: Using feature-based time-series clustering to understand modern intensive longitudinal data (Working Title)}
 \shorttitle{Migration Experience Trajectories}
@@
 \abstract{
-To be written.
+Abstract to be written.
 
@@
     \vspace{1em}
-    \noindent \textbf{Open Science Practices:}\\
-    \noindent \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registration-color.png}} Preregistration+, 
+    \textit{Open Science Practices [?]:}
+    \noindent \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registration-color.png}} Preregistration, 
     \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data, 
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data, \break
     \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
@@
 
-Recent reviews have called for more comprehensive assessment of human experiences and for more longitudinal real-life data, within the psychological sciences more broadly and in migration research in particular \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
-However, while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time.
+\newlength{\mdfmar}
+\setlength{\mdfmar}{1.5em}
+\mdfdefinestyle{mdfbox}{
+    innerleftmargin = +\mdfmar, %
+    innerrightmargin = +2\mdfmar, 
+    innertopmargin = +0\mdfmar, 
+    innerbottommargin = +1.5\mdfmar, 
+    skipabove = -12pt
+}
 
-In essence, the novel extensive longitudinal datasets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for inferential modeling procedures\footnote{For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).}.
-And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the initial developmental data patterns.
+\begin{mdframed}[style=mdfbox]
+\noindent\center\textit{Argument Structure Introduction}:
+\begin{itemize}[noitemsep]
+    \item\ [Relevance]: increase in number and variety of ESM data
+    \item\ [Problem]: no descriptive / understand structure analyses
+    \item\ [Proposal]: feature based clusterting = flexible +  transparent
+    \item\ [Section]: new type of ESM data and issues for analysis:
+    \begin{itemize}[noitemsep]
+        \item incomplete data (complete data models vs. structural missingness)
+        \item non-equidistant
+        \item non-stationary
+        \item changes/trends on different time scales
+        \item interest in multivariate developments
+        \item dimensionality curse
+    \end{itemize}
+    \item\ [Section]: time-series clustering
+    \begin{itemize}[noitemsep]
+        \item shape-based
+        \item model-based
+        \item \textbf{feature-based}
+    \end{itemize}
+    \item\ [Section]: The present Research
+\end{itemize}
+\end{mdframed}
 
-As an example, a recent review of migration experiences has pointed out that despite many complex and dynamic theories, investigations of migrant adaptation have undervalued developmental data --- especially when it comes to the more internal experiences of motivations and emotions \citep[e.g.,][]{Kreienkamp2022d}.
-Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
-To identify which variables are most important in the adaptation of migrants over time, we need methods to break down the data heterogeneity into its core components (in terms of important variables and developments) and we need ways to identify how these core components relate to key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
-There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
-
-In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principle component analysis} (3MPCA).
-We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly \citep[e.g.,][]{Monden2015}.
-We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to decompose social psychological experiences.
-This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
-
-To untangle the heterogeneity in real-life migrant adaptations, we will analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
-All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
-However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences \citep[e.g.,][]{Kreienkamp2022d}, the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
-To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.
-
-ALTERNATIVE:
-
-Recent reviews have called for more comprehensive assessment of human experiences and for more longitudinal real-life data, within the psychological sciences more broadly and in migration research in particular \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
-However, while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time.
-
-In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principle component analysis} (3MPCA).
-We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly \citep[e.g.,][]{Monden2015}.
-We analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
-All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
-However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences \citep[e.g.,][]{Kreienkamp2022d}, the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
-To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.
-
-% Relevance Pragraph
+% Relevance Pragraph and issue
+Recent years have seen a striking increase in the number and variety of research studies using experience sampling data \citep[][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}. 
+With the increased availability of technologies to easily collect large amounts of experience sampling data using mobile devices \citep[e.g.,][]{Keil2020} and web-based applications \citep[e.g.,][]{Arslan2020}, the experience-sampling method (ESM) has been applied to a host of new types of psychological data.
+One example of this phenomenon has been the cultural adaptation literature, where researchers have started using ESM data to follow the daily interactions of migrants with the local majority \citep[e.g.,][]{Kreienkamp2022b}.
+This move toward a broader application of the experience-sampling method, however, stands in stark contrast with recent developments within the psychometric literature on ESM analyses, which have become more specialized and model-focused.
+While generally speaking analytical methods for ESM data have become more readily available \citep[e.g.,][]{ODonnell2021}, flexible descriptive and exploratory analyses have largely been neglected and remain crucially understudied within the experience sampling literature.
 
 % Problem Illustration Paragraph
+Importantly, the understudied `describe and explore' analyses are not only important for contextualizing inferential model tests, but are powerful methods for theoretical and applied users in their own regard.
+Extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables.
+Data-driven models not only describe and summarize this complex data but can help us understand the data by uncovering patterns that might not be detectable otherwise.
+Returning to the case of migration experiences, understanding how people differ in their migration trajectories can be crucial in understanding adaptive and maladaptive patterns.
+By identifying and contextualizing distinct trajectories, we might for example determine which situation- or person-specific differences discern positive from negative developments.
+Outcomes from such an approach might inform both practical early-warning signals as well as academic theory-building efforts.
+There is, thus, a clear need to assess the utility of data-driven/bottom-up analysis procedures that are flexible enough to address the growing variety of intensive longitudinal psychological data.
 
 % Aim / Solution / Proposal Paragraph:
+In this article, we propose to look at recent developments within the wider time-series clustering literature to identify an exploratory data analysis that is flexible enough to apply to the growing variety of ESM data.
+In particular, we propose that feature-based clustering is ideally suited because it offers a flexible variety of validated and intuitive components while also allowing for adequate control by the researcher.
+To showcase the utility of this technique, we apply feature-based clustering to a set of three ESM studies following recent migrants in their daily interactions and relations with the cultural majority group.
 
-ALTERNATIVE:
+\section{Modern ESM Data}
+Two important developments within the psychological sciences have been (1) a shift towards more comprehensive assessments of psychological phenomena and (2) a growing focus on intensive longitudinal real-life data.
+As an example, within the cultural adaptation literature prominent reviews have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time conceptual works have pushed for more comprehensive assessments of human experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
+Such a multivariate, experience-sampling approach allows researchers to test theories more exhaustively and helps build a more embedded understanding of migration processes.
+However, such data also deviates from past intensive longitudinal data sets that have focused on psychological states that are more homogeneous in their developmental characteristics.
+In particular, the new applications of ESM tend to bring new challenges for data dimensionality, -missingness, and time scales.
 
-Two important developments, within the psychological sciences in general and in migration research in particular, have been (1) a shift towards more comprehensive assessment of human experiences and (2) a growing focus on longitudinal real-life data.
-Within the migration research field, recent reviews have been pushing for models that take into account affects, behaviors, cognitions, and desires when it comes to migrant adaptations \citep[e.g.,][]{Kreienkamp2022d, Ward2019}.
-At the same time prominent figures in the field have called for studies that collect longitudinal \citep[e.g.,][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[e.g.,][]{MacInnis2015, McKeown2017, Dixon2005}.
+Concerning dimensionality issues, especially more abstract psychological experiences often need a wider variety of measurements to be captured adequately.
+Today, few clinical conditions are captured with a single symptom measure, emotions are rarely assessed in isolation, and as introduced earlier cultural adaptation has at least four psychological aspects.
+This also means that modern analysis techniques increasingly need be able to accommodate this increased focus on multivariate developments.
+The recent rise in psychological network models has highlighted this need for more multivariate outcome measurements (REFERENCE).
+Additionally, the increased number of variables also heightens the computational load for model estimations and especially when the number of variables and measurements compounds models may not converge --- an issue sometimes referred to as the dimensionality curse.
+A modern descriptive and exploratory analysis technique should consequently be able to summarize and structure the multivariate phenomena without running into computational load issues.
 
-In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
-However, recent technological developments now allow us to easily collect large amounts of experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}.
-And while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time (i.e., multiple aspects).
-In essence, the new extensive longitudinal data comes with new forms of heterogeneity when jointly considering differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for stationary lagged regression models\footnote{meaning models that assume stable means and variances over time; e.g., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses} or basic trajectory models\footnote{e.g., mixed effects models, spline regression models, and latent growth curve modeling.}.
-And while these two approaches are important for inferential model testing, both approaches ignore the more fundamental task of describing and unraveling the developmental data patterns.  
-And while these two approaches are important for inferential model testing, we still miss methods for the more fundamental task of describing and unraveling the developmental data patterns.
+The push for more contextualized models also presents new issues with missing data.
+More event-based assessments tend to encounter issues when models require time intervals between measurements to be identical (i.e., equidistant measurements) and an increased focus on context-specific phenomena has lead to issues when data is only available under certain conditions, including follow-up questions (i.e., structurally missing data).
+Smaller issues of non-equidistant data can be avoided with transformations (e.g., dynamic time warping, REFERENCE) or newer modeling procedures (e.g., continuous-time models, REFERENCE) but it's prevalence is indicative of the structure of real-world issues researchers seek to investigate.
+Structural missingness remains a much more strenuous challenge.
+Many of the most commonly used models need complete data (including, ...) and structurally missing data cannot be imputed as it logically does not exist (REFERENCE).
+This type of missingness becomes a colossal issue when researchers are interested in how two things co-develop and one of the concepts has structural missingness under certain conditions.
+In the example of the migration experiences, researchers might for example be interested in how well-being develops in relations the quality of interactions with the majority group.
+Because rating of interaction quality are only possible when an interaction actually happened, one cannot impute interaction quality ratings for measurements that did not include an interaction.
+The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data (REFERENCE) --- neither option suits a research question that includes a developmental relationship between the two variables.
+In short, adequate descriptive approaches should be able to deal with non-equidistant and structurally missing data in order to address modern ESM data.
 
-Such descriptive methods are not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
-We we need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+Finally, both the push for more multivariate conceptualizations and contextualized measurements have highlighted issues with how temporal developments are conceptualized.
+Firstly, researchers are increasingly interested in how mean levels of variables change over time (i.e., non-stationary), such trend data is often assumed to be non-linear, and with the increased focus on more diverse conceptualizations researchers are interested in relationships between variables that change on fundamentally different time scales.
+Many of the commonly used analyses for ESM data are stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).
+And even fewer analysis techniques allow for substantially different time scales between variables.
+A modern descriptive and exploratory methodology should thus be able to flexibly capture and comparable non-stationary and non-linear trends, also across different time scales.
 
-We recently conducted three experience sampling studies, following the migration experiences of recent migrants to the Netherlands.
-In this manuscript, we aim to assess the utility of an analytical methods that has recently received increasing attention for simultaneously decomposing/considering person-, variable-, and time heterogeneity.
-The proposed three mode principle component analysis is of particular interest because its approach mirrors the famous data box of participants, variables, and measurement occasions (Catell's data box) for understanding psychological data.
+\section{Clustering}
+One promising approach to describing and exploring time-series data has been the idea of structuring the data using clustering analyses.
+Clustering seeks to organize data into groups that are maximally different from one another while minimizing differences within the groups.
+Time-series clustering has been extremely common in other fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, or energy and finance patterns \citep{Aghabozorgi2015}.
+But also for psychological data time-series clustering is not entirely novel \citep[e.g.,][]{ernst2021}.
+However, most clustering introduced for psychological data has faced the dimensionality, -missingness, and time scale issues described above.
 
-Problem: HOWEVER, ...
-
-* unclear how do deal with this more complex data (many variables, persons, and time points).
-  + many different variables could be important.
-  + how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
-At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[][]{ODonnell2021}.
-* heterogeneity between people, over time, and across variables.
-* **unclear how to identify core/important developments.** Especially across multiple variables at he same time (i.e., multiple aspects).
-* unclear which variables, time scales, and methods are useful in practice.
-* curse of dimensionality: clustering algorithms fail
-
-Solution:
-
-* 3MPCA uses the data cube
-
-little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
-Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
-Simultaneously clustering the person-, experience aspect-, and time level of migration experiences, using a three mode PCA, may allow to identify clusters of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
-
-\section{Section Heading}
+Many of the common time-series clustering algorithms used today have been developed within the machine learning literature and have commonly been separated into three different approaches: (1) shape-based, (2) model-based, and (3) feature-based clustering \citep{hautamaki2008, liao2005}.
+Shape-based clustering analyses use the raw time series and seeks to match time series by non-linearly stretching and contracting the time axes.
+The other two approaches do not use the the raw time series directly but convert the data first.
+The model-based clustering has been the most-commonly used method within the psychological literature and clusters the parameters of a parametric model fitted for each participant \citep[e.g.,][]{ernst2021}.
+The flexibility of such clustering is restricted to the assumptions of the model used, so that stationary models, for example, do not capture non-stationary data well.
+Finally, feature-based clustering analyses are the most flexible, as they allow users to summarize the time series data with a wide variety of features that capture the data characteristics of interest.
+The participants are then clustered on basis of the extracted features.
+To showcase the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
+Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time-series, but also linear and non-linear trends, inertia related features such as auto-correlations, or the stability of the time series (e.g., mean absolute change).
+Given the flexibility of feature-based approach, users should be able to apply the analysis across multivariate data, feature extraction should mitigate issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
 
 \section{The Present Research}
-To be written.
 
@@
 
+"The goal of clustering is to identify structure in an unlabeled data set by objectively organizing data into homogeneous groups where the within-group-object dissimilarity is minimized and the between-group-object dissimilarity is maximized." --- \citep[][p.1857]{liao2005}
+
+\section{Draft Graveyard (for scrapping only):}
+
+Extensive longitudinal data sets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
+Yet, past analytical advances have almost exclusively pushed for inferential modeling procedures\footnote{For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).}.
+And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the raw/complex developmental data patterns.
+
+As an example, a recent review of migration experiences has pointed out that despite many complex and dynamic theories, investigations of migrant adaptation have undervalued developmental data --- especially when it comes to the more internal experiences of motivations and emotions \citep[e.g.,][]{Kreienkamp2022d}.
+Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
+To identify which variables are most important in the adaptation of migrants over time, we need methods to break down the data heterogeneity into its core components (in terms of important variables and developments) and we need ways to identify how these core components relate to key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
+
+We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to cluster social psychological trajectories.
+This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
+
+To untangle the heterogeneity in real-life migrant adaptations, we will analyze data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
+All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
+However, given that past investigations of psychological acculturation have under-explored the crucial aspect of motivational experiences \citep[e.g.,][]{Kreienkamp2022d}, the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
+To make full use of all three data sets and to guide future use of clustering procedures with extensive longitudinal psychological data, we will analyze the data in a descriptive and step-wise manner, where we consider the data jointly (and will then assess potential differences between data- and variable sets).
+
+ALTERNATIVE:
+
+In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
+However, recent technological developments now allow us to easily collect large amounts of experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}.
+And while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time (i.e., multiple aspects).
+In essence, the new extensive longitudinal data comes with new forms of heterogeneity when jointly considering differences between people, over time, and across variables.
+Yet, past analytical advances have almost exclusively pushed for stationary lagged regression models\footnote{meaning models that assume stable means and variances over time; e.g., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses} or basic trajectory models\footnote{e.g., mixed effects models, spline regression models, and latent growth curve modeling.}.
+And while these two approaches are important for inferential model testing, we still miss methods for the more fundamental task of describing and unraveling the developmental data patterns.
+
+Such descriptive methods are not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
+We need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
+
+Problem: HOWEVER, ...
+
+\begin{itemize}
+    \item unclear how do deal with this more complex data (many variables, persons, and time points).
+    \begin{itemize}
+        \item many different variables could be important.
+        \item how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
+    \end{itemize}
+    \item \textbf{unclear how to identify core/important developments.} Especially across multiple variables at he same time (i.e., multiple aspects).
+    \item unclear which variables, time scales, and methods are useful in practice.
+    \item curse of dimensionality: clustering algorithms fail 
+\end{itemize}
+
+Solution:
+
+feature-based clustering
+
+little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
+Clustering participants across variables and time-series features may allow to identify groups of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
+
 % Tables
@@
 
+\begin{figure}
+  \caption{Scopus ESM Development}
+  \label{fig:ScopusEsm}
+  \centering\includegraphics[width=\textwidth]{figures/Scopus-ESM-Development.png}
+  \caption*{Note: \\
+  Search Terms: `( "ecological momentary assessment"  OR  "experience sampling"  OR  "intensive longitudinal")' in Title, Abstract, or Keyword.}
+\end{figure}
+
 \printbibliography
@@
 
-\section{Appendix A: Title}
+\section{Placeholder Title}
 \label{app:AppendixTitle}
-To be written.
+To be written, if necessary.
 

### 2022-11-21 methods-and-results.Rmd
m&r init

--- 2022-10-02 Supplemental-Material-A-Annotated-Analysis.Rmd
@@
 ---
-title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
-subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
-author:
-- Jannis Kreienkamp^1,3^
-- Kai Epstude^1,3^
-- Laura F. Bringmann^1,3^
-- Maximilian Agostini^1,3^
-- Peter de Jonge^1,3^
-- Rei Monden^2,3^
-- ^1^University of Groningen, Department of Psychology
-- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
-- ^3^Author order TBD [currently in first name alphabetical order]
-- "Author Information:"
-- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
-E-mail: j.kreienkamp@rug.nl"
-- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
-- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
-- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
-date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
-output: 
-  bookdown::html_document2:
-    fig_caption: yes
-    md_extensions: +footnotes
-    code_folding: hide
-    mathjax: default
-    theme: yeti
-    toc: yes
-    toc_float: yes
-    number_sections: false
-    css: style.css
-    includes:
-      in_header: "_includes/head-custom-rmd.html" 
+output: latex_fragment
 editor_options:
@@
 csl: apa.csl
-header-includes:
-   - \usepackage{amsmath, nccmath}
 ---
-
-```{=html}
-<style type="text/css">
-.main-container {
-  max-width: 1300px;
-  margin-left: auto;
-  margin-right: auto;
-}
-.table {
-  margin-left:auto; 
-  margin-right:auto;
-}
-</style>
-```
 
@@
 # knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings
-knitr::knit_hooks$set(
-  error = function(x, options) {
-    paste('\n\n<div class="alert alert-danger">',
-      gsub("##", "\n", gsub("^##\ Error", "**Error**", x)),
-      "</div>",
-      sep = "\n"
-    )
-  },
-  warning = function(x, options) {
-    paste('\n\n<div class="alert alert-warning">',
-      gsub("##", "\n", gsub("^##\ Warning:", "**Warning**", x)),
-      "</div>",
-      sep = "\n"
-    )
-  },
-  message = function(x, options) {
-    paste('\n\n<div class="alert alert-info">',
-      gsub("##", "\n", x),
-      "</div>",
-      sep = "\n"
-    )
-  }
-)
-htmltools::tagList(rmarkdown::html_dependency_font_awesome())
 
@@
 
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-<i class="fas fa-exclamation-circle"></i> Note.
-Boxplots display the interquartile range (IQR, center box), and the whiskers extend 1.5\*IQR from the lower and upper hinge.
-The white point indicates the mean and the white center line indicates the median.
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-# **Data Preparation**
-
-In an initial preparatory step, we import the data into the R project environment and prepare the variables for further processing and later analyses.
-
-## Data Import
-
-The data were collected using two different survey tools.
-For the study with sojourners (Study 1: worker) we used the survey platform Qualtrics XM, whereas the studies with international students (Study 2: student), and the international medical professionals (Study 3: medical) were conducted using the survey framework FormR.
-This means that the datasets had inconsistent file formats and naming conventions.
-For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*').
-For the two other studies, we import the raw csv files from their respective folders.
-
-```{r formrImport}
+```{r import data, include=FALSE}
 # Load variable name lookup table
@@
 
-# rename function
-reNam <- function(data = NA, names = varNames, study = "S1", survey = "pre") {
-  # for testing
-  # data = dtS1$raw.pre
-  # names = varNames
-  # study = "S1"
-  # survey = "pre"
-
-  var <- paste0("survey", study)
-  varNamOld <- paste0("varNam", study)
-
-  nameTbl <- names %>%
-    filter((!!sym(var)) == survey) %>%
-    select(varNam, varNamOld = (!!sym(varNamOld)))
-
-  data.table::setnames(data, as.character(nameTbl$varNamOld), as.character(nameTbl$varNam), skip_absent=TRUE)
-
-  return(data)
-}
-
-# workers
-# initial data cleaning was done in SPSS (syntax files are available in "")
-dtS1 <- list(
-  raw.pre = read_spss("data/S1_Workers/processed/cleaned/MT - Pre-Measure - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "pre"),
-  raw.post = read_spss("data/S1_Workers/processed/cleaned/MT - Post-Measure - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "post"),
-  raw.morning = read_spss("data/S1_Workers/processed/cleaned/MT - Morning - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "daily"),
-  raw.afternoon = read_spss("data/S1_Workers/processed/cleaned/MT - Afternoon - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "daily")
-)
-
-# students
-dtS2 <- list(
-  raw.pre = read.csv(file = "data/S2_Students/raw/AOTS_Pre.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "pre"),
-  raw.post = read.csv(file = "data/S2_Students/raw/AOTS_Post.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "post"),
-  raw.daily = read.csv(file = "data/S2_Students/raw/AOTS_Daily.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "daily")
-)
-
-# young medical professionals
-dtS3 <- list(
-  raw.eligibility = read.csv("data/S3_Medical/raw/AOTM_Eligibility.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"), # works but should be 'survey = "pre_elig"'
-  raw.pre = read.csv("data/S3_Medical/raw/AOTM_Pre.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"),
-  raw.post = read.csv("data/S3_Medical/raw/AOTM_Post.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "post"),
-  raw.daily = read.csv("data/S3_Medical/raw/AOTM_Daily.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "daily")
-)
-
-# # FROM INITIAL VAR NAME EXPORT [LEGACY]
-# # export var names
-# S1Nam <- rbind(
-#   data.frame(
-#     study = "S1",
-#     survey = "pre",
-#     varNam = names(dtS1$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S1",
-#     survey = "daily",
-#     varNam = names(dtS1$raw.afternoon)
-#   ),
-#   data.frame(
-#     study = "S1",
-#     survey = "post",
-#     varNam = names(dtS1$raw.post)
-#   )
-# )
-# S2Nam <- rbind(
-#   data.frame(
-#     study = "S2",
-#     survey = "pre",
-#     varNam = names(dtS2$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S2",
-#     survey = "daily",
-#     varNam = names(dtS2$raw.daily)
-#   ),
-#   data.frame(
-#     study = "S2",
-#     survey = "post",
-#     varNam = names(dtS2$raw.post)
-#   )
-# )
-# S3Nam <- rbind(
-#   data.frame(
-#     study = "S3",
-#     survey = "pre_elig",
-#     varNam = names(dtS3$raw.eligibility)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "pre_entry",
-#     varNam = names(dtS3$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "daily",
-#     varNam = names(dtS3$raw.daily)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "post",
-#     varNam = names(dtS3$raw.post)
-#   )
-# )
-# write.csv(
-#   x = rbind(
-#     S1Nam,
-#     S2Nam,
-#     S3Nam
-#   ),
-#   file = "preregistration/varNames.csv"
-# )
+# import full datasets
+load("data/s123Full.RData")
+
+# import reduced datasets
+load("data/s123Red.RData")
 ```
 
-## Data Cleaning & Data Exclusions
-
-### Study 1
-
-For the sojourner sample data was collected in four separate surveys: (1) the pre-measurement, (2) the daily morning survey, (3) the daily afternoon survey, as well as (4) a post-measurement.
-We combine the four individual surveys into one cohesive dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We then format the time and date variables and add person- and measurement indices (for easy and meaningful addressing of the data).
-We also exclude our own test data.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS1*' list.
-
-```{r cleanWorker}
-# Create reduced data sets for morning and afternoon
-dat.mo <- dtS1$raw.morning %>%
-  select(-starts_with("t_"))
-#setdiff(names(dtS1$raw.morning), names(dat.mo))
-dat.mo$daytime <- "morning"
-
-dat.af <- dtS1$raw.afternoon %>%
-  select(-starts_with("t_"))
-dat.af$daytime <- "afternoon"
-
-# merge morning and afternoon measurements with indicator [+ clean up]
-daily.dat <- plyr::rbind.fill(dat.mo, dat.af)
-daily.dat <- daily.dat[daily.dat$last_outside_referrer != 55951, ]
-dtS1$daily <- daily.dat
-rm(dat.mo, dat.af, daily.dat)
-
-# reduced data set for pre measurement
-dat.pre.red <- dtS1$raw.pre %>%
-  select(-starts_with("t_"))
-names(dat.pre.red) <- paste(names(dat.pre.red), "pre", sep = ".")
-
-# merge with daily data [+ clean up]
-df.pre <- merge(
-  x = dtS1$daily,
-  y = dat.pre.red,
-  by.x = "last_outside_referrer",
-  by.y = "platformId.pre",
-  all = T
-)
-
-# adjust duplicate names to fit to indicate daily or pre measurement
-names(df.pre) <- gsub("[[:punct:]]x", ".daily", names(df.pre))
-names(df.pre) <- gsub("[[:punct:]]y", ".pre", names(df.pre))
-
-# reduced data set for post-measurement
-dat.post.red <- dtS1$raw.post %>%
-  select(-starts_with("t_"))
-names(dat.post.red) <- paste(names(dat.post.red), "post", sep = ".")
-
-# merge post measurement with pre- and daily data
-df <- merge(
-  x = df.pre,
-  y = dat.post.red,
-  by.x = "last_outside_referrer",
-  by.y = "last_outside_referrer.post",
-  all = T
-)
-
-# adjust duplicate names to indicate pre or post
-names(df) <- gsub("[[:punct:]]x", ".pre", names(df))
-names(df) <- gsub("[[:punct:]]y", ".post", names(df))
-
-# add to list
-dtS1$combined <- df
-rm(df, df.pre, dat.post.red, dat.pre.red)
-
-# create data frame with cleaned data
-df <- dtS1$combined %>%
-  filter(
-    Finished.pre == 1,
-    Finished == 1,
-    !is.na(last_outside_referrer)
-  )
-
-# add running number as measurement ID within participants
-df$measureID <- data.table::rowidv(df, cols = c("last_outside_referrer"))
-
-dtS1$clean <- df
-
-# clean up
-rm(df)
-
-# Export reduced Data
-# write.csv(dtS1$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
-# save(dtS1$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
-```
-
-### Study 2
-
-For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement).
-We also reformat missing values and format core ID variables.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS2*' list.
-
-```{r cleanStudents}
-# our own test IDs
-ownIDs <- c(
-  "beautifulLionfishXXXR5rcgVBzGu8hPvOqrK8UBJBw4owvi9nfRFSFu3lMzYhE",
-  "niceDogoXXXmB8JI5SFu78SF3DVof84mGUPPNUr14p2HYFTtp31a6D1OwAzM6F-K",
-  "amusedQuailXXXmhuc_fpTp8vPkMwDH1BzjaH1d1kHSO1bsPEfsnaEYk4WeVBfPi",
-  "juwGAbtXX0_1kmZtSVqKh3PGaHOICqUyU4iBkrT3nDsI_uifuD1gzKcZerxaM5FL"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS2$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS2$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS2$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = FALSE
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = FALSE
-)
-
-# add to list
-dtS2$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-### Study 3
-
-For the medical professionals sample data was, again, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe.
-We exclude our own test data.
-We also reformat missing values and format core ID variables.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS3*' list.
-
-```{r cleanMedical}
-# our own test IDs
-ownIDs <- c(
-  "test_LeonieXXXSklxecPLW0-FBPM4796o3pUwUhAY5jb9KGw8jQsKxWmGpa1Jiy", 
-  "test_MaxXXXtOp_5dTNefIq0yKXtXt2IN6eEKxeHoPY9mlyvdsqPpLp1B0NGg4UL",
-  "test_JannisXXXBsNqk62fOpX6chbd2tMWPptUdjjnhAqnQ3uBqckZ7gLIEoPlfZ",
-  "quaintLeopardCatXXXAJ9cfSj-_SZLwNwMDxv_xv_iyr1Bg5YFLTlYdrjW0UXZY",
-  "blue-eyedIndianElephantXXXLf5zPMpQCDGS3umFzIj-YVky7ivTItvvozW49m"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS3$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS3$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>% 
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS3$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = F
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = F
-)
-
-# add to list
-dtS3$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-## Calculate needed transformations
-
-### Study 1
-
-For the worker sample, the data transformation stage had three main aims:
-
-1.
-We first corrected time indicators within the surveys.
-In some cases participants completed their daily diary surveys for the afternoon after midnight.
-In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsWorkers}
-df <- dtS1$clean
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S")
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    PID = as.numeric(factor(last_outside_referrer))
-  )
-
-# Time and Date Variables
-df <- df %>%
-  mutate(
-    TID = measureID - 1,
-    # time ID with t0 = 0 for meaningfull intercept interpretations
-    date = substr(created, 1, 10),
-    # awkward way of extracting date (best converted to )
-    time = substr(created, 12, 19),
-    # awkward way of extracting time
-    daynum = as.numeric(factor(dateQualtrics)),
-    # all days as numeric for ordering
-    daycor = ifelse(
-      daytime == "morning" &
-        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
-        daytime == "afternoon" &
-          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
-      daynum - 1,
-      daynum
-    ),
-    daycorTest = ifelse(
-      daytime == "morning" &
-        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
-        daytime == "afternoon" &
-          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
-      1,
-      0
-    ),
-    # correctly identify which date the questionnaire is about
-    daycor.lead = sprintf("%02d", daycor),
-    daytime.lt = ifelse(daytime == "morning", "a", "b"),
-    # morning / afternoon to a / b
-    day_time = paste(daycor.lead, daytime.lt, sep = "_"),
-    # combine day id with morning / afternoon
-    ResponseId = as.numeric(factor(day_time)),
-    # day and time identifier as numeric id
-    SubTime = chron::times(time.0),
-    time.daily = as.character(time),
-    PPDate = as.Date(df$dateQualtrics),
-    number = replace_na(ContactNum, 0),
-    NonDutchNum = replace_na(NonDutchNum, 0)
-  )
-
-# remove seconds from afternoon time
-df$SubTime[df$daytime == "afternoon"] <- paste0(substring(as.character(df$time.0[df$daytime == "afternoon"]), 4, 8), ":00")
-df$time.daily[df$daytime == "afternoon" &
-  !is.na(df$time.daily != "<NA>")] <- paste0(substring(as.character(df$time.daily[df$daytime == "afternoon" &
-  !is.na(df$time.daily != "<NA>")]), 4, 8), ":00")
-
-# Correct morning / afternoon date where survey was collected the day after to indicate the correct date that was targeted
-df$PPDate[df$SubTime < "11:50:00" &
-  df$daytime == "morning"] <- df$PPDate[df$SubTime < "11:50:00" &
-  df$daytime == "morning"] - 1
-df$PPDate[df$SubTime < "18:50:00" &
-  df$daytime == "afternoon"] <- df$PPDate[df$SubTime < "18:50:00" &
-  df$daytime == "afternoon"] - 1
-
-# Make time IDs consistent with later studies
-df$TID <- paste(df$PPDate, str_to_title(df$daytime))
-df$TIDnum <- as.numeric(factor(df$TID %>% gsub(" Morning", "-A", .) %>% gsub(" Afternoon", "-B", .)))
-
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = FALSE),
-    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum),
-    alertness = sum(alertness1, alertness2, na.rm = TRUE),
-    calmness = sum(calmness1, calmness2, na.rm = TRUE),
-    valence = sum(valence1, valence2, na.rm = TRUE)
-  ) %>%
-  ungroup()
-
-# Need scales
-df$keyMotiveFulfilled <- rowSums(df[, c("KeyNeedFulfillment", "DaytimeNeedFulfillment")], na.rm = T)
-df$autonomy.daily.all <- rowSums(df[, c("autonomy_Int", "autonomy_NoInt")], na.rm = T)
-df$competence.daily.all <- rowSums(df[, c("competence_Int", "competence_NoInt")], na.rm = T)
-# cor(df$relatednessOther, df$relatedness_self_1,use="complete.obs")
-df$relatedness.daily.all <- rowMeans(df[, c(
-  "relatednessOther",
-  "relatednessSelf",
-  "relatednessNoInteraction"
-)], na.rm = T)
-
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatedness <- rowMeans(df[, c("relatednessOther", "relatednessSelf")], na.rm = T)
-
-df$autonomy <- df$autonomy.daily.all
-df$competence <- df$competence.daily.all
-df$relatedness <- df$relatedness.daily.all
-
+```{r Preparation of Variable Names}
 varNamIndicesS1 <- c(
@@
 
-df$roommate.pre_calc <- df %>%
-  select(
-    starts_with("roommate")
-  ) %>% 
-  mutate_all(as_factor) %>%
-  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>%
-  pull
-
-df$Reason.pre_calc <- df %>%
-  select(
-    starts_with("Reason"),
-    -Reason_nodesire, 
-    -ReasonOther.pre 
-  ) %>% 
-  mutate_all(as_factor) %>% 
-  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>%
-  pull
-
-df$occupation.pre_calc <- df %>%
-  select(
-    starts_with("occupation"),
-    -occupation_8_TEXT.pre
-  ) %>%
-  mutate_all(as_factor) %>% 
-  unite(., "Comb", sep = ", ", remove = TRUE, na.rm = TRUE) %>% 
-  pull
-
-df$CurrentEducation.pre_calc <-
-  df %>%
-  select(starts_with("CurrentEducation"),
-         -CurrentEducation_6_TEXT.pre) %>%
-  mutate_all(as_factor) %>%
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$groupType.pre_calc <-
-  df %>%
-  select(starts_with("gr_type"),
-          -ends_with("TEXT")) %>%
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$groupContext.pre_calc <-
-  df %>%
-  select(starts_with("gr_context"),
-          -ends_with("TEXT")) %>%
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$dyadType.pre_calc <-
-  df %>%
-  select(starts_with("dyad_type"),
-          -ends_with("TEXT")) %>%
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$context.pre_calc <-
-  df %>%
-  select(starts_with("Context"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$desireType.pre_calc <-
-  df %>%
-  select(starts_with("desire_type"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$desireType.pre_calc <-
-  df %>%
-  select(starts_with("desire_context"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$nonDutchType.pre_calc <-
-  df %>%
-  select(starts_with("NonDutchType"),
-          -ends_with("TEXT")) %>% 
-  mutate_all(as_factor) %>% 
-  unite(.,
-        "Comb",
-        sep = ", ",
-        remove = TRUE,
-        na.rm = TRUE) %>%
-  pull
-
-df$autonomySat.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT0[1-4].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-df$autonomyFrust.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT0[5-8].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-
-df$relatednessSat.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT(09|1[0-2]).pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-df$relatednessFrust.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT1[3-6].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-
-df$competenceSat.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT(1[7-9]|20).pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-df$competenceFrust.pre_calc <-
-  psych::scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(matches("^SDT2[1-4].pre")),
-    min = 1,
-    max = 7
-  )$scores %>% 
-  as.numeric
-
-df$IntGrAnx.pre_calc <-
-  psych::scoreItems(
-    keys = c(rep(1,6),rep(-1,6)),
-    items = df %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
-    min = 1,
-    max = 10
-  )$scores %>% 
-  as.numeric
-
-df$swl.pre_calc <-
-  psych::scoreItems(
-    keys = rep(1,5),
-    items = df %>% select(matches("^SWL[0-9]{2}\\.pre")),
-    min = 1,
-    max = 7,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-
-df$mdmq.alertness.pre_calc <- 
-  psych::scoreItems(
-    keys = c(1, -1, -1, 1),
-    items = df %>% select(MDMQ02.pre, MDMQ05.pre, MDMQ07.pre, MDMQ10.pre),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-df$mdmq.calmness.pre_calc <- 
-  psych::scoreItems(
-    keys = c(-1, 1, -1, 1),
-    items = df %>% select(MDMQ03.pre, MDMQ06.pre, MDMQ09.pre, MDMQ12.pre),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-df$mdmq.valence.pre_calc <- 
-  psych::scoreItems(
-    keys = c(1, -1, 1, -1),
-    items = df %>% select(MDMQ01.pre, MDMQ04.pre, MDMQ08.pre, MDMQ11.pre),
-    min = 1,
-    max = 6,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-
-df$IntGrAnx.post_calc <-
-  psych::scoreItems(
-    keys = c(rep(1,6),rep(-1,6)),
-    items = df %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
-    min = 1,
-    max = 10
-  )$scores %>% 
-  as.numeric
-
-df$swl.post_calc <-
-  psych::scoreItems(
-    keys = rep(1,5),
-    items = df %>% select(matches("^SWL[0-9]{2}\\.post")),
-    min = 1,
-    max = 7,
-    missing = TRUE,
-    impute = "none",
-  )$scores %>% 
-  as.numeric
-
-varNamIndicesS1 <-
-  append(varNamIndicesS1,
-         df %>% select(ends_with("_calc")) %>% names)
-
-#
-df <- df %>%
-  mutate(
-    exWB = scales::rescale(exWB, from = range(exWB, na.rm = TRUE), to = c(-100, 100)),
-    relatedness = relatedness + 50,
-    autonomy = autonomy + 50,
-    competence = competence + 50,
-    KeyNeedFulfillment = KeyNeedFulfillment + 50,
-    KeyNeedDueToPartner = KeyNeedDueToPartner + 50,
-    InteractionContextAccidental = InteractionContextAccidental + 50,
-    InteractionContextvoluntary = InteractionContextvoluntary + 50,
-    InteractionContextCooperative = InteractionContextCooperative + 50,
-    InteractionContextRepresentativeNL = InteractionContextRepresentativeNL + 50,
-    qualityOverall = qualityOverall + 50,
-    qualityMeaning = qualityMeaning + 50,
-    DaytimeNeedFulfillment = DaytimeNeedFulfillment + 50
-  )
-dtS1$full <- df
-rm(df)
-```
-
-### Study 2
-
-For the student sample, the data transformation stage had five main aims:
-
-1.  We first create person, survey type, and measurement ID variables.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  We add information about the interaction partner to the beep during which a person was selected as an interaction partner.\
-4.  We cluster mean-center key variables within participants.\
-5.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsStudents}
-df <- dtS2$clean
-
-# Add ID variables
-df$PID <- as.numeric(factor(df$ResponseId)) # participant ID
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S"),
-    date_period_old = date_period,
-    date_period = ifelse(
-      startTime >= c("12:00") & startTime <= c("18:59"),
-      paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Morning"),
-      ifelse(
-        startTime >= c("19:00") & startTime <= c("23:59"),
-        paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Afternoon"),
-        ifelse(
-          startTime >= c("00:00") & startTime <= c("11:59"),
-          paste(as.Date(created) - 1, "Afternoon"),
-          NA
-        )
-      )
-    )
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    dateOrder = stri_replace_all_regex(
-      date_period,
-      pattern = c(' Morning', ' Afternoon'),
-      replacement = c('_A', '_B'),
-      vectorize = FALSE
-    ),
-    TID = factor(date_period),
-    TIDnum = as.numeric(factor(dateOrder))
-  )
-
-# check whether time ordering worked
-df <- df %>%
-  arrange(PID, TID)
-
-# Interaction as Factor
-df$interaction.f <-
-  factor(df$Interaction,
-    levels = c("no interaction", "Dutch", "Non-Dutch")
-  )
-df$intNL <- ifelse(df$Interaction == "Dutch", 1, 0)
-df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)
-
-df$IntergroupContact <- (df$IntergroupContact-2)*-1
-df$IngroupContact <- (df$IngroupContact-2)*-1
-
-# -------------------------------------------------------------------------------------------------------------
-#                                       Combine Variables
-# -------------------------------------------------------------------------------------------------------------
-# Relatedness
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatednessInteraction <- rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
-df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
-# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; JK: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
-df$relatedness <-
-  rowMeans(df[, c("relatednessInteraction", "relatednessNoInteraction")],
-           na.rm = TRUE) %>%
-  ifelse(is.nan(.), NA, .)
-
-# Core Need
-df$DaytimeNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
-  df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]
-
-df$InteractionNeedFullfillment <- NA 
-df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)] <- 
-  df$KeyNeedFulfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)]
-
-# Randomly selected variables
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    ProSo = mean(c(ProSo1, ProSo2, ProSo3, ProSo4), na.rm = TRUE),
-    AntiSo = mean(c(AntiSo1, AntiSo2, AntiSo3, AntiSo4, AntiSo5, AntiSo6, AntiSo7), na.rm = TRUE)
-  )
-
-# -------------------------------------------------------------------------------------------------------------
-#                                 Add Variables related to interaction partner
-# -------------------------------------------------------------------------------------------------------------
-# create function for later lapply
-createIntPartDf <- function(inp) {
-  # prepare the dataframe so that we can forloop over it later
-  tmp <- data.frame(
-    CC = as.character(inp$CC),
-    NewCC = as.character(inp$NewCC),
-    NewName = as.character(inp$NewName),
-    NewCloseness = inp$NewCloseness,
-    NewGender = inp$NewGender,
-    NewEthnicity = as.character(inp$NewEthnicity),
-    NewRelationship = as.character(inp$NewRelationship)
-  )
-
-  tmp$CC2 <- recode(tmp$CC, "SOMEONE ELSE" = "NA")
-  tmp$CC2 <-
-    ifelse(
-      tmp$CC == 1 |
-        tmp$CC == "SOMEONE ELSE",
-      as.character(tmp$NewName),
-      as.character(tmp$CC2)
-    )
-  # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
-  tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
-  tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)
-
-  # open the variables that will be filled up in the foor-loop
-  tmp$closeness <- rep(NA, nrow(tmp))
-  tmp$gender <- rep(NA, nrow(tmp))
-  tmp$ethnicity <- rep(NA, nrow(tmp))
-  tmp$relationship <- rep(NA, nrow(tmp))
-
-  # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
-  # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
-  for (i in 1:nrow(tmp)) {
-    if (is.na(tmp$CC2[i])) {
-      next
-    } else {
-      tmp$closeness[i] <-
-        na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # find closeness where CC2 matches NewName (na.omit + [1] to get the number)
-      tmp$gender[i] <-
-        na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # (na.omit + [1] to get the number and not the rest of the na.omit list)
-      tmp$ethnicity[i] <-
-        na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] # PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
-      tmp$relationship[i] <-
-        na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
-    }
-  }
-
-  out <- tmp
-  out
-}
-
-# split df per participants and run function
-PP <- split(df, df$PID)
-PP <- lapply(PP, createIntPartDf)
-rm(createIntPartDf)
-
-# add variables back to df
-remergePP <- do.call(rbind.data.frame, PP)
-colnames(remergePP) <-
-  paste(colnames(remergePP), "_Calc", sep = "")
-df <- cbind(df, remergePP)
-rm(remergePP, PP)
-
-dtS2$full <- df
-rm(df)
-```
-
-### Study 3
-
-For the medical professional sample, the data transformation stage had five main aims:
-
-1.  We first create person, survey type, and measurement ID variables.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  We cluster mean-center key variables within participants.\
-4.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsMedical}
-df <- dtS3$clean
-
-# Add ID variables
-df$PID <- as.numeric(factor(df$ResponseId)) # participant ID
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S"),
-    date_period_old = date_period,
-    date_period = ifelse(
-      startTime >= c("12:00") & startTime <= c("18:59"),
-      paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Morning"),
-      ifelse(
-        startTime >= c("19:00") & startTime <= c("23:59"),
-        paste(format(as.POSIXct(created), format = "%Y-%m-%d"), "Afternoon"),
-        ifelse(
-          startTime >= c("00:00") & startTime <= c("11:59"),
-          paste(as.Date(created) - 1, "Afternoon"),
-          NA
-        )
-      )
-    )
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    dateOrder = stri_replace_all_regex(
-      date_period,
-      pattern = c(' Morning', ' Afternoon'),
-      replacement = c('_A', '_B'),
-      vectorize = FALSE
-    ),
-    TID = factor(date_period),
-    TIDnum = as.numeric(factor(dateOrder))
-  )
-
-# check whether time ordering worked
-df <- df %>%
-  arrange(PID, TID)
-
-# Interaction as Factor
-df$interaction.f <-
-  factor(df$Interaction,
-    levels = c("no interaction", "Dutch", "Non-Dutch")
-  )
-df$intNL <- ifelse(df$Interaction == "Dutch", 1, 0)
-df$intNonNL <- ifelse(df$Interaction == "Non-Dutch", 1, 0)
-
-df <- df %>%
-  mutate(
-    NonDutchContact = replace_na(NonDutchNum, 0), # make second non-Dutch countable
-    NonDutchContact = ifelse(NonDutchContact > 1, 1, 0) # recode (yes = 1 -> 1, no = 2 -> 0)
-  ) %>%
-  mutate(
-    OutgroupInteraction = factor(
-      InteractionDumDutch,
-      levels = c(0, 1),
-      labels = c("No", "Yes")
-    ),
-    NonOutgroupInteraction = factor(
-      rowSums(select(., c(InteractionDumNonDutch, NonDutchContact)), na.rm = TRUE), # combine the two non-Dutch Q.,
-      levels = c(0, 1),
-      labels = c("No", "Yes")
-    )
-  )
-
-df$IntergroupContact <- (df$IntergroupContact-2)*-1
-df$IngroupContact <- (df$IngroupContact-2)*-1
-
-# -------------------------------------------------------------------------------------------------------------
-#                                       Combine Variables
-# -------------------------------------------------------------------------------------------------------------
-# Relatedness
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatednessInteraction <-
-  rowMeans(df[c("relatednessSelf", "relatednessOther")], na.rm = TRUE)
-df$relatednessInteraction[df$relatednessInteraction == "NaN"] <- NA
-# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
-df$relatedness <-
-  rowMeans(df[, c("relatednessInteraction", "relatednessNoInteraction")],
-           na.rm = TRUE) %>%
-  ifelse(is.nan(.), NA, .)
-
-df$DaytimeNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)] <-
-  df$KeyNeedFulfillment[df$InteractionDum == 0 & !is.na(df$KeyNeedFulfillment)]
-
-df$InteractionNeedFullfillment <- NA 
-df$InteractionNeedFullfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)] <- 
-  df$KeyNeedFulfillment[df$InteractionDum == 1 & !is.na(df$KeyNeedFulfillment)]
-
-df$InteractionNeedImportance <- NA 
-df$InteractionNeedImportance[df$InteractionDum == 1 & !is.na(df$KeyNeedImp)] <- 
-  df$KeyNeedImp[df$InteractionDum == 1 & !is.na(df$KeyNeedImp)]
-
-# Randomly selected variables
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    ProSo = mean(c(ProSo1, ProSo2, ProSo3, ProSo4), na.rm = TRUE),
-    AntiSo = mean(c(AntiSo1, AntiSo2, AntiSo3, AntiSo4, AntiSo5, AntiSo6, AntiSo7), na.rm = TRUE),
-    agency = mean(c(agency1, agency2, agency3), na.rm = TRUE),
-    autoFrust = mean(c(autoFrust1, autoFrust2, autoFrust3, autoFrust4), na.rm = TRUE),
-    autoSat = mean(c(autoSat1, autoSat2, autoSat3, autoSat4), na.rm = TRUE),
-    relatFrust = mean(c(relatFrust1, relatFrust2, relatFrust3, relatFrust4), na.rm = TRUE),
-    relatSat = mean(c(relatSat1, relatSat2, relatSat3, relatSat4), na.rm = TRUE),
-    compFrust = mean(c(compFrust1, compFrust2, compFrust3, compFrust4), na.rm = TRUE),
-    compSat = mean(c(compSat1, compSat2, compSat3, compSat4), na.rm = TRUE),
-    lonely = mean(c(lonely4, lonely4, lonely4, lonely4), na.rm = TRUE),
-    emotRegPos = mean(c(emotRegPos01, emotRegPos02), na.rm = TRUE),
-    emotRegNeg = mean(c(emotRegNeg01, emotRegNeg02), na.rm = TRUE)
-  )
-
-# Allport's Conditions
-df %>%
-  #filter(OutgroupInteraction == "Yes") %>%
-  select(
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  ) %>%
-  pairs.panels.new
-
-AlportDescr <- df %>%
-  #filter(OutgroupInteraction == "Yes") %>%
-  select(
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  ) %>%
-  psych::describe(., skew=F,ranges=T) %>%
-  as.data.frame() %>%
-  select(-vars) %>%
-  kable(., caption = "Descriptives of Allport's Condition items") %>% 
-  kable_styling("hover", full_width = F, latex_options = "hold_position")
-
-iaWorkerAllport <- 
-  df %>%
-  #filter(OutgroupInteraction == "Yes") %>%
-  select(
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  )
-
-itemScaleAllport01 <- sjPlot::tab_itemscale(iaWorkerAllport)
-
-pca <- parameters::principal_components(iaWorkerAllport)
-factor.groups <- parameters::closest_component(pca)
-
-sjPlot::tab_itemscale(iaWorkerAllport, factor.groups, show.kurtosis = TRUE)
-
-AllportAlpha <- ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)
-
-data <- 
-  df %>%
-  select(
-    PID,
-    TIDnum,
-    InteractionContextEqualStatus,
-    KeyNeedShared,
-    InteractionContextCooperative,
-    InteractionContextvoluntary
-  ) %>%
-  drop_na %>%
-  reshape2::melt(
-    ., 
-    id.vars = c("PID", "TIDnum")
-  )
-
-AllportNestedAlpha <- horst::nestedAlpha(item.level.1 = "value",
-                   level.2      = "TIDnum",
-                   level.3      = "PID",
-                   data         = data)
-rm(data)
-
-iaWorkerAllportScale <- 
-  iaWorkerAllport %>%
-  Scale::Scale() %>%
-  Scale::ItemAnalysis()
-
-df$AllportsCondition <-
-  scoreItems(
-    keys = c(1, 1, 1, 1),
-    items = df %>% select(
-      InteractionContextEqualStatus,
-      KeyNeedShared,
-      InteractionContextCooperative,
-      InteractionContextvoluntary
-    ),
-    min = 0,
-    max = 100
-  )$scores
-
-as.data.frame(psych::describe(df$AllportsCondition, skew=T)) %>%
-  mutate(vars = "Allport's Conditions Index") %>%
-  kable(., caption = "Allport's Conditions: Scale Descriptives", row.names = FALSE) %>% 
-  kable_styling("hover", full_width = F, latex_options = "hold_position")
-
-ggplot(df, aes(x = AllportsCondition)) +
-  geom_histogram() +
-  theme_Publication()
-
-# -------------------------------------------------------------------------------------------------------------
-#                                 Add Variables related to interaction partner
-# -------------------------------------------------------------------------------------------------------------
-# create function for later lapply
-createIntPartDf <- function(inp) {
-  # prepare the dataframe so that we can forloop over it later
-  tmp <- data.frame(
-    CC = as.character(inp$CC),
-    NewCC = as.character(inp$NewCC),
-    NewName = as.character(inp$NewName),
-    NewCloseness = inp$NewCloseness,
-    NewGender = inp$NewGender,
-    NewEthnicity = as.character(inp$NewEthnicity),
-    NewRelationship = as.character(inp$NewRelationship)
-  )
-
-  tmp$CC2 <- recode(tmp$CC, "SOMEONE ELSE" = "NA")
-  tmp$CC2 <-
-    ifelse(
-      tmp$CC == 1 |
-        tmp$CC == "SOMEONE ELSE",
-      as.character(tmp$NewName),
-      as.character(tmp$CC2)
-    )
-  # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
-  tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
-  tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)
-
-  # open the variables that will be filled up in the foor-loop
-  tmp$closeness <- rep(NA, nrow(tmp))
-  tmp$gender <- rep(NA, nrow(tmp))
-  tmp$ethnicity <- rep(NA, nrow(tmp))
-  tmp$relationship <- rep(NA, nrow(tmp))
-
-  # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
-  # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
-  for (i in 1:nrow(tmp)) {
-    if (is.na(tmp$CC2[i])) {
-      next
-    } else {
-      tmp$closeness[i] <-
-        na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # find closeness where CC2 matches NewName (na.omit + [1] to get the number)
-      tmp$gender[i] <-
-        na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] # (na.omit + [1] to get the number and not the rest of the na.omit list)
-      tmp$ethnicity[i] <-
-        na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] # PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
-      tmp$relationship[i] <-
-        na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
-    }
-  }
-
-  out <- tmp
-  out
-}
-
-# split df per participants and run function
-PP <- split(df, df$PID)
-PP <- lapply(PP, createIntPartDf)
-rm(createIntPartDf)
-
-# add variables back to df
-remergePP <- do.call(rbind.data.frame, PP)
-colnames(remergePP) <-
-  paste(colnames(remergePP), "_Calc", sep = "")
-df <- cbind(df, remergePP)
-rm(remergePP, PP)
-
-dtS3$full <- df
-```
-
-## Data Availability and Sample Selection {.tabset .tabset-fade}
-
-As one of our main analyses is a three-mode principal component analysis (3MPCA) we begin by assessing the amount of missing data in each of the ESM studies.
-To assess the missingness in detail, we prepare a data availability table of whether data is available or missing for each possible measurement point (data + morning/afternoon) for all individual participants.
-We then export the data availability tables as comma separated value files (.csv) to be assessed in detail using a spreadsheet program (such as MS Excel).
-
-For our sample selection we address both the selection of time points and participants.
-Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model.
-Thus, we then select the time points for which we have less than 45% missingness and select participants who have less than 45% missingness across the selected time range.
-
-To approximate an ideal sample selection that maximized the number of participants and timepoints that fit the 55% data availability criterion.
-We select sequentially remove the rows or columns with the lowest data availability rate untill all rows and columns have more than 55% data available (i.e., less than 45% missingness).
-
-::: {.alert .alert-info .alert-dismissible .fade .in role="alert"}
-<a class="close" data-dismiss="alert" aria-label="close">×</a> <i class="fas fa-info-circle"></i> <b>Note: </b><br/> This needs to be adjusted current missingness criterion set to more conservative 33%.
-:::
-
-### Study 1
-
-```{r workerDataAvailability}
-dtS1Availability <- dtS1$full %>%
-  select(
-    PID,
-    TIDnum
-  ) %>%
-  arrange(PID, TIDnum) %>%
-  mutate(data = 1)
-dtS1Availability <- reshape::cast(dtS1Availability, PID ~ TIDnum) %>%
-  select(-PID) %>%
-  mutate_all(function(x) ifelse(x>1,1,x))
-# sum(colMeans(dtS1Availability)*100 >= 66)
-# sum(rowMeans(dtS1Availability)*100 >= 66)
-# sum(colMeans(dtS1Availability[-5,])*100 >= 66)
-# sum(rowMeans(dtS1Availability[,-8])*100 >= 66)
-
-rownames(dtS1Availability) <- paste("PP", 1:nrow(dtS1Availability), sep = "_")
-colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
-dtS1AvailabilityKbl <- dtS1Availability
-
-dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)] <- lapply(dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS1AvailabilityKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 1: Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-
-write.csv(dtS1Availability, "data/S1_Workers/processed/workerAvailability.csv")
-```
-
-```{r testOptimizer, include=FALSE}
-# set.seed(42)
-# nr <- 5
-# nc <- 10
-# df <- as.data.table(matrix(rbinom(nr*nc,1,.66), nrow=nr, ncol=nc))
-# df
-# 
-# best.list.row.df = list()
-# for (i in 1:nrow(df)) {
-#   # get best subset for rows based on how many columns have more than 66% data
-#   rowlist = combn(nrow(df), i)
-#   numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
-#   cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
-#   best = which(numobs == max(numobs))[1]
-#   best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
-# }
-# 
-# best.list.col.df = list()
-# for (i in 1:ncol(df)) {
-#   # get best subset for columns based on how many rows have more than 66% data
-#   collist = combn(ncol(df), i)
-#   numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
-#   cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
-#   best = which(numobs == max(numobs))[1]
-#   best.list.col.df = c(best.list.col.df, list(collist[, best]))
-# }
-# rm(df, best.list.row.df, best.list.col.df)
-```
-
-```{r testOptimizer03, include=FALSE}
-# library(dplyr)
-# library(ROI)
-# library(ROI.plugin.glpk)
-# library(ompr)
-# library(ompr.roi)
-# 
-# set.seed(42)
-# tf <- matrix(sample(c(TRUE, FALSE), 1488, replace=TRUE), 31)
-# M3 <- t(replicate(31, 1969:2016, simplify=TRUE))
-# M3[tf] <- NA
-# 
-# m <- +!is.na(M3) # gets logical matrix; 0 if NA else 1    
-# #m <- as.data.frame(dtS1AvailabilityTest)
-# nr <- nrow(m)
-# nc <- ncol(m)
-# n_years <- 15 
-# 
-# 
-# model <- MIPModel() %>% 
-#   # keep[i,j] is 1 if matrix cell [i,j] is to be kept else 0
-#   add_variable(keep[i,j], i = 1:nr, j = 1:nc, typ = "binary") %>% 
-#   # rm_row[i] is 1 if row i is selected for removal else 0
-#   add_variable(rm_row[i], i = 1:nr, type = "binary") %>% 
-#   # rm_col[j] is 1 if column j is selected for removal else 0
-#   add_variable(rm_col[j], j = 1:nc, type = "binary") %>% 
-#   # maximize good cells kept
-#   set_objective(sum_expr(keep[i,j], i = 1:nr, j = 1:nc), "max") %>% 
-#   # cell can be kept only when row is not selected for removal
-#   add_constraint(sum_expr(keep[i,j], j = 1:nc) <= 1 - rm_row[i], i = 1:nr) %>%
-#   # cell can be kept only when column is not selected for removal
-#   add_constraint(sum_expr(keep[i,j], i = 1:nr) <= 1 - rm_col[j], j = 1:nc) %>%
-#   # keep at most n_years columns i.e. remove at least (nc - n_years) columns
-#   # only non-NA values can be kept
-#   add_constraint(m[i,j] + rm_row[i] + rm_col[j] >= 1, i = 1:nr, j = 1:nc) %>% 
-#   #add_constraint(sum_expr(m[,j], j = 1:nc)/nc*100 >= 66) %>% 
-#   #add_constraint(sum_expr(m[i,], i = 1:nr)/nr*100 >= 66) %>% 
-#   # keep cols with more 
-#   #add_constraint(colMeans(m[i,j], i = 1:nr, j = 1:nc) >= 0.66) %>%
-#   # I used >= instead of == to avoid infeasiblity
-#   #add_constraint(sum_expr(rm_col[j], j = 1:nc) >= nc - n_years) %>% 
-#   # solve using free glpk solver
-#   solve_model(with_ROI(solver = "glpk", verbose = TRUE))
-# 
-# solver_status(model)
-# rm_rows <- model %>% 
-#   get_solution(rm_row[i]) %>% 
-#   filter(value > 0) %>% 
-#   pull(i)
-# 
-# rm_cols <- model %>% 
-#   get_solution(rm_col[j]) %>% 
-#   filter(value > 0) %>% 
-#   pull(j) 
-# 
-# result <- m[-rm_rows, -rm_cols, drop = F]
-# result
-```
-
-```{r testOptimizer05}
-cleanM <- function(M, c = 55) { #c = 66
-  check <- list()
-  rowNamesOut <- c()
-  colNamesOut <- c()
-  for (i in 1:length(c(row.names(M), names(M)))) {
-    rowMeans <- rowMeans(M) * 100
-    colMeans <- colMeans(M) * 100
-    rcMeans <- c(rowMeans, colMeans)
-    rm <- which.min(rcMeans) %>% names
-    rc <- ifelse(startsWith(rm, "PP_"), "row", "col")
-
-    check[[i]] <- rcMeans
-
-    if (!all(rcMeans >= c) && rc == "row") {
-      M <- M[!(row.names(M) %in% rm), ]
-      rowNamesOut <- append(rowNamesOut, rm)
-      cat(i, ": Row ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
-    } else if (!all(rcMeans >= c) && rc == "col") {
-      M <- M[, !(names(M) %in% rm)]
-      colNamesOut <- append(colNamesOut, rm)
-      cat(i, ": Column ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
-    } else {
-      cat(i, ": All row- and column means are over ", c, "%. The final matrix has ", nrow(M), " rows and ", ncol(M), " columns.", sep = "")
-      return(
-        list(
-          reducedMatrix = M,
-          rowNamesIn = row.names(M),
-          colNamesIn = names(M),
-          rowNamesOut = rowNamesOut,
-          colNamesOut = colNamesOut,
-          rowMeans = rowMeans,
-          colMeans = colMeans
-        )
-      )
-    }
-  }
-}
-
-dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 55)
-
-PIDout <- gsub("PP_", "", dtS1RedInfo$rowNamesOut) %>% as.numeric
-TIDout <- gsub("t_", "", dtS1RedInfo$colNamesOut) %>% as.numeric
-TIDInRed <- gsub("t_", "", dtS1RedInfo$colNamesIn) %>% as.numeric
-TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
-
-dtS1Red <- dtS1$full %>%
-  filter(
-    !PID %in% PIDout,
-    TIDnum %in% TIDIn
-  ) %>% 
-  mutate(TIDnum = TIDnum - min(TIDnum))
-rm(PIDout, TIDout, TIDInRed, TIDIn)
-
-# change glitch where survey was sent too early
-dtS1Red$TIDnum[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- 10
-dtS1Red$TID[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- "2018-05-18 Morning"
-
-dtS1AvailabilityRedKbl <- dtS1RedInfo$reducedMatrix
-dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)] <- lapply(dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS1AvailabilityRedKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 1: Final Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-### Study 2
-
-```{r studentDataAvailability}
-dtS2Availability <- dtS2$full %>%
-  select(
-    PID,
-    TIDnum
-  ) %>%
-  #na.omit %>%
-  arrange(PID, TIDnum) %>%
-  mutate(data = 1)
-dtS2Availability <- reshape::cast(dtS2Availability, PID ~ TIDnum) %>%
-  select(-PID)
-
-rownames(dtS2Availability) <- paste("PP", 1:nrow(dtS2Availability), sep = "_")
-colnames(dtS2Availability) <- paste("t", 1:ncol(dtS2Availability), sep = "_")
-
-write.csv(dtS2Availability, "data/S2_Students/processed/studentAvailability.csv")
-
-dtS2AvailabilityKbl <- dtS2Availability
-dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)] <- lapply(dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS2AvailabilityKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 2: Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-```{r s2SampleReduction}
-dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 55)
-
-PIDout <- gsub("PP_", "", dtS2RedInfo$rowNamesOut) %>% as.numeric
-TIDout <- gsub("t_", "", dtS2RedInfo$colNamesOut) %>% as.numeric
-TIDInRed <- gsub("t_", "", dtS2RedInfo$colNamesIn) %>% as.numeric
-TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
-
-dtS2Red <- dtS2$full %>%
-  filter(
-    !PID %in% PIDout,
-    TIDnum %in% TIDIn
-  ) %>% 
-  mutate(TIDnum = TIDnum - min(TIDnum))
-rm(PIDout, TIDout, TIDInRed, TIDIn)
-
-# remove glitch entries 
-dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 57 & dtS2Red$ended == "2018-12-20 21:58:31"),]
-dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 59 & dtS2Red$ended == "2018-12-20 21:58:31"),]
-
-dtS2AvailabilityRedKbl <- dtS2RedInfo$reducedMatrix
-dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)] <- lapply(dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS2AvailabilityRedKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 2: Final Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-### Study 3
-
-```{r medicalDataAvailability}
-dtS3Availability <- dtS3$full %>%
-  select(
-    PID,
-    TIDnum
-  ) %>%
-  arrange(PID, TIDnum) %>%
-  mutate(data = 1)
-dtS3Availability <- reshape::cast(dtS3Availability, PID ~ TIDnum, fill = 0) %>%
-  select(-PID)
-
-rownames(dtS3Availability) <- paste("PP", 1:nrow(dtS3Availability), sep = "_")
-colnames(dtS3Availability) <- paste("t", 1:ncol(dtS3Availability), sep = "_")
-
-write.csv(dtS3Availability, "data/S3_Medical/processed/medicalAvailability.csv")
-
-dtS3AvailabilityKbl <- dtS3Availability
-dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)] <- lapply(dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS3AvailabilityKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 2: Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-rm(dtS3AvailabilityKbl)
-```
-
-```{r s3SampleReduction}
-dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 55)
-
-PIDout <- gsub("PP_", "", dtS3RedInfo$rowNamesOut) %>% as.numeric
-TIDout <- gsub("t_", "", dtS3RedInfo$colNamesOut) %>% as.numeric
-TIDInRed <- gsub("t_", "", dtS3RedInfo$colNamesIn) %>% as.numeric
-TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)
-
-dtS3Red <- dtS3$full %>%
-  filter(
-    !PID %in% PIDout,
-    TIDnum %in% TIDIn
-  ) %>% 
-  mutate(TIDnum = TIDnum - min(TIDnum))
-rm(PIDout, TIDout, TIDInRed, TIDIn)
-
-dtS3AvailabilityRedKbl <- dtS3RedInfo$reducedMatrix
-dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)] <- lapply(dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)], function(x) {
-    cell_spec(x,
-              bold = FALSE,
-              color = "white",
-              background = ifelse(x == 1, "green", "red")
-              )
-})
-kbl(
-  dtS3AvailabilityRedKbl,
-  format = "html",
-  escape = FALSE,
-  align = "c",
-  booktabs = TRUE,
-  caption = "Study 3: Final Data Availability" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = FALSE,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  ) %>%
-  scroll_box(width = "100%", height = "500px")
-```
-
-## Time Scales and Timepoint Aggregation
-
-```{r checkVarAvailability}
-varNamesPre <- varNames %>%
-  filter(
-    surveyS1 == "pre",
-    surveyS2 == "pre",
-    str_detect(surveyS3, "^pre")
-  ) %>%
-  select(varNam)
-varNamesPreFull <-
-  data.frame(varNam = Reduce(intersect, list(
-    names(dtS1$full), names(dtS3$full), names(dtS3$full)
-  ))) %>%
-  filter(str_detect(varNam, ".pre$"))
-
-varNamesDaily <- varNames %>%
-  filter(
-    surveyS1 == "daily",
-    surveyS2 == "daily",
-    surveyS3 == "daily"
-  ) %>%
-  select(varNam)
-varNamesDailyFull <-
-  data.frame(varNam = Reduce(intersect, list(
-    names(dtS1Red), names(dtS2Red), names(dtS3Red)
-  ))) %>%
-  filter(!str_detect(varNam, ".pre$|.post$"))
-
-varNamesPost <- varNames %>%
-  filter(
-    surveyS1 == "post",
-    surveyS2 == "post",
-    surveyS3 == "post"
-  ) %>%
-  select(varNam)
-varNamesPostFull <-
-  data.frame(varNam = Reduce(intersect, list(
-    names(dtS1Red), names(dtS3Red), names(dtS3Red)
-  ))) %>%
-  filter(str_detect(varNam, ".post$"))
-```
-
-```{r analysisDFs}
-# Main Analysis: 3MPCA (common variables across studies)
-dropVarPcaMain <- c(
-  # Meta Data
-  "last_outside_referrer",
-  "created",
-  "ended",
-  "ip_address",
-  "date",
-  "ResponseId",
-  "server_language",
-  "Meta_Browser",
-  # Interaction Dummies (non-continuous)
-  "IntergroupContact",
-  "IngroupContact",
-  "InteractionDum",
-  # Interaction Counts (non-continuous)
-  "ContactNum",
-  "NonDutchNum",
-  "groupDutch",
-  # Interaction Descriptives (non-continuous)
-  "duration",
-  "dyadGroup",
-  "groupSize",
-  # Keyneed Freetexts (non-continuous)
-  "KeyNeed",
-  "DaytimeNeed",
-  # Relatedness individual items
-  "relatednessSelf",
-  "relatednessOther",
-  "relatednessNoInteraction"
-)
-
-varNam3mpcaMain <- varNamesDailyFull %>%
-  filter(!varNam %in% dropVarPcaMain) %>%
-  pull
-
-dtMain3mpca <- rbind(
-  dtS1Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
-  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
-  ) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(
-    date = as.Date(gsub(" .*", "", TID)),
-    week = strftime(date, format = "%Y-W%V")
-    ) %>%
-  select(
-    ID,
-    PID,
-    TID,
-    date,
-    week,
-    TIDnum,
-    study,
-    everything()
-  ) %>%
-  arrange(ID, TIDnum)
-
-dtMain3mpcaDaily <- dtMain3mpca %>%
-  group_by(ID, date, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(date))) %>%
-  ungroup %>%
-  select(ID, date, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .))
-
-dtMain3mpcaWeekly <- dtMain3mpca %>%
-  group_by(ID, week, study) %>%
-  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
-  ungroup %>%
-  group_by(study) %>%
-  mutate(TIDnum = as.numeric(factor(week))) %>%
-  ungroup %>%
-  select(ID, week, TIDnum, everything()) %>%
-  mutate_all(~ifelse(is.nan(.), NA, .))
-```
-
-We aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data.
-Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014].
-This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels.
-We chose to select time scales that align with common human cycles.
-We thus compare the variances of bi-daily, daily, and weekly aggregations.
-Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.
-We then chose the time scales that have the most variance.
-
-```{r keyNeedTest}
-
-# WILL USE IMPUTED DATA AND FOR-LOOP ACROSS ALL RELEVANT DVs
-# PLOT EFFECT SIZES
-
-dtMain3mpca <- dtMain3mpca %>%
-  group_by(study) %>%
-  mutate(
-    TIDdate = as.numeric(factor(date)),
-    TIDweek = as.numeric(factor(week))
-  ) %>%
-  ungroup
-
-# Not sure whether TIDs as predictors or empty model?
-keyNeedMlTest01 <-
-  lme4::lmer(KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek + (1 + TIDnum + TIDdate + TIDweek | ID),
-    data = dtMain3mpca
-  )
-
-keyNeedMlTest02 <-
-  lme(
-    KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek,
-    random = ~ 1 + TIDnum + TIDdate + TIDweek | ID,
-    data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
-    control = list(opt = "nlmimb")
-  ) # use optim if it does not converge
-
-# summary(keyNeedMlTest02)
-# summ(keyNeedMlTest01, digits = 3, center = FALSE)
-# anova(keyNeedMlTest02)
-
-```
-
-# **Feature-based Time-Series Clustering**
-
-Feature-based clustering has the advantage of being extremely flexible when it comes to dealing with real-world human data.
-Human data is commonly noisy, incomplete, and non-stationary.
-Many descriptive dimension-reduction, modeling, and raw time-series clustering procedures will require complete or stationary data from the researcher, and are thus currently not well-suited to deal with messy intensive-longitudinal data collected by many social science researchers.
-
-We used the following five-step procedure to arrive at a stable, reproducible, and meaningful differentiation of participants ESM trajectories:
-
-1. Variable selection
-2. Feature extraction
-3. (Feature selection)
-4. Clustering
-5. Visualization, interpretation, and validation
-
-## Variable Selection
-
-We begin by selecting the variables, which we would like to include in the clustering procedure.
-For the clustering of time-series characteristics we focus on variables that are specific to the ESM measurements (i.e., variables collected during the various daily dairy pings).
-This means, that variables from the pre- and post-test are excluded from the clustering procedure.
-While the variables can in theory be added as additional features, they are not the focus of this manuscript.
-
-In our case, we begin by collecting the variables that are in the ESM measurements of all three studies, in order to build the most powerful dataset for the main analysis.
-We then additionally, divide the variables by their response type and by study to check for cluster differences in the validation phase:
-
-1. *Main Analysis*: Variables across all three data sets
-2. *Validation*: Variables by response type (interaction-specific or not)
-3. *Validation*: Variables by study
-
-```{r Preparation of Variable Names}
 varNamS123Aux <- varNames %>%
@@
 
-We also see that a number of variables are specific to interactions.
-These variables will by their very nature have substantially larger missingness (because not every measurement occasion will follow new social interactions) and these variables are also un-imputable in traditional time-series modeling and hypothesis testing procedures.
-
-```{r variable names by response type}
+```{r variable names by response type and study}
 varNames %>%
@@
 
-Please note that not all variables are included in the main analysis, as some variables are only found within specific studies.
-
-## Feature Extraction
-
-Time-series can be described with a plethora of features.
-A good example is the list of 100 time-series features that one can extract with [tsfresh python package](https://tsfresh.readthedocs.io/en/latest/api/tsfresh.feature_extraction.html# "list of features to extract").
-In practice there are usually two steps to the feature engineering process.
-(1) *Feature extraction*, that is summarizing the time series in meaningful descriptors, and (2) *feature selection*, the process of ensuring that only meaningful features are included in the clustering step.
-For the interpretability of our clustering procedure we have made a pre-selection of features, we wish to include to describe our time-series:
-
-1. central tendency
-    + mean ✓
-    + median ✓
-2. variation around the central tendency (i.e., density)
-    + SD ✓
-    + MAD ✓
-3. (in)stability
-    + root mean square of successive differences (rmssd) ✓
-    + mean absolute change (mac) ✓
-4. inertia
-    + lag-1 autocorrelation ✓
-    + periodicity (lag-2 & lag-14 autocorrelation) ✓
-5. linear trend
-    + OLS regression parameter ✓
-6. non-linear trend
-    + GAM edf ✓
-
-(!discuss NA procedures!)
-
-```{r load saved and prepared features}
-load("data/features/featOut.RData")
-```
-
-### Across Studies
-
-We begin by extracting the features for the main analysis that combines the sample of the three studies.
-We do this across all selected variables as well as for the content (non-)specific variables separately.
-During the feature extraction step there can at times be features that cannot be calculated for a small number of participants.
-An example would a participant who never filled in two consecutive measurements that describe an interaction.
-In that case, we cannot extract an auto correlation of lag-1 for some of the interaction specific variables.
-
-```{r Feature Extraction Main Analysis}
-# varNames
-#items <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Joined Dataset of all three studies
-featData <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
+```{r missingness filtering}
+missInfo <- function(full = NA, reduced = NA){
+  missTab <- data.frame(
+    nFull = nrow(full),
+    nRed = nrow(reduced),
+    pptFull = length(unique(full$PID)),
+    pptRed = length(unique(reduced$PID)),
+    timeFull = length(unique(full$TID)),
+    timeRed = length(unique(reduced$TID))
+  ) %>%
     mutate(
-     across(all_of(varNamS123PCA), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS123PCA), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS123PCA), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featFull <-
-#   featureExtractor(
-#     data = featData,
-#     pid = "ID",
-#     tid = "TIDnum",
-#     items = varNamS123
-#   )
-# 
-# featFullContact <-
-#   featureExtractor(
-#     data = featData,
-#     pid = "ID",
-#     tid = "TIDnum",
-#     items = varNamS123Int
-#   )
-# 
-# featFullNoContact <-
-#   featureExtractor(
-#     data = featData,
-#     pid = "ID",
-#     tid = "TIDnum",
-#     items = varNamS123NoInt
-#   )
-# 
-# 
-# featFull$features <- featFull$features %>%
-#     filter(
-#       ID != 19
-#     )
-
-pMissFull <- pMissFeat(
-  all = featFull$features,
-  contact = featFullContact$features,
-  nocontact = featFullNoContact$features,
-  title = "Feature-wise Missingess Across all Studies"
-)
-
-pMissFull$pMissPlot
-
-# featFullImp <- featureImputer(featFull)
-# featFullNoContactImp <- featureImputer(featFullNoContact)
-# featFullContactImp <- featureImputer(featFullContact)
-
-```
-
-For the full dataset across the three studies, we find that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r nrow(featFull$features)` participants and `r ncol(featFull$features)` features.
-These missing features are imputed with a single predictive mean matching imputation using the MICE library.
-
-### Study 1
-
-We then repeat the feature extraction procedure for all selected variables that are specific to Study 1.
-
-```{r Feature Extraction Study 1 specific variables}
-# varNames
-#items <- varNamS1Clust
-
-# Prepare Dataframe
-featDataS1 <- 
-  dtS1Red %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(PID,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS1Clust)) %>%
-   mutate(
-     across(all_of(varNamS1Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS1Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS1Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(PID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featS1 <-
-#   featureExtractor(
-#     data = featDataS1,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS1Clust
-#   )
-# 
-# featS1Contact <-
-#   featureExtractor(
-#     data = featDataS1,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS1Int
-#   )
-# 
-# featS1NoContact <-
-#   featureExtractor(
-#     data = featDataS1,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS1NoInt
-#   )
-
-pMissS1 <- pMissFeat(
-  all = featS1$features,
-  contact = featS1Contact$features,
-  nocontact = featS1NoContact$features,
-  title = "Feature-wise Missingess Study 1"
-)
-
-pMissS1$pMissPlot
-
-# featS1Imp <- featureImputer(featS1)
-# featS1NoContactImp <- featureImputer(featS1NoContact)
-# featS1ContactImp <- featureImputer(featS1Contact)
-
-```
-
-### Study 2
-
-In the same manner, the features are extracted for the variables that are specific to Study 2.
-
-```{r Feature Extraction Study 2 specific variables}
-# varNames
-#items <- varNamS2Clust
-
-# Prepare Dataframe
-featDataS2 <- 
-  dtS2Red %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(PID,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS2Clust)) %>%
-   mutate(
-     across(all_of(varNamS2Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS2Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS2Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(PID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featS2 <-
-#   featureExtractor(
-#     data = featDataS2,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS2Clust
-#   )
-# 
-# featS2Contact <-
-#   featureExtractor(
-#     data = featDataS2,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS2Int
-#   )
-# 
-# featS2NoContact <-
-#   featureExtractor(
-#     data = featDataS2,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS2NoInt
-#   )
-
-pMissS2 <- pMissFeat(
-  all = featS2$features,
-  contact = featS2Contact$features,
-  nocontact = featS2NoContact$features,
-  title = "Feature-wise Missingess Study 2"
-)
-
-pMissS2$pMissPlot
-
-# featS2Imp <- featureImputer(featS2)
-# featS2NoContactImp <- featureImputer(featS2NoContact)
-# featS2ContactImp <- featureImputer(featS2Contact)
-```
-
-### Study 3
-
-And finally, we extract the features for the Study 3 specific variables.
-
-```{r Feature Extraction Study 3 specific variables}
-# varNames
-#items <- varNamS3Clust
-
-# Prepare Dataframe
-featDataS3 <- 
-  dtS3Red %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(PID,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS3Clust)) %>%
-   mutate(
-     across(all_of(varNamS3Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS3Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS3Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(PID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featS3 <-
-#   featureExtractor(
-#     data = featDataS3,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS3Clust
-#   )
-# 
-# featS3Contact <-
-#   featureExtractor(
-#     data = featDataS3,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS3Int
-#   )
-# 
-# featS3NoContact <-
-#   featureExtractor(
-#     data = featDataS3,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS3NoInt
-#   )
-
-pMissS3 <- pMissFeat(
-  all = featS3$features,
-  contact = featS3Contact$features,
-  nocontact = featS3NoContact$features,
-  title = "Feature-wise Missingess Study 3"
-)
-pMissS3$pMissPlot
-
-# featS3Imp <- featureImputer(featS3)
-# featS3NoContactImp <- featureImputer(featS3NoContact)
-# featS3ContactImp <- featureImputer(featS3Contact)
-# 
-# save(list = ls(pattern = "feat"), file = "data/features/featOut.RData")
-```
-
-## Feature Selection
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Here we might have to make a selection of the features we extract because some of the clustering algorithms might not be able to deal with the many different features. See below
-</div>
-
-## Clustering
-
-Cluster analyses have been a staple within a variety of sciences for the better part of a century.
-It's applications are wide and because most procedures fall under the "unsupervised learning" umbrella the methodological developments are fast-paced and new methods or refinements are published constantly.
-Among the many options available four types are found most commonly within the modern literature on clustering analyses:
-
-1. centroid-based
-    + e.g., k-means
-    + no assumptions (+)
-    + fast even with large matrices (+)
-    + non-probabilistic (+/-)
-    + sensitive to local minima (-)
-2. distribution-based
-    + e.g., gaussian mixture model (GMM)
-    + flexible shape (+)
-    + probabilistic (+/-)
-    + (gaussian) distributions assumed (-)
-3. density-based
-    + e.g., dbscan / OPTICS
-    + no shape assumption (+)
-    + does not assign outliers/noise (+)
-4. Hierarchical-based (not relevant)
-    + e.g., agglomerative hierarchical clustering
-    + best with hierarchical data (-)
-    + best with small *n* (-)
-
-Each clustering procedures comes with their own sets of advantages and disadvantages.
-We will focus mainly on the k-means procedure because it has the least assumptions and is fast even with large data sets.
-Additionally, the methods are well-established and the results are easy to interpret.
-As such, the procedure should be applicable to most users within the social sciences and for ESM feature data specifically.
-We offer Gaussian mixture modeling and a common density-based approach as supplementary analyses.
-
-### Main Anlaysis: Across Studies
-
-The main clustering procedure, uses the features calculated for the relevant variables across all three studies.
-
-#### Centroid-based
-
-For our main method, we use the centroid-based k-means clustering.
-The procedure needs the researcher to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution.
-
-```{r kmeans determin k}
-# prepare data for main analysis
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ
-scaled_data <- featFullImp$featuresImpZMat
-
-# load libraries
-library(factoextra)
-library(NbClust)
-
-# Elbow method
-fviz_nbclust(scaled_data, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method")
-# Silhouette method
-fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method")
-# Gap statistic [takes for ever]
-# nboot = 50 to keep the function speedy. 
-# recommended value: nboot= 500 for your analysis.
-# Use verbose = FALSE to hide computing progression.
-# set.seed(123)
-# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
-#   labs(subtitle = "Gap statistic method")
-```
-
-We use this parameter to run the actual k-means cluster analysis, during which the algorithm iteratively seeks to optimally separate the participants into two clusters based on their feature values.
-During the iterative process the algorithm might at times stop at a non-optimal solution (i.e., local minima).
-To avoid this drawback and to find the (globally) optimal solution we let the algorithm run with 100 separate staring positions.
-
-```{r k-means modeling based on selected k}
-kmeansOut <- kmeans(scaled_data, centers = 2, nstart = 100)
-```
-
-Because the clustering happens in a multidimensional space of the `r ncol(featFull$features)` features, we begin by visualizing the data in a two-dimensional space, where we encircle the points of the clusters (i.e., convex hulls) for the first two dimensions of a principle component analysis.
-
-```{r Plot Convex Hulls}
-fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())
-```
-
-We then inspect the cluster-averages of all feature for each variable.
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., how cooperative an interaction was) than for other variables (e.g., Attitudes towards the Dutch).
-Additionally, we see that (2) some within the variables some features are better at distinguishing clusters (e.g., mean and median of well-being) than others (e.g., sd and rmssd of well-being).
-
-```{r features for each variable}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarize_all(mean) %>%
-  ungroup %>%
-  select(-ID) %>% 
-  reshape2::melt(id.vars = "cluster") %>% 
-  mutate(lab = variable) %>%
-  mutate(
-    lab = variable,
-    variable = gsub("\\_.*", "", lab),
-    feature = gsub("^.*?\\_", "", lab)
-  ) %>% 
-  ggplot(., aes(x = feature, y = value, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ variable) +
-  theme_Publication()
-```
-
-Next, we inspect the cluster-averages of all variables for each feature.
-This is the same data as in the plot above.
-However, here we see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median) than others (e.g., auto correlations).
-
-```{r variables for each feature}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarize_all(mean) %>%
-  ungroup %>%
-  select(-ID) %>% 
-  reshape2::melt(id.vars = "cluster") %>% 
-  mutate(
-    lab = variable,
-    variable = gsub("\\_.*", "", lab),
-    feature = gsub("^.*?\\_", "", lab)
-  ) %>% 
-  ggplot(., aes(x = variable, y = value, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ feature) +
-  theme_Publication()
-```
-
-In a final step we visualize the raw time series for each variable and add the average cluster development to see the features in action.
-
-```{r time-series by variable and cluster}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
+      nDif = nFull-nRed,
+      nDifPerc = nDif/nFull*100,
+      pptDif = pptFull-pptRed,
+      pptDifPerc = pptDif/pptFull*100,
+      timeDif = timeFull-timeRed,
+      timeDifPerc = timeDif/timeFull*100
+    ) %>%
     select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ variable, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-```
-
-#### Distribution-based
-
-For the distribution-based clustering we use the commonly used Gaussian mixture modeling (GMM) procedure.
-This procedure assumes  independent Gaussian distributions for each each cluster.
-These distributions are substantially better at identifying clusters if the data comes from multivariate Gaussian distributions (i.e., ellipses in a two dimensional space).
-
-We again begin by determining the optimal number of clusters. For GMMs this is done using common information criteria, such as the BIC.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  This is where we run into trouble.
-These plots are relatively uncommon and according to the internet this is most likely a problem of too many (non-distinguishing or highly correlated) features.
-A good paper on this is <a href="https://proceedings.neurips.cc/paper/2002/file/e58aea67b01fa747687f038dfde066f6-Paper.pdf">Law et al.
-(2002)</a>.
-</div>
-
-```{r gaussian mixture model}
-library(mclust)
-
-scaled_data <- featFullImp$featuresImpZMat
-
-# scaled_data <- featS1Imp$featuresImpZMat  %>% #select(starts_with(varNamS123Int))
-#   select(-any_of(names(which(colSums(is.na(.)) > 0))))
-
-BIC <- mclustBIC(scaled_data)
-plot(BIC)
-summary(BIC)
-mod1 <- Mclust(scaled_data)
-summary(mod1, parameters = FALSE)
-
-# multivariate <- densityMclust(scaled_data, plot=FALSE)
-# summary(multivariate)            
-# plot(multivariate, what = "BIC")
-
-# merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View
-
-#table(class, mod1$classification)
-
-#mod1dr <- MclustDR(mod1)
-#summary(mod1dr, parameters = FALSE)
-#plot(mod1dr, what = "classification")
-```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  The feature selection issue is further highlighted when we do the GMMs by feature. Where we find a reasonable number of clusters for each feature. 
-</div>
-
-```{r MClust feature specific cluster numbers}
-nMClust <- list()
-for(f in unique(gsub("^.*?\\_", "", names(featFullImp$featuresImpZMat)))){
-  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
-  mod1 <- Mclust(scaled_data)
-  #summary(mod1, parameters = FALSE)
-  nMClust[[f]] <- mod1$G
+      nFull, nRed, nDif, nDifPerc,
+      pptFull, pptRed, pptDif, pptDifPerc,
+      timeFull, timeRed, timeDif, timeDifPerc
+    )
+    missTab
 }
 
-nMClust %>%
-  unlist %>%
-  data.frame(feature=names(.), G=., row.names=NULL) %>%
+missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
+missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
+missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)
+
+rbind(
+  missS1,
+  missS2,
+  missS3
+) %>%
+  mutate(study = c(1, 2, 3)) %>%
+  select(study, everything()) %>%
   kbl(.,
       escape = FALSE,
-      booktabs = T,
-      align = c("l", "c"),
+      booktabs = TRUE,
+      align = "c", #c("l", rep("c", ncol(.)-1)),
+      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
       digits=2,
-      col.names = c(
-        "Feature",
-        "Number of Clusters"
-      ),
-      caption = "Number of MClust Clusters if separated by Feature") %>%
+      caption = "Missingness Info by Study") %>%
+  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
   kable_classic(
@@
 
-#### Density-based
-
-For the density-based cluster analysis we focus on the *DBSCAN* algorithm ('Density-Based Spatial Clustering of Applications with Noise') as well as its extension the *OPTICS* procedure ('Ordering Points To Identify the Clustering Structure').
-This procedure is more bottom up than the previous two approaches.
-Instead of providing the number of clusters, a cluster is specified if a minimum number of other points (the MinPts parameter) are within a given radius (the epsilon parameter).
-Together these two parameters determine a density at which a group of data points are considered a cluster.
-While this description is grossly oversimplified, it highlights one of the advantages of the procedures, namely that dense regions can expand to be any shape in the d-dimensional space.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Same feature selection issue as above. 
-</div>
-
-```{r dbscan}
-library(dbscan)
-library(fpc)
-
-#scaled_data <- featFullImp$featuresImpZMat %>% select(starts_with(varNamS123Int))
-scaled_data <- featFullImp$featuresImpZMat  # %>% select(starts_with(varNamS123NoInt))#%>%
-  #select(-any_of(names(which(colSums(is.na(scaled_data)) > 0))))
-
-#is.na(scaled_data) %>% table
-
-# to plot the eps values
-eps_plot = kNNdistplot(scaled_data, k=3)
-# to draw an optimum line
-eps_plot %>% abline(h = 15, lty = 2)
-
-potentialMinPts <- round(SciViews::ln(nrow(scaled_data)))
-eps <- 15
-minPts <- 2
-
-res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0)
-res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
-
-resEqual.check <- all(res.fpc$cluster == res.db$cluster)
-nClustDBSCAN <- length(unique(res.fpc$cluster))-1
-
-fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
-#hullplot(scaled_data, res.fpc)
+\subsection{Data}
+
+To illustrate the
+
+\subsubsection{Participants}
+
+\subsubsection{Timepoints}
+
+\subsubsection{Variables}
+
+\subsection{Procedure}
+
+feature engineering: feature extraction + feature selection
+
+\subsubsection{Feature-Extraction}
+
+use domain knowledge to extract new variables from raw data (summarize)
+
+options: tsfresh \citep[][]{christ2018}, spikes \citep[][]{caro-martin2018}
+
+Our selection: density, inertia, instability, diversity \citep[][]{dejonckheere2019, kuppens2017}
+
+\begin{itemize}
+  \item 
+\end{itemize}
+
+\subsubsection{Feature Selection}
+
+select most consistent, relevant, and non-redundant features.
+
+\textit{k} features means $2^k – 1$ possible models
+
+maximize relevance and minimize redundancy
+
+"However, most of the existing UFS methods primarily focus on the significance of features in maintaining the data structure while ignoring the redundancy among features.
+Moreover, the determination of the proper number of features is another challenge."
+
+\begin{itemize}
+  \item feature selection
+  \begin{itemize}
+    \item filter: evaluate individual features' impact / importance / relevance. Univariate (ranking-based) or multivariate (preferred bc. can handle redundant and irrelevant features)
+    \begin{itemize}
+      \item unvariate, often based on information or spectral analysis (univariate inclusion test \citep[i.e., alpha-corrected t-tests][]{christ2018})
+    \end{itemize}
+    \item wrapper: search problem --- compare different models based on evaluation criterion (i.e., model performance; e.g., $R^2$, accuracy, ...).
+"high computational cost, and they are limited to be used in conjunction with a particular clustering algorithm."
+    \begin{itemize}
+      \item Exhaustive Feature Selection (i.e., brute force)
+      \item forward selection (e.g., greedy Forward Search \citep[][]{wang2006})
+      \item backward elimination
+      \item Bi-directional elimination (Stepwise Selection)
+    \end{itemize}
+    \item hybrid: first filter then wrapper
+    \item [embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance] MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
+  \end{itemize}
+  \item dimension reduction
+\end{itemize}
+
+random forest search \citep[][]{}
+
+```{r}
 
 ```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  And again the number of clusters is reasonable if we consider the features separately: 
-</div>
-
-```{r dbscan feature specific cluister numbers}
-minPts <- 3
-xi <- 0.04 #0.05
-nDbscan <- list()
-for(f in unique(gsub("^.*?\\_", "", names(featFullImp$featuresImpZMat)))){
-  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
-  #res.fpc <- dbscan::dbscan(scaled_data, eps=eps_cl, MinPts=minPts)
-
-  res.optics <- dbscan::optics(scaled_data, minPts = 3)
-
-  #plot(res.optics, col = "grey")
-  #dend <- as.dendrogram(res.optics)
-  #plot(dend, ylab = "Reachability dist.", leaflab = "none")
-  #extractDBSCAN(res.optics, eps_cl = 19.12)
-
-  res <- extractXi(res.optics, xi = xi)
-  res$clusters_xi
-  #plot(res)
-  #hullplot(scaled_data, res)
-
-  #summary(res.optics, parameters = FALSE)
-  nDbscan[[f]] <- nrow(res$clusters_xi)
-}
-
-nDbscan %>%
-  unlist %>%
-  data.frame(feature=names(.), G=., row.names=NULL) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "c"),
-      digits=2,
-      col.names = c(
-        "Feature",
-        "Number of Clusters"
-      ),
-      caption = "Number of OPTICS Clusters if separated by Feature") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-### Follow-up: By Response Type
-
-_**currently removed for faster render until feature selection issue is resolved**_
-
-### Follow-up: By Study
-
-_**currently removed for faster render until feature selection issue is resolved**_
-
-# **Software Information**
-
-The full session information with all relevant system information and all loaded and installed packages is available in the collapsible section below.
-
-<details>
-
-<summary>
-
-System Info
-
-</summary>
-
-\renewcommand{\arraystretch}{0.8}
-
-<!-- decrease line spacing for the table -->
-
-```{r Reproducibility-sessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
-df_session_platform <- devtools::session_info()$platform %>%
-  unlist(.) %>%
-  as.data.frame(.) %>%
-  rownames_to_column(.)
-
-colnames(df_session_platform) <- c("Setting", "Value")
-
-kbl(
-  df_session_platform,
-  booktabs = T,
-  align = "l",
-  caption = "R environment session info for reproducibility of results" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\renewcommand{\arraystretch}{1}
-
-<!-- reset row height/line spacing -->
-
-</details>
-
-<br>
-
-<details>
-
-<summary>
-
-Package Info
-
-</summary>
-
-\renewcommand{\arraystretch}{0.6}
-
-<!-- decrease line spacing for the table -->
-
-```{r Reproducibility-sessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
-df_session_packages <- devtools::session_info()$packages %>%
-  as.data.frame(.) %>%
-  filter(attached == TRUE) %>%
-  dplyr::select(loadedversion, date, source) %>%
-  rownames_to_column()
-
-colnames(df_session_packages) <- c("Package", "Loaded version", "Date", "Source")
-
-kbl(
-  df_session_packages,
-  booktabs = T,
-  align = "l",
-  caption = "Package info for reproducibility of results" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\renewcommand{\arraystretch}{1}
-
-<!-- reset row height/line spacing -->
-
-</details>
-
-<br>
-
-<details>
-
-<summary>
-
-Full Session Info (including loaded but unattached packages --- for troubleshooting only)
-
-</summary>
-
-`r pander(sessionInfo(), compact = FALSE)`
-
-</details>
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-# **References**
-
-::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
-:::

### 2022-11-21 methods-and-results.tex
m&r init

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{varNamIndicesS1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"autonomy"}\NormalTok{,}
  \StringTok{"competence"}\NormalTok{,}
  \StringTok{"relatedness"}\NormalTok{,}
  \StringTok{"education\_level.pre"}\NormalTok{,}
  \StringTok{"associationMerged.pre"}\NormalTok{,}
  \StringTok{"assimilation.pre"}\NormalTok{,}
  \StringTok{"separation.pre"}\NormalTok{,}
  \StringTok{"integration.pre"}\NormalTok{,}
  \StringTok{"marginalization.pre"}\NormalTok{,}
  \StringTok{"VIA\_heritage.pre"}\NormalTok{,}
  \StringTok{"VIA\_Dutch.pre"}\NormalTok{,}
  \StringTok{"SSAS\_surrounding.pre"}\NormalTok{,}
  \StringTok{"SSAS\_privat.pre"}\NormalTok{,}
  \StringTok{"SSAS\_public.pre"}\NormalTok{,}
  \StringTok{"assimilation.post"}\NormalTok{,}
  \StringTok{"separation.post"}\NormalTok{,}
  \StringTok{"integration.post"}\NormalTok{,}
  \StringTok{"marginalization.post"}\NormalTok{,}
  \StringTok{"VIA\_heritage.post"}\NormalTok{,}
  \StringTok{"VIA\_Dutch.post"}\NormalTok{,}
  \StringTok{"rosenberg.post"}\NormalTok{,}
  \StringTok{"social\_support.post"}\NormalTok{,}
  \StringTok{"stress.post"}\NormalTok{,}
  \StringTok{"discrimination.post"}\NormalTok{,}
  \StringTok{"discrimination\_month.post"}\NormalTok{,}
  \StringTok{"NLE\_1month.post"}\NormalTok{,}
  \StringTok{"NLE\_6month.post"}\NormalTok{,}
  \StringTok{"NLE\_12month.post"}
\NormalTok{  )}

\NormalTok{varNamS123Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}\NormalTok{,}
\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}\NormalTok{,}
\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}
\NormalTok{varNamS1Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}
\NormalTok{varNamS2Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}
\NormalTok{varNamS3Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\NormalTok{varNamS123PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}\NormalTok{,}
\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}\NormalTok{,}
\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}
\NormalTok{varNamS1PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}
\NormalTok{varNamS2PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  pull}
\NormalTok{varNamS3PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\NormalTok{varNamOut }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"ResponseId"}\NormalTok{,}
  \StringTok{"relatednessNoInteraction"}\NormalTok{,}
  \StringTok{"relatednessSelf"}\NormalTok{,}
  \StringTok{"relatednessOther"}\NormalTok{,}
  \StringTok{"autonomy\_Int"}\NormalTok{,}
  \StringTok{"autonomy\_NoInt"}\NormalTok{,}
  \StringTok{"competence\_Int"}\NormalTok{,}
  \StringTok{"competence\_NoInt"}\NormalTok{,}
\NormalTok{  varNamS1PCA[}\FunctionTok{grepl}\NormalTok{(}\StringTok{\textquotesingle{}\^{}MDMQ\textquotesingle{}}\NormalTok{, varNamS1PCA)],}
\NormalTok{  varNamS2PCA[}\FunctionTok{grepl}\NormalTok{(}\StringTok{\textquotesingle{}\^{}ProSo|\^{}AntiSo\textquotesingle{}}\NormalTok{, varNamS2PCA)],}
\NormalTok{  varNamS3PCA[}\FunctionTok{grepl}\NormalTok{(}
    \StringTok{\textquotesingle{}\^{}ProSo|\^{}AntiSo|\^{}agency|\^{}autoFrust|\^{}autoSat|\^{}relatFrust|\^{}relatSat|\^{}compFrust|\^{}compSat|\^{}lonely[0{-}9]|\^{}emotRegPos|\^{}emotRegNeg\textquotesingle{}}\NormalTok{,}
\NormalTok{    varNamS3PCA}
\NormalTok{  )]}
\NormalTok{)}
\NormalTok{varNamCore }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"PID"}\NormalTok{,}
  \StringTok{"TID"}\NormalTok{,}
  \StringTok{"TIDnum"}
\NormalTok{)}
\NormalTok{varNamNewAll }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"relatedness"}\NormalTok{,}
  \StringTok{"autonomy"}\NormalTok{,}
  \StringTok{"competence"}
\NormalTok{)}
\NormalTok{varNamNewS1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"relatedness"}\NormalTok{,}
  \StringTok{"autonomy"}\NormalTok{,}
  \StringTok{"competence"}\NormalTok{,}
  \StringTok{"alertness"}\NormalTok{, }
  \StringTok{"calmness"}\NormalTok{, }
  \StringTok{"valence"}
\NormalTok{)}
\NormalTok{varNamNewS2 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"relatedness"}\NormalTok{,}
  \StringTok{"autonomy"}\NormalTok{,}
  \StringTok{"competence"}\NormalTok{,}
  \StringTok{"ProSo"}\NormalTok{, }
  \StringTok{"AntiSo"}
\NormalTok{)}
\NormalTok{varNamNewS3 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"relatedness"}\NormalTok{,}
  \StringTok{"autonomy"}\NormalTok{,}
  \StringTok{"competence"}\NormalTok{,}
  \StringTok{"ProSo"}\NormalTok{, }
  \StringTok{"AntiSo"}\NormalTok{,}
  \StringTok{"agency"}\NormalTok{,}
  \StringTok{"autoFrust"}\NormalTok{,}
  \StringTok{"autoSat"}\NormalTok{,}
  \StringTok{"relatFrust"}\NormalTok{,}
  \StringTok{"relatSat"}\NormalTok{,}
  \StringTok{"compFrust"}\NormalTok{,}
  \StringTok{"compSat"}\NormalTok{,}
  \StringTok{"lonely"}\NormalTok{,}
  \StringTok{"emotRegPos"}\NormalTok{,}
  \StringTok{"emotRegNeg"}
\NormalTok{)}

\NormalTok{varNamNewCat }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \StringTok{"closeness\_Calc"}\NormalTok{,}
  \StringTok{"gender\_Calc"}\NormalTok{,}
  \StringTok{"ethnicity\_Calc"}\NormalTok{,}
  \StringTok{"relationship\_Calc"}
\NormalTok{)}

\NormalTok{varNamIntDep }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
\NormalTok{  varNamS123PCA[}\FunctionTok{grepl}\NormalTok{(}\StringTok{\textquotesingle{}\^{}InteractionContext|AttitudesPartner|KeyNeedDueToPartner|\^{}quality\textquotesingle{}}\NormalTok{, varNamS123PCA)]}
\NormalTok{)}

\NormalTok{varNamIndiceItemsS1 }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    aux }\SpecialCharTok{==} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}
\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{append}\NormalTok{(., }\FunctionTok{gsub}\NormalTok{(}\StringTok{".pre|.post|\_calc"}\NormalTok{, }\StringTok{""}\NormalTok{, varNamIndicesS1)) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  unique}

\NormalTok{varNamS123MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewAll, varNamS123Aux)}
\NormalTok{varNamS1MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamIndicesS1, varNamS1Aux[}\SpecialCharTok{!}\NormalTok{varNamS1Aux }\SpecialCharTok{\%in\%}\NormalTok{ varNamIndiceItemsS1]) }\CommentTok{\#}
\NormalTok{varNamS2MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS2, varNamS2Aux)}
\NormalTok{varNamS3MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS3, varNamS3Aux)}

\NormalTok{varNamS123PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamS123PCA[}\SpecialCharTok{!}\NormalTok{varNamS123PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
\NormalTok{varNamS1PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS1, varNamS1PCA[}\SpecialCharTok{!}\NormalTok{varNamS1PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
\NormalTok{varNamS2PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS2, varNamS2PCA[}\SpecialCharTok{!}\NormalTok{varNamS2PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
\NormalTok{varNamS3PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS3, varNamS3PCA[}\SpecialCharTok{!}\NormalTok{varNamS3PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}

\NormalTok{varNamS123MIPsbl }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
  \FunctionTok{paste}\NormalTok{(varNamS123MI, }\FunctionTok{rep}\NormalTok{(}\StringTok{".pre"}\NormalTok{, }\FunctionTok{length}\NormalTok{(varNamS123MI)), }\AttributeTok{sep =} \StringTok{""}\NormalTok{),}
\NormalTok{  varNamS123MI,}
  \FunctionTok{paste}\NormalTok{(varNamS123MI, }\FunctionTok{rep}\NormalTok{(}\StringTok{".post"}\NormalTok{, }\FunctionTok{length}\NormalTok{(varNamS123MI)), }\AttributeTok{sep =} \StringTok{""}\NormalTok{)}
\NormalTok{)}

\NormalTok{varNamS123MiRed }\OtherTok{\textless{}{-}} \FunctionTok{Reduce}\NormalTok{(}
\NormalTok{  intersect,}
  \FunctionTok{list}\NormalTok{(}
\NormalTok{    dtS1Red }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(PID, TID, TIDnum, }\FunctionTok{any\_of}\NormalTok{(varNamS123MIPsbl)) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ names,}
\NormalTok{    dtS2Red }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(PID, TID, TIDnum, }\FunctionTok{any\_of}\NormalTok{(varNamS123MIPsbl)) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ names,}
\NormalTok{    dtS3Red }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(PID, TID, TIDnum, }\FunctionTok{any\_of}\NormalTok{(varNamS123MIPsbl)) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ names}
\NormalTok{  )}
\NormalTok{)}

\NormalTok{idVars }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"PID"}\NormalTok{, }\StringTok{"TID"}\NormalTok{, }\StringTok{"TIDnum"}\NormalTok{, }\StringTok{"date"}\NormalTok{, }\StringTok{"week"}\NormalTok{, }\StringTok{"study"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# ID Variables (just to re{-}iterate)}
\NormalTok{idVars }\OtherTok{\textless{}{-}}\NormalTok{ idVars}

\CommentTok{\# Common variables across studies }
\NormalTok{varNamS123 }\OtherTok{\textless{}{-}}\NormalTok{ varNamS123PCA[}\SpecialCharTok{!}\NormalTok{varNamS123PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}

\CommentTok{\# Variable Groups}
\CommentTok{\# varNames \%\textgreater{}\%}
\CommentTok{\#   filter(pca != 0) \%\textgreater{}\%}
\CommentTok{\#   mutate(contactSpecific = recode(.$contactSpecific, }
\CommentTok{\#                          \textasciigrave{}{-}1\textasciigrave{}="No Interaction Only",}
\CommentTok{\#                          \textasciigrave{}0\textasciigrave{}="unspecific",}
\CommentTok{\#                          \textasciigrave{}1\textasciigrave{}="Interaction Only")) \%\textgreater{}\%}
\CommentTok{\#   crosstable::crosstable(., varGroup, by=contactSpecific) \%\textgreater{}\%}
\CommentTok{\#   select({-}c(.id, label)) \%\textgreater{}\%}
\CommentTok{\#   kbl(.,}
\CommentTok{\#       escape = FALSE,}
\CommentTok{\#       booktabs = T,}
\CommentTok{\#       align = c("l", "r", "c"),}
\CommentTok{\#       digits=2,}
\CommentTok{\#       caption = "Full Analysis: Variable Groups by Interaction Types") \%\textgreater{}\%}
\CommentTok{\#   kable\_classic(}
\CommentTok{\#     full\_width = F,}
\CommentTok{\#     lightable\_options = "hover",}
\CommentTok{\#     html\_font = "Cambria"}
\CommentTok{\#   )}
\CommentTok{\# }
\CommentTok{\# varNames \%\textgreater{}\%}
\CommentTok{\#   filter(}
\CommentTok{\#     pca != 0}
\CommentTok{\#   ) \%\textgreater{}\%}
\CommentTok{\#   mutate(contactSpecific = recode(.$contactSpecific, }
\CommentTok{\#                          \textasciigrave{}{-}1\textasciigrave{}="No Interaction Only",}
\CommentTok{\#                          \textasciigrave{}0\textasciigrave{}="unspecific",}
\CommentTok{\#                          \textasciigrave{}1\textasciigrave{}="Interaction Only")) \%\textgreater{}\%}
\CommentTok{\#   select(varNam, varGroup, contactSpecific) \%\textgreater{}\%}
\CommentTok{\#   group\_by(varGroup, varNam, contactSpecific) \%\textgreater{}\%}
\CommentTok{\#   summarise(n = n()) \%\textgreater{}\%}
\CommentTok{\#   pivot\_wider(names\_from = c("varGroup", "varNam"), values\_from = n, names\_sep=" {-} ") \%\textgreater{}\%}
\CommentTok{\#   ungroup \%\textgreater{}\%}
\CommentTok{\#   t \%\textgreater{}\%}
\CommentTok{\#   as.data.frame \%\textgreater{}\%}
\CommentTok{\#   janitor::row\_to\_names(1) \%\textgreater{}\%}
\CommentTok{\#   replace(is.na(.), 0) \%\textgreater{}\%}
\CommentTok{\#   tibble::rownames\_to\_column(., var = "variable") \%\textgreater{}\%}
\CommentTok{\#   separate(variable, c("concept", "variable"), " {-} ") \%\textgreater{}\%}
\CommentTok{\#   arrange(concept, unspecific, \textasciigrave{}Interaction Only\textasciigrave{}) \%\textgreater{}\%}
\CommentTok{\#   mutate\_all(\textasciitilde{}recode(.x, \textasciigrave{}0\textasciigrave{}="", \textasciigrave{} 1\textasciigrave{}="✓")) \%\textgreater{}\%}
\CommentTok{\#   kbl(.,}
\CommentTok{\#       escape = FALSE,}
\CommentTok{\#       booktabs = T,}
\CommentTok{\#       align = c("l", "l", rep("c", ncol(.){-}2)),}
\CommentTok{\#       digits=2,}
\CommentTok{\#       caption = "All Possible Variables by Interaction Types") \%\textgreater{}\%}
\CommentTok{\#   kable\_classic(}
\CommentTok{\#     full\_width = F,}
\CommentTok{\#     lightable\_options = "hover",}
\CommentTok{\#     html\_font = "Cambria"}
\CommentTok{\#   ) \%\textgreater{}\%}
\CommentTok{\#   scroll\_box(width = "100\%", height = "500px")}

\CommentTok{\# Non{-}Interaction Specific Variables}
\NormalTok{varNamS123NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS123PCA,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# Interaction Specific Variables}
\NormalTok{varNamS123Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS123PCA,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# S1 variables}
\NormalTok{varNamS1Clust }\OtherTok{\textless{}{-}}\NormalTok{ varNamS1PCA[}\SpecialCharTok{!}\NormalTok{varNamS1PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}

\CommentTok{\# S1 Non{-}Interaction Specific Variables}
\NormalTok{varNamS1NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS1Clust,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# S1 Interaction Specific Variables}
\NormalTok{varNamS1Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS1Clust,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# S2 variables}
\NormalTok{varNamS2Clust }\OtherTok{\textless{}{-}}\NormalTok{ varNamS2PCA[}\SpecialCharTok{!}\NormalTok{varNamS2PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}

\CommentTok{\# S2 Non{-}Interaction Specific Variables}
\NormalTok{varNamS2NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS2Clust,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# S2 Interaction Specific Variables}
\NormalTok{varNamS2Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS2Clust,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# S3 variables}
\NormalTok{varNamS3Clust }\OtherTok{\textless{}{-}}\NormalTok{ varNamS3PCA[}\SpecialCharTok{!}\NormalTok{varNamS3PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}

\CommentTok{\# S3 Non{-}Interaction Specific Variables}
\NormalTok{varNamS3NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS3Clust,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}

\CommentTok{\# S3 Interaction Specific Variables}
\NormalTok{varNamS3Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS3Clust,}
\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  pull}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{varNames }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}
\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
\NormalTok{    varNam }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{contactSpecific =} \FunctionTok{recode}\NormalTok{(}
\NormalTok{      .}\SpecialCharTok{$}\NormalTok{contactSpecific,}
      \StringTok{\textasciigrave{}}\AttributeTok{{-}1}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"No Interaction Only"}\NormalTok{,}
      \StringTok{\textasciigrave{}}\AttributeTok{0}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"unspecific"}\NormalTok{,}
      \StringTok{\textasciigrave{}}\AttributeTok{1}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"Interaction Only"}
\NormalTok{    ),}
    \AttributeTok{study1 =} \FunctionTok{ifelse}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(studyS1), }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \AttributeTok{study2 =} \FunctionTok{ifelse}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(studyS2), }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \AttributeTok{study3 =} \FunctionTok{ifelse}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(studyS3), }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(varNam, varGroup, contactSpecific, study1, study2, study3) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(varGroup, varNam, contactSpecific, study1, study2, study3) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{n =} \FunctionTok{n}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =} \FunctionTok{c}\NormalTok{(}\StringTok{"varGroup"}\NormalTok{, }\StringTok{"varNam"}\NormalTok{, }\StringTok{"study1"}\NormalTok{, }\StringTok{"study2"}\NormalTok{, }\StringTok{"study3"}\NormalTok{), }\AttributeTok{values\_from =}\NormalTok{ n, }\AttributeTok{names\_sep=}\StringTok{" {-} "}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  ungroup }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  t }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  as.data.frame }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  janitor}\SpecialCharTok{::}\FunctionTok{row\_to\_names}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{replace}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(.), }\DecValTok{0}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  tibble}\SpecialCharTok{::}\FunctionTok{rownames\_to\_column}\NormalTok{(., }\AttributeTok{var =} \StringTok{"variable"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{separate}\NormalTok{(variable, }\FunctionTok{c}\NormalTok{(}\StringTok{"concept"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"study1"}\NormalTok{, }\StringTok{"study2"}\NormalTok{, }\StringTok{"study3"}\NormalTok{), }\StringTok{" {-} "}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(study1, study2, study3), as.numeric)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{arrange}\NormalTok{(concept, unspecific, }\StringTok{\textasciigrave{}}\AttributeTok{Interaction Only}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate\_all}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\FunctionTok{recode}\NormalTok{(.x, }\StringTok{\textasciigrave{}}\AttributeTok{0}\StringTok{\textasciigrave{}}\OtherTok{=}\StringTok{""}\NormalTok{, }\StringTok{\textasciigrave{}}\AttributeTok{ 1}\StringTok{\textasciigrave{}}\OtherTok{=}\StringTok{"✓"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{relocate}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"study"}\NormalTok{), }\AttributeTok{.after =} \FunctionTok{last\_col}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rename\_with}\NormalTok{(., }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gsub}\NormalTok{(}\StringTok{"study"}\NormalTok{, }\StringTok{"Study "}\NormalTok{, .x, }\AttributeTok{fixed =} \ConstantTok{TRUE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kbl}\NormalTok{(.,}
      \AttributeTok{escape =} \ConstantTok{FALSE}\NormalTok{,}
      \AttributeTok{booktabs =}\NormalTok{ T,}
      \AttributeTok{align =} \FunctionTok{c}\NormalTok{(}\StringTok{"l"}\NormalTok{, }\StringTok{"l"}\NormalTok{, }\FunctionTok{rep}\NormalTok{(}\StringTok{"c"}\NormalTok{, }\FunctionTok{ncol}\NormalTok{(.)}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{)),}
      \AttributeTok{digits=}\DecValTok{2}\NormalTok{,}
      \AttributeTok{caption =} \StringTok{"Variables by Interaction Types and Study Availability"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_header\_above}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{" "} \OtherTok{=} \DecValTok{2}\NormalTok{, }\StringTok{"Contact Type"} \OtherTok{=} \DecValTok{2}\NormalTok{, }\StringTok{"Study Availability"} \OtherTok{=} \DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable\_classic}\NormalTok{(}
    \AttributeTok{full\_width =}\NormalTok{ F,}
    \AttributeTok{lightable\_options =} \StringTok{"hover"}\NormalTok{,}
    \AttributeTok{html\_font =} \StringTok{"Cambria"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:variable names by response type and study}Variables by Interaction Types and Study Availability}
\centering
\begin{tabular}[t]{llccccc}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Contact Type} & \multicolumn{3}{c}{Study Availability} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-7}
concept & variable & unspecific & Interaction Only & Study 1 & Study 2 & Study 3\\
\midrule
acculturation & AttitudesDutch & ✓ &  & ✓ & ✓ & ✓\\
acculturation & AttitudesPartner &  & ✓ & ✓ & ✓ & ✓\\
contactQuality & InteractionContextAccidental &  & ✓ & ✓ & ✓ & ✓\\
contactQuality & InteractionContextCooperative &  & ✓ & ✓ & ✓ & ✓\\
contactQuality & InteractionContextEqualStatus &  & ✓ &  & ✓ & ✓\\
\addlinespace
contactQuality & InteractionContextRepresentativeNL &  & ✓ & ✓ & ✓ & ✓\\
contactQuality & InteractionContextvoluntary &  & ✓ & ✓ & ✓ & ✓\\
contactQuality & qualityEffective &  & ✓ &  & ✓ & ✓\\
contactQuality & qualityImportance &  & ✓ &  & ✓ & ✓\\
contactQuality & qualityMeaning &  & ✓ & ✓ & ✓ & ✓\\
\addlinespace
contactQuality & qualityOverall &  & ✓ & ✓ & ✓ & ✓\\
contactQuality & qualityStar &  & ✓ & ✓ &  & \\
emotion & afraid & ✓ &  &  &  & ✓\\
emotion & angry & ✓ &  &  &  & ✓\\
emotion & energy & ✓ &  &  &  & ✓\\
\addlinespace
emotion & exWB & ✓ &  & ✓ & ✓ & ✓\\
emotion & lonelyAlways & ✓ &  &  &  & ✓\\
needs & autonomy & ✓ &  &  & ✓ & ✓\\
needs & competence & ✓ &  &  & ✓ & ✓\\
needs & DaytimeNeedFulfillment & ✓ &  & ✓ & ✓ & ✓\\
\addlinespace
needs & DaytimeNeedImportance & ✓ &  &  &  & ✓\\
needs & KeyNeedFulfillment & ✓ &  & ✓ & ✓ & ✓\\
needs & KeyNeedImp & ✓ &  &  &  & ✓\\
needs & StudentGoal01 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal02 & ✓ &  &  & ✓ & ✓\\
\addlinespace
needs & StudentGoal03 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal04 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal05 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal06 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal07 & ✓ &  &  & ✓ & ✓\\
\addlinespace
needs & StudentGoal08 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal09 & ✓ &  &  & ✓ & ✓\\
needs & StudentGoal10 & ✓ &  &  & ✓ & ✓\\
needs & KeyNeedDueToPartner &  & ✓ & ✓ & ✓ & ✓\\
needs & KeyNeedShared &  & ✓ &  & ✓ & ✓\\
\bottomrule
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{missInfo }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{full =} \ConstantTok{NA}\NormalTok{, }\AttributeTok{reduced =} \ConstantTok{NA}\NormalTok{)\{}
\NormalTok{  missTab }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
    \AttributeTok{nFull =} \FunctionTok{nrow}\NormalTok{(full),}
    \AttributeTok{nRed =} \FunctionTok{nrow}\NormalTok{(reduced),}
    \AttributeTok{pptFull =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(full}\SpecialCharTok{$}\NormalTok{PID)),}
    \AttributeTok{pptRed =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(reduced}\SpecialCharTok{$}\NormalTok{PID)),}
    \AttributeTok{timeFull =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(full}\SpecialCharTok{$}\NormalTok{TID)),}
    \AttributeTok{timeRed =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(reduced}\SpecialCharTok{$}\NormalTok{TID))}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}
      \AttributeTok{nDif =}\NormalTok{ nFull}\SpecialCharTok{{-}}\NormalTok{nRed,}
      \AttributeTok{nDifPerc =}\NormalTok{ nDif}\SpecialCharTok{/}\NormalTok{nFull}\SpecialCharTok{*}\DecValTok{100}\NormalTok{,}
      \AttributeTok{pptDif =}\NormalTok{ pptFull}\SpecialCharTok{{-}}\NormalTok{pptRed,}
      \AttributeTok{pptDifPerc =}\NormalTok{ pptDif}\SpecialCharTok{/}\NormalTok{pptFull}\SpecialCharTok{*}\DecValTok{100}\NormalTok{,}
      \AttributeTok{timeDif =}\NormalTok{ timeFull}\SpecialCharTok{{-}}\NormalTok{timeRed,}
      \AttributeTok{timeDifPerc =}\NormalTok{ timeDif}\SpecialCharTok{/}\NormalTok{timeFull}\SpecialCharTok{*}\DecValTok{100}
\NormalTok{    ) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{select}\NormalTok{(}
\NormalTok{      nFull, nRed, nDif, nDifPerc,}
\NormalTok{      pptFull, pptRed, pptDif, pptDifPerc,}
\NormalTok{      timeFull, timeRed, timeDif, timeDifPerc}
\NormalTok{    )}
\NormalTok{    missTab}
\NormalTok{\}}

\NormalTok{missS1 }\OtherTok{\textless{}{-}} \FunctionTok{missInfo}\NormalTok{(}\AttributeTok{full =}\NormalTok{ dtS1}\SpecialCharTok{$}\NormalTok{full, }\AttributeTok{reduced =}\NormalTok{ dtS1Red)}
\NormalTok{missS2 }\OtherTok{\textless{}{-}} \FunctionTok{missInfo}\NormalTok{(}\AttributeTok{full =}\NormalTok{ dtS2}\SpecialCharTok{$}\NormalTok{full, }\AttributeTok{reduced =}\NormalTok{ dtS2Red)}
\NormalTok{missS3 }\OtherTok{\textless{}{-}} \FunctionTok{missInfo}\NormalTok{(}\AttributeTok{full =}\NormalTok{ dtS3}\SpecialCharTok{$}\NormalTok{full, }\AttributeTok{reduced =}\NormalTok{ dtS3Red)}

\FunctionTok{rbind}\NormalTok{(}
\NormalTok{  missS1,}
\NormalTok{  missS2,}
\NormalTok{  missS3}
\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{study =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{select}\NormalTok{(study, }\FunctionTok{everything}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kbl}\NormalTok{(.,}
      \AttributeTok{escape =} \ConstantTok{FALSE}\NormalTok{,}
      \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
      \AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\CommentTok{\#c("l", rep("c", ncol(.){-}1)),}
      \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Study"}\NormalTok{, }\FunctionTok{rep}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"Full"}\NormalTok{,}\StringTok{"Reduced"}\NormalTok{,}\StringTok{"$}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{Delta$"}\NormalTok{, }\StringTok{"\%"}\NormalTok{), }\DecValTok{3}\NormalTok{)),}
      \AttributeTok{digits=}\DecValTok{2}\NormalTok{,}
      \AttributeTok{caption =} \StringTok{"Missingness Info by Study"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{add\_header\_above}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{" "} \OtherTok{=} \DecValTok{1}\NormalTok{, }\StringTok{"Measurements"} \OtherTok{=} \DecValTok{4}\NormalTok{, }\StringTok{"Participants"} \OtherTok{=} \DecValTok{4}\NormalTok{, }\StringTok{"Timepoints"} \OtherTok{=} \DecValTok{4}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable\_classic}\NormalTok{(}
    \AttributeTok{full\_width =}\NormalTok{ F,}
    \AttributeTok{lightable\_options =} \StringTok{"hover"}\NormalTok{,}
    \AttributeTok{html\_font =} \StringTok{"Cambria"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:missingness filtering}Missingness Info by Study}
\centering
\begin{tabular}[t]{ccccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Measurements} & \multicolumn{4}{c}{Participants} & \multicolumn{4}{c}{Timepoints} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13}
Study & Full & Reduced & $\Delta$ & % & Full & Reduced & $\Delta$ & % & Full & Reduced & $\Delta$ & %\\
\midrule
1 & 1225 & 1204 & 21 & 1.71 & 23 & 21 & 2 & 8.70 & 63 & 61 & 2 & 3.17\\
2 & 4965 & 4218 & 747 & 15.05 & 113 & 86 & 27 & 23.89 & 69 & 60 & 9 & 13.04\\
3 & 4107 & 2710 & 1397 & 34.02 & 71 & 50 & 21 & 29.58 & 155 & 67 & 88 & 56.77\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Data}

To illustrate the

\subsubsection{Participants}

\subsubsection{Timepoints}

\subsubsection{Variables}

\subsection{Procedure}

feature engineering: feature extraction + feature selection

\subsubsection{Feature-Extraction}

use domain knowledge to extract new variables from raw data (summarize)

options: tsfresh \citep[][]{christ2018}, spikes
\citep[][]{caro-martin2018}

Our selection: density, inertia, instability, diversity
\citep[][]{dejonckheere2019, kuppens2017}

\begin{itemize}
  \item 
\end{itemize}

\subsubsection{Feature Selection}

select most consistent, relevant, and non-redundant features.

\textit{k} features means \(2^k – 1\) possible models

maximize relevance and minimize redundancy

``However, most of the existing UFS methods primarily focus on the
significance of features in maintaining the data structure while
ignoring the redundancy among features. Moreover, the determination of
the proper number of features is another challenge.''

\begin{itemize}
  \item feature selection
  \begin{itemize}
    \item filter: evaluate individual features' impact / importance / relevance. Univariate (ranking-based) or multivariate (preferred bc. can handle redundant and irrelevant features)
    \begin{itemize}
      \item unvariate, often based on information or spectral analysis (univariate inclusion test \citep[i.e., alpha-corrected t-tests][]{christ2018})
    \end{itemize}
    \item wrapper: search problem --- compare different models based on evaluation criterion (i.e., model performance; e.g., $R^2$, accuracy, ...). "high computational cost, and they are limited to be used in conjunction with a particular clustering algorithm."
    \begin{itemize}
      \item Exhaustive Feature Selection (i.e., brute force)
      \item forward selection (e.g., greedy Forward Search \citep[][]{wang2006})
      \item backward elimination
      \item Bi-directional elimination (Stepwise Selection)
    \end{itemize}
    \item hybrid: first filter then wrapper
    \item [embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance] MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
  \end{itemize}
  \item dimension reduction
\end{itemize}

random forest search \citep[][]{}

### 2022-11-21 methods-and-results.Rmd
hide code

@@
 
-```{r Preparation of Variable Names}
+```{r Preparation of Variable Names, include=FALSE}
 varNamIndicesS1 <- c(
@@
 
-```{r variable names for all analyses}
+```{r variable names for all analyses, include=FALSE}
 # ID Variables (just to re-iterate)
@@
 
-```{r variable names by response type and study}
+```{r variable names by response type and study, include=FALSE}
 varNames %>%
@@
 
-```{r missingness filtering}
+```{r missingness filtering, include=FALSE}
 missInfo <- function(full = NA, reduced = NA){
@@
 
-```{r}
+```{r, include=FALSE}
 

### 2022-11-21 methods-and-results.tex
hide code

@@
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{varNamIndicesS1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"autonomy"}\NormalTok{,}
-  \StringTok{"competence"}\NormalTok{,}
-  \StringTok{"relatedness"}\NormalTok{,}
-  \StringTok{"education\_level.pre"}\NormalTok{,}
-  \StringTok{"associationMerged.pre"}\NormalTok{,}
-  \StringTok{"assimilation.pre"}\NormalTok{,}
-  \StringTok{"separation.pre"}\NormalTok{,}
-  \StringTok{"integration.pre"}\NormalTok{,}
-  \StringTok{"marginalization.pre"}\NormalTok{,}
-  \StringTok{"VIA\_heritage.pre"}\NormalTok{,}
-  \StringTok{"VIA\_Dutch.pre"}\NormalTok{,}
-  \StringTok{"SSAS\_surrounding.pre"}\NormalTok{,}
-  \StringTok{"SSAS\_privat.pre"}\NormalTok{,}
-  \StringTok{"SSAS\_public.pre"}\NormalTok{,}
-  \StringTok{"assimilation.post"}\NormalTok{,}
-  \StringTok{"separation.post"}\NormalTok{,}
-  \StringTok{"integration.post"}\NormalTok{,}
-  \StringTok{"marginalization.post"}\NormalTok{,}
-  \StringTok{"VIA\_heritage.post"}\NormalTok{,}
-  \StringTok{"VIA\_Dutch.post"}\NormalTok{,}
-  \StringTok{"rosenberg.post"}\NormalTok{,}
-  \StringTok{"social\_support.post"}\NormalTok{,}
-  \StringTok{"stress.post"}\NormalTok{,}
-  \StringTok{"discrimination.post"}\NormalTok{,}
-  \StringTok{"discrimination\_month.post"}\NormalTok{,}
-  \StringTok{"NLE\_1month.post"}\NormalTok{,}
-  \StringTok{"NLE\_6month.post"}\NormalTok{,}
-  \StringTok{"NLE\_12month.post"}
-\NormalTok{  )}
-
-\NormalTok{varNamS123Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}\NormalTok{,}
-\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}\NormalTok{,}
-\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-\NormalTok{varNamS1Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-\NormalTok{varNamS2Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-\NormalTok{varNamS3Aux }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    aux }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\NormalTok{varNamS123PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}\NormalTok{,}
-\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}\NormalTok{,}
-\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-\NormalTok{varNamS1PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-\NormalTok{varNamS2PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS2 }\SpecialCharTok{==} \StringTok{"S2"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  pull}
-\NormalTok{varNamS3PCA }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    studyS3 }\SpecialCharTok{==} \StringTok{"S3"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\NormalTok{varNamOut }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"ResponseId"}\NormalTok{,}
-  \StringTok{"relatednessNoInteraction"}\NormalTok{,}
-  \StringTok{"relatednessSelf"}\NormalTok{,}
-  \StringTok{"relatednessOther"}\NormalTok{,}
-  \StringTok{"autonomy\_Int"}\NormalTok{,}
-  \StringTok{"autonomy\_NoInt"}\NormalTok{,}
-  \StringTok{"competence\_Int"}\NormalTok{,}
-  \StringTok{"competence\_NoInt"}\NormalTok{,}
-\NormalTok{  varNamS1PCA[}\FunctionTok{grepl}\NormalTok{(}\StringTok{\textquotesingle{}\^{}MDMQ\textquotesingle{}}\NormalTok{, varNamS1PCA)],}
-\NormalTok{  varNamS2PCA[}\FunctionTok{grepl}\NormalTok{(}\StringTok{\textquotesingle{}\^{}ProSo|\^{}AntiSo\textquotesingle{}}\NormalTok{, varNamS2PCA)],}
-\NormalTok{  varNamS3PCA[}\FunctionTok{grepl}\NormalTok{(}
-    \StringTok{\textquotesingle{}\^{}ProSo|\^{}AntiSo|\^{}agency|\^{}autoFrust|\^{}autoSat|\^{}relatFrust|\^{}relatSat|\^{}compFrust|\^{}compSat|\^{}lonely[0{-}9]|\^{}emotRegPos|\^{}emotRegNeg\textquotesingle{}}\NormalTok{,}
-\NormalTok{    varNamS3PCA}
-\NormalTok{  )]}
-\NormalTok{)}
-\NormalTok{varNamCore }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"PID"}\NormalTok{,}
-  \StringTok{"TID"}\NormalTok{,}
-  \StringTok{"TIDnum"}
-\NormalTok{)}
-\NormalTok{varNamNewAll }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"relatedness"}\NormalTok{,}
-  \StringTok{"autonomy"}\NormalTok{,}
-  \StringTok{"competence"}
-\NormalTok{)}
-\NormalTok{varNamNewS1 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"relatedness"}\NormalTok{,}
-  \StringTok{"autonomy"}\NormalTok{,}
-  \StringTok{"competence"}\NormalTok{,}
-  \StringTok{"alertness"}\NormalTok{, }
-  \StringTok{"calmness"}\NormalTok{, }
-  \StringTok{"valence"}
-\NormalTok{)}
-\NormalTok{varNamNewS2 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"relatedness"}\NormalTok{,}
-  \StringTok{"autonomy"}\NormalTok{,}
-  \StringTok{"competence"}\NormalTok{,}
-  \StringTok{"ProSo"}\NormalTok{, }
-  \StringTok{"AntiSo"}
-\NormalTok{)}
-\NormalTok{varNamNewS3 }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"relatedness"}\NormalTok{,}
-  \StringTok{"autonomy"}\NormalTok{,}
-  \StringTok{"competence"}\NormalTok{,}
-  \StringTok{"ProSo"}\NormalTok{, }
-  \StringTok{"AntiSo"}\NormalTok{,}
-  \StringTok{"agency"}\NormalTok{,}
-  \StringTok{"autoFrust"}\NormalTok{,}
-  \StringTok{"autoSat"}\NormalTok{,}
-  \StringTok{"relatFrust"}\NormalTok{,}
-  \StringTok{"relatSat"}\NormalTok{,}
-  \StringTok{"compFrust"}\NormalTok{,}
-  \StringTok{"compSat"}\NormalTok{,}
-  \StringTok{"lonely"}\NormalTok{,}
-  \StringTok{"emotRegPos"}\NormalTok{,}
-  \StringTok{"emotRegNeg"}
-\NormalTok{)}
-
-\NormalTok{varNamNewCat }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"closeness\_Calc"}\NormalTok{,}
-  \StringTok{"gender\_Calc"}\NormalTok{,}
-  \StringTok{"ethnicity\_Calc"}\NormalTok{,}
-  \StringTok{"relationship\_Calc"}
-\NormalTok{)}
-
-\NormalTok{varNamIntDep }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-\NormalTok{  varNamS123PCA[}\FunctionTok{grepl}\NormalTok{(}\StringTok{\textquotesingle{}\^{}InteractionContext|AttitudesPartner|KeyNeedDueToPartner|\^{}quality\textquotesingle{}}\NormalTok{, varNamS123PCA)]}
-\NormalTok{)}
-
-\NormalTok{varNamIndiceItemsS1 }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    aux }\SpecialCharTok{==} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}
-\NormalTok{    studyS1 }\SpecialCharTok{==} \StringTok{"S1"}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{append}\NormalTok{(., }\FunctionTok{gsub}\NormalTok{(}\StringTok{".pre|.post|\_calc"}\NormalTok{, }\StringTok{""}\NormalTok{, varNamIndicesS1)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  unique}
-
-\NormalTok{varNamS123MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewAll, varNamS123Aux)}
-\NormalTok{varNamS1MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamIndicesS1, varNamS1Aux[}\SpecialCharTok{!}\NormalTok{varNamS1Aux }\SpecialCharTok{\%in\%}\NormalTok{ varNamIndiceItemsS1]) }\CommentTok{\#}
-\NormalTok{varNamS2MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS2, varNamS2Aux)}
-\NormalTok{varNamS3MI }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS3, varNamS3Aux)}
-
-\NormalTok{varNamS123PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamS123PCA[}\SpecialCharTok{!}\NormalTok{varNamS123PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
-\NormalTok{varNamS1PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS1, varNamS1PCA[}\SpecialCharTok{!}\NormalTok{varNamS1PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
-\NormalTok{varNamS2PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS2, varNamS2PCA[}\SpecialCharTok{!}\NormalTok{varNamS2PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
-\NormalTok{varNamS3PCA }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(varNamCore, varNamNewS3, varNamS3PCA[}\SpecialCharTok{!}\NormalTok{varNamS3PCA }\SpecialCharTok{\%in\%}\NormalTok{ varNamOut])}
-
-\NormalTok{varNamS123MIPsbl }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \FunctionTok{paste}\NormalTok{(varNamS123MI, }\FunctionTok{rep}\NormalTok{(}\StringTok{".pre"}\NormalTok{, }\FunctionTok{length}\NormalTok{(varNamS123MI)), }\AttributeTok{sep =} \StringTok{""}\NormalTok{),}
-\NormalTok{  varNamS123MI,}
-  \FunctionTok{paste}\NormalTok{(varNamS123MI, }\FunctionTok{rep}\NormalTok{(}\StringTok{".post"}\NormalTok{, }\FunctionTok{length}\NormalTok{(varNamS123MI)), }\AttributeTok{sep =} \StringTok{""}\NormalTok{)}
-\NormalTok{)}
-
-\NormalTok{varNamS123MiRed }\OtherTok{\textless{}{-}} \FunctionTok{Reduce}\NormalTok{(}
-\NormalTok{  intersect,}
-  \FunctionTok{list}\NormalTok{(}
-\NormalTok{    dtS1Red }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(PID, TID, TIDnum, }\FunctionTok{any\_of}\NormalTok{(varNamS123MIPsbl)) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ names,}
-\NormalTok{    dtS2Red }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(PID, TID, TIDnum, }\FunctionTok{any\_of}\NormalTok{(varNamS123MIPsbl)) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ names,}
-\NormalTok{    dtS3Red }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(PID, TID, TIDnum, }\FunctionTok{any\_of}\NormalTok{(varNamS123MIPsbl)) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ names}
-\NormalTok{  )}
-\NormalTok{)}
-
-\NormalTok{idVars }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"PID"}\NormalTok{, }\StringTok{"TID"}\NormalTok{, }\StringTok{"TIDnum"}\NormalTok{, }\StringTok{"date"}\NormalTok{, }\StringTok{"week"}\NormalTok{, }\StringTok{"study"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# ID Variables (just to re{-}iterate)}
-\NormalTok{idVars }\OtherTok{\textless{}{-}}\NormalTok{ idVars}
-
-\CommentTok{\# Common variables across studies }
-\NormalTok{varNamS123 }\OtherTok{\textless{}{-}}\NormalTok{ varNamS123PCA[}\SpecialCharTok{!}\NormalTok{varNamS123PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}
-
-\CommentTok{\# Variable Groups}
-\CommentTok{\# varNames \%\textgreater{}\%}
-\CommentTok{\#   filter(pca != 0) \%\textgreater{}\%}
-\CommentTok{\#   mutate(contactSpecific = recode(.$contactSpecific, }
-\CommentTok{\#                          \textasciigrave{}{-}1\textasciigrave{}="No Interaction Only",}
-\CommentTok{\#                          \textasciigrave{}0\textasciigrave{}="unspecific",}
-\CommentTok{\#                          \textasciigrave{}1\textasciigrave{}="Interaction Only")) \%\textgreater{}\%}
-\CommentTok{\#   crosstable::crosstable(., varGroup, by=contactSpecific) \%\textgreater{}\%}
-\CommentTok{\#   select({-}c(.id, label)) \%\textgreater{}\%}
-\CommentTok{\#   kbl(.,}
-\CommentTok{\#       escape = FALSE,}
-\CommentTok{\#       booktabs = T,}
-\CommentTok{\#       align = c("l", "r", "c"),}
-\CommentTok{\#       digits=2,}
-\CommentTok{\#       caption = "Full Analysis: Variable Groups by Interaction Types") \%\textgreater{}\%}
-\CommentTok{\#   kable\_classic(}
-\CommentTok{\#     full\_width = F,}
-\CommentTok{\#     lightable\_options = "hover",}
-\CommentTok{\#     html\_font = "Cambria"}
-\CommentTok{\#   )}
-\CommentTok{\# }
-\CommentTok{\# varNames \%\textgreater{}\%}
-\CommentTok{\#   filter(}
-\CommentTok{\#     pca != 0}
-\CommentTok{\#   ) \%\textgreater{}\%}
-\CommentTok{\#   mutate(contactSpecific = recode(.$contactSpecific, }
-\CommentTok{\#                          \textasciigrave{}{-}1\textasciigrave{}="No Interaction Only",}
-\CommentTok{\#                          \textasciigrave{}0\textasciigrave{}="unspecific",}
-\CommentTok{\#                          \textasciigrave{}1\textasciigrave{}="Interaction Only")) \%\textgreater{}\%}
-\CommentTok{\#   select(varNam, varGroup, contactSpecific) \%\textgreater{}\%}
-\CommentTok{\#   group\_by(varGroup, varNam, contactSpecific) \%\textgreater{}\%}
-\CommentTok{\#   summarise(n = n()) \%\textgreater{}\%}
-\CommentTok{\#   pivot\_wider(names\_from = c("varGroup", "varNam"), values\_from = n, names\_sep=" {-} ") \%\textgreater{}\%}
-\CommentTok{\#   ungroup \%\textgreater{}\%}
-\CommentTok{\#   t \%\textgreater{}\%}
-\CommentTok{\#   as.data.frame \%\textgreater{}\%}
-\CommentTok{\#   janitor::row\_to\_names(1) \%\textgreater{}\%}
-\CommentTok{\#   replace(is.na(.), 0) \%\textgreater{}\%}
-\CommentTok{\#   tibble::rownames\_to\_column(., var = "variable") \%\textgreater{}\%}
-\CommentTok{\#   separate(variable, c("concept", "variable"), " {-} ") \%\textgreater{}\%}
-\CommentTok{\#   arrange(concept, unspecific, \textasciigrave{}Interaction Only\textasciigrave{}) \%\textgreater{}\%}
-\CommentTok{\#   mutate\_all(\textasciitilde{}recode(.x, \textasciigrave{}0\textasciigrave{}="", \textasciigrave{} 1\textasciigrave{}="✓")) \%\textgreater{}\%}
-\CommentTok{\#   kbl(.,}
-\CommentTok{\#       escape = FALSE,}
-\CommentTok{\#       booktabs = T,}
-\CommentTok{\#       align = c("l", "l", rep("c", ncol(.){-}2)),}
-\CommentTok{\#       digits=2,}
-\CommentTok{\#       caption = "All Possible Variables by Interaction Types") \%\textgreater{}\%}
-\CommentTok{\#   kable\_classic(}
-\CommentTok{\#     full\_width = F,}
-\CommentTok{\#     lightable\_options = "hover",}
-\CommentTok{\#     html\_font = "Cambria"}
-\CommentTok{\#   ) \%\textgreater{}\%}
-\CommentTok{\#   scroll\_box(width = "100\%", height = "500px")}
-
-\CommentTok{\# Non{-}Interaction Specific Variables}
-\NormalTok{varNamS123NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS123PCA,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# Interaction Specific Variables}
-\NormalTok{varNamS123Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS123PCA,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# S1 variables}
-\NormalTok{varNamS1Clust }\OtherTok{\textless{}{-}}\NormalTok{ varNamS1PCA[}\SpecialCharTok{!}\NormalTok{varNamS1PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}
-
-\CommentTok{\# S1 Non{-}Interaction Specific Variables}
-\NormalTok{varNamS1NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS1Clust,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# S1 Interaction Specific Variables}
-\NormalTok{varNamS1Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS1Clust,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# S2 variables}
-\NormalTok{varNamS2Clust }\OtherTok{\textless{}{-}}\NormalTok{ varNamS2PCA[}\SpecialCharTok{!}\NormalTok{varNamS2PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}
-
-\CommentTok{\# S2 Non{-}Interaction Specific Variables}
-\NormalTok{varNamS2NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS2Clust,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# S2 Interaction Specific Variables}
-\NormalTok{varNamS2Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS2Clust,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# S3 variables}
-\NormalTok{varNamS3Clust }\OtherTok{\textless{}{-}}\NormalTok{ varNamS3PCA[}\SpecialCharTok{!}\NormalTok{varNamS3PCA }\SpecialCharTok{\%in\%}\NormalTok{ idVars]}
-
-\CommentTok{\# S3 Non{-}Interaction Specific Variables}
-\NormalTok{varNamS3NoInt }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS3Clust,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{0}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-
-\CommentTok{\# S3 Interaction Specific Variables}
-\NormalTok{varNamS3Int }\OtherTok{\textless{}{-}}\NormalTok{ varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%}\NormalTok{ varNamS3Clust,}
-\NormalTok{    contactSpecific }\SpecialCharTok{==} \DecValTok{1}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(varNam) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  pull}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{varNames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(}
-\NormalTok{    pca }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{,}
-\NormalTok{    varNam }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{contactSpecific =} \FunctionTok{recode}\NormalTok{(}
-\NormalTok{      .}\SpecialCharTok{$}\NormalTok{contactSpecific,}
-      \StringTok{\textasciigrave{}}\AttributeTok{{-}1}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"No Interaction Only"}\NormalTok{,}
-      \StringTok{\textasciigrave{}}\AttributeTok{0}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"unspecific"}\NormalTok{,}
-      \StringTok{\textasciigrave{}}\AttributeTok{1}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"Interaction Only"}
-\NormalTok{    ),}
-    \AttributeTok{study1 =} \FunctionTok{ifelse}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(studyS1), }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-    \AttributeTok{study2 =} \FunctionTok{ifelse}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(studyS2), }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-    \AttributeTok{study3 =} \FunctionTok{ifelse}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(studyS3), }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{)}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{select}\NormalTok{(varNam, varGroup, contactSpecific, study1, study2, study3) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(varGroup, varNam, contactSpecific, study1, study2, study3) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{n =} \FunctionTok{n}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =} \FunctionTok{c}\NormalTok{(}\StringTok{"varGroup"}\NormalTok{, }\StringTok{"varNam"}\NormalTok{, }\StringTok{"study1"}\NormalTok{, }\StringTok{"study2"}\NormalTok{, }\StringTok{"study3"}\NormalTok{), }\AttributeTok{values\_from =}\NormalTok{ n, }\AttributeTok{names\_sep=}\StringTok{" {-} "}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  ungroup }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  t }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  as.data.frame }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  janitor}\SpecialCharTok{::}\FunctionTok{row\_to\_names}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{replace}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(.), }\DecValTok{0}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  tibble}\SpecialCharTok{::}\FunctionTok{rownames\_to\_column}\NormalTok{(., }\AttributeTok{var =} \StringTok{"variable"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{separate}\NormalTok{(variable, }\FunctionTok{c}\NormalTok{(}\StringTok{"concept"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"study1"}\NormalTok{, }\StringTok{"study2"}\NormalTok{, }\StringTok{"study3"}\NormalTok{), }\StringTok{" {-} "}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{c}\NormalTok{(study1, study2, study3), as.numeric)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{arrange}\NormalTok{(concept, unspecific, }\StringTok{\textasciigrave{}}\AttributeTok{Interaction Only}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate\_all}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\FunctionTok{recode}\NormalTok{(.x, }\StringTok{\textasciigrave{}}\AttributeTok{0}\StringTok{\textasciigrave{}}\OtherTok{=}\StringTok{""}\NormalTok{, }\StringTok{\textasciigrave{}}\AttributeTok{ 1}\StringTok{\textasciigrave{}}\OtherTok{=}\StringTok{"✓"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{relocate}\NormalTok{(}\FunctionTok{starts\_with}\NormalTok{(}\StringTok{"study"}\NormalTok{), }\AttributeTok{.after =} \FunctionTok{last\_col}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{rename\_with}\NormalTok{(., }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gsub}\NormalTok{(}\StringTok{"study"}\NormalTok{, }\StringTok{"Study "}\NormalTok{, .x, }\AttributeTok{fixed =} \ConstantTok{TRUE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{kbl}\NormalTok{(.,}
-      \AttributeTok{escape =} \ConstantTok{FALSE}\NormalTok{,}
-      \AttributeTok{booktabs =}\NormalTok{ T,}
-      \AttributeTok{align =} \FunctionTok{c}\NormalTok{(}\StringTok{"l"}\NormalTok{, }\StringTok{"l"}\NormalTok{, }\FunctionTok{rep}\NormalTok{(}\StringTok{"c"}\NormalTok{, }\FunctionTok{ncol}\NormalTok{(.)}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{)),}
-      \AttributeTok{digits=}\DecValTok{2}\NormalTok{,}
-      \AttributeTok{caption =} \StringTok{"Variables by Interaction Types and Study Availability"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_header\_above}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{" "} \OtherTok{=} \DecValTok{2}\NormalTok{, }\StringTok{"Contact Type"} \OtherTok{=} \DecValTok{2}\NormalTok{, }\StringTok{"Study Availability"} \OtherTok{=} \DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{kable\_classic}\NormalTok{(}
-    \AttributeTok{full\_width =}\NormalTok{ F,}
-    \AttributeTok{lightable\_options =} \StringTok{"hover"}\NormalTok{,}
-    \AttributeTok{html\_font =} \StringTok{"Cambria"}
-\NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{table}
-
-\caption{\label{tab:variable names by response type and study}Variables by Interaction Types and Study Availability}
-\centering
-\begin{tabular}[t]{llccccc}
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Contact Type} & \multicolumn{3}{c}{Study Availability} \\
-\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-7}
-concept & variable & unspecific & Interaction Only & Study 1 & Study 2 & Study 3\\
-\midrule
-acculturation & AttitudesDutch & ✓ &  & ✓ & ✓ & ✓\\
-acculturation & AttitudesPartner &  & ✓ & ✓ & ✓ & ✓\\
-contactQuality & InteractionContextAccidental &  & ✓ & ✓ & ✓ & ✓\\
-contactQuality & InteractionContextCooperative &  & ✓ & ✓ & ✓ & ✓\\
-contactQuality & InteractionContextEqualStatus &  & ✓ &  & ✓ & ✓\\
-\addlinespace
-contactQuality & InteractionContextRepresentativeNL &  & ✓ & ✓ & ✓ & ✓\\
-contactQuality & InteractionContextvoluntary &  & ✓ & ✓ & ✓ & ✓\\
-contactQuality & qualityEffective &  & ✓ &  & ✓ & ✓\\
-contactQuality & qualityImportance &  & ✓ &  & ✓ & ✓\\
-contactQuality & qualityMeaning &  & ✓ & ✓ & ✓ & ✓\\
-\addlinespace
-contactQuality & qualityOverall &  & ✓ & ✓ & ✓ & ✓\\
-contactQuality & qualityStar &  & ✓ & ✓ &  & \\
-emotion & afraid & ✓ &  &  &  & ✓\\
-emotion & angry & ✓ &  &  &  & ✓\\
-emotion & energy & ✓ &  &  &  & ✓\\
-\addlinespace
-emotion & exWB & ✓ &  & ✓ & ✓ & ✓\\
-emotion & lonelyAlways & ✓ &  &  &  & ✓\\
-needs & autonomy & ✓ &  &  & ✓ & ✓\\
-needs & competence & ✓ &  &  & ✓ & ✓\\
-needs & DaytimeNeedFulfillment & ✓ &  & ✓ & ✓ & ✓\\
-\addlinespace
-needs & DaytimeNeedImportance & ✓ &  &  &  & ✓\\
-needs & KeyNeedFulfillment & ✓ &  & ✓ & ✓ & ✓\\
-needs & KeyNeedImp & ✓ &  &  &  & ✓\\
-needs & StudentGoal01 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal02 & ✓ &  &  & ✓ & ✓\\
-\addlinespace
-needs & StudentGoal03 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal04 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal05 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal06 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal07 & ✓ &  &  & ✓ & ✓\\
-\addlinespace
-needs & StudentGoal08 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal09 & ✓ &  &  & ✓ & ✓\\
-needs & StudentGoal10 & ✓ &  &  & ✓ & ✓\\
-needs & KeyNeedDueToPartner &  & ✓ & ✓ & ✓ & ✓\\
-needs & KeyNeedShared &  & ✓ &  & ✓ & ✓\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{missInfo }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{full =} \ConstantTok{NA}\NormalTok{, }\AttributeTok{reduced =} \ConstantTok{NA}\NormalTok{)\{}
-\NormalTok{  missTab }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{nFull =} \FunctionTok{nrow}\NormalTok{(full),}
-    \AttributeTok{nRed =} \FunctionTok{nrow}\NormalTok{(reduced),}
-    \AttributeTok{pptFull =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(full}\SpecialCharTok{$}\NormalTok{PID)),}
-    \AttributeTok{pptRed =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(reduced}\SpecialCharTok{$}\NormalTok{PID)),}
-    \AttributeTok{timeFull =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(full}\SpecialCharTok{$}\NormalTok{TID)),}
-    \AttributeTok{timeRed =} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(reduced}\SpecialCharTok{$}\NormalTok{TID))}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}
-      \AttributeTok{nDif =}\NormalTok{ nFull}\SpecialCharTok{{-}}\NormalTok{nRed,}
-      \AttributeTok{nDifPerc =}\NormalTok{ nDif}\SpecialCharTok{/}\NormalTok{nFull}\SpecialCharTok{*}\DecValTok{100}\NormalTok{,}
-      \AttributeTok{pptDif =}\NormalTok{ pptFull}\SpecialCharTok{{-}}\NormalTok{pptRed,}
-      \AttributeTok{pptDifPerc =}\NormalTok{ pptDif}\SpecialCharTok{/}\NormalTok{pptFull}\SpecialCharTok{*}\DecValTok{100}\NormalTok{,}
-      \AttributeTok{timeDif =}\NormalTok{ timeFull}\SpecialCharTok{{-}}\NormalTok{timeRed,}
-      \AttributeTok{timeDifPerc =}\NormalTok{ timeDif}\SpecialCharTok{/}\NormalTok{timeFull}\SpecialCharTok{*}\DecValTok{100}
-\NormalTok{    ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{select}\NormalTok{(}
-\NormalTok{      nFull, nRed, nDif, nDifPerc,}
-\NormalTok{      pptFull, pptRed, pptDif, pptDifPerc,}
-\NormalTok{      timeFull, timeRed, timeDif, timeDifPerc}
-\NormalTok{    )}
-\NormalTok{    missTab}
-\NormalTok{\}}
-
-\NormalTok{missS1 }\OtherTok{\textless{}{-}} \FunctionTok{missInfo}\NormalTok{(}\AttributeTok{full =}\NormalTok{ dtS1}\SpecialCharTok{$}\NormalTok{full, }\AttributeTok{reduced =}\NormalTok{ dtS1Red)}
-\NormalTok{missS2 }\OtherTok{\textless{}{-}} \FunctionTok{missInfo}\NormalTok{(}\AttributeTok{full =}\NormalTok{ dtS2}\SpecialCharTok{$}\NormalTok{full, }\AttributeTok{reduced =}\NormalTok{ dtS2Red)}
-\NormalTok{missS3 }\OtherTok{\textless{}{-}} \FunctionTok{missInfo}\NormalTok{(}\AttributeTok{full =}\NormalTok{ dtS3}\SpecialCharTok{$}\NormalTok{full, }\AttributeTok{reduced =}\NormalTok{ dtS3Red)}
-
-\FunctionTok{rbind}\NormalTok{(}
-\NormalTok{  missS1,}
-\NormalTok{  missS2,}
-\NormalTok{  missS3}
-\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{study =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(study, }\FunctionTok{everything}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{kbl}\NormalTok{(.,}
-      \AttributeTok{escape =} \ConstantTok{FALSE}\NormalTok{,}
-      \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
-      \AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\CommentTok{\#c("l", rep("c", ncol(.){-}1)),}
-      \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Study"}\NormalTok{, }\FunctionTok{rep}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"Full"}\NormalTok{,}\StringTok{"Reduced"}\NormalTok{,}\StringTok{"$}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{Delta$"}\NormalTok{, }\StringTok{"\%"}\NormalTok{), }\DecValTok{3}\NormalTok{)),}
-      \AttributeTok{digits=}\DecValTok{2}\NormalTok{,}
-      \AttributeTok{caption =} \StringTok{"Missingness Info by Study"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{add\_header\_above}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{" "} \OtherTok{=} \DecValTok{1}\NormalTok{, }\StringTok{"Measurements"} \OtherTok{=} \DecValTok{4}\NormalTok{, }\StringTok{"Participants"} \OtherTok{=} \DecValTok{4}\NormalTok{, }\StringTok{"Timepoints"} \OtherTok{=} \DecValTok{4}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{kable\_classic}\NormalTok{(}
-    \AttributeTok{full\_width =}\NormalTok{ F,}
-    \AttributeTok{lightable\_options =} \StringTok{"hover"}\NormalTok{,}
-    \AttributeTok{html\_font =} \StringTok{"Cambria"}
-\NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{table}
-
-\caption{\label{tab:missingness filtering}Missingness Info by Study}
-\centering
-\begin{tabular}[t]{ccccccccccccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Measurements} & \multicolumn{4}{c}{Participants} & \multicolumn{4}{c}{Timepoints} \\
-\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13}
-Study & Full & Reduced & $\Delta$ & % & Full & Reduced & $\Delta$ & % & Full & Reduced & $\Delta$ & %\\
-\midrule
-1 & 1225 & 1204 & 21 & 1.71 & 23 & 21 & 2 & 8.70 & 63 & 61 & 2 & 3.17\\
-2 & 4965 & 4218 & 747 & 15.05 & 113 & 86 & 27 & 23.89 & 69 & 60 & 9 & 13.04\\
-3 & 4107 & 2710 & 1397 & 34.02 & 71 & 50 & 21 & 29.58 & 155 & 67 & 88 & 56.77\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \subsection{Data}

### 2022-11-21 methods-and-results.Rmd
clustering info added

@@
 \begin{itemize}
+  \item dimension reduction
   \item feature selection
@@
     \end{itemize}
-    \item hybrid: first filter then wrapper
+    \item hybrid: (usually) first filter then wrapper (e.g., Bayesian Filter KMeans)
     \item [embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance] MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
   \end{itemize}
-  \item dimension reduction
 \end{itemize}
@@
 
+\subsection{Clustering}
+
+\begin{itemize}
+  \item centroid-based (e.g., k-means)
+  \begin{itemize}
+    \item non-probabilistic
+    \item no assumptions
+    \item fast
+    \item sensitive to local minima
+  \end{itemize}
+  \item distribution-based (e.g., gaussian mixture model [GMM])
+  \begin{itemize}
+    \item probabilistic
+    \item (gaussian) distributions assumed
+    \item more flexible shape than vanilla centroid-based
+  \end{itemize}
+  \item density-based (e.g., dbscan)
+  \begin{itemize}
+    \item no shape assumption
+    \item do not assign outliers (i.e., outlier detection)
+  \end{itemize}
+  \item Hierarchical-based (e.g., Ward hierarchical clustering)
+  \begin{itemize}
+    \item divisive (top-down) and agglomerative (bottom-up)
+    \item no pre-specified number of clusters
+    \item easy visual inspection (dendogram)
+    \item best for small number of cases
+    \item no reversal of assignments
+  \end{itemize}
+  \item Hybrid (e.g, GMM with kmeans centroids as starting point)
+  \begin{itemize}
+    \item helps avoid individual shortcomings (e.g., local minima avoidance)
+  \end{itemize}
+\end{itemize}
+
 ```{r, include=FALSE}

### 2022-11-21 methods-and-results.tex
clustering info added

@@
 random forest search \citep[][]{}
+
+\subsection{Clustering}
+
+\begin{itemize}
+  \item centroid-based (e.g., k-means)
+  \begin{itemize}
+    \item non-probabilistic
+    \item no assumptions
+    \item fast
+    \item sensitive to local minima
+  \end{itemize}
+  \item distribution-based (e.g., gaussian mixture model [GMM])
+  \begin{itemize}
+    \item probabilistic
+    \item (gaussian) distributions assumed
+    \item more flexible shape than vanilla centroid-based
+  \end{itemize}
+  \item density-based (e.g., dbscan)
+  \begin{itemize}
+    \item no shape assumption
+    \item do not assign outliers (i.e., outlier detection)
+  \end{itemize}
+  \item Hierarchical-based (e.g., Ward hierarchical clustering)
+  \begin{itemize}
+    \item divisive (top-down) and agglomerative (bottom-up)
+    \item no pre-specified number of clusters
+    \item easy visual inspection (dendogram)
+    \item best for small number of cases
+    \item no reversal of assignments
+  \end{itemize}
+  \item Hybrid (e.g, GMM with kmeans centroids as starting point)
+  \begin{itemize}
+    \item helps avoid individual shortcomings (e.g., local minima avoidance)
+  \end{itemize}
+\end{itemize}

### 2022-11-21 methods-and-results.Rmd
feature extraction added

@@
 
-Our selection: density, inertia, instability, diversity \citep[][]{dejonckheere2019, kuppens2017}
+Our selection based on describing time series aspects relevant to human experiences \citep[density, inertia, instability, diversity;][]{dejonckheere2019, kuppens2017}
 
 \begin{itemize}
-  \item 
+  \item diversity: multiple variables
+  \begin{itemize}
+    \item field and concept specific: affect, behavior, cognition, desire
+  \end{itemize}
+  \item (in)stability
+  \begin{itemize}
+    \item mean of sum squared differences
+    \item mean absolute change
+  \end{itemize}
+  \item central tendency
+  \begin{itemize}
+    \item mean
+    \item median
+  \end{itemize}
+  \item variance
+  \begin{itemize}
+    \item variance / standard deviation
+    \item median absolute distance
+    \item (range / min + max)
+    \item (change in SD)
+    \item (variation coefficient (SE/mean))
+  \end{itemize}
+  \item inertia
+  \begin{itemize}
+    \item autocorrelation (lag-1)
+    \item (additional periodicity/seasonality (e.g., day, week))
+  \end{itemize}
+  \item linear trend
+  \begin{itemize}
+    \item OLS regression slope
+  \end{itemize}
+  \item nonlinear trend
+  \begin{itemize}
+    \item GAM spline coefficient: estimated degrees of freedom (edf; how wiggly the line is)
+    \item (number of spikes \citep[e.g.,][]{caro-martin2018})
+    \item (Fourier Transform Coefficients (representation in square- and sine waves))
+    \item (polynomial coefficients (Langevin model))
+    \item (semi-markov model parameters)
+  \end{itemize}
 \end{itemize}
@@
     \item hybrid: (usually) first filter then wrapper (e.g., Bayesian Filter KMeans)
-    \item [embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance] MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
+    \item (embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance) MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
   \end{itemize}

### 2022-11-21 methods-and-results.tex
feature extraction added

@@
 
-Our selection: density, inertia, instability, diversity
-\citep[][]{dejonckheere2019, kuppens2017}
+Our selection based on describing time series aspects relevant to human
+experiences
+\citep[density, inertia, instability, diversity;][]{dejonckheere2019, kuppens2017}
 
 \begin{itemize}
-  \item 
+  \item diversity: multiple variables
+  \begin{itemize}
+    \item field and concept specific: affect, behavior, cognition, desire
+  \end{itemize}
+  \item (in)stability
+  \begin{itemize}
+    \item mean of sum squared differences
+    \item mean absolute change
+  \end{itemize}
+  \item central tendency
+  \begin{itemize}
+    \item mean
+    \item median
+  \end{itemize}
+  \item variance
+  \begin{itemize}
+    \item variance / standard deviation
+    \item median absolute distance
+    \item (range / min + max)
+    \item (change in SD)
+    \item (variation coefficient (SE/mean))
+  \end{itemize}
+  \item inertia
+  \begin{itemize}
+    \item autocorrelation (lag-1)
+    \item (additional periodicity/seasonality (e.g., day, week))
+  \end{itemize}
+  \item linear trend
+  \begin{itemize}
+    \item OLS regression slope
+  \end{itemize}
+  \item nonlinear trend
+  \begin{itemize}
+    \item GAM spline coefficient: estimated degrees of freedom (edf; how wiggly the line is)
+    \item (number of spikes \citep[e.g.,][]{caro-martin2018})
+    \item (Fourier Transform Coefficients (representation in square- and sine waves))
+    \item (polynomial coefficients (Langevin model))
+    \item (semi-markov model parameters)
+  \end{itemize}
 \end{itemize}

### 2022-11-24 manuscript.tex
Updates from Overleaf
flowchart ✅
headings :construction:
extraction intro added :pencil:

@@
 
-\section{The Present Research}
+\section{Feature-based Time Series Clustering}
 
+\subsection{data preparation?}
+data cleaning + importantly, comparable time series:
+\begin{itemize}[noitemsep]
+    \item time-frame
+    \item response scale (e.g., standardize within clusters)
+    \item 
+\end{itemize}
+
+\subsection{variable selection}
+field- and concept specific selection of input variables.
+
+\subsection{feature extraction}
+aim: 
+hundreds of possible features to describe psychological time series.
+One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series.
+However, such an approach is not always advisable for psychological time series.
+For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 180 time series features.
+More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
+This is particularly the case when features have a conceptual/psychological meaning.
+Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
+But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
+We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.
+
+Fortunately, past conceptual and empirical efforts on
+
+\citep[variability, inertia, instability, diversity;][]{dejonckheere2019, kuppens2017}
+
+\subsection{dimensionality reduction}
+
+\subsection{clustering}
+
+\section{Empirical Implementation/Illustration}
 % Methods and Results from RMarkdown render
-%\input{Methods-and-Results}
-\section{Methods}
-To be written.
-
-\section{Results}
-To be written.
+\input{methods-and-results}
 
@@
 
+\begin{figure}
+  \caption{Flowchart Feature-Based Time Series Clustering in Psychology}
+  \label{fig:TSCFlow}
+  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/TimeSeriesClusterFlowSelection.pdf}
+  \caption*{Note: \\
+  Choices selected for illustration in this manuscript are marked in bold.}
+\end{figure}
+
 \printbibliography

### 2022-11-25 manuscript.tex
feature extraction :pencil:

@@
 
-\subsection{data preparation?}
+\subsection{Variable selection}
+field- and concept specific selection of input variables. Diverse
+
+\subsection{Data preparation?}
 data cleaning + importantly, comparable time series:
@@
     \item response scale (e.g., standardize within clusters)
-    \item 
 \end{itemize}
 
-\subsection{variable selection}
-field- and concept specific selection of input variables.
-
-\subsection{feature extraction}
+\subsection{Feature extraction}
 aim: 
@@
 
-Fortunately, past conceptual and empirical efforts on
+Fortunately, past conceptual and empirical efforts offer valuable discussions of common time series features in psychological research.
+While the final selection of features should always be driven by the research questions and field-specific conventions, we can build a practical toolbox of meaningful time series features for psychological data.
+In particular, we propose to focus on six features based on common psychological research questions and recent works on affect dynamics \citep[e.g.,][]{dejonckheere2019, kuppens2017}.
+An overview of the features, their substantive interpretations and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
 
-\citep[variability, inertia, instability, diversity;][]{dejonckheere2019, kuppens2017}
+The first two features are a person's \textit{central tendency} and \textit{variability}.
+Familiar statistics from the probability theory, the two features sit at the heart of many fundamental psychological questions.
+Are some people happier than others (i.e., difference in central tendency), and are people settled in their attitudes towards migrants or do they fluctuate over time (i.e., difference in variability)?
+Mathematically, both features have parametric and robust options to choose from (e.g., parametric: mean and standard deviation; robust: median and median absolute deviation).
 
-\subsection{dimensionality reduction}
+The third and fourth feature describe the structure of the variability within the time series.
+In particular, \textit{(in)stability} captures the average change between two consecutive measurements.
+Do changes tend to be slow and gradual or fast and abrupt?
+Mathematically, (in)stability measurements often look at the average change between two consecutive measurement.
+\textit{Inertia} describes how much a measurement carries over to future measurements.
+There are two main ways in which experiences tend to be connected to future experiences --- resistance to change and seasonality.
+Both forms of inertia relate to different types of research questions.
+Do patients stay in a depressed mood for several measurements (i.e., stability-inertia) and do participants drink more alcohol on Fridays (i.e., periodicity-inertia)?
+Both forms of inertia are often measured using autocorrelations with the either a lag-1 (stability-inertia) or the lag of the seasonality (e.g., seven days).
+Periodicity-inertia can additionally be captured using wave patterns (e.g., fourier or wavelet transformation).
+
+The final two features are linear and nonlinear trends.
+\textit{Linear trends} are a common research interest for longitudinal data.
+Do patient symptoms improve or does worker productivity decline, are familiar types of research questions.
+Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients.
+\textit{Nonlinear trends} are important in two regards.
+Firstly nonlinearity as a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
+Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}.
+The structure of nonlinear trends is often mathematically captured using polinomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to number of spikes \citealp[]{caro-martin2018}).
+
+This selection of the proposed six time series features is in no way exhaustive or imperative.
+Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options.
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice.
+
+\input{tables/esmPsychFeatures}
+
+\subsection{Feature reduction}
 

### 2022-11-25 tables/esmPsychFeatures.tex
feature extraction :pencil:

%\begin{table}%[hbt]
\begin{sidewaystable}
    \centering
    \caption{Examples of Features for Psychological Time Series.}
    \label{tab:esmFeatures} 
    \begin{tabular}{
    >{\raggedright\arraybackslash}p{0.15\linewidth} 
    >{\raggedright\arraybackslash}p{0.35\linewidth} 
    >{\raggedright\arraybackslash}p{0.25\linewidth} 
    >{\raggedright\arraybackslash}p{0.20\linewidth}
    }
        \hline 
        Time Series Feature & Substantive Interpretation & Mathematical Operationalizations & Formulas, refs, other? \\ 
        \hline \\ [-0.5em]
        Diversity \newline \hl{(drop this \& keep in var. selection only?)} & 
        Multivariate measurement of concepts (e.g., affect-behavior-cognition-desire, or bio-psycho-social) \linebreak & 
        -----\linebreak  & 
        {\centering --- ? ---\par} \\

        Central Tendency \linebreak & 
        Average level of the experience across the entire measurement period. \linebreak & 
        mean, \linebreak median, \linebreak mode, ... \linebreak  & 
        {\centering --- ? ---\par} \\

        Variability & 
        Describes the average deviation from the central tendency across the entire measurement period. \linebreak & 
        standard deviation, \linebreak variation coefficient, \linebreak median absolute deviation, ... \linebreak & 
        {\centering --- ? ---\par} \\

        (In)stability & 
        Describes the average change between two consecutive measurements of the experience. \linebreak & 
        mean sum squared differences, \linebreak mean absolute change, \linebreak Ix instability index, ... \linebreak & 
        {\centering --- ? ---\par} \\

        Inertia & 
        Describes how much experiences carry over to the future measurements. This includes resistance to change (i.e., carries over to the next measurement) and periodic or seasonal returns (e.g., self-predictive on a daily or weekly basis). \linebreak & 
        autocorrelation (e.g., lag–1), \linebreak fourier coefficients, \linebreak continuous wavelet transform, ... \linebreak & 
        {\centering --- ? ---\par} \\

        Linear Trend & 
        Describes upwards or downwards linear trend of the experience reports. \linebreak & 
        OLS regression slope, \linebreak avg. piecewise linear reg. slope, ... \linebreak & 
        {\centering --- ? ---\par} \\

        Nonlinearity & 
        Describes the nonlinear structure of the time series. This includes measures that indicate the deviation from the a linear trend as well as nonlinear model parameters. \linebreak & 
        GAM spline edf, \linebreak bicoherence metrics, \linebreak Langevin polinomial coefficient, ... \linebreak & 
        {\centering --- ? ---\par} \\

        \hline \\ [-0.75em]
        \multicolumn{4}{p{\linewidth}}{\footnotesize \textit{Note.} The presented features and operationalizations are neither exhaustive nor necessary for feature-based clustering.}
    \end{tabular}
\end{sidewaystable}

### 2022-11-25 tables/FeatureReduction.tex
feature red tbl :pencil:

%\begin{table}%[hbt]
\begin{sidewaystable}
    \centering
    \caption{Examples of Feature Reduction Approaches and Methods.}
    \label{tab:featureReduction} 
    \begin{tabular}{
    >{\raggedright\arraybackslash}p{0.08\linewidth} 
    >{\raggedright\arraybackslash}p{0.08\linewidth} 
    >{\raggedright\arraybackslash}p{0.54\linewidth} 
    >{\raggedright\arraybackslash}p{0.25\linewidth}
    }
        \hline 
        Approach & Method & Intuitive Description & Algorithm Examples \\ 
        \hline \\ [-0.5em]

        Selection \linebreak & 
        filter \linebreak & 
        Features are selected individually or jointly based selection criteria. One common selection criteria is the amount of information and (unique) variance a feature captures. Univariate methodologies are able identify irrelevant features (i.e., because features do not capture much information), multivariate methods additionally allow to remove redundant features (i.e., because features capture the same information). \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item univariate filter (e.g., Laplacian Score)
            \item multivariate filter (e.g., variance-covariance)
        \end{itemize}
         \linebreak \\

        \linebreak & 
        Wrapper \linebreak & 
        Wrapper methodologies run models with different feature combinations and compare performance. Because the selection process is essentially a search problem this method is computationally intensive. Traditionally wrappers have used forward selections or backwards eliminations but recently alternative approaches have been proposed based such as ant colony and swarm intelligence paradigms. \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item sequential (e.g., forward selection)
            \item bio-inspired (e.g., ELSA)
            \item iterative (e.g., feature salience)
        \end{itemize}
         \linebreak \\

        \linebreak & 
        hybrid \linebreak & 
        Hybrid selection methodological combine filter and wrapper methodologies to avoid the shortcomings of the individual methods. An initial filter step reduces computational effort (efficiency) and the wrapper ensures high performance (effectiveness). \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item filter + wrapper (e.g., Calinski-Harabasz Index)
        \end{itemize}
        \linebreak \\

        Projection \linebreak & 
        linear \linebreak & 
        Linear dimensionality reduction methodologies use linear transformations of the original data to stretch and shift the data in such a way that the data can be `projected' to a lower dimensional space without loosing too much information. These methods are well-established, tend to be fast, and usually do not need much conceptual input from the user. \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item Principal Component Analysis (PCA)
            \item Factor Analysis (FA)
            \item Linear Discriminant Analysis (LDA)
        \end{itemize}
         \linebreak \\

        \linebreak & 
        nonlinear \linebreak & 
        Nonlinear dimensionality reduction methods also seek to map high-dimensional data to a lower dimensional space. However, nonlinear methods have been developed to preserve the local and global structure of more complex multidimensional patterns.
        \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item t-distributed Stochastic Neighbor Embedding (t-SNE)
            \item Multidimensional Scaling (MDS)
            \item Isometric mapping (Isomap)
        \end{itemize}
        \linebreak \\

        \hline \\ [-0.75em]
        \multicolumn{4}{p{\linewidth}}{\footnotesize \textit{Note.} The presented dimensionality reduction methods and -approaches are neither exhaustive nor necessary for feature-based clustering.}
    \end{tabular}
\end{sidewaystable}

### 2022-11-25 manuscript.tex
feature red tbl :pencil:
fig flow n :construction:

@@
 
+\fgrref{fig:TSCFlow}
+
 \subsection{Variable selection}
@@
 In particular, we propose to focus on six features based on common psychological research questions and recent works on affect dynamics \citep[e.g.,][]{dejonckheere2019, kuppens2017}.
-An overview of the features, their substantive interpretations and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
+An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
 
@@
 \subsection{Feature reduction}
+Once researchers researchers have
+
+100 participants * 10 variables * 60 measurements (for an example see \fgrref{fig:TSCFlowN})
+
+\tblref{tab:featureReduction}
+
+\input{tables/FeatureReduction}
 
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
   \caption{Scopus ESM Development}
@@
 
-\begin{figure}
+\begin{figure*}[hbtp]
   \caption{Flowchart Feature-Based Time Series Clustering in Psychology}
@@
   Choices selected for illustration in this manuscript are marked in bold.}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+  \caption{Exemplary Flowchart of Data Points in Feature-Based Time Series Clustering}
+  \label{fig:TSCFlowN}
+  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustFlowN.pdf}
+  \caption*{Note: \\
+  The presented number of participants, variables, and measurement occasions are somewhat arbitrary but generally represent common sample sizes found within the literature.
+Also the number of extracted clusters is presented for illustrative purposes only}
+\end{figure*}
 

### 2022-11-25 manuscript.tex
feature reduction :pencil:

@@
 
-\subsection{Variable selection}
+\subsection{Input Variable}
 field- and concept specific selection of input variables. Diverse
 
-\subsection{Data preparation?}
 data cleaning + importantly, comparable time series:
@@
 \subsection{Feature reduction}
-Once researchers researchers have
+Once a meaningful selection of time series features has been extracted for each of variable and participant, the total number of data points usually remains too large for most clustering algorithms.
+As an example, a relative common sample size of 100 participants and 10 variables of interest, where 8 time series features are extracted, still consists of 80 features per participant (a total of 8,000 data points in this hypothetical example).
+We offer an illustration of the compounding numbers of data points in \fgrref{fig:TSCFlowN}.
+The difficulty of working with such large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
+To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection (a full overview of the approaches, methods, and common algorithms is available in \tblref{tab:featureReduction}).
 
-100 participants * 10 variables * 60 measurements (for an example see \fgrref{fig:TSCFlowN})
+Feature selections seeks to create a subset of the most important features.
+The many available approaches differ in how they seek to determine the importance of the individual features.
+Generally speaking, selection methods can be categorized as `filter', `wrapper', or `hybrid' methods.
+The filter methods, broadly speaking determine important features by identify irrelevant features (e.g., because features do not capture much information), and identify redundant features (e.g., because features capture the same information).
+Wrapper methods on the other hand avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features.
+A wrapper, thus, compares the performance of models with different feature combinations.
+Traditional examples of wrappers are forward selection or backward elimination procedures.
+Because filter methods might not always perform well and wrapper methods are computationally intensive, hybrid methods seek to combine the two methods and find a balance between computational effort (i.e., efficiency) and performance (i.e., effectiveness).
+As an example, some of the common methods use a filter step to reduce the size of features considered in a later wrapper step.
 
-\tblref{tab:featureReduction}
+The second general approach to the dimensionality curse have been the much more general feature projection methods.
+Projections methods are relatively common in psychological research --- particularly, factor- and principal component analyses.
+Generally speaking, projection methods seek to
+transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features (i.e., the data is projected to a lower dimensional space).
+Commonly, this is achieved using linear transformations (e.g., principal component analysis), or more complex nonlinear transformations (e.g., t-SNE).
+Generally speaking, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
+Feature projection methods, on the other had, tend to be more generalized and are more readily available.
 
@@
 
-\subsection{clustering}
+\subsection{Clustering}
 

### 2022-11-25 tables/FeatureReduction.tex
feature reduction :pencil:

@@
         \hline \\ [-0.75em]
-        \multicolumn{4}{p{\linewidth}}{\footnotesize \textit{Note.} The presented dimensionality reduction methods and -approaches are neither exhaustive nor necessary for feature-based clustering.}
+        \multicolumn{4}{p{\linewidth}}{\footnotesize \textit{Note.} The presented dimensionality reduction methods and -approaches are neither exhaustive nor necessary for feature-based clustering.
+Notable additional approaches are `embedded selection methods' that filter as part of the model estimation procedure (e.g., mixture models) and `network-based projection methods' that use neural networks to reduce dimensions (e.g., autoencoders).}
     \end{tabular}

### 2022-11-26 manuscript.tex
clusterTbl :construction:
clustertTbl :sparkles:

@@
 \begin{mdframed}[style=mdfbox]
-\noindent\center\textit{Argument Structure Introduction}:
-\begin{itemize}[noitemsep]
+\noindent\center\textit{Structure}:
+\begin{itemize}[nosep]
+    \item introduction
+    \begin{itemize}[nosep]
         \item\ [Relevance]: increase in number and variety of ESM data
@@
         \item\ [Proposal]: feature based clusterting = flexible +  transparent
-    \item\ [Section]: new type of ESM data and issues for analysis:
-    \begin{itemize}[noitemsep]
+    \end{itemize}
+    \item Modern ESM data: new type of data and issues for analysis:
+    \begin{itemize}[nosep]
         \item incomplete data (complete data models vs. structural missingness)
@@
     \end{itemize}
-    \item\ [Section]: time-series clustering
-    \begin{itemize}[noitemsep]
+    \item Time series clustering
+    \begin{itemize}[nosep]
         \item shape-based
@@
     \end{itemize}
-    \item\ [Section]: The present Research
+    \item Feature-Based Clustering
+    \begin{itemize}[nosep]
+        \item Input variables
+        \item feature extraction
+        \item feature reduction
+        \item feature clustering
+    \end{itemize}
+    \item Empirical Illustration
+    \item Discussion
 \end{itemize}
@@
 
-\section{Clustering}
+\section{Time Series Clustering}
 One promising approach to describing and exploring time-series data has been the idea of structuring the data using clustering analyses.
@@
 Many of the common time-series clustering algorithms used today have been developed within the machine learning literature and have commonly been separated into three different approaches: (1) shape-based, (2) model-based, and (3) feature-based clustering \citep{hautamaki2008, liao2005}.
-Shape-based clustering analyses use the raw time series and seeks to match time series by non-linearly stretching and contracting the time axes.
-The other two approaches do not use the the raw time series directly but convert the data first.
-The model-based clustering has been the most-commonly used method within the psychological literature and clusters the parameters of a parametric model fitted for each participant \citep[e.g.,][]{ernst2021}.
+Shape-based clustering analyses use the raw time series and seek to match time series by non-linearly stretching and contracting the time axes.
+The other two approaches do not use the raw time series directly but convert the data first.
+The model-based clustering has been the most commonly used method within the psychological literature and clusters the parameters of a parametric model fitted for each participant \citep[e.g.,][]{ernst2021}.
 The flexibility of such clustering is restricted to the assumptions of the model used, so that stationary models, for example, do not capture non-stationary data well.
@@
 To showcase the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
-Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time-series, but also linear and non-linear trends, inertia related features such as auto-correlations, or the stability of the time series (e.g., mean absolute change).
-Given the flexibility of feature-based approach, users should be able to apply the analysis across multivariate data, feature extraction should mitigate issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
+Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time series, but also linear and non-linear trends, inertia-related features such as auto-correlations, or the stability of the time series (e.g., mean absolute change).
+Given the flexibility of the feature-based approach, users should be able to apply the analysis across multivariate data, feature extraction should mitigate issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
 
@@
 
-\subsection{Input Variable}
+\subsection{Input variables}
 field- and concept specific selection of input variables. Diverse
@@
 
-The third and fourth feature describe the structure of the variability within the time series.
+The third and fourth features describe the structure of the variability within the time series.
 In particular, \textit{(in)stability} captures the average change between two consecutive measurements.
 Do changes tend to be slow and gradual or fast and abrupt?
-Mathematically, (in)stability measurements often look at the average change between two consecutive measurement.
+Mathematically, (in)stability measurements often look at the average change between two consecutive measurements.
 \textit{Inertia} describes how much a measurement carries over to future measurements.
@@
 Both forms of inertia relate to different types of research questions.
-Do patients stay in a depressed mood for several measurements (i.e., stability-inertia) and do participants drink more alcohol on Fridays (i.e., periodicity-inertia)?
-Both forms of inertia are often measured using autocorrelations with the either a lag-1 (stability-inertia) or the lag of the seasonality (e.g., seven days).
+Do patients stay in a depressed mood for several measurements (i.e., stability-inertia), and do participants drink more alcohol on Fridays (i.e., periodicity-inertia)?
+Both forms of inertia are often measured using autocorrelations with either a lag-1 (stability-inertia) or the lag of the seasonality (e.g., seven days).
 Periodicity-inertia can additionally be captured using wave patterns (e.g., fourier or wavelet transformation).
@@
 \textit{Nonlinear trends} are important in two regards.
-Firstly nonlinearity as a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
+Firstly nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
 Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}.
-The structure of nonlinear trends is often mathematically captured using polinomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to number of spikes \citealp[]{caro-martin2018}).
+The structure of nonlinear trends is often mathematically captured using polynomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
 
-This selection of the proposed six time series features is in no way exhaustive or imperative.
+This selection of the proposed six time-series features is in no way exhaustive or imperative.
 Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options.
@@
 \subsection{Feature reduction}
-Once a meaningful selection of time series features has been extracted for each of variable and participant, the total number of data points usually remains too large for most clustering algorithms.
-As an example, a relative common sample size of 100 participants and 10 variables of interest, where 8 time series features are extracted, still consists of 80 features per participant (a total of 8,000 data points in this hypothetical example).
+Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points usually remains too large for most clustering algorithms.
+As an example, a relatively common sample size of 100 participants and 10 variables of interest, where 8 time series features are extracted, still consists of 80 features per participant (a total of 8,000 data points in this hypothetical example).
 We offer an illustration of the compounding numbers of data points in \fgrref{fig:TSCFlowN}.
-The difficulty of working with such large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
+The difficulty of working with such a large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
 To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection (a full overview of the approaches, methods, and common algorithms is available in \tblref{tab:featureReduction}).
 
-Feature selections seeks to create a subset of the most important features.
+Feature selections seek to create a subset of the most important features.
 The many available approaches differ in how they seek to determine the importance of the individual features.
 Generally speaking, selection methods can be categorized as `filter', `wrapper', or `hybrid' methods.
-The filter methods, broadly speaking determine important features by identify irrelevant features (e.g., because features do not capture much information), and identify redundant features (e.g., because features capture the same information).
+The filter methods, broadly speaking determine important features by identifying irrelevant features (e.g., because features do not capture much information), and identifying redundant features (e.g., because features capture the same information).
 Wrapper methods on the other hand avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features.
@@
 
-The second general approach to the dimensionality curse have been the much more general feature projection methods.
+The second general approach to the dimensionality curse has been feature projection.
 Projections methods are relatively common in psychological research --- particularly, factor- and principal component analyses.
-Generally speaking, projection methods seek to
-transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features (i.e., the data is projected to a lower dimensional space).
+Generally speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features (i.e., the data is projected to a lower dimensional space).
 Commonly, this is achieved using linear transformations (e.g., principal component analysis), or more complex nonlinear transformations (e.g., t-SNE).
 Generally speaking, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
-Feature projection methods, on the other had, tend to be more generalized and are more readily available.
+Feature projection methods, on the other hand, tend to be more generalized and are more readily available.
 
@@
 
-\subsection{Clustering}
+\subsection{Feature clustering}
+For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible.
+The crux of clustering is, thus, to have clearly defined and effective measurements of (dis)similarity.
+Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment (or similarity measurement for qualitative features).
+
+For a more comprehensive overview see \citet{xu2015}
+
+\tblref{tab:clusterApproaches}
+
+\input{tables/clusterApproaches}
 

### 2022-11-26 tables/clusterApproaches.tex
clustertTbl :sparkles:

%\begin{table}%[hbt]
\begin{sidewaystable}
    \centering
    \caption{Common Clustering Approaches.}
    \label{tab:clusterApproaches} 
    \begin{tabular}{
    >{\raggedright\arraybackslash}p{0.15\linewidth} 
    >{\raggedright\arraybackslash}p{0.40\linewidth} 
    >{\raggedright\arraybackslash}p{0.28\linewidth} 
    >{\raggedright\arraybackslash}p{0.12\linewidth}
    }
        \hline 
        Approach & Description & Characteristics & Examples \\ 
        \hline \\ [-0.5em]

        centroid-based \linebreak & 
        Chooses a pre-defined number of potential cluster centers in the feature space and assigns participants to closest center. Then, iteratively, moves the centers until a convergence criterion is met (e.g., all distances to centers minimized).
        \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item[\scriptsize\faPlusCircle] simple and efficient
            \item[\scriptsize\faPlusCircle] no assumptions
            \item[\scriptsize\faPlusCircle] well implemented
            \item[\scriptsize\faMinusCircle] may struggle with complex shapes
            \item[\scriptsize\faMinusCircle] sensitive to initial values and outliers
            %\item[\scriptsize\faMinusCircle] not suitable for non-convex data, relatively sensitive to the outliers, easily drawn into local optimal, the number of clusters needed to be preset, and the clustering result sensitive to the number of clusters \citep{xu2015}.
        \end{itemize}\linebreak & 
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item k-means
            %\item fuzzy c-means
            %\item CLARA
            \item PAM
        \end{itemize}\linebreak \\

        distribution-based \linebreak & 
        Assumes that the data points belong to one of several specific distributions (e.g., Gaussian distributions). Data points that fit to a distribution-based expectation are given higher probability of belonging to that distribution. We can then iteratively check with which model parameters the data points best fit within given number of distributions.
        \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item[\scriptsize\faPlusCircle] probabilistic
            \item[\scriptsize\faPlusCircle] well supported
            \item[\scriptsize\faMinusCircle] time intensive
            \item[\scriptsize\faMinusCircle] distribution and parameter sensitive
        \end{itemize}\linebreak & 
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item GMM
            %\item negative binomial model-based
            \item DBCLASD
        \end{itemize}\linebreak\\

        density-based \linebreak & 
        Regions where several data points are relatively close together (i.e., high density) are assumed to be a cluster. Based on what is considered a dense region (e.g., radius of a region and minimum number of points the radius), points can be either be assigned to one of the clusters or be considered too far away from a dense region. \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item[\scriptsize\faPlusCircle] efficient
            \item[\scriptsize\faPlusCircle] no shape assumption
            \item[\scriptsize\faPlusCircle] do not assign outliers 
            \item[\scriptsize\faMinusCircle] may struggle with uneven densities
            \item[\scriptsize\faMinusCircle] sensitive to high dimensionality
        \end{itemize}\linebreak & 
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item DBSCAN
            \item OPTICS
        \end{itemize}\linebreak\\

        %graph-based \linebreak & 
        %``realized on the graph where the node is regarded as the data point and the edge is regarded as the relationship among data points'' \citep{xu2015} \linebreak &
        %\vspace{-1em}
        %\begin{itemize}[nosep,leftmargin=*,label={--}]
        %    \item[\scriptsize\faPlusCircle] efficient and accurate
        %    \item[\scriptsize\faPlusCircle] no shape assumption
        %    \item[\scriptsize\faMinusCircle] sensitive to graph complexity
        %\end{itemize}\linebreak & 
        %\vspace{-1em}
        %\begin{itemize}[nosep,leftmargin=*,label={--}]
        %    \item MST-based
        %    \item CLICK
        %\end{itemize}\linebreak\\

        hierarchy-based \linebreak & 
        Builds a hierarchy of cluster by step-wise combining the closest two clusters (bottom-up; agglomerative) or top down dividing the data into smaller clusters that maximize distances (top-down; divisive).
        \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item[\scriptsize\faPlusCircle] flexible number of clusters
            \item[\scriptsize\faPlusCircle] no shape assumption 
            \item[\scriptsize\faMinusCircle] small number of cases
            \item[\scriptsize\faMinusCircle] no reversal of assignments
        \end{itemize}\linebreak & 
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item Chameleon
            \item CURE
        \end{itemize}\linebreak \\

        hybrid \linebreak & 
        Usually combines different approaches, which combine the strength of the complementary approaches. Oftentimes the combination also increases efficiency. \linebreak &
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item[\scriptsize\faPlusCircle] avoids individual shortcomings
            \item[\scriptsize\faMinusCircle] less readily available
        \end{itemize}\linebreak & 
        \vspace{-1em}
        \begin{itemize}[nosep,leftmargin=*,label={--}]
            \item DD-means
            \item hybridHclust
        \end{itemize}\linebreak\\

        \hline \\ [-0.75em]
        \multicolumn{4}{p{\linewidth}}{\footnotesize \textit{Note.} The presented clustering approaches and algorithms are neither exhaustive nor necessary for feature-based clustering. Notably, recently innovations have been made based on graph-, fractal-, swarm, and quantum theory \citep[for a more in-depth review see][]{xu2015}.}
    \end{tabular}
\end{sidewaystable}

### 2022-11-26 tables/esmPsychFeatures.tex
clustertTbl :sparkles:

@@
         \hline 
-        Time Series Feature & Substantive Interpretation & Mathematical Operationalizations & Formulas, refs, other? \\ 
+        Time Series Feature & Substantive Interpretation & Example Operationalizations & Formulas, refs, ...? \\ 
         \hline \\ [-0.5em]
@@
         Average level of the experience across the entire measurement period. \linebreak & 
-        mean, \linebreak median, \linebreak mode, ... \linebreak  & 
+        \vspace{-1em}
+        \begin{itemize}[nosep,leftmargin=*,label={--}]
+            \item mean
+            \item median
+            \item mode
+        \end{itemize} \linebreak  & 
         {\centering --- ? ---\par} \\
@@
         Describes the average deviation from the central tendency across the entire measurement period. \linebreak & 
-        standard deviation, \linebreak variation coefficient, \linebreak median absolute deviation, ... \linebreak & 
+        \vspace{-1em}
+        \begin{itemize}[nosep,leftmargin=*,label={--}]
+            \item standard deviation
+            \item variation coefficient
+            \item median absolute deviation
+        \end{itemize} \linebreak & 
         {\centering --- ? ---\par} \\
@@
         Describes the average change between two consecutive measurements of the experience. \linebreak & 
-        mean sum squared differences, \linebreak mean absolute change, \linebreak Ix instability index, ... \linebreak & 
+        \vspace{-1em}
+        \begin{itemize}[nosep,leftmargin=*,label={--}]
+            \item mean sum squared differences
+            \item mean absolute change
+            \item Ix instability index
+        \end{itemize} \linebreak & 
         {\centering --- ? ---\par} \\
@@
 \linebreak &
-        autocorrelation (e.g., lag–1), \linebreak fourier coefficients, \linebreak continuous wavelet transform, ... \linebreak & 
+        \vspace{-1em}
+        \begin{itemize}[nosep,leftmargin=*,label={--}]
+            \item autocorrelation (e.g., lag–1)
+            \item fourier coefficients
+            \item continuous wavelet transform
+        \end{itemize} \linebreak & 
         {\centering --- ? ---\par} \\
@@
         Describes upwards or downwards linear trend of the experience reports. \linebreak & 
-        OLS regression slope, \linebreak avg. piecewise linear reg. slope, ... \linebreak & 
+        \vspace{-1em}
+        \begin{itemize}[nosep,leftmargin=*,label={--}]
+            \item OLS regression slope
+            \item avg. piecewise linear reg. slope
+        \end{itemize} \linebreak & 
         {\centering --- ? ---\par} \\
@@
         Describes the nonlinear structure of the time series. This includes measures that indicate the deviation from the a linear trend as well as nonlinear model parameters. \linebreak & 
-        GAM spline edf, \linebreak bicoherence metrics, \linebreak Langevin polinomial coefficient, ... \linebreak & 
+        \vspace{-1em}
+        \begin{itemize}[nosep,leftmargin=*,label={--}]
+            \item GAM spline edf
+            \item bicoherence metrics
+            \item Langevin polinomial coefficient
+        \end{itemize} \linebreak & 
         {\centering --- ? ---\par} \\

### 2022-11-26 methods-and-results.Rmd
:pencil:

--- 2022-11-21 methods-and-results.Rmd
@@
 
-\subsubsection{Feature-Extraction}
-
 use domain knowledge to extract new variables from raw data (summarize)
 
-options: tsfresh \citep[][]{christ2018}, spikes \citep[][]{caro-martin2018}
-
-Our selection: density, inertia, instability, diversity \citep[][]{dejonckheere2019, kuppens2017}
-
-\begin{itemize}
-  \item 
-\end{itemize}
-
-\subsubsection{Feature Selection}
-
-select most consistent, relevant, and non-redundant features.
+options: tsfresh \citep[][]{christ2018}
 
@@
 
-"However, most of the existing UFS methods primarily focus on the significance of features in maintaining the data structure while ignoring the redundancy among features.
-Moreover, the determination of the proper number of features is another challenge."
-
-\begin{itemize}
-  \item feature selection
-  \begin{itemize}
-    \item filter: evaluate individual features' impact / importance / relevance. Univariate (ranking-based) or multivariate (preferred bc. can handle redundant and irrelevant features)
-    \begin{itemize}
-      \item unvariate, often based on information or spectral analysis (univariate inclusion test \citep[i.e., alpha-corrected t-tests][]{christ2018})
-    \end{itemize}
-    \item wrapper: search problem --- compare different models based on evaluation criterion (i.e., model performance; e.g., $R^2$, accuracy, ...).
-"high computational cost, and they are limited to be used in conjunction with a particular clustering algorithm."
-    \begin{itemize}
-      \item Exhaustive Feature Selection (i.e., brute force)
-      \item forward selection (e.g., greedy Forward Search \citep[][]{wang2006})
-      \item backward elimination
-      \item Bi-directional elimination (Stepwise Selection)
-    \end{itemize}
-    \item hybrid: first filter then wrapper
-    \item [embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance] MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
-  \end{itemize}
-  \item dimension reduction
-\end{itemize}
-
-random forest search \citep[][]{}
-
 ```{r, include=FALSE}

### 2022-11-26 methods-and-results.tex
:pencil:

--- 2022-11-21 methods-and-results.tex
@@
 
-\subsubsection{Feature-Extraction}
-
 use domain knowledge to extract new variables from raw data (summarize)
 
-options: tsfresh \citep[][]{christ2018}, spikes
-\citep[][]{caro-martin2018}
-
-Our selection: density, inertia, instability, diversity
-\citep[][]{dejonckheere2019, kuppens2017}
-
-\begin{itemize}
-  \item 
-\end{itemize}
-
-\subsubsection{Feature Selection}
-
-select most consistent, relevant, and non-redundant features.
+options: tsfresh \citep[][]{christ2018}
 
@@
 maximize relevance and minimize redundancy
-
-``However, most of the existing UFS methods primarily focus on the
-significance of features in maintaining the data structure while
-ignoring the redundancy among features. Moreover, the determination of
-the proper number of features is another challenge.''
-
-\begin{itemize}
-  \item feature selection
-  \begin{itemize}
-    \item filter: evaluate individual features' impact / importance / relevance. Univariate (ranking-based) or multivariate (preferred bc. can handle redundant and irrelevant features)
-    \begin{itemize}
-      \item unvariate, often based on information or spectral analysis (univariate inclusion test \citep[i.e., alpha-corrected t-tests][]{christ2018})
-    \end{itemize}
-    \item wrapper: search problem --- compare different models based on evaluation criterion (i.e., model performance; e.g., $R^2$, accuracy, ...).
-"high computational cost, and they are limited to be used in conjunction with a particular clustering algorithm."
-    \begin{itemize}
-      \item Exhaustive Feature Selection (i.e., brute force)
-      \item forward selection (e.g., greedy Forward Search \citep[][]{wang2006})
-      \item backward elimination
-      \item Bi-directional elimination (Stepwise Selection)
-    \end{itemize}
-    \item hybrid: first filter then wrapper
-    \item [embedded: feature selection as part of the model construction process. E.g., LASSO Regularization L1, or Random Forest Importance] MOSTLY FOR SUPERVISED / LABELED DATA (e.g., regressions)
-  \end{itemize}
-  \item dimension reduction
-\end{itemize}
-
-random forest search \citep[][]{}

### 2022-11-26 manuscript.tex
:construction: variable input

@@
 
-\fgrref{fig:TSCFlow}
+intro-paragraph + \fgrref{fig:TSCFlow}
 
 \subsection{Input variables}
-field- and concept specific selection of input variables. Diverse
+Time series clustering starts with the selection and preparation of the variables of interest.
+While the selection will necessarily be field- and concept-specific, there are a few conceptual and methodological issues that should be considered.
+Conceptually, the included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series.
+There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}, and many health developments are commonly captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}.
+At the same time, however, the added number of variables can become a methodological concern.
+Not only can redundant and irrelevant variables diminish the quality of the analyses, but with intensive longitudinal data the number of data points compounds across participants, measurement occasions, and variables so that additional variables can make many of the following steps substantially more difficult (also see \fgrref{fig:TSCFlowN}).
 
-data cleaning + importantly, comparable time series:
-\begin{itemize}[noitemsep]
-    \item time-frame
-    \item response scale (e.g., standardize within clusters)
-\end{itemize}
+Once the important variables have been selected, the data needs to be prepared for the analysis steps.
+Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
+Two important steps are making the time-frames and response scales comparable across participants.
+Extracting features that describe the development of a month for one participant but 90 days for another participant might not be comparable for some psychological phenomena.
+This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values.
+It is thus generally advised to reduce the data to a common time frame.
+Once the time frames are comparable, also the responses should be comparable within and between participants.
+\hl{It is thus usually advised to standardize all variables for time series clustering.}
+
+\Question{ \textit{Look up how to center/standardize (within- vs. between) and when (prior to extraction vs. prior to clustering)!}}
 
@@
 Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment (or similarity measurement for qualitative features).
+While others have produced excellent overviews of the many clustering approaches \citep[e.g.,][]{xu2015}, we will briefly introduce some of the more readily available approaches suitable for most time series feature data.
+The well-established and readily available clustering approaches can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof (also see \tblref{tab:clusterApproaches}).
 
-For a more comprehensive overview see \citet{xu2015}
+Each of these approaches can be a valuable clustering approach for time series feature data and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency.
+As an example, under ideal conditions, all approaches are likely to suggest very similar clusters.
+However, as an example, when the shape of clusters becomes more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate.
+Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective.
+Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions.
+Yet, in cases where mis-specifications are costly, a method with fewer assumptions might be more advisable.
 
-\tblref{tab:clusterApproaches}
+There is thus, unfortunately, no one-size fits all solution to clustering.
+On the bright side, however, the thriving methodological developments in the clustering literature offer a wide variety of options that offer options even for extremely large and complex data sets.
+We provide a short intuitive explanation for common approaches, together with some of their characteristics and example algorithms in \tblref{tab:clusterApproaches}.
+For our own illustration, we have chosen that centroid-based k-means clustering.
+While k-means comes at the expense of high accuracy with more complex cluster shapes, we specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the data.
+K-means is also well-established within the research community and has been readily implemented in many statistical software packages.
+Additionally, many of the feature selection methods have specifically been designed for the well-established k-means algorithm.
+As such, the k-means offers a good starting point for many psychological researchers and the method should be generalizable across a relatively wide variety of projects.
 

### 2022-11-26 tables/clusterApproaches.tex
:construction: variable input

@@
         density-based \linebreak & 
-        Regions where several data points are relatively close together (i.e., high density) are assumed to be a cluster.
+        Assumes that clusters are regions where several data points are relatively close together (i.e., high density).
 Based on what is considered a dense region (e.g., radius of a region and minimum number of points the radius), points can be either be assigned to one of the clusters or be considered too far away from a dense region.
@@
             \item[\scriptsize\faPlusCircle] no shape assumption 
-            \item[\scriptsize\faMinusCircle] small number of cases
+            \item[\scriptsize\faMinusCircle] best with small number of cases
             \item[\scriptsize\faMinusCircle] no reversal of assignments

### 2022-11-27 manuscript.tex
:pencil: readability improved

@@
 
-\section{Draft Graveyard (for scrapping only):}
-
-Extensive longitudinal data sets come with new forms of heterogeneity, where we have to consider differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for inferential modeling procedures\footnote{For example, stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).}.
-And while inferential model testing is certainly important, we still miss discussions of methods for the more fundamental task of describing, summarizing, and understanding the raw/complex developmental data patterns.
-
-As an example, a recent review of migration experiences has pointed out that despite many complex and dynamic theories, investigations of migrant adaptation have undervalued developmental data --- especially when it comes to the more internal experiences of motivations and emotions \citep[e.g.,][]{Kreienkamp2022d}.
-Yet, understanding how people differ in their migration trajectories, can be crucial in understanding adaptive and maladaptive patterns.
-To identify which variables are most important in the adaptation of migrants over time, we need methods to break down the data heterogeneity into its core components (in terms of important variables and developments) and we need ways to identify how these core components relate to key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
-There is, thus, a clear need to assess the utility of analysis procedures focused on the description and understanding of complex dynamical data.
-
-We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to cluster social psychological trajectories.
-This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects \citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.
-
-To untangle the heterogeneity in real-life migrant adaptations, we will analyze data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands.
-All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation).
-However, given that past investigations of psychological acculturation have under-explored the crucial aspect of motivational experiences \citep[e.g.,][]{Kreienkamp2022d}, the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants.
-To make full use of all three data sets and to guide future use of clustering procedures with extensive longitudinal psychological data, we will analyze the data in a descriptive and step-wise manner, where we consider the data jointly (and will then assess potential differences between data- and variable sets).
-
-ALTERNATIVE:
-
-In the past, such data collections were often unfeasible because they were either physically impractical or too expensive.
-However, recent technological developments now allow us to easily collect large amounts of experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}.
-And while generally speaking analytical methods for such, more complex, data have become more readily available \citep[e.g.,][]{ODonnell2021}, it remains unclear how we should identify key developmental patterns --- especially across multiple variables at the same time (i.e., multiple aspects).
-In essence, the new extensive longitudinal data comes with new forms of heterogeneity when jointly considering differences between people, over time, and across variables.
-Yet, past analytical advances have almost exclusively pushed for stationary lagged regression models\footnote{meaning models that assume stable means and variances over time; e.g., vector autoregression models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses} or basic trajectory models\footnote{e.g., mixed effects models, spline regression models, and latent growth curve modeling.}.
-And while these two approaches are important for inferential model testing, we still miss methods for the more fundamental task of describing and unraveling the developmental data patterns.
-
-Such descriptive methods are not merely important for hypothesis generation, but also for identifying and understanding adaptive and maladaptive patterns.
-We need to identify which variables are most important in the adaptation of migrants, we need methods to identify clusters of similar individuals and developments, and we need to assess whether such clusters differ in key adaptation markers (including, well-being, intergroup anxiety, outgroup trust, or societal participation).
-
-Problem: HOWEVER, ...
-
-\begin{itemize}
-    \item unclear how do deal with this more complex data (many variables, persons, and time points).
-    \begin{itemize}
-        \item many different variables could be important.
-        \item how to consider them jointly (e.g., VAR, DSEM, ARIMA, cross-lagged panel analysis for testing model predictions, limited to specified lag and mean stationarity; Trajectory focused: mixed effects model, spline regressions, latent growth curve modeling, limited: often univariate outcome).
-    \end{itemize}
-    \item \textbf{unclear how to identify core/important developments.} Especially across multiple variables at he same time (i.e., multiple aspects).
-    \item unclear which variables, time scales, and methods are useful in practice.
-    \item curse of dimensionality: clustering algorithms fail 
-\end{itemize}
-
-Solution:
-
-feature-based clustering
-
-little data has thus far investigated the development of migration experiences and no research has assessed the co-development of multiple experience aspects.
-Clustering participants across variables and time-series features may allow to identify groups of similar developments and whether these clusters differ in key adaptation markers (including, well-being, anxiety, trust, or societal inclusion).
-
 % Tables

### 2022-11-27 methods-and-results.Rmd
:construction: data sect

@@
 
+Something about the migration experience literature should go here as part of the illustration intro ...
+
 \subsection{Data}
+We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
+The data set consisted of three individual studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group.
+After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
+The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
+After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
+Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
+For our empirical example we focused on the variables that were collected during the ESM surveys and were available for all three studies.
+Full methodological details are available in the article by \citet[][]{Kreienkamp2022b}.
 
-To illustrate the
+\subsubsection{Sample and time frame (?)}
 
-\subsubsection{Participants}
+filter criteria
+\citep[proportion of missing data and imputations:][]{Madley-Dowd2019}
 
-\subsubsection{Timepoints}
+Total N t.
+Total N ppt.
+
+Study specific exclusions see Online Supplementary Material A.
 
 \subsubsection{Variables}
+We focus on ...
 
-\subsection{Procedure}
+full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:??}
+
+\subsection{Analysis and Results}
+
+\subsubsection{Feature extraction}
+
+\subsubsection{Feature reduction}
+
+\subsubsection{Feature clustering}
 

### 2022-11-27 methods-and-results.tex
:construction: data sect

@@
+Something about the migration experience literature should go here as
+part of the illustration intro \ldots{}
+
 \subsection{Data}
 
-To illustrate the
+We used a data set following migration experiences collected by
+\citet[][]{Kreienkamp2022b}. The data set consisted of three individual
+studies that followed migrants who had recently arrived in the
+Netherlands in their daily interactions with the Dutch majority group.
+After a general migration-focused pre-questionnaire, participants were
+invited twice per day to report on their (potential) interactions with
+majority group members for at least 30 days. The short ESM surveys were
+sent out at around lunch (12pm) and dinner time (7pm). After the 30 day
+study period participants filled in a post-questionnaire that mirrored
+the pre-questionnaire. Participants received either monetary
+compensation or partial course credits based on the number of surveys
+they completed. For our empirical example we focused on the variables
+that were collected during the ESM surveys and were available for all
+three studies. Full methodological details are available in the article
+by \citet[][]{Kreienkamp2022b}.
 
-\subsubsection{Participants}
+\subsubsection{Sample and time frame (?)}
 
-\subsubsection{Timepoints}
+filter criteria
+\citep[proportion of missing data and imputations:][]{Madley-Dowd2019}
+
+Total N t. Total N ppt.
+
+Study specific exclusions see Online Supplementary Material A.
 
@@
 
-\subsection{Procedure}
+We focus on \ldots{}
+
+full methodological details are available in Online Supplemental
+Material A, but basic item information, descriptives, and correlations
+are available in \tblref{tab:??}
+
+\subsection{Analysis and Results}
+
+\subsubsection{Feature extraction}
+
+\subsubsection{Feature reduction}
+
+\subsubsection{Feature clustering}
 

### 2022-11-27 manuscript.tex
:construction: illust intro pos
:construction: expl intro

@@
 \section{Empirical Implementation/Illustration}
+To illustrate the use case of the feature-based time series with psychological ESM data, we ....
+
+Something about the migration experience literature should go here as part of the illustration intro ...
+
 % Methods and Results from RMarkdown render
@@
 \section{Notes}
-
-\citep[][]{Madley-Dowd2019} [proportion of missing data and imputations]
 

### 2022-11-27 methods-and-results.Rmd
:construction: sample Ns

@@
 
-rbind(
+missS123 <- rbind(
   missS1,
@@
   mutate(study = c(1, 2, 3)) %>%
-  select(study, everything()) %>%
+  select(study, everything())
+
+missS123 %>%
   kbl(.,
@@
 
-Something about the migration experience literature should go here as part of the illustration intro ...
-
 \subsection{Data}
 We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consisted of three individual studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group.
+The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
 After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
@@
 Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
-For our empirical example we focused on the variables that were collected during the ESM surveys and were available for all three studies.
-Full methodological details are available in the article by \citet[][]{Kreienkamp2022b}.
+For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
+Full methodological details are available in the empirical article by \citet[][]{Kreienkamp2022b}.
 
-\subsubsection{Sample and time frame (?)}
+\subsubsection{Sample}
+The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$`r missS123$pptFull[missS123$study==1]`, $N_{S2}=$`r missS123$pptFull[missS123$study==2]`, $N_{S3}=$`r missS123$pptFull[missS123$study==3]`) with a total `r format(sum(missS123$nFull), big.mark=",")` measurements.
+The studies differed substantially in the maximum length of participation ($N_{t:S1}=$`r missS123$timeFull[missS123$study==1]`, $N_{t:S2}=$`r missS123$timeFull[missS123$study==2]`, $N_{t:S3}=$`r missS123$timeFull[missS123$study==3]`).
 
-filter criteria
 \citep[proportion of missing data and imputations:][]{Madley-Dowd2019}

### 2022-11-27 methods-and-results.tex
:construction: sample Ns

@@
-Something about the migration experience literature should go here as
-part of the illustration intro \ldots{}
-
 \subsection{Data}
@@
 We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consisted of three individual
-studies that followed migrants who had recently arrived in the
-Netherlands in their daily interactions with the Dutch majority group.
-After a general migration-focused pre-questionnaire, participants were
-invited twice per day to report on their (potential) interactions with
-majority group members for at least 30 days. The short ESM surveys were
-sent out at around lunch (12pm) and dinner time (7pm). After the 30 day
-study period participants filled in a post-questionnaire that mirrored
-the pre-questionnaire. Participants received either monetary
-compensation or partial course credits based on the number of surveys
-they completed. For our empirical example we focused on the variables
-that were collected during the ESM surveys and were available for all
-three studies. Full methodological details are available in the article
-by \citet[][]{Kreienkamp2022b}.
+\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
+followed migrants who had recently arrived in the Netherlands in their
+daily interactions with the Dutch majority group members. After a
+general migration-focused pre-questionnaire, participants were invited
+twice per day to report on their (potential) interactions with majority
+group members for at least 30 days. The short ESM surveys were sent out
+at around lunch (12pm) and dinner time (7pm). After the 30 day study
+period participants filled in a post-questionnaire that mirrored the
+pre-questionnaire. Participants received either monetary compensation or
+partial course credits based on the number of surveys they completed.
+For our empirical example we focus on the variables that were collected
+during the ESM surveys and were available in all three studies. Full
+methodological details are available in the empirical article by
+\citet[][]{Kreienkamp2022b}.
 
-\subsubsection{Sample and time frame (?)}
+\subsubsection{Sample}
 
-filter criteria
+The original studies included 207 participants (\$N\_\{S1\}=\$23,
+\$N\_\{S2\}=\$113,
+\$N\_\{S3\}=\(71) with a total 10,297 measurements. The studies differed substantially in the maximum length of participation (\)N\_\{t:S1\}=\$63,
+\$N\_\{t:S2\}=\$69, \$N\_\{t:S3\}=\$155).
+
 \citep[proportion of missing data and imputations:][]{Madley-Dowd2019}

### 2022-11-27 methods-and-results.Rmd
:bug: fromula render

@@
 \subsubsection{Sample}
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$`r missS123$pptFull[missS123$study==1]`, $N_{S2}=$`r missS123$pptFull[missS123$study==2]`, $N_{S3}=$`r missS123$pptFull[missS123$study==3]`) with a total `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($N_{t:S1}=$`r missS123$timeFull[missS123$study==1]`, $N_{t:S2}=$`r missS123$timeFull[missS123$study==2]`, $N_{t:S3}=$`r missS123$timeFull[missS123$study==3]`).
+The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
+The studies differed substantially in the maximum length of participation ($N_{t:S1}=$`r missS123$timeFull[missS123$study==1]`, $N_{t:S2}=$ `r missS123$timeFull[missS123$study==2]`, $N_{t:S3}=$ `r missS123$timeFull[missS123$study==3]`).
 

### 2022-11-27 methods-and-results.Rmd
:bug: max formula

@@
 The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($N_{t:S1}=$`r missS123$timeFull[missS123$study==1]`, $N_{t:S2}=$ `r missS123$timeFull[missS123$study==2]`, $N_{t:S3}=$ `r missS123$timeFull[missS123$study==3]`).
+The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
 

### 2022-11-27 methods-and-results.Rmd
:pencil: data reduction

@@
 The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-
-\citep[proportion of missing data and imputations:][]{Madley-Dowd2019}
-
-Total N t.
-Total N ppt.
-
-Study specific exclusions see Online Supplementary Material A.
+This was likely due to the option to continue participation without compensation in the later two studies.
+To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45\% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
+Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
+For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
 

### 2022-11-27 methods-and-results.tex
:bug: fromula render
:bug: max formula
:pencil: data reduction

@@
 
-The original studies included 207 participants (\$N\_\{S1\}=\$23,
-\$N\_\{S2\}=\$113,
-\$N\_\{S3\}=\(71) with a total 10,297 measurements. The studies differed substantially in the maximum length of participation (\)N\_\{t:S1\}=\$63,
-\$N\_\{t:S2\}=\$69, \$N\_\{t:S3\}=\$155).
-
-\citep[proportion of missing data and imputations:][]{Madley-Dowd2019}
-
-Total N t. Total N ppt.
-
-Study specific exclusions see Online Supplementary Material A.
+The original studies included 207 participants (\(N_{S1}=\) 23,
+\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
+The studies differed substantially in the maximum length of
+participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
+\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
+continue participation without compensation in the later two studies. To
+make the three studies comparable in participation and time frames, we
+iteratively removed all measurement occasions and participants that had
+more than 45\% missingness
+\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure let to a final sample of 157 participants, who jointly
+produced 8,132 measurements. Importantly, both the participant
+repsonse-patterns and the time frame were now a lot more comparable
+(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
+\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
+procedure and study-specific exclusions see Online Supplemental Material
+A.
 

### 2022-11-27 methods-and-results.Rmd
:sparkles: variables

@@
 
+```{r variables included in main analysis by response type, include=FALSE}
+varNames %>%
+  filter(
+    pca != 0,
+    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
+  ) %>%
+  mutate(
+    contactSpecific = recode(
+      .$contactSpecific,
+      `-1` = "No Interaction Only",
+      `0` = "unspecific",
+      `1` = "Interaction Only"
+    ),
+    study1 = ifelse(!is.na(studyS1), 1, 0),
+    study2 = ifelse(!is.na(studyS2), 1, 0),
+    study3 = ifelse(!is.na(studyS3), 1, 0)
+  ) %>% 
+  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
+  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
+  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
+  ungroup %>%
+  t %>%
+  as.data.frame %>%
+  janitor::row_to_names(1) %>%
+  replace(is.na(.), 0) %>%
+  tibble::rownames_to_column(., var = "variable") %>%
+  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
+  mutate(across(c(study1, study2, study3), as.numeric)) %>%
+  arrange(concept, unspecific, `Interaction Only`) %>% 
+  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  relocate(starts_with("study"), .after = last_col()) %>%
+  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "l", rep("c", ncol(.)-2)),
+      digits=2,
+      caption = "Variables by Interaction Types and Study Availability") %>%
+  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
 ```{r missingness filtering, include=FALSE}
@@
 \subsubsection{Variables}
-We focus on ...
-
-full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:??}
+Included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
+We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, cognitive, and emotional experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
+Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:??}.
 

### 2022-11-27 methods-and-results.tex
:sparkles: variables

@@
 
-We focus on \ldots{}
-
-full methodological details are available in Online Supplemental
-Material A, but basic item information, descriptives, and correlations
-are available in \tblref{tab:??}
+Included 12 variables that were available across all three studies and
+captured information about the participant's interactions, as well as
+cognitive-, emotional-, and motivational self in relationship with the
+majority group. We chose these two aspects in particular because (1) the
+interaction-specific information exemplified the structural missingness
+issue of modern ESM data and (2) the motivational, cognitive, and
+emotional experience offered a diverse conceptualization of migration
+experience (beyond behavioral measurements) that is becoming more common
+in the literature \citep[][]{Kreienkamp2022d}. Full methodological
+details are available in Online Supplemental Material A, but basic item
+information, descriptives, and correlations are available in
+\tblref{tab:??}.
 

### 2022-11-27 methods-and-results.Rmd
:sparkles: descr tbl

--- 2022-11-27 methods-and-results.Rmd
@@
 load("data/s123Red.RData")
+
+# import pre-calculated features for faster renders
+load("data/features/featOut.RData")
 ```
@@
 
-# Variable Groups
-# varNames %>%
-#   filter(pca != 0) %>%
-#   mutate(contactSpecific = recode(.$contactSpecific, 
-#                          `-1`="No Interaction Only",
-#                          `0`="unspecific",
-#                          `1`="Interaction Only")) %>%
-#   crosstable::crosstable(., varGroup, by=contactSpecific) %>%
-#   select(-c(.id, label)) %>%
-#   kbl(.,
-#       escape = FALSE,
-#       booktabs = T,
-#       align = c("l", "r", "c"),
-#       digits=2,
-#       caption = "Full Analysis: Variable Groups by Interaction Types") %>%
-#   kable_classic(
-#     full_width = F,
-#     lightable_options = "hover",
-#     html_font = "Cambria"
-#   )
-# 
-# varNames %>%
-#   filter(
-#     pca != 0
-#   ) %>%
-#   mutate(contactSpecific = recode(.$contactSpecific, 
-#                          `-1`="No Interaction Only",
-#                          `0`="unspecific",
-#                          `1`="Interaction Only")) %>%
-#   select(varNam, varGroup, contactSpecific) %>%
-#   group_by(varGroup, varNam, contactSpecific) %>%
-#   summarise(n = n()) %>%
-#   pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
-#   ungroup %>%
-#   t %>%
-#   as.data.frame %>%
-#   janitor::row_to_names(1) %>%
-#   replace(is.na(.), 0) %>%
-#   tibble::rownames_to_column(., var = "variable") %>%
-#   separate(variable, c("concept", "variable"), " - ") %>%
-#   arrange(concept, unspecific, `Interaction Only`) %>%
-#   mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-#   kbl(.,
-#       escape = FALSE,
-#       booktabs = T,
-#       align = c("l", "l", rep("c", ncol(.)-2)),
-#       digits=2,
-#       caption = "All Possible Variables by Interaction Types") %>%
-#   kable_classic(
-#     full_width = F,
-#     lightable_options = "hover",
-#     html_font = "Cambria"
-#   ) %>%
-#   scroll_box(width = "100%", height = "500px")
-
 # Non-Interaction Specific Variables
@@
 
+```{r variables included in main analysis by response type, include=FALSE}
+varNames %>%
+  filter(
+    pca != 0,
+    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
+  ) %>%
+  mutate(
+    contactSpecific = recode(
+      .$contactSpecific,
+      `-1` = "No Interaction Only",
+      `0` = "unspecific",
+      `1` = "Interaction Only"
+    ),
+    study1 = ifelse(!is.na(studyS1), 1, 0),
+    study2 = ifelse(!is.na(studyS2), 1, 0),
+    study3 = ifelse(!is.na(studyS3), 1, 0)
+  ) %>% 
+  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
+  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
+  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
+  ungroup %>%
+  t %>%
+  as.data.frame %>%
+  janitor::row_to_names(1) %>%
+  replace(is.na(.), 0) %>%
+  tibble::rownames_to_column(., var = "variable") %>%
+  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
+  mutate(across(c(study1, study2, study3), as.numeric)) %>%
+  #select(-study1, -study2, -study3) %>%
+  arrange(concept, unspecific, `Interaction Only`) %>% 
+  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  relocate(starts_with("study"), .after = last_col()) %>%
+  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "l", rep("c", ncol(.)-2)),
+      digits=2,
+      caption = "Variables by Interaction Types and Study Availability") %>%
+  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
+  kable_classic(
+    full_width = F,
+    lightable_options = "hover",
+    html_font = "Cambria"
+  )
+```
+
 ```{r missingness filtering, include=FALSE}
@@
 \subsubsection{Variables}
-We focus on ...
+Included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
+We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
+Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:descrWide}.
 
-full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:??}
+```{r descrWide, include=FALSE}
+dtAll <- rbind(
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
+) %>%
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  mutate(date = as.Date(gsub(" .*", "", TID)),
+         week = strftime(date, format = "%Y-W%V")) %>%
+  select(ID,
+         PID,
+         study,
+         date,
+         week,
+         TIDnum,
+         any_of(varNamS123PCA)) %>%
+  arrange(ID, TIDnum) %>%
+  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
+  as.data.frame
+
+allMlCor <-
+  MlCorMat(
+    data = dtAll,
+    id = "ID",
+    selection = c(
+      "InteractionContextAccidental", 
+      "InteractionContextvoluntary", 
+      "InteractionContextCooperative",
+      "InteractionContextRepresentativeNL",
+      "qualityMeaning",
+      "qualityOverall", 
+      "KeyNeedFulfillment", 
+      "KeyNeedDueToPartner", 
+      "AttitudesPartner",
+      "DaytimeNeedFulfillment",
+      "AttitudesDutch",
+      "exWB"
+    ),
+    labels = c(
+      "Interaction Accidental", 
+      "Interaction Voluntary", 
+      "Interaction Cooperative",
+      "Interaction Representative NL",
+      "Quality Meaningful",
+      "Quality Overall", 
+      "Core Need", 
+      "Core Need Due to Partner", 
+      "Attitudes Partner",
+      "Daytime Core Need",
+      "Attitudes Dutch",
+      "Well-being"
+    )
+  )
+
+# Sample Descriptives Table --- wider version
+allMlCor %>%
+  t %>%
+  kbl(.,
+      format = "latex",
+      caption = "Correlation Table and Descriptive Statistics",
+      booktabs = TRUE,
+      align = c("l", rep("c", ncol(.) - 1))) %>%
+  kable_classic() %>%
+  add_header_above(
+    .,
+    c(
+      "",
+      "Correlations" = ncol(allMlCor),
+      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
+    )
+  ) %>%
+  footnote(
+    general = c(
+      "Upper triangle: Between-person correlations;",
+      "Lower triangle: Within-person correlations;",
+      "*** p < .001, ** p < .01,  * p < .05"
+    )
+  ) %>%
+  kable_styling(latex_options = "scale_down") %>%
+  gsub("\\begin{table}", "\\begin{table}\n\\begin{minipage}[t][\\textheight][t]{\\textwidth}", ., fixed = TRUE) %>% # fix table position
+  gsub("\\end{table}", "\\end{minipage}\n\\end{table}", ., fixed = TRUE) %>% # fix table position
+  save_kable("Tables/descrWide.tex")
+
+```
 

### 2022-11-27 methods-and-results.tex
:sparkles: descr tbl

@@
 interaction-specific information exemplified the structural missingness
-issue of modern ESM data and (2) the motivational, cognitive, and
-emotional experience offered a diverse conceptualization of migration
+issue of modern ESM data and (2) the motivational, emotional, and
+cognitive experience offered a diverse conceptualization of migration
 experience (beyond behavioral measurements) that is becoming more common
@@
 information, descriptives, and correlations are available in
-\tblref{tab:??}.
+\tblref{tab:descrWide}.
 

### 2022-11-27 scripts/functions/MlCorMat.R

@@ -145,13 +145,13 @@ MlCorMat <-
       `ICC(1)` =
         misty::multilevel.icc(
           data %>% select(selection),
-          cluster = data$PID,
+          cluster = data[[id]],
           type = 1
         ),
       `ICC(2)` =
         misty::multilevel.icc(
           data %>% select(selection),
-          cluster = data$PID,
+          cluster = data[[id]],
           type = 2
         )
     )

### 2022-11-27 methods-and-results.Rmd
:bug: tbl rotate

@@
   kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{table}\n\\begin{minipage}[t][\\textheight][t]{\\textwidth}", ., fixed = TRUE) %>% # fix table position
-  gsub("\\end{table}", "\\end{minipage}\n\\end{table}", ., fixed = TRUE) %>% # fix table position
+  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
+  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
   save_kable("Tables/descrWide.tex")

### 2022-11-27 Supplemental-Material-A-Annotated-Analysis.Rmd
m&r init
:pencil: results intro

@@
 
+```{r export cleaned and reduced datasets}
+# save cleaned 
+save(dtS1, dtS2, dtS3, file = "data/s123Full.RData")
+
+# save reduced
+save(list = ls(pattern = "Red$"), file = "data/s123Red.RData")
+```
+
 ## Time Scales and Timepoint Aggregation
@@
 
-## Feature Selection
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Here we might have to make a selection of the features we extract because some of the clustering algorithms might not be able to deal with the many different features. See below
-</div>
-
-Feature selection types:
-
--   manual feature and variable selection
--   wrapper: procedures that add and/or remove predictors to find the optimal combination that maximizes model performance.
--   filter: relevance of the predictors outside of the predictive models and subsequently model only the predictors that pass some criterion.
--   dimension reduction for clustering. No selection more compression of redundant variance.
-
-```{r manual feature selection}
-
-```
-
-```{r automated feature selection}
-
-```
+## Feature Reduction
+
+To avoid the dimensionality curse we reduce the extracted features using a principal component analysis.
+
+MORE INFO NEEDED
 
@@
 
-### Main Anlaysis: Across Studies
-
 The main clustering procedure, uses the features calculated for the relevant variables across all three studies.
-
-#### Centroid-based
-
 For our main method, we use the centroid-based k-means clustering.
@@
 
-#### Distribution-based
-
-For the distribution-based clustering we use the commonly used Gaussian mixture modeling (GMM) procedure.
+<!-- #### Distribution-based -->
+
+<!-- For the distribution-based clustering we use the commonly used Gaussian mixture modeling (GMM) procedure.
 This procedure assumes  independent Gaussian distributions for each each cluster.
 These distributions are substantially better at identifying clusters if the data comes from multivariate Gaussian distributions (i.e., ellipses in a two dimensional space).
-
-We again begin by determining the optimal number of clusters. For GMMs this is done using common information criteria, such as the BIC.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  This is where we run into trouble.
+-->
+
+<!-- We again begin by determining the optimal number of clusters. For GMMs this is done using common information criteria, such as the BIC. -->
+
+<!-- <div class="alert alert-warning alert-dismissible fade in" role="alert"> -->
+<!--   <a class="close" data-dismiss="alert" aria-label="close">&times;</a> -->
+<!--   <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/> -->
+<!--   This is where we run into trouble.
 These plots are relatively uncommon and according to the internet this is most likely a problem of too many (non-distinguishing or highly correlated) features.
@@
 (2002)</a>.
-</div>
-
-```{r gaussian mixture model}
-library(mclust)
-
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-# scaled_data <- featS1Imp$featuresImpZMat  %>% #select(starts_with(varNamS123Int))
-#   select(-any_of(names(which(colSums(is.na(.)) > 0))))
-
-BIC <- mclustBIC(scaled_data)
-plot(BIC)
-summary(BIC)
-mod1 <- Mclust(scaled_data)
-summary(mod1, parameters = FALSE)
-
-# multivariate <- densityMclust(scaled_data, plot=FALSE)
-# summary(multivariate)            
-# plot(multivariate, what = "BIC")
-
-# merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View
-
-#table(class, mod1$classification)
-
-#mod1dr <- MclustDR(mod1)
-#summary(mod1dr, parameters = FALSE)
-#plot(mod1dr, what = "classification")
-```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  The feature selection issue is further highlighted when we do the GMMs by feature. Where we find a reasonable number of clusters for each feature. 
-</div>
-
-```{r MClust feature specific cluster numbers}
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-nMClust <- list()
-for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){
-  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
-  mod1 <- Mclust(scaled_data)
-  #summary(mod1, parameters = FALSE)
-  nMClust[[f]] <- mod1$G
-}
-
-nMClust %>%
-  unlist %>%
-  data.frame(feature=names(.), G=., row.names=NULL) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "c"),
-      digits=2,
-      col.names = c(
-        "Feature",
-        "Number of Clusters"
-      ),
-      caption = "Number of MClust Clusters if separated by Feature") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-
-```
-
-#### Density-based
-
-For the density-based cluster analysis we focus on the *DBSCAN* algorithm ('Density-Based Spatial Clustering of Applications with Noise') as well as its extension the *OPTICS* procedure ('Ordering Points To Identify the Clustering Structure').
+-->
+<!-- </div> -->
+
+<!-- ```{r gaussian mixture model} -->
+<!-- library(mclust) -->
+
+<!-- scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion)) -->
+
+<!-- # scaled_data <- featS1Imp$featuresImpZMat  %>% #select(starts_with(varNamS123Int)) -->
+<!-- #   select(-any_of(names(which(colSums(is.na(.)) > 0)))) -->
+
+<!-- BIC <- mclustBIC(scaled_data) -->
+<!-- plot(BIC) -->
+<!-- summary(BIC) -->
+<!-- mod1 <- Mclust(scaled_data) -->
+<!-- summary(mod1, parameters = FALSE) -->
+
+<!-- # multivariate <- densityMclust(scaled_data, plot=FALSE) -->
+<!-- # summary(multivariate)             -->
+<!-- # plot(multivariate, what = "BIC") -->
+
+<!-- # merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View -->
+
+<!-- #table(class, mod1$classification) -->
+
+<!-- #mod1dr <- MclustDR(mod1) -->
+<!-- #summary(mod1dr, parameters = FALSE) -->
+<!-- #plot(mod1dr, what = "classification") -->
+<!-- ``` -->
+
+<!-- <div class="alert alert-warning alert-dismissible fade in" role="alert"> -->
+<!--   <a class="close" data-dismiss="alert" aria-label="close">&times;</a> -->
+<!--   <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/> -->
+<!--   The feature selection issue is further highlighted when we do the GMMs by feature. Where we find a reasonable number of clusters for each feature.  -->
+<!-- </div> -->
+
+<!-- ```{r MClust feature specific cluster numbers} -->
+<!-- scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion)) -->
+<!-- nMClust <- list() -->
+<!-- for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){ -->
+<!--   scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f)) -->
+<!--   mod1 <- Mclust(scaled_data) -->
+<!--   #summary(mod1, parameters = FALSE) -->
+<!--   nMClust[[f]] <- mod1$G -->
+<!-- } -->
+
+<!-- nMClust %>% -->
+<!--   unlist %>% -->
+<!--   data.frame(feature=names(.), G=., row.names=NULL) %>% -->
+<!--   kbl(., -->
+<!--       escape = FALSE, -->
+<!--       booktabs = T, -->
+<!--       align = c("l", "c"), -->
+<!--       digits=2, -->
+<!--       col.names = c( -->
+<!--         "Feature", -->
+<!--         "Number of Clusters" -->
+<!--       ), -->
+<!--       caption = "Number of MClust Clusters if separated by Feature") %>% -->
+<!--   kable_classic( -->
+<!--     full_width = F, -->
+<!--     lightable_options = "hover", -->
+<!--     html_font = "Cambria" -->
+<!--   ) -->
+
+<!-- ``` -->
+
+<!-- #### Density-based -->
+
+<!-- For the density-based cluster analysis we focus on the *DBSCAN* algorithm ('Density-Based Spatial Clustering of Applications with Noise') as well as its extension the *OPTICS* procedure ('Ordering Points To Identify the Clustering Structure').
 This procedure is more bottom up than the previous two approaches.
@@
 While this description is grossly oversimplified, it highlights one of the advantages of the procedures, namely that dense regions can expand to be any shape in the d-dimensional space.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Same feature selection issue as above. 
-</div>
-
-```{r dbscan}
-library(dbscan)
-library(fpc)
-
-#scaled_data <- featFullImp$featuresImpZMat %>% select(starts_with(varNamS123Int))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))  # %>% select(starts_with(varNamS123NoInt))#%>%
-  #select(-any_of(names(which(colSums(is.na(scaled_data)) > 0))))
-
-#is.na(scaled_data) %>% table
-
-# to plot the eps values
-eps_plot = kNNdistplot(scaled_data, k=3)
-# to draw an optimum line
-eps_plot %>% abline(h = 11, lty = 2)
-
-potentialMinPts <- round(SciViews::ln(nrow(scaled_data)))
-eps <- 11
-minPts <- 2
-
-res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0)
-res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
-
-resEqual.check <- all(res.fpc$cluster == res.db$cluster)
-nClustDBSCAN <- length(unique(res.fpc$cluster))-1
-
-fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
-#hullplot(scaled_data, res.fpc)
-
-```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  And again the number of clusters is reasonable if we consider the features separately: 
-</div>
-
-```{r dbscan feature specific cluister numbers}
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-minPts <- 3
-xi <- 0.04 #0.05
-nDbscan <- list()
-for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){
-  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
-  #res.fpc <- dbscan::dbscan(scaled_data, eps=eps_cl, MinPts=minPts)
-
-  res.optics <- dbscan::optics(scaled_data, minPts = 3)
-
-  #plot(res.optics, col = "grey")
-  #dend <- as.dendrogram(res.optics)
-  #plot(dend, ylab = "Reachability dist.", leaflab = "none")
-  #extractDBSCAN(res.optics, eps_cl = 19.12)
-
-  res <- extractXi(res.optics, xi = xi)
-  res$clusters_xi
-  #plot(res)
-  #hullplot(scaled_data, res)
-
-  #summary(res.optics, parameters = FALSE)
-  nDbscan[[f]] <- nrow(res$clusters_xi)
-}
-
-nDbscan %>%
-  unlist %>%
-  data.frame(feature=names(.), G=., row.names=NULL) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "c"),
-      digits=2,
-      col.names = c(
-        "Feature",
-        "Number of Clusters"
-      ),
-      caption = "Number of OPTICS Clusters if separated by Feature") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-### Follow-up: By Response Type
-
-_**currently removed for faster render until feature selection issue is resolved**_
-
-### Follow-up: By Study
-
-_**currently removed for faster render until feature selection issue is resolved**_
+-->
+
+<!-- <div class="alert alert-warning alert-dismissible fade in" role="alert"> -->
+<!--   <a class="close" data-dismiss="alert" aria-label="close">&times;</a> -->
+<!--   <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/> -->
+<!--   Same feature selection issue as above.  -->
+<!-- </div> -->
+
+<!-- ```{r dbscan} -->
+<!-- library(dbscan) -->
+<!-- library(fpc) -->
+
+<!-- #scaled_data <- featFullImp$featuresImpZMat %>% select(starts_with(varNamS123Int)) -->
+<!-- scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))  # %>% select(starts_with(varNamS123NoInt))#%>% -->
+<!--   #select(-any_of(names(which(colSums(is.na(scaled_data)) > 0)))) -->
+
+<!-- #is.na(scaled_data) %>% table -->
+
+<!-- # to plot the eps values -->
+<!-- eps_plot = kNNdistplot(scaled_data, k=3) -->
+<!-- # to draw an optimum line -->
+<!-- eps_plot %>% abline(h = 11, lty = 2) -->
+
+<!-- potentialMinPts <- round(SciViews::ln(nrow(scaled_data))) -->
+<!-- eps <- 11 -->
+<!-- minPts <- 2 -->
+
+<!-- res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0) -->
+<!-- res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts) -->
+
+<!-- resEqual.check <- all(res.fpc$cluster == res.db$cluster) -->
+<!-- nClustDBSCAN <- length(unique(res.fpc$cluster))-1 -->
+
+<!-- fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication()) -->
+<!-- #hullplot(scaled_data, res.fpc) -->
+
+<!-- ``` -->
+
+<!-- <div class="alert alert-warning alert-dismissible fade in" role="alert"> -->
+<!--   <a class="close" data-dismiss="alert" aria-label="close">&times;</a> -->
+<!--   <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/> -->
+<!--   And again the number of clusters is reasonable if we consider the features separately:  -->
+<!-- </div> -->
+
+<!-- ```{r dbscan feature specific cluister numbers} -->
+<!-- scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion)) -->
+<!-- minPts <- 3 -->
+<!-- xi <- 0.04 #0.05 -->
+<!-- nDbscan <- list() -->
+<!-- for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){ -->
+<!--   scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f)) -->
+<!--   #res.fpc <- dbscan::dbscan(scaled_data, eps=eps_cl, MinPts=minPts) -->
+
+<!--   res.optics <- dbscan::optics(scaled_data, minPts = 3) -->
+
+<!--   #plot(res.optics, col = "grey") -->
+<!--   #dend <- as.dendrogram(res.optics) -->
+<!--   #plot(dend, ylab = "Reachability dist.", leaflab = "none") -->
+<!--   #extractDBSCAN(res.optics, eps_cl = 19.12) -->
+
+<!--   res <- extractXi(res.optics, xi = xi) -->
+<!--   res$clusters_xi -->
+<!--   #plot(res) -->
+<!--   #hullplot(scaled_data, res) -->
+
+<!--   #summary(res.optics, parameters = FALSE) -->
+<!--   nDbscan[[f]] <- nrow(res$clusters_xi) -->
+<!-- } -->
+
+<!-- nDbscan %>% -->
+<!--   unlist %>% -->
+<!--   data.frame(feature=names(.), G=., row.names=NULL) %>% -->
+<!--   kbl(., -->
+<!--       escape = FALSE, -->
+<!--       booktabs = T, -->
+<!--       align = c("l", "c"), -->
+<!--       digits=2, -->
+<!--       col.names = c( -->
+<!--         "Feature", -->
+<!--         "Number of Clusters" -->
+<!--       ), -->
+<!--       caption = "Number of OPTICS Clusters if separated by Feature") %>% -->
+<!--   kable_classic( -->
+<!--     full_width = F, -->
+<!--     lightable_options = "hover", -->
+<!--     html_font = "Cambria" -->
+<!--   ) -->
+<!-- ``` -->
+
+<!-- ### Follow-up: By Response Type -->
+<!-- Potential future project -->
+
+<!-- ### Follow-up: By Study -->
+<!-- Potential future project -->
 

### 2022-11-27 methods-and-results.Rmd
:pencil: results intro

@@
   t %>%
+  as.data.frame %>%
   kbl(.,
@@
 \subsection{Analysis and Results}
+To perform the main analysis on our selected variables, we follow the three extract--reduce--cluster steps in sequential order.
+For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REFERENCES).
 
 \subsubsection{Feature extraction}
+```{r featMiss, include=FALSE}
+pMissFull <- pMissFeat(
+  all = featFull$features,
+  contact = featFullContact$features,
+  nocontact = featFullNoContact$features,
+  title = "Feature-wise Missingess Across all Studies"
+)
+```
+
+For the features, we selected and extracted one of each of the features proposed in \tblref{tab:esmFeatures}.
+We particularly chose the more robust median and median absolute deviation (MAD) for central tendency and variability.
+Within the variability structure, we chose the slightly more intuitive mean absolute change (MAC) for stability and stayed with the common lag--1 autocorrelation for inertia.
+For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
+
+We include a function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
+
+For the full dataset across the three studies, we find that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r ncol(featFull$features)` features per participant.
+These missing features are imputed with a single predictive mean matching imputation using the MICE library.
 

### 2022-11-27 methods-and-results.tex
:pencil: results intro

@@
 
+To perform the main analysis on our selected variables, we follow the
+three extract--reduce--cluster steps in sequential order. For each of
+the steps fully reproducible and annotated code is available in Online
+Supplemental A as well as in the accompanying OSF- and GitHub
+repositories (REFERENCES). This includes a function that automatically
+extracts and prepares all of the time series features presented in
+\tblref{tab:esmFeatures} (see the \texttt{featureExtractor} function in
+our GitHub repository).
+
 \subsubsection{Feature extraction}
+
+For the
+
+For the full dataset across the three studies, we find that about 1.66\%
+of the extracted features are missing across the 145 features per
+participant. These missing features are imputed with a single predictive
+mean matching imputation using the MICE library.
 

### 2022-11-27 methods-and-results.Rmd
:sparkles: feature extraction

@@
 ```{r featMiss, include=FALSE}
+featureExclusion <- c(
+  "mean",
+  "sd",
+  "rmssd",
+  "ar02",
+  "edf_ar",
+  "_n"
+)
+
 pMissFull <- pMissFeat(
-  all = featFull$features,
+  all = featFull$features %>% select(-ends_with(featureExclusion)),
   contact = featFullContact$features,
@@
 For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
+We chose these features in particular because we consider them to be the most broadly applicable features.
+We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
+We include a function that automatically extracts and prepares all features we extracted as well as a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
 
-We include a function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
-
-For the full dataset across the three studies, we find that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r ncol(featFull$features)` features per participant.
-These missing features are imputed with a single predictive mean matching imputation using the MICE library.
+After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across participants across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
+This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
+The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
+The few missing values can however be an issue for some feature reduction or feature clustering algorithms.
+We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
 
 \subsubsection{Feature reduction}
+```{r pca, include=FALSE}
+scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
+
+Rdimtools::do.udfs(scaled_data, k = 5)
+
+res.pca <- prcomp(scaled_data, scale = FALSE)
+fviz_eig(res.pca, addlabels = TRUE)
+summary(res.pca)
+
+var <- get_pca_var(res.pca)
+corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
+fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
+fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)
+
+res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
+# Description of dimension 1
+res.desc$Dim.1
+
+plot_ly(
+  res.pca$rotation %>% as.data.frame,
+  x = ~ PC1,
+  y = ~ PC2,
+  z = ~ PC3,
+  size = 2,
+  type = 'scatter3d',
+  mode = 'markers',
+  text = ~ paste('Variable:', rownames(res.pca$rotation))
+) %>%
+  layout(title = 'Variables in reduced space')
+plot_ly(
+  res.pca$x %>% as.data.frame,
+  x = ~ PC1,
+  y = ~ PC2,
+  z = ~ PC3,
+  size = 2,
+  type = 'scatter3d',
+  mode = 'markers',
+  text = ~ paste('Participant:', rownames(res.pca$x))
+) %>%
+  layout(title = 'Participants in reduced space')
+
+pca.out <- as.data.frame(res.pca$x[,1:3])
+
+```
+
+For the
 

### 2022-11-27 methods-and-results.tex
:sparkles: feature extraction

--- 2022-11-27 methods-and-results.tex
@@
 
+To perform the main analysis on our selected variables, we follow the
+three extract--reduce--cluster steps in sequential order. For each of
+the steps fully reproducible and annotated code is available in Online
+Supplemental A as well as in the accompanying OSF- and GitHub
+repositories (REFERENCES).
+
 \subsubsection{Feature extraction}
+
+For the features, we selected and extracted one of each of the features
+proposed in \tblref{tab:esmFeatures}. We particularly chose the more
+robust median and median absolute deviation (MAD) for central tendency
+and variability. Within the variability structure, we chose the slightly
+more intuitive mean absolute change (MAC) for stability and stayed with
+the common lag--1 autocorrelation for inertia. For the trend summaries,
+we chose the overall linear regression slope to describe the linear
+trend and we summarized the nonlinear trend with the estimated degrees
+of freedom of an empty GAM spline model (edf) --- the edf summarizes the
+\textit{wiggliness} of the spline trend line. We chose these features in
+particular because we consider them to be the most broadly applicable
+features. We also extracted the participant's number of completed ESM
+measurements to ensure that the clusters are comparable in that regard.
+We include a function that automatically extracts and prepares all
+features we extracted as well as a large selection of the time series
+feature operationalizations presented in \tblref{tab:esmFeatures} in our
+GitHub repository (see the \texttt{featureExtractor} function).
+
+After the feature extraction, we found that about 1.40\% of the
+extracted features are missing across participants across the 72
+features per participant. This might, for example, happen if
+participants do not have two subsequent measurements with outgroup
+interactions, so that an autocorrelation with lag-1 cannot be calculated
+for the contact-specific variables. The small number of missing values
+indicates that the feature-based approach indeed largely avoids the
+structural missingness issue. The few missing values can however be an
+issue for some feature reduction or feature clustering algorithms. We,
+thus, impute the missing feature values with a single predictive mean
+matching imputation using the MICE library.
 

### 2022-11-27 methods-and-results.Rmd
:construction: pca code

@@
 
-Rdimtools::do.udfs(scaled_data, k = 5)
-
-res.pca <- prcomp(scaled_data, scale = FALSE)
-fviz_eig(res.pca, addlabels = TRUE)
+res.pca <- prcomp(scaled_data, scale. = FALSE)
 summary(res.pca)
 
-var <- get_pca_var(res.pca)
-corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
-fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
-fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)
+res.varExplained <- data.frame(
+  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
+  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
+)
+fviz_eig(res.pca, addlabels = TRUE)
+ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
+  geom_point() +
+  geom_line() +
+  theme_Publication()
 
-res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
-# Description of dimension 1
-res.desc$Dim.1
+pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= 0.8))
 
@@
 
-pca.out <- as.data.frame(res.pca$x[,1:3])
-
+pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
 ```
 
-For the
+For the feature reduction, we chose the very general \textit{principal component analysis} (PCA).
+While some of the more tailor-made feature selection algorithms can be more accurate and meaningful in reducing the feature dimensionality, PCAs are well-established and can broadly be applied in an automatize manner.
+80% rule of thumb
+
+we select the first `r pca_cutoff` principal components because they explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensions by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
 
 \subsubsection{Feature clustering}
+```{r featureClustering, include=FALSE}
+# load libraries
+library(factoextra)
+library(NbClust)
+
+# prep data
+scaled_data <- pca.out %>% scale
+
+# Elbow method
+fviz_nbclust(scaled_data, kmeans, method = "wss") +
+  geom_vline(xintercept = 2, linetype = 2)+
+  labs(subtitle = "Elbow method")
+# Silhouette method
+fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
+  labs(subtitle = "Silhouette method")
+# Gap statistic [takes for ever]
+# nboot = 50 to keep the function speedy.
+# recommended value: nboot= 500 for your analysis.
+# Use verbose = FALSE to hide computing progression.
+# set.seed(123)
+# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
+#   labs(subtitle = "Gap statistic method")
+```
 

### 2022-11-27 methods-and-results.tex
:construction: pca code

@@
 
+For the feature reduction, we chose the very general
+\textit{principal component analysis} (PCA). While some of the more
+tailor-made feature selection algorithms can be more accurate and
+meaningful in reducing the feature dimensionality, PCAs are
+well-established and can broadly be applied in an automatize manner.
+80\% rule of thumb
+
+we select the first 27 principal components because they explain 80\% of
+the variance in the original 72 features (reducing the dimensions by
+62.50\%).
+
 \subsubsection{Feature clustering}

### 2022-11-27 manuscript.tex
:bug:
Updates from Overleaf

@@
 It is thus generally advised to reduce the data to a common time frame.
-Once the time frames are comparable, also the responses should be comparable within and between participants.
-\hl{It is thus usually advised to standardize all variables for time series clustering.}
-
-\Question{ \textit{Look up how to center/standardize (within- vs. between) and when (prior to extraction vs. prior to clustering)!}}
 
 \subsection{Feature extraction}
-aim: 
+The main aim of feature extraction is 
 hundreds of possible features to describe psychological time series.
@@
 
-\section{Empirical Implementation/Illustration}
+\section{Empirical Illustration}
 To illustrate the use case of the feature-based time series with psychological ESM data, we ....
@@
 \input{methods-and-results}
+\input{tables/descrWide.tex}
 

### 2022-11-27 manuscript.tex
:pencil: feature combination

@@
 The main aim of feature extraction is 
-hundreds of possible features to describe psychological time series.
+hundreds of possible features to describe psychological time series \citep[e.g., tsfresh][]{christ2018}.
 One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series.
@@
 Feature selections seek to create a subset of the most important features.
+Computationally that is a challenging task because \textit{k} features could be considered in \(2^k – 1\) possible combinations --- for the example of 80 features that would allow for \(1.20 \times 10^{24}\) combinations.
 The many available approaches differ in how they seek to determine the importance of the individual features.

### 2022-11-28 methods-and-results.Rmd
:coffin: dead code
:sparkles: feature reduction

@@
 For the feature reduction, we chose the very general \textit{principal component analysis} (PCA).
-While some of the more tailor-made feature selection algorithms can be more accurate and meaningful in reducing the feature dimensionality, PCAs are well-established and can broadly be applied in an automatize manner.
-80% rule of thumb
+Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
+However, PCAs have the distinct benefit that they are well-established within the psychometric literature and can broadly be applied to a wide variety of studies in an automatize manner.
+As our aim is to present a general illustration that can also be adopted for more general data descriptive uses, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
 
-we select the first `r pca_cutoff` principal components because they explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensions by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
+To use the principal component analysis with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
+We then enter all `r ncol(scaled_data)` features into the analysis.
+The PCA will use linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
+The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
+In practice, this means that the PCA decomposes the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
+We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
+A common rule of thumb is to use the principal components that jointly explain 70--90\% of the original variance (i.e., cumulative percentage explained variance).
+For our illustration we we select the first `r pca_cutoff` principal components that explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
+
+We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
+As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' with (highly) correlated variables.
+Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
+While our `r pca_cutoff` principal components is still a relatively high number of variables this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
+As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
+In particular, because the principal components are linear combination of original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
+While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
+Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
+We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
 
@@
   geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method")
+  labs(subtitle = "Elbow method") +
+  theme_Publication()
 # Silhouette method
 fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method")
+  labs(subtitle = "Silhouette method") +
+  theme_Publication()
 # Gap statistic [takes for ever]
@@
 
-feature engineering: feature extraction + feature selection
-
-use domain knowledge to extract new variables from raw data (summarize)
-
-options: tsfresh \citep[][]{christ2018}
-
-\textit{k} features means $2^k – 1$ possible models
-
 maximize relevance and minimize redundancy

### 2022-11-28 methods-and-results.tex
:sparkles: feature reduction

--- 2022-11-27 methods-and-results.tex
@@
 
+For the feature reduction, we chose the very general
+\textit{principal component analysis} (PCA). Some of the more
+tailor-made feature selection algorithms can be more accurate in
+reducing the feature dimensionality and might retain feature importance
+information more directly, depending on the specific data structure.
+However, PCAs have the distinct benefit that they are well-established
+within the psychometric literature and can broadly be applied to a wide
+variety of studies in an automatize manner. As our aim is to present a
+general illustration that can also be adopted for more general data
+descriptive uses, we present the workflow using a PCA here but we
+encourage users to consider more specialized methods as well.
+
+To use the principal component analysis with our extracted time series
+features, we first standardize all features across participants to
+ensure that all features are weighted equally. We then enter all 72
+features into the analysis. The PCA will use linear transformations in
+such a way that the first component captures the most possible variance
+of the original data (e.g., by finding a vector that maximizes the sum
+of squared distances). The following components will then use the same
+method to iteratively explain the most of the remaining variance while
+also ensuring that the components are linearly uncorrelated. In
+practice, this means that the PCA decomposes the 72 features into 72
+principal components but now (because of the uncorrelated linear
+transformations) the first few principal components will capture a
+majority of the variance. We can then decide how much information (i.e.,
+variance) we are willing to sacrifice for a reduced dimensionality. A
+common rule of thumb is to use the principal components that jointly
+explain 70--90\% of the original variance (i.e., cumulative percentage
+explained variance). For our illustration we we select the first 27
+principal components that explain 80\% of the variance in the original
+72 features (reducing the dimensionality by 62.50\%).
+
+We would like to comment on two practical matters when using principal
+components --- the amount of dimensionality reduction and the
+interpretation of the principal components. As for the expected
+dimensionality reduction, given its methodology, PCAs tend to `work
+better' with (highly) correlated variables. Thus, with a set of very
+homogeneous variables and features users will need less principal
+components to explain a large amount of variance, while a more diverse
+set of variables and features will tend to require more principal
+components to capture the same amount of variance. While our 27
+principal components is still a relatively high number of variables this
+is not surprising as we chose a diverse conceptualization and a diverse
+set of time series features. As for the interpretability, PCA allows
+users to extract information on the meaning of the principal components.
+In particular, because the principal components are linear combination
+of original features, users can extract the relative importance of each
+feature for the extracted principal components (i.e., the eigenvectors).
+While this can be useful in understanding the variance in the original
+data or help with manual feature selection, we use the PCA purely to
+reduce the dimensionality for the clustering step. Instead of relying on
+the principal components, we use the original features of interest to
+interpret the later extracted clusters. We particularly advocate for
+such an approach if the original features were chosen for their
+psychological meaning in understanding the time series and broader
+phenomenon of interest.
+
 \subsubsection{Feature clustering}
 
-feature engineering: feature extraction + feature selection
-
-use domain knowledge to extract new variables from raw data (summarize)
-
-options: tsfresh \citep[][]{christ2018}
-
-\textit{k} features means \(2^k – 1\) possible models
-
 maximize relevance and minimize redundancy

### 2022-11-28 methods-and-results.Rmd
:sparkles: feature clustering

@@
 ```{r pca, include=FALSE} 
+# prepare data
+raw_data <- featData
+z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion), ends_with("_n"))
 scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
@@
 
+# extract coordinates for the individuals (PC-scores)
 pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
@@
 For our illustration we we select the first `r pca_cutoff` principal components that explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
+For the extracted principal components we save the `r ncol(scaled_data)` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
 
@@
 
+# check variance differences
+max(pca.out %>% summarize_all(var))/min(pca.out %>% summarize_all(var))
+pca.out %>%
+  summarize_all(sd) %>%
+  pivot_longer(everything(), names_to = "PC", values_to = "sd") %>%
+  ggplot(., aes(x=reorder(PC, sd), y=sd, label=sd)) +
+  geom_bar(stat='identity', width=.5) +
+  coord_flip() +
+  theme_Publication()
+
 # prep data
-scaled_data <- pca.out %>% scale
+scaled_pc <- pca.out #%>% scale
 
 # Elbow method
-fviz_nbclust(scaled_data, kmeans, method = "wss") +
+fviz_nbclust(scaled_pc, kmeans, method = "wss") +
   geom_vline(xintercept = 2, linetype = 2)+
@@
 # Silhouette method
-fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
+fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
   labs(subtitle = "Silhouette method") +
   theme_Publication()
-# Gap statistic [takes for ever]
-# nboot = 50 to keep the function speedy.
-# recommended value: nboot= 500 for your analysis.
-# Use verbose = FALSE to hide computing progression.
-# set.seed(123)
-# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
-#   labs(subtitle = "Gap statistic method")
 ```
 
-maximize relevance and minimize redundancy
+```{r k-means modeling based on selected k, include=FALSE}
+kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
+```
 
-```{r, include=FALSE}
+```{r Plot Convex Hulls, include=FALSE}
+fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
+fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
+```
 
+\Question{It is good practice to standardize the input of the k-means algorithm.
+However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
+If I actually standardize the PC-score, we get really weird k-means results.
+The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2
+I had originally written: 
+"To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
+but I went back to the 'raw' PC-scores now. But we should discuss this.}
+
+For our illustration of the feature clustering procedure, we use the generalizable and efficient centroid-based k-means clustering.
+The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
+We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see \fgrref{fig:??}).
+We then entered the participants' PC-scores into the k-means algorithm.
+We used the recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100 random initial centroid positions to avoid convergence to a sub-optimal solution (i.e., local minima).
+The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
+
+In order to interpret the two clusters, we compared the two clusters in their features as well as in their raw time series.
+We first inspect the clusters with a focus on the included variables (see \fgrref{fig:??}).
+We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., how cooperative an interaction was) than for other variables (e.g., Attitudes towards the Dutch).
+Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being) than others (e.g., MAC of well-being).
+We then inspect the clusters with a focus on the features (see \fgrref{fig:??}).
+While this is the same data as in the plot above, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median) than others (e.g., auto correlations).
+This offers some information on which features were are most important in understanding the two extracted groups.
+
+We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:??}).
+Immediately striking are the mean differences, where participants in the cluster that had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
+The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
+Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
+The 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref{fig:??}).
+We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time.
+
+\Question{Talk about early warning signals here? I.e., divergences in means over time could be invesitigated with targeted studies.}
+
+We can also assess the clusters across any other person-level variable.
+This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
+To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants with more negative experiences on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions.
+While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
+To illustrate the assessment of individual differences, we added a common acculturation measurement that was completed in the pre- and post-questionnaire \hl{(VIA) as well as ...
+.
+We can see that TBA [...].}
+
+```{r average feature by cluster, include=FALSE}
+featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., z_data) %>%
+  group_by(cluster) %>%
+  summarise(across(everything(), list(mean = mean, se = se))) %>%
+  ungroup %>% 
+  select(-starts_with("ID")) %>%
+  pivot_longer(
+    cols = -c(cluster),
+    names_to = c("variable", ".value"),
+    names_pattern = "(.*)_(.*)"
+  ) %>% 
+  mutate(
+    lab = variable,
+    variable = gsub("\\_.*", "", lab),
+    feature = gsub("^.*?\\_", "", lab)
+  ) %>% 
+  select(
+    cluster,
+    lab,
+    variable,
+    feature,
+    everything()
+  )
 ```
+
+```{r clusterVariableGrid, include=FALSE}
+featCluster %>% 
+  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
+  geom_point() +
+  geom_line() +
+  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
+  coord_flip() +
+  labs(
+    x = "Features",
+    y = "Standardized Mean",
+    color = "Cluster"
+  ) +
+  facet_wrap(~ variable) +
+  theme_Publication()
+```
+
+```{r clusterFeatureGrid, include=FALSE}
+featCluster %>% 
+  ggplot(., aes(x = variable, y = mean, group = cluster, color = as.factor(cluster))) +
+  geom_point() +
+  geom_line() +
+  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
+  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
+  coord_flip() +
+  labs(
+    x = "Features",
+    y = "Standardized Mean",
+    color = "Cluster"
+  ) +
+  facet_wrap(~ feature) +
+  theme_Publication()
+```
+
+```{r clusterTS, include=FALSE}
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., raw_data) %>%
+  select(
+    ID,
+    TIDnum,
+    cluster,
+    any_of(varNamS123)
+  ) %>%
+  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
+  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
+  stat_summary(fun=mean, geom="line") +
+  facet_wrap(~ variable, scales = "free_y") +
+  labs(
+    x = "Time ID",
+    y = "Response (Mean Cluster Response Bold)",
+    color = "Cluster"
+  ) +
+  theme_Publication() 
+```
+
+```{r clusterTSTrend, include=FALSE}
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., raw_data) %>%
+  select(
+    ID,
+    TIDnum,
+    cluster,
+    any_of(varNamS123)
+  ) %>%
+  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
+  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
+  geom_smooth(aes(group=cluster)) +
+  #stat_summary(fun=mean, geom="line") +
+  facet_wrap(~ variable, scales = "free_y") +
+  labs(
+    x = "Time ID",
+    y = "Response (Mean Cluster Response Bold)",
+    color = "Cluster"
+  ) +
+  theme_Publication()
+```

### 2022-11-28 methods-and-results.tex
:sparkles: feature clustering

@@
 principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%).
+72 features (reducing the dimensionality by 62.50\%). For the extracted
+principal components we save the 72 PC-scores for each participant
+(i.e., the participants' coordinates in the reduced dimensional space).
 
@@
 
-maximize relevance and minimize redundancy
+\Question{It is good practice to standardize the input of the k-means algorithm.
+However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
+If I actually standardize the PC-score, we get really weird k-means results.
+The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2
+I had originally written: 
+"To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
+but I went back to the 'raw' PC-scores now. But we should discuss this.}
+
+For our illustration of the feature clustering procedure, we use the
+generalizable and efficient centroid-based k-means clustering. The
+procedure needs the user to make a selection on the optimal number of
+clusters in order to run the main algorithm. We use the Elbow and the
+Silhouette method, which both suggest two clusters to be the optimal
+solution (see \fgrref{fig:??}). We then entered the participants'
+PC-scores into the k-means algorithm. We used the recommended Hartigan
+and Wong algorithm \citep{hartigan1979} with 100 random initial centroid
+positions to avoid convergence to a sub-optimal solution (i.e., local
+minima). The k-means analysis assigned 80 participants to cluster 1 and
+76 participants to cluster 2.
+
+In order to interpret the two clusters, we compared the two clusters in
+their features as well as in their raw time series. We first inspect the
+clusters with a focus on the included variables (see \fgrref{fig:??}).
+We (1) see that for some variables the features are generally stronger
+in separating the clusters (e.g., how cooperative an interaction was)
+than for other variables (e.g., Attitudes towards the Dutch).
+Additionally, we see that (2) within the variables some features are
+better at distinguishing clusters (e.g., median of well-being) than
+others (e.g., MAC of well-being). We then inspect the clusters with a
+focus on the features (see \fgrref{fig:??}). While this is the same data
+as in the plot above, we can see more clearly that some features are
+better at distinguishing the clusters across variables (e.g., mean and
+median) than others (e.g., auto correlations). This offers some
+information on which features were are most important in understanding
+the two extracted groups.
+
+We can then combine the two focus approaches to assess the developments
+between the two groups more holistically (see \fgrref{fig:??}).
+Immediately striking are the mean differences, where participants in the
+cluster that had more meaningful and fulfilling outgroup interactions
+also consistently reported more voluntary and cooperative interactions
+but less accidental and involuntary interactions. The same cluster also
+reported an increase in need fulfilling interaction over the 30 day
+period and an increase in interactions that were representative of the
+outgroup. Whereas the other cluster showed a decrease in voluntary,
+cooperative, and positive interactions over the 30 days. The
+`deterioration' cluster also saw a decrease in general need fulfillment
+and experienced well-being over the 30 days (see \fgrref{fig:??}). We
+also see that while interaction representativeness, outgroup attitudes,
+well-being are relatively stable for both clusters, the deteriorating
+cluster also showed substantially higher variablity and instability over
+time.
+
+\Question{Talk about early warning signals here? I.e., divergences in means over time could be invesitigated with targeted studies.}
+
+We can also assess the clusters across any other person-level variable.
+This out-of-feature comparison allows us to check for data artifacts, as
+well as check whether the developmental clusters are associated with
+important social markers and individual differences. To illustrate
+artifact checks, we added the number of measurements into the comparison
+and find that the participants with more negative experiences on average
+completed more ESM surveys and reported on more intergroup interactions
+than the cluster with the more positive interactions. While this
+difference could indicate that the clusters might not entirely be
+comparable in the response patterns, we can find some relief in our data
+exclusion procedures during which we ensured that the general time frame
+and completion rates were not too dissimilar. To illustrate the
+assessment of individual differences, we added a common acculturation
+measurement that was completed in the pre- and post-questionnaire
+\hl{(VIA) as well as ... . We can see that TBA [...].}

### 2022-11-28 methods-and-results.Rmd
:art: text col

@@
 
-\Question{It is good practice to standardize the input of the k-means algorithm.
+\Question{\textcolor}{cyan}{It is good practice to standardize the input of the k-means algorithm.
 However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
 If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2
+The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}
 I had originally written: 
 "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
-but I went back to the 'raw' PC-scores now. But we should discuss this.}
+but I went back to the 'raw' PC-scores now. But we should discuss this.}}
 
@@
 
-\Question{Talk about early warning signals here? I.e., divergences in means over time could be invesitigated with targeted studies.}
+\Question{\textcolor{cyan}{Talk about early warning signals here? I.e., divergences in means over time could be invesitigated with targeted studies.}}
 

### 2022-11-28 methods-and-results.tex
:art: text col

@@
 
-\Question{It is good practice to standardize the input of the k-means algorithm.
+\Question{\textcolor}{cyan}{It is good practice to standardize the input of the k-means algorithm.
 However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
 If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2
+The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}
 I had originally written: 
 "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
-but I went back to the 'raw' PC-scores now. But we should discuss this.}
+but I went back to the 'raw' PC-scores now. But we should discuss this.}\}
 

### 2022-11-28 manuscript.tex
:bug: re-read

--- 2022-11-27 manuscript.tex
@@
 % define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
-\documentclass[man, 12pt, a4paper]{apa7}
+\documentclass[man, 12pt, a4paper, mask]{apa7}
 
@@
     \begin{itemize}[nosep]
-        \item Input variables
+        \item input variables
         \item feature extraction
@@
     \item Empirical Illustration
-    \item Discussion
+    \begin{itemize}[nosep]
+        \item Data
+        \item Analysis and Results
     \end{itemize}
+    \item Discussion (summary, limitation, implications, conclusion, etc.)
+\end{itemize}
+\vspace{1em}
 \end{mdframed}
@@
 With the increased availability of technologies to easily collect large amounts of experience sampling data using mobile devices \citep[e.g.,][]{Keil2020} and web-based applications \citep[e.g.,][]{Arslan2020}, the experience-sampling method (ESM) has been applied to a host of new types of psychological data.
-One example of this phenomenon has been the cultural adaptation literature, where researchers have started using ESM data to follow the daily interactions of migrants with the local majority \citep[e.g.,][]{Kreienkamp2022b}.
 This move toward a broader application of the experience-sampling method, however, stands in stark contrast with recent developments within the psychometric literature on ESM analyses, which have become more specialized and model-focused.
@@
 Data-driven models not only describe and summarize this complex data but can help us understand the data by uncovering patterns that might not be detectable otherwise.
-Returning to the case of migration experiences, understanding how people differ in their migration trajectories can be crucial in understanding adaptive and maladaptive patterns.
+As an example, understanding how people differ in their trajectories can be crucial in understanding adaptive and maladaptive patterns.
 By identifying and contextualizing distinct trajectories, we might for example determine which situation- or person-specific differences discern positive from negative developments.
@@
 In particular, we propose that feature-based clustering is ideally suited because it offers a flexible variety of validated and intuitive components while also allowing for adequate control by the researcher.
-To showcase the utility of this technique, we apply feature-based clustering to a set of three ESM studies following recent migrants in their daily interactions and relations with the cultural majority group.
+To showcase the utility of this technique, we apply feature-based clustering to a set of three ESM studies with complex conceptualizations and missingness patterns.
 
 \section{Modern ESM Data}
+\centerline{<--- \textit{still horrible, probably restructure section} --->}
+
 Two important developments within the psychological sciences have been (1) a shift towards more comprehensive assessments of psychological phenomena and (2) a growing focus on intensive longitudinal real-life data.
-As an example, within the cultural adaptation literature prominent reviews have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time conceptual works have pushed for more comprehensive assessments of human experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-Such a multivariate, experience-sampling approach allows researchers to test theories more exhaustively and helps build a more embedded understanding of migration processes.
+Multivariate experience-sampling approaches allow researchers to test theories more exhaustively and help build a more embedded understanding of human experience processes.
 However, such data also deviates from past intensive longitudinal data sets that have focused on psychological states that are more homogeneous in their developmental characteristics.
@@
 Concerning dimensionality issues, especially more abstract psychological experiences often need a wider variety of measurements to be captured adequately.
-Today, few clinical conditions are captured with a single symptom measure, emotions are rarely assessed in isolation, and as introduced earlier cultural adaptation has at least four psychological aspects.
-This also means that modern analysis techniques increasingly need be able to accommodate this increased focus on multivariate developments.
-The recent rise in psychological network models has highlighted this need for more multivariate outcome measurements (REFERENCE).
-Additionally, the increased number of variables also heightens the computational load for model estimations and especially when the number of variables and measurements compounds models may not converge --- an issue sometimes referred to as the dimensionality curse.
-A modern descriptive and exploratory analysis technique should consequently be able to summarize and structure the multivariate phenomena without running into computational load issues.
+Today, few clinical conditions are captured with a single symptom measure, emotions are rarely assessed in isolation, and socio-cultural experiences are now widely considered to be multimodal.
+This also means that modern analysis techniques increasingly need be able to accommodate an increased focus on multivariate developments.
+The recent rise in psychological network models has highlighted this need for more multivariate outcome measurements \citep[e.g.,][]{decancq2013}.
+Additionally, the increased number of variables also heightens the computational load for model estimations and, especially when the number of variables and measurements compounds, models may not converge --- an issue sometimes referred to as the dimensionality curse.
+A modern descriptive and exploratory analysis technique should consequently be able to summarize and structure multivariate phenomena without running into computational load issues.
 
@@
 More event-based assessments tend to encounter issues when models require time intervals between measurements to be identical (i.e., equidistant measurements) and an increased focus on context-specific phenomena has lead to issues when data is only available under certain conditions, including follow-up questions (i.e., structurally missing data).
-Smaller issues of non-equidistant data can be avoided with transformations (e.g., dynamic time warping, REFERENCE) or newer modeling procedures (e.g., continuous-time models, REFERENCE) but it's prevalence is indicative of the structure of real-world issues researchers seek to investigate.
+Smaller issues of non-equidistant data can be avoided with transformations \citep[e.g., dynamic time warping,][]{berndt1994} or newer modeling procedures \citep[e.g., continuous-time models][]{dehaan-rietdijk2017} but it's prevalence is indicative of the structure of real-world issues researchers seek to investigate.
 Structural missingness remains a much more strenuous challenge.
-Many of the most commonly used models need complete data (including, ...) and structurally missing data cannot be imputed as it logically does not exist (REFERENCE).
+Many of the most commonly used models need complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.
 This type of missingness becomes a colossal issue when researchers are interested in how two things co-develop and one of the concepts has structural missingness under certain conditions.
-In the example of the migration experiences, researchers might for example be interested in how well-being develops in relations the quality of interactions with the majority group.
-Because rating of interaction quality are only possible when an interaction actually happened, one cannot impute interaction quality ratings for measurements that did not include an interaction.
-The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data (REFERENCE) --- neither option suits a research question that includes a developmental relationship between the two variables.
+The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008} --- neither option suits a research question that includes a developmental relationship between the two variables.
 In short, adequate descriptive approaches should be able to deal with non-equidistant and structurally missing data in order to address modern ESM data.
@@
 Firstly, researchers are increasingly interested in how mean levels of variables change over time (i.e., non-stationary), such trend data is often assumed to be non-linear, and with the increased focus on more diverse conceptualizations researchers are interested in relationships between variables that change on fundamentally different time scales.
-Many of the commonly used analyses for ESM data are stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).
+Many of the commonly used analyses for ESM data are stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or use basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).
 And even fewer analysis techniques allow for substantially different time scales between variables.
@@
 \section{Time Series Clustering}
-One promising approach to describing and exploring time-series data has been the idea of structuring the data using clustering analyses.
+One promising approach to describing and exploring time-series data has been the idea of structuring data using clustering analyses.
 Clustering seeks to organize data into groups that are maximally different from one another while minimizing differences within the groups.
-Time-series clustering has been extremely common in other fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, or energy and finance patterns \citep{Aghabozorgi2015}.
+Time-series clustering has been extremely common in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
 But also for psychological data time-series clustering is not entirely novel \citep[e.g.,][]{ernst2021}.
-However, most clustering introduced for psychological data has faced the dimensionality, -missingness, and time scale issues described above.
+However, most clustering introduced for psychological data has faced the dimensionality-, missingness-, and time scale issues described above.
 
@@
 Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time series, but also linear and non-linear trends, inertia-related features such as auto-correlations, or the stability of the time series (e.g., mean absolute change).
-Given the flexibility of the feature-based approach, users should be able to apply the analysis across multivariate data, feature extraction should mitigate issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
+Given the flexibility of the feature-based approach, users are able to apply the analysis across multivariate data, feature extraction mitigates issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
 
@@
 
-intro-paragraph + \fgrref{fig:TSCFlow}
+The feature-based clustering can be structured into four main steps.
+(1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, (3) an optional feature reduction step if there are too many data points for the clustering algorithms, and (4) the actual clustering of the time series features.
+We provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
 
@@
 This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values.
-It is thus generally advised to reduce the data to a common time frame.
+It is thus generally advised to reduce the data to a common and comparable time frame.
 
 \subsection{Feature extraction}
-The main aim of feature extraction is 
-hundreds of possible features to describe psychological time series.
+The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
+There are, however, hundreds of possible features to describe psychological time series \citep[e.g., tsfresh][]{christ2018}.
 One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series.
@@
 More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
-This is particularly the case when features have a conceptual/psychological meaning.
+This is particularly the case when features have a conceptual or psychological meaning.
 Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
@@
 The first two features are a person's \textit{central tendency} and \textit{variability}.
-Familiar statistics from the probability theory, the two features sit at the heart of many fundamental psychological questions.
-Are some people happier than others (i.e., difference in central tendency), and are people settled in their attitudes towards migrants or do they fluctuate over time (i.e., difference in variability)?
+Familiar statistics from probability theory, the two features sit at the heart of many fundamental psychological questions.
+``Are some people happier than others?'' (i.e., difference in central tendency), and "Are people settled in their attitudes towards migrants or do they fluctuate over time?'' (i.e., difference in variability).
 Mathematically, both features have parametric and robust options to choose from (e.g., parametric: mean and standard deviation; robust: median and median absolute deviation).
@@
 Both forms of inertia relate to different types of research questions.
-Do patients stay in a depressed mood for several measurements (i.e., stability-inertia), and do participants drink more alcohol on Fridays (i.e., periodicity-inertia)?
+``Do patients stay in a depressed mood for several measurements?'' (i.e., stability-inertia), and ``Do participants drink more alcohol on Fridays?'' (i.e., periodicity-inertia).
 Both forms of inertia are often measured using autocorrelations with either a lag-1 (stability-inertia) or the lag of the seasonality (e.g., seven days).
@@
 \textit{Linear trends} are a common research interest for longitudinal data.
-Do patient symptoms improve or does worker productivity decline, are familiar types of research questions.
+``Do patient symptoms improve?'' or ``Does worker productivity decline?'', are familiar types of research questions.
 Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients.
 \textit{Nonlinear trends} are important in two regards.
-Firstly nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
+Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
 Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}.
@@
 Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice.
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice (see \tblref{tab:esmFeatures} for a more complete overview).
 
@@
 Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points usually remains too large for most clustering algorithms.
-As an example, a relatively common sample size of 100 participants and 10 variables of interest, where 8 time series features are extracted, still consists of 80 features per participant (a total of 8,000 data points in this hypothetical example).
+As an example, a relatively common scenario would include 10 variables of interest, where 8 time series features are extracted, resulting in 80 features per participant (with a common sample size of 100 participants that would result in a total of 8,000 data points in this hypothetical example).
 We offer an illustration of the compounding numbers of data points in \fgrref{fig:TSCFlowN}.
@@
 Traditional examples of wrappers are forward selection or backward elimination procedures.
-Because filter methods might not always perform well and wrapper methods are computationally intensive, hybrid methods seek to combine the two methods and find a balance between computational effort (i.e., efficiency) and performance (i.e., effectiveness).
-As an example, some of the common methods use a filter step to reduce the size of features considered in a later wrapper step.
+Because filter methods might not always perform well and wrapper methods are computationally intensive\footnote{Computationally \textit{k} features could be considered in \(2^k – 1\) possible combinations --- for the example of 80 features that would allow for \(1.20 \times 10^{24}\) (over one septillion) combinations.}, hybrid methods seek to combine the two methods and find a balance between computational effort (i.e., efficiency) and performance (i.e., effectiveness).
+Methods might, for example, use a filter step to reduce the size of features considered in a later wrapper step.
 
 The second general approach to the dimensionality curse has been feature projection.
-Projections methods are relatively common in psychological research --- particularly, factor- and principal component analyses.
+Projection methods are relatively common in psychological research --- including, factor- and principal component analyses.
 Generally speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features (i.e., the data is projected to a lower dimensional space).
@@
 Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment (or similarity measurement for qualitative features).
-While others have produced excellent overviews of the many clustering approaches \citep[e.g.,][]{xu2015}, we will briefly introduce some of the more readily available approaches suitable for most time series feature data.
-The well-established and readily available clustering approaches can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof (also see \tblref{tab:clusterApproaches}).
+While others have produced excellent overviews of the many clustering approaches available \citep[e.g.,][]{xu2015}, we will briefly introduce some of the more readily available approaches suitable for most time series feature data.
+The well-established and readily available clustering approaches can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof (see \tblref{tab:clusterApproaches}).
 
@@
 As an example, under ideal conditions, all approaches are likely to suggest very similar clusters.
-However, as an example, when the shape of clusters becomes more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate.
+However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate.
 Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective.
@@
 
-There is thus, unfortunately, no one-size fits all solution to clustering.
-On the bright side, however, the thriving methodological developments in the clustering literature offer a wide variety of options that offer options even for extremely large and complex data sets.
+There is thus, unfortunately, no one-size-fits-all solution to clustering.
+On the bright side, however, the thriving methodological developments in the clustering literature offer a wide variety of options that offer prospects even for extremely large and complex data sets.
 We provide a short intuitive explanation for common approaches, together with some of their characteristics and example algorithms in \tblref{tab:clusterApproaches}.
@@
 \section{Empirical Illustration}
-To illustrate the use case of the feature-based time series with psychological ESM data, we ....
+To illustrate common use cases of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected diverse concepts with context-specific measurements.
+The illustration ...
 
-Something about the migration experience literature should go here as part of the illustration intro ...
+\hl{Move migration experience literature background here ...}
 
@@
 "The goal of clustering is to identify structure in an unlabeled data set by objectively organizing data into homogeneous groups where the within-group-object dissimilarity is minimized and the between-group-object dissimilarity is maximized." --- \citep[][p.1857]{liao2005}
+
+\subsection{Migration Parts}
+One example of this phenomenon has been the cultural adaptation literature, where researchers have started using ESM data to follow the daily interactions of migrants with the local majority \citep[e.g.,][]{Kreienkamp2022b}.
+
+As an example, within the cultural adaptation literature, prominent reviews have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time conceptual works have pushed for more comprehensive assessments of human experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
+
+In the example of the migration experiences, researchers might for example be interested in how well-being develops in relations the quality of interactions with the majority group.
+Because rating of interaction quality are only possible when an interaction actually happened, one cannot impute interaction quality ratings for measurements that did not include an interaction.
 

### 2022-11-29 methods-and-results.Rmd
:art: text col
:lipstick: wording

@@
   chunk_output_type: console
-bibliography: references.bib
+bibliography: referencesZotero.bib
 csl: apa.csl
@@
 load("data/features/featOut.RData")
+#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
+#load("data/features/featOutFullImp.RData")
 ```
@@
 \subsubsection{Variables}
-Included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
+We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
 We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
@@
 \subsection{Analysis and Results}
-To perform the main analysis on our selected variables, we follow the three extract--reduce--cluster steps in sequential order.
+To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster steps in sequential order.
 For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REFERENCES).
@@
 We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
-We include a function that automatically extracts and prepares all features we extracted as well as a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
+We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
 
-After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across participants across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
+After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
 This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
 The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
-The few missing values can however be an issue for some feature reduction or feature clustering algorithms.
+The few missing values can, however, be an issue for some feature reduction or feature clustering algorithms.
 We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
@@
 
-For the feature reduction, we chose the very general \textit{principal component analysis} (PCA).
+For the feature reduction, we chose the common \textit{principal component analysis} (PCA).
 Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
@@
 The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
-In practice, this means that the PCA decomposes the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
+In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
 We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
 A common rule of thumb is to use the principal components that jointly explain 70--90\% of the original variance (i.e., cumulative percentage explained variance).
-For our illustration we we select the first `r pca_cutoff` principal components that explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
-For the extracted principal components we save the `r ncol(scaled_data)` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
+For our illustration we select the first `r pca_cutoff` principal components that explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
+For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
 
 We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
-As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' with (highly) correlated variables.
+As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables.
 Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
-While our `r pca_cutoff` principal components is still a relatively high number of variables this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
+Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
 As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
@@
 
-\Question{\textcolor}{cyan}{It is good practice to standardize the input of the k-means algorithm.
+\Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
 However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
@@
 
-In order to interpret the two clusters, we compared the two clusters in their features as well as in their raw time series.
+In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
 We first inspect the clusters with a focus on the included variables (see \fgrref{fig:??}).
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., how cooperative an interaction was) than for other variables (e.g., Attitudes towards the Dutch).
-Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being) than others (e.g., MAC of well-being).
+We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
+Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
+MAC of well-being).
 We then inspect the clusters with a focus on the features (see \fgrref{fig:??}).
-While this is the same data as in the plot above, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median) than others (e.g., auto correlations).
+While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
 This offers some information on which features were are most important in understanding the two extracted groups.
@@
 We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:??}).
-Immediately striking are the mean differences, where participants in the cluster that had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
+Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
 The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
 Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-The 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref{fig:??}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time.
+This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref{fig:??}).
+We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref{fig:??}).
 
-\Question{\textcolor{cyan}{Talk about early warning signals here? I.e., divergences in means over time could be invesitigated with targeted studies.}}
+\Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}}
 
@@
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants with more negative experiences on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions.
+To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions.
 While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.

### 2022-11-29 methods-and-results.tex
:art: text col
:lipstick: wording

@@
 
-Included 12 variables that were available across all three studies and
-captured information about the participant's interactions, as well as
-cognitive-, emotional-, and motivational self in relationship with the
-majority group. We chose these two aspects in particular because (1) the
-interaction-specific information exemplified the structural missingness
-issue of modern ESM data and (2) the motivational, emotional, and
-cognitive experience offered a diverse conceptualization of migration
-experience (beyond behavioral measurements) that is becoming more common
-in the literature \citep[][]{Kreienkamp2022d}. Full methodological
-details are available in Online Supplemental Material A, but basic item
-information, descriptives, and correlations are available in
-\tblref{tab:descrWide}.
+We included 12 variables that were available across all three studies
+and captured information about the participant's interactions, as well
+as cognitive-, emotional-, and motivational self in relationship with
+the majority group. We chose these two aspects in particular because (1)
+the interaction-specific information exemplified the structural
+missingness issue of modern ESM data and (2) the motivational,
+emotional, and cognitive experience offered a diverse conceptualization
+of migration experience (beyond behavioral measurements) that is
+becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
+methodological details are available in Online Supplemental Material A,
+but basic item information, descriptives, and correlations are available
+in \tblref{tab:descrWide}.
 
@@
 
-To perform the main analysis on our selected variables, we follow the
-three extract--reduce--cluster steps in sequential order. For each of
-the steps fully reproducible and annotated code is available in Online
+To perform the main analysis on our selected variables, we showcase the
+extract--reduce--cluster steps in sequential order. For each of the
+steps fully reproducible and annotated code is available in Online
 Supplemental A as well as in the accompanying OSF- and GitHub
@@
 measurements to ensure that the clusters are comparable in that regard.
-We include a function that automatically extracts and prepares all
-features we extracted as well as a large selection of the time series
-feature operationalizations presented in \tblref{tab:esmFeatures} in our
-GitHub repository (see the \texttt{featureExtractor} function).
+We provide an R-function that automatically extracts and prepares a
+large selection of the time series feature operationalizations presented
+in \tblref{tab:esmFeatures} in our GitHub repository (see the
+\texttt{featureExtractor} function).
 
 After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across participants across the 72
-features per participant. This might, for example, happen if
-participants do not have two subsequent measurements with outgroup
-interactions, so that an autocorrelation with lag-1 cannot be calculated
-for the contact-specific variables. The small number of missing values
-indicates that the feature-based approach indeed largely avoids the
-structural missingness issue. The few missing values can however be an
-issue for some feature reduction or feature clustering algorithms. We,
-thus, impute the missing feature values with a single predictive mean
-matching imputation using the MICE library.
+extracted features are missing across the 72 features per participant.
+This might, for example, happen if participants do not have two
+subsequent measurements with outgroup interactions, so that an
+autocorrelation with lag-1 cannot be calculated for the contact-specific
+variables. The small number of missing values indicates that the
+feature-based approach indeed largely avoids the structural missingness
+issue. The few missing values can, however, be an issue for some feature
+reduction or feature clustering algorithms. We, thus, impute the missing
+feature values with a single predictive mean matching imputation using
+the MICE library.
 
@@
 
-For the feature reduction, we chose the very general
+For the feature reduction, we chose the common
 \textit{principal component analysis} (PCA). Some of the more
@@
 also ensuring that the components are linearly uncorrelated. In
-practice, this means that the PCA decomposes the 72 features into 72
+practice, this meant that the PCA decomposed the 72 features into 72
 principal components but now (because of the uncorrelated linear
@@
 explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we we select the first 27
+explained variance). For our illustration we select the first 27
 principal components that explain 80\% of the variance in the original
 72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 72 PC-scores for each participant
+principal components we save the 27 PC-scores for each participant
 (i.e., the participants' coordinates in the reduced dimensional space).
@@
 dimensionality reduction, given its methodology, PCAs tend to `work
-better' with (highly) correlated variables. Thus, with a set of very
-homogeneous variables and features users will need less principal
-components to explain a large amount of variance, while a more diverse
-set of variables and features will tend to require more principal
-components to capture the same amount of variance. While our 27
-principal components is still a relatively high number of variables this
-is not surprising as we chose a diverse conceptualization and a diverse
-set of time series features. As for the interpretability, PCA allows
-users to extract information on the meaning of the principal components.
-In particular, because the principal components are linear combination
-of original features, users can extract the relative importance of each
-feature for the extracted principal components (i.e., the eigenvectors).
-While this can be useful in understanding the variance in the original
-data or help with manual feature selection, we use the PCA purely to
-reduce the dimensionality for the clustering step. Instead of relying on
-the principal components, we use the original features of interest to
-interpret the later extracted clusters. We particularly advocate for
-such an approach if the original features were chosen for their
-psychological meaning in understanding the time series and broader
-phenomenon of interest.
+better' at reducing dimensions with (highly) correlated variables. Thus,
+with a set of very homogeneous variables and features users will need
+less principal components to explain a large amount of variance, while a
+more diverse set of variables and features will tend to require more
+principal components to capture the same amount of variance. Our 27
+principal components are still a relatively high number of variables but
+this is not surprising as we chose a diverse conceptualization and a
+diverse set of time series features. As for the interpretability, PCA
+allows users to extract information on the meaning of the principal
+components. In particular, because the principal components are linear
+combination of original features, users can extract the relative
+importance of each feature for the extracted principal components (i.e.,
+the eigenvectors). While this can be useful in understanding the
+variance in the original data or help with manual feature selection, we
+use the PCA purely to reduce the dimensionality for the clustering step.
+Instead of relying on the principal components, we use the original
+features of interest to interpret the later extracted clusters. We
+particularly advocate for such an approach if the original features were
+chosen for their psychological meaning in understanding the time series
+and broader phenomenon of interest.
 
@@
 "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
-but I went back to the 'raw' PC-scores now. But we should discuss this.}\}
+but I went back to the 'raw' PC-scores now. But we should discuss this.}}
 
@@
 
-In order to interpret the two clusters, we compared the two clusters in
+In order to interpret the two clusters, we compare the two clusters in
 their features as well as in their raw time series. We first inspect the
@@
 We (1) see that for some variables the features are generally stronger
-in separating the clusters (e.g., how cooperative an interaction was)
-than for other variables (e.g., Attitudes towards the Dutch).
-Additionally, we see that (2) within the variables some features are
-better at distinguishing clusters (e.g., median of well-being) than
-others (e.g., MAC of well-being). We then inspect the clusters with a
-focus on the features (see \fgrref{fig:??}). While this is the same data
-as in the plot above, we can see more clearly that some features are
-better at distinguishing the clusters across variables (e.g., mean and
-median) than others (e.g., auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
+in separating the clusters (e.g., `how cooperative an interaction was'
+compared to `attitudes towards the Dutch'). Additionally, we see that
+(2) within the variables some features are better at distinguishing
+clusters (e.g., median of well-being vs.~MAC of well-being). We then
+inspect the clusters with a focus on the features (see \fgrref{fig:??}).
+While this is the same data as for the variable focus, we can see more
+clearly that some features are better at distinguishing the clusters
+across variables (e.g., mean and median compared to auto correlations).
+This offers some information on which features were are most important
+in understanding the two extracted groups.
 
@@
 Immediately striking are the mean differences, where participants in the
-cluster that had more meaningful and fulfilling outgroup interactions
+second cluster had more meaningful and fulfilling outgroup interactions
 also consistently reported more voluntary and cooperative interactions
@@
 outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. The
+cooperative, and positive interactions over the 30 days. This
 `deterioration' cluster also saw a decrease in general need fulfillment
@@
 cluster also showed substantially higher variablity and instability over
-time.
+time (also see \fgrref{fig:??}).
 
-\Question{\textcolor{cyan}{Talk about early warning signals here? I.e., divergences in means over time could be invesitigated with targeted studies.}}
+\Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}}
 
@@
 artifact checks, we added the number of measurements into the comparison
-and find that the participants with more negative experiences on average
+and find that the participants in the detereoration cluster on average
 completed more ESM surveys and reported on more intergroup interactions

### 2022-11-29 methods-and-results.Rmd
:sparkles: cluster plts

@@
 ```{r average feature by cluster, include=FALSE}
+variableLab <- c(
+  "AttitudesDutch" = "Outgroup Attitude",
+  "AttitudesPartner" = "Int: Partner Attitude",
+  "DaytimeNeedFulfillment" = "Need Fulfillment",
+  "exWB" = "Well-Being",
+  "InteractionContextAccidental" = "Int: Accidental",
+  "InteractionContextCooperative" = "Int: Cooperative",
+  "InteractionContextRepresentativeNL" = "Int: Representative",
+  "InteractionContextvoluntary" = "Int: Voluntary",
+  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
+  "KeyNeedFulfillment" = "Int: Need Fulfillment",
+  "qualityMeaning" = "Int: Meaningful",
+  "qualityOverall" = "Int: Quality"
+) %>% enframe(., name = "variable", value = "label")
+
 featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
@@
   mutate(
-    lab = variable,
-    variable = gsub("\\_.*", "", lab),
-    feature = gsub("^.*?\\_", "", lab)
-  ) %>% 
+    nam = variable,
+    variable = gsub("\\_.*", "", nam),
+    feature = gsub("^.*?\\_", "", nam)
+  ) %>% 
+  merge(., variableLab) %>%
   select(
     cluster,
-    lab,
+    nam,
     variable,
+    label,
     feature,
@@
   )
+
+rawCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., raw_data) %>%
+  select(
+    ID,
+    TIDnum,
+    cluster,
+    any_of(varNamS123)
+  ) %>%
+  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+  merge(., variableLab)
 ```
@@
 ```{r clusterVariableGrid, include=FALSE}
-featCluster %>% 
+clusterVariableGrid <- featCluster %>% 
   ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
@@
   ) +
-  facet_wrap(~ variable) +
+  facet_wrap(~ label) +
   theme_Publication()
+clusterVariableGrid
 ```
@@
 ```{r clusterFeatureGrid, include=FALSE}
-featCluster %>% 
-  ggplot(., aes(x = variable, y = mean, group = cluster, color = as.factor(cluster))) +
+clusterFeatureGrid <- featCluster %>% 
+  ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
   geom_point() +
@@
   ) +
-  facet_wrap(~ feature) +
+  facet_wrap(~ feature, ncol=4) +
   theme_Publication()
+clusterFeatureGrid
+```
+
+```{r clusterFeatVarComb, include=FALSE}
+clusterFeatVarComb <- ggpubr::ggarrange(
+  clusterVariableGrid + theme(axis.title.x = element_blank()),
+  clusterFeatureGrid,
+  ncol = 1,
+  labels = c("(A) Variable Focus", "(B) Feature Focus"),
+  common.legend = TRUE,
+  legend = "bottom",
+  align = "v"
+) 
+ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 10)
 ```
@@
 ```{r clusterTS, include=FALSE}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+clusterTS <- rawCluster %>% 
   ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
@@
   stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ variable, scales = "free_y") +
+  facet_wrap(~ label, scales = "free_y") +
   labs(
@@
   theme_Publication() 
+clusterTS
 ```
@@
 ```{r clusterTSTrend, include=FALSE}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
+clusterTSTrend <- rawCluster %>% 
   ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
@@
   #stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ variable, scales = "free_y") +
+  facet_wrap(~ label, scales = "free_y") +
   labs(
@@
   theme_Publication()
-```
+clusterTSTrend
+```
+
+```{r clusterTsComb, include=FALSE}
+clusterTsComb <- ggpubr::ggarrange(
+  clusterTS + theme(axis.title.x = element_blank()),
+  clusterTSTrend,
+  ncol = 1,
+  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
+  common.legend = TRUE,
+  legend = "bottom"
+)
+ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 10)
+```

### 2022-11-30 methods-and-results.Rmd
:art: fig sizes

@@
 ) 
-ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 10)
+ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 12)
 ```
@@
 )
-ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 10)
-```
+ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 12)
+```

### 2022-11-30 methods-and-results.Rmd
:pencil: fig refs

@@
 The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see \fgrref{fig:??}).
+We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see Supplemental Material A).
 We then entered the participants' PC-scores into the k-means algorithm.
@@
 In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
-We first inspect the clusters with a focus on the included variables (see \fgrref{fig:??}).
+We first inspect the clusters with a focus on the included variables (see \fgrref[A]{fig:clusterFeatVar}).
 We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
 Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).
-We then inspect the clusters with a focus on the features (see \fgrref{fig:??}).
+MAC of well-being).We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
 While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
@@
 
-We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:??}).
+We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
 Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
@@
 Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref{fig:??}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref{fig:??}).
+This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
+We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).
 

### 2022-11-30 methods-and-results.tex
:pencil: fig refs

@@
 Silhouette method, which both suggest two clusters to be the optimal
-solution (see \fgrref{fig:??}). We then entered the participants'
-PC-scores into the k-means algorithm. We used the recommended Hartigan
-and Wong algorithm \citep{hartigan1979} with 100 random initial centroid
-positions to avoid convergence to a sub-optimal solution (i.e., local
-minima). The k-means analysis assigned 80 participants to cluster 1 and
-76 participants to cluster 2.
+solution (see Supplemental Material A). We then entered the
+participants' PC-scores into the k-means algorithm. We used the
+recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
+random initial centroid positions to avoid convergence to a sub-optimal
+solution (i.e., local minima). The k-means analysis assigned 80
+participants to cluster 1 and 76 participants to cluster 2.
 
@@
 their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see \fgrref{fig:??}).
-We (1) see that for some variables the features are generally stronger
-in separating the clusters (e.g., `how cooperative an interaction was'
-compared to `attitudes towards the Dutch'). Additionally, we see that
-(2) within the variables some features are better at distinguishing
-clusters (e.g., median of well-being vs.~MAC of well-being). We then
-inspect the clusters with a focus on the features (see \fgrref{fig:??}).
-While this is the same data as for the variable focus, we can see more
-clearly that some features are better at distinguishing the clusters
-across variables (e.g., mean and median compared to auto correlations).
-This offers some information on which features were are most important
-in understanding the two extracted groups.
+clusters with a focus on the included variables (see
+\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
+features are generally stronger in separating the clusters (e.g., `how
+cooperative an interaction was' compared to `attitudes towards the
+Dutch'). Additionally, we see that (2) within the variables some
+features are better at distinguishing clusters (e.g., median of
+well-being vs.~MAC of well-being).We then inspect the clusters with a
+focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
+is the same data as for the variable focus, we can see more clearly that
+some features are better at distinguishing the clusters across variables
+(e.g., mean and median compared to auto correlations). This offers some
+information on which features were are most important in understanding
+the two extracted groups.
 
 We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:??}).
+between the two groups more holistically (see \fgrref{fig:clusterTs}).
 Immediately striking are the mean differences, where participants in the
@@
 `deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see \fgrref{fig:??}). We
-also see that while interaction representativeness, outgroup attitudes,
-well-being are relatively stable for both clusters, the deteriorating
-cluster also showed substantially higher variablity and instability over
-time (also see \fgrref{fig:??}).
+and experienced well-being over the 30 days (see
+\fgrref[B]{fig:clusterTs}). We also see that while interaction
+representativeness, outgroup attitudes, well-being are relatively stable
+for both clusters, the deteriorating cluster also showed substantially
+higher variablity and instability over time (also see
+\fgrref[A]{fig:clusterTs}).
 

### 2022-11-30 methods-and-results.Rmd
:bug: tbl size

@@
 ```{r descrWide, include=FALSE}
+variableLab <- c(
+  "AttitudesDutch" = "Outgroup Attitude",
+  "AttitudesPartner" = "Int: Partner Attitude",
+  "DaytimeNeedFulfillment" = "Need Fulfillment",
+  "exWB" = "Well-Being",
+  "InteractionContextAccidental" = "Int: Accidental",
+  "InteractionContextCooperative" = "Int: Cooperative",
+  "InteractionContextRepresentativeNL" = "Int: Representative",
+  "InteractionContextvoluntary" = "Int: Voluntary",
+  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
+  "KeyNeedFulfillment" = "Int: Need Fulfillment",
+  "qualityMeaning" = "Int: Meaningful",
+  "qualityOverall" = "Int: Quality"
+) %>% enframe(., name = "variable", value = "label")
+
 dtAll <- rbind(
@@
     labels = c(
-      "Interaction Accidental", 
-      "Interaction Voluntary", 
-      "Interaction Cooperative",
-      "Interaction Representative NL",
-      "Quality Meaningful",
-      "Quality Overall", 
+      "Int: Accidental", 
+      "Int: Voluntary", 
+      "Int: Cooperative",
+      "Int: Representative",
+      "Int: Meaningful",
+      "Int: Quality", 
       "Core Need", 
       "Core Need Due to Partner", 
-      "Attitudes Partner",
+      "Attitude Partner",
       "Daytime Core Need",
-      "Attitudes Dutch",
+      "Outgroup Attitude",
       "Well-being"

### 2022-11-30 methods-and-results.Rmd
:sparkles: descr long tbl

@@
   save_kable("Tables/descrWide.tex")
-
+```
+
+```{r descrLong, include=FALSE}
+# Sample Descriptives Table --- longer version
+allMlCor %>%
+  as.data.frame %>%
+  kbl(.,
+      format = "latex",
+      caption = "Correlation Table and Descriptive Statistics",
+      booktabs = TRUE,
+      align = c("l", rep("c", ncol(.) - 1))) %>%
+  kable_classic() %>%
+  # add_header_above(
+  #   .,
+  #   c(
+  #     "",
+  #     "Correlations" = ncol(allMlCor),
+  #     "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
+  #   )
+  # ) %>%
+  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
+  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
+  footnote(
+    general = c(
+      "Upper triangle: Between-person correlations;",
+      "Lower triangle: Within-person correlations;",
+      "*** p < .001, ** p < .01,  * p < .05"
+    )
+  ) %>%
+  kable_styling(latex_options = "scale_down") %>%
+  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
+  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
+  save_kable("Tables/descrWide.tex")
 ```

### 2022-11-30 methods-and-results.Rmd
:bug:

@@
       booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
+      align = c("l", rep("c", ncol(.) - 1)),
+      col.names = c(
+        #"",
+        "Int: \n Accidental", 
+        "Int: \\newline Voluntary", 
+        "Int: \n Cooperative",
+        "Int: \n Representative",
+        "Int: \n Meaningful",
+        "Int: \n Quality", 
+        "Core Need", 
+        "Core Need \n Due to Partner", 
+        "Attitude \n Partner",
+        "Daytime \n Core Need",
+        "Outgroup \n Attitude",
+        "Well-being"
+      )) %>%
   kable_classic() %>%
-  # add_header_above(
-  #   .,
-  #   c(
-  #     "",
-  #     "Correlations" = ncol(allMlCor),
-  #     "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-  #   )
-  # ) %>%
   pack_rows("Correlations", 1, ncol(allMlCor)) %>%
@@
   gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
+  save_kable("Tables/descrLong.tex")
 ```

### 2022-11-30 methods-and-results.Rmd
:art: descr tbl

--- 2022-11-30 methods-and-results.Rmd
@@
       booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
+      align = "c", #c("l", rep("c", ncol(.) - 1)),
+      escape = FALSE,
+      col.names = c(
+        #"",
+        "Int: \\newline Accidental", 
+        "Int: \\newline Voluntary", 
+        "Int: \\newline Cooperative",
+        "Int: \\newline Representative",
+        "Int: \\newline Meaningful",
+        "Int: \\newline Quality", 
+        "Core Need", 
+        "Core Need \\newline Due to Partner", 
+        "Attitude \\newline Partner",
+        "Daytime \\newline Core Need",
+        "Outgroup \\newline Attitude",
+        "Well-being"
+      )) %>%
   kable_classic() %>%
-  # add_header_above(
-  #   .,
-  #   c(
-  #     "",
-  #     "Correlations" = ncol(allMlCor),
-  #     "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-  #   )
-  # ) %>%
   pack_rows("Correlations", 1, ncol(allMlCor)) %>%

### 2022-11-30 methods-and-results.Rmd
:bug: descr tbl header

--- 2022-11-30 methods-and-results.Rmd
@@
       booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
+      align = "c", #c("l", rep("c", ncol(.) - 1)),
+      escape = FALSE,
+      col.names = c(
+        #"",
+        "\\makecell{Int: \\\\ Accidental}",
+        "\\makecell{Int: \\\\ Voluntary}", 
+        "\\makecell{Int: \\\\ Cooperative}",
+        "\\makecell{Int: \\\\ Representative}",
+        "\\makecell{Int: \\\\ Meaningful}",
+        "\\makecell{Int: \\\\ Quality}", 
+        "Core Need", 
+        "\\makecell{Core Need \\\\ Due to Partner}", 
+        "\\makecell{Attitude \\\\ Partner}",
+        "\\makecell{Daytime \\\\ Core Need}",
+        "\\makecell{Outgroup \\\\ Attitude}",
+        "Well-being"
+      )) %>%
   kable_classic() %>%
-  # add_header_above(
-  #   .,
-  #   c(
-  #     "",
-  #     "Correlations" = ncol(allMlCor),
-  #     "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-  #   )
-  # ) %>%
   pack_rows("Correlations", 1, ncol(allMlCor)) %>%

### 2022-11-30 manuscript.tex
:sparkles: figs added
:pnecil2: typo
Updates from Overleaf
:sparkles: migration example intro

@@
 To illustrate common use cases of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected diverse concepts with context-specific measurements.
-The illustration ...
-
-\hl{Move migration experience literature background here ...}
+The illustration specifically uses data from research on migration experiences, where researchers have started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020}.
+This new type of research comes as a response to prominent reviews, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}.
+At the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
+The migration ESM research, thus, offers a good example of the modern ESM data that introduces diverse conceptualizations and more event-specific missingness patterns.
 
@@
 \input{methods-and-results}
-\input{tables/descrWide.tex}
+%\input{tables/descrWide}
+\input{tables/descrLong}
 
@@
 \subsection{Migration Parts}
-One example of this phenomenon has been the cultural adaptation literature, where researchers have started using ESM data to follow the daily interactions of migrants with the local majority \citep[e.g.,][]{Kreienkamp2022b}.
-
-As an example, within the cultural adaptation literature, prominent reviews have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time conceptual works have pushed for more comprehensive assessments of human experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-
 In the example of the migration experiences, researchers might for example be interested in how well-being develops in relations the quality of interactions with the majority group.
@@
   The presented number of participants, variables, and measurement occasions are somewhat arbitrary but generally represent common sample sizes found within the literature.
-Also the number of extracted clusters is presented for illustrative purposes only}
+Also the number of extracted clusters is presented for illustrative purposes only.}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+  \caption{Cluster Group Comparisons based on Features and Variables}
+  \label{fig:clusterFeatVar}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
+  \caption*{Note: \\
+  same data}
+\end{figure*}
+
+\begin{figure*}[hbtp]
+  \caption{Cluster Group Comparisons over time}
+  \label{fig:clusterTs}
+  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
+  \caption*{Note: \\
+  same data}
 \end{figure*}

### 2022-12-01 methods-and-results.Rmd
:sparkles: Out of cluster dif viz

@@
 ```{r average feature by cluster, include=FALSE}
+# CLUSTER VARIABLE LABELS
 variableLab <- c(
@@
 
+# ADD VALIDATION VARS:
+validationNam <- c(
+  "gender.pre",
+  "IntGrAnx01.pre",
+  "IntGrAnx02.pre",
+  "IntGrAnx03.pre",
+  "IntGrAnx04.pre",
+  "IntGrAnx05.pre",
+  "IntGrAnx06.pre",
+  "IntGrAnx07R.pre",
+  "IntGrAnx08R.pre",
+  "IntGrAnx09R.pre",
+  "IntGrAnx10R.pre",
+  "IntGrAnx11R.pre",
+  "IntGrAnx12R.pre",
+  "SWL01.pre",
+  "SWL02.pre",
+  "SWL03.pre",
+  "SWL04.pre",
+  "SWL05.pre",
+  "EvDayDiscr01.post",
+  "EvDayDiscr02.post",
+  "EvDayDiscr03.post",
+  "EvDayDiscr04.post",
+  "EvDayDiscr05.post",
+  "EvDayDiscr06.post",
+  "EvDayDiscr07.post",
+  "EvDayDiscr08.post",
+  "EvDayDiscr09.post"
+)
+
+# EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
+validation <- rbind(
+  dtS1Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S1", gender.pre = (gender.pre-2)*-1),
+  dtS2Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S3")
+) %>% 
+  group_by(study, PID) %>%
+  mutate(ID = cur_group_id()) %>%
+  ungroup %>%
+  group_by(study, ID) %>%
+  summarise(
+    gender = gender.pre,
+    IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
+    SWL.pre = rowMeans(across(starts_with("SWL")), na.rm = TRUE),
+    EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE)
+  ) %>%
+  ungroup %>%
+  distinct %>% 
+  mutate(
+    gender.z = scale(gender, scale = TRUE),
+    IntGrAnx.pre.z = scale(IntGrAnx.pre, scale = TRUE),
+    SWL.pre.z = scale(SWL.pre, scale = TRUE),
+    EvDayDiscr.post.z = scale(EvDayDiscr.post, scale = TRUE)
+  ) %>%
+  select(
+    ID,
+    #gender.z,
+    #IntGrAnx.pre.z,
+    #SWL.pre.z,
+    EvDayDiscr.post.z
+  )
+
+val <- validation %>% select(-ID) %>% names
+validationFeature <- validation
+for(feature in variableLab$variable) {
+  validationFeature <- validationFeature %>% bind_cols(., select(., setNames(val, paste(feature, val, sep = "_"))))
+}
+validationFeature <- validationFeature %>% select(-any_of(val))
+# For checking
+# validation <- rbind(
+#   dtS1Red %>% select(any_of(idVars),any_of(validationNam)) %>% mutate(study = "S1"),
+#   dtS2Red %>% select(any_of(idVars),any_of(validationNam)) %>% mutate(study = "S2"),
+#   dtS3Red %>% select(any_of(idVars),any_of(validationNam)) %>% mutate(study = "S3")
+# ) %>%
+#   mutate(
+#     IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx"))),
+#     SWL.pre = rowMeans(across(starts_with("SWL")))
+#   ) %>% 
+#   group_by(study, PID) %>%
+#   mutate(ID = cur_group_id()) %>%
+#   ungroup
+
+# CLUSTER VARIABLE LABELS
+variableLab <- c(
+  "AttitudesDutch" = "Outgroup Attitude",
+  "AttitudesPartner" = "Int: Partner Attitude",
+  "DaytimeNeedFulfillment" = "Need Fulfillment",
+  "exWB" = "Well-Being",
+  "InteractionContextAccidental" = "Int: Accidental",
+  "InteractionContextCooperative" = "Int: Cooperative",
+  "InteractionContextRepresentativeNL" = "Int: Representative",
+  "InteractionContextvoluntary" = "Int: Voluntary",
+  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
+  "KeyNeedFulfillment" = "Int: Need Fulfillment",
+  "qualityMeaning" = "Int: Meaningful",
+  "qualityOverall" = "Int: Quality"
+) %>% enframe(., name = "variable", value = "label")
+validationLab <- c(
+  "gender.z" = "Val: Gender",
+  "SWL.pre.z" = "Val: Sat. with life",
+  "IntGrAnx.pre.z" = "Val: Intergroup Anxiety",
+  "EvDayDiscr.post.z" = " Discrimination"
+) %>% enframe(., name = "variable", value = "label")
+
+data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., validation) %>%
+  group_by(cluster) %>%
+  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE))
+
 featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
@@
   group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se))) %>%
+  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
   ungroup %>% 
@@
   merge(., variableLab) %>%
+  select(
+    cluster,
+    nam,
+    variable,
+    label,
+    feature,
+    everything()
+  )
+
+featClusterVal <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., z_data) %>%
+  merge(., validationFeature) %>%
+  group_by(cluster) %>%
+  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
+  ungroup %>% 
+  select(-starts_with("ID")) %>%
+  pivot_longer(
+    cols = -c(cluster),
+    names_to = c("variable", ".value"),
+    names_pattern = "(.*)_(.*)"
+  ) %>% 
+  mutate(
+    nam = variable,
+    variable = gsub("\\_.*", "", nam),
+    feature = gsub("^.*?\\_", "", nam)
+  ) %>% 
+  merge(., variableLab) %>% 
+  mutate(feature = stri_replace_all_regex(
+    feature,
+    pattern=validationLab$variable,
+    replacement=validationLab$label,
+    vectorize=FALSE
+  )) %>% 
   select(
@@
 ```{r clusterFeatureGrid, include=FALSE}
-clusterFeatureGrid <- featCluster %>% 
+clusterFeatureGrid <- featClusterVal %>% 
   ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
@@
   ) +
-  facet_wrap(~ feature, ncol=4) +
+  facet_wrap(~ factor(feature, levels = unique(featClusterVal$feature)), ncol=4) +
   theme_Publication()
 clusterFeatureGrid
+
 ```

### 2022-12-01 methods-and-results.Rmd
:pencil2: validation sect. draft

@@
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions.
+To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
 While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
-To illustrate the assessment of individual differences, we added a common acculturation measurement that was completed in the pre- and post-questionnaire \hl{(VIA) as well as ...
-.
-We can see that TBA [...].}
+To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
+\fgrref[B]{fig:clusterTs} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
+Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
 
@@
   )
-
 val <- validation %>% select(-ID) %>% names
@@
 validationFeature <- validationFeature %>% select(-any_of(val))
-# For checking
-# validation <- rbind(
-#   dtS1Red %>% select(any_of(idVars),any_of(validationNam)) %>% mutate(study = "S1"),
-#   dtS2Red %>% select(any_of(idVars),any_of(validationNam)) %>% mutate(study = "S2"),
-#   dtS3Red %>% select(any_of(idVars),any_of(validationNam)) %>% mutate(study = "S3")
-# ) %>%
-#   mutate(
-#     IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx"))),
-#     SWL.pre = rowMeans(across(starts_with("SWL")))
-#   ) %>% 
-#   group_by(study, PID) %>%
-#   mutate(ID = cur_group_id()) %>%
-#   ungroup
 
@@
 
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., validation) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE))
-
 featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
@@
 clusterFeatureGrid
-
 ```

### 2022-12-01 methods-and-results.tex
:pencil2: validation sect. draft

@@
 completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions. While this
-difference could indicate that the clusters might not entirely be
-comparable in the response patterns, we can find some relief in our data
-exclusion procedures during which we ensured that the general time frame
-and completion rates were not too dissimilar. To illustrate the
-assessment of individual differences, we added a common acculturation
-measurement that was completed in the pre- and post-questionnaire
-\hl{(VIA) as well as ... . We can see that TBA [...].}
+than the cluster with the more positive interactions (see
+\fgrref[B]{fig:clusterTs}). While this difference could indicate that
+the clusters might not entirely be comparable in the response patterns,
+we can find some relief in our data exclusion procedures during which we
+ensured that the general time frame and completion rates were not too
+dissimilar. To illustrate the utility of individual differences, we
+compare the two samples in terms of the participants' self-reported
+discrimination experiences in the Netherlands (measured during the
+post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
+participants in the deteriorating cluster reported substantially higher
+levels of everyday discrimination. Thus, both intensive longitudinal and
+cross-sectional variables that were not included in the original
+clustering step can be used to explore and understand the cluster
+differences in more detail.

### 2022-12-03 manuscript.tex
:sparkles: abstract + discussion

@@
 
+% prevent multipage footnotes
+\interfootnotelinepenalty=10000
+
 % language settings
@@
 % Title and header
-\title{Describe and Explore: Using feature-based time-series clustering to understand modern intensive longitudinal data (Working Title)}
-\shorttitle{Migration Experience Trajectories}
+\title{Describe and Explore: Using feature-based time series clustering to understand modern intensive longitudinal data}
+\shorttitle{feature-based time series clustering}
 
@@
 \affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
-\affil[3]{Author order TBD (currently in first name alphabetical order)}
+\affil[3]{Author order TBD}
 
@@
 \abstract{
-Abstract to be written.
-
-\noindent\textbf{Public significance statement}: To be written.
+Psychological time series data has not only become more common but also more diverse and complex.
+Researchers and practitioners are more and more interested in non-stationary developmental patterns of multivariate concepts, across different time scales, while contextualized measurements bring about issues of non-imputable missingness.
+At the same time, clustering analyses offer the potential of reducing these complexities to important and meaningful patterns.
+Such procedures could aid researchers and practitioners in describing and exploring patterns across participants and could promote the creation of more embedded theories and interventions.
+In this manuscript, we propose feature-based time series clustering as a flexible, transparent, and well-grounded clustering approach that addresses the growing dimensionality, missingness, and time scale challenges.
+We introduce the individual feature extraction, -reduction, and -clustering steps and illustrate their utility with an empirical example.
+We show that time series features such as means, autocorrelations, and linear trends are not only familiar to psychological researchers but are also conceptually meaningful in interpreting time series clusters.
+Beyond the conceptual and methodological introduction, we also provide practical algorithm overviews and readily available code for data preparation, analysis, and interpretation.
 
@@
 \keywords{
-    TBD\\
+    time series analysis, feature-based clustering, intensive longitudinal data, ESM\\
     \vspace{1em}
-    \textit{Open Science Practices [?]:}
-    \noindent \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registration-color.png}} Preregistration, 
+    \textit{Open Science Practices:}
     \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data, \break
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
+    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
     \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
@@
 Outcomes from such an approach might inform both practical early-warning signals as well as academic theory-building efforts.
-There is, thus, a clear need to assess the utility of data-driven/bottom-up analysis procedures that are flexible enough to address the growing variety of intensive longitudinal psychological data.
+There is, thus, a clear need to assess the utility of data-driven analysis procedures that are flexible enough to address the growing variety of intensive longitudinal psychological data.
 
@@
 \section{Modern ESM Data}
-\centerline{<--- \textit{still horrible, probably restructure section} --->}
-
-Two important developments within the psychological sciences have been (1) a shift towards more comprehensive assessments of psychological phenomena and (2) a growing focus on intensive longitudinal real-life data.
+Two important developments within the psychological sciences have been a shift towards more comprehensive assessments of psychological phenomena and a growing contextualization of intensive longitudinal real-life data.
 Multivariate experience-sampling approaches allow researchers to test theories more exhaustively and help build a more embedded understanding of human experience processes.
 However, such data also deviates from past intensive longitudinal data sets that have focused on psychological states that are more homogeneous in their developmental characteristics.
-In particular, the new applications of ESM tend to bring new challenges for data dimensionality, -missingness, and time scales.
+It, thus, makes sense to briefly consider which challenges modern ESM data introduce and what qualities a present-day clustering analysis should possess.
+We particularly highlight issues of dimensionality, missingness, and time scales within contemporary ESM data.
 
@@
 The recent rise in psychological network models has highlighted this need for more multivariate outcome measurements \citep[e.g.,][]{decancq2013}.
-Additionally, the increased number of variables also heightens the computational load for model estimations and, especially when the number of variables and measurements compounds, models may not converge --- an issue sometimes referred to as the dimensionality curse.
+Beyond the capability to assess multivariate developments, an increased number of variables also heightens the computational load for model estimations and, especially when the number of variables and measurements compounds, models may not converge --- an issue sometimes referred to as the dimensionality curse.
 A modern descriptive and exploratory analysis technique should consequently be able to summarize and structure multivariate phenomena without running into computational load issues.
 
-The push for more contextualized models also presents new issues with missing data.
-More event-based assessments tend to encounter issues when models require time intervals between measurements to be identical (i.e., equidistant measurements) and an increased focus on context-specific phenomena has lead to issues when data is only available under certain conditions, including follow-up questions (i.e., structurally missing data).
-Smaller issues of non-equidistant data can be avoided with transformations \citep[e.g., dynamic time warping,][]{berndt1994} or newer modeling procedures \citep[e.g., continuous-time models][]{dehaan-rietdijk2017} but it's prevalence is indicative of the structure of real-world issues researchers seek to investigate.
-Structural missingness remains a much more strenuous challenge.
-Many of the most commonly used models need complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.
-This type of missingness becomes a colossal issue when researchers are interested in how two things co-develop and one of the concepts has structural missingness under certain conditions.
-The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008} --- neither option suits a research question that includes a developmental relationship between the two variables.
+The push for more contextualized data also presents two key issues with missing data.
+More event-based assessments tend to (1) encounter issues when models require time intervals between measurements to be identical (i.e., equidistant measurements) and an increased focus on context-specific phenomena has (2) lead to issues when data is only available under certain conditions, including event-based questions (i.e., structurally missing data).
+Smaller issues of non-equidistant data can be avoided with transformations \citep[e.g., dynamic time warping,][]{berndt1994} or newer modeling procedures \citep[e.g., continuous-time models][]{dehaan-rietdijk2017} but for many analyses, including many cluster approaches, non-equidistant measurements remain a prevalent issue.
+Structural missingness remains an even more strenuous challenge.
+Structural missingness occurs when data is missing because it logically cannot be collected.
+It would not be sensible to measure the quality of an event for participants that did not experience the event.
+This type of missingness particular becomes an issue when researchers are interested in how two concepts co-develop and one of the concepts has structural missingness under certain conditions.
+The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008}\footnote{This is the case because the most commonly used models require complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.} --- neither option suits a research question that includes a developmental relationship between the two variables.
 In short, adequate descriptive approaches should be able to deal with non-equidistant and structurally missing data in order to address modern ESM data.
 
-Finally, both the push for more multivariate conceptualizations and contextualized measurements have highlighted issues with how temporal developments are conceptualized.
-Firstly, researchers are increasingly interested in how mean levels of variables change over time (i.e., non-stationary), such trend data is often assumed to be non-linear, and with the increased focus on more diverse conceptualizations researchers are interested in relationships between variables that change on fundamentally different time scales.
-Many of the commonly used analyses for ESM data are stationary lagged regression models that assume stable means and variances over time (incl., vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses) or use basic trajectory models (e.g., mixed effects models, spline regression models, and latent growth curve modeling).
-And even fewer analysis techniques allow for substantially different time scales between variables.
-A modern descriptive and exploratory methodology should thus be able to flexibly capture and comparable non-stationary and non-linear trends, also across different time scales.
+Finally, new approaches to ESM data have highlighted two main issues with how temporal developments are conceptualized.
+Researchers are (1) increasingly interested in how mean-levels of variables change over time, often in a non-linear fashion (i.e., non-stationary), and (2) there is increased interest in relationships between variables that change on fundamentally different time scales.
+Many of the commonly used analyses for ESM data assume that means and variances are stable over time\footnote{These models include the popular vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses.
+Notable time-varying adaption include \citet{haslbeck2021a}.}.
+While several trend modeling methodologies exist (e.g., spline regression models, latent growth curve modeling), their availability is limited and they seldomly extend into the descriptive realm.
+Moreover, among the approaches that allow for means to change, even fewer allow for substantially different time scales between variables, which can be essential in understanding how daily experiences relate to slower psychological phenomena that develop over days, weeks, and months \citep[e.g., see][]{Ram2014}.
+A modern descriptive and exploratory methodology, including a clustering approach, should thus be able to flexibly capture and compare non-stationary and non-linear trends, also across different time scales.
 
@@
 Time-series clustering has been extremely common in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-But also for psychological data time-series clustering is not entirely novel \citep[e.g.,][]{ernst2021}.
+But also for psychological data, time-series clustering is not entirely novel \citep[e.g.,][]{ernst2021}.
 However, most clustering introduced for psychological data has faced the dimensionality-, missingness-, and time scale issues described above.
@@
 The participants are then clustered on basis of the extracted features.
-To showcase the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
+As an illustration of the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
 Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time series, but also linear and non-linear trends, inertia-related features such as auto-correlations, or the stability of the time series (e.g., mean absolute change).
@@
 % aims re-iterated
-To be written.
+We sought to introduce a flexible and transparent analysis that aids researchers in describing, exploring, and understanding psychological time series data.
+Such an analysis needs to address the demands of the growing variety of ESM research.
+In particular, modern describe- and explore analyses should be able to deal with high dimensionality, structural missingness, and diverse time-varying developments.
+We have argued that feature-based time series clustering is one option that meets this challenge.
 
-\subsection{Limitations}
-To be written.
+We have methodologically deconstructed the feature extraction, feature reduction and feature clustering steps and offered methodological as well as conceptual examples for each step.
+Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research practice as both the features as well as the analysis steps are well-established within the field and statistical packages are readily available.
+Time series features (such as means or linear trends) are not only easy to extract but also hold conceptual meaning for psychological data and can be chosen to address specific research questions (also see \tblref{tab:esmFeatures}).
 
-\subsection{Implications}
-To be written.
+To illustrate the practical utility of the approach, we applied the method to real-life empirical data with diverse conceptualizations, missingness, and nonlinear trends.
+We make available the full code, custom functions, and illustrations to aid the approachability of the method.
+In the interpretation step we also show how the clusters can be compared across the original feature set as well as other meaningful situational and person-specific differences (e.g., see \fgrref{fig:clusterFeatVar}).
+We also find that for our example of migration experiences, the method was useful in discerning adaptive from more stressful experiences and helped to contextualize the diverging experiences.
 
-\subsection{Conclusion}
-To be written.
+As with any statistical method, feature-based time series clustering is not without limitations. 
+In particular, feature-based clustering has both a usability as well as robustness limitation in its multiple steps.
+In terms of convenience, each of the three steps requires users to make an informed decision about the choice of method and algorithm.
+These additional steps of decision-making and transparency heighten the initial barrier to entry.
+We hope that our empirical illustration offers a relatively generalizable procedure that showcases the ease of use but clustering unfortunately does not offer a universal one-size-fits-all solution.
 
-\section{Notes}
+In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential uncomparable results across methods.
+While the variety and diversity of methods is helpful in finding options even for more complex types of data, different algorithms often offer different results.
+And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their ideosyncratic shortcomings and might, for example, be sensitive to outliers.
 
-\citep[e.g., see][]{Ram2014} [Processes Across Multiple Time-Scales]
-
-"The goal of clustering is to identify structure in an unlabeled data set by objectively organizing data into homogeneous groups where the within-group-object dissimilarity is minimized and the between-group-object dissimilarity is maximized." --- \citep[][p.1857]{liao2005}
-
-\subsection{Migration Parts}
-In the example of the migration experiences, researchers might for example be interested in how well-being develops in relations the quality of interactions with the majority group.
-Because rating of interaction quality are only possible when an interaction actually happened, one cannot impute interaction quality ratings for measurements that did not include an interaction.
+Notwithstanding the limitations, we believe that feature-based clustering offers an exciting new potential for psychological time series.
+Across a variety of scientific fields, feature-based time series clustering has been valued for its structuring and simplifying utilities.
+Within the growing field of psychological time series, modern ESM data is becoming increasingly more complex in terms of participants, variables, and time points.
+Feature-based time series clusters can aid in reducing these complexities to important and meaningful patterns, an increasingly important task for researchers and practitioners.
 
@@
   \caption*{Note: \\
-  same data}
+  Within the ``(B) Feature Focus'' subplot, the `n' and `Discrimination' comparison variables were not part of the original time series clustering.}
 \end{figure*}
@@
   \caption*{Note: \\
-  same data}
+  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
 \end{figure*}

### 2022-12-03 methods-and-results.Rmd
:bug: tab ref

@@
 We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:descrWide}.
+Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:descrLong}.
 

### 2022-12-03 methods-and-results.tex
:bug: tab ref

@@
 but basic item information, descriptives, and correlations are available
-in \tblref{tab:descrWide}.
+in \tblref{tab:descrLong}.
 

### 2022-12-06 manuscript.tex
:sparkles: intro restr

@@
 Such procedures could aid researchers and practitioners in describing and exploring patterns across participants and could promote the creation of more embedded theories and interventions.
-In this manuscript, we propose feature-based time series clustering as a flexible, transparent, and well-grounded clustering approach that addresses the growing dimensionality, missingness, and time scale challenges.
+In this manuscript, we propose feature-based time series clustering as a flexible, transparent, and well-grounded clustering approach that addresses the growing challenges of dimensionality, missingness, and time scales.
 We introduce the individual feature extraction, -reduction, and -clustering steps and illustrate their utility with an empirical example.
@@
 
+Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g., experience sampling method (ESM),][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}.
+These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, Monden2015}.
+Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across a number of distinct variables \citep[i.e., understand the structure of between-subject differences of within-participant dynamics; e.g.,][]{ernst2021}.
+One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of grouping participants based on the characteristics of their time series dynamics \citep{ariens2020}.
+
+Clustering holds a number of important prospects as it seeks to organize data into groups that maximizes differences between groups while minimizing differences within the groups.
+For time series clustering this means that a participant's dynamics of some variable(s) have maximum similarity with the dynamics of other individuals in the same group, and minimum similarity with individuals in other groups \citep[e.g.,][]{liao2005}.
+As such, time series clustering has a number of methodological and empirical use cases.
+Methodologically, comparing person-specific dynamics between participants allows researchers to investigate whether processes on the individual level generalize to the interindividual level \citep[i.e., ergodicity; e.g.,][]{adolf2019, molenaar2004}.
+Similarly, clusters might allow researchers to identify cases where the meaning of measurements might have changed over time \citep[i.e., measurement invariance; e.g., ][]{adolf2014}.
+For a more detailed of these benefits, see \citet{ernst2021}.
+
+benefits of exploratory separation
+\begin{itemize}[nosep]
+    \item with extremely large data sets models can be estimated on a group level rather than for every individual (REFERENCE for TS clustering as preparation step)
+    \item describe and explore data with reducing complexity: by extracting structure between participants
+    \item phenomena detection \citep[i.e., extract abstract functional principles from phenomena][]{eronen2021a}. Important for building theories and practical interventions.
+    \item check of or avoid ergodicity \citep{adolf2019, molenaar2004} and measurement invariance \citep{adolf2014}
+\end{itemize}
+
+Time-series clustering has been extremely common in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+But also for psychological data, time-series clustering has recently become more common \citep[e.g.,][]{ernst2021}.
+
+Thus far the most common clustering approach for ESM data has been to group individuals based on their model parameters \citep[e.g., intercepts and slopes from vector autoregression models (VAR),][]{ariens2020}.
+While these model parameters are familiar to researchers in the field, the model-based approach has a number of important shortcomings.
+Computationally, the number of parameters compounds with increases in variables and lag orders, which tends to result in efficiency and accuracy reduction for most clustering algorithms and is sometimes referred to as the dimensionality curse \citep{altman2018}\footnote{As an illustration of this scalability issue, consider a VAR model of 8 variables that includes the first three lags (i.e., VAR(3)).
+For each of the 8 variables we would have 25 parameters --- one intercept and 24 slopes (8 variables * 3 lags each).
+That compounds to a total of 200 parameters per person (8 variables * 25 parameters).}.
+But more importantly, using the model parameters means that the clustering is restricted by the assumptions and limitations of the statistical model.
+To take the VAR model as an example, the model assumes that the time intervals between measurements are identical (i.e., equidistant measurements), it does not allow for missing values, and it assumes that the means as well variances of the variables do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
+
+These shortcomings however stand in sharp contrast with common research interests in the field and the types of data researchers commonly collect. 
+Detail on multidimensional nonequidistant structurally missing nonlinear trends.
+
+we thus need a way of dealing with these types of data. 
+While some of these issues can by avoided by transforming the data \citep[e.g., detrending;][]{salles2019} or switching to different time series models \citep[e.g., time-varying VAR;][]{bringmann2018b}, a more flexible alternative would be to expand the clustering inputs to other time series characteristics of interest beyond model parameters.
+
+Importantly, model intercepts and slopes are not the only way in which researchers conceptualize and describe time series.
+
+we propose, ...
+
+-----
+
+Additionally, ``scalability problems, and can lead to performance reductions when the clusters are close to each other''
+
+-----
+\newpage
 % Relevance Pragraph and issue
@@
 Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective.
-Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions.
+Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \hl{(Corduas & Piccollo, 2008)}.
 Yet, in cases where mis-specifications are costly, a method with fewer assumptions might be more advisable.
@@
 Feature-based time series clusters can aid in reducing these complexities to important and meaningful patterns, an increasingly important task for researchers and practitioners.
+
+\section{Notes}
+
+TS feature clustering steps: \citep{rasanen2009, wang2006}
+
+model based clustering: \citep{bulteel2016, stefanovic2022}
+
+sort of feature based clustering in psych \citep{heylen2016}
+
+clustering in reduced space \citep{timmerman2013}
 

### 2022-12-18 manuscript.tex
:bug:

--- 2022-12-03 manuscript.tex
@@
 % Title and header
-\title{Describe and Explore: Using feature-based time series clustering to understand modern intensive longitudinal data}
+\title{Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data}
 \shorttitle{feature-based time series clustering}
@@
 Psychological time series data has not only become more common but also more diverse and complex.
-Researchers and practitioners are more and more interested in non-stationary developmental patterns of multivariate concepts, across different time scales, while contextualized measurements bring about issues of non-imputable missingness.
+Researchers and practitioners are increasingly interested in non-stationary developmental patterns of multivariate concepts, across different time scales, while contextualized measurements bring about issues of non-imputable missingness.
 At the same time, clustering analyses offer the potential of reducing these complexities to important and meaningful patterns.
-Such procedures could aid researchers and practitioners in describing and exploring patterns across participants and could promote the creation of more embedded theories and interventions.
-In this manuscript, we propose feature-based time series clustering as a flexible, transparent, and well-grounded clustering approach that addresses the growing dimensionality, missingness, and time scale challenges.
+Such procedures aid researchers and practitioners in describing and exploring patterns across participants and can promote the creation of more embedded theories and interventions.
+In this manuscript, we propose feature-based time series clustering as a flexible, transparent, and well-grounded clustering approach that addresses the growing challenges of dimensionality, missingness, and time scales.
 We introduce the individual feature extraction, -reduction, and -clustering steps and illustrate their utility with an empirical example.
@@
 
-\begin{mdframed}[style=mdfbox]
-\noindent\center\textit{Structure}:
-\begin{itemize}[nosep]
-    \item introduction
-    \begin{itemize}[nosep]
-        \item\ [Relevance]: increase in number and variety of ESM data
-        \item\ [Problem]: no descriptive / understand structure analyses
-        \item\ [Proposal]: feature based clusterting = flexible +  transparent
-    \end{itemize}
-    \item Modern ESM data: new type of data and issues for analysis:
-    \begin{itemize}[nosep]
-        \item incomplete data (complete data models vs. structural missingness)
-        \item non-equidistant
-        \item non-stationary
-        \item changes/trends on different time scales
-        \item interest in multivariate developments
-        \item dimensionality curse
-    \end{itemize}
-    \item Time series clustering
-    \begin{itemize}[nosep]
-        \item shape-based
-        \item model-based
-        \item \textbf{feature-based}
-    \end{itemize}
-    \item Feature-Based Clustering
-    \begin{itemize}[nosep]
-        \item input variables
-        \item feature extraction
-        \item feature reduction
-        \item feature clustering
-    \end{itemize}
-    \item Empirical Illustration
-    \begin{itemize}[nosep]
-        \item Data
-        \item Analysis and Results
-    \end{itemize}
-    \item Discussion (summary, limitation, implications, conclusion, etc.)
-\end{itemize}
-\vspace{1em}
-\end{mdframed}
+% \begin{mdframed}[style=mdfbox]
+% \noindent\center\textit{Structure}:
+% \begin{itemize}[nosep]
+%     \item introduction
+%     \begin{itemize}[nosep]
+%         \item\ [Relevance]: increase in number and variety of ESM data
+%         \item\ [Problem]: no descriptive / understand structure analyses
+%         \item\ [Proposal]: feature based clusterting = flexible +  transparent
+%     \end{itemize}
+%     \item Modern ESM data: new type of data and issues for analysis:
+%     \begin{itemize}[nosep]
+%         \item incomplete data (complete data models vs. structural missingness)
+%         \item non-equidistant
+%         \item non-stationary
+%         \item changes/trends on different time scales
+%         \item interest in multivariate developments
+%         \item dimensionality curse
+%     \end{itemize}
+%     \item Time series clustering
+%     \begin{itemize}[nosep]
+%         \item shape-based
+%         \item model-based
+%         \item \textbf{feature-based}
+%     \end{itemize}
+%     \item Feature-Based Clustering
+%     \begin{itemize}[nosep]
+%         \item input variables
+%         \item feature extraction
+%         \item feature reduction
+%         \item feature clustering
+%     \end{itemize}
+%     \item Empirical Illustration
+%     \begin{itemize}[nosep]
+%         \item Data
+%         \item Analysis and Results
+%     \end{itemize}
+%     \item Discussion (summary, limitation, implications, conclusion, etc.)
+% \end{itemize}
+% \vspace{1em}
+% \end{mdframed}
 
-% Relevance Pragraph and issue
-Recent years have seen a striking increase in the number and variety of research studies using experience sampling data \citep[][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}. 
-With the increased availability of technologies to easily collect large amounts of experience sampling data using mobile devices \citep[e.g.,][]{Keil2020} and web-based applications \citep[e.g.,][]{Arslan2020}, the experience-sampling method (ESM) has been applied to a host of new types of psychological data.
-This move toward a broader application of the experience-sampling method, however, stands in stark contrast with recent developments within the psychometric literature on ESM analyses, which have become more specialized and model-focused.
-While generally speaking analytical methods for ESM data have become more readily available \citep[e.g.,][]{ODonnell2021}, flexible descriptive and exploratory analyses have largely been neglected and remain crucially understudied within the experience sampling literature.
+Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g., experience sampling method (ESM),][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}.
+These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, Monden2015, wardenaar2013}.
+Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across a number of distinct variables \citep[i.e., understand the structure of between-subject differences of within-participant dynamics; e.g.,][]{ernst2021}.
+One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of grouping participants based on the characteristics of their time series dynamics \citep{ariens2020}.
+In practice, this means that participants are inductively grouped in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{liao2005}.
 
-% Problem Illustration Paragraph
-Importantly, the understudied `describe and explore' analyses are not only important for contextualizing inferential model tests, but are powerful methods for theoretical and applied users in their own regard.
-Extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables.
-Data-driven models not only describe and summarize this complex data but can help us understand the data by uncovering patterns that might not be detectable otherwise.
-As an example, understanding how people differ in their trajectories can be crucial in understanding adaptive and maladaptive patterns.
-By identifying and contextualizing distinct trajectories, we might for example determine which situation- or person-specific differences discern positive from negative developments.
-Outcomes from such an approach might inform both practical early-warning signals as well as academic theory-building efforts.
-There is, thus, a clear need to assess the utility of data-driven analysis procedures that are flexible enough to address the growing variety of intensive longitudinal psychological data.
+Time series clustering has a number of conceptual use cases.
+Specifically, time series clustering aids in reducing the time, variable, and person complexity by extracting structure between participants.
+These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
+Such meaningful psychological principles and phenomena have often been recognized as being imperative for building robust theories and real-world interventions \citep[e.g.,][]{borsboom2021, eronen2020}.
 
-% Aim / Solution / Proposal Paragraph:
-In this article, we propose to look at recent developments within the wider time-series clustering literature to identify an exploratory data analysis that is flexible enough to apply to the growing variety of ESM data.
-In particular, we propose that feature-based clustering is ideally suited because it offers a flexible variety of validated and intuitive components while also allowing for adequate control by the researcher.
-To showcase the utility of this technique, we apply feature-based clustering to a set of three ESM studies with complex conceptualizations and missingness patterns.
+%Time-series clustering has been extremely common in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+But also for psychological data, time-series clustering has recently become more common \citep[e.g.,][]{ernst2021}.
 
-\section{Modern ESM Data}
-Two important developments within the psychological sciences have been a shift towards more comprehensive assessments of psychological phenomena and a growing contextualization of intensive longitudinal real-life data.
-Multivariate experience-sampling approaches allow researchers to test theories more exhaustively and help build a more embedded understanding of human experience processes.
-However, such data also deviates from past intensive longitudinal data sets that have focused on psychological states that are more homogeneous in their developmental characteristics.
-It, thus, makes sense to briefly consider which challenges modern ESM data introduce and what qualities a present-day clustering analysis should possess.
-We particularly highlight issues of dimensionality, missingness, and time scales within contemporary ESM data.
+Thus far, one of the most common clustering approaches for ESM data has been to group individuals based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020}.
+While these model parameters are familiar to researchers in the field, the model-based approach has a number of important shortcomings.
+Computationally, the number of parameters compounds with increases in variables and lag orders, which tends to result in efficiency and accuracy reduction for most clustering algorithms and is sometimes referred to as the dimensionality curse \citep{altman2018}\footnote{As an illustration of this scalability issue, consider a VAR model of 8 variables that includes the first three lags (i.e., VAR(3)).
+For each of the 8 variables, we would have 25 parameters --- one intercept and 24 slopes (8 variables * 3 lags each).
+That compounds to a total of 200 parameters per person (8 variables * 25 parameters).}.
+But more importantly, using model parameters means that the clustering is restricted by the assumptions and limitations of the statistical model.
+To take the common VAR model as an example, the model assumes that the time intervals between measurements are identical (i.e., equidistant measurements), it does not allow for missing values, and it assumes that the means as well as variances of the variables do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
 
-Concerning dimensionality issues, especially more abstract psychological experiences often need a wider variety of measurements to be captured adequately.
-Today, few clinical conditions are captured with a single symptom measure, emotions are rarely assessed in isolation, and socio-cultural experiences are now widely considered to be multimodal.
-This also means that modern analysis techniques increasingly need be able to accommodate an increased focus on multivariate developments.
-The recent rise in psychological network models has highlighted this need for more multivariate outcome measurements \citep[e.g.,][]{decancq2013}.
-Beyond the capability to assess multivariate developments, an increased number of variables also heightens the computational load for model estimations and, especially when the number of variables and measurements compounds, models may not converge --- an issue sometimes referred to as the dimensionality curse.
-A modern descriptive and exploratory analysis technique should consequently be able to summarize and structure multivariate phenomena without running into computational load issues.
+These shortcomings, however, stand in sharp contrast with the types of data researchers commonly collect and common research interests in the field.
+Psychological researchers might, for example, collect data based on the occurrence of natural events, which tends to result in non-equidistant measurements \citep[e.g.,][]{myin-germeys2018, hamaker2017} and context-specific missingness that cannot be imputed, for instance, interaction quality perceptions \citep[e.g.,][]{kivela2022, lavori2008}.
+At the same time, researchers are often specifically interested in non-stationary developments, looking, for example, at how symptom levels and -variations change over time --- often abruptly and non-linearly \citep[][]{bringmann2018b, helmich2020a}.
+We offer a more extensive discussion of the current challenges that clustering approaches should address in \appref{app:ChallengesAppendix}.
+In short, the time series clustering repertoire should be extended to approaches that can more flexibly allow for diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
 
-The push for more contextualized data also presents two key issues with missing data.
-More event-based assessments tend to (1) encounter issues when models require time intervals between measurements to be identical (i.e., equidistant measurements) and an increased focus on context-specific phenomena has (2) lead to issues when data is only available under certain conditions, including event-based questions (i.e., structurally missing data).
-Smaller issues of non-equidistant data can be avoided with transformations \citep[e.g., dynamic time warping,][]{berndt1994} or newer modeling procedures \citep[e.g., continuous-time models][]{dehaan-rietdijk2017} but for many analyses, including many cluster approaches, non-equidistant measurements remain a prevalent issue.
-Structural missingness remains an even more strenuous challenge.
-Structural missingness occurs when data is missing because it logically cannot be collected.
-It would not be sensible to measure the quality of an event for participants that did not experience the event.
-This type of missingness particular becomes an issue when researchers are interested in how two concepts co-develop and one of the concepts has structural missingness under certain conditions.
-The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008}\footnote{This is the case because the most commonly used models require complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.} --- neither option suits a research question that includes a developmental relationship between the two variables.
-In short, adequate descriptive approaches should be able to deal with non-equidistant and structurally missing data in order to address modern ESM data.
+Importantly, model-based clustering is not the only clustering approach, and model intercepts and slopes are not the only way in which researchers conceptualize and describe time series.
+A flexible extension of model-based clustering is feature-based clustering \citep[e.g.,][]{Aghabozorgi2015}\footnote{It should be noted that a third option is shape-based clustering, which uses the raw time series and seeks to match different time series by non-linearly stretching and contracting the time axes.
+Shape-based clustering might be less relevant to many psychological time series because the forced alignment and scaling can, for example, lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
+Such scaling issues are less of a concern for highly regular and comparable time series such as EEG \citep[][]{huang1985}.}.
+Whereas model-based clustering is specifically limited to model parameters as time series characteristics, feature-based clustering extends the time series characteristics to virtually any numerical summary of a time series \citep[e.g.,][]{liao2005}.
+As an illustration of the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
+Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time series, but also linear and non-linear trends, inertia-related features such as auto-correlations, or the stability of the time series \citep[e.g., mean absolute change; e.g.,][]{barandas2020}.
+Given the flexibility of the feature-based approach, users are able to apply the analysis across multivariate data, feature extraction mitigates issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
 
-Finally, new approaches to ESM data have highlighted two main issues with how temporal developments are conceptualized.
-Researchers are (1) increasingly interested in how mean-levels of variables change over time, often in a non-linear fashion (i.e., non-stationary), and (2) there is increased interest in relationships between variables that change on fundamentally different time scales.
-Many of the commonly used analyses for ESM data assume that means and variances are stable over time\footnote{These models include the popular vector autoregressive models, dynamic structural equation models, autoregressive integrated moving average models, and cross-lagged panel analyses.
-Notable time-varying adaption include \citet{haslbeck2021a}.}.
-While several trend modeling methodologies exist (e.g., spline regression models, latent growth curve modeling), their availability is limited and they seldomly extend into the descriptive realm.
-Moreover, among the approaches that allow for means to change, even fewer allow for substantially different time scales between variables, which can be essential in understanding how daily experiences relate to slower psychological phenomena that develop over days, weeks, and months \citep[e.g., see][]{Ram2014}.
-A modern descriptive and exploratory methodology, including a clustering approach, should thus be able to flexibly capture and compare non-stationary and non-linear trends, also across different time scales.
-
-\section{Time Series Clustering}
-One promising approach to describing and exploring time-series data has been the idea of structuring data using clustering analyses.
-Clustering seeks to organize data into groups that are maximally different from one another while minimizing differences within the groups.
-Time-series clustering has been extremely common in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-But also for psychological data, time-series clustering is not entirely novel \citep[e.g.,][]{ernst2021}.
-However, most clustering introduced for psychological data has faced the dimensionality-, missingness-, and time scale issues described above.
-
-Many of the common time-series clustering algorithms used today have been developed within the machine learning literature and have commonly been separated into three different approaches: (1) shape-based, (2) model-based, and (3) feature-based clustering \citep{hautamaki2008, liao2005}.
-Shape-based clustering analyses use the raw time series and seek to match time series by non-linearly stretching and contracting the time axes.
-The other two approaches do not use the raw time series directly but convert the data first.
-The model-based clustering has been the most commonly used method within the psychological literature and clusters the parameters of a parametric model fitted for each participant \citep[e.g.,][]{ernst2021}.
-The flexibility of such clustering is restricted to the assumptions of the model used, so that stationary models, for example, do not capture non-stationary data well.
-Finally, feature-based clustering analyses are the most flexible, as they allow users to summarize the time series data with a wide variety of features that capture the data characteristics of interest.
-The participants are then clustered on basis of the extracted features.
-As an illustration of the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
-Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time series, but also linear and non-linear trends, inertia-related features such as auto-correlations, or the stability of the time series (e.g., mean absolute change).
-Given the flexibility of the feature-based approach, users are able to apply the analysis across multivariate data, feature extraction mitigates issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
+\section{Features of Psychological Time Series}
 
@@
 
-The feature-based clustering can be structured into four main steps.
+------------------------
+
+NOTES:
+
+model based clustering: \citep{bulteel2016, stefanovic2022}
+
+sort of feature-based clustering in psych \citep{heylen2016}
+
+clustering in reduced space \citep{timmerman2013}
+
+------------------------
+
+The feature-based clustering can be structured into four main steps \citep{rasanen2009, wang2006}.
 (1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, (3) an optional feature reduction step if there are too many data points for the clustering algorithms, and (4) the actual clustering of the time series features.
@@
 While the final selection of features should always be driven by the research questions and field-specific conventions, we can build a practical toolbox of meaningful time series features for psychological data.
-In particular, we propose to focus on six features based on common psychological research questions and recent works on affect dynamics \citep[e.g.,][]{dejonckheere2019, kuppens2017}.
+In particular, we propose to focus on six features based on common psychological research questions and recent works on affect dynamics \citep[e.g.,][]{dejonckheere2019, kuppens2017, adya2001}.
 An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
@@
 Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective.
-Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions.
+Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \hl{(Corduas \& Piccollo, 2008)}.
 Yet, in cases where mis-specifications are costly, a method with fewer assumptions might be more advisable.
@@
 As such, the k-means offers a good starting point for many psychological researchers and the method should be generalizable across a relatively wide variety of projects.
+
+\subsection{Cluster Performance Evaluation}
+
+PARAGRAPH ON PERFORMANCE METRICS (incl., cross-validation?)
+
+PARAGRAPH ON REPORTING \citep[e.g.,][]{vandeschoot2017}
 
@@
 
-\section{Placeholder Title}
-\label{app:AppendixTitle}
-To be written, if necessary.
+\section{ESM Data Challenges}
+\label{app:ChallengesAppendix}
+We will briefly consider which challenges modern ESM data introduce and what qualities are called for in an extension of the clustering repertoire.
+We particularly highlight issues of dimensionality, non-equidistant or missing measurements, an interest in non-stationary trends, as well as inconsistent/diverse time scales.
+
+Concerning dimensionality issues, especially more abstract psychological experiences often need a wider variety of measurements to be captured adequately.
+Today, few clinical conditions are captured with a single symptom measure \citep[e.g.,][]{cramer2016}, emotions are rarely assessed in isolation \citep[e.g.,][]{reitsema2022}, and socio-cultural experiences are now widely considered to be multimodal \citep[e.g.,][]{Kreienkamp2022d}.
+This also means that modern analysis techniques increasingly need be able to accommodate an increased focus on multivariate developments.
+At the same time, however, an increase in the number of considered variables tends to come at the expense of computational load for model estimations and clustering models may not converge \citep[the aforementioned dimensionality curse;][]{altman2018}.
+A modern time series clustering technique should consequently be able to summarize and structure multivariate phenomena without running into computational load issues.
+
+Another common type of data that are measurement regiments that collect data in irregular time intervals (i.e., non-equidistant measurements).
+Common are, for example, procedures where participants are asked to respond at random times throughout the day (i.e., signal-contingent) or following specific natural events of interest \citep[i.e., event-contingent; see][]{shiffman2008, myin-germeys2018}.
+Under such conditions data tends to violate the equidistance assumption that is expected by many time series models \citep[][]{hamaker2017}.
+Smaller issues of non-equidistant data can be avoided with transformations \citep[e.g., dynamic time warping,][]{berndt1994} or newer modeling procedures \citep[e.g., continuous-time models;][]{dehaan-rietdijk2017} but for many analyses, including some cluster approaches, non-equidistant measurements remain a prevalent issue.
+
+Structural missingness remains an even more strenuous challenge.
+Structural missingness occurs when data is missing because it logically cannot be collected \citep[as opposed to probabilistically missing data;][]{little2020, mclean2017}.
+Often, however, researchers might want to include variables in their models that are not be available under all conditions.
+Follow-up and event-contingent questions are a common example in ESM studies.
+Researchers, for example, ask about the frequency, intensity, or duration of symptoms --- but only if a symptom was present \citep{kivela2022}.
+Such approaches become specifically critical in cases of sensitive questions such as questions about suicidal ideation or other potentially trauma inducing questions \citep[e.g.,][]{glenn2022}.
+The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008}\footnote{This is the case because the most commonly used models require complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.} --- neither option suits a research question that wishes to include variables with common structural missingness, such as event-specific or follow-up questions.
+In short, new clustering approaches should be able to deal with structurally missing data in order to address modern ESM data.
+
+When it comes to studying developmental trajectories, psychological researchers are often also interested in nonstationary processes because they are more representative of the complex, dynamic behavior of the human mind.
+In psychology, nonstationary processes are typically used to study phenomena such as cognitive development \citep[][]{quartz1997}, decision-making \citep[][]{ratcliff2016}, and emotion dynamics \citep[][]{bringmann2018b}.
+These processes are often characterized by changes in the underlying statistical properties of the data over time, such as changes in the mean or variance \citep[][]{molenaar2009}.
+Especially when considering changes in mean levels, researchers are often interested in nonlinear changes because they describe human functioning better.
+For example, in decision making people might switch between choices \citep[][]{ratcliff2016}, or patients reducing medication might experience mood swings \citep[][]{helmich2020a}.
+Similarly, psychologists are often also interested in how variances change over time.
+This is especially the case because several changes in an individual's variance have been found to be indicative of critical changes, including depression relapses and symptom shifts more generally \citep[e.g.,][]{schreuder2020, wichers2020}.
+There is, thus, also a need for time series clustering algorithms that capture nonstationary processes, including nonlinear trends.
+
+Psychological time series often exhibit complex patterns and relationships that can change over different time scales.
+For example, a time series of daily mood ratings may show a weekly pattern, with higher ratings on the weekends and lower ratings during the week.
+At the same time, the series may also exhibit a longer-term trend, with overall mood levels increasing or decreasing over the course of several months or years \citep[e.g.,][]{Ram2014}.
+\sout{These different time scales can be studied separately or in combination, using different statistical techniques and modeling approaches \citep[][]{bertenthal2007, jeronimus2019a}.} Different time scales can become an even more difficult issue when different variables in a model develop on different time scales \citep{bringmann2022b}.
+Different time scales are thus also a concern clustering approaches should be able to address.
 

### 2022-12-19 manuscript.tex
Updates from Overleaf

@@
 colorlinks   = true, %Colours links instead of ugly boxes
-urlcolor     = blue, %Colour for external hyperlinks
-linkcolor    = blue, %Colour of internal links
-citecolor   = cyan, %Colour of citations
+urlcolor     = blue, %blue,Colour for external hyperlinks
+linkcolor    = black, %blue, Colour of internal links
+citecolor   = black, % cyan, Colour of citations
 pdfstartview={FitV},
@@
 
-\section{Features of Psychological Time Series}
+\section{Feature-based Time Series Clustering}
 
-\section{Feature-based Time Series Clustering}
+\subsection{Features of Psychological Time Series}
+
+\paragraph{Central Tendency}
+
+\paragraph{Variability}
+
+\paragraph{(In)stability}
+
+\paragraph{Inertia and Periodicity}
+
+\paragraph{Linear Trend}
+
+\paragraph{Nonlinearity}
+
+\paragraph{Analysis Steps}
 
@@
 
-\subsection{Input variables}
+\paragraph{Input variables}
 Time series clustering starts with the selection and preparation of the variables of interest.
@@
 
-\subsection{Feature extraction}
+\paragraph{Feature extraction}
 The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
@@
 
-\subsection{Feature reduction}
+\paragraph{Feature reduction}
 Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points usually remains too large for most clustering algorithms.
@@
 
-\subsection{Feature clustering}
+\paragraph{Feature clustering}
 For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible.
@@
 
-\subsection{Cluster Performance Evaluation}
+\paragraph{Cluster Performance Evaluation}
 
 PARAGRAPH ON PERFORMANCE METRICS (incl., cross-validation?)
+
+incl., finding cluster prototypes:
+1. The medoid sequence of the set. 2. The average sequence of the set. 3. The local search prototype
+
+External evaluation measures: These measures require knowledge of the true class labels and are used to compare the cluster assignments to the true class labels.
+Examples include accuracy, F-measure, adjusted Rand index, mutual information, and normalized mutual information.
+
+Internal evaluation measures: These measures do not require knowledge of the true class labels and are used to evaluate the quality of the clusters based on the characteristics of the data itself.
+Examples include silhouette coefficient, Calinski-Harabasz index, Davies-Bouldin index, and the Elbow method.
 

### 2022-12-28 manuscript.tex

@@
 % define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
-\documentclass[man, 12pt, a4paper, mask]{apa7}
+\documentclass[man, 12pt, a4paper, mask, floatsintext]{apa7}
 
@@
 These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, Monden2015, wardenaar2013}.
-Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across a number of distinct variables \citep[i.e., understand the structure of between-subject differences of within-participant dynamics; e.g.,][]{ernst2021}.
-One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of grouping participants based on the characteristics of their time series dynamics \citep{ariens2020}.
-In practice, this means that participants are inductively grouped in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{liao2005}.
+Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across several variables \citep[i.e., understand the structure of between-subject differences of within-participant dynamics; e.g.,][]{ernst2021}.
+One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of grouping participants based on the similarity of their time series \citep{ariens2020}.
 
-Time series clustering has a number of conceptual use cases.
-Specifically, time series clustering aids in reducing the time, variable, and person complexity by extracting structure between participants.
+\begin{figure}[!hbtp] %hbtp
+  \caption{Scopus ESM Development}
+  \label{fig:ScopusEsm}
+  \centering\includegraphics[width=\textwidth]{figures/Scopus-ESM-Development.png}
+  \caption*{Note: \\
+  Search Terms: `( "ecological momentary assessment"  OR  "experience sampling"  OR  "intensive longitudinal")' in Title, Abstract, or Keyword.}
+\end{figure}
+
+The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{liao2005}.
+In practice, this means time series clustering critically depends on identifying similarities or dissimilarities in time series and then accurately distinguishing different groups (REF).
+
+1. phenomena and principle detection
+    a. differences between clusters
+    b. patterns within clusters (\~ early warning signals?)
+2.
+
+Time series clustering has a number of conceptual use cases with psychological data.
+Prime among them is the ability to reduce the time, variable, and person complexity by extracting structure between participants.
 These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
+These phenomena and principles can be (1) meaningful differences between clusters, as well (2) important patterns, trends, and relationships within clusters \citep[e.g.,][]{schrodt2000}.
+
+Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs (REFERENCE).
+
 Such meaningful psychological principles and phenomena have often been recognized as being imperative for building robust theories and real-world interventions \citep[e.g.,][]{borsboom2021, eronen2020}.
+Additionally, time series clustering has been used to inform the development of targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions (REFERENCE).
 
@@
 \subsection{Features of Psychological Time Series}
+There are hundreds of possible ts features
 
-\paragraph{Central Tendency}
+\citet{kuppens2017} originally proposed (1) variability, (2) co-variance or ICC, (3) inertia, and (4) cross-lagged correlations.
+These features were then extended and formalized by \citet{krone2018} into (1) within-person variance, (2) innovation, (3) inertia or autocorrelation, (4) cross-lagged regression parameters, (5) covariances as granularity, and (6) mean intensity.
+\citet{krone2018} were even able to use these features to cluster study participants.
 
-\paragraph{Variability}
+\citet{dejonckheere2019} discuss (1) mean, (2) variance, (3) relative variance, (4) instability, (5) inertia / autoregression, (6) interdependence / network density, (7) granularity or differentiation / ICC, (8) correlation, and (9) diversity / Gini coefficient.
 
-\paragraph{(In)stability}
+Outside the psychological literature, time series features have also included more summaries in terms of measurement distribution, categorical states, and, importantly, nonlinear developments.
 
-\paragraph{Inertia and Periodicity}
+\citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data.
 
-\paragraph{Linear Trend}
-
-\paragraph{Nonlinearity}
-
-\paragraph{Analysis Steps}
-
-------------------------
-
-NOTES:
-
-model based clustering: \citep{bulteel2016, stefanovic2022}
-
-sort of feature-based clustering in psych \citep{heylen2016}
-
-clustering in reduced space \citep{timmerman2013}
-
-------------------------
-
-The feature-based clustering can be structured into four main steps \citep{rasanen2009, wang2006}.
-(1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, (3) an optional feature reduction step if there are too many data points for the clustering algorithms, and (4) the actual clustering of the time series features.
-We provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
-
-\paragraph{Input variables}
-Time series clustering starts with the selection and preparation of the variables of interest.
-While the selection will necessarily be field- and concept-specific, there are a few conceptual and methodological issues that should be considered.
-Conceptually, the included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series.
-There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}, and many health developments are commonly captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}.
-At the same time, however, the added number of variables can become a methodological concern.
-Not only can redundant and irrelevant variables diminish the quality of the analyses, but with intensive longitudinal data the number of data points compounds across participants, measurement occasions, and variables so that additional variables can make many of the following steps substantially more difficult (also see \fgrref{fig:TSCFlowN}).
-
-Once the important variables have been selected, the data needs to be prepared for the analysis steps.
-Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
-Two important steps are making the time-frames and response scales comparable across participants.
-Extracting features that describe the development of a month for one participant but 90 days for another participant might not be comparable for some psychological phenomena.
-This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values.
-It is thus generally advised to reduce the data to a common and comparable time frame.
-
-\paragraph{Feature extraction}
-The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
-There are, however, hundreds of possible features to describe psychological time series \citep[e.g., tsfresh][]{christ2018}.
-One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series.
-However, such an approach is not always advisable for psychological time series.
-For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 180 time series features.
-More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
-This is particularly the case when features have a conceptual or psychological meaning.
-Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
-But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
-We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.
+\citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos.
+This selection was also used by \citep{fulcher2013}.
+\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
 
@@
 The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice (see \tblref{tab:esmFeatures} for a more complete overview).
+
+\paragraph{Central Tendency}
+
+\paragraph{Variability}
+
+\paragraph{(In)stability}
+
+\paragraph{Inertia and Periodicity} self-similarity, seasonality, periodicity, inertia
+
+\paragraph{Linear Trend}
+
+\paragraph{Nonlinearity}
+
+\subsection{Analysis Steps}
+
+------------------------
+
+NOTES:
+
+model based clustering: \citep{bulteel2016, stefanovic2022}
+
+parameter-based clustering: VAR \citep{ernst2021} splines \citep{axen2011}
+
+sort of feature-based clustering in psych \citep{heylen2016} as well as \citep{krone2018}
+
+clustering in reduced space \citep{timmerman2013}
+
+------------------------
+
+The feature-based clustering can be structured into four main steps \citep{rasanen2009, wang2006}.
+(1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, (3) an optional feature reduction step if there are too many data points for the clustering algorithms, and (4) the actual clustering of the time series features.
+We provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
+
+\begin{figure}[!ht] % \begin{figure*}[hbtp] <-- on its own page
+  \caption{Flowchart Feature-Based Time Series Clustering in Psychology}
+  \label{fig:TSCFlow}
+  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/TimeSeriesClusterFlowSelection.pdf}
+  \caption*{Note: \\
+  Choices selected for illustration in this manuscript are marked in bold.}
+\end{figure}
+
+\paragraph{Input variables}
+Time series clustering starts with the selection and preparation of the variables of interest.
+While the selection will necessarily be field- and concept-specific, there are a few conceptual and methodological issues that should be considered.
+Conceptually, the included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series.
+There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}, and many health developments are commonly captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}.
+At the same time, however, the added number of variables can become a methodological concern.
+Not only can redundant and irrelevant variables diminish the quality of the analyses, but with intensive longitudinal data the number of data points compounds across participants, measurement occasions, and variables so that additional variables can make many of the following steps substantially more difficult (also see \fgrref{fig:TSCFlowN}).
+
+\begin{figure}[!ht] %hbtp
+  \caption{Exemplary Flowchart of Data Points in Feature-Based Time Series Clustering}
+  \label{fig:TSCFlowN}
+  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustFlowN.pdf}
+  \caption*{Note: \\
+  The presented number of participants, variables, and measurement occasions are somewhat arbitrary but generally represent common sample sizes found within the literature.
+Also the number of extracted clusters is presented for illustrative purposes only.}
+\end{figure}
+
+Once the important variables have been selected, the data needs to be prepared for the analysis steps.
+Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
+Two important steps are making the time-frames and response scales comparable across participants.
+Extracting features that describe the development of a month for one participant but 90 days for another participant might not be comparable for some psychological phenomena.
+This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values.
+It is thus generally advised to reduce the data to a common and comparable time frame.
+
+\paragraph{Feature extraction}
+The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
+There are, however, hundreds of possible features to describe psychological time series \citep[e.g., tsfresh][]{christ2018}.
+One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series.
+However, such an approach is not always advisable for psychological time series.
+For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 180 time series features.
+More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
+This is particularly the case when features have a conceptual or psychological meaning.
+Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
+But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
+We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.
 
@@
 
-\begin{figure}[hbtp]
-  \caption{Scopus ESM Development}
-  \label{fig:ScopusEsm}
-  \centering\includegraphics[width=\textwidth]{figures/Scopus-ESM-Development.png}
-  \caption*{Note: \\
-  Search Terms: `( "ecological momentary assessment"  OR  "experience sampling"  OR  "intensive longitudinal")' in Title, Abstract, or Keyword.}
-\end{figure}
-
-\begin{figure*}[hbtp]
-  \caption{Flowchart Feature-Based Time Series Clustering in Psychology}
-  \label{fig:TSCFlow}
-  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/TimeSeriesClusterFlowSelection.pdf}
-  \caption*{Note: \\
-  Choices selected for illustration in this manuscript are marked in bold.}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-  \caption{Exemplary Flowchart of Data Points in Feature-Based Time Series Clustering}
-  \label{fig:TSCFlowN}
-  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustFlowN.pdf}
-  \caption*{Note: \\
-  The presented number of participants, variables, and measurement occasions are somewhat arbitrary but generally represent common sample sizes found within the literature.
-Also the number of extracted clusters is presented for illustrative purposes only.}
-\end{figure*}
-
-\begin{figure*}[hbtp]
+\begin{figure*}[!ht] %hbtp
   \caption{Cluster Group Comparisons based on Features and Variables}
@@
 
-\begin{figure*}[hbtp]
+\begin{figure*}[!ht] %hbtp
   \caption{Cluster Group Comparisons over time}

### 2022-12-28 methods-and-results.Rmd
:art: figure placement

@@
 
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons based on Features and Variables}
+  \label{fig:clusterFeatVar}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
+  \caption*{Note: \\
+  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+\end{figure}
+
 We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
@@
 Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
+
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons over time}
+  \label{fig:clusterTs}
+  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
+  \caption*{Note: \\
+  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
+\end{figure}
 

### 2022-12-28 methods-and-results.tex
:art: figure placement

@@
 
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons based on Features and Variables}
+  \label{fig:clusterFeatVar}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
+  \caption*{Note: \\
+  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+\end{figure}
+
 We can then combine the two focus approaches to assess the developments
@@
 differences in more detail.
+
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons over time}
+  \label{fig:clusterTs}
+  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
+  \caption*{Note: \\
+  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
+\end{figure}

### 2022-12-28 manuscript.tex
:art: tbl placement

@@
 
-\begin{figure*}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the ``(B) Feature Focus'' subplot, the `n' and `Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure*}
-
-\begin{figure*}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure*}
-
 \printbibliography
@@
 This also means that modern analysis techniques increasingly need be able to accommodate an increased focus on multivariate developments.
-At the same time, however, an increase in the number of considered variables tends to come at the expense of computational load for model estimations and clustering models may not converge \citep[the aforementioned dimensionality curse;][]{altman2018}.
+At the same time, however, an increase in the number of considered variables tends to come at the expense of computational load for model estimations, and clustering models may not converge \citep[the aforementioned dimensionality curse;][]{altman2018}.
 A modern time series clustering technique should consequently be able to summarize and structure multivariate phenomena without running into computational load issues.
@@
 Structural missingness occurs when data is missing because it logically cannot be collected \citep[as opposed to probabilistically missing data;][]{little2020, mclean2017}.
-Often, however, researchers might want to include variables in their models that are not be available under all conditions.
+Often, however, researchers might want to include variables in their models that are not available under all conditions.
 Follow-up and event-contingent questions are a common example in ESM studies.
 Researchers, for example, ask about the frequency, intensity, or duration of symptoms --- but only if a symptom was present \citep{kivela2022}.
-Such approaches become specifically critical in cases of sensitive questions such as questions about suicidal ideation or other potentially trauma inducing questions \citep[e.g.,][]{glenn2022}.
+Such approaches become specifically critical in cases of sensitive questions such as questions about suicidal ideation or other potentially trauma-inducing questions \citep[e.g.,][]{glenn2022}.
 The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008}\footnote{This is the case because the most commonly used models require complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.} --- neither option suits a research question that wishes to include variables with common structural missingness, such as event-specific or follow-up questions.

### 2022-12-28 methods-and-results.Rmd
:art: tbl placement

@@
   kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
+  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
+  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
   save_kable("Tables/descrLong.tex")

### 2022-12-29 manuscript.tex
:pencil: intro str

@@
 
-Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g., experience sampling method (ESM),][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}.
+Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
+While the terms come for different conceptual backgrounds, they share a focus on collecting data over an extended period of time capturing people's behaviors and experiences as they vary over time and in response to different situations and events.
+In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}.
 These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, Monden2015, wardenaar2013}.
@@
 
-The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{liao2005}.
-In practice, this means time series clustering critically depends on identifying similarities or dissimilarities in time series and then accurately distinguishing different groups (REF).
-
-1. phenomena and principle detection
-    a. differences between clusters
-    b. patterns within clusters (\~ early warning signals?)
-2.
-
 Time series clustering has a number of conceptual use cases with psychological data.
@@
 These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
-These phenomena and principles can be (1) meaningful differences between clusters, as well (2) important patterns, trends, and relationships within clusters \citep[e.g.,][]{schrodt2000}.
-
-Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs (REFERENCE).
-
-Such meaningful psychological principles and phenomena have often been recognized as being imperative for building robust theories and real-world interventions \citep[e.g.,][]{borsboom2021, eronen2020}.
-Additionally, time series clustering has been used to inform the development of targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions (REFERENCE).
+These phenomena and principles can be meaningful differences between participants in different clusters, as well important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
+Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}.
+Once meaningful and robust structures and patterns are detected, time series clustering can, thus, be used to inform the development of robust theories as well as targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions \citep[e.g.,][]{borsboom2021, eronen2020}.
 
@@
 But also for psychological data, time-series clustering has recently become more common \citep[e.g.,][]{ernst2021}.
+
+The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{liao2005}.
+In practice, this means time series clustering critically depends on identifying similarities or dissimilarities in time series and then accurately distinguishing different groups (REF).
 

### 2022-12-30 manuscript.tex
:pencil: fig intr
:pencil: intro str

@@
 Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
-While the terms come for different conceptual backgrounds, they share a focus on collecting data over an extended period of time capturing people's behaviors and experiences as they vary over time and in response to different situations and events.
+While the terms come for different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events.
 In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}.
 These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, Monden2015, wardenaar2013}.
-Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across several variables \citep[i.e., understand the structure of between-subject differences of within-participant dynamics; e.g.,][]{ernst2021}.
-One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of grouping participants based on the similarity of their time series \citep{ariens2020}.
+Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across several variables \citep[e.g.,][]{ernst2021}.
+One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of inductively grouping participants based on the similarity of their time series \citep{ariens2020}.
 
@@
 Time series clustering has a number of conceptual use cases with psychological data.
-Prime among them is the ability to reduce the time, variable, and person complexity by extracting structure between participants.
+Prime among them is the ability to reduce the time, variable, and person complexity by extracting participant-level structures.
 These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
-These phenomena and principles can be meaningful differences between participants in different clusters, as well important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
+These phenomena and principles can be meaningful differences that distinguish participants in different clusters, as well important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
 Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}.
@@
 
-The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{liao2005}.
-In practice, this means time series clustering critically depends on identifying similarities or dissimilarities in time series and then accurately distinguishing different groups (REF).
+The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{Aghabozorgi2015}.
+In practice, this means time series clustering critically depends on identifying similarities or dissimilarities in time series and then accurately distinguishing different groups \citep[e.g.,][]{Aghabozorgi2015}.
+Broadly speaking, time series (dis-)similarity can be approximated by either comparing the raw time series or by comparing some numerical representation of the time series (e.g., \citealp[]{liao2005}; see \fgrref{fig:tsClustTax} for an illustration).
+The (dis-)similarity measurements can then be clustered either on their own or as part of the representation estimation (e.g., \citealp[]{denteuling2021}; see `embedded' in \fgrref{fig:tsClustTax}).
+
+\begin{figure}[!hbtp] %hbtp
+  \caption{Time Series Clustering Taxonomy}
+  \label{fig:tsClustTax}
+  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustTax.pdf}
+  \caption*{Note: \\
+  The taxonomy only exemplifies some of the basic differences between a number of common time series clustering approaches.
+As such, the taxonomy and the notes are neither exhaustive nor complete in distinguishing different approaches.
+Additionally, terms and labels are used inconsistently across different types of literature and are chosen to avoid overlapping labels.}
+\end{figure}
 

### 2023-01-01 manuscript.tex
:sparkles: intro str bullets

@@
 
-% \begin{mdframed}[style=mdfbox]
-% \noindent\center\textit{Structure}:
-% \begin{itemize}[nosep]
-%     \item introduction
-%     \begin{itemize}[nosep]
-%         \item\ [Relevance]: increase in number and variety of ESM data
-%         \item\ [Problem]: no descriptive / understand structure analyses
-%         \item\ [Proposal]: feature based clusterting = flexible +  transparent
-%     \end{itemize}
-%     \item Modern ESM data: new type of data and issues for analysis:
-%     \begin{itemize}[nosep]
-%         \item incomplete data (complete data models vs. structural missingness)
-%         \item non-equidistant
-%         \item non-stationary
-%         \item changes/trends on different time scales
-%         \item interest in multivariate developments
-%         \item dimensionality curse
-%     \end{itemize}
-%     \item Time series clustering
-%     \begin{itemize}[nosep]
-%         \item shape-based
-%         \item model-based
-%         \item \textbf{feature-based}
-%     \end{itemize}
-%     \item Feature-Based Clustering
-%     \begin{itemize}[nosep]
-%         \item input variables
-%         \item feature extraction
-%         \item feature reduction
-%         \item feature clustering
-%     \end{itemize}
-%     \item Empirical Illustration
-%     \begin{itemize}[nosep]
-%         \item Data
-%         \item Analysis and Results
-%     \end{itemize}
-%     \item Discussion (summary, limitation, implications, conclusion, etc.)
-% \end{itemize}
-% \vspace{1em}
-% \end{mdframed}
+\begin{mdframed}[style=mdfbox]
+\noindent\center\textit{Structure}:
+\begin{itemize}[nosep]
+    \item introduction
+    \begin{itemize}[nosep]
+        \item increase in number and variety of ESM data + ts clustering to understand complex developmental differences.
+        \item why cluster time series? 
+        \begin{enumerate}[nosep]
+            \item differences between clusters
+            \item patterns within clusters
+            \item relation to out-of cluster variables
+            \item robust theories and interventions
+        \end{enumerate}
+        \item can't cluster all the raw data (b/c too much data and not directly comparable).
+        \item summarize the data and cluster based on the summaries.
+        \item ideally these summaries are:
+        \begin{enumerate}[nosep]
+            \item good at captuing the data
+            \item meaningful = help us understand how the clustering arrived at the output (i.e., explainable) + helps us make sense of the final cluster output (i.e., interpretable)
+        \end{enumerate}
+        \item currently model parameters as summaries.
+        \item limitations of parameter summaries: assumptions not always good at capturing data and model parameters not always interpretable
+        \item current solution: more complicated models (e.g., address violated assumptions or add different types of parameters)
+        \item proposed solution: features as flexible and meaningful approach of summarizing the time series
+        \item ---------------------------------------------------------------------------
+        \item limitations of model parameters: 
+        \begin{itemize}[nosep]
+            \item strong assumptions = not always a good (dis-)similiarity: 
+            \begin{enumerate}[nosep]
+                \item non-equidistant measurements
+                \item non-imputable missingness
+                \item non-stationary
+                \item non-linear developments
+            \end{enumerate}
+            \item difficult to interpret psychologically 
+            \begin{enumerate}[nosep]
+                \item limited to few ways of psychological effects (e.g., lagged + cross-lagged effects)
+            \end{enumerate}
+        \end{itemize} 
+        \item push for more complicated models aimed at addressing (1) similarity concerns (e.g., amend assumptions; e.g., see den Teuling) or (2) interpretatbility (e.g., more types of effects; e.g., see krone2018)
+        \item proposed alternative: go back to interpretable similarity measurements. General statistics and parameters instead of only parameters.
+    \end{itemize}
+    \item Feature-based Time Series Clustering
+    \begin{itemize}[nosep]
+        \item different types of similarity grouping: five types
+        \item why features: flexible similarity (b/c more variety), less assumptions, easy to calculate/extract, theory-based selection (as per RQ), psychological interpretability, ...
+        \item Features in Psychological Time Series
+        \begin{itemize}[nosep]
+            \item central tendency
+            \item variability
+            \item (in)stability
+            \item self-similarity
+            \item linear trend
+            \item nonlinearity
+        \end{itemize}
+        \item Analysis Steps
+        \begin{itemize}[nosep]
+            \item input variables
+            \item feature extraction
+            \item (feature reduction)
+            \item feature clustering
+            \item cluster evaluation
+        \end{itemize}
+    \end{itemize}
+    \item Empirical Illustration
+    \item Discussion (summary, limitation, implications, conclusion, etc.)
+\end{itemize}
+\vspace{1em}
+\end{mdframed}
 

### 2023-01-02 manuscript.tex
:pencil: intro str

@@
 Time series clustering has a number of conceptual use cases with psychological data.
-Prime among them is the ability to reduce the time, variable, and person complexity by extracting participant-level structures.
+Prime among them is the ability to reduce the time, variable, and person complexity by extracting and organizing participant-level structures.
 These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
@@
 But also for psychological data, time-series clustering has recently become more common \citep[e.g.,][]{ernst2021}.
+
+However, while clustering can be incredibly useful, arriving at these clusters critically depends on two core challenges.
+
+this means time series clustering critically depends on making time series comparable to 
+identify key (dis)similarities and then accurately distinguishing different groups \citep[e.g.,][]{Aghabozorgi2015}.
 

### 2023-01-03 manuscript.tex
:pencil: intro explainability

@@
         \end{enumerate}
-        \item can't cluster all the raw data (b/c too much data and not directly comparable).
-        \item summarize the data and cluster based on the summaries.
+        \item can't cluster all the raw data (b/c too much data and not directly comparable).Instead, summarize the data and cluster based on the summaries.
         \item ideally these summaries are:
@@
 
+%The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{Aghabozorgi2015}.
+
 However, while clustering can be incredibly useful, arriving at these clusters critically depends on two core challenges.
-
-this means time series clustering critically depends on making time series comparable to 
-identify key (dis)similarities and then accurately distinguishing different groups \citep[e.g.,][]{Aghabozorgi2015}.
-
-The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{Aghabozorgi2015}.
-In practice, this means time series clustering critically depends on identifying similarities or dissimilarities in time series and then accurately distinguishing different groups \citep[e.g.,][]{Aghabozorgi2015}.
-Broadly speaking, time series (dis-)similarity can be approximated by either comparing the raw time series or by comparing some numerical representation of the time series (e.g., \citealp[]{liao2005}; see \fgrref{fig:tsClustTax} for an illustration).
-The (dis-)similarity measurements can then be clustered either on their own or as part of the representation estimation (e.g., \citealp[]{denteuling2021}; see `embedded' in \fgrref{fig:tsClustTax}).
+First, time series need to be made comparable in order to identify key (dis)similarities and second, comparable (dis)similarities need to be accurately distinguishing into different groups \citep[e.g.,][]{Aghabozorgi2015}.
+In practice, most psychological time series cannot be compared based on the raw data itself.
+This is the case because in most cases the raw time series include too many time data points for most clustering methods to handle \citep[e.g.,][]{altman2018} and, more importantly, individual time points are oftentimes not directly comparable between participants in psychological data and would lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
+While such issues can be avoided with transformations for highly regular, controlled, and comparable time series such as EEG \citep[e.g.,][]{huang1985}, most ESM psychologists are usually not interested in directly comparing individual timepoints between participants but are interested in developmental patterns and relationships.
+As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{liao2005}; see \fgrref{fig:tsClustTax}).
 
@@
 \end{figure}
+
+The representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[e.g.,][]{vandermaaten2009}.
+Extracting accurate and meaningful representations of the time series can be essential for understanding what goes into the clustering algorithm (i.e., assists with explainability) and can be crucial in making sense of the final cluster output \citep[i.e., assists with interpretability; e.g.,][]{Kennedy2021}.
 

### 2023-01-04 manuscript.tex
:pencil: intro explainability
:pencil: f-b ts clust intr

@@
         \end{enumerate}
-        \item can't cluster all the raw data (b/c too much data and not directly comparable).Instead, summarize the data and cluster based on the summaries.
+        \item can't cluster all the raw data (b/c too much data and not directly comparable). Instead, summarize the data and cluster based on the summaries.
         \item ideally these summaries are:
@@
         \item proposed solution: features as flexible and meaningful approach of summarizing the time series
-        \item ---------------------------------------------------------------------------
-        \item limitations of model parameters: 
-        \begin{itemize}[nosep]
-            \item strong assumptions = not always a good (dis-)similiarity: 
-            \begin{enumerate}[nosep]
-                \item non-equidistant measurements
-                \item non-imputable missingness
-                \item non-stationary
-                \item non-linear developments
-            \end{enumerate}
-            \item difficult to interpret psychologically 
-            \begin{enumerate}[nosep]
-                \item limited to few ways of psychological effects (e.g., lagged + cross-lagged effects)
-            \end{enumerate}
-        \end{itemize} 
-        \item push for more complicated models aimed at addressing (1) similarity concerns (e.g., amend assumptions; e.g., see den Teuling) or (2) interpretatbility (e.g., more types of effects; e.g., see krone2018)
-        \item proposed alternative: go back to interpretable similarity measurements. General statistics and parameters instead of only parameters.
     \end{itemize}
@@
     \begin{itemize}[nosep]
-        \item different types of similarity grouping: five types
         \item why features: flexible similarity (b/c more variety), less assumptions, easy to calculate/extract, theory-based selection (as per RQ), psychological interpretability, ...
@@
 \end{mdframed}
+\newpage
 
@@
 In practice, most psychological time series cannot be compared based on the raw data itself.
-This is the case because in most cases the raw time series include too many time data points for most clustering methods to handle \citep[e.g.,][]{altman2018} and, more importantly, individual time points are oftentimes not directly comparable between participants in psychological data and would lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
-While such issues can be avoided with transformations for highly regular, controlled, and comparable time series such as EEG \citep[e.g.,][]{huang1985}, most ESM psychologists are usually not interested in directly comparing individual timepoints between participants but are interested in developmental patterns and relationships.
-As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{liao2005}; see \fgrref{fig:tsClustTax}).
+This is the case because in most cases the raw time series include too many data points --- sometimes referred to as the dimensionality curse \citep[e.g.,][]{altman2018} --- and, more importantly, individual time points are oftentimes not directly comparable between participants in psychological data and would lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
+While such issues can be avoided with transformations for highly regular, controlled, and comparable time series such as EEG data \citep[e.g.,][]{huang1985}, most ESM researchers are usually not interested in directly comparing individual timepoints between participants but are interested in developmental patterns and relationships.
 
@@
 
-The representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[e.g.,][]{vandermaaten2009}.
+As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{timmerman2013}; see \fgrref{fig:tsClustTax}).
+Ideally, the representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[e.g.,][]{vandermaaten2009}.
 Extracting accurate and meaningful representations of the time series can be essential for understanding what goes into the clustering algorithm (i.e., assists with explainability) and can be crucial in making sense of the final cluster output \citep[i.e., assists with interpretability; e.g.,][]{Kennedy2021}.
 
-Thus far, one of the most common clustering approaches for ESM data has been to group individuals based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020}.
-While these model parameters are familiar to researchers in the field, the model-based approach has a number of important shortcomings.
-Computationally, the number of parameters compounds with increases in variables and lag orders, which tends to result in efficiency and accuracy reduction for most clustering algorithms and is sometimes referred to as the dimensionality curse \citep{altman2018}\footnote{As an illustration of this scalability issue, consider a VAR model of 8 variables that includes the first three lags (i.e., VAR(3)).
+Thus far, one of the most common representation approaches for clustering of ESM data has been to extract person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
+While model parameters are familiar to researchers in the field, the model-based approach has a number of important shortcomings that relate to accuracy and interpretability.
+In terms of accuracy, the validity of model parameters is restricted by the assumptions of the statistical model.
+To take the common VAR model as an example, the model does not accurately capture data with unequal time intervals between measurements (i.e., nonequidistant measurements), it does not allow for missing values, and it does not allow for means or variances to change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
+As for the interpretability of the model parameters, purely data-driven model parameters, such as frequencies from a fourier analysis, are often difficult to interpret conceptually \citep[e.g.,][]{mayor2022}.
+But even more `meaningful' parameters from models used in theory-testing practices, such as VAR parameters, are commonly restricted to lagged partial correlations \citep[e.g.,][]{bringmann2018c}.
+Model parameters might, thus, not always accurately capture the original time series and are often restricted in their conceptual interpretation.
+
+%Computationally, the number of parameters compounds with increases in variables and lag orders, which tends to result in efficiency and accuracy reduction for most clustering algorithms and is sometimes referred to as the dimensionality curse \citep{altman2018}\footnote{As an illustration of this scalability issue, consider a VAR model of 8 variables that includes the first three lags (i.e., VAR(3)).
 For each of the 8 variables, we would have 25 parameters --- one intercept and 24 slopes (8 variables * 3 lags each).
 That compounds to a total of 200 parameters per person (8 variables * 25 parameters).}.
-But more importantly, using model parameters means that the clustering is restricted by the assumptions and limitations of the statistical model.
-To take the common VAR model as an example, the model assumes that the time intervals between measurements are identical (i.e., equidistant measurements), it does not allow for missing values, and it assumes that the means as well as variances of the variables do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
 
-These shortcomings, however, stand in sharp contrast with the types of data researchers commonly collect and common research interests in the field.
+These shortcomings, however, stand in sharp contrast with the types of data researchers commonly collect and do not align common research interests in the field.
 Psychological researchers might, for example, collect data based on the occurrence of natural events, which tends to result in non-equidistant measurements \citep[e.g.,][]{myin-germeys2018, hamaker2017} and context-specific missingness that cannot be imputed, for instance, interaction quality perceptions \citep[e.g.,][]{kivela2022, lavori2008}.
-At the same time, researchers are often specifically interested in non-stationary developments, looking, for example, at how symptom levels and -variations change over time --- often abruptly and non-linearly \citep[][]{bringmann2018b, helmich2020a}.
-We offer a more extensive discussion of the current challenges that clustering approaches should address in \appref{app:ChallengesAppendix}.
-In short, the time series clustering repertoire should be extended to approaches that can more flexibly allow for diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
+At the same time, researchers are often specifically interested in non-stationary developments, looking, for example, at how symptom levels and -variations change over time --- often abruptly and non-linearly \citep[e.g.,][; see \appref{app:ChallengesAppendix} for a more extensive discussion of the current challenges that clustering approaches should address]{bringmann2018b, helmich2020a}.
+There is, thus, a resolute need for more flexible and meaningful time series summaries that can capture diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
 
-Importantly, model-based clustering is not the only clustering approach, and model intercepts and slopes are not the only way in which researchers conceptualize and describe time series.
-A flexible extension of model-based clustering is feature-based clustering \citep[e.g.,][]{Aghabozorgi2015}\footnote{It should be noted that a third option is shape-based clustering, which uses the raw time series and seeks to match different time series by non-linearly stretching and contracting the time axes.
-Shape-based clustering might be less relevant to many psychological time series because the forced alignment and scaling can, for example, lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
-Such scaling issues are less of a concern for highly regular and comparable time series such as EEG \citep[][]{huang1985}.}.
-Whereas model-based clustering is specifically limited to model parameters as time series characteristics, feature-based clustering extends the time series characteristics to virtually any numerical summary of a time series \citep[e.g.,][]{liao2005}.
-As an illustration of the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
-Time series features can be chosen by the user based on their research question and can, for example, include the mean and standard deviation of a time series, but also linear and non-linear trends, inertia-related features such as auto-correlations, or the stability of the time series \citep[e.g., mean absolute change; e.g.,][]{barandas2020}.
-Given the flexibility of the feature-based approach, users are able to apply the analysis across multivariate data, feature extraction mitigates issues of non-equidistant and structurally missing data, and the features can capture and compare across non-linear trends at different time scales.
+% In short, the time series clustering repertoire should be extended to approaches that can more flexibly allow for diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
+
+Recent efforts to address the shortcomings of model parameters for time series clustering, have primarily proposed more complicated models aimed at addressing either accuracy or interpretatbility concerns.
+Accuracy concerns have particularly been addressed by building more advanced models that relax specific assumptions \citep[e.g.,][]{denteuling2021} or by switching to different types of models with less assumptions, including spline models \citep{axen2011}.
+To expand interpretatbility, researchers have proposed to model additional types of parameters, including variability or granularity parameters \citep[e.g., see][]{krone2018}.
+In this manuscript we instead, propose to look at the broader machine learning literature and to consider a wider range of time series features to describe psychological time series.
+In particular, we propose to use an expanded feature-based clustering approach, which allows users to summarize time series based on a research-specific combination of parameters and descriptive statistics.
+We argue that feature-based clustering offers a flexible variety, less strict assumptions, easy and intuitive extraction, theory-based selection, and meaningful psychological interpretability.
 
@@
 
+Feature-based time series clustering broadly speaking means that important features from the time series data are extracted (Ratanamahatana & Keogh, 2004).
+The extracted features can then be used as inputs to a clustering algorithm such as k-means or hierarchical clustering to group the time series into clusters (Xu et al., 2013).
+
+Feature-based time series clustering has been a valued tool in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+
+As an illustration of the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
+
 \subsection{Features of Psychological Time Series}
-There are hundreds of possible ts features
+
+in psychology:
 
@@
 
-The first two features are a person's \textit{central tendency} and \textit{variability}.
-Familiar statistics from probability theory, the two features sit at the heart of many fundamental psychological questions.
-``Are some people happier than others?'' (i.e., difference in central tendency), and "Are people settled in their attitudes towards migrants or do they fluctuate over time?'' (i.e., difference in variability).
-Mathematically, both features have parametric and robust options to choose from (e.g., parametric: mean and standard deviation; robust: median and median absolute deviation).
-
-The third and fourth features describe the structure of the variability within the time series.
-In particular, \textit{(in)stability} captures the average change between two consecutive measurements.
-Do changes tend to be slow and gradual or fast and abrupt?
-Mathematically, (in)stability measurements often look at the average change between two consecutive measurements.
-\textit{Inertia} describes how much a measurement carries over to future measurements.
-There are two main ways in which experiences tend to be connected to future experiences --- resistance to change and seasonality.
-Both forms of inertia relate to different types of research questions.
-``Do patients stay in a depressed mood for several measurements?'' (i.e., stability-inertia), and ``Do participants drink more alcohol on Fridays?'' (i.e., periodicity-inertia).
-Both forms of inertia are often measured using autocorrelations with either a lag-1 (stability-inertia) or the lag of the seasonality (e.g., seven days).
-Periodicity-inertia can additionally be captured using wave patterns (e.g., fourier or wavelet transformation).
-
-The final two features are linear and nonlinear trends.
-\textit{Linear trends} are a common research interest for longitudinal data.
-``Do patient symptoms improve?'' or ``Does worker productivity decline?'', are familiar types of research questions.
-Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients.
-\textit{Nonlinear trends} are important in two regards.
-Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
-Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}.
-The structure of nonlinear trends is often mathematically captured using polynomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
-
 This selection of the proposed six time-series features is in no way exhaustive or imperative.
@@
 
-\paragraph{Central Tendency}
+\paragraph{Central Tendency} Familiar statistics from probability theory, the two features sit at the heart of many fundamental psychological questions.
+``Are some people happier than others?'' (i.e., difference in central tendency).
+Mathematically, the central tendency has parametric and robust options to choose from (e.g., parametric: mean; robust: median).
 
-\paragraph{Variability}
+\paragraph{Variability} "Are people settled in their attitudes towards migrants or do they fluctuate over time?'' (i.e., difference in variability).
+parametric: standard deviation, robust: median absolute deviation.
 
-\paragraph{(In)stability}
+\paragraph{(In)stability} Stability captures the average change between two consecutive measurements.
+Do changes tend to be slow and gradual or fast and abrupt?
+Mathematically, (in)stability measurements often look at the average change between two consecutive measurements \citep[e.g., mean absolute change; e.g.,][]{barandas2020}.
 
-\paragraph{Inertia and Periodicity} self-similarity, seasonality, periodicity, inertia
+\paragraph{Self-similarity} Self-similarity describes how much a measurement carries over to future measurements.
+There are two main ways in which experiences tend to be connected to future experiences --- resistance to change and seasonality.
+Both forms of self-similarity relate to different types of research questions.
+``Do patients stay in a depressed mood for several measurements?'' (i.e., inertia), and ``Do participants drink more alcohol on Fridays?'' (i.e., periodicity).
+Both forms of self-similarity are often measured using autocorrelations with either a lag-1 (inertia) or the lag of the seasonality (e.g., seven days).
+Periodicity can additionally be captured using wave patterns \citep[e.g., fourier or wavelet transformation; e.g.,][]{mayor2022}.
 
-\paragraph{Linear Trend}
+\paragraph{Linear Trend} Linear trends are a common research interest for longitudinal data.
+``Do patient symptoms improve?'' or ``Does worker productivity decline?'', are familiar types of research questions.
+Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients.
 
-\paragraph{Nonlinearity}
+\paragraph{Nonlinearity} Nonlinear trends are important in two regards.
+Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
+Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}.
+The structure of nonlinear trends is often mathematically captured using polynomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
 
@@
 
-model based clustering: \citep{bulteel2016, stefanovic2022}
-
-parameter-based clustering: VAR \citep{ernst2021} splines \citep{axen2011}
-
-sort of feature-based clustering in psych \citep{heylen2016} as well as \citep{krone2018}
-
-clustering in reduced space \citep{timmerman2013}
+\noindent sort of feature-based clustering in psych \citep{heylen2016} as well as \citep{krone2018}
 
@@
 The feature-based clustering can be structured into four main steps \citep{rasanen2009, wang2006}.
-(1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, (3) an optional feature reduction step if there are too many data points for the clustering algorithms, and (4) the actual clustering of the time series features.
+(1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters.
 We provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.

### 2023-01-05 manuscript.tex
:pencil: f-b ts clust intro full draft

@@
 
-Feature-based time series clustering broadly speaking means that important features from the time series data are extracted (Ratanamahatana & Keogh, 2004).
-The extracted features can then be used as inputs to a clustering algorithm such as k-means or hierarchical clustering to group the time series into clusters (Xu et al., 2013).
+The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
+In its most general approach, this can include any statistical summaries of the time series \citep[i.e., features; e.g.,][]{maharaj2019}.
+The extracted features can then be used as inputs to most clustering algorithms, including k-means or hierarchical clustering to group the participants into clusters \citep{wang2006}.
+The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
+However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
+Feature-based time series clustering has been a valued tool in many empirical fields for its flexibility and interpretability.
+As a result, features have been extracted for a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
 
-Feature-based time series clustering has been a valued tool in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
+As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos.
+This selection was also used by \citep{fulcher2013}.
+\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
+A final illustration comes from a commonly used software package for feature extraction `tsfresh', which allows users to extract a total of 794 features to describe all kinds of aspects of a time series \citep[][]{christ2018}.
 
-As an illustration of the flexibility of this approach, a commonly used software package for this approach allows users to extract a total of 794 time series features \citep[][]{christ2018}.
+Given that there are hundreds of possible features to describe psychological time series, one approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series \citep[e.g.,][]{christ2018}.
+However, such an approach is not always advisable for psychological time series.
+For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 794 time series features.
+More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
+This is particularly the case when features have a conceptual or psychological meaning.
+Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
+But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
+We will, thus, propose and describe a set of time series features that have a strong backing within the ESM literature and as such offer a reasonable starting point for most concepts within the field.
 
 \subsection{Features of Psychological Time Series}
-
-in psychology:
-
-\citet{kuppens2017} originally proposed (1) variability, (2) co-variance or ICC, (3) inertia, and (4) cross-lagged correlations.
-These features were then extended and formalized by \citet{krone2018} into (1) within-person variance, (2) innovation, (3) inertia or autocorrelation, (4) cross-lagged regression parameters, (5) covariances as granularity, and (6) mean intensity.
-\citet{krone2018} were even able to use these features to cluster study participants.
-
-\citet{dejonckheere2019} discuss (1) mean, (2) variance, (3) relative variance, (4) instability, (5) inertia / autoregression, (6) interdependence / network density, (7) granularity or differentiation / ICC, (8) correlation, and (9) diversity / Gini coefficient.
-
-Outside the psychological literature, time series features have also included more summaries in terms of measurement distribution, categorical states, and, importantly, nonlinear developments.
-
-\citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data.
-
-\citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos.
-This selection was also used by \citep{fulcher2013}.
-\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
 
@@
 An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
+
+\citet{kuppens2017} originally proposed (1) variability, (2) co-variance or ICC, (3) inertia, and (4) cross-lagged correlations.
+These features were then extended and formalized by \citet{krone2018} into (1) within-person variance, (2) innovation, (3) inertia or autocorrelation, (4) cross-lagged regression parameters, (5) covariances as granularity, and (6) mean intensity.
+\citet{krone2018} even built a parametric model to use a subset of these features to cluster study participants.
+
+\citet{dejonckheere2019} discuss (1) mean, (2) variance, (3) relative variance, (4) instability, (5) inertia / autoregression, (6) interdependence / network density, (7) granularity or differentiation / ICC, (8) correlation, and (9) diversity / Gini coefficient.
+
+Time series features have also come to include more summaries in terms of measurement distribution, categorical states, and, importantly, nonlinear developments.
+As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data.
 
@@
 \paragraph{Feature extraction}
-The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
+more technical description model + descriptive stats
+% The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
 There are, however, hundreds of possible features to describe psychological time series \citep[e.g., tsfresh][]{christ2018}.

### 2023-01-07 manuscript.tex
:pencil: feature refs added

@@
 Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}.
-Once meaningful and robust structures and patterns are detected, time series clustering can, thus, be used to inform the development of robust theories as well as targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions \citep[e.g.,][]{borsboom2021, eronen2020}.
+By detecting meaningful and robust structures and patterns, time series clustering can, thus, be used to inform the development of robust theories as well as targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions \citep[e.g.,][]{borsboom2021, eronen2020}.
 
@@
 The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
-In its most general approach, this can include any statistical summaries of the time series \citep[i.e., features; e.g.,][]{maharaj2019}.
+In its most general approach, this can include any statistical summary of the time series \citep[i.e., feature; e.g.,][]{maharaj2019}.
 The extracted features can then be used as inputs to most clustering algorithms, including k-means or hierarchical clustering to group the participants into clusters \citep{wang2006}.
@@
 With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
-As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos.
-This selection was also used by \citep{fulcher2013}.
+As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos \citep[also see][]{fulcher2013}.
 \citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
@@
 
-\citet{kuppens2017} originally proposed (1) variability, (2) co-variance or ICC, (3) inertia, and (4) cross-lagged correlations.
-These features were then extended and formalized by \citet{krone2018} into (1) within-person variance, (2) innovation, (3) inertia or autocorrelation, (4) cross-lagged regression parameters, (5) covariances as granularity, and (6) mean intensity.
-\citet{krone2018} even built a parametric model to use a subset of these features to cluster study participants.
-
-\citet{dejonckheere2019} discuss (1) mean, (2) variance, (3) relative variance, (4) instability, (5) inertia / autoregression, (6) interdependence / network density, (7) granularity or differentiation / ICC, (8) correlation, and (9) diversity / Gini coefficient.
-
-Time series features have also come to include more summaries in terms of measurement distribution, categorical states, and, importantly, nonlinear developments.
-As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data.
+To understand emotion dynamics, \citet{kuppens2017} originally proposed four features: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
+These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
+\citet{krone2018} even built a parametric model to tentatively cluster study participants.
+From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity (i.e., Gini coefficient)\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
+As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
 
@@
 
-\paragraph{Central Tendency} Familiar statistics from probability theory, the two features sit at the heart of many fundamental psychological questions.
-``Are some people happier than others?'' (i.e., difference in central tendency).
-Mathematically, the central tendency has parametric and robust options to choose from (e.g., parametric: mean; robust: median).
+\paragraph{Central Tendency} Central tendency refers to the statistical measures that represent the ``middle" or ``average" of a set of data.
+The most common measures of central tendency are the mean, median, and mode (Hopkins, 2020).
+As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
+Researchers might, for example, be interested in whether ``Over a one month period, are some people happier than others?''.
+A recent review paper pointed out that the relatively simple mean statistic, is often overlooked in psychological time series analyses but is usually essential for understanding psychological differences \citep{bringmann2018c}.
+As such, several time series modeling approaches now advocate for conscious consideration of the mean \citep[e.g.,][]{bringmann2017a}.
+As with cross-sectional data, for time series data the appropriate measure of central tendency depends on the specific characteristics of the data being analyzed.
+For instance, if the data are normally distributed, the mean is often an appropriate measure.
+If the data are skewed or have outliers, the median may be a better choice \citep{weisberg1992}.
 
-\paragraph{Variability} "Are people settled in their attitudes towards migrants or do they fluctuate over time?'' (i.e., difference in variability).
+\paragraph{Variability} Variability, also referred to as dispersion or spread, captures to the degree to which a set of data differs from the central tendency \citep{weisberg1992}.
+In time series analyses, variability is important because it provides information about the distribution and diversity of the data.
+"Are people settled in their attitudes towards migrants or do they fluctuate over time?'' (i.e., difference in variability).
+There are several measures of variability, including the range, standard deviation, and variance (Hopkins, 2020).
+The appropriate measure to use depends on the specific characteristics of the data being analyzed.
 parametric: standard deviation, robust: median absolute deviation.
 
+``Capturing variability may not just be helpful in identifying a worse psychological state or an inadequate emotion regulation strategy.
+It may also predict future course of symptoms.
+Affective variability at baseline predicted the recurrence of depressive symptoms over a three-year period in remitted depressed patients \citep{timm2017}.
+Similarly, elevated temporal auto-correlation and variance, as well as increased correlation between emotions, increased the probability for an upcoming shift between a normal and a depressed state \citep{vandeleemput2014}.
+''\citep{myin-germeys2018}
+
+EWS: anomalous variance \citep{wichers2019}
+
+increase in variance \citep{scheffer2009}
+
 \paragraph{(In)stability} Stability captures the average change between two consecutive measurements.
-Do changes tend to be slow and gradual or fast and abrupt?
-Mathematically, (in)stability measurements often look at the average change between two consecutive measurements \citep[e.g., mean absolute change; e.g.,][]{barandas2020}.
+While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the `jumpy-ness' of the data over time.
+In other words, instability reflects the fluctuation or inconsistency in a time series data over time, while variability reflects the range or diversity of values in a time series data \citep{trull2008}.
+For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
+Common research questions with psychological time series might, thus, include: ``Do changes tend to be slow and gradual or fast and abrupt?" Mathematically, (in)stability measurements often look at the average change between two consecutive measurements \citep[e.g., mean absolute change; e.g.,][]{ebner-priemer2009,barandas2020}.
 
@@
 Both forms of self-similarity relate to different types of research questions.
-``Do patients stay in a depressed mood for several measurements?'' (i.e., inertia), and ``Do participants drink more alcohol on Fridays?'' (i.e., periodicity).
+
+Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{REF}.
+``Do patients stay in a depressed mood for several measurements?''
+characterize psychological maladjustment \citep{kuppens2010} and can even predict the initial onset of depression \citep{kuppens2012}.
+Mathematically, inertia is often measured using autocorrelations or autoregressions with a lag-1.
+
+Seasonality is the self-similarity of a measurement that returns in a cyclical manner and is often also referred to as periodicity \citep{REF}.
+``Do participants drink more alcohol on Fridays?''.
 Both forms of self-similarity are often measured using autocorrelations with either a lag-1 (inertia) or the lag of the seasonality (e.g., seven days).

### 2023-01-08 manuscript.tex
:sparkles: self-similarity sect

@@
 \citet{krone2018} even built a parametric model to tentatively cluster study participants.
-From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity (i.e., Gini coefficient)\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
+From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
 As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
@@
 
-\paragraph{Variability} Variability, also referred to as dispersion or spread, captures to the degree to which a set of data differs from the central tendency \citep{weisberg1992}.
-In time series analyses, variability is important because it provides information about the distribution and diversity of the data.
-"Are people settled in their attitudes towards migrants or do they fluctuate over time?'' (i.e., difference in variability).
-There are several measures of variability, including the range, standard deviation, and variance (Hopkins, 2020).
+\paragraph{Variability} Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
+In time series analyses, variability is conceptually important because it provides information about the distribution and diversity of the data.
+Especially within-person changes in variability have recently received increasing attention because they can be a potent early warning signal in psychological time series \citep{helmich2020a, vandeleemput2014}.
+A marked within-person increase in variance has, for example, been found to precede crucial transitions --- often considered to be part of a phenonenon know as critical slowing down \citep[i.e., anomalous variances; e.g.,][]{scheffer2009, wichers2019}.
+
+But also more global person-level differences in time series variability have been found to be crucial and relate to important psychological questions.
+Person-specific differences in variability have, for example, been found to be indicative of worse psychological states \citep{myin-germeys2018}.
+And, similarly, person-level differences have also been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
+But also more broadly speaking, psychological researchers and practitioners are theoretical-empirically often interested in person-level differences in variability.
+Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
+
+Methodologically, there are several measures of variability, including the range, standard deviation, and variance.
 The appropriate measure to use depends on the specific characteristics of the data being analyzed.
-parametric: standard deviation, robust: median absolute deviation.
+For normally distributed data variances and standard deviation are often preferred, whereas data with outliers or nonparametric distributions are often better described with robust variability measures such as the median absolute deviation \citep{weisberg1992}.
 
-``Capturing variability may not just be helpful in identifying a worse psychological state or an inadequate emotion regulation strategy.
-It may also predict future course of symptoms.
-Affective variability at baseline predicted the recurrence of depressive symptoms over a three-year period in remitted depressed patients \citep{timm2017}.
-Similarly, elevated temporal auto-correlation and variance, as well as increased correlation between emotions, increased the probability for an upcoming shift between a normal and a depressed state \citep{vandeleemput2014}.
-''\citep{myin-germeys2018}
-
-EWS: anomalous variance \citep{wichers2019}
-
-increase in variance \citep{scheffer2009}
-
-\paragraph{(In)stability} Stability captures the average change between two consecutive measurements.
+\paragraph{(In)stability} Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009}.
 While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the `jumpy-ness' of the data over time.
@@
 For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
-Common research questions with psychological time series might, thus, include: ``Do changes tend to be slow and gradual or fast and abrupt?" Mathematically, (in)stability measurements often look at the average change between two consecutive measurements \citep[e.g., mean absolute change; e.g.,][]{ebner-priemer2009,barandas2020}.
+Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
+Conceptually, the instability feature, thus, relates to a broad range of research questions, including: ``Do changes tend to be slow and gradual or fast and abrupt?" Methodologically, (in)stability measurements often look at the average change between two consecutive measurements, such as the mean absolute change \citep[e.g.,][]{ebner-priemer2009,barandas2020}.
 
-\paragraph{Self-similarity} Self-similarity describes how much a measurement carries over to future measurements.
-There are two main ways in which experiences tend to be connected to future experiences --- resistance to change and seasonality.
+\paragraph{Self-similarity} Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}.
+That is, self-similarity describes how much a measurement carries over to future measurements.
+There are two main ways in which experiences tend to be connected to future experiences in psychological time series --- resistance to change and seasonality.
 Both forms of self-similarity relate to different types of research questions.
 
-Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{REF}.
-``Do patients stay in a depressed mood for several measurements?''
-characterize psychological maladjustment \citep{kuppens2010} and can even predict the initial onset of depression \citep{kuppens2012}.
-Mathematically, inertia is often measured using autocorrelations or autoregressions with a lag-1.
+Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{kuppens2010}.
+If inertia is high a development tends to stay in a certain state or move towards a certain direction.
+Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
+In a similar vein, high inertia at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
+Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?'' One way to measure inertia in time series data is through the use of autocorrelations or autoregressions with a lag-1, which look at the correlation between a given time series and itself at the previous measurement occasion.
+High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}.
 
-Seasonality is the self-similarity of a measurement that returns in a cyclical manner and is often also referred to as periodicity \citep{REF}.
-``Do participants drink more alcohol on Fridays?''.
-Both forms of self-similarity are often measured using autocorrelations with either a lag-1 (inertia) or the lag of the seasonality (e.g., seven days).
-Periodicity can additionally be captured using wave patterns \citep[e.g., fourier or wavelet transformation; e.g.,][]{mayor2022}.
+Seasonality is the self-similarity of a measurement that returns in a cyclical manner and is often also referred to as periodicity \citep{gregson1983}.
+Periodicity in psychological time series data, thus, refers to the presence of recurring patterns at regular intervals.
+In psychological time series, these cyclical patterns are not only dangerous for models that assume stationarity \citep{beal2015} but can be interesting psychological properties in their own regard \citep{epskamp2018, schmittmann2013}.
+Humans generally live in seasonal environments, including, for example, exam periods at universities \citep{fuller2003}.
+But also more internally, personality manifestations, for example, have long been assumed to develop cyclically \citep{cattell1957}.
+The associated research questions can be driven by an understanding of external seasonality factors such as: ``Do participants drink more alcohol on Fridays?'' But research questions can also address periodic patterns as they may be the result of underlying psychological processes: ``Does stress increase with fatigue cycles?'' Similar to inertia, periodicity is often measured using autocorrelations but then with the lag of the seasonality or cycle (e.g., daily, seven days, one month, one semester).
+Periodicity can additionally be captured using spectral analysis or decomposition, such as with wave patterns \citep[e.g., fourier or wavelet transformation; e.g.,][]{mayor2022}.
 

### 2023-01-08 manuscript.tex
:sparkles: lin trnd added

@@
 
-\paragraph{Linear Trend} Linear trends are a common research interest for longitudinal data.
-``Do patient symptoms improve?'' or ``Does worker productivity decline?'', are familiar types of research questions.
-Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients.
+\paragraph{Linear Trend} Linear trends in time series data refer to the presence of a straight-line development of a variable being measured.
+This type of trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
+For psychological time series, researchers have, for example, pointed out linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}.
+Linear trends are often the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}.
+In empirical practice, linear trends are, thus, a common research interest for psychological time series data: ``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions.
+Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients, but also using differential equations \citep{chow2019}.
 
-\paragraph{Nonlinearity} Nonlinear trends are important in two regards.
-Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
-Questions like ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such the bicoherence metrics \citep{cuddy2009}.
-The structure of nonlinear trends is often mathematically captured using polynomial coefficients or more broadly by how wiggly the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
+\paragraph{Nonlinearity} Changes in psychology are not always linear \citep{hayes2007}
+
+Nonlinear trends are important in two regards. Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
+
+``Some examples of nonlinear DEs have already been proposed in studies of ovulatory regulation (Boker, Neale, \& Klump, 2014), circadian rhythms (E.
+N.
+Brown \& Luithardt, 1999), cerebral development (Thatcher, 1998), substance use (Boker & Graham, 1998), cognitive aging (Chow \& Nesselroade, 2004), parent-child interactions (Thomas \& Martin, 1976), dyadic relationships (Chow, Ferrer, \& Nesselroade, 2007); and sudden transitions in attitudes (van der Maas, Kolstein, \& van der Pligt, 2003)." \citep{chow2019}
+
+Questions like: ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.
+
+The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
+One common way is to capture the trends using polynomial coefficients or more broadly by how `wiggly' the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
+
+depression most likely non-linear  \citep{hosenfeld2015}. Patients in recovery showed sudden changes in the improvement of depression \citep{helmich2020a}
 

### 2023-01-08 manuscript.tex
:sparkles: nonlinearity

@@
 Fortunately, past conceptual and empirical efforts offer valuable discussions of common time series features in psychological research.
-While the final selection of features should always be driven by the research questions and field-specific conventions, we can build a practical toolbox of meaningful time series features for psychological data.
-In particular, we propose to focus on six features based on common psychological research questions and recent works on affect dynamics \citep[e.g.,][]{dejonckheere2019, kuppens2017, adya2001}.
-An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
-
 To understand emotion dynamics, \citet{kuppens2017} originally proposed four features: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
@@
 
-This selection of the proposed six time-series features is in no way exhaustive or imperative.
-Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice (see \tblref{tab:esmFeatures} for a more complete overview).
+\input{tables/esmPsychFeatures}
+
+Some of the proposed features are not necessarily well suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the emotion dynamics literature, which has been so formative for the analysis of psychological time series data.
+In particular, covariances and cross-lagged correlations often produce a staggering amount of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}.
+Others such as network density parameters, used to summarize variable interdependence, might not always be meaningful for psychological data \citep{bringmann2019a}.
+At the same time, linear and nonlinear trends are not captured by the features commonly proposed for psychological time series, because the features are often developed for stationary models \citep[e.g.,][]{krone2018}.
+
+Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, we hope to build a practical toolbox of meaningful time series features for psychological data.
+In particular, we propose to focus on six features based on common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
+We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-base clustering of psychological time series data.
+An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
 
@@
 
-\paragraph{Nonlinearity} Changes in psychology are not always linear \citep{hayes2007}
+\paragraph{Nonlinearity} Changes in psychology are not always linear, instead nonlinearity is is a common feature of psychological time series \citep{hayes2007}.
+As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
+Similarly, patients in recovery of depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
+But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
+For psychological researchers, nonlinear developments are, thus, important in two regards.
+Firstly, nonlinearity as a deviation from linear trends is an important feature defining many psychological time series and secondly, the shape and structure of the nonlinear trend can be essential in understanding the important changes and transitions of human experiences \citep{chow2019}.
+Conceptually, researchers might accordingly have research question about the type of the development: ``Is the development of anxiety a nonlinear process?'' as well as the shape and structure of the development: ``Does anxiety change as a sudden or smooth transition?'' Accordingly, researchers can capture the type of development with nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.
+The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
+Common ways to capture the trends include the use of polynomial- and differential equation parameters or more broadly to capture how `wiggly' the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
 
-Nonlinear trends are important in two regards. Firstly, nonlinearity is a deviation from linear trends and secondly, the shape and structure of the nonlinear trend.
-
-``Some examples of nonlinear DEs have already been proposed in studies of ovulatory regulation (Boker, Neale, \& Klump, 2014), circadian rhythms (E.
-N.
-Brown \& Luithardt, 1999), cerebral development (Thatcher, 1998), substance use (Boker & Graham, 1998), cognitive aging (Chow \& Nesselroade, 2004), parent-child interactions (Thomas \& Martin, 1976), dyadic relationships (Chow, Ferrer, \& Nesselroade, 2007); and sudden transitions in attitudes (van der Maas, Kolstein, \& van der Pligt, 2003)." \citep{chow2019}
-
-Questions like: ``Is the development of anxiety a nonlinear process?'' are mathematically captured using nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.
-
-The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
-One common way is to capture the trends using polynomial coefficients or more broadly by how `wiggly' the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
-
-depression most likely non-linear  \citep{hosenfeld2015}. Patients in recovery showed sudden changes in the improvement of depression \citep{helmich2020a}
+It should be noted that this selection of the proposed six time-series features is in no way exhaustive or imperative.
+Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options.
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for a more complete overview).
 
@@
 We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.
-
-\input{tables/esmPsychFeatures}
 

### 2023-01-08 manuscript.tex
:sparkles: Cluster Evaluation sctn added

@@
 To expand interpretatbility, researchers have proposed to model additional types of parameters, including variability or granularity parameters \citep[e.g., see][]{krone2018}.
+
 In this manuscript we instead, propose to look at the broader machine learning literature and to consider a wider range of time series features to describe psychological time series.
@@
 It should be noted that this selection of the proposed six time-series features is in no way exhaustive or imperative.
-Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options.
+Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
 The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for a more complete overview).
@@
 \subsection{Analysis Steps}
-
-------------------------
-
-NOTES:
-
-\noindent sort of feature-based clustering in psych \citep{heylen2016} as well as \citep{krone2018}
-
-------------------------
-
-The feature-based clustering can be structured into four main steps \citep{rasanen2009, wang2006}.
+Armed with a selection of potentially meaningful and accurate time series features to address a given research question, we can consider the practical steps of feature-based time series clustering.
+The feature-based clustering analysis can be structured into four main steps \citep{rasanen2009, wang2006}.
 (1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters.
-We provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
+It is important to note that these steps are just a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
+Nonetheless, below we will briefly describe each of the steps conceptually.
+We also provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
 
@@
 
-\paragraph{Cluster Performance Evaluation}
+\input{tables/clusterApproaches}
 
-PARAGRAPH ON PERFORMANCE METRICS (incl., cross-validation?)
+\paragraph{Cluster Evaluation}
+The final evaluation step includes two main elements, (1) evaluating the performance of the clustering analysis and (2) interpreting the extracted clusters conceptually.
 
-incl., finding cluster prototypes:
-1. The medoid sequence of the set. 2. The average sequence of the set. 3. The local search prototype
+Performance evaluation often means assessing the accuracy, stability, and separation or purity of the clustering \citep{keogh2003}.
+Importantly, any evaluation of the results depends on the research questions, the data, and the methods used.
+However, broadly speaking, evaluation methods can be categorized based on whether the true cluster labels are known or not \citep{saxena2017}.
+If true class labels are known, the cluster assignments can be compared to the true class labels --- using measures such as the F-measure, adjusted Rand index, mutual information, and normalized mutual information \citep[i.e., external evaluation; e.g.,][]{liao2005}.
+However, if the true cluster assignments are unknown, as with most of the exploratory clustering we speak to, the quality of the clusters is assessed based on the characteristics of the data itself \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
+One popular internal evaluation measure is the silhouette score, which measures the degree to which each time series data point is similar to other points within the same cluster, compared to points in other clusters \citep{rousseeuw1987}.
+Another measure is the Calinski-Harabasz index, which assesses the compactness and separation of the clusters \citep{calinski1974}.
+Another common method is to compare different cluster solutions using the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.
+Another important aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}.
+This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}.
+It is important to note that no single evaluation measure is perfect, and different measures may produce different results depending on the characteristics of the data and the research question being addressed \citep{kittler1998}.
+It is therefore important to consider a variety of evaluation measures and to carefully interpret the results in the context of the specific analysis \citep{vinh2009}.
 
-External evaluation measures: These measures require knowledge of the true class labels and are used to compare the cluster assignments to the true class labels.
-Examples include accuracy, F-measure, adjusted Rand index, mutual information, and normalized mutual information.
+The interpretation of feature-based time series clustering in psychology involves understanding the meaning and implications of the clusters obtained from the analysis.
+Here, it is also important to consider the interpretability and practical significance of the clusters.
+While clusters with a high degree of internal cohesion may be more statistically robust, they may not necessarily be meaningful or useful in the context of the specific research question being addressed \citep{kaufman1990}.
+To gain a qualitative understanding of the extracted clusters users can use both the patterns that emerge from the clusters themselves as well as comparisons to between-person differences not included in the initial clustering \citep[e.g.,][]{monden2022}.
+As for the inspection of the clusters themselves, one common approach is comparing the clusters based on the average values of meaningful features \citep{Kennedy2021}.
+Assessing the clusters based on their average features allows users to gain an understanding of the developmental profiles.
+One cluster might, for example, have a relatively strong linear increase in well-being or participants in another cluster might have a relatively high variability of negative emotions.
+Another common option for interpretation is finding an appropriate prototype to represent the cluster's trajectory.
+This is often easiest with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021}.
+For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.
+For most other cases, it is often recommended to use some form of average over time of the responses within the cluster \citep{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.}.
+These prototypes can then, for example, be inspected visually to identify qualitative differences between the cluster prototypes.
+An additional option for interpretations are out-of-cluster comparisons, where users can assess whether the clusters differ on some stable between-person differences, such as personality, age, or situational circumstances \citep[e.g.,][]{monden2022}.
 
-Internal evaluation measures: These measures do not require knowledge of the true class labels and are used to evaluate the quality of the clusters based on the characteristics of the data itself.
-Examples include silhouette coefficient, Calinski-Harabasz index, Davies-Bouldin index, and the Elbow method.
-
-PARAGRAPH ON REPORTING \citep[e.g.,][]{vandeschoot2017}
-
-\input{tables/clusterApproaches}
+Before moving to our empirical illustration, we would like to comment on the reporting of feature-based time series clustering.
+Clustering is inherently an exploratory analysis and, as such, each of the analysis steps involves a set of decisions by the user.
+While this is common for other types of analyses as well, it remains imperative to report the analysis in the most transparent and reproducible manner possible.
+In our own description of the method, we have separated both the time series features as well as the analysis steps into distinct taxonomies in order to facilitate the transparency of the individual steps and decision moments.
+Beyond the structures proposed here, \citet{vandeschoot2017} have proposed an extensive checklist for latent trajectory studies.
+Most of their recommendations and reporting guidelines apply to feature-based clustering as well and might even offer a template for researchers hoping to pre-register their analysis procedures.
 

### 2023-01-15 manuscript.tex
:memo: current version

@@
 
-Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][; also see \fgrref{fig:ScopusEsm}]{hamaker2017}\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
+Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][]{hamaker2017}\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
 While the terms come for different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events.
@@
 Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across several variables \citep[e.g.,][]{ernst2021}.
-One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of inductively grouping participants based on the similarity of their time series \citep{ariens2020}.
-
-\begin{figure}[!hbtp] %hbtp
-  \caption{Scopus ESM Development}
-  \label{fig:ScopusEsm}
-  \centering\includegraphics[width=\textwidth]{figures/Scopus-ESM-Development.png}
-  \caption*{Note: \\
-  Search Terms: `( "ecological momentary assessment"  OR  "experience sampling"  OR  "intensive longitudinal")' in Title, Abstract, or Keyword.}
-\end{figure}
+One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
 
@@
 As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{timmerman2013}; see \fgrref{fig:tsClustTax}).
-Ideally, the representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[e.g.,][]{vandermaaten2009}.
+Ideally, the representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[][]{vandermaaten2009}.
 Extracting accurate and meaningful representations of the time series can be essential for understanding what goes into the clustering algorithm (i.e., assists with explainability) and can be crucial in making sense of the final cluster output \citep[i.e., assists with interpretability; e.g.,][]{Kennedy2021}.
@@
 And, similarly, person-level differences have also been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
-But also more broadly speaking, psychological researchers and practitioners are theoretical-empirically often interested in person-level differences in variability.
+As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
 Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
@@
 While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the `jumpy-ness' of the data over time.
-In other words, instability reflects the fluctuation or inconsistency in a time series data over time, while variability reflects the range or diversity of values in a time series data \citep{trull2008}.
+In other words, variability reflects the range or diversity of values in a time series data, while instability reflects the fluctuation or inconsistency in a time series data over time  \citep{trull2008}.
 For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
@@
 Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{kuppens2010}.
-If inertia is high a development tends to stay in a certain state or move towards a certain direction.
+If inertia is high a development tends to stay in a certain state or continues to move towards a certain direction.
 Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
@@
 Linear trends are often the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}.
-In empirical practice, linear trends are, thus, a common research interest for psychological time series data: ``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions.
+In empirical practice, linear trends are, thus, a common research interest for psychological time series data.
+``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions.
 Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients, but also using differential equations \citep{chow2019}.
@@
 The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
-Common ways to capture the trends include the use of polynomial- and differential equation parameters or more broadly to capture how `wiggly' the line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
+Common ways to capture the trends include the use of polynomial- and differential equation parameters or more broadly to capture how `wiggly' the trend line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
 
@@
 Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for a more complete overview).
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
 
@@
 (1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters.
-It is important to note that these steps are just a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
+It is important to note that these steps are a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
 Nonetheless, below we will briefly describe each of the steps conceptually.
@@
 Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
-Two important steps are making the time-frames and response scales comparable across participants.
-Extracting features that describe the development of a month for one participant but 90 days for another participant might not be comparable for some psychological phenomena.
-This comparability is important on a conceptual level but the difference in data availability might also influence the clustering steps, where the participant with the shorter measurement period would, for example, have a larger number of missing values.
-It is thus generally advised to reduce the data to a common and comparable time frame.
+Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses \citep{liao2005}.
 
 \paragraph{Feature extraction}
-more technical description model + descriptive stats
-% The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
-There are, however, hundreds of possible features to describe psychological time series \citep[e.g., tsfresh][]{christ2018}.
-One approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series.
-However, such an approach is not always advisable for psychological time series.
-For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 180 time series features.
-More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
-This is particularly the case when features have a conceptual or psychological meaning.
-Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
-But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
-We would, thus, strongly advocate for a careful selection of time series features that are meaningful to the field and concepts.
+The main feature extraction step ultimately depends on the choice of the selected time series features but generally speaking the process is comparatively simple.
+For each of the included variables the features are extracted for each participant and replace the original time series \citep{wang2006}.
 
 \paragraph{Feature reduction}
-Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points usually remains too large for most clustering algorithms.
+Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm.
 As an example, a relatively common scenario would include 10 variables of interest, where 8 time series features are extracted, resulting in 80 features per participant (with a common sample size of 100 participants that would result in a total of 8,000 data points in this hypothetical example).

### 2023-01-17 manuscript.tex
:memo: clustering section

@@
 We argue that feature-based clustering offers a flexible variety, less strict assumptions, easy and intuitive extraction, theory-based selection, and meaningful psychological interpretability.
+STRUCTURE SENTENCE: introduction of approach + empirical illustration
 
@@
 The difficulty of working with such a large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
-To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection (a full overview of the approaches, methods, and common algorithms is available in \tblref{tab:featureReduction}).
+To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection \citep[e.g.,][also see \tblref{tab:featureReduction} for a full overview of the approaches, methods, and common algorithms]{erdogmus2008}.
 
 Feature selections seek to create a subset of the most important features.
-The many available approaches differ in how they seek to determine the importance of the individual features.
-Generally speaking, selection methods can be categorized as `filter', `wrapper', or `hybrid' methods.
-The filter methods, broadly speaking determine important features by identifying irrelevant features (e.g., because features do not capture much information), and identifying redundant features (e.g., because features capture the same information).
-Wrapper methods on the other hand avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features.
-A wrapper, thus, compares the performance of models with different feature combinations.
+The many available approaches differ in how they seek to determine the importance of the individual features \citep{alelyani2014}.
+Generally speaking, selection methods can be categorized as `filter', `wrapper', or `hybrid' methods \citep{hoque2014}.
+The filter methods, broadly speaking determine important features by identifying irrelevant features (e.g., because features do not capture much information), and identifying redundant features \citep[e.g., because features capture the same information; e.g., see][]{yu2004}.
+Wrapper methods on the other hand avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features \citep{kohavi1997}.
+A wrapper, thus, compares the performance of models with different feature combinations \citep[e.g., see][]{tang2014}.
 Traditional examples of wrappers are forward selection or backward elimination procedures.
-Because filter methods might not always perform well and wrapper methods are computationally intensive\footnote{Computationally \textit{k} features could be considered in \(2^k – 1\) possible combinations --- for the example of 80 features that would allow for \(1.20 \times 10^{24}\) (over one septillion) combinations.}, hybrid methods seek to combine the two methods and find a balance between computational effort (i.e., efficiency) and performance (i.e., effectiveness).
-Methods might, for example, use a filter step to reduce the size of features considered in a later wrapper step.
+Because filter methods might not always perform well and wrapper methods are computationally intensive\footnote{Computationally \textit{k} features could be considered in \(2^k – 1\) possible combinations --- for the example of 80 features that would allow for \(1.20 \times 10^{24}\) (over one septillion) combinations.}, hybrid methods seek to combine the two methods and find a balance between computational effort (i.e., efficiency) and performance \citep[i.e., effectiveness; e.g.,][]{alelyani2014}.
+Methods might, for example, use a filter step to reduce the size of features considered in a later wrapper step \citep{hsu2011}.
 
 The second general approach to the dimensionality curse has been feature projection.
-Projection methods are relatively common in psychological research --- including, factor- and principal component analyses.
-Generally speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features (i.e., the data is projected to a lower dimensional space).
-Commonly, this is achieved using linear transformations (e.g., principal component analysis), or more complex nonlinear transformations (e.g., t-SNE).
+Generally speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features \citep[i.e., the data is projected to a lower dimensional space; e.g.,][]{carreira-perpinan1997}.
+Projection methods are relatively common in psychological research --- including, factor- and principal component analyses \citep{vandermaaten2009}.
+Commonly, this is achieved using linear transformations \citep[e.g., principal component analysis; see][]{cunningham2015}, or more complex nonlinear transformations \citep[e.g., t-SNE; see]{lee2007}.
 Generally speaking, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
@@
 \paragraph{Feature clustering}
-For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible.
+For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible \citep{liao2005}.
 The crux of clustering is, thus, to have clearly defined and effective measurements of (dis)similarity.
-Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment (or similarity measurement for qualitative features).
+Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment \citep[or similarity measurement for qualitative features; see][]{Aghabozorgi2015}.
 While others have produced excellent overviews of the many clustering approaches available \citep[e.g.,][]{xu2015}, we will briefly introduce some of the more readily available approaches suitable for most time series feature data.
-The well-established and readily available clustering approaches can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof (see \tblref{tab:clusterApproaches}).
+The well-established and readily available clustering approaches can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof \citep[see \tblref{tab:clusterApproaches} for an overview; also see][for a broader review]{jain1999}.
 
 Each of these approaches can be a valuable clustering approach for time series feature data and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency.
-As an example, under ideal conditions, all approaches are likely to suggest very similar clusters.
-However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate.
-Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective.
-Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \hl{(Corduas \& Piccollo, 2008)}.
-Yet, in cases where mis-specifications are costly, a method with fewer assumptions might be more advisable.
+As an example, under ideal conditions, most approaches are likely to provide a similar cluster solution \citep[e.g., for well seprated groups with little noice and few outliers; e.g., see][]{peng2022}.
+However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate \citep[e.g.,][]{langfelder2008}.
+Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective \citep[e.g.,][]{jain2010}.
+Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \citep[e.g.,][]{corduas2008}.
+Yet, in cases where mis-specifications are costly, a method with fewer assumptions might be more advisable \citep[e.g.,][]{ankerst1999}.
 
@@
 For our own illustration, we have chosen that centroid-based k-means clustering.
-While k-means comes at the expense of high accuracy with more complex cluster shapes, we specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the data.
-K-means is also well-established within the research community and has been readily implemented in many statistical software packages.
-Additionally, many of the feature selection methods have specifically been designed for the well-established k-means algorithm.
+While k-means comes at the expense of high accuracy with more complex cluster shapes, we specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the data \citep{jain2010}.
+K-means is also well-established within the research community and has been readily implemented in many statistical software packages \citep{hand2005}.
+Additionally, many of the feature selection methods have specifically been designed for the well-established k-means algorithm \citep[e.g.,][]{boutsidis2010}.
 As such, the k-means offers a good starting point for many psychological researchers and the method should be generalizable across a relatively wide variety of projects.

### 2023-01-17 manuscript.tex
:memo: illustration intro

@@
 At the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-The migration ESM research, thus, offers a good example of the modern ESM data that introduces diverse conceptualizations and more event-specific missingness patterns.
+The migration ESM research, thus, offers a good example of the ESM data that introduces diverse conceptualizations and more event-specific missingness patterns.
 

### 2023-01-17 Supplemental-Material-A-Annotated-Analysis.Rmd
:construction: pca code
:sparkles: feature clustering
:shit: formula test

@@
 
-# Markdown Options
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
@@
 
-## Clustering
+## Feature Clustering
 
@@
 
+IMPORTANT FIX: https://stackoverflow.com/questions/66448838/add-an-interactive-button-to-switch-which-variable-is-mapped-to-fill-in-ggplotly
+
 ```{r features for each variable}

### 2023-01-17 methods-and-results.Rmd
:shit: formula test

@@
 For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in the empirical article by \citet[][]{Kreienkamp2022b}.
+Full methodological details are available in \citet[][]{Kreienkamp2022b}.
 
@@
 We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:descrLong}.
+Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
 
@@
 
+\input{tables/descrLong}
+
 \subsection{Analysis and Results}
-To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster steps in sequential order.
-For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REFERENCES).
+To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
+For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REF).
 
@@
 We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
+
+For the central tendency: $Median$ where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
+
+\begin{equation}
+  Median(X_ij) = \left\{ 
+    \begin{array}{ll}
+      X \left[ \frac{n+1}{2} \right] & if n is odd \\
+      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & if n is even
+    \end{array}
+\end{equation}
 
@@
 
-\Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
+<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
 However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
 If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}
-I had originally written: 
-"To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
-but I went back to the 'raw' PC-scores now. But we should discuss this.}}
+The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
+<!-- I had originally written:  -->
+<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
+<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
 
@@
 
-\Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}}
+<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->
 

### 2023-01-17 methods-and-results.tex
:shit: formula test

@@
 during the ESM surveys and were available in all three studies. Full
-methodological details are available in the empirical article by
-\citet[][]{Kreienkamp2022b}.
+methodological details are available in \citet[][]{Kreienkamp2022b}.
 
@@
 methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are available
-in \tblref{tab:descrLong}.
+but basic item information, descriptives, and correlations are also
+available in \tblref{tab:descrLong}.
+
+\input{tables/descrLong}
 
@@
 To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster steps in sequential order. For each of the
-steps fully reproducible and annotated code is available in Online
-Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REFERENCES).
+extract--reduce--cluster--evaluate steps in sequential order. For each
+of the steps fully reproducible and annotated code is available in
+Online Supplemental A as well as in the accompanying OSF- and GitHub
+repositories (REF).
 
@@
 \texttt{featureExtractor} function).
+
+For the central tendency: \(Median\) where \(X_{ij}\) is the an ordered
+list of values from the time series of variable \(j\) for participant
+\(i\).
+
+\begin{equation}
+  Median(X_ij) = \left\{ 
+    \begin{array}{ll}
+      X \left[ \frac{n+1}{2} \right] & if n is odd \\
+      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & if n is even
+    \end{array}
+\end{equation}
 
@@
 
-\Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
-However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
-If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}
-I had originally written: 
-"To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores." 
-but I went back to the 'raw' PC-scores now. But we should discuss this.}}
-
 For our illustration of the feature clustering procedure, we use the
@@
 
-\Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}}
-
 We can also assess the clusters across any other person-level variable.

### 2023-01-17 methods-and-results.Rmd
:shit: formulas tst
:eyes: render check formulae

@@
 
-\begin{equation}
-  Median(X_ij) = \left\{ 
-    \begin{array}{ll}
-      X \left[ \frac{n+1}{2} \right] & if n is odd \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & if n is even
-    \end{array}
+\begin{equation} \label{eq:median}
+  Median(X_{ij}) = 
+    \begin{cases}
+      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
+      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
+    \end{cases}
+\end{equation}
+
+For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the $Median$ (as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$.
+
+\begin{equation} \label{eq:mad}
+  MAD(x_{ijt}) = Median(| x_{ijt} - \overline{x_{ij}} |)
 \end{equation}

### 2023-01-17 methods-and-results.tex
:shit: formulas tst
:eyes: render check formulae

@@
 
-\begin{equation}
-  Median(X_ij) = \left\{ 
-    \begin{array}{ll}
-      X \left[ \frac{n+1}{2} \right] & if n is odd \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & if n is even
-    \end{array}
+\begin{equation} \label{eq:median}
+  Median(X_{ij}) = 
+    \begin{cases}
+      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
+      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
+    \end{cases}
+\end{equation}
+
+For distribution we chose the \textit{Median Absolute Deviation}
+(\(MAD\)), where we calculate the \(Median\) (as in \eqref{eq:median})
+for the absolute deviations of measurement \(x\) at time point \(t\) for
+participant \(i\) and variable \(j\).
+
+\begin{equation} \label{eq:mad}
+  MAD(x_{ijt}) = Median(| x_{ijt} - \overline{x_{ij}} |)
 \end{equation}

### 2023-01-17 methods-and-results.Rmd
:shit: formula tst

@@
 
-For the central tendency: $Median$ where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
+For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
+The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
 
 \begin{equation} \label{eq:median}
-  Median(X_{ij}) = 
+  M(X_{ij}) = 
     \begin{cases}
@@
 
-For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the $Median$ (as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$.
+For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
 
 \begin{equation} \label{eq:mad}
-  MAD(x_{ijt}) = Median(| x_{ijt} - \overline{x_{ij}} |)
+  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
 \end{equation}
+
+mean absolute change (MAC) for stability
+
+\begin{equation} \label{eq:mac}
+  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
+\end{equation}
+
+lag--1 autocorrelation for inertia: lag $l=1$
+
+\begin{equation} \label{eq:ar}
+  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
+  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
+\end{equation}
+
+overall linear regression slope to describe the linear trend
+
+summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
 

### 2023-01-17 methods-and-results.tex
:shit: formula tst

--- 2023-01-17 methods-and-results.tex
@@
 
-For the central tendency: \(Median\) where \(X_{ij}\) is the an ordered
-list of values from the time series of variable \(j\) for participant
-\(i\).
+For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
+the an ordered list of values from the time series of variable \(j\) for
+participant \(i\). The calculation depends on whether the number of
+measurements in a time series \(n\) is odd or even.
 
-\begin{equation}
-  Median(X_ij) = \left\{ 
-    \begin{array}{ll}
-      X \left[ \frac{n+1}{2} \right] & if n is odd \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & if n is even
-    \end{array}
+\begin{equation} \label{eq:median}
+  M(X_{ij}) = 
+    \begin{cases}
+      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
+      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
+    \end{cases}
 \end{equation}
+
+For distribution we chose the \textit{Median Absolute Deviation}
+(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
+in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
+at time point \(t\) for participant \(i\) and variable \(j\) from the
+median of that time series \(X\).
+
+\begin{equation} \label{eq:mad}
+  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
+\end{equation}
+
+mean absolute change (MAC) for stability
+
+\begin{equation} \label{eq:mac}
+  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
+\end{equation}
+
+lag--1 autocorrelation for inertia: lag \(l=1\)
+
+\begin{equation} \label{eq:ar}
+  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
+  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
+\end{equation}
+
+overall linear regression slope to describe the linear trend
+
+summarized the nonlinear trend with the estimated degrees of freedom of
+an empty GAM spline model (edf) --- the edf summarizes the
+\textit{wiggliness} of the spline trend line.
 

### 2023-01-20 methods-and-results.Rmd
:shit: formula tst

@@
 \begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
+  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ijt-l}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
 \end{equation}
@@
 
+\begin{equation} \label{eq:lin}
+  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
+\end{equation}
+
 summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
+We estimate the number of parametric and non-parametric (or smooth) terms $p$ in the model.
+
+\begin{equation} \label{eq:edf}
+  edf \approx \sum p
+\end{equation}
+
+\textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
+This includes the number of parameters used to estimate each smooth function in the model, as well as any additional parameters used for the overall model.On the other hand, the effective degrees of freedom is a measure of the amount of information in the data that is used to estimate the model.
+It takes into account the smoothness of the estimated functions in the model, and is typically smaller than the estimated degrees of freedom.
+Effective degrees of freedom is used to estimate the uncertainty of the model estimates and to calculate the model's goodness of fit.
+In general, the effective degrees of freedom is a more appropriate measure of the complexity of the model than the estimated degrees of freedom, as it accounts for the smoothness of the estimated functions, which can help to prevent overfitting.
+This is because if the estimated function is smooth, it means that the model is not fitting the noise and has a better generalization.
 

### 2023-01-20 methods-and-results.tex
:shit: formula tst

@@
 
+\begin{equation} \label{eq:lin}
+  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
+\end{equation}
+
 summarized the nonlinear trend with the estimated degrees of freedom of
 an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line.
+\textit{wiggliness} of the spline trend line. We estimate the number of
+parametric and non-parametric (or smooth) terms \(p\) in the model.
+
+\begin{equation} \label{eq:edf}
+  edf \approx \sum p
+\end{equation}
+
+\textit{Notes}: In a Generalized Additive Model (GAM), the estimated
+degrees of freedom refers to the number of parameters used to estimate
+the model. This includes the number of parameters used to estimate each
+smooth function in the model, as well as any additional parameters used
+for the overall model.On the other hand, the effective degrees of
+freedom is a measure of the amount of information in the data that is
+used to estimate the model. It takes into account the smoothness of the
+estimated functions in the model, and is typically smaller than the
+estimated degrees of freedom. Effective degrees of freedom is used to
+estimate the uncertainty of the model estimates and to calculate the
+model's goodness of fit. In general, the effective degrees of freedom is
+a more appropriate measure of the complexity of the model than the
+estimated degrees of freedom, as it accounts for the smoothness of the
+estimated functions, which can help to prevent overfitting. This is
+because if the estimated function is smooth, it means that the model is
+not fitting the noise and has a better generalization.
 

### 2023-01-28 manuscript.tex
:bug: duplicate tbls
:shit: intro restr

@@
 % Title and header
-\title{Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data}
+%\title{Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data}
+\title{A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series}
+
 \shorttitle{feature-based time series clustering}
@@
             \item increase in number and variety of ESM data + ts clustering to understand complex developmental differences.
-        \item why cluster time series? 
-        \begin{enumerate}[nosep]
-            \item differences between clusters
-            \item patterns within clusters
-            \item relation to out-of cluster variables
-            \item robust theories and interventions
-        \end{enumerate}
-        \item can't cluster all the raw data (b/c too much data and not directly comparable). Instead, summarize the data and cluster based on the summaries.
-        \item ideally these summaries are:
-        \begin{enumerate}[nosep]
-            \item good at captuing the data
-            \item meaningful = help us understand how the clustering arrived at the output (i.e., explainable) + helps us make sense of the final cluster output (i.e., interpretable)
-        \end{enumerate}
-        \item currently model parameters as summaries.
-        \item limitations of parameter summaries: assumptions not always good at capturing data and model parameters not always interpretable
-        \item current solution: more complicated models (e.g., address violated assumptions or add different types of parameters)
-        \item proposed solution: features as flexible and meaningful approach of summarizing the time series
+            \item ESM literature: features to capture meaningful aspects
+            \item current clustering based on common model parameters, which only include a small selection of features and often have strong assumptions
+            \item Instead of building more complicated models (e.g., address violated assumptions or add different types of parameters), we propose to cluster based on important and relevant features directly.
+            \item Called feature-based clustering. Common procedure in digital phenotyping and the broader ml literature. Used in many fields
+            \item in this paper we provide a practical introduction to the method and illustrate its utility with real-world ESM data.
         \end{itemize}
+    \item Data Used for Illustration
+    \item Analysis Steps + Application
+        \begin{itemize}[nosep]
+            \item input variables
+            \item feature extraction
+            \item (feature reduction)
+            \item feature clustering
+            \item cluster evaluation
+        \end{itemize}
+    \item Discussion (summary, limitation, implications, conclusion, etc.)
+    \newpage
+    \item\ [...]
     \item Feature-based Time Series Clustering
@@
 In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}.
-These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, Monden2015, wardenaar2013}.
-Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their temporal developments across several variables \citep[e.g.,][]{ernst2021}.
-One relatively recent addition to the toolbox of ESM researchers has been time series clustering --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
+These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
+Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
+Researchers and practitioners are, for example, asking: ``Do patients differ in their development of several key symptoms?'' \citep{Monden2015} or ``How do migrants differ in the development of their psychological needs as they arrive in the new environment?'' \citep{Kreienkamp2022d}.
+Recently, one promising way of identifying developmental patterns that differ between participants has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
+Such clustering, however, depends on comparing participants based on meaningful (dis)similarities in their time series \citep[][]{Aghabozorgi2015}.
 
-Time series clustering has a number of conceptual use cases with psychological data.
-Prime among them is the ability to reduce the time, variable, and person complexity by extracting and organizing participant-level structures.
-These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
-These phenomena and principles can be meaningful differences that distinguish participants in different clusters, as well important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
-Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}.
-By detecting meaningful and robust structures and patterns, time series clustering can, thus, be used to inform the development of robust theories as well as targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions \citep[e.g.,][]{borsboom2021, eronen2020}.
+Fortunately, past conceptual and empirical works in the ESM literature have collected a number of measures that capture meaningful aspects of psychological time series.
+Originally, \citet{kuppens2017} proposed four features to capture psychological dynamics: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
+These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
+From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}.
 
-%Time-series clustering has been extremely common in other empirical fields, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-But also for psychological data, time-series clustering has recently become more common \citep[e.g.,][]{ernst2021}.
+Yet, despite this rich diversity of meaningful time series features in psychology, most clustering of ESM data has only focused on a small and restrictive selection of time series characteristics.
+Thus far, one of the most common approaches has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
+While such model parameters are familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c}.
+Additionally, the validity of model parameters is traditionally restricted by the assumptions of the particular statistical model.
+To take the common VAR model as an example, the model explicitly assumes that the time intervals between measurements are consistent (i.e., equidistant measurement assumption), it does not allow for missing values, and it assumes that means or variances do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
+All of these are types of data researchers commonly collect and contradict important researcher interests in the field \citep[][; see \appref{app:ChallengesAppendix} for a more extensive discussion of the current challenges that clustering approaches should address]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
+Model parameters might, thus, not always accurately capture the original time series and are often restricted in the features they capture.
 
-%The main aim of time series clustering is to inductively group participants in such a way that the dynamics of some variable(s) have maximum similarity for individuals in the same group, and minimum similarity for individuals in different groups \citep[e.g.,][]{Aghabozorgi2015}.
+% There is, thus, a resolute need for more flexible and meaningful time series summaries that can capture diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
 
-However, while clustering can be incredibly useful, arriving at these clusters critically depends on two core challenges.
-First, time series need to be made comparable in order to identify key (dis)similarities and second, comparable (dis)similarities need to be accurately distinguishing into different groups \citep[e.g.,][]{Aghabozorgi2015}.
-In practice, most psychological time series cannot be compared based on the raw data itself.
-This is the case because in most cases the raw time series include too many data points --- sometimes referred to as the dimensionality curse \citep[e.g.,][]{altman2018} --- and, more importantly, individual time points are oftentimes not directly comparable between participants in psychological data and would lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
-While such issues can be avoided with transformations for highly regular, controlled, and comparable time series such as EEG data \citep[e.g.,][]{huang1985}, most ESM researchers are usually not interested in directly comparing individual timepoints between participants but are interested in developmental patterns and relationships.
+%Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models aimed at addressing either accuracy or interpretability concerns.
+Accuracy concerns have particularly been addressed by building more advanced models that relax specific assumptions \citep[e.g.,][]{denteuling2021} or by switching to different types of models with fewer assumptions, including spline models \citep{axen2011}.
+To expand interpretability, researchers have proposed to model additional types of parameters, including variability or granularity parameters \citep[e.g., see][]{krone2018}.
+%\citet{krone2018} even built a parametric model to tentatively cluster study participants.
 
-\begin{figure}[!hbtp] %hbtp
-  \caption{Time Series Clustering Taxonomy}
-  \label{fig:tsClustTax}
-  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustTax.pdf}
-  \caption*{Note: \\
-  The taxonomy only exemplifies some of the basic differences between a number of common time series clustering approaches.
-As such, the taxonomy and the notes are neither exhaustive nor complete in distinguishing different approaches.
-Additionally, terms and labels are used inconsistently across different types of literature and are chosen to avoid overlapping labels.}
-\end{figure}
+Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models, which either seek to relax specific assumptions \citep[e.g.,][]{denteuling2021} or include additional time series features \citep[e.g., see][]{krone2018}.
+In this manuscript, we instead, propose to cluster psychological time series based on important and relevant features directly.
+Such an approach is a common procedure in digital phenotyping \citep[][]{REF} and the broader machine learning literature \citep[][]{maharaj2019}.
+Adequately named \textit{feature-based time series clustering}, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
 
-As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{timmerman2013}; see \fgrref{fig:tsClustTax}).
-Ideally, the representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[][]{vandermaaten2009}.
-Extracting accurate and meaningful representations of the time series can be essential for understanding what goes into the clustering algorithm (i.e., assists with explainability) and can be crucial in making sense of the final cluster output \citep[i.e., assists with interpretability; e.g.,][]{Kennedy2021}.
+In this paper, we provide a practical introduction to the method and illustrate its utility with real-world ESM data.
+We argue that feature-based clustering offers a flexible variety, less strict assumptions, easy and intuitive extraction, theory-based selection, and meaningful psychological interpretability.
+Below, we introduce the empirical data we use for the illustration and then delineate the individual analysis steps.
 
-Thus far, one of the most common representation approaches for clustering of ESM data has been to extract person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
-While model parameters are familiar to researchers in the field, the model-based approach has a number of important shortcomings that relate to accuracy and interpretability.
-In terms of accuracy, the validity of model parameters is restricted by the assumptions of the statistical model.
-To take the common VAR model as an example, the model does not accurately capture data with unequal time intervals between measurements (i.e., nonequidistant measurements), it does not allow for missing values, and it does not allow for means or variances to change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
-As for the interpretability of the model parameters, purely data-driven model parameters, such as frequencies from a fourier analysis, are often difficult to interpret conceptually \citep[e.g.,][]{mayor2022}.
-But even more `meaningful' parameters from models used in theory-testing practices, such as VAR parameters, are commonly restricted to lagged partial correlations \citep[e.g.,][]{bringmann2018c}.
-Model parameters might, thus, not always accurately capture the original time series and are often restricted in their conceptual interpretation.
+\section{Data Used for Illustration}
 
-%Computationally, the number of parameters compounds with increases in variables and lag orders, which tends to result in efficiency and accuracy reduction for most clustering algorithms and is sometimes referred to as the dimensionality curse \citep{altman2018}\footnote{As an illustration of this scalability issue, consider a VAR model of 8 variables that includes the first three lags (i.e., VAR(3)).
-For each of the 8 variables, we would have 25 parameters --- one intercept and 24 slopes (8 variables * 3 lags each).
-That compounds to a total of 200 parameters per person (8 variables * 25 parameters).}.
-
-These shortcomings, however, stand in sharp contrast with the types of data researchers commonly collect and do not align common research interests in the field.
-Psychological researchers might, for example, collect data based on the occurrence of natural events, which tends to result in non-equidistant measurements \citep[e.g.,][]{myin-germeys2018, hamaker2017} and context-specific missingness that cannot be imputed, for instance, interaction quality perceptions \citep[e.g.,][]{kivela2022, lavori2008}.
-At the same time, researchers are often specifically interested in non-stationary developments, looking, for example, at how symptom levels and -variations change over time --- often abruptly and non-linearly \citep[e.g.,][; see \appref{app:ChallengesAppendix} for a more extensive discussion of the current challenges that clustering approaches should address]{bringmann2018b, helmich2020a}.
-There is, thus, a resolute need for more flexible and meaningful time series summaries that can capture diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
-
-% In short, the time series clustering repertoire should be extended to approaches that can more flexibly allow for diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.
-
-Recent efforts to address the shortcomings of model parameters for time series clustering, have primarily proposed more complicated models aimed at addressing either accuracy or interpretatbility concerns.
-Accuracy concerns have particularly been addressed by building more advanced models that relax specific assumptions \citep[e.g.,][]{denteuling2021} or by switching to different types of models with less assumptions, including spline models \citep{axen2011}.
-To expand interpretatbility, researchers have proposed to model additional types of parameters, including variability or granularity parameters \citep[e.g., see][]{krone2018}.
-
-In this manuscript we instead, propose to look at the broader machine learning literature and to consider a wider range of time series features to describe psychological time series.
-In particular, we propose to use an expanded feature-based clustering approach, which allows users to summarize time series based on a research-specific combination of parameters and descriptive statistics.
-We argue that feature-based clustering offers a flexible variety, less strict assumptions, easy and intuitive extraction, theory-based selection, and meaningful psychological interpretability.
-STRUCTURE SENTENCE: introduction of approach + empirical illustration
+\section{Analysis Steps and Application}
 
@@
 The main feature extraction step ultimately depends on the choice of the selected time series features but generally speaking the process is comparatively simple.
-For each of the included variables the features are extracted for each participant and replace the original time series \citep{wang2006}.
+For each of the included variables the features are extracted for each participant and replace the original time series \citep{wang2006}\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
+As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
 
@@
 %\input{tables/descrWide} 
-\input{tables/descrLong}
+%\input{tables/descrLong}
 
@@
 Time series features (such as means or linear trends) are not only easy to extract but also hold conceptual meaning for psychological data and can be chosen to address specific research questions (also see \tblref{tab:esmFeatures}).
+
+\begin{figure}[!hbtp] %hbtp
+  \caption{Time Series Clustering Taxonomy}
+  \label{fig:tsClustTax}
+  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustTax.pdf}
+  \caption*{Note: \\
+  The taxonomy only exemplifies some of the basic differences between a number of common time series clustering approaches.
+As such, the taxonomy and the notes are neither exhaustive nor complete in distinguishing different approaches.
+Additionally, terms and labels are used inconsistently across different types of literature and are chosen to avoid overlapping labels.}
+\end{figure}
 
@@
 
-\section{ESM Data Challenges}
+\section{ESM Data Challenges and Promises}
 \label{app:ChallengesAppendix}
+
+\subsection{Promises}
+
+Time series clustering has a number of conceptual use cases with psychological data.
+Prime among them is the ability to reduce the time, variable, and person complexity by extracting and organizing participant-level structures.
+These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
+These phenomena and principles can be meaningful differences that distinguish participants in different clusters, as well important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
+Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}.
+By detecting meaningful and robust structures and patterns, time series clustering can, thus, be used to inform the development of robust theories as well as targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions \citep[e.g.,][]{borsboom2021, eronen2020}.
+
+However, while clustering can be incredibly useful, arriving at these clusters critically depends on two core challenges.
+First, time series need to be made comparable in order to identify key (dis)similarities and second, comparable (dis)similarities need to be accurately distinguishing into different groups \citep[e.g.,][]{Aghabozorgi2015}.
+In practice, most psychological time series cannot be compared based on the raw data itself.
+This is the case because in most cases the raw time series include too many data points --- sometimes referred to as the dimensionality curse \citep[e.g.,][]{altman2018} --- and, more importantly, individual time points are oftentimes not directly comparable between participants in psychological data and would lead to misspecifications \citep[e.g., ][]{faloutsos1994}.
+While such issues can be avoided with transformations for highly regular, controlled, and comparable time series such as EEG data \citep[e.g.,][]{huang1985}, most ESM researchers are usually not interested in directly comparing individual timepoints between participants but are interested in developmental patterns and relationships.
+
+As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{timmerman2013}; see \fgrref{fig:tsClustTax}).
+Ideally, the representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[][]{vandermaaten2009}.
+Extracting accurate and meaningful representations of the time series can be essential for understanding what goes into the clustering algorithm (i.e., assists with explainability) and can be crucial in making sense of the final cluster output \citep[i.e., assists with interpretability; e.g.,][]{Kennedy2021}.
+
+\subsection{Challenges}
+
 We will briefly consider which challenges modern ESM data introduce and what qualities are called for in an extension of the clustering repertoire.
@@
 
+%These shortcomings, however, stand in sharp contrast with the types of data researchers commonly collect and do not align with common research interests in the field.
+Psychological researchers might, for example, collect data based on the occurrence of natural events, which tends to result in non-equidistant measurements \citep[e.g.,][]{myin-germeys2018, hamaker2017} and context-specific missingness that cannot be imputed, for instance, interaction quality perceptions \citep[e.g.,][]{kivela2022, lavori2008}.
+At the same time, researchers are often specifically interested in non-stationary developments, looking, for example, at how symptom levels and -variations change over time --- often abruptly and non-linearly \citep[e.g.,][]{bringmann2018b, helmich2020a}.
+
 \end{document}

### 2023-01-30 manuscript.tex
:shit: intro re-str

@@
 
-\begin{mdframed}[style=mdfbox]
-\noindent\center\textit{Structure}:
-\begin{itemize}[nosep]
-    \item introduction
-        \begin{itemize}[nosep]
-            \item increase in number and variety of ESM data + ts clustering to understand complex developmental differences.
-            \item ESM literature: features to capture meaningful aspects
-            \item current clustering based on common model parameters, which only include a small selection of features and often have strong assumptions
-            \item Instead of building more complicated models (e.g., address violated assumptions or add different types of parameters), we propose to cluster based on important and relevant features directly.
-            \item Called feature-based clustering. Common procedure in digital phenotyping and the broader ml literature. Used in many fields
-            \item in this paper we provide a practical introduction to the method and illustrate its utility with real-world ESM data.
-        \end{itemize}
-    \item Data Used for Illustration
-    \item Analysis Steps + Application
-        \begin{itemize}[nosep]
-            \item input variables
-            \item feature extraction
-            \item (feature reduction)
-            \item feature clustering
-            \item cluster evaluation
-        \end{itemize}
-    \item Discussion (summary, limitation, implications, conclusion, etc.)
-    \newpage
-    \item\ [...]
-    \item Feature-based Time Series Clustering
-    \begin{itemize}[nosep]
-        \item why features: flexible similarity (b/c more variety), less assumptions, easy to calculate/extract, theory-based selection (as per RQ), psychological interpretability, ...
-        \item Features in Psychological Time Series
-        \begin{itemize}[nosep]
-            \item central tendency
-            \item variability
-            \item (in)stability
-            \item self-similarity
-            \item linear trend
-            \item nonlinearity
-        \end{itemize}
-        \item Analysis Steps
-        \begin{itemize}[nosep]
-            \item input variables
-            \item feature extraction
-            \item (feature reduction)
-            \item feature clustering
-            \item cluster evaluation
-        \end{itemize}
-    \end{itemize}
-    \item Empirical Illustration
-    \item Discussion (summary, limitation, implications, conclusion, etc.)
-\end{itemize}
-\vspace{1em}
-\end{mdframed}
-\newpage
+% \begin{mdframed}[style=mdfbox]
+% \noindent\center\textit{Structure}:
+% \begin{itemize}[nosep]
+%     \item introduction
+%         \begin{itemize}[nosep]
+%             \item increase in number and variety of ESM data + ts clustering to understand complex developmental differences.
+%             \item ESM literature: features to capture meaningful aspects
+%             \item current clustering based on common model parameters, which only include a small selection of features and often have strong assumptions
+%             \item Instead of building more complicated models (e.g., address violated assumptions or add different types of parameters), we propose to cluster based on important and relevant features directly.
+%             \item Called feature-based clustering. Common procedure in digital phenotyping and the broader ml literature. Used in many fields
+%             \item in this paper we provide a practical introduction to the method and illustrate its utility with real-world ESM data.
+%         \end{itemize}
+%     \item Data Used for Illustration
+%     \item Analysis Steps + Application
+%         \begin{itemize}[nosep]
+%             \item input variables
+%             \item feature extraction
+%             \item (feature reduction)
+%             \item feature clustering
+%             \item cluster evaluation
+%         \end{itemize}
+%     \item Discussion (summary, limitation, implications, conclusion, etc.)
+%     \newpage
+%     \item\ [...]
+%     \item Feature-based Time Series Clustering
+%     \begin{itemize}[nosep]
+%         \item why features: flexible similarity (b/c more variety), less assumptions, easy to calculate/extract, theory-based selection (as per RQ), psychological interpretability, ...
+%         \item Features in Psychological Time Series
+%         \begin{itemize}[nosep]
+%             \item central tendency
+%             \item variability
+%             \item (in)stability
+%             \item self-similarity
+%             \item linear trend
+%             \item nonlinearity
+%         \end{itemize}
+%         \item Analysis Steps
+%         \begin{itemize}[nosep]
+%             \item input variables
+%             \item feature extraction
+%             \item (feature reduction)
+%             \item feature clustering
+%             \item cluster evaluation
+%         \end{itemize}
+%     \end{itemize}
+%     \item Empirical Illustration
+%     \item Discussion (summary, limitation, implications, conclusion, etc.)
+% \end{itemize}
+% \vspace{1em}
+% \end{mdframed}
+% \newpage
 
@@
 Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
-Researchers and practitioners are, for example, asking: ``Do patients differ in their development of several key symptoms?'' \citep{Monden2015} or ``How do migrants differ in the development of their psychological needs as they arrive in the new environment?'' \citep{Kreienkamp2022d}.
-Recently, one promising way of identifying developmental patterns that differ between participants has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
+Researchers and practitioners are, for example, asking: ``Do patients differ in their development of several key symptoms?'' \citep{Monden2015} or ``How do migrants differ in the development of their core psychological needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
+Recently, one promising way of identifying person-level developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
 Such clustering, however, depends on comparing participants based on meaningful (dis)similarities in their time series \citep[][]{Aghabozorgi2015}.
@@
 Fortunately, past conceptual and empirical works in the ESM literature have collected a number of measures that capture meaningful aspects of psychological time series.
-Originally, \citet{kuppens2017} proposed four features to capture psychological dynamics: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
+One commonly used set of features comes from \citet{kuppens2017}, who proposed four features to understand psychological dynamics: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
 These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
-From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}.
+From a slightly different perspective \citet{dejonckheere2019} later added three additional features of (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}.
+Each of these statistics not only captures a distinct aspect of psychological time series, but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.
 
 Yet, despite this rich diversity of meaningful time series features in psychology, most clustering of ESM data has only focused on a small and restrictive selection of time series characteristics.
-Thus far, one of the most common approaches has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
-While such model parameters are familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c}.
+Thus far, the most common approaches has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
+While such model parameters are familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c} --- only two of the many potentially important time series features.
 Additionally, the validity of model parameters is traditionally restricted by the assumptions of the particular statistical model.
 To take the common VAR model as an example, the model explicitly assumes that the time intervals between measurements are consistent (i.e., equidistant measurement assumption), it does not allow for missing values, and it assumes that means or variances do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
-All of these are types of data researchers commonly collect and contradict important researcher interests in the field \citep[][; see \appref{app:ChallengesAppendix} for a more extensive discussion of the current challenges that clustering approaches should address]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
-Model parameters might, thus, not always accurately capture the original time series and are often restricted in the features they capture.
+All of these violated assumptions reflect types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
+Model parameters might, thus, not always accurately capture the time series and are often restricted in the features they capture.
 
@@
 Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models, which either seek to relax specific assumptions \citep[e.g.,][]{denteuling2021} or include additional time series features \citep[e.g., see][]{krone2018}.
-In this manuscript, we instead, propose to cluster psychological time series based on important and relevant features directly.
-Such an approach is a common procedure in digital phenotyping \citep[][]{REF} and the broader machine learning literature \citep[][]{maharaj2019}.
-Adequately named \textit{feature-based time series clustering}, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+In this manuscript, we instead, propose to directly cluster based on important and relevant features.
+Adequately named \textit{feature-based time series clustering}, such an approach has been a common procedure in digital phenotyping \citep[][]{loftus2022} and the broader machine learning literature \citep[][]{maharaj2019}.
+As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+We argue that for psychological time series data, feature-based clustering offers a flexible variety, theory-based selection, fewer strict assumptions, easy and intuitive analysis, and meaningful psychological interpretability.
 
-In this paper, we provide a practical introduction to the method and illustrate its utility with real-world ESM data.
-We argue that feature-based clustering offers a flexible variety, less strict assumptions, easy and intuitive extraction, theory-based selection, and meaningful psychological interpretability.
-Below, we introduce the empirical data we use for the illustration and then delineate the individual analysis steps.
+In the sections below we particularly seek to provide a practical introduction to the method.
+To do so, we illustrate the utility of the method with real-world ESM data.
+We use this data to discuss which psychological time series features are well-suited for a clustering approach, introduce the individual analysis steps, and provide practical guidance on common algorithms and analysis code.
+We first introduce the empirical case study data and then follow the individual analysis steps to discuss key decisions and approaches.
 
@@
 
+[...] \textit{move the introduction of the case study and data here} [...]
+
 \section{Analysis Steps and Application}
+
+[...] \textit{move the analysis steps and results here} [...]
 

### 2023-02-10 manuscript.tex
:construction: re-structuring

@@
 
-Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][]{hamaker2017}\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
+Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][]{hamaker2017}.
+These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
+Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
+Researchers and practitioners are, for example, asking: ``Do the key symptoms of different patients develop in contrasting ways?'' \citep{Monden2015} or ``How do migrants differ in the development of their core psychological needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
+There is, thus, a clear need for analysis techniques that identify person-level differences in developmental patterns for psychological data.
+
+Recently, one promising way of identifying person-level developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
+This analysis type essentially seeks to capture comparable within-person developments --- often with some form of numeric summary --- and then groups the persons based on the extracted summaries \citep[][]{liao2005}\footnote{It should be noted in some cases the time series do not need to be summarized and can be compared directly.
+Such analyses are however only possible with highly regular and controlled data, such as EEG data \citep{huang1985}, or when clustering variables within the person rather than identifying differences between persons \citep{haslbeck2022}.
+Multivariate data from extensive psychological survey studies, as we discuss here, are seldomly directly comparable across persons \citep[e.g., ][]{faloutsos1994}.
+We provide a broader embedding of the current methods for psychological time series in the discussion section.}.
+Time series clustering, thus, crucially depends on identifying meaningful summaries of the time series developments, which can be used to compare participants \citep[][]{Aghabozorgi2015}.
+
+Fortunately, past conceptual and empirical works in the experience sampling (ESM) literature have collected a number of meaningful aspects of psychological time series\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
 While the terms come for different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events.
 In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}.
-These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
-Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
-Researchers and practitioners are, for example, asking: ``Do patients differ in their development of several key symptoms?'' \citep{Monden2015} or ``How do migrants differ in the development of their core psychological needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
-Recently, one promising way of identifying person-level developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
-Such clustering, however, depends on comparing participants based on meaningful (dis)similarities in their time series \citep[][]{Aghabozorgi2015}.
-
-Fortunately, past conceptual and empirical works in the ESM literature have collected a number of measures that capture meaningful aspects of psychological time series.
-One commonly used set of features comes from \citet{kuppens2017}, who proposed four features to understand psychological dynamics: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
-These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
-From a slightly different perspective \citet{dejonckheere2019} later added three additional features of (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}.
+Important aspects might include concerns over whether a symptom consistently stays at a certain level without much variability or whether some emotions develop together.
+For many of the most important developmental aspects, researchers have assembled numeric measures that capture these patterns.
+These summary statistics are often called ``dynamic measures'', ``principles of change'', or ``dynamic features'' of the psychological time series \citep{dejonckheere2019, kuppens2017, krone2018}.
+For most conceptualizations, the proposed number of aspects ranges from four to twelve key features \citep[][]{wang2006, dejonckheere2019}.
 Each of these statistics not only captures a distinct aspect of psychological time series, but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.
@@
 Yet, despite this rich diversity of meaningful time series features in psychology, most clustering of ESM data has only focused on a small and restrictive selection of time series characteristics.
-Thus far, the most common approaches has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
+Thus far, the most common approach has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
 While such model parameters are familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c} --- only two of the many potentially important time series features.
@@
 
-[...] \textit{move the introduction of the case study and data here} [...]
+To illustrate the functioning and utility of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected data from research on migration experiences.
+Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020} and this type of research exemplifies some of the real-world data issues that ESM data commonly faces.
+The new type of research comes as a response to prominent reviews within the migration literature, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
+The migration ESM research, thus, offers a good example of the ESM data that introduces a multivariate conceptualization with event-specific missingness patterns.
+
+[...]
 
@@
 
-[...] \textit{move the analysis steps and results here} [...]
+Introduction on the structure ...
 
-\section{Feature-based Time Series Clustering}
-
-The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
-In its most general approach, this can include any statistical summary of the time series \citep[i.e., feature; e.g.,][]{maharaj2019}.
-The extracted features can then be used as inputs to most clustering algorithms, including k-means or hierarchical clustering to group the participants into clusters \citep{wang2006}.
-The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
-However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
-Feature-based time series clustering has been a valued tool in many empirical fields for its flexibility and interpretability.
-As a result, features have been extracted for a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-
-With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
-As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos \citep[also see][]{fulcher2013}.
-\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
-A final illustration comes from a commonly used software package for feature extraction `tsfresh', which allows users to extract a total of 794 features to describe all kinds of aspects of a time series \citep[][]{christ2018}.
-
-Given that there are hundreds of possible features to describe psychological time series, one approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series \citep[e.g.,][]{christ2018}.
-However, such an approach is not always advisable for psychological time series.
-For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 794 time series features.
-More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
-This is particularly the case when features have a conceptual or psychological meaning.
-Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
-But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
-We will, thus, propose and describe a set of time series features that have a strong backing within the ESM literature and as such offer a reasonable starting point for most concepts within the field.
-
-\subsection{Features of Psychological Time Series}
-
-Fortunately, past conceptual and empirical efforts offer valuable discussions of common time series features in psychological research.
-To understand emotion dynamics, \citet{kuppens2017} originally proposed four features: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
-These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
-\citet{krone2018} even built a parametric model to tentatively cluster study participants.
-From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
-As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
-
-\input{tables/esmPsychFeatures}
-
-Some of the proposed features are not necessarily well suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the emotion dynamics literature, which has been so formative for the analysis of psychological time series data.
-In particular, covariances and cross-lagged correlations often produce a staggering amount of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}.
-Others such as network density parameters, used to summarize variable interdependence, might not always be meaningful for psychological data \citep{bringmann2019a}.
-At the same time, linear and nonlinear trends are not captured by the features commonly proposed for psychological time series, because the features are often developed for stationary models \citep[e.g.,][]{krone2018}.
-
-Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, we hope to build a practical toolbox of meaningful time series features for psychological data.
-In particular, we propose to focus on six features based on common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
-We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-base clustering of psychological time series data.
-An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
-
-\paragraph{Central Tendency} Central tendency refers to the statistical measures that represent the ``middle" or ``average" of a set of data.
-The most common measures of central tendency are the mean, median, and mode (Hopkins, 2020).
-As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
-Researchers might, for example, be interested in whether ``Over a one month period, are some people happier than others?''.
-A recent review paper pointed out that the relatively simple mean statistic, is often overlooked in psychological time series analyses but is usually essential for understanding psychological differences \citep{bringmann2018c}.
-As such, several time series modeling approaches now advocate for conscious consideration of the mean \citep[e.g.,][]{bringmann2017a}.
-As with cross-sectional data, for time series data the appropriate measure of central tendency depends on the specific characteristics of the data being analyzed.
-For instance, if the data are normally distributed, the mean is often an appropriate measure.
-If the data are skewed or have outliers, the median may be a better choice \citep{weisberg1992}.
-
-\paragraph{Variability} Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
-In time series analyses, variability is conceptually important because it provides information about the distribution and diversity of the data.
-Especially within-person changes in variability have recently received increasing attention because they can be a potent early warning signal in psychological time series \citep{helmich2020a, vandeleemput2014}.
-A marked within-person increase in variance has, for example, been found to precede crucial transitions --- often considered to be part of a phenonenon know as critical slowing down \citep[i.e., anomalous variances; e.g.,][]{scheffer2009, wichers2019}.
-
-But also more global person-level differences in time series variability have been found to be crucial and relate to important psychological questions.
-Person-specific differences in variability have, for example, been found to be indicative of worse psychological states \citep{myin-germeys2018}.
-And, similarly, person-level differences have also been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
-As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
-Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
-
-Methodologically, there are several measures of variability, including the range, standard deviation, and variance.
-The appropriate measure to use depends on the specific characteristics of the data being analyzed.
-For normally distributed data variances and standard deviation are often preferred, whereas data with outliers or nonparametric distributions are often better described with robust variability measures such as the median absolute deviation \citep{weisberg1992}.
-
-\paragraph{(In)stability} Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009}.
-While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the `jumpy-ness' of the data over time.
-In other words, variability reflects the range or diversity of values in a time series data, while instability reflects the fluctuation or inconsistency in a time series data over time  \citep{trull2008}.
-For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
-Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
-Conceptually, the instability feature, thus, relates to a broad range of research questions, including: ``Do changes tend to be slow and gradual or fast and abrupt?" Methodologically, (in)stability measurements often look at the average change between two consecutive measurements, such as the mean absolute change \citep[e.g.,][]{ebner-priemer2009,barandas2020}.
-
-\paragraph{Self-similarity} Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}.
-That is, self-similarity describes how much a measurement carries over to future measurements.
-There are two main ways in which experiences tend to be connected to future experiences in psychological time series --- resistance to change and seasonality.
-Both forms of self-similarity relate to different types of research questions.
-
-Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{kuppens2010}.
-If inertia is high a development tends to stay in a certain state or continues to move towards a certain direction.
-Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
-In a similar vein, high inertia at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
-Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?'' One way to measure inertia in time series data is through the use of autocorrelations or autoregressions with a lag-1, which look at the correlation between a given time series and itself at the previous measurement occasion.
-High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}.
-
-Seasonality is the self-similarity of a measurement that returns in a cyclical manner and is often also referred to as periodicity \citep{gregson1983}.
-Periodicity in psychological time series data, thus, refers to the presence of recurring patterns at regular intervals.
-In psychological time series, these cyclical patterns are not only dangerous for models that assume stationarity \citep{beal2015} but can be interesting psychological properties in their own regard \citep{epskamp2018, schmittmann2013}.
-Humans generally live in seasonal environments, including, for example, exam periods at universities \citep{fuller2003}.
-But also more internally, personality manifestations, for example, have long been assumed to develop cyclically \citep{cattell1957}.
-The associated research questions can be driven by an understanding of external seasonality factors such as: ``Do participants drink more alcohol on Fridays?'' But research questions can also address periodic patterns as they may be the result of underlying psychological processes: ``Does stress increase with fatigue cycles?'' Similar to inertia, periodicity is often measured using autocorrelations but then with the lag of the seasonality or cycle (e.g., daily, seven days, one month, one semester).
-Periodicity can additionally be captured using spectral analysis or decomposition, such as with wave patterns \citep[e.g., fourier or wavelet transformation; e.g.,][]{mayor2022}.
-
-\paragraph{Linear Trend} Linear trends in time series data refer to the presence of a straight-line development of a variable being measured.
-This type of trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
-For psychological time series, researchers have, for example, pointed out linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}.
-Linear trends are often the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}.
-In empirical practice, linear trends are, thus, a common research interest for psychological time series data.
-``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions.
-Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients, but also using differential equations \citep{chow2019}.
-
-\paragraph{Nonlinearity} Changes in psychology are not always linear, instead nonlinearity is is a common feature of psychological time series \citep{hayes2007}.
-As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
-Similarly, patients in recovery of depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
-But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
-For psychological researchers, nonlinear developments are, thus, important in two regards.
-Firstly, nonlinearity as a deviation from linear trends is an important feature defining many psychological time series and secondly, the shape and structure of the nonlinear trend can be essential in understanding the important changes and transitions of human experiences \citep{chow2019}.
-Conceptually, researchers might accordingly have research question about the type of the development: ``Is the development of anxiety a nonlinear process?'' as well as the shape and structure of the development: ``Does anxiety change as a sudden or smooth transition?'' Accordingly, researchers can capture the type of development with nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.
-The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
-Common ways to capture the trends include the use of polynomial- and differential equation parameters or more broadly to capture how `wiggly' the trend line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
-
-It should be noted that this selection of the proposed six time-series features is in no way exhaustive or imperative.
-Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
-
-\subsection{Analysis Steps}
 Armed with a selection of potentially meaningful and accurate time series features to address a given research question, we can consider the practical steps of feature-based time series clustering.
@@
 As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
+
+The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
+In its most general approach, this can include any statistical summary of the time series \citep[i.e., feature; e.g.,][]{maharaj2019}.
+The extracted features can then be used as inputs to most clustering algorithms, including k-means or hierarchical clustering to group the participants into clusters \citep{wang2006}.
+The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
+However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
+Feature-based time series clustering has been a valued tool in many empirical fields for its flexibility and interpretability.
+As a result, features have been extracted for a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
+
+With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
+As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos \citep[also see][]{fulcher2013}.
+\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
+A final illustration comes from a commonly used software package for feature extraction `tsfresh', which allows users to extract a total of 794 features to describe all kinds of aspects of a time series \citep[][]{christ2018}.
+
+Given that there are hundreds of possible features to describe psychological time series, one approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series \citep[e.g.,][]{christ2018}.
+However, such an approach is not always advisable for psychological time series.
+For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 794 time series features.
+More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
+This is particularly the case when features have a conceptual or psychological meaning.
+Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
+But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
+We will, thus, propose and describe a set of time series features that have a strong backing within the ESM literature and as such offer a reasonable starting point for most concepts within the field.
+
+Fortunately, past conceptual and empirical efforts offer valuable discussions of common time series features in psychological research.
+To understand emotion dynamics, \citet{kuppens2017} originally proposed four features: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
+These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
+\citet{krone2018} even built a parametric model to tentatively cluster study participants.
+From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
+As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
+
+\input{tables/esmPsychFeatures}
+
+Some of the proposed features are not necessarily well suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the emotion dynamics literature, which has been so formative for the analysis of psychological time series data.
+In particular, covariances and cross-lagged correlations often produce a staggering amount of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}.
+Others such as network density parameters, used to summarize variable interdependence, might not always be meaningful for psychological data \citep{bringmann2019a}.
+At the same time, linear and nonlinear trends are not captured by the features commonly proposed for psychological time series, because the features are often developed for stationary models \citep[e.g.,][]{krone2018}.
+
+Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, we hope to build a practical toolbox of meaningful time series features for psychological data.
+In particular, we propose to focus on six features based on common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
+We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-base clustering of psychological time series data.
+An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
+
+It should be noted that this selection of the proposed six time-series features is in no way exhaustive or imperative.
+Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
 
@@
 Most of their recommendations and reporting guidelines apply to feature-based clustering as well and might even offer a template for researchers hoping to pre-register their analysis procedures.
-
-\section{Empirical Illustration}
-To illustrate common use cases of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected diverse concepts with context-specific measurements.
-The illustration specifically uses data from research on migration experiences, where researchers have started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020}.
-This new type of research comes as a response to prominent reviews, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}.
-At the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-The migration ESM research, thus, offers a good example of the ESM data that introduces diverse conceptualizations and more event-specific missingness patterns.
 
@@
 
+\section{Background of the Proposed Psychological Time Series Features}
+\label{app:FeatureBackground}
+
+\paragraph{Central Tendency} Central tendency refers to the statistical measures that represent the ``middle" or ``average" of a set of data.
+The most common measures of central tendency are the mean, median, and mode (Hopkins, 2020).
+As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
+Researchers might, for example, be interested in whether ``Over a one month period, are some people happier than others?''.
+A recent review paper pointed out that the relatively simple mean statistic, is often overlooked in psychological time series analyses but is usually essential for understanding psychological differences \citep{bringmann2018c}.
+As such, several time series modeling approaches now advocate for conscious consideration of the mean \citep[e.g.,][]{bringmann2017a}.
+As with cross-sectional data, for time series data the appropriate measure of central tendency depends on the specific characteristics of the data being analyzed.
+For instance, if the data are normally distributed, the mean is often an appropriate measure.
+If the data are skewed or have outliers, the median may be a better choice \citep{weisberg1992}.
+
+\paragraph{Variability} Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
+In time series analyses, variability is conceptually important because it provides information about the distribution and diversity of the data.
+Especially within-person changes in variability have recently received increasing attention because they can be a potent early warning signal in psychological time series \citep{helmich2020a, vandeleemput2014}.
+A marked within-person increase in variance has, for example, been found to precede crucial transitions --- often considered to be part of a phenonenon know as critical slowing down \citep[i.e., anomalous variances; e.g.,][]{scheffer2009, wichers2019}.
+
+But also more global person-level differences in time series variability have been found to be crucial and relate to important psychological questions.
+Person-specific differences in variability have, for example, been found to be indicative of worse psychological states \citep{myin-germeys2018}.
+And, similarly, person-level differences have also been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
+As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
+Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
+
+Methodologically, there are several measures of variability, including the range, standard deviation, and variance.
+The appropriate measure to use depends on the specific characteristics of the data being analyzed.
+For normally distributed data variances and standard deviation are often preferred, whereas data with outliers or nonparametric distributions are often better described with robust variability measures such as the median absolute deviation \citep{weisberg1992}.
+
+\paragraph{(In)stability} Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009}.
+While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the `jumpy-ness' of the data over time.
+In other words, variability reflects the range or diversity of values in a time series data, while instability reflects the fluctuation or inconsistency in a time series data over time  \citep{trull2008}.
+For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
+Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
+Conceptually, the instability feature, thus, relates to a broad range of research questions, including: ``Do changes tend to be slow and gradual or fast and abrupt?" Methodologically, (in)stability measurements often look at the average change between two consecutive measurements, such as the mean absolute change \citep[e.g.,][]{ebner-priemer2009,barandas2020}.
+
+\paragraph{Self-similarity} Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}.
+That is, self-similarity describes how much a measurement carries over to future measurements.
+There are two main ways in which experiences tend to be connected to future experiences in psychological time series --- resistance to change and seasonality.
+Both forms of self-similarity relate to different types of research questions.
+
+Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{kuppens2010}.
+If inertia is high a development tends to stay in a certain state or continues to move towards a certain direction.
+Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
+In a similar vein, high inertia at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
+Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?'' One way to measure inertia in time series data is through the use of autocorrelations or autoregressions with a lag-1, which look at the correlation between a given time series and itself at the previous measurement occasion.
+High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}.
+
+Seasonality is the self-similarity of a measurement that returns in a cyclical manner and is often also referred to as periodicity \citep{gregson1983}.
+Periodicity in psychological time series data, thus, refers to the presence of recurring patterns at regular intervals.
+In psychological time series, these cyclical patterns are not only dangerous for models that assume stationarity \citep{beal2015} but can be interesting psychological properties in their own regard \citep{epskamp2018, schmittmann2013}.
+Humans generally live in seasonal environments, including, for example, exam periods at universities \citep{fuller2003}.
+But also more internally, personality manifestations, for example, have long been assumed to develop cyclically \citep{cattell1957}.
+The associated research questions can be driven by an understanding of external seasonality factors such as: ``Do participants drink more alcohol on Fridays?'' But research questions can also address periodic patterns as they may be the result of underlying psychological processes: ``Does stress increase with fatigue cycles?'' Similar to inertia, periodicity is often measured using autocorrelations but then with the lag of the seasonality or cycle (e.g., daily, seven days, one month, one semester).
+Periodicity can additionally be captured using spectral analysis or decomposition, such as with wave patterns \citep[e.g., fourier or wavelet transformation; e.g.,][]{mayor2022}.
+
+\paragraph{Linear Trend} Linear trends in time series data refer to the presence of a straight-line development of a variable being measured.
+This type of trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
+For psychological time series, researchers have, for example, pointed out linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}.
+Linear trends are often the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}.
+In empirical practice, linear trends are, thus, a common research interest for psychological time series data.
+``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions.
+Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients, but also using differential equations \citep{chow2019}.
+
+\paragraph{Nonlinearity} Changes in psychology are not always linear, instead nonlinearity is is a common feature of psychological time series \citep{hayes2007}.
+As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
+Similarly, patients in recovery of depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
+But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
+For psychological researchers, nonlinear developments are, thus, important in two regards.
+Firstly, nonlinearity as a deviation from linear trends is an important feature defining many psychological time series and secondly, the shape and structure of the nonlinear trend can be essential in understanding the important changes and transitions of human experiences \citep{chow2019}.
+Conceptually, researchers might accordingly have research question about the type of the development: ``Is the development of anxiety a nonlinear process?'' as well as the shape and structure of the development: ``Does anxiety change as a sudden or smooth transition?'' Accordingly, researchers can capture the type of development with nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.
+The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
+Common ways to capture the trends include the use of polynomial- and differential equation parameters or more broadly to capture how `wiggly' the trend line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
+
 \end{document}

### 2023-02-17 IllustrationResultsRenderReport.Rmd
:construction: section-based re-render

--- 2022-11-02 Supplemental-Material-A-Annotated-Analysis.Rmd
@@
 ---
-title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
-subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
+title: "Report: Illustration Results"
+subtitle: "Render report for 'A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series'"
 author:
-- Jannis Kreienkamp^1,3^
-- Kai Epstude^1,3^
-- Laura F. Bringmann^1,3^
-- Maximilian Agostini^1,3^
-- Peter de Jonge^1,3^
-- Rei Monden^2,3^
-- ^1^University of Groningen, Department of Psychology
-- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
-- ^3^Author order TBD [currently in first name alphabetical order]
+- ██████ ██████████^1^
+- ██████████ ████████^1^
+- █████ █████████^1^
+- █████ ██ █████^1^
+- ███ ███████^1^
+- ^1^██████████ ██ █████████, Department of Psychology
 - "Author Information:"
-- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
-E-mail: j.kreienkamp@rug.nl"
+- "Correspondence concerning this article should be addressed to ██████ ██████████, ██████████ ██ ██████████, █████████ ██ ██████████, █████ ██████████ ███, ████ ██ █████████ ████████████████.
+E-mail: █████████████████"
 - 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
-- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
+- 'The data repository for this manuscript is available at <a href="https://osf.io/" target="_blank">osf.io/TBD</a>'
 - 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
@@
 
-```{=html}
-<style type="text/css">
-.main-container {
-  max-width: 1300px;
-  margin-left: auto;
-  margin-right: auto;
-}
-.table {
-  margin-left:auto; 
-  margin-right:auto;
-}
-</style>
-```
-
-```{r setup, include=FALSE}
+```{r}
+#| label: setup
+#| include: false
+
 # R Studio Clean-Up
@@
 )
-```
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-<i class="fas fa-exclamation-circle"></i> Note.
-Boxplots display the interquartile range (IQR, center box), and the whiskers extend 1.5\*IQR from the lower and upper hinge.
-The white point indicates the mean and the white center line indicates the median.
-
-</br>
-
-------------------------------------------------------------------------
-
-</br>
-
-# **Data Preparation**
-
-In an initial preparatory step, we import the data into the R project environment and prepare the variables for further processing and later analyses.
-
-## Data Import
-
-The data were collected using two different survey tools.
-For the study with sojourners (Study 1: worker) we used the survey platform Qualtrics XM, whereas the studies with international students (Study 2: student), and the international medical professionals (Study 3: medical) were conducted using the survey framework FormR.
-This means that the datasets had inconsistent file formats and naming conventions.
-For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*').
-For the two other studies, we import the raw csv files from their respective folders.
-
-```{r formrImport}
+
+outputDir <- "latex-results"
+```
+
+```{r}
+#| label: import data
+#| include: true
+
 # Load variable name lookup table
@@
 
-# rename function
-reNam <- function(data = NA, names = varNames, study = "S1", survey = "pre") {
-  # for testing
-  # data = dtS1$raw.pre
-  # names = varNames
-  # study = "S1"
-  # survey = "pre"
-
-  var <- paste0("survey", study)
-  varNamOld <- paste0("varNam", study)
-
-  nameTbl <- names %>%
-    filter((!!sym(var)) == survey) %>%
-    select(varNam, varNamOld = (!!sym(varNamOld)))
-
-  data.table::setnames(data, as.character(nameTbl$varNamOld), as.character(nameTbl$varNam), skip_absent=TRUE)
-
-  return(data)
-}
-
-# workers
-# initial data cleaning was done in SPSS (syntax files are available in "")
-dtS1 <- list(
-  raw.pre = read_spss("data/S1_Workers/processed/cleaned/MT - Pre-Measure - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "pre"),
-  raw.post = read_spss("data/S1_Workers/processed/cleaned/MT - Post-Measure - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "post"),
-  raw.morning = read_spss("data/S1_Workers/processed/cleaned/MT - Morning - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "daily"),
-  raw.afternoon = read_spss("data/S1_Workers/processed/cleaned/MT - Afternoon - 06-15-2018.sav") %>% 
-    reNam(data = ., names = varNames, study = "S1", survey = "daily")
-)
-
-# students
-dtS2 <- list(
-  raw.pre = read.csv(file = "data/S2_Students/raw/AOTS_Pre.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "pre"),
-  raw.post = read.csv(file = "data/S2_Students/raw/AOTS_Post.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "post"),
-  raw.daily = read.csv(file = "data/S2_Students/raw/AOTS_Daily.csv", header = T, sep = ",") %>% 
-    reNam(data = ., names = varNames, study = "S2", survey = "daily")
-)
-
-# young medical professionals
-dtS3 <- list(
-  raw.eligibility = read.csv("data/S3_Medical/raw/AOTM_Eligibility.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"), # works but should be 'survey = "pre_elig"'
-  raw.pre = read.csv("data/S3_Medical/raw/AOTM_Pre.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "pre_entry"),
-  raw.post = read.csv("data/S3_Medical/raw/AOTM_Post.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "post"),
-  raw.daily = read.csv("data/S3_Medical/raw/AOTM_Daily.csv") %>% 
-    reNam(data = ., names = varNames, study = "S3", survey = "daily")
-)
-
-# # FROM INITIAL VAR NAME EXPORT [LEGACY]
-# # export var names
-# S1Nam <- rbind(
-#   data.frame(
-#     study = "S1",
-#     survey = "pre",
-#     varNam = names(dtS1$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S1",
-#     survey = "daily",
-#     varNam = names(dtS1$raw.afternoon)
-#   ),
-#   data.frame(
-#     study = "S1",
-#     survey = "post",
-#     varNam = names(dtS1$raw.post)
-#   )
-# )
-# S2Nam <- rbind(
-#   data.frame(
-#     study = "S2",
-#     survey = "pre",
-#     varNam = names(dtS2$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S2",
-#     survey = "daily",
-#     varNam = names(dtS2$raw.daily)
-#   ),
-#   data.frame(
-#     study = "S2",
-#     survey = "post",
-#     varNam = names(dtS2$raw.post)
-#   )
-# )
-# S3Nam <- rbind(
-#   data.frame(
-#     study = "S3",
-#     survey = "pre_elig",
-#     varNam = names(dtS3$raw.eligibility)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "pre_entry",
-#     varNam = names(dtS3$raw.pre)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "daily",
-#     varNam = names(dtS3$raw.daily)
-#   ),
-#   data.frame(
-#     study = "S3",
-#     survey = "post",
-#     varNam = names(dtS3$raw.post)
-#   )
-# )
-# write.csv(
-#   x = rbind(
-#     S1Nam,
-#     S2Nam,
-#     S3Nam
-#   ),
-#   file = "preregistration/varNames.csv"
-# )
-```
-
-## Data Cleaning & Data Exclusions
-
-### Study 1
-
-For the sojourner sample data was collected in four separate surveys: (1) the pre-measurement, (2) the daily morning survey, (3) the daily afternoon survey, as well as (4) a post-measurement.
-We combine the four individual surveys into one cohesive dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We then format the time and date variables and add person- and measurement indices (for easy and meaningful addressing of the data).
-We also exclude our own test data.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS1*' list.
-
-```{r cleanWorker}
-# Create reduced data sets for morning and afternoon
-dat.mo <- dtS1$raw.morning %>%
-  select(-starts_with("t_"))
-#setdiff(names(dtS1$raw.morning), names(dat.mo))
-dat.mo$daytime <- "morning"
-
-dat.af <- dtS1$raw.afternoon %>%
-  select(-starts_with("t_"))
-dat.af$daytime <- "afternoon"
-
-# merge morning and afternoon measurements with indicator [+ clean up]
-daily.dat <- plyr::rbind.fill(dat.mo, dat.af)
-daily.dat <- daily.dat[daily.dat$last_outside_referrer != 55951, ]
-dtS1$daily <- daily.dat
-rm(dat.mo, dat.af, daily.dat)
-
-# reduced data set for pre measurement
-dat.pre.red <- dtS1$raw.pre %>%
-  select(-starts_with("t_"))
-names(dat.pre.red) <- paste(names(dat.pre.red), "pre", sep = ".")
-
-# merge with daily data [+ clean up]
-df.pre <- merge(
-  x = dtS1$daily,
-  y = dat.pre.red,
-  by.x = "last_outside_referrer",
-  by.y = "platformId.pre",
-  all = T
-)
-
-# adjust duplicate names to fit to indicate daily or pre measurement
-names(df.pre) <- gsub("[[:punct:]]x", ".daily", names(df.pre))
-names(df.pre) <- gsub("[[:punct:]]y", ".pre", names(df.pre))
-
-# reduced data set for post-measurement
-dat.post.red <- dtS1$raw.post %>%
-  select(-starts_with("t_"))
-names(dat.post.red) <- paste(names(dat.post.red), "post", sep = ".")
-
-# merge post measurement with pre- and daily data
-df <- merge(
-  x = df.pre,
-  y = dat.post.red,
-  by.x = "last_outside_referrer",
-  by.y = "last_outside_referrer.post",
-  all = T
-)
-
-# adjust duplicate names to indicate pre or post
-names(df) <- gsub("[[:punct:]]x", ".pre", names(df))
-names(df) <- gsub("[[:punct:]]y", ".post", names(df))
-
-# add to list
-dtS1$combined <- df
-rm(df, df.pre, dat.post.red, dat.pre.red)
-
-# create data frame with cleaned data
-df <- dtS1$combined %>%
-  filter(
-    Finished.pre == 1,
-    Finished == 1,
-    !is.na(last_outside_referrer)
-  )
-
-# add running number as measurement ID within participants
-df$measureID <- data.table::rowidv(df, cols = c("last_outside_referrer"))
-
-dtS1$clean <- df
-
-# clean up
-rm(df)
-
-# Export reduced Data
-# write.csv(dtS1$clean, "data/processed/MT_clean-merged_07-05-2018.csv", row.names = F)
-# save(dtS1$clean, file = "data/processed/MT_clean-merged_07-05-2018.RData")
-```
-
-### Study 2
-
-For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
-We exclude our own test data as well as one participant who entered the study twice (but gave different responses during the pre-measurement).
-We also reformat missing values and format core ID variables.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS2*' list.
-
-```{r cleanStudents}
-# our own test IDs
-ownIDs <- c(
-  "beautifulLionfishXXXR5rcgVBzGu8hPvOqrK8UBJBw4owvi9nfRFSFu3lMzYhE",
-  "niceDogoXXXmB8JI5SFu78SF3DVof84mGUPPNUr14p2HYFTtp31a6D1OwAzM6F-K",
-  "amusedQuailXXXmhuc_fpTp8vPkMwDH1BzjaH1d1kHSO1bsPEfsnaEYk4WeVBfPi",
-  "juwGAbtXX0_1kmZtSVqKh3PGaHOICqUyU4iBkrT3nDsI_uifuD1gzKcZerxaM5FL"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS2$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!email %in% .$email[duplicated(.$email)]) %>% # remove all who did the pre questionnaire multiple times (b/c inconsistent ratings scales)
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS2$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS2$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  filter(!is.na(ended)) %>% # remove all who never finished
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = FALSE
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = FALSE
-)
-
-# add to list
-dtS2$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-### Study 3
-
-For the medical professionals sample data was, again, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
-We combine the three individual surveys into one large dataframe.
-We exclude our own test data.
-We also reformat missing values and format core ID variables.\
-<i class="fas fa-info-circle"></i> *Note:* All data preparation steps are saved in the '*dtS3*' list.
-
-```{r cleanMedical}
-# our own test IDs
-ownIDs <- c(
-  "test_LeonieXXXSklxecPLW0-FBPM4796o3pUwUhAY5jb9KGw8jQsKxWmGpa1Jiy", 
-  "test_MaxXXXtOp_5dTNefIq0yKXtXt2IN6eEKxeHoPY9mlyvdsqPpLp1B0NGg4UL",
-  "test_JannisXXXBsNqk62fOpX6chbd2tMWPptUdjjnhAqnQ3uBqckZ7gLIEoPlfZ",
-  "quaintLeopardCatXXXAJ9cfSj-_SZLwNwMDxv_xv_iyr1Bg5YFLTlYdrjW0UXZY",
-  "blue-eyedIndianElephantXXXLf5zPMpQCDGS3umFzIj-YVky7ivTItvvozW49m"
-)
-
-# Prepare dfs for Cleaning
-df.pre <- dtS3$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!is.na(ended)) %>% # remove all who did not finish
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.pre) <- paste(names(df.pre), "pre", sep = ".")
-
-df.post <- dtS3$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>% 
-  filter(!is.na(ResponseId)) %>% # remove own test runs
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  filter(!ResponseId %in% .$ResponseId[duplicated(.$ResponseId)]) %>% # remove all duplicate ResponseIds
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-names(df.post) <- paste(names(df.post), "post", sep = ".")
-
-df.daily <- dtS3$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
-  filter(!ResponseId %in% ownIDs) %>% # remove our own test
-  filter(ResponseId %in% df.pre$ResponseId) %>% # remove anyone who wasn't in the pre
-  #filter(!is.na(ended)) %>% # remove all who never finished [disabled because only relevant if data is missing]
-  mutate(ResponseId = as.character(ResponseId)) # turn factor into character strings (probably just precaution)
-
-# merge daily with pre
-dfPreDaily <- merge(
-  x = df.daily,
-  y = df.pre,
-  by.x = "ResponseId",
-  by.y = "ResponseId.pre", 
-  suffixes = c(".daily", ".pre"),
-  all = F
-)
-
-# merge daily with post
-dfCombined <- merge(
-  x = dfPreDaily,
-  y = df.post,
-  by.x = "ResponseId",
-  by.y = "ResponseId.post", 
-  suffixes = c(".pre", ".post"),
-  all = F
-)
-
-# add to list
-dtS3$clean <- dfCombined
-
-# clean up workspace
-rm(df.pre, df.daily, df.post, dfPreDaily, dfCombined, ownIDs)
-```
-
-## Calculate needed transformations
-
-### Study 1
-
-For the worker sample, the data transformation stage had three main aims:
-
-1.
-We first corrected time indicators within the surveys.
-In some cases participants completed their daily diary surveys for the afternoon after midnight.
-In these cases the measurement still is in reference to the previous day and is indicated in the corrected variable.\
-2.
-We then created indices of scales.
-Some indices were multi-item scales while some indices combine equivalent measurement for different situational circumstances (e.g., competence perceptions after interactions and at measurement occasions without interactions).\
-3.  Finally, we calculated several basic participant summaries (averages across all measurement occasions).
-
-```{r newVarsWorkers}
-df <- dtS1$clean
-
-# Correct date_period ()
-df <- df %>%
-  mutate(
-    startDate = as.Date(created),
-    startTime = format(as.POSIXct(created), format = "%H:%M:%S")
-  )
-
-# order time
-# df$TID <- factor(df$date_period, levels = unique(dtS3$raw.daily$date_period))
-# df$TIDnum <- as.numeric(df$TID) # get numeric TID
-df <- df %>%
-  mutate(
-    PID = as.numeric(factor(last_outside_referrer))
-  )
-
-# Time and Date Variables
-df <- df %>%
-  mutate(
-    TID = measureID - 1,
-    # time ID with t0 = 0 for meaningfull intercept interpretations
-    date = substr(created, 1, 10),
-    # awkward way of extracting date (best converted to )
-    time = substr(created, 12, 19),
-    # awkward way of extracting time
-    daynum = as.numeric(factor(dateQualtrics)),
-    # all days as numeric for ordering
-    daycor = ifelse(
-      daytime == "morning" &
-        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
-        daytime == "afternoon" &
-          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
-      daynum - 1,
-      daynum
-    ),
-    daycorTest = ifelse(
-      daytime == "morning" &
-        period_to_seconds(hms(time)) < period_to_seconds(hms("12:00:00")) |
-        daytime == "afternoon" &
-          period_to_seconds(hms(time)) < period_to_seconds(hms("19:00:00")),
-      1,
-      0
-    ),
-    # correctly identify which date the questionnaire is about
-    daycor.lead = sprintf("%02d", daycor),
-    daytime.lt = ifelse(daytime == "morning", "a", "b"),
-    # morning / afternoon to a / b
-    day_time = paste(daycor.lead, daytime.lt, sep = "_"),
-    # combine day id with morning / afternoon
-    ResponseId = as.numeric(factor(day_time)),
-    # day and time identifier as numeric id
-    SubTime = chron::times(time.0),
-    time.daily = as.character(time),
-    PPDate = as.Date(df$dateQualtrics),
-    number = replace_na(ContactNum, 0),
-    NonDutchNum = replace_na(NonDutchNum, 0)
-  )
-
-# remove seconds from afternoon time
-df$SubTime[df$daytime == "afternoon"] <- paste0(substring(as.character(df$time.0[df$daytime == "afternoon"]), 4, 8), ":00")
-df$time.daily[df$daytime == "afternoon" &
-  !is.na(df$time.daily != "<NA>")] <- paste0(substring(as.character(df$time.daily[df$daytime == "afternoon" &
-  !is.na(df$time.daily != "<NA>")]), 4, 8), ":00")
-
-# Correct morning / afternoon date where survey was collected the day after to indicate the correct date that was targeted
-df$PPDate[df$SubTime < "11:50:00" &
-  df$daytime == "morning"] <- df$PPDate[df$SubTime < "11:50:00" &
-  df$daytime == "morning"] - 1
-df$PPDate[df$SubTime < "18:50:00" &
-  df$daytime == "afternoon"] <- df$PPDate[df$SubTime < "18:50:00" &
-  df$daytime == "afternoon"] - 1
-
-# Make time IDs consistent with later studies
-df$TID <- paste(df$PPDate, str_to_title(df$daytime))
-df$TIDnum <- as.numeric(factor(df$TID %>% gsub(" Morning", "-A", .) %>% gsub(" Afternoon", "-B", .)))
-
-df <- df %>%
-  rowwise() %>%
-  mutate(
-    InteractionDum = sum(IntergroupContact, IngroupContact, na.rm = FALSE),
-    InteractionDum = if_else(InteractionDum > 0, 1, InteractionDum),
-    alertness = sum(alertness1, alertness2, na.rm = TRUE),
-    calmness = sum(calmness1, calmness2, na.rm = TRUE),
-    valence = sum(valence1, valence2, na.rm = TRUE)
-  ) %>%
-  ungroup()
-
-# Need scales
-df$keyMotiveFulfilled <- rowSums(df[, c("KeyNeedFulfillment", "DaytimeNeedFulfillment")], na.rm = T)
-df$autonomy.daily.all <- rowSums(df[, c("autonomy_Int", "autonomy_NoInt")], na.rm = T)
-df$competence.daily.all <- rowSums(df[, c("competence_Int", "competence_NoInt")], na.rm = T)
-# cor(df$relatednessOther, df$relatedness_self_1,use="complete.obs")
-df$relatedness.daily.all <- rowMeans(df[, c(
-  "relatednessOther",
-  "relatednessSelf",
-  "relatednessNoInteraction"
-)], na.rm = T)
-
-pairs.panels.new(
-  df[c("relatednessSelf", "relatednessOther")],
-  labels = c(
-    "I shared information about myself.",
-    "X shared information about themselves."
-  )
-)
-df$relatedness <- rowMeans(df[, c("relatednessOther", "relatednessSelf")], na.rm = T)
-
-df$autonomy <- df$autonomy.daily.all
-df$competence <- df$competence.daily.all
-df$relatedness <- df$relatedness.daily.all
+# import full datasets
+load("data/s123Full.RData")
+
+# import reduced datasets
+load("data/s123Red.RData")
+
+# import pre-calculated features for faster renders
+load("data/features/featOut.RData")
+#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
+#load("data/features/featOutFullImp.RData")
+```
+
+```{r}
+#| label: Preparation of Variable Names
+#| include: true
 
@@
 
-```{r variable names for all analyses}
+```{r}
+#| label: variable names for all analyses
+#| include: true
+
 # ID Variables (just to re-iterate)
@@
 
-We also see that a number of variables are specific to interactions.
-These variables will by their very nature have substantially larger missingness (because not every measurement occasion will follow new social interactions) and these variables are also un-imputable in traditional time-series modeling and hypothesis testing procedures.
-
-```{r variable names by response type and study}
+```{r}
+#| label: variable names by response type and study
+#| include: true
+
 varNames %>%
@@
 
-Please note that not all variables are included in the main analysis, as some variables are only found within specific studies.
-
-## Feature Extraction
-
-Time-series can be described with a plethora of features.
-A good example is the list of 100 time-series features that one can extract with [tsfresh python package](https://tsfresh.readthedocs.io/en/latest/api/tsfresh.feature_extraction.html# "list of features to extract").
-In practice there are usually two steps to the feature engineering process.
-(1) *Feature extraction*, that is summarizing the time series in meaningful descriptors, and (2) *feature selection*, the process of ensuring that only meaningful features are included in the clustering step.
-For the interpretability of our clustering procedure we have made a pre-selection of features, we wish to include to describe our time-series:
-
-1. central tendency
-    + mean ✓
-    + median ✓
-2. variation around the central tendency (i.e., density)
-    + SD ✓
-    + MAD ✓
-3. (in)stability
-    + root mean square of successive differences (rmssd) ✓
-    + mean absolute change (mac) ✓
-4. inertia
-    + lag-1 autocorrelation ✓
-    + periodicity (lag-2 & lag-14 autocorrelation) ✓
-5. linear trend
-    + OLS regression parameter ✓
-6. non-linear trend
-    + GAM edf ✓
-
-(!discuss NA procedures!)
-
-```{r load saved and prepared features}
-load("data/features/featOut.RData")
-```
-
-### Across Studies
-
-We begin by extracting the features for the main analysis that combines the sample of the three studies.
-We do this across all selected variables as well as for the content (non-)specific variables separately.
-During the feature extraction step there can at times be features that cannot be calculated for a small number of participants.
-An example would a participant who never filled in two consecutive measurements that describe an interaction.
-In that case, we cannot extract an auto correlation of lag-1 for some of the interaction specific variables.
-
-```{r Feature Extraction Main Analysis}
-# varNames
-#items <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Joined Dataset of all three studies
-featData <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
+```{r}
+#| label: variables included in main analysis by response type
+#| include: true
+
+varNames %>%
+  filter(
+    pca != 0,
+    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
+  ) %>%
+  mutate(
+    contactSpecific = recode(
+      .$contactSpecific,
+      `-1` = "No Interaction Only",
+      `0` = "unspecific",
+      `1` = "Interaction Only"
+    ),
+    study1 = ifelse(!is.na(studyS1), 1, 0),
+    study2 = ifelse(!is.na(studyS2), 1, 0),
+    study3 = ifelse(!is.na(studyS3), 1, 0)
+  ) %>% 
+  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
+  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
+  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
   ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
-   mutate(
-     across(all_of(varNamS123PCA), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS123PCA), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS123PCA), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-apply(dtS1Red, 2, sum(is.na(.))/nrow(.))
-
-your_dataframe %>%
-  group_by(Your_Participantid) %>%
-  summarise(across(everything(), funs(100*mean(is.na(.)))))
-
-# Don't re-run unless you have a few hours to spare!
-# featFull <-
-#   featureExtractor(
-#     data = featData,
-#     pid = "ID",
-#     tid = "TIDnum",
-#     items = varNamS123
-#   )
-# 
-# featFullContact <-
-#   featureExtractor(
-#     data = featData,
-#     pid = "ID",
-#     tid = "TIDnum",
-#     items = varNamS123Int
-#   )
-# 
-# featFullNoContact <-
-#   featureExtractor(
-#     data = featData,
-#     pid = "ID",
-#     tid = "TIDnum",
-#     items = varNamS123NoInt
-#   )
-# 
-# 
-# featFull$features <- featFull$features %>%
-#     filter(
-#       ID != 19
-#     )
-
-pMissFull <- pMissFeat(
-  all = featFull$features,
-  contact = featFullContact$features,
-  nocontact = featFullNoContact$features,
-  title = "Feature-wise Missingess Across all Studies"
-)
-
-pMissFull$pMissPlot
-
-# featFullImp <- featureImputer(featFull)
-# featFullNoContactImp <- featureImputer(featFullNoContact)
-# featFullContactImp <- featureImputer(featFullContact)
-
-```
-
-For the full dataset across the three studies, we find that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r nrow(featFull$features)` participants and `r ncol(featFull$features)` features.
-These missing features are imputed with a single predictive mean matching imputation using the MICE library.
-
-### Study 1
-
-We then repeat the feature extraction procedure for all selected variables that are specific to Study 1.
-
-```{r Feature Extraction Study 1 specific variables}
-# varNames
-#items <- varNamS1Clust
-
-# Prepare Dataframe
-featDataS1 <- 
-  dtS1Red %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(PID,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS1Clust)) %>%
-   mutate(
-     across(all_of(varNamS1Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS1Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS1Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(PID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featS1 <-
-#   featureExtractor(
-#     data = featDataS1,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS1Clust
-#   )
-# 
-# featS1Contact <-
-#   featureExtractor(
-#     data = featDataS1,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS1Int
-#   )
-# 
-# featS1NoContact <-
-#   featureExtractor(
-#     data = featDataS1,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS1NoInt
-#   )
-
-pMissS1 <- pMissFeat(
-  all = featS1$features,
-  contact = featS1Contact$features,
-  nocontact = featS1NoContact$features,
-  title = "Feature-wise Missingess Study 1"
-)
-
-pMissS1$pMissPlot
-
-# featS1Imp <- featureImputer(featS1)
-# featS1NoContactImp <- featureImputer(featS1NoContact)
-# featS1ContactImp <- featureImputer(featS1Contact)
-
-```
-
-### Study 2
-
-In the same manner, the features are extracted for the variables that are specific to Study 2.
-
-```{r Feature Extraction Study 2 specific variables}
-# varNames
-#items <- varNamS2Clust
-
-# Prepare Dataframe
-featDataS2 <- 
-  dtS2Red %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(PID,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS2Clust)) %>%
-   mutate(
-     across(all_of(varNamS2Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS2Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS2Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(PID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featS2 <-
-#   featureExtractor(
-#     data = featDataS2,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS2Clust
-#   )
-# 
-# featS2Contact <-
-#   featureExtractor(
-#     data = featDataS2,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS2Int
-#   )
-# 
-# featS2NoContact <-
-#   featureExtractor(
-#     data = featDataS2,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS2NoInt
-#   )
-
-pMissS2 <- pMissFeat(
-  all = featS2$features,
-  contact = featS2Contact$features,
-  nocontact = featS2NoContact$features,
-  title = "Feature-wise Missingess Study 2"
-)
-
-pMissS2$pMissPlot
-
-# featS2Imp <- featureImputer(featS2)
-# featS2NoContactImp <- featureImputer(featS2NoContact)
-# featS2ContactImp <- featureImputer(featS2Contact)
-```
-
-### Study 3
-
-And finally, we extract the features for the Study 3 specific variables.
-
-```{r Feature Extraction Study 3 specific variables}
-# varNames
-#items <- varNamS3Clust
-
-# Prepare Dataframe
-featDataS3 <- 
-  dtS3Red %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(PID,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS3Clust)) %>%
-   mutate(
-     across(all_of(varNamS3Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
-     across(all_of(varNamS3Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
-     across(all_of(varNamS3Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
-   ) %>%
-  arrange(PID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-# Don't re-run unless you have a few hours to spare!
-# featS3 <-
-#   featureExtractor(
-#     data = featDataS3,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS3Clust
-#   )
-# 
-# featS3Contact <-
-#   featureExtractor(
-#     data = featDataS3,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS3Int
-#   )
-# 
-# featS3NoContact <-
-#   featureExtractor(
-#     data = featDataS3,
-#     pid = "PID",
-#     tid = "TIDnum",
-#     items = varNamS3NoInt
-#   )
-
-pMissS3 <- pMissFeat(
-  all = featS3$features,
-  contact = featS3Contact$features,
-  nocontact = featS3NoContact$features,
-  title = "Feature-wise Missingess Study 3"
-)
-pMissS3$pMissPlot
-
-# featS3Imp <- featureImputer(featS3)
-# featS3NoContactImp <- featureImputer(featS3NoContact)
-# featS3ContactImp <- featureImputer(featS3Contact)
-# 
-# save(list = ls(pattern = "feat"), file = "data/features/featOut.RData")
-```
-
-## Feature Selection
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Here we might have to make a selection of the features we extract because some of the clustering algorithms might not be able to deal with the many different features. See below
-</div>
-
-Feature selection types:
-
--   manual feature and variable selection
--   wrapper: procedures that add and/or remove predictors to find the optimal combination that maximizes model performance.
--   filter: relevance of the predictors outside of the predictive models and subsequently model only the predictors that pass some criterion.
--   dimension reduction for clustering. No selection more compression of redundant variance.
-
-```{r manual feature selection}
-
-```
-
-```{r automated feature selection}
-
-```
-
-```{r dimension reduction}
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ
-scaled_data <- featFullImp$featuresImpZMat
-
-res.pca <- prcomp(scaled_data, scale = FALSE)
-fviz_eig(res.pca, addlabels = TRUE)
-
-var <- get_pca_var(res.pca)
-corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
-fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
-fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)
-
-res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
-# Description of dimension 1
-res.desc$Dim.1
-
-plot_ly(
-  res.pca$rotation %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Variable:', rownames(res.pca$rotation))
-) %>%
-  layout(title = 'Variables in reduced space')
-plot_ly(
-  res.pca$x %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Participant:', rownames(res.pca$x))
-) %>%
-  layout(title = 'Participants in reduced space')
-
-pca.out <- as.data.frame(res.pca$x[,1:3])
-
-# load libraries
-library(factoextra)
-library(NbClust)
-# elbow method
-fviz_nbclust(pca.out, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method")
-# Silhouette method
-fviz_nbclust(pca.out, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method")
-# Gap statistic [takes for ever]
-# nboot = 50 to keep the function speedy.
-# recommended value: nboot= 500 for your analysis.
-# Use verbose = FALSE to hide computing progression.
-# set.seed(123)
-# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
-#   labs(subtitle = "Gap statistic method")
-
-kmeansOut <- kmeans(pca.out, centers = 3, nstart = 100)
-fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())
-
-# MCLUST
-library(mclust)
-BIC <- mclustBIC(pca.out)
-plot(BIC)
-summary(BIC)
-mod1 <- Mclust(pca.out)
-summary(mod1, parameters = FALSE)
-
-# DBSCAN
-library(dbscan)
-library(fpc)
-
-# to plot the eps values
-eps_plot = kNNdistplot(pca.out, k=3)
-# to draw an optimum line
-eps_plot %>% abline(h = 3.5, lty = 2)
-
-potentialMinPts <- round(SciViews::ln(nrow(pca.out)))
-eps <- 3.5
-minPts <- 2
-
-res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0)
-res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
-
-resEqual.check <- all(res.fpc$cluster == res.db$cluster)
-nClustDBSCAN <- length(unique(res.fpc$cluster))-1
-
-fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
-
-```
-
-## Clustering
-
-Cluster analyses have been a staple within a variety of sciences for the better part of a century.
-It's applications are wide and because most procedures fall under the "unsupervised learning" umbrella the methodological developments are fast-paced and new methods or refinements are published constantly.
-Among the many options available four types are found most commonly within the modern literature on clustering analyses:
-
-1. centroid-based
-    + e.g., k-means
-    + no assumptions (+)
-    + fast even with large matrices (+)
-    + non-probabilistic (+/-)
-    + sensitive to local minima (-)
-2. distribution-based
-    + e.g., gaussian mixture model (GMM)
-    + flexible shape (+)
-    + probabilistic (+/-)
-    + (gaussian) distributions assumed (-)
-3. density-based
-    + e.g., dbscan / OPTICS
-    + no shape assumption (+)
-    + does not assign outliers/noise (+)
-4. Hierarchical-based (not relevant)
-    + e.g., agglomerative hierarchical clustering
-    + best with hierarchical data (-)
-    + best with small *n* (-)
-
-Each clustering procedures comes with their own sets of advantages and disadvantages.
-We will focus mainly on the k-means procedure because it has the least assumptions and is fast even with large data sets.
-Additionally, the methods are well-established and the results are easy to interpret.
-As such, the procedure should be applicable to most users within the social sciences and for ESM feature data specifically.
-We offer Gaussian mixture modeling and a common density-based approach as supplementary analyses.
-
-### Main Anlaysis: Across Studies
-
-The main clustering procedure, uses the features calculated for the relevant variables across all three studies.
-
-#### Centroid-based
-
-For our main method, we use the centroid-based k-means clustering.
-The procedure needs the researcher to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution.
-
-```{r kmeans determin k}
-
-featureExclusion <- c(
-  "mean",
-  "sd",
-  "rmssd",
-  "ar02",
-  "edf_ar"
-)
-
-# prepare data for main analysis
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-# load libraries
-library(factoextra)
-library(NbClust)
-
-# Elbow method
-fviz_nbclust(scaled_data, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method")
-# Silhouette method
-fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method")
-# Gap statistic [takes for ever]
-# nboot = 50 to keep the function speedy. 
-# recommended value: nboot= 500 for your analysis.
-# Use verbose = FALSE to hide computing progression.
-# set.seed(123)
-# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
-#   labs(subtitle = "Gap statistic method")
-```
-
-We use this parameter to run the actual k-means cluster analysis, during which the algorithm iteratively seeks to optimally separate the participants into two clusters based on their feature values.
-During the iterative process the algorithm might at times stop at a non-optimal solution (i.e., local minima).
-To avoid this drawback and to find the (globally) optimal solution we let the algorithm run with 100 separate staring positions.
-
-```{r k-means modeling based on selected k}
-kmeansOut <- kmeans(scaled_data, centers = 2, nstart = 100)
-```
-
-Because the clustering happens in a multidimensional space of the `r ncol(featFull$features)` features, we begin by visualizing the data in a two-dimensional space, where we encircle the points of the clusters (i.e., convex hulls) for the first two dimensions of a principle component analysis.
-
-```{r Plot Convex Hulls}
-fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())
-```
-
-We then inspect the cluster-averages of all feature for each variable.
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., how cooperative an interaction was) than for other variables (e.g., Attitudes towards the Dutch).
-Additionally, we see that (2) some within the variables some features are better at distinguishing clusters (e.g., mean and median of well-being) than others (e.g., sd and rmssd of well-being).
-
-```{r features for each variable}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarize_all(mean) %>%
-  ungroup %>%
-  select(-ID) %>% 
-  reshape2::melt(id.vars = "cluster") %>% 
-  mutate(lab = variable) %>%
-  mutate(
-    lab = variable,
-    variable = gsub("\\_.*", "", lab),
-    feature = gsub("^.*?\\_", "", lab)
-  ) %>% 
-  ggplot(., aes(x = feature, y = value, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ variable) +
-  theme_Publication()
-```
-
-Next, we inspect the cluster-averages of all variables for each feature.
-This is the same data as in the plot above.
-However, here we see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median) than others (e.g., auto correlations).
-
-```{r variables for each feature}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarize_all(mean) %>%
-  ungroup %>%
-  select(-ID) %>% 
-  reshape2::melt(id.vars = "cluster") %>% 
-  mutate(
-    lab = variable,
-    variable = gsub("\\_.*", "", lab),
-    feature = gsub("^.*?\\_", "", lab)
-  ) %>% 
-  ggplot(., aes(x = variable, y = value, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ feature) +
-  theme_Publication()
-```
-
-In a final step we visualize the raw time series for each variable and add the average cluster development to see the features in action.
-
-```{r time-series by variable and cluster}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ variable, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-```
-
-```{r time-series linear trends by variable and cluster}
-data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
-  #stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ variable, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-```
-
-#### Distribution-based
-
-For the distribution-based clustering we use the commonly used Gaussian mixture modeling (GMM) procedure.
-This procedure assumes  independent Gaussian distributions for each each cluster.
-These distributions are substantially better at identifying clusters if the data comes from multivariate Gaussian distributions (i.e., ellipses in a two dimensional space).
-
-We again begin by determining the optimal number of clusters. For GMMs this is done using common information criteria, such as the BIC.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  This is where we run into trouble.
-These plots are relatively uncommon and according to the internet this is most likely a problem of too many (non-distinguishing or highly correlated) features.
-A good paper on this is <a href="https://proceedings.neurips.cc/paper/2002/file/e58aea67b01fa747687f038dfde066f6-Paper.pdf">Law et al.
-(2002)</a>.
-</div>
-
-```{r gaussian mixture model}
-library(mclust)
-
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-# scaled_data <- featS1Imp$featuresImpZMat  %>% #select(starts_with(varNamS123Int))
-#   select(-any_of(names(which(colSums(is.na(.)) > 0))))
-
-BIC <- mclustBIC(scaled_data)
-plot(BIC)
-summary(BIC)
-mod1 <- Mclust(scaled_data)
-summary(mod1, parameters = FALSE)
-
-# multivariate <- densityMclust(scaled_data, plot=FALSE)
-# summary(multivariate)            
-# plot(multivariate, what = "BIC")
-
-# merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View
-
-#table(class, mod1$classification)
-
-#mod1dr <- MclustDR(mod1)
-#summary(mod1dr, parameters = FALSE)
-#plot(mod1dr, what = "classification")
-```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  The feature selection issue is further highlighted when we do the GMMs by feature. Where we find a reasonable number of clusters for each feature. 
-</div>
-
-```{r MClust feature specific cluster numbers}
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-nMClust <- list()
-for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){
-  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
-  mod1 <- Mclust(scaled_data)
-  #summary(mod1, parameters = FALSE)
-  nMClust[[f]] <- mod1$G
-}
-
-nMClust %>%
-  unlist %>%
-  data.frame(feature=names(.), G=., row.names=NULL) %>%
+  t %>%
+  as.data.frame %>%
+  janitor::row_to_names(1) %>%
+  replace(is.na(.), 0) %>%
+  tibble::rownames_to_column(., var = "variable") %>%
+  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
+  mutate(across(c(study1, study2, study3), as.numeric)) %>%
+  #select(-study1, -study2, -study3) %>%
+  arrange(concept, unspecific, `Interaction Only`) %>% 
+  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  relocate(starts_with("study"), .after = last_col()) %>%
+  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
   kbl(.,
@@
       booktabs = T,
-      align = c("l", "c"),
+      align = c("l", "l", rep("c", ncol(.)-2)),
       digits=2,
-      col.names = c(
-        "Feature",
-        "Number of Clusters"
-      ),
-      caption = "Number of MClust Clusters if separated by Feature") %>%
+      caption = "Variables by Interaction Types and Study Availability") %>%
+  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
   kable_classic(
@@
   )
-
-```
-
-#### Density-based
-
-For the density-based cluster analysis we focus on the *DBSCAN* algorithm ('Density-Based Spatial Clustering of Applications with Noise') as well as its extension the *OPTICS* procedure ('Ordering Points To Identify the Clustering Structure').
-This procedure is more bottom up than the previous two approaches.
-Instead of providing the number of clusters, a cluster is specified if a minimum number of other points (the MinPts parameter) are within a given radius (the epsilon parameter).
-Together these two parameters determine a density at which a group of data points are considered a cluster.
-While this description is grossly oversimplified, it highlights one of the advantages of the procedures, namely that dense regions can expand to be any shape in the d-dimensional space.
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  Same feature selection issue as above. 
-</div>
-
-```{r dbscan}
-library(dbscan)
-library(fpc)
-
-#scaled_data <- featFullImp$featuresImpZMat %>% select(starts_with(varNamS123Int))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))  # %>% select(starts_with(varNamS123NoInt))#%>%
-  #select(-any_of(names(which(colSums(is.na(scaled_data)) > 0))))
-
-#is.na(scaled_data) %>% table
-
-# to plot the eps values
-eps_plot = kNNdistplot(scaled_data, k=3)
-# to draw an optimum line
-eps_plot %>% abline(h = 11, lty = 2)
-
-potentialMinPts <- round(SciViews::ln(nrow(scaled_data)))
-eps <- 11
-minPts <- 2
-
-res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, showplot=0)
-res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
-
-resEqual.check <- all(res.fpc$cluster == res.db$cluster)
-nClustDBSCAN <- length(unique(res.fpc$cluster))-1
-
-fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
-#hullplot(scaled_data, res.fpc)
-
-```
-
-<div class="alert alert-warning alert-dismissible fade in" role="alert">
-  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
-  <i class="fas fa-exclamation-triangle"></i> <b>Note: </b><br/>
-  And again the number of clusters is reasonable if we consider the features separately: 
-</div>
-
-```{r dbscan feature specific cluister numbers}
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-minPts <- 3
-xi <- 0.04 #0.05
-nDbscan <- list()
-for(f in unique(gsub("^.*?\\_", "", names(scaled_data)))){
-  scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(f))
-  #res.fpc <- dbscan::dbscan(scaled_data, eps=eps_cl, MinPts=minPts)
-
-  res.optics <- dbscan::optics(scaled_data, minPts = 3)
-
-  #plot(res.optics, col = "grey")
-  #dend <- as.dendrogram(res.optics)
-  #plot(dend, ylab = "Reachability dist.", leaflab = "none")
-  #extractDBSCAN(res.optics, eps_cl = 19.12)
-
-  res <- extractXi(res.optics, xi = xi)
-  res$clusters_xi
-  #plot(res)
-  #hullplot(scaled_data, res)
-
-  #summary(res.optics, parameters = FALSE)
-  nDbscan[[f]] <- nrow(res$clusters_xi)
-}
-
-nDbscan %>%
-  unlist %>%
-  data.frame(feature=names(.), G=., row.names=NULL) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "c"),
-      digits=2,
-      col.names = c(
-        "Feature",
-        "Number of Clusters"
-      ),
-      caption = "Number of OPTICS Clusters if separated by Feature") %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-### Follow-up: By Response Type
-
-_**currently removed for faster render until feature selection issue is resolved**_
-
-### Follow-up: By Study
-
-_**currently removed for faster render until feature selection issue is resolved**_
-
-# **Software Information**
+```
+
+# Data Description
+
+```{r}
+#| label: render data description
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/01DataDescription.Rmd",
+  output_dir = outputDir
+)
+```
+
+# Input Variables
+
+```{r}
+#| label: render input variables
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/02InputVariables.Rmd",
+  output_dir = outputDir
+)
+```
+
+# Feature Extraction
+
+```{r}
+#| label: render feature extraction
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/03FeatureExtraction.Rmd",
+  output_dir = outputDir
+)
+```
+
+# Feature Reduction
+
+```{r}
+#| label: render feature reduction
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/04FeatureReduction.Rmd",
+  output_dir = outputDir
+)
+```
+
+# Feature Clustering
+
+```{r}
+#| label: render feature clustering
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/05FeatureClustering.Rmd",
+  output_dir = outputDir
+)
+```
+
+# Cluster Evaluation
+
+```{r}
+#| label: render cluster evaluation
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/06ClusterEvaluation.Rmd",
+  output_dir = outputDir
+)
+```
+
+# Software Information
 
@@
 
-# **References**
+# References
 

### 2023-02-17 docs/methods-and-results.tex
:construction: section-based re-render

--- 2023-01-20 methods-and-results.tex
@@
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
+
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
+
+\author{}
+\date{}
+
+\begin{document}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
+
 \subsection{Data}
@@
 \end{figure}
+
+\end{document}

### 2023-02-17 latex-results/01DataDescription.tex
:construction: section-based re-render

--- 2023-01-17 methods-and-results.tex
@@
-\subsection{Data}
-
 We used a data set following migration experiences collected by
@@
 
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies
-and captured information about the participant's interactions, as well
-as cognitive-, emotional-, and motivational self in relationship with
-the majority group. We chose these two aspects in particular because (1)
-the interaction-specific information exemplified the structural
-missingness issue of modern ESM data and (2) the motivational,
-emotional, and cognitive experience offered a diverse conceptualization
-of migration experience (beyond behavioral measurements) that is
-becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
-methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are also
-available in \tblref{tab:descrLong}.
-
 \input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster--evaluate steps in sequential order. For each
-of the steps fully reproducible and annotated code is available in
-Online Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REF).
-
-\subsubsection{Feature extraction}
-
-For the features, we selected and extracted one of each of the features
-proposed in \tblref{tab:esmFeatures}. We particularly chose the more
-robust median and median absolute deviation (MAD) for central tendency
-and variability. Within the variability structure, we chose the slightly
-more intuitive mean absolute change (MAC) for stability and stayed with
-the common lag--1 autocorrelation for inertia. For the trend summaries,
-we chose the overall linear regression slope to describe the linear
-trend and we summarized the nonlinear trend with the estimated degrees
-of freedom of an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We chose these features in
-particular because we consider them to be the most broadly applicable
-features. We also extracted the participant's number of completed ESM
-measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a
-large selection of the time series feature operationalizations presented
-in \tblref{tab:esmFeatures} in our GitHub repository (see the
-\texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
-the an ordered list of values from the time series of variable \(j\) for
-participant \(i\). The calculation depends on whether the number of
-measurements in a time series \(n\) is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation}
-(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
-in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag \(l=1\)
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of
-an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across the 72 features per participant.
-This might, for example, happen if participants do not have two
-subsequent measurements with outgroup interactions, so that an
-autocorrelation with lag-1 cannot be calculated for the contact-specific
-variables. The small number of missing values indicates that the
-feature-based approach indeed largely avoids the structural missingness
-issue. The few missing values can, however, be an issue for some feature
-reduction or feature clustering algorithms. We, thus, impute the missing
-feature values with a single predictive mean matching imputation using
-the MICE library.
-
-\subsubsection{Feature reduction}
-
-For the feature reduction, we chose the common
-\textit{principal component analysis} (PCA). Some of the more
-tailor-made feature selection algorithms can be more accurate in
-reducing the feature dimensionality and might retain feature importance
-information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established
-within the psychometric literature and can broadly be applied to a wide
-variety of studies in an automatize manner. As our aim is to present a
-general illustration that can also be adopted for more general data
-descriptive uses, we present the workflow using a PCA here but we
-encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series
-features, we first standardize all features across participants to
-ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
-method to iteratively explain the most of the remaining variance while
-also ensuring that the components are linearly uncorrelated. In
-practice, this meant that the PCA decomposed the 72 features into 72
-principal components but now (because of the uncorrelated linear
-transformations) the first few principal components will capture a
-majority of the variance. We can then decide how much information (i.e.,
-variance) we are willing to sacrifice for a reduced dimensionality. A
-common rule of thumb is to use the principal components that jointly
-explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we select the first 27
-principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 27 PC-scores for each participant
-(i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal
-components --- the amount of dimensionality reduction and the
-interpretation of the principal components. As for the expected
-dimensionality reduction, given its methodology, PCAs tend to `work
-better' at reducing dimensions with (highly) correlated variables. Thus,
-with a set of very homogeneous variables and features users will need
-less principal components to explain a large amount of variance, while a
-more diverse set of variables and features will tend to require more
-principal components to capture the same amount of variance. Our 27
-principal components are still a relatively high number of variables but
-this is not surprising as we chose a diverse conceptualization and a
-diverse set of time series features. As for the interpretability, PCA
-allows users to extract information on the meaning of the principal
-components. In particular, because the principal components are linear
-combination of original features, users can extract the relative
-importance of each feature for the extracted principal components (i.e.,
-the eigenvectors). While this can be useful in understanding the
-variance in the original data or help with manual feature selection, we
-use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original
-features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-For our illustration of the feature clustering procedure, we use the
-generalizable and efficient centroid-based k-means clustering. The
-procedure needs the user to make a selection on the optimal number of
-clusters in order to run the main algorithm. We use the Elbow and the
-Silhouette method, which both suggest two clusters to be the optimal
-solution (see Supplemental Material A). We then entered the
-participants' PC-scores into the k-means algorithm. We used the
-recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
-random initial centroid positions to avoid convergence to a sub-optimal
-solution (i.e., local minima). The k-means analysis assigned 80
-participants to cluster 1 and 76 participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in
-their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see
-\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
-features are generally stronger in separating the clusters (e.g., `how
-cooperative an interaction was' compared to `attitudes towards the
-Dutch'). Additionally, we see that (2) within the variables some
-features are better at distinguishing clusters (e.g., median of
-well-being vs.~MAC of well-being).We then inspect the clusters with a
-focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
-is the same data as for the variable focus, we can see more clearly that
-some features are better at distinguishing the clusters across variables
-(e.g., mean and median compared to auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the
-second cluster had more meaningful and fulfilling outgroup interactions
-also consistently reported more voluntary and cooperative interactions
-but less accidental and involuntary interactions. The same cluster also
-reported an increase in need fulfilling interaction over the 30 day
-period and an increase in interactions that were representative of the
-outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. This
-`deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see
-\fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, well-being are relatively stable
-for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability over time (also see
-\fgrref[A]{fig:clusterTs}).
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as
-well as check whether the developmental clusters are associated with
-important social markers and individual differences. To illustrate
-artifact checks, we added the number of measurements into the comparison
-and find that the participants in the detereoration cluster on average
-completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions (see
-\fgrref[B]{fig:clusterTs}). While this difference could indicate that
-the clusters might not entirely be comparable in the response patterns,
-we can find some relief in our data exclusion procedures during which we
-ensured that the general time frame and completion rates were not too
-dissimilar. To illustrate the utility of individual differences, we
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
-participants in the deteriorating cluster reported substantially higher
-levels of everyday discrimination. Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}

### 2023-02-17 latex-results/02InputVariables.tex
:construction: section-based re-render

--- 2023-01-17 methods-and-results.tex
@@
-\subsection{Data}
-
-We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
-followed migrants who had recently arrived in the Netherlands in their
-daily interactions with the Dutch majority group members. After a
-general migration-focused pre-questionnaire, participants were invited
-twice per day to report on their (potential) interactions with majority
-group members for at least 30 days. The short ESM surveys were sent out
-at around lunch (12pm) and dinner time (7pm). After the 30 day study
-period participants filled in a post-questionnaire that mirrored the
-pre-questionnaire. Participants received either monetary compensation or
-partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Full
-methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included 207 participants (\(N_{S1}=\) 23,
-\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
-The studies differed substantially in the maximum length of
-participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
-\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
-continue participation without compensation in the later two studies. To
-make the three studies comparable in participation and time frames, we
-iteratively removed all measurement occasions and participants that had
-more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now a lot more comparable
-(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
-
-\subsubsection{Variables}
-
 We included 12 variables that were available across all three studies
@@
 \input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster--evaluate steps in sequential order. For each
-of the steps fully reproducible and annotated code is available in
-Online Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REF).
-
-\subsubsection{Feature extraction}
-
-For the features, we selected and extracted one of each of the features
-proposed in \tblref{tab:esmFeatures}. We particularly chose the more
-robust median and median absolute deviation (MAD) for central tendency
-and variability. Within the variability structure, we chose the slightly
-more intuitive mean absolute change (MAC) for stability and stayed with
-the common lag--1 autocorrelation for inertia. For the trend summaries,
-we chose the overall linear regression slope to describe the linear
-trend and we summarized the nonlinear trend with the estimated degrees
-of freedom of an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We chose these features in
-particular because we consider them to be the most broadly applicable
-features. We also extracted the participant's number of completed ESM
-measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a
-large selection of the time series feature operationalizations presented
-in \tblref{tab:esmFeatures} in our GitHub repository (see the
-\texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
-the an ordered list of values from the time series of variable \(j\) for
-participant \(i\). The calculation depends on whether the number of
-measurements in a time series \(n\) is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation}
-(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
-in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag \(l=1\)
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of
-an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across the 72 features per participant.
-This might, for example, happen if participants do not have two
-subsequent measurements with outgroup interactions, so that an
-autocorrelation with lag-1 cannot be calculated for the contact-specific
-variables. The small number of missing values indicates that the
-feature-based approach indeed largely avoids the structural missingness
-issue. The few missing values can, however, be an issue for some feature
-reduction or feature clustering algorithms. We, thus, impute the missing
-feature values with a single predictive mean matching imputation using
-the MICE library.
-
-\subsubsection{Feature reduction}
-
-For the feature reduction, we chose the common
-\textit{principal component analysis} (PCA). Some of the more
-tailor-made feature selection algorithms can be more accurate in
-reducing the feature dimensionality and might retain feature importance
-information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established
-within the psychometric literature and can broadly be applied to a wide
-variety of studies in an automatize manner. As our aim is to present a
-general illustration that can also be adopted for more general data
-descriptive uses, we present the workflow using a PCA here but we
-encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series
-features, we first standardize all features across participants to
-ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
-method to iteratively explain the most of the remaining variance while
-also ensuring that the components are linearly uncorrelated. In
-practice, this meant that the PCA decomposed the 72 features into 72
-principal components but now (because of the uncorrelated linear
-transformations) the first few principal components will capture a
-majority of the variance. We can then decide how much information (i.e.,
-variance) we are willing to sacrifice for a reduced dimensionality. A
-common rule of thumb is to use the principal components that jointly
-explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we select the first 27
-principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 27 PC-scores for each participant
-(i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal
-components --- the amount of dimensionality reduction and the
-interpretation of the principal components. As for the expected
-dimensionality reduction, given its methodology, PCAs tend to `work
-better' at reducing dimensions with (highly) correlated variables. Thus,
-with a set of very homogeneous variables and features users will need
-less principal components to explain a large amount of variance, while a
-more diverse set of variables and features will tend to require more
-principal components to capture the same amount of variance. Our 27
-principal components are still a relatively high number of variables but
-this is not surprising as we chose a diverse conceptualization and a
-diverse set of time series features. As for the interpretability, PCA
-allows users to extract information on the meaning of the principal
-components. In particular, because the principal components are linear
-combination of original features, users can extract the relative
-importance of each feature for the extracted principal components (i.e.,
-the eigenvectors). While this can be useful in understanding the
-variance in the original data or help with manual feature selection, we
-use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original
-features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-For our illustration of the feature clustering procedure, we use the
-generalizable and efficient centroid-based k-means clustering. The
-procedure needs the user to make a selection on the optimal number of
-clusters in order to run the main algorithm. We use the Elbow and the
-Silhouette method, which both suggest two clusters to be the optimal
-solution (see Supplemental Material A). We then entered the
-participants' PC-scores into the k-means algorithm. We used the
-recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
-random initial centroid positions to avoid convergence to a sub-optimal
-solution (i.e., local minima). The k-means analysis assigned 80
-participants to cluster 1 and 76 participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in
-their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see
-\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
-features are generally stronger in separating the clusters (e.g., `how
-cooperative an interaction was' compared to `attitudes towards the
-Dutch'). Additionally, we see that (2) within the variables some
-features are better at distinguishing clusters (e.g., median of
-well-being vs.~MAC of well-being).We then inspect the clusters with a
-focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
-is the same data as for the variable focus, we can see more clearly that
-some features are better at distinguishing the clusters across variables
-(e.g., mean and median compared to auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the
-second cluster had more meaningful and fulfilling outgroup interactions
-also consistently reported more voluntary and cooperative interactions
-but less accidental and involuntary interactions. The same cluster also
-reported an increase in need fulfilling interaction over the 30 day
-period and an increase in interactions that were representative of the
-outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. This
-`deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see
-\fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, well-being are relatively stable
-for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability over time (also see
-\fgrref[A]{fig:clusterTs}).
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as
-well as check whether the developmental clusters are associated with
-important social markers and individual differences. To illustrate
-artifact checks, we added the number of measurements into the comparison
-and find that the participants in the detereoration cluster on average
-completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions (see
-\fgrref[B]{fig:clusterTs}). While this difference could indicate that
-the clusters might not entirely be comparable in the response patterns,
-we can find some relief in our data exclusion procedures during which we
-ensured that the general time frame and completion rates were not too
-dissimilar. To illustrate the utility of individual differences, we
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
-participants in the deteriorating cluster reported substantially higher
-levels of everyday discrimination. Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}

### 2023-02-17 latex-results/03FeatureExtraction.tex
:construction: section-based re-render

--- 2023-01-20 methods-and-results.tex
@@
-\subsection{Data}
-
-We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
-followed migrants who had recently arrived in the Netherlands in their
-daily interactions with the Dutch majority group members. After a
-general migration-focused pre-questionnaire, participants were invited
-twice per day to report on their (potential) interactions with majority
-group members for at least 30 days. The short ESM surveys were sent out
-at around lunch (12pm) and dinner time (7pm). After the 30 day study
-period participants filled in a post-questionnaire that mirrored the
-pre-questionnaire. Participants received either monetary compensation or
-partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Full
-methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included 207 participants (\(N_{S1}=\) 23,
-\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
-The studies differed substantially in the maximum length of
-participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
-\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
-continue participation without compensation in the later two studies. To
-make the three studies comparable in participation and time frames, we
-iteratively removed all measurement occasions and participants that had
-more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now a lot more comparable
-(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies
-and captured information about the participant's interactions, as well
-as cognitive-, emotional-, and motivational self in relationship with
-the majority group. We chose these two aspects in particular because (1)
-the interaction-specific information exemplified the structural
-missingness issue of modern ESM data and (2) the motivational,
-emotional, and cognitive experience offered a diverse conceptualization
-of migration experience (beyond behavioral measurements) that is
-becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
-methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are also
-available in \tblref{tab:descrLong}.
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster--evaluate steps in sequential order. For each
-of the steps fully reproducible and annotated code is available in
-Online Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REF).
-
-\subsubsection{Feature extraction}
-
 For the features, we selected and extracted one of each of the features
@@
 the MICE library.
-
-\subsubsection{Feature reduction}
-
-For the feature reduction, we chose the common
-\textit{principal component analysis} (PCA). Some of the more
-tailor-made feature selection algorithms can be more accurate in
-reducing the feature dimensionality and might retain feature importance
-information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established
-within the psychometric literature and can broadly be applied to a wide
-variety of studies in an automatize manner. As our aim is to present a
-general illustration that can also be adopted for more general data
-descriptive uses, we present the workflow using a PCA here but we
-encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series
-features, we first standardize all features across participants to
-ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
-method to iteratively explain the most of the remaining variance while
-also ensuring that the components are linearly uncorrelated. In
-practice, this meant that the PCA decomposed the 72 features into 72
-principal components but now (because of the uncorrelated linear
-transformations) the first few principal components will capture a
-majority of the variance. We can then decide how much information (i.e.,
-variance) we are willing to sacrifice for a reduced dimensionality. A
-common rule of thumb is to use the principal components that jointly
-explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we select the first 27
-principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 27 PC-scores for each participant
-(i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal
-components --- the amount of dimensionality reduction and the
-interpretation of the principal components. As for the expected
-dimensionality reduction, given its methodology, PCAs tend to `work
-better' at reducing dimensions with (highly) correlated variables. Thus,
-with a set of very homogeneous variables and features users will need
-less principal components to explain a large amount of variance, while a
-more diverse set of variables and features will tend to require more
-principal components to capture the same amount of variance. Our 27
-principal components are still a relatively high number of variables but
-this is not surprising as we chose a diverse conceptualization and a
-diverse set of time series features. As for the interpretability, PCA
-allows users to extract information on the meaning of the principal
-components. In particular, because the principal components are linear
-combination of original features, users can extract the relative
-importance of each feature for the extracted principal components (i.e.,
-the eigenvectors). While this can be useful in understanding the
-variance in the original data or help with manual feature selection, we
-use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original
-features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-For our illustration of the feature clustering procedure, we use the
-generalizable and efficient centroid-based k-means clustering. The
-procedure needs the user to make a selection on the optimal number of
-clusters in order to run the main algorithm. We use the Elbow and the
-Silhouette method, which both suggest two clusters to be the optimal
-solution (see Supplemental Material A). We then entered the
-participants' PC-scores into the k-means algorithm. We used the
-recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
-random initial centroid positions to avoid convergence to a sub-optimal
-solution (i.e., local minima). The k-means analysis assigned 80
-participants to cluster 1 and 76 participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in
-their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see
-\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
-features are generally stronger in separating the clusters (e.g., `how
-cooperative an interaction was' compared to `attitudes towards the
-Dutch'). Additionally, we see that (2) within the variables some
-features are better at distinguishing clusters (e.g., median of
-well-being vs.~MAC of well-being).We then inspect the clusters with a
-focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
-is the same data as for the variable focus, we can see more clearly that
-some features are better at distinguishing the clusters across variables
-(e.g., mean and median compared to auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the
-second cluster had more meaningful and fulfilling outgroup interactions
-also consistently reported more voluntary and cooperative interactions
-but less accidental and involuntary interactions. The same cluster also
-reported an increase in need fulfilling interaction over the 30 day
-period and an increase in interactions that were representative of the
-outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. This
-`deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see
-\fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, well-being are relatively stable
-for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability over time (also see
-\fgrref[A]{fig:clusterTs}).
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as
-well as check whether the developmental clusters are associated with
-important social markers and individual differences. To illustrate
-artifact checks, we added the number of measurements into the comparison
-and find that the participants in the detereoration cluster on average
-completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions (see
-\fgrref[B]{fig:clusterTs}). While this difference could indicate that
-the clusters might not entirely be comparable in the response patterns,
-we can find some relief in our data exclusion procedures during which we
-ensured that the general time frame and completion rates were not too
-dissimilar. To illustrate the utility of individual differences, we
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
-participants in the deteriorating cluster reported substantially higher
-levels of everyday discrimination. Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}

### 2023-02-17 latex-results/04FeatureReduction.tex
:construction: section-based re-render

--- 2023-01-17 methods-and-results.tex
@@
-\subsection{Data}
-
-We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
-followed migrants who had recently arrived in the Netherlands in their
-daily interactions with the Dutch majority group members. After a
-general migration-focused pre-questionnaire, participants were invited
-twice per day to report on their (potential) interactions with majority
-group members for at least 30 days. The short ESM surveys were sent out
-at around lunch (12pm) and dinner time (7pm). After the 30 day study
-period participants filled in a post-questionnaire that mirrored the
-pre-questionnaire. Participants received either monetary compensation or
-partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Full
-methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included 207 participants (\(N_{S1}=\) 23,
-\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
-The studies differed substantially in the maximum length of
-participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
-\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
-continue participation without compensation in the later two studies. To
-make the three studies comparable in participation and time frames, we
-iteratively removed all measurement occasions and participants that had
-more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now a lot more comparable
-(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies
-and captured information about the participant's interactions, as well
-as cognitive-, emotional-, and motivational self in relationship with
-the majority group. We chose these two aspects in particular because (1)
-the interaction-specific information exemplified the structural
-missingness issue of modern ESM data and (2) the motivational,
-emotional, and cognitive experience offered a diverse conceptualization
-of migration experience (beyond behavioral measurements) that is
-becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
-methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are also
-available in \tblref{tab:descrLong}.
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster--evaluate steps in sequential order. For each
-of the steps fully reproducible and annotated code is available in
-Online Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REF).
-
-\subsubsection{Feature extraction}
-
-For the features, we selected and extracted one of each of the features
-proposed in \tblref{tab:esmFeatures}. We particularly chose the more
-robust median and median absolute deviation (MAD) for central tendency
-and variability. Within the variability structure, we chose the slightly
-more intuitive mean absolute change (MAC) for stability and stayed with
-the common lag--1 autocorrelation for inertia. For the trend summaries,
-we chose the overall linear regression slope to describe the linear
-trend and we summarized the nonlinear trend with the estimated degrees
-of freedom of an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We chose these features in
-particular because we consider them to be the most broadly applicable
-features. We also extracted the participant's number of completed ESM
-measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a
-large selection of the time series feature operationalizations presented
-in \tblref{tab:esmFeatures} in our GitHub repository (see the
-\texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
-the an ordered list of values from the time series of variable \(j\) for
-participant \(i\). The calculation depends on whether the number of
-measurements in a time series \(n\) is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation}
-(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
-in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag \(l=1\)
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of
-an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across the 72 features per participant.
-This might, for example, happen if participants do not have two
-subsequent measurements with outgroup interactions, so that an
-autocorrelation with lag-1 cannot be calculated for the contact-specific
-variables. The small number of missing values indicates that the
-feature-based approach indeed largely avoids the structural missingness
-issue. The few missing values can, however, be an issue for some feature
-reduction or feature clustering algorithms. We, thus, impute the missing
-feature values with a single predictive mean matching imputation using
-the MICE library.
-
-\subsubsection{Feature reduction}
-
 For the feature reduction, we chose the common
@@
 and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-For our illustration of the feature clustering procedure, we use the
-generalizable and efficient centroid-based k-means clustering. The
-procedure needs the user to make a selection on the optimal number of
-clusters in order to run the main algorithm. We use the Elbow and the
-Silhouette method, which both suggest two clusters to be the optimal
-solution (see Supplemental Material A). We then entered the
-participants' PC-scores into the k-means algorithm. We used the
-recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
-random initial centroid positions to avoid convergence to a sub-optimal
-solution (i.e., local minima). The k-means analysis assigned 80
-participants to cluster 1 and 76 participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in
-their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see
-\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
-features are generally stronger in separating the clusters (e.g., `how
-cooperative an interaction was' compared to `attitudes towards the
-Dutch'). Additionally, we see that (2) within the variables some
-features are better at distinguishing clusters (e.g., median of
-well-being vs.~MAC of well-being).We then inspect the clusters with a
-focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
-is the same data as for the variable focus, we can see more clearly that
-some features are better at distinguishing the clusters across variables
-(e.g., mean and median compared to auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the
-second cluster had more meaningful and fulfilling outgroup interactions
-also consistently reported more voluntary and cooperative interactions
-but less accidental and involuntary interactions. The same cluster also
-reported an increase in need fulfilling interaction over the 30 day
-period and an increase in interactions that were representative of the
-outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. This
-`deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see
-\fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, well-being are relatively stable
-for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability over time (also see
-\fgrref[A]{fig:clusterTs}).
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as
-well as check whether the developmental clusters are associated with
-important social markers and individual differences. To illustrate
-artifact checks, we added the number of measurements into the comparison
-and find that the participants in the detereoration cluster on average
-completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions (see
-\fgrref[B]{fig:clusterTs}). While this difference could indicate that
-the clusters might not entirely be comparable in the response patterns,
-we can find some relief in our data exclusion procedures during which we
-ensured that the general time frame and completion rates were not too
-dissimilar. To illustrate the utility of individual differences, we
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
-participants in the deteriorating cluster reported substantially higher
-levels of everyday discrimination. Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}

### 2023-02-17 latex-results/05FeatureClustering.tex
:construction: section-based re-render

--- 2023-01-17 methods-and-results.tex
@@
-\subsection{Data}
-
-We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
-followed migrants who had recently arrived in the Netherlands in their
-daily interactions with the Dutch majority group members. After a
-general migration-focused pre-questionnaire, participants were invited
-twice per day to report on their (potential) interactions with majority
-group members for at least 30 days. The short ESM surveys were sent out
-at around lunch (12pm) and dinner time (7pm). After the 30 day study
-period participants filled in a post-questionnaire that mirrored the
-pre-questionnaire. Participants received either monetary compensation or
-partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Full
-methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included 207 participants (\(N_{S1}=\) 23,
-\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
-The studies differed substantially in the maximum length of
-participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
-\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
-continue participation without compensation in the later two studies. To
-make the three studies comparable in participation and time frames, we
-iteratively removed all measurement occasions and participants that had
-more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now a lot more comparable
-(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies
-and captured information about the participant's interactions, as well
-as cognitive-, emotional-, and motivational self in relationship with
-the majority group. We chose these two aspects in particular because (1)
-the interaction-specific information exemplified the structural
-missingness issue of modern ESM data and (2) the motivational,
-emotional, and cognitive experience offered a diverse conceptualization
-of migration experience (beyond behavioral measurements) that is
-becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
-methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are also
-available in \tblref{tab:descrLong}.
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster--evaluate steps in sequential order. For each
-of the steps fully reproducible and annotated code is available in
-Online Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REF).
-
-\subsubsection{Feature extraction}
-
-For the features, we selected and extracted one of each of the features
-proposed in \tblref{tab:esmFeatures}. We particularly chose the more
-robust median and median absolute deviation (MAD) for central tendency
-and variability. Within the variability structure, we chose the slightly
-more intuitive mean absolute change (MAC) for stability and stayed with
-the common lag--1 autocorrelation for inertia. For the trend summaries,
-we chose the overall linear regression slope to describe the linear
-trend and we summarized the nonlinear trend with the estimated degrees
-of freedom of an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We chose these features in
-particular because we consider them to be the most broadly applicable
-features. We also extracted the participant's number of completed ESM
-measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a
-large selection of the time series feature operationalizations presented
-in \tblref{tab:esmFeatures} in our GitHub repository (see the
-\texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
-the an ordered list of values from the time series of variable \(j\) for
-participant \(i\). The calculation depends on whether the number of
-measurements in a time series \(n\) is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation}
-(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
-in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag \(l=1\)
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of
-an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across the 72 features per participant.
-This might, for example, happen if participants do not have two
-subsequent measurements with outgroup interactions, so that an
-autocorrelation with lag-1 cannot be calculated for the contact-specific
-variables. The small number of missing values indicates that the
-feature-based approach indeed largely avoids the structural missingness
-issue. The few missing values can, however, be an issue for some feature
-reduction or feature clustering algorithms. We, thus, impute the missing
-feature values with a single predictive mean matching imputation using
-the MICE library.
-
-\subsubsection{Feature reduction}
-
-For the feature reduction, we chose the common
-\textit{principal component analysis} (PCA). Some of the more
-tailor-made feature selection algorithms can be more accurate in
-reducing the feature dimensionality and might retain feature importance
-information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established
-within the psychometric literature and can broadly be applied to a wide
-variety of studies in an automatize manner. As our aim is to present a
-general illustration that can also be adopted for more general data
-descriptive uses, we present the workflow using a PCA here but we
-encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series
-features, we first standardize all features across participants to
-ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
-method to iteratively explain the most of the remaining variance while
-also ensuring that the components are linearly uncorrelated. In
-practice, this meant that the PCA decomposed the 72 features into 72
-principal components but now (because of the uncorrelated linear
-transformations) the first few principal components will capture a
-majority of the variance. We can then decide how much information (i.e.,
-variance) we are willing to sacrifice for a reduced dimensionality. A
-common rule of thumb is to use the principal components that jointly
-explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we select the first 27
-principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 27 PC-scores for each participant
-(i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal
-components --- the amount of dimensionality reduction and the
-interpretation of the principal components. As for the expected
-dimensionality reduction, given its methodology, PCAs tend to `work
-better' at reducing dimensions with (highly) correlated variables. Thus,
-with a set of very homogeneous variables and features users will need
-less principal components to explain a large amount of variance, while a
-more diverse set of variables and features will tend to require more
-principal components to capture the same amount of variance. Our 27
-principal components are still a relatively high number of variables but
-this is not surprising as we chose a diverse conceptualization and a
-diverse set of time series features. As for the interpretability, PCA
-allows users to extract information on the meaning of the principal
-components. In particular, because the principal components are linear
-combination of original features, users can extract the relative
-importance of each feature for the extracted principal components (i.e.,
-the eigenvectors). While this can be useful in understanding the
-variance in the original data or help with manual feature selection, we
-use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original
-features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
 For our illustration of the feature clustering procedure, we use the
@@
 participants to cluster 1 and 76 participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in
-their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see
-\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
-features are generally stronger in separating the clusters (e.g., `how
-cooperative an interaction was' compared to `attitudes towards the
-Dutch'). Additionally, we see that (2) within the variables some
-features are better at distinguishing clusters (e.g., median of
-well-being vs.~MAC of well-being).We then inspect the clusters with a
-focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
-is the same data as for the variable focus, we can see more clearly that
-some features are better at distinguishing the clusters across variables
-(e.g., mean and median compared to auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the
-second cluster had more meaningful and fulfilling outgroup interactions
-also consistently reported more voluntary and cooperative interactions
-but less accidental and involuntary interactions. The same cluster also
-reported an increase in need fulfilling interaction over the 30 day
-period and an increase in interactions that were representative of the
-outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. This
-`deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see
-\fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, well-being are relatively stable
-for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability over time (also see
-\fgrref[A]{fig:clusterTs}).
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as
-well as check whether the developmental clusters are associated with
-important social markers and individual differences. To illustrate
-artifact checks, we added the number of measurements into the comparison
-and find that the participants in the detereoration cluster on average
-completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions (see
-\fgrref[B]{fig:clusterTs}). While this difference could indicate that
-the clusters might not entirely be comparable in the response patterns,
-we can find some relief in our data exclusion procedures during which we
-ensured that the general time frame and completion rates were not too
-dissimilar. To illustrate the utility of individual differences, we
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
-participants in the deteriorating cluster reported substantially higher
-levels of everyday discrimination. Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}

### 2023-02-17 methods-and-results.qmd
:construction: section-based re-render

--- 2023-01-20 methods-and-results.Rmd
@@
 output: latex_fragment
-editor_options:
-  chunk_output_type: console
+# format: 
+#   latex:
+#     standalone: false
+#format: context
 bibliography: referencesZotero.bib
 csl: apa.csl
+execute:
+  include: false
+  #echo: false
+  #eval: false
+  warning: false
+  error: false
+  #cache: true
+  freeze: auto
 ---
 
-```{r setup, include=FALSE}
+```{r}
+#| label: setup
+#| include: false
+
 # R Studio Clean-Up
@@
 
-```{r import data, include=FALSE}
+```{r}
+#| label: import data
+#| include: false
+
 # Load variable name lookup table
@@
 
-```{r Preparation of Variable Names, include=FALSE}
+```{r}
+#| label: Preparation of Variable Names
+#| include: false
+
 varNamIndicesS1 <- c(
@@
 
-```{r variable names for all analyses, include=FALSE}
+```{r}
+#| label: variable names for all analyses
+#| include: false
+
 # ID Variables (just to re-iterate)
@@
 
-```{r variable names by response type and study, include=FALSE}
+```{r}
+#| label: variable names by response type and study
+#| include: false
+
 varNames %>%
@@
 
-```{r variables included in main analysis by response type, include=FALSE}
+```{r}
+#| label: variables included in main analysis by response type
+#| include: false
+
 varNames %>%
@@
 
-```{r missingness filtering, include=FALSE}
+```{r}
+#| label: missingness filtering
+
 missInfo <- function(full = NA, reduced = NA){
@@
 \subsection{Data}
+
 We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
@@
 \subsubsection{Sample}
+
 The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
@@
 This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45\% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
 This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
@@
 \subsubsection{Variables}
+
 We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
@@
 
-```{r descrWide, include=FALSE}
+```{r}
+#| label: descrWide
+
 variableLab <- c(
@@
 
-```{r descrLong, include=FALSE}
+```{r}
+#| label: descrLong
+
 # Sample Descriptives Table --- longer version
@@
 \subsection{Analysis and Results}
+
 To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
@@
 \subsubsection{Feature extraction}
-```{r featMiss, include=FALSE}
+
+```{r}
+#| label: featMiss
+
 featureExclusion <- c(
@@
 
+```{=tex}
 \begin{equation} \label{eq:median}
@@
 \end{equation}
-
+```
 For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
 
+```{=tex}
 \begin{equation} \label{eq:mad}
@@
 \end{equation}
-
+```
 mean absolute change (MAC) for stability
 
+```{=tex}
 \begin{equation} \label{eq:mac}
@@
 \end{equation}
-
+```
 lag--1 autocorrelation for inertia: lag $l=1$
 
+```{=tex}
 \begin{equation} \label{eq:ar}
@@
 \end{equation}
-
+```
 overall linear regression slope to describe the linear trend
 
+```{=tex}
 \begin{equation} \label{eq:lin}
@@
 \end{equation}
-
+```
 summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
@@
 
+```{=tex}
 \begin{equation} \label{eq:edf}
@@
 \end{equation}
-
+```
 \textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
@@
 
-After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
+After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
 This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
@@
 \subsubsection{Feature reduction}
-```{r pca, include=FALSE} 
+
+```{r}
+#| label: pca
+
 # prepare data
@@
 We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
-A common rule of thumb is to use the principal components that jointly explain 70--90\% of the original variance (i.e., cumulative percentage explained variance).
-For our illustration we select the first `r pca_cutoff` principal components that explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
+A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance (i.e., cumulative percentage explained variance).
+For our illustration we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
 For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
@@
 \subsubsection{Feature clustering}
-```{r featureClustering, include=FALSE}
+
+```{r}
+#| label: featureClustering
+
 # load libraries
@@
 
-```{r k-means modeling based on selected k, include=FALSE}
+```{r}
+#| label: k-means modeling based on selected k
+
 kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
@@
 
-```{r Plot Convex Hulls, include=FALSE}
+```{r}
+#| label: Plot Convex Hulls
+
 fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
@@
 The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
+
 <!-- I had originally written:  -->
+
 <!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
+
 <!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
@@
 
+```{=tex}
 \begin{figure}[!ht] %hbtp
@@
 \end{figure}
-
+```
 We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
@@
 \end{figure}
-
-```{r average feature by cluster, include=FALSE}
+```
+
+```{r}
+#| label: average feature by cluster
+
 # CLUSTER VARIABLE LABELS
@@
 
-```{r clusterVariableGrid, include=FALSE}
+```{r}
+#| label: clusterVariableGrid
+
 clusterVariableGrid <- featCluster %>% 
@@
 
-```{r clusterFeatureGrid, include=FALSE}
+```{r}
+#| label: clusterFeatureGrid
+
 clusterFeatureGrid <- featClusterVal %>% 
@@
 
-```{r clusterFeatVarComb, include=FALSE}
+```{r}
+#| label: clusterFeatVarComb
+
 clusterFeatVarComb <- ggpubr::ggarrange(

### 2023-02-17 sections/01DataDescription.Rmd
:construction: section-based re-render

--- 2023-02-17 methods-and-results.qmd
@@
 output: latex_fragment
-# format: 
-#   latex:
-#     standalone: false
-#format: context
-bibliography: referencesZotero.bib
-csl: apa.csl
-execute:
-  include: false
-  #echo: false
-  #eval: false
-  warning: false
-  error: false
-  #cache: true
-  freeze: auto
+bibliography: ../referencesZotero.bib
+csl: ../apa.csl
 ---
@@
 ```{r}
-#| label: setup
+#| label: section setup
 #| include: false
-
-# R Studio Clean-Up
-cat("\014") # clear console
-# rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
-gc() # garbage collector
-
-# Install and Load Packages
-# !IMPORTANT!
-# BEFORE FIRST RENDER:
-# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
-# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
-lib <- c(
-  "rmarkdown",
-  "knitr",
-  "remedy",
-  "bookdown",
-  #"tidyverse",
-  "MASS",
-  "brms",
-  "psych",
-  "ggplot2",
-  "ggthemes",
-  "haven",
-  "RColorBrewer",
-  "plotly",
-  "grid",
-  "gridExtra",
-  "ggpattern",
-  "lme4",
-  "nlme",
-  "jtools",
-  "gtsummary",
-  "sessioninfo",
-  "tibble",
-  "pander",
-  "devtools",
-  "mada",
-  "data.table",
-  "plyr",
-  "dplyr",
-  "tidyr",
-  "Hmisc",
-  "kableExtra",
-  "papaja",
-  "stringr",
-  "stringi",
-  "reshape2",
-  "anytime",
-  "lubridate",
-  "purrr",
-  "metafor",
-  "dygraphs",
-  "readxl", 
-  "reshape",
-  "factoextra",
-  "Amelia", 
-  "ThreeWay",
-  "DescTools",
-  # GAM
-  "gratia",
-  #"MASS",
-  "mgcv",
-  # For moving window models
-  "Kendall",
-  # For the Hamed & Rao correction of tau
-  "modifiedmk", 
-  "mgcViz",
-  "shiny"
-)
-invisible(lapply(lib, library, character.only = TRUE))
-rm(lib)
-
-anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
-anytime::addFormats("%Y-%m-%d %H:%M:%S")
-anytime::addFormats("%d-%m-%Y")  
-#options(rgl.useNULL = TRUE)
-
-# Load Custom Packages
-source("./scripts/functions/fun.panel.R")
-source("./scripts/functions/themes.R")
-source("./scripts/functions/binaryCor.R")
-source("./scripts/functions/MlCorMat.R")
-source("./scripts/functions/MlTbl.R")
-source("./scripts/functions/metaLmer.R")
-source("./scripts/functions/meanViz.R")
-source("./scripts/functions/gam.R")
-source("./scripts/functions/tsFeatureExtractor.R")
-
-# Markdown Options
-knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
-knitr::opts_knit$get("root.dir") # check working directory
-options(
-  scipen = 999,
-  digits = 4,
-  width = 400
-) # removes scientific quotation
-# knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings
 
@@
   fig.height = 8,
-  fig.path = "Figures/",
-  echo = TRUE,
+  fig.path = "../Figures/",
+  include = FALSE,
   warning = FALSE,
   message = FALSE
-)
-```
-
-```{r}
-#| label: import data
-#| include: false
-
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
-
-# import full datasets
-load("data/s123Full.RData")
-
-# import reduced datasets
-load("data/s123Red.RData")
-
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
-
-```{r}
-#| label: Preparation of Variable Names
-#| include: false
-
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r}
-#| label: variable names for all analyses
-#| include: false
-
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r}
-#| label: variable names by response type and study
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: variables included in main analysis by response type
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
 )
@@
 
-\subsection{Data}
-
 We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
@@
 For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
-We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
 
@@
 \input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
-For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REF).
-
-\subsubsection{Feature extraction}
-
-```{r}
-#| label: featMiss
-
-featureExclusion <- c(
-  "mean",
-  "sd",
-  "rmssd",
-  "ar02",
-  "edf_ar",
-  "_n"
-)
-
-pMissFull <- pMissFeat(
-  all = featFull$features %>% select(-ends_with(featureExclusion)),
-  contact = featFullContact$features,
-  nocontact = featFullNoContact$features,
-  title = "Feature-wise Missingess Across all Studies"
-)
-```
-
-For the features, we selected and extracted one of each of the features proposed in \tblref{tab:esmFeatures}.
-We particularly chose the more robust median and median absolute deviation (MAD) for central tendency and variability.
-Within the variability structure, we chose the slightly more intuitive mean absolute change (MAC) for stability and stayed with the common lag--1 autocorrelation for inertia.
-For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We chose these features in particular because we consider them to be the most broadly applicable features.
-We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
-The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
-
-```{=tex}
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-```
-For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
-
-```{=tex}
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-```
-mean absolute change (MAC) for stability
-
-```{=tex}
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-```
-lag--1 autocorrelation for inertia: lag $l=1$
-
-```{=tex}
-\begin{equation} \label{eq:ar}
-  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ijt-l}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
-\end{equation}
-```
-overall linear regression slope to describe the linear trend
-
-```{=tex}
-\begin{equation} \label{eq:lin}
-  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
-\end{equation}
-```
-summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We estimate the number of parametric and non-parametric (or smooth) terms $p$ in the model.
-
-```{=tex}
-\begin{equation} \label{eq:edf}
-  edf \approx \sum p
-\end{equation}
-```
-\textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
-This includes the number of parameters used to estimate each smooth function in the model, as well as any additional parameters used for the overall model.On the other hand, the effective degrees of freedom is a measure of the amount of information in the data that is used to estimate the model.
-It takes into account the smoothness of the estimated functions in the model, and is typically smaller than the estimated degrees of freedom.
-Effective degrees of freedom is used to estimate the uncertainty of the model estimates and to calculate the model's goodness of fit.
-In general, the effective degrees of freedom is a more appropriate measure of the complexity of the model than the estimated degrees of freedom, as it accounts for the smoothness of the estimated functions, which can help to prevent overfitting.
-This is because if the estimated function is smooth, it means that the model is not fitting the noise and has a better generalization.
-
-After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
-This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
-The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
-The few missing values can, however, be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
-
-\subsubsection{Feature reduction}
-
-```{r}
-#| label: pca
-
-# prepare data
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion), ends_with("_n"))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-res.pca <- prcomp(scaled_data, scale. = FALSE)
-summary(res.pca)
-
-res.varExplained <- data.frame(
-  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
-  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
-)
-fviz_eig(res.pca, addlabels = TRUE)
-ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
-  geom_point() +
-  geom_line() +
-  theme_Publication()
-
-pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= 0.8))
-
-plot_ly(
-  res.pca$rotation %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Variable:', rownames(res.pca$rotation))
-) %>%
-  layout(title = 'Variables in reduced space')
-plot_ly(
-  res.pca$x %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Participant:', rownames(res.pca$x))
-) %>%
-  layout(title = 'Participants in reduced space')
-
-# extract coordinates for the individuals (PC-scores)
-pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
-```
-
-For the feature reduction, we chose the common \textit{principal component analysis} (PCA).
-Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established within the psychometric literature and can broadly be applied to a wide variety of studies in an automatize manner.
-As our aim is to present a general illustration that can also be adopted for more general data descriptive uses, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
-We then enter all `r ncol(scaled_data)` features into the analysis.
-The PCA will use linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
-The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
-In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
-We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
-A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance (i.e., cumulative percentage explained variance).
-For our illustration we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
-For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
-As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables.
-Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
-Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
-As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
-In particular, because the principal components are linear combination of original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
-While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-```{r}
-#| label: featureClustering
-
-# load libraries
-library(factoextra)
-library(NbClust)
-
-# check variance differences
-max(pca.out %>% summarize_all(var))/min(pca.out %>% summarize_all(var))
-pca.out %>%
-  summarize_all(sd) %>%
-  pivot_longer(everything(), names_to = "PC", values_to = "sd") %>%
-  ggplot(., aes(x=reorder(PC, sd), y=sd, label=sd)) +
-  geom_bar(stat='identity', width=.5) +
-  coord_flip() +
-  theme_Publication()
-
-# prep data
-scaled_pc <- pca.out #%>% scale
-
-# Elbow method
-fviz_nbclust(scaled_pc, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method") +
-  theme_Publication()
-# Silhouette method
-fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method") +
-  theme_Publication()
-```
-
-```{r}
-#| label: k-means modeling based on selected k
-
-kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
-```
-
-```{r}
-#| label: Plot Convex Hulls
-
-fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
-fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
-```
-
-<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
-However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
-If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
-
-<!-- I had originally written:  -->
-
-<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
-
-<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
-
-For our illustration of the feature clustering procedure, we use the generalizable and efficient centroid-based k-means clustering.
-The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see Supplemental Material A).
-We then entered the participants' PC-scores into the k-means algorithm.
-We used the recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100 random initial centroid positions to avoid convergence to a sub-optimal solution (i.e., local minima).
-The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
-We first inspect the clusters with a focus on the included variables (see \fgrref[A]{fig:clusterFeatVar}).
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
-Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
-This offers some information on which features were are most important in understanding the two extracted groups.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-```
-We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
-The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).
-
-<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
-While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
-To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterTs} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
-Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}
-```
-
-```{r}
-#| label: average feature by cluster
-
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-# ADD VALIDATION VARS:
-validationNam <- c(
-  "gender.pre",
-  "IntGrAnx01.pre",
-  "IntGrAnx02.pre",
-  "IntGrAnx03.pre",
-  "IntGrAnx04.pre",
-  "IntGrAnx05.pre",
-  "IntGrAnx06.pre",
-  "IntGrAnx07R.pre",
-  "IntGrAnx08R.pre",
-  "IntGrAnx09R.pre",
-  "IntGrAnx10R.pre",
-  "IntGrAnx11R.pre",
-  "IntGrAnx12R.pre",
-  "SWL01.pre",
-  "SWL02.pre",
-  "SWL03.pre",
-  "SWL04.pre",
-  "SWL05.pre",
-  "EvDayDiscr01.post",
-  "EvDayDiscr02.post",
-  "EvDayDiscr03.post",
-  "EvDayDiscr04.post",
-  "EvDayDiscr05.post",
-  "EvDayDiscr06.post",
-  "EvDayDiscr07.post",
-  "EvDayDiscr08.post",
-  "EvDayDiscr09.post"
-)
-
-# EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
-validation <- rbind(
-  dtS1Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S1", gender.pre = (gender.pre-2)*-1),
-  dtS2Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S3")
-) %>% 
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  group_by(study, ID) %>%
-  summarise(
-    gender = gender.pre,
-    IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
-    SWL.pre = rowMeans(across(starts_with("SWL")), na.rm = TRUE),
-    EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE)
-  ) %>%
-  ungroup %>%
-  distinct %>% 
-  mutate(
-    gender.z = scale(gender, scale = TRUE),
-    IntGrAnx.pre.z = scale(IntGrAnx.pre, scale = TRUE),
-    SWL.pre.z = scale(SWL.pre, scale = TRUE),
-    EvDayDiscr.post.z = scale(EvDayDiscr.post, scale = TRUE)
-  ) %>%
-  select(
-    ID,
-    #gender.z,
-    #IntGrAnx.pre.z,
-    #SWL.pre.z,
-    EvDayDiscr.post.z
-  )
-val <- validation %>% select(-ID) %>% names
-validationFeature <- validation
-for(feature in variableLab$variable) {
-  validationFeature <- validationFeature %>% bind_cols(., select(., setNames(val, paste(feature, val, sep = "_"))))
-}
-validationFeature <- validationFeature %>% select(-any_of(val))
-
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-validationLab <- c(
-  "gender.z" = "Val: Gender",
-  "SWL.pre.z" = "Val: Sat. with life",
-  "IntGrAnx.pre.z" = "Val: Intergroup Anxiety",
-  "EvDayDiscr.post.z" = " Discrimination"
-) %>% enframe(., name = "variable", value = "label")
-
-featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>%
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-featClusterVal <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  merge(., validationFeature) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>% 
-  mutate(feature = stri_replace_all_regex(
-    feature,
-    pattern=validationLab$variable,
-    replacement=validationLab$label,
-    vectorize=FALSE
-  )) %>% 
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-rawCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  merge(., variableLab)
-```
-
-```{r}
-#| label: clusterVariableGrid
-
-clusterVariableGrid <- featCluster %>% 
-  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ label) +
-  theme_Publication()
-clusterVariableGrid
-```
-
-```{r}
-#| label: clusterFeatureGrid
-
-clusterFeatureGrid <- featClusterVal %>% 
-  ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ factor(feature, levels = unique(featClusterVal$feature)), ncol=4) +
-  theme_Publication()
-clusterFeatureGrid
-```
-
-```{r}
-#| label: clusterFeatVarComb
-
-clusterFeatVarComb <- ggpubr::ggarrange(
-  clusterVariableGrid + theme(axis.title.x = element_blank()),
-  clusterFeatureGrid,
-  ncol = 1,
-  labels = c("(A) Variable Focus", "(B) Feature Focus"),
-  common.legend = TRUE,
-  legend = "bottom",
-  align = "v"
-) 
-ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 12)
-```
-
-```{r clusterTS, include=FALSE}
-clusterTS <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication() 
-clusterTS
-```
-
-```{r clusterTSTrend, include=FALSE}
-clusterTSTrend <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
-  #stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-clusterTSTrend
-```
-
-```{r clusterTsComb, include=FALSE}
-clusterTsComb <- ggpubr::ggarrange(
-  clusterTS + theme(axis.title.x = element_blank()),
-  clusterTSTrend,
-  ncol = 1,
-  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
-  common.legend = TRUE,
-  legend = "bottom"
-)
-ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 12)
-```

### 2023-02-17 sections/02InputVariables.Rmd
:construction: section-based re-render

--- 2023-02-17 methods-and-results.qmd
@@
   fig.height = 8,
-  fig.path = "Figures/",
-  echo = TRUE,
+  fig.path = "../Figures/",
+  include = FALSE,
   warning = FALSE,
@@
 ```
-
-```{r}
-#| label: import data
-#| include: false
-
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
-
-# import full datasets
-load("data/s123Full.RData")
-
-# import reduced datasets
-load("data/s123Red.RData")
-
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
-
-```{r}
-#| label: Preparation of Variable Names
-#| include: false
-
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r}
-#| label: variable names for all analyses
-#| include: false
-
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r}
-#| label: variable names by response type and study
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: variables included in main analysis by response type
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: missingness filtering
-
-missInfo <- function(full = NA, reduced = NA){
-  missTab <- data.frame(
-    nFull = nrow(full),
-    nRed = nrow(reduced),
-    pptFull = length(unique(full$PID)),
-    pptRed = length(unique(reduced$PID)),
-    timeFull = length(unique(full$TID)),
-    timeRed = length(unique(reduced$TID))
-  ) %>%
-    mutate(
-      nDif = nFull-nRed,
-      nDifPerc = nDif/nFull*100,
-      pptDif = pptFull-pptRed,
-      pptDifPerc = pptDif/pptFull*100,
-      timeDif = timeFull-timeRed,
-      timeDifPerc = timeDif/timeFull*100
-    ) %>%
-    select(
-      nFull, nRed, nDif, nDifPerc,
-      pptFull, pptRed, pptDif, pptDifPerc,
-      timeFull, timeRed, timeDif, timeDifPerc
-    )
-    missTab
-}
-
-missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
-missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
-missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)
-
-missS123 <- rbind(
-  missS1,
-  missS2,
-  missS3
-) %>%
-  mutate(study = c(1, 2, 3)) %>%
-  select(study, everything())
-
-missS123 %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.)-1)),
-      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
-      digits=2,
-      caption = "Missingness Info by Study") %>%
-  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\subsection{Data}
-
-We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
-After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
-The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
-After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
-Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-\subsubsection{Variables}
 

### 2023-02-17 sections/03FeatureExtraction.Rmd
:construction: section-based re-render

--- 2023-02-17 methods-and-results.qmd
@@
 ```
-
-```{r}
-#| label: import data
-#| include: false
-
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
-
-# import full datasets
-load("data/s123Full.RData")
-
-# import reduced datasets
-load("data/s123Red.RData")
-
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
-
-```{r}
-#| label: Preparation of Variable Names
-#| include: false
-
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r}
-#| label: variable names for all analyses
-#| include: false
-
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r}
-#| label: variable names by response type and study
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: variables included in main analysis by response type
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: missingness filtering
-
-missInfo <- function(full = NA, reduced = NA){
-  missTab <- data.frame(
-    nFull = nrow(full),
-    nRed = nrow(reduced),
-    pptFull = length(unique(full$PID)),
-    pptRed = length(unique(reduced$PID)),
-    timeFull = length(unique(full$TID)),
-    timeRed = length(unique(reduced$TID))
-  ) %>%
-    mutate(
-      nDif = nFull-nRed,
-      nDifPerc = nDif/nFull*100,
-      pptDif = pptFull-pptRed,
-      pptDifPerc = pptDif/pptFull*100,
-      timeDif = timeFull-timeRed,
-      timeDifPerc = timeDif/timeFull*100
-    ) %>%
-    select(
-      nFull, nRed, nDif, nDifPerc,
-      pptFull, pptRed, pptDif, pptDifPerc,
-      timeFull, timeRed, timeDif, timeDifPerc
-    )
-    missTab
-}
-
-missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
-missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
-missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)
-
-missS123 <- rbind(
-  missS1,
-  missS2,
-  missS3
-) %>%
-  mutate(study = c(1, 2, 3)) %>%
-  select(study, everything())
-
-missS123 %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.)-1)),
-      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
-      digits=2,
-      caption = "Missingness Info by Study") %>%
-  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\subsection{Data}
-
-We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
-After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
-The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
-After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
-Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
-We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
-
-```{r}
-#| label: descrWide
-
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-dtAll <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-allMlCor <-
-  MlCorMat(
-    data = dtAll,
-    id = "ID",
-    selection = c(
-      "InteractionContextAccidental", 
-      "InteractionContextvoluntary", 
-      "InteractionContextCooperative",
-      "InteractionContextRepresentativeNL",
-      "qualityMeaning",
-      "qualityOverall", 
-      "KeyNeedFulfillment", 
-      "KeyNeedDueToPartner", 
-      "AttitudesPartner",
-      "DaytimeNeedFulfillment",
-      "AttitudesDutch",
-      "exWB"
-    ),
-    labels = c(
-      "Int: Accidental", 
-      "Int: Voluntary", 
-      "Int: Cooperative",
-      "Int: Representative",
-      "Int: Meaningful",
-      "Int: Quality", 
-      "Core Need", 
-      "Core Need Due to Partner", 
-      "Attitude Partner",
-      "Daytime Core Need",
-      "Outgroup Attitude",
-      "Well-being"
-    )
-  )
-
-# Sample Descriptives Table --- wider version
-allMlCor %>%
-  t %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
-  kable_classic() %>%
-  add_header_above(
-    .,
-    c(
-      "",
-      "Correlations" = ncol(allMlCor),
-      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-    )
-  ) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
-```
-
-```{r}
-#| label: descrLong
-
-# Sample Descriptives Table --- longer version
-allMlCor %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.) - 1)),
-      escape = FALSE,
-      col.names = c(
-        #"",
-        "\\makecell{Int: \\\\ Accidental}",
-        "\\makecell{Int: \\\\ Voluntary}", 
-        "\\makecell{Int: \\\\ Cooperative}",
-        "\\makecell{Int: \\\\ Representative}",
-        "\\makecell{Int: \\\\ Meaningful}",
-        "\\makecell{Int: \\\\ Quality}", 
-        "Core Need", 
-        "\\makecell{Core Need \\\\ Due to Partner}", 
-        "\\makecell{Attitude \\\\ Partner}",
-        "\\makecell{Daytime \\\\ Core Need}",
-        "\\makecell{Outgroup \\\\ Attitude}",
-        "Well-being"
-      )) %>%
-  kable_classic() %>%
-  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
-  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrLong.tex")
-```
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
-For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REF).
-
-\subsubsection{Feature extraction}
 
@@
 ```
+
 \textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
@@
 We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
-
-\subsubsection{Feature reduction}
-
-```{r}
-#| label: pca
-
-# prepare data
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion), ends_with("_n"))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-res.pca <- prcomp(scaled_data, scale. = FALSE)
-summary(res.pca)
-
-res.varExplained <- data.frame(
-  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
-  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
-)
-fviz_eig(res.pca, addlabels = TRUE)
-ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
-  geom_point() +
-  geom_line() +
-  theme_Publication()
-
-pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= 0.8))
-
-plot_ly(
-  res.pca$rotation %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Variable:', rownames(res.pca$rotation))
-) %>%
-  layout(title = 'Variables in reduced space')
-plot_ly(
-  res.pca$x %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Participant:', rownames(res.pca$x))
-) %>%
-  layout(title = 'Participants in reduced space')
-
-# extract coordinates for the individuals (PC-scores)
-pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
-```
-
-For the feature reduction, we chose the common \textit{principal component analysis} (PCA).
-Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established within the psychometric literature and can broadly be applied to a wide variety of studies in an automatize manner.
-As our aim is to present a general illustration that can also be adopted for more general data descriptive uses, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
-We then enter all `r ncol(scaled_data)` features into the analysis.
-The PCA will use linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
-The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
-In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
-We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
-A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance (i.e., cumulative percentage explained variance).
-For our illustration we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
-For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
-As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables.
-Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
-Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
-As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
-In particular, because the principal components are linear combination of original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
-While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-```{r}
-#| label: featureClustering
-
-# load libraries
-library(factoextra)
-library(NbClust)
-
-# check variance differences
-max(pca.out %>% summarize_all(var))/min(pca.out %>% summarize_all(var))
-pca.out %>%
-  summarize_all(sd) %>%
-  pivot_longer(everything(), names_to = "PC", values_to = "sd") %>%
-  ggplot(., aes(x=reorder(PC, sd), y=sd, label=sd)) +
-  geom_bar(stat='identity', width=.5) +
-  coord_flip() +
-  theme_Publication()
-
-# prep data
-scaled_pc <- pca.out #%>% scale
-
-# Elbow method
-fviz_nbclust(scaled_pc, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method") +
-  theme_Publication()
-# Silhouette method
-fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method") +
-  theme_Publication()
-```
-
-```{r}
-#| label: k-means modeling based on selected k
-
-kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
-```
-
-```{r}
-#| label: Plot Convex Hulls
-
-fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
-fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
-```
-
-<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
-However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
-If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
-
-<!-- I had originally written:  -->
-
-<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
-
-<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
-
-For our illustration of the feature clustering procedure, we use the generalizable and efficient centroid-based k-means clustering.
-The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see Supplemental Material A).
-We then entered the participants' PC-scores into the k-means algorithm.
-We used the recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100 random initial centroid positions to avoid convergence to a sub-optimal solution (i.e., local minima).
-The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
-We first inspect the clusters with a focus on the included variables (see \fgrref[A]{fig:clusterFeatVar}).
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
-Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
-This offers some information on which features were are most important in understanding the two extracted groups.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-```
-We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
-The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).
-
-<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
-While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
-To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterTs} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
-Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}
-```
-
-```{r}
-#| label: average feature by cluster
-
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-# ADD VALIDATION VARS:
-validationNam <- c(
-  "gender.pre",
-  "IntGrAnx01.pre",
-  "IntGrAnx02.pre",
-  "IntGrAnx03.pre",
-  "IntGrAnx04.pre",
-  "IntGrAnx05.pre",
-  "IntGrAnx06.pre",
-  "IntGrAnx07R.pre",
-  "IntGrAnx08R.pre",
-  "IntGrAnx09R.pre",
-  "IntGrAnx10R.pre",
-  "IntGrAnx11R.pre",
-  "IntGrAnx12R.pre",
-  "SWL01.pre",
-  "SWL02.pre",
-  "SWL03.pre",
-  "SWL04.pre",
-  "SWL05.pre",
-  "EvDayDiscr01.post",
-  "EvDayDiscr02.post",
-  "EvDayDiscr03.post",
-  "EvDayDiscr04.post",
-  "EvDayDiscr05.post",
-  "EvDayDiscr06.post",
-  "EvDayDiscr07.post",
-  "EvDayDiscr08.post",
-  "EvDayDiscr09.post"
-)
-
-# EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
-validation <- rbind(
-  dtS1Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S1", gender.pre = (gender.pre-2)*-1),
-  dtS2Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S3")
-) %>% 
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  group_by(study, ID) %>%
-  summarise(
-    gender = gender.pre,
-    IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
-    SWL.pre = rowMeans(across(starts_with("SWL")), na.rm = TRUE),
-    EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE)
-  ) %>%
-  ungroup %>%
-  distinct %>% 
-  mutate(
-    gender.z = scale(gender, scale = TRUE),
-    IntGrAnx.pre.z = scale(IntGrAnx.pre, scale = TRUE),
-    SWL.pre.z = scale(SWL.pre, scale = TRUE),
-    EvDayDiscr.post.z = scale(EvDayDiscr.post, scale = TRUE)
-  ) %>%
-  select(
-    ID,
-    #gender.z,
-    #IntGrAnx.pre.z,
-    #SWL.pre.z,
-    EvDayDiscr.post.z
-  )
-val <- validation %>% select(-ID) %>% names
-validationFeature <- validation
-for(feature in variableLab$variable) {
-  validationFeature <- validationFeature %>% bind_cols(., select(., setNames(val, paste(feature, val, sep = "_"))))
-}
-validationFeature <- validationFeature %>% select(-any_of(val))
-
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-validationLab <- c(
-  "gender.z" = "Val: Gender",
-  "SWL.pre.z" = "Val: Sat. with life",
-  "IntGrAnx.pre.z" = "Val: Intergroup Anxiety",
-  "EvDayDiscr.post.z" = " Discrimination"
-) %>% enframe(., name = "variable", value = "label")
-
-featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>%
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-featClusterVal <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  merge(., validationFeature) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>% 
-  mutate(feature = stri_replace_all_regex(
-    feature,
-    pattern=validationLab$variable,
-    replacement=validationLab$label,
-    vectorize=FALSE
-  )) %>% 
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-rawCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  merge(., variableLab)
-```
-
-```{r}
-#| label: clusterVariableGrid
-
-clusterVariableGrid <- featCluster %>% 
-  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ label) +
-  theme_Publication()
-clusterVariableGrid
-```
-
-```{r}
-#| label: clusterFeatureGrid
-
-clusterFeatureGrid <- featClusterVal %>% 
-  ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ factor(feature, levels = unique(featClusterVal$feature)), ncol=4) +
-  theme_Publication()
-clusterFeatureGrid
-```
-
-```{r}
-#| label: clusterFeatVarComb
-
-clusterFeatVarComb <- ggpubr::ggarrange(
-  clusterVariableGrid + theme(axis.title.x = element_blank()),
-  clusterFeatureGrid,
-  ncol = 1,
-  labels = c("(A) Variable Focus", "(B) Feature Focus"),
-  common.legend = TRUE,
-  legend = "bottom",
-  align = "v"
-) 
-ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 12)
-```
-
-```{r clusterTS, include=FALSE}
-clusterTS <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication() 
-clusterTS
-```
-
-```{r clusterTSTrend, include=FALSE}
-clusterTSTrend <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
-  #stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-clusterTSTrend
-```
-
-```{r clusterTsComb, include=FALSE}
-clusterTsComb <- ggpubr::ggarrange(
-  clusterTS + theme(axis.title.x = element_blank()),
-  clusterTSTrend,
-  ncol = 1,
-  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
-  common.legend = TRUE,
-  legend = "bottom"
-)
-ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 12)
-```

### 2023-02-17 sections/04FeatureReduction.Rmd
:construction: section-based re-render

--- 2023-01-17 methods-and-results.Rmd
@@
 output: latex_fragment
-editor_options:
-  chunk_output_type: console
-bibliography: referencesZotero.bib
-csl: apa.csl
+bibliography: ../referencesZotero.bib
+csl: ../apa.csl
 ---
 
-```{r setup, include=FALSE}
-# R Studio Clean-Up
-cat("\014") # clear console
-# rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
-gc() # garbage collector
-
-# Install and Load Packages
-# !IMPORTANT!
-# BEFORE FIRST RENDER:
-# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
-# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
-lib <- c(
-  "rmarkdown",
-  "knitr",
-  "remedy",
-  "bookdown",
-  #"tidyverse",
-  "MASS",
-  "brms",
-  "psych",
-  "ggplot2",
-  "ggthemes",
-  "haven",
-  "RColorBrewer",
-  "plotly",
-  "grid",
-  "gridExtra",
-  "ggpattern",
-  "lme4",
-  "nlme",
-  "jtools",
-  "gtsummary",
-  "sessioninfo",
-  "tibble",
-  "pander",
-  "devtools",
-  "mada",
-  "data.table",
-  "plyr",
-  "dplyr",
-  "tidyr",
-  "Hmisc",
-  "kableExtra",
-  "papaja",
-  "stringr",
-  "stringi",
-  "reshape2",
-  "anytime",
-  "lubridate",
-  "purrr",
-  "metafor",
-  "dygraphs",
-  "readxl", 
-  "reshape",
-  "factoextra",
-  "Amelia", 
-  "ThreeWay",
-  "DescTools",
-  # GAM
-  "gratia",
-  #"MASS",
-  "mgcv",
-  # For moving window models
-  "Kendall",
-  # For the Hamed & Rao correction of tau
-  "modifiedmk", 
-  "mgcViz",
-  "shiny"
-)
-invisible(lapply(lib, library, character.only = TRUE))
-rm(lib)
-
-anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
-anytime::addFormats("%Y-%m-%d %H:%M:%S")
-anytime::addFormats("%d-%m-%Y")  
-#options(rgl.useNULL = TRUE)
-
-# Load Custom Packages
-source("./scripts/functions/fun.panel.R")
-source("./scripts/functions/themes.R")
-source("./scripts/functions/binaryCor.R")
-source("./scripts/functions/MlCorMat.R")
-source("./scripts/functions/MlTbl.R")
-source("./scripts/functions/metaLmer.R")
-source("./scripts/functions/meanViz.R")
-source("./scripts/functions/gam.R")
-source("./scripts/functions/tsFeatureExtractor.R")
-
-# Markdown Options
-knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
-knitr::opts_knit$get("root.dir") # check working directory
-options(
-  scipen = 999,
-  digits = 4,
-  width = 400
-) # removes scientific quotation
-# knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings
+```{r}
+#| label: section setup
+#| include: false
 
@@
 
-```{r import data, include=FALSE}
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
+```{r}
+#| label: pca
 
-# import full datasets
-load("data/s123Full.RData")
-
-# import reduced datasets
-load("data/s123Red.RData")
-
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
-
-```{r Preparation of Variable Names, include=FALSE}
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r variable names for all analyses, include=FALSE}
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r variable names by response type and study, include=FALSE}
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r variables included in main analysis by response type, include=FALSE}
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r missingness filtering, include=FALSE}
-missInfo <- function(full = NA, reduced = NA){
-  missTab <- data.frame(
-    nFull = nrow(full),
-    nRed = nrow(reduced),
-    pptFull = length(unique(full$PID)),
-    pptRed = length(unique(reduced$PID)),
-    timeFull = length(unique(full$TID)),
-    timeRed = length(unique(reduced$TID))
-  ) %>%
-    mutate(
-      nDif = nFull-nRed,
-      nDifPerc = nDif/nFull*100,
-      pptDif = pptFull-pptRed,
-      pptDifPerc = pptDif/pptFull*100,
-      timeDif = timeFull-timeRed,
-      timeDifPerc = timeDif/timeFull*100
-    ) %>%
-    select(
-      nFull, nRed, nDif, nDifPerc,
-      pptFull, pptRed, pptDif, pptDifPerc,
-      timeFull, timeRed, timeDif, timeDifPerc
-    )
-    missTab
-}
-
-missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
-missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
-missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)
-
-missS123 <- rbind(
-  missS1,
-  missS2,
-  missS3
-) %>%
-  mutate(study = c(1, 2, 3)) %>%
-  select(study, everything())
-
-missS123 %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.)-1)),
-      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
-      digits=2,
-      caption = "Missingness Info by Study") %>%
-  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\subsection{Data}
-We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
-After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
-The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
-After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
-Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45\% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-\subsubsection{Variables}
-We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
-We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
-
-```{r descrWide, include=FALSE}
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-dtAll <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-allMlCor <-
-  MlCorMat(
-    data = dtAll,
-    id = "ID",
-    selection = c(
-      "InteractionContextAccidental", 
-      "InteractionContextvoluntary", 
-      "InteractionContextCooperative",
-      "InteractionContextRepresentativeNL",
-      "qualityMeaning",
-      "qualityOverall", 
-      "KeyNeedFulfillment", 
-      "KeyNeedDueToPartner", 
-      "AttitudesPartner",
-      "DaytimeNeedFulfillment",
-      "AttitudesDutch",
-      "exWB"
-    ),
-    labels = c(
-      "Int: Accidental", 
-      "Int: Voluntary", 
-      "Int: Cooperative",
-      "Int: Representative",
-      "Int: Meaningful",
-      "Int: Quality", 
-      "Core Need", 
-      "Core Need Due to Partner", 
-      "Attitude Partner",
-      "Daytime Core Need",
-      "Outgroup Attitude",
-      "Well-being"
-    )
-  )
-
-# Sample Descriptives Table --- wider version
-allMlCor %>%
-  t %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
-  kable_classic() %>%
-  add_header_above(
-    .,
-    c(
-      "",
-      "Correlations" = ncol(allMlCor),
-      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-    )
-  ) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
-```
-
-```{r descrLong, include=FALSE}
-# Sample Descriptives Table --- longer version
-allMlCor %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.) - 1)),
-      escape = FALSE,
-      col.names = c(
-        #"",
-        "\\makecell{Int: \\\\ Accidental}",
-        "\\makecell{Int: \\\\ Voluntary}", 
-        "\\makecell{Int: \\\\ Cooperative}",
-        "\\makecell{Int: \\\\ Representative}",
-        "\\makecell{Int: \\\\ Meaningful}",
-        "\\makecell{Int: \\\\ Quality}", 
-        "Core Need", 
-        "\\makecell{Core Need \\\\ Due to Partner}", 
-        "\\makecell{Attitude \\\\ Partner}",
-        "\\makecell{Daytime \\\\ Core Need}",
-        "\\makecell{Outgroup \\\\ Attitude}",
-        "Well-being"
-      )) %>%
-  kable_classic() %>%
-  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
-  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrLong.tex")
-```
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
-For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REF).
-
-\subsubsection{Feature extraction}
-```{r featMiss, include=FALSE}
-featureExclusion <- c(
-  "mean",
-  "sd",
-  "rmssd",
-  "ar02",
-  "edf_ar",
-  "_n"
-)
-
-pMissFull <- pMissFeat(
-  all = featFull$features %>% select(-ends_with(featureExclusion)),
-  contact = featFullContact$features,
-  nocontact = featFullNoContact$features,
-  title = "Feature-wise Missingess Across all Studies"
-)
-```
-
-For the features, we selected and extracted one of each of the features proposed in \tblref{tab:esmFeatures}.
-We particularly chose the more robust median and median absolute deviation (MAD) for central tendency and variability.
-Within the variability structure, we chose the slightly more intuitive mean absolute change (MAC) for stability and stayed with the common lag--1 autocorrelation for inertia.
-For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We chose these features in particular because we consider them to be the most broadly applicable features.
-We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
-The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag $l=1$
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
-This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
-The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
-The few missing values can, however, be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
-
-\subsubsection{Feature reduction}
-```{r pca, include=FALSE} 
 # prepare data
@@
 We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-```{r featureClustering, include=FALSE}
-# load libraries
-library(factoextra)
-library(NbClust)
-
-# check variance differences
-max(pca.out %>% summarize_all(var))/min(pca.out %>% summarize_all(var))
-pca.out %>%
-  summarize_all(sd) %>%
-  pivot_longer(everything(), names_to = "PC", values_to = "sd") %>%
-  ggplot(., aes(x=reorder(PC, sd), y=sd, label=sd)) +
-  geom_bar(stat='identity', width=.5) +
-  coord_flip() +
-  theme_Publication()
-
-# prep data
-scaled_pc <- pca.out #%>% scale
-
-# Elbow method
-fviz_nbclust(scaled_pc, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method") +
-  theme_Publication()
-# Silhouette method
-fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method") +
-  theme_Publication()
-```
-
-```{r k-means modeling based on selected k, include=FALSE}
-kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
-```
-
-```{r Plot Convex Hulls, include=FALSE}
-fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
-fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
-```
-
-<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
-However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
-If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
-<!-- I had originally written:  -->
-<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
-<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
-
-For our illustration of the feature clustering procedure, we use the generalizable and efficient centroid-based k-means clustering.
-The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see Supplemental Material A).
-We then entered the participants' PC-scores into the k-means algorithm.
-We used the recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100 random initial centroid positions to avoid convergence to a sub-optimal solution (i.e., local minima).
-The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
-We first inspect the clusters with a focus on the included variables (see \fgrref[A]{fig:clusterFeatVar}).
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
-Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
-This offers some information on which features were are most important in understanding the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
-The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).
-
-<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
-While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
-To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterTs} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
-Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}
-
-```{r average feature by cluster, include=FALSE}
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-# ADD VALIDATION VARS:
-validationNam <- c(
-  "gender.pre",
-  "IntGrAnx01.pre",
-  "IntGrAnx02.pre",
-  "IntGrAnx03.pre",
-  "IntGrAnx04.pre",
-  "IntGrAnx05.pre",
-  "IntGrAnx06.pre",
-  "IntGrAnx07R.pre",
-  "IntGrAnx08R.pre",
-  "IntGrAnx09R.pre",
-  "IntGrAnx10R.pre",
-  "IntGrAnx11R.pre",
-  "IntGrAnx12R.pre",
-  "SWL01.pre",
-  "SWL02.pre",
-  "SWL03.pre",
-  "SWL04.pre",
-  "SWL05.pre",
-  "EvDayDiscr01.post",
-  "EvDayDiscr02.post",
-  "EvDayDiscr03.post",
-  "EvDayDiscr04.post",
-  "EvDayDiscr05.post",
-  "EvDayDiscr06.post",
-  "EvDayDiscr07.post",
-  "EvDayDiscr08.post",
-  "EvDayDiscr09.post"
-)
-
-# EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
-validation <- rbind(
-  dtS1Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S1", gender.pre = (gender.pre-2)*-1),
-  dtS2Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S3")
-) %>% 
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  group_by(study, ID) %>%
-  summarise(
-    gender = gender.pre,
-    IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
-    SWL.pre = rowMeans(across(starts_with("SWL")), na.rm = TRUE),
-    EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE)
-  ) %>%
-  ungroup %>%
-  distinct %>% 
-  mutate(
-    gender.z = scale(gender, scale = TRUE),
-    IntGrAnx.pre.z = scale(IntGrAnx.pre, scale = TRUE),
-    SWL.pre.z = scale(SWL.pre, scale = TRUE),
-    EvDayDiscr.post.z = scale(EvDayDiscr.post, scale = TRUE)
-  ) %>%
-  select(
-    ID,
-    #gender.z,
-    #IntGrAnx.pre.z,
-    #SWL.pre.z,
-    EvDayDiscr.post.z
-  )
-val <- validation %>% select(-ID) %>% names
-validationFeature <- validation
-for(feature in variableLab$variable) {
-  validationFeature <- validationFeature %>% bind_cols(., select(., setNames(val, paste(feature, val, sep = "_"))))
-}
-validationFeature <- validationFeature %>% select(-any_of(val))
-
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-validationLab <- c(
-  "gender.z" = "Val: Gender",
-  "SWL.pre.z" = "Val: Sat. with life",
-  "IntGrAnx.pre.z" = "Val: Intergroup Anxiety",
-  "EvDayDiscr.post.z" = " Discrimination"
-) %>% enframe(., name = "variable", value = "label")
-
-featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>%
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-featClusterVal <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  merge(., validationFeature) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>% 
-  mutate(feature = stri_replace_all_regex(
-    feature,
-    pattern=validationLab$variable,
-    replacement=validationLab$label,
-    vectorize=FALSE
-  )) %>% 
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-rawCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  merge(., variableLab)
-```
-
-```{r clusterVariableGrid, include=FALSE}
-clusterVariableGrid <- featCluster %>% 
-  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ label) +
-  theme_Publication()
-clusterVariableGrid
-```
-
-```{r clusterFeatureGrid, include=FALSE}
-clusterFeatureGrid <- featClusterVal %>% 
-  ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ factor(feature, levels = unique(featClusterVal$feature)), ncol=4) +
-  theme_Publication()
-clusterFeatureGrid
-```
-
-```{r clusterFeatVarComb, include=FALSE}
-clusterFeatVarComb <- ggpubr::ggarrange(
-  clusterVariableGrid + theme(axis.title.x = element_blank()),
-  clusterFeatureGrid,
-  ncol = 1,
-  labels = c("(A) Variable Focus", "(B) Feature Focus"),
-  common.legend = TRUE,
-  legend = "bottom",
-  align = "v"
-) 
-ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 12)
-```
-
-```{r clusterTS, include=FALSE}
-clusterTS <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication() 
-clusterTS
-```
-
-```{r clusterTSTrend, include=FALSE}
-clusterTSTrend <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
-  #stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-clusterTSTrend
-```
-
-```{r clusterTsComb, include=FALSE}
-clusterTsComb <- ggpubr::ggarrange(
-  clusterTS + theme(axis.title.x = element_blank()),
-  clusterTSTrend,
-  ncol = 1,
-  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
-  common.legend = TRUE,
-  legend = "bottom"
-)
-ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 12)
-```

### 2023-02-17 sections/05FeatureClustering.Rmd
:construction: section-based re-render

--- 2023-01-17 methods-and-results.Rmd
@@
 
-```{r import data, include=FALSE}
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
+```{r}
+#| label: featureClustering
 
-# import full datasets
-load("data/s123Full.RData")
-
-# import reduced datasets
-load("data/s123Red.RData")
-
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
-
-```{r Preparation of Variable Names, include=FALSE}
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r variable names for all analyses, include=FALSE}
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r variable names by response type and study, include=FALSE}
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r variables included in main analysis by response type, include=FALSE}
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r missingness filtering, include=FALSE}
-missInfo <- function(full = NA, reduced = NA){
-  missTab <- data.frame(
-    nFull = nrow(full),
-    nRed = nrow(reduced),
-    pptFull = length(unique(full$PID)),
-    pptRed = length(unique(reduced$PID)),
-    timeFull = length(unique(full$TID)),
-    timeRed = length(unique(reduced$TID))
-  ) %>%
-    mutate(
-      nDif = nFull-nRed,
-      nDifPerc = nDif/nFull*100,
-      pptDif = pptFull-pptRed,
-      pptDifPerc = pptDif/pptFull*100,
-      timeDif = timeFull-timeRed,
-      timeDifPerc = timeDif/timeFull*100
-    ) %>%
-    select(
-      nFull, nRed, nDif, nDifPerc,
-      pptFull, pptRed, pptDif, pptDifPerc,
-      timeFull, timeRed, timeDif, timeDifPerc
-    )
-    missTab
-}
-
-missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
-missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
-missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)
-
-missS123 <- rbind(
-  missS1,
-  missS2,
-  missS3
-) %>%
-  mutate(study = c(1, 2, 3)) %>%
-  select(study, everything())
-
-missS123 %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.)-1)),
-      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
-      digits=2,
-      caption = "Missingness Info by Study") %>%
-  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\subsection{Data}
-We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
-After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
-The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
-After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
-Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45\% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-\subsubsection{Variables}
-We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
-We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
-
-```{r descrWide, include=FALSE}
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-dtAll <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-allMlCor <-
-  MlCorMat(
-    data = dtAll,
-    id = "ID",
-    selection = c(
-      "InteractionContextAccidental", 
-      "InteractionContextvoluntary", 
-      "InteractionContextCooperative",
-      "InteractionContextRepresentativeNL",
-      "qualityMeaning",
-      "qualityOverall", 
-      "KeyNeedFulfillment", 
-      "KeyNeedDueToPartner", 
-      "AttitudesPartner",
-      "DaytimeNeedFulfillment",
-      "AttitudesDutch",
-      "exWB"
-    ),
-    labels = c(
-      "Int: Accidental", 
-      "Int: Voluntary", 
-      "Int: Cooperative",
-      "Int: Representative",
-      "Int: Meaningful",
-      "Int: Quality", 
-      "Core Need", 
-      "Core Need Due to Partner", 
-      "Attitude Partner",
-      "Daytime Core Need",
-      "Outgroup Attitude",
-      "Well-being"
-    )
-  )
-
-# Sample Descriptives Table --- wider version
-allMlCor %>%
-  t %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
-  kable_classic() %>%
-  add_header_above(
-    .,
-    c(
-      "",
-      "Correlations" = ncol(allMlCor),
-      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-    )
-  ) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
-```
-
-```{r descrLong, include=FALSE}
-# Sample Descriptives Table --- longer version
-allMlCor %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.) - 1)),
-      escape = FALSE,
-      col.names = c(
-        #"",
-        "\\makecell{Int: \\\\ Accidental}",
-        "\\makecell{Int: \\\\ Voluntary}", 
-        "\\makecell{Int: \\\\ Cooperative}",
-        "\\makecell{Int: \\\\ Representative}",
-        "\\makecell{Int: \\\\ Meaningful}",
-        "\\makecell{Int: \\\\ Quality}", 
-        "Core Need", 
-        "\\makecell{Core Need \\\\ Due to Partner}", 
-        "\\makecell{Attitude \\\\ Partner}",
-        "\\makecell{Daytime \\\\ Core Need}",
-        "\\makecell{Outgroup \\\\ Attitude}",
-        "Well-being"
-      )) %>%
-  kable_classic() %>%
-  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
-  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrLong.tex")
-```
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
-For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REF).
-
-\subsubsection{Feature extraction}
-```{r featMiss, include=FALSE}
-featureExclusion <- c(
-  "mean",
-  "sd",
-  "rmssd",
-  "ar02",
-  "edf_ar",
-  "_n"
-)
-
-pMissFull <- pMissFeat(
-  all = featFull$features %>% select(-ends_with(featureExclusion)),
-  contact = featFullContact$features,
-  nocontact = featFullNoContact$features,
-  title = "Feature-wise Missingess Across all Studies"
-)
-```
-
-For the features, we selected and extracted one of each of the features proposed in \tblref{tab:esmFeatures}.
-We particularly chose the more robust median and median absolute deviation (MAD) for central tendency and variability.
-Within the variability structure, we chose the slightly more intuitive mean absolute change (MAC) for stability and stayed with the common lag--1 autocorrelation for inertia.
-For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We chose these features in particular because we consider them to be the most broadly applicable features.
-We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
-The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag $l=1$
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`\% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
-This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
-The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
-The few missing values can, however, be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
-
-\subsubsection{Feature reduction}
-```{r pca, include=FALSE} 
-# prepare data
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion), ends_with("_n"))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-res.pca <- prcomp(scaled_data, scale. = FALSE)
-summary(res.pca)
-
-res.varExplained <- data.frame(
-  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
-  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
-)
-fviz_eig(res.pca, addlabels = TRUE)
-ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
-  geom_point() +
-  geom_line() +
-  theme_Publication()
-
-pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= 0.8))
-
-plot_ly(
-  res.pca$rotation %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Variable:', rownames(res.pca$rotation))
-) %>%
-  layout(title = 'Variables in reduced space')
-plot_ly(
-  res.pca$x %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Participant:', rownames(res.pca$x))
-) %>%
-  layout(title = 'Participants in reduced space')
-
-# extract coordinates for the individuals (PC-scores)
-pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
-```
-
-For the feature reduction, we chose the common \textit{principal component analysis} (PCA).
-Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established within the psychometric literature and can broadly be applied to a wide variety of studies in an automatize manner.
-As our aim is to present a general illustration that can also be adopted for more general data descriptive uses, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
-We then enter all `r ncol(scaled_data)` features into the analysis.
-The PCA will use linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
-The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
-In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
-We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
-A common rule of thumb is to use the principal components that jointly explain 70--90\% of the original variance (i.e., cumulative percentage explained variance).
-For our illustration we select the first `r pca_cutoff` principal components that explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).
-For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
-As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables.
-Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
-Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
-As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
-In particular, because the principal components are linear combination of original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
-While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-```{r featureClustering, include=FALSE}
 # load libraries
@@
 The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
+
 <!-- I had originally written:  -->
@@
 The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
-
-In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
-We first inspect the clusters with a focus on the included variables (see \fgrref[A]{fig:clusterFeatVar}).
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
-Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
-This offers some information on which features were are most important in understanding the two extracted groups.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-
-We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
-Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
-The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).
-
-<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->
-
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
-While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
-To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterTs} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
-Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
-
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
-\end{figure}
-
-```{r average feature by cluster, include=FALSE}
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-# ADD VALIDATION VARS:
-validationNam <- c(
-  "gender.pre",
-  "IntGrAnx01.pre",
-  "IntGrAnx02.pre",
-  "IntGrAnx03.pre",
-  "IntGrAnx04.pre",
-  "IntGrAnx05.pre",
-  "IntGrAnx06.pre",
-  "IntGrAnx07R.pre",
-  "IntGrAnx08R.pre",
-  "IntGrAnx09R.pre",
-  "IntGrAnx10R.pre",
-  "IntGrAnx11R.pre",
-  "IntGrAnx12R.pre",
-  "SWL01.pre",
-  "SWL02.pre",
-  "SWL03.pre",
-  "SWL04.pre",
-  "SWL05.pre",
-  "EvDayDiscr01.post",
-  "EvDayDiscr02.post",
-  "EvDayDiscr03.post",
-  "EvDayDiscr04.post",
-  "EvDayDiscr05.post",
-  "EvDayDiscr06.post",
-  "EvDayDiscr07.post",
-  "EvDayDiscr08.post",
-  "EvDayDiscr09.post"
-)
-
-# EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
-validation <- rbind(
-  dtS1Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S1", gender.pre = (gender.pre-2)*-1),
-  dtS2Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(idVars), any_of(validationNam)) %>% mutate(study = "S3")
-) %>% 
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  group_by(study, ID) %>%
-  summarise(
-    gender = gender.pre,
-    IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
-    SWL.pre = rowMeans(across(starts_with("SWL")), na.rm = TRUE),
-    EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE)
-  ) %>%
-  ungroup %>%
-  distinct %>% 
-  mutate(
-    gender.z = scale(gender, scale = TRUE),
-    IntGrAnx.pre.z = scale(IntGrAnx.pre, scale = TRUE),
-    SWL.pre.z = scale(SWL.pre, scale = TRUE),
-    EvDayDiscr.post.z = scale(EvDayDiscr.post, scale = TRUE)
-  ) %>%
-  select(
-    ID,
-    #gender.z,
-    #IntGrAnx.pre.z,
-    #SWL.pre.z,
-    EvDayDiscr.post.z
-  )
-val <- validation %>% select(-ID) %>% names
-validationFeature <- validation
-for(feature in variableLab$variable) {
-  validationFeature <- validationFeature %>% bind_cols(., select(., setNames(val, paste(feature, val, sep = "_"))))
-}
-validationFeature <- validationFeature %>% select(-any_of(val))
-
-# CLUSTER VARIABLE LABELS
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-validationLab <- c(
-  "gender.z" = "Val: Gender",
-  "SWL.pre.z" = "Val: Sat. with life",
-  "IntGrAnx.pre.z" = "Val: Intergroup Anxiety",
-  "EvDayDiscr.post.z" = " Discrimination"
-) %>% enframe(., name = "variable", value = "label")
-
-featCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>%
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-featClusterVal <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., z_data) %>%
-  merge(., validationFeature) %>%
-  group_by(cluster) %>%
-  summarise(across(everything(), list(mean = mean, se = se), na.rm=TRUE)) %>%
-  ungroup %>% 
-  select(-starts_with("ID")) %>%
-  pivot_longer(
-    cols = -c(cluster),
-    names_to = c("variable", ".value"),
-    names_pattern = "(.*)_(.*)"
-  ) %>% 
-  mutate(
-    nam = variable,
-    variable = gsub("\\_.*", "", nam),
-    feature = gsub("^.*?\\_", "", nam)
-  ) %>% 
-  merge(., variableLab) %>% 
-  mutate(feature = stri_replace_all_regex(
-    feature,
-    pattern=validationLab$variable,
-    replacement=validationLab$label,
-    vectorize=FALSE
-  )) %>% 
-  select(
-    cluster,
-    nam,
-    variable,
-    label,
-    feature,
-    everything()
-  )
-
-rawCluster <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
-  merge(., raw_data) %>%
-  select(
-    ID,
-    TIDnum,
-    cluster,
-    any_of(varNamS123)
-  ) %>%
-  reshape2::melt(id.vars = c("cluster", "ID","TIDnum")) %>% 
-  merge(., variableLab)
-```
-
-```{r clusterVariableGrid, include=FALSE}
-clusterVariableGrid <- featCluster %>% 
-  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ label) +
-  theme_Publication()
-clusterVariableGrid
-```
-
-```{r clusterFeatureGrid, include=FALSE}
-clusterFeatureGrid <- featClusterVal %>% 
-  ggplot(., aes(x = label, y = mean, group = cluster, color = as.factor(cluster))) +
-  geom_point() +
-  geom_line() +
-  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
-  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
-  coord_flip() +
-  labs(
-    x = "Features",
-    y = "Standardized Mean",
-    color = "Cluster"
-  ) +
-  facet_wrap(~ factor(feature, levels = unique(featClusterVal$feature)), ncol=4) +
-  theme_Publication()
-clusterFeatureGrid
-```
-
-```{r clusterFeatVarComb, include=FALSE}
-clusterFeatVarComb <- ggpubr::ggarrange(
-  clusterVariableGrid + theme(axis.title.x = element_blank()),
-  clusterFeatureGrid,
-  ncol = 1,
-  labels = c("(A) Variable Focus", "(B) Feature Focus"),
-  common.legend = TRUE,
-  legend = "bottom",
-  align = "v"
-) 
-ggsave("Figures/clusterFeatVarComb.pdf", clusterFeatVarComb, width = 12, height = 12)
-```
-
-```{r clusterTS, include=FALSE}
-clusterTS <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication() 
-clusterTS
-```
-
-```{r clusterTSTrend, include=FALSE}
-clusterTSTrend <- rawCluster %>% 
-  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
-  #stat_summary(fun=mean, geom="line") +
-  facet_wrap(~ label, scales = "free_y") +
-  labs(
-    x = "Time ID",
-    y = "Response (Mean Cluster Response Bold)",
-    color = "Cluster"
-  ) +
-  theme_Publication()
-clusterTSTrend
-```
-
-```{r clusterTsComb, include=FALSE}
-clusterTsComb <- ggpubr::ggarrange(
-  clusterTS + theme(axis.title.x = element_blank()),
-  clusterTSTrend,
-  ncol = 1,
-  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
-  common.legend = TRUE,
-  legend = "bottom"
-)
-ggsave("Figures/clusterTsComb.pdf", clusterTsComb, width = 12, height = 12)
-```

### 2023-02-17 sections/06ClusterEvaluation.Rmd
:construction: section-based re-render

--- 2023-02-17 methods-and-results.qmd
@@
 ```
-
-```{r}
-#| label: import data
-#| include: false
-
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
-
-# import full datasets
-load("data/s123Full.RData")
-
-# import reduced datasets
-load("data/s123Red.RData")
-
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
-
-```{r}
-#| label: Preparation of Variable Names
-#| include: false
-
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r}
-#| label: variable names for all analyses
-#| include: false
-
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r}
-#| label: variable names by response type and study
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: variables included in main analysis by response type
-#| include: false
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-```{r}
-#| label: missingness filtering
-
-missInfo <- function(full = NA, reduced = NA){
-  missTab <- data.frame(
-    nFull = nrow(full),
-    nRed = nrow(reduced),
-    pptFull = length(unique(full$PID)),
-    pptRed = length(unique(reduced$PID)),
-    timeFull = length(unique(full$TID)),
-    timeRed = length(unique(reduced$TID))
-  ) %>%
-    mutate(
-      nDif = nFull-nRed,
-      nDifPerc = nDif/nFull*100,
-      pptDif = pptFull-pptRed,
-      pptDifPerc = pptDif/pptFull*100,
-      timeDif = timeFull-timeRed,
-      timeDifPerc = timeDif/timeFull*100
-    ) %>%
-    select(
-      nFull, nRed, nDif, nDifPerc,
-      pptFull, pptRed, pptDif, pptDifPerc,
-      timeFull, timeRed, timeDif, timeDifPerc
-    )
-    missTab
-}
-
-missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
-missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
-missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)
-
-missS123 <- rbind(
-  missS1,
-  missS2,
-  missS3
-) %>%
-  mutate(study = c(1, 2, 3)) %>%
-  select(study, everything())
-
-missS123 %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.)-1)),
-      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
-      digits=2,
-      caption = "Missingness Info by Study") %>%
-  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-\subsection{Data}
-
-We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
-After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
-The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
-After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
-Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
-We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
-
-```{r}
-#| label: descrWide
-
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-dtAll <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-allMlCor <-
-  MlCorMat(
-    data = dtAll,
-    id = "ID",
-    selection = c(
-      "InteractionContextAccidental", 
-      "InteractionContextvoluntary", 
-      "InteractionContextCooperative",
-      "InteractionContextRepresentativeNL",
-      "qualityMeaning",
-      "qualityOverall", 
-      "KeyNeedFulfillment", 
-      "KeyNeedDueToPartner", 
-      "AttitudesPartner",
-      "DaytimeNeedFulfillment",
-      "AttitudesDutch",
-      "exWB"
-    ),
-    labels = c(
-      "Int: Accidental", 
-      "Int: Voluntary", 
-      "Int: Cooperative",
-      "Int: Representative",
-      "Int: Meaningful",
-      "Int: Quality", 
-      "Core Need", 
-      "Core Need Due to Partner", 
-      "Attitude Partner",
-      "Daytime Core Need",
-      "Outgroup Attitude",
-      "Well-being"
-    )
-  )
-
-# Sample Descriptives Table --- wider version
-allMlCor %>%
-  t %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
-  kable_classic() %>%
-  add_header_above(
-    .,
-    c(
-      "",
-      "Correlations" = ncol(allMlCor),
-      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-    )
-  ) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
-```
-
-```{r}
-#| label: descrLong
-
-# Sample Descriptives Table --- longer version
-allMlCor %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.) - 1)),
-      escape = FALSE,
-      col.names = c(
-        #"",
-        "\\makecell{Int: \\\\ Accidental}",
-        "\\makecell{Int: \\\\ Voluntary}", 
-        "\\makecell{Int: \\\\ Cooperative}",
-        "\\makecell{Int: \\\\ Representative}",
-        "\\makecell{Int: \\\\ Meaningful}",
-        "\\makecell{Int: \\\\ Quality}", 
-        "Core Need", 
-        "\\makecell{Core Need \\\\ Due to Partner}", 
-        "\\makecell{Attitude \\\\ Partner}",
-        "\\makecell{Daytime \\\\ Core Need}",
-        "\\makecell{Outgroup \\\\ Attitude}",
-        "Well-being"
-      )) %>%
-  kable_classic() %>%
-  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
-  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrLong.tex")
-```
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the extract--reduce--cluster--evaluate steps in sequential order.
-For each of the steps fully reproducible and annotated code is available in Online Supplemental A as well as in the accompanying OSF- and GitHub repositories (REF).
-
-\subsubsection{Feature extraction}
-
-```{r}
-#| label: featMiss
-
-featureExclusion <- c(
-  "mean",
-  "sd",
-  "rmssd",
-  "ar02",
-  "edf_ar",
-  "_n"
-)
-
-pMissFull <- pMissFeat(
-  all = featFull$features %>% select(-ends_with(featureExclusion)),
-  contact = featFullContact$features,
-  nocontact = featFullNoContact$features,
-  title = "Feature-wise Missingess Across all Studies"
-)
-```
-
-For the features, we selected and extracted one of each of the features proposed in \tblref{tab:esmFeatures}.
-We particularly chose the more robust median and median absolute deviation (MAD) for central tendency and variability.
-Within the variability structure, we chose the slightly more intuitive mean absolute change (MAC) for stability and stayed with the common lag--1 autocorrelation for inertia.
-For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We chose these features in particular because we consider them to be the most broadly applicable features.
-We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
-The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
-
-```{=tex}
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-```
-For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
-
-```{=tex}
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-```
-mean absolute change (MAC) for stability
-
-```{=tex}
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-```
-lag--1 autocorrelation for inertia: lag $l=1$
-
-```{=tex}
-\begin{equation} \label{eq:ar}
-  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ijt-l}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
-\end{equation}
-```
-overall linear regression slope to describe the linear trend
-
-```{=tex}
-\begin{equation} \label{eq:lin}
-  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
-\end{equation}
-```
-summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We estimate the number of parametric and non-parametric (or smooth) terms $p$ in the model.
-
-```{=tex}
-\begin{equation} \label{eq:edf}
-  edf \approx \sum p
-\end{equation}
-```
-\textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
-This includes the number of parameters used to estimate each smooth function in the model, as well as any additional parameters used for the overall model.On the other hand, the effective degrees of freedom is a measure of the amount of information in the data that is used to estimate the model.
-It takes into account the smoothness of the estimated functions in the model, and is typically smaller than the estimated degrees of freedom.
-Effective degrees of freedom is used to estimate the uncertainty of the model estimates and to calculate the model's goodness of fit.
-In general, the effective degrees of freedom is a more appropriate measure of the complexity of the model than the estimated degrees of freedom, as it accounts for the smoothness of the estimated functions, which can help to prevent overfitting.
-This is because if the estimated function is smooth, it means that the model is not fitting the noise and has a better generalization.
-
-After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
-This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables.
-The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
-The few missing values can, however, be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
-
-\subsubsection{Feature reduction}
-
-```{r}
-#| label: pca
-
-# prepare data
-raw_data <- featData
-z_data <- featFullImp$featuresImpZ %>% select(-ends_with(featureExclusion), ends_with("_n"))
-scaled_data <- featFullImp$featuresImpZMat %>% select(-ends_with(featureExclusion))
-
-res.pca <- prcomp(scaled_data, scale. = FALSE)
-summary(res.pca)
-
-res.varExplained <- data.frame(
-  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
-  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
-)
-fviz_eig(res.pca, addlabels = TRUE)
-ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
-  geom_point() +
-  geom_line() +
-  theme_Publication()
-
-pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= 0.8))
-
-plot_ly(
-  res.pca$rotation %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Variable:', rownames(res.pca$rotation))
-) %>%
-  layout(title = 'Variables in reduced space')
-plot_ly(
-  res.pca$x %>% as.data.frame,
-  x = ~ PC1,
-  y = ~ PC2,
-  z = ~ PC3,
-  size = 2,
-  type = 'scatter3d',
-  mode = 'markers',
-  text = ~ paste('Participant:', rownames(res.pca$x))
-) %>%
-  layout(title = 'Participants in reduced space')
-
-# extract coordinates for the individuals (PC-scores)
-pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
-```
-
-For the feature reduction, we chose the common \textit{principal component analysis} (PCA).
-Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established within the psychometric literature and can broadly be applied to a wide variety of studies in an automatize manner.
-As our aim is to present a general illustration that can also be adopted for more general data descriptive uses, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
-We then enter all `r ncol(scaled_data)` features into the analysis.
-The PCA will use linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
-The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
-In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
-We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
-A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance (i.e., cumulative percentage explained variance).
-For our illustration we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
-For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
-As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables.
-Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
-Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
-As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
-In particular, because the principal components are linear combination of original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
-While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-```{r}
-#| label: featureClustering
-
-# load libraries
-library(factoextra)
-library(NbClust)
-
-# check variance differences
-max(pca.out %>% summarize_all(var))/min(pca.out %>% summarize_all(var))
-pca.out %>%
-  summarize_all(sd) %>%
-  pivot_longer(everything(), names_to = "PC", values_to = "sd") %>%
-  ggplot(., aes(x=reorder(PC, sd), y=sd, label=sd)) +
-  geom_bar(stat='identity', width=.5) +
-  coord_flip() +
-  theme_Publication()
-
-# prep data
-scaled_pc <- pca.out #%>% scale
-
-# Elbow method
-fviz_nbclust(scaled_pc, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method") +
-  theme_Publication()
-# Silhouette method
-fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method") +
-  theme_Publication()
-```
-
-```{r}
-#| label: k-means modeling based on selected k
-
-kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
-```
-
-```{r}
-#| label: Plot Convex Hulls
-
-fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
-fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
-```
-
-<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
-However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
-If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
-
-<!-- I had originally written:  -->
-
-<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
-
-<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
-
-For our illustration of the feature clustering procedure, we use the generalizable and efficient centroid-based k-means clustering.
-The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see Supplemental Material A).
-We then entered the participants' PC-scores into the k-means algorithm.
-We used the recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100 random initial centroid positions to avoid convergence to a sub-optimal solution (i.e., local minima).
-The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
 

### 2023-02-17 sections/01DataDescription.Rmd
:construction: figure folder fix

@@
   fig.height = 8,
-  fig.path = "../Figures/",
+  fig.path = "../figures/",
   include = FALSE,

### 2023-02-21 manuscript.tex
:construction: R sections added
:construction: input variables cleaned up

@@
 Recently, one promising way of identifying person-level developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
-This analysis type essentially seeks to capture comparable within-person developments --- often with some form of numeric summary --- and then groups the persons based on the extracted summaries \citep[][]{liao2005}\footnote{It should be noted in some cases the time series do not need to be summarized and can be compared directly.
+This analysis type essentially seeks to capture comparable within-person developments --- such as whether a variable remains stable over time, consistently increases, or exhibits cyclical patterns --- and then groups the persons based on these patterns  \citep[][]{liao2005}\footnote{It should be noted in some cases the time series do not need to be summarized and can be compared directly.
 Such analyses are however only possible with highly regular and controlled data, such as EEG data \citep{huang1985}, or when clustering variables within the person rather than identifying differences between persons \citep{haslbeck2022}.
@@
 To take the common VAR model as an example, the model explicitly assumes that the time intervals between measurements are consistent (i.e., equidistant measurement assumption), it does not allow for missing values, and it assumes that means or variances do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
-All of these violated assumptions reflect types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
+These assumptions however stand incontrast to types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
 Model parameters might, thus, not always accurately capture the time series and are often restricted in the features they capture.
@@
 As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-We argue that for psychological time series data, feature-based clustering offers a flexible variety, theory-based selection, fewer strict assumptions, easy and intuitive analysis, and meaningful psychological interpretability.
+ESM APPLICATION.
+We argue that for psychological time series data, feature-based clustering offers a flexible variety, theory-based selection, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.
 
@@
 To illustrate the functioning and utility of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected data from research on migration experiences.
-Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020} and this type of research exemplifies some of the real-world data issues that ESM data commonly faces.
-The new type of research comes as a response to prominent reviews within the migration literature, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-The migration ESM research, thus, offers a good example of the ESM data that introduces a multivariate conceptualization with event-specific missingness patterns.
+Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020}.
+This type of research comes as a response to prominent reviews within the migration literature, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
+Importantly for our illustration, the migration ESM research, thus, exemplifies some of the real-world data issues that ESM data commonly faces, including a multivariate conceptualization with event-specific missingness patterns.
 
-[...]
+\input{latex-results/01DataDescription.tex}
 
@@
 
-Introduction on the structure ...
-
-Armed with a selection of potentially meaningful and accurate time series features to address a given research question, we can consider the practical steps of feature-based time series clustering.
-The feature-based clustering analysis can be structured into four main steps \citep{rasanen2009, wang2006}.
+To introduce and illustrate the feature-based clustering analysis, we will follow the conceptual steps of the procedure in sequential order and discuss key issues for each step.
+To do so we will follow the common separation that has structured feature-based clustering into four main steps \citep{rasanen2009, wang2006}.
 (1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters.
-It is important to note that these steps are a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
-Nonetheless, below we will briefly describe each of the steps conceptually.
+While this is a common conceptual separation of procedural elements, it is important to note that these steps are a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
+Nonetheless, the conceptual nature of these steps allows us to introduce the mayor elements of the analysis.
 We also provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
@@
 
-\paragraph{Input variables}
+\subsection{Input variables}
 Time series clustering starts with the selection and preparation of the variables of interest.
@@
 Conceptually, the included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series.
-There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}, and many health developments are commonly captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}.
+One of the advantages of feature-based clustering is that it is inherently adept to accomodating multi-variate concepts --- a common aim in ESM research.
+There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, many health developments are captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}, and migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}.
 At the same time, however, the added number of variables can become a methodological concern.
@@
 
-Once the important variables have been selected, the data needs to be prepared for the analysis steps.
-Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
-Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses \citep{liao2005}.
+\input{latex-results/02InputVariables.tex}
 
-\paragraph{Feature extraction}
+\subsection{Feature extraction}
 The main feature extraction step ultimately depends on the choice of the selected time series features but generally speaking the process is comparatively simple.
@@
 
-\paragraph{Feature reduction}
+\input{latex-results/03FeatureExtraction.tex}
+
+\subsubsection{Feature reduction}
 Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm.
@@
 
-\paragraph{Feature clustering}
+\input{latex-results/04FeatureReduction.tex}
+
+\subsection{Feature clustering}
 For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible \citep{liao2005}.
@@
 
-\paragraph{Cluster Evaluation}
+\input{latex-results/05FeatureClustering.tex}
+
+\subsection{Cluster Evaluation}
 The final evaluation step includes two main elements, (1) evaluating the performance of the clustering analysis and (2) interpreting the extracted clusters conceptually.
@@
 
-% Methods and Results from RMarkdown render
-\input{methods-and-results}
-%\input{tables/descrWide} 
-%\input{tables/descrLong}
+\input{latex-results/06ClusterEvaluation.tex}
 

### 2023-02-21 latex-results/01DataDescription.tex
:memo: input variables adjusted

@@
-We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
-followed migrants who had recently arrived in the Netherlands in their
-daily interactions with the Dutch majority group members. After a
+The data set we use consists of three studies that followed migrants who
+had recently arrived in the Netherlands in their daily interactions with
+the Dutch majority group members \citet[][]{Kreienkamp2022b}. After a
 general migration-focused pre-questionnaire, participants were invited
@@
 partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Full
-methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
 
 The original studies included 207 participants (\(N_{S1}=\) 23,
-\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
-The studies differed substantially in the maximum length of
-participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
-\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
-continue participation without compensation in the later two studies. To
-make the three studies comparable in participation and time frames, we
-iteratively removed all measurement occasions and participants that had
-more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now a lot more comparable
-(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
-
-\input{tables/descrLong}
+\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 ESM
+measurements. Each of the studies focused on recently arrived
+first-generation migrants and each study included a number of
+ideosyncratic variables relevant for the broader research collective.
+For our empirical example we focus on the variables that were collected
+during the ESM surveys and were available in all three studies. Variable
+selection and preparation is described as part of the illustration below
+but additional methodological details about the study setup are
+available in \citet[][]{Kreienkamp2022b}.

### 2023-02-21 latex-results/02InputVariables.tex
:memo: input variables adjusted

@@
-We included 12 variables that were available across all three studies
-and captured information about the participant's interactions, as well
-as cognitive-, emotional-, and motivational self in relationship with
-the majority group. We chose these two aspects in particular because (1)
-the interaction-specific information exemplified the structural
-missingness issue of modern ESM data and (2) the motivational,
-emotional, and cognitive experience offered a diverse conceptualization
-of migration experience (beyond behavioral measurements) that is
-becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
-methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are also
-available in \tblref{tab:descrLong}.
+For our illustration, we include 12 variables that were available across
+all three studies and captured information about the participant's
+interactions, as well as cognitive-, emotional-, and motivational self
+in relationship with the majority group. We chose these aspects in
+particular because (1) the interaction-specific information exemplified
+the structural missingness issue of modern ESM data and (2) the
+motivational, emotional, and cognitive experience offered a diverse
+conceptualization of migration experience (beyond behavioral
+measurements) that is becoming more common in the literature
+\citep[][]{Kreienkamp2022d}. Full methodological details are available
+in Online Supplemental Material A, but basic item information,
+descriptives, and correlations are also available in
+\tblref{tab:descrLong}.
+
+Once the important variables have been selected, the data needs to be
+prepared for the analysis steps. Importantly, this not only means
+validating and cleaning the data (e.g., re-coding, removing duplicate or
+unwanted measurements) but also making the time series comparable. Two
+important steps are making the time-frames and response scales
+comparable across participants --- for example, by choosing a time frame
+that is common to most participants and standardizing the participants'
+responses \citep{liao2005}.
+
+In our illustration data set, the studies differed substantially in the
+maximum length of participation (\(\text{max}(t_{S1})=\) 63,
+\(\text{max}(t_{S2})=\) 69, \(\text{max}(t_{S3})=\) 155). This was
+likely due to the option to continue participation without compensation
+in the later two studies. To make the three studies comparable in
+participation and time frames, we iteratively removed all measurement
+occasions and participants that had more than 45\% missingness
+\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure let to a final sample of 157 participants, who jointly
+produced 8,132 measurements. Importantly, both the participant
+repsonse-patterns and the time frame were now substantially more
+comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
+\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
+procedure and study-specific exclusions see Online Supplemental Material
+A.
 

### 2023-02-21 sections/01DataDescription.Rmd
:memo: input variables adjusted

@@
 
-We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}.
-The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members.
+The data set we use consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members \citet[][]{Kreienkamp2022b}.
 After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
@@
 Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
+
+The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` ESM measurements.
+Each of the studies focused on recently arrived first-generation migrants and each study included a number of ideosyncratic variables relevant for the broader research collective.
 For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Full methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements.
-The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
-
-```{r}
-#| label: descrWide
-
-variableLab <- c(
-  "AttitudesDutch" = "Outgroup Attitude",
-  "AttitudesPartner" = "Int: Partner Attitude",
-  "DaytimeNeedFulfillment" = "Need Fulfillment",
-  "exWB" = "Well-Being",
-  "InteractionContextAccidental" = "Int: Accidental",
-  "InteractionContextCooperative" = "Int: Cooperative",
-  "InteractionContextRepresentativeNL" = "Int: Representative",
-  "InteractionContextvoluntary" = "Int: Voluntary",
-  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
-  "KeyNeedFulfillment" = "Int: Need Fulfillment",
-  "qualityMeaning" = "Int: Meaningful",
-  "qualityOverall" = "Int: Quality"
-) %>% enframe(., name = "variable", value = "label")
-
-dtAll <- rbind(
-  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
-    mutate(across(!TID & !study, as.numeric)),
-  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
-  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
-) %>%
-  group_by(study, PID) %>%
-  mutate(ID = cur_group_id()) %>%
-  ungroup %>%
-  mutate(date = as.Date(gsub(" .*", "", TID)),
-         week = strftime(date, format = "%Y-W%V")) %>%
-  select(ID,
-         PID,
-         study,
-         date,
-         week,
-         TIDnum,
-         any_of(varNamS123PCA)) %>%
-  arrange(ID, TIDnum) %>%
-  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
-
-allMlCor <-
-  MlCorMat(
-    data = dtAll,
-    id = "ID",
-    selection = c(
-      "InteractionContextAccidental", 
-      "InteractionContextvoluntary", 
-      "InteractionContextCooperative",
-      "InteractionContextRepresentativeNL",
-      "qualityMeaning",
-      "qualityOverall", 
-      "KeyNeedFulfillment", 
-      "KeyNeedDueToPartner", 
-      "AttitudesPartner",
-      "DaytimeNeedFulfillment",
-      "AttitudesDutch",
-      "exWB"
-    ),
-    labels = c(
-      "Int: Accidental", 
-      "Int: Voluntary", 
-      "Int: Cooperative",
-      "Int: Representative",
-      "Int: Meaningful",
-      "Int: Quality", 
-      "Core Need", 
-      "Core Need Due to Partner", 
-      "Attitude Partner",
-      "Daytime Core Need",
-      "Outgroup Attitude",
-      "Well-being"
-    )
-  )
-
-# Sample Descriptives Table --- wider version
-allMlCor %>%
-  t %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = c("l", rep("c", ncol(.) - 1))) %>%
-  kable_classic() %>%
-  add_header_above(
-    .,
-    c(
-      "",
-      "Correlations" = ncol(allMlCor),
-      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
-    )
-  ) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable}\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
-```
-
-```{r}
-#| label: descrLong
-
-# Sample Descriptives Table --- longer version
-allMlCor %>%
-  as.data.frame %>%
-  kbl(.,
-      format = "latex",
-      caption = "Correlation Table and Descriptive Statistics",
-      booktabs = TRUE,
-      align = "c", #c("l", rep("c", ncol(.) - 1)),
-      escape = FALSE,
-      col.names = c(
-        #"",
-        "\\makecell{Int: \\\\ Accidental}",
-        "\\makecell{Int: \\\\ Voluntary}", 
-        "\\makecell{Int: \\\\ Cooperative}",
-        "\\makecell{Int: \\\\ Representative}",
-        "\\makecell{Int: \\\\ Meaningful}",
-        "\\makecell{Int: \\\\ Quality}", 
-        "Core Need", 
-        "\\makecell{Core Need \\\\ Due to Partner}", 
-        "\\makecell{Attitude \\\\ Partner}",
-        "\\makecell{Daytime \\\\ Core Need}",
-        "\\makecell{Outgroup \\\\ Attitude}",
-        "Well-being"
-      )) %>%
-  kable_classic() %>%
-  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
-  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
-  footnote(
-    general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
-    )
-  ) %>%
-  kable_styling(latex_options = "scale_down") %>%
-  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
-  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrLong.tex")
-```
-
-\input{tables/descrLong}
+Variable selection and preparation is described as part of the illustration below but additional methodological details about the study setup are available in \citet[][]{Kreienkamp2022b}.

### 2023-02-21 sections/02InputVariables.Rmd
:memo: input variables adjusted

@@
 
-We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
-We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
+For our illustration, we include 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
+We chose these aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
 Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
+
+Once the important variables have been selected, the data needs to be prepared for the analysis steps.
+Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
+Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses \citep{liao2005}.
+
+In our illustration data set, the studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
+This was likely due to the option to continue participation without compensation in the later two studies.
+To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
+Importantly, both the participant repsonse-patterns and the time frame were now substantially more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
+For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
 

### 2023-02-21 tables/esmPsychFeatures.tex
:sparkles: abstract + discussion
:art: tbl placement
:memo: literature references added

@@
 %\begin{table}%[hbt]
-\begin{sidewaystable}
+\begin{sidewaystable*}[!hbtp]
     \centering
@@
         \hline 
-        Time Series Feature & Substantive Interpretation & Example Operationalizations & Formulas, refs, ...? \\ 
+        Time Series Feature & Substantive Interpretation & Example Operationalizations & Further Reading \\ 
         \hline \\ [-0.5em]
-        Diversity \newline \hl{(drop this \& keep in var. selection only?)} & 
-        Multivariate measurement of concepts (e.g., affect-behavior-cognition-desire, or bio-psycho-social) \linebreak & 
-        -----\linebreak  & 
-        {\centering --- ? ---\par} \\
+
+        %Diversity \newline \hl{(drop this \& keep in var. selection only?)} & 
+        %Multivariate measurement of concepts (e.g., affect-behavior-cognition-desire, or bio-psycho-social) \linebreak & 
+        %-----\linebreak  & 
+        %{\centering --- ? ---\par} \\
 
@@
         \end{itemize} \linebreak  & 
-        {\centering --- ? ---\par} \\
+        \citet{bringmann2018c}; \citet{weisberg1992} \\
 
@@
         \end{itemize} \linebreak & 
-        {\centering --- ? ---\par} \\
+        \citet{helmich2020a}; \linebreak\citet{vandeleemput2014} \\
 
@@
         \end{itemize} \linebreak & 
-        {\centering --- ? ---\par} \\
+        \citet{kivela2022}; \linebreak\citet{wichers2019} \\
 
-        Inertia & 
+        Self-similarity & 
         Describes how much experiences carry over to the future measurements.
@@
         \end{itemize} \linebreak & 
-        {\centering --- ? ---\par} \\
+        \citet{epskamp2018}; \linebreak\citet{kuppens2010} \\
 
@@
         \end{itemize} \linebreak & 
-        {\centering --- ? ---\par} \\
+        \citet{gottman1969}; \linebreak\citet{oravecz2016} \\
 
@@
         \end{itemize} \linebreak & 
-        {\centering --- ? ---\par} \\
+        \citet{caro-martin2018}; \linebreak\citet{castro-alvarez2022a} \\
 
@@
     \end{tabular}
-\end{sidewaystable}
+\end{sidewaystable*}

### 2023-02-21 IllustrationResultsRenderReport.Rmd
:construction: figure folder fix
:construction: tbl lab fix

@@
 subtitle: "Render report for 'A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series'"
-author:
-- ██████ ██████████^1^
-- ██████████ ████████^1^
-- █████ █████████^1^
-- █████ ██ █████^1^
-- ███ ███████^1^
-- ^1^██████████ ██ █████████, Department of Psychology
-- "Author Information:"
-- "Correspondence concerning this article should be addressed to ██████ ██████████, ██████████ ██ ██████████, █████████ ██ ██████████, █████ ██████████ ███, ████ ██ █████████ ████████████████.
-E-mail: █████████████████"
-- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
-- 'The data repository for this manuscript is available at <a href="https://osf.io/" target="_blank">osf.io/TBD</a>'
-- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
 date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
-output: 
-  bookdown::html_document2:
-    fig_caption: yes
-    md_extensions: +footnotes
-    code_folding: hide
-    mathjax: default
-    theme: yeti
-    toc: yes
-    toc_float: yes
-    number_sections: false
-    css: style.css
-    includes:
-      in_header: "_includes/head-custom-rmd.html" 
+output: pdf_document
 editor_options:
@@
   fig.height = 8,
-  fig.path = "Figures/",
+  fig.path = "figures/",
   echo = TRUE,
@@
   arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
+  mutate_all(~recode(.x, `0`="", ` 1`="X")) %>%
   relocate(starts_with("study"), .after = last_col()) %>%
@@
       caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
+  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3))
 ```
@@
 
-<details>
+<!-- <details> -->
 
-<summary>
+<!-- <summary> -->
 
@@
 
-</summary>
+<!-- </summary> -->
 
-\renewcommand{\arraystretch}{0.8}
+<!-- \renewcommand{\arraystretch}{0.8} -->
 
-<!-- decrease line spacing for the table -->
+<!-- <!-- decrease line spacing for the table --> -->
 
@@
   caption = "R environment session info for reproducibility of results" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
 )
@@
 
-\renewcommand{\arraystretch}{1}
+<!-- \renewcommand{\arraystretch}{1} -->
 
-<!-- reset row height/line spacing -->
+<!-- <!-- reset row height/line spacing --> -->
 
-</details>
+<!-- </details> -->
 
-<br>
+<!-- <br> -->
 
-<details>
+<!-- <details> -->
 
-<summary>
+<!-- <summary> -->
 
@@
 
-</summary>
+<!-- </summary> -->
 
-\renewcommand{\arraystretch}{0.6}
+<!-- \renewcommand{\arraystretch}{0.6} -->
 
-<!-- decrease line spacing for the table -->
+<!-- <!-- decrease line spacing for the table --> -->
 
@@
   caption = "Package info for reproducibility of results" # complete caption for main document
-) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
 )
@@
 
-</summary>
+<!-- </summary> -->
 
@@
 
-</details>
+<!-- </details> -->
 
-</br>
+<!-- </br> -->
 
@@
 
-# References
-
 ::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}

### 2023-02-21 latex-results/02InputVariables.tex
:construction: tbl lab fix

@@
 measurements) that is becoming more common in the literature
-\citep[][]{Kreienkamp2022d}. Full methodological details are available
-in Online Supplemental Material A, but basic item information,
-descriptives, and correlations are also available in
-\tblref{tab:descrLong}.
+\citep[][]{Kreienkamp2022d}. The breadth of the included variables also
+showcases the utility of the method for a growing body of literature
+that considers heterogeneous and complex concepts. As a result, the
+number of included variables is also on the higher end for psychological
+concepts and additionally allows us to showcase the efficiency benefits
+of the method and offers a reasonable use case for the optional feature
+reduction step.
 
@@
 comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
+\(\text{max}(t_{S3})=\) 67). Full methodological details are available
+in Online Supplemental Material A, but basic item information,
+descriptives, and correlations are also available in
+\tblref{tab:descrLong}.
 

### 2023-02-21 latex-results/03FeatureExtraction.tex
:construction: tbl lab fix

@@
 smooth function in the model, as well as any additional parameters used
-for the overall model.On the other hand, the effective degrees of
+for the overall model. On the other hand, the effective degrees of
 freedom is a measure of the amount of information in the data that is

### 2023-02-21 methods-and-results.qmd
:construction: tbl lab fix

@@
 #     standalone: false
-#format: context
 bibliography: referencesZotero.bib

### 2023-02-21 sections/02InputVariables.Rmd
:construction: tbl lab fix

@@
 We chose these aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
-Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
+The breadth of the included variables also showcases the utility of the method for a growing body of literature that considers heterogeneous and complex concepts.
+As a result, the number of included variables is also on the higher end for psychological concepts and additionally allows us to showcase the efficiency benefits of the method and offers a reasonable use case for the optional feature reduction step.
 
@@
 Importantly, both the participant repsonse-patterns and the time frame were now substantially more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
-For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.
+Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
 
 ```{r}
-#| label: descrWide
+#| label: descriptive-tbl-prep
 
@@
   )
+```
 
+```{r}
 # Sample Descriptives Table --- wider version
@@
       caption = "Correlation Table and Descriptive Statistics",
+      label = "descrWide",
       booktabs = TRUE,
@@
 ```{r}
-#| label: descrLong
-
 # Sample Descriptives Table --- longer version
@@
       escape = FALSE,
+      label = "descrLong",
       col.names = c(

### 2023-02-21 sections/03FeatureExtraction.Rmd
:construction: tbl lab fix

@@
 \textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
-This includes the number of parameters used to estimate each smooth function in the model, as well as any additional parameters used for the overall model.On the other hand, the effective degrees of freedom is a measure of the amount of information in the data that is used to estimate the model.
+This includes the number of parameters used to estimate each smooth function in the model, as well as any additional parameters used for the overall model.
+On the other hand, the effective degrees of freedom is a measure of the amount of information in the data that is used to estimate the model.
 It takes into account the smoothness of the estimated functions in the model, and is typically smaller than the estimated degrees of freedom.

### 2023-02-23 manuscript.tex
:construction: feature extraction intro

@@
 
-\subsection{Input variables}
+\subsection{Input Variables}
 Time series clustering starts with the selection and preparation of the variables of interest.
@@
 
-\subsection{Feature extraction}
-The main feature extraction step ultimately depends on the choice of the selected time series features but generally speaking the process is comparatively simple.
-For each of the included variables the features are extracted for each participant and replace the original time series \citep{wang2006}\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
-As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
+\subsection{Feature Extraction}
+Armed with a relevant selection of key variables, the main aim of the feature extraction is to describe the most important and meaningful aspects of a time series.
+In its most general approach, feature extraction can include any numeric summary of the time series \citep[e.g.,][]{maharaj2019}.
+Given this flexibility, a staggering variety of time series features have been proposed across different disciplines.
+For example, \citet{wang2006} proposed 9 structural characteristics \citep[also see][]{fulcher2013}, \citet{adya2001} collected 28 features relevant for forecasting, and a commonly used software package for feature extraction `tsfresh' allows users to extract a total of 794 features of a time series \citep[][]{christ2018}.
 
-The main aim of feature extraction is to describe the most important and meaningful aspects of a time series.
-In its most general approach, this can include any statistical summary of the time series \citep[i.e., feature; e.g.,][]{maharaj2019}.
-The extracted features can then be used as inputs to most clustering algorithms, including k-means or hierarchical clustering to group the participants into clusters \citep{wang2006}.
-The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
-However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
-Feature-based time series clustering has been a valued tool in many empirical fields for its flexibility and interpretability.
-As a result, features have been extracted for a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-
-With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
-As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos \citep[also see][]{fulcher2013}.
-\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
-A final illustration comes from a commonly used software package for feature extraction `tsfresh', which allows users to extract a total of 794 features to describe all kinds of aspects of a time series \citep[][]{christ2018}.
-
-Given that there are hundreds of possible features to describe psychological time series, one approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series \citep[e.g.,][]{christ2018}.
-However, such an approach is not always advisable for psychological time series.
-For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 794 time series features.
-More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
-This is particularly the case when features have a conceptual or psychological meaning.
-Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
-But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
-We will, thus, propose and describe a set of time series features that have a strong backing within the ESM literature and as such offer a reasonable starting point for most concepts within the field.
+However, not all time series features might be relevant to psychological time series or any particular research question.
+For example, a psychologist interested in well-being might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
+Instead, we advocate that we look at time series features that have a strong backing within the ESM literature and offer meaningful interpretability.
 
@@
 \citet{krone2018} even built a parametric model to tentatively cluster study participants.
-From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}\footnote{It should be noted that also within the psychological literature alternative summaries have been proposed that for example include measurement distribution, nonlinear developments, or categorical states.
+From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}\footnote{It should be noted that also within the psychological literature, alternative summaries have been proposed that, for example, include measurement distribution, nonlinear developments, or categorical states.
 As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.
+
+Some of the features found in the psychological literature are not necessarily well-suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the literature.
+In particular, covariances and cross-lagged correlations often produce a large number of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}.
+Others such as network density parameters, used to summarize variable interdependence, might not always be meaningful for psychological data \citep{bringmann2019a}.
+At the same time, linear and nonlinear trends are not captured by the features commonly proposed for psychological time series, because the features are often developed for stationary models \citep[e.g.,][]{krone2018}.
+
+Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, for our illustration we chose six features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
+We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-based clustering of psychological time series data.
+An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
+
+It is important to reiterate that the six selected time-series features are in no way exhaustive or imperative.
+Both using a more data-driven approach to features or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
 
@@
 
-Some of the proposed features are not necessarily well suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the emotion dynamics literature, which has been so formative for the analysis of psychological time series data.
-In particular, covariances and cross-lagged correlations often produce a staggering amount of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}.
-Others such as network density parameters, used to summarize variable interdependence, might not always be meaningful for psychological data \citep{bringmann2019a}.
-At the same time, linear and nonlinear trends are not captured by the features commonly proposed for psychological time series, because the features are often developed for stationary models \citep[e.g.,][]{krone2018}.
-
-Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, we hope to build a practical toolbox of meaningful time series features for psychological data.
-In particular, we propose to focus on six features based on common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
-We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-base clustering of psychological time series data.
-An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
-
-It should be noted that this selection of the proposed six time-series features is in no way exhaustive or imperative.
-Both using a purely data-driven approach or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
-
 \input{latex-results/03FeatureExtraction.tex}
 
-\subsubsection{Feature reduction}
+\subsection{Feature Reduction}
 Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm.
-As an example, a relatively common scenario would include 10 variables of interest, where 8 time series features are extracted, resulting in 80 features per participant (with a common sample size of 100 participants that would result in a total of 8,000 data points in this hypothetical example).
+As an example, a relatively common scenario would include 10 variables of interest, where eight time series features are extracted, resulting in 80 features per participant (with a common sample size of 100 participants that would result in a total of 8,000 data points in this hypothetical example).
 We offer an illustration of the compounding numbers of data points in \fgrref{fig:TSCFlowN}.
@@
 
-\subsection{Feature clustering}
+\subsection{Feature Clustering}
 For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible \citep{liao2005}.
@@
 Each of these approaches can be a valuable clustering approach for time series feature data and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency.
-As an example, under ideal conditions, most approaches are likely to provide a similar cluster solution \citep[e.g., for well seprated groups with little noice and few outliers; e.g., see][]{peng2022}.
+As an example, under ideal conditions, most approaches are likely to provide a similar cluster solution \citep[e.g., for well-separated groups with little noise and few outliers; e.g., see][]{peng2022}.
 However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate \citep[e.g.,][]{langfelder2008}.
@@
 Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \citep[e.g.,][]{corduas2008}.
-Yet, in cases where mis-specifications are costly, a method with fewer assumptions might be more advisable \citep[e.g.,][]{ankerst1999}.
+Yet, in cases where misspecifications are costly, a method with fewer assumptions might be more advisable \citep[e.g.,][]{ankerst1999}.
 
@@
 Assessing the clusters based on their average features allows users to gain an understanding of the developmental profiles.
-One cluster might, for example, have a relatively strong linear increase in well-being or participants in another cluster might have a relatively high variability of negative emotions.
+One cluster might, for example, have a relatively strong linear increase in well-being, or participants in another cluster might have a relatively high variability of negative emotions.
 Another common option for interpretation is finding an appropriate prototype to represent the cluster's trajectory.
@@
 \section{Discussion}
+
+NOTES:
+The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
+However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
+
+---
+
 % aims re-iterated
@@
 
-We have methodologically deconstructed the feature extraction, feature reduction and feature clustering steps and offered methodological as well as conceptual examples for each step.
+We have methodologically deconstructed the feature extraction, feature reduction, and feature clustering steps and offered methodological as well as conceptual examples for each step.
 Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research practice as both the features as well as the analysis steps are well-established within the field and statistical packages are readily available.
@@
 We make available the full code, custom functions, and illustrations to aid the approachability of the method.
-In the interpretation step we also show how the clusters can be compared across the original feature set as well as other meaningful situational and person-specific differences (e.g., see \fgrref{fig:clusterFeatVar}).
+In the interpretation step, we also show how the clusters can be compared across the original feature set as well as other meaningful situational and person-specific differences (e.g., see \fgrref{fig:clusterFeatVar}).
 We also find that for our example of migration experiences, the method was useful in discerning adaptive from more stressful experiences and helped to contextualize the diverging experiences.
@@
 
-In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential uncomparable results across methods.
-While the variety and diversity of methods is helpful in finding options even for more complex types of data, different algorithms often offer different results.
-And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their ideosyncratic shortcomings and might, for example, be sensitive to outliers.
+In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential incomparable results across methods.
+While the variety and diversity of methods are helpful in finding options even for more complex types of data, different algorithms often offer different results.
+And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their idiosyncratic shortcomings and might, for example, be sensitive to outliers.
 
@@
 
+With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
+As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos \citep[also see][]{fulcher2013}.
+\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
+A final illustration comes from a commonly used software package for feature extraction `tsfresh', which allows users to extract a total of 794 features to describe all kinds of aspects of a time series \citep[][]{christ2018}.
+
+Given that there are hundreds of possible features to describe psychological time series, one approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series \citep[e.g.,][]{christ2018}.
+However, such an approach is not always advisable for psychological time series.
+For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 794 time series features.
+More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
+This is particularly the case when features have a conceptual or psychological meaning.
+Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
+But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
+We will, thus, propose and describe a set of time series features that have a strong backing within the ESM literature and as such offer a reasonable starting point for most concepts within the field.
+
 \paragraph{Central Tendency} Central tendency refers to the statistical measures that represent the ``middle" or ``average" of a set of data.

### 2023-02-23 manuscript.tex
:construction: feature extraction link to appB

@@
 Both using a more data-driven approach to features or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
+The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview and \appref{app:FeatureBackground} for additional information on the theoretical relevance of the selected features).
 

### 2023-02-23 manuscript.tex
:construction: paragraph str and order

--- 2023-02-23 manuscript.tex
@@
 Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, for our illustration we chose six features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
+An overview of the selected features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
 We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-based clustering of psychological time series data.
-An overview of the proposed features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
-
-It is important to reiterate that the six selected time-series features are in no way exhaustive or imperative.
-Both using a more data-driven approach to features or selecting other aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
-The list seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions (see \tblref{tab:esmFeatures} for an overview).
+We provide additional background on the theoretical relevance of the selected features in \appref{app:FeatureBackground}.
 
@@
 \input{latex-results/03FeatureExtraction.tex}
+
+It is important to reiterate that the six selected time-series features are in no way exhaustive or imperative.
+Both using a more data-driven approach to the selection of features or selecting entirely different aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
+Our choice seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions.
 

### 2023-02-23 manuscript.tex
:construction: extraction results transition

@@
 \newcommand{\appref}[2][]{\hyperref[#2]{Appendix \ref*{#2}#1}}
+\newcommand{\equatref}[2][]{\hyperref[#2]{Equation \ref*{#2}#1}}
 
@@
 An overview of the selected features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
-We offer a short introduction of each of these six features below as a practical starting point for researchers and practitioners interested in feature-based clustering of psychological time series data.
 We provide additional background on the theoretical relevance of the selected features in \appref{app:FeatureBackground}.
+For each of the six time series features we chose, we chose a mathematical representation that was appropriate for our type of data.
+We provide a short introduction of each feature below.
 

### 2023-02-23 manuscript.tex
:construction: feature reduction str

@@
 We provide a short introduction of each feature below.
+Beyond the operationalizations we chose for our case study, we provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
 
@@
 The filter methods, broadly speaking determine important features by identifying irrelevant features (e.g., because features do not capture much information), and identifying redundant features \citep[e.g., because features capture the same information; e.g., see][]{yu2004}.
-Wrapper methods on the other hand avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features \citep{kohavi1997}.
+Wrapper methods, on the other hand, avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features \citep{kohavi1997}.
 A wrapper, thus, compares the performance of models with different feature combinations \citep[e.g., see][]{tang2014}.
@@
 Methods might, for example, use a filter step to reduce the size of features considered in a later wrapper step \citep{hsu2011}.
+Feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
 
 The second general approach to the dimensionality curse has been feature projection.
-Generally speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features \citep[i.e., the data is projected to a lower dimensional space; e.g.,][]{carreira-perpinan1997}.
+Broadly speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features \citep[i.e., the data is projected to a lower dimensional space; e.g.,][]{carreira-perpinan1997}.
 Projection methods are relatively common in psychological research --- including, factor- and principal component analyses \citep{vandermaaten2009}.
 Commonly, this is achieved using linear transformations \citep[e.g., principal component analysis; see][]{cunningham2015}, or more complex nonlinear transformations \citep[e.g., t-SNE; see]{lee2007}.
-Generally speaking, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
-Feature projection methods, on the other hand, tend to be more generalized and are more readily available.
+Feature projection methods have been popular because of their widespread availability as well as their high efficiency.
 
@@
 
-\paragraph{Nonlinearity} Changes in psychology are not always linear, instead nonlinearity is is a common feature of psychological time series \citep{hayes2007}.
+\paragraph{Nonlinearity} Changes in psychology are not always linear, instead nonlinearity is a common feature of psychological time series \citep{hayes2007}.
 As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
-Similarly, patients in recovery of depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
+Similarly, patients in recovery from depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
 But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.

### 2023-02-23 latex-results/03FeatureExtraction.tex
:sparkles: feature extraction illustration choices added

@@
-For the features, we selected and extracted one of each of the features
-proposed in \tblref{tab:esmFeatures}. We particularly chose the more
-robust median and median absolute deviation (MAD) for central tendency
-and variability. Within the variability structure, we chose the slightly
-more intuitive mean absolute change (MAC) for stability and stayed with
-the common lag--1 autocorrelation for inertia. For the trend summaries,
-we chose the overall linear regression slope to describe the linear
-trend and we summarized the nonlinear trend with the estimated degrees
-of freedom of an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We chose these features in
-particular because we consider them to be the most broadly applicable
-features. We also extracted the participant's number of completed ESM
-measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a
-large selection of the time series feature operationalizations presented
-in \tblref{tab:esmFeatures} in our GitHub repository (see the
-\texttt{featureExtractor} function).
+\paragraph{Central tendency.}
 
-For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
-the an ordered list of values from the time series of variable \(j\) for
-participant \(i\). The calculation depends on whether the number of
-measurements in a time series \(n\) is odd or even.
+The central tendency refers to the statistical measures that represent
+the ``typical'' or ``average'' of a set of data. The most common
+measures of central tendency are the mean, median, and mode
+\citep{weisberg1992}. As a familiar statistic from probability theory,
+the central tendency sits at the heart of many fundamental questions
+about psychological time series. Researchers might, for example, be
+interested in whether ``Over a one month period, are some people happier
+than others?''. For our central tendency feature we chose the robust
+\textit{median}, which can avoid potential issues with non-normally
+distributed time series responses or outliers \citep{weisberg1992}. To
+calculate the \textit{median} (\(M\)), we let \(X_{ij}\) be the ordered
+list of values from the time series of variable \(j\) for participant
+\(i\). The calculation depends on whether the number of measurements in
+a time series \(n\) is odd or even.
 
@@
 
-For distribution we chose the \textit{Median Absolute Deviation}
-(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
-in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
+\paragraph{Variability.}
+
+Variability captures the degree to which a set of data differs from the
+central tendency and is sometimes also referred to as the dispersion or
+spread of the data \citep{weisberg1992}. In time series analyses,
+variability is conceptually important because information about the
+distribution and diversity of the data has been found to be indicative
+of worse psychological states \citep{myin-germeys2018}. Person level
+differences of ESM measurements have, for example, been associated with
+higher levels of psycho-pathological recurrences among depression
+patients \citep{timm2017}. For our illustration data, we chose to
+capture the time series variability with the
+\textit{Median Absolute Deviation} (\(MAD\)), where we calculate the
+\textit{median} (\(M\); calculated as in \equatref{eq:median}) for the
+absolute deviations of measurement \(x\) at time point \(t\) for
+participant \(i\) and variable \(j\) from the median of that time series
+\(X\).
 
 \begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
+  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
 \end{equation}
 
-mean absolute change (MAC) for stability
+\paragraph{Instability.}
+
+Instability captures the average change between two consecutive
+measurements \citep{ebner-priemer2009}. While instability is
+conceptually related to the variability feature, variability does not
+take into account temporal dependency, whereas instability looks at the
+`jumpy-ness' of the data over time. In other words, variability reflects
+the range or diversity of values in a time series data, while
+instability reflects the fluctuation or inconsistency in a time series
+data over time \citep{trull2008}. For example, if a person has rapid and
+extreme changes in mood their mood is highly unstable, while if a
+person's mood responses span a wide range over the entire study period,
+their mood is highly variable \citep{jahng2008}. Within psychological
+time series, instability measurements have especially been important in
+the research of borderline personality disorder \citep{trull2008} and
+suicidality \citep{kivela2022}, but also in understanding early warning
+signals more generally \citep{wichers2019}. Conceptually, the
+instability feature, thus, relates to a broad range of research
+questions, including: ``Do changes tend to be slow and gradual or fast
+and abrupt?'' For our data we chose the \textit{mean absolute change}
+\citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
+the average absolute difference of measurements \(x\) at time points
+\(t\) and \(t-1\), for each time series \(X\) of participant \(i\) and
+variable \(j\).
 
 \begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
+  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t}-x_{t-1}\right|
 \end{equation}
 
-lag--1 autocorrelation for inertia: lag \(l=1\)
+\paragraph{Self-similarity.}
+
+Self-similarity in time series data refers to the property of a time
+series to exhibit similar patterns of behavior over different time
+scales \citep{dmello2021}. That is, self-similarity describes how much a
+measurement carries over to future measurements. One important
+self-similarity in psychological time series is \texit{inertia} --- how
+much a measurement carries over to its next measurement
+\citep{kuppens2010}. If inertia is high a development tends to stay in a
+certain state or continues to move towards a certain direction. Because
+high inertia is resistant to change, in emotion dynamics high inertia
+has been found to be indicative of underreactive systems and to be
+characteristic of psychological maladjustment \citep{kuppens2010}. In a
+similar vein, high inertia at baseline was even predictive of the
+initial onset of depression \citep{kuppens2012}. Conceptually, inertia
+is more broadly connected to research questions such as: ``Do patients
+stay in a depressed mood for several measurements?'\,' For our
+illustration case, we chose the commonly used autocorrelation or
+autoregression with a lag-1 to capture the inertia. High autocorrelation
+values can indicate high levels of inertia, while low autocorrelation
+values may indicate a more unpredictable or volatile time series
+\citep{dejonckheere2019}. The lag--1 autocorrelation \(r_{ij,1}\) looks
+at the average correlation between a measurement \(x\) and the
+proceeding measurement \(x_{t-1}\) for the time series \(X\) of
+participant \(i\) and variable \(j\).
 
 \begin{equation} \label{eq:ar}
-  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ijt-l}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
 \end{equation}
 
-overall linear regression slope to describe the linear trend
+\paragraph{Linear trend.}
+
+In non-stationary time series a linear trend can be observed when there
+is a consistent increase or decrease in the data over time
+\citep{nyblom1986}. For psychological time series, researchers have, for
+example, pointed out the importance of linear trends in interpersonal
+communications \citep{vasileiadou2014}, and emotion dynamics
+\citep{oravecz2016}. Theoretically, linear trends are often considered
+the simplest way of assessing whether a psychological theory of change
+is appropriate \citep{gottman1969}. In empirical practice, linear trends
+are, thus, commonly exemplified by research questions such as ``Do
+patient symptoms improve consistently?'' or ``Does worker productivity
+decline continuously?'' For the variables in our illustration data set
+we chose an overall linear regression slope to capture the linear trend.
 
@@
 
-summarized the nonlinear trend with the estimated degrees of freedom of
-an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We estimate the number of
-parametric and non-parametric (or smooth) terms \(p\) in the model.
+\paragraph{Nonlinearity.}
+
+Changes in psychology are not always linear, instead nonlinearity is a
+common feature of psychological time series \citep{hayes2007}. As an
+example, episodic disorders, such as depression, are most likely best
+described as non-linear systems \citep{hosenfeld2015}. Similarly,
+patients in recovery from depression showed sudden changes in the
+improvement of depression \citep{helmich2020a}. But also substance abuse
+\citep{boker1998} or attitude changes rarely develop linearly
+\citep{vandermaas2003}. Conceptually, researchers might have research
+question about the type of the development: ``Is the development of
+anxiety a nonlinear process?'' as well as the shape and structure of the
+development: ``Does anxiety change as a sudden or smooth transition?''
+We summarized the nonlinear trend with the
+\textit{estimated degrees of freedom} of an empty GAM spline model. The
+\(edf\) summarizes the \textit{wiggliness} of a spline trend line by
+estimating the number of parametric and non-parametric (or smooth) terms
+\(p\) in the model. An edf of 1 would be equivalent to a linear
+relationship (i.e., one linear slope parameter), whereas a higher edf
+(particularly an edf \textgreater{} 2) is indicative of a non-linear
+trend.
 
@@
 
-\textit{Notes}: In a Generalized Additive Model (GAM), the estimated
-degrees of freedom refers to the number of parameters used to estimate
-the model. This includes the number of parameters used to estimate each
-smooth function in the model, as well as any additional parameters used
-for the overall model. On the other hand, the effective degrees of
-freedom is a measure of the amount of information in the data that is
-used to estimate the model. It takes into account the smoothness of the
-estimated functions in the model, and is typically smaller than the
-estimated degrees of freedom. Effective degrees of freedom is used to
-estimate the uncertainty of the model estimates and to calculate the
-model's goodness of fit. In general, the effective degrees of freedom is
-a more appropriate measure of the complexity of the model than the
-estimated degrees of freedom, as it accounts for the smoothness of the
-estimated functions, which can help to prevent overfitting. This is
-because if the estimated function is smooth, it means that the model is
-not fitting the noise and has a better generalization.
-
-After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across the 72 features per participant.
-This might, for example, happen if participants do not have two
-subsequent measurements with outgroup interactions, so that an
-autocorrelation with lag-1 cannot be calculated for the contact-specific
-variables. The small number of missing values indicates that the
-feature-based approach indeed largely avoids the structural missingness
-issue. The few missing values can, however, be an issue for some feature
-reduction or feature clustering algorithms. We, thus, impute the missing
-feature values with a single predictive mean matching imputation using
-the MICE library.
+Beyond our main features of interest, we also extracted the
+participant's number of completed ESM measurements to ensure that the
+clusters are comparable in that regard. After the feature extraction, we
+found that about 1.40\% of the extracted features are missing across the
+72 features per participant. This might, for example, happen if
+participants do not have two subsequent measurements with outgroup
+interactions, so that an autocorrelation with lag-1 cannot be calculated
+for the contact-specific variables. The small number of missing values
+indicates that the feature-based approach indeed largely avoids the
+structural missingness issue. The few missing values can, however, be an
+issue for some feature reduction or feature clustering algorithms. We,
+thus, impute the missing feature values with a single predictive mean
+matching imputation using the MICE library.

### 2023-02-23 sections/03FeatureExtraction.Rmd
:sparkles: feature extraction illustration choices added

@@
 
-For the features, we selected and extracted one of each of the features proposed in \tblref{tab:esmFeatures}.
-We particularly chose the more robust median and median absolute deviation (MAD) for central tendency and variability.
-Within the variability structure, we chose the slightly more intuitive mean absolute change (MAC) for stability and stayed with the common lag--1 autocorrelation for inertia.
-For the trend summaries, we chose the overall linear regression slope to describe the linear trend and we summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We chose these features in particular because we consider them to be the most broadly applicable features.
-We also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.
+\paragraph{Central tendency.}
+The central tendency refers to the statistical measures that represent the "typical" or "average" of a set of data.
+The most common measures of central tendency are the mean, median, and mode \citep{weisberg1992}.
+As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
+Researchers might, for example, be interested in whether "Over a one month period, are some people happier than others?".
+For our central tendency feature we chose the robust \textit{median}, which can avoid potential issues with non-normally distributed time series responses or outliers \citep{weisberg1992}.
+To calculate the \textit{median} ($M$), we let $X_{ij}$ be the ordered list of values from the time series of variable $j$ for participant $i$.
 The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
@@
 ```
-For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
+
+\paragraph{Variability.}
+Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
+In time series analyses, variability is conceptually important because information about the distribution and diversity of the data has been found to be indicative of worse psychological states \citep{myin-germeys2018}.
+Person level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
+For our illustration data, we chose to capture the time series variability with the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \equatref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
 
@@
 \begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
+  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
 \end{equation}
 ```
-mean absolute change (MAC) for stability
+
+\paragraph{Instability.}
+Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009}.
+While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the 'jumpy-ness' of the data over time.
+In other words, variability reflects the range or diversity of values in a time series data, while instability reflects the fluctuation or inconsistency in a time series data over time  \citep{trull2008}.
+For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
+Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
+Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "Do changes tend to be slow and gradual or fast and abrupt?"
+For our data we chose the \textit{mean absolute change} \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at the average absolute difference of measurements $x$ at time points $t$ and $t-1$, for each time series $X$ of participant $i$ and variable $j$.
 
@@
 \begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
+  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t}-x_{t-1}\right|
 \end{equation}
 ```
-lag--1 autocorrelation for inertia: lag $l=1$
+
+\paragraph{Self-similarity.}
+Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}.
+That is, self-similarity describes how much a measurement carries over to future measurements.
+One important self-similarity in psychological time series is \texit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010}.
+If inertia is high a development tends to stay in a certain state or continues to move towards a certain direction.
+Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
+In a similar vein, high inertia at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
+Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?''
+For our illustration case, we chose the commonly used autocorrelation or autoregression with a lag-1 to capture the inertia.
+High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}.
+The lag--1 autocorrelation $r_{ij,1}$ looks at the average correlation between a measurement $x$ and the proceeding measurement $x_{t-1}$ for the time series $X$ of participant $i$ and variable $j$.
 
@@
 \begin{equation} \label{eq:ar}
-  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ijt-l}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
 \end{equation}
 ```
-overall linear regression slope to describe the linear trend
+
+\paragraph{Linear trend.}
+In non-stationary time series a linear trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
+For psychological time series, researchers have, for example, pointed out the importance of linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}.
+Theoretically, linear trends are often considered the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}.
+In empirical practice, linear trends are, thus, commonly exemplified by research questions such as "Do patient symptoms improve consistently?" or "Does worker productivity decline continuously?"
+For the variables in our illustration data set we chose an overall linear regression slope to capture the linear trend.
 
@@
 ```
-summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.
-We estimate the number of parametric and non-parametric (or smooth) terms $p$ in the model.
+
+\paragraph{Nonlinearity.}
+Changes in psychology are not always linear, instead nonlinearity is a common feature of psychological time series \citep{hayes2007}.
+As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
+Similarly, patients in recovery from depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
+But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
+Conceptually, researchers might have research question about the type of the development: "Is the development of anxiety a nonlinear process?" as well as the shape and structure of the development: "Does anxiety change as a sudden or smooth transition?"
+We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model.
+The $edf$ summarizes the \textit{wiggliness} of a spline trend line by estimating the number of parametric and non-parametric (or smooth) terms $p$ in the model.
+An edf of 1 would be equivalent to a linear relationship (i.e., one linear slope parameter), whereas a higher edf (particularly an edf > 2) is indicative of a non-linear trend.
 
@@
 
-\textit{Notes}: In a Generalized Additive Model (GAM), the estimated degrees of freedom refers to the number of parameters used to estimate the model.
-This includes the number of parameters used to estimate each smooth function in the model, as well as any additional parameters used for the overall model.
-On the other hand, the effective degrees of freedom is a measure of the amount of information in the data that is used to estimate the model.
-It takes into account the smoothness of the estimated functions in the model, and is typically smaller than the estimated degrees of freedom.
-Effective degrees of freedom is used to estimate the uncertainty of the model estimates and to calculate the model's goodness of fit.
-In general, the effective degrees of freedom is a more appropriate measure of the complexity of the model than the estimated degrees of freedom, as it accounts for the smoothness of the estimated functions, which can help to prevent overfitting.
-This is because if the estimated function is smooth, it means that the model is not fitting the noise and has a better generalization.
-
+Beyond our main features of interest, we also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
 After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.

### 2023-02-23 manuscript.tex
:construction: clustering intro str

@@
 \subsection{Feature Clustering}
-For feature-based time-series clustering, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible \citep{liao2005}.
+For the actual clustering of the time-series features, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible \citep{liao2005}.
 The crux of clustering is, thus, to have clearly defined and effective measurements of (dis)similarity.
 Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment \citep[or similarity measurement for qualitative features; see][]{Aghabozorgi2015}.
-While others have produced excellent overviews of the many clustering approaches available \citep[e.g.,][]{xu2015}, we will briefly introduce some of the more readily available approaches suitable for most time series feature data.
-The well-established and readily available clustering approaches can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof \citep[see \tblref{tab:clusterApproaches} for an overview; also see][for a broader review]{jain1999}.
+While others have produced excellent overviews of the many clustering approaches available \citep[e.g.,][]{xu2015}, the more readily available approaches suitable for most time series feature data can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof \citep[see \tblref{tab:clusterApproaches} for an overview; also see][for a broader review]{jain1999}.
 
-Each of these approaches can be a valuable clustering approach for time series feature data and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency.
-As an example, under ideal conditions, most approaches are likely to provide a similar cluster solution \citep[e.g., for well-separated groups with little noise and few outliers; e.g., see][]{peng2022}.
-However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate \citep[e.g.,][]{langfelder2008}.
-Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective \citep[e.g.,][]{jain2010}.
-Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \citep[e.g.,][]{corduas2008}.
-Yet, in cases where misspecifications are costly, a method with fewer assumptions might be more advisable \citep[e.g.,][]{ankerst1999}.
-
-There is thus, unfortunately, no one-size-fits-all solution to clustering.
-On the bright side, however, the thriving methodological developments in the clustering literature offer a wide variety of options that offer prospects even for extremely large and complex data sets.
+There is, unfortunately, no one-size-fits-all solution to clustering and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency.
 We provide a short intuitive explanation for common approaches, together with some of their characteristics and example algorithms in \tblref{tab:clusterApproaches}.
@@
 As such, the k-means offers a good starting point for many psychological researchers and the method should be generalizable across a relatively wide variety of projects.
+
+%Each of these approaches can be a valuable clustering approach for time series feature data and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency.
+As an example, under ideal conditions, most approaches are likely to provide a similar cluster solution \citep[e.g., for well-separated groups with little noise and few outliers; e.g., see][]{peng2022}.
+However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate \citep[e.g.,][]{langfelder2008}.
+Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective \citep[e.g.,][]{jain2010}.
+Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \citep[e.g.,][]{corduas2008}.
+Yet, in cases where misspecifications are costly, a method with fewer assumptions might be more advisable \citep[e.g.,][]{ankerst1999}.
 

### 2023-02-23 latex-results/03FeatureExtraction.tex
:memo: feature reduction results structure

@@
 interested in whether ``Over a one month period, are some people happier
-than others?''. For our central tendency feature we chose the robust
+than others?''
+
+For the central tendency feature of our illustration we chose the robust
 \textit{median}, which can avoid potential issues with non-normally
@@
 higher levels of psycho-pathological recurrences among depression
-patients \citep{timm2017}. For our illustration data, we chose to
-capture the time series variability with the
-\textit{Median Absolute Deviation} (\(MAD\)), where we calculate the
-\textit{median} (\(M\); calculated as in \equatref{eq:median}) for the
-absolute deviations of measurement \(x\) at time point \(t\) for
-participant \(i\) and variable \(j\) from the median of that time series
-\(X\).
+patients \citep{timm2017}. As such, psychological researchers and
+practitioners are often empirically interested in person-level
+differences in variability. Researchers on polarization and
+radicalization might for example ask: ``Are people settled in their
+attitudes towards migrants or vary across the measurement period?'\,'
+
+For our illustration data, we chose to capture the time series
+variability with the \textit{Median Absolute Deviation} (\(MAD\)), where
+we calculate the \textit{median} (\(M\); calculated as in
+\equatref{eq:median}) for the absolute deviations of measurement \(x\)
+at time point \(t\) for participant \(i\) and variable \(j\) from the
+median of that time series \(X\).
 
@@
 questions, including: ``Do changes tend to be slow and gradual or fast
-and abrupt?'' For our data we chose the \textit{mean absolute change}
+and abrupt?''
+
+For our data we chose the \textit{mean absolute change}
 \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
@@
 is more broadly connected to research questions such as: ``Do patients
-stay in a depressed mood for several measurements?'\,' For our
-illustration case, we chose the commonly used autocorrelation or
+stay in a depressed mood for several measurements?'\,'
+
+For our illustration case, we chose the commonly used autocorrelation or
 autoregression with a lag-1 to capture the inertia. High autocorrelation
@@
 patient symptoms improve consistently?'' or ``Does worker productivity
-decline continuously?'' For the variables in our illustration data set
-we chose an overall linear regression slope to capture the linear trend.
+decline continuously?''
+
+For the variables in our illustration data set we chose an overall
+linear regression slope to capture the linear trend. The regression
+slope provides the average change from one time point \(t\) to the next
+across all measurements \(x\) of a time series \(X\) of participant
+\(i\) and variable \(j\).
 

### 2023-02-23 latex-results/04FeatureReduction.tex
:memo: feature reduction results structure

@@
 ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
+features into the analysis. The PCA uses linear transformations in such
+a way that the first component captures the most possible variance of
+the original data (e.g., by finding a vector that maximizes the sum of
+squared distances). The following components will then use the same
 method to iteratively explain the most of the remaining variance while
@@
 features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
+particularly advocate for such an approach if all original features are
+considered psychologically meaningful in understanding the time series
+and users would like to retain the features for inpterpretation
+(irrespective of the features importance).

### 2023-02-23 sections/03FeatureExtraction.Rmd
:memo: feature reduction results structure

@@
 As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
-Researchers might, for example, be interested in whether "Over a one month period, are some people happier than others?".
-For our central tendency feature we chose the robust \textit{median}, which can avoid potential issues with non-normally distributed time series responses or outliers \citep{weisberg1992}.
+Researchers might, for example, be interested in whether "Over a one month period, are some people happier than others?"
+
+For the central tendency feature of our illustration we chose the robust \textit{median}, which can avoid potential issues with non-normally distributed time series responses or outliers \citep{weisberg1992}.
 To calculate the \textit{median} ($M$), we let $X_{ij}$ be the ordered list of values from the time series of variable $j$ for participant $i$.
@@
 Person level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
+As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
+Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
+
 For our illustration data, we chose to capture the time series variability with the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \equatref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
@@
 Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "Do changes tend to be slow and gradual or fast and abrupt?"
+
 For our data we chose the \textit{mean absolute change} \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at the average absolute difference of measurements $x$ at time points $t$ and $t-1$, for each time series $X$ of participant $i$ and variable $j$.
@@
 Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?''
+
 For our illustration case, we chose the commonly used autocorrelation or autoregression with a lag-1 to capture the inertia.
@@
 In empirical practice, linear trends are, thus, commonly exemplified by research questions such as "Do patient symptoms improve consistently?" or "Does worker productivity decline continuously?"
+
 For the variables in our illustration data set we chose an overall linear regression slope to capture the linear trend.
+The regression slope provides the average change from one time point $t$ to the next across all measurements $x$ of a time series $X$ of participant $i$ and variable $j$.
 

### 2023-02-23 sections/04FeatureReduction.Rmd
:memo: feature reduction results structure

@@
 We then enter all `r ncol(scaled_data)` features into the analysis.
-The PCA will use linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
+The PCA uses linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
 The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
@@
 Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if the original features were chosen for their psychological meaning in understanding the time series and broader phenomenon of interest.
+We particularly advocate for such an approach if all original features are considered psychologically meaningful in understanding the time series and users would like to retain the features for inpterpretation (irrespective of the features importance).

### 2023-02-24 latex-results/05FeatureClustering.tex
:construction: cluster explanation

@@
-For our illustration of the feature clustering procedure, we use the
-generalizable and efficient centroid-based k-means clustering. The
-procedure needs the user to make a selection on the optimal number of
-clusters in order to run the main algorithm. We use the Elbow and the
-Silhouette method, which both suggest two clusters to be the optimal
-solution (see Supplemental Material A). We then entered the
-participants' PC-scores into the k-means algorithm. We used the
-recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
-random initial centroid positions to avoid convergence to a sub-optimal
-solution (i.e., local minima). The k-means analysis assigned 80
+For the actual k-means clustering, the analysis seeks to minimize the
+total within-cluster variation. We used the Hartigan and Wong algorithm,
+which is a widely used approach in k-means clustering of time series
+features \citep{hartigan1979}. This algorithm is designed to optimize
+the clustering of the feature data into \(k\) groups, where \(k\) is a
+pre-defined number of clusters. The algorithm starts by randomly
+separating the data points into k clusters and then iteratively updates
+the assignment of each point to the nearest cluster center until
+convergence. The Hartigan and Wong algorithm specifically calculates the
+within-cluster variation (\(W\)) of cluster \(C_i\) as the summed
+squared Euclidean distances of the feature \(x\) to its assigned cluster
+centroid \(\mu_i\):
+
+\begin{equation} \label{eq:kWCi}
+  W(C_i) = \sum_{x \in C_i}(x-\mu_i)^2
+\end{equation}
+
+By summing the within cluster sum of squares from all \(k\) clusters, we
+can then derive the total within-cluster sum of square \(WCSS\):
+
+\begin{equation} \label{eq:kWCSS}
+  WCSS = \sum_{i=1}^k W(C_i) = \sum_{i=1}^k \sum_{x \in C_i} (x - \mu_i)^2
+\end{equation}
+
+It is this \(WCSS\) that becomes the objective function to be minimized,
+by iteratively moving features from one cluster to another
+\citep{hartigan1979}. In particular, the algorithm (1) calculates the
+cluster centroids of the initial partitioning, (2) checks whether any
+feature has a centroid that is closer than that of the currently
+assigned cluster (3) updates the centroids based on any reassigned
+features, and then iterates between steps two and three until \(WCSS\)
+is minimized (i.e., locally optimal convergence) or a maximum number of
+iterations is reached. Given the iterative nature of the algorithm, the
+initial partitioning is often important because the algorithm might
+arrive at a suboptimal clustering where the \(WCSS\) cannot be further
+reduced by moving any feature to another cluster, despite a better
+solution existing (i.e., a local minimum). It is, therefore often
+recommended to run the k-means clustering with several different
+starting positions.
+
+In our case we used the Elbow and the Silhouette method to determine the
+recommended number of \(k\) clusters, which for both methods suggested
+two clusters (see Supplemental Material A). We then entered the
+participants' PC-scores from the previous step into the k-means
+algorithm. To avoid local minima we used 100 random initial centroid
+positions. In the final solution the k-means analysis assigned 80
 participants to cluster 1 and 76 participants to cluster 2.

### 2023-02-24 sections/05FeatureClustering.Rmd
:construction: cluster explanation

@@
 
-For our illustration of the feature clustering procedure, we use the generalizable and efficient centroid-based k-means clustering.
-The procedure needs the user to make a selection on the optimal number of clusters in order to run the main algorithm.
-We use the Elbow and the Silhouette method, which both suggest two clusters to be the optimal solution (see Supplemental Material A).
-We then entered the participants' PC-scores into the k-means algorithm.
-We used the recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100 random initial centroid positions to avoid convergence to a sub-optimal solution (i.e., local minima).
-The k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
+For the actual k-means clustering, the analysis seeks to minimize the total within-cluster variation.
+We used the Hartigan and Wong algorithm, which is a widely used approach in k-means clustering of time series features \citep{hartigan1979}.
+This algorithm is designed to optimize the clustering of the feature data into $k$ groups, where $k$ is a pre-defined number of clusters.
+The algorithm starts by randomly separating the data points into k clusters and then iteratively updates the assignment of each point to the nearest cluster center until convergence.
+The Hartigan and Wong algorithm specifically calculates the within-cluster variation ($W$) of cluster $C_i$ as the summed squared Euclidean distances of the feature $x$ to its assigned cluster centroid $\mu_i$:
+
+```{=tex}
+\begin{equation} \label{eq:kWCi}
+  W(C_i) = \sum_{x \in C_i}(x-\mu_i)^2
+\end{equation}
+```
+
+By summing the within cluster sum of squares from all $k$ clusters, we can then derive the total within-cluster sum of square $WCSS$:
+
+```{=tex}
+\begin{equation} \label{eq:kWCSS}
+  WCSS = \sum_{i=1}^k W(C_i) = \sum_{i=1}^k \sum_{x \in C_i} (x - \mu_i)^2
+\end{equation}
+```
+
+It is this $WCSS$ that becomes the objective function to be minimized, by iteratively moving features from one cluster to another \citep{hartigan1979}.
+In particular, the algorithm (1) calculates the cluster centroids of the initial partitioning, (2) checks whether any feature has a centroid that is closer than that of the currently assigned cluster (3) updates the centroids based on any reassigned features, and then iterates between steps two and three until $WCSS$ is minimized (i.e., locally optimal convergence) or a maximum number of iterations is reached.
+Given the iterative nature of the algorithm, the initial partitioning is often important because the algorithm might arrive at a suboptimal clustering where the $WCSS$ cannot be further reduced by moving any feature to another cluster, despite a better solution existing (i.e., a local minimum).
+It is, therefore often recommended to run the k-means clustering with several different starting positions.
+
+In our case we used the Elbow and the Silhouette method to determine the recommended number of $k$ clusters, which for both methods suggested two clusters (see Supplemental Material A).
+We then entered the participants' PC-scores from the previous step into the k-means algorithm.
+To avoid local minima we used 100 random initial centroid positions.
+In the final solution the k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.

### 2023-02-24 manuscript.tex
:construction: cluster evaluation lang

@@
 \subsection{Cluster Evaluation}
-The final evaluation step includes two main elements, (1) evaluating the performance of the clustering analysis and (2) interpreting the extracted clusters conceptually.
+Now that the participants have been assigned to their respective clusters based on the similarity of their time series features, the final evaluation step includes two main elements, (1) evaluating the performance of the clustering analysis and (2) interpreting the extracted clusters conceptually.
 
@@
 If true class labels are known, the cluster assignments can be compared to the true class labels --- using measures such as the F-measure, adjusted Rand index, mutual information, and normalized mutual information \citep[i.e., external evaluation; e.g.,][]{liao2005}.
-However, if the true cluster assignments are unknown, as with most of the exploratory clustering we speak to, the quality of the clusters is assessed based on the characteristics of the data itself \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
+However, if the true cluster assignments are unknown, as with most of the psychological time series we speak to, the quality of the clusters is assessed based on the characteristics of the data itself \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
 One popular internal evaluation measure is the silhouette score, which measures the degree to which each time series data point is similar to other points within the same cluster, compared to points in other clusters \citep{rousseeuw1987}.
@@
 It is therefore important to consider a variety of evaluation measures and to carefully interpret the results in the context of the specific analysis \citep{vinh2009}.
+
+In our own illustration example, we focused on the readily available silhouette score, and Calinski-Harabasz index to judge the performance of the clustering step.
 
@@
 
-Before moving to our empirical illustration, we would like to comment on the reporting of feature-based time series clustering.
+\input{latex-results/06ClusterEvaluation.tex}
+
+Before we move to the final discussion and conclusions, we would like to comment on the reporting of feature-based time series clustering.
 Clustering is inherently an exploratory analysis and, as such, each of the analysis steps involves a set of decisions by the user.
@@
 Most of their recommendations and reporting guidelines apply to feature-based clustering as well and might even offer a template for researchers hoping to pre-register their analysis procedures.
-
-\input{latex-results/06ClusterEvaluation.tex}
 

### 2023-02-24 latex-results/05FeatureClustering.tex
:memo: clustering wording

@@
-For the actual k-means clustering, the analysis seeks to minimize the
-total within-cluster variation. We used the Hartigan and Wong algorithm,
-which is a widely used approach in k-means clustering of time series
-features \citep{hartigan1979}. This algorithm is designed to optimize
-the clustering of the feature data into \(k\) groups, where \(k\) is a
-pre-defined number of clusters. The algorithm starts by randomly
-separating the data points into k clusters and then iteratively updates
-the assignment of each point to the nearest cluster center until
-convergence. The Hartigan and Wong algorithm specifically calculates the
+During the k-means clustering itself, the analysis seeks to minimize the
+total within-cluster variation. The analysis is designed to optimize the
+clustering of the feature data into \(k\) groups, where \(k\) is a
+pre-defined number of clusters. We used the Hartigan and Wong algorithm,
+which is a widely used algorithm in k-means clustering
+\citep{hartigan1979}. The algorithm starts by randomly separating the
+data points into k clusters and then iteratively updates the assignment
+of each point to the nearest cluster center until convergence. To do so,
+the Hartigan and Wong algorithm specifically calculates the
 within-cluster variation (\(W\)) of cluster \(C_i\) as the summed
-squared Euclidean distances of the feature \(x\) to its assigned cluster
+squared Euclidean distances of the feature \(x\) to the closest cluster
 centroid \(\mu_i\):
@@
 
-By summing the within cluster sum of squares from all \(k\) clusters, we
+By summing the within-cluster sum of squares from all \(k\) clusters, we
 can then derive the total within-cluster sum of square \(WCSS\):
@@
 In our case we used the Elbow and the Silhouette method to determine the
-recommended number of \(k\) clusters, which for both methods suggested
-two clusters (see Supplemental Material A). We then entered the
-participants' PC-scores from the previous step into the k-means
+recommended number of \(k\) clusters, which for both methods was two
+clusters (see Supplemental Material A). We then entered the
+participants' PC-scores from the feature reduction step into the k-means
 algorithm. To avoid local minima we used 100 random initial centroid

### 2023-02-24 sections/05FeatureClustering.Rmd
:memo: clustering wording

@@
 
-For the actual k-means clustering, the analysis seeks to minimize the total within-cluster variation.
-We used the Hartigan and Wong algorithm, which is a widely used approach in k-means clustering of time series features \citep{hartigan1979}.
-This algorithm is designed to optimize the clustering of the feature data into $k$ groups, where $k$ is a pre-defined number of clusters.
+During the k-means clustering itself, the analysis seeks to minimize the total within-cluster variation.
+The analysis is designed to optimize the clustering of the feature data into $k$ groups, where $k$ is a pre-defined number of clusters.
+We used the Hartigan and Wong algorithm, which is a widely used algorithm in k-means clustering \citep{hartigan1979}.
 The algorithm starts by randomly separating the data points into k clusters and then iteratively updates the assignment of each point to the nearest cluster center until convergence.
-The Hartigan and Wong algorithm specifically calculates the within-cluster variation ($W$) of cluster $C_i$ as the summed squared Euclidean distances of the feature $x$ to its assigned cluster centroid $\mu_i$:
+To do so, the Hartigan and Wong algorithm specifically calculates the within-cluster variation ($W$) of cluster $C_i$ as the summed squared Euclidean distances of the feature $x$ to the closest cluster centroid $\mu_i$:
 
@@
 
-By summing the within cluster sum of squares from all $k$ clusters, we can then derive the total within-cluster sum of square $WCSS$:
+By summing the within-cluster sum of squares from all $k$ clusters, we can then derive the total within-cluster sum of square $WCSS$:
 
@@
 
-In our case we used the Elbow and the Silhouette method to determine the recommended number of $k$ clusters, which for both methods suggested two clusters (see Supplemental Material A).
-We then entered the participants' PC-scores from the previous step into the k-means algorithm.
+In our case we used the Elbow and the Silhouette method to determine the recommended number of $k$ clusters, which for both methods was two clusters (see Supplemental Material A).
+We then entered the participants' PC-scores from the feature reduction step into the k-means algorithm.
 To avoid local minima we used 100 random initial centroid positions.

### 2023-02-24 manuscript.tex
:memo: cluster eval intro shortened

--- 2023-02-23 manuscript.tex
@@
 \subsection{Cluster Evaluation}
-The final evaluation step includes two main elements, (1) evaluating the performance of the clustering analysis and (2) interpreting the extracted clusters conceptually.
+Now that the participants have been assigned to their respective clusters based on the similarity of their time series features, the final evaluation step includes two main elements, (1) evaluating the performance of the clustering analyses to choose an optimal solution and (2) interpreting the extracted clusters conceptually.
 
@@
 If true class labels are known, the cluster assignments can be compared to the true class labels --- using measures such as the F-measure, adjusted Rand index, mutual information, and normalized mutual information \citep[i.e., external evaluation; e.g.,][]{liao2005}.
-However, if the true cluster assignments are unknown, as with most of the exploratory clustering we speak to, the quality of the clusters is assessed based on the characteristics of the data itself \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
-One popular internal evaluation measure is the silhouette score, which measures the degree to which each time series data point is similar to other points within the same cluster, compared to points in other clusters \citep{rousseeuw1987}.
-Another measure is the Calinski-Harabasz index, which assesses the compactness and separation of the clusters \citep{calinski1974}.
-Another common method is to compare different cluster solutions using the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.
-Another important aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}.
-This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}.
-It is important to note that no single evaluation measure is perfect, and different measures may produce different results depending on the characteristics of the data and the research question being addressed \citep{kittler1998}.
-It is therefore important to consider a variety of evaluation measures and to carefully interpret the results in the context of the specific analysis \citep{vinh2009}.
+However, if the true cluster assignments are unknown, as with our psychological time series, the quality of the clusters is assessed based on the characteristics of the data itself \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
+
+\input{latex-results/06ClusterEvaluationPerformance.tex}
 
@@
 
-Before moving to our empirical illustration, we would like to comment on the reporting of feature-based time series clustering.
+\input{latex-results/06ClusterEvaluationInterpretation.tex}
+
+Before we move to the final discussion and conclusions, we would like to comment on the reporting of feature-based time series clustering.
 Clustering is inherently an exploratory analysis and, as such, each of the analysis steps involves a set of decisions by the user.

### 2023-02-24 IllustrationResultsRenderReport.Rmd
:sparkles: cluster and performance results expl added

@@
 
+## CLuster Performance
+
 ```{r}
-#| label: render cluster evaluation
+#| label: render cluster performance
 #| include: true
@@
 rmarkdown::render(
-  input = "sections/06ClusterEvaluation.Rmd",
+  input = "sections/06ClusterEvaluationPerformance.Rmd",
+  output_dir = outputDir
+)
+```
+
+## CLuster Interpretation
+
+```{r}
+#| label: render cluster interpretation
+#| include: true
+#| results: 'asis'
+
+rmarkdown::render(
+  input = "sections/07ClusterEvaluationInterpretation.Rmd",
   output_dir = outputDir

### 2023-02-24 latex-results/05FeatureClustering.tex
:sparkles: cluster and performance results expl added

@@
 is minimized (i.e., locally optimal convergence) or a maximum number of
-iterations is reached. Given the iterative nature of the algorithm, the
-initial partitioning is often important because the algorithm might
-arrive at a suboptimal clustering where the \(WCSS\) cannot be further
-reduced by moving any feature to another cluster, despite a better
-solution existing (i.e., a local minimum). It is, therefore often
+iterations is reached \citep{jain2010}. Given the iterative nature of
+the algorithm, the initial partitioning is often important because the
+algorithm might arrive at a suboptimal clustering where the \(WCSS\)
+cannot be further reduced by moving any feature to another cluster,
+despite a better solution existing
+\citep[i.e., a local minimum;][]{timmerman2013}. It is, therefore often
 recommended to run the k-means clustering with several different
@@
 
-In our case we used the Elbow and the Silhouette method to determine the
-recommended number of \(k\) clusters, which for both methods was two
-clusters (see Supplemental Material A). We then entered the
-participants' PC-scores from the feature reduction step into the k-means
-algorithm. To avoid local minima we used 100 random initial centroid
-positions. In the final solution the k-means analysis assigned 80
-participants to cluster 1 and 76 participants to cluster 2.
+In our case we entered the participants' PC-scores from the feature
+reduction step into the k-means algorithm. Because we did know the
+underlying number of clusters within our sample, we calculated the
+cluster solutions for \(k =\) 2 through 10. To avoid local minima we
+used 100 random initial centroid positions for each run. Each of the 9
+cluster solutions converged within the iteration limit. In the next step
+we will then evaluate which of the extracted cluster solutions offers
+the best fit with the data.

### 2023-02-24 latex-results/06ClusterEvaluationPerformance.tex
:sparkles: cluster and performance results expl added

In our own illustration example, we used the \texttt{cluster.stats()}
function from the \texttt{fpc} \textsf{R} package, which calculates a
wide variety of internal cluster validity statistics for each of the
extracted clustering solutions. With real-world data no single
evaluation measure is likely perfect, and different measures may produce
different results depending on the characteristics of the data and the
research question being addressed \citep{kittler1998}. It is therefore
important to consider a variety of evaluation measures and to carefully
interpret the results in the context of the specific analysis
\citep{vinh2009}. We found that across most indices, the analysis with
\(k=2\) clusters performed the best. Three commonly reported indices we
would like to highlight are the comparison of within clusters sum of
squares, the average silhouette score, and the Calinski-Harabasz index.
The first statistic we looked at was the a total within-cluster sum of
square \(WCSS\) (also see \equatref{eq:kWCSS}). While the within cluster
variation will naturally decrease with (more) smaller clusters, we
observed that the decrease in \(WCSS\) was largest until \(k=2\) after
which the decrease was much smaller. This method is also called the
\textit{elbow method} \citep{syakur2018}. We then looked at a second,
commonly used measure, the average silhouette score. This statistic
measures the degree to which each time feature data point is similar to
other points within the same cluster, compared to points in other
clusters \citep{rousseeuw1987}. In our case, the \(k=2\) solution
maximized the silhouette coefficient (\(s_2=\) 0.09). Finally, the
Calinski-Harabasz index assesses the compactness and separation of the
clusters by asssessing the ratio of the sum of between-clusters
dispersion and of inter-cluster dispersion for all clusters --- thus,
the higher the score the better the performances \citep{calinski1974}.
In our case the \(k=2\) solution also showed the highest
Calinski-Harabasz index (\(CH_2=\) 16.38; a full table of all extracted
validity statistics is available in Supplemental Material
A)\footnote{It is important to note that another commonly assessed aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}. This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}. Especially when comparing different clustering algorithms one common index is the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.}.
In the final \(k=2\) solution the k-means analysis also assigned a
relatively even number of participants to cluster 1
(\$n\_\{C\_1\}=\(80) and cluster 2 (\)n\_\{C\_1\}=\$76).

### 2023-02-24 latex-results/07ClusterEvaluationInterpretation.tex
:sparkles: cluster and performance results expl added

--- 2023-01-17 methods-and-results.tex
@@
-\subsection{Data}
-
-We used a data set following migration experiences collected by
-\citet[][]{Kreienkamp2022b}. The data set consists of three studies that
-followed migrants who had recently arrived in the Netherlands in their
-daily interactions with the Dutch majority group members. After a
-general migration-focused pre-questionnaire, participants were invited
-twice per day to report on their (potential) interactions with majority
-group members for at least 30 days. The short ESM surveys were sent out
-at around lunch (12pm) and dinner time (7pm). After the 30 day study
-period participants filled in a post-questionnaire that mirrored the
-pre-questionnaire. Participants received either monetary compensation or
-partial course credits based on the number of surveys they completed.
-For our empirical example we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Full
-methodological details are available in \citet[][]{Kreienkamp2022b}.
-
-\subsubsection{Sample}
-
-The original studies included 207 participants (\(N_{S1}=\) 23,
-\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 measurements.
-The studies differed substantially in the maximum length of
-participation (\(\text{max}(t_{S1})=\) 63, \(\text{max}(t_{S2})=\) 69,
-\(\text{max}(t_{S3})=\) 155). This was likely due to the option to
-continue participation without compensation in the later two studies. To
-make the three studies comparable in participation and time frames, we
-iteratively removed all measurement occasions and participants that had
-more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now a lot more comparable
-(\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
-procedure and study-specific exclusions see Online Supplemental Material
-A.
-
-\subsubsection{Variables}
-
-We included 12 variables that were available across all three studies
-and captured information about the participant's interactions, as well
-as cognitive-, emotional-, and motivational self in relationship with
-the majority group. We chose these two aspects in particular because (1)
-the interaction-specific information exemplified the structural
-missingness issue of modern ESM data and (2) the motivational,
-emotional, and cognitive experience offered a diverse conceptualization
-of migration experience (beyond behavioral measurements) that is
-becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full
-methodological details are available in Online Supplemental Material A,
-but basic item information, descriptives, and correlations are also
-available in \tblref{tab:descrLong}.
-
-\input{tables/descrLong}
-
-\subsection{Analysis and Results}
-
-To perform the main analysis on our selected variables, we showcase the
-extract--reduce--cluster--evaluate steps in sequential order. For each
-of the steps fully reproducible and annotated code is available in
-Online Supplemental A as well as in the accompanying OSF- and GitHub
-repositories (REF).
-
-\subsubsection{Feature extraction}
-
-For the features, we selected and extracted one of each of the features
-proposed in \tblref{tab:esmFeatures}. We particularly chose the more
-robust median and median absolute deviation (MAD) for central tendency
-and variability. Within the variability structure, we chose the slightly
-more intuitive mean absolute change (MAC) for stability and stayed with
-the common lag--1 autocorrelation for inertia. For the trend summaries,
-we chose the overall linear regression slope to describe the linear
-trend and we summarized the nonlinear trend with the estimated degrees
-of freedom of an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line. We chose these features in
-particular because we consider them to be the most broadly applicable
-features. We also extracted the participant's number of completed ESM
-measurements to ensure that the clusters are comparable in that regard.
-We provide an R-function that automatically extracts and prepares a
-large selection of the time series feature operationalizations presented
-in \tblref{tab:esmFeatures} in our GitHub repository (see the
-\texttt{featureExtractor} function).
-
-For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
-the an ordered list of values from the time series of variable \(j\) for
-participant \(i\). The calculation depends on whether the number of
-measurements in a time series \(n\) is odd or even.
-
-\begin{equation} \label{eq:median}
-  M(X_{ij}) = 
-    \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
-    \end{cases}
-\end{equation}
-
-For distribution we chose the \textit{Median Absolute Deviation}
-(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
-in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
-
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
-\end{equation}
-
-mean absolute change (MAC) for stability
-
-\begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
-\end{equation}
-
-lag--1 autocorrelation for inertia: lag \(l=1\)
-
-\begin{equation} \label{eq:ar}
-  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
-  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
-\end{equation}
-
-overall linear regression slope to describe the linear trend
-
-summarized the nonlinear trend with the estimated degrees of freedom of
-an empty GAM spline model (edf) --- the edf summarizes the
-\textit{wiggliness} of the spline trend line.
-
-After the feature extraction, we found that about 1.40\% of the
-extracted features are missing across the 72 features per participant.
-This might, for example, happen if participants do not have two
-subsequent measurements with outgroup interactions, so that an
-autocorrelation with lag-1 cannot be calculated for the contact-specific
-variables. The small number of missing values indicates that the
-feature-based approach indeed largely avoids the structural missingness
-issue. The few missing values can, however, be an issue for some feature
-reduction or feature clustering algorithms. We, thus, impute the missing
-feature values with a single predictive mean matching imputation using
-the MICE library.
-
-\subsubsection{Feature reduction}
-
-For the feature reduction, we chose the common
-\textit{principal component analysis} (PCA). Some of the more
-tailor-made feature selection algorithms can be more accurate in
-reducing the feature dimensionality and might retain feature importance
-information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established
-within the psychometric literature and can broadly be applied to a wide
-variety of studies in an automatize manner. As our aim is to present a
-general illustration that can also be adopted for more general data
-descriptive uses, we present the workflow using a PCA here but we
-encourage users to consider more specialized methods as well.
-
-To use the principal component analysis with our extracted time series
-features, we first standardize all features across participants to
-ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
-method to iteratively explain the most of the remaining variance while
-also ensuring that the components are linearly uncorrelated. In
-practice, this meant that the PCA decomposed the 72 features into 72
-principal components but now (because of the uncorrelated linear
-transformations) the first few principal components will capture a
-majority of the variance. We can then decide how much information (i.e.,
-variance) we are willing to sacrifice for a reduced dimensionality. A
-common rule of thumb is to use the principal components that jointly
-explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we select the first 27
-principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 27 PC-scores for each participant
-(i.e., the participants' coordinates in the reduced dimensional space).
-
-We would like to comment on two practical matters when using principal
-components --- the amount of dimensionality reduction and the
-interpretation of the principal components. As for the expected
-dimensionality reduction, given its methodology, PCAs tend to `work
-better' at reducing dimensions with (highly) correlated variables. Thus,
-with a set of very homogeneous variables and features users will need
-less principal components to explain a large amount of variance, while a
-more diverse set of variables and features will tend to require more
-principal components to capture the same amount of variance. Our 27
-principal components are still a relatively high number of variables but
-this is not surprising as we chose a diverse conceptualization and a
-diverse set of time series features. As for the interpretability, PCA
-allows users to extract information on the meaning of the principal
-components. In particular, because the principal components are linear
-combination of original features, users can extract the relative
-importance of each feature for the extracted principal components (i.e.,
-the eigenvectors). While this can be useful in understanding the
-variance in the original data or help with manual feature selection, we
-use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original
-features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
-
-\subsubsection{Feature clustering}
-
-For our illustration of the feature clustering procedure, we use the
-generalizable and efficient centroid-based k-means clustering. The
-procedure needs the user to make a selection on the optimal number of
-clusters in order to run the main algorithm. We use the Elbow and the
-Silhouette method, which both suggest two clusters to be the optimal
-solution (see Supplemental Material A). We then entered the
-participants' PC-scores into the k-means algorithm. We used the
-recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
-random initial centroid positions to avoid convergence to a sub-optimal
-solution (i.e., local minima). The k-means analysis assigned 80
-participants to cluster 1 and 76 participants to cluster 2.
-
 In order to interpret the two clusters, we compare the two clusters in

### 2023-02-24 sections/05FeatureClustering.Rmd
:sparkles: cluster and performance results expl added

@@
 scaled_pc <- pca.out #%>% scale
-
-# Elbow method
-fviz_nbclust(scaled_pc, kmeans, method = "wss") +
-  geom_vline(xintercept = 2, linetype = 2)+
-  labs(subtitle = "Elbow method") +
-  theme_Publication()
-# Silhouette method
-fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
-  labs(subtitle = "Silhouette method") +
-  theme_Publication()
 ```
@@
 
-kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
+kmeansRes <- list()
+for (i in 2:10) {
+  kmeansRes[[i-1]] <- kmeans(scaled_pc, centers = i, nstart = 100)
+}
 ```
-
-```{r}
-#| label: Plot Convex Hulls
-
-fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
-fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
-```
-
-<!-- \Question{\textcolor{cyan}{It is good practice to standardize the input of the k-means algorithm.
-However, I think that does not make sense for PCs because the PCA is based on the idea of decomposing variances (and capturing the most variance in PC1, ...).
-If I actually standardize the PC-score, we get really weird k-means results.
-The only discussion of this I could find was a tutorial that DID NOT standardize the PCs: \url{https://medium.com/more-python-less-problems/principal-component-analysis-and-k-means-clustering-to-visualize-a-high-dimensional-dataset-577b2a7a5fe2}  -->
-
-<!-- I had originally written:  -->
-<!-- "To ensure that the clustering is weighing the principal components equally (as opposed to their variance), we standardize the particpants' PC-scores."  -->
-<!-- but I went back to the 'raw' PC-scores now. But we should discuss this.}} -->
 
@@
 It is this $WCSS$ that becomes the objective function to be minimized, by iteratively moving features from one cluster to another \citep{hartigan1979}.
-In particular, the algorithm (1) calculates the cluster centroids of the initial partitioning, (2) checks whether any feature has a centroid that is closer than that of the currently assigned cluster (3) updates the centroids based on any reassigned features, and then iterates between steps two and three until $WCSS$ is minimized (i.e., locally optimal convergence) or a maximum number of iterations is reached.
-Given the iterative nature of the algorithm, the initial partitioning is often important because the algorithm might arrive at a suboptimal clustering where the $WCSS$ cannot be further reduced by moving any feature to another cluster, despite a better solution existing (i.e., a local minimum).
+In particular, the algorithm (1) calculates the cluster centroids of the initial partitioning, (2) checks whether any feature has a centroid that is closer than that of the currently assigned cluster (3) updates the centroids based on any reassigned features, and then iterates between steps two and three until $WCSS$ is minimized (i.e., locally optimal convergence) or a maximum number of iterations is reached \citep{jain2010}.
+Given the iterative nature of the algorithm, the initial partitioning is often important because the algorithm might arrive at a suboptimal clustering where the $WCSS$ cannot be further reduced by moving any feature to another cluster, despite a better solution existing \citep[i.e., a local minimum;][]{timmerman2013}.
 It is, therefore often recommended to run the k-means clustering with several different starting positions.
 
-In our case we used the Elbow and the Silhouette method to determine the recommended number of $k$ clusters, which for both methods was two clusters (see Supplemental Material A).
-We then entered the participants' PC-scores from the feature reduction step into the k-means algorithm.
-To avoid local minima we used 100 random initial centroid positions.
-In the final solution the k-means analysis assigned `r table(kmeansOut$cluster)[1]` participants to cluster 1 and `r table(kmeansOut$cluster)[2]` participants to cluster 2.
+In our case we entered the participants' PC-scores from the feature reduction step into the k-means algorithm.
+Because we did know the underlying number of clusters within our sample, we calculated the cluster solutions for $k =$ 2 through 10.
+To avoid local minima we used 100 random initial centroid positions for each run.
+Each of the 9 cluster solutions converged within the iteration limit.
+In the next step we will then evaluate which of the extracted cluster solutions offers the best fit with the data.

### 2023-02-24 sections/06ClusterEvaluationPerformance.Rmd
:sparkles: cluster and performance results expl added

---
output: latex_fragment
bibliography: ../referencesZotero.bib
csl: ../apa.csl
---

```{r}
#| label: section setup
#| include: false

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "../figures/",
  include = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r}
#| label: cluster perfomance indices

kmeansPerf <- list()
for(i in 2:10){
  kmeansPerf[[i-1]] <- fpc::cluster.stats(d = dist(scaled_pc), clustering = kmeansRes[[i-1]]$cluster, G2 = TRUE, G3 = TRUE, aggregateonly = TRUE) %>% unlist
}
kmeansPerfComp <- data.frame(Reduce(rbind, kmeansPerf))

kmeansStatNames <- c(
  "n" = "number of cases", 
  "cluster.number" = "number of points", 
  "min.cluster.size" = "size of smallest cluster", 
  "noisen" = "number of noise points", # Not relevant for kmeans 
  "average.between" = "average distance between clusters", 
  "average.within" = "average distance within clusters", # (reweighted so that every observation, rather than every distance, has the same weight)
  "max.diameter" = "maximum cluster diameter",
  "min.separation" = "minimum cluster separation", 
  "ave.within.cluster.ss" = "within clusters sum of squares", # generalisation
  "avg.silwidth" = "average silhouette width", 
  "g2" = "Goodman Kruskal's Gamma coefficient", # See Milligan and Cooper (1985), Gordon (1999, p. 62). 
  "g3" = "G3 coefficient", # See Gordon (1999, p. 62)
  "pearsongamma" = "Normalized gamma", # correlation between distances and a 0-1-vector where 0 means same cluster, 1 means different clusters. see Halkidi et al. (2001)
  "dunn" = "Dunn index", # minimum separation / maximum diameter, Dunn index, see Halkidi et al. (2002).
  "dunn2" = "Dunn index 2", # minimum average dissimilarity between two cluster / maximum average within cluster dissimilarity, another version of the family of Dunn indexes
  "entropy" = "entropy of distribution of cluster memberships", # see Meila(2007)
  "wb.ratio" = "average within / average between", 
  "ch" = "Calinski and Harabasz index", # (Calinski and Harabasz 1974, optimal in Milligan and Cooper 1985; generalised for dissimilarites in Hennig and Liao 2013)
  "widestgap" = "widest within-cluster gap", 
  "sindex" = "separation index"
)

kmeansPerfComp %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = "c",
      digits = 2,
      row.names = FALSE,
      col.names = kmeansStatNames,
      caption = "Variables by Interaction Types and Study Availability") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
```

```{r}
#| label: Plot Performance

# Elbow method
fviz_nbclust(scaled_pc, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Elbow method") +
  theme_Publication()
# Silhouette method
fviz_nbclust(scaled_pc, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method") +
  theme_Publication()

fviz_cluster(kmeansOut, geom = "point", data = scaled_pc, ggtheme = theme_Publication())
fviz_cluster(kmeansOut, geom = "point", data = z_data, ggtheme = theme_Publication())
```

In our own illustration example, we used the \texttt{cluster.stats()} function from the \texttt{fpc} \textsf{R} package, which calculates a wide variety of internal cluster validity statistics for each of the extracted clustering solutions. With real-world data no single evaluation measure is likely perfect, and different measures may produce different results depending on the characteristics of the data and the research question being addressed \citep{kittler1998}. It is therefore important to consider a variety of evaluation measures and to carefully interpret the results in the context of the specific analysis \citep{vinh2009}. We found that across most indices, the analysis with $k=2$ clusters performed the best. Three commonly reported indices we would like to highlight are the comparison of within clusters sum of squares, the average silhouette score, and the Calinski-Harabasz index. The first statistic we looked at was the a total within-cluster sum of square $WCSS$ (also see \equatref{eq:kWCSS}). While the within cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$ after which the decrease was much smaller. This method is also called the \textit{elbow method} \citep{syakur2018}. We then looked at a second, commonly used measure, the average silhouette score. This statistic measures the degree to which each time feature data point is similar to other points within the same cluster, compared to points in other clusters \citep{rousseeuw1987}. In our case, the $k=2$ solution maximized the silhouette coefficient ($s_2=$ `r format(round(kmeansPerfComp$avg.silwidth[kmeansPerfComp$cluster.number==2], 2), nsmall = 2)`). Finally, the Calinski-Harabasz index assesses the compactness and separation of the clusters by asssessing the ratio of the sum of between-clusters dispersion and of inter-cluster dispersion for all clusters --- thus, the higher the score the better the performances \citep{calinski1974}. In our case the $k=2$ solution also showed the highest Calinski-Harabasz index ($CH_2=$ `r format(round(kmeansPerfComp$ch[kmeansPerfComp$cluster.number==2], 2), nsmall = 2)`; a full table of all extracted validity statistics is available in Supplemental Material A)\footnote{It is important to note that another commonly assessed aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}. This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}. Especially when comparing different clustering algorithms one common index is the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.}. In the final $k=2$ solution the k-means analysis also assigned a relatively even number of participants to cluster 1 ($n_{C_1}=$`r table(kmeansOut$cluster)[1]`) and cluster 2 ($n_{C_1}=$`r table(kmeansOut$cluster)[2]`).

### 2023-02-24 latex-results/06ClusterEvaluationPerformance.tex
:construction: typo fix

@@
 In the final \(k=2\) solution the k-means analysis also assigned a
-relatively even number of participants to cluster 1
-(\$n\_\{C\_1\}=\(80) and cluster 2 (\)n\_\{C\_1\}=\$76).
+relatively even number of participants to cluster 1 (\(n_{C_1}=\) 80)
+and cluster 2 (\(n_{C_1}=\) 76).

### 2023-02-24 sections/06ClusterEvaluationPerformance.Rmd
:construction: typo fix

@@
 Especially when comparing different clustering algorithms one common index is the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.}.
-In the final $k=2$ solution the k-means analysis also assigned a relatively even number of participants to cluster 1 ($n_{C_1}=$`r table(kmeansOut$cluster)[1]`) and cluster 2 ($n_{C_1}=$`r table(kmeansOut$cluster)[2]`).
+In the final $k=2$ solution the k-means analysis also assigned a relatively even number of participants to cluster 1 ($n_{C_1}=$ `r table(kmeansOut$cluster)[1]`) and cluster 2 ($n_{C_1}=$ `r table(kmeansOut$cluster)[2]`).

### 2023-02-24 manuscript.tex
:memo: cluster interpretation txt moved to Rmd

@@
 
+\subsubsection{Performance}
 Performance evaluation often means assessing the accuracy, stability, and separation or purity of the clustering \citep{keogh2003}.
@@
 
-The interpretation of feature-based time series clustering in psychology involves understanding the meaning and implications of the clusters obtained from the analysis.
-Here, it is also important to consider the interpretability and practical significance of the clusters.
-While clusters with a high degree of internal cohesion may be more statistically robust, they may not necessarily be meaningful or useful in the context of the specific research question being addressed \citep{kaufman1990}.
-To gain a qualitative understanding of the extracted clusters users can use both the patterns that emerge from the clusters themselves as well as comparisons to between-person differences not included in the initial clustering \citep[e.g.,][]{monden2022}.
-As for the inspection of the clusters themselves, one common approach is comparing the clusters based on the average values of meaningful features \citep{Kennedy2021}.
-Assessing the clusters based on their average features allows users to gain an understanding of the developmental profiles.
-One cluster might, for example, have a relatively strong linear increase in well-being, or participants in another cluster might have a relatively high variability of negative emotions.
-Another common option for interpretation is finding an appropriate prototype to represent the cluster's trajectory.
-This is often easiest with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021}.
-For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.
-For most other cases, it is often recommended to use some form of average over time of the responses within the cluster \citep{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.}.
-These prototypes can then, for example, be inspected visually to identify qualitative differences between the cluster prototypes.
-An additional option for interpretations are out-of-cluster comparisons, where users can assess whether the clusters differ on some stable between-person differences, such as personality, age, or situational circumstances \citep[e.g.,][]{monden2022}.
+\subsubsection{Interpretation}
+The interpretation of feature-based time series clustering in psychology involves understanding the meaning and implications of the obtained clusters.
+In order to make sense of the clustering results, we here focus on three general aspects of the results \citep{kaufman1990}.
+(1) Assessing differences between the clusters in the original time series features, (2) comparing the clusters based on prototype developments, (3) comparing the clusters based on between-person differences that were not included in the initial clustering.
 
-\input{latex-results/06ClusterEvaluationInterpretation.tex}
+\input{latex-results/07ClusterEvaluationInterpretation.tex}
 

### 2023-02-24 manuscript.tex
:memo: discussion notes

@@
 
+\section{Discussion}
+
+\begin{center}
+\rule{0.4\textwidth}{.4pt}
+\end{center}
+
+\textbf{NOTES:}
+
+The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
+However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
+
 Before we move to the final discussion and conclusions, we would like to comment on the reporting of feature-based time series clustering.
@@
 
-\section{Discussion}
-
-NOTES:
-The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
-However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
-
----
+\begin{center}
+\rule{0.4\textwidth}{.4pt}
+\end{center}
 

### 2023-02-24 latex-results/07ClusterEvaluationInterpretation.tex
:sparkles: feature interpretation integrated

@@
-In order to interpret the two clusters, we compare the two clusters in
-their features as well as in their raw time series. We first inspect the
-clusters with a focus on the included variables (see
-\fgrref[A]{fig:clusterFeatVar}). We (1) see that for some variables the
-features are generally stronger in separating the clusters (e.g., `how
-cooperative an interaction was' compared to `attitudes towards the
-Dutch'). Additionally, we see that (2) within the variables some
-features are better at distinguishing clusters (e.g., median of
-well-being vs.~MAC of well-being).We then inspect the clusters with a
-focus on the features (see \fgrref[B]{fig:clusterFeatVar}). While this
-is the same data as for the variable focus, we can see more clearly that
-some features are better at distinguishing the clusters across variables
-(e.g., mean and median compared to auto correlations). This offers some
-information on which features were are most important in understanding
-the two extracted groups.
+We first inspect the clusters based on the average values of meaningful
+features
+\citep[see \fgrref[A]{fig:clusterFeatVar}{]}{[}{]}\{Kennedy2021\}. We
+(1) see that for some variables the features are generally stronger in
+separating the clusters (e.g., `how cooperative an interaction was'
+compared to `attitudes towards the Dutch'). Additionally, we see that
+(2) within the variables some features are better at distinguishing
+clusters (e.g., median of well-being vs.~MAC of well-being).We then
+inspect the clusters with a focus on the features (see
+\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
+variable focus, we can see more clearly that some features are better at
+distinguishing the clusters across variables (e.g., mean and median
+compared to auto correlations). This offers some information on which
+features were are most important in understanding the two extracted
+groups.
 
@@
 
-We can then combine the two focus approaches to assess the developments
-between the two groups more holistically (see \fgrref{fig:clusterTs}).
+In the second step, we look at prototypical trajectories of the
+clusters. For k-means clustering it is often recommended to use the
+average over time of the responses within the cluster
+\citep[see \fgrref{fig:clusterTs}][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021}.
+For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
 Immediately striking are the mean differences, where participants in the
@@
 
-We can also assess the clusters across any other person-level variable.
-This out-of-feature comparison allows us to check for data artifacts, as
-well as check whether the developmental clusters are associated with
-important social markers and individual differences. To illustrate
-artifact checks, we added the number of measurements into the comparison
-and find that the participants in the detereoration cluster on average
-completed more ESM surveys and reported on more intergroup interactions
-than the cluster with the more positive interactions (see
-\fgrref[B]{fig:clusterTs}). While this difference could indicate that
-the clusters might not entirely be comparable in the response patterns,
-we can find some relief in our data exclusion procedures during which we
-ensured that the general time frame and completion rates were not too
-dissimilar. To illustrate the utility of individual differences, we
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). \fgrref[B]{fig:clusterTs} illustrates that
-participants in the deteriorating cluster reported substantially higher
-levels of everyday discrimination. Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
+Finally, we can also assess the clusters across other person-level
+variable \citep[e.g.,][]{monden2022}. This out-of-feature comparison
+allows us to check for data artifacts, as well as check whether the
+developmental clusters are associated with important social markers and
+individual differences. To illustrate artifact checks, we added the
+number of measurements into the comparison and find that the
+participants in the detereoration cluster on average completed more ESM
+surveys and reported on more intergroup interactions than the cluster
+with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
+While this difference could indicate that the clusters might not
+entirely be comparable in the response patterns, we can find some relief
+in our data exclusion procedures during which we ensured that the
+general time frame and completion rates were not too dissimilar. To
+illustrate the utility of individual differences, we compare the two
+samples in terms of the participants' self-reported discrimination
+experiences in the Netherlands (measured during the post-measurement).
+\fgrref[B]{fig:clusterTs} illustrates that participants in the
+deteriorating cluster reported substantially higher levels of everyday
+discrimination. Thus, both intensive longitudinal and cross-sectional
+variables that were not included in the original clustering step can be
+used to explore and understand the cluster differences in more detail.
 

### 2023-02-24 sections/07ClusterEvaluationInterpretation.Rmd
:sparkles: feature interpretation integrated

--- 2023-02-17 sections/06ClusterEvaluation.Rmd
@@
 
-In order to interpret the two clusters, we compare the two clusters in their features as well as in their raw time series.
-We first inspect the clusters with a focus on the included variables (see \fgrref[A]{fig:clusterFeatVar}).
+We first inspect the clusters based on the average values of meaningful features \citep[see \fgrref[A]{fig:clusterFeatVar}][]{Kennedy2021}.
 We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
@@
 ```
-We can then combine the two focus approaches to assess the developments between the two groups more holistically (see \fgrref{fig:clusterTs}).
+
+In the second step, we look at prototypical trajectories of the clusters.
+For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs}][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021}.
+For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
 Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
@@
 
-We can also assess the clusters across any other person-level variable.
+Finally, we can also assess the clusters across other person-level variable \citep[e.g.,][]{monden2022}.
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.

### 2023-02-24 manuscript.tex
:memo: dim red sct shortened

@@
 The difficulty of working with such a large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
-To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection \citep[e.g.,][also see \tblref{tab:featureReduction} for a full overview of the approaches, methods, and common algorithms]{erdogmus2008}.
 
-Feature selections seek to create a subset of the most important features.
+To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection \citep[e.g.,][]{erdogmus2008}.
+While feature selection refers to the process of identifying and selecting a subset of relevant features from the original feature set \citep{alelyani2014}, feature projection refers to the process of transforming the original feature set into a new feature set of lower dimensionality \citep[][]{carreira-perpinan1997}.
+In general, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
+Feature projection methods, on the other hand, have been popular because of their widespread availability as well as their high efficiency.
+We provide an overview of the common approaches, an intuitive introduction to common methods, and exemplar algorithms in \tblref{tab:featureReduction}.
+
+%Feature selections seek to create a subset of the most important features.
 The many available approaches differ in how they seek to determine the importance of the individual features \citep{alelyani2014}.
@@
 
-The second general approach to the dimensionality curse has been feature projection.
+%The second general approach to the dimensionality curse has been feature projection.
 Broadly speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features \citep[i.e., the data is projected to a lower dimensional space; e.g.,][]{carreira-perpinan1997}.
@@
 
+\newpage
 \section{Discussion}
@@
 
-\textbf{NOTES:}
+\noindent\textbf{NOTES:}
 

### 2023-02-24 tables/FeatureReduction.tex
:art: tbl placement
:memo: dim red sct shortened

@@
 One common selection criteria is the amount of information and (unique) variance a feature captures.
-Univariate methodologies are able identify irrelevant features (i.e., because features do not capture much information), multivariate methods additionally allow to remove redundant features (i.e., because features capture the same information).
+Univariate methodologies are able identify irrelevant features (i.e., because features do not capture much information), multivariate methods additionally allow to remove redundant features \citep[i.e., because features capture the same information][]{yu2004}.
 \linebreak &
@@
 Because the selection process is essentially a search problem this method is computationally intensive.
-Traditionally wrappers have used forward selections or backwards eliminations but recently alternative approaches have been proposed based such as ant colony and swarm intelligence paradigms.
+Traditionally wrappers have used forward selections or backwards eliminations but recently alternative approaches have been proposed based such as ant colony and swarm intelligence paradigms \citep[e.g., see][]{tang2014}.
 \linebreak &
@@
         Hybrid selection methodological combine filter and wrapper methodologies to avoid the shortcomings of the individual methods.
-An initial filter step reduces computational effort (efficiency) and the wrapper ensures high performance (effectiveness).
+An initial filter step reduces computational effort (efficiency) and the wrapper ensures high performance \citep[effectiveness; e.g.,][]{alelyani2014}.
 \linebreak &
@@
         Linear dimensionality reduction methodologies use linear transformations of the original data to stretch and shift the data in such a way that the data can be `projected' to a lower dimensional space without loosing too much information.
-These methods are well-established, tend to be fast, and usually do not need much conceptual input from the user.
+These methods are well-established, tend to be fast, and usually do not need much conceptual input from the user \citep[][]{cunningham2015}.
 \linebreak &
@@
         Nonlinear dimensionality reduction methods also seek to map high-dimensional data to a lower dimensional space.
-However, nonlinear methods have been developed to preserve the local and global structure of more complex multidimensional patterns.
+However, nonlinear methods have been developed to preserve the local and global structure of more complex multidimensional patterns \citep[e.g.,][]{lee2007}.
         \linebreak &

### 2023-02-24 sections/04FeatureReduction.Rmd
:memo: feat red adj

@@
 
-For the feature reduction, we chose the common \textit{principal component analysis} (PCA).
+For our own illustration data we chose a feature projection method to reduce the dimensionality of our extracted features.
+In particular, we selected the commonly used \textit{principal component analysis} (PCA).
 Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.

### 2023-06-12 manuscript.tex
Updates from Overleaf

@@
 Fortunately, past conceptual and empirical works in the experience sampling (ESM) literature have collected a number of meaningful aspects of psychological time series\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
-While the terms come for different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events.
+While the terms come from different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events.
 In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}.
@@
 For most conceptualizations, the proposed number of aspects ranges from four to twelve key features \citep[][]{wang2006, dejonckheere2019}.
-Each of these statistics not only captures a distinct aspect of psychological time series, but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.
+Each of these statistics not only captures a distinct aspect of psychological time series but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.
 
@@
 To take the common VAR model as an example, the model explicitly assumes that the time intervals between measurements are consistent (i.e., equidistant measurement assumption), it does not allow for missing values, and it assumes that means or variances do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}.
-These assumptions however stand incontrast to types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
+These assumptions however stand in contrast to types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
 Model parameters might, thus, not always accurately capture the time series and are often restricted in the features they capture.
@@
 As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-ESM APPLICATION.
+RE-ADD ESM APPLICATIONS.
 We argue that for psychological time series data, feature-based clustering offers a flexible variety, theory-based selection, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.

### 2023-06-24 latex-results/01DataDescription.tex
typos

@@
 had recently arrived in the Netherlands in their daily interactions with
-the Dutch majority group members \citet[][]{Kreienkamp2022b}. After a
+the Dutch majority group members \citep[for the data set see][]{Kreienkamp2022b}. After a
 general migration-focused pre-questionnaire, participants were invited
@@
 selection and preparation is described as part of the illustration below
-but additional methodological details about the study setup are
-available in \citet[][]{Kreienkamp2022b}.
+but for additional methodological details about the study setup see \citet[][]{Kreienkamp2022b}.

### 2023-06-24 latex-results/03FeatureExtraction.tex
typos

@@
 radicalization might for example ask: ``Are people settled in their
-attitudes towards migrants or vary across the measurement period?'\,'
+attitudes towards migrants or do they vary across the measurement period?'\,'
 

### 2023-06-24 latex-results/07ClusterEvaluationInterpretation.tex
:memo: feat red adj
typos

@@
 We first inspect the clusters based on the average values of meaningful
-features
-\citep[see \fgrref[A]{fig:clusterFeatVar}{]}{[}{]}\{Kennedy2021\}. We
+features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We
 (1) see that for some variables the features are generally stronger in
@@
 (2) within the variables some features are better at distinguishing
-clusters (e.g., median of well-being vs.~MAC of well-being).We then
+clusters (e.g., median of well-being vs.~MAC of well-being). We then
 inspect the clusters with a focus on the features (see
@@
 \citep[see \fgrref{fig:clusterTs}][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}.
 For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
@@
 experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterTs} illustrates that participants in the
+\fgrref[B]{fig:clusterFeatVar} illustrates that participants in the
 deteriorating cluster reported substantially higher levels of everyday

### 2023-06-24 manuscript.tex
typos

@@
 % formatting links in the PDF file
+\definecolor{myurlcolor}{HTML}{1a0dab}
 \hypersetup{
@@
 colorlinks   = true, %Colours links instead of ugly boxes
-urlcolor     = blue, %blue,Colour for external hyperlinks
+urlcolor     = myurlcolor, %blue,Colour for external hyperlinks
 linkcolor    = black, %blue, Colour of internal links
@@
 Psychological time series data has not only become more common but also more diverse and complex.
-Researchers and practitioners are increasingly interested in non-stationary developmental patterns of multivariate concepts, across different time scales, while contextualized measurements bring about issues of non-imputable missingness.
-At the same time, clustering analyses offer the potential of reducing these complexities to important and meaningful patterns.
-Such procedures aid researchers and practitioners in describing and exploring patterns across participants and can promote the creation of more embedded theories and interventions.
-In this manuscript, we propose feature-based time series clustering as a flexible, transparent, and well-grounded clustering approach that addresses the growing challenges of dimensionality, missingness, and time scales.
-We introduce the individual feature extraction, -reduction, and -clustering steps and illustrate their utility with an empirical example.
-We show that time series features such as means, autocorrelations, and linear trends are not only familiar to psychological researchers but are also conceptually meaningful in interpreting time series clusters.
-Beyond the conceptual and methodological introduction, we also provide practical algorithm overviews and readily available code for data preparation, analysis, and interpretation.
+Researchers and practitioners are increasingly interested in identifying person-level differences in the developments of their patients, or participants.
+Past research has proposed a number of `dynamic measures' that describe meaningful developmental patterns for psychological data (e.g., instability, inertia, linear trend).
+Yet, commonly used clustering approaches are often not able to include these meaningful measures (e.g., due to model assumptions).
+We propose feature-based time series clustering as a flexible, transparent, and well-grounded approach that clusters participants based on the dynamic measures directly using common clustering algorithms.
+We introduce the approach and illustrate the utility of the method with real-world empirical data that highlight common ESM challenges of multivariate conceptualizations, structural missingness, and nonlinear trends.
+We use the data to showcase the main steps of input selection, feature extraction, feature reduction, feature clustering, and cluster evaluation.
+We also provide practical algorithm overviews and readily available code for data preparation, analysis, and interpretation.
 
@@
 % **CHEAT SHEET / LEGEND**
-%
+
 % Comments:
@@
 % '\Question' question note indicator in PDF (triangle with question mark)
-%
+
 % Citation (with Natbib citation style):
@@
 % '\citeauthor' and '\citeyear' only year or author
-% 
+
 % Headings:
@@
 % '\subparagraph{}' heading level 5
-%
+
 % formatting:
@@
 % 'begin{center}...end{center}' center text or object
-%
+
 % Math mode:
@@
 % '\sum_{k=1}^n' sum for 1 through n
-%
+
 % Insert things from elsewhere:
@@
 % '\item' item in list section
-%
+
 % Symbols:
@@
 Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][]{hamaker2017}.
-These extensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
+These intensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
 Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
-Researchers and practitioners are, for example, asking: ``Do the key symptoms of different patients develop in contrasting ways?'' \citep{Monden2015} or ``How do migrants differ in the development of their core psychological needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
+Researchers and practitioners are, for example, asking: ``Do the symptoms of different patients develop in contrasting ways?'' \citep{Monden2015} or ``How do migrants differ in the development of their self-reported needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
 There is, thus, a clear need for analysis techniques that identify person-level differences in developmental patterns for psychological data.
@@
 Such analyses are however only possible with highly regular and controlled data, such as EEG data \citep{huang1985}, or when clustering variables within the person rather than identifying differences between persons \citep{haslbeck2022}.
-Multivariate data from extensive psychological survey studies, as we discuss here, are seldomly directly comparable across persons \citep[e.g., ][]{faloutsos1994}.
+Multivariate data from intensive psychological survey studies, as we discuss here, are seldomly directly comparable across persons \citep[e.g., ][]{faloutsos1994}.
 We provide a broader embedding of the current methods for psychological time series in the discussion section.}.
@@
 
-Fortunately, past conceptual and empirical works in the experience sampling (ESM) literature have collected a number of meaningful aspects of psychological time series\footnote{In psychology, extensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
+Fortunately, past conceptual and empirical works in the experience sampling (ESM) literature have collected a number of meaningful aspects of psychological time series\footnote{In psychology, intensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies.
 While the terms come from different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events.
@@
 
-Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models, which either seek to relax specific assumptions \citep[e.g.,][]{denteuling2021} or include additional time series features \citep[e.g., see][]{krone2018}.
+Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models, which either seek to relax specific assumptions \citep[e.g.,][]{denteuling2021} or include additional time series features \citep[e.g., see][]{krone2018, gates2017}.
 In this manuscript, we instead, propose to directly cluster based on important and relevant features.
@@
 As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-RE-ADD ESM APPLICATIONS.
 We argue that for psychological time series data, feature-based clustering offers a flexible variety, theory-based selection, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.
@@
 
-To illustrate the functioning and utility of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected data from research on migration experiences.
+To illustrate the functioning and utility of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected data on migration experiences.
 Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020}.
@@
   The presented number of participants, variables, and measurement occasions are somewhat arbitrary but generally represent common sample sizes found within the literature.
-Also the number of extracted clusters is presented for illustrative purposes only.}
+Also, the number of extracted clusters is presented for illustrative purposes only.}
 \end{figure}
@@
 An overview of the selected features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
-We provide additional background on the theoretical relevance of the selected features in \appref{app:FeatureBackground}.
-For each of the six time series features we chose, we chose a mathematical representation that was appropriate for our type of data.
+For each of the six time series features we chose, we selected a mathematical representation that was appropriate for our type of data.
 We provide a short introduction of each feature below.
@@
 We offer an illustration of the compounding numbers of data points in \fgrref{fig:TSCFlowN}.
-The difficulty of working with such a large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
+The difficulty of finding stable clusters for data with a large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.
 
@@
 \section{Discussion}
+% aims re-iterated
+The purpose of this article was to introduce feature-based time series clustering as an amenable and transparent approach to understanding person-level differences in developmental patterns of psychological time series data.
+Rather than relying on person-specific model parameters, which can be restrictive and assumption-bound, we argue for the more flexible and theoretically grounded approach of directly clustering on important and relevant features of the time series data.
+By leveraging the rich array of dynamic measures, our approach offers the advantages of flexibility, theory-based selection, fewer strict assumptions, and improved interpretability, thus potentially enriching our understanding of heterogeneous psychological processes in intensive longitudinal studies.
 
-\begin{center}
-\rule{0.4\textwidth}{.4pt}
-\end{center}
+% summary
+To illustrate the practical utility of the approach, we applied the method to real-world empirical data that highlight common ESM issues of multivariate conceptualizations, structural missingness, and nonlinear trends \citep[e.g.,][]{ariens2020}.
+With the real-world data, we followed a stepwise approach to discuss key issues during input selection, feature extraction, feature reduction, feature clustering, and cluster evaluation.
+Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research practice, as both the features and the analysis steps are well established within the field, and statistical packages are readily available.
+Time series features (such as means or linear trends) are not only easy to extract, but also hold conceptual meaning for psychological data and can be chosen to address specific research questions (also see \tblref{tab:esmFeatures}).
 
-\noindent\textbf{NOTES:}
+Importantly, we show that feature-based clustering is not only approachable but provides interpretable and transparent insights about the grouped patterns.
+For our example of migration experiences, the method was useful to discern adaptive from more stressful experiences and helped to contextualize divergent experiences.
+We were able to show which variables and which dynamic features were most impactful in discerning the trajectory clusters and were even able to further contextualize the results with out-of-feature comparisons (see, e.g., \fgrref{fig:clusterFeatVar}).
 
-The clustering of model parameters is, thus, technically a restrictive sub-form of feature-based clustering --- where lagged regression coefficients, for example, can be considered to be one type of feature capturing a time series.
-However, beyond model parameters, time series might also be described with any other statistical feature \citep{tiano2021}.
+\subsection{Limitations}
+While feature-based time series clustering offers a promising approach to understanding psychological time series data, it is not without limitations.
+In particular, feature-based clustering has both usability- and robustness limitations across its multiple steps.
 
-Before we move to the final discussion and conclusions, we would like to comment on the reporting of feature-based time series clustering.
-Clustering is inherently an exploratory analysis and, as such, each of the analysis steps involves a set of decisions by the user.
-While this is common for other types of analyses as well, it remains imperative to report the analysis in the most transparent and reproducible manner possible.
+In terms of convenience, each of the steps requires users to make an informed decision about the choice of method and algorithm.
+These additional steps of decision-making and transparency increase the initial barrier to entry.
+We hope that our empirical illustration offers a relatively generalizable procedure that showcases the ease of use, but clustering unfortunately does not offer a universal one-size-fits-all solution.
+
+In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential for incomparable results between methods \citep[e.g.,][]{bastiaansen2019}.
+While the variety and diversity of methods are helpful in finding options even for more complex types of data, different algorithms often offer different results \citep[e.g.,][]{keogh2005}.
+And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their idiosyncratic shortcomings and might, for example, be sensitive to outliers \citep{xu2015}.
+
+As an illustration, the choice of features to extract from the time series data is a critical step that can significantly influence the results of the clustering process.
+In the current example, we chose to extract features such as means, autocorrelations, and linear trends, which are psychologically and conceptually meaningful in interpreting our time series clusters.
+However, this selection is not exhaustive and may not capture all relevant aspects of the time series data.
+For example, we did not consider attributes like periodicity or spectral density, which could shed light on the data's cyclical patterns.
+The choice of features largely hinges on the researcher's specific research question and assumptions about the data, thereby injecting a level of subjectivity into the process.
+Similar challenges arise with the choice of the clustering algorithm or the cluster illustration.
+These challenges are not unique to feature-based clustering, rather they are common to all clustering approaches \citep{liao2005}.
+However, it is important to remember that multi-stepped data-driven approaches are particularly vulnerable to the impact of the researchers' degrees of freedom.
+
+One potential remedy to many of the limitations of feature-based clustering lies in transparently and reproducibly reporting the decisions by the user for each of the analysis steps.
 In our own description of the method, we have separated both the time series features as well as the analysis steps into distinct taxonomies in order to facilitate the transparency of the individual steps and decision moments.
 Beyond the structures proposed here, \citet{vandeschoot2017} have proposed an extensive checklist for latent trajectory studies.
-Most of their recommendations and reporting guidelines apply to feature-based clustering as well and might even offer a template for researchers hoping to pre-register their analysis procedures.
+Most of their recommendations and reporting guidelines also apply to feature-based clustering and might even offer a template for researchers who hope to preregister their analysis procedures \citep[also see][]{kirtley2021}.
 
-\begin{center}
-\rule{0.4\textwidth}{.4pt}
-\end{center}
+\subsection{Implications}
+Notwithstanding the limitations, we believe that feature-based clustering offers exciting new potential for psychological time series.
+Across a variety of scientific fields, feature-based time series clustering has been valued for its structuring and simplifying utilities.
+Within the growing field of psychological time series, ESM data is becoming increasingly more complex in terms of participants, variables, and time points.
+Feature-based time series clusters can help reduce these complexities to important and meaningful patterns, an increasingly important task for researchers and practitioners.
 
-% aims re-iterated
-We sought to introduce a flexible and transparent analysis that aids researchers in describing, exploring, and understanding psychological time series data.
-Such an analysis needs to address the demands of the growing variety of ESM research.
-In particular, modern describe- and explore analyses should be able to deal with high dimensionality, structural missingness, and diverse time-varying developments.
-We have argued that feature-based time series clustering is one option that meets this challenge.
+For researchers, the feature-based time series clustering approach offers a number of compelling implications.
+The flexibility and interpretability mean that feature-based time series clustering can be applied to a wide range of data types and research questions.
+The method can be used to contextualize preexisting groups by extracting their time series features and comparing a data-driven approach with existing group labels.
+Furthermore, the feature-based approach can also be used as an exploratory, descriptive, or predictive approach to intensive longitudinal data.
+By reducing the complexities of ESM data to important and meaningful patterns, a bottom-up approach can aid in the creation of more embedded theories and interventions, or simply in describing the often complex and heterogeneous data researchers collect during ESM studies.
 
-We have methodologically deconstructed the feature extraction, feature reduction, and feature clustering steps and offered methodological as well as conceptual examples for each step.
-Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research practice as both the features as well as the analysis steps are well-established within the field and statistical packages are readily available.
-Time series features (such as means or linear trends) are not only easy to extract but also hold conceptual meaning for psychological data and can be chosen to address specific research questions (also see \tblref{tab:esmFeatures}).
+Looking ahead, the feature-based time series clustering approach opens up new avenues for future research.
+While the approach has shown promise in dealing with the challenges of dimensionality, missingness, and time scales, there is potential for further refinement and expansion.
+To showcase the exciting potential for future methodological integrations, we will briefly consider the broad range of alternative approaches to time series clustering (see \fgrref{fig:tsClustTax}).
+
+For instance, given that the approach does not assume the stationarity restrictions of many model-based approaches, future research may want to integrate the research on capturing nonlinear trends from the shape-based approaches into the time series feature world.
+Bicoherence metrics, polynomial-, and differential equation parameters may be used to capture the type and structure of nonlinear developments \citep[e.g.,][]{caro-martin2018, mayor2022}.
+All of which would add to the under-studied trend features of ESM data.
+
+Similarly, iterative or embedded approaches have a number of upsides, such as extracting a cluster-level model that captures the prototypical patterns of the group \citep[e.g.,][]{denteuling2021, lane2019}.
+To provide similar functionality, researchers may evaluate the utility of using recurrent neural networks, such as long short-term memory (LSTM) models to generate cluster-level trajectories \citep[e.g., by training the LSTM based on the feature matrix and the original time series as the target; e.g., see][]{fraley2002, nagin1999}.
+Alternatively, rule-based classifiers could extract rule sets that narrow down the conditions under which a given feature set is assigned to a specific cluster \citep[][]{benard2019}.
 
@@
 
-To illustrate the practical utility of the approach, we applied the method to real-life empirical data with diverse conceptualizations, missingness, and nonlinear trends.
-We make available the full code, custom functions, and illustrations to aid the approachability of the method.
-In the interpretation step, we also show how the clusters can be compared across the original feature set as well as other meaningful situational and person-specific differences (e.g., see \fgrref{fig:clusterFeatVar}).
-We also find that for our example of migration experiences, the method was useful in discerning adaptive from more stressful experiences and helped to contextualize the diverging experiences.
+Beyond the direct academic use, the feature-based time series clustering approach also addresses practical and applied uses.
+For practitioners, the approach offers a practical and grounded method for dealing with the challenges of complex and messy patient-, customer-, or user data.
+Not only does the approach directly deal with dimensionality, missingness, and time scales in the time series, but the interpretability and transparency aspects offer particular utility in applied settings, where the costs of misspecification are high.
+Additionally, the approach is also more readily accessible to practitioners who may not have extensive training in complex data analysis techniques.
+We provide practical algorithm overviews and readily available code for data preparation, analysis, and interpretation.
+The ability to identify and interpret meaningful patterns in time series data can have significant implications for practice, particularly in fields such as clinical, organizational, or social psychology, where understanding individual differences and developmental patterns can inform interventions and decision-processes.
 
-As with any statistical method, feature-based time series clustering is not without limitations. 
-In particular, feature-based clustering has both a usability as well as robustness limitation in its multiple steps.
-In terms of convenience, each of the three steps requires users to make an informed decision about the choice of method and algorithm.
-These additional steps of decision-making and transparency heighten the initial barrier to entry.
-We hope that our empirical illustration offers a relatively generalizable procedure that showcases the ease of use but clustering unfortunately does not offer a universal one-size-fits-all solution.
-
-In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential incomparable results across methods.
-While the variety and diversity of methods are helpful in finding options even for more complex types of data, different algorithms often offer different results.
-And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their idiosyncratic shortcomings and might, for example, be sensitive to outliers.
-
-Notwithstanding the limitations, we believe that feature-based clustering offers an exciting new potential for psychological time series.
-Across a variety of scientific fields, feature-based time series clustering has been valued for its structuring and simplifying utilities.
-Within the growing field of psychological time series, modern ESM data is becoming increasingly more complex in terms of participants, variables, and time points.
-Feature-based time series clusters can aid in reducing these complexities to important and meaningful patterns, an increasingly important task for researchers and practitioners.
+% conclusion
+In conclusion, we show that feature-based time series clustering can effectively reduce the complexities of psychological time series data to important and meaningful patterns.
+It does so with more flexibility, versatility, and less strict assumptions than many of the commonly used approaches to date.
+As such the feature-based time series clustering approach addresses key challenges in the field and aids researchers and practitioners in describing and exploring patterns across participants.
+We hope that the method adds to the methodological toolkit of ESM researchers and promotes the creation of more embedded methods, theories, and interventions.
 
@@
 These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}.
-These phenomena and principles can be meaningful differences that distinguish participants in different clusters, as well important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
+These phenomena and principles can be meaningful differences that distinguish participants in different clusters, as well as important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}.
 Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}.
@@
 
-Another common type of data that are measurement regiments that collect data in irregular time intervals (i.e., non-equidistant measurements).
+Another common type of data are measurement regiments that collect data in irregular time intervals (i.e., non-equidistant measurements).
 Common are, for example, procedures where participants are asked to respond at random times throughout the day (i.e., signal-contingent) or following specific natural events of interest \citep[i.e., event-contingent; see][]{shiffman2008, myin-germeys2018}.
@@
 
-When it comes to studying developmental trajectories, psychological researchers are often also interested in nonstationary processes because they are more representative of the complex, dynamic behavior of the human mind.
+When it comes to studying developmental trajectories, psychological researchers are often also interested in nonstationary processes because they are more representative of the complex, dynamic patterns of the human mind.
 In psychology, nonstationary processes are typically used to study phenomena such as cognitive development \citep[][]{quartz1997}, decision-making \citep[][]{ratcliff2016}, and emotion dynamics \citep[][]{bringmann2018b}.
@@
 At the same time, the series may also exhibit a longer-term trend, with overall mood levels increasing or decreasing over the course of several months or years \citep[e.g.,][]{Ram2014}.
-\sout{These different time scales can be studied separately or in combination, using different statistical techniques and modeling approaches \citep[][]{bertenthal2007, jeronimus2019a}.} Different time scales can become an even more difficult issue when different variables in a model develop on different time scales \citep{bringmann2022b}.
+These different time scales can be studied separately or in combination, using different statistical techniques and modeling approaches \citep[][]{bertenthal2007, jeronimus2019a}.
+Different time scales can become an even more difficult issue when different variables in a model develop on different time scales \citep{bringmann2022b}.
 Different time scales are thus also a concern clustering approaches should be able to address.
+
+It is this background of the common challenges of current ESM data, upon which we propose to consider feature-based clustering.
+The flexibility of using a wide variety of features that represent the important developmental patterns allows users to circumvent many of the issues with multi-dimensionality, non-equidistant or missing measurements, non-stationary trends, as well as diverse time scales.
 
@@
 
-\section{Background of the Proposed Psychological Time Series Features}
-\label{app:FeatureBackground}
-
-With its rich history, a staggering variety of time series features have been proposed, not all of which are meaningful for psychological time series.
-As an example, \citet{wang2006} propose a set of nine statistical features for describing a trajectory: (1) trend, (2) seasonality, (3) periodicity, (4) skewness, (5) kurtosis, (6) serial correlation, (7) non-linearity, (8) self-similarity, and (9) chaos \citep[also see][]{fulcher2013}.
-\citet{adya2001} proposed a total of 28 features relevant for forecasting, which largely fit within the selection proposed by \citet{wang2006}.
-A final illustration comes from a commonly used software package for feature extraction `tsfresh', which allows users to extract a total of 794 features to describe all kinds of aspects of a time series \citep[][]{christ2018}.
-
-Given that there are hundreds of possible features to describe psychological time series, one approach to choosing relevant features would be to extract a large number of features and then assess which features are most effective at capturing differences in the time series \citep[e.g.,][]{christ2018}.
-However, such an approach is not always advisable for psychological time series.
-For one, features should reduce the data dimensionality --- it would thus not necessarily be advisable to describe 60 ESM measurements with 794 time series features.
-More importantly, however, careful feature extraction can be crucial in the interpretability and explainability of the results.
-This is particularly the case when features have a conceptual or psychological meaning.
-Taking the concept of well-being as an example, psychologists might be interested in whether certain participants tend to have higher well-being in general (i.e., mean) or whether some participants fluctuate between extremely high and low well-being (i.e., variance).
-But psychologists might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density).
-We will, thus, propose and describe a set of time series features that have a strong backing within the ESM literature and as such offer a reasonable starting point for most concepts within the field.
-
-\paragraph{Central Tendency} Central tendency refers to the statistical measures that represent the ``middle" or ``average" of a set of data.
-The most common measures of central tendency are the mean, median, and mode (Hopkins, 2020).
-As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
-Researchers might, for example, be interested in whether ``Over a one month period, are some people happier than others?''.
-A recent review paper pointed out that the relatively simple mean statistic, is often overlooked in psychological time series analyses but is usually essential for understanding psychological differences \citep{bringmann2018c}.
-As such, several time series modeling approaches now advocate for conscious consideration of the mean \citep[e.g.,][]{bringmann2017a}.
-As with cross-sectional data, for time series data the appropriate measure of central tendency depends on the specific characteristics of the data being analyzed.
-For instance, if the data are normally distributed, the mean is often an appropriate measure.
-If the data are skewed or have outliers, the median may be a better choice \citep{weisberg1992}.
-
-\paragraph{Variability} Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
-In time series analyses, variability is conceptually important because it provides information about the distribution and diversity of the data.
-Especially within-person changes in variability have recently received increasing attention because they can be a potent early warning signal in psychological time series \citep{helmich2020a, vandeleemput2014}.
-A marked within-person increase in variance has, for example, been found to precede crucial transitions --- often considered to be part of a phenonenon know as critical slowing down \citep[i.e., anomalous variances; e.g.,][]{scheffer2009, wichers2019}.
-
-But also more global person-level differences in time series variability have been found to be crucial and relate to important psychological questions.
-Person-specific differences in variability have, for example, been found to be indicative of worse psychological states \citep{myin-germeys2018}.
-And, similarly, person-level differences have also been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
-As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
-Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
-
-Methodologically, there are several measures of variability, including the range, standard deviation, and variance.
-The appropriate measure to use depends on the specific characteristics of the data being analyzed.
-For normally distributed data variances and standard deviation are often preferred, whereas data with outliers or nonparametric distributions are often better described with robust variability measures such as the median absolute deviation \citep{weisberg1992}.
-
-\paragraph{(In)stability} Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009}.
-While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the `jumpy-ness' of the data over time.
-In other words, variability reflects the range or diversity of values in a time series data, while instability reflects the fluctuation or inconsistency in a time series data over time  \citep{trull2008}.
-For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
-Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
-Conceptually, the instability feature, thus, relates to a broad range of research questions, including: ``Do changes tend to be slow and gradual or fast and abrupt?" Methodologically, (in)stability measurements often look at the average change between two consecutive measurements, such as the mean absolute change \citep[e.g.,][]{ebner-priemer2009,barandas2020}.
-
-\paragraph{Self-similarity} Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}.
-That is, self-similarity describes how much a measurement carries over to future measurements.
-There are two main ways in which experiences tend to be connected to future experiences in psychological time series --- resistance to change and seasonality.
-Both forms of self-similarity relate to different types of research questions.
-
-Resistance to change is the self-similarity of a measurement with its next measurement and is often referred to as inertia \citep{kuppens2010}.
-If inertia is high a development tends to stay in a certain state or continues to move towards a certain direction.
-Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
-In a similar vein, high inertia at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
-Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?'' One way to measure inertia in time series data is through the use of autocorrelations or autoregressions with a lag-1, which look at the correlation between a given time series and itself at the previous measurement occasion.
-High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}.
-
-Seasonality is the self-similarity of a measurement that returns in a cyclical manner and is often also referred to as periodicity \citep{gregson1983}.
-Periodicity in psychological time series data, thus, refers to the presence of recurring patterns at regular intervals.
-In psychological time series, these cyclical patterns are not only dangerous for models that assume stationarity \citep{beal2015} but can be interesting psychological properties in their own regard \citep{epskamp2018, schmittmann2013}.
-Humans generally live in seasonal environments, including, for example, exam periods at universities \citep{fuller2003}.
-But also more internally, personality manifestations, for example, have long been assumed to develop cyclically \citep{cattell1957}.
-The associated research questions can be driven by an understanding of external seasonality factors such as: ``Do participants drink more alcohol on Fridays?'' But research questions can also address periodic patterns as they may be the result of underlying psychological processes: ``Does stress increase with fatigue cycles?'' Similar to inertia, periodicity is often measured using autocorrelations but then with the lag of the seasonality or cycle (e.g., daily, seven days, one month, one semester).
-Periodicity can additionally be captured using spectral analysis or decomposition, such as with wave patterns \citep[e.g., fourier or wavelet transformation; e.g.,][]{mayor2022}.
-
-\paragraph{Linear Trend} Linear trends in time series data refer to the presence of a straight-line development of a variable being measured.
-This type of trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
-For psychological time series, researchers have, for example, pointed out linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}.
-Linear trends are often the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}.
-In empirical practice, linear trends are, thus, a common research interest for psychological time series data.
-``Do patient symptoms improve consistently?'' or ``Does worker productivity decline continuously?'', are familiar types of research questions.
-Mathematically, such linear trends are commonly captured using (piecewise) linear regression coefficients, but also using differential equations \citep{chow2019}.
-
-\paragraph{Nonlinearity} Changes in psychology are not always linear, instead nonlinearity is a common feature of psychological time series \citep{hayes2007}.
-As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
-Similarly, patients in recovery from depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
-But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
-For psychological researchers, nonlinear developments are, thus, important in two regards.
-Firstly, nonlinearity as a deviation from linear trends is an important feature defining many psychological time series and secondly, the shape and structure of the nonlinear trend can be essential in understanding the important changes and transitions of human experiences \citep{chow2019}.
-Conceptually, researchers might accordingly have research question about the type of the development: ``Is the development of anxiety a nonlinear process?'' as well as the shape and structure of the development: ``Does anxiety change as a sudden or smooth transition?'' Accordingly, researchers can capture the type of development with nonlinearity parameters such as the bicoherence metrics \citep{cuddy2009}.
-The structure of nonlinear trends is often mathematically more difficult to capture \citep{castro-alvarez2022a}.
-Common ways to capture the trends include the use of polynomial- and differential equation parameters or more broadly to capture how `wiggly' the trend line is (e.g., estimated degrees of freedom of GAM spline models; similar to the number of spikes \citealp[]{caro-martin2018}).
-
 \end{document}

### 2023-06-24 sections/01DataDescription.Rmd
typos

@@
 
-The data set we use consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members \citet[][]{Kreienkamp2022b}.
+The data set we use consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members \citep[for the data set see][]{Kreienkamp2022b}.
 After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
@@
 For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Variable selection and preparation is described as part of the illustration below but additional methodological details about the study setup are available in \citet[][]{Kreienkamp2022b}.
+Variable selection and preparation is described as part of the illustration below but for additional methodological details about the study setup see \citet[][]{Kreienkamp2022b}.

### 2023-06-24 sections/03FeatureExtraction.Rmd
typos

@@
 As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
-Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or vary across the measurement period?''
+Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or do they vary across the measurement period?''
 

### 2023-06-24 sections/07ClusterEvaluationInterpretation.Rmd
:memo: feat red adj
typos

@@
 
-We first inspect the clusters based on the average values of meaningful features \citep[see \fgrref[A]{fig:clusterFeatVar}][]{Kennedy2021}.
+We first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
 We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
 Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
+MAC of well-being).
+We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
 While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
@@
 For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs}][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}.
 For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
@@
 To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterTs} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
+\fgrref[B]{fig:clusterFeatVar} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
 Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.

### 2023-06-24 IllustrationResultsRenderReport.Rmd
feedback LFB

@@
   "mada",
+  # "fpc",
   "data.table",

### 2023-06-24 latex-results/01DataDescription.tex
feedback LFB

--- 2023-02-21 latex-results/01DataDescription.tex
@@
 had recently arrived in the Netherlands in their daily interactions with
-the Dutch majority group members \citet[][]{Kreienkamp2022b}. After a
-general migration-focused pre-questionnaire, participants were invited
-twice per day to report on their (potential) interactions with majority
-group members for at least 30 days. The short ESM surveys were sent out
-at around lunch (12pm) and dinner time (7pm). After the 30 day study
-period participants filled in a post-questionnaire that mirrored the
+the Dutch majority group members
+\citep[for the data set see][]{Kreienkamp2022b}. After a general
+migration-focused pre-questionnaire, participants were invited twice per
+day to report on their (potential) interactions with majority group
+members for at least 30 days. The short ESM surveys were sent out at
+around lunch (12pm) and dinner time (7pm). After the 30 day study period
+participants filled in a post-questionnaire that mirrored the
 pre-questionnaire. Participants received either monetary compensation or
@@
 selection and preparation is described as part of the illustration below
-but additional methodological details about the study setup are
-available in \citet[][]{Kreienkamp2022b}.
+but for additional methodological details about the study setup see
+\citet[][]{Kreienkamp2022b}.

### 2023-06-24 latex-results/03FeatureExtraction.tex
feedback LFB

@@
 distribution and diversity of the data has been found to be indicative
-of worse psychological states \citep{myin-germeys2018}. Person level
-differences of ESM measurements have, for example, been associated with
-higher levels of psycho-pathological recurrences among depression
-patients \citep{timm2017}. As such, psychological researchers and
-practitioners are often empirically interested in person-level
+of worse psychological states \citep{myin-germeys2018, helmich2021}.
+Person level differences of ESM measurements have, for example, been
+associated with higher levels of psycho-pathological recurrences among
+depression patients \citep{timm2017}. As such, psychological researchers
+and practitioners are often empirically interested in person-level
 differences in variability. Researchers on polarization and
 radicalization might for example ask: ``Are people settled in their
-attitudes towards migrants or do they vary across the measurement period?'\,'
+attitudes towards migrants or do they vary across the measurement
+period?'\,'
 
@@
 at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\).
+median of that time series \(X\). We again chose the robust statistic
+because the Median-based measure is less affected by non-normal
+distributions and extreme values or outliers compared to other measures
+of variability like the standard deviation \citep{weisberg1992}
 
@@
 instability feature, thus, relates to a broad range of research
-questions, including: ``Do changes tend to be slow and gradual or fast
-and abrupt?''
+questions, including: ``What is the nature of daily mood swings in
+individuals with bipolar disorder?'' or ``How does the pattern of
+day-to-day (i.e., measurement-to-measurement) fluctuations in
+self-esteem during adolescence reflect the process of identity formation
+and self-concept development?''
 
@@
 \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
-the average absolute difference of measurements \(x\) at time points
-\(t\) and \(t-1\), for each time series \(X\) of participant \(i\) and
-variable \(j\).
+the average absolute difference of two consecutive measurements \(x\) at
+time points \(t\) and \(t-1\), for each time series \(X\) of participant
+\(i\) and variable \(j\).
 
@@
 \end{equation}
+
+Another common measurement of instability is the
+\textit{Mean of the Squared Successive Differences} (\(MSSD\)), which is
+often preferred where differences in magnitude are more important than
+the frequency of those changes, for example, when big shifts in time
+series are considered more impactful or when outliers are meaningful and
+need to be taken into account.
 
@@
 much a measurement carries over to its next measurement
-\citep{kuppens2010}. If inertia is high a development tends to stay in a
-certain state or continues to move towards a certain direction. Because
-high inertia is resistant to change, in emotion dynamics high inertia
-has been found to be indicative of underreactive systems and to be
-characteristic of psychological maladjustment \citep{kuppens2010}. In a
-similar vein, high inertia at baseline was even predictive of the
+\citep{kuppens2010, suls1998}. If inertia is high a development tends to
+stay in a certain state. Because high inertia is resistant to change, in
+emotion dynamics high inertia of negative affect has been found to be
+indicative of underreactive systems and to be characteristic of
+psychological maladjustment \citep{kuppens2010}. In a similar vein, high
+inertia in negative affect at baseline was even predictive of the
 initial onset of depression \citep{kuppens2012}. Conceptually, inertia
 is more broadly connected to research questions such as: ``Do patients
-stay in a depressed mood for several measurements?'\,'
+stay in a negative mood for several measurements?'\,'
 
@@
 linear regression slope to capture the linear trend. The regression
-slope provides the average change from one time point \(t\) to the next
-across all measurements \(x\) of a time series \(X\) of participant
-\(i\) and variable \(j\).
+slope \(b_{ij}\) provides the average change from one time point \(t\)
+to the next across all measurements \(x\) of a time series \(X\) of
+participant \(i\) and variable \(j\). The specific form of the OLS slope
+formula we provide below, calculates \(b_{ij}\) as the sum across all
+time points of the product of the deviation of time \(t\) from its mean
+\(\overline{t}\) and the deviation of \(x_{ij}\) from its mean
+\(\overline{x}_{ij}\) at each time point, divided by the sum across all
+time points of the square of the deviation of time from its mean
+(\(\sum(t-\overline{t})^2\)). Intuitively, the formula captures the rate
+of change of variable \(x_{ij}\) with respect to time. This slope will
+indicate how the variable \(x_{ij}\) changes over time, controlling for
+its mean value and the mean of time. If the slope is positive,
+\(x_{ij}\) increases over time; if it's negative, \(x_{ij}\) decreases
+over time.
 
@@
 anxiety a nonlinear process?'' as well as the shape and structure of the
-development: ``Does anxiety change as a sudden or smooth transition?''
-We summarized the nonlinear trend with the
+development: ``How many spikes in anxiety did a patient experience?'' We
+summarized the nonlinear trend with the
 \textit{estimated degrees of freedom} of an empty GAM spline model. The
@@
 participant's number of completed ESM measurements to ensure that the
-clusters are comparable in that regard. After the feature extraction, we
-found that about 1.40\% of the extracted features are missing across the
-72 features per participant. This might, for example, happen if
+clusters are comparable in that regard (i.e., to exclude spurious
+explanations for the cluster assignments). After the feature extraction,
+we found that about 1.40\% of the extracted features are missing across
+the 72 features per participant. This might, for example, happen if
 participants do not have two subsequent measurements with outgroup
@@
 indicates that the feature-based approach indeed largely avoids the
-structural missingness issue. The few missing values can, however, be an
-issue for some feature reduction or feature clustering algorithms. We,
-thus, impute the missing feature values with a single predictive mean
-matching imputation using the MICE library.
+structural missingness issue. However, even the few missing values can
+be an issue for some feature reduction or feature clustering algorithms.
+We, thus, impute the missing feature values with a single predictive
+mean matching imputation using the MICE library \citep[][]{buuren2011}.

### 2023-06-24 latex-results/04FeatureReduction.tex
:memo: feat red adj
feedback LFB

--- 2023-02-17 latex-results/04FeatureReduction.tex
@@
-For the feature reduction, we chose the common
+For our own illustration data we chose a feature projection method to
+reduce the dimensionality of our extracted features. We particularly
+chose the feature projection method for its broad applicability. We,
+specifically, selected the commonly used
 \textit{principal component analysis} (PCA). Some of the more
@@
 However, PCAs have the distinct benefit that they are well-established
-within the psychometric literature and can broadly be applied to a wide
-variety of studies in an automatize manner. As our aim is to present a
-general illustration that can also be adopted for more general data
-descriptive uses, we present the workflow using a PCA here but we
-encourage users to consider more specialized methods as well.
+within the psychometric literature \citep{jolliffe2011} and can broadly
+be applied to a wide variety of studies in an automatized manner
+\citep{abdi2010}. As our aim is to present a general illustration that
+can also be adopted across use-cases, we present the workflow using a
+PCA here but we encourage users to consider more specialized methods as
+well.
 
@@
 ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA will use linear transformations in
-such a way that the first component captures the most possible variance
-of the original data (e.g., by finding a vector that maximizes the sum
-of squared distances). The following components will then use the same
-method to iteratively explain the most of the remaining variance while
-also ensuring that the components are linearly uncorrelated. In
-practice, this meant that the PCA decomposed the 72 features into 72
-principal components but now (because of the uncorrelated linear
-transformations) the first few principal components will capture a
-majority of the variance. We can then decide how much information (i.e.,
-variance) we are willing to sacrifice for a reduced dimensionality. A
-common rule of thumb is to use the principal components that jointly
-explain 70--90\% of the original variance (i.e., cumulative percentage
-explained variance). For our illustration we select the first 27
-principal components that explain 80\% of the variance in the original
-72 features (reducing the dimensionality by 62.50\%). For the extracted
-principal components we save the 27 PC-scores for each participant
-(i.e., the participants' coordinates in the reduced dimensional space).
+features into the analysis. The PCA uses linear transformations in such
+a way that the first component captures the most possible variance of
+the original data
+\citep[e.g., by finding a vector that maximizes the sum of squared distances][]{jolliffe2002, abdi2010}.
+The following components will then use the same method to iteratively
+explain the most of the remaining variance while also ensuring that the
+components are linearly uncorrelated \citep{shlens2014}. In practice,
+this meant that the PCA decomposed the 72 features into 72 principal
+components but now (because of the uncorrelated linear transformations)
+the first few principal components will capture a majority of the
+variance. We can then decide how much information (i.e., variance) we
+are willing to sacrifice for a reduced dimensionality. A common rule of
+thumb is to use the principal components that jointly explain 70--90\%
+of the original variance
+\citep[i.e., cumulative percentage explained variance; e.g.,][]{jackson2003}.
+For our illustration we select the first 27 principal components that
+explain 80\% of the variance in the original 72 features (reducing the
+dimensionality by 62.50\%). For the extracted principal components we
+save the 27 PC-scores for each participant (i.e., the participants'
+coordinates in the reduced dimensional space).
 
@@
 dimensionality reduction, given its methodology, PCAs tend to `work
-better' at reducing dimensions with (highly) correlated variables. Thus,
-with a set of very homogeneous variables and features users will need
-less principal components to explain a large amount of variance, while a
-more diverse set of variables and features will tend to require more
-principal components to capture the same amount of variance. Our 27
+better' at reducing dimensions with (highly) correlated variables
+\citep[e.g.,][]{jolliffe2002}. Thus, with a set of very homogeneous
+variables and features users will need less principal components to
+explain a large amount of variance, while a more diverse set of
+variables and features will tend to require more principal components to
+capture the same amount of variance \citep[e.g.,][]{abdi2010}. Our 27
 principal components are still a relatively high number of variables but
@@
 features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if the original features were
-chosen for their psychological meaning in understanding the time series
-and broader phenomenon of interest.
+particularly advocate for such an approach if all original features are
+considered meaningful in understanding the time series and users would
+like to retain the features for inpterpretation (irrespective of the
+features importance).

### 2023-06-24 latex-results/06ClusterEvaluationPerformance.tex
:sparkles: feature interpretation integrated
feedback LFB

@@
 The first statistic we looked at was the a total within-cluster sum of
-square \(WCSS\) (also see \equatref{eq:kWCSS}). While the within cluster
+square \(WCSS\) (see also \equatref{eq:kWCSS}). While the within cluster
 variation will naturally decrease with (more) smaller clusters, we
 observed that the decrease in \(WCSS\) was largest until \(k=2\) after
-which the decrease was much smaller. This method is also called the
-\textit{elbow method} \citep{syakur2018}. We then looked at a second,
-commonly used measure, the average silhouette score. This statistic
-measures the degree to which each time feature data point is similar to
-other points within the same cluster, compared to points in other
-clusters \citep{rousseeuw1987}. In our case, the \(k=2\) solution
+which the decrease was much smaller. This method is also sometimes
+referred to as the `elbow method' \citep{syakur2018}. We then looked at
+a second, commonly used measure, the average silhouette score. This
+statistic measures the degree to which each time feature data point is
+similar to other points within the same cluster, compared to points in
+other clusters \citep{rousseeuw1987}. In our case, the \(k=2\) solution
 maximized the silhouette coefficient (\(s_2=\) 0.09). Finally, the
@@
 In the final \(k=2\) solution the k-means analysis also assigned a
-relatively even number of participants to cluster 1 (\(n_{C_1}=\) 80)
-and cluster 2 (\(n_{C_1}=\) 76).
+relatively even number of participants to cluster 1 (\(n_{C_1}=\) 76)
+and cluster 2 (\(n_{C_1}=\) 80).

### 2023-06-24 latex-results/07ClusterEvaluationInterpretation.tex
feedback LFB

@@
 features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We
-(1) see that for some variables the features are generally stronger in
-separating the clusters (e.g., `how cooperative an interaction was'
-compared to `attitudes towards the Dutch'). Additionally, we see that
-(2) within the variables some features are better at distinguishing
-clusters (e.g., median of well-being vs.~MAC of well-being). We then
-inspect the clusters with a focus on the features (see
-\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
-variable focus, we can see more clearly that some features are better at
-distinguishing the clusters across variables (e.g., mean and median
-compared to auto correlations). This offers some information on which
+see that for some variables the features are generally stronger in
+separating the clusters. We, for example see that the item on `how
+cooperative the interaction was' distinguishes the two clusters across
+almost all seven features (except for the auto-correlation, see
+\fgrref[A]{fig:clusterFeatVar}). Compare this to the `outgroup
+attitudes' item where the differences between the clusters are much
+smaller almost all features. We then inspect the clusters with a focus
+on the features (see \fgrref[B]{fig:clusterFeatVar}). While this is the
+same data as for the variable focus, we can see more clearly that some
+features are better at distinguishing the clusters across variables. For
+example, mad and median distinguish the two clusters across almost all
+variables (except for the item of whether the interaction was
+representative of the outgroup). These two features stand in stark
+contrast to other features, such as the lag-1 auto correlations, which
+showed much smaller differences between the two clusters (see
+\fgrref[B]{fig:clusterFeatVar}). This offers some information on which
 features were are most important in understanding the two extracted
-groups.
+groups. Taken these two perspectives together, we can also focus on
+individual features or variables in particular. We, for example, see a
+strong difference in the average well-being, where participants in
+cluster showed a much lower median well-being over the time series. At
+the same time, in terms of stability, both groups have virtually
+identical averages MAC statistics for well-being (see
+\fgrref[A]{fig:clusterFeatVar}). There are, thus, variables and features
+that distinguish the clusters better than others and a combination of
+variables and features lets us explore meaningful group differences in
+more detail.
 
@@
 average over time of the responses within the cluster
-\citep[see \fgrref{fig:clusterTs}][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
+\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
 Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}.
@@
 surveys and reported on more intergroup interactions than the cluster
-with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
-While this difference could indicate that the clusters might not
-entirely be comparable in the response patterns, we can find some relief
-in our data exclusion procedures during which we ensured that the
-general time frame and completion rates were not too dissimilar. To
-illustrate the utility of individual differences, we compare the two
-samples in terms of the participants' self-reported discrimination
-experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterFeatVar} illustrates that participants in the
-deteriorating cluster reported substantially higher levels of everyday
-discrimination. Thus, both intensive longitudinal and cross-sectional
-variables that were not included in the original clustering step can be
-used to explore and understand the cluster differences in more detail.
+with the more positive interactions (see \(n\) in
+\fgrref[B]{fig:clusterFeatVar}). While this difference could indicate
+that the clusters might not entirely be comparable in the response
+patterns, we can find some relief in our data exclusion procedures
+during which we ensured that the general time frame and completion rates
+are similar for all participants. To illustrate the utility of
+individual differences, we compare the two samples in terms of the
+participants' self-reported discrimination experiences in the
+Netherlands (measured during the post-measurement). When looking at the
+group comparison, find that participants in the deteriorating cluster
+reported substantially higher levels of everyday discrimination
+(\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal and
+cross-sectional variables that were not included in the original
+clustering step can be used to explore and understand the cluster
+differences in more detail.
+
+In short, we find that that the feature-based clustering discerned two
+meaningfully different groups of participants. We find an adaptive group
+(cluster 2) that reports higher well-being and more positive outgroup
+attiudes (median) that are also stable over time (MAD, MAC) and tend to
+increase over the 30 day test period (linear trend). This group also
+reported consistently more meaningful, need-fulfilling, and cooperative
+outgroup interactions (median). This group with overwhelmingly positive
+experiences stands in contrast with a more detrimental group (cluster
+1). This cluster, on average, reported much less positive, less
+meaningful, and less fulfilling interactions and interaction patterns
+(median). This group also reported less positive outgroup attitudes,
+lower well-being, and more discrimination experiences (median). At the
+same time, for members of this detrimental cluster conditions seemed to
+deteriorate over time (linear trend), and there was generally less
+consistency in the experiences they were able to have (MAC, MAD, edf).
+
+This cluster separation, then, has a number of empirical and practical
+applications. Firstly, the clusters are descriptive. With tens of
+variables, hundreds of participants, and thousands of measurements,
+singular descriptive statistics are often not able to capture the
+complex patterns that describe the data set. The feature-based
+clustering offers some direct insight into the complexity within the
+data set. Secondly, the clusters identify important groups. The adaptive
+and deteriorating groups offer starting points for empirical exploration
+as well as practical interventions. Researchers can start probing what
+exactly distinguishes the two groups further and generate new bottom-up
+hypotheses. Practitioners, can use the group separation to identify
+individuals in need of assistance and can explore contextual factors
+that might contribute to the difficulties some might face. Thirdly, the
+feature-based approach is flexible and meaningful. We were able to use a
+wide range of time series features that have been central in the ESM
+literature and were able to use them directly to identify meaningful
+groups.
 

### 2023-06-24 sections/03FeatureExtraction.Rmd
feedback LFB

@@
 Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
-In time series analyses, variability is conceptually important because information about the distribution and diversity of the data has been found to be indicative of worse psychological states \citep{myin-germeys2018}.
+In time series analyses, variability is conceptually important because information about the distribution and diversity of the data has been found to be indicative of worse psychological states \citep{myin-germeys2018, helmich2021}.
 Person level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
@@
 For our illustration data, we chose to capture the time series variability with the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \equatref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
+We again chose the robust statistic because the Median-based measure is less affected by non-normal distributions and extreme values or outliers compared to other measures of variability like the standard deviation \citep{weisberg1992}
 
@@
 Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
-Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "Do changes tend to be slow and gradual or fast and abrupt?"
+Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "What is the nature of daily mood swings in individuals with bipolar disorder?" or "How does the pattern of day-to-day (i.e., measurement-to-measurement) fluctuations in self-esteem during adolescence reflect the process of identity formation and self-concept development?"
 
-For our data we chose the \textit{mean absolute change} \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at the average absolute difference of measurements $x$ at time points $t$ and $t-1$, for each time series $X$ of participant $i$ and variable $j$.
+For our data we chose the \textit{mean absolute change} \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at the average absolute difference of two consecutive measurements $x$ at time points $t$ and $t-1$, for each time series $X$ of participant $i$ and variable $j$.
 
@@
 
+Another common measurement of instability is the \textit{Mean of the Squared Successive Differences} ($MSSD$), which is often preferred where differences in magnitude are more important than the frequency of those changes, for example, when big shifts in time series are considered more impactful or when outliers are meaningful and need to be taken into account.
+
 \paragraph{Self-similarity.}
@@
 That is, self-similarity describes how much a measurement carries over to future measurements.
-One important self-similarity in psychological time series is \texit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010}.
-If inertia is high a development tends to stay in a certain state or continues to move towards a certain direction.
-Because high inertia is resistant to change, in emotion dynamics high inertia has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
-In a similar vein, high inertia at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
-Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a depressed mood for several measurements?''
+One important self-similarity in psychological time series is \texit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}.
+If inertia is high a development tends to stay in a certain state.
+Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
+In a similar vein, high inertia in negative affect at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
+Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a negative mood for several measurements?''
 
@@
 For the variables in our illustration data set we chose an overall linear regression slope to capture the linear trend.
-The regression slope provides the average change from one time point $t$ to the next across all measurements $x$ of a time series $X$ of participant $i$ and variable $j$.
+The regression slope $b_{ij}$ provides the average change from one time point $t$ to the next across all measurements $x$ of a time series $X$ of participant $i$ and variable $j$.
+The specific form of the OLS slope formula we provide below, calculates $b_{ij}$ as the sum across all time points of the product of the deviation of time $t$ from its mean $\overline{t}$ and the deviation of $x_{ij}$ from its mean $\overline{x}_{ij}$ at each time point, divided by the sum across all time points of the square of the deviation of time from its mean ($\sum(t-\overline{t})^2$).
+Intuitively, the formula captures the rate of change of variable $x_{ij}$ with respect to time.
+This slope will indicate how the variable $x_{ij}$ changes over time, controlling for its mean value and the mean of time.
+If the slope is positive, $x_{ij}$ increases over time; if it's negative, $x_{ij}$ decreases over time.
 
@@
 But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
-Conceptually, researchers might have research question about the type of the development: "Is the development of anxiety a nonlinear process?" as well as the shape and structure of the development: "Does anxiety change as a sudden or smooth transition?"
+Conceptually, researchers might have research question about the type of the development: "Is the development of anxiety a nonlinear process?" as well as the shape and structure of the development: "How many spikes in anxiety did a patient experience?"
 We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model.
@@
 
-Beyond our main features of interest, we also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard.
+Beyond our main features of interest, we also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard (i.e., to exclude spurious explanations for the cluster assignments).
 After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant.
@@
 The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.
-The few missing values can, however, be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library.
+However, even the few missing values can be an issue for some feature reduction or feature clustering algorithms.
+We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library \citep[][]{buuren2011}.

### 2023-06-24 sections/04FeatureReduction.Rmd
feedback LFB

@@
 For our own illustration data we chose a feature projection method to reduce the dimensionality of our extracted features.
-In particular, we selected the commonly used \textit{principal component analysis} (PCA).
+We particularly chose the feature projection method for its broad applicability.
+We, specifically, selected the commonly used \textit{principal component analysis} (PCA).
 Some of the more tailor-made feature selection algorithms can be more accurate in reducing the feature dimensionality and might retain feature importance information more directly, depending on the specific data structure.
-However, PCAs have the distinct benefit that they are well-established within the psychometric literature and can broadly be applied to a wide variety of studies in an automatize manner.
-As our aim is to present a general illustration that can also be adopted for more general data descriptive uses, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
+However, PCAs have the distinct benefit that they are well-established within the psychometric literature \citep{jolliffe2011} and can broadly be applied to a wide variety of studies in an automatized manner \citep{abdi2010}.
+As our aim is to present a general illustration that can also be adopted across use-cases, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
 
@@
 We then enter all `r ncol(scaled_data)` features into the analysis.
-The PCA uses linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances).
-The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated.
+The PCA uses linear transformations in such a way that the first component captures the most possible variance of the original data \citep[e.g., by finding a vector that maximizes the sum of squared distances][]{jolliffe2002, abdi2010}.
+The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated \citep{shlens2014}.
 In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance.
 We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality.
-A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance (i.e., cumulative percentage explained variance).
+A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance \citep[i.e., cumulative percentage explained variance; e.g.,][]{jackson2003}.
 For our illustration we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
@@
 We would like to comment on two practical matters when using principal components --- the amount of dimensionality reduction and the interpretation of the principal components.
-As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables.
-Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance.
+As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables \citep[e.g.,][]{jolliffe2002}.
+Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance \citep[e.g.,][]{abdi2010}.
 Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
@@
 Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if all original features are considered psychologically meaningful in understanding the time series and users would like to retain the features for inpterpretation (irrespective of the features importance).
+We particularly advocate for such an approach if all original features are considered meaningful in understanding the time series and users would like to retain the features for inpterpretation (irrespective of the features importance).

### 2023-06-24 sections/06ClusterEvaluationPerformance.Rmd
:sparkles: feature interpretation integrated
feedback LFB

@@
 #| label: cluster perfomance indices
+
+library(fpc)
 
@@
 Three commonly reported indices we would like to highlight are the comparison of within clusters sum of squares, the average silhouette score, and the Calinski-Harabasz index.
-The first statistic we looked at was the a total within-cluster sum of square $WCSS$ (also see \equatref{eq:kWCSS}).
+The first statistic we looked at was the a total within-cluster sum of square $WCSS$ (see also \equatref{eq:kWCSS}).
 While the within cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$ after which the decrease was much smaller.
-This method is also called the \textit{elbow method} \citep{syakur2018}.
+This method is also sometimes referred to as the 'elbow method' \citep{syakur2018}.
 We then looked at a second, commonly used measure, the average silhouette score.

### 2023-06-24 sections/07ClusterEvaluationInterpretation.Rmd
feedback LFB

@@
 We first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
-We (1) see that for some variables the features are generally stronger in separating the clusters (e.g., 'how cooperative an interaction was' compared to 'attitudes towards the Dutch').
-Additionally, we see that (2) within the variables some features are better at distinguishing clusters (e.g., median of well-being vs.
-MAC of well-being).
+We see that for some variables the features are generally stronger in separating the clusters.
+We, for example see that the item on 'how cooperative the interaction was' distinguishes the two clusters across almost all seven features (except for the auto-correlation, see \fgrref[A]{fig:clusterFeatVar}).
+Compare this to the 'outgroup attitudes' item where the differences between the clusters are much smaller almost all features.
 We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables (e.g., mean and median compared to auto correlations).
+While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables.
+For example, mad and median distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
+These two features stand in stark contrast to other features, such as the lag-1 auto correlations, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}).
 This offers some information on which features were are most important in understanding the two extracted groups.
+Taken these two perspectives together, we can also focus on individual features or variables in particular.
+We, for example, see a strong difference in the average well-being, where participants in cluster showed a much lower median well-being over the time series.
+At the same time, in terms of stability, both groups have virtually identical averages MAC statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
+There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
 
@@
 In the second step, we look at prototypical trajectories of the clusters.
-For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs}][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
+For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
 Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}.
@@
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see \fgrref[B]{fig:clusterTs}).
-While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates were not too dissimilar.
+To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see $n$ in \fgrref[B]{fig:clusterFeatVar}).
+While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates are similar for all participants.
 To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-\fgrref[B]{fig:clusterFeatVar} illustrates that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination.
+When looking at the group comparison, find that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
 Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
+
+In short, we find that that the feature-based clustering discerned two meaningfully different groups of participants.
+We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attiudes (median) that are also stable over time (MAD, MAC) and tend to increase over the 30 day test period (linear trend).
+This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (median).
+This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 1).
+This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (median).
+This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (median).
+At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (linear trend), and there was generally less consistency in the experiences they were able to have (MAC, MAD, edf).
+
+This cluster separation, then, has a number of empirical and practical applications.
+Firstly, the clusters are descriptive.
+With tens of variables, hundreds of participants, and thousands of measurements, singular descriptive statistics are often not able to capture the complex patterns that describe the data set.
+The feature-based clustering offers some direct insight into the complexity within the data set.
+Secondly, the clusters identify important groups.
+The adaptive and deteriorating groups offer starting points for empirical exploration as well as practical interventions.
+Researchers can start probing what exactly distinguishes the two groups further and generate new bottom-up hypotheses.
+Practitioners, can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
+Thirdly, the feature-based approach is flexible and meaningful.
+We were able to use a wide range of time series features that have been central in the ESM literature and were able to use them directly to identify meaningful groups.
 

### 2023-06-24 scripts/functions/MlCorMat.R

@@ -124,7 +124,7 @@ MlCorMat <-
     
     rMlComb <- matrix("", nrow = length(selection), ncol = length(selection))
     rMlComb[upper.tri(rMlComb, diag = FALSE)] <- rMlWitBtw[upper.tri(rMlWitBtw, diag = FALSE)]
-    rMlComb[lower.tri(rMlComb, diag = FALSE)] <- rMlWitStar[upper.tri(rMlWitStar, diag = FALSE)]
+    rMlComb[lower.tri(rMlComb, diag = FALSE)] <- rMlWitStar[lower.tri(rMlWitStar, diag = FALSE)]
     rownames(rMlComb) <- labels
     colnames(rMlComb) <- paste(labels, "", sep="")
     
@@ -155,6 +155,7 @@ MlCorMat <-
           type = 2
         )
     )
+    
     
     for (i in 1:length(selection)) {
       dataRed <- 
@@ -212,7 +213,9 @@ MlTraitState <-
       #mutate_if(is.factor, as.numeric) %>% # let's hope this doesn't cause any problems :-D
       mutate_at(selection, as.numeric) %>% # let's hope this works 
       mutate_at(selection, list(gm = ~mean(., na.rm=TRUE))) %>% # grand mean
+      mutate_at(selection, list(gsd = ~sd(., na.rm=TRUE))) %>% # grand sd
       mutate_at(selection, list(gmc = ~.-mean(., na.rm=TRUE))) %>% # grand mean centered
+      mutate_at(selection, list(gmz = ~(.-mean(., na.rm=TRUE))/sd(., na.rm=TRUE))) %>% # grand mean standardized
       group_by_at(vars(matches(id))) %>%
       mutate_at(selection, list(cm = ~mean(., na.rm=TRUE))) %>% # cluster mean
       mutate_at(selection, list(csd = ~sd(., na.rm=TRUE))) %>% # cluster sd
@@ -243,43 +246,6 @@ MlTraitState <-
       mutate_at(paste0(selection, "_cm"), list(c = ~.-mean(., na.rm=TRUE))) # grand mean
     
     data
-  }
-
-PCAnorm <-
-  function(data,
-           pid,
-           tid,
-           selection) {
-    
-    # for testing
-    # data <- dtMain3mpca
-    # pid <- "PID"
-    # tid <- "TIDnum"
-    # selection <- varNam3mpcaMainAnalysis
-    
-    data02 <- data %>%
-      mutate_at(selection, as.numeric) %>%
[…]

### 2023-06-29 latex-results/07ClusterEvaluationInterpretation.tex
typose

@@
 see that for some variables the features are generally stronger in
-separating the clusters. We, for example see that the item on `how
-cooperative the interaction was' distinguishes the two clusters across
+separating the clusters. We, for example see that the item on `\textit{how
+cooperative the interaction was}' distinguishes the two clusters across
 almost all seven features (except for the auto-correlation, see
-\fgrref[A]{fig:clusterFeatVar}). Compare this to the `outgroup
-attitudes' item where the differences between the clusters are much
-smaller almost all features. We then inspect the clusters with a focus
+\fgrref[A]{fig:clusterFeatVar}). Compare this to the `\textit{outgroup
+attitudes}' item where the differences between the clusters are much
+smaller for almost all features. We then inspect the clusters with a focus
 on the features (see \fgrref[B]{fig:clusterFeatVar}). While this is the
@@
 features are better at distinguishing the clusters across variables. For
-example, mad and median distinguish the two clusters across almost all
+example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all
 variables (except for the item of whether the interaction was
 representative of the outgroup). These two features stand in stark
-contrast to other features, such as the lag-1 auto correlations, which
+contrast to other features, such as the \textit{lag-1 auto correlations}, which
 showed much smaller differences between the two clusters (see
 \fgrref[B]{fig:clusterFeatVar}). This offers some information on which
-features were are most important in understanding the two extracted
+features were most important in understanding the two extracted
 groups. Taken these two perspectives together, we can also focus on
@@
 strong difference in the average well-being, where participants in
-cluster showed a much lower median well-being over the time series. At
+cluster 2 showed a much lower median well-being over the time series. At
 the same time, in terms of stability, both groups have virtually
-identical averages MAC statistics for well-being (see
+identical averages \textit{MAC} statistics for well-being (see
 \fgrref[A]{fig:clusterFeatVar}). There are, thus, variables and features
@@
 more detail.
+In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
+We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a}.
 
@@
   \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+  Within the "(B) Feature Focus" subplot, the '\textit{n}' and '\textit{Discrimination}' comparison variables were not part of the original time series clustering.}
 \end{figure}
@@
 \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
 For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
@@
 but less accidental and involuntary interactions. The same cluster also
-reported an increase in need fulfilling interaction over the 30 day
+reported an increase in need-fulfilling interaction over the 30-day
 period and an increase in interactions that were representative of the
@@
 \fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, well-being are relatively stable
+representativeness, outgroup attitudes, and well-being are relatively stable
 for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability over time (also see
+higher variablity and instability on most of the other variables (also see
 \fgrref[A]{fig:clusterTs}).
@@
 number of measurements into the comparison and find that the
-participants in the detereoration cluster on average completed more ESM
+participants in the deterioration cluster on average completed more ESM
 surveys and reported on more intergroup interactions than the cluster

### 2023-06-29 manuscript.tex
typose

@@
 \author[*,1,3]{Jannis Kreienkamp}
+\author[1,3]{Maximilian Agostini}
+\author[1,3]{Kai Epstude}
+\author[2,3]{Rei Monden}
+\author[1,3]{Peter de Jonge}
 \author[1,3]{Laura F. Bringmann}
-\author[1,3]{Kai Epstude}
-\author[1,3]{Maximilian Agostini}
-\author[1,3]{Peter de Jonge}
-\author[2,3]{Rei Monden}
 \affiliation{\hfill}
@@
 In general, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.
-Feature projection methods, on the other hand, have been popular because of their widespread availability as well as their high efficiency.
+Feature projection methods, on the other hand, have been popular because they are efficient, widely available, and applicable to a wide range of data types.
 We provide an overview of the common approaches, an intuitive introduction to common methods, and exemplar algorithms in \tblref{tab:featureReduction}.
@@
 We provide a short intuitive explanation for common approaches, together with some of their characteristics and example algorithms in \tblref{tab:clusterApproaches}.
-For our own illustration, we have chosen that centroid-based k-means clustering.
-While k-means comes at the expense of high accuracy with more complex cluster shapes, we specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the data \citep{jain2010}.
+For our own illustration, we have chosen the centroid-based k-means clustering.
+While k-means comes at the expense of high accuracy with more complex cluster shapes, we specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the clusters \citep{jain2010}.
 K-means is also well-established within the research community and has been readily implemented in many statistical software packages \citep{hand2005}.
@@
 If true class labels are known, the cluster assignments can be compared to the true class labels --- using measures such as the F-measure, adjusted Rand index, mutual information, and normalized mutual information \citep[i.e., external evaluation; e.g.,][]{liao2005}.
-However, if the true cluster assignments are unknown, as with our psychological time series, the quality of the clusters is assessed based on the characteristics of the data itself \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
+However, if the true cluster assignments are unknown, as with our psychological time series, the quality of the clusters is assessed based on the characteristics of the data itself, such as separation and homogeneity of the clusters, or goodness of fit indices \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.
 
@@
 For our example of migration experiences, the method was useful to discern adaptive from more stressful experiences and helped to contextualize divergent experiences.
-We were able to show which variables and which dynamic features were most impactful in discerning the trajectory clusters and were even able to further contextualize the results with out-of-feature comparisons (see, e.g., \fgrref{fig:clusterFeatVar}).
+We found that some variables, such as interaction quality perceptions or need fulfillment were particularly important in distinguishing the groups (see \fgrref[A]{fig:clusterFeatVar}).
+Similarly, we found that the central tendency (median), variability (MAD), and linear trend (slope) were the most impactful dynamic features in discerning the trajectory clusters (see \fgrref[B]{fig:clusterFeatVar}).
+Jointly these two approaches allowed us to identify a cluster that had generally positive and improving experiences while the other cluster had more negative and deteriorating experiences.
+We were even able to further contextualize the results with out-of-feature comparisons, where we found that the group with the more difficult experiences also reported substantially more discrimination experiences during the post-test (see, e.g., \fgrref[B]{fig:clusterFeatVar}).
+In short, the feature-based approach allowed us to identify directly interpretable and meaningful groups, where we transparently know what data input the clusters are based on.
 
@@
 These additional steps of decision-making and transparency increase the initial barrier to entry.
-We hope that our empirical illustration offers a relatively generalizable procedure that showcases the ease of use, but clustering unfortunately does not offer a universal one-size-fits-all solution.
+We hope that our empirical illustration, the sample code, and the custom functions, offer a relatively generalizable procedure that showcases the ease of use, but clustering, unfortunately, does not offer a universal one-size-fits-all solution.
 
-In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential for incomparable results between methods \citep[e.g.,][]{bastiaansen2019}.
+In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential inconsistent results between methods \citep[e.g.,][]{bastiaansen2019}.
+A different set of variables, features, or a different clustering algorithm, might have resulted in substantially different cluster assignments.
 While the variety and diversity of methods are helpful in finding options even for more complex types of data, different algorithms often offer different results \citep[e.g.,][]{keogh2005}.
-And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their idiosyncratic shortcomings and might, for example, be sensitive to outliers \citep{xu2015}.
+And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their idiosyncratic shortcomings \citep{xu2015}.
 
@@
 
-One potential remedy to many of the limitations of feature-based clustering lies in transparently and reproducibly reporting the decisions by the user for each of the analysis steps.
-In our own description of the method, we have separated both the time series features as well as the analysis steps into distinct taxonomies in order to facilitate the transparency of the individual steps and decision moments.
+One potential remedy to many of the limitations of feature-based clustering lies in transparently and reproducibly reporting the decisions of the user for each of the analysis steps.
+In our own description of the method, we have provided a range of options and have motivated our own choices to facilitate the transparency of the individual steps and decision moments.
 Beyond the structures proposed here, \citet{vandeschoot2017} have proposed an extensive checklist for latent trajectory studies.
@@
 \subsection{Implications}
-Notwithstanding the limitations, we believe that feature-based clustering offers exciting new potential for psychological time series.
-Across a variety of scientific fields, feature-based time series clustering has been valued for its structuring and simplifying utilities.
-Within the growing field of psychological time series, ESM data is becoming increasingly more complex in terms of participants, variables, and time points.
-Feature-based time series clusters can help reduce these complexities to important and meaningful patterns, an increasingly important task for researchers and practitioners.
+Notwithstanding the limitations, we believe that feature-based clustering offers exciting new potential for researchers and practitioners assessing psychological time series.
 
@@
 
-For instance, given that the approach does not assume the stationarity restrictions of many model-based approaches, future research may want to integrate the research on capturing nonlinear trends from the shape-based approaches into the time series feature world.
-Bicoherence metrics, polynomial-, and differential equation parameters may be used to capture the type and structure of nonlinear developments \citep[e.g.,][]{caro-martin2018, mayor2022}.
-All of which would add to the under-studied trend features of ESM data.
+For instance, given that the approach does not assume the stationarity restrictions of many model-based approaches, future research can now more easily integrate many of the (non-)linear trend features.
+Research on capturing nonlinear trends has been growing over the past years and there are exciting possibilities to bring these developments to ESM data \citep{bringmann2023}.
+For example, bicoherence metrics, polynomial-, and differential equation parameters may be used to capture the type and structure of nonlinear developments \citep[e.g.
+shape-based approaches in \fgrref{fig:tsClustTax}; see also][]{caro-martin2018, mayor2022}.
+New features capturing nonlinear structures would add to the under-studied trend features of ESM data.
 
-Similarly, iterative or embedded approaches have a number of upsides, such as extracting a cluster-level model that captures the prototypical patterns of the group \citep[e.g.,][]{denteuling2021, lane2019}.
-To provide similar functionality, researchers may evaluate the utility of using recurrent neural networks, such as long short-term memory (LSTM) models to generate cluster-level trajectories \citep[e.g., by training the LSTM based on the feature matrix and the original time series as the target; e.g., see][]{fraley2002, nagin1999}.
+% Additionally, feature-based clustering algorithms traditionally do not provide a cluster-level prototype trajectory.
+Instead, users have to extract, often suboptimal, approximations of the group-level patterns (e.g., see the section on \textit{Cluster Evaluation}).
+Iterative or embedded clustering approaches, such as mixture models or group-based trajectory models, often estimate a group-based model and variance around that group-based model \citep[see \fgrref{fig:tsClustTax} and \tblref{tab:clusterApproaches}; e.g.,][]{denteuling2021, lane2019}.
+To provide similar functionality for feature-based models, researchers may evaluate the utility of using recurrent neural networks, such as long short-term memory (LSTM) models to generate cluster-level trajectories.
+One could, for example, train an LSTM based on the feature matrix and the original time series as the target \citep[e.g., see][]{fraley2002, nagin1999}.
 Alternatively, rule-based classifiers could extract rule sets that narrow down the conditions under which a given feature set is assigned to a specific cluster \citep[][]{benard2019}.
+This could potentially allow us to generate cluster-level trajectories, giving us an estimate of the typical trend for each group.
 

### 2023-06-29 tables/FeatureReduction.tex
typose

@@
         filter \linebreak & 
-        Features are selected individually or jointly based selection criteria.
-One common selection criteria is the amount of information and (unique) variance a feature captures.
-Univariate methodologies are able identify irrelevant features (i.e., because features do not capture much information), multivariate methods additionally allow to remove redundant features \citep[i.e., because features capture the same information][]{yu2004}.
+        Features are selected individually or jointly based on selection criteria.
+One common selection criterion is the amount of information and (unique) variance a feature captures.
+Univariate methodologies are able to identify irrelevant features (i.e., because features do not capture much information), multivariate methods additionally allow to remove redundant features \citep[i.e., because features capture the same information][]{yu2004}.
 \linebreak &
@@
 Because the selection process is essentially a search problem this method is computationally intensive.
-Traditionally wrappers have used forward selections or backwards eliminations but recently alternative approaches have been proposed based such as ant colony and swarm intelligence paradigms \citep[e.g., see][]{tang2014}.
+Traditionally wrappers have used forward selections or backward eliminations but recently alternative approaches have been proposed based such as ant colony and swarm intelligence paradigms \citep[e.g., see][]{tang2014}.
 \linebreak &
@@
         hybrid \linebreak & 
-        Hybrid selection methodological combine filter and wrapper methodologies to avoid the shortcomings of the individual methods.
-An initial filter step reduces computational effort (efficiency) and the wrapper ensures high performance \citep[effectiveness; e.g.,][]{alelyani2014}.
+        Hybrid selection methodologies combine filter and wrapper methodologies to avoid the shortcomings of the individual methods.
+An initial filter step selects candidate features (efficiency) and a subsequent wrapper ensures high performance \citep[effectiveness; e.g.,][]{alelyani2014}.
 \linebreak &

### 2023-06-29 latex-results/07ClusterEvaluationInterpretation.tex
feedback LFB

@@
 see that for some variables the features are generally stronger in
-separating the clusters. We, for example see that the item on `\textit{how
-cooperative the interaction was}' distinguishes the two clusters across
-almost all seven features (except for the auto-correlation, see
-\fgrref[A]{fig:clusterFeatVar}). Compare this to the `\textit{outgroup
-attitudes}' item where the differences between the clusters are much
-smaller for almost all features. We then inspect the clusters with a focus
-on the features (see \fgrref[B]{fig:clusterFeatVar}). While this is the
-same data as for the variable focus, we can see more clearly that some
-features are better at distinguishing the clusters across variables. For
-example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all
+separating the clusters. We, for example see that the item on
+`\textit{how cooperative the interaction was}' distinguishes the two
+clusters across almost all seven features (except for the
+\textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}). Compare
+this to the `\textit{outgroup attitudes}' item where the differences
+between the clusters are much smaller for almost all features. We then
+inspect the clusters with a focus on the features (see
+\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
+variable focus, we can see more clearly that some features are better at
+distinguishing the clusters across variables. For example, \textit{MAD}
+and \textit{median} distinguish the two clusters across almost all
 variables (except for the item of whether the interaction was
 representative of the outgroup). These two features stand in stark
-contrast to other features, such as the \textit{lag-1 auto correlations}, which
-showed much smaller differences between the two clusters (see
-\fgrref[B]{fig:clusterFeatVar}). This offers some information on which
-features were most important in understanding the two extracted
-groups. Taken these two perspectives together, we can also focus on
-individual features or variables in particular. We, for example, see a
-strong difference in the average well-being, where participants in
-cluster 2 showed a much lower median well-being over the time series. At
-the same time, in terms of stability, both groups have virtually
-identical averages \textit{MAC} statistics for well-being (see
-\fgrref[A]{fig:clusterFeatVar}). There are, thus, variables and features
-that distinguish the clusters better than others and a combination of
-variables and features lets us explore meaningful group differences in
-more detail.
-In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
-We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a}.
+contrast to other features, such as the
+\textit{lag-1 auto correlations}, which showed much smaller differences
+between the two clusters (see \fgrref[B]{fig:clusterFeatVar}). This
+offers some information on which features were most important in
+understanding the two extracted groups. Taken these two perspectives
+together, we can also focus on individual features or variables in
+particular. We, for example, see a strong difference in the average
+well-being, where participants in cluster 2 showed a much lower median
+well-being over the time series. At the same time, in terms of
+stability, both groups have virtually identical averages \textit{MAC}
+statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There
+are, thus, variables and features that distinguish the clusters better
+than others and a combination of variables and features lets us explore
+meaningful group differences in more detail. In our case, we see that
+the central tendency, variability, and linear trend are best at
+distinguishing a group with mainly positive experiences (cluster 1) from
+a group with a more negative experience (cluster 2). We also see that
+our clusters line up with the past literature on the importance of
+focusing on simpler and more meaningful statistics
+\citep{bringmann2018c, eronen2021a}.
 
@@
   \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the '\textit{n}' and '\textit{Discrimination}' comparison variables were not part of the original time series clustering.}
+  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
 \end{figure}
@@
 For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
-Immediately striking are the mean differences, where participants in the
-second cluster had more meaningful and fulfilling outgroup interactions
-also consistently reported more voluntary and cooperative interactions
-but less accidental and involuntary interactions. The same cluster also
-reported an increase in need-fulfilling interaction over the 30-day
-period and an increase in interactions that were representative of the
-outgroup. Whereas the other cluster showed a decrease in voluntary,
-cooperative, and positive interactions over the 30 days. This
-`deterioration' cluster also saw a decrease in general need fulfillment
-and experienced well-being over the 30 days (see
+Immediately striking are the mean differences, where participants in
+cluster 1 had more meaningful and fulfilling outgroup interactions also
+consistently reported more voluntary and cooperative interactions but
+less accidental and involuntary interactions. The same cluster (cluster
+1) also reported an increase in need-fulfilling interaction over the
+30-day period and an increase in interactions that were representative
+of the outgroup. Whereas the other cluster (cluster 2) showed a decrease
+in voluntary, cooperative, and positive interactions over the 30 days.
+This `deterioration' cluster (cluster 2) also saw a decrease in general
+need fulfillment and experienced well-being over the 30 days (see
 \fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, and well-being are relatively stable
-for both clusters, the deteriorating cluster also showed substantially
-higher variablity and instability on most of the other variables (also see
-\fgrref[A]{fig:clusterTs}).
+representativeness, outgroup attitudes, and well-being are relatively
+stable for both clusters, the deteriorating cluster (cluster 2) also
+showed substantially higher variablity and instability on most of the
+other variables (also see \fgrref[A]{fig:clusterTs}).
 
@@
 individual differences. To illustrate artifact checks, we added the
-number of measurements into the comparison and find that the
-participants in the deterioration cluster on average completed more ESM
-surveys and reported on more intergroup interactions than the cluster
-with the more positive interactions (see \(n\) in
-\fgrref[B]{fig:clusterFeatVar}). While this difference could indicate
-that the clusters might not entirely be comparable in the response
-patterns, we can find some relief in our data exclusion procedures
-during which we ensured that the general time frame and completion rates
-are similar for all participants. To illustrate the utility of
-individual differences, we compare the two samples in terms of the
-participants' self-reported discrimination experiences in the
-Netherlands (measured during the post-measurement). When looking at the
-group comparison, find that participants in the deteriorating cluster
-reported substantially higher levels of everyday discrimination
-(\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal and
-cross-sectional variables that were not included in the original
-clustering step can be used to explore and understand the cluster
-differences in more detail.
+number of ESM measurements into the comparison and find that the
+participants in the deterioration cluster (cluster 2) on average
+completed slightly more ESM surveys in general and reported on more
+intergroup interactions in particular (see \(n\) in
+\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures we
+ensured that the general time frame and completion rates are similar for
+all participants and indeed the numbers in ESM measurements generally
+are largely similar (e.g., see \(n\) for well-being and outgroup
+attitudes). However, the difference in reported number of interactions
+might indicate either a clustering artifact or a meaningful difference.
+The higher average number of interactions in cluster 2 could, for
+example, indicate a clustering artifact if variances are more inflated
+due to the larger samples \citep{kogan2006}. At the same time, however,
+the difference in the number of experienced interactions might also
+indicate a meaningful difference, where the deteriorating cluster
+(cluster 2) on average reported more outgroup interactions, but these
+interactions were less voluntary, less meaningful, and less positive.
+Such a finding is in line with past research highlighting the role of
+negative intergroup interations in explaining intergroup relations
+\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. To further illustrate
+the utility of assessing out of feature individual differences, we also
+compare the two samples in terms of the participants' self-reported
+discrimination experiences in the Netherlands (measured during the
+post-measurement). When looking at the group comparison, we find that
+participants in the deteriorating cluster (cluster 2) reported
+substantially higher levels of everyday discrimination
+(\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
+(e.g., sum of specific ESM measurements) and cross-sectional variables
+(e.g., general discrimination differences) that were not included in the
+original clustering step can be used to explore and understand the
+cluster differences in more detail.
 
@@
 (cluster 2) that reports higher well-being and more positive outgroup
-attiudes (median) that are also stable over time (MAD, MAC) and tend to
-increase over the 30 day test period (linear trend). This group also
-reported consistently more meaningful, need-fulfilling, and cooperative
-outgroup interactions (median). This group with overwhelmingly positive
-experiences stands in contrast with a more detrimental group (cluster
-1). This cluster, on average, reported much less positive, less
-meaningful, and less fulfilling interactions and interaction patterns
-(median). This group also reported less positive outgroup attitudes,
-lower well-being, and more discrimination experiences (median). At the
-same time, for members of this detrimental cluster conditions seemed to
-deteriorate over time (linear trend), and there was generally less
-consistency in the experiences they were able to have (MAC, MAD, edf).
+attitudes (\textit{median}) that are also stable over time
+(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
+period (\textit{linear trend}). This group also reported consistently
+more meaningful, need-fulfilling, and cooperative outgroup interactions
+(\textit{median}). This group with overwhelmingly positive experiences
+stands in contrast with a more detrimental group (cluster 1). This
+cluster, on average, reported much less positive, less meaningful, and
+less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower
+well-being, and more discrimination experiences (\textit{median}). At
+the same time, for members of this detrimental cluster conditions seemed
+to deteriorate over time (\textit{linear trend}), and there was
+generally less consistency in the experiences they were able to have
+(\textit{MAC}, \textit{MAD}, \textit{edf}).
 
@@
 clustering offers some direct insight into the complexity within the
-data set. Secondly, the clusters identify important groups. The adaptive
-and deteriorating groups offer starting points for empirical exploration
-as well as practical interventions. Researchers can start probing what
-exactly distinguishes the two groups further and generate new bottom-up
-hypotheses. Practitioners, can use the group separation to identify
-individuals in need of assistance and can explore contextual factors
-that might contribute to the difficulties some might face. Thirdly, the
+data set. In our empirical example, we for example see that participants
+are meaningfully distinguished by a combination of high (vs.~low)
+central tendency, variability, and linear trend. Secondly, the clusters
+identify important groups. The adaptive and deteriorating groups offer
+starting points for empirical exploration as well as practical
+interventions. Researchers can start probing what exactly distinguishes
+the two groups further and generate new bottom-up hypotheses.
+Practitioners, can use the group separation to identify individuals in
+need of assistance and can explore contextual factors that might
+contribute to the difficulties some might face. Thirdly, the
 feature-based approach is flexible and meaningful. We were able to use a
@@
 literature and were able to use them directly to identify meaningful
-groups.
+groups. For our empirical illustration we, among others for example,
+chose to focus on whether participants differed in their their average
+well-being (i.e., \textit{median}), how much their well-being would vary
+over time (i.e, \textit{MAD}), and whether their well-being would on
+average increase or decrease over time (i.e., \textit{linear trend}).
+Alternatively, for others cyclical patterns might be more important ---
+for example, whether well-being was higher on weekends. Importantly, in
+any case, we did not need to translate these dynamic features into
+probabilistic inference models (e.g., VAR models) to cluster the
+participants.
 

### 2023-06-29 sections/07ClusterEvaluationInterpretation.Rmd
feedback LFB

@@
 We see that for some variables the features are generally stronger in separating the clusters.
-We, for example see that the item on 'how cooperative the interaction was' distinguishes the two clusters across almost all seven features (except for the auto-correlation, see \fgrref[A]{fig:clusterFeatVar}).
-Compare this to the 'outgroup attitudes' item where the differences between the clusters are much smaller almost all features.
+We, for example see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
+Compare this to the '\textit{outgroup attitudes}' item where the differences between the clusters are much smaller for almost all features.
 We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
 While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables.
-For example, mad and median distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
-These two features stand in stark contrast to other features, such as the lag-1 auto correlations, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}).
-This offers some information on which features were are most important in understanding the two extracted groups.
+For example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
+These two features stand in stark contrast to other features, such as the \textit{lag-1 auto correlations}, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}).
+This offers some information on which features were most important in understanding the two extracted groups.
 Taken these two perspectives together, we can also focus on individual features or variables in particular.
-We, for example, see a strong difference in the average well-being, where participants in cluster showed a much lower median well-being over the time series.
-At the same time, in terms of stability, both groups have virtually identical averages MAC statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
+We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series.
+At the same time, in terms of stability, both groups have virtually identical averages \textit{MAC} statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
 There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
+In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
+We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a}.
 
@@
 For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or S-GIMME \citep[e.g.][]{lane2019}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
 For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
-Immediately striking are the mean differences, where participants in the second cluster had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
-The same cluster also reported an increase in need fulfilling interaction over the 30 day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This 'deterioration' cluster also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, well-being are relatively stable for both clusters, the deteriorating cluster also showed substantially higher variablity and instability over time (also see \fgrref[A]{fig:clusterTs}).
+Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
+The same cluster (cluster 1) also reported an increase in need-fulfilling interaction over the 30-day period and an increase in interactions that were representative of the outgroup.
+Whereas the other cluster (cluster 2) showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
+This `deterioration' cluster (cluster 2) also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
+We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variablity and instability on most of the other variables (also see \fgrref[A]{fig:clusterTs}).
 
@@
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of measurements into the comparison and find that the participants in the detereoration cluster on average completed more ESM surveys and reported on more intergroup interactions than the cluster with the more positive interactions (see $n$ in \fgrref[B]{fig:clusterFeatVar}).
-While this difference could indicate that the clusters might not entirely be comparable in the response patterns, we can find some relief in our data exclusion procedures during which we ensured that the general time frame and completion rates are similar for all participants.
-To illustrate the utility of individual differences, we compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-When looking at the group comparison, find that participants in the deteriorating cluster reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
-Thus, both intensive longitudinal and cross-sectional variables that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
+To illustrate artifact checks, we added the number of ESM measurements into the comparison and find that the participants in the deterioration cluster (cluster 2) on average completed slightly more ESM surveys in general and reported on more intergroup interactions in particular (see $n$ in \fgrref[B]{fig:clusterFeatVar}).
+In our data exclusion procedures we ensured that the general time frame and completion rates are similar for all participants and indeed the numbers in ESM measurements generally are largely similar (e.g., see $n$ for well-being and outgroup attitudes).
+However, the difference in reported number of interactions might indicate either a clustering artifact or a meaningful difference.
+The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are more inflated due to the larger samples \citep{kogan2006}.
+At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions, but these interactions were less voluntary, less meaningful, and less positive.
+Such a finding is in line with past research highlighting the role of negative intergroup interations in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+To further illustrate the utility of assessing out of feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
+When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
+Thus, both intensive longitudinal (e.g., sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
 
 In short, we find that that the feature-based clustering discerned two meaningfully different groups of participants.
-We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attiudes (median) that are also stable over time (MAD, MAC) and tend to increase over the 30 day test period (linear trend).
-This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (median).
+We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
+This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
 This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 1).
-This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (median).
-This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (median).
-At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (linear trend), and there was generally less consistency in the experiences they were able to have (MAC, MAD, edf).
+This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
+At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
 
@@
 The feature-based clustering offers some direct insight into the complexity within the data set.
+In our empirical example, we for example see that participants are meaningfully distinguished by a combination of high (vs.
+low) central tendency, variability, and linear trend.
 Secondly, the clusters identify important groups.
@@
 We were able to use a wide range of time series features that have been central in the ESM literature and were able to use them directly to identify meaningful groups.
+For our empirical illustration we, among others for example, chose to focus on whether participants differed in their their average well-being (i.e., \textit{median}), how much their well-being would vary over time (i.e, \textit{MAD}), and whether their well-being would on average increase or decrease over time (i.e., \textit{linear trend}).
+Alternatively, for others cyclical patterns might be more important --- for example, whether well-being was higher on weekends.
+Importantly, in any case, we did not need to translate these dynamic features into probabilistic inference models (e.g., VAR models) to cluster the participants.
 

### 2023-07-22 manuscript.tex
feedback Rei

@@
 This type of research comes as a response to prominent reviews within the migration literature, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-Importantly for our illustration, the migration ESM research, thus, exemplifies some of the real-world data issues that ESM data commonly faces, including a multivariate conceptualization with event-specific missingness patterns.
+Importantly for our illustration, the migration ESM research, thus, exemplifies some of the real-world data issues that ESM data commonly face, including a multivariate conceptualization with event-specific missingness patterns.
 
@@
 While this is a common conceptual separation of procedural elements, it is important to note that these steps are a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
-Nonetheless, the conceptual nature of these steps allows us to introduce the mayor elements of the analysis.
+Nonetheless, the conceptual nature of these steps allows us to introduce the major elements of the analysis.
 We also provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.
@@
 We found that some variables, such as interaction quality perceptions or need fulfillment were particularly important in distinguishing the groups (see \fgrref[A]{fig:clusterFeatVar}).
-Similarly, we found that the central tendency (median), variability (MAD), and linear trend (slope) were the most impactful dynamic features in discerning the trajectory clusters (see \fgrref[B]{fig:clusterFeatVar}).
+Similarly, we found that the central tendency (\textit{median}), variability (\textit{MAD}), and linear trend (\textit{slope}) were the most impactful dynamic features in discerning the trajectory clusters (see \fgrref[B]{fig:clusterFeatVar}).
 Jointly these two approaches allowed us to identify a cluster that had generally positive and improving experiences while the other cluster had more negative and deteriorating experiences.
@@
 Similar challenges arise with the choice of the clustering algorithm or the cluster illustration.
-These challenges are not unique to feature-based clustering, rather they are common to all clustering approaches \citep{liao2005}.
+These challenges are not unique to feature-based clustering, rather they are common to all clustering approaches \citep{liao2005, horne2020}.
 However, it is important to remember that multi-stepped data-driven approaches are particularly vulnerable to the impact of the researchers' degrees of freedom.
@@
 shape-based approaches in \fgrref{fig:tsClustTax}; see also][]{caro-martin2018, mayor2022}.
-New features capturing nonlinear structures would add to the under-studied trend features of ESM data.
+New features capturing nonlinear structures would add to the under-studied (non-)linear trend features of ESM data.
 
@@
 Beyond the direct academic use, the feature-based time series clustering approach also addresses practical and applied uses.
-For practitioners, the approach offers a practical and grounded method for dealing with the challenges of complex and messy patient-, customer-, or user data.
+For practitioners, the approach offers a practical and grounded method for dealing with the challenges of complex and messy data from multiple patients, customers, or users.
 Not only does the approach directly deal with dimensionality, missingness, and time scales in the time series, but the interpretability and transparency aspects offer particular utility in applied settings, where the costs of misspecification are high.

### 2023-07-22 sections/01DataDescription.Rmd
feedback Rei

@@
 After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.
-The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm).
-After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
+The short ESM surveys were sent out at around lunch (12pm) and dinner time (7 pm).
+After the 30-day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire.
 Participants received either monetary compensation or partial course credits based on the number of surveys they completed.
@@
 The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` ESM measurements.
-Each of the studies focused on recently arrived first-generation migrants and each study included a number of ideosyncratic variables relevant for the broader research collective.
-For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies.
-Variable selection and preparation is described as part of the illustration below but for additional methodological details about the study setup see \citet[][]{Kreienkamp2022b}.
+Each of the studies focused on recently arrived first-generation migrants and each study included a number of idiosyncratic variables relevant for the broader research collective.
+For our empirical example, we focus on the variables that were collected during the ESM surveys and were available in all three studies.
+Variable selection and preparation are described as part of the illustration below but for additional methodological details about the study setup see \citet[][]{Kreienkamp2022b}.

### 2023-07-22 sections/02InputVariables.Rmd
feedback Rei

@@
 Once the important variables have been selected, the data needs to be prepared for the analysis steps.
-Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time series comparable.
-Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses \citep{liao2005}.
+Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time-series comparable.
+Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses \citep['data exclusion' and 'data transformation' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.
 
 In our illustration data set, the studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the later two studies.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant repsonse-patterns and the time frame were now substantially more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
+This was likely due to the option to continue participation without compensation in the latter two studies.
+To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure led to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
+Importantly, both the participant response-patterns and the time frame were now substantially more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
 Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.

### 2023-07-22 sections/03FeatureExtraction.Rmd
feedback Rei

@@
 As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series.
-Researchers might, for example, be interested in whether "Over a one month period, are some people happier than others?"
+Researchers might, for example, be interested in whether "Over a one-month period, are some people happier than others?"
 
@@
 If inertia is high a development tends to stay in a certain state.
-Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of underreactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
+Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of under-reactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
 In a similar vein, high inertia in negative affect at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
@@
 \paragraph{Linear trend.}
-In non-stationary time series a linear trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
+In non-stationary time series, a linear trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}.
 For psychological time series, researchers have, for example, pointed out the importance of linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}.
@@
 
-For the variables in our illustration data set we chose an overall linear regression slope to capture the linear trend.
+For the variables in our illustration data set, we chose an overall linear regression slope to capture the linear trend.
 The regression slope $b_{ij}$ provides the average change from one time point $t$ to the next across all measurements $x$ of a time series $X$ of participant $i$ and variable $j$.
-The specific form of the OLS slope formula we provide below, calculates $b_{ij}$ as the sum across all time points of the product of the deviation of time $t$ from its mean $\overline{t}$ and the deviation of $x_{ij}$ from its mean $\overline{x}_{ij}$ at each time point, divided by the sum across all time points of the square of the deviation of time from its mean ($\sum(t-\overline{t})^2$).
+The specific form of the OLS slope formula we provide below calculates $b_{ij}$ as the sum across all time points of the product of the deviation of time $t$ from its mean $\overline{t}$ and the deviation of $x_{ij}$ from its mean $\overline{x}_{ij}$ at each time point, divided by the sum across all time points of the square of the deviation of time from its mean ($\sum(t-\overline{t})^2$).
 Intuitively, the formula captures the rate of change of variable $x_{ij}$ with respect to time.
@@
 \paragraph{Nonlinearity.}
-Changes in psychology are not always linear, instead nonlinearity is a common feature of psychological time series \citep{hayes2007}.
+Changes in psychology are not always linear, instead, nonlinearity is a common feature of psychological time series \citep{hayes2007}.
 As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
@@
 But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
-Conceptually, researchers might have research question about the type of the development: "Is the development of anxiety a nonlinear process?" as well as the shape and structure of the development: "How many spikes in anxiety did a patient experience?"
+Conceptually, researchers might have research questions about the type of the development: "Is the development of anxiety a nonlinear process?" as well as the shape and structure of the development: "How many spikes in anxiety did a patient experience?"
 We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model.
@@
 We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library \citep[][]{buuren2011}.
+Note again that with this procedure we only need to impute an extremely small number of missing values as most feature calculations can use the available data instead.

### 2023-07-22 sections/04FeatureReduction.Rmd
feedback Rei

@@
 
-For our own illustration data we chose a feature projection method to reduce the dimensionality of our extracted features.
+For our own illustration data, we chose a feature projection method to reduce the dimensionality of our extracted features.
 We particularly chose the feature projection method for its broad applicability.
@@
 However, PCAs have the distinct benefit that they are well-established within the psychometric literature \citep{jolliffe2011} and can broadly be applied to a wide variety of studies in an automatized manner \citep{abdi2010}.
-As our aim is to present a general illustration that can also be adopted across use-cases, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
+As our aim is to present a general illustration that can also be adopted across use cases, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
 
@@
 A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance \citep[i.e., cumulative percentage explained variance; e.g.,][]{jackson2003}.
-For our illustration we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
-For the extracted principal components we save the `r pca_cutoff` PC-scores for each participant (i.e., the participants' coordinates in the reduced dimensional space).
+For our illustration, we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).
+For the extracted principal components we save the `r pca_cutoff` principal component scores for each participant (i.e., the participants' coordinates in the reduced dimensional space; PC-scores).
 
@@
 As for the expected dimensionality reduction, given its methodology, PCAs tend to 'work better' at reducing dimensions with (highly) correlated variables \citep[e.g.,][]{jolliffe2002}.
-Thus, with a set of very homogeneous variables and features users will need less principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance \citep[e.g.,][]{abdi2010}.
+Thus, with a set of very homogeneous variables and features users will need fewer principal components to explain a large amount of variance, while a more diverse set of variables and features will tend to require more principal components to capture the same amount of variance \citep[e.g.,][]{abdi2010}.
 Our `r pca_cutoff` principal components are still a relatively high number of variables but this is not surprising as we chose a diverse conceptualization and a diverse set of time series features.
-As for the interpretability, PCA allows users to extract information on the meaning of the principal components.
-In particular, because the principal components are linear combination of original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
+As for interpretability, PCA allows users to extract information on the meaning of the principal components.
+In particular, because the principal components are linear combinations of the original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
 While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
 Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.
-We particularly advocate for such an approach if all original features are considered meaningful in understanding the time series and users would like to retain the features for inpterpretation (irrespective of the features importance).
+We particularly advocate for such an approach if all original features are considered meaningful in understanding the time series and users would like to retain the features for interpretation (irrespective of the features' importance).

### 2023-07-22 sections/05FeatureClustering.Rmd
:memo: feat red adj
typos
feedback Rei

@@
 
-In our case we entered the participants' PC-scores from the feature reduction step into the k-means algorithm.
-Because we did know the underlying number of clusters within our sample, we calculated the cluster solutions for $k =$ 2 through 10.
+In our case, we entered the participants' PC-scores from the feature reduction step into the k-means algorithm.
+Because we did not know the underlying number of clusters within our sample, we calculated the cluster solutions for $k=\{2, \dots , 10\}$ .
 To avoid local minima we used 100 random initial centroid positions for each run.
 Each of the 9 cluster solutions converged within the iteration limit.
-In the next step we will then evaluate which of the extracted cluster solutions offers the best fit with the data.
+In the next step, we will then evaluate which of the extracted cluster solutions offers the best fit with the data.

### 2023-07-22 sections/06ClusterEvaluationPerformance.Rmd
feedback Rei

@@
 Three commonly reported indices we would like to highlight are the comparison of within clusters sum of squares, the average silhouette score, and the Calinski-Harabasz index.
-The first statistic we looked at was the a total within-cluster sum of square $WCSS$ (see also \equatref{eq:kWCSS}).
-While the within cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$ after which the decrease was much smaller.
+The first statistic we looked at was the total within-cluster sum of square $WCSS$ (see also \equatref{eq:kWCSS}).
+While the within-cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$ after which the decrease was much smaller.
 This method is also sometimes referred to as the 'elbow method' \citep{syakur2018}.
@@
 In our case, the $k=2$ solution maximized the silhouette coefficient ($s_2=$ `r format(round(kmeansPerfComp$avg.silwidth[kmeansPerfComp$cluster.number==2], 2), nsmall = 2)`).
-Finally, the Calinski-Harabasz index assesses the compactness and separation of the clusters by asssessing the ratio of the sum of between-clusters dispersion and of inter-cluster dispersion for all clusters --- thus, the higher the score the better the performances \citep{calinski1974}.
-In our case the $k=2$ solution also showed the highest Calinski-Harabasz index ($CH_2=$ `r format(round(kmeansPerfComp$ch[kmeansPerfComp$cluster.number==2], 2), nsmall = 2)`; a full table of all extracted validity statistics is available in Supplemental Material A)\footnote{It is important to note that another commonly assessed aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}.
+Finally, the Calinski-Harabasz index assesses the compactness and separation of the clusters by assessing the ratio of the sum of between-clusters dispersion and of inter-cluster dispersion for all clusters --- thus, the higher the score the better the performances \citep{calinski1974}.
+In our case, the $k=2$ solution also showed the highest Calinski-Harabasz index ($CH_2=$ `r format(round(kmeansPerfComp$ch[kmeansPerfComp$cluster.number==2], 2), nsmall = 2)`; a full table of all extracted validity statistics is available in Supplemental Material A)\footnote{It is important to note that another commonly assessed aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}.
 This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}.

### 2023-07-22 sections/07ClusterEvaluationInterpretation.Rmd
feedback Rei

@@
 We see that for some variables the features are generally stronger in separating the clusters.
-We, for example see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
+We, for example, see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
 Compare this to the '\textit{outgroup attitudes}' item where the differences between the clusters are much smaller for almost all features.
@@
 This offers some information on which features were most important in understanding the two extracted groups.
-Taken these two perspectives together, we can also focus on individual features or variables in particular.
+Taking these two perspectives together, we can also focus on individual features or variables in particular.
 We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series.
@@
   \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+  "Int." = interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
 \end{figure}
@@
 
-In the second step, we look at prototypical trajectories of the clusters.
+In the second step, we look at the prototypical trajectories of the clusters.
 For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
@@
 For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
-Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions also consistently reported more voluntary and cooperative interactions but less accidental and involuntary interactions.
-The same cluster (cluster 1) also reported an increase in need-fulfilling interaction over the 30-day period and an increase in interactions that were representative of the outgroup.
+Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions and also consistently reported more voluntary and cooperative interactions but fewer accidental and involuntary interactions.
+The same cluster (cluster 1) also reported an increase in need-fulfilling interactions over the 30-day period and an increase in interactions that were representative of the outgroup.
 Whereas the other cluster (cluster 2) showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
 This `deterioration' cluster (cluster 2) also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variablity and instability on most of the other variables (also see \fgrref[A]{fig:clusterTs}).
+We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variability and instability on most of the other variables (also see \fgrref[A]{fig:clusterTs}).
 
-<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be invesitigated with targeted studies.}} -->
+<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be investigated with targeted studies.}} -->
 
-Finally, we can also assess the clusters across other person-level variable \citep[e.g.,][]{monden2022}.
+Finally, we can also assess the clusters across other person-level variables \citep[e.g.,][]{monden2022}.
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
 To illustrate artifact checks, we added the number of ESM measurements into the comparison and find that the participants in the deterioration cluster (cluster 2) on average completed slightly more ESM surveys in general and reported on more intergroup interactions in particular (see $n$ in \fgrref[B]{fig:clusterFeatVar}).
-In our data exclusion procedures we ensured that the general time frame and completion rates are similar for all participants and indeed the numbers in ESM measurements generally are largely similar (e.g., see $n$ for well-being and outgroup attitudes).
-However, the difference in reported number of interactions might indicate either a clustering artifact or a meaningful difference.
+In our data exclusion procedures, we ensured that the general time frame and completion rates are similar for all participants and indeed the numbers in ESM measurements generally are largely similar (e.g., see $n$ for well-being and outgroup attitudes).
+However, the difference in the reported number of interactions might indicate either a clustering artifact or a meaningful difference.
 The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are more inflated due to the larger samples \citep{kogan2006}.
 At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions, but these interactions were less voluntary, less meaningful, and less positive.
-Such a finding is in line with past research highlighting the role of negative intergroup interations in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
-To further illustrate the utility of assessing out of feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
+Such a finding is in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+To further illustrate the utility of assessing out-of-feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
 When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
-Thus, both intensive longitudinal (e.g., sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
+Thus, both intensive longitudinal (e.g., the sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
 
-In short, we find that that the feature-based clustering discerned two meaningfully different groups of participants.
+In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
 We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
@@
 The feature-based clustering offers some direct insight into the complexity within the data set.
-In our empirical example, we for example see that participants are meaningfully distinguished by a combination of high (vs.
+In our empirical example, we, for example, see that participants are meaningfully distinguished by a combination of high (vs.
 low) central tendency, variability, and linear trend.
@@
 Researchers can start probing what exactly distinguishes the two groups further and generate new bottom-up hypotheses.
-Practitioners, can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
+Practitioners can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
 Thirdly, the feature-based approach is flexible and meaningful.
 We were able to use a wide range of time series features that have been central in the ESM literature and were able to use them directly to identify meaningful groups.
-For our empirical illustration we, among others for example, chose to focus on whether participants differed in their their average well-being (i.e., \textit{median}), how much their well-being would vary over time (i.e, \textit{MAD}), and whether their well-being would on average increase or decrease over time (i.e., \textit{linear trend}).
+For our empirical illustration we, among others for example, chose to focus on whether participants differed in their average well-being (i.e., \textit{median}), how much their well-being would vary over time (i.e, \textit{MAD}), and whether their well-being would on average increase or decrease over time (i.e., \textit{linear trend}).
 Alternatively, for others cyclical patterns might be more important --- for example, whether well-being was higher on weekends.

### 2023-07-22 latex-results/01DataDescription.tex
feedback rm and pdj

@@
 members for at least 30 days. The short ESM surveys were sent out at
-around lunch (12pm) and dinner time (7pm). After the 30 day study period
-participants filled in a post-questionnaire that mirrored the
+around lunch (12pm) and dinner time (7 pm). After the 30-day study
+period participants filled in a post-questionnaire that mirrored the
 pre-questionnaire. Participants received either monetary compensation or
@@
 first-generation migrants and each study included a number of
-ideosyncratic variables relevant for the broader research collective.
-For our empirical example we focus on the variables that were collected
+idiosyncratic variables relevant for the broader research collective.
+For our empirical example, we focus on the variables that were collected
 during the ESM surveys and were available in all three studies. Variable
-selection and preparation is described as part of the illustration below
-but for additional methodological details about the study setup see
-\citet[][]{Kreienkamp2022b}.
+selection and preparation are described as part of the illustration
+below but for additional methodological details about the study setup
+see \citet[][]{Kreienkamp2022b}.

### 2023-07-22 latex-results/02InputVariables.tex
feedback rm and pdj

@@
 validating and cleaning the data (e.g., re-coding, removing duplicate or
-unwanted measurements) but also making the time series comparable. Two
+unwanted measurements) but also making the time-series comparable. Two
 important steps are making the time-frames and response scales
@@
 that is common to most participants and standardizing the participants'
-responses \citep{liao2005}.
+responses
+\citep['data exclusion' and 'data transformation' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.
 
@@
 likely due to the option to continue participation without compensation
-in the later two studies. To make the three studies comparable in
+in the latter two studies. To make the three studies comparable in
 participation and time frames, we iteratively removed all measurement
 occasions and participants that had more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
+\citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure led to a final sample of 157 participants, who jointly
 produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now substantially more
+response-patterns and the time frame were now substantially more
 comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,

### 2023-07-22 latex-results/03FeatureExtraction.tex
feedback rm and pdj

@@
 about psychological time series. Researchers might, for example, be
-interested in whether ``Over a one month period, are some people happier
+interested in whether ``Over a one-month period, are some people happier
 than others?''
@@
 emotion dynamics high inertia of negative affect has been found to be
-indicative of underreactive systems and to be characteristic of
+indicative of under-reactive systems and to be characteristic of
 psychological maladjustment \citep{kuppens2010}. In a similar vein, high
@@
 
-In non-stationary time series a linear trend can be observed when there
+In non-stationary time series, a linear trend can be observed when there
 is a consistent increase or decrease in the data over time
@@
 
-For the variables in our illustration data set we chose an overall
+For the variables in our illustration data set, we chose an overall
 linear regression slope to capture the linear trend. The regression
@@
 participant \(i\) and variable \(j\). The specific form of the OLS slope
-formula we provide below, calculates \(b_{ij}\) as the sum across all
+formula we provide below calculates \(b_{ij}\) as the sum across all
 time points of the product of the deviation of time \(t\) from its mean
@@
 
-Changes in psychology are not always linear, instead nonlinearity is a
+Changes in psychology are not always linear, instead, nonlinearity is a
 common feature of psychological time series \citep{hayes2007}. As an
@@
 \citep{vandermaas2003}. Conceptually, researchers might have research
-question about the type of the development: ``Is the development of
+questions about the type of the development: ``Is the development of
 anxiety a nonlinear process?'' as well as the shape and structure of the
@@
 mean matching imputation using the MICE library \citep[][]{buuren2011}.
+Note again that with this procedure we only need to impute an extremely
+small number of missing values as most feature calculations can use the
+available data instead.

### 2023-07-22 latex-results/04FeatureReduction.tex
feedback rm and pdj

@@
-For our own illustration data we chose a feature projection method to
+For our own illustration data, we chose a feature projection method to
 reduce the dimensionality of our extracted features. We particularly
@@
 \citep{abdi2010}. As our aim is to present a general illustration that
-can also be adopted across use-cases, we present the workflow using a
+can also be adopted across use cases, we present the workflow using a
 PCA here but we encourage users to consider more specialized methods as
@@
 \citep[i.e., cumulative percentage explained variance; e.g.,][]{jackson2003}.
-For our illustration we select the first 27 principal components that
+For our illustration, we select the first 27 principal components that
 explain 80\% of the variance in the original 72 features (reducing the
 dimensionality by 62.50\%). For the extracted principal components we
-save the 27 PC-scores for each participant (i.e., the participants'
-coordinates in the reduced dimensional space).
+save the 27 principal component scores for each participant (i.e., the
+participants' coordinates in the reduced dimensional space; PC-scores).
 
@@
 \citep[e.g.,][]{jolliffe2002}. Thus, with a set of very homogeneous
-variables and features users will need less principal components to
+variables and features users will need fewer principal components to
 explain a large amount of variance, while a more diverse set of
@@
 this is not surprising as we chose a diverse conceptualization and a
-diverse set of time series features. As for the interpretability, PCA
-allows users to extract information on the meaning of the principal
-components. In particular, because the principal components are linear
-combination of original features, users can extract the relative
-importance of each feature for the extracted principal components (i.e.,
-the eigenvectors). While this can be useful in understanding the
-variance in the original data or help with manual feature selection, we
-use the PCA purely to reduce the dimensionality for the clustering step.
-Instead of relying on the principal components, we use the original
-features of interest to interpret the later extracted clusters. We
-particularly advocate for such an approach if all original features are
-considered meaningful in understanding the time series and users would
-like to retain the features for inpterpretation (irrespective of the
-features importance).
+diverse set of time series features. As for interpretability, PCA allows
+users to extract information on the meaning of the principal components.
+In particular, because the principal components are linear combinations
+of the original features, users can extract the relative importance of
+each feature for the extracted principal components (i.e., the
+eigenvectors). While this can be useful in understanding the variance in
+the original data or help with manual feature selection, we use the PCA
+purely to reduce the dimensionality for the clustering step. Instead of
+relying on the principal components, we use the original features of
+interest to interpret the later extracted clusters. We particularly
+advocate for such an approach if all original features are considered
+meaningful in understanding the time series and users would like to
+retain the features for interpretation (irrespective of the features'
+importance).

### 2023-07-22 latex-results/05FeatureClustering.tex
:memo: feat red adj
typos
feedback rm and pdj

@@
 
-In our case we entered the participants' PC-scores from the feature
-reduction step into the k-means algorithm. Because we did know the
+In our case, we entered the participants' PC-scores from the feature
+reduction step into the k-means algorithm. Because we did not know the
 underlying number of clusters within our sample, we calculated the
-cluster solutions for \(k =\) 2 through 10. To avoid local minima we
+cluster solutions for \(k=\{2, \dots , 10\}\) . To avoid local minima we
 used 100 random initial centroid positions for each run. Each of the 9
-cluster solutions converged within the iteration limit. In the next step
-we will then evaluate which of the extracted cluster solutions offers
-the best fit with the data.
+cluster solutions converged within the iteration limit. In the next
+step, we will then evaluate which of the extracted cluster solutions
+offers the best fit with the data.

### 2023-07-22 latex-results/06ClusterEvaluationPerformance.tex
feedback rm and pdj

@@
 squares, the average silhouette score, and the Calinski-Harabasz index.
-The first statistic we looked at was the a total within-cluster sum of
-square \(WCSS\) (see also \equatref{eq:kWCSS}). While the within cluster
+The first statistic we looked at was the total within-cluster sum of
+square \(WCSS\) (see also \equatref{eq:kWCSS}). While the within-cluster
 variation will naturally decrease with (more) smaller clusters, we
@@
 Calinski-Harabasz index assesses the compactness and separation of the
-clusters by asssessing the ratio of the sum of between-clusters
+clusters by assessing the ratio of the sum of between-clusters
 dispersion and of inter-cluster dispersion for all clusters --- thus,
 the higher the score the better the performances \citep{calinski1974}.
-In our case the \(k=2\) solution also showed the highest
+In our case, the \(k=2\) solution also showed the highest
 Calinski-Harabasz index (\(CH_2=\) 16.38; a full table of all extracted

### 2023-07-22 latex-results/07ClusterEvaluationInterpretation.tex
feedback rm and pdj

@@
 see that for some variables the features are generally stronger in
-separating the clusters. We, for example see that the item on
+separating the clusters. We, for example, see that the item on
 `\textit{how cooperative the interaction was}' distinguishes the two
@@
 offers some information on which features were most important in
-understanding the two extracted groups. Taken these two perspectives
+understanding the two extracted groups. Taking these two perspectives
 together, we can also focus on individual features or variables in
@@
   \caption*{Note: \\
-  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+  "Int." = interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
 \end{figure}
 
-In the second step, we look at prototypical trajectories of the
+In the second step, we look at the prototypical trajectories of the
 clusters. For k-means clustering it is often recommended to use the
@@
 Immediately striking are the mean differences, where participants in
-cluster 1 had more meaningful and fulfilling outgroup interactions also
-consistently reported more voluntary and cooperative interactions but
-less accidental and involuntary interactions. The same cluster (cluster
-1) also reported an increase in need-fulfilling interaction over the
-30-day period and an increase in interactions that were representative
-of the outgroup. Whereas the other cluster (cluster 2) showed a decrease
-in voluntary, cooperative, and positive interactions over the 30 days.
-This `deterioration' cluster (cluster 2) also saw a decrease in general
-need fulfillment and experienced well-being over the 30 days (see
-\fgrref[B]{fig:clusterTs}). We also see that while interaction
-representativeness, outgroup attitudes, and well-being are relatively
-stable for both clusters, the deteriorating cluster (cluster 2) also
-showed substantially higher variablity and instability on most of the
-other variables (also see \fgrref[A]{fig:clusterTs}).
+cluster 1 had more meaningful and fulfilling outgroup interactions and
+also consistently reported more voluntary and cooperative interactions
+but fewer accidental and involuntary interactions. The same cluster
+(cluster 1) also reported an increase in need-fulfilling interactions
+over the 30-day period and an increase in interactions that were
+representative of the outgroup. Whereas the other cluster (cluster 2)
+showed a decrease in voluntary, cooperative, and positive interactions
+over the 30 days. This `deterioration' cluster (cluster 2) also saw a
+decrease in general need fulfillment and experienced well-being over the
+30 days (see \fgrref[B]{fig:clusterTs}). We also see that while
+interaction representativeness, outgroup attitudes, and well-being are
+relatively stable for both clusters, the deteriorating cluster (cluster
+2) also showed substantially higher variability and instability on most
+of the other variables (also see \fgrref[A]{fig:clusterTs}).
 
 Finally, we can also assess the clusters across other person-level
-variable \citep[e.g.,][]{monden2022}. This out-of-feature comparison
+variables \citep[e.g.,][]{monden2022}. This out-of-feature comparison
 allows us to check for data artifacts, as well as check whether the
@@
 intergroup interactions in particular (see \(n\) in
-\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures we
+\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures, we
 ensured that the general time frame and completion rates are similar for
@@
 are largely similar (e.g., see \(n\) for well-being and outgroup
-attitudes). However, the difference in reported number of interactions
-might indicate either a clustering artifact or a meaningful difference.
-The higher average number of interactions in cluster 2 could, for
-example, indicate a clustering artifact if variances are more inflated
-due to the larger samples \citep{kogan2006}. At the same time, however,
-the difference in the number of experienced interactions might also
-indicate a meaningful difference, where the deteriorating cluster
+attitudes). However, the difference in the reported number of
+interactions might indicate either a clustering artifact or a meaningful
+difference. The higher average number of interactions in cluster 2
+could, for example, indicate a clustering artifact if variances are more
+inflated due to the larger samples \citep{kogan2006}. At the same time,
+however, the difference in the number of experienced interactions might
+also indicate a meaningful difference, where the deteriorating cluster
 (cluster 2) on average reported more outgroup interactions, but these
@@
 Such a finding is in line with past research highlighting the role of
-negative intergroup interations in explaining intergroup relations
+negative intergroup interactions in explaining intergroup relations
 \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. To further illustrate
-the utility of assessing out of feature individual differences, we also
+the utility of assessing out-of-feature individual differences, we also
 compare the two samples in terms of the participants' self-reported
@@
 (\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
-(e.g., sum of specific ESM measurements) and cross-sectional variables
-(e.g., general discrimination differences) that were not included in the
-original clustering step can be used to explore and understand the
-cluster differences in more detail.
+(e.g., the sum of specific ESM measurements) and cross-sectional
+variables (e.g., general discrimination differences) that were not
+included in the original clustering step can be used to explore and
+understand the cluster differences in more detail.
 
-In short, we find that that the feature-based clustering discerned two
+In short, we find that the feature-based clustering discerned two
 meaningfully different groups of participants. We find an adaptive group
@@
 clustering offers some direct insight into the complexity within the
-data set. In our empirical example, we for example see that participants
-are meaningfully distinguished by a combination of high (vs.~low)
-central tendency, variability, and linear trend. Secondly, the clusters
-identify important groups. The adaptive and deteriorating groups offer
-starting points for empirical exploration as well as practical
-interventions. Researchers can start probing what exactly distinguishes
-the two groups further and generate new bottom-up hypotheses.
-Practitioners, can use the group separation to identify individuals in
-need of assistance and can explore contextual factors that might
-contribute to the difficulties some might face. Thirdly, the
+data set. In our empirical example, we, for example, see that
+participants are meaningfully distinguished by a combination of high
+(vs.~low) central tendency, variability, and linear trend. Secondly, the
+clusters identify important groups. The adaptive and deteriorating
+groups offer starting points for empirical exploration as well as
+practical interventions. Researchers can start probing what exactly
+distinguishes the two groups further and generate new bottom-up
+hypotheses. Practitioners can use the group separation to identify
+individuals in need of assistance and can explore contextual factors
+that might contribute to the difficulties some might face. Thirdly, the
 feature-based approach is flexible and meaningful. We were able to use a
@@
 groups. For our empirical illustration we, among others for example,
-chose to focus on whether participants differed in their their average
+chose to focus on whether participants differed in their average
 well-being (i.e., \textit{median}), how much their well-being would vary

### 2023-07-23 manuscript.tex
internal feedback

@@
 Psychological time series data has not only become more common but also more diverse and complex.
-Researchers and practitioners are increasingly interested in identifying person-level differences in the developments of their patients, or participants.
+Researchers and practitioners are increasingly interested in identifying important differences in the developments of their patients, or participants.
 Past research has proposed a number of `dynamic measures' that describe meaningful developmental patterns for psychological data (e.g., instability, inertia, linear trend).
@@
 Researchers and practitioners are, for example, asking: ``Do the symptoms of different patients develop in contrasting ways?'' \citep{Monden2015} or ``How do migrants differ in the development of their self-reported needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
-There is, thus, a clear need for analysis techniques that identify person-level differences in developmental patterns for psychological data.
+There is, thus, a clear need for analysis techniques that identify between-subject differences in developmental patterns for psychological data.
 
-Recently, one promising way of identifying person-level developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
+Recently, one promising way of identifying between-subject developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}.
 This analysis type essentially seeks to capture comparable within-person developments --- such as whether a variable remains stable over time, consistently increases, or exhibits cyclical patterns --- and then groups the persons based on these patterns  \citep[][]{liao2005}\footnote{It should be noted in some cases the time series do not need to be summarized and can be compared directly.
@@
 As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-We argue that for psychological time series data, feature-based clustering offers a flexible variety, theory-based selection, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.
+We argue that for psychological time series data, feature-based clustering offers a flexible variety, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.
 
@@
 To illustrate the functioning and utility of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected data on migration experiences.
-Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups \citep[e.g.,][]{Keil2020}.
-This type of research comes as a response to prominent reviews within the migration literature, which have called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}, while at the same time, conceptual works have pushed for more diverse assessments of migration experiences, including motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
-Importantly for our illustration, the migration ESM research, thus, exemplifies some of the real-world data issues that ESM data commonly face, including a multivariate conceptualization with event-specific missingness patterns.
+Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups.
+We have seen both new technological advances to capture such interactions \citep[e.g.,][]{Keil2020} as well as a rise in empirical studies that look at aspects such as the well-being of migrants \citep[e.g.,][]{hendriks2016} and the distress of intergroup contact \citep[e.g.,][]{doucerain2023}.
+
+This type of research comes as a response to a long-standing theoretical tradition highlighting the dynamic and developmental nature of cultural adaptation \citep[e.g.,][]{berry1986}.
+Importantly, such developmental trajectories are often difficult and stressful \citep{Ward2016} --- so that some may have an adaptive trajectory while others face a more difficult and grievous trajectory \citep{kim2017}.
+Prominent reviews within the migration literature have, thus, called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}.
+By the same token, there is a crucial need to distinguish (mal-)adaptive clusters within these developmental trajectories and to relate these clusters to individual differences and contextual variables to make them understandable \citep[e.g.,][]{choi2009}.
+
+At the same time, conceptual works on cultural adaptation have highlighted the multidimensional perspective necessary to understand migration experiences.
+A recent scoping review has particularly highlighted that cultural adaptation is best understood as a joint process of motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}.
+The data on migration experiences are, thus, explicitly multidimensional, mirroring the increase of complex data within the ESM literature \citep[][]{wardenaar2013}.
+
+In short, the data on migration experience we are using for this illustration, address a pressing societal issue of identifying and understanding diverging trajectories.
+And importantly for our illustration, the migration ESM research, also, exemplifies the real-world data issues that ESM data commonly face, including a multivariate conceptualization with event-specific missingness patterns (also see \appref[]{app:ChallengesAppendix} for an expanded discussion of the current challenges within ESM data that are addressed within the data).
 
@@
 % aims re-iterated
-The purpose of this article was to introduce feature-based time series clustering as an amenable and transparent approach to understanding person-level differences in developmental patterns of psychological time series data.
+The purpose of this article was to introduce feature-based time series clustering as an amenable and transparent approach to understanding between-person differences in developmental patterns of psychological time series data.
 Rather than relying on person-specific model parameters, which can be restrictive and assumption-bound, we argue for the more flexible and theoretically grounded approach of directly clustering on important and relevant features of the time series data.
-By leveraging the rich array of dynamic measures, our approach offers the advantages of flexibility, theory-based selection, fewer strict assumptions, and improved interpretability, thus potentially enriching our understanding of heterogeneous psychological processes in intensive longitudinal studies.
+By leveraging the rich array of dynamic measures, our approach offers the advantages of flexibility, fewer strict assumptions, and improved interpretability, thus potentially enriching our understanding of heterogeneous psychological processes in intensive longitudinal studies.
 

### 2023-07-23 sections/01DataDescription.Rmd
internal feedback

@@
 
-The data set we use consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members \citep[for the data set see][]{Kreienkamp2022b}.
+Matching these requirements, the data set we use consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members \citep[for the data set see][]{Kreienkamp2022b}.
 After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days.

### 2023-07-23 sections/03FeatureExtraction.Rmd
internal feedback

@@
 Person level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
-As such, psychological researchers and practitioners are often empirically interested in person-level differences in variability.
+As such, psychological researchers and practitioners are often empirically interested in between-person differences in variability.
 Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or do they vary across the measurement period?''
@@
 Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
-Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "What is the nature of daily mood swings in individuals with bipolar disorder?" or "How does the pattern of day-to-day (i.e., measurement-to-measurement) fluctuations in self-esteem during adolescence reflect the process of identity formation and self-concept development?"
+Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "What is the nature of the identification changes in those who start working in a new country?" or "How does the pattern of day-to-day (i.e., measurement-to-measurement) fluctuations in self-esteem during adolescence reflect the process of identity formation and self-concept development?"
 
@@
 In a similar vein, high inertia in negative affect at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
-Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a negative mood for several measurements?''
+Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a negative mood for several measurements?'' or ``Do migrants stay with their language practice for several days at a time?''
 
@@
 But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}.
-Conceptually, researchers might have research questions about the type of the development: "Is the development of anxiety a nonlinear process?" as well as the shape and structure of the development: "How many spikes in anxiety did a patient experience?"
+Conceptually, researchers might have research questions about the type of the development: "Is the development of well-being a nonlinear process?" as well as the shape and structure of the development: "How many spikes in well-being did a migrant experience?"
 We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model.

### 2023-07-23 sections/07ClusterEvaluationInterpretation.Rmd
internal feedback

@@
 
-We first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
+In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
+We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
+This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
+This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 1).
+This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
+At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
+
+To identify these patterns, we first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
 We see that for some variables the features are generally stronger in separating the clusters.
@@
   \caption*{Note: \\
-  "Int." = interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
 \end{figure}
@@
 
-<!-- \Question{\textcolor{cyan}{Talk about early warning signals here? E.g., divergences in means over time could be investigated with targeted studies.}} -->
-
-Finally, we can also assess the clusters across other person-level variables \citep[e.g.,][]{monden2022}.
+Finally, we can also assess the clusters across other individual difference variables \citep[e.g.,][]{monden2022}.
 This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
@@
 
-In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
-We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
-This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
-This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 1).
-This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
-This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
-At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
-
 This cluster separation, then, has a number of empirical and practical applications.
@@
 Researchers can start probing what exactly distinguishes the two groups further and generate new bottom-up hypotheses.
-Practitioners can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
+Practitioners in the resettlement field can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
 Thirdly, the feature-based approach is flexible and meaningful.

### 2023-07-23 sections/07ClusterEvaluationInterpretation.Rmd
internal feedback

@@
   geom_line() +
+  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
   #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
@@
   #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
+  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
   coord_flip() +
@@
   stat_summary(fun=mean, geom="line") +
+  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
   facet_wrap(~ label, scales = "free_y") +
@@
   geom_smooth(aes(group=cluster)) +
+  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
   #stat_summary(fun=mean, geom="line") +

### 2023-07-28 Supplemental-Material-A-Annotated-Analysis.Rmd
adj SI: A

@@
   bookdown::html_document2:
+    highlight: pygments
     fig_caption: yes
@@
 }
+
+body
+  { counter-reset: source-line 0; }
+pre.numberSource code
+  { counter-reset: none; }
+
 </style>
@@
   warning = FALSE,
-  message = FALSE
+  message = FALSE,
+  class.source = "numberLines lineAnchors"
 )
@@
 This means that the datasets had inconsistent file formats and naming conventions.
-For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*').
+For the Qualtrics study we pre-processed some variables to ease the import process (for the syntax files see the SPSS files in '*data/S1_Workers/processed/cleaned*' and for the raw data files see '*data/S1_Workers/raw*').
 For the two other studies, we import the raw csv files from their respective folders.
@@
 
-For the student sample data was, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
+For the student sample data were, similarly, collected in three separate surveys: (1) the pre-measurement, (2) the daily survey sent out at lunch and dinner time, and (3) a post-measurement.
 We combine the three individual surveys into one large dataframe and drop superfluous variables that are not relevant to the analyses relevant here.
@@
 df.pre <- dtS2$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
+  mutate(across(where(is.character), na_if, "")) %>%
+  mutate(across(where(is.character), na_if, "NA")) %>%
   filter(!is.na(ended)) %>% # remove all who did not finish
@@
 df.post <- dtS2$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
+  mutate(across(where(is.character), na_if, "")) %>%
+  mutate(across(where(is.character), na_if, "NA")) %>%
   filter(!is.na(ResponseId)) %>% # remove own test runs
@@
 df.daily <- dtS2$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
+  mutate(across(where(is.character), na_if, "")) %>%
+  mutate(across(where(is.character), na_if, "NA")) %>%
   filter(!ResponseId %in% ownIDs) %>% # remove our own test
@@
 df.pre <- dtS3$raw.pre %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
+  mutate(across(where(is.character), na_if, "")) %>%
+  mutate(across(where(is.character), na_if, "NA")) %>%
   filter(!is.na(ended)) %>% # remove all who did not finish
@@
 df.post <- dtS3$raw.post %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>% 
+  mutate(across(where(is.character), na_if, "")) %>%
+  mutate(across(where(is.character), na_if, "NA")) %>%
   filter(!is.na(ResponseId)) %>% # remove own test runs
@@
 df.daily <- dtS3$raw.daily %>%
-  mutate_all(na_if, "") %>%
-  mutate_all(na_if, "NA") %>%
+  mutate(across(where(is.character), na_if, "")) %>%
+  mutate(across(where(is.character), na_if, "NA")) %>%
   filter(!ResponseId %in% ownIDs) %>% # remove our own test
@@
 2. Feature extraction
-3. (Feature selection)
+3. Feature reduction
 4. Clustering
@@
 
-We begin by selecting the variables, which we would like to include in the clustering procedure.
-For the clustering of time-series characteristics we focus on variables that are specific to the ESM measurements (i.e., variables collected during the various daily dairy pings).
-This means, that variables from the pre- and post-test are excluded from the clustering procedure.
-While the variables can in theory be added as additional features, they are not the focus of this manuscript.
+We included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
+We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature [@Kreienkamp2022d].
+
+FILTER BY main
 
@@
 
-apply(dtS1Red, 2, sum(is.na(.))/nrow(.))
-
-your_dataframe %>%
-  group_by(Your_Participantid) %>%
-  summarise(across(everything(), funs(100*mean(is.na(.)))))
+# apply(dtS1Red, 2, sum(is.na(.))/nrow(.))
+# 
+# your_dataframe %>%
+#   group_by(Your_Participantid) %>%
+#   summarise(across(everything(), funs(100*mean(is.na(.)))))
 
@@
 
-var <- get_pca_var(res.pca)
-corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
+# var <- get_pca_var(res.pca)
+# corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
 fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
@@
 
-res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
+res.desc <-
+  FactoMineR::dimdesc(
+    FactoMineR::PCA(scaled_data, scale.unit = FALSE, graph = FALSE),
+    axes = c(1, 2),
+    proba = 0.05
+  )
 # Description of dimension 1
-res.desc$Dim.1
+# res.desc$Dim.1
 

### 2023-07-28 manuscript.tex
feedback lfb

@@
 \abstract{
-Psychological time series data has not only become more common but also more diverse and complex.
-Researchers and practitioners are increasingly interested in identifying important differences in the developments of their patients, or participants.
+Psychological researchers and practitioners collect increasingly complex time series data aimed at identifying differences between the developments of participants or patients.
 Past research has proposed a number of `dynamic measures' that describe meaningful developmental patterns for psychological data (e.g., instability, inertia, linear trend).
@@
 Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][]{hamaker2017}.
-These intensive longitudinal data sets come with new forms of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
-Oftentimes, however, researchers are interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
+These intensive longitudinal data sets come with different sources of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}.
+However, despite its complexity, researchers are often interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}.
 Researchers and practitioners are, for example, asking: ``Do the symptoms of different patients develop in contrasting ways?'' \citep{Monden2015} or ``How do migrants differ in the development of their self-reported needs as they arrive in a new country?'' \citep{Kreienkamp2022d}.
@@
 In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}.
+Such aspects can be indicative of adaptive and maladaptive developments within the individual, can identify crucial transitions, or more generally are helpful in understanding a psychological time series.
 Important aspects might include concerns over whether a symptom consistently stays at a certain level without much variability or whether some emotions develop together.
@@
 These summary statistics are often called ``dynamic measures'', ``principles of change'', or ``dynamic features'' of the psychological time series \citep{dejonckheere2019, kuppens2017, krone2018}.
-For most conceptualizations, the proposed number of aspects ranges from four to twelve key features \citep[][]{wang2006, dejonckheere2019}.
-Each of these statistics not only captures a distinct aspect of psychological time series but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.
+Most research groups working on these time series features have proposed an overlapping number anywhere between four and twelve key features relevant to psychologists \citep[][]{wang2006, dejonckheere2019}.
+Each of these time series features not only captures a distinct aspect of psychological time series but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.
 
-Yet, despite this rich diversity of meaningful time series features in psychology, most clustering of ESM data has only focused on a small and restrictive selection of time series characteristics.
+Yet, despite this diversity of meaningful time series features in psychology, most clustering of ESM data has only focused on a small and restrictive selection of time series characteristics.
 Thus far, the most common approach has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}.
-While such model parameters are familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c} --- only two of the many potentially important time series features.
+While such model parameters have the advantage of being familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c} --- only two of the many potentially important time series features.
 Additionally, the validity of model parameters is traditionally restricted by the assumptions of the particular statistical model.
@@
 These assumptions however stand in contrast to types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}.
-Model parameters might, thus, not always accurately capture the time series and are often restricted in the features they capture.
+Model parameters might, thus, not always accurately capture the time series and are often restricted in the time series features they capture.
 
@@
 Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models, which either seek to relax specific assumptions \citep[e.g.,][]{denteuling2021} or include additional time series features \citep[e.g., see][]{krone2018, gates2017}.
-In this manuscript, we instead, propose to directly cluster based on important and relevant features.
+In this manuscript, we instead, propose to directly cluster based on important and relevant time series features.
 Adequately named \textit{feature-based time series clustering}, such an approach has been a common procedure in digital phenotyping \citep[][]{loftus2022} and the broader machine learning literature \citep[][]{maharaj2019}.
 As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}.
-We argue that for psychological time series data, feature-based clustering offers a flexible variety, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.
+We argue that for psychological time series data, feature-based clustering offers a flexible approach, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.
 
-In the sections below we particularly seek to provide a practical introduction to the method.
+In the sections below we seek to provide a practical introduction to the method.
 To do so, we illustrate the utility of the method with real-world ESM data.
 We use this data to discuss which psychological time series features are well-suited for a clustering approach, introduce the individual analysis steps, and provide practical guidance on common algorithms and analysis code.
-We first introduce the empirical case study data and then follow the individual analysis steps to discuss key decisions and approaches.
 
@@
 To do so we will follow the common separation that has structured feature-based clustering into four main steps \citep{rasanen2009, wang2006}.
-(1) The selection and preparation of the input variables, (2) the extraction of the features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters.
+(1) The selection and preparation of the input variables, (2) the extraction of the time series features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters.
 While this is a common conceptual separation of procedural elements, it is important to note that these steps are a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed.
@@
 Fortunately, past conceptual and empirical efforts offer valuable discussions of common time series features in psychological research.
-To understand emotion dynamics, \citet{kuppens2017} originally proposed four features: (1) within-person variability, (2) co-variance or ICC, (3) inertia or autocorrelation, and (4) cross-lagged correlations.
+To understand emotion dynamics, \citet{kuppens2017} originally proposed four dynamic features: (1) within-person variability, (2) co-variance or intraclass coefficient (ICC), (3) inertia or autocorrelation, and (4) cross-lagged correlations.
 These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity.
@@
 
-Some of the features found in the psychological literature are not necessarily well-suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the literature.
+Some of the time series features found in the psychological literature are not necessarily well-suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the literature.
 In particular, covariances and cross-lagged correlations often produce a large number of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}.
@@
 
-Thus, while the final selection of features should always be driven by the research questions and field-specific conventions, for our illustration we chose six features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
-An overview of the selected features, their substantive interpretations, and mathematical operationalizations is also available in \tblref{tab:esmFeatures}.
-For each of the six time series features we chose, we selected a mathematical representation that was appropriate for our type of data.
+Thus, while the final selection of time series features should always be driven by the research questions and field-specific conventions, for our illustration we chose six time series features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
+An overview of the selected time series features, their substantive interpretations, and mathematical operationalizations is available in \tblref{tab:esmFeatures}.
+For each of the six time series features, we selected a mathematical representation that was appropriate for our type of data.
 We provide a short introduction of each feature below.
@@
 It is important to reiterate that the six selected time-series features are in no way exhaustive or imperative.
-Both using a more data-driven approach to the selection of features or selecting entirely different aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
-Our choice seeks to offer a practical toolbox of features that are common and meaningful to psychological research questions and -practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions.
+Both using a more data-driven approach to the selection of time series features or selecting entirely different aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}.
+Our choice seeks to offer a practical toolbox of time series features that are common and meaningful to psychological research questions and practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions.
 
 \subsection{Feature Reduction}
-Once a meaningful selection of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm.
+Once a meaningful set of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm.
 As an example, a relatively common scenario would include 10 variables of interest, where eight time series features are extracted, resulting in 80 features per participant (with a common sample size of 100 participants that would result in a total of 8,000 data points in this hypothetical example).
@@
 For our own illustration, we have chosen the centroid-based k-means clustering.
-While k-means comes at the expense of high accuracy with more complex cluster shapes, we specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the clusters \citep{jain2010}.
+Although k-means sacrifices some level of accuracy, it offers certain advantages.
+We specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the clusters \citep{jain2010}.
 K-means is also well-established within the research community and has been readily implemented in many statistical software packages \citep{hand2005}.
@@
 The purpose of this article was to introduce feature-based time series clustering as an amenable and transparent approach to understanding between-person differences in developmental patterns of psychological time series data.
-Rather than relying on person-specific model parameters, which can be restrictive and assumption-bound, we argue for the more flexible and theoretically grounded approach of directly clustering on important and relevant features of the time series data.
+Rather than relying on person-specific model parameters, which can be restrictive and assumption-bound, we argue for the more flexible and theoretically grounded approach of directly clustering on relevant features of the time series data.
 By leveraging the rich array of dynamic measures, our approach offers the advantages of flexibility, fewer strict assumptions, and improved interpretability, thus potentially enriching our understanding of heterogeneous psychological processes in intensive longitudinal studies.
@@
 With the real-world data, we followed a stepwise approach to discuss key issues during input selection, feature extraction, feature reduction, feature clustering, and cluster evaluation.
-Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research practice, as both the features and the analysis steps are well established within the field, and statistical packages are readily available.
+Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research, as both the time series features and the analysis steps are well established within the field, and statistical packages are readily available.
 Time series features (such as means or linear trends) are not only easy to extract, but also hold conceptual meaning for psychological data and can be chosen to address specific research questions (also see \tblref{tab:esmFeatures}).
@@
 In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential inconsistent results between methods \citep[e.g.,][]{bastiaansen2019}.
-A different set of variables, features, or a different clustering algorithm, might have resulted in substantially different cluster assignments.
+A different set of variables, time series features, or a different clustering algorithm, might have resulted in substantially different cluster assignments.
 While the variety and diversity of methods are helpful in finding options even for more complex types of data, different algorithms often offer different results \citep[e.g.,][]{keogh2005}.
@@
 
-As an illustration, the choice of features to extract from the time series data is a critical step that can significantly influence the results of the clustering process.
-In the current example, we chose to extract features such as means, autocorrelations, and linear trends, which are psychologically and conceptually meaningful in interpreting our time series clusters.
+As an illustration, the choice of time series features to extract from the time series data is a critical step that can significantly influence the results of the clustering process.
+In the current example, we chose to extract time series features such as means, autocorrelations, and linear trends, which are psychologically and conceptually meaningful in interpreting our time series clusters.
 However, this selection is not exhaustive and may not capture all relevant aspects of the time series data.
 For example, we did not consider attributes like periodicity or spectral density, which could shed light on the data's cyclical patterns.
-The choice of features largely hinges on the researcher's specific research question and assumptions about the data, thereby injecting a level of subjectivity into the process.
+The choice of time series features largely hinges on the researcher's specific research question and assumptions about the data, thereby injecting a level of subjectivity into the process.
 Similar challenges arise with the choice of the clustering algorithm or the cluster illustration.
@@
 shape-based approaches in \fgrref{fig:tsClustTax}; see also][]{caro-martin2018, mayor2022}.
-New features capturing nonlinear structures would add to the under-studied (non-)linear trend features of ESM data.
+New time series features capturing nonlinear structures would add to the under-studied (non-)linear trend features of ESM data.
 
@@
 It is this background of the common challenges of current ESM data, upon which we propose to consider feature-based clustering.
-The flexibility of using a wide variety of features that represent the important developmental patterns allows users to circumvent many of the issues with multi-dimensionality, non-equidistant or missing measurements, non-stationary trends, as well as diverse time scales.
+The flexibility of using a wide variety of time series features that represent the important developmental patterns allows users to circumvent many of the issues with multi-dimensionality, non-equidistant or missing measurements, non-stationary trends, as well as diverse time scales.
 

### 2023-07-28 sections/02InputVariables.Rmd
feedback lfb

@@
 
-For our illustration, we include 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group.
+For our illustration, we include 12 variables that were measured as part of the ESM surveys in all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group (see \tblref{tab:descrLong} for an overview).
 We chose these aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
@@
 ```{r}
+# ADJUST LABELS!! (and add question texts?)
+
 # Sample Descriptives Table --- longer version
@@
     general = c(
-      "Upper triangle: Between-person correlations;",
-      "Lower triangle: Within-person correlations;",
-      "*** p < .001, ** p < .01,  * p < .05"
+      '"Int." = interaction.',
+      'Upper triangle: Between-person correlations;',
+      'Lower triangle: Within-person correlations;',
+      '*** p < .001, ** p < .01,  * p < .05'
     )

### 2023-07-28 sections/03FeatureExtraction.Rmd
feedback lfb

@@
 In time series analyses, variability is conceptually important because information about the distribution and diversity of the data has been found to be indicative of worse psychological states \citep{myin-germeys2018, helmich2021}.
-Person level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
+Person-level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}.
 As such, psychological researchers and practitioners are often empirically interested in between-person differences in variability.
-Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or do they vary across the measurement period?''
+Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or do they vary across the measurement period?"
 
@@
 Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
-Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "What is the nature of the identification changes in those who start working in a new country?" or "How does the pattern of day-to-day (i.e., measurement-to-measurement) fluctuations in self-esteem during adolescence reflect the process of identity formation and self-concept development?"
+Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "What is the nature of the identification changes in those who start working in a new country?" or "Do strong daily fluctuations in self-esteem reflect the process of identity formation in adolescents?"
 
@@
 \begin{equation} \label{eq:mac}
-  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t}-x_{t-1}\right|
+  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=2, \ldots, t}\left|x_{t}-x_{t-1}\right|
 \end{equation}
@@
 
-Another common measurement of instability is the \textit{Mean of the Squared Successive Differences} ($MSSD$), which is often preferred where differences in magnitude are more important than the frequency of those changes, for example, when big shifts in time series are considered more impactful or when outliers are meaningful and need to be taken into account.
+Another common measurement of instability is the \textit{Mean of the Squared Successive Differences} ($MSSD$), which is often preferred where differences in magnitude are more important than the frequency of those changes, for example, when big shifts in time series are considered more impactful or when outliers are meaningful and need to be taken into account \citep{chatfield2003}.
 
@@
 High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}.
-The lag--1 autocorrelation $r_{ij,1}$ looks at the average correlation between a measurement $x$ and the proceeding measurement $x_{t-1}$ for the time series $X$ of participant $i$ and variable $j$.
+The lag--1 autocorrelation $r_{ij,1}$ looks at the average correlation between a measurement $x$ and the preceding measurement $x_{t-1}$ for the time series $X$ of participant $i$ and variable $j$ with $n$ measurements.
 
@@
 \begin{equation} \label{eq:ar}
-  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+  r_{ij,1} = \frac{\sum_{t=2}^{n}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+\end{equation}
+```
+
+Where $\overline{x}_{ij}$ is the mean of the time series $x_{ij}$, calculated as:
+
+```{=tex}
+\begin{equation} \label{eq:mean_for_ar1}
+  \overline{x}_{ij} = \frac{1}{n} \sum_{t=1}^{n} x_{ijt}
 \end{equation}
@@
 Conceptually, researchers might have research questions about the type of the development: "Is the development of well-being a nonlinear process?" as well as the shape and structure of the development: "How many spikes in well-being did a migrant experience?"
+
 We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model.
-The $edf$ summarizes the \textit{wiggliness} of a spline trend line by estimating the number of parametric and non-parametric (or smooth) terms $p$ in the model.
-An edf of 1 would be equivalent to a linear relationship (i.e., one linear slope parameter), whereas a higher edf (particularly an edf > 2) is indicative of a non-linear trend.
+The $edf$ summarizes the \textit{wiggliness} of a spline trend line \citep{wood2017}.
+The degrees of freedom of a spline model are primarily determined by the number of knots and the order of the spline.
+For instance, a cubic spline with $k$ knots has $k$+3 degrees of freedom \citep{faraway2016}.
+However, in a penalized spline framework, which is commonly used for GAMs, the effective degrees of freedom can be less than $k$+3.
+This is because the model employs a smoothing parameter to control the trade-off between the complexity (flexibility) of the model and its fit to the data, thereby penalizing overly complex models and potentially reducing the effective degrees of freedom \citep{marx1998}.
+Intuitively then an edf of 1 would be equivalent to a linear relationship (i.e., one linear slope parameter), whereas a higher edf (particularly an edf > 2) is indicative of a non-linear trend.
+The estimated degrees of freedom are commonly based on a concept called 'effective degrees of freedom' and can be represented as the trace $tr$ (i.e., the sum of the diagonal elements) of the smoother matrix $S$, a symmetric matrix that maps from the raw data to the smooth estimates \citep{wood2017}.
+
+```{=tex}
+\begin{equation} \label{eq:df}
+  edf = tr(S)
+\end{equation}
+```
+
+<!--Effective degrees of freedom for individual smooth terms are obtained by summing the Fii values corresponding to the coefficients βi of the smooth term (e.g., if the coefficients of a particular smooth are β5, .
+.
+.
+, β11 then the smooth’s EDF is $\sum_{i=5}^{11} = F_{ii}).--!>
+
+<!-- An alternative approach to understanding the edf is to approximate the number of parametric and non-parametric (or smooth) terms $p$ in the penalized GAM.
+Both formulas are measuring the same concept (the model's complexity or flexibility) and both account for the penalization introduced to control overfitting in GAMs \citep{hastie1986}.
 
@@
 ```
+--!>
 

### 2023-07-28 sections/04FeatureReduction.Rmd
feedback lfb

@@
 
-To use the principal component analysis with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
+To use the PCA with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
 We then enter all `r ncol(scaled_data)` features into the analysis.

### 2023-07-28 sections/07ClusterEvaluationInterpretation.Rmd
feedback lfb

@@
 In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
-We find an adaptive group (cluster 2) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
+We find an adaptive group (cluster 1) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
 This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
-This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 1).
+This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 2).
 This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
 This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
-At the same time, for members of this detrimental cluster conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
+At the same time, for members of this detrimental cluster (cluster 2) conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
 
@@
 We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series.
-At the same time, in terms of stability, both groups have virtually identical averages \textit{MAC} statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
+At the same time, in terms of stability, both groups have virtually identical average \textit{MAC} statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
 There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
@@
 Such a finding is in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+
 To further illustrate the utility of assessing out-of-feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
@@
   \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
+  Subplot (A) displays the variable cluster means at every measurement occasion.
+Subplot (B) shows the GAM spline for each cluster across the measurement occasions.
+The thinner lines present all individual time series.}
 \end{figure}

### 2023-07-28 latex-results/01DataDescription.tex
feedback lfb

@@
-The data set we use consists of three studies that followed migrants who
-had recently arrived in the Netherlands in their daily interactions with
-the Dutch majority group members
-\citep[for the data set see][]{Kreienkamp2022b}. After a general
+Matching these requirements, the data set we use consists of three
+studies that followed migrants who had recently arrived in the
+Netherlands in their daily interactions with the Dutch majority group
+members \citep[for the data set see][]{Kreienkamp2022b}. After a general
 migration-focused pre-questionnaire, participants were invited twice per

### 2023-07-28 latex-results/02InputVariables.tex
feedback lfb

@@
-For our illustration, we include 12 variables that were available across
-all three studies and captured information about the participant's
-interactions, as well as cognitive-, emotional-, and motivational self
-in relationship with the majority group. We chose these aspects in
+For our illustration, we include 12 variables that were measured as part
+of the ESM surveys in all three studies and captured information about
+the participant's interactions, as well as cognitive-, emotional-, and
+motivational self in relationship with the majority group (see
+\tblref{tab:descrLong} for an overview). We chose these aspects in
 particular because (1) the interaction-specific information exemplified

### 2023-07-28 latex-results/03FeatureExtraction.tex
feedback lfb

@@
 of worse psychological states \citep{myin-germeys2018, helmich2021}.
-Person level differences of ESM measurements have, for example, been
+Person-level differences of ESM measurements have, for example, been
 associated with higher levels of psycho-pathological recurrences among
 depression patients \citep{timm2017}. As such, psychological researchers
-and practitioners are often empirically interested in person-level
+and practitioners are often empirically interested in between-person
 differences in variability. Researchers on polarization and
@@
 attitudes towards migrants or do they vary across the measurement
-period?'\,'
+period?''
 
@@
 instability feature, thus, relates to a broad range of research
-questions, including: ``What is the nature of daily mood swings in
-individuals with bipolar disorder?'' or ``How does the pattern of
-day-to-day (i.e., measurement-to-measurement) fluctuations in
-self-esteem during adolescence reflect the process of identity formation
-and self-concept development?''
+questions, including: ``What is the nature of the identification changes
+in those who start working in a new country?'' or ``Do strong daily
+fluctuations in self-esteem reflect the process of identity formation in
+adolescents?''
 
@@
 series are considered more impactful or when outliers are meaningful and
-need to be taken into account.
+need to be taken into account \citep{chatfield2003}.
 
@@
 initial onset of depression \citep{kuppens2012}. Conceptually, inertia
-is more broadly connected to research questions such as: ``Do patients
-stay in a negative mood for several measurements?'\,'
+is more broadly connected to research questions such as:
+\texttt{Do\ patients\ stay\ in\ a\ negative\ mood\ for\ several\ measurements?\textquotesingle{}\textquotesingle{}\ or}Do
+migrants stay with their language practice for several days at a
+time?'\,'
 
@@
 \citep{dejonckheere2019}. The lag--1 autocorrelation \(r_{ij,1}\) looks
-at the average correlation between a measurement \(x\) and the
-proceeding measurement \(x_{t-1}\) for the time series \(X\) of
-participant \(i\) and variable \(j\).
+at the average correlation between a measurement \(x\) and the preceding
+measurement \(x_{t-1}\) for the time series \(X\) of participant \(i\)
+and variable \(j\) with \(n\) measurements.
 
 \begin{equation} \label{eq:ar}
-  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+  r_{ij,1} = \frac{\sum_{t=2}^{n}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+\end{equation}
+
+Where \(\overline{x}_{ij}\) is the mean of the time series \(x_{ij}\),
+calculated as:
+
+\begin{equation} \label{eq:mean_for_ar1}
+  \overline{x}_{ij} = \frac{1}{n} \sum_{t=1}^{n} x_{ijt}
 \end{equation}
@@
 questions about the type of the development: ``Is the development of
-anxiety a nonlinear process?'' as well as the shape and structure of the
-development: ``How many spikes in anxiety did a patient experience?'' We
-summarized the nonlinear trend with the
+well-being a nonlinear process?'' as well as the shape and structure of
+the development: ``How many spikes in well-being did a migrant
+experience?''
+
+We summarized the nonlinear trend with the
 \textit{estimated degrees of freedom} of an empty GAM spline model. The
-\(edf\) summarizes the \textit{wiggliness} of a spline trend line by
-estimating the number of parametric and non-parametric (or smooth) terms
-\(p\) in the model. An edf of 1 would be equivalent to a linear
+\(edf\) summarizes the \textit{wiggliness} of a spline trend line
+\citep{wood2017}. The degrees of freedom of a spline model are primarily
+determined by the number of knots and the order of the spline. For
+instance, a cubic spline with \(k\) knots has \(k\)+3 degrees of freedom
+\citep{faraway2016}. However, in a penalized spline framework, which is
+commonly used for GAMs, the effective degrees of freedom can be less
+than \(k\)+3. This is because the model employs a smoothing parameter to
+control the trade-off between the complexity (flexibility) of the model
+and its fit to the data, thereby penalizing overly complex models and
+potentially reducing the effective degrees of freedom \citep{marx1998}.
+Intuitively then an edf of 1 would be equivalent to a linear
 relationship (i.e., one linear slope parameter), whereas a higher edf
 (particularly an edf \textgreater{} 2) is indicative of a non-linear
-trend.
+trend. The estimated degrees of freedom are commonly based on a concept
+called `effective degrees of freedom' and can be represented as the
+trace \(tr\) (i.e., the sum of the diagonal elements) of the smoother
+matrix \(S\), a symmetric matrix that maps from the raw data to the
+smooth estimates \citep{wood2017}.
+
+\begin{equation} \label{eq:df}
+  edf = tr(S)
+\end{equation}
+
+\textless!--Effective degrees of freedom for individual smooth terms are
+obtained by summing the Fii values corresponding to the coefficients βi
+of the smooth term (e.g., if the coefficients of a particular smooth are
+β5, . . . , β11 then the smooth's EDF is \$\sum\emph{\{i=5\}\^{}\{11\} =
+F}\{ii\}).--!\textgreater{}
+
+\textless!-- An alternative approach to understanding the edf is to
+approximate the number of parametric and non-parametric (or smooth)
+terms \(p\) in the penalized GAM. Both formulas are measuring the same
+concept (the model's complexity or flexibility) and both account for the
+penalization introduced to control overfitting in GAMs
+\citep{hastie1986}.
 
@@
 \end{equation}
+
+--!\textgreater{}
 

### 2023-07-28 latex-results/04FeatureReduction.tex
feedback lfb

@@
 
-To use the principal component analysis with our extracted time series
-features, we first standardize all features across participants to
-ensure that all features are weighted equally. We then enter all 72
-features into the analysis. The PCA uses linear transformations in such
-a way that the first component captures the most possible variance of
-the original data
+To use the PCA with our extracted time series features, we first
+standardize all features across participants to ensure that all features
+are weighted equally. We then enter all 72 features into the analysis.
+The PCA uses linear transformations in such a way that the first
+component captures the most possible variance of the original data
 \citep[e.g., by finding a vector that maximizes the sum of squared distances][]{jolliffe2002, abdi2010}.

### 2023-07-28 latex-results/07ClusterEvaluationInterpretation.tex
feedback lfb

@@
-We first inspect the clusters based on the average values of meaningful
-features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We
-see that for some variables the features are generally stronger in
-separating the clusters. We, for example, see that the item on
+In short, we find that the feature-based clustering discerned two
+meaningfully different groups of participants. We find an adaptive group
+(cluster 1) that reports higher well-being and more positive outgroup
+attitudes (\textit{median}) that are also stable over time
+(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
+period (\textit{linear trend}). This group also reported consistently
+more meaningful, need-fulfilling, and cooperative outgroup interactions
+(\textit{median}). This group with overwhelmingly positive experiences
+stands in contrast with a more detrimental group (cluster 2). This
+cluster, on average, reported much less positive, less meaningful, and
+less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower
+well-being, and more discrimination experiences (\textit{median}). At
+the same time, for members of this detrimental cluster (cluster 2)
+conditions seemed to deteriorate over time (\textit{linear trend}), and
+there was generally less consistency in the experiences they were able
+to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
+
+To identify these patterns, we first inspect the clusters based on the
+average values of meaningful features (see
+\fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We see that for
+some variables the features are generally stronger in separating the
+clusters. We, for example, see that the item on
 `\textit{how cooperative the interaction was}' distinguishes the two
@@
 well-being over the time series. At the same time, in terms of
-stability, both groups have virtually identical averages \textit{MAC}
+stability, both groups have virtually identical average \textit{MAC}
 statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There
@@
 
-Finally, we can also assess the clusters across other person-level
-variables \citep[e.g.,][]{monden2022}. This out-of-feature comparison
-allows us to check for data artifacts, as well as check whether the
-developmental clusters are associated with important social markers and
-individual differences. To illustrate artifact checks, we added the
-number of ESM measurements into the comparison and find that the
-participants in the deterioration cluster (cluster 2) on average
+Finally, we can also assess the clusters across other individual
+difference variables \citep[e.g.,][]{monden2022}. This out-of-feature
+comparison allows us to check for data artifacts, as well as check
+whether the developmental clusters are associated with important social
+markers and individual differences. To illustrate artifact checks, we
+added the number of ESM measurements into the comparison and find that
+the participants in the deterioration cluster (cluster 2) on average
 completed slightly more ESM surveys in general and reported on more
@@
 negative intergroup interactions in explaining intergroup relations
-\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. To further illustrate
-the utility of assessing out-of-feature individual differences, we also
-compare the two samples in terms of the participants' self-reported
-discrimination experiences in the Netherlands (measured during the
-post-measurement). When looking at the group comparison, we find that
-participants in the deteriorating cluster (cluster 2) reported
-substantially higher levels of everyday discrimination
-(\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
-(e.g., the sum of specific ESM measurements) and cross-sectional
-variables (e.g., general discrimination differences) that were not
-included in the original clustering step can be used to explore and
-understand the cluster differences in more detail.
+\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
 
-In short, we find that the feature-based clustering discerned two
-meaningfully different groups of participants. We find an adaptive group
-(cluster 2) that reports higher well-being and more positive outgroup
-attitudes (\textit{median}) that are also stable over time
-(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
-period (\textit{linear trend}). This group also reported consistently
-more meaningful, need-fulfilling, and cooperative outgroup interactions
-(\textit{median}). This group with overwhelmingly positive experiences
-stands in contrast with a more detrimental group (cluster 1). This
-cluster, on average, reported much less positive, less meaningful, and
-less fulfilling interactions and interaction patterns (\textit{median}).
-This group also reported less positive outgroup attitudes, lower
-well-being, and more discrimination experiences (\textit{median}). At
-the same time, for members of this detrimental cluster conditions seemed
-to deteriorate over time (\textit{linear trend}), and there was
-generally less consistency in the experiences they were able to have
-(\textit{MAC}, \textit{MAD}, \textit{edf}).
+To further illustrate the utility of assessing out-of-feature individual
+differences, we also compare the two samples in terms of the
+participants' self-reported discrimination experiences in the
+Netherlands (measured during the post-measurement). When looking at the
+group comparison, we find that participants in the deteriorating cluster
+(cluster 2) reported substantially higher levels of everyday
+discrimination (\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive
+longitudinal (e.g., the sum of specific ESM measurements) and
+cross-sectional variables (e.g., general discrimination differences)
+that were not included in the original clustering step can be used to
+explore and understand the cluster differences in more detail.
 
@@
 distinguishes the two groups further and generate new bottom-up
-hypotheses. Practitioners can use the group separation to identify
-individuals in need of assistance and can explore contextual factors
-that might contribute to the difficulties some might face. Thirdly, the
-feature-based approach is flexible and meaningful. We were able to use a
-wide range of time series features that have been central in the ESM
-literature and were able to use them directly to identify meaningful
-groups. For our empirical illustration we, among others for example,
-chose to focus on whether participants differed in their average
-well-being (i.e., \textit{median}), how much their well-being would vary
-over time (i.e, \textit{MAD}), and whether their well-being would on
-average increase or decrease over time (i.e., \textit{linear trend}).
-Alternatively, for others cyclical patterns might be more important ---
-for example, whether well-being was higher on weekends. Importantly, in
-any case, we did not need to translate these dynamic features into
-probabilistic inference models (e.g., VAR models) to cluster the
-participants.
+hypotheses. Practitioners in the resettlement field can use the group
+separation to identify individuals in need of assistance and can explore
+contextual factors that might contribute to the difficulties some might
+face. Thirdly, the feature-based approach is flexible and meaningful. We
+were able to use a wide range of time series features that have been
+central in the ESM literature and were able to use them directly to
+identify meaningful groups. For our empirical illustration we, among
+others for example, chose to focus on whether participants differed in
+their average well-being (i.e., \textit{median}), how much their
+well-being would vary over time (i.e, \textit{MAD}), and whether their
+well-being would on average increase or decrease over time (i.e.,
+\textit{linear trend}). Alternatively, for others cyclical patterns
+might be more important --- for example, whether well-being was higher
+on weekends. Importantly, in any case, we did not need to translate
+these dynamic features into probabilistic inference models (e.g., VAR
+models) to cluster the participants.
 

### 2023-07-29 IllustrationResultsRenderReport.Rmd
internal feedback
feedback lfb

@@
   ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
+  select(varNam, question, varGroup, contactSpecific, study1, study2, study3) %>%
   filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
+  group_by(varGroup, varNam, question, contactSpecific, study1, study2, study3) %>%
   summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
+  pivot_wider(names_from = c("varNam", "question", "varGroup", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
   ungroup %>%
@@
   tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
+  separate(variable, c("Variable", "Question", "Aspect", "study1", "study2", "study3"), " - ") %>%
   mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="X")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
+  select(-study1, -study2, -study3) %>%
+  arrange(`Interaction Only`, Variable) %>% 
+  mutate(ESM = 1) %>%
+  mutate_all(~recode(.x, `0`="", ` 1`="✔")) %>%
+  relocate(starts_with("ESM"), .after = "Aspect") %>%
+  # relocate(starts_with("study"), .after = last_col()) %>%
+  # rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
   kbl(.,
@@
       booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
+      align = c("l", "l", "l", rep("c", ncol(.)-3)),
       digits=2,
       caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
+  add_header_above(c(" " = 4, "Contact Specific" = 2)) %>%
   kable_classic(
@@
 
-## CLuster Performance
+## Cluster Performance
 
@@
 
-## CLuster Interpretation
+## Cluster Interpretation
 

### 2023-07-29 latex-results/07ClusterEvaluationInterpretation.tex
feedback lfb

@@
 could, for example, indicate a clustering artifact if variances are more
-inflated due to the larger samples \citep{kogan2006}. At the same time,
-however, the difference in the number of experienced interactions might
-also indicate a meaningful difference, where the deteriorating cluster
-(cluster 2) on average reported more outgroup interactions, but these
-interactions were less voluntary, less meaningful, and less positive.
-Such a finding is in line with past research highlighting the role of
-negative intergroup interactions in explaining intergroup relations
+inflated due to the larger samples \citep{kogan2006}. In our case this
+seems less likely because one out of four variables did not differ in
+terms of the MAD (see \fgrref{fig:cluster_comparison_mad}). At the same
+time, however, the difference in the number of experienced interactions
+might also indicate a meaningful difference, where the deteriorating
+cluster (cluster 2) on average reported more outgroup interactions
+(\(difference\) = 1.028, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
+\(95%CI [0.76, 1.30]), but these interactions were less voluntary (
+\)difference\$ = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
+\(95%CI [-1.31, -0.77]), less meaningful (
+\)difference\$ = -1.003, \(t\)(136.40) = -7.16, \(p\) \textless{} .001,
+\(95%CI [-1.28, -0.73]), and less positive (
+\)difference\$ = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{} .001,
+\$95\%CI {[}-1.61, -1.15{]}). Such a finding is in line with past
+research highlighting the role of negative intergroup interactions in
+explaining intergroup relations
 \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster comparison of Median Absolute Deviation for all variables}
+  \label{fig:cluster_comparison_mad}
+  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_mad.pdf}
+\end{figure}
 
@@
 contextual factors that might contribute to the difficulties some might
-face. Thirdly, the feature-based approach is flexible and meaningful. We
-were able to use a wide range of time series features that have been
-central in the ESM literature and were able to use them directly to
-identify meaningful groups. For our empirical illustration we, among
-others for example, chose to focus on whether participants differed in
-their average well-being (i.e., \textit{median}), how much their
-well-being would vary over time (i.e, \textit{MAD}), and whether their
-well-being would on average increase or decrease over time (i.e.,
+face. In our illustration we, for example, found that participants in
+the deteriorating cluster (cluster 2) reported less need fulfilling
+interactions over time. Thirdly, the feature-based approach is flexible
+and meaningful. We were able to use a wide range of time series features
+that have been central in the ESM literature and were able to use them
+directly to identify meaningful groups. For our empirical illustration
+we, among others for example, chose to focus on whether participants
+differed in their average well-being (i.e., \textit{median}), how much
+their well-being would vary over time (i.e, \textit{MAD}), and whether
+their well-being would on average increase or decrease over time (i.e.,
 \textit{linear trend}). Alternatively, for others cyclical patterns

### 2023-07-29 sections/02InputVariables.Rmd
feedback lfb

@@
 As a result, the number of included variables is also on the higher end for psychological concepts and additionally allows us to showcase the efficiency benefits of the method and offers a reasonable use case for the optional feature reduction step.
+
+```{r}
+
+varNames %>%
+  filter(
+    pca != 0,
+    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
+  ) %>%
+  mutate(
+    contactSpecific = recode(
+      .$contactSpecific,
+      `-1` = "No Interaction Only",
+      `0` = "unspecific",
+      `1` = "Interaction Only"
+    ),
+    study1 = ifelse(!is.na(studyS1), 1, 0),
+    study2 = ifelse(!is.na(studyS2), 1, 0),
+    study3 = ifelse(!is.na(studyS3), 1, 0)
+  ) %>% 
+  select(varNam, question, varGroup, contactSpecific, study1, study2, study3) %>%
+  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
+  group_by(varGroup, varNam, question, contactSpecific, study1, study2, study3) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = c("varNam", "question", "varGroup", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
+  ungroup %>%
+  t %>%
+  as.data.frame %>%
+  janitor::row_to_names(1) %>%
+  replace(is.na(.), 0) %>%
+  tibble::rownames_to_column(., var = "variable") %>%
+  separate(variable, c("Variable", "Question", "Aspect", "study1", "study2", "study3"), " - ") %>%
+  mutate(across(c(study1, study2, study3), as.numeric)) %>%
+  select(-study1, -study2, -study3) %>%
+  arrange(`Interaction Only`, Variable) %>% 
+  mutate(ESM = 1) %>%
+  mutate_all(~recode(.x, `0`="", ` 1`="✔")) %>%
+  relocate(starts_with("ESM"), .after = "Aspect") %>%
+  # relocate(starts_with("study"), .after = last_col()) %>%
+  # rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
+  kbl(.,
+      escape = FALSE,
+      booktabs = T,
+      align = c("l", "l", "l", rep("c", ncol(.)-3)),
+      digits = 2,
+      format = "latex",
+      label = "var_selection",
+      caption = "Variable Selection") %>%
+  add_header_above(c(" " = 4, "Contact Specific" = 2)) %>%
+  kable_styling(latex_options = "scale_down") %>%
+  gsub("\\begin{table}", "\\begin{sidewaystable*}[!hbtp]\n\\centering", ., fixed = TRUE) %>%
+  gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
+  save_kable("tables/var_selection.tex")
+
+```
+
+\input{tables/var_selection}
 
@@
       "Int: Quality", 
-      "Core Need", 
-      "Core Need Due to Partner", 
+      "Int: Need Fulfillment", 
+      "Int: Need Fulfillment Partner", 
       "Attitude Partner",
-      "Daytime Core Need",
+      "Daytime Need Fulfillment",
       "Outgroup Attitude",
@@
   gsub("\\end{table}", "\\end{sidewaystable}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrWide.tex")
+  save_kable("tables/descrWide.tex")
 ```
@@
         "\\makecell{Int: \\\\ Quality}", 
-        "Core Need", 
-        "\\makecell{Core Need \\\\ Due to Partner}", 
+        "\\makecell{Int: \\\\ Need Fulfil.}", 
+        "\\makecell{Int: \\\\ Need Fulfil. Partner}", 
         "\\makecell{Attitude \\\\ Partner}",
@@
   gsub("\\end{table}", "\\end{sidewaystable*}", ., fixed = TRUE) %>%
-  save_kable("Tables/descrLong.tex")
+  save_kable("tables/descrLong.tex")
 ```

### 2023-07-29 sections/07ClusterEvaluationInterpretation.Rmd
feedback lfb

@@
 )
-```
-
-In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
-We find an adaptive group (cluster 1) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
-This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
-This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 2).
-This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
-This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
-At the same time, for members of this detrimental cluster (cluster 2) conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
-
-To identify these patterns, we first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
-We see that for some variables the features are generally stronger in separating the clusters.
-We, for example, see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
-Compare this to the '\textit{outgroup attitudes}' item where the differences between the clusters are much smaller for almost all features.
-We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables.
-For example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
-These two features stand in stark contrast to other features, such as the \textit{lag-1 auto correlations}, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}).
-This offers some information on which features were most important in understanding the two extracted groups.
-Taking these two perspectives together, we can also focus on individual features or variables in particular.
-We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series.
-At the same time, in terms of stability, both groups have virtually identical average \textit{MAC} statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
-There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
-In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
-We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a}.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-```
-
-In the second step, we look at the prototypical trajectories of the clusters.
-For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
-For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
-Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions and also consistently reported more voluntary and cooperative interactions but fewer accidental and involuntary interactions.
-The same cluster (cluster 1) also reported an increase in need-fulfilling interactions over the 30-day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster (cluster 2) showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This `deterioration' cluster (cluster 2) also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variability and instability on most of the other variables (also see \fgrref[A]{fig:clusterTs}).
-
-Finally, we can also assess the clusters across other individual difference variables \citep[e.g.,][]{monden2022}.
-This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
-To illustrate artifact checks, we added the number of ESM measurements into the comparison and find that the participants in the deterioration cluster (cluster 2) on average completed slightly more ESM surveys in general and reported on more intergroup interactions in particular (see $n$ in \fgrref[B]{fig:clusterFeatVar}).
-In our data exclusion procedures, we ensured that the general time frame and completion rates are similar for all participants and indeed the numbers in ESM measurements generally are largely similar (e.g., see $n$ for well-being and outgroup attitudes).
-However, the difference in the reported number of interactions might indicate either a clustering artifact or a meaningful difference.
-The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are more inflated due to the larger samples \citep{kogan2006}.
-At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions, but these interactions were less voluntary, less meaningful, and less positive.
-Such a finding is in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
-
-To further illustrate the utility of assessing out-of-feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
-Thus, both intensive longitudinal (e.g., the sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
-
-This cluster separation, then, has a number of empirical and practical applications.
-Firstly, the clusters are descriptive.
-With tens of variables, hundreds of participants, and thousands of measurements, singular descriptive statistics are often not able to capture the complex patterns that describe the data set.
-The feature-based clustering offers some direct insight into the complexity within the data set.
-In our empirical example, we, for example, see that participants are meaningfully distinguished by a combination of high (vs.
-low) central tendency, variability, and linear trend.
-Secondly, the clusters identify important groups.
-The adaptive and deteriorating groups offer starting points for empirical exploration as well as practical interventions.
-Researchers can start probing what exactly distinguishes the two groups further and generate new bottom-up hypotheses.
-Practitioners in the resettlement field can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
-Thirdly, the feature-based approach is flexible and meaningful.
-We were able to use a wide range of time series features that have been central in the ESM literature and were able to use them directly to identify meaningful groups.
-For our empirical illustration we, among others for example, chose to focus on whether participants differed in their average well-being (i.e., \textit{median}), how much their well-being would vary over time (i.e, \textit{MAD}), and whether their well-being would on average increase or decrease over time (i.e., \textit{linear trend}).
-Alternatively, for others cyclical patterns might be more important --- for example, whether well-being was higher on weekends.
-Importantly, in any case, we did not need to translate these dynamic features into probabilistic inference models (e.g., VAR models) to cluster the participants.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons over time}
-  \label{fig:clusterTs}
-  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
-  \caption*{Note: \\
-  Subplot (A) displays the variable cluster means at every measurement occasion.
-Subplot (B) shows the GAM spline for each cluster across the measurement occasions.
-The thinner lines present all individual time series.}
-\end{figure}
 ```
@@
 
+In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
+We find an adaptive group (cluster 1) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
+This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
+This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 2).
+This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
+At the same time, for members of this detrimental cluster (cluster 2) conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
+
+To identify these patterns, we first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
+We see that for some variables the features are generally stronger in separating the clusters.
+We, for example, see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
+Compare this to the '\textit{outgroup attitudes}' item where the differences between the clusters are much smaller for almost all features.
+We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
+While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables.
+For example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
+These two features stand in stark contrast to other features, such as the \textit{lag-1 auto correlations}, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}).
+This offers some information on which features were most important in understanding the two extracted groups.
+Taking these two perspectives together, we can also focus on individual features or variables in particular.
+We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series.
+At the same time, in terms of stability, both groups have virtually identical average \textit{MAC} statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
+There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
+In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
+We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a}.
+
+```{=tex}
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons based on Features and Variables}
+  \label{fig:clusterFeatVar}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
+  \caption*{Note: \\
+  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+\end{figure}
+```
+
+In the second step, we look at the prototypical trajectories of the clusters.
+For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
+For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
+Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions and also consistently reported more voluntary and cooperative interactions but fewer accidental and involuntary interactions.
+The same cluster (cluster 1) also reported an increase in need-fulfilling interactions over the 30-day period and an increase in interactions that were representative of the outgroup.
+Whereas the other cluster (cluster 2) showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
+This `deterioration' cluster (cluster 2) also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
+We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variability and instability on most of the other variables (also see \fgrref[A]{fig:clusterTs}).
+
+```{r}
+#| label: formal cluster comparison
+#| echo: false
+
+raw_feat_clust <- data.frame(ID = as.numeric(names(kmeansOut$cluster)), cluster = kmeansOut$cluster) %>%
+  merge(., z_data) %>%
+  merge(., validationFeature)
+
+concepts_and_feature <- str_match(colnames(raw_feat_clust), "(.*)_(.*)")
+concepts <- na.omit(unique(concepts_and_feature[, 2]))
+ts_features <- na.omit(unique(concepts_and_feature[, 3]))
+
+feat_t_test <- function(df, cluster_var, concept, ts_feature) {
+  column_name <- paste0(concept, "_", ts_feature)
+
+  group1 <- df[df[[cluster_var]] == unique(df[[cluster_var]])[1], column_name]
+  group2 <- df[df[[cluster_var]] == unique(df[[cluster_var]])[2], column_name]
+
+  t.test(group1, group2)
+}
+
+results <- list()
+for (feature in ts_features) {
+  for (concept in concepts) {
+    results[[feature]][[concept]] <- feat_t_test(raw_feat_clust, "cluster", concept, feature)
+  }
+}
+
+results$ar01$AttitudesDutch$conf.int
+
+results_df <- data.frame(concept = character(),
+                         feature = character(),
+                         difference = numeric(),
+                         df = numeric(),
+                         t_value = numeric(),
+                         p_value = numeric(),
+                         lwr = numeric(),
+                         upr = numeric(),
+                         stringsAsFactors = FALSE)
+
+for (feature in ts_features) {
+  for (concept in concepts) {
+    # Run t-test
+    test_result <- results[[feature]][[concept]]
+
+    # Extract required values from t-test result
+    difference <- test_result$estimate[1] - test_result$estimate[2]  # difference between group means
+    t_value <- test_result$statistic  # t-value
+    df <- test_result$parameter  # df
+    p_value <- test_result$p.value  # p-value
+    lwr <- test_result$conf.int[1] # lower CI
+    upr <- test_result$conf.int[2] # upper CI
+
+    # Add a row to the results dataframe
+    results_df <- rbind(results_df, data.frame(concept = concept,
+                                               feature = feature,
+                                               difference = difference,
+                                               t_value = t_value,
+                                               df = df,
+                                               p_value = p_value,
+                                               lwr = lwr,
+                                               upr = upr,
+                                               stringsAsFactors = FALSE))
+  }
+}
+
+results_df <- results_df %>%
+      mutate(star = ifelse(p_value < 0.001, "***", ifelse(p_value < 0.01, "**", ifelse(p_value < 0.05, "*", ""))))
+
+feature_plot_list <- list()
+
+library(forestplot)
+for (feature in unique(results_df$feature)) {
+  # Subset the data for the current feature
+  feature_data <- results_df[results_df$feature == feature, ]
+
+  # Create a matrix for the forestplot function
+  plot_matrix <- rbind(c("", "Difference", "Lower CI", "Upper CI"),
+                       cbind(feature_data$concept,
+                             feature_data$difference,
+                             feature_data$lwr,
+                             feature_data$upr))
+
+  # Create the forest plot
+  feature_plot_list[[feature]] <- forestplot::forestplot(labeltext = plot_matrix[, 1], 
+             mean = as.numeric(plot_matrix[, 2]),
+             lower = as.numeric(plot_matrix[, 3]),
+             upper = as.numeric(plot_matrix[, 4]),
+             xlab = "Difference",
+             title = paste0("Forest plot for ", feature))
+}
+
+concept_plot_list <- list()
+for (concept in unique(results_df$concept)) {
+  # Subset the data for the current feature
+  concept_data <- results_df[results_df$concept == concept, ]
+
+  # Create a matrix for the forestplot function
+  plot_matrix <- rbind(c("", "Difference", "Lower CI", "Upper CI"),
+                       cbind(feature_data$concept,
+                             feature_data$difference,
+                             feature_data$lwr,
+                             feature_data$upr))
+
+  # Create the forest plot
+  concept_plot_list[[concept]] <- forestplot::forestplot(labeltext = plot_matrix[, 1], 
+             mean = as.numeric(plot_matrix[, 2]),
+             lower = as.numeric(plot_matrix[, 3]),
+             upper = as.numeric(plot_matrix[, 4]),
+             xlab = "Difference",
+             title = paste0("Forest plot for ", concept))
+}
+
+dev.off()
+pdf("figures/feature_comparison_mad.pdf")
+feature_plot_list$mad
+dev.off()
+png("figures/feature_comparison_mad.png")
+feature_plot_list$mad
+dev.off()
+
+feat_ttest_apa <- function(results_df, concept_in, feature_in) {
+  # concept_in <- "InteractionContextvoluntary"
+  # feature_in  <- "median"
+
+  diff <- results_df %>% 
+    filter(concept == concept_in, feature == feature_in) %>% 
+    select(difference) %>% 
+    pull %>% 
+    format(round(., 2), nsmall=2)
+  t <- results_df %>% 
+    filter(concept == concept_in, feature == feature_in) %>% 
+    select(t_value) %>% 
+    pull %>% 
+    round(., 2) %>% 
+    format(., nsmall=2)
+  df <- results_df %>% 
+    filter(concept == concept_in, feature == feature_in) %>% 
+    select(df) %>% 
+    pull %>% 
+    round(., 2) %>%
+    format(., nsmall=2)
+  p <- results_df %>% 
+    filter(concept == concept_in, feature == feature_in) %>% 
+    select(p_value) %>% 
+    pull
+  p <- ifelse(p < 0.001, "< .001", 
+              paste0("= ", format(round(p, 3), nsmall=3)))
+  lwr <- results_df %>% 
+    filter(concept == concept_in, feature == feature_in) %>% 
+    select(lwr) %>% 
+    pull %>% 
+    round(., 2) %>%
+    format(., nsmall=2)
+  upr <- results_df %>% 
+    filter(concept == concept_in, feature == feature_in) %>% 
+    select(upr) %>% 
+    pull %>% 
+    round(., 2) %>%
+    format(., nsmall=2)
+
+  paste("$difference$ = ", diff,
+        ", $t$(", df, ") = ", t,
+        ", $p$ ", p , 
+        ", $95%CI [", lwr,
+        ", ", upr, "]",
+        sep = "")
+}
+```
+
+Finally, we can also assess the clusters across other individual difference variables \citep[e.g.,][]{monden2022}.
+This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.
+To illustrate artifact checks, we added the number of ESM measurements into the comparison and find that the participants in the deterioration cluster (cluster 2) on average completed slightly more ESM surveys in general and reported on more intergroup interactions in particular (see $n$ in \fgrref[B]{fig:clusterFeatVar}).
+In our data exclusion procedures, we ensured that the general time frame and completion rates are similar for all participants and indeed the numbers in ESM measurements generally are largely similar (e.g., see $n$ for well-being and outgroup attitudes).
+However, the difference in the reported number of interactions might indicate either a clustering artifact or a meaningful difference.
+The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are more inflated due to the larger samples \citep{kogan2006}.
+In our case this seems less likely because one out of four variables did not differ in terms of the MAD (see \fgrref{fig:cluster_comparison_mad}).
+At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions (`r feat_ttest_apa(results_df, "InteractionContextvoluntary", "n")`), but these interactions were less voluntary (`r feat_ttest_apa(results_df, "InteractionContextvoluntary", "median")`), less meaningful (`r feat_ttest_apa(results_df, "qualityMeaning", "median")`), and less positive (`r feat_ttest_apa(results_df, "qualityOverall", "median")`).
+Such a finding is in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+
+```{=tex}
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster comparison of Median Absolute Deviation for all variables}
+  \label{fig:cluster_comparison_mad}
+  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_mad.pdf}
+\end{figure}
+```
+
+To further illustrate the utility of assessing out-of-feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
+When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
+Thus, both intensive longitudinal (e.g., the sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.
+
+This cluster separation, then, has a number of empirical and practical applications.
+Firstly, the clusters are descriptive.
+With tens of variables, hundreds of participants, and thousands of measurements, singular descriptive statistics are often not able to capture the complex patterns that describe the data set.
+The feature-based clustering offers some direct insight into the complexity within the data set.
+In our empirical example, we, for example, see that participants are meaningfully distinguished by a combination of high (vs.
+low) central tendency, variability, and linear trend.
+Secondly, the clusters identify important groups.
+The adaptive and deteriorating groups offer starting points for empirical exploration as well as practical interventions.
+Researchers can start probing what exactly distinguishes the two groups further and generate new bottom-up hypotheses.
+Practitioners in the resettlement field can use the group separation to identify individuals in need of assistance and can explore contextual factors that might contribute to the difficulties some might face.
+In our illustration we, for example, found that participants in the deteriorating cluster (cluster 2) reported less need fulfilling interactions over time.
+Thirdly, the feature-based approach is flexible and meaningful.
+We were able to use a wide range of time series features that have been central in the ESM literature and were able to use them directly to identify meaningful groups.
+For our empirical illustration we, among others for example, chose to focus on whether participants differed in their average well-being (i.e., \textit{median}), how much their well-being would vary over time (i.e, \textit{MAD}), and whether their well-being would on average increase or decrease over time (i.e., \textit{linear trend}).
+Alternatively, for others cyclical patterns might be more important --- for example, whether well-being was higher on weekends.
+Importantly, in any case, we did not need to translate these dynamic features into probabilistic inference models (e.g., VAR models) to cluster the participants.
+
+```{=tex}
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons over time}
+  \label{fig:clusterTs}
+  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
+  \caption*{Note: \\
+  Subplot (A) displays the variable cluster means at every measurement occasion.
+Subplot (B) shows the GAM spline for each cluster across the measurement occasions.
+The thinner lines present all individual time series.}
+\end{figure}
+```
+
 ```{r}

### 2023-07-29 tables/var_selection.tex
feedback lfb

\begin{sidewaystable*}[!hbtp]
\centering

\caption{\label{tab:var_selection}Variable Selection}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllccc}
\toprule
\multicolumn{4}{c}{ } & \multicolumn{2}{c}{Contact Specific} \\
\cmidrule(l{3pt}r{3pt}){5-6}
Variable & Question & Aspect & ESM & unspecific & Interaction Only\\
\midrule
AttitudesPartner & At the moment, how favorably do you feel towards -NAME- & cognition & ✔ &  & ✔\\
InteractionContextAccidental & The interaction with -NAME- was accidental. & cognition & ✔ &  & ✔\\
InteractionContextCooperative & The interaction with -NAME- was cooperative. & cognition & ✔ &  & ✔\\
InteractionContextRepresentativeNL & The interaction with -NAME- was representative of the Dutch. & cognition & ✔ &  & ✔\\
InteractionContextvoluntary & The interaction with -NAME- was voluntary. & cognition & ✔ &  & ✔\\
\addlinespace
KeyNeedDueToPartner & -NAME- helped fulfill your goal (-GOAL-) & needs & ✔ &  & ✔\\
qualityMeaning & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & ✔ &  & ✔\\
qualityOverall & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & ✔ &  & ✔\\
AttitudesDutch & At the moment, how favorably do you feel towards the Dutch. & cognition & ✔ & ✔ & \\
DaytimeNeedFulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & ✔ & ✔ & \\
\addlinespace
KeyNeedFulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & ✔ & ✔ & \\
exWB & How do you feel right now? very sad --- very happy & emotion & ✔ & ✔ & \\
\bottomrule
\end{tabular}}
\end{sidewaystable*}

### 2023-07-29 scripts/functions/tsFeatureExtractor.R

@@ -1,5 +1,29 @@
 
-
+#' Feature Extraction Function
+#'
+#' This function extracts features from a given dataset, grouping it by an identifier variable, and summarizing it by a set of measures 
+#' for each item in the given list of items. It generates univariate features such as mean, median, standard deviation, 
+#' mean absolute deviation, root mean square of successive differences, and mean absolute change.
+#' It also generates multivariate and time-based features, applying linear and non-linear trends using generalized additive models (GAMs),
+#' and calculates auto-regressive components and their confidence intervals.
+#' It standardizes the features, and outputs a list containing raw and standardized features, as well as the GAM models.
+#'
+#' @param data A data frame from which to extract features.
+#' @param items A character vector of the names of the variables for which to compute features.
+#' @param pid A character string specifying the name of the person identifier variable.
+#' @param tid A character string specifying the name of the time identifier variable.
+#'
+#' @return A list containing:
+#' features: A data frame of raw features.
+#' featuresZ: A data frame of standardized features.
+#' featuresZMat: A data frame of standardized features with person identifiers as row names.
+#' gam: A list of fitted GAM models for each person.
+#'
+#' @examples
+#' \dontrun{
+#' featureExtractor(df, c("item1", "item2"), "ID", "time")
+#' }
+#' @export
 featureExtractor <- function(data, items, pid, tid) {
   # for testing:
   # data <- featData
@@ -7,34 +31,34 @@ featureExtractor <- function(data, items, pid, tid) {
   # tid <- "TIDnum"
   # items <- varNamS123PCA[!varNamS123PCA %in% idVars]
   
+  # Change the names of the identifier columns to "ID" and "TIDnum"
   names(data)[names(data) == pid] <- "ID"
   names(data)[names(data) == tid] <- "TIDnum"
   
-  # univariate features
+  # Generate univariate features
   featUnivar <- data %>%
     group_by(ID) %>%
     summarise(across(
       any_of(items),
       list(
-        # central tendency
         mean = ~ mean(.x, na.rm = TRUE),
         median = ~ stats::median(.x, na.rm = TRUE),
-        # variance
[…]

### 2023-07-29 sections/02InputVariables.Rmd
checkmark

@@
   mutate(ESM = 1) %>%
-  mutate_all(~recode(.x, `0`="", ` 1`="✔")) %>%
+  mutate_all(~recode(.x, `0`="", ` 1`="\\Checkmark")) %>%
   relocate(starts_with("ESM"), .after = "Aspect") %>%

### 2023-07-29 tables/var_selection.tex
checkmark

@@
 \midrule
-AttitudesPartner & At the moment, how favorably do you feel towards -NAME- & cognition & ✔ &  & ✔\\
-InteractionContextAccidental & The interaction with -NAME- was accidental. & cognition & ✔ &  & ✔\\
-InteractionContextCooperative & The interaction with -NAME- was cooperative. & cognition & ✔ &  & ✔\\
-InteractionContextRepresentativeNL & The interaction with -NAME- was representative of the Dutch. & cognition & ✔ &  & ✔\\
-InteractionContextvoluntary & The interaction with -NAME- was voluntary. & cognition & ✔ &  & ✔\\
+AttitudesPartner & At the moment, how favorably do you feel towards -NAME- & cognition & \Checkmark &  & \Checkmark\\
+InteractionContextAccidental & The interaction with -NAME- was accidental. & cognition & \Checkmark &  & \Checkmark\\
+InteractionContextCooperative & The interaction with -NAME- was cooperative. & cognition & \Checkmark &  & \Checkmark\\
+InteractionContextRepresentativeNL & The interaction with -NAME- was representative of the Dutch. & cognition & \Checkmark &  & \Checkmark\\
+InteractionContextvoluntary & The interaction with -NAME- was voluntary. & cognition & \Checkmark &  & \Checkmark\\
 \addlinespace
-KeyNeedDueToPartner & -NAME- helped fulfill your goal (-GOAL-) & needs & ✔ &  & ✔\\
-qualityMeaning & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & ✔ &  & ✔\\
-qualityOverall & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & ✔ &  & ✔\\
-AttitudesDutch & At the moment, how favorably do you feel towards the Dutch. & cognition & ✔ & ✔ & \\
-DaytimeNeedFulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & ✔ & ✔ & \\
+KeyNeedDueToPartner & -NAME- helped fulfill your goal (-GOAL-) & needs & \Checkmark &  & \Checkmark\\
+qualityMeaning & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & \Checkmark &  & \Checkmark\\
+qualityOverall & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & \Checkmark &  & \Checkmark\\
+AttitudesDutch & At the moment, how favorably do you feel towards the Dutch. & cognition & \Checkmark & \Checkmark & \\
+DaytimeNeedFulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\
 \addlinespace
-KeyNeedFulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & ✔ & ✔ & \\
-exWB & How do you feel right now? very sad --- very happy & emotion & ✔ & ✔ & \\
+KeyNeedFulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\
+exWB & How do you feel right now? very sad --- very happy & emotion & \Checkmark & \Checkmark & \\
 \bottomrule

### 2023-07-29 tables/var_selection.tex
format error

@@
 InteractionContextvoluntary & The interaction with -NAME- was voluntary. & cognition & \Checkmark &  & \Checkmark\\
-\addlinespace
 KeyNeedDueToPartner & -NAME- helped fulfill your goal (-GOAL-) & needs & \Checkmark &  & \Checkmark\\
@@
 DaytimeNeedFulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\
-\addlinespace
 KeyNeedFulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\

### 2023-07-29 sections/02InputVariables.Rmd
format error

--- 2023-07-29 sections/02InputVariables.Rmd
@@
       format = "latex", 
+      linesep = "",
       label = "var_selection",

### 2023-07-29 sections/03FeatureExtraction.Rmd
writing adj.

@@
 
-<!--Effective degrees of freedom for individual smooth terms are obtained by summing the Fii values corresponding to the coefficients βi of the smooth term (e.g., if the coefficients of a particular smooth are β5, .
+```{r}
+#| include: False
+
+# Effective degrees of freedom for individual smooth terms are obtained by summing the Fii values corresponding to the coefficients βi of the smooth term (e.g., if the coefficients of a particular smooth are β5, .
 .
 .
-, β11 then the smooth’s EDF is $\sum_{i=5}^{11} = F_{ii}).--!>
-
-<!-- An alternative approach to understanding the edf is to approximate the number of parametric and non-parametric (or smooth) terms $p$ in the penalized GAM.
+, β11 then the smooth’s EDF is $\sum_{i=5}^{11} = F_{ii}).
+# 
+# An alternative approach to understanding the edf is to approximate the number of parametric and non-parametric (or smooth) terms $p$ in the penalized GAM.
 Both formulas are measuring the same concept (the model's complexity or flexibility) and both account for the penalization introduced to control overfitting in GAMs \citep{hastie1986}.
-
-```{=tex}
-\begin{equation} \label{eq:edf}
-  edf \approx \sum p
-\end{equation}
+# 
+# ```{=tex}
+# \begin{equation} \label{eq:edf}
+#   edf \approx \sum p
+# \end{equation}
+# ```
 ```
---!>
 

### 2023-07-29 latex-results/03FeatureExtraction.tex
feedback lfb

@@
 
-\textless!--Effective degrees of freedom for individual smooth terms are
-obtained by summing the Fii values corresponding to the coefficients βi
-of the smooth term (e.g., if the coefficients of a particular smooth are
-β5, . . . , β11 then the smooth's EDF is \$\sum\emph{\{i=5\}\^{}\{11\} =
-F}\{ii\}).--!\textgreater{}
-
-\textless!-- An alternative approach to understanding the edf is to
-approximate the number of parametric and non-parametric (or smooth)
-terms \(p\) in the penalized GAM. Both formulas are measuring the same
-concept (the model's complexity or flexibility) and both account for the
-penalization introduced to control overfitting in GAMs
-\citep{hastie1986}.
-
-\begin{equation} \label{eq:edf}
-  edf \approx \sum p
-\end{equation}
-
---!\textgreater{}
-
 Beyond our main features of interest, we also extracted the

### 2023-07-29 latex-results/07ClusterEvaluationInterpretation.tex
feedback lfb

@@
 cluster (cluster 2) on average reported more outgroup interactions
-(\(difference\) = 1.028, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
-\(95%CI [0.76, 1.30]), but these interactions were less voluntary (
-\)difference\$ = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
-\(95%CI [-1.31, -0.77]), less meaningful (
-\)difference\$ = -1.003, \(t\)(136.40) = -7.16, \(p\) \textless{} .001,
-\(95%CI [-1.28, -0.73]), and less positive (
-\)difference\$ = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{} .001,
-\$95\%CI {[}-1.61, -1.15{]}). Such a finding is in line with past
-research highlighting the role of negative intergroup interactions in
-explaining intergroup relations
-\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+(\(difference\) = 1.03, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
+\(95%CI
+\) {[}0.76, 1.30{]}), but these interactions were less voluntary
+(\(difference\) = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
+\(95%CI
+\) {[}-1.31, -0.77{]}), less meaningful (\(difference\) = -1.00,
+\(t\)(136.40) = -7.16, \(p\) \textless{} .001, \(95%CI
+\) {[}-1.28, -0.73{]}), and less positive (\(difference\) = -1.38,
+\(t\)(152.31) = -11.94, \(p\) \textless{} .001, \(95%CI
+\) {[}-1.61, -1.15{]}). Such a finding is in line with past research
+highlighting the role of negative intergroup interactions in explaining
+intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
 

### 2023-07-29 sections/07ClusterEvaluationInterpretation.Rmd
feedback lfb

@@
     pull %>% 
-    format(round(., 2), nsmall=2)
+    round(., 2) %>% 
+    format(., nsmall=2)
   t <- results_df %>% 
@@
         ", $p$ ", p , 
-        ", $95%CI [", lwr,
+        ", $95%CI$ [", lwr,
         ", ", upr, "]",

### 2023-07-29 latex-results/07ClusterEvaluationInterpretation.tex
feedback lfb

--- 2023-07-29 latex-results/07ClusterEvaluationInterpretation.tex
@@
 cluster (cluster 2) on average reported more outgroup interactions
-(\(difference\) = 1.028, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
-\(95%CI [0.76, 1.30]), but these interactions were less voluntary (
-\)difference\$ = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
-\(95%CI [-1.31, -0.77]), less meaningful (
-\)difference\$ = -1.003, \(t\)(136.40) = -7.16, \(p\) \textless{} .001,
-\(95%CI [-1.28, -0.73]), and less positive (
-\)difference\$ = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{} .001,
-\$95\%CI {[}-1.61, -1.15{]}). Such a finding is in line with past
-research highlighting the role of negative intergroup interactions in
-explaining intergroup relations
+(\textit{difference} = 1.03, \(t\)(150.83) = 7.50, \(p\) \textless{}
+.001, \textit{95\%CI} {[}0.76, 1.30{]}), but these interactions were
+less voluntary (\textit{difference} = -1.04, \(t\)(108.89) = -7.71,
+\(p\) \textless{} .001, \textit{95\%CI} {[}-1.31, -0.77{]}), less
+meaningful (\textit{difference} = -1.00, \(t\)(136.40) = -7.16, \(p\)
+\textless{} .001, \textit{95\%CI} {[}-1.28, -0.73{]}), and less positive
+(\textit{difference} = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{}
+.001, \textit{95\%CI} {[}-1.61, -1.15{]}). Such a finding is in line
+with past research highlighting the role of negative intergroup
+interactions in explaining intergroup relations
 \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.

### 2023-07-29 sections/07ClusterEvaluationInterpretation.Rmd
feedback lfb

@@
              upper = as.numeric(plot_matrix[, 4]),
-             xlab = "Difference",
-             title = paste0("Forest plot for ", concept))
+             xlab = "Difference")
 }
@@
 
-  paste("$difference$ = ", diff,
+  paste("\\textit{difference} = ", diff,
         ", $t$(", df, ") = ", t,
         ", $p$ ", p , 
-        ", $95%CI$ [", lwr,
+        ", \\textit{95\\%CI} [", lwr,
         ", ", upr, "]",

### 2023-08-02 sections/02InputVariables.Rmd
internal feedback lfb

--- 2023-07-29 sections/02InputVariables.Rmd
@@
 
-For our illustration, we include 12 variables that were measured as part of the ESM surveys in all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group (see \tblref{tab:descrLong} for an overview).
+For our illustration, we include 12 variables that were measured as part of the ESM surveys in all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group (see \tblref{tab:var_selection} for an overview).
 We chose these aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
@@
   add_header_above(c(" " = 4, "Contact Specific" = 2)) %>%
+  footnote(
+    general = c('All items used a continuous slider and were rescaled to a range of 0--100.')
+  ) %>%
   kable_styling(latex_options = "scale_down") %>%
@@
 In our illustration data set, the studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`).
-This was likely due to the option to continue participation without compensation in the latter two studies.
+This was likely due to the option to continue participation without compensation in the latter study.
 To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.

### 2023-08-02 sections/07ClusterEvaluationInterpretation.Rmd
internal feedback lfb

@@
 However, the difference in the reported number of interactions might indicate either a clustering artifact or a meaningful difference.
-The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are more inflated due to the larger samples \citep{kogan2006}.
-In our case this seems less likely because one out of four variables did not differ in terms of the MAD (see \fgrref{fig:cluster_comparison_mad}).
+The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are substantially larger due to the larger samples \citep[e.g., restriction of range in the smaller sample][]{kogan2006}.
+In our case, this seems less likely because one out of four variables did not differ in terms of the MAD (i.e., our selected measurement of the time series variance; see \fgrref{fig:cluster_comparison_mad} for an illustration).
 At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions (`r feat_ttest_apa(results_df, "InteractionContextvoluntary", "n")`), but these interactions were less voluntary (`r feat_ttest_apa(results_df, "InteractionContextvoluntary", "median")`), less meaningful (`r feat_ttest_apa(results_df, "qualityMeaning", "median")`), and less positive (`r feat_ttest_apa(results_df, "qualityOverall", "median")`).
-Such a finding is in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+Thus, while more research is needed for a conclusive test, our data seems to suggest that the differences in reported interactions are a meaningful difference between the clusters.
+Such a finding would also be in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
 
@@
 To further illustrate the utility of assessing out-of-feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\fgrref[B]{fig:clusterFeatVar}).
+When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\hl{T-TEST HERE!!}; \fgrref[B]{fig:clusterFeatVar}).
 Thus, both intensive longitudinal (e.g., the sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.

### 2023-08-02 latex-results/07ClusterEvaluationInterpretation.tex
internal feedback lfb

--- 2023-07-29 latex-results/07ClusterEvaluationInterpretation.tex
@@
 difference. The higher average number of interactions in cluster 2
-could, for example, indicate a clustering artifact if variances are more
-inflated due to the larger samples \citep{kogan2006}. In our case this
-seems less likely because one out of four variables did not differ in
-terms of the MAD (see \fgrref{fig:cluster_comparison_mad}). At the same
-time, however, the difference in the number of experienced interactions
-might also indicate a meaningful difference, where the deteriorating
-cluster (cluster 2) on average reported more outgroup interactions
-(\(difference\) = 1.028, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
-\(95%CI [0.76, 1.30]), but these interactions were less voluntary (
-\)difference\$ = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
-\(95%CI [-1.31, -0.77]), less meaningful (
-\)difference\$ = -1.003, \(t\)(136.40) = -7.16, \(p\) \textless{} .001,
-\(95%CI [-1.28, -0.73]), and less positive (
-\)difference\$ = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{} .001,
-\$95\%CI {[}-1.61, -1.15{]}). Such a finding is in line with past
-research highlighting the role of negative intergroup interactions in
-explaining intergroup relations
+could, for example, indicate a clustering artifact if variances are
+substantially larger due to the larger samples
+\citep[e.g., restriction of range in the smaller sample][]{kogan2006}.
+In our case, this seems less likely because one out of four variables
+did not differ in terms of the MAD (i.e., our selected measurement of
+the time series variance; see \fgrref{fig:cluster_comparison_mad} for an
+illustration). At the same time, however, the difference in the number
+of experienced interactions might also indicate a meaningful difference,
+where the deteriorating cluster (cluster 2) on average reported more
+outgroup interactions (\textit{difference} = 1.03, \(t\)(150.83) = 7.50,
+\(p\) \textless{} .001, \textit{95\%CI} {[}0.76, 1.30{]}), but these
+interactions were less voluntary (\textit{difference} = -1.04,
+\(t\)(108.89) = -7.71, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.31,
+-0.77{]}), less meaningful (\textit{difference} = -1.00, \(t\)(136.40) =
+-7.16, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.28, -0.73{]}), and
+less positive (\textit{difference} = -1.38, \(t\)(152.31) = -11.94,
+\(p\) \textless{} .001, \textit{95\%CI} {[}-1.61, -1.15{]}). Thus, while
+more research is needed for a conclusive test, our data seems to suggest
+that the differences in reported interactions are a meaningful
+difference between the clusters. Such a finding would also be in line
+with past research highlighting the role of negative intergroup
+interactions in explaining intergroup relations
 \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
@@
 (cluster 2) reported substantially higher levels of everyday
-discrimination (\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive
-longitudinal (e.g., the sum of specific ESM measurements) and
-cross-sectional variables (e.g., general discrimination differences)
-that were not included in the original clustering step can be used to
-explore and understand the cluster differences in more detail.
+discrimination (\textit{difference} = 0.40, \(t\)(151.71) = 2.56, \(p\)
+= 0.011, \textit{95\%CI} {[}0.09, 0.71{]};
+\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
+(e.g., the sum of specific ESM measurements) and cross-sectional
+variables (e.g., general discrimination differences) that were not
+included in the original clustering step can be used to explore and
+understand the cluster differences in more detail.
 

### 2023-08-02 sections/07ClusterEvaluationInterpretation.Rmd
internal feedback lfb

@@
 
-results$ar01$AttitudesDutch$conf.int
-
 results_df <- data.frame(concept = character(),
@@
 results_df <- results_df %>%
-      mutate(star = ifelse(p_value < 0.001, "***", ifelse(p_value < 0.01, "**", ifelse(p_value < 0.05, "*", ""))))
+      mutate(
+        star = ifelse(p_value < 0.001, "***", ifelse(p_value < 0.01, "**", ifelse(p_value < 0.05, "*", "")))
+      ) %>% 
+ left_join(
+   variableLab,
+   by = c('concept' = 'variable')
+ )
 
 feature_plot_list <- list()
-
-library(forestplot)
 for (feature in unique(results_df$feature)) {
@@
 
-  # Create a matrix for the forestplot function
-  plot_matrix <- rbind(c("", "Difference", "Lower CI", "Upper CI"),
-                       cbind(feature_data$concept,
-                             feature_data$difference,
-                             feature_data$lwr,
-                             feature_data$upr))
+  df_plot <- data.frame(label = feature_data$label,
+                        difference = feature_data$difference,
+                        lwr = feature_data$lwr,
+                        upr = feature_data$upr)
 
-  # Create the forest plot
-  feature_plot_list[[feature]] <- forestplot::forestplot(labeltext = plot_matrix[, 1], 
-             mean = as.numeric(plot_matrix[, 2]),
-             lower = as.numeric(plot_matrix[, 3]),
-             upper = as.numeric(plot_matrix[, 4]),
-             xlab = "Difference",
-             title = paste0("Forest plot for ", feature))
+  feature_plot_list[[feature]] <- ggplot(df_plot, aes(y = label)) +
+    geom_point(aes(x = difference), shape=15, size=2) +
+    geom_linerange(aes(xmin = lwr, xmax = upr)) +
+    geom_vline(xintercept = 0, linetype="dashed") +
+    scale_x_continuous(expand = expansion(mult = 0)) +
+    theme_Publication() + 
+    theme(panel.grid.major = element_blank(), 
+          panel.grid.minor = element_blank(), 
+          axis.text.y = element_text(hjust=0), 
+          axis.line.y = element_blank(),
+          axis.ticks.y = element_blank(),
+          axis.title.y = element_blank())
 }
 
-concept_plot_list <- list()
-for (concept in unique(results_df$concept)) {
-  # Subset the data for the current feature
-  concept_data <- results_df[results_df$concept == concept, ]
-
-  # Create a matrix for the forestplot function
-  plot_matrix <- rbind(c("", "Difference", "Lower CI", "Upper CI"),
-                       cbind(feature_data$concept,
-                             feature_data$difference,
-                             feature_data$lwr,
-                             feature_data$upr))
-
-  # Create the forest plot
-  concept_plot_list[[concept]] <- forestplot::forestplot(labeltext = plot_matrix[, 1], 
-             mean = as.numeric(plot_matrix[, 2]),
-             lower = as.numeric(plot_matrix[, 3]),
-             upper = as.numeric(plot_matrix[, 4]),
-             xlab = "Difference")
+# Check if any graphics devices are open
+if (length(dev.list()) > 0) {
+  # Close the graphics device
+  dev.off()
 }
-
-dev.off()
 pdf("figures/feature_comparison_mad.pdf")
@@
 To further illustrate the utility of assessing out-of-feature individual differences, we also compare the two samples in terms of the participants' self-reported discrimination experiences in the Netherlands (measured during the post-measurement).
-When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (\hl{T-TEST HERE!!}; \fgrref[B]{fig:clusterFeatVar}).
+When looking at the group comparison, we find that participants in the deteriorating cluster (cluster 2) reported substantially higher levels of everyday discrimination (`r feat_ttest_apa(results_df, "exWB", "EvDayDiscr.post.z")`; \fgrref[B]{fig:clusterFeatVar}).
 Thus, both intensive longitudinal (e.g., the sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be used to explore and understand the cluster differences in more detail.

### 2023-08-02 sections/02InputVariables.Rmd
tbl labs adj

@@
 
+variableLab <- c(
+  "AttitudesDutch" = "Outgroup Attitude",
+  "AttitudesPartner" = "Int: Partner Attitude",
+  "DaytimeNeedFulfillment" = "Need Fulfillment",
+  "exWB" = "Well-Being",
+  "InteractionContextAccidental" = "Int: Accidental",
+  "InteractionContextCooperative" = "Int: Cooperative",
+  "InteractionContextRepresentativeNL" = "Int: Representative",
+  "InteractionContextvoluntary" = "Int: Voluntary",
+  "KeyNeedDueToPartner" = "Int: Need Fulfillment Partner",
+  "KeyNeedFulfillment" = "Int: Need Fulfillment",
+  "qualityMeaning" = "Int: Meaningful",
+  "qualityOverall" = "Int: Quality"
+) %>% enframe(., name = "varNam", value = "label")
+
 varNames %>%
@@
     varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
+  ) %>%
+  left_join(
+    variableLab,
+    by = 'varNam'
   ) %>% 
@@
   ) %>% 
-  select(varNam, question, varGroup, contactSpecific, study1, study2, study3) %>%
+  select(label, question, varGroup, contactSpecific, study1, study2, study3) %>%
   filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, question, contactSpecific, study1, study2, study3) %>%
+  group_by(varGroup, label, question, contactSpecific, study1, study2, study3) %>%
   summarise(n = n()) %>%
-  pivot_wider(names_from = c("varNam", "question", "varGroup", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
+  pivot_wider(names_from = c("label", "question", "varGroup", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
   ungroup %>%

### 2023-08-02 tables/var_selection.tex
internal feedback lfb
tbl labs adj

--- 2023-07-29 tables/var_selection.tex
@@
 \cmidrule(l{3pt}r{3pt}){5-6}
-Variable & Question & Aspect & ESM & unspecific & Interaction Only\\
+Variable & Question & Aspect & ESM & Interaction Only & unspecific\\
 \midrule
-AttitudesPartner & At the moment, how favorably do you feel towards -NAME- & cognition & ✔ &  & ✔\\
-InteractionContextAccidental & The interaction with -NAME- was accidental. & cognition & ✔ &  & ✔\\
-InteractionContextCooperative & The interaction with -NAME- was cooperative. & cognition & ✔ &  & ✔\\
-InteractionContextRepresentativeNL & The interaction with -NAME- was representative of the Dutch. & cognition & ✔ &  & ✔\\
-InteractionContextvoluntary & The interaction with -NAME- was voluntary. & cognition & ✔ &  & ✔\\
-\addlinespace
-KeyNeedDueToPartner & -NAME- helped fulfill your goal (-GOAL-) & needs & ✔ &  & ✔\\
-qualityMeaning & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & ✔ &  & ✔\\
-qualityOverall & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & ✔ &  & ✔\\
-AttitudesDutch & At the moment, how favorably do you feel towards the Dutch. & cognition & ✔ & ✔ & \\
-DaytimeNeedFulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & ✔ & ✔ & \\
-\addlinespace
-KeyNeedFulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & ✔ & ✔ & \\
-exWB & How do you feel right now? very sad --- very happy & emotion & ✔ & ✔ & \\
+Int: Accidental & The interaction with -NAME- was accidental. & cognition & \Checkmark & \Checkmark & \\
+Int: Cooperative & The interaction with -NAME- was cooperative. & cognition & \Checkmark & \Checkmark & \\
+Int: Meaningful & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & \Checkmark & \Checkmark & \\
+Int: Need Fulfillment Partner & -NAME- helped fulfill your goal (-GOAL-) & needs & \Checkmark & \Checkmark & \\
+Int: Partner Attitude & At the moment, how favorably do you feel towards -NAME- & cognition & \Checkmark & \Checkmark & \\
+Int: Quality & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & \Checkmark & \Checkmark & \\
+Int: Representative & The interaction with -NAME- was representative of the Dutch. & cognition & \Checkmark & \Checkmark & \\
+Int: Voluntary & The interaction with -NAME- was voluntary. & cognition & \Checkmark & \Checkmark & \\
+Int: Need Fulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark &  & \Checkmark\\
+Need Fulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & \Checkmark &  & \Checkmark\\
+Outgroup Attitude & At the moment, how favorably do you feel towards the Dutch. & cognition & \Checkmark &  & \Checkmark\\
+Well-Being & How do you feel right now? very sad --- very happy & emotion & \Checkmark &  & \Checkmark\\
 \bottomrule
+\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{6}{l}{\rule{0pt}{1em}All items used a continuous slider and were rescaled to a range of 0--100.}\\
 \end{tabular}}

### 2023-08-07 manuscript.tex
typos

@@
 \usepackage{tabu}
+\usepackage{bbding}
 
@@
    \addORCIDlink{* Jannis Kreienkamp}{0000-0002-1831-5604}\\
-   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}\\
+   \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
+   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}\\
    \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
-   \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
    \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
-   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}
+   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}
 
@@
 Conceptually, the included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series.
-One of the advantages of feature-based clustering is that it is inherently adept to accomodating multi-variate concepts --- a common aim in ESM research.
+One of the advantages of feature-based clustering is that it is inherently adept at accommodating multi-variate concepts --- a common aim in ESM research.
 There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, many health developments are captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}, and migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}.

### 2023-08-07 manuscript.tex
typos

@@
 The authors received no specific funding for this work.
-Materials and  software is available at \url{https://janniscodes.github.io/migration-trajectories/}  \citep{KreienkampTBD}.
-Protocols, materials, data, and code are available at \url{https://osf.io/TBA} \citep{KreienkampTBD}.
-The preregistration of our analysis can be accessed as part of our Open Science Framework repository \citep{KreienkampTBD}.
+Materials and  software are available at \url{https://janniscodes.github.io/migration-trajectories/}  \citep{KreienkampTBD}.
+Materials, data, and code are available at \url{https://osf.io/TBA} \citep{KreienkampTBD}.
 

### 2023-08-07 Supplemental-Material-A-Annotated-Analysis.Rmd
var info updated

@@
 
+### Data Export for OSF
+
+```{r}
+#| label: export-osf-datae
+
+export_vars <- c(
+  "PID",
+  "TIDnum",
+  "TID",
+  "created",
+  "ended",
+  "KeyNeedFulfillment",
+  "KeyNeedDueToPartner",
+  "InteractionContextAccidental",
+  "InteractionContextvoluntary",
+  "InteractionContextCooperative",
+  "InteractionContextRepresentativeNL",
+  "qualityOverall",
+  "qualityMeaning",
+  "DaytimeNeedFulfillment",
+  "AttitudesDutch",
+  "AttitudesPartner",
+  "exWB",
+  "EvDayDiscr01.post",
+  "EvDayDiscr02.post",
+  "EvDayDiscr03.post",
+  "EvDayDiscr04.post",
+  "EvDayDiscr05.post",
+  "EvDayDiscr06.post",
+  "EvDayDiscr07.post",
+  "EvDayDiscr08.post",
+  "EvDayDiscr09.post"
+)
+
+dtS1_osf <- dtS1$full %>%
+  select(all_of(export_vars)) %>%
+  mutate(EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE),
+         study = "S1") %>%
+  select(-starts_with("EvDayDiscr0"))
+
+dtS2_osf <- dtS2$full %>%
+  select(all_of(export_vars)) %>%
+  mutate(EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE),
+         study = "S2") %>%
+  select(-starts_with("EvDayDiscr0"))
+
+dtS3_osf <- dtS3$full %>%
+  select(all_of(export_vars)) %>%
+  mutate(EvDayDiscr.post = rowMeans(across(starts_with("EvDayDiscr")), na.rm = TRUE),
+         study = "S3") %>%
+  select(-starts_with("EvDayDiscr0"))
+
+dt_osf <- rbind(
+  dtS1_osf,
+  dtS2_osf,
+  dtS3_osf
+) %>% 
+  mutate(exWB = scales::rescale(exWB, from = range(exWB, na.rm = TRUE), to = c(0, 100)))
+
+saveRDS(dt_osf, file="data/osf_mini.Rda")
+```
+
 ## Data Availability and Sample Selection {.tabset .tabset-fade}

### 2023-08-07 sections/02InputVariables.Rmd
var info updated

@@
   select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
-  as.data.frame
+  as.data.frame %>% 
+  mutate(exWB = scales::rescale(exWB, from = range(exWB, na.rm = TRUE), to = c(0, 100)))
 

### 2023-08-07 sections/07ClusterEvaluationInterpretation.Rmd
var info updated

@@
   labs(
-    x = "Features",
+    x = "Variable",
     y = "Standardized Mean",

### 2023-08-07 supplemental-material-a-tutorial.qmd
var info updated

---
title: "Tutorial: Feature-base time series clustering"
format:
  html:
    theme: cosmo
    # css: style.css
    toc: true
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
#| echo: true

1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-08-07 tables/var_selection.tex
var info updated

@@
 Int: Meaningful & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & \Checkmark & \Checkmark & \\
+Int: Need Fulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\
 Int: Need Fulfillment Partner & -NAME- helped fulfill your goal (-GOAL-) & needs & \Checkmark & \Checkmark & \\
@@
 Int: Voluntary & The interaction with -NAME- was voluntary. & cognition & \Checkmark & \Checkmark & \\
-Int: Need Fulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark &  & \Checkmark\\
 Need Fulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & \Checkmark &  & \Checkmark\\

### 2023-08-07 tutorial/01-input-variables.qmd
var info updated

---
title: "1. Input Variables"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "readxl",
  "tidyverse",
  "kableExtra",
  "reshape",
  "knitcitations"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
anytime::addFormats("%Y-%m-%d %H:%M:%S")
anytime::addFormats("%d-%m-%Y")

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

source("scripts/load_data.R")

```

Time series clustering starts with the selection and preparation of the variables of interest. While the selection will necessarily be field- and concept-specific, there are a few conceptual and methodological issues that should be considered. Conceptually, we will need to make a choice of key variables that comprehensively capture the concept you would like to assess over time. Once the key variables are selected, we will need make sure that the variables are ready to be summarized using the time series features and that the format fits the chosen clustering procedure.

## Variable Selection

The included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series. One of the advantages of feature-based clustering is that it is inherently adept at accommodating multi-variate concepts --- a common aim in ESM research. In our illustration, we apply the clustering process to a recent set of studies that collected data on migration experiences. The data seek to capture the process of cultural adaptation and aims to identify diverging trajectories. Importantly for our illustration, the migration ESM research, also, exemplifies the real-world data issues that ESM data commonly face, including a multivariate conceptualization with event-specific missingness patterns.

For our illustration, we include 12 variables that were measured as part of the ESM surveys in all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group [@Kreienkamp2022d].

### Import the raw data

We begin by importing the raw data as well as the variable meta data. Both of these data files are available as part of <a href="https://www.osf.io" target="_blank">our OSF repository</a>. To ensure the privacy and confidentiality of the study participants, we took a number of steps to pre-process the data:

- consistent naming
- key variables only (participant privacy)
- post-test variable pre-calculated
- long format ESM data of all three studies combined (indicator variable: `study`)

This means that we import the [minimal reproducible dataset](https://doi.org/10.1002/cpet.32) of the key variables (`dt_raw` using the _data/osf_mini.Rda_ file) as well as a Microsoft Excel sheet providing a range of supplementary information, including the survey questions and variable labels (`var_meta` using the _data/osf_var_meta.xlsx_ file).

```{r}
#| label: import data
#| eval: false

dt_raw <- readRDS(file="data/osf_mini.Rda")
var_meta <- readxl::read_excel("data/osf_var_meta.xlsx")
```

### Variable Overview

To get an overview of the variables we display the variable names, labels, item descriptions, as well as the survey and analysis context (see [@tbl-var-overview]).

```{r}
#| label: tbl-var-overview
#| tbl-cap: "Variable Overview"
#| code-fold: true

var_meta %>%
  transmute(
    "Variable Name" = name,
    "Label" = label,
    "Description" = description,
    "Type" = type,
    "Contact Specific" = recode(contact, `0`="", ` 1`="✔"),
    "ESM" = recode(esm, `0`="", ` 1`="✔"),
    "Post" = recode(post, `0`="", ` 1`="✔"),
    "Cluster" = recode(cluster, `0`="", ` 1`="✔"),
    "Validation" = recode(interpretation, `0`="", ` 1`="✔")
  ) %>%
  kbl(.,
      escape = FALSE,
      booktabs = TRUE,
      label = "var_overview",
      format = "html",
      align = c(rep("l", 4), rep("c", ncol(.)-4)),
      digits = 2,
      linesep = "",
      caption = "Variable Overview") %>%
  add_header_above(c(" " = 5, "Survey" = 2, "Analysis" = 2)) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px") %>%
  footnote(
    general = c('All ESM items used a continuous slider and were rescaled to a range of 0—100.')
  )

```

## Variable Preparation

the important variables have been selected, the data needs to be prepared for the analysis steps. Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time-series comparable. Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses [@liao2005].

### Data Availability

In our illustration data set, the studies differed substantially in the maximum length of participation ($max(t_{S1})$ = 63, $max(t_{S2})$ = 69, $max(t_{S3})$ = 155). To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness. This procedure is in line with the general recommendation for data that might still need to rely on imputations for later model testing [@Madley-Dowd2019].

#### Study 1

For the Study 1 we first assess the data availability pattern. We do this visually by representing participants as rows and measurement occasions as columns:

```{r}
#| label: Data Availability Study 1
#| echo: false
#| warning: false
#| error: false

dtS1Availability <- dt_raw %>%
  filter(study == "S1") %>%
  select(
    PID,
    TIDnum
  ) %>%
  arrange(PID, TIDnum) %>%
  mutate(data = 1)
dtS1Availability <- reshape::cast(dtS1Availability, PID ~ TIDnum) %>%
  select(-PID) %>%
  mutate_all(function(x) ifelse(x>1,1,x))

rownames(dtS1Availability) <- paste("PP", 1:nrow(dtS1Availability), sep = "_")
colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
dtS1AvailabilityKbl <- dtS1Availability

dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)] <- lapply(dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS1AvailabilityKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 1: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

We then run the `cleanM` function, where sequentially remove the row or column with the most missing data until only rows and columns remain that have less than 45% missingness.

:::{.scrolling}

```{r}
#| label: Data Availability Adjustment Study 1

source("scripts/cleanM.R")
dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 55)
```

:::

With the adjusted dataset we again visualize the data availability.

```{r}
#| label: Plot Data Availability Adjustment Study 1

PIDout <- gsub("PP_", "", dtS1RedInfo$rowNamesOut) %>% as.numeric
TIDout <- gsub("t_", "", dtS1RedInfo$colNamesOut) %>% as.numeric
TIDInRed <- gsub("t_", "", dtS1RedInfo$colNamesIn) %>% as.numeric
TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)

dtS1Red <- dt_raw %>%
  filter(study == "S1") %>%
  filter(
    !PID %in% PIDout,
    TIDnum %in% TIDIn
  ) %>% 
  mutate(TIDnum = TIDnum - min(TIDnum))
rm(PIDout, TIDout, TIDInRed, TIDIn)

# change glitch where survey was sent too early
dtS1Red$TIDnum[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- 10
dtS1Red$TID[dtS1Red$PID == 7 & as.character(dtS1Red$created) == "2018-05-18 11:46:40"] <- "2018-05-18 Morning"

dtS1AvailabilityRedKbl <- dtS1RedInfo$reducedMatrix
dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)] <- lapply(dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS1AvailabilityRedKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 1: Final Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

#### Study 2

For Study 2 we follow the same procedure, where we first visualize the original data availability:

```{r}
#| label: Data Availability Study 2
#| echo: false
#| warning: false
#| error: false

dtS2Availability <- dt_raw %>%
  filter(study == "S2") %>%
  select(
    PID,
    TIDnum
  ) %>%
  arrange(PID, TIDnum) %>%
  mutate(data = 1)
dtS2Availability <- reshape::cast(dtS2Availability, PID ~ TIDnum) %>%
  select(-PID)

rownames(dtS2Availability) <- paste("PP", 1:nrow(dtS2Availability), sep = "_")
colnames(dtS2Availability) <- paste("t", 1:ncol(dtS2Availability), sep = "_")

dtS2AvailabilityKbl <- dtS2Availability
dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)] <- lapply(dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS2AvailabilityKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

Again run the `cleanM` function to harmonize the missingness:

:::{.scrolling}

```{r}
#| label: Data Availability Adjustment Study 2

dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 55)

```

:::

And visualize the adjusted data availability.

```{r}
#| label: Plot Data Availability Adjustment Study 2

PIDout <- gsub("PP_", "", dtS2RedInfo$rowNamesOut) %>% as.numeric
TIDout <- gsub("t_", "", dtS2RedInfo$colNamesOut) %>% as.numeric
TIDInRed <- gsub("t_", "", dtS2RedInfo$colNamesIn) %>% as.numeric
TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)

dtS2Red <- dt_raw %>%
  filter(study == "S2") %>%
  filter(
    !PID %in% PIDout,
    TIDnum %in% TIDIn
  ) %>% 
  mutate(TIDnum = TIDnum - min(TIDnum))
rm(PIDout, TIDout, TIDInRed, TIDIn)

# remove glitch entries 
dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 57 & dtS2Red$ended == "2018-12-20 21:58:31"),]
dtS2Red <- dtS2Red[!(dtS2Red$PID == 46 & dtS2Red$TIDnum == 59 & dtS2Red$ended == "2018-12-20 21:58:31"),]

dtS2AvailabilityRedKbl <- dtS2RedInfo$reducedMatrix
dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)] <- lapply(dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS2AvailabilityRedKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Final Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

#### Study 3

The data availability procedure for Study 3 again mirrors that of the previous two studies, where we assess the original availability (now with much more heterogeneity):

```{r}
#| label: Data Availability Study 3
#| echo: false
#| warning: false
#| error: false

dtS3Availability <- dt_raw %>%
  filter(study == "S3") %>%
  select(
    PID,
    TIDnum
  ) %>%
  arrange(PID, TIDnum) %>%
  mutate(data = 1)
dtS3Availability <- reshape::cast(dtS3Availability, PID ~ TIDnum, fill = 0) %>%
  select(-PID)

rownames(dtS3Availability) <- paste("PP", 1:nrow(dtS3Availability), sep = "_")
colnames(dtS3Availability) <- paste("t", 1:ncol(dtS3Availability), sep = "_")

dtS3AvailabilityKbl <- dtS3Availability
dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)] <- lapply(dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS3AvailabilityKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
rm(dtS3AvailabilityKbl)
```

Run the sequential missingess reduction function:

:::{.scrolling}

```{r}
#| label: Data Availability Adjustment Study 3

dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 55)
```

:::

And re-assess the data availability.

```{r}
#| label: PlotData Availability Adjustment Study 3

PIDout <- gsub("PP_", "", dtS3RedInfo$rowNamesOut) %>% as.numeric
TIDout <- gsub("t_", "", dtS3RedInfo$colNamesOut) %>% as.numeric
TIDInRed <- gsub("t_", "", dtS3RedInfo$colNamesIn) %>% as.numeric
TIDIn <- seq(min(TIDInRed), max(TIDInRed), 1)

dtS3Red <- dt_raw %>%
  filter(study == "S3") %>%
  filter(
    !PID %in% PIDout,
    TIDnum %in% TIDIn
  ) %>% 
  mutate(TIDnum = TIDnum - min(TIDnum))
rm(PIDout, TIDout, TIDInRed, TIDIn)

dtS3AvailabilityRedKbl <- dtS3RedInfo$reducedMatrix
dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)] <- lapply(dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS3AvailabilityRedKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 3: Final Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

We then assess the missingness within three adjusted datasets by comparing how many measurements, participants, and timepoints were removed for each study. In general, we also find that the number of time points are now much more comparable ($max(t_{S1})$ = 61, $max(t_{S2})$ = 60, $max(t_{S3})$ = 67).

```{r}
#| label: missingness filtering

source("scripts/missInfo.R")

missS1 <- missInfo(full = dt_raw %>% filter(study == "S1"), reduced = dtS1Red)
missS2 <- missInfo(full = dt_raw %>% filter(study == "S2"), reduced = dtS2Red)
missS3 <- missInfo(full = dt_raw %>% filter(study == "S3"), reduced = dtS3Red)

missS123 <- rbind(
  missS1,
  missS2,
  missS3
) %>%
  mutate(study = c(1, 2, 3)) %>%
  select(study, everything())

missS123 %>%
  kbl(.,
      escape = FALSE,
      format = "html",
      booktabs = TRUE,
      align = "c", #c("l", rep("c", ncol(.)-1)),
      col.names = c("Study", rep(c("Full","Reduced","&Delta;", "%"), 3)),
      digits = 2,
      caption = "Missingness Info by Study") %>%
  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

```

## Data Descriptives

To get a better understanding of the resulting dataset, we as look at the variable descriptives and inter-variable relationships. We combine the harmonized datasets into a single data frame `dtAll` and use our custom `MlCorMat` function to create a multilevel correlation and descriptives object: `allMlCor`.

```{r}
#| label: descriptive preparation
#| warning: false
#| error: false

dtAll <- rbind(
  dtS1Red %>% select(-c("created", "ended")) %>% mutate(study = "S1") %>%
    mutate(across(!TID & !study, as.numeric)),
  dtS2Red %>% select(-c("created", "ended")) %>% mutate(study = "S2"),
  dtS3Red %>% select(-c("created", "ended")) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(ID,
         PID,
         study,
         date,
         week,
         TIDnum,
         everything()) %>%
  arrange(ID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  as.data.frame

source("scripts/MlCorMat.R")
allMlCor <-
  MlCorMat(
    data = dtAll,
    id = "ID",
    selection = var_meta$name[var_meta$cluster == 1],
    labels = var_meta$label[var_meta$cluster == 1]
  )
```

We can then use the `kableExtra` package [@R-kableExtra] to display the descriptives table, which includes between-participant and within-participant bi-variate relationship information (correlations) as well as general descriptive statistics (mean, sd, ICC).

```{r}
#| label: descriptives table

allMlCor %>%
  as.data.frame %>%
  kbl(.,
      format = "html",
      caption = "Correlation Table and Descriptive Statistics",
      booktabs = TRUE,
      align = "c", #c("l", rep("c", ncol(.) - 1)),
      escape = FALSE,
      label = "descrLong",
      col.names = c(
        #"",
        "Int: Accidental",
        "Int: Voluntary", 
        "Int: Cooperative",
        "Int: Representative",
        "Int: Meaningful",
        "Int: Quality", 
        "Int: Need Fulfil.", 
        "Int: Need Fulfil. Partner", 
        "Int: Attitude Partner",
        "Daytime Core Need",
        "Outgroup Attitude",
        "Well-being"
      )) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  pack_rows("Correlations", 1, ncol(allMlCor)) %>%
  pack_rows("Descriptives", ncol(allMlCor)+1, nrow(allMlCor)) %>%
  footnote(
    general = c(
      '"Int." = interaction.',
      'Upper triangle: Between-person correlations;',
      'Lower triangle: Within-person correlations;',
      '*** p < .001, ** p < .01,  * p < .05'
    )
  )

```

Prepared with a cleaned and harmonized set of data we can move on to the feature extraction step.

## References

::: {#refs}
:::

### 2023-08-07 tutorial/02-feature-extraction.qmd
var info updated

---
title: "2. Feature Extraction"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "tidyverse",
  "kableExtra",
  "tsFeatureExtracR"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

# import sourced data for faster renders
source("scripts/load_data.R")
```

Armed with a relevant selection of key variables, the main aim of the feature extraction is to describe the most important and meaningful aspects of a time series. In its most general approach, feature extraction can include any numeric summary of the time series [@maharaj2019]. Given this flexibility, a staggering variety of time series features have been proposed across different disciplines [@wang2006;@fulcher2013;@adya2001].

## Feature Choice

With such a large number of options, choosing meaningful and relevant time series features is not an easy feat. However, not all time series features might be relevant to psychological time series or any particular research question. For example, a psychologist interested in well-being might not necessarily be interested in the exact time point after which 50% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density). Instead, we advocate that we look at time series features that have a strong backing within the experience sampling (i.e., ESM) literature and offer meaningful interpretability.

## Extraction Extraction

For our illustration, we rely on the work of researchers in the ESM literature, where different research teams have proposed a number meaningful measures to capture time series features that are relevant to psychological time series [@kuppens2017;@krone2018;dejonckheere2019;@wendt2020]. The main article discusses the individual time series features, their conceptual interpretation, as well as the technical calculation in more detail. Here we focus on the code to extract the features for the included variables.

The selection of time series features should always be driven by the research questions and field-specific conventions, but for our illustration we chose six time series features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity. We provide the `featureExtractor()` function to extract a number of measurements for each time series aspect. The time series feature operationalizations we chose for our illustration here are marked in bold.

1. central tendency
    + mean
    + **median**
2. variation around the central tendency (i.e., variability)
    + SD
    + **MAD**
3. (in)stability
    + root mean square of successive differences (rmssd)
    + **mean absolute change (mac)**
4. self-similarity
    + **lag-1 autocorrelation**
    + periodicity (lag-2 & lag-14 autocorrelation)
5. linear trend
    + **OLS regression parameter**
6. non-linear trend
    + **GAM edf**

_Note:_ Beyond the time series features described in this illustration, the [`tsfesh` python package](https://tsfresh.readthedocs.io/en/latest/) allows users to extract almost 800 time series features [@christ2018]. The package is well maintained and offers clear documentation of how each time series feature is calculated.

We, thus, use enter our cleaned variables into the `featureExtractor()` function, where we additionally specify the participant id `pid` and the numeric time id `tid` as well as a vector with the column names of the variables for which we want to extract the features `items`.

```{r}
#| label: run feature extraction
#| eval: false

# devtools::install_github("JannisCodes/tsFeatureExtracR")
library(tsFeatureExtracR)

# just in case check that only variables that have any data (i.e., not all NA)
featData <- dtAll %>%
  arrange(ID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% 
  as.data.frame

# Don't re-run unless you have a few hours to spare!
featFull <-
  featureExtractor(
    data = featData,
    pid = "ID",
    tid = "TIDnum",
    items = var_meta$name[var_meta$cluster == 1]
  )

# remove internal test account we missed earlier
featFull$features <- featFull$features %>%
    filter(
      ID != 19
    )

```

## Feature Missingness

After the feature extraction, again check how much missingess the data now entails. This check is important for two reasons. Firstly, the feature extraction should largely reduce the context-specific missingness as part of the summarization. However, secondly, even small numbers of missingness might not be accepted by some feature reduction or feature clustering algorithms. We this assess the missingess and impute the missing features based on the remaining data (provided that the missingness is small).

### Check Missingness

To assess the missingness, we enter the extracted features into our `feature_missing()` function, which assess global and feature=specific missingess (including a graphical representation of frequency of different amounts of missingess across the extracted features).

```{r}
#| label: feature missinginess

feature_selection <- c(
  "ar01",
  "edf",
  "lin",
  "mac",
  "mad",
  "median"
)

perc_miss_features <-
  feature_missing(featFull$features %>% 
                    select(ends_with(feature_selection)), 
                  title = "Feature-wise Missingness")

perc_miss_features$plt_miss_per_feature
```

For the time series features across the three studies, we find that about `r format(round(perc_miss_features$miss_overall, 2), nsmall=2)`% of the extracted features are missing across the `r nrow(featFull$features)` participants and `r ncol(featFull$features)` features. This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables. The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue.

### Impute Missing Features

For all missing features, we impute a single predictive mean matching imputation using a custom wrapper (i.e., the `featureImputer()` function) of the `MICE` package [@buuren2011]. Note again that with this procedure we only need to impute an extremely small number of missing values as most feature calculations can use the available data instead.

```{r}
#| label: feature imputer
#| eval: false

featFullImp <- featureImputer(featFull)
```

With the extracted and imputed features we will next have to check whether the feature set is still too large for a clustering analysis.

## References

::: {#refs}
:::

### 2023-08-07 tutorial/03-feature-reduction.qmd
var info updated

---
title: "3. Feature Reduction"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "tidyverse",
  "kableExtra",
  "tsFeatureExtracR",
  "stats",
  "factoextra",
  "plotly",
  "htmltools"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

# import sourced data for faster renders
source("scripts/load_data.R")

feature_selection <- c(
  "ar01",
  "edf",
  "lin",
  "mac",
  "mad",
  "median"
)
```

Once a meaningful set of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm [@altman2018]. In such cases, it is often recommended to perform an (optional) feature reduction. The commmonly used feature reduction procedures can commonly be separated into _feature selection_ and _feature projection_. We provide a more detailed discussion of the pros and cons of the different approaches. For this illustration we focus on feature projections, which have been popular because they are efficient, widely available, and applicable to a wide range of data types. That being said, because feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample [@carreira-perpinan1997], we recommend users to carefully consider the most useful option for their use case.

It is important to note again that a feature reduction step is not always necessary. Especially methods such as k-means are well equipped to deal with relatively large data sets [@jain2010]. It is important to check whether the full feature set leads to convergence issues or unstable cluster solutions. We go through the feature reduction steps as an exemplary illustration only.

## Principal Component Analysis

 We, specifically, selected the commonly used _principal component analysis_ (PCA). PCAs have the distinct benefit that they are well-established within the psychometric literature [@jolliffe2011] and can broadly be applied to a wide variety of studies in an automatized manner [@abdi2010]. As our aim is to present a general illustration that can also be adopted across use cases, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.

### Running the PCA

To use the PCA with our extracted time series features, we focus on a transformed feature set where the features are standardized across participants to ensure that all features are weighted equally. This transformation was already calculated as part of the `featureExtractor()` function and we now save it as a data frame `scaled_data`.

```{r}
#| label: pca data preparation

scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(feature_selection)) 
```

We then enter the standardized features into the principal component analysis using the `prcomp()` function from the `stats` package [@R-base].

```{r}
#| label: run pca

res.pca <- prcomp(scaled_data, scale. = FALSE)
pca_summary <- summary(res.pca)

# extract explained variances
res.varExplained <- data.frame(
  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
)
```

The PCA uses linear transformations in such a way that the first component captures the most possible variance of the original data (e.g., by finding a vector that maximizes the sum of squared distances; see @jolliffe2002, @abdi2010). The following components will then use the same method to iteratively explain the most of the remaining variance while also ensuring that the components are linearly uncorrelated [@shlens2014]. In practice, this meant that the PCA decomposed the `r ncol(scaled_data)` features into `r ncol(scaled_data)` principal components but now (because of the uncorrelated linear transformations) the first few principal components will capture a majority of the variance (see [@tbl-variance-decomposition]).

```{r}
#| label: tbl-variance-decomposition
#| tbl-cap: "Variance Decomposition"

# Creating a data frame to hold variance details
variance_df <- data.frame(
  Component = 1:length(pca_summary$sdev),
  `Standard Deviation` = pca_summary$sdev,
  `Proportion of Variance` = pca_summary$importance[2, ],
  `Cumulative Proportion` = pca_summary$importance[3, ]
)

# Producing the kable output
variance_df %>%
 kbl(.,
      escape = FALSE,
      booktabs = TRUE,
      label = "pca_summary",
      format = "html",
      digits = 2,
      linesep = "",
      col.names = c("Component", "Standard Deviation", "Proportion of Variance", "Cumulative Proportion"),
      caption = "PCA Component Summary") %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

### Component Selection

We can then decide how much information (i.e., variance) we are willing to sacrifice for a reduced dimensionality. A common rule of thumb is to use the principal components that jointly explain 70--90% of the original variance [@jackson2003]. For our illustration, we chose a 80% of the cumulative percentage explained variance as the cutoff (i.e., `cutoff_percentage`)

```{r}
#| label: pca variance cut off

cutoff_percentage <- 0.8
pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= cutoff_percentage))
```

This means that we select the first `r pca_cutoff` principal components that explain 80% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensionality by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`%).

```{r}
#| label: visualize variance explained

fviz_eig(res.pca, 
         addlabels = TRUE, 
         choice = "variance",
         ncp = pca_cutoff,
         barfill = "#a6a6a6",
         barcolor = "transparent",
         main = "Scree Plot of Explained Variance per Principal Component",
         xlab = "Principal Components") +
         theme_Publication() +
  theme(text = element_text(size = 14, color = "#333333"),
        axis.title = element_text(size = 16, face = "bold", color = "#555555"),
        plot.title = element_text(size = 18, hjust = 0.5, face = "bold", color = "#444444"),
        legend.position = "none",
        panel.grid.minor = element_blank())

res.varExplained <- data.frame(
  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
)

ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
  geom_point() +
  geom_line() +
  geom_vline(mapping = aes(xintercept = pca_cutoff), linetype = "longdash") +
  geom_hline(mapping = aes(yintercept = cutoff_percentage), linetype = "longdash") +
  geom_text(aes(x = pca_cutoff, y = min(Explained_Variance) + 0.01), 
            label = paste0(pca_cutoff, " components"), hjust = -0.1, vjust = 0, size = 3) +
  geom_text(aes(x = min(Components), y = cutoff_percentage), 
            label = paste0(cutoff_percentage*100, "%"), hjust = 0, vjust = -1, size = 3) +
  labs(title = "Cumulative Proportion of Variance Explained",
       x = "Number of Components",
       y = "Proportion of Variance Explained") +
  tsFeatureExtracR::theme_Publication()
```

We can then also get a slightly better of an understanding of how the data is distributed by visualizing the features and the participants within the lower dimensional space. We here chose the first three principal components to allow for human undestandable dimensionality.

```{r}
#| label: plot particpants and features in reduced space

plot1 <- plot_ly(
  res.pca$rotation %>% as.data.frame,
  x = ~ PC1,
  y = ~ PC2,
  z = ~ PC3,
  size = 2,
  type = 'scatter3d',
  mode = 'markers',
  text = ~ paste('Variable:', rownames(res.pca$rotation))
) %>%
  layout(title = 'Variables in reduced space')
plot2 <- plot_ly(
  res.pca$x %>% as.data.frame,
  x = ~ PC1,
  y = ~ PC2,
  z = ~ PC3,
  size = 2,
  type = 'scatter3d',
  mode = 'markers',
  text = ~ paste('Participant:', rownames(res.pca$x))
) %>%
  layout(title = 'Participants in reduced space')

# Display side by side using htmltools
div(
  style = "display: flex; justify-content: space-between;",
  list(
    div(plot1, style = "width: 50%;"),
    div(plot2, style = "width: 50%;")
  )
)
```

### Save Principal Component Scores

For the extracted principal components we save the `r pca_cutoff` principal component scores for each participant (i.e., the participants' coordinates in the reduced dimensional space; PC-scores) in the `pca_scores` data frame.

```{r}
#| label: save PC scores

# extract coordinates for the individuals (PC-scores)
pca_scores <- as.data.frame(res.pca$x[,1:pca_cutoff])
```

It is important to note that we do not standardize the principal component scores again before adding them into the clustering algorithm. With the PCs the differences in variances are crucial to the dimensionality reduction and standardizing the PC-scores would remove this information.

```{r}
#| label: visualize the standard deviations of each extracted component

pca_scores %>%
  summarize_all(sd) %>%
  pivot_longer(everything(), names_to = "PC", values_to = "sd") %>%
  ggplot(., aes(x=reorder(PC, sd), y=sd, label=sd)) +
  geom_bar(stat='identity', width=.5) +
  coord_flip() +
  labs(
    x = "Principal Component",
    y = "Standard Deviation",
    title = "Standard Deviation of each Principal Component"
  ) +
  theme_Publication()
```

With the reduced variable space of the PC-scores we can now move on to the clustering of the data.

```{r}
#| label: save pca.out for following qmd files
#| include: false

saveRDS(pca_scores, file = "data/pca_scores.Rds")
```

## References

::: {#refs}
:::

### 2023-08-07 tutorial/04-clustering.qmd
var info updated

---
title: "4. Clustering"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "tidyverse",
  "kableExtra",
  "tsFeatureExtracR",
  "factoextra",
  "NbClust"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

# import sourced data for faster renders
source("scripts/load_data.R")

pca_scores <- readRDS(file = "data/pca_scores.Rds")

feature_selection <- c(
  "ar01",
  "edf",
  "lin",
  "mac",
  "mad",
  "median"
)
```

For the actual clustering of the time-series features, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible [@liao2005]. There are a many different ways in which (time series) features can be clustered into homogeneous and well separated groups. We provide a more in-depth discussion of the different options in the main manuscript and we recommend the excellent overview of @xu2015 for a more general description of the different available approaches. Briefly speaking, the more readily available approaches suitable for most time series feature data can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof [@jain1999]. There is, unfortunately, no one-size-fits-all solution to clustering and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency. For our own illustration, we have chosen the centroid-based k-means clustering.

## K-means Clustering

While k-means may not always deliver the highest accuracy, it brings forth distinct benefits. We selected k-means primarily due to its notable efficiency, which handles vast numbers of participants and features without imposing numerous limiting assumptions on cluster shapes [@jain2010]. The k-means technique has garnered recognition in the research world and has found its way into numerous statistical software tools [@hand2005]. Moreover, a significant number of feature selection approaches have been tailored explicitly for the reputable k-means algorithm [@boutsidis2010]. Therefore, k-means serves as an excellent foundation for many in the field of psychological research, and its applicability spans a diverse array of projects.

In practice, we entered the participants' PC-scores from the feature reduction step into the k-means algorithm. Because we did not know the underlying number of clusters within our sample, we calculated the cluster solutions for $k=\{2, \dots , 10\}$ . To avoid local minima we used 100 random initial centroid positions for each run. We here use a for-loop to arrive at these cluster solutions using the `kmeans()` function from the `stats` package [@R-base] and save the results in the `kmeans_results` list.

```{r}
#| label: k-means modeling based on selected k

kmeans_results <- list()
for (i in 2:10) {
  kmeans_results[[i-1]] <- kmeans(pca_scores, centers = i, nstart = 100)
}
```

Each of the 9 cluster solutions converged within the iteration limit. In the next step, we will then evaluate which of the extracted cluster solutions offers the best fit with the data.

```{r}
#| label: save kmeans results
#| include: false

saveRDS(kmeans_results, file = "data/kmeans_results.Rds")
```

## References

::: {#refs}
:::

### 2023-08-07 tutorial/05-evaluation.qmd
var info updated

---
title: "5. Cluster Evaluation"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "tidyverse",
  "kableExtra",
  "tsFeatureExtracR",
  "factoextra",
  "fpc"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

# import sourced data for faster renders
source("scripts/load_data.R")

pca_scores <- readRDS(file = "data/pca_scores.Rds")
kmeans_results <- readRDS(file = "data/kmeans_results.Rds")

feature_selection <- c(
  "ar01",
  "edf",
  "lin",
  "mac",
  "mad",
  "median",
  "_n",
  "EvDayDiscr.post"
)

raw_data <- featData
z_data <- featFullImp$featuresImpZ %>% select("ID", ends_with(feature_selection))
scaled_data <- featFullImp$featuresImpZMat %>% select(ends_with(feature_selection))

```

Now that the participants have been assigned to their respective clusters based on the similarity of their time series features, the final evaluation step includes two main elements, (1) evaluating the performance of the clustering analyses to choose an optimal solution and (2) interpreting the extracted clusters conceptually.

## Performance

In the performance evaluation step, we are mainly concerned with finding the optimal clustering solution. In the previous step, we extracted 9 possible solutions that might all be good at separating the participants. We thus have to evaluate and compare the performance of each solution. In practical terms, performance evaluation often means assessing the accuracy, stability, and separation or purity of the clustering [@keogh2003]. In our own illustration example, we used the `cluster.stats()` function from the `fpc` package, which calculates a wide variety of internal cluster validity statistics for each of the extracted clustering solutions.

```{r}
#| label: tbl-cluster-performance

kmeans_performance <- list()
for (i in 2:10) {
  kmeans_performance[[i - 1]] <-
    fpc::cluster.stats(
      d = dist(pca_scores),
      clustering = kmeans_results[[i - 1]]$cluster,
      G2 = TRUE,
      G3 = TRUE,
      aggregateonly = TRUE
    ) %>% unlist
}
df_kmeans_performance <-
  data.frame(Reduce(rbind, kmeans_performance))

kmeans_metrics_names <- c(
  "n" = "number of cases", 
  "cluster.number" = "number of points", 
  "min.cluster.size" = "size of smallest cluster", 
  "noisen" = "number of noise points", # Not relevant for kmeans 
  "average.between" = "average distance between clusters", 
  "average.within" = "average distance within clusters", # (reweighted so that every observation, rather than every distance, has the same weight)
  "max.diameter" = "maximum cluster diameter",
  "min.separation" = "minimum cluster separation", 
  "ave.within.cluster.ss" = "within clusters sum of squares", # generalisation
  "avg.silwidth" = "average silhouette width", 
  "g2" = "Goodman Kruskal's Gamma coefficient", # See Milligan and Cooper (1985), Gordon (1999, p. 62). 
  "g3" = "G3 coefficient", # See Gordon (1999, p. 62)
  "pearsongamma" = "Normalized gamma", # correlation between distances and a 0-1-vector where 0 means same cluster, 1 means different clusters. see Halkidi et al. (2001)
  "dunn" = "Dunn index", # minimum separation / maximum diameter, Dunn index, see Halkidi et al. (2002).
  "dunn2" = "Dunn index 2", # minimum average dissimilarity between two cluster / maximum average within cluster dissimilarity, another version of the family of Dunn indexes
  "entropy" = "entropy of distribution of cluster memberships", # see Meila(2007)
  "wb.ratio" = "average within / average between", 
  "ch" = "Calinski and Harabasz index", # (Calinski and Harabasz 1974, optimal in Milligan and Cooper 1985; generalised for dissimilarites in Hennig and Liao 2013)
  "widestgap" = "widest within-cluster gap", 
  "sindex" = "separation index"
)

df_kmeans_performance %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = "c",
      digits = 2,
      row.names = FALSE,
      col.names = kmeans_metrics_names,
      caption = "Cluster Performance for the $k$ = 2–10") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

With real-world data no single evaluation measure is likely perfect, and different measures may produce different results depending on the characteristics of the data and the research question being addressed [@kittler1998]. We found that across most indices, the analysis with $k=2$ clusters performed the best (see [@tbl-cluster-performance]).

### Visual Illustration

Two commonly reported performance measures showed this visually as well. The first statistic is the total within-cluster sum of square $WCSS$. While the within-cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$ after which the decrease was much smaller. This method is also sometimes referred to as the 'elbow method' [@syakur2018].

```{r}
#| label: Plot Elbow method Performance

fviz_nbclust(pca_scores, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Performance Metric: Elbow method") +
  theme_Publication()
```

A second, commonly used measure is the average silhouette score. This statistic measures the degree to which each time feature data point is similar to other points within the same cluster, compared to points in other clusters [@rousseeuw1987]. In our case, the $k=2$ solution maximized the silhouette coefficient ($s_2=$ `r format(round(df_kmeans_performance$avg.silwidth[df_kmeans_performance$cluster.number==2], 2), nsmall = 2)`).

```{r}
#| label: Plot Silhouette method Performance

fviz_nbclust(pca_scores, kmeans, method = "silhouette")+
  labs(subtitle = "Performance Metric: Silhouette method") +
  theme_Publication()
```

### Best-performing Solution

We thus, save the $k=2$ solution as the final model `kmeans_final`.

```{r}
#| label: kmeans choice

set.seed(42)
kmeans_final <- kmeans(pca_scores, centers = 2, nstart = 100)

```

We can then use the cluster assignments to visualize the two clusters. This is commonly done using a projection method where the group separation is visualized on a two-dimensional pane. This is not ideal (because the clustering takes a multi-dimensional space into account), but can give a first indivation of the cluster separation.

```{r}
# label: visualize the cluster assignment

fviz_cluster(kmeans_final, geom = "point", data = z_data, ggtheme = theme_Publication())
```

## Interpretation

The interpretation of feature-based time series clustering in psychology involves understanding the meaning and implications of the obtained clusters. In order to make sense of the clustering results, we here focus on three general aspects of the results [@kaufman1990]. (1) Assessing differences between the clusters in the original time series features, (2) comparing the clusters based on prototype developments, (3) comparing the clusters based on between-person differences that were not included in the initial clustering.

### TS Feature Comparison

We first inspect the clusters based on the average values of meaningful features. To do so, we merge the cluster assignments from the `kmeans_final` object back to the original set of standardized time series features `z_data` (same as `scaled_data` earlier). We then are able to group the features by the clusters and calculate average feature values for each cluster. We also pivot the dataframe into long format for easier visualizations, separate the variable and the feature identifiers, and add more descriptive labels for the visualizations.

```{r}
#| label: Extract cluster assignments and summarize
#| message: false
#| warning: false

featCluster <- data.frame(ID = as.numeric(names(kmeans_final$cluster)), 
                            cluster = kmeans_final$cluster) %>%
  left_join(z_data, by = "ID") %>%
  left_join(
    featData %>% 
      select(ID, EvDayDiscr.post) %>% 
      distinct(ID, EvDayDiscr.post, .keep_all = TRUE) %>%
      mutate(EvDayDiscr.post_z = scale(EvDayDiscr.post, scale = TRUE)) %>%
      replace_na(list(EvDayDiscr.post_z = NA)) %>%
      select(
        ID,
        EvDayDiscr.post_z
      ), 
    by = "ID"
  ) %>%
  select(-ID) %>%
  group_by(cluster) %>%
  summarise(across(everything(), list(mean = mean), na.rm=TRUE)) %>%
  ungroup %>% 
  select(-starts_with("ID")) %>%
  pivot_longer(
    cols = -c(cluster),
    names_to = c("variable", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>% 
  mutate(
    nam = variable,
    variable = gsub("\\_.*", "", nam),
    feature = gsub("^.*?\\_", "", nam)
  ) %>% 
  mutate(feature = ifelse(feature == "z", "mean", feature)) %>%
  left_join(var_meta %>% select(name, label), by = c("variable" = "name")) %>%
  select(
    cluster,
    nam,
    variable,
    label,
    feature,
    everything()
  )
```

We can then plot the average features and for each variable and flipping the axes allows us to assess the same data once with a focus on the variable (see [@fig-clusterFeatVarComb] A) and once with a focus on the features (see [@fig-clusterFeatVarComb] B).

```{r}
#| label: fig-clusterFeatVarComb
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 18
#| fig-cap: Note that n and discrimination are out of cluster variables.

# clusterVariableGrid
clusterVariableGrid <- featCluster %>% 
  filter(variable != "EvDayDiscr.post",
         feature != "n") %>%
  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
  coord_flip() +
  labs(
    x = "Features",
    y = "Standardized Mean",
    color = "Cluster"
  ) +
  facet_wrap(~ label) +
  theme_Publication()

# clusterFeatureGrid
clusterFeatureGrid <- featCluster %>% 
  ggplot(., aes(x = reorder(label, -as.numeric(label == "Post: Discrimination")), y = mean, group = cluster, color = as.factor(cluster))) +
  geom_point() +
  geom_line() +
  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
  coord_flip() +
  labs(
    x = "Variable",
    y = "Standardized Mean",
    color = "Cluster"
  ) +
  facet_wrap(~ factor(feature, levels = unique(featCluster$feature)), ncol=4) +
  theme_Publication()

ggpubr::ggarrange(
  clusterVariableGrid + theme(axis.title.x = element_blank()),
  clusterFeatureGrid,
  ncol = 1,
  labels = c("(A) Variable Focus", "(B) Feature Focus"),
  common.legend = TRUE,
  legend = "bottom",
  align = "v"
) 
```

We discuss the individual differences and their theoretical relevance as part of the main manuscript. However, in terms of methodological insight, we find variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail. In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2). We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics [@bringmann2018c;eronen2021a].

### Average Time Series

In the second step, we look at the prototypical trajectories of the clusters. For k-means clustering it is often recommended to use the average over time of the responses within the cluster [@niennattrakul2007].  We visualize the cluster-specific time point averages (see [@fig-cluster-timeseries] A) as well as an average spline per cluster (see [@fig-cluster-timeseries] A).

```{r}
#| label: fig-cluster-timeseries
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 18

rawCluster <- data.frame(ID = as.numeric(names(kmeans_final$cluster)), 
                            cluster = kmeans_final$cluster) %>%
  left_join(raw_data, by = "ID") %>%
  select(
    ID,
    TIDnum,
    cluster,
    everything()
  ) %>%
  reshape2::melt(id.vars = c("cluster", "ID", "TIDnum")) %>% 
  filter(
    !variable %in% c("PID", "study", "date", "week", "TID", "EvDayDiscr.post"),
    !is.na(value)
  ) %>%
  left_join(var_meta %>% select(name, label), by = c("variable" = "name")) %>%
  mutate(value = as.numeric(value))

clusterTS <- rawCluster %>% 
  ggplot(., aes(x=TIDnum, y=as.numeric(value), group=cluster, color=as.factor(cluster))) +
  geom_line(aes(x=TIDnum, y=as.numeric(value), group=ID, color=as.factor(cluster)), alpha=0.05) +
  stat_summary(fun=mean, geom="line") +
  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
  facet_wrap(. ~ label) +
  labs(
    x = "Time ID",
    y = "Response (Mean Cluster Response Bold)",
    color = "Cluster"
  ) +
  theme_Publication()

clusterTSTrend <- rawCluster %>% 
  ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
  geom_smooth(aes(group=cluster)) +
  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
  #stat_summary(fun=mean, geom="line") +
  facet_wrap(~ label) +
  labs(
    x = "Time ID",
    y = "Response (Mean Cluster Response Bold)",
    color = "Cluster"
  ) +
  theme_Publication()

ggpubr::ggarrange(
  clusterTS + theme(axis.title.x = element_blank()),
  clusterTSTrend,
  ncol = 1,
  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
  common.legend = TRUE,
  legend = "bottom"
)
```

While we discuss the interpretations in more detail in the manuscript, the average cluster trajectories offer clearer insights into the mean differences, the variances, as well as the trends over time.

### Out-of-cluster Comparison

Finally, we can also assess the clusters across other individual difference variables [@monden2022]. This out-of-feature comparison allows us to check for data artifacts, as well as check whether the developmental clusters are associated with important social markers and individual differences.

To illustrate artifact checks, we added the number of ESM measurements into the comparison and find that the participants in the deterioration cluster (cluster 2) on average completed slightly more ESM surveys in general and reported on more intergroup interactions in particular (see $n$ in [@fig-clusterFeatVarComb] B). This difference in the reported number of interactions might indicate either a clustering artifact or a meaningful difference. The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are substantially larger due to the larger samples (e.g., restriction of range in the smaller sample; @kogan2006). In our case, this seems less likely because one out of four variables did not differ in terms of the MAD (i.e., our selected measurement of the time series variance; see [@fig-mad-cluster-comparison] below).

```{r}
#| label: fig-mad-cluster-comparison
#| message: false
#| warning: false

# ESM features
raw_feat_clust <-
  data.frame(ID = as.numeric(names(kmeans_final$cluster)),
             cluster = kmeans_final$cluster) %>%
  left_join(z_data, by = "ID")

concepts_and_feature <- str_match(colnames(raw_feat_clust), "(.*)_(.*)")
concepts <- as.character(na.omit(unique(concepts_and_feature[, 2])))
ts_features <- as.character(na.omit(unique(concepts_and_feature[, 3])))

results <- run_all_tests(
  df = raw_feat_clust,
  cluster_var = "cluster",
  ts_features = ts_features,
  concepts = concepts
)

results_df <- convert_test_to_dataframe(
  results = results,
  ts_features = ts_features, 
  concept = concepts
)

# validation feature
validation_feature <-
  featData %>% 
  select(ID, EvDayDiscr.post) %>% 
  distinct(ID, EvDayDiscr.post, .keep_all = TRUE) %>%
  mutate(EvDayDiscr.post_z = scale(EvDayDiscr.post, scale = TRUE)) %>%
  replace_na(list(EvDayDiscr.post_z = NA)) %>%
  select(
    ID,
    EvDayDiscr.post_z
  ) %>%
  left_join(data.frame(ID = as.numeric(names(kmeans_final$cluster)),
             cluster = kmeans_final$cluster), by = "ID")

t_discrimination <- t.test(
  validation_feature$EvDayDiscr.post_z[validation_feature$cluster == 1],
  validation_feature$EvDayDiscr.post_z[validation_feature$cluster == 2],
  na.action = na.omit
)

results_df <- rbind(
  results_df,
  data.frame(
    concept = "EvDayDiscr.post",
    feature = "mean",
    difference = t_discrimination$estimate[1] - t_discrimination$estimate[2],
    t_value = t_discrimination$statistic,
    df = t_discrimination$parameter,
    p_value = t_discrimination$p.value,
    lwr = t_discrimination$conf.int[1],
    upr = t_discrimination$conf.int[2]
  )
) %>%
  annotate_significance(., 
                        variableLab = var_meta %>% select(name, label), 
                        join_by = c('concept' = 'name'))

feature_plot_list <- generate_feature_plots(results_df)

feature_plot_list$mad

feature_ttest_apa(results_df, "mean", "EvDayDiscr.post", "html")
```

At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions (`r feature_ttest_apa(results_df, "InteractionContextvoluntary", "n", "html")`), but these interactions were less voluntary (`r feature_ttest_apa(results_df, "InteractionContextvoluntary", "median", "html")`), less meaningful (`r feature_ttest_apa(results_df, "qualityMeaning", "median", "html")`), and less positive (`r feature_ttest_apa(results_df, "qualityOverall", "median", "html")`).

To underscore the value of examining out-of-feature individual variations, we also contrasted the two samples based on the participants' self-reported experiences of discrimination in the Netherlands, captured during the post-measurement phase. Upon analyzing the group differences, it was evident that those in the deteriorating cluster (cluster 2) indicated notably elevated instances of daily discrimination (`r feature_ttest_apa(results_df, "EvDayDiscr.post", "mean", "html")`; also see [@fig-clusterFeatVarComb] B). In short, both intensive longitudinal measurement (e.g., the sum of specific ESM measurements) and cross-sectional variables (e.g., general discrimination differences) that were not included in the original clustering step can be leveraged to delve deeper into understanding cluster disparities.

## References

::: {#refs}
:::

### 2023-08-07 tutorial/about.qmd
var info updated

--- 2023-02-24 IllustrationResultsRenderReport.Rmd
@@
 ---
-title: "Report: Illustration Results"
-subtitle: "Render report for 'A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series'"
-date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
-output: pdf_document
-editor_options:
-  chunk_output_type: console
-bibliography: referencesZotero.bib
-csl: apa.csl
-header-includes:
-   - \usepackage{amsmath, nccmath}
+title: "About"
+bibliography: [packages.bib, ../referencesZotero.bib]
+csl: ../apa.csl
 ---
@@
 #| label: setup
-#| include: false
+#| include: FALSE
 
-# R Studio Clean-Up
-cat("\014") # clear console
-# rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
-gc() # garbage collector
-
-# Install and Load Packages
-# !IMPORTANT!
-# BEFORE FIRST RENDER:
-# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
-# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
 lib <- c(
@@
   "knitr",
-  "remedy",
-  "bookdown",
-  #"tidyverse",
-  "MASS",
-  "brms",
-  "psych",
-  "ggplot2",
-  "ggthemes",
-  "haven",
-  "RColorBrewer",
-  "plotly",
-  "grid",
-  "gridExtra",
-  "ggpattern",
-  "lme4",
-  "nlme",
-  "jtools",
-  "gtsummary",
-  "sessioninfo",
-  "tibble",
-  "pander",
-  "devtools",
-  "mada",
-  "data.table",
-  "plyr",
-  "dplyr",
-  "tidyr",
-  "Hmisc",
+  "tidyverse",
   "kableExtra",
-  "papaja",
-  "stringr",
-  "stringi",
-  "reshape2",
-  "anytime",
-  "lubridate",
-  "purrr",
-  "metafor",
-  "dygraphs",
+  "tsFeatureExtracR",
+  "factoextra",
+  "fpc",
   "readxl",
   "reshape",
-  "factoextra",
-  "Amelia", 
-  "ThreeWay",
-  "DescTools",
-  # GAM
-  "gratia",
-  #"MASS",
-  "mgcv",
-  # For moving window models
-  "Kendall",
-  # For the Hamed & Rao correction of tau
-  "modifiedmk", 
-  "mgcViz",
-  "shiny"
+  "tsFeatureExtracR",
+  "stats",
+  "plotly",
+  "htmltools",
+  "NbClust",
+  "devtools",
+  "pander",
+  "knitcitations",
+  "grateful",
+  "bibtex"
 )
 invisible(lapply(lib, library, character.only = TRUE))
-rm(lib)
+knitr::write_bib(file = 'packages.bib')
 
-anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
-anytime::addFormats("%Y-%m-%d %H:%M:%S")
-anytime::addFormats("%d-%m-%Y")  
-#options(rgl.useNULL = TRUE)
-
-# Load Custom Packages
-source("./scripts/functions/fun.panel.R")
-source("./scripts/functions/themes.R")
-source("./scripts/functions/binaryCor.R")
-source("./scripts/functions/MlCorMat.R")
-source("./scripts/functions/MlTbl.R")
-source("./scripts/functions/metaLmer.R")
-source("./scripts/functions/meanViz.R")
-source("./scripts/functions/gam.R")
-source("./scripts/functions/tsFeatureExtractor.R")
-
-# Markdown Options
-knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
-knitr::opts_knit$get("root.dir") # check working directory
-options(
-  scipen = 999,
-  digits = 4,
-  width = 400
-) # removes scientific quotation
-# knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings
-
-# Global Chunk Options
 knitr::opts_chunk$set(
-  fig.width = 12,
-  fig.height = 8,
-  fig.path = "figures/",
-  echo = TRUE,
-  warning = FALSE,
-  message = FALSE
+  class.source = 'numberLines lineAnchors'
 )
 
-outputDir <- "latex-results"
 ```
 
-```{r}
-#| label: import data
-#| include: true
+This illustration accompanies the publication "A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series". As such, the illustration is not ...
 
-# Load variable name lookup table
-varNames <- readxl::read_excel("preregistration/varNames.xlsx")
+## Open Data
 
-# import full datasets
-load("data/s123Full.RData")
+short open data description here. Link to OSF
 
-# import reduced datasets
-load("data/s123Red.RData")
+## Open Code
 
-# import pre-calculated features for faster renders
-load("data/features/featOut.RData")
-#save(featData, featFullImp, file = "data/features/featOutFullImp.RData") # speed up attempt 
-#load("data/features/featOutFullImp.RData")
-```
+short open code description here. Link to GitHub
 
-```{r}
-#| label: Preparation of Variable Names
-#| include: true
-
-varNamIndicesS1 <- c(
-  "autonomy",
-  "competence",
-  "relatedness",
-  "education_level.pre",
-  "associationMerged.pre",
-  "assimilation.pre",
-  "separation.pre",
-  "integration.pre",
-  "marginalization.pre",
-  "VIA_heritage.pre",
-  "VIA_Dutch.pre",
-  "SSAS_surrounding.pre",
-  "SSAS_privat.pre",
-  "SSAS_public.pre",
-  "assimilation.post",
-  "separation.post",
-  "integration.post",
-  "marginalization.post",
-  "VIA_heritage.post",
-  "VIA_Dutch.post",
-  "rosenberg.post",
-  "social_support.post",
-  "stress.post",
-  "discrimination.post",
-  "discrimination_month.post",
-  "NLE_1month.post",
-  "NLE_6month.post",
-  "NLE_12month.post"
-  )
-
-varNamS123Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS3Aux <- varNames %>%
-  filter(
-    aux != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamS123PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1",
-    studyS2 == "S2",
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS1PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull
-varNamS2PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS2 == "S2"
-  ) %>%
-  select(varNam) %>% 
-  pull
-varNamS3PCA <- varNames %>%
-  filter(
-    pca != 0,
-    studyS3 == "S3"
-  ) %>%
-  select(varNam) %>%
-  pull
-
-varNamOut <- c(
-  "ResponseId",
-  "relatednessNoInteraction",
-  "relatednessSelf",
-  "relatednessOther",
-  "autonomy_Int",
-  "autonomy_NoInt",
-  "competence_Int",
-  "competence_NoInt",
-  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
-  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
-  varNamS3PCA[grepl(
-    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
-    varNamS3PCA
-  )]
-)
-varNamCore <- c(
-  "PID",
-  "TID",
-  "TIDnum"
-)
-varNamNewAll <- c(
-  "relatedness",
-  "autonomy",
-  "competence"
-)
-varNamNewS1 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "alertness", 
-  "calmness", 
-  "valence"
-)
-varNamNewS2 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo"
-)
-varNamNewS3 <- c(
-  "relatedness",
-  "autonomy",
-  "competence",
-  "ProSo", 
-  "AntiSo",
-  "agency",
-  "autoFrust",
-  "autoSat",
-  "relatFrust",
-  "relatSat",
-  "compFrust",
-  "compSat",
-  "lonely",
-  "emotRegPos",
-  "emotRegNeg"
-)
-
-varNamNewCat <- c(
-  "closeness_Calc",
-  "gender_Calc",
-  "ethnicity_Calc",
-  "relationship_Calc"
-)
-
-varNamIntDep <- c(
-  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
-)
-
-varNamIndiceItemsS1 <- varNames %>%
-  filter(
-    aux == -1,
-    studyS1 == "S1"
-  ) %>%
-  select(varNam) %>%
-  pull %>%
-  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
-  unique
-
-varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
-varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
-varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
-varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)
-
-varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
-varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
-varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
-varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])
-
-varNamS123MIPsbl <- c(
-  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
-  varNamS123MI,
-  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
-)
-
-varNamS123MiRed <- Reduce(
-  intersect,
-  list(
-    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
-    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
-  )
-)
-
-idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
-```
-
-```{r}
-#| label: variable names for all analyses
-#| include: true
-
-# ID Variables (just to re-iterate)
-idVars <- idVars
-
-# Common variables across studies 
-varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]
-
-# Non-Interaction Specific Variables
-varNamS123NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# Interaction Specific Variables
-varNamS123Int <- varNames %>%
-  filter(
-    varNam %in% varNamS123PCA,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 variables
-varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]
-
-# S1 Non-Interaction Specific Variables
-varNamS1NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S1 Interaction Specific Variables
-varNamS1Int <- varNames %>%
-  filter(
-    varNam %in% varNamS1Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 variables
-varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]
-
-# S2 Non-Interaction Specific Variables
-varNamS2NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S2 Interaction Specific Variables
-varNamS2Int <- varNames %>%
-  filter(
-    varNam %in% varNamS2Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 variables
-varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]
-
-# S3 Non-Interaction Specific Variables
-varNamS3NoInt <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 0
-  ) %>%
-  select(varNam) %>%
-  pull
-
-# S3 Interaction Specific Variables
-varNamS3Int <- varNames %>%
-  filter(
-    varNam %in% varNamS3Clust,
-    contactSpecific == 1
-  ) %>%
-  select(varNam) %>%
-  pull
-```
-
-```{r}
-#| label: variable names by response type and study
-#| include: true
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="X")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3))
-```
-
-```{r}
-#| label: variables included in main analysis by response type
-#| include: true
-
-varNames %>%
-  filter(
-    pca != 0,
-    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
-  ) %>%
-  mutate(
-    contactSpecific = recode(
-      .$contactSpecific,
-      `-1` = "No Interaction Only",
-      `0` = "unspecific",
-      `1` = "Interaction Only"
-    ),
-    study1 = ifelse(!is.na(studyS1), 1, 0),
-    study2 = ifelse(!is.na(studyS2), 1, 0),
-    study3 = ifelse(!is.na(studyS3), 1, 0)
-  ) %>% 
-  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
-  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
-  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
-  summarise(n = n()) %>%
-  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
-  ungroup %>%
-  t %>%
-  as.data.frame %>%
-  janitor::row_to_names(1) %>%
-  replace(is.na(.), 0) %>%
-  tibble::rownames_to_column(., var = "variable") %>%
-  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
-  mutate(across(c(study1, study2, study3), as.numeric)) %>%
-  #select(-study1, -study2, -study3) %>%
-  arrange(concept, unspecific, `Interaction Only`) %>% 
-  mutate_all(~recode(.x, `0`="", ` 1`="X")) %>%
-  relocate(starts_with("study"), .after = last_col()) %>%
-  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
-  kbl(.,
-      escape = FALSE,
-      booktabs = T,
-      align = c("l", "l", rep("c", ncol(.)-2)),
-      digits=2,
-      caption = "Variables by Interaction Types and Study Availability") %>%
-  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
-  kable_classic(
-    full_width = F,
-    lightable_options = "hover",
-    html_font = "Cambria"
-  )
-```
-
-# Data Description
-
-```{r}
-#| label: render data description
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/01DataDescription.Rmd",
-  output_dir = outputDir
-)
-```
-
-# Input Variables
-
-```{r}
-#| label: render input variables
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/02InputVariables.Rmd",
-  output_dir = outputDir
-)
-```
-
-# Feature Extraction
-
-```{r}
-#| label: render feature extraction
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/03FeatureExtraction.Rmd",
-  output_dir = outputDir
-)
-```
-
-# Feature Reduction
-
-```{r}
-#| label: render feature reduction
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/04FeatureReduction.Rmd",
-  output_dir = outputDir
-)
-```
-
-# Feature Clustering
-
-```{r}
-#| label: render feature clustering
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/05FeatureClustering.Rmd",
-  output_dir = outputDir
-)
-```
-
-# Cluster Evaluation
-
-## CLuster Performance
-
-```{r}
-#| label: render cluster performance
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/06ClusterEvaluationPerformance.Rmd",
-  output_dir = outputDir
-)
-```
-
-## CLuster Interpretation
-
-```{r}
-#| label: render cluster interpretation
-#| include: true
-#| results: 'asis'
-
-rmarkdown::render(
-  input = "sections/07ClusterEvaluationInterpretation.Rmd",
-  output_dir = outputDir
-)
-```
-
-# Software Information
+## Software Information
 
@@
 
-<!-- <details> -->
+<details>
 
-<!-- <summary> -->
+<summary>
 
@@
 
-<!-- </summary> -->
+</summary>
 
-<!-- \renewcommand{\arraystretch}{0.8} -->
+\renewcommand{\arraystretch}{0.8}
 
-<!-- <!-- decrease line spacing for the table --> -->
+<!-- decrease line spacing for the table -->
 
@@
 
-<!-- \renewcommand{\arraystretch}{1} -->
+\renewcommand{\arraystretch}{1}
 
-<!-- <!-- reset row height/line spacing --> -->
+<!-- reset row height/line spacing -->
 
-<!-- </details> -->
+</details>
 
-<!-- <br> -->
+<details>
 
-<!-- <details> -->
-
-<!-- <summary> -->
+<summary>
 
@@
 
-<!-- </summary> -->
+</summary>
 
-<!-- \renewcommand{\arraystretch}{0.6} -->
+\renewcommand{\arraystretch}{0.6}
 
-<!-- <!-- decrease line spacing for the table --> -->
+<!-- decrease line spacing for the table -->
 
@@
 
-<!-- </summary> -->
+</summary>
 
@@
 
-<!-- </details> -->
+</details>
 
-<!-- </br> -->
+## References
 
-------------------------------------------------------------------------
+```{r}
+#| echo: false
+#| results: asis
 
-</br>
+bib_entries <- read.bib("packages.bib")
+all_keys <- names(bib_entries)
 
-::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
+# Create a hidden div with all the references
+cat('<div style="display:none;">')
+for (key in all_keys) {
+  cat(sprintf('[@%s] ', key))
+}
+cat('</div>')
+```
+
+::: {#refs}
 :::

### 2023-08-07 tutorial/index.qmd
var info updated

---
title: "Illustration: Feature Based Time Series Clustering in Psychology [tsFeatureExtracR]"
author:
  - name: Jannis Kreienkamp
    orcid: 0000-0002-1831-5604
    email: j.kreienkamp@rug.nl
    corresponding: true
    # affiliations:
    #   - ref: rug
  - name: Maximilian Agostini
    orcid: 0000-0001-6435-7621
    # affiliations:
    #   - ref: rug
  - name: Rei Monden
    orcid: 0000-0003-1744-5447
    # affiliations:
    #   - ref: hu
  - name: Kai Epstude
    orcid: 0000-0001-9817-3847
    # affiliations:
    #   - ref: rug
  - name: Peter de Jonge
    orcid: 0000-0002-0866-6929
    # affiliations:
    #   - ref: rug
    #   - ref: icpe
  - name: Laura F. Bringmann
    orcid: 0000-0002-8091-9935
    # affiliations:
    #   - ref: rug
    #   - ref: icpe
# affiliations:
#   - id: rug
#     name: University of Groningen
#   - id: hu
#     name: Hiroshima University
#   - id: icpe
#     name: ICPE
---

```{r}
#| label: setup
#| include: FALSE

knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors"
  )
```

<br>

Psychological researchers and practitioners collect increasingly complex time series data aimed at identifying differences between the developments of participants or patients. Past research has proposed a number of `dynamic features' that describe meaningful developmental patterns for psychological data (e.g., instability, inertia, linear trend). We propose that feature-based clustering offers an ideal fit for this type of data. We cluster participants based on the dynamic measures directly using common clustering algorithms, which means that the approach is flexible, transparent, and well-grounded.

## The Illustration

This illustration accompanies the publication "A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series" and provides a hands-on tutorial-style code documentation. We provide annotated R code of all analysis steps as well as an associated R package that collects several of the functions we created along the way.

### The Structure

We introduce the feature-based clustering approach and illustrate the utility of the method with real-world empirical data that highlight common ESM challenges of multivariate conceptualizations, structural missingness, and nonlinear trends. We use the data to showcase the main steps of input selection, feature extraction, feature reduction, feature clustering, and cluster evaluation.

1. [Input variables](01-input-variables.qmd)
2. [Feature extraction](02-feature-extraction.qmd)
3. [Feature reduction](03-feature-reduction.qmd)
4. [Clustering](04-clustering.qmd)
5. [Evaluation](05-evaluation.qmd)
6. [About](about.qmd)

You can navigate through this illustration using the sidebar on the left or go through the individual steps using the forward and backward buttons at the bottom of each page. We close this illustration we a short about section that links important resources and software information.

### 2023-08-07 scripts/functions/tsFeatureExtractor.R

@@ -1,10 +1,10 @@
 
 #' Feature Extraction Function
 #'
-#' This function extracts features from a given dataset, grouping it by an identifier variable, and summarizing it by a set of measures 
+#' This function extracts features from a given dataset, grouping it by an id variables, and summarizing it into the time series features 
 #' for each item in the given list of items. It generates univariate features such as mean, median, standard deviation, 
 #' mean absolute deviation, root mean square of successive differences, and mean absolute change.
-#' It also generates multivariate and time-based features, applying linear and non-linear trends using generalized additive models (GAMs),
+#' It also generates multivariate, applying linear and non-linear trends using generalized additive models (GAMs),
 #' and calculates auto-regressive components and their confidence intervals.
 #' It standardizes the features, and outputs a list containing raw and standardized features, as well as the GAM models.
 #'
@@ -14,10 +14,10 @@
 #' @param tid A character string specifying the name of the time identifier variable.
 #'
 #' @return A list containing:
-#' features: A data frame of raw features.
-#' featuresZ: A data frame of standardized features.
-#' featuresZMat: A data frame of standardized features with person identifiers as row names.
-#' gam: A list of fitted GAM models for each person.
+#'    * features: A data frame of raw features.
+#'    * featuresZ: A data frame of standardized features.
+#'    * featuresZMat: A data frame of standardized features with person identifiers as row names.
+#'    * gam: A list of fitted GAM models for each person.
 #'
 #' @examples
 #' \dontrun{
@@ -25,11 +25,6 @@
 #' }
 #' @export
 featureExtractor <- function(data, items, pid, tid) {
-  # for testing:
-  # data <- featData
-  # pid <- "ID" #c("ID", "PID")
-  # tid <- "TIDnum"
-  # items <- varNamS123PCA[!varNamS123PCA %in% idVars]
   
   # Change the names of the identifier columns to "ID" and "TIDnum"
   names(data)[names(data) == pid] <- "ID"
@@ -214,38 +209,153 @@ featureExtractor <- function(data, items, pid, tid) {
   return(out)
 }
 
+#' Feature Imputation Function
+#'
+#' This function performs imputation on the features data derived from the featureExtractor function. 
+#' It uses the Multiple Imputation by Chained Equations (MICE) algorithm to handle missing data. 
[…]

### 2023-08-07 Supplemental Material A - Illustration/Supplemental-Material-A-Illustration.tex
osf and github added

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper]{apa7}

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome}
\usepackage[normalem]{ulem}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{afterpage}
\usepackage{float}
\usepackage{titling}
\usepackage{censor}
\usepackage{tcolorbox}

% formatting links in the PDF file
\hypersetup{
pdfpagemode={UseOutlines},
bookmarksopen=true,
bookmarksopenlevel=0,
hypertexnames=false,
colorlinks   = true, %Colours links instead of ugly boxes
urlcolor     = blue, %Colour for external hyperlinks
linkcolor    = blue, %Colour of internal links
citecolor   = cyan, %Colour of citations
pdfstartview={FitV},
unicode,
breaklinks=true,
}

% language settings
\DeclareLanguageMapping{american}{american-apa}

% add reference library file
\addbibresource{../references.bib}

% Title and header
\title{Supplemental Information A: Illustration}
\shorttitle{SI A: Illustration}
%\author{Jannis Kreienkamp, Laura F. Bringmann, Raili F. Engler, Peter de Jonge, Kai Epstude}
\author{[authors masked for peer review]}

% set indentation size
\setlength\parindent{1.27cm}

% adapt table and figure labels
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

% Start of the main document:
\begin{document}

% add title information (incl. title page and abstract)
\begin{titlepage}
	{\noindent\Large Supplementary Information for \par}
	\vspace{0.5cm}
	{\noindent\Large Need Fulfillment During Intergroup Contact: Three Experience Sampling Studies\par}
	\vspace{1.5cm}
	{\noindent\LARGE\bfseries \thetitle \par}
	\vspace{2cm}
	{\noindent\Large\itshape \theauthor \par}
	\vfill
	%\noindent Corresponding Author: Jannis Kreienkamp\par
	%\noindent E-mail: j.kreienkamp@rug.nl\par
	\noindent Corresponding Author: [masked for peer review]\par
	\noindent E-mail: [masked for peer review]\par
	\vfill

    % Bottom of the page
	{\noindent Last updated: \today\par}
\end{titlepage}

% add title again on page 1 (after title page)
\begin{center}
   \textbf{\thetitle} 
\end{center}

This supplementary information documents the methodology and results of all three studies. As part of our open supplemental materials, we share the full RMarkdown file which offers an annotated version of our all analysis steps. This file offers a transparent and reproducible analysis code, which includes full author commentary. We rendered the RMarkdown as an interactive HTML file, which we host as part of our open GitHub repository. We recommend the rendered version for almost all readers (full R code is also available via the rendered version).

\vspace{.5cm}
\begin{tcolorbox}
    \vspace{0.2cm} \centering 
    \href{https://janniscodes.github.io/ts-feature-clustering-illustration/}{https://janniscodes.github.io/ts-feature-clustering-illustration/}
    \vspace{0.2cm} 
\end{tcolorbox}

For readers interested in the raw files, the raw RMarkdown file is available in our OSF repository \citep[see][]{KreienkampMasked-clust-osf} and can also be accessed as part of the full GitHub repository \citep[][]{KreienkampMasked-clust-github}.

\printbibliography

\end{document}

### 2023-08-07 manuscript.tex
osf and github added

@@
 % Authors
-\author[*,1,3]{Jannis Kreienkamp}
-\author[1,3]{Maximilian Agostini}
-\author[1,3]{Kai Epstude}
-\author[2,3]{Rei Monden}
+\author[*,1]{Jannis Kreienkamp}
+\author[1]{Maximilian Agostini}
+\author[1]{Kai Epstude}
+\author[2]{Rei Monden}
 \author[1,3]{Peter de Jonge}
@@
 \affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
-\affil[3]{Author order TBD}
+\affil[3]{ICPE}
 
@@
 The authors received no specific funding for this work.
-Materials and  software are available at \url{https://janniscodes.github.io/migration-trajectories/}  \citep{KreienkampTBD}.
-Materials, data, and code are available at \url{https://osf.io/TBA} \citep{KreienkampTBD}.
+Materials and  software are available at \url{https://janniscodes.github.io/migration-trajectories/} \citep{KreienkampMasked-clust-github}.
+Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{KreienkampMasked-clust-osf}.
 
@@
     \textit{Open Science Practices:}
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
-    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements\\
+    \vspace{1em}
+    Materials and software are available as part of our project GitHub repository \citep{KreienkampMasked-clust-github}.
+The full illustration code is also available as a tutorial-style website \citep{KreienkampMasked-illustration-github}.
+Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{KreienkampMasked-clust-osf}.
 }

### 2023-08-07 sections/07ClusterEvaluationInterpretation.Rmd
reference fixes

@@
 clusterVariableGrid <- featCluster %>% 
+  filter(feature != "n") %>%
   ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +

### 2023-08-07 tutorial/about.qmd
tutorial about updated

@@
 
-This illustration accompanies the publication "A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series". As such, the illustration is not ...
+This illustration accompanies the publication "A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series".
+The aim of this illustration website is to offer a transparent, open, and accessible repository providing annotated code for all analysis steps.
 
@@
 
-short open data description here. Link to OSF
+We provide all data necessary to replicate the full results of the main manuscript as well as this illustration.
+We specifically provide a minial reproducible dataset that includes all variables necessary to perform the analysis, while still protecting the privacy and confidentiality of the participants.
+All data is available as part of [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].
 
@@
 
-short open code description here. Link to GitHub
+The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com) [@KreienkampMasked-clust-github].
+A minimal version of the code as well as an archive file are available in [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].
+For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package.
+This package is available as its own GitHub repository and can be downloaded with the `devtools` package as `tsFeatureExtracR`: `devtools::install_github("JannisCodes/tsFeatureExtracR")` [@KreienkampBlinded2023-tsFeatureExtracR].
 

### 2023-08-07 Supplemental Material A - Illustration/Supplemental-Material-A-Illustration.tex

@@
 % add reference library file
-\addbibresource{../references.bib}
+\addbibresource{../referencesZotero.bib}
 
 % Title and header
-\title{Supplemental Information A: Illustration}
-\shorttitle{SI A: Illustration}
+\title{Supplemental Information A: R Code Illustration}
+\shorttitle{SI A: R Code Illustration}
 %\author{Jannis Kreienkamp, Laura F. Bringmann, Raili F. Engler, Peter de Jonge, Kai Epstude}
@@
 
-This supplementary information documents the methodology and results of all three studies.
-As part of our open supplemental materials, we share the full RMarkdown file which offers an annotated version of our all analysis steps.
-This file offers a transparent and reproducible analysis code, which includes full author commentary.
-We rendered the RMarkdown as an interactive HTML file, which we host as part of our open GitHub repository.
+This supplementary information documents our tutorial-style illustration website.
+As part of our open supplemental materials, we share the full illustration as an interactive tutorial-style website.
+This file offers transparent and reproducible analysis code, which includes full author commentary.
+We rendered the Quarto project as an interactive HTML website, which we host as part of our open GitHub repository.
 We recommend the rendered version for almost all readers (full R code is also available via the rendered version).
@@
 
-For readers interested in the raw files, the raw RMarkdown file is available in our OSF repository \citep[see][]{KreienkampMasked-clust-osf} and can also be accessed as part of the full GitHub repository \citep[][]{KreienkampMasked-clust-github}.
+For readers interested in the raw files, the raw Quarto files are available in our OSF repository \citep[see][]{KreienkampMasked-clust-osf} and can also be accessed as part of the full GitHub repository \citep[][]{KreienkampMasked-clust-github}.
 

### 2023-08-07 manuscript.tex

@@
 We provide a short introduction of each feature below.
-Beyond the operationalizations we chose for our case study, we provide an R-function that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures} in our GitHub repository (see the \texttt{featureExtractor} function).
+Beyond the operationalizations we chose for our case study, we collected the R-functions we created for the analyses as an R package that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures}.
+All functions are available as part of the package GitHub repository (see the \texttt{featureExtractor()} function; \citealp{KreienkampBlinded2023-tsFeatureExtracR}) and are annotated as part of our tutorial-style illustration (see Supplemental Material A).
 

### 2023-08-07 tutorial/05-evaluation.qmd
tutorial updated

@@
 #| label: tbl-cluster-performance
+#| fig-cap: Cluster Performance for the k = 2–10
 
@@
       row.names = FALSE,
-      col.names = kmeans_metrics_names,
-      caption = "Cluster Performance for the $k$ = 2–10") %>%
+      col.names = kmeans_metrics_names) %>%
   kable_classic(
@@
   ) %>%
-  scroll_box(width = "100%", height = "500px")
+  scroll_box(width = "100%")
 ```

### 2023-08-07 tutorial/about.qmd
tutorial updated

@@
 For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package.
-This package is available as its own GitHub repository and can be downloaded with the `devtools` package as `tsFeatureExtracR`: `devtools::install_github("JannisCodes/tsFeatureExtracR")` [@KreienkampBlinded2023-tsFeatureExtracR].
+This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz) [@KreienkampBlinded2023-tsFeatureExtracR].
 

### 2023-08-07 tutorial/index.qmd
tutorial updated

@@
 author:
-  - name: Jannis Kreienkamp
-    orcid: 0000-0002-1831-5604
-    email: j.kreienkamp@rug.nl
-    corresponding: true
-    # affiliations:
-    #   - ref: rug
-  - name: Maximilian Agostini
-    orcid: 0000-0001-6435-7621
-    # affiliations:
-    #   - ref: rug
-  - name: Rei Monden
-    orcid: 0000-0003-1744-5447
-    # affiliations:
-    #   - ref: hu
-  - name: Kai Epstude
-    orcid: 0000-0001-9817-3847
-    # affiliations:
-    #   - ref: rug
-  - name: Peter de Jonge
-    orcid: 0000-0002-0866-6929
-    # affiliations:
-    #   - ref: rug
-    #   - ref: icpe
-  - name: Laura F. Bringmann
-    orcid: 0000-0002-8091-9935
-    # affiliations:
-    #   - ref: rug
-    #   - ref: icpe
+  - name: Masked for Peer Review
+  - name: ██████ ██████████
+    orcid: 0000-0000-0000-0000
+  - name: ██████████ ████████
+    orcid: 0000-0000-0000-0000
+  - name: ███ ██████
+    orcid: 0000-0000-0000-0000
+  - name: ███ ███████
+    orcid: 0000-0000-0000-0000
+  - name: █████ ██ █████
+    orcid: 0000-0000-0000-0000
+  - name: █████ ██ █████████
+    orcid: 0000-0000-0000-0000
+  # - name: Jannis Kreienkamp
+  #   orcid: 0000-0002-1831-5604
+  #   email: j.kreienkamp@rug.nl
+  #   corresponding: true
+  #   # affiliations:
+  #   #   - ref: rug
+  # - name: Maximilian Agostini
+  #   orcid: 0000-0001-6435-7621
+  #   # affiliations:
+  #   #   - ref: rug
+  # - name: Rei Monden
+  #   orcid: 0000-0003-1744-5447
+  #   # affiliations:
+  #   #   - ref: hu
+  # - name: Kai Epstude
+  #   orcid: 0000-0001-9817-3847
+  #   # affiliations:
+  #   #   - ref: rug
+  # - name: Peter de Jonge
+  #   orcid: 0000-0002-0866-6929
+  #   # affiliations:
+  #   #   - ref: rug
+  #   #   - ref: icpe
+  # - name: Laura F. Bringmann
+  #   orcid: 0000-0002-8091-9935
+  #   # affiliations:
+  #   #   - ref: rug
+  #   #   - ref: icpe
 # affiliations:

### 2023-08-07 Supplemental Material A - Illustration/Supplemental-Material-A-Illustration.tex

@@
 linkcolor    = blue, %Colour of internal links
-citecolor   = cyan, %Colour of citations
+citecolor   = black, %Colour of citations
 pdfstartview={FitV},
@@
 % Title and header
-\title{Supplemental Information A: R Code Illustration}
-\shorttitle{SI A: R Code Illustration}
+\title{Supplemental Information A: Illustration R Code}
+\shorttitle{SI A: Illustration R Code}
 %\author{Jannis Kreienkamp, Laura F. Bringmann, Raili F. Engler, Peter de Jonge, Kai Epstude}
@@
 	\vspace{0.5cm}
-	{\noindent\Large Need Fulfillment During Intergroup Contact: Three Experience Sampling Studies\par}
+	{\noindent\Large A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series\par}
 	\vspace{1.5cm}

### 2023-08-07 tutorial/02-feature-extraction.qmd
tutorial updated

@@
 
-For our illustration, we rely on the work of researchers in the ESM literature, where different research teams have proposed a number meaningful measures to capture time series features that are relevant to psychological time series [@kuppens2017;@krone2018;dejonckheere2019;@wendt2020].
+For our illustration, we rely on the work of researchers in the ESM literature, where different research teams have proposed a number meaningful measures to capture time series features that are relevant to psychological time series [@kuppens2017;@krone2018;@dejonckheere2019;@wendt2020].
 The main article discusses the individual time series features, their conceptual interpretation, as well as the technical calculation in more detail.
@@
 We provide the `featureExtractor()` function to extract a number of measurements for each time series aspect.
+The function is also available as part of the R package that bundles a number of functions we used for the illustration ([Download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz)).
 The time series feature operationalizations we chose for our illustration here are marked in bold.

### 2023-08-07 tutorial/05-evaluation.qmd
tutorial updated

@@
 #| label: tbl-cluster-performance
-#| fig-cap: Cluster Performance for the k = 2–10
+#| tbl-cap: Cluster Performance for k = 2–10
 
@@
       escape = FALSE,
-      booktabs = T,
+      booktabs = TRUE,
       align = "c",
@@
 #| fig-height: 18
-#| fig-cap: Note that n and discrimination are out of cluster variables.
+#| fig-cap: Cluster comparison by feature and variable. Note that n and discrimination are out of cluster variables.
 
@@
 In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
-We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics [@bringmann2018c;eronen2021a].
+We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics [@bringmann2018c;@eronen2021a].
 
@@
 #| fig-height: 18
+#| fig-cap: Average cluster trajectories (using LM and GAM)
 
@@
 #| warning: false
+#| fig-cap: Cluster comparison MAD across variables
 
@@
 feature_plot_list$mad
-
-feature_ttest_apa(results_df, "mean", "EvDayDiscr.post", "html")
 ```

### 2023-08-07 tutorial/about.qmd
tutorial updated

@@
 We specifically provide a minial reproducible dataset that includes all variables necessary to perform the analysis, while still protecting the privacy and confidentiality of the participants.
-All data is available as part of [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].
+All data is available as part of [our Open Science repository](https://osf.io/j8dzv/?view_only=c765efa51e6545e588b29999611eb789) [@KreienkampMasked-clust-osf].
 
@@
 The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com) [@KreienkampMasked-clust-github].
-A minimal version of the code as well as an archive file are available in [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].
+A minimal version of the code as well as an archive file are available in [our Open Science repository](https://osf.io/j8dzv/?view_only=c765efa51e6545e588b29999611eb789) [@KreienkampMasked-clust-osf].
 For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package.

### 2023-08-09 tutorial/02-feature-extraction.qmd
typo

@@
 
-_Note:_ Beyond the time series features described in this illustration, the [`tsfesh` python package](https://tsfresh.readthedocs.io/en/latest/) allows users to extract almost 800 time series features [@christ2018].
+_Note:_ Beyond the time series features described in this illustration, the [`tsfresh` python package](https://tsfresh.readthedocs.io/en/latest/) allows users to extract almost 800 time series features [@christ2018].
 The package is well maintained and offers clear documentation of how each time series feature is calculated.

### 2023-08-11 tutorial/02-feature-extraction.qmd
tutorial update

@@
 
+::: {.grid}
+
+::: {.g-col-4}
+[![](images/logo_tsfeatureextractr.png)](https://github.com/JannisCodes/tsFeatureExtracR)
+[Download the R package here](https://tsfeatureclustr.com/assets/tsFeatureExtracR_0.1.0.tar.gz)
+:::
+
+::: {.g-col-8}
 The selection of time series features should always be driven by the research questions and field-specific conventions, but for our illustration we chose six time series features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity.
 We provide the `featureExtractor()` function to extract a number of measurements for each time series aspect.
-The function is also available as part of the R package that bundles a number of functions we used for the illustration ([Download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz)).
+The function is also available as part of the R package that bundles a number of functions we used for the illustration.
+:::
+
+:::
+
 The time series feature operationalizations we chose for our illustration here are marked in bold.

### 2023-08-11 tutorial/about.qmd
tutorial update

@@
 For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package.
-This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz) [@KreienkampBlinded2023-tsFeatureExtracR].
+This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://tsfeatureclustr.com/assets/tsFeatureExtracR_0.1.0.tar.gz) [@KreienkampBlinded2023-tsFeatureExtracR].
 

### 2023-08-11 tutorial/index.qmd
tutorial update

@@
 ---
-title: "Illustration: Feature Based Time Series Clustering in Psychology [tsFeatureExtracR]"
+title: "Illustration: Feature Based Time Series Clustering in Psychology [tsFeatureClustR]"
 author:

### 2023-08-11 manuscript.tex

@@
 % formatting links in the PDF file
-\definecolor{myurlcolor}{HTML}{1a0dab}
+\definecolor{googleurlcolor}{HTML}{1a0dab}
 \hypersetup{
@@
 colorlinks   = true, %Colours links instead of ugly boxes
-urlcolor     = myurlcolor, %blue,Colour for external hyperlinks
+urlcolor     = googleurlcolor, %blue,Colour for external hyperlinks
 linkcolor    = black, %blue, Colour of internal links
@@
 \author[1]{Maximilian Agostini}
+\author[2]{Rei Monden}
 \author[1]{Kai Epstude}
-\author[2]{Rei Monden}
 \author[1,3]{Peter de Jonge}

### 2023-08-11 Supplemental Material A - Illustration/Supplemental-Material-A-Illustration.tex
references updated

@@
 % formatting links in the PDF file
+\definecolor{googleurlcolor}{HTML}{1a0dab}
 \hypersetup{
@@
 colorlinks   = true, %Colours links instead of ugly boxes
-urlcolor     = blue, %Colour for external hyperlinks
-linkcolor    = blue, %Colour of internal links
-citecolor   = black, %Colour of citations
+urlcolor     = googleurlcolor, %blue,Colour for external hyperlinks
+linkcolor    = black, %blue, Colour of internal links
+citecolor   = black, % cyan, Colour of citations
 pdfstartview={FitV},
@@
 This file offers transparent and reproducible analysis code, which includes full author commentary.
-We rendered the Quarto project as an interactive HTML website, which we host as part of our open GitHub repository.
+We rendered the Quarto project as an interactive HTML website, which we host as part of our open GitHub repository \citep{KreienkampMasked-illustration-github}.
 We recommend the rendered version for almost all readers (full R code is also available via the rendered version).
@@
     \vspace{0.2cm} \centering 
-    \href{https://janniscodes.github.io/ts-feature-clustering-illustration/}{https://janniscodes.github.io/ts-feature-clustering-illustration/}
+    \href{https://janniscodes.github.io/ts-feature-clustering-illustration/}{www.tsFeatureClustR.com}
     \vspace{0.2cm} 
@@
 
-For readers interested in the raw files, the raw Quarto files are available in our OSF repository \citep[see][]{KreienkampMasked-clust-osf} and can also be accessed as part of the full GitHub repository \citep[][]{KreienkampMasked-clust-github}.
+For readers interested in the raw files, the raw Quarto files are available in our OSF repository \citep[see][]{KreienkampMasked-clust-osf} and can also be accessed as part of the project GitHub repository \citep[][]{KreienkampMasked-clust-github}.
 

### 2023-08-11 TitlePage.tex
references updated

--- 2022-05-12 TitlePage.tex
@@
     colorlinks   = true, %Colours links instead of ugly boxes
-urlcolor     = blue, %Colour for external hyperlinks
-linkcolor    = blue, %Colour of internal links
-citecolor   = cyan, %Colour of citations
+    urlcolor     = googleurlcolor, %blue,Colour for external hyperlinks
+    linkcolor    = black, %blue, Colour of internal links
+    citecolor   = black, % cyan, Colour of citations
     pdfstartview={FitV},
@@
 
-% add reference library file
-\addbibresource{references.bib}
+% custom open science badge height
+\newlength{\badgeheight}
+\setlength{\badgeheight}{1em}
 
 % add reference library file
-\addbibresource{references.bib}
+\addbibresource{referencesZotero.bib}
 
 % Title and header
-\title{Migration Experience Trajectories: Subtitle TBD}
-\shorttitle{Migration Experience Trajectories}
+\title{A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series\\}
+\shorttitle{feature-based time series clustering}
 
 % Authors
-\author[*,1,3]{Jannis Kreienkamp}
+\author[*,1]{Jannis Kreienkamp}
+\author[1]{Maximilian Agostini}
+\author[2]{Rei Monden}
+\author[1]{Kai Epstude}
+\author[1,3]{Peter de Jonge}
 \author[1,3]{Laura F. Bringmann}
-\author[1,3]{Kai Epstude}
-\author[1,3]{Maximilian Agostini}
-\author[1,3]{Peter de Jonge}
-\author[2,3]{Rei Tendeiro-Monden}
 \affiliation{\hfill}
@@
 \affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
-\affil[3]{Author order TBD [currently in first name alphabetical order]}
+\affil[3]{Interdisciplinary Center Psychopathology and Emotion Regulation}
 
@@
    \addORCIDlink{* Jannis Kreienkamp}{0000-0002-1831-5604}\\
-   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}\\
    \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
+   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}\\
    \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
    \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
-   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}
+   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}
 
@@
 The authors received no specific funding for this work.
-Materials and  software is available at \url{https://janniscodes.github.io/migration-trajectories/}  \citep{KreienkampTBD}.
-Protocols, materials, data, and code are available at \url{https://osf.io/TBA} \citep{KreienkampTBD}.
-The preregistration of our analysis can be accessed as part of our Open Science Framework repository \citep{KreienkampTBD}.
+Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{Kreienkamp-clust-osf}.
+Materials and software are also available as part of our project GitHub repository \citep{Kreienkamp-clust-github}.
+The full illustration code is also available as a tutorial-style website at \href{https://tsfeatureclustr.com}{www.tsFeatureClustR.com} \citep{Kreienkamp-illustration-github}.
 
 Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).
-E-mail: j.kreienkamp@rug.nl}
+E-mail: \href{mailto:j.kreienkamp@rug.nl}{j.kreienkamp@rug.nl}
+
+\vspace{1em}
+\textit{Open Science Practices:}
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
+
+}
 
@@
 % add title information (incl. title page and abstract)
-\maketitle
+%\maketitle
+
+\makeatletter
+\vspace*{2em}
+\begin{center}
+\noindent \textbf\@title
+\vspace{1em}
+\noindent \@author
+\end{center}
+\vspace{2em}
+\begin{center}%
+    \textbf\acksname%
+\end{center}%
+\vspace{-1em}
+\protect\raggedright
+\setlength{\parindent}{0.5in}
+\indent\par\@acks%
 

### 2023-08-11 manuscript.tex
references updated

@@
     \vspace{1em}
-    Materials and software are available as part of our project GitHub repository \citep{KreienkampMasked-clust-github}.
-The full illustration code is also available as a tutorial-style website \citep{KreienkampMasked-illustration-github}.
     Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{KreienkampMasked-clust-osf}.
+Materials and software are also available as part of our project GitHub repository \citep{KreienkampMasked-clust-github}.
+The full illustration code is also available as a tutorial-style website at \href{https://tsfeatureclustr.com}{www.tsFeatureClustR.com} \citep{KreienkampMasked-illustration-github}.
 }

### 2023-08-11 sections/03FeatureExtraction.Rmd
typo fixed

@@
 That is, self-similarity describes how much a measurement carries over to future measurements.
-One important self-similarity in psychological time series is \texit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}.
+One important self-similarity in psychological time series is \textit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}.
 If inertia is high a development tends to stay in a certain state.

### 2023-08-11 manuscript_unmasked.tex
unmasked version added

--- 2023-08-11 manuscript.tex
@@
 % define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
-\documentclass[man, 12pt, a4paper, mask, floatsintext]{apa7}
+\documentclass[man, 12pt, a4paper, floatsintext]{apa7}
 
@@
 \affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
-\affil[3]{ICPE}
+\affil[3]{Interdisciplinary Center Psychopathology and Emotion Regulation}
 
@@
 The authors received no specific funding for this work.
-Materials and  software are available at \url{https://janniscodes.github.io/migration-trajectories/} \citep{KreienkampMasked-clust-github}.
-Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{KreienkampMasked-clust-osf}.
+Materials and  software are available at \url{https://janniscodes.github.io/migration-trajectories/} \citep{Kreienkamp-clust-github}.
+Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{Kreienkamp-clust-osf}.
 
@@
     \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
-    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements\\
-    \vspace{1em}
-    Materials and software are available as part of our project GitHub repository \citep{KreienkampMasked-clust-github}.
-The full illustration code is also available as a tutorial-style website \citep{KreienkampMasked-illustration-github}.
-Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{KreienkampMasked-clust-osf}.
+    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
 }
@@
 
-\input{latex-results/01DataDescription.tex}
+Matching these requirements, the data set we use consists of three
+studies that followed migrants who had recently arrived in the
+Netherlands in their daily interactions with the Dutch majority group
+members \citep[for the data set see][]{Kreienkamp2022b}. After a general
+migration-focused pre-questionnaire, participants were invited twice per
+day to report on their (potential) interactions with majority group
+members for at least 30 days. The short ESM surveys were sent out at
+around lunch (12pm) and dinner time (7 pm). After the 30-day study
+period participants filled in a post-questionnaire that mirrored the
+pre-questionnaire. Participants received either monetary compensation or
+partial course credits based on the number of surveys they completed.
+
+The original studies included 207 participants (\(N_{S1}=\) 23,
+\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 ESM
+measurements. Each of the studies focused on recently arrived
+first-generation migrants and each study included a number of
+idiosyncratic variables relevant for the broader research collective.
+For our empirical example, we focus on the variables that were collected
+during the ESM surveys and were available in all three studies. Variable
+selection and preparation are described as part of the illustration
+below but for additional methodological details about the study setup
+see \citet[][]{Kreienkamp2022b}. Each study was approved by the ethics
+board of the university of origin and all participants gave informed
+consent.
 
@@
 
-\input{latex-results/02InputVariables.tex}
+For our illustration, we include 12 variables that were measured as part
+of the ESM surveys in all three studies and captured information about
+the participant's interactions, as well as cognitive-, emotional-, and
+motivational self in relationship with the majority group (see
+\tblref{tab:var_selection} for an overview). We chose these aspects in
+particular because (1) the interaction-specific information exemplified
+the structural missingness issue of modern ESM data and (2) the
+motivational, emotional, and cognitive experience offered a diverse
+conceptualization of migration experience (beyond behavioral
+measurements) that is becoming more common in the literature
+\citep[][]{Kreienkamp2022d}. The breadth of the included variables also
+showcases the utility of the method for a growing body of literature
+that considers heterogeneous and complex concepts. As a result, the
+number of included variables is also on the higher end for psychological
+concepts and additionally allows us to showcase the efficiency benefits
+of the method and offers a reasonable use case for the optional feature
+reduction step.
+
+\input{tables/var_selection}
+
+Once the important variables have been selected, the data needs to be
+prepared for the analysis steps. Importantly, this not only means
+validating and cleaning the data (e.g., re-coding, removing duplicate or
+unwanted measurements) but also making the time-series comparable. Two
+important steps are making the time-frames and response scales
+comparable across participants --- for example, by choosing a time frame
+that is common to most participants and standardizing the participants'
+responses
+\citep['data exclusion' and 'data transformation' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.
+
+In our illustration data set, the studies differed substantially in the
+maximum length of participation (\(\text{max}(t_{S1})=\) 63,
+\(\text{max}(t_{S2})=\) 69, \(\text{max}(t_{S3})=\) 155). This was
+likely due to the option to continue participation without compensation
+in the latter study. To make the three studies comparable in
+participation and time frames, we iteratively removed all measurement
+occasions and participants that had more than 45\% missingness
+\citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
+This procedure led to a final sample of 157 participants, who jointly
+produced 8,132 measurements. Importantly, both the participant
+response-patterns and the time frame were now substantially more
+comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
+\(\text{max}(t_{S3})=\) 67). Full methodological details are available
+in Online Supplemental Material A, but basic item information,
+descriptives, and correlations are also available in
+\tblref{tab:descrLong}.
+
+\input{tables/descrLong}
 
@@
 Beyond the operationalizations we chose for our case study, we collected the R-functions we created for the analyses as an R package that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures}.
-All functions are available as part of the package GitHub repository (see the \texttt{featureExtractor()} function; \citealp{KreienkampBlinded2023-tsFeatureExtracR}) and are annotated as part of our tutorial-style illustration (see Supplemental Material A).
+All functions are available as part of the package GitHub repository (see the \texttt{featureExtractor()} function; \citealp{Kreienkamp2023-tsFeatureExtracR}) and are annotated as part of our tutorial-style illustration (see Supplemental Material A).
 
@@
 
-\input{latex-results/03FeatureExtraction.tex}
+\paragraph{Central tendency.}
+
+The central tendency refers to the statistical measures that represent
+the ``typical'' or ``average'' of a set of data. The most common
+measures of central tendency are the mean, median, and mode
+\citep{weisberg1992}. As a familiar statistic from probability theory,
+the central tendency sits at the heart of many fundamental questions
+about psychological time series. Researchers might, for example, be
+interested in whether ``Over a one-month period, are some people happier
+than others?''
+
+For the central tendency feature of our illustration we chose the robust
+\textit{median}, which can avoid potential issues with non-normally
+distributed time series responses or outliers \citep{weisberg1992}. To
+calculate the \textit{median} (\(M\)), we let \(X_{ij}\) be the ordered
+list of values from the time series of variable \(j\) for participant
+\(i\). The calculation depends on whether the number of measurements in
+a time series \(n\) is odd or even.
+
+\begin{equation} \label{eq:median}
+  M(X_{ij}) = 
+    \begin{cases}
+      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
+      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
+    \end{cases}
+\end{equation}
+
+\paragraph{Variability.}
+
+Variability captures the degree to which a set of data differs from the
+central tendency and is sometimes also referred to as the dispersion or
+spread of the data \citep{weisberg1992}. In time series analyses,
+variability is conceptually important because information about the
+distribution and diversity of the data has been found to be indicative
+of worse psychological states \citep{myin-germeys2018, helmich2021}.
+Person-level differences of ESM measurements have, for example, been
+associated with higher levels of psycho-pathological recurrences among
+depression patients \citep{timm2017}. As such, psychological researchers
+and practitioners are often empirically interested in between-person
+differences in variability. Researchers on polarization and
+radicalization might for example ask: ``Are people settled in their
+attitudes towards migrants or do they vary across the measurement
+period?''
+
+For our illustration data, we chose to capture the time series
+variability with the \textit{Median Absolute Deviation} (\(MAD\)), where
+we calculate the \textit{median} (\(M\); calculated as in
+\equatref{eq:median}) for the absolute deviations of measurement \(x\)
+at time point \(t\) for participant \(i\) and variable \(j\) from the
+median of that time series \(X\). We again chose the robust statistic
+because the Median-based measure is less affected by non-normal
+distributions and extreme values or outliers compared to other measures
+of variability like the standard deviation \citep{weisberg1992}
+
+\begin{equation} \label{eq:mad}
+  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
+\end{equation}
+
+\paragraph{Instability.}
+
+Instability captures the average change between two consecutive
+measurements \citep{ebner-priemer2009}. While instability is
+conceptually related to the variability feature, variability does not
+take into account temporal dependency, whereas instability looks at the
+`jumpy-ness' of the data over time. In other words, variability reflects
+the range or diversity of values in a time series data, while
+instability reflects the fluctuation or inconsistency in a time series
+data over time \citep{trull2008}. For example, if a person has rapid and
+extreme changes in mood their mood is highly unstable, while if a
+person's mood responses span a wide range over the entire study period,
+their mood is highly variable \citep{jahng2008}. Within psychological
+time series, instability measurements have especially been important in
+the research of borderline personality disorder \citep{trull2008} and
+suicidality \citep{kivela2022}, but also in understanding early warning
+signals more generally \citep{wichers2019}. Conceptually, the
+instability feature, thus, relates to a broad range of research
+questions, including: ``What is the nature of the identification changes
+in those who start working in a new country?'' or ``Do strong daily
+fluctuations in self-esteem reflect the process of identity formation in
+adolescents?''
+
+For our data we chose the \textit{mean absolute change}
+\citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
+the average absolute difference of two consecutive measurements \(x\) at
+time points \(t\) and \(t-1\), for each time series \(X\) of participant
+\(i\) and variable \(j\).
+
+\begin{equation} \label{eq:mac}
+  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=2, \ldots, t}\left|x_{t}-x_{t-1}\right|
+\end{equation}
+
+Another common measurement of instability is the
+\textit{Mean of the Squared Successive Differences} (\(MSSD\)), which is
+often preferred where differences in magnitude are more important than
+the frequency of those changes, for example, when big shifts in time
+series are considered more impactful or when outliers are meaningful and
+need to be taken into account \citep{chatfield2003}.
+
+\paragraph{Self-similarity.}
+
+Self-similarity in time series data refers to the property of a time
+series to exhibit similar patterns of behavior over different time
+scales \citep{dmello2021}. That is, self-similarity describes how much a
+measurement carries over to future measurements. One important
+self-similarity in psychological time series is \textit{inertia} --- how
+much a measurement carries over to its next measurement
+\citep{kuppens2010, suls1998}. If inertia is high a development tends to
+stay in a certain state. Because high inertia is resistant to change, in
+emotion dynamics high inertia of negative affect has been found to be
+indicative of under-reactive systems and to be characteristic of
+psychological maladjustment \citep{kuppens2010}. In a similar vein, high
+inertia in negative affect at baseline was even predictive of the
+initial onset of depression \citep{kuppens2012}. Conceptually, inertia
+is more broadly connected to research questions such as:
+\texttt{Do\ patients\ stay\ in\ a\ negative\ mood\ for\ several\ measurements?\textquotesingle{}\textquotesingle{}\ or}Do
+migrants stay with their language practice for several days at a
+time?'\,'
+
+For our illustration case, we chose the commonly used autocorrelation or
+autoregression with a lag-1 to capture the inertia. High autocorrelation
+values can indicate high levels of inertia, while low autocorrelation
+values may indicate a more unpredictable or volatile time series
+\citep{dejonckheere2019}. The lag--1 autocorrelation \(r_{ij,1}\) looks
+at the average correlation between a measurement \(x\) and the preceding
+measurement \(x_{t-1}\) for the time series \(X\) of participant \(i\)
+and variable \(j\) with \(n\) measurements.
+
+\begin{equation} \label{eq:ar}
+  r_{ij,1} = \frac{\sum_{t=2}^{n}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
+\end{equation}
+
+Where \(\overline{x}_{ij}\) is the mean of the time series \(x_{ij}\),
+calculated as:
+
+\begin{equation} \label{eq:mean_for_ar1}
+  \overline{x}_{ij} = \frac{1}{n} \sum_{t=1}^{n} x_{ijt}
+\end{equation}
+
+\paragraph{Linear trend.}
+
+In non-stationary time series, a linear trend can be observed when there
+is a consistent increase or decrease in the data over time
+\citep{nyblom1986}. For psychological time series, researchers have, for
+example, pointed out the importance of linear trends in interpersonal
+communications \citep{vasileiadou2014}, and emotion dynamics
+\citep{oravecz2016}. Theoretically, linear trends are often considered
+the simplest way of assessing whether a psychological theory of change
+is appropriate \citep{gottman1969}. In empirical practice, linear trends
+are, thus, commonly exemplified by research questions such as ``Do
+patient symptoms improve consistently?'' or ``Does worker productivity
+decline continuously?''
+
+For the variables in our illustration data set, we chose an overall
+linear regression slope to capture the linear trend. The regression
+slope \(b_{ij}\) provides the average change from one time point \(t\)
+to the next across all measurements \(x\) of a time series \(X\) of
+participant \(i\) and variable \(j\). The specific form of the OLS slope
+formula we provide below calculates \(b_{ij}\) as the sum across all
+time points of the product of the deviation of time \(t\) from its mean
+\(\overline{t}\) and the deviation of \(x_{ij}\) from its mean
+\(\overline{x}_{ij}\) at each time point, divided by the sum across all
+time points of the square of the deviation of time from its mean
+(\(\sum(t-\overline{t})^2\)). Intuitively, the formula captures the rate
+of change of variable \(x_{ij}\) with respect to time. This slope will
+indicate how the variable \(x_{ij}\) changes over time, controlling for
+its mean value and the mean of time. If the slope is positive,
+\(x_{ij}\) increases over time; if it's negative, \(x_{ij}\) decreases
+over time.
+
+\begin{equation} \label{eq:lin}
+  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
+\end{equation}
+
+\paragraph{Nonlinearity.}
+
+Changes in psychology are not always linear, instead, nonlinearity is a
+common feature of psychological time series \citep{hayes2007}. As an
+example, episodic disorders, such as depression, are most likely best
+described as non-linear systems \citep{hosenfeld2015}. Similarly,
+patients in recovery from depression showed sudden changes in the
+improvement of depression \citep{helmich2020a}. But also substance abuse
+\citep{boker1998} or attitude changes rarely develop linearly
+\citep{vandermaas2003}. Conceptually, researchers might have research
+questions about the type of the development: ``Is the development of
+well-being a nonlinear process?'' as well as the shape and structure of
+the development: ``How many spikes in well-being did a migrant
+experience?''
+
+We summarized the nonlinear trend with the
+\textit{estimated degrees of freedom} of an empty GAM spline model. The
+\(edf\) summarizes the \textit{wiggliness} of a spline trend line
+\citep{wood2017}. The degrees of freedom of a spline model are primarily
+determined by the number of knots and the order of the spline. For
+instance, a cubic spline with \(k\) knots has \(k\)+3 degrees of freedom
+\citep{faraway2016}. However, in a penalized spline framework, which is
+commonly used for GAMs, the effective degrees of freedom can be less
+than \(k\)+3. This is because the model employs a smoothing parameter to
+control the trade-off between the complexity (flexibility) of the model
+and its fit to the data, thereby penalizing overly complex models and
+potentially reducing the effective degrees of freedom \citep{marx1998}.
+Intuitively then an edf of 1 would be equivalent to a linear
+relationship (i.e., one linear slope parameter), whereas a higher edf
+(particularly an edf \textgreater{} 2) is indicative of a non-linear
+trend. The estimated degrees of freedom are commonly based on a concept
+called `effective degrees of freedom' and can be represented as the
+trace \(tr\) (i.e., the sum of the diagonal elements) of the smoother
+matrix \(S\), a symmetric matrix that maps from the raw data to the
+smooth estimates \citep{wood2017}.
+
+\begin{equation} \label{eq:df}
+  edf = tr(S)
+\end{equation}
+
+Beyond our main features of interest, we also extracted the
+participant's number of completed ESM measurements to ensure that the
+clusters are comparable in that regard (i.e., to exclude spurious
+explanations for the cluster assignments). After the feature extraction,
+we found that about 1.40\% of the extracted features are missing across
+the 72 features per participant. This might, for example, happen if
+participants do not have two subsequent measurements with outgroup
+interactions, so that an autocorrelation with lag-1 cannot be calculated
+for the contact-specific variables. The small number of missing values
+indicates that the feature-based approach indeed largely avoids the
+structural missingness issue. However, even the few missing values can
+be an issue for some feature reduction or feature clustering algorithms.
+We, thus, impute the missing feature values with a single predictive
+mean matching imputation using the MICE library \citep[][]{buuren2011}.
+Note again that with this procedure we only need to impute an extremely
+small number of missing values as most feature calculations can use the
+available data instead.
 
@@
 
-\input{latex-results/04FeatureReduction.tex}
+For our own illustration data, we chose a feature projection method to
+reduce the dimensionality of our extracted features. We particularly
+chose the feature projection method for its broad applicability. We,
+specifically, selected the commonly used
+\textit{principal component analysis} (PCA). Some of the more
+tailor-made feature selection algorithms can be more accurate in
+reducing the feature dimensionality and might retain feature importance
+information more directly, depending on the specific data structure.
+However, PCAs have the distinct benefit that they are well-established
+within the psychometric literature \citep{jolliffe2011} and can broadly
+be applied to a wide variety of studies in an automatized manner
+\citep{abdi2010}. As our aim is to present a general illustration that
+can also be adopted across use cases, we present the workflow using a
+PCA here but we encourage users to consider more specialized methods as
+well.
+
+To use the PCA with our extracted time series features, we first
+standardize all features across participants to ensure that all features
+are weighted equally. We then enter all 72 features into the analysis.
+The PCA uses linear transformations in such a way that the first
+component captures the most possible variance of the original data
+\citep[e.g., by finding a vector that maximizes the sum of squared distances][]{jolliffe2002, abdi2010}.
+The following components will then use the same method to iteratively
+explain the most of the remaining variance while also ensuring that the
+components are linearly uncorrelated \citep{shlens2014}. In practice,
+this meant that the PCA decomposed the 72 features into 72 principal
+components but now (because of the uncorrelated linear transformations)
+the first few principal components will capture a majority of the
+variance. We can then decide how much information (i.e., variance) we
+are willing to sacrifice for a reduced dimensionality. A common rule of
+thumb is to use the principal components that jointly explain 70--90\%
+of the original variance
+\citep[i.e., cumulative percentage explained variance; e.g.,][]{jackson2003}.
+For our illustration, we select the first 27 principal components that
+explain 80\% of the variance in the original 72 features (reducing the
+dimensionality by 62.50\%). For the extracted principal components we
+save the 27 principal component scores for each participant (i.e., the
+participants' coordinates in the reduced dimensional space; PC-scores).
+
+We would like to comment on two practical matters when using principal
+components --- the amount of dimensionality reduction and the
+interpretation of the principal components. As for the expected
+dimensionality reduction, given its methodology, PCAs tend to `work
+better' at reducing dimensions with (highly) correlated variables
+\citep[e.g.,][]{jolliffe2002}. Thus, with a set of very homogeneous
+variables and features users will need fewer principal components to
+explain a large amount of variance, while a more diverse set of
+variables and features will tend to require more principal components to
+capture the same amount of variance \citep[e.g.,][]{abdi2010}. Our 27
+principal components are still a relatively high number of variables but
+this is not surprising as we chose a diverse conceptualization and a
+diverse set of time series features. As for interpretability, PCA allows
+users to extract information on the meaning of the principal components.
+In particular, because the principal components are linear combinations
+of the original features, users can extract the relative importance of
+each feature for the extracted principal components (i.e., the
+eigenvectors). While this can be useful in understanding the variance in
+the original data or help with manual feature selection, we use the PCA
+purely to reduce the dimensionality for the clustering step. Instead of
+relying on the principal components, we use the original features of
+interest to interpret the later extracted clusters. We particularly
+advocate for such an approach if all original features are considered
+meaningful in understanding the time series and users would like to
+retain the features for interpretation (irrespective of the features'
+importance).
 
@@
 
-\input{latex-results/05FeatureClustering.tex}
+During the k-means clustering itself, the analysis seeks to minimize the
+total within-cluster variation. The analysis is designed to optimize the
+clustering of the feature data into \(k\) groups, where \(k\) is a
+pre-defined number of clusters. We used the Hartigan and Wong algorithm,
+which is a widely used algorithm in k-means clustering
+\citep{hartigan1979}. The algorithm starts by randomly separating the
+data points into k clusters and then iteratively updates the assignment
+of each point to the nearest cluster center until convergence. To do so,
+the Hartigan and Wong algorithm specifically calculates the
+within-cluster variation (\(W\)) of cluster \(C_i\) as the summed
+squared Euclidean distances of the feature \(x\) to the closest cluster
+centroid \(\mu_i\):
+
+\begin{equation} \label{eq:kWCi}
+  W(C_i) = \sum_{x \in C_i}(x-\mu_i)^2
+\end{equation}
+
+By summing the within-cluster sum of squares from all \(k\) clusters, we
+can then derive the total within-cluster sum of square \(WCSS\):
+
+\begin{equation} \label{eq:kWCSS}
+  WCSS = \sum_{i=1}^k W(C_i) = \sum_{i=1}^k \sum_{x \in C_i} (x - \mu_i)^2
+\end{equation}
+
+It is this \(WCSS\) that becomes the objective function to be minimized,
+by iteratively moving features from one cluster to another
+\citep{hartigan1979}. In particular, the algorithm (1) calculates the
+cluster centroids of the initial partitioning, (2) checks whether any
+feature has a centroid that is closer than that of the currently
+assigned cluster (3) updates the centroids based on any reassigned
+features, and then iterates between steps two and three until \(WCSS\)
+is minimized (i.e., locally optimal convergence) or a maximum number of
+iterations is reached \citep{jain2010}. Given the iterative nature of
+the algorithm, the initial partitioning is often important because the
+algorithm might arrive at a suboptimal clustering where the \(WCSS\)
+cannot be further reduced by moving any feature to another cluster,
+despite a better solution existing
+\citep[i.e., a local minimum;][]{timmerman2013}. It is, therefore often
+recommended to run the k-means clustering with several different
+starting positions.
+
+In our case, we entered the participants' PC-scores from the feature
+reduction step into the k-means algorithm. Because we did not know the
+underlying number of clusters within our sample, we calculated the
+cluster solutions for \(k=\{2, \dots , 10\}\) . To avoid local minima we
+used 100 random initial centroid positions for each run. Each of the 9
+cluster solutions converged within the iteration limit. In the next
+step, we will then evaluate which of the extracted cluster solutions
+offers the best fit with the data.
 
@@
 
-\input{latex-results/06ClusterEvaluationPerformance.tex}
+In our own illustration example, we used the \texttt{cluster.stats()}
+function from the \texttt{fpc} \textsf{R} package, which calculates a
+wide variety of internal cluster validity statistics for each of the
+extracted clustering solutions. With real-world data no single
+evaluation measure is likely perfect, and different measures may produce
+different results depending on the characteristics of the data and the
+research question being addressed \citep{kittler1998}. It is therefore
+important to consider a variety of evaluation measures and to carefully
+interpret the results in the context of the specific analysis
+\citep{vinh2009}. We found that across most indices, the analysis with
+\(k=2\) clusters performed the best. Three commonly reported indices we
+would like to highlight are the comparison of within clusters sum of
+squares, the average silhouette score, and the Calinski-Harabasz index.
+The first statistic we looked at was the total within-cluster sum of
+square \(WCSS\) (see also \equatref{eq:kWCSS}). While the within-cluster
+variation will naturally decrease with (more) smaller clusters, we
+observed that the decrease in \(WCSS\) was largest until \(k=2\) after
+which the decrease was much smaller. This method is also sometimes
+referred to as the `elbow method' \citep{syakur2018}. We then looked at
+a second, commonly used measure, the average silhouette score. This
+statistic measures the degree to which each time feature data point is
+similar to other points within the same cluster, compared to points in
+other clusters \citep{rousseeuw1987}. In our case, the \(k=2\) solution
+maximized the silhouette coefficient (\(s_2=\) 0.09). Finally, the
+Calinski-Harabasz index assesses the compactness and separation of the
+clusters by assessing the ratio of the sum of between-clusters
+dispersion and of inter-cluster dispersion for all clusters --- thus,
+the higher the score the better the performances \citep{calinski1974}.
+In our case, the \(k=2\) solution also showed the highest
+Calinski-Harabasz index (\(CH_2=\) 16.38; a full table of all extracted
+validity statistics is available in Supplemental Material
+A)\footnote{It is important to note that another commonly assessed aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}.
+This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}.
+Especially when comparing different clustering algorithms one common index is the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.}.
+In the final \(k=2\) solution the k-means analysis also assigned a
+relatively even number of participants to cluster 1 (\(n_{C_1}=\) 76)
+and cluster 2 (\(n_{C_1}=\) 80).
 
@@
 
-\input{latex-results/07ClusterEvaluationInterpretation.tex}
+In short, we find that the feature-based clustering discerned two
+meaningfully different groups of participants. We find an adaptive group
+(cluster 1) that reports higher well-being and more positive outgroup
+attitudes (\textit{median}) that are also stable over time
+(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
+period (\textit{linear trend}). This group also reported consistently
+more meaningful, need-fulfilling, and cooperative outgroup interactions
+(\textit{median}). This group with overwhelmingly positive experiences
+stands in contrast with a more detrimental group (cluster 2). This
+cluster, on average, reported much less positive, less meaningful, and
+less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower
+well-being, and more discrimination experiences (\textit{median}). At
+the same time, for members of this detrimental cluster (cluster 2)
+conditions seemed to deteriorate over time (\textit{linear trend}), and
+there was generally less consistency in the experiences they were able
+to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
+
+To identify these patterns, we first inspect the clusters based on the
+average values of meaningful features (see
+\fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We see that for
+some variables the features are generally stronger in separating the
+clusters. We, for example, see that the item on
+`\textit{how cooperative the interaction was}' distinguishes the two
+clusters across almost all seven features (except for the
+\textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}). Compare
+this to the `\textit{outgroup attitudes}' item where the differences
+between the clusters are much smaller for almost all features. We then
+inspect the clusters with a focus on the features (see
+\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
+variable focus, we can see more clearly that some features are better at
+distinguishing the clusters across variables. For example, \textit{MAD}
+and \textit{median} distinguish the two clusters across almost all
+variables (except for the item of whether the interaction was
+representative of the outgroup). These two features stand in stark
+contrast to other features, such as the
+\textit{lag-1 auto correlations}, which showed much smaller differences
+between the two clusters (see \fgrref[B]{fig:clusterFeatVar}). This
+offers some information on which features were most important in
+understanding the two extracted groups. Taking these two perspectives
+together, we can also focus on individual features or variables in
+particular. We, for example, see a strong difference in the average
+well-being, where participants in cluster 2 showed a much lower median
+well-being over the time series. At the same time, in terms of
+stability, both groups have virtually identical average \textit{MAC}
+statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There
+are, thus, variables and features that distinguish the clusters better
+than others and a combination of variables and features lets us explore
+meaningful group differences in more detail. In our case, we see that
+the central tendency, variability, and linear trend are best at
+distinguishing a group with mainly positive experiences (cluster 1) from
+a group with a more negative experience (cluster 2). We also see that
+our clusters line up with the past literature on the importance of
+focusing on simpler and more meaningful statistics
+\citep{bringmann2018c, eronen2021a}.
+
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons based on Features and Variables}
+  \label{fig:clusterFeatVar}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
+  \caption*{Note: \\
+  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+\end{figure}
+
+In the second step, we look at the prototypical trajectories of the
+clusters. For k-means clustering it is often recommended to use the
+average over time of the responses within the cluster
+\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
+Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
+For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
+Immediately striking are the mean differences, where participants in
+cluster 1 had more meaningful and fulfilling outgroup interactions and
+also consistently reported more voluntary and cooperative interactions
+but fewer accidental and involuntary interactions. The same cluster
+(cluster 1) also reported an increase in need-fulfilling interactions
+over the 30-day period and an increase in interactions that were
+representative of the outgroup. Whereas the other cluster (cluster 2)
+showed a decrease in voluntary, cooperative, and positive interactions
+over the 30 days. This `deterioration' cluster (cluster 2) also saw a
+decrease in general need fulfillment and experienced well-being over the
+30 days (see \fgrref[B]{fig:clusterTs}). We also see that while
+interaction representativeness, outgroup attitudes, and well-being are
+relatively stable for both clusters, the deteriorating cluster (cluster
+2) also showed substantially higher variability and instability on most
+of the other variables (also see \fgrref[A]{fig:clusterTs}).
+
+Finally, we can also assess the clusters across other individual
+difference variables \citep[e.g.,][]{monden2022}. This out-of-feature
+comparison allows us to check for data artifacts, as well as check
+whether the developmental clusters are associated with important social
+markers and individual differences. To illustrate artifact checks, we
+added the number of ESM measurements into the comparison and find that
+the participants in the deterioration cluster (cluster 2) on average
+completed slightly more ESM surveys in general and reported on more
+intergroup interactions in particular (see \(n\) in
+\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures, we
+ensured that the general time frame and completion rates are similar for
+all participants and indeed the numbers in ESM measurements generally
+are largely similar (e.g., see \(n\) for well-being and outgroup
+attitudes). However, the difference in the reported number of
+interactions might indicate either a clustering artifact or a meaningful
+difference. The higher average number of interactions in cluster 2
+could, for example, indicate a clustering artifact if variances are
+substantially larger due to the larger samples
+\citep[e.g., restriction of range in the smaller sample][]{kogan2006}.
+In our case, this seems less likely because one out of four variables
+did not differ in terms of the MAD (i.e., our selected measurement of
+the time series variance; see \fgrref{fig:cluster_comparison_mad} for an
+illustration). At the same time, however, the difference in the number
+of experienced interactions might also indicate a meaningful difference,
+where the deteriorating cluster (cluster 2) on average reported more
+outgroup interactions (\textit{difference} = 1.03, \(t\)(150.83) = 7.50,
+\(p\) \textless{} .001, \textit{95\%CI} {[}0.76, 1.30{]}), but these
+interactions were less voluntary (\textit{difference} = -1.04,
+\(t\)(108.89) = -7.71, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.31,
+-0.77{]}), less meaningful (\textit{difference} = -1.00, \(t\)(136.40) =
+-7.16, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.28, -0.73{]}), and
+less positive (\textit{difference} = -1.38, \(t\)(152.31) = -11.94,
+\(p\) \textless{} .001, \textit{95\%CI} {[}-1.61, -1.15{]}). Thus, while
+more research is needed for a conclusive test, our data seems to suggest
+that the differences in reported interactions are a meaningful
+difference between the clusters. Such a finding would also be in line
+with past research highlighting the role of negative intergroup
+interactions in explaining intergroup relations
+\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster comparison of Median Absolute Deviation for all variables}
+  \label{fig:cluster_comparison_mad}
+  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_mad.pdf}
+\end{figure}
+
+To further illustrate the utility of assessing out-of-feature individual
+differences, we also compare the two samples in terms of the
+participants' self-reported discrimination experiences in the
+Netherlands (measured during the post-measurement). When looking at the
+group comparison, we find that participants in the deteriorating cluster
+(cluster 2) reported substantially higher levels of everyday
+discrimination (\textit{difference} = 0.40, \(t\)(151.71) = 2.56, \(p\)
+= 0.011, \textit{95\%CI} {[}0.09, 0.71{]};
+\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
+(e.g., the sum of specific ESM measurements) and cross-sectional
+variables (e.g., general discrimination differences) that were not
+included in the original clustering step can be used to explore and
+understand the cluster differences in more detail.
+
+This cluster separation, then, has a number of empirical and practical
+applications. Firstly, the clusters are descriptive. With tens of
+variables, hundreds of participants, and thousands of measurements,
+singular descriptive statistics are often not able to capture the
+complex patterns that describe the data set. The feature-based
+clustering offers some direct insight into the complexity within the
+data set. In our empirical example, we, for example, see that
+participants are meaningfully distinguished by a combination of high
+(vs.~low) central tendency, variability, and linear trend. Secondly, the
+clusters identify important groups. The adaptive and deteriorating
+groups offer starting points for empirical exploration as well as
+practical interventions. Researchers can start probing what exactly
+distinguishes the two groups further and generate new bottom-up
+hypotheses. Practitioners in the resettlement field can use the group
+separation to identify individuals in need of assistance and can explore
+contextual factors that might contribute to the difficulties some might
+face. In our illustration we, for example, found that participants in
+the deteriorating cluster (cluster 2) reported less need fulfilling
+interactions over time. Thirdly, the feature-based approach is flexible
+and meaningful. We were able to use a wide range of time series features
+that have been central in the ESM literature and were able to use them
+directly to identify meaningful groups. For our empirical illustration
+we, among others for example, chose to focus on whether participants
+differed in their average well-being (i.e., \textit{median}), how much
+their well-being would vary over time (i.e, \textit{MAD}), and whether
+their well-being would on average increase or decrease over time (i.e.,
+\textit{linear trend}). Alternatively, for others cyclical patterns
+might be more important --- for example, whether well-being was higher
+on weekends. Importantly, in any case, we did not need to translate
+these dynamic features into probabilistic inference models (e.g., VAR
+models) to cluster the participants.
+
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons over time}
+  \label{fig:clusterTs}
+  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
+  \caption*{Note: \\
+  Subplot (A) displays the variable cluster means at every measurement occasion.
+Subplot (B) shows the GAM spline for each cluster across the measurement occasions.
+The thinner lines present all individual time series.}
+\end{figure}
 

### 2024-11-11 latex-results/01DataDescription.tex
typos fixed
config updated

@@
 For our empirical example, we focus on the variables that were collected
-during the ESM surveys and were available in all three studies. Variable
+during ESM surveys and were available in all three studies. Variable
 selection and preparation are described as part of the illustration
 below but for additional methodological details about the study setup
-see \citet[][]{Kreienkamp2022b}.
+see \citet[][]{Kreienkamp2022b}. Each study was approved by the ethics
+board of the university of origin and all participants gave informed
+consent.

### 2024-11-11 latex-results/02InputVariables.tex
feedback lfb
internal feedback lfb
config updated

--- 2023-02-21 latex-results/02InputVariables.tex
@@
-For our illustration, we include 12 variables that were available across
-all three studies and captured information about the participant's
-interactions, as well as cognitive-, emotional-, and motivational self
-in relationship with the majority group. We chose these aspects in
+For our illustration, we include 12 variables that were measured as part
+of the ESM surveys in all three studies and captured information about
+the participant's interactions, as well as the cognitive-, emotional-,
+and motivational self in relationship with the majority group (see
+\tblref{tab:var_selection} for an overview). We chose these aspects in
 particular because (1) the interaction-specific information exemplified
@@
 concepts and additionally allows us to showcase the efficiency benefits
-of the method and offers a reasonable use case for the optional feature
-reduction step.
+of the method and offers a reasonable use case for the feature reduction
+step.
+
+\input{tables/var_selection}
 
@@
 prepared for the analysis steps. Importantly, this not only means
-validating and cleaning the data (e.g., re-coding, removing duplicate or
-unwanted measurements) but also making the time series comparable. Two
-important steps are making the time-frames and response scales
-comparable across participants --- for example, by choosing a time frame
-that is common to most participants and standardizing the participants'
-responses \citep{liao2005}.
+validating and cleaning the data (e.g., re-coding, combining scale
+items) but also making the time-series comparable. Making the
+time-frames and response scales comparable across participants, for
+example, includes choosing a time frame that is common to most
+participants
+\citep['data preparation' and 'data cleaning' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.
 
@@
 likely due to the option to continue participation without compensation
-in the later two studies. To make the three studies comparable in
+in the latter study. To make the three studies comparable in
 participation and time frames, we iteratively removed all measurement
 occasions and participants that had more than 45\% missingness
-\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure let to a final sample of 157 participants, who jointly
-produced 8,132 measurements. Importantly, both the participant
-repsonse-patterns and the time frame were now substantially more
-comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
-\(\text{max}(t_{S3})=\) 67). Full methodological details are available
-in Online Supplemental Material A, but basic item information,
-descriptives, and correlations are also available in
+\citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing; see][]{Madley-Dowd2019}\footnote{Please note that for cases where the clustering is the main analysis, this high missingness threshold may be too conservative.
+As part of our validation analyses in \appref[]{app:ValidationAnalyses} we compare the model presented here with varying levels of missing data allowed.}.
+This procedure led to a final sample of 157 participants, who jointly
+produced 8,132 survey responses. Importantly, both the participant
+response patterns and the time frame were now substantially more
+comparable (number of measurement occasions per person: \(t_{S1}\): min
+= 40, max = 61, mean = 57.33, sd = 4.69; \(t_{S2}\): min = 33, max = 60,
+mean = 49.05, sd = 6.73; \(t_{S3}\): min = 36, max = 65, mean = 54.20,
+sd = 7.04). It is important to consider that some time series features
+may be less reliable when the number of measurement occasions per person
+is low (e.g., below 30 measurements per person), and this should be
+taken into account when conducting similar analyses. Full methodological
+details are available in Online Supplemental Material A, but basic item
+information, descriptives, and correlations are also available in
 \tblref{tab:descrLong}.

### 2024-11-11 latex-results/03FeatureExtraction.tex
typo fixed
config updated

@@
 
-For the central tendency feature of our illustration we chose the robust
-\textit{median}, which can avoid potential issues with non-normally
-distributed time series responses or outliers \citep{weisberg1992}. To
-calculate the \textit{median} (\(M\)), we let \(X_{ij}\) be the ordered
-list of values from the time series of variable \(j\) for participant
-\(i\). The calculation depends on whether the number of measurements in
-a time series \(n\) is odd or even.
+For the central tendency feature of our illustration, we chose the
+median (\(M\)), which effectively addresses potential complications
+arising from non-normally distributed responses or outliers within time
+series datasets \citep{weisberg1992}. To compute the median, it is
+imperative to differentiate between two types of time series
+representations for a given variable \(j\) related to participant \(i\):
+the chronological series and the ordered series. The chronological time
+series, denoted by \(X_{ij}\), encapsulates the sequence of observations
+\(\{x_{ij1}, x_{ij2}, ..., x_{ijn}\}\) for variable \(j\) concerning
+participant \(i\), organized by their temporal occurrence. Here,
+\(x_{ijt}\) signifies a specific observation at time \(t\) within this
+sequence. In contrast, the ordered time series, represented as
+\(\mathbf{X}_{ij}\), is derived from \(X_{ij}\) by sorting the
+observations in ascending order of magnitude. This ordered set is
+expressed as
+\(\{\mathbf{x}_{ij1}, \mathbf{x}_{ij2}, ..., \mathbf{x}_{ijn}\}\), with
+each \(\mathbf{x}_{ijk}\) corresponding to the \(k\)-th element in the
+reordered series \(\mathbf{X}_{ij}\).
+
+The median \(M(\mathbf{X}_{ij})\) is then the value located precisely at
+the center of the ordered time series \(\mathbf{X}_{ij}\). Depending on
+whether the total number of observations (\(n\)) is odd or even, the
+median is either the middel \(k\)-th element if \(n\) is odd, or the
+average of the two middle values if \(n\) is even:
 
 \begin{equation} \label{eq:median}
-  M(X_{ij}) = 
+  M(\mathbf{X}_{ij}) = 
   \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
+    \mathbf{x}_{ij\left(\frac{n+1}{2}\right)} & \text{if $n$ is odd} \\
+    \frac{\mathbf{x}_{ij\left(\frac{n}{2}\right)} + \mathbf{x}_{ij\left(\frac{n}{2} + 1\right)}}{2} & \text{if $n$ is even}
   \end{cases}
 \end{equation}
+
+This approach ensures that the median is a reliable indicator of central
+tendency in time series analysis, unaffected by data distribution
+asymmetries or the presence of outliers.
 
@@
 central tendency and is sometimes also referred to as the dispersion or
-spread of the data \citep{weisberg1992}. In time series analyses,
-variability is conceptually important because information about the
-distribution and diversity of the data has been found to be indicative
-of worse psychological states \citep{myin-germeys2018, helmich2021}.
-Person-level differences of ESM measurements have, for example, been
-associated with higher levels of psycho-pathological recurrences among
-depression patients \citep{timm2017}. As such, psychological researchers
-and practitioners are often empirically interested in between-person
+spread of the data \citep{weisberg1992}. Common measurements of the
+variability are the variance or standard deviation as well as their
+robust counter parts. In time series analyses, variability is
+conceptually important because information about the distribution and
+diversity of the data has been found to be indicative of worse
+psychological states \citep{myin-germeys2018, helmich2021}. Person-level
+differences of ESM measurements have, for example, been associated with
+higher levels of psycho-pathological recurrences among depression
+patients \citep{timm2017}. As such, psychological researchers and
+practitioners are often empirically interested in between-person
 differences in variability. Researchers on polarization and
@@
 
-For our illustration data, we chose to capture the time series
-variability with the \textit{Median Absolute Deviation} (\(MAD\)), where
-we calculate the \textit{median} (\(M\); calculated as in
-\equatref{eq:median}) for the absolute deviations of measurement \(x\)
-at time point \(t\) for participant \(i\) and variable \(j\) from the
-median of that time series \(X\). We again chose the robust statistic
-because the Median-based measure is less affected by non-normal
-distributions and extreme values or outliers compared to other measures
-of variability like the standard deviation \citep{weisberg1992}
+For our illustration data, we chose the
+\textit{Median Absolute Deviation} (\(MAD\)) to gauge the variability
+within our time series data. This choice is motivated by the robustness
+of MAD, particularly its resilience to the effects of non-normal
+distributions and outliers, which can significantly skew traditional
+variability measures like the standard deviation \citep{weisberg1992}.
+For a given variable \(j\) and participant \(i\), the MAD is calculated
+by first determining the median (\(M\)) of the ordered time series
+\(\mathbf{X}_{ij}\) as outlined in \equatref{eq:median}. We then compute
+the absolute deviations of each observation in the time series
+\(X_{ij}\) from this median value. Specifically, for each time point
+\(t\), we calculate the absolute difference between \(x_{ijt}\) and the
+series median \(M(\mathbf{X}_{ij})\). The MAD is then the median of
+these absolute deviations:
 
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
-\end{equation}
+\begin{subequations}\label{eq:mad}
+    \begin{align} 
+      MAD(X_{ij}) &= M(\left| x_{ijt} - M(\mathbf{X}_{ij}) \right|) \label{eq:mad_general} \\
+                  &= M(\{ |x_{ij1} - M(\mathbf{X}_{ij})|, |x_{ij2} - M(\mathbf{X}_{ij})|, \ldots, |x_{ijn} - M(\mathbf{X}_{ij})| \}) \label{eq:mad_detailed}
+    \end{align}
+\end{subequations}
+
+The calculation of MAD focuses on the magnitudes of deviations, ensuring
+it provides a robust measure of dispersion that reflects the inherent
+variability in the time series data.
 
@@
 Instability captures the average change between two consecutive
-measurements \citep{ebner-priemer2009}. While instability is
+measurements \citep{ebner-priemer2009, jahng2008}. While instability is
 conceptually related to the variability feature, variability does not
@@
 `jumpy-ness' of the data over time. In other words, variability reflects
-the range or diversity of values in a time series data, while
-instability reflects the fluctuation or inconsistency in a time series
-data over time \citep{trull2008}. For example, if a person has rapid and
-extreme changes in mood their mood is highly unstable, while if a
-person's mood responses span a wide range over the entire study period,
-their mood is highly variable \citep{jahng2008}. Within psychological
-time series, instability measurements have especially been important in
-the research of borderline personality disorder \citep{trull2008} and
-suicidality \citep{kivela2022}, but also in understanding early warning
-signals more generally \citep{wichers2019}. Conceptually, the
-instability feature, thus, relates to a broad range of research
-questions, including: ``What is the nature of the identification changes
-in those who start working in a new country?'' or ``Do strong daily
-fluctuations in self-esteem reflect the process of identity formation in
-adolescents?''
+the range or diversity of values in the un-ordered time series data,
+while instability reflects the fluctuation or inconsistency in a time
+series data over time \citep{trull2008, houben2015, koval2013}. For
+example, if a person has rapid and extreme changes in mood their mood is
+highly unstable, while if a person's mood responses span a wide range
+over the entire study period, their mood is highly variable
+\citep[note that this does not need to be rapidly changing or instable, e.g., when there is linear increase over time; also see][]{jahng2008}.
+Within psychological time series, instability measurements have
+especially been important in the research of borderline personality
+disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also
+in understanding early warning signals more generally
+\citep{wichers2019}. Conceptually, the instability feature, thus,
+relates to a broad range of research questions, including: ``What is the
+nature of the identification changes in those who start working in a new
+country?'' or ``Do strong daily fluctuations in self-esteem reflect the
+process of identity formation in adolescents?''
 
@@
 series are considered more impactful or when outliers are meaningful and
-need to be taken into account \citep{chatfield2003}.
+need to be taken into account \citep{chatfield2003, bos2019}. For
+psychological ESM data, some research suggests that amplitude and
+frequency might predict different health outcomes and can be
+investigated jointly \citep{wang2012a, jahng2008}.
 
-\paragraph{Self-similarity.}
+\paragraph{Temporal dependence.}
 
-Self-similarity in time series data refers to the property of a time
-series to exhibit similar patterns of behavior over different time
-scales \citep{dmello2021}. That is, self-similarity describes how much a
-measurement carries over to future measurements. One important
-self-similarity in psychological time series is \texit{inertia} --- how
-much a measurement carries over to its next measurement
+Temporal dependence in time series data refers to the degree to which a
+time series is influenced by its past values, exhibiting patterns of
+behavior that may be regular over different time scales
+\citep{dmello2021}. In the context of psychological time series, an
+important aspect of temporal dependence is \textit{inertia} --- how much
+a measurement carries over to its next measurement
 \citep{kuppens2010, suls1998}. If inertia is high a development tends to
@@
 psychological maladjustment \citep{kuppens2010}. In a similar vein, high
-inertia in negative affect at baseline was even predictive of the
-initial onset of depression \citep{kuppens2012}. Conceptually, inertia
-is more broadly connected to research questions such as:
-\texttt{Do\ patients\ stay\ in\ a\ negative\ mood\ for\ several\ measurements?\textquotesingle{}\textquotesingle{}\ or}Do
-migrants stay with their language practice for several days at a
-time?'\,'
+inertia in negative affect at baseline was predictive of the initial
+onset of depression \citep{kuppens2012}. Conceptually, inertia is more
+broadly connected to research questions such as: ``Do patients stay in a
+negative mood for several measurements?'' or ``Do migrants stay with
+their language practice for several days at a time?''
 
@@
 \end{equation}
+
+While inertia captures the simplest case of temporal dynamics, lag-1, we
+acknowledge that temporal dependence in psychological time series may
+also exhibit more complex relationships, including higher lagged auto
+correlations or cyclical relationships (fourier coefficients, or
+continuous wavelet transforms are often used to capture such
+relationships).
 
@@
 common feature of psychological time series \citep{hayes2007}. As an
-example, episodic disorders, such as depression, are most likely best
+example, episodic disorders, such as depression, are often best
 described as non-linear systems \citep{hosenfeld2015}. Similarly,
@@
 \(edf\) summarizes the \textit{wiggliness} of a spline trend line
-\citep{wood2017}. The degrees of freedom of a spline model are primarily
-determined by the number of knots and the order of the spline. For
-instance, a cubic spline with \(k\) knots has \(k\)+3 degrees of freedom
-\citep{faraway2016}. However, in a penalized spline framework, which is
-commonly used for GAMs, the effective degrees of freedom can be less
-than \(k\)+3. This is because the model employs a smoothing parameter to
-control the trade-off between the complexity (flexibility) of the model
-and its fit to the data, thereby penalizing overly complex models and
-potentially reducing the effective degrees of freedom \citep{marx1998}.
-Intuitively then an edf of 1 would be equivalent to a linear
-relationship (i.e., one linear slope parameter), whereas a higher edf
-(particularly an edf \textgreater{} 2) is indicative of a non-linear
-trend. The estimated degrees of freedom are commonly based on a concept
-called `effective degrees of freedom' and can be represented as the
-trace \(tr\) (i.e., the sum of the diagonal elements) of the smoother
-matrix \(S\), a symmetric matrix that maps from the raw data to the
-smooth estimates \citep{wood2017}.
+\citep{wood2017, bringmann2017}. The degrees of freedom of a spline
+model are primarily determined by the number of knots and the order of
+the spline. For instance, a cubic spline with \(k\) knots has \(k\)+3
+degrees of freedom
+\citep{faraway2016, haslbeck2021a, castro-alvarez2024}. However, in a
+penalized spline framework, which is commonly used for GAMs, the
+effective degrees of freedom can be less than \(k\)+3. This is because
+the model employs a smoothing parameter to control the trade-off between
+the complexity (flexibility) of the model and its fit to the data,
+thereby penalizing overly complex models and potentially reducing the
+effective degrees of freedom \citep{marx1998}. Intuitively then an edf
+of 1 would be equivalent to a linear relationship (i.e., one linear
+slope parameter), whereas a higher edf (particularly an edf
+\textgreater{} 2) is indicative of a non-linear trend. The estimated
+degrees of freedom are commonly based on a concept called `effective
+degrees of freedom' and can be represented as the trace \(tr\) (i.e.,
+the sum of the diagonal elements) of the smoother matrix \(S\), a
+symmetric matrix that maps from the raw data to the smooth estimates
+\citep{wood2017}.
 
@@
 \end{equation}
+
+\paragraph{Additional considerations.}
 
@@
 be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive
-mean matching imputation using the MICE library \citep[][]{buuren2011}.
-Note again that with this procedure we only need to impute an extremely
-small number of missing values as most feature calculations can use the
+We, thus, impute missing feature values via predictive mean matching
+(PMM) with the MICE package in R, employing a single imputation and
+specifying a maximum of 50 iterations and a fixed seed for convergence
+and reproducibility \citep[][]{buuren2011}. We chose PMM for its ability
+to preserve the original data distribution without assuming normality
+and robustly handling multiple data types \citep{vanbuuren2006}. Note
+again that with this procedure we only need to impute an extremely small
+number of missing values as most feature calculations can use the
 available data instead.

### 2024-11-11 latex-results/04FeatureReduction.tex
config updated

@@
 PCA here but we encourage users to consider more specialized methods as
-well.
+well (we provide an example decision guide in \sicontext).
 
@@
 standardize all features across participants to ensure that all features
-are weighted equally. We then enter all 72 features into the analysis.
-The PCA uses linear transformations in such a way that the first
-component captures the most possible variance of the original data
+are weighted equally \citep[][]{horne2020}. We then enter all 72
+features into the analysis. The PCA uses linear transformations in such
+a way that the first component captures the most possible variance of
+the original data
 \citep[e.g., by finding a vector that maximizes the sum of squared distances][]{jolliffe2002, abdi2010}.
@@
 the original data or help with manual feature selection, we use the PCA
-purely to reduce the dimensionality for the clustering step. Instead of
-relying on the principal components, we use the original features of
-interest to interpret the later extracted clusters. We particularly
-advocate for such an approach if all original features are considered
-meaningful in understanding the time series and users would like to
-retain the features for interpretation (irrespective of the features'
-importance).
+here purely to reduce the dimensionality for the clustering step.
+Instead of relying on the principal components, we use the original
+features of interest to interpret the later extracted clusters. We
+particularly advocate for such an approach if all original features are
+considered meaningful in understanding the time series and users would
+like to retain the features for interpretation (irrespective of the
+features' importance).

### 2024-11-11 latex-results/06ClusterEvaluationPerformance.tex
config updated

@@
 wide variety of internal cluster validity statistics for each of the
-extracted clustering solutions. With real-world data no single
-evaluation measure is likely perfect, and different measures may produce
-different results depending on the characteristics of the data and the
-research question being addressed \citep{kittler1998}. It is therefore
-important to consider a variety of evaluation measures and to carefully
-interpret the results in the context of the specific analysis
-\citep{vinh2009}. We found that across most indices, the analysis with
-\(k=2\) clusters performed the best. Three commonly reported indices we
-would like to highlight are the comparison of within clusters sum of
-squares, the average silhouette score, and the Calinski-Harabasz index.
-The first statistic we looked at was the total within-cluster sum of
-square \(WCSS\) (see also \equatref{eq:kWCSS}). While the within-cluster
+extracted clustering solutions. With real-world data, no single
+evaluation measure is likely to be perfect. Different measures may yield
+varying results based on the data characteristics and the research
+question at hand \citep{kittler1998}. It is therefore important to
+consider a variety of evaluation measures and to carefully interpret the
+results in the context of the specific analysis \citep{vinh2009}. We
+found that across most indices, the analysis with \(k=2\) clusters
+performed the best. Three commonly reported indices we would like to
+highlight are the comparison of within clusters sum of squares, the
+average silhouette score, and the Calinski-Harabasz index. The first
+statistic we looked at was the total within-cluster sum of square
+\(WCSS\) (see also \equatref{eq:kWCSS}). While the within-cluster
 variation will naturally decrease with (more) smaller clusters, we
-observed that the decrease in \(WCSS\) was largest until \(k=2\) after
+observed that the decrease in \(WCSS\) was largest until \(k=2\), after
 which the decrease was much smaller. This method is also sometimes
@@
 and cluster 2 (\(n_{C_1}=\) 80).
+
+To ensure that the clustering is necessary in the first place, we also
+compare the performance to a single cluster solution (i.e., a single
+centroid). The comparison to this \(k=1\) solution is slightly different
+because metrics like the between-cluster separation are not available.
+Nonetheless, comparing the within-cluster sums of squares (SS) and the
+explained variance, we find that two clusters indeed outperform a single
+cluster solution. Specifically, the total within-cluster SS decreased
+from 8940.21 for one cluster to 8080.67 for two clusters. Additionally,
+the variance explained increased from \textless.001 to 0.096 when the
+cluster count was increased to two
+\citep[e.g., ][; also see \situtorial\ for full results]{beijers2022}.

### 2024-11-11 latex-results/07ClusterEvaluationInterpretation.tex
config updated

@@
 meaningfully different groups of participants. We find an adaptive group
-(cluster 1) that reports higher well-being and more positive outgroup
-attitudes (\textit{median}) that are also stable over time
-(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
-period (\textit{linear trend}). This group also reported consistently
-more meaningful, need-fulfilling, and cooperative outgroup interactions
-(\textit{median}). This group with overwhelmingly positive experiences
+(cluster 1) that reports higher well-being (\textit{median}:
+\textit{difference} = -0.52, \(t\)(153.87) = -3.34, \(p\) = 0.001,
+\textit{95\%CI} {[}-0.82, -0.21{]}; also see
+\fgrref[A]{fig:cluster_comparison_features}) and more positive outgroup
+interactions (\textit{median}: \textit{difference} = -1.38,
+\(t\)(152.31) = -11.94, \(p\) \textless{} .001, \textit{95\%CI}
+{[}-1.61, -1.15{]}), which are also stable over time (\textit{MAC}:
+\textit{difference} = 0.54, \(t\)(153.98) = 3.49, \(p\) \textless{}
+.001, \textit{95\%CI} {[}0.23, 0.84{]}) and tend to increase more over
+the 30 day test period (\textit{linear trend}: \textit{difference} =
+-0.55, \(t\)(149.90) = -3.55, \(p\) \textless{} .001, \textit{95\%CI}
+{[}-0.85, -0.24{]}; also see
+\fgrref[C]{fig:cluster_comparison_features}). This group also reported
+consistently more meaningful (\textit{median}: \textit{difference} =
+-1.00, \(t\)(136.40) = -7.16, \(p\) \textless{} .001, \textit{95\%CI}
+{[}-1.28, -0.73{]}), need-fulfilling (\textit{median}:
+\textit{difference} = -0.99, \(t\)(135.30) = -7.17, \(p\) \textless{}
+.001, \textit{95\%CI} {[}-1.26, -0.72{]}), and cooperative outgroup
+interactions (\textit{median}: \textit{difference} = -1.33,
+\(t\)(120.36) = -11.28, \(p\) \textless{} .001, \textit{95\%CI}
+{[}-1.56, -1.10{]}). This group with overwhelmingly positive experiences
 stands in contrast with a more detrimental group (cluster 2). This
@@
 there was generally less consistency in the experiences they were able
-to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
+to have (\textit{MAC}, \textit{MAD}, \textit{edf}; also see
+\fgrref[]{fig:cluster_comparison_features}; for a full and interactive
+comparison of all features see \situtorial).
 
@@
 representative of the outgroup). These two features stand in stark
-contrast to other features, such as the
-\textit{lag-1 auto correlations}, which showed much smaller differences
-between the two clusters (see \fgrref[B]{fig:clusterFeatVar}). This
-offers some information on which features were most important in
-understanding the two extracted groups. Taking these two perspectives
-together, we can also focus on individual features or variables in
-particular. We, for example, see a strong difference in the average
-well-being, where participants in cluster 2 showed a much lower median
-well-being over the time series. At the same time, in terms of
-stability, both groups have virtually identical average \textit{MAC}
-statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There
-are, thus, variables and features that distinguish the clusters better
-than others and a combination of variables and features lets us explore
-meaningful group differences in more detail. In our case, we see that
-the central tendency, variability, and linear trend are best at
-distinguishing a group with mainly positive experiences (cluster 1) from
-a group with a more negative experience (cluster 2). We also see that
-our clusters line up with the past literature on the importance of
-focusing on simpler and more meaningful statistics
-\citep{bringmann2018c, eronen2021a}.
+contrast to other features, such as the \textit{lag-1 auto correlations}
+or the \textit{GAM edf}, which showed much smaller differences between
+the two clusters (see \fgrref[B]{fig:clusterFeatVar} and
+\fgrref[]{fig:cluster_comparison_features}; please note that we offer
+readers an interactive tool to assess the cluster differences for all
+features in \situtorial). This offers some information on which features
+were most important in differentiating the two extracted groups but also
+shows that with real-world data, not all features will have enough range
+to distinguish people on all variables (e.g., see the non-linearity
+patterns in \fgrref[]{fig:cluster_comparison_features}; for a more
+direct illustration of GAM edf differences see \citealp{bringmann2017}).
+Taking these two perspectives together, we can also focus on individual
+features or variables in particular. We, for example, see a strong
+difference in the average well-being, where participants in cluster 2
+showed a much lower median well-being over the time series
+(\textit{difference} = -0.52, \(t\)(153.87) = -3.34, \(p\) = 0.001,
+\textit{95\%CI} {[}-0.82, -0.21{]}). At the same time, in terms of
+well-being stability, both groups have virtually identical average
+\textit{MAC} statistics for well-being (\textit{difference} = -0.01,
+\(t\)(153.96) = -0.04, \(p\) = 0.968, \textit{95\%CI} {[}-0.32, 0.31{]};
+also see \fgrref[A]{fig:clusterFeatVar}). There are, thus, variables and
+features that distinguish the clusters better than others and a
+combination of variables and features lets us explore meaningful group
+differences in more detail. In our case, we see that the central
+tendency, variability, and linear trend are best at distinguishing a
+group with mainly positive experiences (cluster 1) from a group with a
+more negative experience (cluster 2). We also see that our clusters line
+up with the past literature on the importance of focusing on simpler and
+more meaningful statistics
+\citep{bringmann2018c, eronen2021a, dejonckheere2019}.
 
@@
   \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb_tutorial.pdf}
   \caption*{Note: \\
-  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
+  "Int." = outgroup interaction, "mad" = median absolute deviation, "mac" = mean absolute change, "lin" = linear slope, "edf" = estimated degrees of freedom of an empty GAM spline model, "ar01" = lag-1 autocorrelation, "OCC"/"occ" = out-of-cluster comparison\\
+  Within the "(B) Feature Focus" subplot, the 'n (within ooc)' is an out-of-cluster comparison of the within-person available measurements for each variable; the 'between ooc (mean)' are also out-of-cluster comparisons but on a between person level.
+'Measurements removed' is the person-specific count of measurement occasions removed during the missingness handling and 'Discrimination' is the scale mean of daily discimination experiences (measured during the final survey).}
 \end{figure}
@@
 average over time of the responses within the cluster
-\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
-For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
+\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.}.
 Immediately striking are the mean differences, where participants in
@@
 over the 30 days. This `deterioration' cluster (cluster 2) also saw a
-decrease in general need fulfillment and experienced well-being over the
-30 days (see \fgrref[B]{fig:clusterTs}). We also see that while
-interaction representativeness, outgroup attitudes, and well-being are
-relatively stable for both clusters, the deteriorating cluster (cluster
-2) also showed substantially higher variability and instability on most
-of the other variables (also see \fgrref[A]{fig:clusterTs}).
+decrease in general need fulfillment but not experienced well-being over
+the 30 days (see \fgrref[C]{fig:cluster_comparison_features}). We also
+see that while interaction representativeness, outgroup attitudes, and
+well-being are relatively stable for both clusters, the deteriorating
+cluster (cluster 2) also showed substantially higher variability and
+instability on most of the other variables (although these effects are
+much smaller; see \fgrref[A]{fig:clusterTs}).
 
@@
 did not differ in terms of the MAD (i.e., our selected measurement of
-the time series variance; see \fgrref{fig:cluster_comparison_mad} for an
-illustration). At the same time, however, the difference in the number
-of experienced interactions might also indicate a meaningful difference,
-where the deteriorating cluster (cluster 2) on average reported more
-outgroup interactions (\textit{difference} = 1.03, \(t\)(150.83) = 7.50,
-\(p\) \textless{} .001, \textit{95\%CI} {[}0.76, 1.30{]}), but these
-interactions were less voluntary (\textit{difference} = -1.04,
-\(t\)(108.89) = -7.71, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.31,
--0.77{]}), less meaningful (\textit{difference} = -1.00, \(t\)(136.40) =
--7.16, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.28, -0.73{]}), and
-less positive (\textit{difference} = -1.38, \(t\)(152.31) = -11.94,
-\(p\) \textless{} .001, \textit{95\%CI} {[}-1.61, -1.15{]}). Thus, while
-more research is needed for a conclusive test, our data seems to suggest
-that the differences in reported interactions are a meaningful
-difference between the clusters. Such a finding would also be in line
-with past research highlighting the role of negative intergroup
-interactions in explaining intergroup relations
-\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+the time series variance; see \fgrref{fig:cluster_comparison_features}
+for an illustration). At the same time, however, the difference in the
+number of experienced interactions might also indicate a meaningful
+difference, where the deteriorating cluster (cluster 2) on average
+reported more outgroup interactions (\textit{difference} = 1.03,
+\(t\)(150.83) = 7.50, \(p\) \textless{} .001, \textit{95\%CI} {[}0.76,
+1.30{]}), but these interactions were less voluntary
+(\textit{difference} = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{}
+.001, \textit{95\%CI} {[}-1.31, -0.77{]}), less meaningful
+(\textit{difference} = -1.00, \(t\)(136.40) = -7.16, \(p\) \textless{}
+.001, \textit{95\%CI} {[}-1.28, -0.73{]}), and less positive
+(\textit{difference} = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{}
+.001, \textit{95\%CI} {[}-1.61, -1.15{]}). Thus, while more research is
+needed for a conclusive test, our data seems to suggest that the
+differences in reported interactions are a meaningful difference between
+the clusters. Such a finding would also be in line with past research
+highlighting the role of negative intergroup interactions in explaining
+intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. A
+related validity check was the inclusion of the missingness handling,
+where we compared the two clusters on the average number of measurements
+removed as part of the missingness handling. We find that the clusters
+did not significantly differ on this metric suggesting that the
+missingness handling did not affect the cluster separation (also see
+\appref[]{app:ValidationAnalyses} and \situtorial).
 
 \begin{figure}[!ht] %hbtp
-  \caption{Cluster comparison of Median Absolute Deviation for all variables}
-  \label{fig:cluster_comparison_mad}
-  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_mad.pdf}
+  \caption{Comparison Cluster Differences by Features and Variables.}
+  \label{fig:cluster_comparison_features}
+  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_combined.pdf}
+  \caption*{Note:\\
+  The figure shows the differences between the clusters in the standardized features that were entered into the dimensionality reduction (for each input variable).
+We display the median (panel A), the median absolute deviation (MAD, panel B), the univariate linear slope (panel C), as well as the estimated degrees of freedom of the generalized additive model splines (GAM edf, panel D).
+Please also note that as part of \situtorial, we provide readers with an interactive selection tool to compare cluster differences on all variables and features.}
 \end{figure}
@@
   Subplot (A) displays the variable cluster means at every measurement occasion.
+The thinner lines present all individual time series.
 Subplot (B) shows the GAM spline for each cluster across the measurement occasions.
-The thinner lines present all individual time series.}
+The thinner lines present all individual GAM Splines.}
 \end{figure}

### 2024-11-11 sections/01DataDescription.Rmd
typos fixed
config updated

@@
 Each of the studies focused on recently arrived first-generation migrants and each study included a number of idiosyncratic variables relevant for the broader research collective.
-For our empirical example, we focus on the variables that were collected during the ESM surveys and were available in all three studies.
+For our empirical example, we focus on the variables that were collected during ESM surveys and were available in all three studies.
 Variable selection and preparation are described as part of the illustration below but for additional methodological details about the study setup see \citet[][]{Kreienkamp2022b}.
+Each study was approved by the ethics board of the university of origin and all participants gave informed consent.

### 2024-11-11 sections/02InputVariables.Rmd
config updated

@@
 
-For our illustration, we include 12 variables that were measured as part of the ESM surveys in all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group (see \tblref{tab:var_selection} for an overview).
+For our illustration, we include 12 variables that were measured as part of the ESM surveys in all three studies and captured information about the participant's interactions, as well as the cognitive-, emotional-, and motivational self in relationship with the majority group (see \tblref{tab:var_selection} for an overview).
 We chose these aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}.
 The breadth of the included variables also showcases the utility of the method for a growing body of literature that considers heterogeneous and complex concepts.
-As a result, the number of included variables is also on the higher end for psychological concepts and additionally allows us to showcase the efficiency benefits of the method and offers a reasonable use case for the optional feature reduction step.
+As a result, the number of included variables is also on the higher end for psychological concepts and additionally allows us to showcase the efficiency benefits of the method and offers a reasonable use case for the feature reduction step.
 
@@
 
+```{r}
+
+descr_N_all <- rbind(
+  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
+    mutate(across(!TID & !study, as.numeric)),
+  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
+  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
+)
+
+# Step 1: Count the number of measurements per person within each study
+measurement_counts <- descr_N_all %>%
+  group_by(study, PID) %>%                # Group by study and PID
+  summarise(measurement_count = n()) %>%  # Count the number of rows per PID (this is the number of measurements)
+  ungroup()                               # Ungroup the data
+
+# Step 2: Calculate the mean and standard deviation of measurements per study
+measurement_summary <- measurement_counts %>%
+  group_by(study) %>%                              # Group by study
+  summarise(
+    min_measurements = min(measurement_count),     # Calculate min of measurements per study
+    max_measurements = max(measurement_count),     # Calculate max of measurements per study
+    mean_measurements = mean(measurement_count),   # Calculate mean of measurements per study
+    sd_measurements = sd(measurement_count)        # Calculate standard deviation of measurements per study
+  )
+
+generate_descriptive_report <- function(study_id, measurement_summary) {
+  # Extract relevant statistics for the given study
+  min_measurements <- measurement_summary$min_measurements[measurement_summary$study == study_id]
+  max_measurements <- measurement_summary$max_measurements[measurement_summary$study == study_id]
+  mean_measurements <- format(round(measurement_summary$mean_measurements[measurement_summary$study == study_id], 2), nsmall=2)
+  sd_measurements <- format(round(measurement_summary$sd_measurements[measurement_summary$study == study_id], 2), nsmall=2)
+
+  # Create the LaTeX formatted string
+  latex_string <- paste0(
+    "$t_{", study_id, "}$: min = ", min_measurements,
+    ", max = ", max_measurements,
+    ", mean = ", mean_measurements,
+    ", sd = ", sd_measurements
+  )
+
+  return(latex_string)
+}
+
+latex_S1 <- generate_descriptive_report("S1", measurement_summary)
+latex_S2 <- generate_descriptive_report("S2", measurement_summary)
+latex_S3 <- generate_descriptive_report("S3", measurement_summary)
+
+```
+
 Once the important variables have been selected, the data needs to be prepared for the analysis steps.
-Importantly, this not only means validating and cleaning the data (e.g., re-coding, removing duplicate or unwanted measurements) but also making the time-series comparable.
-Two important steps are making the time-frames and response scales comparable across participants --- for example, by choosing a time frame that is common to most participants and standardizing the participants' responses \citep['data exclusion' and 'data transformation' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.
+Importantly, this not only means validating and cleaning the data (e.g., re-coding, combining scale items) but also making the time-series comparable.
+Making the time-frames and response scales comparable across participants, for example, includes choosing a time frame that is common to most participants \citep['data preparation' and 'data cleaning' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.
 
@@
 This was likely due to the option to continue participation without compensation in the latter study.
-To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
-This procedure led to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements.
-Importantly, both the participant response-patterns and the time frame were now substantially more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`).
+To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness \citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing; see][]{Madley-Dowd2019}\footnote{Please note that for cases where the clustering is the main analysis, this high missingness threshold may be too conservative.
+As part of our validation analyses in \appref[]{app:ValidationAnalyses} we compare the model presented here with varying levels of missing data allowed.}.
+This procedure led to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` survey responses.
+Importantly, both the participant response patterns and the time frame were now substantially more comparable (number of measurement occasions per person: `r latex_S1`; `r latex_S2`; `r latex_S3`).
+It is important to consider that some time series features may be less reliable when the number of measurement occasions per person is low (e.g., below 30 measurements per person), and this should be taken into account when conducting similar analyses.
 Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are also available in \tblref{tab:descrLong}.
@@
     general = c(
+      '"Int." = outrgroup interaction, "-GOAL-" = self-identified goal, "-NAME-" = provided name of the interaction partner',
       "Upper triangle: Between-person correlations;",
@@
     general = c(
-      '"Int." = interaction.',
+      '"Int." = outgroup interaction, "ICC" = intraclass correlation coefficient, "SD" = standard deviation',
       'Upper triangle: Between-person correlations;',

### 2024-11-11 sections/03FeatureExtraction.Rmd
config updated

--- 2023-07-29 sections/03FeatureExtraction.Rmd
@@
 
-For the central tendency feature of our illustration we chose the robust \textit{median}, which can avoid potential issues with non-normally distributed time series responses or outliers \citep{weisberg1992}.
-To calculate the \textit{median} ($M$), we let $X_{ij}$ be the ordered list of values from the time series of variable $j$ for participant $i$.
-The calculation depends on whether the number of measurements in a time series $n$ is odd or even.
+For the central tendency feature of our illustration, we chose the median (\(M\)), which effectively addresses potential complications arising from non-normally distributed responses or outliers within time series datasets \citep{weisberg1992}.
+To compute the median, it is imperative to differentiate between two types of time series representations for a given variable \( j \) related to participant \( i \): the chronological series and the ordered series.
+The chronological time series, denoted by \( X_{ij} \), encapsulates the sequence of observations \( \{x_{ij1}, x_{ij2}, ..., x_{ijn}\} \) for variable \( j \) concerning participant \( i \), organized by their temporal occurrence.
+Here, \( x_{ijt} \) signifies a specific observation at time \( t \) within this sequence.
+In contrast, the ordered time series, represented as \( \mathbf{X}_{ij} \), is derived from \( X_{ij} \) by sorting the observations in ascending order of magnitude.
+This ordered set is expressed as \( \{\mathbf{x}_{ij1}, \mathbf{x}_{ij2}, ..., \mathbf{x}_{ijn}\} \), with each \( \mathbf{x}_{ijk} \) corresponding to the \(k\)-th element in the reordered series \( \mathbf{X}_{ij} \).
+
+The median \( M(\mathbf{X}_{ij}) \) is then the value located precisely at the center of the ordered time series \( \mathbf{X}_{ij} \).
+Depending on whether the total number of observations (\( n \)) is odd or even, the median is either the middel \(k\)-th element if \( n \) is odd, or the average of the two middle values if \( n \) is even:
 
@@
 \begin{equation} \label{eq:median}
-  M(X_{ij}) = 
+  M(\mathbf{X}_{ij}) = 
   \begin{cases}
-      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
-      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
+    \mathbf{x}_{ij\left(\frac{n+1}{2}\right)} & \text{if $n$ is odd} \\
+    \frac{\mathbf{x}_{ij\left(\frac{n}{2}\right)} + \mathbf{x}_{ij\left(\frac{n}{2} + 1\right)}}{2} & \text{if $n$ is even}
   \end{cases}
@@
 
+This approach ensures that the median is a reliable indicator of central tendency in time series analysis, unaffected by data distribution asymmetries or the presence of outliers.
+
 \paragraph{Variability.}
 Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}.
+Common measurements of the variability are the variance or standard deviation as well as their robust counter parts.
 In time series analyses, variability is conceptually important because information about the distribution and diversity of the data has been found to be indicative of worse psychological states \citep{myin-germeys2018, helmich2021}.
@@
 
-For our illustration data, we chose to capture the time series variability with the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \equatref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.
-We again chose the robust statistic because the Median-based measure is less affected by non-normal distributions and extreme values or outliers compared to other measures of variability like the standard deviation \citep{weisberg1992}
+For our illustration data, we chose the \textit{Median Absolute Deviation} ($MAD$) to gauge the variability within our time series data.
+This choice is motivated by the robustness of MAD, particularly its resilience to the effects of non-normal distributions and outliers, which can significantly skew traditional variability measures like the standard deviation \citep{weisberg1992}.
+For a given variable \( j \) and participant \( i \), the MAD is calculated by first determining the median (\( M \)) of the ordered time series \( \mathbf{X}_{ij} \) as outlined in \equatref{eq:median}.
+We then compute the absolute deviations of each observation in the time series \( X_{ij} \) from this median value.
+Specifically, for each time point \( t \), we calculate the absolute difference between \( x_{ijt} \) and the series median \( M(\mathbf{X}_{ij}) \).
+The MAD is then the median of these absolute deviations:
 
 ```{=tex}
-\begin{equation} \label{eq:mad}
-  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
-\end{equation}
+\begin{subequations}\label{eq:mad}
+    \begin{align} 
+      MAD(X_{ij}) &= M(\left| x_{ijt} - M(\mathbf{X}_{ij}) \right|) \label{eq:mad_general} \\
+                  &= M(\{ |x_{ij1} - M(\mathbf{X}_{ij})|, |x_{ij2} - M(\mathbf{X}_{ij})|, \ldots, |x_{ijn} - M(\mathbf{X}_{ij})| \}) \label{eq:mad_detailed}
+    \end{align}
+\end{subequations}
 ```
 
+The calculation of MAD focuses on the magnitudes of deviations, ensuring it provides a robust measure of dispersion that reflects the inherent variability in the time series data.
+
 \paragraph{Instability.}
-Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009}.
+Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009, jahng2008}.
 While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the 'jumpy-ness' of the data over time.
-In other words, variability reflects the range or diversity of values in a time series data, while instability reflects the fluctuation or inconsistency in a time series data over time  \citep{trull2008}.
-For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep{jahng2008}.
+In other words, variability reflects the range or diversity of values in the un-ordered time series data, while instability reflects the fluctuation or inconsistency in a time series data over time \citep{trull2008, houben2015, koval2013}.
+For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep[note that this does not need to be rapidly changing or instable, e.g., when there is linear increase over time; also see][]{jahng2008}.
 Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}.
@@
 
-Another common measurement of instability is the \textit{Mean of the Squared Successive Differences} ($MSSD$), which is often preferred where differences in magnitude are more important than the frequency of those changes, for example, when big shifts in time series are considered more impactful or when outliers are meaningful and need to be taken into account \citep{chatfield2003}.
+Another common measurement of instability is the \textit{Mean of the Squared Successive Differences} (\(MSSD\)), which is often preferred where differences in magnitude are more important than the frequency of those changes, for example, when big shifts in time series are considered more impactful or when outliers are meaningful and need to be taken into account \citep{chatfield2003, bos2019}.
+For psychological ESM data, some research suggests that amplitude and frequency might predict different health outcomes and can be investigated jointly \citep{wang2012a, jahng2008}.
 
-\paragraph{Self-similarity.}
-Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}.
-That is, self-similarity describes how much a measurement carries over to future measurements.
-One important self-similarity in psychological time series is \texit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}.
+\paragraph{Temporal dependence.}
+Temporal dependence in time series data refers to the degree to which a time series is influenced by its past values, exhibiting patterns of behavior that may be regular over different time scales \citep{dmello2021}.
+In the context of psychological time series, an important aspect of temporal dependence is \textit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}.
 If inertia is high a development tends to stay in a certain state.
 Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of under-reactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}.
-In a similar vein, high inertia in negative affect at baseline was even predictive of the initial onset of depression \citep{kuppens2012}.
-Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a negative mood for several measurements?'' or ``Do migrants stay with their language practice for several days at a time?''
+In a similar vein, high inertia in negative affect at baseline was predictive of the initial onset of depression \citep{kuppens2012}.
+Conceptually, inertia is more broadly connected to research questions such as: "Do patients stay in a negative mood for several measurements?" or "Do migrants stay with their language practice for several days at a time?"
 
@@
 ```
+
+While inertia captures the simplest case of temporal dynamics, lag-1, we acknowledge that temporal dependence in psychological time series may also exhibit more complex relationships, including higher lagged auto correlations or cyclical relationships (fourier coefficients, or continuous wavelet transforms are often used to capture such relationships).
 
@@
 Changes in psychology are not always linear, instead, nonlinearity is a common feature of psychological time series \citep{hayes2007}.
-As an example, episodic disorders, such as depression, are most likely best described as non-linear systems \citep{hosenfeld2015}.
+As an example, episodic disorders, such as depression, are often best described as non-linear systems \citep{hosenfeld2015}.
 Similarly, patients in recovery from depression showed sudden changes in the improvement of depression \citep{helmich2020a}.
@@
 We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model.
-The $edf$ summarizes the \textit{wiggliness} of a spline trend line \citep{wood2017}.
+The $edf$ summarizes the \textit{wiggliness} of a spline trend line \citep{wood2017, bringmann2017}.
 The degrees of freedom of a spline model are primarily determined by the number of knots and the order of the spline.
-For instance, a cubic spline with $k$ knots has $k$+3 degrees of freedom \citep{faraway2016}.
+For instance, a cubic spline with $k$ knots has $k$+3 degrees of freedom \citep{faraway2016, haslbeck2021a, castro-alvarez2024}.
 However, in a penalized spline framework, which is commonly used for GAMs, the effective degrees of freedom can be less than $k$+3.
@@
 
+\paragraph{Additional considerations.}
 Beyond our main features of interest, we also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard (i.e., to exclude spurious explanations for the cluster assignments).
@@
 However, even the few missing values can be an issue for some feature reduction or feature clustering algorithms.
-We, thus, impute the missing feature values with a single predictive mean matching imputation using the MICE library \citep[][]{buuren2011}.
+We, thus, impute missing feature values via predictive mean matching (PMM) with the MICE package in R, employing a single imputation and specifying a maximum of 50 iterations and a fixed seed for convergence and reproducibility \citep[][]{buuren2011}.
+We chose PMM for its ability to preserve the original data distribution without assuming normality and robustly handling multiple data types \citep{vanbuuren2006}.
 Note again that with this procedure we only need to impute an extremely small number of missing values as most feature calculations can use the available data instead.

### 2024-11-11 sections/04FeatureReduction.Rmd
config updated

@@
 However, PCAs have the distinct benefit that they are well-established within the psychometric literature \citep{jolliffe2011} and can broadly be applied to a wide variety of studies in an automatized manner \citep{abdi2010}.
-As our aim is to present a general illustration that can also be adopted across use cases, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well.
+As our aim is to present a general illustration that can also be adopted across use cases, we present the workflow using a PCA here but we encourage users to consider more specialized methods as well (we provide an example decision guide in \sicontext).
 
-To use the PCA with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally.
+To use the PCA with our extracted time series features, we first standardize all features across participants to ensure that all features are weighted equally \citep[][]{horne2020}.
 We then enter all `r ncol(scaled_data)` features into the analysis.
@@
 In particular, because the principal components are linear combinations of the original features, users can extract the relative importance of each feature for the extracted principal components (i.e., the eigenvectors).
-While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA purely to reduce the dimensionality for the clustering step.
+While this can be useful in understanding the variance in the original data or help with manual feature selection, we use the PCA here purely to reduce the dimensionality for the clustering step.
 Instead of relying on the principal components, we use the original features of interest to interpret the later extracted clusters.

### 2024-11-11 sections/06ClusterEvaluationPerformance.Rmd
config updated

@@
 for(i in 2:10){
-  kmeansPerf[[i-1]] <- fpc::cluster.stats(d = dist(scaled_pc), clustering = kmeansRes[[i-1]]$cluster, G2 = TRUE, G3 = TRUE, aggregateonly = TRUE) %>% unlist
+  kmeansPerf[[i-1]] <- fpc::cluster.stats(d = dist(scaled_pc), clustering = kmeansRes[[i]]$cluster, G2 = TRUE, G3 = TRUE, aggregateonly = TRUE) %>% unlist
 }
@@
   "dunn2" = "Dunn index 2", # minimum average dissimilarity between two cluster / maximum average within cluster dissimilarity, another version of the family of Dunn indexes
-  "entropy" = "entropy of distribution of cluster memberships", # see Meila(2007)
+  "entropy" = "entropy of distribution of cluster memberships", # see Meila (2007)
   "wb.ratio" = "average within / average between", 
@@
 
+set.seed(12345)
 kmeansOut <- kmeans(scaled_pc, centers = 2, nstart = 100)
@@
 In our own illustration example, we used the \texttt{cluster.stats()} function from the \texttt{fpc} \textsf{R} package, which calculates a wide variety of internal cluster validity statistics for each of the extracted clustering solutions.
-With real-world data no single evaluation measure is likely perfect, and different measures may produce different results depending on the characteristics of the data and the research question being addressed \citep{kittler1998}.
+With real-world data, no single evaluation measure is likely to be perfect.
+Different measures may yield varying results based on the data characteristics and the research question at hand \citep{kittler1998}.
 It is therefore important to consider a variety of evaluation measures and to carefully interpret the results in the context of the specific analysis \citep{vinh2009}.
@@
 The first statistic we looked at was the total within-cluster sum of square $WCSS$ (see also \equatref{eq:kWCSS}).
-While the within-cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$ after which the decrease was much smaller.
+While the within-cluster variation will naturally decrease with (more) smaller clusters, we observed that the decrease in $WCSS$ was largest until $k=2$, after which the decrease was much smaller.
 This method is also sometimes referred to as the 'elbow method' \citep{syakur2018}.
@@
 In the final $k=2$ solution the k-means analysis also assigned a relatively even number of participants to cluster 1 ($n_{C_1}=$ `r table(kmeansOut$cluster)[1]`) and cluster 2 ($n_{C_1}=$ `r table(kmeansOut$cluster)[2]`).
+
+```{r}
+#| label: test single cluster solution
+
+library(cluster)
+
+numeric_data <- scaled_pc[, sapply(scaled_pc, is.numeric)]
+
+# also calculate k=1 solution
+kmeans_one_cluster <- kmeans(numeric_data, centers = 1, nstart = 25)
+kmeans_two_clusters <- kmeans(numeric_data, centers = 2, nstart = 25)
+
+# Compare the total within-cluster sum of squares
+total_withinss_one_cluster <- kmeans_one_cluster$tot.withinss
+total_withinss_two_clusters <- kmeans_two_clusters$tot.withinss
+
+# Calculate the total sum of squares
+total_ss <- sum(sapply(numeric_data, function(x) sum((x - mean(x))^2)))
+var_explained_one_cluster <- 1 - (total_withinss_one_cluster / total_ss)
+var_explained_two_clusters <- 1 - (total_withinss_two_clusters / total_ss)
+
+# create table output
+singleClusterPerf <- data.frame(
+  Clusters = c("One Cluster", "Two Clusters"),
+  TotalWithinSS = c(total_withinss_one_cluster, total_withinss_two_clusters),
+  VarianceExplained = c(var_explained_one_cluster, var_explained_two_clusters)
+)
+singleClusterPerf$VarianceExplained <-
+  ifelse(
+    singleClusterPerf$VarianceExplained < 0.001,
+    "<.001",
+    as.character(round(singleClusterPerf$VarianceExplained, 3))
+  )
+singleClusterPerfNames <- c("Cluster Count", "Total Within-Cluster SS", "Variance Explained")
+
+# singleClusterPerf %>%
+#   kbl(.,
+#       escape = TRUE,
+#       booktabs = TRUE,
+#       align = c("l", "c", "c"),
+#       digits = 3,
+#       row.names = FALSE,
+#       col.names = singleClusterPerfNames,
+#       caption = "K-Means Single Cluster Performance Comparison") %>%
+#   kable_classic(
+#     full_width = FALSE,
+#     lightable_options = "hover",
+#     html_font = "Cambria"
+#   )
+```
+
+To ensure that the clustering is necessary in the first place, we also compare the performance to a single cluster solution (i.e., a single centroid).
+The comparison to this \(k=1\) solution is slightly different because metrics like the between-cluster separation are not available.
+Nonetheless, comparing the within-cluster sums of squares (SS) and the explained variance, we find that two clusters indeed outperform a single cluster solution.
+Specifically, the total within-cluster SS decreased from `r format(filter(singleClusterPerf, Clusters == "One Cluster")$TotalWithinSS, digits=2, nsmall=2)` for one cluster to `r format(filter(singleClusterPerf, Clusters == "Two Clusters")$TotalWithinSS, digits=2, nsmall=2)` for two clusters.
+Additionally, the variance explained increased from `r format(filter(singleClusterPerf, Clusters == "One Cluster")$VarianceExplained, digits=3, nsmall=3)` to `r format(filter(singleClusterPerf, Clusters == "Two Clusters")$VarianceExplained, digits=3, nsmall=3)` when the cluster count was increased to two \citep[e.g., ][; also see \situtorial\ for full results]{beijers2022}.

### 2024-11-11 sections/07ClusterEvaluationInterpretation.Rmd
config updated

@@
 # EXTRACT OUT-OF-CLUSTER VALIDATION VARIABLES
+dt_raw <- readRDS(file="tutorial/data/osf_mini.Rda")
+missingness_ooc <- full_join(
+  featData %>%
+    group_by(study, PID) %>%
+    summarise(n_feature = n()) %>%
+    ungroup(),
+  dt_raw %>%
+    group_by(study, PID) %>%
+    summarise(n_raw = n()) %>%
+    ungroup(),
+  by = c("study", "PID")
+) %>%
+  mutate(
+    nrm = n_raw - n_feature
+  ) %>%
+  filter(
+    !is.na(n_feature)
+  ) %>%
+  mutate(nrm.z = as.numeric(scale(nrm))) %>%
+  select(study, PID, nrm.z)
+
 validation <- rbind(
@@
 ) %>% 
+  left_join(., missingness_ooc, by = c("study", "PID")) %>%
   group_by(study, PID) %>%
@@
     gender = gender.pre,
+    nrm.z = nrm.z,
     IntGrAnx.pre = rowMeans(across(starts_with("IntGrAnx")), na.rm = TRUE),
@@
     #SWL.pre.z,
-    EvDayDiscr.post.z
+    EvDayDiscr.post.z,
+    nrm.z
   )
@@
 
-In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
-We find an adaptive group (cluster 1) that reports higher well-being and more positive outgroup attitudes (\textit{median}) that are also stable over time (\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test period (\textit{linear trend}).
-This group also reported consistently more meaningful, need-fulfilling, and cooperative outgroup interactions (\textit{median}).
-This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 2).
-This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
-This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
-At the same time, for members of this detrimental cluster (cluster 2) conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}).
-
-To identify these patterns, we first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
-We see that for some variables the features are generally stronger in separating the clusters.
-We, for example, see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
-Compare this to the '\textit{outgroup attitudes}' item where the differences between the clusters are much smaller for almost all features.
-We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
-While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables.
-For example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
-These two features stand in stark contrast to other features, such as the \textit{lag-1 auto correlations}, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar}).
-This offers some information on which features were most important in understanding the two extracted groups.
-Taking these two perspectives together, we can also focus on individual features or variables in particular.
-We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series.
-At the same time, in terms of stability, both groups have virtually identical average \textit{MAC} statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}).
-There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
-In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
-We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a}.
-
-```{=tex}
-\begin{figure}[!ht] %hbtp
-  \caption{Cluster Group Comparisons based on Features and Variables}
-  \label{fig:clusterFeatVar}
-  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
-  \caption*{Note: \\
-  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
-\end{figure}
-```
-
-In the second step, we look at the prototypical trajectories of the clusters.
-For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.
-Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}.
-For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
-Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions and also consistently reported more voluntary and cooperative interactions but fewer accidental and involuntary interactions.
-The same cluster (cluster 1) also reported an increase in need-fulfilling interactions over the 30-day period and an increase in interactions that were representative of the outgroup.
-Whereas the other cluster (cluster 2) showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
-This `deterioration' cluster (cluster 2) also saw a decrease in general need fulfillment and experienced well-being over the 30 days (see \fgrref[B]{fig:clusterTs}).
-We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variability and instability on most of the other variables (also see \fgrref[A]{fig:clusterTs}).
-
 ```{r}
@@
 
+plot_feature_selection <- c(
+  "median",
+  "mad",
+  "lin",
+  "edf"
+)
+feature_limits <- results_df %>%
+  filter(feature %in% plot_feature_selection) %>% 
+  summarise(
+    min_lwr = min(lwr, na.rm = TRUE),
+    max_upr = max(upr, na.rm = TRUE)
+  ) %>%
+  as.numeric()
+
+feature_comparison_combined <-
+  ggpubr::ggarrange(
+    feature_plot_list$median +
+      scale_x_continuous(limits = feature_limits) +
+      theme(
+        axis.title.x = element_blank(),
+        plot.margin = margin(
+          t = 20,
+          r = 1,
+          b = 20,
+          l = 20
+        )
+      ),
+    feature_plot_list$mad +
+      scale_x_continuous(limits = feature_limits) +
+      theme(
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(
+          t = 20,
+          r = 20,
+          b = 20,
+          l = 20
+        )
+      ),
+    feature_plot_list$lin +
+      scale_x_continuous(limits = feature_limits) +
+      theme(# aspect.ratio = 1,
+        plot.margin = margin(
+          t = 20,
+          r = 1,
+          b = 10,
+          l = 20
+        )),
+    feature_plot_list$edf +
+      scale_x_continuous(limits = feature_limits) +
+      theme(
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(
+          t = 20,
+          r = 20,
+          b = 10,
+          l = 20
+        )
+      ),
+    ncol = 2,
+    nrow = 2,
+    widths = c(1.325, 1),
+    # align = "hv",
+    labels = c("(A) Median", "(B) MAD", "(C) Linear Slope", "(D) GAM edf"),
+    hjust = c(-0.25, -0.27, -0.17, -0.22),
+    common.legend = TRUE,
+    legend = "bottom"
+  ) 
+ggsave("Figures/feature_comparison_combined.pdf", feature_comparison_combined, width = 12, height = 12)
+
 # Check if any graphics devices are open
@@
 }
+
 ```
+
+In short, we find that the feature-based clustering discerned two meaningfully different groups of participants.
+We find an adaptive group (cluster 1) that reports higher well-being (\textit{median}: `r feat_ttest_apa(results_df, "exWB", "median")`; also see \fgrref[A]{fig:cluster_comparison_features}) and more positive outgroup interactions (\textit{median}: `r feat_ttest_apa(results_df, "qualityOverall", "median")`), which are also stable over time (\textit{MAC}: `r feat_ttest_apa(results_df, "qualityOverall", "mac")`) and tend to increase more over the 30 day test period (\textit{linear trend}: `r feat_ttest_apa(results_df, "qualityOverall", "lin")`; also see \fgrref[C]{fig:cluster_comparison_features}).
+This group also reported consistently more meaningful (\textit{median}: `r feat_ttest_apa(results_df, "qualityMeaning", "median")`), need-fulfilling (\textit{median}: `r feat_ttest_apa(results_df, "KeyNeedFulfillment", "median")`), and cooperative outgroup interactions (\textit{median}: `r feat_ttest_apa(results_df, "InteractionContextCooperative", "median")`).
+This group with overwhelmingly positive experiences stands in contrast with a more detrimental group (cluster 2).
+This cluster, on average, reported much less positive, less meaningful, and less fulfilling interactions and interaction patterns (\textit{median}).
+This group also reported less positive outgroup attitudes, lower well-being, and more discrimination experiences (\textit{median}).
+At the same time, for members of this detrimental cluster (cluster 2) conditions seemed to deteriorate over time (\textit{linear trend}), and there was generally less consistency in the experiences they were able to have (\textit{MAC}, \textit{MAD}, \textit{edf}; also see \fgrref[]{fig:cluster_comparison_features}; for a full and interactive comparison of all features see \situtorial).
+
+To identify these patterns, we first inspect the clusters based on the average values of meaningful features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}).
+We see that for some variables the features are generally stronger in separating the clusters.
+We, for example, see that the item on '\textit{how cooperative the interaction was}' distinguishes the two clusters across almost all seven features (except for the \textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}).
+Compare this to the '\textit{outgroup attitudes}' item where the differences between the clusters are much smaller for almost all features.
+We then inspect the clusters with a focus on the features (see \fgrref[B]{fig:clusterFeatVar}).
+While this is the same data as for the variable focus, we can see more clearly that some features are better at distinguishing the clusters across variables.
+For example, \textit{MAD} and \textit{median} distinguish the two clusters across almost all variables (except for the item of whether the interaction was representative of the outgroup).
+These two features stand in stark contrast to other features, such as the \textit{lag-1 auto correlations} or the \textit{GAM edf}, which showed much smaller differences between the two clusters (see \fgrref[B]{fig:clusterFeatVar} and \fgrref[]{fig:cluster_comparison_features}; please note that we offer readers an interactive tool to assess the cluster differences for all features in \situtorial).
+This offers some information on which features were most important in differentiating the two extracted groups but also shows that with real-world data, not all features will have enough range to distinguish people on all variables (e.g., see the non-linearity patterns in \fgrref[]{fig:cluster_comparison_features}; for a more direct illustration of GAM edf differences see \citealp{bringmann2017}).
+Taking these two perspectives together, we can also focus on individual features or variables in particular.
+We, for example, see a strong difference in the average well-being, where participants in cluster 2 showed a much lower median well-being over the time series (`r feat_ttest_apa(results_df, "exWB", "median")`).
+At the same time, in terms of well-being stability, both groups have virtually identical average \textit{MAC} statistics for well-being (`r feat_ttest_apa(results_df, "exWB", "mac")`; also see \fgrref[A]{fig:clusterFeatVar}).
+There are, thus, variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail.
+In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2).
+We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics \citep{bringmann2018c, eronen2021a, dejonckheere2019}.
+
+```{=tex}
+\begin{figure}[!ht] %hbtp
+  \caption{Cluster Group Comparisons based on Features and Variables}
+  \label{fig:clusterFeatVar}
+  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb_tutorial.pdf}
+  \caption*{Note: \\
+  "Int." = outgroup interaction, "mad" = median absolute deviation, "mac" = mean absolute change, "lin" = linear slope, "edf" = estimated degrees of freedom of an empty GAM spline model, "ar01" = lag-1 autocorrelation, "OCC"/"occ" = out-of-cluster comparison\\
+  Within the "(B) Feature Focus" subplot, the 'n (within ooc)' is an out-of-cluster comparison of the within-person available measurements for each variable; the 'between ooc (mean)' are also out-of-cluster comparisons but on a between person level.
+'Measurements removed' is the person-specific count of measurement occasions removed during the missingness handling and 'Discrimination' is the scale mean of daily discimination experiences (measured during the final survey).}
+\end{figure}
+```
+
+In the second step, we look at the prototypical trajectories of the clusters.
+For k-means clustering it is often recommended to use the average over time of the responses within the cluster \citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.}.
+Immediately striking are the mean differences, where participants in cluster 1 had more meaningful and fulfilling outgroup interactions and also consistently reported more voluntary and cooperative interactions but fewer accidental and involuntary interactions.
+The same cluster (cluster 1) also reported an increase in need-fulfilling interactions over the 30-day period and an increase in interactions that were representative of the outgroup.
+Whereas the other cluster (cluster 2) showed a decrease in voluntary, cooperative, and positive interactions over the 30 days.
+This `deterioration' cluster (cluster 2) also saw a decrease in general need fulfillment but not experienced well-being over the 30 days (see \fgrref[C]{fig:cluster_comparison_features}).
+We also see that while interaction representativeness, outgroup attitudes, and well-being are relatively stable for both clusters, the deteriorating cluster (cluster 2) also showed substantially higher variability and instability on most of the other variables (although these effects are much smaller; see \fgrref[A]{fig:clusterTs}).
 
@@
 The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are substantially larger due to the larger samples \citep[e.g., restriction of range in the smaller sample][]{kogan2006}.
-In our case, this seems less likely because one out of four variables did not differ in terms of the MAD (i.e., our selected measurement of the time series variance; see \fgrref{fig:cluster_comparison_mad} for an illustration).
+In our case, this seems less likely because one out of four variables did not differ in terms of the MAD (i.e., our selected measurement of the time series variance; see \fgrref{fig:cluster_comparison_features} for an illustration).
 At the same time, however, the difference in the number of experienced interactions might also indicate a meaningful difference, where the deteriorating cluster (cluster 2) on average reported more outgroup interactions (`r feat_ttest_apa(results_df, "InteractionContextvoluntary", "n")`), but these interactions were less voluntary (`r feat_ttest_apa(results_df, "InteractionContextvoluntary", "median")`), less meaningful (`r feat_ttest_apa(results_df, "qualityMeaning", "median")`), and less positive (`r feat_ttest_apa(results_df, "qualityOverall", "median")`).
@@
 Such a finding would also be in line with past research highlighting the role of negative intergroup interactions in explaining intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.
+A related validity check was the inclusion of the missingness handling, where we compared the two clusters on the average number of measurements removed as part of the missingness handling.
+We find that the clusters did not significantly differ on this metric suggesting that the missingness handling did not affect the cluster separation (also see \appref[]{app:ValidationAnalyses} and \situtorial).
 
@@
 \begin{figure}[!ht] %hbtp
-  \caption{Cluster comparison of Median Absolute Deviation for all variables}
-  \label{fig:cluster_comparison_mad}
-  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_mad.pdf}
+  \caption{Comparison Cluster Differences by Features and Variables.}
+  \label{fig:cluster_comparison_features}
+  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_combined.pdf}
+  \caption*{Note:\\
+  The figure shows the differences between the clusters in the standardized features that were entered into the dimensionality reduction (for each input variable).
+We display the median (panel A), the median absolute deviation (MAD, panel B), the univariate linear slope (panel C), as well as the estimated degrees of freedom of the generalized additive model splines (GAM edf, panel D).
+Please also note that as part of \situtorial, we provide readers with an interactive selection tool to compare cluster differences on all variables and features.}
 \end{figure}
@@
   ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
+  geom_line(stat="smooth", method = "gam", aes(group=ID, color=as.factor(cluster)), se = FALSE, alpha=0.1) +
+  geom_smooth(aes(group=cluster), method = "gam", se = FALSE) +
   scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
@@
   ncol = 1,
-  labels = c("(A) Timepoint Means", "(B) Average Spline Trends"),
+  labels = c("(A) Timepoint Means", "(B) Spline Trends"),
   common.legend = TRUE,

### 2024-11-11 tutorial/05-evaluation.qmd
tutorial update
config updated

@@
   "factoextra",
-  "fpc"
+  "fpc",
+  "shiny",
+  "shinyWidgets",
+  "shinyjs",
+  "bslib"
 )
@@
 
-set.seed(42)
+set.seed(12345)
 kmeans_final <- kmeans(pca_scores, centers = 2, nstart = 100)
+kmeans_final$cluster
 
+saveRDS(kmeans_final, file = "data/kmeans_final.Rds")
 ```
@@
 
+::: scrolling
 ```{r}
@@
 
-featCluster <- data.frame(ID = as.numeric(names(kmeans_final$cluster)), 
-                            cluster = kmeans_final$cluster) %>%
-  left_join(z_data, by = "ID") %>%
-  left_join(
+# calculate number of removed datapoints
+missingness_ooc_id <- full_join(
   featData %>%
+    mutate(uid = paste(study, PID, sep = "_")) %>%
+    group_by(uid, ID) %>%
+    summarise(n_feature = n()) %>%
+    ungroup(),
+  dt_raw %>%
+    mutate(uid = paste(study, PID, sep = "_")) %>%
+    group_by(uid) %>%
+    summarise(n_raw = n()) %>%
+    ungroup(),
+  by = "uid"
+) %>%
+  mutate(
+    n_rm = n_raw - n_feature
+  ) %>%
+  filter(
+    !is.na(ID)
+  ) %>%
+  mutate(nrm_z = scale(n_rm)) %>%
+  select(ID, nrm_z)
+discrimination_ooc_id <- featData %>% 
   select(ID, EvDayDiscr.post) %>% 
@@
     EvDayDiscr.post_z
-      ), 
-    by = "ID"
-  ) %>%
+  )
+
+featCluster <- data.frame(ID = as.numeric(names(kmeans_final$cluster)), 
+                            cluster = kmeans_final$cluster) %>%
+  left_join(z_data, by = "ID") %>%
+  left_join(discrimination_ooc_id, by = "ID") %>% 
+  left_join(missingness_ooc_id, by = "ID") %>%
   select(-ID) %>%
@@
   mutate(feature = ifelse(feature == "z", "mean", feature)) %>%
-  left_join(var_meta %>% select(name, label), by = c("variable" = "name")) %>%
+  left_join(var_meta %>% select(name, label) %>% rbind(c("nrm", "OOC: Measurements removed")), by = c("variable" = "name")) %>%
   select(
@@
     everything()
-  )
+  ) %>% 
+  mutate(feature = gsub("^n$", "n (within ooc)", feature)) %>% 
+  mutate(feature = gsub("^mean$", "between ooc (mean)", feature))
+
 ```
+:::
 
@@
   filter(variable != "EvDayDiscr.post",
-         feature != "n") %>%
+         variable != "nrm",
+         feature != "n (within ooc)") %>%
   ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
@@
 clusterFeatureGrid <- featCluster %>% 
-  ggplot(., aes(x = reorder(label, -as.numeric(label == "Post: Discrimination")), y = mean, group = cluster, color = as.factor(cluster))) +
+  ggplot(., aes(x = reorder(label, as.numeric(!label %in% c("OOC: Discrimination", "OOC: Measurements removed"))), y = mean, group = cluster, color = as.factor(cluster))) +
   geom_point() +
@@
 
-ggpubr::ggarrange(
+clusterFeatVarComb <- ggpubr::ggarrange(
   clusterVariableGrid + theme(axis.title.x = element_blank()),
@@
 ) 
+ggsave("../Figures/clusterFeatVarComb_tutorial.pdf", clusterFeatVarComb, width = 12, height = 12)
+clusterFeatVarComb
 ```
@@
   ggplot(., aes(x=TIDnum, y=value, group=cluster, color=as.factor(cluster))) +
-  geom_line(aes(x=TIDnum, y=value, group=ID, color=as.factor(cluster)), alpha=0.05) +
-  geom_smooth(aes(group=cluster)) +
+  geom_line(stat="smooth", method = "gam", aes(group=ID, color=as.factor(cluster)), se = FALSE, alpha=0.1) +
+  geom_smooth(aes(group=cluster), method = "gam", se = FALSE) +
   scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
-  #stat_summary(fun=mean, geom="line") +
   facet_wrap(~ label) +
+  scale_y_continuous(limits = c(0, 100)) +
   labs(
@@
 
-### Out-of-cluster Comparison
+### (Out-of-)Cluster Comparison
 
@@
 The higher average number of interactions in cluster 2 could, for example, indicate a clustering artifact if variances are substantially larger due to the larger samples (e.g., restriction of range in the smaller sample; @kogan2006).
-In our case, this seems less likely because one out of four variables did not differ in terms of the MAD (i.e., our selected measurement of the time series variance; see [@fig-mad-cluster-comparison] below).
+In our case, this seems less likely because one out of four variables did not differ in terms of the MAD (i.e., our selected measurement of the time series variance; see plot selection below).
 
+::: scrolling
 ```{r}
@@
 
-# validation feature
-validation_feature <-
+# validation features
+missingness_ooc <- full_join(
+  featData %>%
+    mutate(uid = paste(study, PID, sep = "_")) %>%
+    group_by(uid, ID) %>%
+    summarise(n_feature = n()) %>%
+    ungroup(),
+  dt_raw %>%
+    mutate(uid = paste(study, PID, sep = "_")) %>%
+    group_by(uid) %>%
+    summarise(n_raw = n()) %>%
+    ungroup(),
+  by = "uid"
+) %>%
+  mutate(
+    n_rm = n_raw - n_feature
+  ) %>%
+  filter(
+    !is.na(ID)
+  ) %>%
+  select(ID, n_rm) %>%
+  mutate(n_rm_z = scale(n_rm)) %>%
+  left_join(data.frame(ID = as.numeric(names(kmeans_final$cluster)),
+             cluster = kmeans_final$cluster), by = "ID")
+t_missingness <- t.test(
+  missingness_ooc$n_rm_z[missingness_ooc$cluster == 1],
+  missingness_ooc$n_rm_z[missingness_ooc$cluster == 2],
+  na.action = na.omit
+)
+
+discrimination_ooc <-
   featData %>% 
@@
 t_discrimination <- t.test(
-  validation_feature$EvDayDiscr.post_z[validation_feature$cluster == 1],
-  validation_feature$EvDayDiscr.post_z[validation_feature$cluster == 2],
+  discrimination_ooc$EvDayDiscr.post_z[discrimination_ooc$cluster == 1],
+  discrimination_ooc$EvDayDiscr.post_z[discrimination_ooc$cluster == 2],
   na.action = na.omit
@@
   data.frame(
-    concept = "EvDayDiscr.post",
-    feature = "mean",
-    difference = t_discrimination$estimate[1] - t_discrimination$estimate[2],
-    t_value = t_discrimination$statistic,
-    df = t_discrimination$parameter,
-    p_value = t_discrimination$p.value,
-    lwr = t_discrimination$conf.int[1],
-    upr = t_discrimination$conf.int[2]
+    concept = c("EvDayDiscr.post", "missingness"),
+    feature = c("mean", "mean"),
+    difference = c(t_discrimination$estimate[1] - t_discrimination$estimate[2], t_missingness$estimate[1] - t_missingness$estimate[2]),
+    t_value = c(t_discrimination$statistic, t_missingness$statistic),
+    df = c(t_discrimination$parameter, t_missingness$parameter),
+    p_value = c(t_discrimination$p.value, t_missingness$p.value),
+    lwr = c(t_discrimination$conf.int[1], t_missingness$conf.int[1]),
+    upr = c(t_discrimination$conf.int[2], t_missingness$conf.int[2])
   )
@@
 
+results_df[results_df$concept=="missingness",colnames(results_df)=="label"] <- "Missingness (N removed)"
+
 feature_plot_list <- generate_feature_plots(results_df)
+```
+:::
 
-feature_plot_list$mad
+```{r}
+#| label: render hidden plots
+#| echo: false
+#| results: asis
+
+for (name in names(feature_plot_list)) {
+  # Only hide plots that are not plot_mad
+  cat(paste0('<div id="plot_', name, '" style="display: ', ifelse(name == "mad", 'block', 'none'), ';">'))
+  print(feature_plot_list[[name]])
+  cat('</div>')
+}
 ```
+
+<select id="plot_selector" onchange="showPlot()">
+  <option value="plot_mad" selected>MAD</option>
+  <option value="plot_median">Median</option>
+  <option value="plot_n">Measurements</option>
+  <option value="plot_lin">linear slope</option>
+  <option value="plot_mac">MAC</option>
+  <option value="plot_edf">GAM edf</option>
+  <option value="plot_ar01">AR lag1</option>
+  <option value="plot_mean">Validation</option>
+</select>
+
+<script>
+function showPlot() {
+  var selectedPlot = document.getElementById("plot_selector").value;
+  var plots = document.querySelectorAll('div[id^="plot_"]');
+
+  plots.forEach(function(plot) {
+    if (plot.id === selectedPlot) {
+      plot.style.display = 'block';
+    } else {
+      plot.style.display = 'none';
+    }
+  });
+}
+
+// Call showPlot on page load to display the default plot
+document.addEventListener('DOMContentLoaded', showPlot);
+</script>
 

### 2024-11-11 tutorial/06-validation-analyses.qmd
config updated

---
title: "6. Validation Analyses"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "readxl",
  "tidyverse",
  "kableExtra",
  "reshape",
  "knitcitations",
  "factoextra",
  "fpc"
)

invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
anytime::addFormats("%Y-%m-%d %H:%M:%S")
anytime::addFormats("%d-%m-%Y")

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

source("scripts/load_data.R")

```

```{r}
#| label: import data
#| include: FALSE

dt_raw <- readRDS(file="data/osf_mini.Rda")
var_meta <- readxl::read_excel("data/osf_var_meta.xlsx")

scaled_pc <- readRDS(file = "data/pca_scores.Rds")
```

To further understand our cluster solution and ensure the validity of our results, we ran a number of validation analyses. Within these additional analyses, we assess a single cluster solution, the impact of the missingness handling as well as an alternative analysis with a reduced feature set focusing on non-dynamic features. As an additional resource, we also developed an interactive web application to try out different method and parameter options.

## Single Cluster Solution

To ensure that the clustering is necessary in the first place, we also compare the performance to a single cluster solution (i.e., a single centroid). The comparison to this $k=1$ solution is slightly different because metrics like the between cluster separation are not available. Nonetheless, comparing the within-cluster sums of squares and the explained variance, we find that two clusters indeed are outperforming a single cluster solution.

```{r}
#| label: test single cluster solution

library(cluster)

numeric_data <- scaled_pc[, sapply(scaled_pc, is.numeric)]

# also calculate k=1 solution
kmeans_one_cluster <- kmeans(numeric_data, centers = 1, nstart = 25)
kmeans_two_clusters <- kmeans(numeric_data, centers = 2, nstart = 25)

# Compare the total within-cluster sum of squares
total_withinss_one_cluster <- kmeans_one_cluster$tot.withinss
total_withinss_two_clusters <- kmeans_two_clusters$tot.withinss

# Calculate the total sum of squares
total_ss <- sum(sapply(numeric_data, function(x) sum((x - mean(x))^2)))
var_explained_one_cluster <- 1 - (total_withinss_one_cluster / total_ss)
var_explained_two_clusters <- 1 - (total_withinss_two_clusters / total_ss)

# create table output
singleClusterPerf <- data.frame(
  Clusters = c("One Cluster", "Two Clusters"),
  TotalWithinSS = c(total_withinss_one_cluster, total_withinss_two_clusters),
  VarianceExplained = c(var_explained_one_cluster, var_explained_two_clusters)
)
singleClusterPerf$VarianceExplained <-
  ifelse(
    singleClusterPerf$VarianceExplained < 0.001,
    "<.001",
    as.character(round(singleClusterPerf$VarianceExplained, 3))
  )
singleClusterPerfNames <- c("Cluster Count", "Total Within-Cluster SS", "Variance Explained")
singleClusterPerf %>%
  kbl(.,
      escape = TRUE,
      booktabs = TRUE,
      align = c("l", "c", "c"),
      digits = 3,
      row.names = FALSE,
      col.names = singleClusterPerfNames,
      caption = "K-Means Single Cluster Performance Comparison") %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

## Missingness Handling

```{r}
#| label: Load data for steps that takes forever to run
#| include: FALSE

combined_data <- readRDS("data/missingness_thresholds/01_raw_combined/combined_data.rds")
feature_results <- readRDS("data/missingness_thresholds/02_features/features_all.rds")
analysis_results <- readRDS("data/missingness_thresholds/03_analysis/analysis_results_all.rds")

```

In our illustration data set, the studies differed substantially in the maximum length of participation ($max(t_{S1})$ = 63, $max(t_{S2})$ = 69, $max(t_{S3})$ = 155). To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45% missingness in the main text. While this procedure works well for users who wish to use the clustering in combination with other parametric models, the 45% threshold might be too conservative if the analysis stands on its own. We here show how the results would have differed if the missingness threshold is set more liberally.

### Sample Extraction

We now run the `cleanM` function for the completeness thresholds of 0%, 5%, 10%, 15%, 20%, 25%, 30%, 35%, 40%, 45%, 50%, as well as the original 55% (i.e., allowing up to 100% missingness). The non-conforming rows and colums are again sequentially remove staring with the row or column with the most missing data until only rows and columns remain that fit the threshold. The resulting missingess information and reduced datasets are stores in a new list called `missingness_samples`.

::: scrolling
```{r}
#| label: Data Availability Extraction
#| eval: false

source("scripts/cleanM.R")

# Function to prepare availability data for a given study
prepareAvailabilityData <- function(dt_raw, study_id) {
  dtAvailability <- dt_raw %>%
    filter(study == study_id) %>%
    select(PID, TIDnum) %>%
    arrange(PID, TIDnum) %>%
    mutate(data = 1)
  dtAvailability <- reshape::cast(dtAvailability, PID ~ TIDnum, value = "data", fun.aggregate = length) %>%
    select(-PID) %>%
    mutate_all(function(x) ifelse(x > 1, 1, x))
  return(dtAvailability)
}

# Function to process data for a given study and threshold
processStudyData <- function(dtAvailability, study_id, threshold) {
  dtRedInfo <- cleanM(M = as.data.frame(dtAvailability), c = threshold)
  PIDout <- gsub("PP_", "", dtRedInfo$rowNamesOut) %>% as.numeric
  TIDInRed <- gsub("t_", "", dtRedInfo$colNamesIn) %>% as.numeric

  # Check if TIDInRed has valid numeric values
  if (length(TIDInRed) > 0 && all(!is.na(TIDInRed))) {
    TIDIn <- seq(min(TIDInRed, na.rm = TRUE), max(TIDInRed, na.rm = TRUE), 1)
  } else {
    # Handle the case where TIDInRed is empty or invalid
    TIDIn <- numeric(0)  # An empty numeric vector
  }

  dtRed <- dt_raw %>%
    filter(study == study_id) %>%
    filter(
      !PID %in% PIDout,
      TIDnum %in% TIDIn
    ) %>%
    mutate(TIDnum = TIDnum - min(TIDnum, na.rm = TRUE))

  return(list(dtRedInfo = dtRedInfo, dtRed = dtRed))
}

# Initialize lists to store results
missingness_samples <- list()

# Define the studies and thresholds
studies <- c("S1", "S2", "S3")
completeness_thresholds <- seq(from = 0, to = 55, by = 5)

# Iterate over studies and thresholds
for (study in studies) {
  dtAvailability <- prepareAvailabilityData(dt_raw, study)
  rownames(dtAvailability) <- paste("PP", 1:nrow(dtAvailability), sep = "_")
  colnames(dtAvailability) <- paste("t", 1:ncol(dtAvailability), sep = "_")

  for (threshold in completeness_thresholds) {
    result_key <- paste(study, threshold, sep = "_")
    cat("\n\nSTUDY: ", study, "; THRESHOLD: ", threshold, "\n")
    missingness_samples[[result_key]] <- processStudyData(dtAvailability, study, threshold)
  }
}

```
:::

This allows us to store all the missingess infomation for each study. From this we can extract the reduced data sets for each threshold and combine them into a single reduced data set per threshold. We also save the compute intensive steps for these validation analyses.

::: scrolling
```{r}
#| label: Combine data sets by threshold
#| eval: false

# Initialize a list to store the combined data for each threshold
combined_data <- list()

# Unique thresholds are extracted from the list keys
unique_thresholds <- unique(sub(".*_", "", names(missingness_samples)))

# Loop through each threshold
for (threshold in unique_thresholds) {
  # Initialize an empty list to store data for each study at the current threshold
  study_data_list <- list()

  # Loop through each study and extract the corresponding dtRed data
  for (study in c("S1", "S2", "S3")) {
    key <- paste(study, threshold, sep = "_")
    dtRed <- missingness_samples[[key]]$dtRed

    # Apply the necessary transformations
    dtRed_transformed <- dtRed %>%
      select(-c("created", "ended")) %>%
      mutate(study = study) # %>%
      # mutate(across(!TID & !study, as.numeric))

    # Store the transformed data in the list
    study_data_list[[study]] <- dtRed_transformed
  }

  # Combine the data from all three studies for the current threshold
  dtAll <- bind_rows(study_data_list) %>%
    group_by(study, PID) %>%
    mutate(ID = cur_group_id()) %>%
    ungroup %>%
    mutate(date = as.Date(gsub(" .*", "", TID)),
           week = strftime(date, format = "%Y-W%V")) %>%
    select(ID, PID, study, date, week, TIDnum, everything()) %>%
    arrange(ID, TIDnum) %>%
    select_if(~ sum(!is.na(.)) > 1) %>%  # only include variables that have any data
    as.data.frame

  # Store the combined data frame in the list
  combined_data[[threshold]] <- dtAll
}

# Save so that we do not have to run this compute intensive step again
saveRDS(combined_data, "data/missingness_thresholds/01_raw_combined/combined_data.rds")
```
:::

### Feature Extraction

With the reduced data sets for each threshold extracted, we can extract the features for each participant within the reduced data sets. The process is exactly the same as for the main analysis just looping over the `thresholds` in the `combined_data` list. We store the resulting feature lists in the `feature_results` elment which keeps all relevant matrices and data.frames for each threshold dataset.

::: scrolling
```{r}
#| label: run feature extraction
#| eval: false

# devtools::install_github("JannisCodes/tsFeatureExtracR")
library(tsFeatureExtracR)

# Define the directory to store the results
base_dir <- "data/missingness_thresholds/02_features"

# Check if the directory exists; if not, create it
if (!dir.exists(base_dir)) {
  dir.create(base_dir, recursive = TRUE)
}

# Initialize a list to store the feature extraction results for each threshold
feature_results <- list()

feature_selection <- c(
  "ar01",
  "edf",
  "lin",
  "mac",
  "mad",
  "median"
)

# Loop through each dataset in combined_data
for (threshold in names(combined_data)) {
  dtAll <- combined_data[[threshold]]

  # Ensure the dataset only contains variables with any data
  featData <- dtAll %>%
    arrange(ID, TIDnum) %>%
    select_if(~ sum(!is.na(.)) > 1) %>% 
    as.data.frame

  # Check for duplicates and remove all occurrences
  duplicate_indices <- duplicated(featData[c("ID", "TIDnum")]) | duplicated(featData[c("ID", "TIDnum")], fromLast = TRUE)
  if (any(duplicate_indices)) {
    warning("Duplicate rows based on ID and TIDnum have been found. All instances will be removed.")
    # Remove all instances of duplicates
    featData <- featData[!duplicate_indices, ]
  }

  cat("\n\nExtracting Features for threshold: ", threshold, "\n")
  # Extract features for the current threshold dataset
  featFull <- featureExtractor(
    data = featData,
    pid = "ID",
    tid = "TIDnum",
    items = var_meta$name[var_meta$cluster == 1]
  )

  # Filter out unwanted data (e.g., internal test accounts)
  featFull$features <- featFull$features %>%
    filter(ID != 19)

  # Handle missing features
  perc_miss_features <- feature_missing(featFull$features %>% 
                    select(ends_with(feature_selection)), 
                  title = paste("Feature-wise Missingness for threshold:", threshold, "percent"))

  cat("Imputing missing features for threshold: ", threshold, "\n")
  # Impute missing features
  featFullImp <- featureImputer(featFull)

  # Store the results
  feature_results[[threshold]] <- list(
    features = featFullImp,
    missingness_plot = perc_miss_features$plt_miss_per_feature
  )
  cat("Added features for threshold: ", threshold, ".\n")

  # Save locally 
  features_path <- file.path(base_dir, paste0("features_", threshold, ".rds"))
  missingness_plot_path <- file.path(base_dir, paste0("missingness_plot_", threshold, ".png"))

  # Save the feature results
  saveRDS(feature_results[[threshold]]$features, features_path)

  # Save the missingness plot
  png(missingness_plot_path)
  print(feature_results[[threshold]]$missingness_plot)
  dev.off()
  cat("Feature results and missingness plots for threshold: ", threshold, " have been saved in", base_dir, "\n")
}

# Save so that we do not have to run this compute intensive step again
saveRDS(feature_results, "data/missingness_thresholds/02_features/features_all.rds")
cat("All feature results have been saved in", base_dir, ". \n\n*Done!*")

```
:::

### Run Analyses

We run the same set of analyses (i.e., a PCA and a k-means clustering) for each of the feature sets that we extracted based on the different thresholds. To full automate this process, we again select the principal components to capture 80% of the original variance and we use a number of different performance metrics for different $k$s in the k-means analysis to select the $k$ that is optimal according to the most metrics. We also set a random seed `set.seed(12345)` for strict reproducibility.

::: scrolling
```{r}
#| label: run pca and kmeans
#| eval: false

analysis_results <- list()

for (threshold in names(feature_results)) {

  # Set seed for reproducibility
  set.seed(12345)

  cat("\n\nRunning Analyses for", threshold)
  scaled_data <- feature_results[[threshold]]$features$featuresImpZMat %>%
    select(ends_with(feature_selection))

  # Check for and handle NA values
  if (anyNA(scaled_data)) {
    warning(paste("NA values found in scaled data for threshold", threshold))
    next  # Skip this iteration if NA values are found
  }

  # PCA 
  res.pca <- prcomp(scaled_data, scale. = FALSE)
  cutoff_percentage <- 0.8
  pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= cutoff_percentage))
  pca_scores <- as.data.frame(res.pca$x[,1:pca_cutoff])

  # kmeans with NbClust to find the best number of clusters
  kmeans_performance <- NbClust::NbClust(data = pca_scores, diss = NULL, distance = "euclidean",
                                  min.nc = 2, max.nc = 15, method = "kmeans", index = "all")
  kmeans_best_k <- kmeans_performance$Best.nc["Number_clusters", ] %>%
    as.numeric() %>%
    table() %>%
    which.max() %>%
    names() %>%
    as.numeric()
  kmeans_results <- kmeans(pca_scores, centers = kmeans_best_k, nstart = 100)

  analysis_results[[threshold]] <- list(
    pca_scores = pca_scores,
    kmeans_results = kmeans_results
  )
}

# Save so that we do not have to run this compute intensive step again
saveRDS(analysis_results, "data/missingness_thresholds/03_analysis/analysis_results_all.rds")
cat("All analysis results have been saved in", base_dir, ". \n\n*Done!*")
```
:::

### Compare Results

To compare the results from the different missingness thresholds we look at the optimal number of clusters ($k$) as well as the similarity of the extracted cluster solutions.

We find that with almost all missingness thresholds the optimal number of k-means clusters is 2. The only exceptions are at the thresholds of 0% = 3, 10% = 5, and 25% = 3 (i.e., 0% threshold = 100% missingness allowed for each).

```{r}
#| label: Compare Cluster Assignments
#| warning: false
#| error: false

# Initialize an empty list to store the data frames for each threshold
threshold_dfs <- list()

for (threshold in names(analysis_results)) {

  # extract unique ids from data reduction
  red_ids <- combined_data[[threshold]] %>%
    mutate(uid = paste(study, PID, sep = "_")) %>%
    select(ID, uid) %>% 
    distinct()

  # Extract the data frame for the current threshold
  cluster_assignments <- feature_results[[threshold]]$features$features %>%
    select(ID) %>%
    mutate(!!paste("cluster", threshold, sep = "_") := analysis_results[[threshold]]$kmeans_results$cluster)

  # join cluster assignments and uids
  df <- red_ids %>%
    full_join(cluster_assignments, by = "ID") %>%
    select(-ID)

  # Store the data frame in the list
  threshold_dfs[[threshold]] <- df
}

# Combine all the data frames into a single data frame
threshold_cluster_assignments <- reduce(threshold_dfs, full_join, by = "uid")

# Apply max function to each column, excluding 'uid'
max_values <- apply(threshold_cluster_assignments %>% select(-uid), 2, max, na.rm = TRUE)

# Convert the max values into a data frame
data.frame(k = max_values) %>%
  rownames_to_column("threshold") %>%
  mutate(threshold = paste(gsub("cluster_", "", threshold), "%", sep = ""))%>%
  kbl(.,
      escape = FALSE,
      booktabs = TRUE,
      label = "optimal_k_threshold",
      format = "html",
      digits = 2,
      linesep = "",
      align = c("c", "c"),
      col.names = c("Missingeness Threshold", "Optimal k"),
      caption = "Optimal K based on Missingess Approach") %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  )

```

We then compare the clustering results obtained at different thresholds using the Adjusted Rand Index (ARI), which quantifies the similarity between two data clustering assignments. By calculating the ARI for every pair of threshold-based clusterings, we can assess how consistent the cluster assignments are across varying thresholds, even when the number of clusters or their composition changes. This comparison helps us understand the stability of our clustering solution and identify which thresholds yield similar or distinct grouping patterns, providing valuable insights into the robustness of our clustering approach against parameter variations. The ARI is normalized such that -1 indicates perfect disagreement, 0 indicates random (or chance) clustering, and 1 indicates perfect agreement. As a rule of thumb, let us consider that scores close to 0 or negative indicate bad or no similarity (worse than chance), values between 0 and 0.5 reflect reasonable to moderate similarity, scores above 0.5 suggest good similarity, and values close to 1 represent excellent similarity, indicating nearly identical cluster assignments [@hubert1985].

```{r}
#| label: ARI comparison
#| warning: false
#| error: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "ARI Comparison Plot"

# Calculate ARI values
library(mclust)  # For ARI calculation

# Initialize a matrix to store the ARI values
n <- length(threshold_dfs)
comparison_matrix <- matrix(NA, n, n, dimnames = list(names(threshold_dfs), names(threshold_dfs)))

# Double loop to compare each pair of thresholds
for (i in 1:(n - 1)) {
  for (j in (i + 1):n) {
    # Extract the cluster assignments for the two thresholds being compared
    threshold_i <- threshold_dfs[[i]]
    threshold_j <- threshold_dfs[[j]]

    # Merge the two data frames on uid to find common participants
    merged_df <- merge(threshold_i, threshold_j, by = "uid", all = FALSE)  # all = FALSE ensures only common uids

    # Ensure there are no NAs before calculating ARI
    valid_entries <- complete.cases(merged_df)
    merged_df <- merged_df[valid_entries,]

    # Check if there are enough common participants to compare
    if (nrow(merged_df) > 1) {
      # Calculate ARI for common participants
      ari_value <- adjustedRandIndex(merged_df[, paste("cluster", names(threshold_dfs)[i], sep = "_")],
                                     merged_df[, paste("cluster", names(threshold_dfs)[j], sep = "_")])
    } else {
      # Not enough data to compare
      ari_value <- NA
    }

    # Store the ARI value in the matrix
    comparison_matrix[i, j] <- ari_value
    comparison_matrix[j, i] <- ari_value  # ARI is symmetric
  }
}

# calculate the average ARI for description
avg_ari <- mean(comparison_matrix[upper.tri(comparison_matrix)]) %>%
  format(., digits = 3, nsmall = 3)
remove <- c("0", "10", "25")
avg_ari_k2 <- comparison_matrix[!rownames(comparison_matrix) %in% remove, !colnames(comparison_matrix) %in% remove] %>%
  .[upper.tri(.)] %>% 
  mean() %>%
  format(., digits = 3, nsmall = 3)

# Convert the matrix to a long format suitable for ggplot
comparison_matrix_long <- comparison_matrix %>%
  as.data.frame() %>%
  rownames_to_column("row") %>%
  mutate(row = as.numeric(row)) %>%
  gather(key = "column", value = "value", -row) %>%
  mutate(
    column = as.numeric(column),
    value = as.numeric(format(value, digits = 3, nsmall = 3))
  )

# Create the plot
ggplot(comparison_matrix_long, aes(x = row, y = column, fill = value)) +
  geom_tile(data = subset(comparison_matrix_long, row > column), aes(fill = value), color = "white") +
  scale_fill_gradient(low = "blue", high = "red", na.value = "white", limits = range(comparison_matrix_long$value, na.rm = TRUE), name = "ARI value") +
  geom_text(data = subset(comparison_matrix_long, row < column), aes(label = sprintf("%.3f", as.numeric(value))), color = "black") +
  scale_y_reverse(breaks = seq(0, 55, 5), labels = seq(0, 55, 5)) +
  scale_x_continuous(breaks = seq(0, 55, 5), labels = seq(0, 55, 5)) +
  coord_fixed() +
  theme_minimal() +
  theme(axis.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_line(color = "grey", size = 0.5),
        panel.border = element_rect(color = "black", fill = NA))

```

We find that with the exception of when the optimal solution is larger than 2 the cluster similarity is extremely high (see threshold = 10% as an example). In short, the number of clusters seems to be a much bigger decision than the actual cluster assignment itself. At the same time however, the ARIs seem to suggest that the differences in features extracted based on the different missingness thresholds is negligible (as long as the number of clusters is the same). In fact the mean ARI is `r avg_ari` (and the mean ARI for all $k=2$ is `r avg_ari_k2`).

## Simplified Model

One question that remains is whether a much simpler model with only central tendency and variance would perform similarly well and would result in a similar separation of the clusters. We here adjust the code we used for the main analysis (just on a smaller feature set). Wherever the code is functionally unchanged from the main analysis we do not rehash it here. If you would like to inspect the full code, you can do this via our OSF and GitHub repositories (referenced in the [about page](about.qmd)).

```{r}
#| label: load data
#| include: false

rm(list = ls())

source("scripts/load_data.R")
source("scripts/theme_Publication.R")

kmeans_final <- readRDS("data/kmeans_final.Rds")
pca_scores <- readRDS(file = "data/pca_scores.Rds")

```

### Prepare Data

We follow the same data preparation steps as during the main analysis.

```{r}
#| label: pca data preparation

feature_selection_full <- c(
  "ar01",
  "edf",
  "lin",
  "mac",
  "mad",
  "median"
)

feature_selection <- c(
  "mad",
  "median"
)

scaled_data_simple <- featFullImp$featuresImpZMat %>% select(ends_with(feature_selection)) 
z_data_simple <- featFullImp$featuresImpZ %>% select("ID", ends_with(feature_selection))

z_data <- featFullImp$featuresImpZ %>% select("ID", ends_with(feature_selection_full))

```

### Run Analyses

Also our analysis steps are identical to the main analysis. We find that in the reduced feature set, $k=2$ are the most optimal solutions. We extract the k-means cluster assignments for this model.

```{r}
#| label: run analyses
#| output: false

res.pca_simple <- prcomp(scaled_data_simple, scale. = FALSE)
cutoff_percentage <- 0.8
pca_cutoff_simple <- min(which(cumsum(summary(res.pca_simple)[["importance"]]['Proportion of Variance', ]) >= cutoff_percentage))
pca_scores_simple <- as.data.frame(res.pca_simple$x[, 1:pca_cutoff_simple])

# kmeans with NbClust to find the best number of clusters
kmeans_performance_simple <-
  NbClust::NbClust(
    data = pca_scores_simple,
    diss = NULL,
    distance = "euclidean",
    min.nc = 2,
    max.nc = 15,
    method = "kmeans",
    index = "all"
  )
kmeans_best_k_simple <-
  kmeans_performance_simple$Best.nc["Number_clusters",] %>%
  as.numeric() %>%
  table() %>%
  which.max() %>%
  names() %>%
  as.numeric()

set.seed(42)
kmeans_results_simple <- kmeans(pca_scores_simple, centers = kmeans_best_k_simple, nstart = 100)

```

### Compare Results

We begin by comparing the simpler model to the model from the main analysis in terms of performance and similarity.

#### Performance

To assess the performance of the clustering algorithm with the parameters you have selected, we display a number performance metrics. These metrics can be divided into 'general performance' (looking at separation and cohesion) and 'cluster structure' (distribution of points across clusters and the compactness of clusters). We have selected three common metrics for each of the categories. For the general performance we inspect:

::: {.callout-note collapse="true"}
## Average Silhouette Width

This metric provides a measure of how close each point in one cluster is to points in the neighboring clusters. Higher silhouette width indicates better separation and cohesion.
:::

::: {.callout-note collapse="true"}
## Calinski and Harabasz Index

This index is a ratio of between-cluster variance to within-cluster variance. Higher values generally indicate clusters are well separated and well defined.
:::

::: {.callout-note collapse="true"}
## Dunn Index

Measures the ratio of the smallest distance between observations not in the same cluster to the largest intra-cluster distance. Higher values indicate better clustering by maximizing inter-cluster distances while minimizing intra-cluster distances.
:::

For the cluster structure, we display:

::: {.callout-note collapse="true"}
## Entropy of Distribution of Cluster Memberships

Provides a measure of how evenly data points are spread across the clusters. Lower entropy indicates a more definitive classification of points into clusters.
:::

::: {.callout-note collapse="true"}
## Minimum Cluster Size

Indicates the size of the smallest cluster. This is useful for identifying if the clustering algorithm is producing any very small clusters, which might be outliers or noise.
:::

::: {.callout-note collapse="true"}
## Average Distance Within Clusters

This metric offers insight into the compactness of the clusters. Lower values indicate that points within a cluster are closer to each other, suggesting better cluster cohesion.
:::

```{r}
#| label: compare simplified performance
#| echo: false

kmeans_performance_simplified <- list(
  og <- fpc::cluster.stats(
      d = dist(pca_scores),
      clustering = kmeans_final$cluster,
      G2 = TRUE,
      G3 = TRUE,
      aggregateonly = TRUE
    ) %>% unlist,
  simple_kopt <- fpc::cluster.stats(
      d = dist(pca_scores_simple),
      clustering = kmeans_results_simple$cluster,
      G2 = TRUE,
      G3 = TRUE,
      aggregateonly = TRUE
    ) %>% unlist
)

df_kmeans_performance_simplified <-
  data.frame(approach = c("main", "simplified"), Reduce(rbind, kmeans_performance_simplified))

selected_metrics_ids <- c("approach", "n", "cluster.number", "avg.silwidth", "ch", "dunn", "entropy", "min.cluster.size", "average.within")
metrics_selected <- df_kmeans_performance_simplified %>%
  select(all_of(selected_metrics_ids))

metrics_names <- c(
  "approach" = "Analysis Label",
  "n" = "number of cases", 
  "cluster.number" = "number of points", 
  "min.cluster.size" = "size of smallest cluster", 
  "noisen" = "number of noise points", # Not relevant for kmeans 
  "average.between" = "average distance between clusters", 
  "average.within" = "average distance within clusters", # (reweighted so that every observation, rather than every distance, has the same weight)
  "max.diameter" = "maximum cluster diameter",
  "min.separation" = "minimum cluster separation", 
  "ave.within.cluster.ss" = "within clusters sum of squares", # generalisation
  "avg.silwidth" = "average silhouette width", 
  "g2" = "Goodman Kruskal's Gamma coefficient", # See Milligan and Cooper (1985), Gordon (1999, p. 62). 
  "g3" = "G3 coefficient", # See Gordon (1999, p. 62)
  "pearsongamma" = "Normalized gamma", # correlation between distances and a 0-1-vector where 0 means same cluster, 1 means different clusters. see Halkidi et al. (2001)
  "dunn" = "Dunn index", # minimum separation / maximum diameter, Dunn index, see Halkidi et al. (2002).
  "dunn2" = "Dunn index 2", # minimum average dissimilarity between two cluster / maximum average within cluster dissimilarity, another version of the family of Dunn indexes
  "entropy" = "entropy of distribution of cluster memberships", # see Meila(2007)
  "wb.ratio" = "average within / average between", 
  "ch" = "Calinski and Harabasz index", # (Calinski and Harabasz 1974, optimal in Milligan and Cooper 1985; generalised for dissimilarites in Hennig and Liao 2013)
  "widestgap" = "widest within-cluster gap", 
  "sindex" = "separation index"
)

metrics_selected %>%
  kbl(.,
      format = "html",  # Ensure format is explicitly set to "html"
      escape = FALSE,
      booktabs = TRUE,  # Use booktabs for better line quality
      digits = 2,
      align = "c",
      row.names = FALSE,
      col.names = metrics_names[selected_metrics_ids]
    ) %>%
    kable_styling(bootstrap_options = c("condensed"), full_width = FALSE, html_font = "Cambria") %>%
    add_header_above(c(
      " " = 3,
      "General Clustering Performance" = 3,
      "Cluster Structure Insight" = 3
    ))

```

We find that both models perform well across the performance metrics.

#### Similarity

Similar to the missingness threshold models above, we also look at the Adjusted Rand Index (ARI) to compare the similarity of the models.

```{r}
#| label: simplified similarity 
#| echo: false
#| warning: false
#| error: false

simplified_kmeans_df <- data.frame(
  cluster_og = kmeans_final$cluster,
  cluster_simple = kmeans_results_simple$cluster
)

ari_value <- adjustedRandIndex(simplified_kmeans_df[, 1], simplified_kmeans_df[,2])

```

We find a relatively high Adjusted Rand Index: $ARI=$ `r format(ari_value, digits=3, nsmall=3)` --- indicating that the simplified model separates the two clusters in a similar manner.

#### Full TS Feature Comparison

```{r}
#| label: Extract cluster assignments and summarize
#| message: false
#| warning: false  
#| echo: false

featCluster_simple <- data.frame(ID = as.numeric(names(kmeans_results_simple$cluster)), 
                            cluster = kmeans_results_simple$cluster) %>%
  left_join(z_data, by = "ID") %>%
  left_join(
    featData %>% 
      select(ID, EvDayDiscr.post) %>% 
      distinct(ID, EvDayDiscr.post, .keep_all = TRUE) %>%
      mutate(EvDayDiscr.post_z = scale(EvDayDiscr.post, scale = TRUE)) %>%
      replace_na(list(EvDayDiscr.post_z = NA)) %>%
      select(
        ID,
        EvDayDiscr.post_z
      ), 
    by = "ID"
  ) %>%
  select(-ID) %>%
  group_by(cluster) %>%
  summarise(across(everything(), list(mean = mean), na.rm=TRUE)) %>%
  ungroup %>% 
  select(-starts_with("ID")) %>%
  pivot_longer(
    cols = -c(cluster),
    names_to = c("variable", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>% 
  mutate(
    nam = variable,
    variable = gsub("\\_.*", "", nam),
    feature = gsub("^.*?\\_", "", nam)
  ) %>% 
  mutate(feature = ifelse(feature == "z", "mean", feature)) %>%
  left_join(var_meta %>% select(name, label), by = c("variable" = "name")) %>%
  select(
    cluster,
    nam,
    variable,
    label,
    feature,
    everything()
  )
```

We can then plot the average features and for each variable and flipping the axes allows us to assess the same data once with a focus on the variable (see [@fig-clusterFeatVarComb_simple] A) and once with a focus on the features (see [@fig-clusterFeatVarComb_simple] B).

```{r}
#| label: fig-clusterFeatVarComb_simple
#| message: false
#| warning: false
#| echo: false
#| fig-width: 12
#| fig-height: 18
#| fig-cap: Cluster comparison by feature and variable. Note that n and discrimination are out of cluster variables.

# clusterVariableGrid
clusterVariableGrid_simple <- featCluster_simple %>% 
  filter(variable != "EvDayDiscr.post",
         feature != "n") %>%
  ggplot(., aes(x = feature, y = mean, group = cluster, color = as.factor(cluster))) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
  coord_flip() +
  labs(
    x = "Features",
    y = "Standardized Mean",
    color = "Cluster"
  ) +
  facet_wrap(~ label) +
  theme_Publication()

# clusterFeatureGrid
clusterFeatureGrid_simple <- featCluster_simple %>% 
  ggplot(., aes(x = reorder(label, -as.numeric(label == "Post: Discrimination")), y = mean, group = cluster, color = as.factor(cluster))) +
  geom_point() +
  geom_line() +
  #geom_ribbon(aes(ymin = mean-se, ymax = mean+se), alpha = 0.2) +
  #geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=.1, position=position_dodge(0.1), size=0.5, alpha=0.4) +
  scale_colour_manual(values = c("#384B6B", "#E2892B")) + 
  coord_flip() +
  labs(
    x = "Variable",
    y = "Standardized Mean",
    color = "Cluster"
  ) +
  facet_wrap(~ factor(feature, levels = unique(featCluster_simple$feature)), ncol=4) +
  theme_Publication()

ggpubr::ggarrange(
  clusterVariableGrid_simple + theme(axis.title.x = element_blank()),
  clusterFeatureGrid_simple,
  ncol = 1,
  labels = c("(A) Variable Focus", "(B) Feature Focus"),
  common.legend = TRUE,
  legend = "bottom",
  align = "v"
) 
```

As was already indicated by the $ARI$, the clusters differences look very similar to that of the main analysis, with slightly larger Median and MAD differences but slightly smaller differences in most of the other features (across the variables).

## Alternative Models

As a final resource, we have developed a web application offers an interactive dashboard for exploring and understanding time series feature clustering based on the data set of the main analyses. This web application is designed to facilitate the exploration of different clustering algorithms and dimensionality reduction techniques. As the web application accompanies this article, we focus on the time series features from the main analysis data set. The web application is designed to be relevant for several target groups, including researchers, students, or data enthusiasts. The tsFeatureClustR Web App offers a hands-on experience that aims to enhance the understanding of time series data clustering and its underlying processes.

We allow users to try out different combinations of dimensionality reduction algorithms (PCA, t-SNE, Autoencoder, UMAP) and clustering algorithms (k-means, DBSCAN, Hierarchical agglomerative clustering). For each of the methods, we have developed an interface that lets users explore the key parameter settings of the algorithms. To provide an introduction to the diversity of possible combinations, we have pre-calculated the performance of the algorithm combinations for common parameter values (showing users a comparison of up to 18,557 model combinations). On a second panel, readers then have the chance to explore the cluster results based on their own interaction with the different parameters.

If you would like to explore the different alternative models, you can follow the following link:

<div class="d-flex justify-content-center">
  <a class="btn btn-primary btn-lg" href="https://jannis-kreienkamp.shinyapps.io/tsFeatureClustR/" target="_blank" role="button">To the Web App</a>
</div>

## References

::: {#refs}
:::

### 2024-11-11 tutorial/scripts/cleanM.R

@@ -22,6 +22,7 @@
 #' @ImportFrom dplyr %>%
 
 cleanM <- function(M, c = 55) {
+  # M <- as.data.frame(dtAvailability)
   
   check <- list()
   rowNamesOut <- c()
@@ -35,6 +36,7 @@ cleanM <- function(M, c = 55) {
     
     # Determine which row/column has the lowest mean and should be removed
     rm <- which.min(rcMeans) %>% names
+    # cat(rm)
     rc <- ifelse(startsWith(rm, "PP_"), "row", "col")
     
     check[[i]] <- rcMeans

### 2024-11-13 tutorial/06-validation-analyses.qmd
de-anonymized for publication

@@
 <div class="d-flex justify-content-center">
-  <a class="btn btn-primary btn-lg" href="https://jannis-kreienkamp.shinyapps.io/tsFeatureClustR/" target="_blank" role="button">To the Web App</a>
+  <a class="btn btn-primary btn-lg" href="https://tsfeatureclustr.shinyapps.io/webapp/" target="_blank" role="button">To the Web App</a>
 </div>

### 2024-11-13 tutorial/about.qmd
de-anonymized for publication

--- 2023-08-07 tutorial/about.qmd
@@
 We specifically provide a minial reproducible dataset that includes all variables necessary to perform the analysis, while still protecting the privacy and confidentiality of the participants.
-All data is available as part of [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].
+All data is available as part of [our Open Science repository](https://doi.org/10.17605/OSF.IO/J8DZV) [@Kreienkamp-clust-osf].
 
@@
 
-The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com) [@KreienkampMasked-clust-github].
-A minimal version of the code as well as an archive file are available in [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].
+The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com/JannisCodes/migration-trajectories) [@Kreienkamp-clust-github].
+A minimal version of the code as well as an archive file are available in [our Open Science repository](https://doi.org/10.17605/OSF.IO/J8DZV) [@Kreienkamp-clust-osf].
 For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package.
-This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz) [@KreienkampBlinded2023-tsFeatureExtracR].
+This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://tsfeatureclustr.com/assets/tsFeatureExtracR_0.1.0.tar.gz) [@Kreienkamp2023-tsFeatureExtracR].
 

### 2024-11-13 tutorial/index.qmd
config updated
de-anonymized for publication

--- 2023-08-07 tutorial/index.qmd
@@
 ---
-title: "Illustration: Feature Based Time Series Clustering in Psychology [tsFeatureExtracR]"
+title: "Illustration: Feature Based Time Series Clustering in Psychology [tsFeatureClustR]"
 author:
+  # - name: Masked for Peer Review
+  # - name: ██████ ██████████
+  #   orcid: 0000-0000-0000-0000
+  # - name: ██████████ ████████
+  #   orcid: 0000-0000-0000-0000
+  # - name: ███ ██████
+  #   orcid: 0000-0000-0000-0000
+  # - name: ███ ███████
+  #   orcid: 0000-0000-0000-0000
+  # - name: █████ ██ █████
+  #   orcid: 0000-0000-0000-0000
+  # - name: █████ ██ █████████
+  #   orcid: 0000-0000-0000-0000
 - name: Jannis Kreienkamp
   orcid: 0000-0002-1831-5604
-    email: j.kreienkamp@rug.nl
+  email: jannis@thedataflowcompany.com
   corresponding: true
@@
 5. [Evaluation](05-evaluation.qmd)
-6. [About](about.qmd)
+6. [Validation Analyses](06-validation-analyses.qmd)
+7. [About](about.qmd)
 
